### 2021-02-18 paper/linkagecluster_bk_paper.Rmd
added paper draft

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
header-includes: \usepackage{hyperref}
linestretch: 1
link-citations: yes
linkcolor: blue
fontfamily: mathpazo
fontsize: 12pt
---
\begin{flushright} 
	\end{flushright}
	\begin{center} \textbf{Linkage Cluster Paper Draft}

	Brian Kundinger

	\end{center}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Related Work

Sadinle (2017) proposed a Bayesian method for bipartite record linkage that provides a foundation for much of our proposed method. In order to ensure one-to-one matching, his Gibbs sampler conditions on the linkage status of all other records when making each linkage assignment. While this elegantly ensures transitivity, the probability distribution of the linkage vector $\mathbf{Z}$ is somewhat complicated, and does not easily allow for inclusion of additional structure among records. Our method however, motivated by recent decouple/recouple methods, initially relaxes the one-to-one matching assumption in order to incorporate linkage information differing by subgroups.

There has been some work done in the literature on conducting record linkage in situations where feild accuracy differs by subgroups of the data, but these are hampered by technical issues that make them difficult to use in practice. Larson (2005) proposed Bayesian heirarchical record linkage model that explicitly modelled the differences in parameters across blocks of the data. In relying on blocking however, he assumes that the blocking variable is error-free, when in practice such a procedure could result in many missed linkages. In contrast, our proposed method induces high probabilities of matching within a linkage cluster while still allowing the possibility to match across clusters (when the clustering field is recorded with error. Hu et al (2015) proposed a method that used a logistic regression submodel within a broader record linkage procedure to account for "expected changes" in the record of the underlying entity based on subgroup membership. This method however specified for use in temporal settings when the true attributes of the record change with time, and it is unclear how such a method would be implemented when records are collected syncronously with error, or when time information is unavailable. Additionally, the method is semi-supervised, requiring training data to implement the logistic regression, and is that not well suited for record linkage in practice. In contrast, the proposed method is fully unsupervised and is shown to not require extensive tuning of prior distributions, allowing for ease of implementation in a variety of settings.

## Notation and Assumptions

Our notation and assumptions closely follow that of Sadinle (2017). Denote two files as $\mathbf{X}_A$ and $\mathbf{X}_B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $\mathbf{X}_A$ and $j \in \{1, \ldots, n_B\}$ in $\mathbf{X}_B$. Without loss of generality, label the files such that $n_A \geq n_B$. We also assume there are no duplicates within files, only across. For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each takes $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating level agreement between the two records on a specified field.

Following the Fellegi Sunter framework, we define $m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)$ to be the probability of observing agreement level $l$ in field $f$ for records $i$ and $j$ given that the records are a match, and similarly define $u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)$, for non-matches. We denote $\lambda$ to be the (marginal) probability that some record $j \in \mathbf{X}_B$ has a match in $\mathbf{X}_A$.

Lastly, we adopt from Sadinle 2017 the matching labelling $\mathbf{Z} = (Z_1, \ldots, Z_{n_A})$, defined as
$$Z_j=\begin{cases} 
    i,  & \text{if records } i\in \mathbf{X}_A \text{ and } j\in \mathbf{X}_B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in \mathbf{X}_B \text{ does not have a match in file} \mathbf{X}_B; \\
\end{cases}$$

Wherever possible, we reserve superscripts for denoting field, level, and linkage cluster, and reserve subscripts for record indices. For example, $\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})$ is the probability distribution governing field $f$ for matching records, and $\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)$ is product of the relevant of the appropriate $\mathbf{m}$ parameters for record pair $(i,j)$. We hope that these conventions avoid overloaded notation in the likelihood and subsequent derivations.

# Model Specification

Additionally, our model differs from that of Sadinle 2017 through its explicit dependence on a beta random variable $\lambda$ that models the rate of matching across records. Sadinle marginalizes over such a random variable in his derivations of  the "beta prior for bipartite matching," but here we provide derivations without marginalizing in order to specify differing rates of matching for different linkage clusters.

## Prior Distributions and Likelihood

For fields $f \in \{1, \ldots, F\}$ and levels $l\in \{1, \ldots, L_f\}$ we adopt the following likelihood and prior distributions. Note that likelihood is related to that of Sadinle (2017), but includes dependence on the $\lambda$ random variable. It also makes it explict that units of observation under this model are the $n_B$ records in $\mathbf{X}_B$, a crucial distinction when compared to the likelihood over the $n_A \times n_B$ record pairs in the original Fellegi Sunter model.
<!-- $$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}}(1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}}\right]$$ -->

$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}} (1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]$$

$$\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})$$
$$\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})$$
$$Z_j | \lambda =
\begin{cases} 
    \frac{1}{n_A}\lambda  & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}$$

$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) $$
The prior for $Z_j$ has equal probability of matching to all records $i\in \mathbf{X}_A$, and non-matching probability governed by $\lambda$. Observe two reasonable choices for $p(Z, \lambda)$:

If $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$, then the prior for $Z$, integrated over $\lambda$, becomes

$$=\begin{cases} 
    1  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}$$
Thus this is a uniform prior on the labellings of $Z$. A second useful prior would be $\lambda \sim \text{Beta}(1, 1)$, which yields
$$=\begin{cases} 
    1  & z_j \leq n_A; \\
    n_A &  z_j  = n_A + 1 \\
\end{cases}$$
Here, a match and a non-match are equaly likely, and among the matches, all records are likely to be a match. There is little difference between the the two priors in practice, as the data should overpower the prior.

## Posterior Sampling

The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have 
$$\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))$$
$$\mathbf{u^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))$$
where $\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}$ and $\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}$.

Since $\lambda$ provides the probability that a record in $\mathbf{X}_B$ has match in $\mathbf{X}_A$, it has a full conditional distribution

$$\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})$$
where $D = \sum_1^{n_B} (Z_j < n_A + 1)$ is the number of records identified as matches in the previous Gibbs iteration.

Lastly, we sample each $Z_j$ according to these posterior distributions. To do so, note that as a function of $Z_j$, the likelihood is a discrete distibution with probabilities proportional to

$$p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}$$
$$\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)$$

$$\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}$$
where $w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$. The interested reader should note that these are precisely the likelihood ratios used in the Fellegi-Sunter model to classify matches and non-matches, and we therefore refer to $w_{ij}$ as the *Fellegi Sunter weights*.

The full conditional for $Z_j$ is thus given by
$$p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}$$
$$\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)$$
$$= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} $$
$$ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
\begin{cases} 
    \frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}$$

Here, one should note that if we choose to integrate over the posterior of $\lambda$ and rearrange terms, we produce the following sampling mechanism:

$$Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
\end{cases}$$

The relationship between this sampling mechanism and that of Sadinle 2017 warrants discussion.

## Bayes Estimate

Sadinle 2017

## Decouple/Recouple

One difficulty in record linkage is conducting linkage while taking appropriate dependencies into account. In the context of this paper in which we assume that each file is recorded without duplicates, this means our linkage procedure should respect one-to-one matching. The posterior sampling of $\mathbf{Z}$ outlined above resolves half of this issue; since we only sample one potential match for record $j$ in each Gibbs iteration, it is impossible to match $j$ to two records at once. This in itself is an improvement over the classic Fellegi-Sunter method that labels record pairs as matches or non matches independently of all other pairs. However, our posterior sampling of $\mathbf{Z}$ does allow for the possibility that $Z_j = i = Z_{j'}$ for some $j' \neq j$, which cleary violates the assumption that $\mathbf{X}_B$ is recorded without duplicates. Sadinle 2017 resolves this issue by adapting the set of candidate matches for every $Z_j$, but we forgo this rigidity and instead "decouple" the linkage vector $\mathbf{Z}$ for the computational advantantages of sampling $Z_j$ independently. In our formulation, such illegal matches occasionally occur within individual Gibbs iterations due to simple randomness, but they do not occur consistently enough between the same records to measurably influence the eventual Bayes estimate of the linkage structure.

However, if we want to ensure the theoretic purity of each MCMC sample of the linkage structure, we can "recouple" $\mathbf{Z}$ by simply resolving these illegal matchings. Explictly, define clusters $C(i) := \{j | Z_j = i\}$, and define all clusters such that $|C(i)| > 1$ as *conflict clusters*.  Within each cluster, resample the candidate matches according to their Fellegi-Sunter weights. We show this step to be computationally inexpensive and to be a reasonable replication of more organic one-to-one matching. This procedure would be most useful in cases where there are many "near matches" across files, but in most cases, we find this this step to be unnecessary.

## Linkage cluster model

The primary contribution of this paper is the establishment of linkage clusters, which allow the reliability parameters $\mathbf{m}^f$, the discernment parameters $\mathbf{u}^f$, and also the linkage probability parameter $\lambda$ to vary according to certain attributes in the record pairs. We shall see that this set up upweights linkage within clusters and downweights linkage across clusters, thus using attribute information in way not yet explored in the literture.

Assign each record in both files to a linkage cluster $S_k \in \{S_1, \ldots, S_K\}$ based on some attribute in the data. We see that the comparison vector

<!-- The classic Fellegi Sunter model operates through weights $w_{ij} = \frac{P(\boldsymbol{\gamma}_{ij}| \text{Records match})}{P(\boldsymbol{\gamma}_{ij}| \text{Records do not match})}$. In our notation, these weights are given as follows: -->
<!-- $$w_{ij} = \frac{P(\boldsymbol{\gamma}_{ij}| Z_j = i)}{P(\boldsymbol{\gamma}_{ij}| Z_j \neq i)} = \frac{\prod_{f=1}^F \prod_{l=1}^{L_f} m_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}{\prod_{f=1}^F \prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}$$ -->

<!-- $$p(\Gamma_{.j}|Z_{j}) = \frac{\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}{\sum_{z_j = i}^{n_A}\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A +1}(1-\lambda)\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}$$ -->
<!-- We can divide the numerator and denominator by $\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}$ to obtain -->

<!-- $$p(\Gamma_{.j}|Z_{j}) =\frac{\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{\mathbf{1}_{z_j = i} \times \mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)}{\sum_{z_j = 1}^{n_A}\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{\mathbf{1}_{z_j = i} \times \mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)}$$ -->

<!-- Our model decouples the draws of the elements of the linkage structure, sampling each $Z_j$ independently from all other $Z_{j'}$. This decoupling allows the incorporation of additional information on each record $j$, thus enabling the linkage cluster formulation proposed in Section XXXX. We clearly admit that this does allow the possibility that $Z_j = i = Z_{j'}$ for some $i\in \mathbf{X}_A$, breaking one-to-one matching requirements in individual Gibbs iterations. However, if our model assumptions are correct and there are no duplicates within files, these sorts of illegal matches are rare and evenly distrubted across records, and thus do not seriously influence our eventual Bayes estimate of the linkage structure. Additionally, these illegal matches can easily be resolved in a post-hoc conflict resolution step, or "recoupling", as described in Section XXXX. -->

### 2021-02-18 paper/linkagecluster_bk_paper.tex
added paper draft

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{hyperref}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Linkage Cluster Paper Draft}

    Brian Kundinger

    \end{center}

\hypertarget{related-work}{%
\subsection{Related Work}\label{related-work}}

Sadinle (2017) proposed a Bayesian method for bipartite record linkage
that provides a foundation for much of our proposed method. In order to
ensure one-to-one matching, his Gibbs sampler conditions on the linkage
status of all other records when making each linkage assignment. While
this elegantly ensures transitivity, the probability distribution of the
linkage vector \(\mathbf{Z}\) is somewhat complicated, and does not
easily allow for inclusion of additional structure among records. Our
method however, motivated by recent decouple/recouple methods, initially
relaxes the one-to-one matching assumption in order to incorporate
linkage information differing by subgroups.

There has been some work done in the literature on conducting record
linkage in situations where feild accuracy differs by subgroups of the
data, but these are hampered by technical issues that make them
difficult to use in practice. Larson (2005) proposed Bayesian
heirarchical record linkage model that explicitly modelled the
differences in parameters across blocks of the data. In relying on
blocking however, he assumes that the blocking variable is error-free,
when in practice such a procedure could result in many missed linkages.
In contrast, our proposed method induces high probabilities of matching
within a linkage cluster while still allowing the possibility to match
across clusters (when the clustering field is recorded with error. Hu et
al (2015) proposed a method that used a logistic regression submodel
within a broader record linkage procedure to account for ``expected
changes'' in the record of the underlying entity based on subgroup
membership. This method however specified for use in temporal settings
when the true attributes of the record change with time, and it is
unclear how such a method would be implemented when records are
collected syncronously with error, or when time information is
unavailable. Additionally, the method is semi-supervised, requiring
training data to implement the logistic regression, and is that not well
suited for record linkage in practice. In contrast, the proposed method
is fully unsupervised and is shown to not require extensive tuning of
prior distributions, allowing for ease of implementation in a variety of
settings.

\hypertarget{notation-and-assumptions}{%
\subsection{Notation and Assumptions}\label{notation-and-assumptions}}

Our notation and assumptions closely follow that of Sadinle (2017).
Denote two files as \(\mathbf{X}_A\) and \(\mathbf{X}_B\), with \(n_A\)
and \(n_B\) records respectively, and with records indexed as
\(i \in \{1, \ldots, n_A\}\) in \(\mathbf{X}_A\) and
\(j \in \{1, \ldots, n_B\}\) in \(\mathbf{X}_B\). Without loss of
generality, label the files such that \(n_A \geq n_B\). We also assume
there are no duplicates within files, only across. For each record pair
under consideration, we generate a comparison vector
\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
where \(F\) is the number of fields used in the linkage and each takes
\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
indicating level agreement between the two records on a specified field.

Following the Fellegi Sunter framework, we define
\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
observing agreement level \(l\) in field \(f\) for records \(i\) and
\(j\) given that the records are a match, and similarly define
\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We
denote \(\lambda\) to be the (marginal) probability that some record
\(j \in \mathbf{X}_B\) has a match in \(\mathbf{X}_A\).

Lastly, we adopt from Sadinle 2017 the matching labelling
\(\mathbf{Z} = (Z_1, \ldots, Z_{n_A})\), defined as \[Z_j=\begin{cases} 
    i,  & \text{if records } i\in \mathbf{X}_A \text{ and } j\in \mathbf{X}_B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in \mathbf{X}_B \text{ does not have a match in file} \mathbf{X}_B; \\
\end{cases}\]

Wherever possible, we reserve superscripts for denoting field, level,
and linkage cluster, and reserve subscripts for record indices. For
example, \(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the
probability distribution governing field \(f\) for matching records, and
\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
is product of the relevant of the appropriate \(\mathbf{m}\) parameters
for record pair \((i,j)\). We hope that these conventions avoid
overloaded notation in the likelihood and subsequent derivations.

\hypertarget{model-specification}{%
\section{Model Specification}\label{model-specification}}

Additionally, our model differs from that of Sadinle 2017 through its
explicit dependence on a beta random variable \(\lambda\) that models
the rate of matching across records. Sadinle marginalizes over such a
random variable in his derivations of the ``beta prior for bipartite
matching,'' but here we provide derivations without marginalizing in
order to specify differing rates of matching for different linkage
clusters.

\hypertarget{prior-distributions-and-likelihood}{%
\subsection{Prior Distributions and
Likelihood}\label{prior-distributions-and-likelihood}}

For fields \(f \in \{1, \ldots, F\}\) and levels
\(l\in \{1, \ldots, L_f\}\) we adopt the following likelihood and prior
distributions. Note that likelihood is related to that of Sadinle
(2017), but includes dependence on the \(\lambda\) random variable. It
also makes it explict that units of observation under this model are the
\(n_B\) records in \(\mathbf{X}_B\), a crucial distinction when compared
to the likelihood over the \(n_A \times n_B\) record pairs in the
original Fellegi Sunter model.

\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}} (1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]\]

\[\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})\]
\[\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})\]
\[Z_j | \lambda =
\begin{cases} 
    \frac{1}{n_A}\lambda  & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}\]

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
prior for \(Z_j\) has equal probability of matching to all records
\(i\in \mathbf{X}_A\), and non-matching probability governed by
\(\lambda\). Observe two reasonable choices for \(p(Z, \lambda)\):

If \(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\), then the prior for
\(Z\), integrated over \(\lambda\), becomes

\[=\begin{cases} 
    1  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}\] Thus this is a uniform prior on the labellings of \(Z\). A
second useful prior would be \(\lambda \sim \text{Beta}(1, 1)\), which
yields \[=\begin{cases} 
    1  & z_j \leq n_A; \\
    n_A &  z_j  = n_A + 1 \\
\end{cases}\] Here, a match and a non-match are equaly likely, and among
the matches, all records are likely to be a match. There is little
difference between the the two priors in practice, as the data should
overpower the prior.

\hypertarget{posterior-sampling}{%
\subsection{Posterior Sampling}\label{posterior-sampling}}

The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
standard multinomial-dirichlet mechanics. Thus we have
\[\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
\[\mathbf{u^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
where
\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
and
\(\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).

Since \(\lambda\) provides the probability that a record in
\(\mathbf{X}_B\) has match in \(\mathbf{X}_A\), it has a full
conditional distribution

\[\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
where \(D = \sum_1^{n_B} (Z_j < n_A + 1)\) is the number of records
identified as matches in the previous Gibbs iteration.

Lastly, we sample each \(Z_j\) according to these posterior
distributions. To do so, note that as a function of \(Z_j\), the
likelihood is a discrete distibution with probabilities proportional to

\[p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}\]
\[\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\]

\[\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}\]
where
\(w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
The interested reader should note that these are precisely the
likelihood ratios used in the Fellegi-Sunter model to classify matches
and non-matches, and we therefore refer to \(w_{ij}\) as the
\emph{Fellegi Sunter weights}.

The full conditional for \(Z_j\) is thus given by
\[p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}\]
\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
\begin{cases} 
    \frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}\]

Here, one should note that if we choose to integrate over the posterior
of \(\lambda\) and rearrange terms, we produce the following sampling
mechanism:

\[Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
\end{cases}\]

The relationship between this sampling mechanism and that of Sadinle
2017 warrants discussion.

\hypertarget{bayes-estimate}{%
\subsection{Bayes Estimate}\label{bayes-estimate}}

Sadinle 2017

\hypertarget{decouplerecouple}{%
\subsection{Decouple/Recouple}\label{decouplerecouple}}

One difficulty in record linkage is conducting linkage while taking
appropriate dependencies into account. In the context of this paper in
which we assume that each file is recorded without duplicates, this
means our linkage procedure should respect one-to-one matching. The
posterior sampling of \(\mathbf{Z}\) outlined above resolves half of
this issue; since we only sample one potential match for record \(j\) in
each Gibbs iteration, it is impossible to match \(j\) to two records at
once. This in itself is an improvement over the classic Fellegi-Sunter
method that labels record pairs as matches or non matches independently
of all other pairs. However, our posterior sampling of \(\mathbf{Z}\)
does allow for the possibility that \(Z_j = i = Z_{j'}\) for some
\(j' \neq j\), which cleary violates the assumption that
\(\mathbf{X}_B\) is recorded without duplicates. Sadinle 2017 resolves
this issue by adapting the set of candidate matches for every \(Z_j\),
but we forgo this rigidity and instead ``decouple'' the linkage vector
\(\mathbf{Z}\) for the computational advantantages of sampling \(Z_j\)
independently. In our formulation, such illegal matches occasionally
occur within individual Gibbs iterations due to simple randomness, but
they do not occur consistently enough between the same records to
measurably influence the eventual Bayes estimate of the linkage
structure.

However, if we want to ensure the theoretic purity of each MCMC sample
of the linkage structure, we can ``recouple'' \(\mathbf{Z}\) by simply
resolving these illegal matchings. Explictly, define clusters
\(C(i) := \{j | Z_j = i\}\), and define all clusters such that
\(|C(i)| > 1\) as \emph{conflict clusters}. Within each cluster,
resample the candidate matches according to their Fellegi-Sunter
weights. We show this step to be computationally inexpensive and to be a
reasonable replication of more organic one-to-one matching. This
procedure would be most useful in cases where there are many ``near
matches'' across files, but in most cases, we find this this step to be
unnecessary.

\hypertarget{linkage-cluster-model}{%
\subsection{Linkage cluster model}\label{linkage-cluster-model}}

The primary contribution of this paper is the establishment of linkage
clusters, which allow the reliability parameters \(\mathbf{m}^f\), the
discernment parameters \(\mathbf{u}^f\), and also the linkage
probability parameter \(\lambda\) to vary according to certain
attributes in the record pairs. We shall see that this set up upweights
linkage within clusters and downweights linkage across clusters, thus
using attribute information in way not yet explored in the literture.

Assign each record in both files to a linkage cluster
\(S_k \in \{S_1, \ldots, S_K\}\) based on some attribute in the data. We
see that the comparison vector

\end{document}

### 2021-02-20 paper/linkagecluster_bk_paper.Rmd
Likelihood and full conditional of linkage cluster model, first attempt

@@
 However, if we want to ensure the theoretic purity of each MCMC sample of the linkage structure, we can "recouple" $\mathbf{Z}$ by simply resolving these illegal matchings.
-Explictly, define clusters $C(i) := \{j | Z_j = i\}$, and define all clusters such that $|C(i)| > 1$ as *conflict clusters*.
+Explictly, define clusters $R(i) := \{j | Z_j = i\}$, and define all clusters such that $|R(i)| > 1$ as *resolution clusters*.
 Within each cluster, resample the candidate matches according to their Fellegi-Sunter weights.
@@
 The primary contribution of this paper is the establishment of linkage clusters, which allow the reliability parameters $\mathbf{m}^f$, the discernment parameters $\mathbf{u}^f$, and also the linkage probability parameter $\lambda$ to vary according to certain attributes in the record pairs.
-We shall see that this set up upweights linkage within clusters and downweights linkage across clusters, thus using attribute information in way not yet explored in the literture.
+We shall see that this set up upweights linkage within clusters and downweights linkage across clusters, thus using attribute information in a way not yet explored in the literture.
 
-Assign each record in both files to a linkage cluster $S_k \in \{S_1, \ldots, S_K\}$ based on some attribute in the data. We see that the comparison vector
+Linkage cluster assignment occurs at both at the level of the records and of the record pairs.
+Assign each record in each file an appropriate linkage cluster $S_k \in \{S_1, \ldots, S_k\}$, denoting the membership record $j$ in the $k^{th}$ cluster by $j \in S_k$.
+We assign the record pair comparison vector $\gamma_{ij}$ to a cluster only when both records belong to that cluster; otherwise we assign it to $S_0$.
+For example
+$$i \in S_1, j \in S_1 \implies \gamma_{ij} \in S_1$$
+$$i \in S_1, j \in S_2 \implies \gamma_{ij} \in S_0$$
+Cluster assignments at the record pair level are necessary for the updating the $\mathbf{m}$ and $\mathbf{u}$ parameters.
+We denote the reliability and discernment parameters assoicated with each field for each cluster as $\mathbf{m}^{fk}$ and $\mathbf{u}^{fk}$ respectively.
+To improve readability, we modify the definition of $\mathbf{m}_{ij}$ to incorporate the linkage cluster assignment.
+Explicitly,
+$$\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\left(m^{fl0}\right)^{\mathbf{1}_{\gamma_{ij} \in S_0}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}$$
+with an analogous definition for $\mathbf{u}_{ij}$.
 
-<!-- The classic Fellegi Sunter model operates through weights $w_{ij} = \frac{P(\boldsymbol{\gamma}_{ij}| \text{Records match})}{P(\boldsymbol{\gamma}_{ij}| \text{Records do not match})}$.
-In our notation, these weights are given as follows: -->
-<!-- $$w_{ij} = \frac{P(\boldsymbol{\gamma}_{ij}| Z_j = i)}{P(\boldsymbol{\gamma}_{ij}| Z_j \neq i)} = \frac{\prod_{f=1}^F \prod_{l=1}^{L_f} m_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}{\prod_{f=1}^F \prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}$$ -->
+Additionally, define the vector of linkage probabilities $\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})$, where $\lambda_{k1} = P(z_{j\in S_k} \in S_k)$ is the probability that a record in $\mathbf{X}_{B_k}$ has a match within its cluster in $\mathbf{X}_{A_k}$, $\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})$ is the probability that such a record $\mathbf{X}_{B_k}$ has a match outside of its cluster $\mathbf{X}_{A_{k'}}$ for some $k' \neq k$, and $\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)$ is the probability of a non-match.
+This is a multinomial distribution where we necessarily have $\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}$.
+These parameters use linkage cluster assignment at the record level.
 
-<!-- $$p(\Gamma_{.j}|Z_{j}) = \frac{\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}{\sum_{z_j = i}^{n_A}\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A +1}(1-\lambda)\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}$$ -->
-<!-- We can divide the numerator and denominator by $\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}$ to obtain -->
+Our new model specification becomes
 
-<!-- $$p(\Gamma_{.j}|Z_{j}) =\frac{\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{\mathbf{1}_{z_j = i} \times \mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)}{\sum_{z_j = 1}^{n_A}\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{\mathbf{1}_{z_j = i} \times \mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)}$$ -->
+$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \boldsymbol{\theta}) =\prod_{k=1}^K \prod_{j \in \mathbf{X}_B^k}  \left[\lambda_{k1}^{\mathbf{1}_{z_j \in S_k}}\lambda_{k2}^{\mathbf{1}_{z_j \in S_{k'}}} \lambda_{k3}^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]$$
 
-<!-- Our model decouples the draws of the elements of the linkage structure, sampling each $Z_j$ independently from all other $Z_{j'}$.
-This decoupling allows the incorporation of additional information on each record $j$, thus enabling the linkage cluster formulation proposed in Section XXXX.
-We clearly admit that this does allow the possibility that $Z_j = i = Z_{j'}$ for some $i\in \mathbf{X}_A$, breaking one-to-one matching requirements in individual Gibbs iterations.
-However, if our model assumptions are correct and there are no duplicates within files, these sorts of illegal matches are rare and evenly distrubted across records, and thus do not seriously influence our eventual Bayes estimate of the linkage structure.
-Additionally, these illegal matches can easily be resolved in a post-hoc conflict resolution step, or "recoupling", as described in Section XXXX.
+$$\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})$$
+$$\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})$$
+$$Z_{j \in S_k} | \Lambda_k =
+\begin{cases} 
+    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
+    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}$$
+
+$$\Lambda \sim \text{Dirichlet}(\Lambda_0)$$
+where $\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).$
+
+**THIS WOULD GO IN APPENDIX**: Through similar derivations, this yields full conditionals:
+The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have 
+$$\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))$$
+
+$$\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))$$
+where $$\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}$$
+
+$$\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij}^f = l}$$
+
+We also have straightforward posterior distritutions for the linkage probability vector $\Lambda_k$:
+
+$$\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)$$
+
+where $\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})$ with 
+$$D_{k1} = \sum_{j \in \mathbf{X}_{B_k}} \mathbf{1}_{z_j \in S_k}$$ $$D_{k2} = \sum_{j \in \mathbf{X}_{B_k}} \mathbf{1}_{z_j \in S_{k'}}$$ $$\text{and } D_{k3} = \sum_{j \in \mathbf{X}_{B_k}} \mathbf{1}_{z_j = n_A + 1}$$
+
+Our full conditional for $Z_j$ follows a familiar form:
+
+$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
+\begin{cases} 
+    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
+    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}$$
+
+And after integrating over $\Lambda_k$ and rearraning terms,
+
+$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
+\begin{cases} 
+    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
+    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
+     1 & z_j = n_A + 1 \\
+\end{cases}$$
+
+<!-- While the base model tracked matches and non-matches, our linkage cluster requires us additionally to be able to track matches within and across linkage clusters.
+Define the match parameter vector $\boldsymbol{\lambda}^k = (\lambda^{k1}, \lambda^{k0}, 1 - \lambda^{k1} - \lambda^{k0})$ where $\lambda^{k1}$ is the probability of matching within cluster $k$ and $\lambda^{k0}$ is probability of matching across cluster.
+(Man this is hard to write explicitly!
+Will need to revise).
 -->
+
+<!-- Associate each record in both files with linkage cluster $k \in \{1, \ldots, K\}$, and represent the association of record $j$ in with cluster $k$ by saying $C(j) = k$.
+Denote the records in $\mathbf{X}_A$ and $\mathbf{X}_B$ associated with cluster $k$ by $\mathbf{X}_A^k$ and $\mathbf{X}_B^k$ respectively.
+We assign record pairs to cluster $k$ only if both records are assigned to that cluster, ie when $C(i) = k = C(j)$.
+Otherwise, the pair is assigned to cluster $0$.
+-->

### 2021-02-20 paper/linkagecluster_bk_paper.tex
Likelihood and full conditional of linkage cluster model, first attempt

@@
 resolving these illegal matchings. Explictly, define clusters
-\(C(i) := \{j | Z_j = i\}\), and define all clusters such that
-\(|C(i)| > 1\) as \emph{conflict clusters}. Within each cluster,
+\(R(i) := \{j | Z_j = i\}\), and define all clusters such that
+\(|R(i)| > 1\) as \emph{resolution clusters}. Within each cluster,
 resample the candidate matches according to their Fellegi-Sunter
@@
 linkage within clusters and downweights linkage across clusters, thus
-using attribute information in way not yet explored in the literture.
+using attribute information in a way not yet explored in the literture.
 
-Assign each record in both files to a linkage cluster
-\(S_k \in \{S_1, \ldots, S_K\}\) based on some attribute in the data. We
-see that the comparison vector
+Linkage cluster assignment occurs at both at the level of the records
+and of the record pairs. Assign each record in each file an appropriate
+linkage cluster \(S_k \in \{S_1, \ldots, S_k\}\), denoting the
+membership record \(j\) in the \(k^{th}\) cluster by \(j \in S_k\). We
+assign the record pair comparison vector \(\gamma_{ij}\) to a cluster
+only when both records belong to that cluster; otherwise we assign it to
+\(S_0\). For example
+\[i \in S_1, j \in S_1 \implies \gamma_{ij} \in S_1\]
+\[i \in S_1, j \in S_2 \implies \gamma_{ij} \in S_0\] Cluster
+assignments at the record pair level are necessary for the updating the
+\(\mathbf{m}\) and \(\mathbf{u}\) parameters. We denote the reliability
+and discernment parameters assoicated with each field for each cluster
+as \(\mathbf{m}^{fk}\) and \(\mathbf{u}^{fk}\) respectively. To improve
+readability, we modify the definition of \(\mathbf{m}_{ij}\) to
+incorporate the linkage cluster assignment. Explicitly,
+\[\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\left(m^{fl0}\right)^{\mathbf{1}_{\gamma_{ij} \in S_0}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}\]
+with an analogous definition for \(\mathbf{u}_{ij}\).
+
+Additionally, define the vector of linkage probabilities
+\(\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})\), where
+\(\lambda_{k1} = P(z_{j\in S_k} \in S_k)\) is the probability that a
+record in \(\mathbf{X}_{B_k}\) has a match within its cluster in
+\(\mathbf{X}_{A_k}\), \(\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})\) is
+the probability that such a record \(\mathbf{X}_{B_k}\) has a match
+outside of its cluster \(\mathbf{X}_{A_{k'}}\) for some \(k' \neq k\),
+and \(\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)\) is the probability of a
+non-match. This is a multinomial distribution where we necessarily have
+\(\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}\). These parameters use
+linkage cluster assignment at the record level.
+
+(THIS WOULD GO IN APPENDIX) Our new model specification becomes
+
+\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \boldsymbol{\theta}) =\prod_{k=1}^K \prod_{j \in \mathbf{X}_B^k}  \left[\lambda_{k1}^{\mathbf{1}_{z_j \in S_k}}\lambda_{k2}^{\mathbf{1}_{z_j \in S_{k'}}} \lambda_{k3}^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]\]
+
+\[\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})\]
+\[\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})\]
+\[Z_{j \in S_k} | \Lambda_k =
+\begin{cases} 
+    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
+    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}\]
+
+\[\Lambda \sim \text{Dirichlet}(\Lambda_0)\] where
+\(\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).\)
+
+\textbf{THIS WOULD GO IN APPENDIX}: Through similar derivations, this
+yields full conditionals: The \(\mathbf{m}\) and \(\mathbf{u}\)
+parameters are updated through standard multinomial-dirichlet mechanics.
+Thus we have
+\[\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))\]
+
+\[\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))\]
+where
+\[\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}\]
+
+\[\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij}^f = l}\]
+
+We also have straightforward posterior distritutions for the linkage
+probability vector \(\Lambda_k\):
+
+\[\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)\]
+
+where \(\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})\) with
+\[D_{k1} = \sum_{j \in \mathbf{X}_{B_k}} \mathbf{1}_{z_j \in S_k}\]
+\[D_{k2} = \sum_{j \in \mathbf{X}_{B_k}} \mathbf{1}_{z_j \in S_{k'}}\]
+\[\text{and } D_{k3} = \sum_{j \in \mathbf{X}_{B_k}} \mathbf{1}_{z_j = n_A + 1}\]
+
+Our full conditional for \(Z_j\) follows a familiar form:
+
+\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
+\begin{cases} 
+    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
+    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}\]
+
+And after integrating over \(\Lambda_k\) and rearraning terms,
+
+\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
+\begin{cases} 
+    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
+    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
+     1 & z_j = n_A + 1 \\
+\end{cases}\]
 

### 2021-02-25 paper/linkagecluster_bk_paper.Rmd
updated derivations

@@
 
-$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}} (1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]$$
+$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
 
@@
 $$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) $$
-The prior for $Z_j$ has equal probability of matching to all records $i\in \mathbf{X}_A$, and non-matching probability governed by $\lambda$. Observe two reasonable choices for $p(Z, \lambda)$:
-
-If $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$, then the prior for $Z$, integrated over $\lambda$, becomes
-
-$$=\begin{cases} 
-    1  & z_j \leq n_A; \\
-    1 &  z_j  = n_A + 1 \\
-\end{cases}$$
-Thus this is a uniform prior on the labellings of $Z$. A second useful prior would be $\lambda \sim \text{Beta}(1, 1)$, which yields
-$$=\begin{cases} 
-    1  & z_j \leq n_A; \\
-    n_A &  z_j  = n_A + 1 \\
-\end{cases}$$
-Here, a match and a non-match are equaly likely, and among the matches, all records are likely to be a match.
-There is little difference between the the two priors in practice, as the data should overpower the prior.
+The prior for $Z_j$ has equal probability of matching to all records $i\in \mathbf{X}_A$, and non-matching probability governed by $\lambda$.
+Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$ prior corresponds to a uniform prior on the labelling of $\mathbf{Z}$.
 
@@
 
-The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have 
+We work with the following factorization of the joint distribution:
+
+$$p(\Gamma, \mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) = p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)$$
+
+This factorization leads to following Gibbs Sampler:
+
+\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:} The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics.
+Thus we have
+
 $$\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))$$
@@
 
-Since $\lambda$ provides the probability that a record in $\mathbf{X}_B$ has match in $\mathbf{X}_A$, it has a full conditional distribution
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{s}$} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes and failures, and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
+Thus we have
 
-$$\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})$$
-where $D = \sum_1^{n_B} (Z_j < n_A + 1)$ is the number of records identified as matches in the previous Gibbs iteration.
+$$p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)$$
+$$\propto \lambda^D (1-\lambda)^{n_B - D} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}$$
+$$ \propto \lambda^{D + \alpha_{\lambda} - 1} (1-\lambda)^{n_B - D + \beta_{\lambda} -1}$$
+$$\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})$$
 
-Lastly, we sample each $Z_j$ according to these posterior distributions. To do so, note that as a function of $Z_j$, the likelihood is a discrete distibution with probabilities proportional to
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
+Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities given by
+$$p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) = \frac{\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}}{\sum_{z_j = 1}^{n_B} \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}} = \frac{\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}}{\sum_{z_j = 1}^{n_B}\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}} $$
+$$\propto 
+\begin{cases} 
+    w_{ij}  & z_j \leq n_A; \\
+    1 &  z_j  = n_A + 1 \\
+\end{cases}$$
 
-$$p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}$$
-$$\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)$$
-
-$$\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}$$
-where $w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
+where where $w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
 The interested reader should note that these are precisely the likelihood ratios used in the Fellegi-Sunter model to classify matches and non-matches, and we therefore refer to $w_{ij}$ as the *Fellegi Sunter weights*.
 
-The full conditional for $Z_j$ is thus given by
-$$p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}$$
+With the likelihood in this form, we can derive the full conditional
+$$p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)$$
+
 $$\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)$$
@@
 
-The relationship between this sampling mechanism and that of Sadinle 2017 warrants discussion.
+\newpage
 
@@
 This procedure would be most useful in cases where there are many "near matches" across files, but in most cases, we find this this step to be unnecessary.
+
+\newpage
 
@@
 
+\newpage 
+Observe two reasonable choices for $p(Z, \lambda)$:
+
+If $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$, then the prior for $Z$, integrated over $\lambda$, becomes
+
+$$=\begin{cases} 
+    1  & z_j \leq n_A; \\
+    1 &  z_j  = n_A + 1 \\
+\end{cases}$$
+Thus this is a uniform prior on the labellings of $Z$. A second useful prior would be $\lambda \sim \text{Beta}(1, 1)$, which yields
+$$=\begin{cases} 
+    1  & z_j \leq n_A; \\
+    n_A &  z_j  = n_A + 1 \\
+\end{cases}$$
+Here, a match and a non-match are equaly likely, and among the matches, all records are likely to be a match.
+There is little difference between the the two priors in practice, as the data should overpower the prior.
+
+The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have
+
+$$\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))$$
+$$\mathbf{u^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))$$
+where $\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}$ and $\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}$.
+
+Since $\lambda$ provides the probability that a record in $\mathbf{X}_B$ has match in $\mathbf{X}_A$, it has a full conditional distribution
+
+$$\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})$$
+where $D = \sum_1^{n_B} (Z_j < n_A + 1)$ is the number of records identified as matches in the previous Gibbs iteration.
+
+Lastly, we sample each $Z_j$ according to these posterior distributions. To do so, note that as a function of $Z_j$, the likelihood is a discrete distibution with probabilities proportional to
+
+$$p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}$$
+$$\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)$$
+
+$$\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}$$
+where $w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
+The interested reader should note that these are precisely the likelihood ratios used in the Fellegi-Sunter model to classify matches and non-matches, and we therefore refer to $w_{ij}$ as the *Fellegi Sunter weights*.
+
+The full conditional for $Z_j$ is thus given by
+$$p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}$$
+$$\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)$$
+$$= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} $$
+$$ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
+\begin{cases} 
+    \frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
+     1-\lambda &  z_j  = n_A + 1 \\
+\end{cases}$$
+
+Here, one should note that if we choose to integrate over the posterior of $\lambda$ and rearrange terms, we produce the following sampling mechanism:
+
+$$Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\begin{cases} 
+    w_{ij}  & z_j \leq n_A; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
+\end{cases}$$
+
+The relationship between this sampling mechanism and that of Sadinle 2017 warrants discussion.
+
 <!-- While the base model tracked matches and non-matches, our linkage cluster requires us additionally to be able to track matches within and across linkage clusters.

### 2021-02-25 paper/linkagecluster_bk_paper.tex
updated derivations

@@
 
-\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}} (1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]\]
+\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
 
@@
 \(i\in \mathbf{X}_A\), and non-matching probability governed by
-\(\lambda\). Observe two reasonable choices for \(p(Z, \lambda)\):
-
-If \(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\), then the prior for
-\(Z\), integrated over \(\lambda\), becomes
-
-\[=\begin{cases} 
-    1  & z_j \leq n_A; \\
-    1 &  z_j  = n_A + 1 \\
-\end{cases}\] Thus this is a uniform prior on the labellings of \(Z\). A
-second useful prior would be \(\lambda \sim \text{Beta}(1, 1)\), which
-yields \[=\begin{cases} 
-    1  & z_j \leq n_A; \\
-    n_A &  z_j  = n_A + 1 \\
-\end{cases}\] Here, a match and a non-match are equaly likely, and among
-the matches, all records are likely to be a match. There is little
-difference between the the two priors in practice, as the data should
-overpower the prior.
+\(\lambda\). Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds
+to a prior belief that nonmatches and matches are equally likely, and a
+\(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\) prior corresponds to a
+uniform prior on the labelling of \(\mathbf{Z}\).
 
@@
 
+We work with the following factorization of the joint distribution:
+
+\[p(\Gamma, \mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) = p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)\]
+
+This factorization leads to following Gibbs Sampler:
+
+\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
 The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
 standard multinomial-dirichlet mechanics. Thus we have
+
 \[\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
@@
 
-Since \(\lambda\) provides the probability that a record in
-\(\mathbf{X}_B\) has match in \(\mathbf{X}_A\), it has a full
-conditional distribution
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{s}$} As a function of
+\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
+of successes and failures, and therefore
+\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) determined
+only by the number of dupliates
+\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
+\(\mathbf{Z}\). Thus we have
 
-\[\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
-where \(D = \sum_1^{n_B} (Z_j < n_A + 1)\) is the number of records
-identified as matches in the previous Gibbs iteration.
+\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
+\[\propto \lambda^D (1-\lambda)^{n_B - D} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
+\[ \propto \lambda^{D + \alpha_{\lambda} - 1} (1-\lambda)^{n_B - D + \beta_{\lambda} -1}\]
+\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
 
-Lastly, we sample each \(Z_j\) according to these posterior
-distributions. To do so, note that as a function of \(Z_j\), the
-likelihood is a discrete distibution with probabilities proportional to
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$}
+Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
+only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
+denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
+that as a function of \(Z_j\), the likelihood
+\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})\) is a discrete
+distribution with probalities given by
+\[p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) = \frac{\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}}{\sum_{z_j = 1}^{n_B} \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}} = \frac{\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}}{\sum_{z_j = 1}^{n_B}\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}} \]
+\[\propto 
+\begin{cases} 
+    w_{ij}  & z_j \leq n_A; \\
+    1 &  z_j  = n_A + 1 \\
+\end{cases}\]
 
-\[p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}\]
-\[\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\]
-
-\[\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}\]
-where
+where where
 \(w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
@@
 
-The full conditional for \(Z_j\) is thus given by
-\[p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}\]
+With the likelihood in this form, we can derive the full conditional
+\[p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)\]
+
 \[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
@@
 
-The relationship between this sampling mechanism and that of Sadinle
-2017 warrants discussion.
+\newpage
 
@@
 
+\newpage
+
 \hypertarget{linkage-cluster-model}{%
@@
 
-(THIS WOULD GO IN APPENDIX) Our new model specification becomes
+Our new model specification becomes
 
@@
 
+\newpage
+
+Observe two reasonable choices for \(p(Z, \lambda)\):
+
+If \(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\), then the prior for
+\(Z\), integrated over \(\lambda\), becomes
+
+\[=\begin{cases} 
+    1  & z_j \leq n_A; \\
+    1 &  z_j  = n_A + 1 \\
+\end{cases}\] Thus this is a uniform prior on the labellings of \(Z\). A
+second useful prior would be \(\lambda \sim \text{Beta}(1, 1)\), which
+yields \[=\begin{cases} 
+    1  & z_j \leq n_A; \\
+    n_A &  z_j  = n_A + 1 \\
+\end{cases}\] Here, a match and a non-match are equaly likely, and among
+the matches, all records are likely to be a match. There is little
+difference between the the two priors in practice, as the data should
+overpower the prior.
+
+The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
+standard multinomial-dirichlet mechanics. Thus we have
+
+\[\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
+\[\mathbf{u^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
+where
+\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
+and
+\(\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).
+
+Since \(\lambda\) provides the probability that a record in
+\(\mathbf{X}_B\) has match in \(\mathbf{X}_A\), it has a full
+conditional distribution
+
+\[\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
+where \(D = \sum_1^{n_B} (Z_j < n_A + 1)\) is the number of records
+identified as matches in the previous Gibbs iteration.
+
+Lastly, we sample each \(Z_j\) according to these posterior
+distributions. To do so, note that as a function of \(Z_j\), the
+likelihood is a discrete distibution with probabilities proportional to
+
+\[p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}\]
+\[\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\]
+
+\[\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}\]
+where
+\(w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
+The interested reader should note that these are precisely the
+likelihood ratios used in the Fellegi-Sunter model to classify matches
+and non-matches, and we therefore refer to \(w_{ij}\) as the
+\emph{Fellegi Sunter weights}.
+
+The full conditional for \(Z_j\) is thus given by
+\[p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}\]
+\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
+\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
+\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
+\begin{cases} 
+    \frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
+     1-\lambda &  z_j  = n_A + 1 \\
+\end{cases}\]
+
+Here, one should note that if we choose to integrate over the posterior
+of \(\lambda\) and rearrange terms, we produce the following sampling
+mechanism:
+
+\[Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\begin{cases} 
+    w_{ij}  & z_j \leq n_A; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
+\end{cases}\]
+
+The relationship between this sampling mechanism and that of Sadinle
+2017 warrants discussion.
+
 \end{document}

### 2021-02-27 paper/base_model_specification.Rmd
Updated derivations

--- 2021-02-18 paper/linkagecluster_bk_paper.Rmd
@@
 	\end{flushright}
-	\begin{center} \textbf{Linkage Cluster Paper Draft}
+	\begin{center} \textbf{Base Model Specification}
 
@@
 ```
-
-## Related Work
-
-Sadinle (2017) proposed a Bayesian method for bipartite record linkage that provides a foundation for much of our proposed method.
-In order to ensure one-to-one matching, his Gibbs sampler conditions on the linkage status of all other records when making each linkage assignment.
-While this elegantly ensures transitivity, the probability distribution of the linkage vector $\mathbf{Z}$ is somewhat complicated, and does not easily allow for inclusion of additional structure among records.
-Our method however, motivated by recent decouple/recouple methods, initially relaxes the one-to-one matching assumption in order to incorporate linkage information differing by subgroups.
-
-There has been some work done in the literature on conducting record linkage in situations where feild accuracy differs by subgroups of the data, but these are hampered by technical issues that make them difficult to use in practice.
-Larson (2005) proposed Bayesian heirarchical record linkage model that explicitly modelled the differences in parameters across blocks of the data.
-In relying on blocking however, he assumes that the blocking variable is error-free, when in practice such a procedure could result in many missed linkages.
-In contrast, our proposed method induces high probabilities of matching within a linkage cluster while still allowing the possibility to match across clusters (when the clustering field is recorded with error.
-Hu et al (2015) proposed a method that used a logistic regression submodel within a broader record linkage procedure to account for "expected changes" in the record of the underlying entity based on subgroup membership.
-This method however specified for use in temporal settings when the true attributes of the record change with time, and it is unclear how such a method would be implemented when records are collected syncronously with error, or when time information is unavailable.
-Additionally, the method is semi-supervised, requiring training data to implement the logistic regression, and is that not well suited for record linkage in practice.
-In contrast, the proposed method is fully unsupervised and is shown to not require extensive tuning of prior distributions, allowing for ease of implementation in a variety of settings.
 
@@
 It also makes it explict that units of observation under this model are the $n_B$ records in $\mathbf{X}_B$, a crucial distinction when compared to the likelihood over the $n_A \times n_B$ record pairs in the original Fellegi Sunter model.
-<!-- $$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}}(1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}}\right]$$ -->
 
-$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}} (1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]$$
+$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
 
@@
 
-Since $\lambda$ provides the probability that a record in $\mathbf{X}_B$ has match in $\mathbf{X}_A$, it has a full conditional distribution
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$ As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
+Thus we have
 
-$$\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})$$
-where $D = \sum_1^{n_B} (Z_j < n_A + 1)$ is the number of records identified as matches in the previous Gibbs iteration.
+$$p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)$$
+$$\propto \lambda^D (1-\lambda)^{n_B - D} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}$$
+$$ \propto \lambda^{D + \alpha_{\lambda} - 1} (1-\lambda)^{n_B - D + \beta_{\lambda} -1}$$
+$$\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})$$
 
-Lastly, we sample each $Z_j$ according to these posterior distributions. To do so, note that as a function of $Z_j$, the likelihood is a discrete distibution with probabilities proportional to
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
+Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to
+$$p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
+$$\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{Divide through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}$$
+$$=
+\begin{cases} 
+    w_{ij}  & z_j \leq n_A; \\
+    1 &  z_j  = n_A + 1 \\
+\end{cases}$$
 
-$$p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}$$
-$$\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)$$
-
-$$\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}$$
 where $w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
@@
 
-The full conditional for $Z_j$ is thus given by
-$$p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}$$
+With the likelihood in this form, we can derive the full conditional
+$$p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)$$
+
 $$\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)$$
@@
 \end{cases}$$
-
-The relationship between this sampling mechanism and that of Sadinle 2017 warrants discussion.
-
-## Bayes Estimate
-
-Sadinle 2017
-
-## Decouple/Recouple
-
-One difficulty in record linkage is conducting linkage while taking appropriate dependencies into account.
-In the context of this paper in which we assume that each file is recorded without duplicates, this means our linkage procedure should respect one-to-one matching.
-The posterior sampling of $\mathbf{Z}$ outlined above resolves half of this issue; since we only sample one potential match for record $j$ in each Gibbs iteration, it is impossible to match $j$ to two records at once.
-This in itself is an improvement over the classic Fellegi-Sunter method that labels record pairs as matches or non matches independently of all other pairs.
-However, our posterior sampling of $\mathbf{Z}$ does allow for the possibility that $Z_j = i = Z_{j'}$ for some $j' \neq j$, which cleary violates the assumption that $\mathbf{X}_B$ is recorded without duplicates.
-Sadinle 2017 resolves this issue by adapting the set of candidate matches for every $Z_j$, but we forgo this rigidity and instead "decouple" the linkage vector $\mathbf{Z}$ for the computational advantantages of sampling $Z_j$ independently.
-In our formulation, such illegal matches occasionally occur within individual Gibbs iterations due to simple randomness, but they do not occur consistently enough between the same records to measurably influence the eventual Bayes estimate of the linkage structure.
-
-However, if we want to ensure the theoretic purity of each MCMC sample of the linkage structure, we can "recouple" $\mathbf{Z}$ by simply resolving these illegal matchings.
-Explictly, define clusters $C(i) := \{j | Z_j = i\}$, and define all clusters such that $|C(i)| > 1$ as *conflict clusters*.
-Within each cluster, resample the candidate matches according to their Fellegi-Sunter weights.
-We show this step to be computationally inexpensive and to be a reasonable replication of more organic one-to-one matching.
-This procedure would be most useful in cases where there are many "near matches" across files, but in most cases, we find this this step to be unnecessary.
-
-## Linkage cluster model
-
-The primary contribution of this paper is the establishment of linkage clusters, which allow the reliability parameters $\mathbf{m}^f$, the discernment parameters $\mathbf{u}^f$, and also the linkage probability parameter $\lambda$ to vary according to certain attributes in the record pairs.
-We shall see that this set up upweights linkage within clusters and downweights linkage across clusters, thus using attribute information in way not yet explored in the literture.
-
-Assign each record in both files to a linkage cluster $S_k \in \{S_1, \ldots, S_K\}$ based on some attribute in the data. We see that the comparison vector
-
-<!-- The classic Fellegi Sunter model operates through weights $w_{ij} = \frac{P(\boldsymbol{\gamma}_{ij}| \text{Records match})}{P(\boldsymbol{\gamma}_{ij}| \text{Records do not match})}$.
-In our notation, these weights are given as follows: -->
-<!-- $$w_{ij} = \frac{P(\boldsymbol{\gamma}_{ij}| Z_j = i)}{P(\boldsymbol{\gamma}_{ij}| Z_j \neq i)} = \frac{\prod_{f=1}^F \prod_{l=1}^{L_f} m_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}{\prod_{f=1}^F \prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}$$ -->
-
-<!-- $$p(\Gamma_{.j}|Z_{j}) = \frac{\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}{\sum_{z_j = i}^{n_A}\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m_{fl}^{\mathbf{1}_{z_j = i}}u_{fl}^{\mathbf{1}_{z_j \neq i}}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A +1}(1-\lambda)\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}}$$ -->
-<!-- We can divide the numerator and denominator by $\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\mathbf{1}_{\gamma_{ij}^f = l}}$ to obtain -->
-
-<!-- $$p(\Gamma_{.j}|Z_{j}) =\frac{\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{\mathbf{1}_{z_j = i} \times \mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)}{\sum_{z_j = 1}^{n_A}\mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{\mathbf{1}_{z_j = i} \times \mathbf{1}_{\gamma_{ij}^f = l}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)}$$ -->
-
-<!-- Our model decouples the draws of the elements of the linkage structure, sampling each $Z_j$ independently from all other $Z_{j'}$.
-This decoupling allows the incorporation of additional information on each record $j$, thus enabling the linkage cluster formulation proposed in Section XXXX.
-We clearly admit that this does allow the possibility that $Z_j = i = Z_{j'}$ for some $i\in \mathbf{X}_A$, breaking one-to-one matching requirements in individual Gibbs iterations.
-However, if our model assumptions are correct and there are no duplicates within files, these sorts of illegal matches are rare and evenly distrubted across records, and thus do not seriously influence our eventual Bayes estimate of the linkage structure.
-Additionally, these illegal matches can easily be resolved in a post-hoc conflict resolution step, or "recoupling", as described in Section XXXX.
--->

### 2021-02-27 paper/base_model_specification.tex
Updated derivations

--- 2021-02-18 paper/linkagecluster_bk_paper.tex
@@
     \end{flushright}
-    \begin{center} \textbf{Linkage Cluster Paper Draft}
+    \begin{center} \textbf{Base Model Specification}
 
@@
     \end{center}
-
-\hypertarget{related-work}{%
-\subsection{Related Work}\label{related-work}}
-
-Sadinle (2017) proposed a Bayesian method for bipartite record linkage
-that provides a foundation for much of our proposed method. In order to
-ensure one-to-one matching, his Gibbs sampler conditions on the linkage
-status of all other records when making each linkage assignment. While
-this elegantly ensures transitivity, the probability distribution of the
-linkage vector \(\mathbf{Z}\) is somewhat complicated, and does not
-easily allow for inclusion of additional structure among records. Our
-method however, motivated by recent decouple/recouple methods, initially
-relaxes the one-to-one matching assumption in order to incorporate
-linkage information differing by subgroups.
-
-There has been some work done in the literature on conducting record
-linkage in situations where feild accuracy differs by subgroups of the
-data, but these are hampered by technical issues that make them
-difficult to use in practice. Larson (2005) proposed Bayesian
-heirarchical record linkage model that explicitly modelled the
-differences in parameters across blocks of the data. In relying on
-blocking however, he assumes that the blocking variable is error-free,
-when in practice such a procedure could result in many missed linkages.
-In contrast, our proposed method induces high probabilities of matching
-within a linkage cluster while still allowing the possibility to match
-across clusters (when the clustering field is recorded with error. Hu et
-al (2015) proposed a method that used a logistic regression submodel
-within a broader record linkage procedure to account for ``expected
-changes'' in the record of the underlying entity based on subgroup
-membership. This method however specified for use in temporal settings
-when the true attributes of the record change with time, and it is
-unclear how such a method would be implemented when records are
-collected syncronously with error, or when time information is
-unavailable. Additionally, the method is semi-supervised, requiring
-training data to implement the logistic regression, and is that not well
-suited for record linkage in practice. In contrast, the proposed method
-is fully unsupervised and is shown to not require extensive tuning of
-prior distributions, allowing for ease of implementation in a variety of
-settings.
 
@@
 
-Since \(\lambda\) provides the probability that a record in
-\(\mathbf{X}_B\) has match in \(\mathbf{X}_A\), it has a full
-conditional distribution
+\textbackslash underline\{Sample \(\lambda^{(s+1)}|\mathbf{Z}^{(s)}\) As
+a function of \(\lambda\), the linkage structure parameter
+\(\mathbf{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and
+failures (when \(z_j = n_A + 1\)), and therefore
+\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
+determined only by the number of dupliates
+\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
+\(\mathbf{Z}\). Thus we have
 
-\[\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
-where \(D = \sum_1^{n_B} (Z_j < n_A + 1)\) is the number of records
-identified as matches in the previous Gibbs iteration.
+\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
+\[\propto \lambda^D (1-\lambda)^{n_B - D} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
+\[ \propto \lambda^{D + \alpha_{\lambda} - 1} (1-\lambda)^{n_B - D + \beta_{\lambda} -1}\]
+\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
 
-Lastly, we sample each \(Z_j\) according to these posterior
-distributions. To do so, note that as a function of \(Z_j\), the
-likelihood is a discrete distibution with probabilities proportional to
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$}
+Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
+only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
+denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
+that as a function of \(Z_j\), the likelihood
+\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})\) is a discrete
+distribution with probalities proportional to
+\[p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
+\[\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{Divide through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\]
+\[=
+\begin{cases} 
+    w_{ij}  & z_j \leq n_A; \\
+    1 &  z_j  = n_A + 1 \\
+\end{cases}\]
 
-\[p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}\]
-\[\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\]
-
-\[\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}\]
 where
@@
 
-The relationship between this sampling mechanism and that of Sadinle
-2017 warrants discussion.
-
-\hypertarget{bayes-estimate}{%
-\subsection{Bayes Estimate}\label{bayes-estimate}}
-
-Sadinle 2017
-
-\hypertarget{decouplerecouple}{%
-\subsection{Decouple/Recouple}\label{decouplerecouple}}
-
-One difficulty in record linkage is conducting linkage while taking
-appropriate dependencies into account. In the context of this paper in
-which we assume that each file is recorded without duplicates, this
-means our linkage procedure should respect one-to-one matching. The
-posterior sampling of \(\mathbf{Z}\) outlined above resolves half of
-this issue; since we only sample one potential match for record \(j\) in
-each Gibbs iteration, it is impossible to match \(j\) to two records at
-once. This in itself is an improvement over the classic Fellegi-Sunter
-method that labels record pairs as matches or non matches independently
-of all other pairs. However, our posterior sampling of \(\mathbf{Z}\)
-does allow for the possibility that \(Z_j = i = Z_{j'}\) for some
-\(j' \neq j\), which cleary violates the assumption that
-\(\mathbf{X}_B\) is recorded without duplicates. Sadinle 2017 resolves
-this issue by adapting the set of candidate matches for every \(Z_j\),
-but we forgo this rigidity and instead ``decouple'' the linkage vector
-\(\mathbf{Z}\) for the computational advantantages of sampling \(Z_j\)
-independently. In our formulation, such illegal matches occasionally
-occur within individual Gibbs iterations due to simple randomness, but
-they do not occur consistently enough between the same records to
-measurably influence the eventual Bayes estimate of the linkage
-structure.
-
-However, if we want to ensure the theoretic purity of each MCMC sample
-of the linkage structure, we can ``recouple'' \(\mathbf{Z}\) by simply
-resolving these illegal matchings. Explictly, define clusters
-\(C(i) := \{j | Z_j = i\}\), and define all clusters such that
-\(|C(i)| > 1\) as \emph{conflict clusters}. Within each cluster,
-resample the candidate matches according to their Fellegi-Sunter
-weights. We show this step to be computationally inexpensive and to be a
-reasonable replication of more organic one-to-one matching. This
-procedure would be most useful in cases where there are many ``near
-matches'' across files, but in most cases, we find this this step to be
-unnecessary.
-
-\hypertarget{linkage-cluster-model}{%
-\subsection{Linkage cluster model}\label{linkage-cluster-model}}
-
-The primary contribution of this paper is the establishment of linkage
-clusters, which allow the reliability parameters \(\mathbf{m}^f\), the
-discernment parameters \(\mathbf{u}^f\), and also the linkage
-probability parameter \(\lambda\) to vary according to certain
-attributes in the record pairs. We shall see that this set up upweights
-linkage within clusters and downweights linkage across clusters, thus
-using attribute information in way not yet explored in the literture.
-
-Assign each record in both files to a linkage cluster
-\(S_k \in \{S_1, \ldots, S_K\}\) based on some attribute in the data. We
-see that the comparison vector
-
 \end{document}

### 2021-02-27 paper/linkagecluster_bk_paper.Rmd
Updated derivations

@@
 \underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
-Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities given by
-$$p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) = \frac{\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}}{\sum_{z_j = 1}^{n_B} \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}} = \frac{\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}}{\sum_{z_j = 1}^{n_B}\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}} $$
-$$\propto 
+Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to
+$$p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
+$$\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{Divide through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}$$
+$$=
 \begin{cases} 
@@
 
-where where $w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
+where $w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
 The interested reader should note that these are precisely the likelihood ratios used in the Fellegi-Sunter model to classify matches and non-matches, and we therefore refer to $w_{ij}$ as the *Fellegi Sunter weights*.

### 2021-02-27 R/NLC_functions.R

@@ -1,4 +1,4 @@
-BKSimple <- function(comparisons, m.prior, u.prior,
+BKSimple <- function(comparisons, m_prior, u_prior,
                      alpha, beta, S, burn){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
@@ -38,8 +38,8 @@ BKSimple <- function(comparisons, m.prior, u.prior,
       sum(x == 1 & Z.temp == 0)
     })
 
-    m.post <- m.prior + aZ
-    u.post <- u.prior + bZ
+    m.post <- m_prior + aZ
+    u.post <- u_prior + bZ
 
     m.post <- split(m.post, parameter.indicators)
     m <- as.vector(unlist(sapply(m.post, function(x){
@@ -62,6 +62,110 @@ BKSimple <- function(comparisons, m.prior, u.prior,
     Z <- unname(sapply(weights, function(x){
       sample(c(candidates, n1 + 1), 1, prob = c(x, offset))
     }))
+
+    Z.temp <- sapply(Z, function(x){
+      if(x < n1 + 1){
+        vec <- rep(0, n1)
+        vec[x] <- 1
+        vec
+      }else{
+        rep(0, n1)
+      }
+    })
+    L <- sum(Z < n1 + 1)
+
+    Z.SAMPS[,s] <- Z
+    #M.SAMPS[,s] <- m
+    #U.SAMPS[,s] <- u
+    L.SAMPS[s] <- L
+
+    if (s %% (S / 100) == 0) {
+      flush.console()
+      cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+    }
+  }
+
+  Z.SAMPS <- Z.SAMPS[,-(1:burn)]
+  #L.SAMPS <- L.SAMPS[-(1:burn)]
+  #M.SAMPS <- M.SAMPS[,-(1:burn)]
+  #U.SAMPS <- U.SAMPS[,-(1:burn)]
+
+  Z.SAMPS
+}
+
+
+Sadinle17 <- function(comparisons, m_prior, uprior,
+                     alpha, beta, S, burn){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRecords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
[…]

### 2021-02-27 R/NLCpar_functions.R
played with parallelization

@@ -1,0 +1,94 @@
+BKSimplePar <- function(comparisons, m_prior, u_prior,
+                     alpha, beta, S, burn){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRecords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
+  # alpha = first parameter of prior for linkage probability
+  # beta = second parameter of prior for linkage probability
+  # S = number of Gibbs iterations
+  # burn = number of iterations to be discarded as burn-in
+
+  fields <- length(comparisons[[4]])
+  n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
+  indicators <- comparisons[[1]]
+
+  parameter.indicators <- as.vector(unlist(sapply(1:fields, function(x){
+    rep(x, comparisons[[4]][x])
+  })))
+
+  ids <- expand.grid(1:n1, 1:n2)
+  candidates <- 1:n1
+  Z.SAMPS <- matrix(NA, nrow = n2, ncol = S)
+  M.SAMPS <- matrix(NA, nrow = dim(indicators)[2], ncol = S)
+  U.SAMPS <- matrix(NA, nrow = dim(indicators)[2], ncol = S)
+  L.SAMPS <- vector(length = S)
+  Z.temp <- rep(0, n1*n2)
+  Z <- rep(n1+1, n2)
+  L <- 0
+
+  # Gibbs
+  for(s in 1:S){
+    aZ <- apply(indicators, 2, function(x){
+      sum(x == 1 & Z.temp == 1)
+    })
+    bZ <- apply(indicators, 2, function(x){
+      sum(x == 1 & Z.temp == 0)
+    })
+
+    m.post <- m_prior + aZ
+    u.post <- u_prior + bZ
+
+    m.post <- split(m.post, parameter.indicators)
+    m <- as.vector(unlist(sapply(m.post, function(x){
+      prob <- MCMCpack::rdirichlet(1, x)
+      prob/sum(prob)
+    })))
+
+    u.post <- split(u.post, parameter.indicators)
+    u <- as.vector(unlist(sapply(u.post, function(x){
+      prob <- MCMCpack::rdirichlet(1, x)
[…]

### 2021-02-27 R/NLCpar_functions.R
experiment with parallel

@@ -59,10 +59,11 @@ BKSimplePar <- function(comparisons, m_prior, u_prior,
     weights <- split(weights, ids[,2])
     offset <- n1*(n2 - L + beta)/(L+ alpha)
 
-    Z <- unname(parSapply(cl, weights, FUN = function(x){
+    system.time(
+    Z <- unlist(mclapply(weights, function(x){
       sample(c(candidates, n1 + 1), 1, prob = c(x, offset))
     }))
-
+)
     Z.temp <- sapply(Z, function(x){
       if(x < n1 + 1){
         vec <- rep(0, n1)

### 2021-02-27 paper/base_model_specification.Rmd
Base model specification revisions

@@
 Our notation and assumptions closely follow that of Sadinle (2017).
-Denote two files as $\mathbf{X}_A$ and $\mathbf{X}_B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $\mathbf{X}_A$ and $j \in \{1, \ldots, n_B\}$ in $\mathbf{X}_B$.
+Denote two files as $A$ and $B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $A$ and $j \in \{1, \ldots, n_B\}$ in $B$.
 Without loss of generality, label the files such that $n_A \geq n_B$.
 We also assume there are no duplicates within files, only across.
-For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each takes $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating level agreement between the two records on a specified field.
+For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating level agreement between the two records on a specified field.
+
+To indicate matching status, we adopt the *linkage structure parameter* $\mathbf{Z} = (Z_1, \ldots, Z_{n_A})$ from Sadinle 2017, defined as
+$$Z_j=\begin{cases} 
+    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
+    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
+\end{cases}$$
+This provides more memory efficient storage for the linkage information than a $n_A \times n_B$ sparse matrix of indicators.
 
 Following the Fellegi Sunter framework, we define $m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)$ to be the probability of observing agreement level $l$ in field $f$ for records $i$ and $j$ given that the records are a match, and similarly define $u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)$, for non-matches.
-We denote $\lambda$ to be the (marginal) probability that some record $j \in \mathbf{X}_B$ has a match in $\mathbf{X}_A$.
+We denote $\lambda$ to be the (marginal) probability that some record $j \in B$ has a match in $A$.
 
-Lastly, we adopt from Sadinle 2017 the matching labelling $\mathbf{Z} = (Z_1, \ldots, Z_{n_A})$, defined as
-$$Z_j=\begin{cases} 
-    i,  & \text{if records } i\in \mathbf{X}_A \text{ and } j\in \mathbf{X}_B \text{ refer to the same entity}; \\
-    n_A + 1,  & \text{if record } j\in \mathbf{X}_B \text{ does not have a match in file} \mathbf{X}_B; \\
-\end{cases}$$
-
-Wherever possible, we reserve superscripts for denoting field, level, and linkage cluster, and reserve subscripts for record indices.
+Wherever possible, we reserve superscripts for denoting field and level, while reserving subscripts for record indices.
 For example, $\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})$ is the probability distribution governing field $f$ for matching records, and $\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)$ is product of the relevant of the appropriate $\mathbf{m}$ parameters for record pair $(i,j)$.
@@
 
-Additionally, our model differs from that of Sadinle 2017 through its explicit dependence on a beta random variable $\lambda$ that models the rate of matching across records.
+Our model differs from that of Sadinle 2017 through its explicit dependence on a beta random variable $\lambda$ that models the rate of matching across records.
 Sadinle marginalizes over such a random variable in his derivations of  the "beta prior for bipartite matching," but here we provide derivations without marginalizing in order to specify differing rates of matching for different linkage clusters.
@@
 For fields $f \in \{1, \ldots, F\}$ and levels $l\in \{1, \ldots, L_f\}$ we adopt the following likelihood and prior distributions.
-Note that likelihood is related to that of Sadinle (2017), but includes dependence on the $\lambda$ random variable.
-It also makes it explict that units of observation under this model are the $n_B$ records in $\mathbf{X}_B$, a crucial distinction when compared to the likelihood over the $n_A \times n_B$ record pairs in the original Fellegi Sunter model.
 
@@
 $$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) $$
-The prior for $Z_j$ has equal probability of matching to all records $i\in \mathbf{X}_A$, and non-matching probability governed by $\lambda$.
+The prior for $Z_j$ has equal probability of matching to all records $i\in A$, and non-matching probability governed by $\lambda$.
 Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$ prior corresponds to a uniform prior on the labelling of $\mathbf{Z}$.
@@
 
-$$\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))$$
-$$\mathbf{u^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))$$
+$$\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))$$
+$$\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))$$
 where $\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}$ and $\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}$.
 
-\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$ As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
 Thus we have
@@
 \underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
-Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to
+Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to
 $$p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
-$$\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{Divide through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}$$
+$$\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}$$
 $$=
@@
 
-Here, one should note that if we choose to integrate over the posterior of $\lambda$ and rearrange terms, we produce the following sampling mechanism:
+Here, we integrate over the posterior of $\lambda$ and rearrange terms to produce the final full conditional:
 

### 2021-02-27 paper/base_model_specification.tex
Base model specification revisions

@@
 Our notation and assumptions closely follow that of Sadinle (2017).
-Denote two files as \(\mathbf{X}_A\) and \(\mathbf{X}_B\), with \(n_A\)
-and \(n_B\) records respectively, and with records indexed as
-\(i \in \{1, \ldots, n_A\}\) in \(\mathbf{X}_A\) and
-\(j \in \{1, \ldots, n_B\}\) in \(\mathbf{X}_B\). Without loss of
+Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
+respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
+in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
 generality, label the files such that \(n_A \geq n_B\). We also assume
@@
 \(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
-where \(F\) is the number of fields used in the linkage and each takes
+where \(F\) is the number of fields used in the linkage and each
 \(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
 indicating level agreement between the two records on a specified field.
+
+To indicate matching status, we adopt the \emph{linkage structure
+parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_A})\) from Sadinle 2017,
+defined as \[Z_j=\begin{cases} 
+    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
+    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
+\end{cases}\] This provides more memory efficient storage for the
+linkage information than a \(n_A \times n_B\) sparse matrix of
+indicators.
 
@@
 denote \(\lambda\) to be the (marginal) probability that some record
-\(j \in \mathbf{X}_B\) has a match in \(\mathbf{X}_A\).
+\(j \in B\) has a match in \(A\).
 
-Lastly, we adopt from Sadinle 2017 the matching labelling
-\(\mathbf{Z} = (Z_1, \ldots, Z_{n_A})\), defined as \[Z_j=\begin{cases} 
-    i,  & \text{if records } i\in \mathbf{X}_A \text{ and } j\in \mathbf{X}_B \text{ refer to the same entity}; \\
-    n_A + 1,  & \text{if record } j\in \mathbf{X}_B \text{ does not have a match in file} \mathbf{X}_B; \\
-\end{cases}\]
-
-Wherever possible, we reserve superscripts for denoting field, level,
-and linkage cluster, and reserve subscripts for record indices. For
-example, \(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the
-probability distribution governing field \(f\) for matching records, and
+Wherever possible, we reserve superscripts for denoting field and level,
+while reserving subscripts for record indices. For example,
+\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
+distribution governing field \(f\) for matching records, and
 \(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
@@
 
-Additionally, our model differs from that of Sadinle 2017 through its
-explicit dependence on a beta random variable \(\lambda\) that models
-the rate of matching across records. Sadinle marginalizes over such a
-random variable in his derivations of the ``beta prior for bipartite
+Our model differs from that of Sadinle 2017 through its explicit
+dependence on a beta random variable \(\lambda\) that models the rate of
+matching across records. Sadinle marginalizes over such a random
+variable in his derivations of the ``beta prior for bipartite
 matching,'' but here we provide derivations without marginalizing in
@@
 \(l\in \{1, \ldots, L_f\}\) we adopt the following likelihood and prior
-distributions. Note that likelihood is related to that of Sadinle
-(2017), but includes dependence on the \(\lambda\) random variable. It
-also makes it explict that units of observation under this model are the
-\(n_B\) records in \(\mathbf{X}_B\), a crucial distinction when compared
-to the likelihood over the \(n_A \times n_B\) record pairs in the
-original Fellegi Sunter model.
+distributions.
 
@@
 prior for \(Z_j\) has equal probability of matching to all records
-\(i\in \mathbf{X}_A\), and non-matching probability governed by
-\(\lambda\). Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds
-to a prior belief that nonmatches and matches are equally likely, and a
+\(i\in A\), and non-matching probability governed by \(\lambda\).
+Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds to a prior
+belief that nonmatches and matches are equally likely, and a
 \(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\) prior corresponds to a
@@
 
-\[\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
-\[\mathbf{u^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
+\[\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
+\[\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
 where
@@
 
-\textbackslash underline\{Sample \(\lambda^{(s+1)}|\mathbf{Z}^{(s)}\) As
-a function of \(\lambda\), the linkage structure parameter
-\(\mathbf{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and
-failures (when \(z_j = n_A + 1\)), and therefore
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$} As a function of
+\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
+of successes (when \(z_j < n_A + 1\)) and failures (when
+\(z_j = n_A + 1\)), and therefore
 \(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
@@
 that as a function of \(Z_j\), the likelihood
-\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u})\) is a discrete
-distribution with probalities proportional to
+\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
+is a discrete distribution with probalities proportional to
 \[p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
-\[\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{Divide through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\]
+\[\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\]
 \[=
@@
 
-Here, one should note that if we choose to integrate over the posterior
-of \(\lambda\) and rearrange terms, we produce the following sampling
-mechanism:
+Here, we integrate over the posterior of \(\lambda\) and rearrange terms
+to produce the final full conditional:
 

### 2021-02-27 R/BRACS.R
experimented with savings

@@ -1,5 +1,5 @@
 BRACS_BK <- function(comparisons, m_prior, u_prior,
-                       alpha, beta, S, burn){
+                       alpha, beta, S, burn, show_progress = T){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -126,12 +126,13 @@ BRACS_BK <- function(comparisons, m_prior, u_prior,
     #U.SAMPS[,s] <- u
     #L.SAMPS[s] <- L
 
-    if (s %% (S / 100) == 0) {
-      flush.console()
-      cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+    if(show_progress){
+      if (s %% (S / 100) == 0) {
+        flush.console()
+        cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+      }
     }
   }
-
   Z.SAMPS <- Z.SAMPS[,-(1:burn)]
   #L.SAMPS <- L.SAMPS[-(1:burn)]
   #M.SAMPS <- M.SAMPS[,-(1:burn)]
@@ -141,7 +142,7 @@ BRACS_BK <- function(comparisons, m_prior, u_prior,
 }
 
 BRACS_Heck <- function(comparisons, m_prior, u_prior,
-                     alpha, beta, S, burn){
+                     alpha, beta, S, burn, show_progress = T){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -269,12 +270,13 @@ BRACS_Heck <- function(comparisons, m_prior, u_prior,
     #U.SAMPS[,s] <- u
     #L.SAMPS[s] <- L
 
-    if (s %% (S / 100) == 0) {
-      flush.console()
-      cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+    if(show_progress){
+      if (s %% (S / 100) == 0) {
+        flush.console()
+        cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+      }
     }
   }
-
   Z.SAMPS <- Z.SAMPS[,-(1:burn)]
   #L.SAMPS <- L.SAMPS[-(1:burn)]
   #M.SAMPS <- M.SAMPS[,-(1:burn)]

### 2021-02-27 R/NLC_functions.R

@@ -1,5 +1,5 @@
 BKSimple <- function(comparisons, m_prior, u_prior,
-                     alpha, beta, S, burn){
+                     alpha, beta, S, burn, show_progress = T){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -10,6 +10,7 @@ BKSimple <- function(comparisons, m_prior, u_prior,
   # beta = second parameter of prior for linkage probability
   # S = number of Gibbs iterations
   # burn = number of iterations to be discarded as burn-in
+  # show_progress = set to false to show simulation progress
 
   fields <- length(comparisons[[4]])
   n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
@@ -78,13 +79,13 @@ BKSimple <- function(comparisons, m_prior, u_prior,
     #M.SAMPS[,s] <- m
     #U.SAMPS[,s] <- u
     L.SAMPS[s] <- L
-
-    if (s %% (S / 100) == 0) {
-      flush.console()
-      cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+    if(show_progress){
+      if (s %% (S / 100) == 0) {
+        flush.console()
+        cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+      }
     }
   }
-
   Z.SAMPS <- Z.SAMPS[,-(1:burn)]
   #L.SAMPS <- L.SAMPS[-(1:burn)]
   #M.SAMPS <- M.SAMPS[,-(1:burn)]
@@ -95,7 +96,7 @@ BKSimple <- function(comparisons, m_prior, u_prior,
 
 
 Sadinle17 <- function(comparisons, m_prior, uprior,
-                     alpha, beta, S, burn){
+                     alpha, beta, S, burn, show_progress = F){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -183,12 +184,13 @@ Sadinle17 <- function(comparisons, m_prior, uprior,
     #U.SAMPS[,s] <- u
     L.SAMPS[s] <- L
 
-    if (s %% (S / 100) == 0) {
-      flush.console()
-      cat("\r", paste("Simulation", ": ", s / (S / 100), "% complete", sep = ""))
+    if(show_progress){
+      if (s %% (S / 100) == 0) {
[…]

### 2021-03-03 R/BRACS.R
BRACS2 - faster linkage cluster method. Still working out bugs

@@ -14,7 +14,7 @@ BRACS_BK <- function(comparisons, m_prior, u_prior,
   fields <- length(comparisons[[4]])
   n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
   ids <- expand.grid(1:n1, 1:n2)
-  indicators <- comparisons[[1]]
+  indicators <- data.frame(comparisons[[1]])
 
   parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
     rep(x, comparisons[[4]][x])
@@ -34,7 +34,7 @@ BRACS_BK <- function(comparisons, m_prior, u_prior,
     to_change <- which(ids[, 2] %in% n2_indices[[k]])
     linkage_cluster[to_change] <- k
   }
-
+#indicators <- data.frame(indicators, linkage_cluster)
 
   candidates <- 1:n1
   Z.SAMPS <- matrix(NA, nrow = n2, ncol = S)
@@ -50,12 +50,46 @@ BRACS_BK <- function(comparisons, m_prior, u_prior,
   # Gibbs
   for(s in 1:S){
 
-    AZ[-varying_fields, ] <- t(apply(indicators[, -varying_fields], 2, function(x){
-      sum(x == 1 & Z.temp == 1)
-    }))
-    BZ[-varying_fields, ] <- t(apply(indicators[, -varying_fields], 2, function(x){
-      sum(x == 1 & Z.temp == 0)
-    }))
+    AZ[-varying_fields, ]  <- indicators %>%
+      select(-varying_fields) %>%
+      filter(Z.temp == 1) %>%
+      colSums() %>%
+      unname()
+
+
+    BZ[-varying_fields, ]  <- indicators %>%
+      select(-varying_fields) %>%
+      filter(Z.temp == 0) %>%
+      colSums() %>%
+      unname()
+
+    # AZ[varying_fields, ] <-
+    #   indicators %>%
+    #   select(varying_fields, linkage_cluster) %>%
+    #   filter(Z.temp == 1) %>%
+    #   split(x= ., .$linkage_cluster)%>%
+    #   map(colSums)%>%
+    #   map(~.x[-length(.x)])%>%
+    #   do.call(rbind, .)
+
+    # thing <- indicators %>%
+    #   select(varying_fields, linkage_cluster) %>%
+    #   filter(Z.temp == 1) %>%
+    #   nest_by(linkage_cluster)
+    #
+    # unlist(lapply(thing$data, function(x){
[…]

### 2021-03-03 R/NLC_functions.R

@@ -14,7 +14,7 @@ BKSimple <- function(comparisons, m_prior, u_prior,
 
   fields <- length(comparisons[[4]])
   n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
-  indicators <- comparisons[[1]]
+  indicators <- data.frame(comparisons[[1]])
 
   parameter.indicators <- as.vector(unlist(sapply(1:fields, function(x){
     rep(x, comparisons[[4]][x])
@@ -32,12 +32,18 @@ BKSimple <- function(comparisons, m_prior, u_prior,
 
   # Gibbs
   for(s in 1:S){
-    aZ <- apply(indicators, 2, function(x){
-      sum(x == 1 & Z.temp == 1)
-    })
-    bZ <- apply(indicators, 2, function(x){
-      sum(x == 1 & Z.temp == 0)
-    })
+    aZ <- indicators %>%
+      filter(Z.temp == 1) %>%
+      colSums()
+    bZ <- indicators %>%
+      filter(Z.temp == 0) %>%
+      colSums()
+    # aZ <- apply(indicators, 2, function(x){
+    #   sum(x == 1 & Z.temp == 1)
+    # })
+    # bZ <- apply(indicators, 2, function(x){
+    #   sum(x == 1 & Z.temp == 0)
+    # })
 
     m.post <- m_prior + aZ
     u.post <- u_prior + bZ
@@ -64,7 +70,7 @@ BKSimple <- function(comparisons, m_prior, u_prior,
       sample(c(candidates, n1 + 1), 1, prob = c(x, offset))
     }))
 
-    Z.temp <- sapply(Z, function(x){
+    Z.temp <- as.vector(sapply(Z, function(x){
       if(x < n1 + 1){
         vec <- rep(0, n1)
         vec[x] <- 1
@@ -72,7 +78,7 @@ BKSimple <- function(comparisons, m_prior, u_prior,
       }else{
         rep(0, n1)
       }
-    })
+    }))
     L <- sum(Z < n1 + 1)
 
     Z.SAMPS[,s] <- Z
@@ -110,7 +116,7 @@ Sadinle17 <- function(comparisons, m_prior, uprior,
 
   fields <- length(comparisons[[4]])
   n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
-  indicators <- comparisons[[1]]
+  indicators <- data.frame(comparisons[[1]])
 
[…]

### 2021-03-03 R/SimulateComparisons.R

@@ -88,3 +88,95 @@ SimulateComparisonsLC <- function(m, u, levels, varying_fields,
   Ztrue[df2matches] <- df1matches
   list(indicators, n1, n2, levels, Ztrue, varying_fields, n1_vec, n2_vec)
 }
+
+
+SimulateComparisonsLC2 <- function(m, u, levels, varying.fields,
+                                  n1, n2, n1.vec, n2.vec, overlap.vec){
+  #
+  #
+  #
+  #
+  #
+  N <- n1*n2
+  ids <- expand.grid(1:n1, 1:n2)
+  parameter.indicators <- as.vector(unlist(sapply(1:length(levels), function(x){
+    rep(x, levels[x])
+  })))
+
+  K <- dim(m)[1]
+  N.vec <- n2.vec * n1;
+  overlap <- sum(overlap.vec)
+
+  # n1.breaks <- c(0, cumsum(n1.vec))
+  n2.breaks <- c(0, cumsum(n2.vec))
+  # n1.lc <- lapply(1:K, function(x){
+  #   (n1.breaks[x]+1):n1.breaks[x+1]
+  # })
+  n2.lc <- lapply(1:K, function(x){
+    (n2.breaks[x]+1):n2.breaks[x+1]
+  })
+
+  lc <- as.vector(sapply(1:K, function(x){
+    rep(x, n2.vec[x])
+  }))
+
+  lc.ids <- split(1:n2, lc)
+  lc.info <- NULL
+  for(k in 1:K){
+    lc.info[[k]] <- list(lc.ids[[k]], overlap.vec[k])
+  }
+
+  lc.long <- as.vector(sapply(1:K, function(x){
+    rep(rep(x, n2.vec[x]), n1)
+  }))
+
+  match.vectors <- nonmatch.vectors <- NULL
+  for(k in 1:K){
+    temp.match <- matrix(NA, nrow = overlap.vec[k], ncol = dim(m)[2])
+    temp.nonmatch <- matrix(NA, nrow = N.vec[k] - overlap.vec[k], ncol = dim(m)[2])
+    m.list <- split(m[k, ], parameter.indicators)
+    u.list <- split(u[k, ], parameter.indicators)
+
+
+    for(i in 1:overlap.vec[k]){
+      temp.match[i,]<- unname(unlist(lapply(m.list, function(x){
+        rmultinom(1, 1, x)
+      })))
+    }
+
+    for(i in 1:(N.vec[k] - overlap.vec[k])){
[…]

### 2021-03-19 notes/Notes_3_19_21.Rmd
Made new set of distributions for simulation

---
title: "Notes 3-18-19"
subtitle: "New Distributions for BRACS Paper"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.height = "80%", out.width = "80%",
                      fig.align = 'center')

library(knitr)
```

## New Distributions

I tried out some new distributions to more closely mirror the voter registration context. Here, I am using first name, last name, race, and party. The $u$ parameters for all of these are somewhat fixed, but the $m$ can be adjusted to weaken or strengthen the variable. When I tried a 5th variable, I ran into the same problem where there was too much information in the data, so you couldn't clearly see the effects of BRACS.

```{r}
include_graphics("figures/new_dist_table.png")
```

## Results

As mentioned before, we see a substantial increase in precision, which also leads to an increase in fmeasure. For the average link probabilities, I split the records by their linkage cluster, looked at the declared matches, and took an average of those posterior match probabilities. On their own, these probabilities do not tell us much, since they could just indicate we are more confident in wrong decisions. However, since we see an increase in fmeasure, I think these numbers are useful. Also, there may be better ways to report or calculate this information, so I'll keep thinking about it.

```{r}
include_graphics("figures/bracs_eval_plot.png")
```

### 2021-03-19 R/BRACS2.R

@@ -47,7 +47,16 @@ BRACS_BK2 <- function(comparisons, m_prior, u_prior,
   L <- 0
   AZ <- BZ <- matrix(0, nrow = length(parameter_split), ncol = K)
   m <- u <- matrix(0, nrow = length(parameter_split), ncol = K)
+
   indicators <- data.frame(indicators_raw, linkage_cluster, Z.temp)
+  varnames <- str_c("X", 1:dim(indicators_raw)[2])
+  names(indicators[1:dim(indicators_raw)[2], ]) <- varnames
+  indicators <- indicators %>%
+    rename_with(.fn = ~stringr::str_replace(., "X", "V"),
+                .cols = varying_fields)
+
+
+
 
   # Gibbs
   for(s in 1:S){
@@ -66,7 +75,7 @@ BRACS_BK2 <- function(comparisons, m_prior, u_prior,
     #   unname()
 
     counts <- indicators %>%
-      select(-varying_fields, Z.temp) %>%
+      select(contains("X"), Z.temp) %>%
       group_by(Z.temp, .drop = FALSE) %>%
       summarise(across(.cols = contains("X"),
                        .fns = sum))
@@ -83,27 +92,27 @@ BRACS_BK2 <- function(comparisons, m_prior, u_prior,
 
 
     lc_counts <- indicators %>%
-      select(varying_fields, Z.temp, linkage_cluster) %>%
+      select(contains("V"), Z.temp, linkage_cluster) %>%
       group_by(linkage_cluster, Z.temp, .drop = FALSE) %>%
-      summarise(across(.cols = contains("X"),
+      summarise(across(.cols = contains("V"),
                        .fns = sum),
                 .groups = "drop")
 
     AZ[varying_fields, ] <- lc_counts %>%
       filter(Z.temp == 1) %>%
-      select(contains("X")) %>%
+      select(contains("V")) %>%
       as.matrix() %>%
       unname() %>%
       t()
 
     BZ[varying_fields, ] <- lc_counts %>%
       filter(Z.temp == 0) %>%
-      select(contains("X")) %>%
+      select(contains("V")) %>%
       as.matrix() %>%
       unname() %>%
       t()
 
-    # AZ[varying_fields, ] <-
+
     #   indicators %>%
[…]

### 2021-03-21 paper/base_model_specification.Rmd
Updated base_model_specification and implemented hashing

@@
 	\end{flushright}
-	\begin{center} \textbf{Base Model Specification}
+	\begin{center} \textbf{Base Model and New Linkage Cluster Model Specification}
 
@@
 
-To indicate matching status, we adopt the *linkage structure parameter* $\mathbf{Z} = (Z_1, \ldots, Z_{n_A})$ from Sadinle 2017, defined as
+To indicate matching status, we adopt the *linkage structure parameter* $\mathbf{Z} = (Z_1, \ldots, Z_{n_B})$ from Sadinle 2017, defined as
 $$Z_j=\begin{cases} 
@@
 # Model Specification
+
+(This will be revised to talk more expliticly about sampling $Z_j$ independently)
 
@@
 $$\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))$$
-where $\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}$ and $\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}$.
+where $\alpha_{fl}(\mathbf{Z})= \sum_{i,j} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}$ and $\beta_{fl}(\mathbf{Z})= \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}$.
 
-\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
 Thus we have
@@
 
-\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
 Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to
-$$p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
-$$\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}$$
-$$=
+
+\begin{align*}
+p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\\
+&\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} && \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\\
+&=
 \begin{cases} 
@@
     1 &  z_j  = n_A + 1 \\
-\end{cases}$$
+\end{cases}\\
+\end{align*}
 
@@
 
-Here, we integrate over the posterior of $\lambda$ and rearrange terms to produce the final full conditional:
+In order to make fair comparisons against the Sadinle 2017 model, we integrate over the posterior of $\lambda$ and rearrange terms to produce the final full conditoinal:
 
@@
 \end{cases}$$
+
+## Decouple/Recouple
+
+(This will be revised; I'm not sure if I really believe the "Recouple" step anymore! I'll need look into it more on larger datasets)
+
+One difficulty in record linkage is conducting linkage while taking appropriate dependencies into account.
+In the context of this paper in which we assume that each file is recorded without duplicates, this means our linkage procedure should respect one-to-one matching.
+The posterior sampling of $\mathbf{Z}$ outlined above resolves half of this issue; since we only sample one potential match for record $j$ in each Gibbs iteration, it is impossible to match $j$ to two records at once.
+This in itself is an improvement over the classic Fellegi-Sunter method that labels record pairs as matches or non matches independently of all other pairs.
+However, our posterior sampling of $\mathbf{Z}$ does allow for the possibility that $Z_j = i = Z_{j'}$ for some $j' \neq j$, which cleary violates the assumption that $\mathbf{X}_B$ is recorded without duplicates.
+Sadinle 2017 resolves this issue by adapting the set of candidate matches for every $Z_j$, but we forgo this rigidity and instead "decouple" the linkage vector $\mathbf{Z}$ for the computational advantantages of sampling $Z_j$ independently.
+In our formulation, such illegal matches occasionally occur within individual Gibbs iterations due to simple randomness, but they do not occur consistently enough between the same records to measurably influence the eventual Bayes estimate of the linkage structure.
+
+However, if we want to ensure the theoretic purity of each MCMC sample of the linkage structure, we can "recouple" $\mathbf{Z}$ by simply resolving these illegal matchings.
+Explictly, define sets $R(i) := \{j | Z_j = i\}$, and define all sets such that $|R(i)| > 1$ as *resolution sets*.
+Within each set, resample the candidate matches according to their Fellegi-Sunter weights.
+We show this step to be computationally inexpensive and to be a reasonable replication of more organic one-to-one matching.
+This procedure would be most useful in cases where there are many "near matches" across files, but in most cases, we find this this step to be unnecessary.
+
+## Bayes Estimate
+
+Same as Sadinle.
+
+## Efficient Computation
+
+Adopting some from Enamorado 2019, some steps my own. very promising, stay tuned!
+
+\newpage
+
+## Proposed Linkage Cluster Model
+
+(Still thinking through best notation and terminology)
+
+The primary contribution of this paper is the establishment of linkage clusters, which allow the reliability parameters $\mathbf{m}^f$, the discernment parameters $\mathbf{u}^f$, and also the linkage probability parameter $\lambda$ to vary according to certain attributes in the record pairs.
+We shall see that this set up upweights linkage within clusters and downweights linkage across clusters, thus using attribute information in a way not yet explored in the literture.
+
+Partition files $A$ and $B$ into *record clusters* $\{A_1, \ldots, A_k\}$ and $\{B_1, \ldots, B_k\}$ according to some attribute in the data.
+We assign the record pair  $(i,j)$ to a *linkage cluster* $S_k \in \{S_1, \ldots, S_K\}$ only when both records belong the corresponding record clusters; otherwise we assign it to $S_0$.
+For example
+$$i \in A_1, j \in B_1 \implies \gamma_{ij} \in S_1$$
+$$i \in A_1, j \in B_2 \implies \gamma_{ij} \in S_0$$
+Cluster assignments at the record pair level are necessary for the updating the $\mathbf{m}$ and $\mathbf{u}$ parameters.
+We denote the reliability and discernment parameters assoicated with each field for each cluster as $\mathbf{m}^{fk}$ and $\mathbf{u}^{fk}$ respectively.
+To improve readability, we modify the definition of $\mathbf{m}_{ij}$ to incorporate the linkage cluster assignment.
+Explicitly,
+$$\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\prod_{k=0}^{K}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}$$
+with an analogous definition for $\mathbf{u}_{ij}$.
+
+Additionally, define the vector of linkage probabilities $\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})$, where $\lambda_{k1} = P(z_{j\in S_k} \in S_k)$ is the probability that a record in $B_k$ has a match within its cluster in $A_k$, $\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})$ is the probability that such a record $B_k$ has a match outside of its cluster, in  $\mathbf{X}_{A_{k'}}$ for some $k' \neq k$, and $\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)$ is the probability of a non-match.
+This is a multinomial distribution where we necessarily have $\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}$.
+These parameters use linkage cluster assignment at the record level.
+
+Our new model specification becomes
+
+$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
+
+$$\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})$$
+
+$$\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})$$
+$$Z_{j \in S_k} | \Lambda_k =
+\begin{cases} 
+    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
+    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}$$
+
+$$\Lambda \sim \text{Dirichlet}(\Lambda_0)$$
+where $\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).$
+
+**THIS WOULD GO IN APPENDIX**: Through similar derivations, this yields full conditionals:
+The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have 
+$$\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))$$
+
+$$\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))$$
+where $$\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}$$
+
+$$\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij} \in S_k}$$
+
+We also have straightforward posterior distritutions for the linkage probability vector $\Lambda_k$:
+
+$$\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)$$
+
+where $\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})$ with 
+$$D_{k1} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_k}$$ $$D_{k2} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_{k'}}$$ $$\text{and } D_{k3} = \sum_{j \in B_k} \mathbf{1}_{z_j = n_A + 1}$$
+
+Our full conditional for $Z_j$ follows a familiar form:
+
+$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
+\begin{cases} 
+    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
+    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}$$
+
+And after integrating over $\Lambda_k$ and rearranging terms,
+
+$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
+\begin{cases} 
+    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
+    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
+     1 & z_j = n_A + 1 \\
+\end{cases}$$

### 2021-03-21 paper/base_model_specification.tex
Updated base_model_specification and implemented hashing

@@
     \end{flushright}
-    \begin{center} \textbf{Base Model Specification}
+    \begin{center} \textbf{Base Model and New Linkage Cluster Model Specification}
 
@@
 To indicate matching status, we adopt the \emph{linkage structure
-parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_A})\) from Sadinle 2017,
+parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_B})\) from Sadinle 2017,
 defined as \[Z_j=\begin{cases} 
@@
 \section{Model Specification}\label{model-specification}}
+
+(This will be revised to talk more expliticly about sampling \(Z_j\)
+independently)
 
@@
 where
-\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
+\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
 and
-\(\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).
+\(\beta_{fl}(\mathbf{Z})= \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).
 
-\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$} As a function of
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
 \(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
@@
 
-\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$}
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
 Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
@@
 is a discrete distribution with probalities proportional to
-\[p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
-\[\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\]
-\[=
+
+\begin{align*}
+p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\\
+&\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} && \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\\
+&=
 \begin{cases} 
@@
     1 &  z_j  = n_A + 1 \\
-\end{cases}\]
+\end{cases}\\
+\end{align*}
 
@@
 
-Here, we integrate over the posterior of \(\lambda\) and rearrange terms
-to produce the final full conditional:
+In order to make fair comparisons against the Sadinle 2017 model, we
+integrate over the posterior of \(\lambda\) and rearrange terms to
+produce the final full conditoinal:
 
@@
 
+\hypertarget{decouplerecouple}{%
+\subsection{Decouple/Recouple}\label{decouplerecouple}}
+
+(This will be revised; I'm not sure if I really believe the ``Recouple''
+step anymore! I'll need look into it more on larger datasets)
+
+One difficulty in record linkage is conducting linkage while taking
+appropriate dependencies into account. In the context of this paper in
+which we assume that each file is recorded without duplicates, this
+means our linkage procedure should respect one-to-one matching. The
+posterior sampling of \(\mathbf{Z}\) outlined above resolves half of
+this issue; since we only sample one potential match for record \(j\) in
+each Gibbs iteration, it is impossible to match \(j\) to two records at
+once. This in itself is an improvement over the classic Fellegi-Sunter
+method that labels record pairs as matches or non matches independently
+of all other pairs. However, our posterior sampling of \(\mathbf{Z}\)
+does allow for the possibility that \(Z_j = i = Z_{j'}\) for some
+\(j' \neq j\), which cleary violates the assumption that
+\(\mathbf{X}_B\) is recorded without duplicates. Sadinle 2017 resolves
+this issue by adapting the set of candidate matches for every \(Z_j\),
+but we forgo this rigidity and instead ``decouple'' the linkage vector
+\(\mathbf{Z}\) for the computational advantantages of sampling \(Z_j\)
+independently. In our formulation, such illegal matches occasionally
+occur within individual Gibbs iterations due to simple randomness, but
+they do not occur consistently enough between the same records to
+measurably influence the eventual Bayes estimate of the linkage
+structure.
+
+However, if we want to ensure the theoretic purity of each MCMC sample
+of the linkage structure, we can ``recouple'' \(\mathbf{Z}\) by simply
+resolving these illegal matchings. Explictly, define sets
+\(R(i) := \{j | Z_j = i\}\), and define all sets such that
+\(|R(i)| > 1\) as \emph{resolution sets}. Within each set, resample the
+candidate matches according to their Fellegi-Sunter weights. We show
+this step to be computationally inexpensive and to be a reasonable
+replication of more organic one-to-one matching. This procedure would be
+most useful in cases where there are many ``near matches'' across files,
+but in most cases, we find this this step to be unnecessary.
+
+\hypertarget{bayes-estimate}{%
+\subsection{Bayes Estimate}\label{bayes-estimate}}
+
+Same as Sadinle.
+
+\hypertarget{efficient-computation}{%
+\subsection{Efficient Computation}\label{efficient-computation}}
+
+Adopting some from Enamorado 2019, some steps my own. very promising,
+stay tuned!
+
+\newpage
+
+\hypertarget{proposed-linkage-cluster-model}{%
+\subsection{Proposed Linkage Cluster
+Model}\label{proposed-linkage-cluster-model}}
+
+(Still thinking through best notation and terminology)
+
+The primary contribution of this paper is the establishment of linkage
+clusters, which allow the reliability parameters \(\mathbf{m}^f\), the
+discernment parameters \(\mathbf{u}^f\), and also the linkage
+probability parameter \(\lambda\) to vary according to certain
+attributes in the record pairs. We shall see that this set up upweights
+linkage within clusters and downweights linkage across clusters, thus
+using attribute information in a way not yet explored in the literture.
+
+Partition files \(A\) and \(B\) into \emph{record clusters}
+\(\{A_1, \ldots, A_k\}\) and \(\{B_1, \ldots, B_k\}\) according to some
+attribute in the data. We assign the record pair \((i,j)\) to a
+\emph{linkage cluster} \(S_k \in \{S_1, \ldots, S_K\}\) only when both
+records belong the corresponding record clusters; otherwise we assign it
+to \(S_0\). For example
+\[i \in A_1, j \in B_1 \implies \gamma_{ij} \in S_1\]
+\[i \in A_1, j \in B_2 \implies \gamma_{ij} \in S_0\] Cluster
+assignments at the record pair level are necessary for the updating the
+\(\mathbf{m}\) and \(\mathbf{u}\) parameters. We denote the reliability
+and discernment parameters assoicated with each field for each cluster
+as \(\mathbf{m}^{fk}\) and \(\mathbf{u}^{fk}\) respectively. To improve
+readability, we modify the definition of \(\mathbf{m}_{ij}\) to
+incorporate the linkage cluster assignment. Explicitly,
+\[\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\prod_{k=0}^{K}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}\]
+with an analogous definition for \(\mathbf{u}_{ij}\).
+
+Additionally, define the vector of linkage probabilities
+\(\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})\), where
+\(\lambda_{k1} = P(z_{j\in S_k} \in S_k)\) is the probability that a
+record in \(B_k\) has a match within its cluster in \(A_k\),
+\(\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})\) is the probability that
+such a record \(B_k\) has a match outside of its cluster, in
+\(\mathbf{X}_{A_{k'}}\) for some \(k' \neq k\), and
+\(\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)\) is the probability of a
+non-match. This is a multinomial distribution where we necessarily have
+\(\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}\). These parameters use
+linkage cluster assignment at the record level.
+
+Our new model specification becomes
+
+\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
+
+\[\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})\]
+
+\[\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})\]
+\[Z_{j \in S_k} | \Lambda_k =
+\begin{cases} 
+    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
+    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}\]
+
+\[\Lambda \sim \text{Dirichlet}(\Lambda_0)\] where
+\(\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).\)
+
+\textbf{THIS WOULD GO IN APPENDIX}: Through similar derivations, this
+yields full conditionals: The \(\mathbf{m}\) and \(\mathbf{u}\)
+parameters are updated through standard multinomial-dirichlet mechanics.
+Thus we have
+\[\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))\]
+
+\[\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))\]
+where
+\[\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}\]
+
+\[\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij} \in S_k}\]
+
+We also have straightforward posterior distritutions for the linkage
+probability vector \(\Lambda_k\):
+
+\[\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)\]
+
+where \(\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})\) with
+\[D_{k1} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_k}\]
+\[D_{k2} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_{k'}}\]
+\[\text{and } D_{k3} = \sum_{j \in B_k} \mathbf{1}_{z_j = n_A + 1}\]
+
+Our full conditional for \(Z_j\) follows a familiar form:
+
+\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
+\begin{cases} 
+    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
+    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
+     \lambda_{k3} & z_j = n_A + 1 \\
+\end{cases}\]
+
+And after integrating over \(\Lambda_k\) and rearranging terms,
+
+\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
+\begin{cases} 
+    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
+    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
+     1 & z_j = n_A + 1 \\
+\end{cases}\]
+
 \end{document}

### 2021-03-21 R/GetUniquePatterns.R

@@ -1,0 +1,92 @@
+GetUniquePatterns <- function(comparisons, fast = F){
+
+  indicators <- comparisons[[1]]
+  N <- dim(indicators)[1]
+  levels <- comparisons[[4]]
+  fields <- length(levels)
+  gk <- unlist(lapply(1:fields, function(x){
+    0:(levels[x] -1)
+  }))
+  Lf_vec <- unlist(lapply(1:fields, function(x){
+    rep(levels[x], levels[x])
+  }))
+
+  fields_vec <- unlist(lapply(1:fields, function(x){
+    rep(seq_len(fields)[x], levels[x])
+  }))
+
+
+  #gk_ij <- sweep(indicators, MARGIN = 2, gk, FUN = "*")
+  # Lf_long <- sapply(Lf_vec, function(x){
+  #   rep(x, N)
+  # })
+  # fields_long <- sapply(fields_vec, function(x){
+  #   rep(x, N)
+  # })
+
+  level_sum <- cumsum(levels)
+  level_sum_vec <-  unlist(lapply(1:fields, function(x){
+    rep(level_sum[x], levels[x])
+  }))
+
+  # level_sum_long <- sapply(level_sum_vec, function(x){
+  #   rep(x, N)
+  # })
+
+  #hash <- rowSums((gk_ij >0) * 2 ^ (gk_ij + (fields_long > 1) * level_sum_long))
+  if(fast){
+    numCores <- parallel::detectCores()
+    cl <- parallel::makeCluster(numCores)
+
+    parallel::clusterExport(cl, c("gk", "fields_vec", "level_sum_vec"))
+    hash_temp <- parallel::parApply(cl, indicators, 1, FUN = function(x){
+      gk_ij <- x * gk
+      (gk_ij >0) * 2 ^ (gk_ij + (fields_vec > 1) * level_sum_vec)
+    })
+
+    parallel::stopCluster(cl)
+
+  } else {
+
+    hash_temp <- apply(indicators, 1, function(x){
+      gk_ij <- x * gk
+      (gk_ij >0) * 2 ^ (gk_ij + (fields_vec > 1) * level_sum_vec)
+    })
+
+
+    # thing <- lapply(seq_len(nrow(indicators)), function(x){
+    #   indicators[x, ]
+    # })
+    #
+    # hash_temp <- lapply(thing, function(x){
+    #   gk_ij <- x * gk
[…]

### 2021-03-21 R/NLC_hash.R

@@ -1,0 +1,150 @@
+BKSimple_hash <- function(comparisons, m_prior, u_prior,
+                      alpha, beta, S, burn, show_progress = T, fast =F){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRecords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
+  # alpha = first parameter of prior for linkage probability
+  # beta = second parameter of prior for linkage probability
+  # S = number of Gibbs iterations
+  # burn = number of iterations to be discarded as burn-in
+  # show_progress = set to false to show simulation progress
+
+  fields <- length(comparisons[[4]])
+  n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
+  indicators_raw <-comparisons[[1]]
+
+  if (fast) {
+    numCores <- parallel::detectCores()
+    cl <- parallel::makeCluster(numCores)
+    }
+
+  patterns <- GetUniquePatterns(comparisons, fast)
+
+  parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
+    rep(x, comparisons[[4]][x])
+  })))
+
+  ids <- expand.grid(1:n1, 1:n2)
+  candidates <- 1:n1
+  Z.SAMPS <- matrix(NA, nrow = n2, ncol = S)
+  M.SAMPS <- matrix(NA, nrow = dim(indicators_raw)[2], ncol = S)
+  U.SAMPS <- matrix(NA, nrow = dim(indicators_raw)[2], ncol = S)
+  L.SAMPS <- vector(length = S)
+  Z.temp <- rep(0, n1*n2)
+  Z <- rep(n1+1, n2)
+  L <- 0
+  #AZ <- BZ <- rep(0, length(parameter_split))
+  m <- u <- rep(0, length(parameter_split))
+
+  #indicators <- data.frame(indicators_raw, Z.temp)
+
+
+  hash_id <- patterns[[1]]
+  hash <- data.frame(hash_id)
+  unique_patterns <- patterns[[2]]
+  pattern_counts <- patterns[[3]]
+  P <- dim(unique_patterns)[1]
+
+
+  # Gibbs
+  for(s in 1:S){
+
+    hash$Z.temp <- Z.temp
[…]

### 2021-03-21 paper/linkagecluster_bk_paper.Rmd
hashing is awesome

@@
 However, if we want to ensure the theoretic purity of each MCMC sample of the linkage structure, we can "recouple" $\mathbf{Z}$ by simply resolving these illegal matchings.
-Explictly, define clusters $R(i) := \{j | Z_j = i\}$, and define all clusters such that $|R(i)| > 1$ as *resolution clusters*.
-Within each cluster, resample the candidate matches according to their Fellegi-Sunter weights.
+Explictly, define sets $R(i) := \{j | Z_j = i\}$, and define all sets such that $|R(i)| > 1$ as *resolution sets*.
+Within each set, resample the candidate matches according to their Fellegi-Sunter weights.
 We show this step to be computationally inexpensive and to be a reasonable replication of more organic one-to-one matching.

### 2021-03-23 notes/Notes_3_26_21.Rmd
Beginning speed and accuracy tests, and uploaded 3/26 meeting notes

---
title: "Unveiling parlr"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
library(RecordLinkage)
library(tidyverse)
```

## Overview

- Computational Speed-ups
- Next Step

## parlr

- I have been playing around with calling my method "\underline{Pa}rallelized \underline{R}ecord \underline{L}inkage in \underline{R}", or parlr for short.

- Broadly speaking, I use two strategies for computational speedup:
  - Performing operations on the **unique aggreement patterns** rather than the individual record pairs. Rigorously, sums over the unique patterns are *sufficient statistics*.
  - Relaxation of one to one matching to allow for parallel sampling of $Z$

Together, these have made **huge** improvements on code

## Identifying Unique Patterns

- Enamorado uses the following hashing function to identify unique aggreement patterns

$$H(i,j) = \sum_{f=1}^F \mathbf{1}_{\gamma_f (i,j) >0}2^{\gamma_f (i,j) + \mathbf{1}_{f>1} \times \sum_{a=1}^{f-1}(L_a -1)}$$

- I identify the unique patterns, the number of each of of those patterns in the data as a whole, and the number of those patterns for each record $j \in B$.

- Define $P$ to be the number of unique agreement patterns, and note it is bounded by $\prod_{f =1}^F L_f$.

## Identifying Unique Patterns

- Each agreement pattern corresponds to a particular type of contribution to the posteriors of the $m$ and $u$ parameters. 
- To update $m$, I used to sum over each variable for all $N$ record pairs. I simply identify the number of matchings in the $Z$ vector corresponding to each agreement of $P$ patterns.

- The nonmatches are even easier: just substract this number from the total present in the data!

## Calculating FS weights

- Again, I used to do this for each of $N$ records, but now I calculate this for the $P$ unique patterns.

- Just these two changes sped up the code considerably. However, I thought of one other clever trick to really push this idea to the max.

## Breaking up the Z sampler

- Four observations
  - Sampling many objects with varying probabilities difficult
  - Sampling *fewer* objects with varying probabilities is easy
  - Sampling many objects with *uniform* probabilities is also easy
  - For a record $j \in B$, each record $i\in A$ with a given agreement pattern has the same probability of matching. That is, the unique patterns form *equivalence classes*, and within a class, each record is equally likely.

## Breaking up the Z sampler

- Rather than sampling $Z_j$ from $N$ many options, I instead sample it from the $P$ unique patterns, each one multiplied by its prevalance for record $j$.

- After sampling the unique pattern, I sample the record associated with that pattern uniformly at random.

- Since all calculations within the Gibbs Sampler depend only on the unique pattern, and since my model samples the $Z_j$ independently,  I can sample for the records *at the end of the entire Gibbs procedure*

## Summary

- We have *almost entirely* removed dependence on $N$ from our Gibbs procedure, and replaced them with steps that depend on $P$.

- Filling in the Gibbs sampler post-hoc technically depends on $N$, but its a very inexpensive step that only has to be done once

- Only major hurdle with $N$ is calculating the comparison vectors, which is just unavoidable in the Fellegi Sunter Framework

## Speed up

```{r}
knitr::include_graphics("figures/speed_plot.png")
```

## Speed up

- When $n_A$ and $n_B$ are small, the computational savings are minimal, and Sadinle's `BRL` runs faster just because its in C.

- As the datasets grow bigger, computation for `BRL` grows *quadratically*, while `parlr` seems to grow *linearly*

- The plot shown is with **no parallel computing**

- If my method was programmed in C, I would expect even greater computational savings

## Accuracy

```{r}
knitr::include_graphics("figures/acc_plot.png")
```

## Accuracy

```{r}
knitr::include_graphics("figures/acc_plot_pairwise.png")
```

## Accuracy

- Overal, the `parlr` method peforms comparably to Sadinle17. There is slight evidence that recall is systematically lower for `parlr` at high error levels, but it is *very* slight. I'll need to run this at different level of overlap to explore more.

## Next Steps

- I have not yet implemented any parallelization. We will cover this soon is Shawns class. The steps above however provide *the vast majority* of the computational speed up.

- Later on, I'll be able to use this same hashing function for linkage clusters by replacing 2 with the $k^{th}$ prime number for the $k^{th}$ linkage cluster. (For Jody's paper, I'll present just the standard method)

## Next Steps

- This model and these computational savings really inform everything everything I'm doing. Apart from the BRACS paper with Jody, I'm really going to focus on getting this done

- Larger versions of the simulations above, with different levels of overlap.

- \textcolor{blue}{Question:} What other kinds of simulations or evaluations would I need?

- \textcolor{blue}{Question:} Is this still a good fit for Bayesian Analysis? Or should it go to a journal more concerned with computation and scalability?

### 2021-03-23 R/CleanGibbs.R

@@ -1,0 +1,5 @@
+sample_with_1 <- function(x, size, replace = F, prob = NULL) {
+  if (length(x) == 1) return(x)
+  base::sample(x, size = size, replace = replace, prob = prob)
+}
+

### 2021-03-23 R/NLC_function2.R

@@ -112,4 +112,5 @@ BKSimple2 <- function(comparisons, m_prior, u_prior,
   #U.SAMPS <- U.SAMPS[,-(1:burn)]
 
   Z.SAMPS
+
 }

### 2021-03-23 R/NLC_hash2.R

@@ -1,0 +1,188 @@
+BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
+                          alpha = 1, beta = 1, S = 1000, burn = 100, show_progress = T, fast =F){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRecords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
+  # alpha = first parameter of prior for linkage probability
+  # beta = second parameter of prior for linkage probability
+  # S = number of Gibbs iterations
+  # burn = number of iterations to be discarded as burn-in
+  # show_progress = set to false to show simulation progress
+
+  fields <- length(comparisons[[4]])
+  n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
+  indicators_raw <-comparisons[[1]]
+
+  if (fast) {
+    numCores <- parallel::detectCores()
+    cl <- parallel::makeCluster(numCores)
+  }
+
+  patterns <- GetUniquePatterns2(comparisons, fast)
+
+  parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
+    rep(x, comparisons[[4]][x])
+  })))
+
+  ids <- expand.grid(1:n1, 1:n2)
+  #rec2 <- ids[,2]
+
+  hash_id <- patterns[[1]]
+  hash <- data.frame(hash_id)
+  #hash_split <- split(hash_id, ids[,2])
+  unique_patterns <- patterns[[2]]
+  pattern_counts <- patterns[[3]]
+  P <- dim(unique_patterns)[1]
+  counts_by_rec <- patterns[[4]]
+  hash_to_rec1 <- patterns[[5]]
+  candidates <- 1:n1
+  candidates_P <- 1:(P+1)
+  Z.SAMPS <- matrix(NA, nrow = n2, ncol = S)
+  M.SAMPS <- matrix(NA, nrow = dim(indicators_raw)[2], ncol = S)
+  U.SAMPS <- matrix(NA, nrow = dim(indicators_raw)[2], ncol = S)
+  L.SAMPS <- vector(length = S)
+  Z.temp <- rep(0, n1*n2)
+  Z <- rep(n1+1, n2)
+  L <- 0
[…]

### 2021-03-23 R/SimulateComparisons.R

@@ -29,7 +29,11 @@ SimulateComparisons <- function(m, u, levels, n1, n2, overlap){
     rmultinom(N - overlap, 1, x)
   })
   indicators[-match_index,] <- t(do.call(rbind, gamma_nonmatch))
-  list(indicators, n1, n2, levels, Ztrue)
+  list(comparisons = indicators,
+       n1 = n1,
+       n2 = n2,
+       nDisagLevs = levels,
+       Ztrue = Ztrue)
 }
 
 SimulateComparisonsLC <- function(m, u, levels, varying_fields,

### 2021-03-24 notes/Notes_3_26_21.Rmd
plot with constant nB, sampling demonstration

@@
 
-- Enamorado uses the following hashing function to identify unique aggreement patterns
+- Enamorado uses the following hashing function to identify unique aggreement patterns. I've also figured out how to do this efficiently in `dplyr`.
 
@@
 
+## Calculating FS weights
+
+- Within a Gibbs iteration, all records with the same agreement pattern will have the same FS weight.
+So instead of calculating the weight for each of $n_A \times n_B$ records, I only need to do it for $P$ unique patterns.
+
+- This step is taken directly from `fastlink`.
+
 ## Identifying Unique Patterns
 
+- `fastlink` doesn't have posterior updates for Dirichlet distributions, so I came up with this step on my own.
+
 - Each agreement pattern corresponds to a particular type of contribution to the posteriors of the $m$ and $u$ parameters.
-- To update $m$, I used to sum over each variable for all $N$ record pairs. I simply identify the number of matchings in the $Z$ vector corresponding to each agreement of $P$ patterns.
+
+- To update $m$, I used to sum over each variable for all $N$ record pairs. Now, I simply identify the number of matchings in the $Z$ vector corresponding to each of $P$ patterns.
 
 - The nonmatches are even easier: just substract this number from the total present in the data!
-
-## Calculating FS weights
-
-- Again, I used to do this for each of $N$ records, but now I calculate this for the $P$ unique patterns.
-
-- Just these two changes sped up the code considerably. However, I thought of one other clever trick to really push this idea to the max.
 
@@
 
-- Rather than sampling $Z_j$ from $N$ many options, I instead sample it from the $P$ unique patterns, each one multiplied by its prevalance for record $j$.
+- Rather than sampling $Z_j$ sequentially from $N$ many records, I can sample it in parallel from $P$ many equivalence classes of unique patterns, each one multiplied by its prevalance for record $j$.
 
@@
 
+- Only major hurdle with $N$ is calculating the comparison vectors, which is just unavoidable in the Fellegi Sunter Framework. However, whatever Ted does for `fastlink` would work here too.
+
 - Filling in the Gibbs sampler post-hoc technically depends on $N$, but its a very inexpensive step that only has to be done once
-
-- Only major hurdle with $N$ is calculating the comparison vectors, which is just unavoidable in the Fellegi Sunter Framework
 
@@
 - If my method was programmed in C, I would expect even greater computational savings
+
+## More on Speed Up
+
+- Within the Gibbs sampler, calculations only depend on $n_B$ (the number samples I need to take to fill a $Z$ vector), and $P$ (the number of unique patterns).
+$n_A$ only influences the scale of the counts for the unique patterns.
+
+- $n_A$ only comes to play in forming the comparison vectors and in back-filling the samples of $Z$ at the end of the sampler.
+
+- This suggests that the sampler should be *almost constant* in $n_A$ (ie, the Gibbs sampling for a $4000 \times 500$ problem should not be substantially longer than for a $500 \times 500$ problem)
+
+- Particularly useful for linking a smaller dataset to a much larger dataset. One example would be the DNC data in BRACS!
+
+## More on Speed Up
+
+```{r}
+knitr::include_graphics("figures/speed_plot_fixed_nB.png")
+```
 
@@
 
+## Limitations
+
+- The original FS method is $O(n_A \times n_B)$ hard.
+With hashing methods, `fastlink` makes the problem only $O(P)$ hard, since record pairs are considered indepenently and classification is based *only* on the unique pattern.
+
+- Hashing methods in my model make the problem $O(n_B \times P)$ hard.
+No matter what, I need to sample $n_B$ many times to fill a linkage vector $Z$.
+In this sense, `fastlink` is inherently faster than this method
+
+- However, method takes into consider much more of the dependency structure in the data, making this computational burden worthwhile
+
+- \textcolor{blue}{Note:} If this line of thinking is useful for the paper, I'll need some help making this rigorous!
+
 ## Next Steps

### 2021-04-10 notes/Notes_4_16_21.Rmd
Conflict Resolution and Meeting Notes

---
title: "Slight Setback: One-to-one Matching"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
library(RecordLinkage)
library(tidyverse)
```

## Overview

- Completed Sadinle Simulation
- Revisiting $n$
- Conflict Resolution

## Sadinle Simulation

- I have completely replicated the Sadinle simulation using `parlr` and `BRL`
- Three levels of error, three levels of overlap, 100 simultations each. Total of 900 simulations

## Sadinle Simulation

```{r}
knitr::include_graphics("figures/sadinle_sim_plot.png")
```

## Sadinle Simulation

- `parlr` shows notably weaker performance only in very high error levels and very high overlap

- This makes sense: the discrepancy between `parlr` and `BRL` will be greatest with high overlap, because this is when `BRL` is removing the most records from the set of potential matches

## Revisiting n

- Last meeting, I had a plot showing the computational savings of `parlr` for different sizes of the data

- Jerry asked whether `parlr` was maintaining comparable performance for the larger datasets. I prematurely said yes.

## Revisiting n

```{r}
knitr::include_graphics("figures/sadinle_speed_plot.png")
```

## Revisiting n

```{r}
knitr::include_graphics("figures/acc_plot_with_size.png")
```

## Whats going on?

- Combing through the results, I realized that this discrepancy was due to pairs that violate one-to-one matching. Such pairings are not possible under `BRL`, but are possible with `parlr`

- Worse, the rate of such matchings occuring *increases* with the size of the data

## Violations of One-to-One

```{r}
knitr::include_graphics("figures/violations_plot.png")
```

## Violations of One-to-One

- This makes sense: all of the nonmatching pairs are generated from the $\mathbf{u}$ distriutions. Simulating "strong" comparison vectors is rare under $\mathbf{u}$, but the chance of simulating *at least one* strong comparison vector to compete with the true match increases with $n_A$

- I also think that this phenomenon might not occur as strongly with real data. Generating comparison vectors themselves is not very close to reality.

- This might be something I want to bring up with Jody/DNC at some point though. (All of that has been pretty quiet recently)

## Conflict Resolution

- The Fellegi-Sunter methods all have a method of conflict resolve to create one-to-one matching posthoc.

- Here, this can simply be done by accepting the match with the highest posterior probability. Rigorously, this indeed minimizes the Bayes risk, since we are assuming only one match can be true. (Might need to make this more rigorous)

- I expect that the number of conflicts to be resolved under this method is *far lower* than under `fastlink`.

## Conflict Resolution

```{r}
knitr::include_graphics("figures/acc_plot_with_size_resolved.png")
```

## Next Steps

- Eventually, I should redo the Sadinle Simulation with the conflict resolution step

- An alternative approach to the conflict resolution step might be an **active learning** method, where the user detects pairs that violate one-to-one matchings while the process is going on.

- Recieved El Salvador data from Beka, haven't looked at it yet

- Using these ideas for my final project in Eric Laber's class

### 2021-04-10 paper/parlr_paper_1.Rmd
Conflict Resolution and Meeting Notes

--- 2021-02-27 paper/base_model_specification.Rmd
@@
 \end{cases}$$
+
+## Bayes Estimate
+
+Our Gibbs sampler provides posterior samples of $\mathbf{Z}$ which we use to make our final decisions about the linkage structure.
+In the case where we false matches and missed matches contribute the same loss, we declare $(i,j)$ to be a match whenever $P(Z_j = i) > \frac{1}{2}$ according to these posterior samples.
+We can create more elaborate decisions by attributing different loss values to different kinds of error, and allowing pairings with middling posterior probabilities to be left unlabled by the algorithm so that the modeller can address those manually.
+For further discussion, see Sadinle 2017.
+
+# Efficient Computation
+
+Broadly speakings, we increase our computational efficiency by tecognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
+Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns present in $\Gamma$, and note that $P$ is bounded above by $\prod_{f=1}^F L_f$.
+We index these agreement patterns by $p \in \{1, \ldots, P\}$, and say $\gamma_{ij} \in h_p$ when the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern.
+Wherever possible, we conduct calculcations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.
+
+## Representation of the Data
+
+The classic Fellegi Sunter represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$.
+To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicators of lengths $L_f$.
+For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework.
+This is a bijective transformive that does not change the meaning of the data, but this representation will ease calculations and posterior updates.
+
+## Hashing
+
+In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair.
+We however do not store these comparison vectors themselves, but instead only the a hashed value indicating the unique agreement pattern.
+We store this information in a nested list.
+For each $p$, we also calculate $|h_p|$ total instances of agreement pattern $h_p$ in the entire $\Gamma$ matrix, and then for each $j$, we also calculate $|h_p|_j$ the total instances agreement pattern for each $\Gamma_{.j}$
+
+## Updating m and u
+
+After receiving matching statuses from $\mathbf{Z}$, the Sadinle method calculates $\alpha_{fl}(\mathbf{Z})$ and $\beta_{fl}(\mathbf{Z})$ for each field and level.
+This constitutes $2 \times \sum L_f$ many summations over $n_A \times n_B$ quantities, and becomes computionally burdensome with large data.
+In contrast, we recognize that each unique agreement pattern contributes to the posterior $\alpha(\mathbf{Z})$ and $\beta(\mathbf{Z})$ vectors in the same way.
+In fact, if we denote $\mathbf{Z}(h_p)$ to be the number of record pairs with agreement pattern $h_p$, then the contribution of pairs of pattern $h_p$ to $\alpha(\mathbf{Z})$ is simply $\mathbf{Z}(h_p) \times h_p$.
+Thus our posterior update is simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p$.
+Then $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)$
+
+## Updating Z
+
+Sadinle uses the following full conditional for $Z_j$ in order to strictly enforce one-to-one matching.
+
+The indicator removes previously matched records from the set of candidate records, and $L_{-j}$ updates iteratively after each matching decision, $n_B$ many times per Gibbs iteration.
+This process is *inherently serial* and becomes computationally burdensome for large data.
+By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel.
+More importantly, since only the aggrement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns.
+This boosts computation time far greater than parallelization.
+
+To do this
+
+##

### 2021-04-10 R/LinkRecordsBK.R

@@ -1,0 +1,74 @@
+LinkRecordsBK<- function(Zchain, n1, lFNM=1, lFM1=1, lFM2=2, lR=Inf){
+  #
+  # This is a complete copy of "linkrecords" from BRL, only modified
+  # so that it passes on the posterior link probabilities
+  #
+  #
+  #
+
+  # control the input
+  if(!is.matrix(Zchain)) stop("Zchain should be a matrix")
+  n2 <- nrow(Zchain)
+  # make sure the labels in Zchain are within the expected range
+  if(max(Zchain) > n1 + n2) stop("Labels in Zchain exceed n1+n2")
+  # - positive losses
+  C0 <- (lFNM > 0) & (lFM1 > 0) & (lFM2 > 0) & (lR > 0)
+  # - conditions of Theorem 1 of Sadinle (2017)
+  C1 <- (lR == Inf) & (lFNM <= lFM1) & (lFNM + lFM1 <= lFM2)
+  # - conditions of Theorem 2 of Sadinle (2017)
+  C2 <- ((lFM2 >= lFM1) & (lFM1 >= 2*lR)) | ((lFM1 >= lFNM) & (lFM2 >= lFM1 + lFNM))
+  # - conditions of Theorem 3 of Sadinle (2017)
+  C3 <- (lFM2 >= lFM1) & (lFM1 >= 2*lR) & (lFNM >= 2*lR)
+  # check we can handle the specified losses
+  if(!C0) stop("Losses need to be positive")
+  if(!any(c(C1,C2,C3))) stop("Invalid configuration of losses")
+
+  # temporarily replace all nonlink labels by n1+1
+  Zchain[Zchain > n1+1] <- n1+1
+  tableLabels <- apply(Zchain, 1, tabulate, nbins=max(Zchain))
+  tableLabels <- tableLabels/ncol(Zchain)
+  probNoLink <- tableLabels[n1+1,]
+  # find marginal best option for each record based only on probability
+  maxProbOption <- apply(tableLabels, 2, which.max)
+  maxProbOption[maxProbOption==n1+1] <- (n1+1:n2)[maxProbOption==n1+1]
+  probMaxProbOption <- apply(tableLabels, 2, max)
+  maxProbOptionIsLink <- maxProbOption <= n1
+
+  if(C1){# if not using reject option and conditions of Theorem 1
+
+    Zhat <- (n1+1):(n1+n2)
+    tholdLink <- lFM1/(lFM1+lFNM) +
[…]

### 2021-04-10 R/ResolveConflicts.R

@@ -1,0 +1,14 @@
+ResolveConflicts <- function(Zhat){
+
+  post_probs <- Zhat[[2]]
+  Zhat <- Zhat[[1]]
+  double_matches <- Zhat[duplicated(Zhat)]
+  to_resolve <- unlist(lapply(double_matches, function(x){
+    dfB_options <- which(Zhat == x)
+    dfB_probs <- post_probs[dfB_options]
+    non_matches <- dfB_options[-which.max(dfB_probs)]
+    non_matches
+  }))
+  Zhat[to_resolve] <- max(Zhat)
+  Zhat
+}

### 2021-04-12 paper/parlr_paper_1.Rmd
First draft of Efficient Computation

@@
 	\end{flushright}
-	\begin{center} \textbf{Base Model and New Linkage Cluster Model Specification}
+	\begin{center} \textbf{Parlr: Parallelized Record Linkage in R}
 
@@
 ```
+
+## Introduction
+
+Most methods for conducting record linkage, the task of identifying duplicate records across datasets, are derived from the foundational article by Fellegi and Sunter in 1969.
+Within this family of methods, two methods have notable strengths: `fastlink`, proposed by Enamorado et al (2019), continues a is able to conduct linkage on incredibly large datasets
 
@@
 
-(This will be revised to talk more expliticly about sampling $Z_j$ independently)
+(This will be revised to talk more explicitly about sampling $Z_j$ independently)
 
@@
 The prior for $Z_j$ has equal probability of matching to all records $i\in A$, and non-matching probability governed by $\lambda$.
-Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$ prior corresponds to a uniform prior on the labelling of $\mathbf{Z}$.
+Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$ prior corresponds to a uniform prior on the labeling of $\mathbf{Z}$.
 
@@
 
-\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of duplicates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$.
 Thus we have
@@
 \underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
-Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to
+Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probabilities proportional to
 
@@
 
-In order to make fair comparisons against the Sadinle 2017 model, we integrate over the posterior of $\lambda$ and rearrange terms to produce the final full conditoinal:
+In order to make fair comparisons against the Sadinle 2017 model, we integrate over the posterior of $\lambda$ and rearrange terms to produce the final full conditional:
 
-$$Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+$$p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
 \begin{cases} 
-    w_{ij}  & z_j \leq n_A; \\
-     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
+    w_{ij}  & i \leq n_A; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
 \end{cases}$$
@@
 In the case where we false matches and missed matches contribute the same loss, we declare $(i,j)$ to be a match whenever $P(Z_j = i) > \frac{1}{2}$ according to these posterior samples.
-We can create more elaborate decisions by attributing different loss values to different kinds of error, and allowing pairings with middling posterior probabilities to be left unlabled by the algorithm so that the modeller can address those manually.
+We can create more elaborate decisions by attributing different loss values to different kinds of error, and allowing pairings with middling posterior probabilities to be left unlabeled by the algorithm so that the modeler can address those manually.
 For further discussion, see Sadinle 2017.
@@
 
-Broadly speakings, we increase our computational efficiency by tecognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns present in $\Gamma$, and note that $P$ is bounded above by $\prod_{f=1}^F L_f$.
-We index these agreement patterns by $p \in \{1, \ldots, P\}$, and say $\gamma_{ij} \in h_p$ when the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern.
-Wherever possible, we conduct calculcations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.
+Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
+Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns.
+Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$.
+We index these agreement patterns by $p \in \{1, \ldots, P\}$, and say $(i,j) \in h_p$ when the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern.
+Wherever possible, we conduct calculations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.
 
-## Representation of the Data
-
-The classic Fellegi Sunter represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$.
-To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicators of lengths $L_f$.
-For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework.
-This is a bijective transformive that does not change the meaning of the data, but this representation will ease calculations and posterior updates.
-
-## Hashing
+## Data Representation and Hashing
 
 In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair.
-We however do not store these comparison vectors themselves, but instead only the a hashed value indicating the unique agreement pattern.
-We store this information in a nested list.
-For each $p$, we also calculate $|h_p|$ total instances of agreement pattern $h_p$ in the entire $\Gamma$ matrix, and then for each $j$, we also calculate $|h_p|_j$ the total instances agreement pattern for each $\Gamma_{.j}$
+We however do not store these comparison vectors themselves, but instead only the a hashed value corresponding to the agreement pattern of the $(i, j)$ pair.
+Enamorado et al (2019) provided the hashing function
+$$\sum_{f=1}^F \mathbf{1}_{\gamma_{(i,j)}^f >0}2^{\gamma_{(i,j)}^f + \mathbf{1}_{f>1} \times \sum_{e=1}^{f-1}(L_e-1)}$$
+to map each agreement pattern to a unique value, but packages like `dplyr` in `R` are capable of this as well.
 
-## Updating m and u
+We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $h_p$ with record $j \in B$.
+For each $p$, we also calculate $|h_p| = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$ total instances of agreement pattern $h_p$ throughout the data, and also for each $j$, we  calculate $|h_p|_j = \sum_{i=1}^{n_A} \mathbf{1}_{(i,j) \in h_p}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.
+
+For large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\} for more manageable computations.
+On a single machine, we can read-in data sequentially, conduct hashing, collect results, and delete the original data from memory before continuing with the next chunk of data.
+With multiple cores or multiple machines, this can be done in parallel.
+Storing this hashed information still becomes burdensome for large data, but this hashing method greatly expands the capabilities of the Fellegi-Sunter framework.
+
+Lastly, the classic Fellegi Sunter method represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$.
+To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicator vectors of lengths $L_f$.
+For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework.
+This is a bijective transformation that does not change the meaning of the data, but this representation will ease calculations and posterior updates.
+
+## Gibbs Sampling
 
 After receiving matching statuses from $\mathbf{Z}$, the Sadinle method calculates $\alpha_{fl}(\mathbf{Z})$ and $\beta_{fl}(\mathbf{Z})$ for each field and level.
-This constitutes $2 \times \sum L_f$ many summations over $n_A \times n_B$ quantities, and becomes computionally burdensome with large data.
+This constitutes $2 \times \sum L_f$ many summations over $n_A \times n_B$ quantities, and becomes computationally burdensome with large data.
 In contrast, we recognize that each unique agreement pattern contributes to the posterior $\alpha(\mathbf{Z})$ and $\beta(\mathbf{Z})$ vectors in the same way.
-In fact, if we denote $\mathbf{Z}(h_p)$ to be the number of record pairs with agreement pattern $h_p$, then the contribution of pairs of pattern $h_p$ to $\alpha(\mathbf{Z})$ is simply $\mathbf{Z}(h_p) \times h_p$.
-Thus our posterior update is simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p$.
-Then $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)$
+In fact, if we denote $\mathbf{Z}(h_p) = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $h_p$, then the contribution of pairs of pattern $h_p$ to the $\alpha(\mathbf{Z})$ vector is simply $\mathbf{Z}(h_p) \times h_p$.
+Thus our posterior updates are simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p$ and  $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)$.
+Note that this constitutes $P$ many summations over $n_B$ quantities, and thus avoids the $n_A \times n_B$ summation from the original method.
 
-## Updating Z
+<!-- Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching, given by  -->
 
-Sadinle uses the following full conditional for $Z_j$ in order to strictly enforce one-to-one matching.
+<!-- $$p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto -->
+<!-- \begin{cases}  -->
+<!--     w_{ij} \mathbf{1}_{i \neq Z_{j'}; \; j' \neq j} & i \leq n_A; \\ -->
+<!--      D_{-j} \frac{n_B - D_{-j} + \beta_{\lambda}}{D_{-j} + \alpha_{\lambda}} & i  = n_A + 1 \\ -->
+<!-- \end{cases}$$ -->
 
-The indicator removes previously matched records from the set of candidate records, and $L_{-j}$ updates iteratively after each matching decision, $n_B$ many times per Gibbs iteration.
+<!-- The indicator removes previously matched records from the set of candidate records, and $D_{-j}:=\sum_{B \setminus j} $ updates iteratively after each matching decision, $n_B$ many times per Gibbs iteration.
 This process is *inherently serial* and becomes computationally burdensome for large data.
-By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel.
-More importantly, since only the aggrement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns.
+-->
+Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching.
+Particular, this sampler removed previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*.By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel.
+More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns.
 This boosts computation time far greater than parallelization.
 
-To do this
+To do this we calculate the Fellegi Sunter weight $w_{h_p}$ for each unique pattern and sample the agreement pattern  $h(Z_j)$ following the modified full conditional:
 
-##
+$$h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\begin{cases} 
+    w_{h_p}\times |h_p|_j  & h(z_j) \in H; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
+\end{cases}$$
+
+We complete the entire Gibbs procedure at the level of the $P$ agreement patterns.
+After, we can back-fill the records corresponding to the agreement patterns by sampling uniformly at random among candidate records stored in $\tilde{\Gamma}$.
+
+## Enforcing Bipartite Matching
+
+Add this
+
+# Simulation Studies
+
+- Replicating the Sadinle Study
+
+- `parlr` and `BRL` speed comparison
+
+# Data Analysis
+
+- Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?

### 2021-04-12 paper/parlr_paper_1.tex
First draft of Efficient Computation

--- 2021-02-27 paper/base_model_specification.tex
@@
     \end{flushright}
-    \begin{center} \textbf{Base Model Specification}
+    \begin{center} \textbf{Parlr: Parallelized Record Linkage in R}
 
@@
     \end{center}
+
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
+
+Most methods for conducting record linkage, the task of identifying
+duplicate records across datasets, are derived from the foundational
+article by Fellegi and Sunter in 1969. Within this family of methods,
+two methods have notable strengths: \texttt{fastlink}, proposed by
+Enamorado et al (2019), continues a is able to conduct linkage on
+incredibly large datasets
 
@@
 \section{Model Specification}\label{model-specification}}
+
+(This will be revised to talk more explicitly about sampling \(Z_j\)
+independently)
 
@@
 \(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\) prior corresponds to a
-uniform prior on the labelling of \(\mathbf{Z}\).
+uniform prior on the labeling of \(\mathbf{Z}\).
 
@@
 \(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
-determined only by the number of dupliates
+determined only by the number of duplicates
 \(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
@@
 \(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
-is a discrete distribution with probalities proportional to
-\[p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
-\[\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\]
-\[=
+is a discrete distribution with probabilities proportional to
+
+\begin{align*}
+p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\\
+&\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} && \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\\
+&=
 \begin{cases} 
@@
 
-Here, we integrate over the posterior of \(\lambda\) and rearrange terms
-to produce the final full conditional:
+In order to make fair comparisons against the Sadinle 2017 model, we
+integrate over the posterior of \(\lambda\) and rearrange terms to
+produce the final full conditional:
 
-\[Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
 \begin{cases} 
-    w_{ij}  & z_j \leq n_A; \\
+    w_{ij}  & i \leq n_A; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
+\end{cases}\]
+
+\hypertarget{bayes-estimate}{%
+\subsection{Bayes Estimate}\label{bayes-estimate}}
+
+Our Gibbs sampler provides posterior samples of \(\mathbf{Z}\) which we
+use to make our final decisions about the linkage structure. In the case
+where we false matches and missed matches contribute the same loss, we
+declare \((i,j)\) to be a match whenever \(P(Z_j = i) > \frac{1}{2}\)
+according to these posterior samples. We can create more elaborate
+decisions by attributing different loss values to different kinds of
+error, and allowing pairings with middling posterior probabilities to be
+left unlabeled by the algorithm so that the modeler can address those
+manually. For further discussion, see Sadinle 2017.
+
+\hypertarget{efficient-computation}{%
+\section{Efficient Computation}\label{efficient-computation}}
+
+Broadly speaking, we increase our computational efficiency by
+recognizing that record pairs contribute to posterior calculations only
+through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(H\)
+be the set of unique agreement patterns in the data, let \(P\) denote
+the total number of unique agreement patterns. Note that \(P\) is
+bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not
+scale with \(n_A\) or \(n_B\). We index these agreement patterns by
+\(p \in \{1, \ldots, P\}\), and say \((i,j) \in h_p\) when the \((i,j)\)
+pair exhibits the \(p^{th}\) agreement pattern. Wherever possible, we
+conduct calculations over these \(P\) agreement patterns rather than the
+\(n_A \times n_B\) record pairs.
+
+\hypertarget{data-representation-and-hashing}{%
+\subsection{Data Representation and
+Hashing}\label{data-representation-and-hashing}}
+
+In the classic Fellegi Sunter framework, \(\Gamma\) is a
+\(n_A n_B \times F\) matrix, with each row providing the comparison
+vector for a different \((i,j)\) pair. We however do not store these
+comparison vectors themselves, but instead only the a hashed value
+corresponding to the agreement pattern of the \((i, j)\) pair. Enamorado
+et al (2019) provided the hashing function
+\[\sum_{f=1}^F \mathbf{1}_{\gamma_{(i,j)}^f >0}2^{\gamma_{(i,j)}^f + \mathbf{1}_{f>1} \times \sum_{e=1}^{f-1}(L_e-1)}\]
+to map each agreement pattern to a unique value, but packages like
+\texttt{dplyr} in \texttt{R} are capable of this as well.
+
+We store this information in a nested list \(\tilde{\Gamma}\) where the
+\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
+in \(A\) that share agreement pattern \(h_p\) with record \(j \in B\).
+For each \(p\), we also calculate
+\(|h_p| = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\)
+total instances of agreement pattern \(h_p\) throughout the data, and
+also for each \(j\), we calculate
+\(|h_p|_j = \sum_{i=1}^{n_A} \mathbf{1}_{(i,j) \in h_p}\) the instances
+of agreement pattern \(p\) among the comparison vectors between record
+\(j \in B\) and each of the \(n_A\) records in \(A\).
+
+For large data, we can partition the two datasets \(A\) and \(B\) into
+smaller blocks \(\{A_m\}\) and \$\{B\_m\} for more manageable
+computations. On a single machine, we can read-in data sequentially,
+conduct hashing, collect results, and delete the original data from
+memory before continuing with the next chunk of data. With multiple
+cores or multiple machines, this can be done in parallel. Storing this
+hashed information still becomes burdensome for large data, but this
+hashing method greatly expands the capabilities of the Fellegi-Sunter
+framework.
+
+Lastly, the classic Fellegi Sunter method represents the \(\gamma_{ij}\)
+comparison vector as vector of length \(F\), with each component
+\(\gamma_{ij}^f\) taking on values in \(\{0, \ldots, L_f - 1\}\). To
+ease computations, we instead represent the comparison as a
+concatenation of \(F\) many binary indicator vectors of lengths \(L_f\).
+For example, if \(L_1 = L_2 = 2\) and \(L_3 = 3\), then
+\(\gamma_{ij} = (1, 0, 2)\) under the classical framework becomes
+\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
+bijective transformation that does not change the meaning of the data,
+but this representation will ease calculations and posterior updates.
+
+\hypertarget{gibbs-sampling}{%
+\subsection{Gibbs Sampling}\label{gibbs-sampling}}
+
+After receiving matching statuses from \(\mathbf{Z}\), the Sadinle
+method calculates \(\alpha_{fl}(\mathbf{Z})\) and
+\(\beta_{fl}(\mathbf{Z})\) for each field and level. This constitutes
+\(2 \times \sum L_f\) many summations over \(n_A \times n_B\)
+quantities, and becomes computationally burdensome with large data. In
+contrast, we recognize that each unique agreement pattern contributes to
+the posterior \(\alpha(\mathbf{Z})\) and \(\beta(\mathbf{Z})\) vectors
+in the same way. In fact, if we denote
+\(\mathbf{Z}(h_p) = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to
+be the number of matching record pairs with agreement pattern \(h_p\),
+then the contribution of pairs of pattern \(h_p\) to the
+\(\alpha(\mathbf{Z})\) vector is simply \(\mathbf{Z}(h_p) \times h_p\).
+Thus our posterior updates are simply
+\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p\)
+and
+\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)\).
+Note that this constitutes \(P\) many summations over \(n_B\)
+quantities, and thus avoids the \(n_A \times n_B\) summation from the
+original method.
+
+Sadinle uses a prior for \(\mathbf{Z}\) that induces the a full
+conditional for \(Z_j\) that strictly enforces one-to-one matching.
+Particular, this sampler removed previously matches records from the set
+of candidate records when sampling \(Z_j\), creating a dependency that
+makes the sampler \emph{inherently serial}.By weakening the one-to-one
+requirement, our full conditional for \(Z\) does not depend on the rest
+of the \(\mathbf{Z}_{-j}\) vector, and thus can be computed in parallel.
+More importantly, since only the agreement pattern of \(Z_j\) is used
+for calculations within the Gibbs sampler, and not the particular record
+label, we can conduct this sampling only at the level of the unique
+agreement patterns. This boosts computation time far greater than
+parallelization.
+
+To do this we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
+unique pattern and sample the agreement pattern \(h(Z_j)\) following the
+modified full conditional:
+
+\[h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\begin{cases} 
+    w_{h_p}\times |h_p|_j  & h(z_j) \in H; \\
      n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
@@
 
+We complete the entire Gibbs procedure at the level of the \(P\)
+agreement patterns. After, we can back-fill the records corresponding to
+the agreement patterns by sampling uniformly at random among candidate
+records stored in \(\tilde{\Gamma}\).
+
+\hypertarget{enforcing-bipartite-matching}{%
+\subsection{Enforcing Bipartite
+Matching}\label{enforcing-bipartite-matching}}
+
+Add this
+
+\hypertarget{simulation-studies}{%
+\section{Simulation Studies}\label{simulation-studies}}
+
+\begin{itemize}
+\item
+  Replicating the Sadinle Study
+\item
+  \texttt{parlr} and \texttt{BRL} speed comparison
+\end{itemize}
+
+\hypertarget{data-analysis}{%
+\section{Data Analysis}\label{data-analysis}}
+
+\begin{itemize}
+\tightlist
+\item
+  Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?
+\end{itemize}
+
 \end{document}

### 2021-04-17 notes/Notes_4_23_21.Rmd
random indexing

---
title: "Random Indexing"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
library(RecordLinkage)
library(tidyverse)
```

## Overview

- Final piece: Random Indexing

## Review of Indexing

- In the literature, *indexing* is when we remove certain agreement patterns from consideration for matching after making the comparison vectors

- Jody indexed on first and last name, meaning that she chose a threshold for name similarity, and then set all record pairs with name agreement below that threshold to be nonmatching

- Used to reduce comparison space and avoid false matches

## BK's Thoughts on Indexing

- If your model isn't able to avoid these types of matchings, you probably need a different model

- In Jody's code, indexing gave modest computational savings. Instead of summing over $n_A \times n_B$ record pairs to update $m$ and $u$, she only had to sum over record pairs that met the indexing threshold

- Uses human judgement

## Random Indexing

- Let $P$ be the number of unique agreement patterns, and denote a particular agreement pattern by $h_p$ for $p \in \{1, \ldots, P\}$. Let $|h_p|$ be the total number of pairs of $p^{th}$ agreement pattern in the data, and let $|h_p|_j$ be the number of record pairs of the $p^{th}$ agreement pattern between $j\in B$ and the records in $A$.

- Remember that when sampling $Z_j$, all record pairs of agreement pattern $h_p$ are equally likely. Without knowing anything about the agreement pattern itself, this tells us that any agreement pattern for which $|h_p|_j$ is large is unlikely to be a match.

## Data Storage

- Since the overwhelming majority of record pairs are of agreement patterns with mostly $0's$, these pairs contribute the most the the data storage burden.

- All contributions to posterior udpates themselves are governed by $|h_p|_j$, not the record labels themselves. So instead of storing arbitrarily large number of record labels, lets store only small number $R$ of them.

- Even if I remove certain records from consideration, I maintain their probabilities and weights through $|h_p|_j$. This is similar to the *u correction* concept from the McVeigh paper (I think!)

- Choosing $R$ is up to the modeller, but much less room for error than choosing fields and thresholds to index on. Also, the goal is not avoid false matches (the model takes care of that), but instead data storage.

## Example

- I simulated comparison vectors for a $4000 \times 4000$ linkage problem

- The comparison matrix $\Gamma$ produced by `BRL` is 610 MB.

- After hashing, the object is only 68 MB. Pretty good.

- After running random indexing, the object was only **9 MB.** I ran `parlr`, and got the exact same results!

## How does this help?

- I still need to create the comparison vectors in the first place. This is the unavoidable downsize of the Fellegi-Sunter methods, and there is no way to avoid this.

- However, I can create comparison vectors for smaller chunks of the data, reduce the peices through hashing and random indexing, combining results as I go along, without ever needing to store the entire $n_A \times n_B$ matrix of comparison vectors. In fact, no matter how large $n_A$ is, we store at most $n_B \times P \times R$ record pairs.

- Processing these smaller chunks can be done in parallel.

- Should allow us to conduct *significantly* larger record linkage tasks.

## Naming

- Since I'm removing records from consideration, I thought it might be nice to invoke the idea of indexing. Since the records I remove are chosen at random (within each pattern/equivalence class), I thought *random indexing* might be a nice word.

- Might be clearer to just call this *thinning* or something else though.

## Next Steps

- This whole concept is ready to go! I will have a (very rough) first draft ready for you by the next meeting.

- These mechanics will lay the foundation (with slight modifications) for the linkage cluster model I've been developing. This is on hold until the first `parlr` paper is done.

### 2021-04-17 paper/FAQ_STA561_Project.Rmd
random indexing

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
header-includes: \usepackage{hyperref}
linestretch: 1
link-citations: yes
linkcolor: blue
fontfamily: mathpazo
fontsize: 12pt
---
\begin{flushright} 
	\end{flushright}
	\begin{center} \textbf{Parlr: Frequently Asked Questions}

	Brian Kundinger

	\end{center}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### What are parlr's competitors?

Most record linkage techniques are derived from the seminal 1969 paper by Fellegi and Sunter: "A Theory for Record Linkage". The defining characteristic of their model was to transform the sets of records, which constitute text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated. Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates a $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets. We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f$ providing their similarity score on the $f^{th}$ field. For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeller. With notation $\gamma_{ij}^f \in \{0, \ldots, L_f - 1\}$ for the number of possible agreement levels for field $f$, we use binary 0-1 variables to indicate exact matching, and 0-1-2 variables to provide an option for partial matching. For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model.

Two modern adaptations of the Fellegi Sunter model are important for understanding `parlr` and its contribution. In 2019, Enamorado et al proposed `fastlink`, a method that closely followed the modelling assumptions of Fellegi and Sunter, but used innovating hashing techniques to greatly enhance its speed and scale. Importantly, they maintained Fellegi and Sunter's independent matching assumption, in which the matching status of any $(i,j)$ pair is made independently of the matching of any other $(i', j')$. This leads to many matchings that violate "one-to-one" considerations that are often implicit in the data, which need to be resolved in a post-processing step. In contrast, Sadinle (2017) proposed "Beta Record Linkage for Bipartite Matching" and the accompanying `BRL` package which strictly enforces one-to-one matching. Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, then samples a potential link. However, accounting for these dependencies throughout the linkage process is computionally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.

### How is parlr different?

While `fastlink` makes $n_A \times n_B$ many indepedent binary linkage decisions and Sadinle makes $n_B$ many *dependent* linkage decisions among $n_A$ many options, `parlr` strikes middle ground by making $n_B$ many *independent* linkage decisions among $n_A$ many options. That is, in each iteration of the Gibbs sampler, we make one linkage decision per $j \in B$, but do not remove previously matched records from consideration. While this does mean that a particular iteration of the Gibbs sampler may contain some matchings that violate one-to-one requirements, these violations happen infreqently and randomly across all record pairing, so that they do not seriously harm the final Bayes estimate of the linkage structure we make from all the Gibbs samples. In cases where our Bayes estimate does contain matches that violate one-to-one, we note that such cases far less frequently than under `fastlink`, and that these cases are easily resolved by simply taking the match pair with the highest posterior match probability.

Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns. Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$. We index these agreement patterns by $p \in \{1, \ldots, P\}$, and say $(i,j) \in h_p$ when the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern. Wherever possible, we conduct calculations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.

By making linkage decisions for each $j\in B$ independently, we do not need to concern ourselves with the precise record label for $j's$ potential pairing within the Gibbs sampler, only the agreement pattern of the $\gamma$ vector. While the original Fellegi Sunter framework has computational complexity $O(n_A \times n_B)$, these modifications completely remove dependence on $n_A$ throughout the Gibbs sampler, reducing the complexity to $O(n_B)$

### How accurate is it?

In his 2017 paper, Sadinle demonstrated the strength of his method by running `BRL` on simulated pairs of files with differing levels of errors and overlap. Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakend performance in the extreme scenario of very high errors and very high overlap across files.

$$\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}$$
$$\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}$$
$$\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)$$

```{r fig.align='center'}
knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
```

### How fast is it?

We simulate $\Gamma$ matrices of comparison vectors from multinomial distributions meant to emulate the behavior of similarity scores across first name, last name, and day, month, and and year of birth. Distributions provided below. We simulate these data for different values of $n_A and n_B$, and compare the runtime of `parlr` against `BRL`. We note here that `BRL` is coded in C, which makes for unfair comparison against `parlr`, currently only built in R. We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data. In particular, runtime for `BRL` seems to grow quadratically while runtime for `parlr` seems to grow linearly. Additionally, although `parlr` is amenable to parallezation, this simulation was run on a single core. Running `parlr` in C or C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more drastic results.

```{r fig.align='center'}
knitr::include_graphics("../notes/figures/sadinle_speed_plot.png")
```

The above discussion suggests that for fixed $n_B$, computation time should remain mostly constant with growing $n_A$. Shockingly, this seems to be true. In the plot below, fixing $n_B = 500$, we see linear growth for the runtime under `BRL` as $n_A$ increases, with much more static runtime under `parlr`. The slight increases in runtime that we do see are due primarily to the hashing step, which again can be run in parallel for large data.

```{r fig.align='center'}
knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
```

### What about deduplication of a single file

At the moment, `parlr` (and its `BRL` competitor) are only able to handle bipartite matching across two files, meaning there is duplication across, but not within, files. Though it seems similar to linkage across files, deduplication within one file is suprisingly much more difficult to model probabalistically. In particular, `parlr` assumes  that the linkage decision for record $j \in B$ is independent from all other $j' \in B$, allowing us to propose a prior $\lambda \sim \text{Beta}(\alpha, \beta)$ for the probability of $j$ having some match in $A$, and update that prior through standard beta-binomial posterior updates. With only one file however, if the sampler decides that record $1$ matches with record $5$, there is no way to determine the linkage of record $5$ independently; we do not even have $n_B$ independent linkage decisions to work with! Constructing fast, scalable methods for deduplication is one avenue of future work.

### Anything else?

Kundinger is particularly interested in linkage scenarios when the reliability of the data (the level of error in the fields), and the rates of matching, differ by subgroups in the data. The simplest case of this occurs when matching historical records for women; due to marriage conventions, the last name is a much less reliable identifier for women than for men. In an upcoming extention, we propose *linkage clusters* that account these difficulties, while still maintaining the computational advances from `parlr`.

### 2021-04-17 paper/FAQ_STA561_Project.tex
random indexing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{hyperref}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Parlr: Frequently Asked Questions}

    Brian Kundinger

    \end{center}

\hypertarget{what-are-parlrs-competitors}{%
\subsubsection{What are parlr's
competitors?}\label{what-are-parlrs-competitors}}

Most record linkage techniques are derived from the seminal 1969 paper
by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
characteristic of their model was to transform the sets of records,
which constitute text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates a
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f\) providing their similarity score on the
\(f^{th}\) field. For ease of modeling and computation, we restrict
these similarity scores to be discrete, ordinal variables, and the
construction of these is left to the modeller. With notation
\(\gamma_{ij}^f \in \{0, \ldots, L_f - 1\}\) for the number of possible
agreement levels for field \(f\), we use binary 0-1 variables to
indicate exact matching, and 0-1-2 variables to provide an option for
partial matching. For text data, we calculate similarity based on
Levenstein distance or some other text similarity score, and bin these
scores to integers for use in the model.

Two modern adaptations of the Fellegi Sunter model are important for
understanding \texttt{parlr} and its contribution. In 2019, Enamorado et
al proposed \texttt{fastlink}, a method that closely followed the
modelling assumptions of Fellegi and Sunter, but used innovating hashing
techniques to greatly enhance its speed and scale. Importantly, they
maintained Fellegi and Sunter's independent matching assumption, in
which the matching status of any \((i,j)\) pair is made independently of
the matching of any other \((i', j')\). This leads to many matchings
that violate ``one-to-one'' considerations that are often implicit in
the data, which need to be resolved in a post-processing step. In
contrast, Sadinle (2017) proposed ``Beta Record Linkage for Bipartite
Matching'' and the accompanying \texttt{BRL} package which strictly
enforces one-to-one matching. Specifically, in each iteration of his
Gibbs sampler, he considers each record \(j\in B\), removes from
consideration the records \(i\in A\) that have already been matched,
then samples a potential link. However, accounting for these
dependencies throughout the linkage process is computionally burdensome,
leaving \texttt{BRL} only suitable for small to moderate linkage
problems.

\hypertarget{how-is-parlr-different}{%
\subsubsection{How is parlr different?}\label{how-is-parlr-different}}

While \texttt{fastlink} makes \(n_A \times n_B\) many indepedent binary
linkage decisions and Sadinle makes \(n_B\) many \emph{dependent}
linkage decisions among \(n_A\) many options, \texttt{parlr} strikes
middle ground by making \(n_B\) many \emph{independent} linkage
decisions among \(n_A\) many options. That is, in each iteration of the
Gibbs sampler, we make one linkage decision per \(j \in B\), but do not
remove previously matched records from consideration. While this does
mean that a particular iteration of the Gibbs sampler may contain some
matchings that violate one-to-one requirements, these violations happen
infreqently and randomly across all record pairing, so that they do not
seriously harm the final Bayes estimate of the linkage structure we make
from all the Gibbs samples. In cases where our Bayes estimate does
contain matches that violate one-to-one, we note that such cases far
less frequently than under \texttt{fastlink}, and that these cases are
easily resolved by simply taking the match pair with the highest
posterior match probability.

Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(H\)
be the set of unique agreement patterns in the data, let \(P\) denote
the total number of unique agreement patterns. Note that \(P\) is
bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not
scale with \(n_A\) or \(n_B\). We index these agreement patterns by
\(p \in \{1, \ldots, P\}\), and say \((i,j) \in h_p\) when the \((i,j)\)
pair exhibits the \(p^{th}\) agreement pattern. Wherever possible, we
conduct calculations over these \(P\) agreement patterns rather than the
\(n_A \times n_B\) record pairs.

By making linkage decisions for each \(j\in B\) independently, we do not
need to concern ourselves with the precise record label for \(j's\)
potential pairing within the Gibbs sampler, only the agreement pattern
of the \(\gamma\) vector. While the original Fellegi Sunter framework
has computational complexity \(O(n_A \times n_B)\), these modifications
completely remove dependence on \(n_A\) throughout the Gibbs sampler,
reducing the complexity to \(O(n_B)\)

\hypertarget{how-accurate-is-it}{%
\subsubsection{How accurate is it?}\label{how-accurate-is-it}}

In his 2017 paper, Sadinle demonstrated the strength of his method by
running \texttt{BRL} on simulated pairs of files with differing levels
of errors and overlap. Comparing our method against \texttt{BRL} on
these same simulated datasets, we find that our method only has weakend
performance in the extreme scenario of very high errors and very high
overlap across files.
\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]

\begin{center}\includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot} \end{center}

\hypertarget{how-fast-is-it}{%
\subsubsection{How fast is it?}\label{how-fast-is-it}}

We simulate \(\Gamma\) matrices of comparison vectors from multinomial
distributions meant to emulate the behavior of similarity scores across
first name, last name, and day, month, and and year of birth.
Distributions provided below. We simulate these data for different
values of \(n_A and n_B\), and compare the runtime of \texttt{parlr}
against \texttt{BRL}. We note here that \texttt{BRL} is coded in C,
which makes for unfair comparison against \texttt{parlr}, currently only
built in R. We see that at low data size, \texttt{BRL} outperforms, but
that \texttt{parlr} is significantly faster at handling larger data. In
particular, runtime for \texttt{BRL} seems to grow quadratically while
runtime for \texttt{parlr} seems to grow linearly. Additionally,
although \texttt{parlr} is amenable to parallezation, this simulation
was run on a single core. Running \texttt{parlr} in C or C++ with
paralellization for the hashing step and sampling the matching status of
the record pairs should lead to even more drastic results.

\begin{center}\includegraphics[width=23.6in]{../notes/figures/sadinle_speed_plot} \end{center}

The above discussion suggests that for fixed \(n_B\), computation time
should remain mostly constant with growing \(n_A\). Shockingly, this
seems to be true. In the plot below, fixing \(n_B = 500\), we see linear
growth for the runtime under \texttt{BRL} as \(n_A\) increases, with
much more static runtime under \texttt{parlr}. The slight increases in
runtime that we do see are due primarily to the hashing step, which
again can be run in parallel for large data.

\begin{center}\includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB} \end{center}

\hypertarget{what-about-deduplication-of-a-single-file}{%
\subsubsection{What about deduplication of a single
file}\label{what-about-deduplication-of-a-single-file}}

At the moment, \texttt{parlr} (and its \texttt{BRL} competitor) are only
able to handle bipartite matching across two files, meaning there is
duplication across, but not within, files. Though it seems similar to
linkage across files, deduplication within one file is suprisingly much
more difficult to model probabalistically. In particular, \texttt{parlr}
assumes that the linkage decision for record \(j \in B\) is independent
from all other \(j' \in B\), allowing us to propose a prior
\(\lambda \sim \text{Beta}(\alpha, \beta)\) for the probability of \(j\)
having some match in \(A\), and update that prior through standard
beta-binomial posterior updates. With only one file however, if the
sampler decides that record \(1\) matches with record \(5\), there is no
way to determine the linkage of record \(5\) independently; we do not
even have \(n_B\) independent linkage decisions to work with!
Constructing fast, scalable methods for deduplication is one avenue of
future work.

\hypertarget{anything-else}{%
\subsubsection{Anything else?}\label{anything-else}}

Kundinger is particularly interested in linkage scenarios when the
reliability of the data (the level of error in the fields), and the
rates of matching, differ by subgroups in the data. The simplest case of
this occurs when matching historical records for women; due to marriage
conventions, the last name is a much less reliable identifier for women
than for men. In an upcoming extention, we propose \emph{linkage
clusters} that account these difficulties, while still maintaining the
computational advances from \texttt{parlr}.

\end{document}

### 2021-04-17 paper/parlr_paper_1.Rmd
random indexing

@@
 Following the Fellegi Sunter framework, we define $m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)$ to be the probability of observing agreement level $l$ in field $f$ for records $i$ and $j$ given that the records are a match, and similarly define $u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)$, for non-matches.
-We denote $\lambda$ to be the (marginal) probability that some record $j \in B$ has a match in $A$.
+We also adopt Fellegi and Sunter's conditionally independent fields assumption that the level of agreement on one field is independent of the level of agreement on another.
+Though this assumption is often not reasonable (for example, first name and gender are two clearly dependent fields), but it is common within the record linkage literature and generally leads to models that perform well in practice; see discussion for further remarks.
+Lastly, we define $\lambda$ to be the (marginal) probability that some record $j \in B$ has a match in $A$.
 
@@
 In contrast, we recognize that each unique agreement pattern contributes to the posterior $\alpha(\mathbf{Z})$ and $\beta(\mathbf{Z})$ vectors in the same way.
-In fact, if we denote $\mathbf{Z}(h_p) = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $h_p$, then the contribution of pairs of pattern $h_p$ to the $\alpha(\mathbf{Z})$ vector is simply $\mathbf{Z}(h_p) \times h_p$.
-Thus our posterior updates are simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p$ and  $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)$.
-Note that this constitutes $P$ many summations over $n_B$ quantities, and thus avoids the $n_A \times n_B$ summation from the original method.
+In fact, if we denote $|h_p^m| = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $h_p$, then the contribution of pairs of pattern $h_p$ to the $\alpha(\mathbf{Z})$ vector is simply $\mathbf{Z}(h_p) \times h_p$.
+Thus our posterior update for $\alpha$ is simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P |h_p^m| \times h_p$.
+Then, we can easily calculate $|h_p^u|$, the number of nonmatching record pairs of agreement pattern $p$, by subtracting the number of matching pairs from the total present in the data; that is $|h_p^u| = |h_p| - |h_p^m|$.
+From this, we can update our $\beta$ parameter through  $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p^u| \times h_p$.
+Note that these constitute $P$ many summations over $n_B$ quantities, and thus avoid the $n_A \times n_B$ summation from the original method.
 
@@
 -->
+
 Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching.
@@
 
-To do this we calculate the Fellegi Sunter weight $w_{h_p}$ for each unique pattern and sample the agreement pattern  $h(Z_j)$ following the modified full conditional:
+To do this we calculate the Fellegi Sunter weight $w_{h_p}$ for each unique pattern,  sample the agreement pattern between $j$ and its potential match, and then sample the record label uniformly among viable records.
+More concretely, define $H(Z_j)$ to be the agreement pattern between $j$ and its potential match, and say $H(Z_j) = h_{P+1}$ when $Z_j = n_A + 1$.
+Then,
 
-$$h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+$$H\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
 \begin{cases} 
@@
 \end{cases}$$
+<!-- $$p\left(Z_j^{(s+1)} = i\right) = \frac{1}{\left|H\left(Z_j^{(s+1)}\right)\right|}; \;\;i \text{ such that } (i, j) \in H\left(Z_j^{(s+1)}\right)$$ -->
 
@@
 - Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?
+
+# Discussion
+
+We note that this often not a reasonable assumption; for example if we know that two records differ on gender, this does not give any information on whether their year of births are the same, but it does suggest that its more likely that their first names are different.

### 2021-04-17 paper/parlr_paper_1.tex
random indexing

@@
 \(j\) given that the records are a match, and similarly define
-\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We
-denote \(\lambda\) to be the (marginal) probability that some record
-\(j \in B\) has a match in \(A\).
+\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We also
+adopt Fellegi and Sunter's conditionally independent fields assumption
+that the level of agreement on one field is independent of the level of
+agreement on another. Though this assumption is often not reasonable
+(for example, first name and gender are two clearly dependent fields),
+but it is common within the record linkage literature and generally
+leads to models that perform well in practice; see discussion for
+further remarks. Lastly, we define \(\lambda\) to be the (marginal)
+probability that some record \(j \in B\) has a match in \(A\).
 
@@
 in the same way. In fact, if we denote
-\(\mathbf{Z}(h_p) = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to
-be the number of matching record pairs with agreement pattern \(h_p\),
-then the contribution of pairs of pattern \(h_p\) to the
-\(\alpha(\mathbf{Z})\) vector is simply \(\mathbf{Z}(h_p) \times h_p\).
-Thus our posterior updates are simply
-\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p\)
-and
-\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)\).
-Note that this constitutes \(P\) many summations over \(n_B\)
-quantities, and thus avoids the \(n_A \times n_B\) summation from the
-original method.
+\(|h_p^m| = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
+number of matching record pairs with agreement pattern \(h_p\), then the
+contribution of pairs of pattern \(h_p\) to the \(\alpha(\mathbf{Z})\)
+vector is simply \(\mathbf{Z}(h_p) \times h_p\). Thus our posterior
+update for \(\alpha\) is simply
+\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P |h_p^m| \times h_p\).
+Then, we can easily calculate \(|h_p^u|\), the number of nonmatching
+record pairs of agreement pattern \(p\), by subtracting the number of
+matching pairs from the total present in the data; that is
+\(|h_p^u| = |h_p| - |h_p^m|\). From this, we can update our \(\beta\)
+parameter through
+\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p^u| \times h_p\). Note
+that these constitute \(P\) many summations over \(n_B\) quantities, and
+thus avoids the \(n_A \times n_B\) summation from the original method.
 
@@
 To do this we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
-unique pattern and sample the agreement pattern \(h(Z_j)\) following the
-modified full conditional:
+unique pattern, sample the agreement pattern between \(j\) and its
+potential match, and then sample the record label uniformly among viable
+records. More concretely, define \(H(Z_j)\) to be the agreement pattern
+between \(j\) and its potential match, and say \(H(Z_j) = h_{P+1}\) when
+\(Z_j = n_A + 1\). Then,
 
-\[h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\[H\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
 \begin{cases} 
@@
 
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+We note that this often not a reasonable assumption; for example if we
+know that two records differ on gender, this does not give any
+information on whether their year of births are the same, but it does
+suggest that its more likely that their first names are different.
+
 \end{document}

### 2021-04-17 R/GetUniquePatterns2.R

@@ -1,4 +1,4 @@
-GetUniquePatterns2 <- function(cd, fast = F){
+GetUniquePatterns2 <- function(cd, fast = F, R = NULL){
 
   indicators <- cd[[1]]
   N <- dim(indicators)[1]
@@ -111,8 +111,14 @@ pattern_df <- df %>%
     append(x, list(n1 +1))
   })
 
+  if(!is.null(R)){
+    hash_to_rec1 <- lapply(hash_to_rec1, function(z){
+    map(z, ~RandomIndexing(.x, R))
+  })
+  }
 
-  patterns <- list(hash_id,
+
+  patterns <- list(#hash_id,
                    unique_patterns,
                    pattern_counts,
                    counts_by_rec,

### 2021-04-17 R/NLC_hash2.R

@@ -1,5 +1,6 @@
 BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
-                          alpha = 1, beta = 1, S = 1000, burn = 100, show_progress = T, fast =F){
+                          alpha = 1, beta = 1, S = 1000, burn = 100,
+                          show_progress = T, fast = F, R = NULL){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -21,7 +22,7 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
     cl <- parallel::makeCluster(numCores)
   }
 
-  patterns <- GetUniquePatterns2(comparisons, fast)
+  patterns <- GetUniquePatterns2(comparisons, fast, R)
 
   parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
     rep(x, comparisons[[4]][x])
@@ -30,14 +31,14 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   ids <- expand.grid(1:n1, 1:n2)
   #rec2 <- ids[,2]
 
-  hash_id <- patterns[[1]]
-  hash <- data.frame(hash_id)
+  #hash_id <- patterns[[1]]
+  #hash <- data.frame(hash_id)
   #hash_split <- split(hash_id, ids[,2])
-  unique_patterns <- patterns[[2]]
-  pattern_counts <- patterns[[3]]
+  unique_patterns <- patterns[[1]]
+  pattern_counts <- patterns[[2]]
   P <- dim(unique_patterns)[1]
-  counts_by_rec <- patterns[[4]]
-  hash_to_rec1 <- patterns[[5]]
+  counts_by_rec <- patterns[[3]]
+  hash_to_rec1 <- patterns[[4]]
   candidates <- 1:n1
   candidates_P <- 1:(P+1)
   Z.SAMPS <- matrix(NA, nrow = n2, ncol = S)

### 2021-04-19 paper/FAQ_STA561_Project.Rmd
GetPossiblePatterns function

@@
 For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeller.
-With notation $\gamma_{ij}^f \in \{0, \ldots, L_f - 1\}$ for the number of possible agreement levels for field $f$, we use binary 0-1 variables to indicate exact matching, and 0-1-2 variables to provide an option for partial matching.
+It is common to use use binary 0-1 variables to indicate exact matching, and 0-1-2 variables to provide an option for partial matching.
 For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model.
@@
 Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns.
-Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$.
+Let $P$ denote the total number of unique agreement patterns; note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$.
 We index these agreement patterns by $p \in \{1, \ldots, P\}$, and say $(i,j) \in h_p$ when the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern.
 Wherever possible, we conduct calculations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.
-
-By making linkage decisions for each $j\in B$ independently, we do not need to concern ourselves with the precise record label for $j's$ potential pairing within the Gibbs sampler, only the agreement pattern of the $\gamma$ vector.
 While the original Fellegi Sunter framework has computational complexity $O(n_A \times n_B)$, these modifications completely remove dependence on $n_A$ throughout the Gibbs sampler, reducing the complexity to $O(n_B)$
@@
 Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakend performance in the extreme scenario of very high errors and very high overlap across files.
-
-$$\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}$$
-$$\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}$$
-$$\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)$$
-
-```{r fig.align='center'}
-knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
-```
+See plots in technical appendix.
 
@@
 We simulate $\Gamma$ matrices of comparison vectors from multinomial distributions meant to emulate the behavior of similarity scores across first name, last name, and day, month, and and year of birth.
-Distributions provided below.
-We simulate these data for different values of $n_A and n_B$, and compare the runtime of `parlr` against `BRL`.
+We simulate these data for different values of $n_A$ and $n_B$, and compare the runtime of `parlr` against `BRL`.
+We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data.
+See plots in technical appendix.
+
+### How fast can it be?
+
 We note here that `BRL` is coded in C, which makes for unfair comparison against `parlr`, currently only built in R.
-We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data.
-In particular, runtime for `BRL` seems to grow quadratically while runtime for `parlr` seems to grow linearly.
-Additionally, although `parlr` is amenable to parallezation, this simulation was run on a single core.
-Running `parlr` in C or C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more drastic results.
-
-```{r fig.align='center'}
-knitr::include_graphics("../notes/figures/sadinle_speed_plot.png")
-```
-
-The above discussion suggests that for fixed $n_B$, computation time should remain mostly constant with growing $n_A$.
-Shockingly, this seems to be true.
-In the plot below, fixing $n_B = 500$, we see linear growth for the runtime under `BRL` as $n_A$ increases, with much more static runtime under `parlr`.
-The slight increases in runtime that we do see are due primarily to the hashing step, which again can be run in parallel for large data.
-
-```{r fig.align='center'}
-knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
-```
+Additionally, although `parlr` is amenable to parallezation, our simulations thus far have only been run on a single core.
+Running `parlr` in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more drastic results.
 
@@
 
-At the moment, `parlr` (and its `BRL` competitor) are only able to handle bipartite matching across two files, meaning there is duplication across, but not within, files.
+At the moment, `parlr` (and its `BRL` competitor) is only able to handle bipartite matching across two files, meaning there is duplication across, but not within, files.
 Though it seems similar to linkage across files, deduplication within one file is suprisingly much more difficult to model probabalistically.
 In particular, `parlr` assumes  that the linkage decision for record $j \in B$ is independent from all other $j' \in B$, allowing us to propose a prior $\lambda \sim \text{Beta}(\alpha, \beta)$ for the probability of $j$ having some match in $A$, and update that prior through standard beta-binomial posterior updates.
-With only one file however, if the sampler decides that record $1$ matches with record $5$, there is no way to determine the linkage of record $5$ independently; we do not even have $n_B$ independent linkage decisions to work with!
+However, with only one file of $n$ records, we do not even have $n$ independent linkage decisions to work with!
+For example, if the sampler decides that record $1$ matches with record $5$, there is no way to determine the linkage of record $5$ independently.
 Constructing fast, scalable methods for deduplication is one avenue of future work.

### 2021-04-19 paper/FAQ_STA561_Project.tex
GetPossiblePatterns function

@@
 these similarity scores to be discrete, ordinal variables, and the
-construction of these is left to the modeller. With notation
-\(\gamma_{ij}^f \in \{0, \ldots, L_f - 1\}\) for the number of possible
-agreement levels for field \(f\), we use binary 0-1 variables to
-indicate exact matching, and 0-1-2 variables to provide an option for
-partial matching. For text data, we calculate similarity based on
-Levenstein distance or some other text similarity score, and bin these
-scores to integers for use in the model.
+construction of these is left to the modeller. It is common to use use
+binary 0-1 variables to indicate exact matching, and 0-1-2 variables to
+provide an option for partial matching. For text data, we calculate
+similarity based on Levenstein distance or some other text similarity
+score, and bin these scores to integers for use in the model.
 
@@
 recognizing that record pairs contribute to posterior calculations only
-through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(H\)
-be the set of unique agreement patterns in the data, let \(P\) denote
-the total number of unique agreement patterns. Note that \(P\) is
+through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(P\)
+denote the total number of unique agreement patterns; note that \(P\) is
 bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not
@@
 conduct calculations over these \(P\) agreement patterns rather than the
-\(n_A \times n_B\) record pairs.
-
-By making linkage decisions for each \(j\in B\) independently, we do not
-need to concern ourselves with the precise record label for \(j's\)
-potential pairing within the Gibbs sampler, only the agreement pattern
-of the \(\gamma\) vector. While the original Fellegi Sunter framework
-has computational complexity \(O(n_A \times n_B)\), these modifications
-completely remove dependence on \(n_A\) throughout the Gibbs sampler,
-reducing the complexity to \(O(n_B)\)
+\(n_A \times n_B\) record pairs. While the original Fellegi Sunter
+framework has computational complexity \(O(n_A \times n_B)\), these
+modifications completely remove dependence on \(n_A\) throughout the
+Gibbs sampler, reducing the complexity to \(O(n_B)\)
 
@@
 performance in the extreme scenario of very high errors and very high
-overlap across files.
-\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
-\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
-\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
-
-\begin{center}\includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot} \end{center}
+overlap across files. See plots in technical appendix.
 
@@
 distributions meant to emulate the behavior of similarity scores across
-first name, last name, and day, month, and and year of birth.
-Distributions provided below. We simulate these data for different
-values of \(n_A and n_B\), and compare the runtime of \texttt{parlr}
-against \texttt{BRL}. We note here that \texttt{BRL} is coded in C,
-which makes for unfair comparison against \texttt{parlr}, currently only
-built in R. We see that at low data size, \texttt{BRL} outperforms, but
-that \texttt{parlr} is significantly faster at handling larger data. In
-particular, runtime for \texttt{BRL} seems to grow quadratically while
-runtime for \texttt{parlr} seems to grow linearly. Additionally,
-although \texttt{parlr} is amenable to parallezation, this simulation
-was run on a single core. Running \texttt{parlr} in C or C++ with
-paralellization for the hashing step and sampling the matching status of
-the record pairs should lead to even more drastic results.
+first name, last name, and day, month, and and year of birth. We
+simulate these data for different values of \(n_A\) and \(n_B\), and
+compare the runtime of \texttt{parlr} against \texttt{BRL}. We see that
+at low data size, \texttt{BRL} outperforms, but that \texttt{parlr} is
+significantly faster at handling larger data. See plots in technical
+appendix.
 
-\begin{center}\includegraphics[width=23.6in]{../notes/figures/sadinle_speed_plot} \end{center}
+\hypertarget{how-fast-can-it-be}{%
+\subsubsection{How fast can it be?}\label{how-fast-can-it-be}}
 
-The above discussion suggests that for fixed \(n_B\), computation time
-should remain mostly constant with growing \(n_A\). Shockingly, this
-seems to be true. In the plot below, fixing \(n_B = 500\), we see linear
-growth for the runtime under \texttt{BRL} as \(n_A\) increases, with
-much more static runtime under \texttt{parlr}. The slight increases in
-runtime that we do see are due primarily to the hashing step, which
-again can be run in parallel for large data.
-
-\begin{center}\includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB} \end{center}
+We note here that \texttt{BRL} is coded in C, which makes for unfair
+comparison against \texttt{parlr}, currently only built in R.
+Additionally, although \texttt{parlr} is amenable to parallezation, our
+simulations thus far have only been run on a single core. Running
+\texttt{parlr} in C++ with paralellization for the hashing step and
+sampling the matching status of the record pairs should lead to even
+more drastic results.
 
@@
 
-At the moment, \texttt{parlr} (and its \texttt{BRL} competitor) are only
+At the moment, \texttt{parlr} (and its \texttt{BRL} competitor) is only
 able to handle bipartite matching across two files, meaning there is
@@
 having some match in \(A\), and update that prior through standard
-beta-binomial posterior updates. With only one file however, if the
-sampler decides that record \(1\) matches with record \(5\), there is no
-way to determine the linkage of record \(5\) independently; we do not
-even have \(n_B\) independent linkage decisions to work with!
-Constructing fast, scalable methods for deduplication is one avenue of
-future work.
+beta-binomial posterior updates. However, with only one file of \(n\)
+records, we do not even have \(n\) independent linkage decisions to work
+with! For example, if the sampler decides that record \(1\) matches with
+record \(5\), there is no way to determine the linkage of record \(5\)
+independently. Constructing fast, scalable methods for deduplication is
+one avenue of future work.
 

### 2021-04-19 paper/appendix_STA561_project.Rmd
GetPossiblePatterns function

--- 2021-04-17 paper/parlr_paper_1.Rmd
@@
 	\end{flushright}
-	\begin{center} \textbf{Parlr: Parallelized Record Linkage in R}
+	\begin{center} \textbf{Parlr: Parallelized Record Linkage in R} \linebreak
+	\textbf{Technical Appendix}
 
 	Brian Kundinger
+
+	April 23, 2021
+
+	STA 561
 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
 
-## Introduction
+<!-- ## Introduction -->
 
-Most methods for conducting record linkage, the task of identifying duplicate records across datasets, are derived from the foundational article by Fellegi and Sunter in 1969.
-Within this family of methods, two methods have notable strengths: `fastlink`, proposed by Enamorado et al (2019), continues a is able to conduct linkage on incredibly large datasets
+<!-- Most methods for conducting record linkage, the task of identifying duplicate records across datasets, are derived from the foundational article by Fellegi and Sunter in 1969.
+Within this family of methods, two methods have notable strengths: `fastlink`, proposed by Enamorado et al (2019), continues a is able to conduct linkage on incredibly large datasets -->
 
@@
 
-Our notation and assumptions closely follow that of Sadinle (2017).
-Denote two files as $A$ and $B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $A$ and $j \in \{1, \ldots, n_B\}$ in $B$.
+Denote two files as $A$ and $B$, with $n_A$ and $n_B$ records respectively, with records indexed as $i \in \{1, \ldots, n_A\}$ in $A$ and $j \in \{1, \ldots, n_B\}$ in $B$.
 Without loss of generality, label the files such that $n_A \geq n_B$.
 We also assume there are no duplicates within files, only across.
-For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating level agreement between the two records on a specified field.
+
+For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating the level agreement between the two records on a specified field.
+For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeler.
+With notation $\gamma_{ij}^f \in \{0, \ldots, L_f - 1\}$ for the number of possible agreement levels for field $f$, we use binary 0-1 variables to indicate exact matching, and 0-1-2 variables to provide an option for partial matching.
+For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model.
 
@@
 Wherever possible, we reserve superscripts for denoting field and level, while reserving subscripts for record indices.
-For example, $\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})$ is the probability distribution governing field $f$ for matching records, and $\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)$ is product of the relevant of the appropriate $\mathbf{m}$ parameters for record pair $(i,j)$.
+For example, $\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})$ is the probability distribution governing field $f$ for matching records, and $\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)$ is the product of the appropriate $\mathbf{m}$ parameters for record pair $(i,j)$.
 We hope that these conventions avoid overloaded notation in the likelihood and subsequent derivations.
@@
 
-(This will be revised to talk more explicitly about sampling $Z_j$ independently)
+<!-- Our model differs from that of Sadinle 2017 through its explicit dependence on a beta random variable $\lambda$ that models the rate of matching across records.
+Sadinle marginalizes over such a random variable in his derivations of  the "beta prior for bipartite matching," but here we provide derivations without marginalizing in order to specify differing rates of matching for different linkage clusters.
+-->
 
-Our model differs from that of Sadinle 2017 through its explicit dependence on a beta random variable $\lambda$ that models the rate of matching across records.
-Sadinle marginalizes over such a random variable in his derivations of  the "beta prior for bipartite matching," but here we provide derivations without marginalizing in order to specify differing rates of matching for different linkage clusters.
-
-## Prior Distributions and Likelihood
+<!-- ## Prior Distributions and Likelihood -->
 
@@
 
+Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for the final Bayes estimate to link multiple records in $B$ to one record in $A$.
+The modeler can either report both such matches (with their respective posterior match probabilities), or resolve these conflicts by accepting only the match with highest posterior probability.
+Rigorous proof of this resolution procedure is deferred to future work.
+
 # Efficient Computation
@@
 Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns.
+Let $\mathcal{H}$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns.
 Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$.
@@
 
-We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $h_p$ with record $j \in B$.
-For each $p$, we also calculate $|h_p| = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$ total instances of agreement pattern $h_p$ throughout the data, and also for each $j$, we  calculate $|h_p|_j = \sum_{i=1}^{n_A} \mathbf{1}_{(i,j) \in h_p}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.
+We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $p$ with record $j \in B$.
+For each $p$, we also calculate $H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$, the total instances of agreement pattern $p$ throughout the data, and also for each $j$, we  calculate $H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.
 
-For large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\} for more manageable computations.
+For large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\}$ for more manageable computations.
 On a single machine, we can read-in data sequentially, conduct hashing, collect results, and delete the original data from memory before continuing with the next chunk of data.
@@
 In contrast, we recognize that each unique agreement pattern contributes to the posterior $\alpha(\mathbf{Z})$ and $\beta(\mathbf{Z})$ vectors in the same way.
-In fact, if we denote $|h_p^m| = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $h_p$, then the contribution of pairs of pattern $h_p$ to the $\alpha(\mathbf{Z})$ vector is simply $\mathbf{Z}(h_p) \times h_p$.
-Thus our posterior update for $\alpha$ is simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P |h_p^m| \times h_p$.
-Then, we can easily calculate $|h_p^u|$, the number of nonmatching record pairs of agreement pattern $p$, by subtracting the number of matching pairs from the total present in the data; that is $|h_p^u| = |h_p| - |h_p^m|$.
-From this, we can update our $\beta$ parameter through  $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p^u| \times h_p$.
+In fact, if we denote $H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$, then the contribution of pairs of pattern $p$ to the $\alpha(\mathbf{Z})$ vector is simply $H_p^m \times h_p$.
+Thus our posterior update for $\alpha$ is simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p$.
+Then, we can easily calculate $H_p^u$, the number of nonmatching record pairs of agreement pattern $p$, by subtracting the number of matching pairs from the total present in the data; that is $H_p^u = H_p - H_p^m$.
+From this, we can update our $\beta$ parameter through  $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p$.
 Note that these constitute $P$ many summations over $n_B$ quantities, and thus avoid the $n_A \times n_B$ summation from the original method.
@@
 Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching.
-Particular, this sampler removed previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*.By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel.
+In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*.
+By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel.
 More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns.
@@
 
-To do this we calculate the Fellegi Sunter weight $w_{h_p}$ for each unique pattern,  sample the agreement pattern between $j$ and its potential match, and then sample the record label uniformly among viable records.
-More concretely, define $H(Z_j)$ to be the agreement pattern between $j$ and its potential match, and say $H(Z_j) = h_{P+1}$ when $Z_j = n_A + 1$.
+To do this, we calculate the Fellegi Sunter weight $w_{h_p}$ for each unique pattern,  sample the agreement pattern between $j$ and its potential match, and then sample the record label uniformly among viable records.
+More concretely, define $h(Z_j)$ to be the agreement pattern between $j$ and its potential match, and say $h(Z_j) = h_{P+1}$ when $Z_j = n_A + 1$.
 Then,
 
-$$H\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+$$h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
 \begin{cases} 
-    w_{h_p}\times |h_p|_j  & h(z_j) \in H; \\
-     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
+    w_{h_p}\times H_{p_j}  & p \leq P; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
 \end{cases}$$
@@
 
-## Enforcing Bipartite Matching
-
-Add this
-
 # Simulation Studies
 
-- Replicating the Sadinle Study
+In his 2017 paper, Sadinle demonstrated the strength of his method by running `BRL` on simulated pairs of files with differing levels of errors and overlap.
+Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakened performance in the extreme scenario of very high errors and very high overlap across files.
+We use the following measures of accuracy in our analysis: although using only recall and precision is common in the literature, we find it useful to have a single metric the gauge overall performance, and thus also include F measure.
 
-- `parlr` and `BRL` speed comparison
+$$\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}$$
+$$\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}$$
+$$\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)$$
 
-# Data Analysis
+```{r fig.align='center'}
+knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
+```
 
-- Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?
+To demonstrate speed, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
+Distributions are meant to emulate the behavior of similarity scores across first name, last name, and day, month, and and year of birth, with 1 indicating an exact match on a particular field and 0 indicating a nonmatch.
+We simulate these data for different values of $n_A$ and $n_B$, and compare the runtime of `parlr` against `BRL`.
+We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data.
 
-# Discussion
+In particular, runtime for `BRL` seems to grow quadratically while runtime for `parlr` seems to grow linearly.
 
-We note that this often not a reasonable assumption; for example if we know that two records differ on gender, this does not give any information on whether their year of births are the same, but it does suggest that its more likely that their first names are different.
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrr}
+  \hline
+ & m & u \\ 
+  \hline
+fname\_0 & 0.05 & 0.99 \\ 
+  fname\_1 & 0.95 & 0.01 \\ 
+  lname\_0 & 0.05 & 0.99 \\ 
+  lname\_1 & 0.95 & 0.01 \\ 
+  day\_0 & 0.05 & 0.97 \\ 
+  day\_1 & 0.95 & 0.03 \\ 
+  month\_0 & 0.05 & 0.92 \\ 
+  month\_1 & 0.95 & 0.08 \\ 
+  year\_0 & 0.05 & 0.93 \\ 
+  year\_1 & 0.95 & 0.07 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+<!-- ```{r fig.align='center'} -->
+<!-- knitr::include_graphics("../notes/figures/speed_distributions.png") -->
+<!-- ``` -->
+
+```{r fig.align='center'}
+knitr::include_graphics("../notes/figures/sadinle_speed_plot.png")
+```
+
+The above discussion suggests that for fixed $n_B$, computation time should remain mostly constant with growing $n_A$.
+Shockingly, this seems to be true.
+In the plot below, fixing $n_B = 500$, we see linear growth for the runtime under `BRL` as $n_A$ increases, with much more static runtime under `parlr`.
+The slight increases in runtime that we do see are due primarily to the hashing step, which again can be run in parallel for large data.
+
+```{r fig.align='center'}
+knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
+```
+
+We note here that `BRL` is coded in C, which makes for unfair comparison against `parlr`, currently only built in R.
+Additionally, although `parlr` is amenable to parallezation, this simulation was run on a single core.
+Running `parlr` in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more drastic results.

### 2021-04-19 paper/appendix_STA561_project.tex
GetPossiblePatterns function

--- 2021-04-12 paper/parlr_paper_1.tex
@@
     \end{flushright}
-    \begin{center} \textbf{Parlr: Parallelized Record Linkage in R}
+    \begin{center} \textbf{Parlr: Parallelized Record Linkage in R} \linebreak
+    \textbf{Technical Appendix}
 
@@
 
+    April 23, 2021
+
+    STA 561
+
     \end{center}
-
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
-
-Most methods for conducting record linkage, the task of identifying
-duplicate records across datasets, are derived from the foundational
-article by Fellegi and Sunter in 1969. Within this family of methods,
-two methods have notable strengths: \texttt{fastlink}, proposed by
-Enamorado et al (2019), continues a is able to conduct linkage on
-incredibly large datasets
 
@@
 
-Our notation and assumptions closely follow that of Sadinle (2017).
 Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
-respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
-in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
+respectively, with records indexed as \(i \in \{1, \ldots, n_A\}\) in
+\(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
 generality, label the files such that \(n_A \geq n_B\). We also assume
-there are no duplicates within files, only across. For each record pair
-under consideration, we generate a comparison vector
+there are no duplicates within files, only across.
+
+For each record pair under consideration, we generate a comparison
+vector
 \(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
@@
 \(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
-indicating level agreement between the two records on a specified field.
+indicating the level agreement between the two records on a specified
+field. For ease of modeling and computation, we restrict these
+similarity scores to be discrete, ordinal variables, and the
+construction of these is left to the modeler. With notation
+\(\gamma_{ij}^f \in \{0, \ldots, L_f - 1\}\) for the number of possible
+agreement levels for field \(f\), we use binary 0-1 variables to
+indicate exact matching, and 0-1-2 variables to provide an option for
+partial matching. For text data, we calculate similarity based on
+Levenstein distance or some other text similarity score, and bin these
+scores to integers for use in the model.
 
@@
 \(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
-is product of the relevant of the appropriate \(\mathbf{m}\) parameters
-for record pair \((i,j)\). We hope that these conventions avoid
-overloaded notation in the likelihood and subsequent derivations.
+is the product of the appropriate \(\mathbf{m}\) parameters for record
+pair \((i,j)\). We hope that these conventions avoid overloaded notation
+in the likelihood and subsequent derivations.
 
@@
 \section{Model Specification}\label{model-specification}}
-
-(This will be revised to talk more explicitly about sampling \(Z_j\)
-independently)
-
-Our model differs from that of Sadinle 2017 through its explicit
-dependence on a beta random variable \(\lambda\) that models the rate of
-matching across records. Sadinle marginalizes over such a random
-variable in his derivations of the ``beta prior for bipartite
-matching,'' but here we provide derivations without marginalizing in
-order to specify differing rates of matching for different linkage
-clusters.
-
-\hypertarget{prior-distributions-and-likelihood}{%
-\subsection{Prior Distributions and
-Likelihood}\label{prior-distributions-and-likelihood}}
 
@@
 
+Since our Gibbs procedure does not strictly enforce one-to-one matching,
+it is possible for the final Bayes estimate to link multiple records in
+\(B\) to one record in \(A\). The modeler can either report both such
+matches (with their respective posterior match probabilities), or
+resolve these conflicts by accepting only the match with highest
+posterior probability. Rigorous proof of this resolution procedure is
+deferred to future work.
+
 \hypertarget{efficient-computation}{%
@@
 recognizing that record pairs contribute to posterior calculations only
-through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(H\)
-be the set of unique agreement patterns in the data, let \(P\) denote
-the total number of unique agreement patterns. Note that \(P\) is
-bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not
-scale with \(n_A\) or \(n_B\). We index these agreement patterns by
-\(p \in \{1, \ldots, P\}\), and say \((i,j) \in h_p\) when the \((i,j)\)
-pair exhibits the \(p^{th}\) agreement pattern. Wherever possible, we
-conduct calculations over these \(P\) agreement patterns rather than the
-\(n_A \times n_B\) record pairs.
+through the agreement pattern of the \(\gamma_{ij}\) vector. Let
+\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
+\(P\) denote the total number of unique agreement patterns. Note that
+\(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound
+does not scale with \(n_A\) or \(n_B\). We index these agreement
+patterns by \(p \in \{1, \ldots, P\}\), and say \((i,j) \in h_p\) when
+the \((i,j)\) pair exhibits the \(p^{th}\) agreement pattern. Wherever
+possible, we conduct calculations over these \(P\) agreement patterns
+rather than the \(n_A \times n_B\) record pairs.
 
@@
 \(p^{th}\) component of the \(j^{th}\) list contains a vector of records
-in \(A\) that share agreement pattern \(h_p\) with record \(j \in B\).
-For each \(p\), we also calculate
-\(|h_p| = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\)
-total instances of agreement pattern \(h_p\) throughout the data, and
+in \(A\) that share agreement pattern \(p\) with record \(j \in B\). For
+each \(p\), we also calculate
+\(H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\),
+the total instances of agreement pattern \(p\) throughout the data, and
 also for each \(j\), we calculate
-\(|h_p|_j = \sum_{i=1}^{n_A} \mathbf{1}_{(i,j) \in h_p}\) the instances
-of agreement pattern \(p\) among the comparison vectors between record
-\(j \in B\) and each of the \(n_A\) records in \(A\).
+\(H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}\) the
+instances of agreement pattern \(p\) among the comparison vectors
+between record \(j \in B\) and each of the \(n_A\) records in \(A\).
 
 For large data, we can partition the two datasets \(A\) and \(B\) into
-smaller blocks \(\{A_m\}\) and \$\{B\_m\} for more manageable
+smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more manageable
 computations. On a single machine, we can read-in data sequentially,
@@
 in the same way. In fact, if we denote
-\(\mathbf{Z}(h_p) = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to
-be the number of matching record pairs with agreement pattern \(h_p\),
-then the contribution of pairs of pattern \(h_p\) to the
-\(\alpha(\mathbf{Z})\) vector is simply \(\mathbf{Z}(h_p) \times h_p\).
-Thus our posterior updates are simply
-\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p\)
-and
-\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)\).
-Note that this constitutes \(P\) many summations over \(n_B\)
-quantities, and thus avoids the \(n_A \times n_B\) summation from the
-original method.
+\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
+number of matching record pairs with agreement pattern \(p\), then the
+contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
+vector is simply \(H_p^m \times h_p\). Thus our posterior update for
+\(\alpha\) is simply
+\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
+we can easily calculate \(H_p^u\), the number of nonmatching record
+pairs of agreement pattern \(p\), by subtracting the number of matching
+pairs from the total present in the data; that is
+\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
+parameter through
+\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
+that these constitute \(P\) many summations over \(n_B\) quantities, and
+thus avoid the \(n_A \times n_B\) summation from the original method.
 
 Sadinle uses a prior for \(\mathbf{Z}\) that induces the a full
-conditional for \(Z_j\) that strictly enforces one-to-one matching.
-Particular, this sampler removed previously matches records from the set
+conditional for \(Z_j\) that strictly enforces one-to-one matching. In
+particular, this sampler removes previously matches records from the set
 of candidate records when sampling \(Z_j\), creating a dependency that
-makes the sampler \emph{inherently serial}.By weakening the one-to-one
+makes the sampler \emph{inherently serial}. By weakening the one-to-one
 requirement, our full conditional for \(Z\) does not depend on the rest
@@
 
-To do this we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
-unique pattern and sample the agreement pattern \(h(Z_j)\) following the
-modified full conditional:
+To do this, we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
+unique pattern, sample the agreement pattern between \(j\) and its
+potential match, and then sample the record label uniformly among viable
+records. More concretely, define \(h(Z_j)\) to be the agreement pattern
+between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when
+\(Z_j = n_A + 1\). Then,
 
@@
 \begin{cases} 
-    w_{h_p}\times |h_p|_j  & h(z_j) \in H; \\
-     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
+    w_{h_p}\times H_{p_j}  & p \leq P; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
 \end{cases}\]
@@
 
-\hypertarget{enforcing-bipartite-matching}{%
-\subsection{Enforcing Bipartite
-Matching}\label{enforcing-bipartite-matching}}
-
-Add this
-
 \hypertarget{simulation-studies}{%
@@
 
-\begin{itemize}
-\item
-  Replicating the Sadinle Study
-\item
-  \texttt{parlr} and \texttt{BRL} speed comparison
-\end{itemize}
+In his 2017 paper, Sadinle demonstrated the strength of his method by
+running \texttt{BRL} on simulated pairs of files with differing levels
+of errors and overlap. Comparing our method against \texttt{BRL} on
+these same simulated datasets, we find that our method only has weakened
+performance in the extreme scenario of very high errors and very high
+overlap across files. We use the following measures of accuracy in our
+analysis: although using only recall and precision is common in the
+literature, we find it useful to have a single metric the gauge overall
+performance, and thus also include F measure.
 
-\hypertarget{data-analysis}{%
-\section{Data Analysis}\label{data-analysis}}
+\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
+\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
+\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
 
-\begin{itemize}
-\tightlist
-\item
-  Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?
-\end{itemize}
+\begin{center}\includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot} \end{center}
+
+To demonstrate speed, we generate comparison vectors from pre-specified
+distributions so that we can easily increase the size of the linkage
+problem. Distributions are meant to emulate the behavior of similarity
+scores across first name, last name, and day, month, and and year of
+birth, with 1 indicating an exact match on a particular field and 0
+indicating a nonmatch. We simulate these data for different values of
+\(n_A\) and \(n_B\), and compare the runtime of \texttt{parlr} against
+\texttt{BRL}. We see that at low data size, \texttt{BRL} outperforms,
+but that \texttt{parlr} is significantly faster at handling larger data.
+
+In particular, runtime for \texttt{BRL} seems to grow quadratically
+while runtime for \texttt{parlr} seems to grow linearly.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrr}
+  \hline
+ & m & u \\ 
+  \hline
+fname\_0 & 0.05 & 0.99 \\ 
+  fname\_1 & 0.95 & 0.01 \\ 
+  lname\_0 & 0.05 & 0.99 \\ 
+  lname\_1 & 0.95 & 0.01 \\ 
+  day\_0 & 0.05 & 0.97 \\ 
+  day\_1 & 0.95 & 0.03 \\ 
+  month\_0 & 0.05 & 0.92 \\ 
+  month\_1 & 0.95 & 0.08 \\ 
+  year\_0 & 0.05 & 0.93 \\ 
+  year\_1 & 0.95 & 0.07 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{center}\includegraphics[width=23.6in]{../notes/figures/sadinle_speed_plot} \end{center}
+
+The above discussion suggests that for fixed \(n_B\), computation time
+should remain mostly constant with growing \(n_A\). Shockingly, this
+seems to be true. In the plot below, fixing \(n_B = 500\), we see linear
+growth for the runtime under \texttt{BRL} as \(n_A\) increases, with
+much more static runtime under \texttt{parlr}. The slight increases in
+runtime that we do see are due primarily to the hashing step, which
+again can be run in parallel for large data.
+
+\begin{center}\includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB} \end{center}
+
+We note here that \texttt{BRL} is coded in C, which makes for unfair
+comparison against \texttt{parlr}, currently only built in R.
+Additionally, although \texttt{parlr} is amenable to parallezation, this
+simulation was run on a single core. Running \texttt{parlr} in C++ with
+paralellization for the hashing step and sampling the matching status of
+the record pairs should lead to even more drastic results.
 

### 2021-04-19 R/PossiblePatterns.R

@@ -1,0 +1,22 @@
+FS_to_Sadinle <- function(gamma, levels){
+  unname(unlist(mapply(function(z, y){
+    gamma_f <- rep(0, y)
+    gamma_f[z] <- 1
+    gamma_f
+  }, z = gamma, y = levels)))
+}
+
+GetPossiblePatterns <- function(levels){
+  possible_values <- lapply(levels, function(x){
+    seq_len(x)
+  })
+  possible_patterns <- data.frame(do.call(expand.grid, possible_values))
+
+  thing <- data.frame(t(apply(possible_patterns, 1, function(x){
+  FS_to_Sadinle(x, levels)
+  })))
+  thing %>%
+    unite(patterns, 1:dim(.)[2], sep = "") %>%
+    unlist() %>%
+    unname()
+}

### 2021-04-25 notes/Notes_4_23_21.Rmd
GetPossiblePatterns function
faster SimulateComparisons function

@@
 
+- Example in this [HRDAG post](https://hrdag.org/2013/03/07/hrdags-record-de-duplication/).
+
 - Processing these smaller chunks can be done in parallel.
 
-- Should allow us to conduct *significantly* larger record linkage tasks.
+- Should allow us to conduct **much** larger record linkage tasks.
 
@@
 - These mechanics will lay the foundation (with slight modifications) for the linkage cluster model I've been developing. This is on hold until the first `parlr` paper is done.
+
+- Plan on meeting next week (Friday, April 30th), but two weeks of vacation after that.

### 2021-04-25 paper/FAQ_STA561_Project.Rmd
faster SimulateComparisons function

@@
 Most record linkage techniques are derived from the seminal 1969 paper by Fellegi and Sunter: "A Theory for Record Linkage".
-The defining characteristic of their model was to transform the sets of records, which constitute text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated.
+The defining characteristic of their model was to transform the sets of records, which often contain text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated.
 Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates a $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets.

### 2021-04-25 paper/appendix_STA561_project.Rmd
faster SimulateComparisons function

@@
 In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair.
-We however do not store these comparison vectors themselves, but instead only the a hashed value corresponding to the agreement pattern of the $(i, j)$ pair.
+We however do not store these comparison vectors themselves, but instead only a hashed value corresponding to the agreement pattern of the $(i, j)$ pair.
 Enamorado et al (2019) provided the hashing function

### 2021-04-25 paper/parlr_paper_1.Rmd
faster SimulateComparisons function

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
 
-## Introduction
+## Related Work
 
-Most methods for conducting record linkage, the task of identifying duplicate records across datasets, are derived from the foundational article by Fellegi and Sunter in 1969.
-Within this family of methods, two methods have notable strengths: `fastlink`, proposed by Enamorado et al (2019), continues a is able to conduct linkage on incredibly large datasets
+Most record linkage techniques are derived from the seminal 1969 paper by Fellegi and Sunter: "A Theory for Record Linkage".
+The defining characteristic of their model was to transform the sets of records, which often contain text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated.
+Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates a $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets.
+We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f$ providing their similarity score on the $f^{th}$ field.
+For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeller.
+It is common to use use binary 0-1 variables to indicate exact matching, and 0-1-2 variables to provide an option for partial matching.
+For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model.
+
+Two modern adaptations of the Fellegi Sunter model are important for understanding `parlr` and its contribution.
+In 2019, Enamorado et al proposed `fastlink`, a method that closely followed the modelling assumptions of Fellegi and Sunter, but used innovating hashing techniques to greatly enhance its speed and scale.
+Importantly, they maintained Fellegi and Sunter's independent matching assumption, in which the matching status of any $(i,j)$ pair is made independently of the matching of any other $(i', j')$.
+This leads to many matchings that violate "one-to-one" considerations that are often implicit in the data, which need to be resolved in a post-processing step.
+In contrast, Sadinle (2017) proposed "Beta Record Linkage for Bipartite Matching" and the accompanying `BRL` package which strictly enforces one-to-one matching.
+Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, then samples a potential link.
+However, accounting for these dependencies throughout the linkage process is computionally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.
+
+Outside of the Fellegi-Sunter framework, some researchers have developed methods that model the data directly, not a derived set of comparison vectors.
+As an early example, Steorts et al 2016 produced `blink`, which had the advantage of a computational complexity that grew linearly with the data (as opposed to quadratically within Fellegi Sunter), but was plagued with slow mixing times and was unable to incorportate text data.
+Later, Steorts et al 2020 improved on this method with `dblink`, which was able to simulate text data within its Gibbs sampler by drawing from an empirical distribution, and remedied its slow mixing through a probabalistic blocking approach.
+However, this method remains computationally intensive and requires distributed computing to manage even moderate linkage problems in acceptable time.
 
@@
 We also assume there are no duplicates within files, only across.
-For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating level agreement between the two records on a specified field.
+For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating the level agreement between the two records on a specified field.
 
@@
 
+Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for the final Bayes estimate to link mulitple records in $B$ to one record in $A$.
+The modeller can either report both such matches (with their respective posterior match probabilities), or resolve these conflicts by accepting only the match with highest posterior probability.
+Rigorous proof of this resolution procedure is deferred to future work.
+
 # Efficient Computation
@@
 Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns.
+Let $\mathcal{H}$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns.
 Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$.
-We index these agreement patterns by $p \in \{1, \ldots, P\}$, and say $(i,j) \in h_p$ when the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern.
+Prior to processing the data, we identify all $P$ patterns in $\mathcal{H}$ and enumerate them $h_1, \ldots, h_P$.
+When the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern, we say $(i,j) \in h_p$.
 Wherever possible, we conduct calculations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.
 
-## Data Representation and Hashing
+## Data Representation, Hashing, and Storage
+
+The classic Fellegi Sunter method represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$.
+To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicator vectors of lengths $L_f$.
+For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework.
+This is a bijective transformation that does not change the meaning of the data, but this representation will ease calculations and posterior updates.
 
 In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair.
-We however do not store these comparison vectors themselves, but instead only the a hashed value corresponding to the agreement pattern of the $(i, j)$ pair.
-Enamorado et al (2019) provided the hashing function
-$$\sum_{f=1}^F \mathbf{1}_{\gamma_{(i,j)}^f >0}2^{\gamma_{(i,j)}^f + \mathbf{1}_{f>1} \times \sum_{e=1}^{f-1}(L_e-1)}$$
-to map each agreement pattern to a unique value, but packages like `dplyr` in `R` are capable of this as well.
+We however do not store these comparison vectors themselves, but instead only a hashed value corresponding to the agreement pattern of the $(i, j)$ pair.
+We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $p$ with record $j \in B$.
+For each $p$, we also calculate $H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$, the total instances of agreement pattern $p$ throughout the data, and also for each $j$, we  calculate $H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.
 
-We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $h_p$ with record $j \in B$.
-For each $p$, we also calculate $|h_p| = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$ total instances of agreement pattern $h_p$ throughout the data, and also for each $j$, we  calculate $|h_p|_j = \sum_{i=1}^{n_A} \mathbf{1}_{(i,j) \in h_p}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.
-
-For large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\} for more manageable computations.
+For large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\}$ for more manageable computations.
 On a single machine, we can read-in data sequentially, conduct hashing, collect results, and delete the original data from memory before continuing with the next chunk of data.
@@
 
-Lastly, the classic Fellegi Sunter method represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$.
-To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicator vectors of lengths $L_f$.
-For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework.
-This is a bijective transformation that does not change the meaning of the data, but this representation will ease calculations and posterior updates.
+The hashing procedure described above considerably reduces the memory needed to store the comparison information: instead of storing $n_A \times n_B$ comparison vectors, which are relatively long under either the Fellegi Sunter or our modified framework, we only store the $P$ unique vectors, and then $n_A \times n_B$ scalar quantities relating record pairs to those vectors.
+However, even storing these $n_A \times n_B$ scalar labels can become burdensome with large data.
+Worse, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
+
+To address this, we propose a new strategy called *random indexing*.
+In standard indexing, the modeller decides a certain criteria that they expect all true matching pairs to satisfy, and a priori label any record pairs that do not meet that criteria as non-matching.
+For example, one might only consider pairs with a certain similarity score on a field deemed to be important (like first name), or only pairs with exact matching on a specified number of fields.
+While generally chosen to be be quite loose, establishing these criteria requires knowledge of the problem and invites room for human error.
+We propose a method of reducing the comparison space (and reducing storage requirements) without these drawbacks.
+Note that all records of the same agreement pattern have the same probality when sampling $Z_j$.
+Therefore we know that records belonging to an $h_p$ such that $H_{p_j}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate, even without considering the form of the agreement pattern itself.
+In random indexing, rather than store all of these unlikely record labels, we choose to store only a small number $R$ of them.
+Posterior calculations still attribute the appropiate weight to all records through the summary statistics $H_{p_j}, H_p^m$, and $H_p^u$; this is similar to the *u-correction* concept introduced by ________.
+Rather than storing $n_A \times n_B$ record labels, random indexing allows us to store at most $n_B \times P \times R$, regardless of how large $n_A$ is.
+Coupled with the partitioning strategy for creating comparison vectors, this greatly expands the size of the linkage tasks we can undertake.
 
@@
 
-## Enforcing Bipartite Matching
-
-Add this
-
 # Simulation Studies
 
-- Replicating the Sadinle Study
+In his 2017 paper, Sadinle demonstrated the strength of his method by running `BRL` on simulated pairs of files with differing levels of errors and overlap.
+Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakened performance in the extreme scenario of very high errors and very high overlap across files.
+We use the following measures of accuracy in our analysis: although using only recall and precision is common in the literature, we find it useful to have a single metric the gauge overall performance, and thus also include F measure.
 
-- `parlr` and `BRL` speed comparison
+$$\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}$$
+$$\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}$$
+$$\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)$$
+
+```{r fig.align='center'}
+knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
+```
+
+To demonstrate speed, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
+Distributions are meant to emulate the behavior of similarity scores across first name, last name, and day, month, and and year of birth, with 1 indicating an exact match on a particular field and 0 indicating a nonmatch.
+We simulate these data for different values of $n_A$ and $n_B$, and compare the runtime of `parlr` against `BRL`.
+We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data.
+
+In particular, runtime for `BRL` seems to grow quadratically while runtime for `parlr` seems to grow linearly.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrr}
+  \hline
+ & m & u \\ 
+  \hline
+fname\_0 & 0.05 & 0.99 \\ 
+  fname\_1 & 0.95 & 0.01 \\ 
+  lname\_0 & 0.05 & 0.99 \\ 
+  lname\_1 & 0.95 & 0.01 \\ 
+  day\_0 & 0.05 & 0.97 \\ 
+  day\_1 & 0.95 & 0.03 \\ 
+  month\_0 & 0.05 & 0.92 \\ 
+  month\_1 & 0.95 & 0.08 \\ 
+  year\_0 & 0.05 & 0.93 \\ 
+  year\_1 & 0.95 & 0.07 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+<!-- ```{r fig.align='center'} -->
+<!-- knitr::include_graphics("../notes/figures/speed_distributions.png") -->
+<!-- ``` -->
+
+```{r fig.align='center'}
+knitr::include_graphics("../notes/figures/sadinle_speed_plot.png")
+```
+
+The above discussion suggests that for fixed $n_B$, computation time should remain mostly constant with growing $n_A$.
+Shockingly, this seems to be true.
+In the plot below, fixing $n_B = 500$, we see linear growth for the runtime under `BRL` as $n_A$ increases, with much more static runtime under `parlr`.
+The slight increases in runtime that we do see are due primarily to the hashing step, which again can be run in parallel for large data.
+
+```{r fig.align='center'}
+knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
+```
+
+We note here that `BRL` is coded in C, which makes for unfair comparison against `parlr`, currently only built in R.
+Additionally, although `parlr` is amenable to parallezation, this simulation was run on a single core.
+Running `parlr` in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more drastic results.
 

### 2021-04-25 paper/parlr_paper_1.tex
faster SimulateComparisons function

@@
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\hypertarget{related-work}{%
+\subsection{Related Work}\label{related-work}}
 
-Most methods for conducting record linkage, the task of identifying
-duplicate records across datasets, are derived from the foundational
-article by Fellegi and Sunter in 1969. Within this family of methods,
-two methods have notable strengths: \texttt{fastlink}, proposed by
-Enamorado et al (2019), continues a is able to conduct linkage on
-incredibly large datasets
+Most record linkage techniques are derived from the seminal 1969 paper
+by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
+characteristic of their model was to transform the sets of records,
+which often contain text data that is difficult to model, into sets of
+comparison vectors governed by parameters that can be more easily
+estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
+records respectively, and if the files share \(F\) fields in common upon
+which to base the linkage, the Fellegi and Sunter approach generates a
+\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
+between each pair of records across datasets. We say \(\gamma_{ij}\) is
+the comparison vector for record \(i \in A\) and record \(j \in B\),
+with \(\gamma_{ij}^f\) providing their similarity score on the
+\(f^{th}\) field. For ease of modeling and computation, we restrict
+these similarity scores to be discrete, ordinal variables, and the
+construction of these is left to the modeller. It is common to use use
+binary 0-1 variables to indicate exact matching, and 0-1-2 variables to
+provide an option for partial matching. For text data, we calculate
+similarity based on Levenstein distance or some other text similarity
+score, and bin these scores to integers for use in the model.
+
+Two modern adaptations of the Fellegi Sunter model are important for
+understanding \texttt{parlr} and its contribution. In 2019, Enamorado et
+al proposed \texttt{fastlink}, a method that closely followed the
+modelling assumptions of Fellegi and Sunter, but used innovating hashing
+techniques to greatly enhance its speed and scale. Importantly, they
+maintained Fellegi and Sunter's independent matching assumption, in
+which the matching status of any \((i,j)\) pair is made independently of
+the matching of any other \((i', j')\). This leads to many matchings
+that violate ``one-to-one'' considerations that are often implicit in
+the data, which need to be resolved in a post-processing step. In
+contrast, Sadinle (2017) proposed ``Beta Record Linkage for Bipartite
+Matching'' and the accompanying \texttt{BRL} package which strictly
+enforces one-to-one matching. Specifically, in each iteration of his
+Gibbs sampler, he considers each record \(j\in B\), removes from
+consideration the records \(i\in A\) that have already been matched,
+then samples a potential link. However, accounting for these
+dependencies throughout the linkage process is computionally burdensome,
+leaving \texttt{BRL} only suitable for small to moderate linkage
+problems.
+
+Outside of the Fellegi-Sunter framework, some researchers have developed
+methods that model the data directly, not a derived set of comparison
+vectors. As an early example, Steorts et al 2016 produced
+\texttt{blink}, which had the advantage of a computational complexity
+that grew linearly with the data (as opposed to quadratically within
+Fellegi Sunter), but was plagued with slow mixing times and was unable
+to incorportate text data. Later, Steorts et al 2020 improved on this
+method with \texttt{dblink}, which was able to simulate text data within
+its Gibbs sampler by drawing from an empirical distribution, and
+remedied its slow mixing through a probabalistic blocking approach.
+However, this method remains computationally intensive and requires
+distributed computing to manage even moderate linkage problems in
+acceptable time.
 
@@
 \(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
-indicating level agreement between the two records on a specified field.
+indicating the level agreement between the two records on a specified
+field.
 
@@
 
+Since our Gibbs procedure does not strictly enforce one-to-one matching,
+it is possible for the final Bayes estimate to link mulitple records in
+\(B\) to one record in \(A\). The modeller can either report both such
+matches (with their respective posterior match probabilities), or
+resolve these conflicts by accepting only the match with highest
+posterior probability. Rigorous proof of this resolution procedure is
+deferred to future work.
+
 \hypertarget{efficient-computation}{%
@@
 recognizing that record pairs contribute to posterior calculations only
-through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(H\)
-be the set of unique agreement patterns in the data, let \(P\) denote
-the total number of unique agreement patterns. Note that \(P\) is
-bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not
-scale with \(n_A\) or \(n_B\). We index these agreement patterns by
-\(p \in \{1, \ldots, P\}\), and say \((i,j) \in h_p\) when the \((i,j)\)
-pair exhibits the \(p^{th}\) agreement pattern. Wherever possible, we
+through the agreement pattern of the \(\gamma_{ij}\) vector. Let
+\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
+\(P\) denote the total number of unique agreement patterns. Note that
+\(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound
+does not scale with \(n_A\) or \(n_B\). Prior to processing the data, we
+identify all \(P\) patterns in \(\mathcal{H}\) and enumerate them
+\(h_1, \ldots, h_P\). When the \((i,j)\) pair exhibits the \(p^{th}\)
+agreement pattern, we say \((i,j) \in h_p\). Wherever possible, we
 conduct calculations over these \(P\) agreement patterns rather than the
@@
 
-\hypertarget{data-representation-and-hashing}{%
-\subsection{Data Representation and
-Hashing}\label{data-representation-and-hashing}}
+\hypertarget{data-representation-hashing-and-storage}{%
+\subsection{Data Representation, Hashing, and
+Storage}\label{data-representation-hashing-and-storage}}
 
-In the classic Fellegi Sunter framework, \(\Gamma\) is a
-\(n_A n_B \times F\) matrix, with each row providing the comparison
-vector for a different \((i,j)\) pair. We however do not store these
-comparison vectors themselves, but instead only the a hashed value
-corresponding to the agreement pattern of the \((i, j)\) pair. Enamorado
-et al (2019) provided the hashing function
-\[\sum_{f=1}^F \mathbf{1}_{\gamma_{(i,j)}^f >0}2^{\gamma_{(i,j)}^f + \mathbf{1}_{f>1} \times \sum_{e=1}^{f-1}(L_e-1)}\]
-to map each agreement pattern to a unique value, but packages like
-\texttt{dplyr} in \texttt{R} are capable of this as well.
-
-We store this information in a nested list \(\tilde{\Gamma}\) where the
-\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
-in \(A\) that share agreement pattern \(h_p\) with record \(j \in B\).
-For each \(p\), we also calculate
-\(|h_p| = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\)
-total instances of agreement pattern \(h_p\) throughout the data, and
-also for each \(j\), we calculate
-\(|h_p|_j = \sum_{i=1}^{n_A} \mathbf{1}_{(i,j) \in h_p}\) the instances
-of agreement pattern \(p\) among the comparison vectors between record
-\(j \in B\) and each of the \(n_A\) records in \(A\).
-
-For large data, we can partition the two datasets \(A\) and \(B\) into
-smaller blocks \(\{A_m\}\) and \$\{B\_m\} for more manageable
-computations. On a single machine, we can read-in data sequentially,
-conduct hashing, collect results, and delete the original data from
-memory before continuing with the next chunk of data. With multiple
-cores or multiple machines, this can be done in parallel. Storing this
-hashed information still becomes burdensome for large data, but this
-hashing method greatly expands the capabilities of the Fellegi-Sunter
-framework.
-
-Lastly, the classic Fellegi Sunter method represents the \(\gamma_{ij}\)
+The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
 comparison vector as vector of length \(F\), with each component
@@
 but this representation will ease calculations and posterior updates.
+
+In the classic Fellegi Sunter framework, \(\Gamma\) is a
+\(n_A n_B \times F\) matrix, with each row providing the comparison
+vector for a different \((i,j)\) pair. We however do not store these
+comparison vectors themselves, but instead only a hashed value
+corresponding to the agreement pattern of the \((i, j)\) pair. We store
+this information in a nested list \(\tilde{\Gamma}\) where the
+\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
+in \(A\) that share agreement pattern \(p\) with record \(j \in B\). For
+each \(p\), we also calculate
+\(H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\),
+the total instances of agreement pattern \(p\) throughout the data, and
+also for each \(j\), we calculate
+\(H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}\) the
+instances of agreement pattern \(p\) among the comparison vectors
+between record \(j \in B\) and each of the \(n_A\) records in \(A\).
+
+For large data, we can partition the two datasets \(A\) and \(B\) into
+smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more manageable
+computations. On a single machine, we can read-in data sequentially,
+conduct hashing, collect results, and delete the original data from
+memory before continuing with the next chunk of data. With multiple
+cores or multiple machines, this can be done in parallel. Storing this
+hashed information still becomes burdensome for large data, but this
+hashing method greatly expands the capabilities of the Fellegi-Sunter
+framework.
+
+The hashing procedure described above considerably reduces the memory
+needed to store the comparison information: instead of storing
+\(n_A \times n_B\) comparison vectors, which are relatively long under
+either the Fellegi Sunter or our modified framework, we only store the
+\(P\) unique vectors, and then \(n_A \times n_B\) scalar quantities
+relating record pairs to those vectors. However, even storing these
+\(n_A \times n_B\) scalar labels can become burdensome with large data.
+Worse, the overwhelming majority of these labels relate to record pairs
+that are clear non-matches.
+
+To address this, we propose a new strategy called \emph{random
+indexing}. In standard indexing, the modeller decides a certain criteria
+that they expect all true matching pairs to satisfy, and a priori label
+any record pairs that do not meet that criteria as non-matching. For
+example, one might only consider pairs with a certain similarity score
+on a field deemed to be important (like first name), or only pairs with
+exact matching on a specified number of fields. While generally chosen
+to be be quite loose, establishing these criteria requires knowledge of
+the problem and invites room for human error. We propose a method of
+reducing the comparison space (and reducing storage requirements)
+without these drawbacks. Note that all records of the same agreement
+pattern have the same probality when sampling \(Z_j\). Therefore we know
+that records belonging to an \(h_p\) such that \(H_{p_j}\) is large are
+very unlikely to be sampled consistently enough to be deemed a match
+through the Bayes estimate, even without considering the form of the
+agreement pattern itself. In random indexing, rather than store all of
+these unlikely record labels, we choose to store only a small number
+\(R\) of them. Posterior calculations still attribute the appropiate
+weight to all records through the summary statistics \(H_{p_j}, H_p^m\),
+and \(H_p^u\); this is similar to the \emph{u-correction} concept
+introduced by \_\_\_\_\_\_\_\_. Rather than storing \(n_A \times n_B\)
+record labels, random indexing allows us to store at most
+\(n_B \times P \times R\), regardless of how large \(n_A\) is. Coupled
+with the partitioning strategy for creating comparison vectors, this
+greatly expands the size of the linkage tasks we can undertake.
 
@@
 in the same way. In fact, if we denote
-\(|h_p^m| = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
-number of matching record pairs with agreement pattern \(h_p\), then the
-contribution of pairs of pattern \(h_p\) to the \(\alpha(\mathbf{Z})\)
-vector is simply \(\mathbf{Z}(h_p) \times h_p\). Thus our posterior
-update for \(\alpha\) is simply
-\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P |h_p^m| \times h_p\).
-Then, we can easily calculate \(|h_p^u|\), the number of nonmatching
-record pairs of agreement pattern \(p\), by subtracting the number of
-matching pairs from the total present in the data; that is
-\(|h_p^u| = |h_p| - |h_p^m|\). From this, we can update our \(\beta\)
+\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
+number of matching record pairs with agreement pattern \(p\), then the
+contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
+vector is simply \(H_p^m \times h_p\). Thus our posterior update for
+\(\alpha\) is simply
+\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
+we can easily calculate \(H_p^u\), the number of nonmatching record
+pairs of agreement pattern \(p\), by subtracting the number of matching
+pairs from the total present in the data; that is
+\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
 parameter through
-\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p^u| \times h_p\). Note
+\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
 that these constitute \(P\) many summations over \(n_B\) quantities, and
-thus avoids the \(n_A \times n_B\) summation from the original method.
+thus avoid the \(n_A \times n_B\) summation from the original method.
 
 Sadinle uses a prior for \(\mathbf{Z}\) that induces the a full
-conditional for \(Z_j\) that strictly enforces one-to-one matching.
-Particular, this sampler removed previously matches records from the set
+conditional for \(Z_j\) that strictly enforces one-to-one matching. In
+particular, this sampler removes previously matches records from the set
 of candidate records when sampling \(Z_j\), creating a dependency that
-makes the sampler \emph{inherently serial}.By weakening the one-to-one
+makes the sampler \emph{inherently serial}. By weakening the one-to-one
 requirement, our full conditional for \(Z\) does not depend on the rest
@@
 
-To do this we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
+To do this, we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
 unique pattern, sample the agreement pattern between \(j\) and its
 potential match, and then sample the record label uniformly among viable
-records. More concretely, define \(H(Z_j)\) to be the agreement pattern
-between \(j\) and its potential match, and say \(H(Z_j) = h_{P+1}\) when
+records. More concretely, define \(h(Z_j)\) to be the agreement pattern
+between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when
 \(Z_j = n_A + 1\). Then,
 
-\[H\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\[h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
 \begin{cases} 
-    w_{h_p}\times |h_p|_j  & h(z_j) \in H; \\
-     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
+    w_{h_p}\times H_{p_j}  & p \leq P; \\
+     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
 \end{cases}\]
@@
 
-\begin{itemize}
-\item
-  Replicating the Sadinle Study
-\item
-  \texttt{parlr} and \texttt{BRL} speed comparison
-\end{itemize}
+In his 2017 paper, Sadinle demonstrated the strength of his method by
+running \texttt{BRL} on simulated pairs of files with differing levels
+of errors and overlap. Comparing our method against \texttt{BRL} on
+these same simulated datasets, we find that our method only has weakened
+performance in the extreme scenario of very high errors and very high
+overlap across files. We use the following measures of accuracy in our
+analysis: although using only recall and precision is common in the
+literature, we find it useful to have a single metric the gauge overall
+performance, and thus also include F measure.
+
+\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
+\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
+\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
+
+\begin{center}\includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot} \end{center}
+
+To demonstrate speed, we generate comparison vectors from pre-specified
+distributions so that we can easily increase the size of the linkage
+problem. Distributions are meant to emulate the behavior of similarity
+scores across first name, last name, and day, month, and and year of
+birth, with 1 indicating an exact match on a particular field and 0
+indicating a nonmatch. We simulate these data for different values of
+\(n_A\) and \(n_B\), and compare the runtime of \texttt{parlr} against
+\texttt{BRL}. We see that at low data size, \texttt{BRL} outperforms,
+but that \texttt{parlr} is significantly faster at handling larger data.
+
+In particular, runtime for \texttt{BRL} seems to grow quadratically
+while runtime for \texttt{parlr} seems to grow linearly.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrr}
+  \hline
+ & m & u \\ 
+  \hline
+fname\_0 & 0.05 & 0.99 \\ 
+  fname\_1 & 0.95 & 0.01 \\ 
+  lname\_0 & 0.05 & 0.99 \\ 
+  lname\_1 & 0.95 & 0.01 \\ 
+  day\_0 & 0.05 & 0.97 \\ 
+  day\_1 & 0.95 & 0.03 \\ 
+  month\_0 & 0.05 & 0.92 \\ 
+  month\_1 & 0.95 & 0.08 \\ 
+  year\_0 & 0.05 & 0.93 \\ 
+  year\_1 & 0.95 & 0.07 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{center}\includegraphics[width=23.6in]{../notes/figures/sadinle_speed_plot} \end{center}
+
+The above discussion suggests that for fixed \(n_B\), computation time
+should remain mostly constant with growing \(n_A\). Shockingly, this
+seems to be true. In the plot below, fixing \(n_B = 500\), we see linear
+growth for the runtime under \texttt{BRL} as \(n_A\) increases, with
+much more static runtime under \texttt{parlr}. The slight increases in
+runtime that we do see are due primarily to the hashing step, which
+again can be run in parallel for large data.
+
+\begin{center}\includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB} \end{center}
+
+We note here that \texttt{BRL} is coded in C, which makes for unfair
+comparison against \texttt{parlr}, currently only built in R.
+Additionally, although \texttt{parlr} is amenable to parallezation, this
+simulation was run on a single core. Running \texttt{parlr} in C++ with
+paralellization for the hashing step and sampling the matching status of
+the record pairs should lead to even more drastic results.
 

### 2021-04-25 R/MakeComparisons.R

@@ -1,0 +1,12 @@
+# n <- 50
+# ids <- expand.grid(1:n, 1:n)
+# df1 <- sample(1:5, n, replace = TRUE)
+# df2 <- sample(1:5, n, replace = TRUE)
+# df1[ids[,1]] == df2[ids[,2]]
+
+PossiblePatternsFS <- function(levels){
+  levels %>%
+    purrr::map(seq_len) %>%
+    do.call(expand.grid, .) %>%
+    unite(patterns_FS, seq_along(levels), sep = "")
+}

### 2021-04-25 R/PossiblePatterns.R

@@ -6,6 +6,14 @@ FS_to_Sadinle <- function(gamma, levels){
   }, z = gamma, y = levels)))
 }
 
+FS_to_Sadinle2 <- function(gamma_f, Lf){
+  new_gamma <- matrix(0, nrow = length(gamma_f), ncol = Lf)
+  for(i in seq_along(gamma_f)){
+    new_gamma[i, gamma_f[i]+1] <- 1
+  }
+  new_gamma
+}
+
 GetPossiblePatterns <- function(levels){
   possible_values <- lapply(levels, function(x){
     seq_len(x)

### 2021-04-25 R/SimulateComparisons.R

@@ -36,6 +36,46 @@ SimulateComparisons <- function(m, u, levels, n1, n2, overlap){
        Ztrue = Ztrue)
 }
 
+SimulateComparisonsFast <- function(m, u, levels, n1, n2, overlap){
+  parameter_split <- unlist(lapply(1:length(levels), function(x){
+    rep(x, levels[x])
+  }))
+
+  N <- n1 * n2
+  ids <- expand.grid(1:n1, 1:n2)
+  indicators <- matrix(NA, nrow = N, ncol = length(levels))
+
+  df1matches <- df2matches <- seq_len(overlap)
+  #pairs <- cbind(df1matches, df2matches)
+
+  Ztrue <- rep(n1 + 1, n2)
+  Ztrue[df2matches] <- df1matches
+
+  match_index <- which(ids[,1] == ids[,2])[seq_len(overlap)]
+
+  m.list <- split(m, parameter_split)
+  u.list <- split(u, parameter_split)
+
+  gamma_match <- sapply(m.list, function(x){
+    sample(seq_along(x) - 1, overlap, replace = T, x)
+  })
+
+  indicators[match_index,] <- gamma_match
+
+  gamma_nonmatch <- sapply(u.list, function(x){
+    sample(seq_along(x) - 1, N - overlap, replace = T, x)
+  })
+  indicators[-match_index,] <- gamma_nonmatch
+  Sadinle_indicators <- map2(data.frame(indicators), levels, ~FS_to_Sadinle2(.x, .y)) %>%
+    do.call(cbind, .)
+
+  list(comparisons = Sadinle_indicators,
+       n1 = n1,
+       n2 = n2,
+       nDisagLevs = levels,
+       Ztrue = Ztrue)
+}
+
 SimulateComparisonsLC <- function(m, u, levels, varying_fields,
                                    n1, n2, n1_vec, n2_vec, overlap_vec){
   parameter_split <- unlist(lapply(1:length(levels), function(x){

### 2021-04-28 paper/parlr_paper_1.Rmd
finished big simulation. updated parlr paper

@@
 
+## Introduction (or Abstract?)
+
+Record linkage is the task of identifying duplicate records across multiple data sources.
+This task is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error.
+Additionally, record linkage is inherently computationally intensive, and many existing methods are not feasible for many applications.
+In this paper, we propose a Bayesian model for record linkage that is amenable to parallel computing, hashing techniques and efficient storage mechanisms, that is scalable to large linkage tasks.
+
 ## Related Work
@@
 We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f$ providing their similarity score on the $f^{th}$ field.
-For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeller.
+For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeler.
 It is common to use use binary 0-1 variables to indicate exact matching, and 0-1-2 variables to provide an option for partial matching.
@@
 Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, then samples a potential link.
-However, accounting for these dependencies throughout the linkage process is computionally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.
+However, accounting for these dependencies throughout the linkage process is computationally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.
 
 Outside of the Fellegi-Sunter framework, some researchers have developed methods that model the data directly, not a derived set of comparison vectors.
-As an early example, Steorts et al 2016 produced `blink`, which had the advantage of a computational complexity that grew linearly with the data (as opposed to quadratically within Fellegi Sunter), but was plagued with slow mixing times and was unable to incorportate text data.
-Later, Steorts et al 2020 improved on this method with `dblink`, which was able to simulate text data within its Gibbs sampler by drawing from an empirical distribution, and remedied its slow mixing through a probabalistic blocking approach.
+As an early example, Steorts et al 2016 produced `blink`, which had the advantage of a computational complexity that grew linearly with the data (as opposed to quadratically within Fellegi Sunter), but was plagued with slow mixing times and was unable to incorporate text data.
+Later, Steorts et al 2020 improved on this method with `dblink`, which was able to simulate text data within its Gibbs sampler by drawing from an empirical distribution, and remedied its slow mixing through a probabilistic blocking approach.
 However, this method remains computationally intensive and requires distributed computing to manage even moderate linkage problems in acceptable time.
@@
 
-(This will be revised to talk more explicitly about sampling $Z_j$ independently)
+<!-- (This will be revised to talk more explicitly about sampling $Z_j$ independently) -->
 
-Our model differs from that of Sadinle 2017 through its explicit dependence on a beta random variable $\lambda$ that models the rate of matching across records.
+<!-- Our model differs from that of Sadinle 2017 through its explicit dependence on a beta random variable $\lambda$ that models the rate of matching across records.
 Sadinle marginalizes over such a random variable in his derivations of  the "beta prior for bipartite matching," but here we provide derivations without marginalizing in order to specify differing rates of matching for different linkage clusters.
+-->
 
-## Prior Distributions and Likelihood
+<!-- ## Prior Distributions and Likelihood -->
 
@@
 Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$ prior corresponds to a uniform prior on the labeling of $\mathbf{Z}$.
+
+Here the reader should the relationship between our proposed model and that of Sadinle 2017.
+In his model, Sadinle constructs a prior distribution on the entire $\mathbf{Z}$ vector, which induces a Gibbs sampler that strictly enforces one-to-one matching.
+In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*.
+We however use independent priors for each $Z_j$, thereby weakening the one-to-one requirement; our sampler does ensure that each record in $B$ can be matched to at most one record in $A$, but allows for the possiblity that multiple records in $B$ match to the same record in $A$.
+With this change, components of the $\mathbf{Z}$ vector can be computed in parallel, allowing for significant computational gains.
+More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns, offering even more computational savings.
+This will be explained further in \textcolor{red}{Section XX}.
 
@@
 
-Our Gibbs sampler provides posterior samples of $\mathbf{Z}$ which we use to make our final decisions about the linkage structure.
+<!-- Our Gibbs sampler provides posterior samples of $\mathbf{Z}$ which we use to make our final decisions about the linkage structure.
 In the case where we false matches and missed matches contribute the same loss, we declare $(i,j)$ to be a match whenever $P(Z_j = i) > \frac{1}{2}$ according to these posterior samples.
@@
 For further discussion, see Sadinle 2017.
+-->
 
-Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for the final Bayes estimate to link mulitple records in $B$ to one record in $A$.
-The modeller can either report both such matches (with their respective posterior match probabilities), or resolve these conflicts by accepting only the match with highest posterior probability.
+(Adapt from Sadinle)
+
+Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for the final Bayes estimate to link multiple records in $B$ to one record in $A$.
+The modeler can either report both such matches (with their respective posterior match probabilities), or resolve these conflicts by accepting only the match with highest posterior probability.
 Rigorous proof of this resolution procedure is deferred to future work.
@@
 
-To address this, we propose a new strategy called *random indexing*.
-In standard indexing, the modeller decides a certain criteria that they expect all true matching pairs to satisfy, and a priori label any record pairs that do not meet that criteria as non-matching.
+To address this, we propose a new strategy called *storage efficient indexing* (SEI).
+In standard indexing, the modeler decides a certain criteria that they expect all true matching pairs to satisfy, and a priori label any record pairs that do not meet that criteria as non-matching.
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important (like first name), or only pairs with exact matching on a specified number of fields.
@@
 We propose a method of reducing the comparison space (and reducing storage requirements) without these drawbacks.
-Note that all records of the same agreement pattern have the same probality when sampling $Z_j$.
+Note that all records of the same agreement pattern have the same probability when sampling $Z_j$.
 Therefore we know that records belonging to an $h_p$ such that $H_{p_j}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate, even without considering the form of the agreement pattern itself.
 In random indexing, rather than store all of these unlikely record labels, we choose to store only a small number $R$ of them.
-Posterior calculations still attribute the appropiate weight to all records through the summary statistics $H_{p_j}, H_p^m$, and $H_p^u$; this is similar to the *u-correction* concept introduced by ________.
+Posterior calculations still attribute the appropriate weight to all records through the summary statistics $H_{p_j}, H_p^m$, and $H_p^u$; this is similar to the *u-correction* concept introduced by ________.
 Rather than storing $n_A \times n_B$ record labels, random indexing allows us to store at most $n_B \times P \times R$, regardless of how large $n_A$ is.
@@
 
-After receiving matching statuses from $\mathbf{Z}$, the Sadinle method calculates $\alpha_{fl}(\mathbf{Z})$ and $\beta_{fl}(\mathbf{Z})$ for each field and level.
+\underline{Updating $\mathbf{m}$ and $\mathbf{u}$:} After receiving matching statuses from $\mathbf{Z}$, the Sadinle method calculates $\alpha_{fl}(\mathbf{Z})$ and $\beta_{fl}(\mathbf{Z})$ for each field and level.
 This constitutes $2 \times \sum L_f$ many summations over $n_A \times n_B$ quantities, and becomes computationally burdensome with large data.
@@
 
-Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching.
+<!-- Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching.
 In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*.
@@
 This boosts computation time far greater than parallelization.
+-->
 
-To do this, we calculate the Fellegi Sunter weight $w_{h_p}$ for each unique pattern,  sample the agreement pattern between $j$ and its potential match, and then sample the record label uniformly among viable records.
+\underline{Updating $\mathbf{Z}$:} Although sampling $Z_j$ from a the full conditional provided earlier is conceptually straightforward, it becomes computational burdensome when $n_A$ is larger.
+The reader can confirm that sampling a value from a large set of unequal probabilities becomes difficult in most programming languages.
+To speed up computation, we break this sampling step into two simpler steps.
+First, we calculate the Fellegi Sunter weight $w_{h_p}$ associated with each unique pattern and sample the agreement pattern between $j$ and its potential match.
+Second, we sample the record label uniformly among records associated with that agreement pattern for that particular $j\in B$.
 More concretely, define $h(Z_j)$ to be the agreement pattern between $j$ and its potential match, and say $h(Z_j) = h_{P+1}$ when $Z_j = n_A + 1$.
@@
 
-We complete the entire Gibbs procedure at the level of the $P$ agreement patterns.
+Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
+Thuse We complete the entire Gibbs procedure at the level of the $P$ agreement patterns.
 After, we can back-fill the records corresponding to the agreement patterns by sampling uniformly at random among candidate records stored in $\tilde{\Gamma}$.
+Sampling uniformly is computational simple even for large sets of candidate records, but this step can also be parallelized when working with large data.
 
@@
 Distributions are meant to emulate the behavior of similarity scores across first name, last name, and day, month, and and year of birth, with 1 indicating an exact match on a particular field and 0 indicating a nonmatch.
-We simulate these data for different values of $n_A$ and $n_B$, and compare the runtime of `parlr` against `BRL`.
+We simulate these data for different values of $n_A$ and $n_B$, and compare the run-time of `parlr` against `BRL`.
 We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data.
 
-In particular, runtime for `BRL` seems to grow quadratically while runtime for `parlr` seems to grow linearly.
+In particular, run-time for `BRL` seems to grow quadratically while run-time for `parlr` seems to grow linearly.
 
@@
 Shockingly, this seems to be true.
-In the plot below, fixing $n_B = 500$, we see linear growth for the runtime under `BRL` as $n_A$ increases, with much more static runtime under `parlr`.
-The slight increases in runtime that we do see are due primarily to the hashing step, which again can be run in parallel for large data.
+In the plot below, fixing $n_B = 500$, we see linear growth for the run-time under `BRL` as $n_A$ increases, with much more static run-time under `parlr`.
+The slight increases in run-time that we do see are due primarily to the hashing step, which again can be run in parallel for large data.
 
@@
 We note here that `BRL` is coded in C, which makes for unfair comparison against `parlr`, currently only built in R.
-Additionally, although `parlr` is amenable to parallezation, this simulation was run on a single core.
+Additionally, although `parlr` is amenable to parallelization, this simulation was run on a single core.
 Running `parlr` in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more drastic results.
@@
 
-- Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?
+- Thinking of re-running the El Salvador linkage from the Sadinle 2017 paper, but doing so with without the formal blocking and indexing that he uses.
+Instead, I'd use hashing and storage efficient indexing -- I should be able to do linkage on the all the data at once!
 
@@
 
-We note that this often not a reasonable assumption; for example if we know that two records differ on gender, this does not give any information on whether their year of births are the same, but it does suggest that its more likely that their first names are different.
+<!-- We note that this often not a reasonable assumption; for example if we know that two records differ on gender, this does not give any information on whether their year of births are the same, but it does suggest that its more likely that their first names are different.
+-->

### 2021-04-28 paper/parlr_paper_1.tex
finished big simulation. updated parlr paper

@@
 
+\hypertarget{introduction-or-abstract}{%
+\subsection{Introduction (or
+Abstract?)}\label{introduction-or-abstract}}
+
+Record linkage is the task of identifying duplicate records across
+multiple data sources. This task is straightforward when provided unique
+identifiers or highly reliable information, but becomes difficult when
+data sources are plagued with error. Additionally, record linkage is
+inherently computationally intensive, and many existing methods are not
+feasible for many applications. In this paper, we propose a Bayesian
+model for record linkage that is amenable to parallel computing, hashing
+techniques and efficient storage mechanisms, that is scalable to large
+linkage tasks.
+
 \hypertarget{related-work}{%
@@
 these similarity scores to be discrete, ordinal variables, and the
-construction of these is left to the modeller. It is common to use use
+construction of these is left to the modeler. It is common to use use
 binary 0-1 variables to indicate exact matching, and 0-1-2 variables to
@@
 then samples a potential link. However, accounting for these
-dependencies throughout the linkage process is computionally burdensome,
-leaving \texttt{BRL} only suitable for small to moderate linkage
-problems.
+dependencies throughout the linkage process is computationally
+burdensome, leaving \texttt{BRL} only suitable for small to moderate
+linkage problems.
 
@@
 Fellegi Sunter), but was plagued with slow mixing times and was unable
-to incorportate text data. Later, Steorts et al 2020 improved on this
+to incorporate text data. Later, Steorts et al 2020 improved on this
 method with \texttt{dblink}, which was able to simulate text data within
 its Gibbs sampler by drawing from an empirical distribution, and
-remedied its slow mixing through a probabalistic blocking approach.
+remedied its slow mixing through a probabilistic blocking approach.
 However, this method remains computationally intensive and requires
@@
 
-(This will be revised to talk more explicitly about sampling \(Z_j\)
-independently)
-
-Our model differs from that of Sadinle 2017 through its explicit
-dependence on a beta random variable \(\lambda\) that models the rate of
-matching across records. Sadinle marginalizes over such a random
-variable in his derivations of the ``beta prior for bipartite
-matching,'' but here we provide derivations without marginalizing in
-order to specify differing rates of matching for different linkage
-clusters.
-
-\hypertarget{prior-distributions-and-likelihood}{%
-\subsection{Prior Distributions and
-Likelihood}\label{prior-distributions-and-likelihood}}
-
 For fields \(f \in \{1, \ldots, F\}\) and levels
@@
 uniform prior on the labeling of \(\mathbf{Z}\).
+
+Here the reader should the relationship between our proposed model and
+that of Sadinle 2017. In his model, Sadinle constructs a prior
+distribution on the entire \(\mathbf{Z}\) vector, which induces a Gibbs
+sampler that strictly enforces one-to-one matching. In particular, this
+sampler removes previously matches records from the set of candidate
+records when sampling \(Z_j\), creating a dependency that makes the
+sampler \emph{inherently serial}. We however use independent priors for
+each \(Z_j\), thereby weakening the one-to-one requirement; our sampler
+does ensure that each record in \(B\) can be matched to at most one
+record in \(A\), but allows for the possiblity that multiple records in
+\(B\) match to the same record in \(A\). With this change, components of
+the \(\mathbf{Z}\) vector can be computed in parallel, allowing for
+significant computational gains. More importantly, since only the
+agreement pattern of \(Z_j\) is used for calculations within the Gibbs
+sampler, and not the particular record label, we can conduct this
+sampling only at the level of the unique agreement patterns, offering
+even more computational savings. This will be explained further in
+\textcolor{red}{Section XX}.
 
@@
 
-Our Gibbs sampler provides posterior samples of \(\mathbf{Z}\) which we
-use to make our final decisions about the linkage structure. In the case
-where we false matches and missed matches contribute the same loss, we
-declare \((i,j)\) to be a match whenever \(P(Z_j = i) > \frac{1}{2}\)
-according to these posterior samples. We can create more elaborate
-decisions by attributing different loss values to different kinds of
-error, and allowing pairings with middling posterior probabilities to be
-left unlabeled by the algorithm so that the modeler can address those
-manually. For further discussion, see Sadinle 2017.
+(Adapt from Sadinle)
 
 Since our Gibbs procedure does not strictly enforce one-to-one matching,
-it is possible for the final Bayes estimate to link mulitple records in
-\(B\) to one record in \(A\). The modeller can either report both such
+it is possible for the final Bayes estimate to link multiple records in
+\(B\) to one record in \(A\). The modeler can either report both such
 matches (with their respective posterior match probabilities), or
@@
 
-To address this, we propose a new strategy called \emph{random
-indexing}. In standard indexing, the modeller decides a certain criteria
-that they expect all true matching pairs to satisfy, and a priori label
-any record pairs that do not meet that criteria as non-matching. For
-example, one might only consider pairs with a certain similarity score
-on a field deemed to be important (like first name), or only pairs with
-exact matching on a specified number of fields. While generally chosen
-to be be quite loose, establishing these criteria requires knowledge of
-the problem and invites room for human error. We propose a method of
-reducing the comparison space (and reducing storage requirements)
-without these drawbacks. Note that all records of the same agreement
-pattern have the same probality when sampling \(Z_j\). Therefore we know
-that records belonging to an \(h_p\) such that \(H_{p_j}\) is large are
-very unlikely to be sampled consistently enough to be deemed a match
-through the Bayes estimate, even without considering the form of the
-agreement pattern itself. In random indexing, rather than store all of
-these unlikely record labels, we choose to store only a small number
-\(R\) of them. Posterior calculations still attribute the appropiate
-weight to all records through the summary statistics \(H_{p_j}, H_p^m\),
-and \(H_p^u\); this is similar to the \emph{u-correction} concept
-introduced by \_\_\_\_\_\_\_\_. Rather than storing \(n_A \times n_B\)
-record labels, random indexing allows us to store at most
-\(n_B \times P \times R\), regardless of how large \(n_A\) is. Coupled
-with the partitioning strategy for creating comparison vectors, this
-greatly expands the size of the linkage tasks we can undertake.
+To address this, we propose a new strategy called \emph{storage
+efficient indexing} (SEI). In standard indexing, the modeler decides a
+certain criteria that they expect all true matching pairs to satisfy,
+and a priori label any record pairs that do not meet that criteria as
+non-matching. For example, one might only consider pairs with a certain
+similarity score on a field deemed to be important (like first name), or
+only pairs with exact matching on a specified number of fields. While
+generally chosen to be be quite loose, establishing these criteria
+requires knowledge of the problem and invites room for human error. We
+propose a method of reducing the comparison space (and reducing storage
+requirements) without these drawbacks. Note that all records of the same
+agreement pattern have the same probability when sampling \(Z_j\).
+Therefore we know that records belonging to an \(h_p\) such that
+\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
+to be deemed a match through the Bayes estimate, even without
+considering the form of the agreement pattern itself. In random
+indexing, rather than store all of these unlikely record labels, we
+choose to store only a small number \(R\) of them. Posterior
+calculations still attribute the appropriate weight to all records
+through the summary statistics \(H_{p_j}, H_p^m\), and \(H_p^u\); this
+is similar to the \emph{u-correction} concept introduced by
+\_\_\_\_\_\_\_\_. Rather than storing \(n_A \times n_B\) record labels,
+random indexing allows us to store at most \(n_B \times P \times R\),
+regardless of how large \(n_A\) is. Coupled with the partitioning
+strategy for creating comparison vectors, this greatly expands the size
+of the linkage tasks we can undertake.
 
@@
 
-After receiving matching statuses from \(\mathbf{Z}\), the Sadinle
-method calculates \(\alpha_{fl}(\mathbf{Z})\) and
-\(\beta_{fl}(\mathbf{Z})\) for each field and level. This constitutes
-\(2 \times \sum L_f\) many summations over \(n_A \times n_B\)
-quantities, and becomes computationally burdensome with large data. In
-contrast, we recognize that each unique agreement pattern contributes to
-the posterior \(\alpha(\mathbf{Z})\) and \(\beta(\mathbf{Z})\) vectors
-in the same way. In fact, if we denote
+\underline{Updating $\mathbf{m}$ and $\mathbf{u}$:} After receiving
+matching statuses from \(\mathbf{Z}\), the Sadinle method calculates
+\(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each
+field and level. This constitutes \(2 \times \sum L_f\) many summations
+over \(n_A \times n_B\) quantities, and becomes computationally
+burdensome with large data. In contrast, we recognize that each unique
+agreement pattern contributes to the posterior \(\alpha(\mathbf{Z})\)
+and \(\beta(\mathbf{Z})\) vectors in the same way. In fact, if we denote
 \(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
@@
 
-Sadinle uses a prior for \(\mathbf{Z}\) that induces the a full
-conditional for \(Z_j\) that strictly enforces one-to-one matching. In
-particular, this sampler removes previously matches records from the set
-of candidate records when sampling \(Z_j\), creating a dependency that
-makes the sampler \emph{inherently serial}. By weakening the one-to-one
-requirement, our full conditional for \(Z\) does not depend on the rest
-of the \(\mathbf{Z}_{-j}\) vector, and thus can be computed in parallel.
-More importantly, since only the agreement pattern of \(Z_j\) is used
-for calculations within the Gibbs sampler, and not the particular record
-label, we can conduct this sampling only at the level of the unique
-agreement patterns. This boosts computation time far greater than
-parallelization.
-
-To do this, we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
-unique pattern, sample the agreement pattern between \(j\) and its
-potential match, and then sample the record label uniformly among viable
-records. More concretely, define \(h(Z_j)\) to be the agreement pattern
-between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when
-\(Z_j = n_A + 1\). Then,
+\underline{Updating $\mathbf{Z}$:} Although sampling \(Z_j\) from a the
+full conditional provided earlier is conceptually straightforward, it
+becomes computational burdensome when \(n_A\) is larger. The reader can
+confirm that sampling a value from a large set of unequal probabilities
+becomes difficult in most programming languages. To speed up
+computation, we break this sampling step into two simpler steps. First,
+we calculate the Fellegi Sunter weight \(w_{h_p}\) associated with each
+unique pattern and sample the agreement pattern between \(j\) and its
+potential match. Second, we sample the record label uniformly among
+records associated with that agreement pattern for that particular
+\(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement
+pattern between \(j\) and its potential match, and say
+\(h(Z_j) = h_{P+1}\) when \(Z_j = n_A + 1\). Then,
 
@@
 
-We complete the entire Gibbs procedure at the level of the \(P\)
-agreement patterns. After, we can back-fill the records corresponding to
-the agreement patterns by sampling uniformly at random among candidate
-records stored in \(\tilde{\Gamma}\).
+Lastly, we recognize that all posterior updates are governed by the
+agreement patterns of the record pairs rather than the record labels
+themselves. Thuse We complete the entire Gibbs procedure at the level of
+the \(P\) agreement patterns. After, we can back-fill the records
+corresponding to the agreement patterns by sampling uniformly at random
+among candidate records stored in \(\tilde{\Gamma}\). Sampling uniformly
+is computational simple even for large sets of candidate records, but
+this step can also be parallelized when working with large data.
 
@@
 indicating a nonmatch. We simulate these data for different values of
-\(n_A\) and \(n_B\), and compare the runtime of \texttt{parlr} against
+\(n_A\) and \(n_B\), and compare the run-time of \texttt{parlr} against
 \texttt{BRL}. We see that at low data size, \texttt{BRL} outperforms,
@@
 
-In particular, runtime for \texttt{BRL} seems to grow quadratically
-while runtime for \texttt{parlr} seems to grow linearly.
+In particular, run-time for \texttt{BRL} seems to grow quadratically
+while run-time for \texttt{parlr} seems to grow linearly.
 
@@
 seems to be true. In the plot below, fixing \(n_B = 500\), we see linear
-growth for the runtime under \texttt{BRL} as \(n_A\) increases, with
-much more static runtime under \texttt{parlr}. The slight increases in
-runtime that we do see are due primarily to the hashing step, which
+growth for the run-time under \texttt{BRL} as \(n_A\) increases, with
+much more static run-time under \texttt{parlr}. The slight increases in
+run-time that we do see are due primarily to the hashing step, which
 again can be run in parallel for large data.
@@
 comparison against \texttt{parlr}, currently only built in R.
-Additionally, although \texttt{parlr} is amenable to parallezation, this
-simulation was run on a single core. Running \texttt{parlr} in C++ with
-paralellization for the hashing step and sampling the matching status of
-the record pairs should lead to even more drastic results.
+Additionally, although \texttt{parlr} is amenable to parallelization,
+this simulation was run on a single core. Running \texttt{parlr} in C++
+with paralellization for the hashing step and sampling the matching
+status of the record pairs should lead to even more drastic results.
 
@@
 \item
-  Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?
+  Thinking of re-running the El Salvador linkage from the Sadinle 2017
+  paper, but doing so with without the formal blocking and indexing that
+  he uses. Instead, I'd use hashing and storage efficient indexing -- I
+  should be able to do linkage on the all the data at once!
 \end{itemize}
@@
 
-We note that this often not a reasonable assumption; for example if we
-know that two records differ on gender, this does not give any
-information on whether their year of births are the same, but it does
-suggest that its more likely that their first names are different.
-
 \end{document}

### 2021-04-28 R/GetUniquePatterns2.R

@@ -9,7 +9,7 @@ GetUniquePatterns2 <- function(cd, fast = F, R = NULL){
   ids <- expand.grid(1:n1, 1:n2)
   rec1 <- ids[,1]
   rec2 <- ids[,2]
-  # levels <- cd[[4]]
+  levels <- cd[[4]]
   # fields <- length(levels)
   # gk <- unlist(lapply(1:fields, function(x){
   #   0:(levels[x] -1)
@@ -90,7 +90,7 @@ pattern_df <- df %>%
 
   temp <- data.frame(indicators, rec1, rec2, hash_id)
   pattern_counts <- temp %>%
-    group_by(hash_id) %>%
+    group_by(hash_id, .drop = F) %>%
     count() %>%
     pull()
 
@@ -125,3 +125,119 @@ pattern_df <- df %>%
                    hash_to_rec1)
 
 }
+
+
+ProcessHash <- function(cd, fast = F, R = NULL){
+
+  hash_id <- cd[[1]]
+  N <- length(hash_id)
+  #fields <- dim(indicators)[2]
+  n1 <- cd[[2]]
+  n2 <- cd[[3]]
+
+  ids <- expand.grid(1:n1, 1:n2)
+  rec1 <- ids[,1]
+  rec2 <- ids[,2]
+  levels <- cd[[4]]
+
+
+  unique_patterns <- GetPossiblePatternsSad_sep(levels)
+  P <- dim(unique_patterns)[1]
+
+  hash_id <- factor(hash_id, levels = 1:P)
+
+
+  temp <- data.frame(rec1, rec2, hash_id)
+  pattern_counts <- temp %>%
+    group_by(hash_id, .drop = F) %>%
+    count() %>%
+    pull()
+
+
+  counts_by_rec <- temp %>%
+    group_by(rec2, hash_id, .drop = F) %>%
+    count()
+
+  counts_by_rec <- split(counts_by_rec$n, counts_by_rec$rec2)
+
+  thing <- split(temp, rec2) #change to groupby
+  hash_to_rec1 <- map(thing, ~ lapply(1:P, function(y){
+    which(.x$hash_id == y)
+  })
+  )
+  hash_to_rec1 <- lapply(hash_to_rec1, function(x){
+    append(x, list(n1 +1))
+  })
+
+  if(!is.null(R)){
+    hash_to_rec1 <- lapply(hash_to_rec1, function(z){
[…]

### 2021-04-28 R/MakeComparisons.R

@@ -4,9 +4,4 @@
 # df2 <- sample(1:5, n, replace = TRUE)
 # df1[ids[,1]] == df2[ids[,2]]
 
-PossiblePatternsFS <- function(levels){
-  levels %>%
-    purrr::map(seq_len) %>%
-    do.call(expand.grid, .) %>%
-    unite(patterns_FS, seq_along(levels), sep = "")
-}
+

### 2021-04-28 R/SimulateBig.R

@@ -1,0 +1,96 @@
+SimulateBig <- function(m, u, levels, n1, n2_list, overlap){
+  parameter_split <- unlist(lapply(1:length(levels), function(x){
+    rep(x, levels[x])
+  }))
+  overlap_per_chunk <- overlap/length(n2_list)
+  possible_patterns <- data.frame(GetPossiblePatternsFS_sep(levels))
+
+
+  patterns_big <- lapply(n2_list, function(z){
+
+  N <- n1 * length(z)
+  ids <- expand.grid(1:n1, z)
+  hash_id <- integer(N)
+
+  df1matches <- df2matches <- z[1:(length(z)/2)]
+  #pairs <- cbind(df1matches, df2matches)
+
+  Ztrue <- rep(n1 + 1, length(z))
+  Ztrue[1:(length(z)/2)] <- df1matches
+
+  match_index <- which(ids[,1] == ids[,2])[seq_len(overlap_per_chunk)]
+
+  m.list <- split(m, parameter_split)
+  u.list <- split(u, parameter_split)
+
+
+  P <- dim(possible_patterns)[1]
+
+  m_probs <- map2(possible_patterns, m.list, ~{
+    .y[.x]
+  }) %>%
+    do.call(cbind, .) %>%
+    apply(1, prod)
+
+  u_probs <- map2(possible_patterns, u.list, ~{
+    .y[.x]
+  }) %>%
+    do.call(cbind, .) %>%
+    apply(1, prod)
+
+  gamma_match <- sample(1:P, overlap_per_chunk, replace = T, m_probs)
+  gamma_nonmatch <- sample(1:P, N - overlap_per_chunk, replace = T, u_probs)
+
+  hash_id[match_index] <- gamma_match
+  hash_id[-match_index] <- gamma_nonmatch
+
+  # gamma_nonmatch <- sapply(u.list, function(x){
+  #   sample(seq_along(x) - 1, N - overlap, replace = T, x)
+  # })
+  # indicators[-match_index,] <- gamma_nonmatch
+  # Sadinle_indicators <- map2(data.frame(indicators), levels, ~FS_to_Sadinle2(.x, .y)) %>%
+  #   do.call(cbind, .)
+
+
+  cd <- list(comparisons = hash_id,
+       n1 = n1,
+       n2_index = z,
+       nDisagLevs = levels,
+       Ztrue = Ztrue)
+
+  patterns_chunk <- ProcessHashBig(cd, fast = F, R = 10, Ztrue)
[…]

### 2021-05-02 R/NLC_hash2.R
first push of el salvador data

@@ -42,8 +42,8 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   candidates <- 1:n1
   candidates_P <- 1:(P+1)
   Z.SAMPS <- matrix(NA, nrow = n2, ncol = S)
-  M.SAMPS <- matrix(NA, nrow = dim(indicators_raw)[2], ncol = S)
-  U.SAMPS <- matrix(NA, nrow = dim(indicators_raw)[2], ncol = S)
+  M.SAMPS <- matrix(NA, nrow = length(parameter_split), ncol = S)
+  U.SAMPS <- matrix(NA, nrow = length(parameter_split), ncol = S)
   L.SAMPS <- vector(length = S)
   Z.temp <- rep(0, n1*n2)
   Z <- rep(n1+1, n2)
@@ -161,8 +161,8 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
     # hash$Z.temp <- Z.temp
 
     Z.SAMPS[,s] <- Z
-    #M.SAMPS[,s] <- m
-    #U.SAMPS[,s] <- u
+    M.SAMPS[,s] <- m
+    U.SAMPS[,s] <- u
     L.SAMPS[s] <- L
 
     if(show_progress){
@@ -176,15 +176,17 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
     parallel::stopCluster(cl)
   }
   Z.SAMPS <- Z.SAMPS[,-(1:burn)]
-  #L.SAMPS <- L.SAMPS[-(1:burn)]
-  #M.SAMPS <- M.SAMPS[,-(1:burn)]
+  L.SAMPS <- L.SAMPS[-(1:burn)]
+  M.SAMPS <- M.SAMPS[,-(1:burn)]
   #U.SAMPS <- U.SAMPS[,-(1:burn)]
 
   final_gibbs <- apply(Z.SAMPS, 2, function(z){
     unlist(imap(z, ~sample_with_1(hash_to_rec1[[.y]][[.x]], 1)))
   })
 
-  final_gibbs
+  list(Z = final_gibbs,
+       m = M.SAMPS,
+       u = U.SAMPS)
 
 }
 

### 2021-05-03 notes/el_salvador_analysis_1.Rmd
Notes for El Salvador analysis

---
title: "Notes 3-18-19"
subtitle: "New Distributions for BRACS Paper"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.height = "80%", out.width = "80%",
                      fig.align = 'center')

library(knitr)
```

# El Salvador Data: Parlr vs Sadinle17

I've attempted to reproduce the results from Sadinle's 2017 paper, which has been a little bit of a mixed bag. The code that Beka gave me does not seem to be the same that is used in the actual 2017 paper. I think this for two reasons:

- Code involves three datasets (ER, CHDES, and UNTC), while paper only links two (ER and CHDES)
- Code creates comparison matrix that is not compatible with BRL

Nevertheless, I proceeded. For a first run-through, I tried to build the comparison vectors in the exact same way that Sadinle did. He uses 4 levels for many fields, which I'll discuss more later. Also, his paper calls for three agreement levels for location of death (exact municipality, adjacent municipality, nonmatch), but the code I have doesn't do that, and that would be pretty difficult to implement. So I just use binary comparisons.

My results are close, but not identical to what's provided in the Sadinle paper.

- After data cleaning, my ER dataset has 1323 records, while the paper says there should be 1324. 
- The paper has a Bayes estimate with 187 links, but my implementation (with `BRL`) finds 178. This might be because the paper detects adjecent municipalities, while I do not
- Distribution of $m$ parameters for given name substantially different from whats in the paper.

I can go more into details about these things, but for the moment, I'm more concerned with comparing the results of `BRL` and `parlr` in my own implementation, rather than reproducing Sadinle17 exactly.

## Bad News: More Patterns More Problems

The table below shows the amount of overlap detected by each method, as well as the time it took for the linkage to run. You'll notice that `parlr` is *significantly* slower than `BRL` here.

```{r}
include_graphics("../notes/figures/el_salvador/time_table_big_P.png")
```

However, this is largely because Sadinle uses 4 level of agreement for most of the fields, leading to $P = 4096$. Given that $n_A = 4420$, `parlr` will give almost no computational savings in this context. Still however, lets examine the results

Below is the posterior distribution of the overlap between the files. We see that `parlr` typically matches me, which is to be expected; at minimum, there are the 13 matches that `parlr` made that violated one-to-one matching which Sadinle avoids. For two reasons, I do not consider this to be a problem.

```{r}
include_graphics("../notes/figures/el_salvador/overlap_distribution_bigP.png")
```

- Within the Sadinle framework, it is typical that Gibbs iterations have more matches than the eventualy Bayes estimate. This is because many of the matches are due to randomness, and they only make it through to the Bayes estimate if that same matching occurs consistently 50% of the time. In Sadinle's paper, the overlap has a mean of 241, and a 90% credible interval of [227, 256], but a Bayes estimate of only 187.

- Severe overmatching within the Gibbs iterations can possibly lead to biased postiors for the $m$ parameters. There is an interesting conversation to be had here. When Enamorado compared `fastlink` to `BRL`, he notes that the first estimate under `fastlink` has way too many matches, but that `fastlink` produces similar results to `BRL` after resolving all the matches that violate one-to-one. While this is true, the overmatching within his EM algorithm lead to biases in the $m$ parameters, and these can not be cleaned up posthoc. In the Sadinle paper, he notes that the FS method leads to nonsensicle estimates of the $m$ parameters, and I believe it is for this reason.

This led to me worry that parameter estimates under `parlr` would be similarly biased. However, good news, they are not! This is because the overmatching under `parlr` is very slight compared to `fastlink`

```{r}
include_graphics("../notes/figures/el_salvador/m_posterior_bigP.png")
```

This comparison of $m$ parameters also makes me more convinced that Sadinle is using way too many levels per field. So now we're going to run `parlr` with fewer levels and see how things change.

## Good News: Less Patterns Less Time

## Questions

Is still a good case study to choose? Consider that:

- I haven't been able to exactly reproduce the results from the Sadinle paper
- This doesn't show off the computational gains of `parlr`

I think its nice that it demonstrates the influence of the $P$ patterns, but I'm not sure if thats enough. Let me know your thoughts on this!

### 2021-05-03 notes/el_salvador_analysis_1.Rmd
finished small P analysis

@@
 - After data cleaning, my ER dataset has 1323 records, while the paper says there should be 1324. 
-- The paper has a Bayes estimate with 187 links, but my implementation (with `BRL`) finds 178. This might be because the paper detects adjecent municipalities, while I do not
+- The paper has a Bayes estimate with 187 links, but my implementation (with `BRL`) finds 178. This might be because the paper detects adjacent municipalities, while I do not
 - Distribution of $m$ parameters for given name substantially different from whats in the paper.
@@
 
-- Within the Sadinle framework, it is typical that Gibbs iterations have more matches than the eventualy Bayes estimate.
+- Within the Sadinle framework, it is typical that Gibbs iterations have more matches than the eventual Bayes estimate.
 This is because many of the matches are due to randomness, and they only make it through to the Bayes estimate if that same matching occurs consistently 50% of the time.
@@
 
-- Severe overmatching within the Gibbs iterations can possibly lead to biased postiors for the $m$ parameters.
+- Severe overmatching within the Gibbs iterations can possibly lead to biased posteriors for the $m$ parameters.
 There is an interesting conversation to be had here.
@@
 While this is true, the overmatching within his EM algorithm lead to biases in the $m$ parameters, and these can not be cleaned up posthoc.
-In the Sadinle paper, he notes that the FS method leads to nonsensicle estimates of the $m$ parameters, and I believe it is for this reason.
+In the Sadinle paper, he notes that the FS method leads to nonsensical estimates of the $m$ parameters, and I believe it is for this reason.
 
@@
 
-## Good News: Less Patterns Less Time
+## Good News: Less Patterns Less Problems
+
+I reran everything with only 3 levels for first name, last name, and day, and only 2 levels for everything else. This yields $P = 432$.
+
+```{r}
+include_graphics("../notes/figures/el_salvador/time_table_small_P.png")
+```
+
+```{r}
+include_graphics("../notes/figures/el_salvador/overlap_distribution_smallP.png")
+```
+
+```{r}
+include_graphics("../notes/figures/el_salvador/m_posterior_smallP.png")
+```
+
+## Limitation of parlr
+
+Doing this analysis made me realize a limitation of this `parlr` method.
+When you have many many fields and levels, you can have situations where $P > n_A$, which means that doing calculations over all possible patterns will actually be slower than the standard method.
+Of course, there will only be at most $n_A$ realized patterns (and probably less than that ), but computational saving will only be significant when $P << n_A$.
+
+One other issue regarding the size of $P$: Numerically, missing values lead to another unique pattern.
+Consider the previous examples with 5 fields with only 2 agreement levels, yeilding $P = 2^5 =32$ unique patterns.
+If there is at even one missing value for each of the fields, this would lead to $P = 3^5 =243$ unique patterns.
+For large $n_A$ this is still a big savings, but its important to note how quickly $P$ can grow, especially with regards to missing values.
+
+(Sadinle's `BRL` incorrectly classifies all missing values as disagreements. Adopting this would help keep $P$ small, but is bad for tons of different reasons.)
 
@@
 
-I think its nice that it demonstrates the influence of the $P$ patterns, but I'm not sure if thats enough. Let me know your thoughts on this!
+I think its nice that it demonstrates the influence of the $P$ patterns, but I'm not sure if that's enough. Let me know your thoughts on this!

### 2021-05-05 notes/el_salvador_analysis_1.Rmd
big and small p
bad mixing plot

@@
 ---
-title: "Notes 3-18-19"
-subtitle: "New Distributions for BRACS Paper"
+title: "El Salvador Linkage Analysis"
+
 output:
@@
 - Code involves three datasets (ER, CHDES, and UNTC), while paper only links two (ER and CHDES)
-- Code creates comparison matrix that is not compatible with BRL
+- There is code for priors for "truncation points" which are used in Sadinle's 2014 model.
 
@@
 - After data cleaning, my ER dataset has 1323 records, while the paper says there should be 1324. 
-- The paper has a Bayes estimate with 187 links, but my implementation (with `BRL`) finds 178. This might be because the paper detects adjacent municipalities, while I do not
-- Distribution of $m$ parameters for given name substantially different from whats in the paper.
+- The paper has a Bayes estimate with 187 links, but my implementation (with `BRL`) finds 180. This might be because the paper detects adjacent municipalities, while I do not.
+- Distribution of $m$ parameters for given name is a bit different from whats in the paper.
 
@@
 
-## Bad News: More Patterns More Problems
+## Using Sadinle's Levels
 
-The table below shows the amount of overlap detected by each method, as well as the time it took for the linkage to run. You'll notice that `parlr` is *significantly* slower than `BRL` here.
+The table below shows the amount of overlap detected by each method, as well as the time it took for the linkage to run.
+Interestingly, since Sadinle uses so many levels per field, we have $4^5 \times 2^2 = 4096$ possible unique patterns, so I was worried that `parlr` wouldn't give the desired speedups.
+However, only 352 patterns are actually realized, so `parlr` goes pretty fast!
 
@@
 ```
-
-However, this is largely because Sadinle uses 4 level of agreement for most of the fields, leading to $P = 4096$.
-Given that $n_A = 4420$, `parlr` will give almost no computational savings in this context.
-Still however, lets examine the results
 
@@
 However, good news, they are not!
-This is because the overmatching under `parlr` is very slight compared to `fastlink`
+This is because the overmatching under `parlr` is very slight compared to `fastlink`.
 
@@
 
-This comparison of $m$ parameters also makes me more convinced that Sadinle is using way too many levels per field. So now we're going to run `parlr` with fewer levels and see how things change.
+Looking at these $m$ parameters, we see that many levels are redundant.
+Also, since less that 10% of the possible patterns are realized, it really does seems like that many levels are unneccesary.
+So now we're going to run `parlr` with fewer levels and see how things change.
 
-## Good News: Less Patterns Less Problems
+## Using Less Patterns
 
-I reran everything with only 3 levels for first name, last name, and day, and only 2 levels for everything else. This yields $P = 432$.
+I reran everything with only 3 levels for first name, last name, and day, and only 2 levels for everything else.
+I probably could use just two levels for the names since the code puts so much work into standardizing names and using a modified Levenstein distance, but for now, I'll stick with 3 levels.
+Here $P$ is bounded by 216, and only 159 unique patterns are realized.
+Results stay basically the same, but see significant speed up for `parlr`.
 
@@
 
-## Limitation of parlr
+<!-- ## Limitations of parlr -->
 
-Doing this analysis made me realize a limitation of this `parlr` method.
+<!-- Doing this analysis made me realize a limitation of this `parlr` method.
 When you have many many fields and levels, you can have situations where $P > n_A$, which means that doing calculations over all possible patterns will actually be slower than the standard method.
 Of course, there will only be at most $n_A$ realized patterns (and probably less than that ), but computational saving will only be significant when $P << n_A$.
+-->
 
-One other issue regarding the size of $P$: Numerically, missing values lead to another unique pattern.
+<!-- One other issue regarding the size of $P$: Numerically, missing values lead to another unique pattern.
 Consider the previous examples with 5 fields with only 2 agreement levels, yeilding $P = 2^5 =32$ unique patterns.
@@
 For large $n_A$ this is still a big savings, but its important to note how quickly $P$ can grow, especially with regards to missing values.
+-->
 
-(Sadinle's `BRL` incorrectly classifies all missing values as disagreements. Adopting this would help keep $P$ small, but is bad for tons of different reasons.)
+<!-- (Sadinle's `BRL` incorrectly classifies all missing values as disagreements. Adopting this would help keep $P$ small, but is bad for tons of different reasons.) -->
+
+## Mixing and Multiple Possible Matches
+
+Something interesting about this comparison is that this dataset is specifically an example where `parlr` and Sadinle would differ the most.
+In cases were there multiple plausible matches, each method produces undesirable behavior.
+For example, examine the records of Rosa and Rosa Maria in the CDHES dataset.
+
+```{r}
+include_graphics("../notes/figures/el_salvador/rosa_records.png")
+```
+
+- In `parlr` both records in $B$ match to the same record in $A$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
+
+- In Sadinle, the Gibbs process creates one matching status and stays there for a while, but if one pair "unmatches," then the other record has a chance to latch on.
+Then the Gibbs process is stuck with that matching status for a while.
+This seems to be poor mixing.
+
+```{r}
+include_graphics("../notes/figures/el_salvador/bad_mixing.png")
+```
+
+As a result, `parlr` provides an initial Bayes Estimate (without conflict resolution) with two matches, each with high posterior probability (90% +) The modeller can algorithmically take the one with highest posterior, or examine the records manually.
+Sadinle provides only one match (consistent with 1-1 matching), with low posterior probability (a bit above 50%).
+
+I used to think this behavior made the El Salvador data a bad option for my case study, but I think its actually instructive to examine this closely. Let me know your thoughts on this!
+
+(PS, it took me forever to figure out how to represent this visually!)
 
@@
 
-Is still a good case study to choose? Consider that:
+Is still a good case study to choose? The first thing that worries is that I have't been able to 100% replicate Sadinle's results. Let me know if you think this is an issue.
 
-- I haven't been able to exactly reproduce the results from the Sadinle paper
-- This doesn't show off the computational gains of `parlr`
-
-I think its nice that it demonstrates the influence of the $P$ patterns, but I'm not sure if that's enough. Let me know your thoughts on this!
+The second thing that worries me is this dataset is precisely one where Sadinle and `parlr` would have the largest discrepancy; the material discussed in the section above.

### 2021-05-05 R/ResolveConflicts.R

@@ -12,3 +12,4 @@ ResolveConflicts <- function(Zhat){
   Zhat[to_resolve] <- max(Zhat)
   Zhat
 }
+

### 2021-05-05 notes/el_salvador_analysis_1.Rmd
done with first sadinle analyis

@@
 I've attempted to reproduce the results from Sadinle's 2017 paper, which has been a little bit of a mixed bag.
-The code that Beka gave me does not seem to be the same that is used in the actual 2017 paper.
+The code we have does not seem to be the same that is used in the actual 2017 paper.
 I think this for two reasons:
@@
 - Code involves three datasets (ER, CHDES, and UNTC), while paper only links two (ER and CHDES)
-- There is code for priors for "truncation points" which are used in Sadinle's 2014 model.
+- There is code for priors for "truncation points" which are elements of Sadinle's 2014 model.
 
@@
 For a first run-through, I tried to build the comparison vectors in the exact same way that Sadinle did.
-He uses 4 levels for many fields, which I'll discuss more later.
-Also, his paper calls for three agreement levels for location of death (exact municipality, adjacent municipality, nonmatch), but the code I have doesn't do that, and that would be pretty difficult to implement.
+One difference in implementation is that his paper calls for three agreement levels for location of death (exact municipality, adjacent municipality, nonmatch), but the code I have doesn't do that, and that would be pretty difficult to implement.
 So I just use binary comparisons.
@@
 - After data cleaning, my ER dataset has 1323 records, while the paper says there should be 1324. 
-- The paper has a Bayes estimate with 187 links, but my implementation (with `BRL`) finds 180. This might be because the paper detects adjacent municipalities, while I do not.
+- The paper has a Bayes estimate with 187 links, but my implementation (with `BRL`) finds 179. This might be because the paper detects adjacent municipalities, while I do not.
 - Distribution of $m$ parameters for given name is a bit different from whats in the paper.
@@
 Below is the posterior distribution of the overlap between the files.
-We see that `parlr` typically matches me, which is to be expected; at minimum, there are the 13 matches that `parlr` made that violated one-to-one matching which Sadinle avoids.
+We see that `parlr` typically matches more, which is to be expected; at minimum, there are the 13 matches that `parlr` made that violated one-to-one matching which Sadinle avoids.
 For two reasons, I do not consider this to be a problem.
@@
 
-- Severe overmatching within the Gibbs iterations can possibly lead to biased posteriors for the $m$ parameters.
-There is an interesting conversation to be had here.
-When Enamorado compared `fastlink` to `BRL`, he notes that the first estimate under `fastlink` has way too many matches, but that `fastlink` produces similar results to `BRL` after resolving all the matches that violate one-to-one.
+- Severe overmatching within the Gibbs iterations can possibly lead to biased posteriors for the $m$ parameters (this is Sadinle's critique of Fellegi Sunter).
+This led to me worry that parameter estimates under `parlr` would be similarly biased.
+However, good news, they are not!
+This is because the overmatching under `parlr` is very slight compared to FS and `fastlink`.
+
+<!-- When Enamorado compared `fastlink` to `BRL`, he notes that the first estimate under `fastlink` has way too many matches, but that `fastlink` produces similar results to `BRL` after resolving all the matches that violate one-to-one.
 While this is true, the overmatching within his EM algorithm lead to biases in the $m$ parameters, and these can not be cleaned up posthoc.
 In the Sadinle paper, he notes that the FS method leads to nonsensical estimates of the $m$ parameters, and I believe it is for this reason.
-
-This led to me worry that parameter estimates under `parlr` would be similarly biased.
-However, good news, they are not!
-This is because the overmatching under `parlr` is very slight compared to `fastlink`.
+-->
 
@@
 I probably could use just two levels for the names since the code puts so much work into standardizing names and using a modified Levenstein distance, but for now, I'll stick with 3 levels.
-Here $P$ is bounded by 216, and only 159 unique patterns are realized.
+Here $P$ is bounded by $3^3 \times 2^3 = 216$, and only 159 unique patterns are realized.
 Results stay basically the same, but see significant speed up for `parlr`.
@@
 
-## Mixing and Multiple Possible Matches
+\newpage
+
+## Mixing and Multiple Plausible Matches
 
 Something interesting about this comparison is that this dataset is specifically an example where `parlr` and Sadinle would differ the most.
-In cases were there multiple plausible matches, each method produces undesirable behavior.
+Interestingly, in cases were there multiple plausible matches, each method produces undesirable behavior.
 For example, examine the records of Rosa and Rosa Maria in the CDHES dataset.
@@
 
-(PS, it took me forever to figure out how to represent this visually!)
-
 ## Questions

### 2021-05-31 notes/fabl_wss.Rmd
wss slides first draft

---
title: "fabl: Fast Beta Linkage"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
library(RecordLinkage)
library(tidyverse)
```

## Overview

- Many record linkage methods are derived from Fellegi and Sunter's "A Theory for Record Linkage" (1969)

- Transform the $F$ fields of information among the $n_A$ records in File $A$ and the $n_B$ records in File $B$ into a $n_A n_B \times F$ matrix of comparison vectors $\Gamma$.

- Comparisons for each field $f$ are binned to integers $\{1, \ldots, L_f\}$, with 1 being the highest level of agreement
  - Common to use $\{1, 2\}$ to reflect exact agreement or disagreement, and $\{1, 2, 3\}$ to allow for partial agreement

## Comparison Vector Example

```{r}
knitr::include_graphics("figures/gamma_example_1.png")
```

## Comparison Vector Example

```{r}
knitr::include_graphics("figures/gamma_example_2.png")
```

## Modelling Comparison Vectors

- Comparison vectors are modelled though *reliability* parameters $\mathbf{m} = (\mathbf{m}^1, \ldots, \mathbf{m}^{F})$ and *discernment* parameters $\mathbf{u} = (\mathbf{u}^1, \ldots, \mathbf{u}^{F})$

- $m^{fl} = P(\gamma^f = l|\text{Match})$ and $u^{fl} = P(\gamma^f = l|\text{Nonmatch})$

- $m_{ij} = P(\boldsymbol{\gamma}_{ij}|\text{Match}) =  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m^{fl} \times I(\gamma_{ij}^f = l)$

- $u_{ij}$ defined similarly

- $w_{ij} = \frac{m_{ij}}{u_{ij}}$, the *Fellegi-Sunter weight*

## Modelling Linkage Structure

- To indicate matching status, we adopt the *linkage structure parameter* $\mathbf{Z} = (Z_1, \ldots, Z_{n_B})$ defined as
$$Z_j=\begin{cases} 
    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
\end{cases}$$

- More efficient than a $n_A \times n_B$ matrix

## Modelling Linkage Structure

- Fellegi-Sunter (1969) made *independent* classification decisions for all $n_A n_B$ record pairs.

- Sadinle (2017) made *dependent* matching decisions for the $n_B$ records in $B$, strictly enforcing one-to-one matching

- With Fast Beta Linkage (`fabl`, pronounced "fable") We make *independent* matching decisions for the $n_B$ records in $B$, which sometimes violates one-to-one, but allows for much faster computation

<!-- - $\lambda:= P(Z_j < n_A + 1)$ is the marginal probability that a record in $B$ has some match in $A$ -->

## Model Specification

For fields $f \in \{1, \ldots, F\}$ and levels $l\in \{1, \ldots, L_f\}$ we adopt the following likelihood and prior distributions.

$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$

$$\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})$$
$$\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})$$
$$Z_j | \lambda =
\begin{cases} 
    \frac{1}{n_A}\lambda  & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}$$

$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) $$

## Gibbs Sampling

- $\mathbf{m}$, $\mathbf{u}$, and $\lambda$ are standard multinomial-dirichlet and beta-binomial updates

- We sample $Z_j$ independently from all other $Z_{j'}$, but dependent on the $\mathbf{Z}$ vector from the previous iteration. The full conditional is

$$P\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
    w_{ij}  & i \leq n_A; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
\end{cases}$$

## Efficient Computation

- Let $\mathcal{H} = \{h_1, \ldots, h_P\}$ be the set of unique agreement patterns in the data. Note that $P \leq \prod_{f = 1}^F L_f$.

- When the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern, we say $(i,j) \in h_p$

- Wherever possible, utilize $P << n_A$ to speed up computation

## Efficient Computation

- Rather than compute $w_{ij}$ for all $n_A n_B$ record pairs, we can compute $w_p$ for each agreement pattern.

- $H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$, the total instances of agreement pattern $p$ throughout the data. Use this to speed up posterior updates for $\mathbf{m}$ and $\mathbf{u}$.

- For each $j$, calculate $H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$. Use this to speed up sampling of $\mathbf{Z}$

## Breaking Up the Z Sampler

- Sampling $n_A$ objects with varying probabilities difficult
- Sampling $P$ objects with varying probabilities is easy
- Sampling $H_{p_j}$ (can be very large) objects with *uniform* probabilities is also easy
- For a record $j \in B$, each record $i\in A$ with a given agreement pattern has the same probability of matching. That is, the unique patterns form *equivalence classes*, and within a class, each record is equally likely.

## Breaking Up the Z Sampler

- More concretely, define $h(Z_j)$ to be the agreement pattern between $j$ and its potential match, and say $h(Z_j) = h_{P+1}$ when $Z_j = n_A + 1$. We sample the agreement pattern through:

$$h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{p}\times H_{p_j}  & p \leq P; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
\end{cases}$$

- Then, we sample the record for the match uniformly among candidate records

## Speed Gains

```{r}
knitr::include_graphics("figures/sadinle_speed_plot2.png")
```

### 2021-05-31 paper/parlr_paper_1.Rmd
wss slides first draft

@@
 
-## Introduction (or Abstract?)
+## Introduction
 
@@
 This task is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error.
-Additionally, record linkage is inherently computationally intensive, and many existing methods are not feasible for many applications.
+Additionally, many record linkage tasks are large in nature, and many existing methods are too computational intensive to be used in practice.
 In this paper, we propose a Bayesian model for record linkage that is amenable to parallel computing, hashing techniques and efficient storage mechanisms, that is scalable to large linkage tasks.
+
+The proposed method is inspired by the "Beta Record Linkage" method proposed by Mauricio Sadinle's 2017 "Bayesian Estimation of Bipartite Matching for Record Linkage".
+As such, we demonstrate our method through a comparison of our methods against Sadinle's on the simulation study he designed and the El Salvador homocide case study he utilized.
+We include one additional simulation study to demonstrate the power of our method on large datasets.
+In what follows, Section 2 describes previous work in record linkage that lays the foundation for our method, Section 3 provides our model and the Gibbs sampler used for posterior inference, and Section 4 describes strategies used for more efficient and scalable implementation.
+Then, Sections 5 and 6 demonstrate the method through simulations and a case study.
+Lastly, our discussion highlights open questions in the record linkage liturature and motivates future work.
 
@@
 Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates a $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets.
-We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f$ providing their similarity score on the $f^{th}$ field.
+We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f \in \{1, \ldots, L_f\}$ providing their similarity score on the $f^{th}$ field.
 For ease of modeling and computation, we restrict these similarity scores to be discrete, ordinal variables, and the construction of these is left to the modeler.
-It is common to use use binary 0-1 variables to indicate exact matching, and 0-1-2 variables to provide an option for partial matching.
+We adopt the convention that $\gamma_{ij}^f = 1$ corresponds to the highest level of agreement and $\gamma_{ij}^f = L_f$ corresponds to the lowest.
+It is common to use use binary 1-2 variables to indicate exact matching, and 1-2-3 variables to provide an option for partial matching.
 For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model.
 
-Two modern adaptations of the Fellegi Sunter model are important for understanding `parlr` and its contribution.
+The likelihood in the original Fellegi Sunter model is a mixture model through which each record pair is independently classified as a match or nonmatch.
+This independent classification often leads to sets of matches that break expectations of transisitivity, and Jaro devised an procedure to reduce a set of conflicting matches to the mostly likely set of one-to-one matchings.
+One key strength of Sadinle's `BRL` is that it is garunteed to produce a set of bipartite matchings that respect one-to-one assumptions--without the need for any post-hoc procedures.
+Specifically, in each iteration of his Gibbs sampler, he considers each record $j\in B$, removes from consideration the records $i\in A$ that have already been matched, and then samples a potential link.
+However, accounting for these dependencies throughout the linkage process is computationally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.
+
+<!-- Two modern adaptations of the Fellegi Sunter model are important for understanding `parlr` and its contribution.
 In 2019, Enamorado et al proposed `fastlink`, a method that closely followed the modelling assumptions of Fellegi and Sunter, but used innovating hashing techniques to greatly enhance its speed and scale.
@@
 However, accounting for these dependencies throughout the linkage process is computationally burdensome, leaving `BRL` only suitable for small to moderate linkage problems.
+-->
 
@@
 However, this method remains computationally intensive and requires distributed computing to manage even moderate linkage problems in acceptable time.
+Because these methods fall outside of the Fellegi-Sunter framework, we do not consider them further in this paper.
 
@@
 
-Our notation and assumptions closely follow that of Sadinle (2017).
+Our notation and assumptions closely follow those of Sadinle.
 Denote two files as $A$ and $B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $A$ and $j \in \{1, \ldots, n_B\}$ in $B$.
@@
 
-# Model Specification
+## Model Specification
 
@@
 We however use independent priors for each $Z_j$, thereby weakening the one-to-one requirement; our sampler does ensure that each record in $B$ can be matched to at most one record in $A$, but allows for the possiblity that multiple records in $B$ match to the same record in $A$.
+We explore the ramifications of this distinction with the El Salvador case study.
 With this change, components of the $\mathbf{Z}$ vector can be computed in parallel, allowing for significant computational gains.
@@
 
+We calculate a Bayes estimate $\hat{\mathbf{Z}}$ for the linkage parameter $\mathbf{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior loss.
+Specifically, we minimize the quantity $L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)$ where
+
+$$L(\hat{Z_j}, Z_j)=\begin{cases} 
+    0  & \text{if } Z_j = \hat{Z_j}; \\
+    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
+\end{cases}$$
+
 <!-- Our Gibbs sampler provides posterior samples of $\mathbf{Z}$ which we use to make our final decisions about the linkage structure.
@@
 In random indexing, rather than store all of these unlikely record labels, we choose to store only a small number $R$ of them.
-Posterior calculations still attribute the appropriate weight to all records through the summary statistics $H_{p_j}, H_p^m$, and $H_p^u$; this is similar to the *u-correction* concept introduced by ________.
-Rather than storing $n_A \times n_B$ record labels, random indexing allows us to store at most $n_B \times P \times R$, regardless of how large $n_A$ is.
+Posterior calculations still attribute the appropriate weight to all records through the summary statistics $H_{p_j}, H_p^m$, and $H_p^u$.
+Rather than storing $n_A \times n_B$ record labels, SEI allows us to store at most $n_B \times P \times R$ labels, regardless of how large $n_A$ is.
 Coupled with the partitioning strategy for creating comparison vectors, this greatly expands the size of the linkage tasks we can undertake.
@@
 
-# Simulation Studies
+# Simulation Study 1 - Accuracy
 
-In his 2017 paper, Sadinle demonstrated the strength of his method by running `BRL` on simulated pairs of files with differing levels of errors and overlap.
-Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakened performance in the extreme scenario of very high errors and very high overlap across files.
+In his 2017 paper, Sadinle demonstrated the accuracy of his method by running `BRL` on simulated pairs of files with differing levels of errors and overlap.
+Each file consists of first name, last name, gender, postal code, age, and occupation, and following Sadinle's implementation, we use only four fields for the linkage process: both names, age, and occupation.
+Similarity scores for the names are binned into three levels according the default settings for the `compareRecords` function in the `BRL` package, and age and occupation similarity scores are binary for exact matching.
+We use flat priors for all $\mathbf{m}$ and $\mathbf{u}$ parameters, and run the Gibbs sampler for 1000 iterations, with the first 100 discarded as burn-in.
+
+Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakened performance in the most extreme scenario of very high errors and very high overlap across files.
 We use the following measures of accuracy in our analysis: although using only recall and precision is common in the literature, we find it useful to have a single metric the gauge overall performance, and thus also include F measure.
@@
 
+# Simulation Study 2 - Speed
+
 To demonstrate speed, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
@@
 We simulate these data for different values of $n_A$ and $n_B$, and compare the run-time of `parlr` against `BRL`.
+Note that the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in the larger simulations.
+
 We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data.
-
-In particular, run-time for `BRL` seems to grow quadratically while run-time for `parlr` seems to grow linearly.
+In particular, run-time for `BRL` seems to grow quadratically (or linearly with the size of both $A$ and $B$) while run-time for `parlr` seems to grow linearly (in the size of $B$).
 
@@
 
-<!-- ```{r fig.align='center'} -->
-<!-- knitr::include_graphics("../notes/figures/speed_distributions.png") -->
-<!-- ``` -->
-
 ```{r fig.align='center'}
@@
 The above discussion suggests that for fixed $n_B$, computation time should remain mostly constant with growing $n_A$.
-Shockingly, this seems to be true.
+Simulation study suggests that this is true.
 In the plot below, fixing $n_B = 500$, we see linear growth for the run-time under `BRL` as $n_A$ increases, with much more static run-time under `parlr`.
@@
 
-- Thinking of re-running the El Salvador linkage from the Sadinle 2017 paper, but doing so with without the formal blocking and indexing that he uses.
-Instead, I'd use hashing and storage efficient indexing -- I should be able to do linkage on the all the data at once!
+The country of El Salvador was immersed in civil war from 1980 to 1991, and throughout the time, several organizations attempted to document casualties of the conflict.
+When estimating the total number of casualities, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casuality lists.
+To obtain a more accurate estimate of the casualties then, we follow Sadinle's 2017 paper and link records of El Salvadoran civilian casualties from two sources: El Rescate - Tutela Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).
+The ERTL dataset consists of digitized reports that had been published throughout the conflict.
+The CDHES dataset consists of casualties that had been reported directly to the organization, and later digitized.
+
+There are several challenges with working with such data.
+Firstly, both datasets have been automatically digitized, which inherently leads to some degree of typographical error.
+Secondly, the CDHES records are all second hand accounts reported by individuals, which can result in additional errors.
+Lastly, the only fields recorded are given name, last name, date of death, and place of death; it is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
+This last point nearly breaks the above mentioned assumption that there are no duplicates within files, and reveals a key difference between `BRL` and our proposed method.
+
+We only utilized records with nonmissing entries for given and last name, results in $n_A = 4420$ files in CHDES and $n_B = 1323$ files in ERTL.
+The names were standardized to account for common misspellings in the Spanish language, and then compared using a modified Levenstein distance to account for the fact that second names are often ommited.
+Place of birth is recorded by municipality and department within that municipality; however, since  department was missing in 95% of records in CHDES and 80% of records in ERTL, we excluded it from our linkage process.
+Thus we conduct linkage using given name, last name, municipality, and day, month, and year of death.
+We again use flat priors for the $\mathbf{m}$ and $\mathbf{u}$ parameters, but use 2000 Gibbs iterations with 200 discarded as burn-in to mirror the implementation in Sadinle 2017.
+
+```{r}
+include_graphics("../notes/figures/el_salvador/overlap_distribution_smallP.png")
+```
+
+```{r}
+include_graphics("../notes/figures/el_salvador/m_posterior_smallP.png")
+```
+
+To mirror the original implementation, we constructed the comparison vectors using 4 levels of agreement for each field, according to the thresholds provided in Figure XX.
+This took ____ seconds for our proposed method and ____ for `BRL`, already demonstrating the increased efficiency of our method.
+However, we note that posterior distributions of several levels of the $\mathbf{m}$ and $\mathbf{u}$ parameters were nearly identical, and that of the $4^5 \times 2 = 2048$ possible agreement patterns, only ___ are realized in the data.
+This leads us to believe that such high number of aggreement levels creates unnecesary distictions in the data and makes the comparison vectors less interpretable.
+Therefore we re-ran our analysis with fewer agreement levels for each field (see Figure XX), and obtained analogous results.
+With ____ possible agreement patterns, ___ were realized in the data, and our proposed method became much faster, finishing in ___ seconds.
+Meanwhile `BRL` took ___ seconds, relatively unchanged from the first implementation.
+This demonstrates the way that the computational complexity of our method depends on the number of unique agreement patterns, and how significant computional gains can be made by simplifying the construction of the comparison vectors.
+We also note that estimates of the $\mathbf{m}$ and $\mathbf{u}$ paramters under each method are very similar, as shown in Figure XX.
+
+## Violations of One-to-One Matching
+
+Figure XX shows the posterior distribution for $D$, the number of duplicates found across file for both `fabl` and `BRL`.
+We see that `fabl` consistently overmatches within each Gibbs iteration when compared to `BRL`, which is to be expected because `BRL` explicitly prevents matches that violate one-to-one matching throughout the entire sampler.
+Most of such matchings are due to the randomness in the sampling procedure, and they occur sporadically throughout the sampler in such a way that does not measurably influence the eventual Bayes estimate.
+
+Overall, both methods presented similar results.
+`fabl` yielded an initial Bayes estimate of ____ matches found across files, and after resolving matches that violated one-to-one requirements, yeilded ___ matches across files.
+This is acceptably close to the ___ matches found by `BRL`.
+The main reason for this discrepancy is the difference in how each model handles situations in which one file in $B$ has multiple plausible matches in $A$.
+
+```{r}
+include_graphics("../notes/figures/el_salvador/rosa_records.png")
+```
+
+```{r}
+include_graphics("../notes/figures/el_salvador/bad_mixing.png")
+```
+
+The records shown in Figure XX provide one such instance.
+Note these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records ___ and ____ in ERTL correspond to different individuals, but their records are nearly identical.
+Additionally, using the modified Levenstein distance from Sadinle 2017, the comparison vectors $\gamma_{2776, 825}$ and $\gamma_{2776, 826}$ are exactly identical.
+
+Figure XX shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrate how each method handles this situation.
+Under `parlr`, both records in $B$ match to the same record in $A$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
+Under `BRL`, the Gibbs process creates one matching configuration stays there for a while.
+However, if one pair "unmatches," then the other record has a chance to latch on.
+Then,  Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
+Additionally, `parlr` allows the modeller to inspect records with mulitple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
+`BRL` in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
 
@@
 
-<!-- We note that this often not a reasonable assumption; for example if we know that two records differ on gender, this does not give any information on whether their year of births are the same, but it does suggest that its more likely that their first names are different.
--->
+We have presented a method for Bayesian record linkage that is feasible for large datasets.
+In particular, our hashing procedure and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making this method particularly powerful in linking records when one datafile is substantially smaller than the other.
+
+In our case study, we included an exploration of how to conduct record linkage when modelling assumptions are not met in practice. We explored "one-to-many" scenarios in which

### 2021-05-31 paper/parlr_paper_1.tex
wss slides first draft

@@
 
-\hypertarget{introduction-or-abstract}{%
-\subsection{Introduction (or
-Abstract?)}\label{introduction-or-abstract}}
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
@@
 identifiers or highly reliable information, but becomes difficult when
-data sources are plagued with error. Additionally, record linkage is
-inherently computationally intensive, and many existing methods are not
-feasible for many applications. In this paper, we propose a Bayesian
-model for record linkage that is amenable to parallel computing, hashing
-techniques and efficient storage mechanisms, that is scalable to large
-linkage tasks.
+data sources are plagued with error. Additionally, many record linkage
+tasks are large in nature, and many existing methods are too
+computational intensive to be used in practice. In this paper, we
+propose a Bayesian model for record linkage that is amenable to parallel
+computing, hashing techniques and efficient storage mechanisms, that is
+scalable to large linkage tasks.
+
+The proposed method is inspired by the ``Beta Record Linkage'' method
+proposed by Mauricio Sadinle's 2017 ``Bayesian Estimation of Bipartite
+Matching for Record Linkage''. As such, we demonstrate our method
+through a comparison of our methods against Sadinle's on the simulation
+study he designed and the El Salvador homocide case study he utilized.
+We include one additional simulation study to demonstrate the power of
+our method on large datasets. In what follows, Section 2 describes
+previous work in record linkage that lays the foundation for our method,
+Section 3 provides our model and the Gibbs sampler used for posterior
+inference, and Section 4 describes strategies used for more efficient
+and scalable implementation. Then, Sections 5 and 6 demonstrate the
+method through simulations and a case study. Lastly, our discussion
+highlights open questions in the record linkage liturature and motivates
+future work.
 
@@
 
-Two modern adaptations of the Fellegi Sunter model are important for
-understanding \texttt{parlr} and its contribution. In 2019, Enamorado et
-al proposed \texttt{fastlink}, a method that closely followed the
-modelling assumptions of Fellegi and Sunter, but used innovating hashing
-techniques to greatly enhance its speed and scale. Importantly, they
-maintained Fellegi and Sunter's independent matching assumption, in
-which the matching status of any \((i,j)\) pair is made independently of
-the matching of any other \((i', j')\). This leads to many matchings
-that violate ``one-to-one'' considerations that are often implicit in
-the data, which need to be resolved in a post-processing step. In
-contrast, Sadinle (2017) proposed ``Beta Record Linkage for Bipartite
-Matching'' and the accompanying \texttt{BRL} package which strictly
-enforces one-to-one matching. Specifically, in each iteration of his
-Gibbs sampler, he considers each record \(j\in B\), removes from
-consideration the records \(i\in A\) that have already been matched,
-then samples a potential link. However, accounting for these
-dependencies throughout the linkage process is computationally
-burdensome, leaving \texttt{BRL} only suitable for small to moderate
-linkage problems.
+The likelihood in the original Fellegi Sunter model is a mixture model
+through which each record pair is independently classified as a match or
+nonmatch. This independent classification often leads to sets of matches
+that break expectations of transisitivity, and Jaro devised an procedure
+to reduce a set of conflicting matches to the mostly likely set of
+one-to-one matchings. One key strength of Sadinle's \texttt{BRL} is that
+it is garunteed to produce a set of bipartite matchings that respect
+one-to-one assumptions--without the need for any post-hoc procedures.
+Specifically, in each iteration of his Gibbs sampler, he considers each
+record \(j\in B\), removes from consideration the records \(i\in A\)
+that have already been matched, and then samples a potential link.
+However, accounting for these dependencies throughout the linkage
+process is computationally burdensome, leaving \texttt{BRL} only
+suitable for small to moderate linkage problems.
 
@@
 distributed computing to manage even moderate linkage problems in
-acceptable time.
+acceptable time. Because these methods fall outside of the
+Fellegi-Sunter framework, we do not consider them further in this paper.
 
@@
 
-Our notation and assumptions closely follow that of Sadinle (2017).
-Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
-respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
-in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
-generality, label the files such that \(n_A \geq n_B\). We also assume
-there are no duplicates within files, only across. For each record pair
-under consideration, we generate a comparison vector
+Our notation and assumptions closely follow those of Sadinle. Denote two
+files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records respectively,
+and with records indexed as \(i \in \{1, \ldots, n_A\}\) in \(A\) and
+\(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of generality, label
+the files such that \(n_A \geq n_B\). We also assume there are no
+duplicates within files, only across. For each record pair under
+consideration, we generate a comparison vector
 \(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
@@
 \hypertarget{model-specification}{%
-\section{Model Specification}\label{model-specification}}
+\subsection{Model Specification}\label{model-specification}}
 
@@
 record in \(A\), but allows for the possiblity that multiple records in
-\(B\) match to the same record in \(A\). With this change, components of
-the \(\mathbf{Z}\) vector can be computed in parallel, allowing for
-significant computational gains. More importantly, since only the
-agreement pattern of \(Z_j\) is used for calculations within the Gibbs
-sampler, and not the particular record label, we can conduct this
-sampling only at the level of the unique agreement patterns, offering
-even more computational savings. This will be explained further in
-\textcolor{red}{Section XX}.
+\(B\) match to the same record in \(A\). We explore the ramifications of
+this distinction with the El Salvador case study. With this change,
+components of the \(\mathbf{Z}\) vector can be computed in parallel,
+allowing for significant computational gains. More importantly, since
+only the agreement pattern of \(Z_j\) is used for calculations within
+the Gibbs sampler, and not the particular record label, we can conduct
+this sampling only at the level of the unique agreement patterns,
+offering even more computational savings. This will be explained further
+in \textcolor{red}{Section XX}.
 
@@
 calculations still attribute the appropriate weight to all records
-through the summary statistics \(H_{p_j}, H_p^m\), and \(H_p^u\); this
-is similar to the \emph{u-correction} concept introduced by
-\_\_\_\_\_\_\_\_. Rather than storing \(n_A \times n_B\) record labels,
-random indexing allows us to store at most \(n_B \times P \times R\),
-regardless of how large \(n_A\) is. Coupled with the partitioning
-strategy for creating comparison vectors, this greatly expands the size
-of the linkage tasks we can undertake.
+through the summary statistics \(H_{p_j}, H_p^m\), and \(H_p^u\). Rather
+than storing \(n_A \times n_B\) record labels, SEI allows us to store at
+most \(n_B \times P \times R\) labels, regardless of how large \(n_A\)
+is. Coupled with the partitioning strategy for creating comparison
+vectors, this greatly expands the size of the linkage tasks we can
+undertake.
 
@@
 
-\hypertarget{simulation-studies}{%
-\section{Simulation Studies}\label{simulation-studies}}
+\hypertarget{simulation-study-1---accuracy}{%
+\section{Simulation Study 1 -
+Accuracy}\label{simulation-study-1---accuracy}}
 
-In his 2017 paper, Sadinle demonstrated the strength of his method by
+In his 2017 paper, Sadinle demonstrated the accuracy of his method by
 running \texttt{BRL} on simulated pairs of files with differing levels
-of errors and overlap. Comparing our method against \texttt{BRL} on
-these same simulated datasets, we find that our method only has weakened
-performance in the extreme scenario of very high errors and very high
-overlap across files. We use the following measures of accuracy in our
-analysis: although using only recall and precision is common in the
-literature, we find it useful to have a single metric the gauge overall
-performance, and thus also include F measure.
+of errors and overlap. Each file consists of first name, last name,
+gender, postal code, age, and occupation, and following Sadinle's
+implementation, we use only four fields for the linkage process: both
+names, age, and occupation. Similarity scores for the names are binned
+into three levels according the default settings for the
+\texttt{compareRecords} function in the \texttt{BRL} package, and age
+and occupation similarity scores are binary for exact matching. We use
+flat priors for all \(\mathbf{m}\) and \(\mathbf{u}\) parameters, and
+run the Gibbs sampler for 1000 iterations, with the first 100 discarded
+as burn-in.
+
+Comparing our method against \texttt{BRL} on these same simulated
+datasets, we find that our method only has weakened performance in the
+most extreme scenario of very high errors and very high overlap across
+files. We use the following measures of accuracy in our analysis:
+although using only recall and precision is common in the literature, we
+find it useful to have a single metric the gauge overall performance,
+and thus also include F measure.
 
@@
 \begin{center}\includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot} \end{center}
+
+\hypertarget{simulation-study-2---speed}{%
+\section{Simulation Study 2 - Speed}\label{simulation-study-2---speed}}
 
@@
 \(n_A\) and \(n_B\), and compare the run-time of \texttt{parlr} against
-\texttt{BRL}. We see that at low data size, \texttt{BRL} outperforms,
-but that \texttt{parlr} is significantly faster at handling larger data.
+\texttt{BRL}. Note that the number of unique patterns \(P\) is bounded
+above by \(2^5 = 32\), a bound which is consistently attained in the
+larger simulations.
 
-In particular, run-time for \texttt{BRL} seems to grow quadratically
-while run-time for \texttt{parlr} seems to grow linearly.
+We see that at low data size, \texttt{BRL} outperforms, but that
+\texttt{parlr} is significantly faster at handling larger data. In
+particular, run-time for \texttt{BRL} seems to grow quadratically (or
+linearly with the size of both \(A\) and \(B\)) while run-time for
+\texttt{parlr} seems to grow linearly (in the size of \(B\)).
 
@@
 The above discussion suggests that for fixed \(n_B\), computation time
-should remain mostly constant with growing \(n_A\). Shockingly, this
-seems to be true. In the plot below, fixing \(n_B = 500\), we see linear
-growth for the run-time under \texttt{BRL} as \(n_A\) increases, with
-much more static run-time under \texttt{parlr}. The slight increases in
-run-time that we do see are due primarily to the hashing step, which
-again can be run in parallel for large data.
+should remain mostly constant with growing \(n_A\). Simulation study
+suggests that this is true. In the plot below, fixing \(n_B = 500\), we
+see linear growth for the run-time under \texttt{BRL} as \(n_A\)
+increases, with much more static run-time under \texttt{parlr}. The
+slight increases in run-time that we do see are due primarily to the
+hashing step, which again can be run in parallel for large data.
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Thinking of re-running the El Salvador linkage from the Sadinle 2017
-  paper, but doing so with without the formal blocking and indexing that
-  he uses. Instead, I'd use hashing and storage efficient indexing -- I
-  should be able to do linkage on the all the data at once!
-\end{itemize}
+the country of El Salvador was immersed in civil war from 1980 to 1991,
+and throughout the time, several organizations attempted to document
+casualties of the conflict. When estimating the total number of
+casualities, one cannot simply sum the numbers recorded by each
+organization, as it is likely that the same individuals are recorded in
+multiple casuality lists. To obtain a more accurate estimate of the
+casualties then, we follow Sadinle's 2017 paper and link records of El
+Salvadoran civilian casualties from two sources: El Rescate - Tutela
+Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
+acronym in Spanish). The ERTL dataset consists of digitized reports that
+had been published throughout the conflict. The CDHES dataset consists
+of casualties that had been reported directly to the organization, and
+later digitized.
+
+There are several challenges with working with such data. Firstly, both
+datasets have been automatically digitized, which inherently leads to
+some degree of typographical error. Secondly, the CDHES records are all
+second hand accounts reported by individuals, which can result in
+additional errors. Lastly, the only fields recorded are given name, last
+name, date of death, and place of death; it is relatively common for a
+parent and child to share the same given name, resulting in
+indistinguishable records for two different individuals. This last point
+nearly breaks the above mentioned assumption that there are no
+duplicates within files, and reveals a key difference between
+\texttt{BRL} and our proposed method.
+
+We only utilized records with nonmissing entries for given and last
+name, results in \(n_A = 4420\) files in CHDES and \(n_B = 1323\) files
+in ERTL. The names were standardized to account for common misspellings
+in the Spanish language, and then compared using a modified Levenstein
+distance to account for the fact that second names are often ommited.
+Place of birth is recorded by municipality and department within that
+municipality; however, since department was missing in 95\% of records
+in CHDES and 80\% of records in ERTL, we excluded it from our linkage
+process. Thus we conduct linkage using given name, last name,
+municipality, and day, month, and year of death. We again use flat
+priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters, but use
+2000 Gibbs iterations with 200 discarded as burn-in to mirror the
+implementation in Sadinle 2017.
 

### 2021-05-31 R/NLC_hash2.R

@@ -22,7 +22,9 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
     cl <- parallel::makeCluster(numCores)
   }
 
+  ptm <- proc.time()
   patterns <- GetUniquePatterns2(comparisons, fast, R)
+  elapsed_hash <- proc.time() - ptm
 
   parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
     rep(x, comparisons[[4]][x])
@@ -186,7 +188,8 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
 
   list(Z = final_gibbs,
        m = M.SAMPS,
-       u = U.SAMPS)
+       u = U.SAMPS,
+       hash_time = elapsed_hash[3])
 
 }
 

### 2021-06-04 notes/fabl_wss.Rmd
wss second draft

@@
 - Comparisons for each field $f$ are binned to integers $\{1, \ldots, L_f\}$, with 1 being the highest level of agreement
+
   - Common to use $\{1, 2\}$ to reflect exact agreement or disagreement, and $\{1, 2, 3\}$ to allow for partial agreement
@@
 $$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
-
 $$\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})$$
@@
 \end{cases}$$
-
 $$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) $$

### 2021-06-04 notes/fabl_wss_2.Rmd
wss second draft

---
title: "fabl: Fast Beta Linkage"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
library(RecordLinkage)
library(tidyverse)
```

## El Salvadoran Civil War

- El Salvadoran Civil War (1979 - 1992)

- Multiple organizations have documented casualties, and we need to find duplicates across separate datafiles to accurately estimate *documented identifiable deaths* (DID)

- Not much information, only first name, last name, and date and place of death
- Spanish language: very common for parent and child to share the same name, "second" names often omitted

## Foundations of Method

- Many record linkage methods are derived from Fellegi and Sunter's "A Theory for Record Linkage" (1969)

- Transform the $F$ fields of information among the $n_A$ records in File $A$ and the $n_B$ records in File $B$ into a $n_A n_B \times F$ matrix of comparison vectors $\Gamma$.

- Comparisons for each field $f$ are binned to integers $\{1, \ldots, L_f\}$, with 1 being the highest level of agreement

  - Common to use $\{1, 2\}$ to reflect exact agreement or disagreement, and $\{1, 2, 3\}$ to allow for partial agreement

## Comparison Vector Example

```{r}
knitr::include_graphics("figures/gamma_example_1.png")
```

## Comparison Vector Example

```{r}
knitr::include_graphics("figures/gamma_example_2.png")
```
<!-- ## Challenges -->

<!-- ```{r} -->
<!-- knitr::include_graphics("figures/el_salvador/rosa_records.png") -->
<!-- ``` -->

<!-- - Not much information, only first name, last name, and date and place of death -->
<!-- - Spanish language: very common for parent and child to share the same name, "second" names often omitted -->

## Previous Work

- Fellegi-Sunter (1969) made matching decisions on all $n_A n_B$ pairs independently. Fast, but can lead to severe overmatching.

- Sadinle (2017) proposed *Beta Record Linkage* (`BRL`), a Bayesian method that strictly enforces one-to-one matching. Matching proposals are *dependent*, resulting in a slow Gibbs sampler.

## Contribution

- *Fast Beta Linkage* (`fabl`, pronounced "fable") makes one proposal per record in $B$, but makes these proposals *independently*. Allows for much quicker Gibbs sampler

- Hash comparison vectors to unique patterns, perform calculations over $P$ patterns rather than $n_An_B$ record pairs

- *Storage efficient indexing* to greatly use storage requirements of comparison vectors, allowing for larger linkage tasks.

- Gibbs sampler under `fabl` has computational complexity $O(n_B \times P \times F)$, significant savings compared to `BRL`'s $O(n_B \times n_A \times F)$

## Simulation

```{r}
knitr::include_graphics("figures/sadinle_speed_plot2.png")
```

## El Salvador Data

```{r}
knitr::include_graphics("figures/el_salvador/overlap_distribution_smallP.png")
```

## El Salvador Data

```{r}
knitr::include_graphics("figures/el_salvador/DID_distribution_smallP.png")
```

### 2021-06-04 R/GetUniquePatterns2.R

@@ -10,72 +10,11 @@ GetUniquePatterns2 <- function(cd, fast = F, R = NULL){
   rec1 <- ids[,1]
   rec2 <- ids[,2]
   levels <- cd[[4]]
-  # fields <- length(levels)
-  # gk <- unlist(lapply(1:fields, function(x){
-  #   0:(levels[x] -1)
-  # }))
-  # Lf_vec <- unlist(lapply(1:fields, function(x){
-  #   rep(levels[x], levels[x])
-  # }))
-  #
-  # fields_vec <- unlist(lapply(1:fields, function(x){
-  #   rep(seq_len(fields)[x], levels[x])
-  # }))
 
-
-  #gk_ij <- sweep(indicators, MARGIN = 2, gk, FUN = "*")
-  # Lf_long <- sapply(Lf_vec, function(x){
-  #   rep(x, N)
-  # })
-  # fields_long <- sapply(fields_vec, function(x){
-  #   rep(x, N)
-  # })
-
-  # level_sum <- cumsum(levels)
-  # level_sum_vec <-  unlist(lapply(1:fields, function(x){
-  #   rep(level_sum[x], levels[x])
-  # }))
 df <- data.frame(indicators, rec1, rec2)
 
 pattern_df <- df %>%
   unite(pattern, 1:fields, sep = "")
-  # level_sum_long <- sapply(level_sum_vec, function(x){
-  #   rep(x, N)
-  # })
-
-  #hash <- rowSums((gk_ij >0) * 2 ^ (gk_ij + (fields_long > 1) * level_sum_long))
-  # if(fast){
-  #   numCores <- parallel::detectCores()
-  #   cl <- parallel::makeCluster(numCores)
-  #
-  #   parallel::clusterExport(cl, c("gk", "fields_vec", "level_sum_vec"))
-  #   hash_temp <- parallel::parApply(cl, indicators, 1, FUN = function(x){
-  #     gk_ij <- x * gk
-  #     (gk_ij >0) * 2 ^ (gk_ij + (fields_vec > 1) * level_sum_vec)
-  #   })
-  #
-  #   parallel::stopCluster(cl)
-  #
-  # } else {
-  #
-  #   hash_temp <- apply(indicators, 1, function(x){
-  #     gk_ij <- x * gk
-  #     (gk_ij >0) * 2 ^ (gk_ij + (fields_vec > 1) * level_sum_vec)
-  #   })
-
-
-    # thing <- lapply(seq_len(nrow(indicators)), function(x){
[…]

### 2021-06-06 notes/fabl_wss_2.Rmd
first draft of paper

@@
 
-- Multiple organizations have documented casualties, and we need to find duplicates across separate datafiles to accurately estimate *documented identifiable deaths* (DID)
+- Multiple organizations have documented casualties. El Rescate - Tutela Legal (ERTL) contains $n_A = 4420$ records, and  Salvadoran Human Rights Commission (CDHES) contains $n_B =  1323$ records.
+
+- We need to find duplicates across separate datafiles to accurately estimate *documented identifiable deaths* (DID)
 
 - Not much information, only first name, last name, and date and place of death
+
 - Spanish language: very common for parent and child to share the same name, "second" names often omitted

### 2021-06-06 paper/parlr_paper_1.Rmd
first draft of paper

@@
 	\end{flushright}
-	\begin{center} \textbf{Parlr: Parallelized Record Linkage in R}
+	\begin{center} \textbf{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}
 
-	Brian Kundinger
+	Brian Kundinger, Rebecca Steorts, and Jerome Reiter
 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, fig.align='center')
 ```
@@
 
-Record linkage is the task of identifying duplicate records across multiple data sources.
+Entity resolution (or record linkage) is the task of identifying duplicate records across multiple data sources.
 This task is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error.
-Additionally, many record linkage tasks are large in nature, and many existing methods are too computational intensive to be used in practice.
-In this paper, we propose a Bayesian model for record linkage that is amenable to parallel computing, hashing techniques and efficient storage mechanisms, that is scalable to large linkage tasks.
+Additionally, many entity resolution tasks are large in nature, and many existing methods are too computational intensive to be used in practice.
+In this paper, we propose a Bayesian model for record linkage that uses hashing techniques and efficient storage mechanisms, and is amenable to parallelization, making it scalable to large linkage tasks.
 
-The proposed method is inspired by the "Beta Record Linkage" method proposed by Mauricio Sadinle's 2017 "Bayesian Estimation of Bipartite Matching for Record Linkage".
-As such, we demonstrate our method through a comparison of our methods against Sadinle's on the simulation study he designed and the El Salvador homocide case study he utilized.
-We include one additional simulation study to demonstrate the power of our method on large datasets.
+The proposed method is inspired by the *Beta Record Linkage* (`BRL`) method proposed in Mauricio Sadinle's 2017 "Bayesian Estimation of Bipartite Matching for Record Linkage".
+To recognize this lineage, we name our method *Fast Beta Linkage* (`fabl`, pronounced "fable").
 In what follows, Section 2 describes previous work in record linkage that lays the foundation for our method, Section 3 provides our model and the Gibbs sampler used for posterior inference, and Section 4 describes strategies used for more efficient and scalable implementation.
-Then, Sections 5 and 6 demonstrate the method through simulations and a case study.
+Then, Section 5  demonstrates the accuracy and speed of out method through two simulation studies, and Section 6 reveals interesting properties of the method through a case study of El Salvadoran casuality records.
 Lastly, our discussion highlights open questions in the record linkage liturature and motivates future work.
@@
 Most record linkage techniques are derived from the seminal 1969 paper by Fellegi and Sunter: "A Theory for Record Linkage".
-The defining characteristic of their model was to transform the sets of records, which often contain text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated.
-Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates a $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets.
+The defining characteristic of their approach was to transform the sets of records, which often contain text data that is difficult to model, into sets of comparison vectors governed by parameters that can be more easily estimated.
+Concretely, if files $A$ and $B$ have $n_A$ and $n_B$ records respectively, and if the files share $F$ fields in common upon which to base the linkage, the Fellegi and Sunter approach generates an $n_A n_B \times F$ matrix $\Gamma$, which contains similarity scores between each pair of records across datasets.
 We say $\gamma_{ij}$ is the comparison vector for record $i \in A$ and record $j \in B$, with $\gamma_{ij}^f \in \{1, \ldots, L_f\}$ providing their similarity score on the $f^{th}$ field.
@@
 For text data, we calculate similarity based on Levenstein distance or some other text similarity score, and bin these scores to integers for use in the model.
+Jaro (1989) provided guidelines for appropriate binning for the comparison vectors, but details of implementation are generally context specific.
 
 The likelihood in the original Fellegi Sunter model is a mixture model through which each record pair is independently classified as a match or nonmatch.
-This independent classification often leads to sets of matches that break expectations of transisitivity, and Jaro devised an procedure to reduce a set of conflicting matches to the mostly likely set of one-to-one matchings.
+This independent classification often leads to sets of matches that break expectations of transitivity, which is undesirable.
+Although Jaro devised an procedure to reduce a set of conflicting matches to the mostly likely set of one-to-one matchings, the abundance of false matches can bias the estimation of other parameters of the model.
+
 One key strength of Sadinle's `BRL` is that it is garunteed to produce a set of bipartite matchings that respect one-to-one assumptions--without the need for any post-hoc procedures.
@@
 
-<!-- Two modern adaptations of the Fellegi Sunter model are important for understanding `parlr` and its contribution.
+<!-- Two modern adaptations of the Fellegi Sunter model are important for understanding `fabl` and its contribution.
 In 2019, Enamorado et al proposed `fastlink`, a method that closely followed the modelling assumptions of Fellegi and Sunter, but used innovating hashing techniques to greatly enhance its speed and scale.
@@
 
-Our notation and assumptions closely follow those of Sadinle.
 Denote two files as $A$ and $B$, with $n_A$ and $n_B$ records respectively, and with records indexed as $i \in \{1, \ldots, n_A\}$ in $A$ and $j \in \{1, \ldots, n_B\}$ in $B$.
 Without loss of generality, label the files such that $n_A \geq n_B$.
-We also assume there are no duplicates within files, only across.
+We assume there are no duplicates within files, only across.
 For each record pair under consideration, we generate a comparison vector $\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}$, where  $F$ is the number of fields used in the linkage and each $\gamma_{ij}^f$ takes on a value $l \in \{1, \ldots, L_f\}$ indicating the level agreement between the two records on a specified field.
@@
 The prior for $Z_j$ has equal probability of matching to all records $i\in A$, and non-matching probability governed by $\lambda$.
-Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}(1, \frac{1}{n_A})$ prior corresponds to a uniform prior on the labeling of $\mathbf{Z}$.
+Therefore a  $\lambda \sim \text{Beta}(1, 1)$ corresponds to a prior belief that nonmatches and matches are equally likely, and a $\lambda \sim \text{Beta}\left(1, \frac{1}{n_A}\right)$ prior corresponds to a uniform prior on the labeling of $\mathbf{Z}$.
 
@@
 In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*.
-We however use independent priors for each $Z_j$, thereby weakening the one-to-one requirement; our sampler does ensure that each record in $B$ can be matched to at most one record in $A$, but allows for the possiblity that multiple records in $B$ match to the same record in $A$.
+We however use independent priors for each $Z_j$, creating a sampler that is *perfectly parallelizable*, allowing for significant computational gains.
+More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns, offering even more computational savings.
+In doing so however, we thereby weaken the one-to-one requirement from `BRL`; our sampler does ensure that each record in $B$ can be matched to at most one record in $A$, but allows for the possiblity that multiple records in $B$ match to the same record in $A$.
 We explore the ramifications of this distinction with the El Salvador case study.
-With this change, components of the $\mathbf{Z}$ vector can be computed in parallel, allowing for significant computational gains.
-More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns, offering even more computational savings.
-This will be explained further in \textcolor{red}{Section XX}.
 
@@
 
-We calculate a Bayes estimate $\hat{\mathbf{Z}}$ for the linkage parameter $\mathbf{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior loss.
+We calculate a Bayes estimate $\hat{\mathbf{Z}}$ for the linkage parameter $\mathbf{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
+We adopt the loss function proposed in Sadinle 2017, in which $\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}$, with $R$ representing the option to leave the matching undetermined by the model.
 Specifically, we minimize the quantity $L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)$ where
@@
     0  & \text{if } Z_j = \hat{Z_j}; \\
-    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
+    \theta_R,  & \text{if } \hat{Z_j} = R; \\
+    \theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
+    \theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
+    \theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
+\end{cases}$$
+Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false non-match, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
+This loss function leads to closed form decision rules for minimizing posterior expected loss.
+In this paper, we adopt losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule
+
+$$\hat{Z}_j =\begin{cases} 
+    i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
+    0,  & \text{otherwise} ; \\
 \end{cases}$$
 
-<!-- Our Gibbs sampler provides posterior samples of $\mathbf{Z}$ which we use to make our final decisions about the linkage structure.
-In the case where we false matches and missed matches contribute the same loss, we declare $(i,j)$ to be a match whenever $P(Z_j = i) > \frac{1}{2}$ according to these posterior samples.
-We can create more elaborate decisions by attributing different loss values to different kinds of error, and allowing pairings with middling posterior probabilities to be left unlabeled by the algorithm so that the modeler can address those manually.
-For further discussion, see Sadinle 2017.
--->
-
-(Adapt from Sadinle)
+For a more indepth explanation of this function and the induced Bayes estimate, see Sadinle 2017.
 
@@
 The modeler can either report both such matches (with their respective posterior match probabilities), or resolve these conflicts by accepting only the match with highest posterior probability.
-Rigorous proof of this resolution procedure is deferred to future work.
+Such a resolution procedure indeed is equivalent to minimizing posterior risk under the restriction of one-to-one matching, and as thus theoretically justified within the Bayesian framework.
 
@@
 Broadly speaking, we increase our computational efficiency by recognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-Let $\mathcal{H}$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns.
+Let $\mathcal{H}$ be the set of unique agreement patterns in the data, let $P = |\mathcal{H}|$ denote the total number of unique agreement patterns.
 Note that $P$ is bounded above by $\prod_{f=1}^F L_f$, and that this bound does not scale with $n_A$ or $n_B$.
@@
 
-The classic Fellegi Sunter method represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$.
+The classic Fellegi Sunter method represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{1, \ldots, L_f \}$.
 To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicator vectors of lengths $L_f$.
-For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework.
+For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (2, 1, 3)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework.
 This is a bijective transformation that does not change the meaning of the data, but this representation will ease calculations and posterior updates.
@@
 In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair.
-We however do not store these comparison vectors themselves, but instead only a hashed value corresponding to the agreement pattern of the $(i, j)$ pair.
+In contrast, we do not store these comparison vectors themselves, but instead only a hashed value corresponding to the agreement pattern of the $(i, j)$ pair.
 We store this information in a nested list $\tilde{\Gamma}$ where the $p^{th}$ component of the $j^{th}$ list contains a vector of records in $A$ that share agreement pattern $p$ with record $j \in B$.
 For each $p$, we also calculate $H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}$, the total instances of agreement pattern $p$ throughout the data, and also for each $j$, we  calculate $H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}$ the instances of agreement pattern $p$ among the comparison vectors between record $j \in B$ and each of the $n_A$ records in $A$.
-
-For large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\}$ for more manageable computations.
-On a single machine, we can read-in data sequentially, conduct hashing, collect results, and delete the original data from memory before continuing with the next chunk of data.
-With multiple cores or multiple machines, this can be done in parallel.
-Storing this hashed information still becomes burdensome for large data, but this hashing method greatly expands the capabilities of the Fellegi-Sunter framework.
 
@@
 Therefore we know that records belonging to an $h_p$ such that $H_{p_j}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate, even without considering the form of the agreement pattern itself.
-In random indexing, rather than store all of these unlikely record labels, we choose to store only a small number $R$ of them.
-Posterior calculations still attribute the appropriate weight to all records through the summary statistics $H_{p_j}, H_p^m$, and $H_p^u$.
+
+In SEI, rather than store all of these unlikely record labels, we choose to store only a small number $R$ of them.
+Posterior calculations still attribute the appropriate weight to all records through the summary statistics $H_p$, and $H_{p_j}$.
 Rather than storing $n_A \times n_B$ record labels, SEI allows us to store at most $n_B \times P \times R$ labels, regardless of how large $n_A$ is.
-Coupled with the partitioning strategy for creating comparison vectors, this greatly expands the size of the linkage tasks we can undertake.
+
+Laslty, for large data, we can partition the two datasets $A$ and $B$ into smaller blocks $\{A_m\}$ and $\{B_m\}$ for more manageable computations.
+On a single machine, we can read-in data sequentially, conduct hashing, compress information through SEI, and delete the original data from memory before continuing with the next chunk of data.
+With multiple cores or multiple machines, this can be done in parallel.
+Thus the combination of hashing, SEI, and partitioning allows us to conduct linkage tasks over much larger datasets.
 
@@
 
-<!-- Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching, given by  -->
-
-<!-- $$p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto -->
-<!-- \begin{cases}  -->
-<!--     w_{ij} \mathbf{1}_{i \neq Z_{j'}; \; j' \neq j} & i \leq n_A; \\ -->
-<!--      D_{-j} \frac{n_B - D_{-j} + \beta_{\lambda}}{D_{-j} + \alpha_{\lambda}} & i  = n_A + 1 \\ -->
-<!-- \end{cases}$$ -->
-
-<!-- The indicator removes previously matched records from the set of candidate records, and $D_{-j}:=\sum_{B \setminus j} $ updates iteratively after each matching decision, $n_B$ many times per Gibbs iteration.
-This process is *inherently serial* and becomes computationally burdensome for large data.
--->
-
-<!-- Sadinle uses a prior for $\mathbf{Z}$ that induces the a full conditional for $Z_j$ that strictly enforces one-to-one matching.
-In particular, this sampler removes previously matches records from the set of candidate records when sampling $Z_j$, creating a dependency that makes the sampler *inherently serial*.
-By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel.
-More importantly, since only the agreement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns.
-This boosts computation time far greater than parallelization.
--->
-
 \underline{Updating $\mathbf{Z}$:} Although sampling $Z_j$ from a the full conditional provided earlier is conceptually straightforward, it becomes computational burdensome when $n_A$ is larger.
@@
 To speed up computation, we break this sampling step into two simpler steps.
-First, we calculate the Fellegi Sunter weight $w_{h_p}$ associated with each unique pattern and sample the agreement pattern between $j$ and its potential match.
+First, we calculate the Fellegi Sunter weight $w_{p}$ associated with each unique pattern and sample the agreement pattern between $j$ and its potential match.
 Second, we sample the record label uniformly among records associated with that agreement pattern for that particular $j\in B$.
@@
 
-$$h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+$$P\left(h\left(Z_j^{(s+1)}\right) = p | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
 \begin{cases} 
-    w_{h_p}\times H_{p_j}  & p \leq P; \\
+    w_{p}\times H_{p_j}  & p \leq P; \\
      n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
 \end{cases}$$
-<!-- $$p\left(Z_j^{(s+1)} = i\right) = \frac{1}{\left|H\left(Z_j^{(s+1)}\right)\right|}; \;\;i \text{ such that } (i, j) \in H\left(Z_j^{(s+1)}\right)$$ -->
+$$P\left(Z_j^{(s+1)} = i\bigg|\;\; h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
+\frac{1}{H_{p_j}} & (i, j) \in h_p \\
+0 & \text{otherwise} \\
+\end{cases}$$
 
 Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
-Thuse We complete the entire Gibbs procedure at the level of the $P$ agreement patterns.
-After, we can back-fill the records corresponding to the agreement patterns by sampling uniformly at random among candidate records stored in $\tilde{\Gamma}$.
+Thus we complete the entire Gibbs procedure first at the level of the $P$ agreement patterns with the first equation above.
+After, we can back-fill the records corresponding to the agreement patterns through the second equation.
 Sampling uniformly is computational simple even for large sets of candidate records, but this step can also be parallelized when working with large data.
@@
 Comparing our method against `BRL` on these same simulated datasets, we find that our method only has weakened performance in the most extreme scenario of very high errors and very high overlap across files.
-We use the following measures of accuracy in our analysis: although using only recall and precision is common in the literature, we find it useful to have a single metric the gauge overall performance, and thus also include F measure.
+We use the following measures of accuracy in our analysis: although using only recall and precision is common in the literature, we find it useful to have a single metric to gauge overall performance, and thus also include F measure.
 
@@
 
-```{r fig.align='center'}
+```{r fig.cap="Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
+For each level of overlap and each level of error, we have 100 paired sets of 500 records."}
 knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
@@
 To demonstrate speed, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-Distributions are meant to emulate the behavior of similarity scores across first name, last name, and day, month, and and year of birth, with 1 indicating an exact match on a particular field and 0 indicating a nonmatch.
-We simulate these data for different values of $n_A$ and $n_B$, and compare the run-time of `parlr` against `BRL`.
+Distributions are meant to emulate the behavior of similarity scores across first name, last name, and day, month.
+For simplicity, we consider only exact matching, so a vector (1, 0) corresponds to agreeement and (0, 1) to disagreement.
+We simulate these data for different values of $n_A$ and $n_B$, and compare the run-time of `fabl` against `BRL`.
 Note that the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in the larger simulations.
 
-We see that at low data size, `BRL` outperforms, but that `parlr` is significantly faster at handling larger data.
-In particular, run-time for `BRL` seems to grow quadratically (or linearly with the size of both $A$ and $B$) while run-time for `parlr` seems to grow linearly (in the size of $B$).
+We see that at low data size, `BRL` outperforms, but that `fabl` is significantly faster at handling larger data.
+In particular, run-time for `BRL` seems to grow quadratically (or linearly with the size of both $A$ and $B$) while run-time for `fabl` seems to grow linearly (in the size of $B$).
 
@@
   \hline
-fname\_0 & 0.05 & 0.99 \\ 
-  fname\_1 & 0.95 & 0.01 \\ 
-  lname\_0 & 0.05 & 0.99 \\ 
-  lname\_1 & 0.95 & 0.01 \\ 
-  day\_0 & 0.05 & 0.97 \\ 
-  day\_1 & 0.95 & 0.03 \\ 
-  month\_0 & 0.05 & 0.92 \\ 
-  month\_1 & 0.95 & 0.08 \\ 
-  year\_0 & 0.05 & 0.93 \\ 
-  year\_1 & 0.95 & 0.07 \\ 
+  fname & (0.95, 0.05) & (0.01, 0.99) \\ 
+  lname & (0.95, 0.05) & (0.01, 0.99) \\ 
+  day & (0.95, 0.05) & (0.03, 0.97) \\ 
+  month & (0.95, 0.05) & (0.08, 0.92) \\ 
+  year & (0.95, 0.05) & (0.07, 0.93) \\ 
    \hline
@@
 
-```{r fig.align='center'}
-knitr::include_graphics("../notes/figures/sadinle_speed_plot.png")
+```{r fig.cap="Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$"}
+knitr::include_graphics("../notes/figures/sadinle_speed_plot2.png")
 ```
@@
 Simulation study suggests that this is true.
-In the plot below, fixing $n_B = 500$, we see linear growth for the run-time under `BRL` as $n_A$ increases, with much more static run-time under `parlr`.
+In the plot below, fixing $n_B = 500$, we see linear growth for the run-time under `BRL` as $n_A$ increases, with much more static run-time under `fabl`.
 The slight increases in run-time that we do see are due primarily to the hashing step, which again can be run in parallel for large data.
 
-```{r fig.align='center'}
+```{r fig.cap="Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500"}
 knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
@@
 
-We note here that `BRL` is coded in C, which makes for unfair comparison against `parlr`, currently only built in R.
-Additionally, although `parlr` is amenable to parallelization, this simulation was run on a single core.
-Running `parlr` in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more drastic results.
+We note here that `BRL` is coded in C, which makes for unfair comparison against `fabl`, currently only built in R.
+Additionally, although `fabl` is amenable to parallelization, this simulation was run on a single core.
+Running `fabl` in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
 
@@
 Lastly, the only fields recorded are given name, last name, date of death, and place of death; it is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-This last point nearly breaks the above mentioned assumption that there are no duplicates within files, and reveals a key difference between `BRL` and our proposed method.
+This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between `BRL` and our proposed method.
 
@@
 Thus we conduct linkage using given name, last name, municipality, and day, month, and year of death.
-We again use flat priors for the $\mathbf{m}$ and $\mathbf{u}$ parameters, but use 2000 Gibbs iterations with 200 discarded as burn-in to mirror the implementation in Sadinle 2017.
+We again use flat priors for the $\mathbf{m}$ and $\mathbf{u}$ parameters.
 
-```{r}
-include_graphics("../notes/figures/el_salvador/overlap_distribution_smallP.png")
-```
-
-```{r}
-include_graphics("../notes/figures/el_salvador/m_posterior_smallP.png")
+```{r fig.cap="Posterior estimates of m parameters with 95\\% credible intervals"}
+knitr::include_graphics("../notes/figures/el_salvador/m_posterior_smallP.png")
 ```
@@
 To mirror the original implementation, we constructed the comparison vectors using 4 levels of agreement for each field, according to the thresholds provided in Figure XX.
-This took ____ seconds for our proposed method and ____ for `BRL`, already demonstrating the increased efficiency of our method.
-However, we note that posterior distributions of several levels of the $\mathbf{m}$ and $\mathbf{u}$ parameters were nearly identical, and that of the $4^5 \times 2 = 2048$ possible agreement patterns, only ___ are realized in the data.
+This took 422 seconds for our proposed method and 240 for `BRL`.
+However, we observed that posterior distributions of several levels of the $\mathbf{m}$ and $\mathbf{u}$ parameters were nearly identical, and that of the $4^5 \times 2 = 2048$ possible agreement patterns, only 1173 are realized in the data.
 This leads us to believe that such high number of aggreement levels creates unnecesary distictions in the data and makes the comparison vectors less interpretable.
 Therefore we re-ran our analysis with fewer agreement levels for each field (see Figure XX), and obtained analogous results.
-With ____ possible agreement patterns, ___ were realized in the data, and our proposed method became much faster, finishing in ___ seconds.
-Meanwhile `BRL` took ___ seconds, relatively unchanged from the first implementation.
+With 216 possible agreement patterns, 159 were realized in the data, and our proposed method became much faster, finishing in 124 seconds.
+Meanwhile `BRL` took 239 seconds, relatively unchanged from the first implementation.
 This demonstrates the way that the computational complexity of our method depends on the number of unique agreement patterns, and how significant computional gains can be made by simplifying the construction of the comparison vectors.
@@
 
+```{r fig.cap="Posterior distribution of overlap across the two files.
+The solid lines show the Bayes estimate for the amonut of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching"}
+knitr::include_graphics("../notes/figures/el_salvador/overlap_distribution_smallP_bayes.png")
+```
+
 Overall, both methods presented similar results.
-`fabl` yielded an initial Bayes estimate of ____ matches found across files, and after resolving matches that violated one-to-one requirements, yeilded ___ matches across files.
-This is acceptably close to the ___ matches found by `BRL`.
+`fabl` yielded an initial Bayes estimate of 195 matches found across files, and after resolving matches that violated one-to-one requirements, yeilded 180 matches across files.
+This is acceptably close to the 181 matches found by `BRL`.
 The main reason for this discrepancy is the difference in how each model handles situations in which one file in $B$ has multiple plausible matches in $A$.
 
-```{r}
-include_graphics("../notes/figures/el_salvador/rosa_records.png")
+```{r fig.cap="Example of linkage situation with multiple plausible matches"}
+knitr::include_graphics("../notes/figures/el_salvador/rosa_records.png")
 ```
 
-```{r}
-include_graphics("../notes/figures/el_salvador/bad_mixing.png")
+```{r fig.cap="Gibbs sampling in situation with multiple plausible matches."}
+knitr::include_graphics("../notes/figures/el_salvador/bad_mixing.png")
 ```
@@
 The records shown in Figure XX provide one such instance.
-Note these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records ___ and ____ in ERTL correspond to different individuals, but their records are nearly identical.
+Note these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals, but their records are nearly identical.
 Additionally, using the modified Levenstein distance from Sadinle 2017, the comparison vectors $\gamma_{2776, 825}$ and $\gamma_{2776, 826}$ are exactly identical.
 
-Figure XX shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrate how each method handles this situation.
-Under `parlr`, both records in $B$ match to the same record in $A$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
+Figure XX shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
+Under `fabl`, both records in $B$ match to the same record in $A$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
 Under `BRL`, the Gibbs process creates one matching configuration stays there for a while.
 However, if one pair "unmatches," then the other record has a chance to latch on.
-Then,  Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
-Additionally, `parlr` allows the modeller to inspect records with mulitple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
+Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
+Additionally, `fabl` allows the modeller to inspect records with mulitple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
 `BRL` in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
@@
 
-In our case study, we included an exploration of how to conduct record linkage when modelling assumptions are not met in practice. We explored "one-to-many" scenarios in which
+In our case study, we included an exploration of how to conduct record linkage when modelling assumptions are not met in practice.
+We explored "one-to-many" scenarios in which one record in $A$ has multiple plausible matches in $B$, and showed how both `fabl` and `BRL` demonstrated undesirable qualities.
+Other issues arise under "many-to-one" scenarios, where one record in $B$ has multiple plausible matches in $A$, and "many-to-many" scenarios in which there is duplication both across and within datasets.
+Tuning `fabl` for use in these scenarious is one potential avenue for future work.

### 2021-06-06 paper/parlr_paper_1.tex
first draft of paper

@@
     \end{flushright}
-    \begin{center} \textbf{Parlr: Parallelized Record Linkage in R}
+    \begin{center} \textbf{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}
 
-    Brian Kundinger
+    Brian Kundinger, Rebecca Steorts, and Jerome Reiter
 
@@
 
-Record linkage is the task of identifying duplicate records across
-multiple data sources. This task is straightforward when provided unique
-identifiers or highly reliable information, but becomes difficult when
-data sources are plagued with error. Additionally, many record linkage
-tasks are large in nature, and many existing methods are too
-computational intensive to be used in practice. In this paper, we
-propose a Bayesian model for record linkage that is amenable to parallel
-computing, hashing techniques and efficient storage mechanisms, that is
-scalable to large linkage tasks.
+Entity resolution (or record linkage) is the task of identifying
+duplicate records across multiple data sources. This task is
+straightforward when provided unique identifiers or highly reliable
+information, but becomes difficult when data sources are plagued with
+error. Additionally, many entity resolution tasks are large in nature,
+and many existing methods are too computational intensive to be used in
+practice. In this paper, we propose a Bayesian model for record linkage
+that uses hashing techniques and efficient storage mechanisms, and is
+amenable to parallelization, making it scalable to large linkage tasks.
 
-The proposed method is inspired by the ``Beta Record Linkage'' method
-proposed by Mauricio Sadinle's 2017 ``Bayesian Estimation of Bipartite
-Matching for Record Linkage''. As such, we demonstrate our method
-through a comparison of our methods against Sadinle's on the simulation
-study he designed and the El Salvador homocide case study he utilized.
-We include one additional simulation study to demonstrate the power of
-our method on large datasets. In what follows, Section 2 describes
-previous work in record linkage that lays the foundation for our method,
-Section 3 provides our model and the Gibbs sampler used for posterior
-inference, and Section 4 describes strategies used for more efficient
-and scalable implementation. Then, Sections 5 and 6 demonstrate the
-method through simulations and a case study. Lastly, our discussion
-highlights open questions in the record linkage liturature and motivates
-future work.
+The proposed method is inspired by the \emph{Beta Record Linkage}
+(\texttt{BRL}) method proposed in Mauricio Sadinle's 2017 ``Bayesian
+Estimation of Bipartite Matching for Record Linkage''. To recognize this
+lineage, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
+pronounced ``fable''). In what follows, Section 2 describes previous
+work in record linkage that lays the foundation for our method, Section
+3 provides our model and the Gibbs sampler used for posterior inference,
+and Section 4 describes strategies used for more efficient and scalable
+implementation. Then, Section 5 demonstrates the accuracy and speed of
+out method through two simulation studies, and Section 6 reveals
+interesting properties of the method through a case study of El
+Salvadoran casuality records. Lastly, our discussion highlights open
+questions in the record linkage liturature and motivates future work.
 
@@
 by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
-characteristic of their model was to transform the sets of records,
+characteristic of their approach was to transform the sets of records,
 which often contain text data that is difficult to model, into sets of
@@
 records respectively, and if the files share \(F\) fields in common upon
-which to base the linkage, the Fellegi and Sunter approach generates a
+which to base the linkage, the Fellegi and Sunter approach generates an
 \(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
@@
 the comparison vector for record \(i \in A\) and record \(j \in B\),
-with \(\gamma_{ij}^f\) providing their similarity score on the
-\(f^{th}\) field. For ease of modeling and computation, we restrict
-these similarity scores to be discrete, ordinal variables, and the
-construction of these is left to the modeler. It is common to use use
-binary 0-1 variables to indicate exact matching, and 0-1-2 variables to
-provide an option for partial matching. For text data, we calculate
-similarity based on Levenstein distance or some other text similarity
-score, and bin these scores to integers for use in the model.
+with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
+score on the \(f^{th}\) field. For ease of modeling and computation, we
+restrict these similarity scores to be discrete, ordinal variables, and
+the construction of these is left to the modeler. We adopt the
+convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
+of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
+is common to use use binary 1-2 variables to indicate exact matching,
+and 1-2-3 variables to provide an option for partial matching. For text
+data, we calculate similarity based on Levenstein distance or some other
+text similarity score, and bin these scores to integers for use in the
+model. Jaro (1989) provided guidelines for appropriate binning for the
+comparison vectors, but details of implementation are generally context
+specific.
 
@@
 nonmatch. This independent classification often leads to sets of matches
-that break expectations of transisitivity, and Jaro devised an procedure
-to reduce a set of conflicting matches to the mostly likely set of
-one-to-one matchings. One key strength of Sadinle's \texttt{BRL} is that
-it is garunteed to produce a set of bipartite matchings that respect
-one-to-one assumptions--without the need for any post-hoc procedures.
-Specifically, in each iteration of his Gibbs sampler, he considers each
-record \(j\in B\), removes from consideration the records \(i\in A\)
-that have already been matched, and then samples a potential link.
-However, accounting for these dependencies throughout the linkage
-process is computationally burdensome, leaving \texttt{BRL} only
-suitable for small to moderate linkage problems.
+that break expectations of transitivity, which is undesirable. Although
+Jaro devised an procedure to reduce a set of conflicting matches to the
+mostly likely set of one-to-one matchings, the abundance of false
+matches can bias the estimation of other parameters of the model.
+
+One key strength of Sadinle's \texttt{BRL} is that it is garunteed to
+produce a set of bipartite matchings that respect one-to-one
+assumptions--without the need for any post-hoc procedures. Specifically,
+in each iteration of his Gibbs sampler, he considers each record
+\(j\in B\), removes from consideration the records \(i\in A\) that have
+already been matched, and then samples a potential link. However,
+accounting for these dependencies throughout the linkage process is
+computationally burdensome, leaving \texttt{BRL} only suitable for small
+to moderate linkage problems.
 
@@
 
-Our notation and assumptions closely follow those of Sadinle. Denote two
-files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records respectively,
-and with records indexed as \(i \in \{1, \ldots, n_A\}\) in \(A\) and
-\(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of generality, label
-the files such that \(n_A \geq n_B\). We also assume there are no
-duplicates within files, only across. For each record pair under
+Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
+respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
+in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
+generality, label the files such that \(n_A \geq n_B\). We assume there
+are no duplicates within files, only across. For each record pair under
 consideration, we generate a comparison vector
@@
 belief that nonmatches and matches are equally likely, and a
-\(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\) prior corresponds to a
-uniform prior on the labeling of \(\mathbf{Z}\).
+\(\lambda \sim \text{Beta}\left(1, \frac{1}{n_A}\right)\) prior
+corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
 
@@
 sampler \emph{inherently serial}. We however use independent priors for
-each \(Z_j\), thereby weakening the one-to-one requirement; our sampler
-does ensure that each record in \(B\) can be matched to at most one
-record in \(A\), but allows for the possiblity that multiple records in
-\(B\) match to the same record in \(A\). We explore the ramifications of
-this distinction with the El Salvador case study. With this change,
-components of the \(\mathbf{Z}\) vector can be computed in parallel,
-allowing for significant computational gains. More importantly, since
-only the agreement pattern of \(Z_j\) is used for calculations within
-the Gibbs sampler, and not the particular record label, we can conduct
-this sampling only at the level of the unique agreement patterns,
-offering even more computational savings. This will be explained further
-in \textcolor{red}{Section XX}.
+each \(Z_j\), creating a sampler that is \emph{perfectly
+parallelizable}, allowing for significant computational gains. More
+importantly, since only the agreement pattern of \(Z_j\) is used for
+calculations within the Gibbs sampler, and not the particular record
+label, we can conduct this sampling only at the level of the unique
+agreement patterns, offering even more computational savings. In doing
+so however, we thereby weaken the one-to-one requirement from
+\texttt{BRL}; our sampler does ensure that each record in \(B\) can be
+matched to at most one record in \(A\), but allows for the possiblity
+that multiple records in \(B\) match to the same record in \(A\). We
+explore the ramifications of this distinction with the El Salvador case
+study.
 
@@
 
-(Adapt from Sadinle)
+We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
+parameter \(\mathbf{Z}\) by assigning different positive losses to
+different types of errors, and minimizing posterior expected loss. We
+adopt the loss function proposed in Sadinle 2017, in which
+\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
+option to leave the matching undetermined by the model. Specifically, we
+minimize the quantity
+\(L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)\)
+where
+
+\[L(\hat{Z_j}, Z_j)=\begin{cases} 
+    0  & \text{if } Z_j = \hat{Z_j}; \\
+    \theta_R,  & \text{if } \hat{Z_j} = R; \\
+    \theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
+    \theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
+    \theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
+\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
+on the linkage status, \(\theta_{10}\) is the loss from a false
+non-match, \(\theta_{01}\) is the loss from a false match, and
+\(\theta_{11}\) is the loss from the special case of a false match in
+which the record has a true match other than the one estimated by the
+model. This loss function leads to closed form decision rules for
+minimizing posterior expected loss. In this paper, we adopt losses
+\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\),
+inducing the intuitive decision rule
+
+\[\hat{Z}_j =\begin{cases} 
+    i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
+    0,  & \text{otherwise} ; \\
+\end{cases}\]
+
+For a more indepth explanation of this function and the induced Bayes
+estimate, see Sadinle 2017.
 
@@
 resolve these conflicts by accepting only the match with highest
-posterior probability. Rigorous proof of this resolution procedure is
-deferred to future work.
+posterior probability. Such a resolution procedure indeed is equivalent
+to minimizing posterior risk under the restriction of one-to-one
+matching, and as thus theoretically justified within the Bayesian
+framework.
 
@@
 \(\mathcal{H}\) be the set of unique agreement patterns in the data, let
-\(P\) denote the total number of unique agreement patterns. Note that
-\(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound
-does not scale with \(n_A\) or \(n_B\). Prior to processing the data, we
-identify all \(P\) patterns in \(\mathcal{H}\) and enumerate them
-\(h_1, \ldots, h_P\). When the \((i,j)\) pair exhibits the \(p^{th}\)
-agreement pattern, we say \((i,j) \in h_p\). Wherever possible, we
-conduct calculations over these \(P\) agreement patterns rather than the
-\(n_A \times n_B\) record pairs.
+\(P = |\mathcal{H}|\) denote the total number of unique agreement
+patterns. Note that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
+that this bound does not scale with \(n_A\) or \(n_B\). Prior to
+processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
+and enumerate them \(h_1, \ldots, h_P\). When the \((i,j)\) pair
+exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
+Wherever possible, we conduct calculations over these \(P\) agreement
+patterns rather than the \(n_A \times n_B\) record pairs.
 
@@
 comparison vector as vector of length \(F\), with each component
-\(\gamma_{ij}^f\) taking on values in \(\{0, \ldots, L_f - 1\}\). To
-ease computations, we instead represent the comparison as a
-concatenation of \(F\) many binary indicator vectors of lengths \(L_f\).
-For example, if \(L_1 = L_2 = 2\) and \(L_3 = 3\), then
-\(\gamma_{ij} = (1, 0, 2)\) under the classical framework becomes
+\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). To ease
+computations, we instead represent the comparison as a concatenation of
+\(F\) many binary indicator vectors of lengths \(L_f\). For example, if
+\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
+under the classical framework becomes
 \(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
@@
 \(n_A n_B \times F\) matrix, with each row providing the comparison
-vector for a different \((i,j)\) pair. We however do not store these
-comparison vectors themselves, but instead only a hashed value
+vector for a different \((i,j)\) pair. In contrast, we do not store
+these comparison vectors themselves, but instead only a hashed value
 corresponding to the agreement pattern of the \((i, j)\) pair. We store
@@
 between record \(j \in B\) and each of the \(n_A\) records in \(A\).
-
-For large data, we can partition the two datasets \(A\) and \(B\) into
-smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more manageable
-computations. On a single machine, we can read-in data sequentially,
-conduct hashing, collect results, and delete the original data from
-memory before continuing with the next chunk of data. With multiple
-cores or multiple machines, this can be done in parallel. Storing this
-hashed information still becomes burdensome for large data, but this
-hashing method greatly expands the capabilities of the Fellegi-Sunter
-framework.
 
@@
 to be deemed a match through the Bayes estimate, even without
-considering the form of the agreement pattern itself. In random
-indexing, rather than store all of these unlikely record labels, we
-choose to store only a small number \(R\) of them. Posterior
-calculations still attribute the appropriate weight to all records
-through the summary statistics \(H_{p_j}, H_p^m\), and \(H_p^u\). Rather
-than storing \(n_A \times n_B\) record labels, SEI allows us to store at
-most \(n_B \times P \times R\) labels, regardless of how large \(n_A\)
-is. Coupled with the partitioning strategy for creating comparison
-vectors, this greatly expands the size of the linkage tasks we can
-undertake.
+considering the form of the agreement pattern itself.
+
+In SEI, rather than store all of these unlikely record labels, we choose
+to store only a small number \(R\) of them. Posterior calculations still
+attribute the appropriate weight to all records through the summary
+statistics \(H_p\), and \(H_{p_j}\). Rather than storing
+\(n_A \times n_B\) record labels, SEI allows us to store at most
+\(n_B \times P \times R\) labels, regardless of how large \(n_A\) is.
+
+Laslty, for large data, we can partition the two datasets \(A\) and
+\(B\) into smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more
+manageable computations. On a single machine, we can read-in data
+sequentially, conduct hashing, compress information through SEI, and
+delete the original data from memory before continuing with the next
+chunk of data. With multiple cores or multiple machines, this can be
+done in parallel. Thus the combination of hashing, SEI, and partitioning
+allows us to conduct linkage tasks over much larger datasets.
 
@@
 computation, we break this sampling step into two simpler steps. First,
-we calculate the Fellegi Sunter weight \(w_{h_p}\) associated with each
+we calculate the Fellegi Sunter weight \(w_{p}\) associated with each
 unique pattern and sample the agreement pattern between \(j\) and its
@@
 
-\[h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
+\[P\left(h\left(Z_j^{(s+1)}\right) = p | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
 \begin{cases} 
-    w_{h_p}\times H_{p_j}  & p \leq P; \\
+    w_{p}\times H_{p_j}  & p \leq P; \\
      n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
+\end{cases}\]
+\[P\left(Z_j^{(s+1)} = i\bigg|\;\; h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
+\frac{1}{H_{p_j}} & (i, j) \in h_p \\
+0 & \text{otherwise} \\
 \end{cases}\]
@@
 agreement patterns of the record pairs rather than the record labels
-themselves. Thuse We complete the entire Gibbs procedure at the level of
-the \(P\) agreement patterns. After, we can back-fill the records
-corresponding to the agreement patterns by sampling uniformly at random
-among candidate records stored in \(\tilde{\Gamma}\). Sampling uniformly
-is computational simple even for large sets of candidate records, but
-this step can also be parallelized when working with large data.
+themselves. Thus we complete the entire Gibbs procedure first at the
+level of the \(P\) agreement patterns with the first equation above.
+After, we can back-fill the records corresponding to the agreement
+patterns through the second equation. Sampling uniformly is
+computational simple even for large sets of candidate records, but this
+step can also be parallelized when working with large data.
 
@@
 although using only recall and precision is common in the literature, we
-find it useful to have a single metric the gauge overall performance,
-and thus also include F measure.
+find it useful to have a single metric to gauge overall performance, and
+thus also include F measure.
 
@@
 
-\begin{center}\includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot} \end{center}
+\begin{figure}
+
+{\centering \includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot}
+
+}
+
+\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
+For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:unnamed-chunk-1}
+\end{figure}
 
@@
 problem. Distributions are meant to emulate the behavior of similarity
-scores across first name, last name, and day, month, and and year of
-birth, with 1 indicating an exact match on a particular field and 0
-indicating a nonmatch. We simulate these data for different values of
-\(n_A\) and \(n_B\), and compare the run-time of \texttt{parlr} against
-\texttt{BRL}. Note that the number of unique patterns \(P\) is bounded
-above by \(2^5 = 32\), a bound which is consistently attained in the
-larger simulations.
+scores across first name, last name, and day, month. For simplicity, we
+consider only exact matching, so a vector (1, 0) corresponds to
+agreeement and (0, 1) to disagreement. We simulate these data for
+different values of \(n_A\) and \(n_B\), and compare the run-time of
+\texttt{fabl} against \texttt{BRL}. Note that the number of unique
+patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is
+consistently attained in the larger simulations.
 
 We see that at low data size, \texttt{BRL} outperforms, but that
-\texttt{parlr} is significantly faster at handling larger data. In
+\texttt{fabl} is significantly faster at handling larger data. In
 particular, run-time for \texttt{BRL} seems to grow quadratically (or
 linearly with the size of both \(A\) and \(B\)) while run-time for
-\texttt{parlr} seems to grow linearly (in the size of \(B\)).
+\texttt{fabl} seems to grow linearly (in the size of \(B\)).
 
@@
 
-\begin{center}\includegraphics[width=23.6in]{../notes/figures/sadinle_speed_plot} \end{center}
+\begin{figure}
+
+{\centering \includegraphics[width=17.75in]{../notes/figures/sadinle_speed_plot2}
+
+}
+
+\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:unnamed-chunk-2}
+\end{figure}
 
@@
 see linear growth for the run-time under \texttt{BRL} as \(n_A\)
-increases, with much more static run-time under \texttt{parlr}. The
+increases, with much more static run-time under \texttt{fabl}. The
 slight increases in run-time that we do see are due primarily to the
@@
 
-\begin{center}\includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB} \end{center}
+\begin{figure}
+
+{\centering \includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB}
+
+}
+
+\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:unnamed-chunk-3}
+\end{figure}
 
 We note here that \texttt{BRL} is coded in C, which makes for unfair
-comparison against \texttt{parlr}, currently only built in R.
-Additionally, although \texttt{parlr} is amenable to parallelization,
-this simulation was run on a single core. Running \texttt{parlr} in C++
+comparison against \texttt{fabl}, currently only built in R.
+Additionally, although \texttt{fabl} is amenable to parallelization,
+this simulation was run on a single core. Running \texttt{fabl} in C++
 with paralellization for the hashing step and sampling the matching
-status of the record pairs should lead to even more drastic results.
+status of the record pairs should lead to even more computational gains.
 
@@
 
-the country of El Salvador was immersed in civil war from 1980 to 1991,
+The country of El Salvador was immersed in civil war from 1980 to 1991,
 and throughout the time, several organizations attempted to document
@@
 indistinguishable records for two different individuals. This last point
-nearly breaks the above mentioned assumption that there are no
+nearly breaks the earlier mentioned assumption that there are no
 duplicates within files, and reveals a key difference between
@@
 municipality, and day, month, and year of death. We again use flat
-priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters, but use
-2000 Gibbs iterations with 200 discarded as burn-in to mirror the
-implementation in Sadinle 2017.
+priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters.
+
+\begin{figure}
+
+{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/m_posterior_smallP}
+
+}
+
+\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+To mirror the original implementation, we constructed the comparison
+vectors using 4 levels of agreement for each field, according to the
+thresholds provided in Figure XX. This took 422 seconds for our proposed
+method and 240 for \texttt{BRL}. However, we observed that posterior
+distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
+parameters were nearly identical, and that of the
+\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
+realized in the data. This leads us to believe that such high number of
+aggreement levels creates unnecesary distictions in the data and makes
+the comparison vectors less interpretable. Therefore we re-ran our
+analysis with fewer agreement levels for each field (see Figure XX), and
+obtained analogous results. With 216 possible agreement patterns, 159
+were realized in the data, and our proposed method became much faster,
+finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
+relatively unchanged from the first implementation. This demonstrates
+the way that the computational complexity of our method depends on the
+number of unique agreement patterns, and how significant computional
+gains can be made by simplifying the construction of the comparison
+vectors. We also note that estimates of the \(\mathbf{m}\) and
+\(\mathbf{u}\) paramters under each method are very similar, as shown in
+Figure XX.
+
+\hypertarget{violations-of-one-to-one-matching}{%
+\subsection{Violations of One-to-One
+Matching}\label{violations-of-one-to-one-matching}}
+
+Figure XX shows the posterior distribution for \(D\), the number of
+duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
+see that \texttt{fabl} consistently overmatches within each Gibbs
+iteration when compared to \texttt{BRL}, which is to be expected because
+\texttt{BRL} explicitly prevents matches that violate one-to-one
+matching throughout the entire sampler. Most of such matchings are due
+to the randomness in the sampling procedure, and they occur sporadically
+throughout the sampler in such a way that does not measurably influence
+the eventual Bayes estimate.
+
+\begin{figure}
+
+{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+
+}
+
+\caption{Posterior distribution of overlap across the two files.
+The solid lines show the Bayes estimate for the amonut of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:unnamed-chunk-5}
+\end{figure}
+
+Overall, both methods presented similar results. \texttt{fabl} yielded
+an initial Bayes estimate of 195 matches found across files, and after
+resolving matches that violated one-to-one requirements, yeilded 180
+matches across files. This is acceptably close to the 181 matches found
+by \texttt{BRL}. The main reason for this discrepancy is the difference
+in how each model handles situations in which one file in \(B\) has
+multiple plausible matches in \(A\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=6.51in]{../notes/figures/el_salvador/rosa_records}
+
+}
+
+\caption{Example of linkage situation with multiple plausible matches}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/bad_mixing}
+
+}
+
+\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:unnamed-chunk-7}
+\end{figure}
+
+The records shown in Figure XX provide one such instance. Note these
+records present a near violation of our assumption that there are no
+duplications within files; we continue to assume that records 825 and
+826 in ERTL correspond to different individuals, but their records are
+nearly identical. Additionally, using the modified Levenstein distance
+from Sadinle 2017, the comparison vectors \(\gamma_{2776, 825}\) and
+\(\gamma_{2776, 826}\) are exactly identical.
+
+Figure XX shows the values that \(Z_{825}\) and \(Z_{826}\) take on
+throughout the Gibbs sampler, and demonstrates how each method handles
+this situation. Under \texttt{fabl}, both records in \(B\) match to the
+same record in \(A\) throughout the Gibbs process, creating consistent
+violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
+creates one matching configuration stays there for a while. However, if
+one pair ``unmatches,'' then the other record has a chance to latch on.
+Then, the Gibbs process is stuck with that matching status for a while,
+resulting in a Gibbs process with poor mixing. Additionally,
+\texttt{fabl} allows the modeller to inspect records with mulitple
+plausible matches, and if they desire, to then choose the record pairing
+with the highest posterior probability. \texttt{BRL} in contrast, in
+strictly enforcing one-to-one matching throughout the sampler, can lead
+to situations where none of the plausible matches reach the threshold to
+be identified through the Bayes estimate.
 
@@
 
+We have presented a method for Bayesian record linkage that is feasible
+for large datasets. In particular, our hashing procedure and model
+assumptions allow for a linkage procedure whose computational complexity
+does not scale with the size of the larger dataset, making this method
+particularly powerful in linking records when one datafile is
+substantially smaller than the other.
+
+In our case study, we included an exploration of how to conduct record
+linkage when modelling assumptions are not met in practice. We explored
+``one-to-many'' scenarios in which one record in \(A\) has multiple
+plausible matches in \(B\), and showed how both \texttt{fabl} and
+\texttt{BRL} demonstrated undesirable qualities. Other issues arise
+under ``many-to-one'' scenarios, where one record in \(B\) has multiple
+plausible matches in \(A\), and ``many-to-many'' scenarios in which
+there is duplication both across and within datasets. Tuning
+\texttt{fabl} for use in these scenarious is one potential avenue for
+future work.
+
 \end{document}

### 2021-06-07 template-applied/aoas_revision_2017_09_18_FINAL_rcs_ANSHU_FINAL.tex
adding two latex templates.

\documentclass[aoas]{imsart}
%\documentclass{imsart}

\usepackage{natbib}
\usepackage{comment}

\usepackage{amsmath,bm,amsfonts,color,amsthm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{subcaption}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{float}
\restylefloat{table}
\usepackage[switch]{lineno}
\linenumbers
\input{symbols_journal}
\usepackage{booktabs}

\newcounter{itemnum}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\usepackage{mathtools}	
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[table,x11names]{xcolor}

%\usepackage[utf]{arabxetex}
%
%\newfontfamily\arabicfont[Script=Arabic]{Amiri}
%\SetKw{KwSet}{Set}

%\usepackage{tikz}
%\usetikzlibrary{decorations.pathmorphing} % noisy shapes
%\usetikzlibrary{fit}         % fitting shapes to coordinates
%\usetikzlibrary{backgrounds} % drawing the background after the
%\usetikzlibrary{positioning}
%\usetikzlibrary{shadows}
%\usepgflibrary{shapes}
%\tikzstyle{observed}=[circle, thick, minimum size=0.7cm, draw=black!100, fill=black!20]
%\tikzstyle{latent}=[circle, thick, minimum size=0.7cm, draw=black!100]
%\tikzstyle{plate}=[rectangle, thick, inner sep=0.3cm, draw=black!100]
%\tikzstyle{shadeplate}=[rectangle, thick, inner sep=0.3cm, draw=black!100, fill=black!10]

\usepackage[switch]{lineno}
%\usepackage{arabtex}
%\usepackage{utf8}
%\usepackage{fontspec}
%\newfontfamily{\F}{Geeza Pro} % Farsi
\linenumbers
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
%\usepackage[ruled,lined]{algorithm2e}
%\SetKw{KwSet}{Set}
%\usepackage[numbers,square]{natbib}
%\usepackage[left=1in,top=1in,right=1in,bottom=1in,nohead]{geometry}
%\usepackage[all]{xy}
%\usepackage{pdfsync}
%\usepackage{setspace}
%\usepackage{hyperref}
%\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}

% Bibliography
%\bibliographystyle{mybibsty}

%\def\algorithmautorefname{Algorithm}

%\graphicspath{{images/}}

\begin{document}
\begin{frontmatter}
\title{ Unique Entity Estimation with Application to the Syrian Conflict}
\runtitle{Entity Estimation with Application to the Syrian Conflict}
\begin{aug}
  \author{\fnms{Beidi} \snm{Chen}\ead[label=e1]{beidi.chen@rice.edu}},
  \author{\fnms{Anshumali} \snm{Shrivastava}\ead[label=e2]{anshumali@rice.edu}}, \and
    \author{\fnms{Rebecca C.} \snm{Steorts}\corref{}\ead[label=e3]{beka@stat.duke.edu}}
  \runauthor{Chen, Shrivastava, and Steorts}
    \address{Department of Computer Science \\
   Rice University, Houston, TX USA\\ \printead{e1,e2}}
      \address{Department of Statistical Science and Computer Science\\ Duke University \\
  Durham, NC USA\\ \printead{e3}}
\end{aug}

\begin{abstract}
Our work is motivated to estimate the number of documented identifiable deaths in the ongoing Syrian conflict.
Entity resolution identifies and removes duplicate entities in large, noisy databases and has grown in both usage and new developments as a result of increased data availability. Nevertheless, entity resolution has tradeoffs regarding assumptions of the data generation process, error rates, and computational scalability that make it a difficult task for real applications. In this paper, we focus on a related problem of unique entity estimation, which is the task of estimating the unique number of entities and associated standard errors in a data set with duplicate entities. Unique entity estimation shares many fundamental challenges of entity resolution, namely, that the computational cost of all-to-all entity comparisons is intractable for large databases. To circumvent this computational barrier, we propose an efficient (near-linear time) estimation algorithm based on locality sensitive hashing. Our estimator, under realistic assumptions, is unbiased and has provably low variance compared to existing random sampling based approaches. In addition, we empirically show its superiority over the state-of-the-art estimators on three real applications. Furthermore, our methodology, when applied to the Syrian data set, provides an estimate of $190,369 \pm 207$ casualties, which is very close to the Human Rights Data Analysis Group (HRDAG) estimate of 190,000. Our work provides an example of challenges and efforts involved in solving a real noisy challenging problem where modeling assumptions may not hold.
\end{abstract}

\begin{keyword}
\kwd{Syrian conflict}
\kwd{unique entity estimation}
\kwd{entity resolution}
\kwd{clustering}
\kwd{dimension reduction}
\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{sec:intro}

Our work is motivated by a real estimation problem associated with the ongoing conflict in Syria. While tremendously well documented, it is hard to know how many individuals have been killed from conflict-related violence in Syria. Since March 2011, increasing reports of deaths have appeared in both the national and international news. There are many inconsistencies from various media sources, which is inherent from the data collection process and the fact that reported victims are documented from multiple sources. Consequently our ultimate goal is to find a death estimate (with associated standard errors) because such information may contribute to future transitional justice and accountability measures. For instance, statistical estimates of death counts have been introduced as evidence in national court cases and international tribunals investigating the responsibility of state leaders for crimes against humanity \citep{hrdag-2017}. The task shows a classical example of noisy and unreliable information in big data, which often brings a new set of challenges for statistical estimation.

The main challenge with reliable death estimation of the Syrian data set is the fact that individuals that are documented as dead are often duplicated in the data sets. Due to the task at hand, one could turn to entity resolution  (de-duplication or record linkage), which refers to the task of removing duplicated records in noisy datasets that refer to the same entity  \citep{liseo_2011, sadinle_2014, getoor_2006, gutman_2013,  McCallumWellner04, fellegi_1969}. Entity resolution has become intrinsically important given it is fundamental in many large data processing applications. Informally, let us assume that each entity (records) is a vector in $\mathbb{R}^D$. Then given a data set of $M$ records aggregated from many data sources with possibly numerous duplicated entities perturbed by noise, the task of entity resolution is to identify and remove the duplicate entities. For a review of entity resolution see \citep{winkler_2006, christen_2012, liseo2013some}.

One important subtask of entity resolution is estimating the number of unique entities (records) $n$ out of $M > n$ duplicated entities, which we call \emph{unique entity estimation}. Entity resolution is a more difficult problem because it requires one to link each entity to its associated duplicate entities. To obtain high accuracy entity resolution, the algorithms must at least evaluate a significant amount of pairs for potential duplicates to ensure a link is not missed. Due to this and to the best of our knowledge, accurate entity resolution algorithms scale quadratically or higher ($> O(M^2))$ making them computationally intractable for large data sets.  Reducing the computational cost in entity resolution is known as blocking, which via deterministic or probabilistic algorithms places similar records into blocks or bins \citep{christen_2012, steorts14comparison}. The computational efficiency comes at the cost of missed links and reduced accuracy for entity resolution. It is further not clear if we can use these crude but cheap entity resolution sub-routines for unbiased estimation of unique entities with strong statistical guarantees.

%{\bf Our Focus:}
The primary focus of this paper is on a \emph{unique entity estimation} algorithm that is motivated by the ongoing conflict in Syria which has the following properties:
\begin{enumerate}
  \item The estimation cost should be significantly less than quadratic ($O(M^2)$). In particular, any methodology requiring one to evaluate all pairs for linkage is not suitable. This is crucial for the Syrian data set and other large, noisy data sets (Section~\ref{sec:SyriaChallenges}).
  \item To ensure accountability regarding uniquely estimating the number of documented identifiable victims in the Syrian conflict, it is essential to understand the statistical properties of any proposed estimator. Such a requirement eliminates many heuristics and rule-based entity resolution tasks, where the estimates may be very far from the true value.
  \item  In most real entity resolution tasks, duplicated data can occur with arbitrary large changes including missing information, which we observe in the Syrian data set, and standard modeling assumptions may not hold due to the noise inherent in the data. Due to this, we prefer to not make strong modeling assumptions regarding the data generation process.
\end{enumerate}

\subsection{Related Work for Unique Entity Estimation}
\label{sec:RelatedWork}

The three aforementioned properties that we desire eliminates all but random sampling based approaches. In this section, we review them briefly.

To our knowledge, only two random sampling based methodologies satisfy such requirements. \cite{1978paper} proposed sampling a large enough subgraph to estimate the total number of connected components based on the properties of the sub-sampled subgraph. Also, \cite{chazelle2005approximating} proposed finding connected components with high probability by sampling random vertices and then visiting their associated components using breadth-first search (BFS). One major issue with random sampling is that most sampled pairs are unlikely to be matches (no edge) providing nearly no information, as the underlying graph is generally very sparse in practice. Randomly sampling vertices and running BFS required by~\cite{chazelle2005approximating} is very likely to result in singleton vertices because many records are themselves unique in entity resolution data sets. In addition, finding all possible connections of a given vertex would require $O(M)$ query for edges. A query for edges corresponds  to the query for actual link between two records. Sub-sampling a sub-graph, as in~\cite{1978paper}, of size $s$ requires $O(s^2)$ edge queries to completely observe it. Thus, $s$ should be reasonably small in order to scale. Unfortunately, requiring a small $s$ hurts the variance of the estimator. We show that the accuracy of both aforementioned methodologies is similar to the non-adaptive variant of our estimator which has provably large variance. In addition, we show both theoretically and empirically that the methodologies based on random sampling lead to poor estimators.

While some methods have recently been proposed for accurate estimation of unique records, they belong to the Bayesian literature and have difficulty scaling due to the curse of dimensionality with Markov chain Monte Carlo \cite{steorts??bayesian, sadinle_2014, liseo_2011}. The evaluation of the likelihood itself is quadratic. Furthermore, they rely on a strong assumption about the specified generative models for the duplicate records. Given such computational challenges with the current state of the methods in the literature, we take a simple approach, especially given the large and constantly growing  data sets that we seek to analyze. Our focus in on practical methodologies which can easily scale to large volume of data sets with minimal assumptions. Specifically, we propose a unique entity estimation algorithm with sub-quadratic cost, which can be reduced to approximating the number of connected components in a graph with sub-quadratic queries for edges (Section~\ref{form}).

The rest of the paper proceeds as follows. Section \ref{syriadata} provides our motivational application from the Syrian conflict 
and Section \ref{sec:SyriaChallenges} remarks on the main challenges of the Syrian data set and our proposed methodology. Section \ref{sec:lsh} provides background on variants of locality sensitive hashing (LSH), which is paramount to our proposed methodology.  Section \ref{proposal} provides our proposed methodology for unique entity estimation, which is the first formalism of using efficient adaptive LSH on edges to estimate the connected components with sub-quadratic computational time. (An example of our approach is given in section \ref{sec:toy}). More specifically, we draw connections to our methodology to random and adaptive sampling in section \ref{analysis}, where we show under realistic assumptions that our estimator is theoretically unbiased and has provably low variance. In addition, in section \ref{sec:randVsAdaptive}, we compare random and adaptive sampling for the Syrian data set, illustrating the strengths of adaptive sampling. We introduce the variant of LSH that we use in section \ref{sec:missingVariant}. Finally, we provide our complete algorithm for unique entity estimation in section \ref{putit}. 
Section \ref{experiments} provides evaluations of all the related estimation methods on three real data sets from the music and food industry as well as official statistics; section \ref{syria} reports our estimation of the number of causalities in the Syrian conflict.

\subsection{The Syrian Conflict}
\label{syriadata}
Thanks to Human Rights Data Analysis Group (HRDAG), we have access to four databases from the Syrian conflict which roughly cover the same period, namely March 2011 -- November 2013. These respectively come from the Violation Documentation Centre (VDC), Syrian Center for Statistics and Research (CSR-SY), Syrian Network for Human Rights (SNHR), and Syria Shuhada website (SS). Each database lists a different number of recorded victims killed in the Syrian conflict, along with available identifying information including full Arabic name, date of death, death location, gender, among others.\footnote{These databases include documented identifiable victims and not those who are missing in the conflict, hence, any estimate reported only refers to the data at hand.}

%This list does not contain individuals that are missing, and a victim is listed as killed if they are identifiable.

Since the above information is collected via indirect ways such as through friends and religious leaders, or traditional media resources, it naturally comes with many challenges. The data set has biases, spelling errors, and missing values. In addition, it is well known that there are duplicate entities present in the data sets, making estimation more difficult.
%In addition, it is well known that many records refer to the same entity (individual), a problem known as entity resolution (record linkage or de-duplication).
The ambiguities in Arabic names make the situation significantly worse as there can be a large textual difference between the full and short names in Arabic.

Such ambiguities and lack of any additional information make entity resolution (ER) on this data set considerably challenging~\citep{price2014updated}. Owing to the significance of the problem, HRDAG has provided labels for a large subset of the data set. More specifically, four different human experts from the HRDAG manually reviewed pairs of records in the four data sets, classifying them as matches if referred to the same entity and non-matches otherwise. {\em Our first {\bf goal} is to accurately estimate the number of unique victims.} Obtaining a match or non-match label of a given record pair may require momentous cost such as manual human supervision or involving sophisticated machine learning. Given that coming up with hand-matched data is a costly process,  {\em our second {\bf goal}} is providing a proxy, automated mechanism to create labeled data. (More information regarding the Syrian data set can be found in Appendix \ref{sec:syrian}).

\subsection{Challenges and Proposed Solutions}
\label{sec:SyriaChallenges}
%Evaluating all pairs for match/mismatch seems to solve the problem.
Consider evaluating the Syrian data set using all-to-all records comparisons to remove duplicate entities. With approximately 300,000 records from the Syrian data set, we have around 45 billion pairs ($4.5 \times 10^{10}$). Therefore, it is impractical to classify all these pairs as matches/non-matches reliably. We cannot expect few expert (four in our case) to manually label 45 billion pairs. Human supervision cannot be tremendously escalated. A simple computation of all pairwise similarity (45 billion) takes more than 6 days on a heavyweight machine that can run 56 threads in parallel (28 cores in total). In general, this quadratic computational cost is widely considered infeasible for large data sets. Algorithmic labeling of every pair, even if possible for relatively small datasets, is neither reliable nor efficient. Furthermore, it is hard to understand the statistical properties of algorithmic labelling of pairs. Such challenges, therefore, motivates us to focus on the estimation algorithm with constraints mentioned in Section~\ref{sec:intro}.

{\bf Our Contributions:} We formalize unique entity estimation as approximating the number of connected components in a graph with sub-quadratic $\ll O(M^2)$ computational time. We then propose a generic methodology that provides an estimate in sample (with standard errors). Our proposal leverages locality sensitive hashing (LSH), in a novel way, for the estimation process with the required computational complexity, that is less than quadratic. Our proposed estimator is unbiased and has provably low variance compared to random sampling based approaches. To the best of our knowledge this is the first use of locality sensitive hashing for unique entity estimation in record linkage setting.
%
More broadly, our unique entity estimation procedure is applicable to many applications, and we illustrate this on three additional real, fully labelled, entity resolution data sets, which include the food industry, the music industry, and an application from official statistics. In the absence of ground truth information, We give our best efforts estimation for the Syrian conflict, where we show an estimation of 190,369, with standard deviation of 207, reported casualties, which is very close to the 2014 HRDAG estimate of 190,000. This clearly demonstrates the power of our efficient estimator in practice which does not rely on any strong modelling assumptions. Out of 45 billion possible pairs, our estimator only queries around 1 million adaptively sampled pairs for labels, yielding a 450\% reduction. The labelling was done using support vector machines (SVMs) trained on a small hand matched labelled examples provided by four domain experts. Our work is an example of the efforts required to solve a real noisy challenging problem where modeling assumptions may not hold.

%Our contributions in this paper are four-fold. First, we propose a methodology that provides an estimate in sample (with standard errors) using adaptive locality sensitive hashing, which has computational complexity that is less than quadratic. Second, we propose an estimator for unique entity estimation that is unbiased and has provably low variance. Third, in the absence of ground truth information, our methodology suggests that it can provide a proxy for such labels. Finally, we illustrate the success of our framework on a case study from the Syrian conflict in addition to three other real data sets, making comparisons to the current state-of-the-art in the literature.

\section{Variants of Locality Sensitive Hashing (LSH)}
\label{sec:lsh}
%\textcolor{red}{Redone: AS please check. I have tried to make it clear, accurate and minimal!}
In this section, we first provide a review of locality sensitive hashing (LSH) and min-wise hashing, which is crucial to our proposed methodology. We then introduce a variant of LSH --- Densified One Permutation Hashing (DOPH), which is paramount to our proposed algorithm for unique entity estimation in terms of scalability.

\subsection{Shingling}
%In many entity resolution tasks, one often transforms the set of all records into sets by a shingle or gram representation\citep{rajaraman_2012}.  That is, each record is treated as a string and is replaced by a ``bag'' (or ``multi-set'') of length-$k$ contiguous sub-strings that it contains. This representation is commonly referred to as a k-gram or k-shingle of the record string. As an example, the string ``TORONTO'' yields the bag of length-two shingles ``TO'', ``OR'', ``RO'', ``ON'', ``NT'', ``TO'', where ``TO" appears twice. To better preserve the structure of a string or the grammar, one might use a bag of words (BoW) representation, in which one counts the number of times each words appears in the string. Finally, one might use a combination of shingling and a bags of words representation. In this, approach, one shingles words (instead of characters) in a string to try and preserve the grammatical structure.

In entity resolution tasks, each record can be represented as a string of information. For example, record in the Syrian data set can be represented as a short \textit{text} description of the person who died in the conflict. In this paper, we use a k-grams based shingle representation, which is the most common representation of text data and naturally gives a set token (or k-grams).
That is, each record is treated as a string and is replaced by a ``bag'' (or ``multi-set'') of length-$k$ contiguous sub-strings that it contains. Since we will use a k-gram based approach to transform the records, our representation of each record will also be a set, which consist of all the $k$-contiguous characters occurring in record string. As an illustration, for the record \text{BAKER, TED}, we separate it into a 2-gram representation. The resulting set is the following: $$\text{BA, AK, KE, ER, ER, TE, ED}.$$ In another example, consider \text{Sammy, Smith}, whose 2-gram set representation is $$\text{SA, AM, MM, MY, MS, SM, MI, IT, TH}.$$ We now have two records that have been transformed into a 2-gram representation. Thus, for every record (string) we obtain a set $\subset \mathcal{U}$, where the universe $\mathcal{U}$ is the set of all possible $k$-contiguous characters.
% k-grams based shingle representation is the most common representation of text data, which naturally gives a set token (or k-grams). Since we will use a k-gram based approach to transform the records, our representation of each record will also be a set. The set will consist of all the $k$-contiguous characters occurring in record string. As an illustration, for the record \text{BAKER, TED}, we separate it into a 2-gram representation. The resulting set is the following: $$\text{BA, AK, KE, ER, ER, TE, ED}.$$ In another example, consider \text{Sammy, Smith}, whose 2-gram set representation is $$\text{SA, AM, MM, MY, MS, SM, MI, IT, TH}.$$ We now have two records that have been transformed into a 2-gram representation. Thus, for every record (string) we obtain a set $\subset \mathcal{U}$, where the universe $\mathcal{U}$ is the set of all possible $k$-contiguous characters.

\subsection{Locality Sensitive Hashing}
LSH---a two-decade old probabilistic technique and method for dimension reduction---comes with sound mathematical formalism and guarantees. LSH is widely used in computer science and database engineering as a way of rapidly finding approximate nearest neighbors \citep{Proc:Indyk_STOC98,gionis_1999}. Specifically, the variant of LSH that we utilize is scalable to large databases, and allows for similarity based sampling of entities in less than a quadratic amount of time.

In LSH, a hash function is defined as $y = h(x),$ where $y$ is the \emph{hash code} and $h(\cdot)$ the \emph{hash function}. A \emph{hash table} is a data structure that is composed of \emph{buckets} (not to be confused with blocks), each of which is indexed by a \emph{hash code}. Each reference item $x$ is placed into a bucket $h(x).$

More precisely, LSH is a family of function that map vectors to a discrete set, namely, $h:\mathbb{R}^D \rightarrow \{1, \ 2,\cdots , M\}$, where $M$ is in finite range. Given this family of functions, similar points (entities) are likely to have the same hash value compared to dissimilar points (entities). The notion of similarity is specified by comparing two vectors of points (entities), $x$ and $y.$ We will denote a general notion of similarity by $\text{SIM}({x,y}).$ In this paper, we only require a relaxed version LSH, and we define this below. Formally, a LSH is defined by the following definition below:
%LSH are family of functions mapping vectors to a discrete set, i.e., $h:\mathbb{R}^D \rightarrow \{1, \ 2,\cdot , M\}$ where $M$ is in finite range. In this family, similar points are likely to have the same hash value compare to dissimilar points. The notion of similarity is specified.  The exact definition is complicated and involved; we give a relaxed version of the definition for the sake of clarity.
\begin{definition} (Locality Sensitive Hashing (LSH))\ Under a given similarity metrix $\text{SIM}(x,y)$, a family $\mathcal{H}$ is locality sensitive if for any four points $x_1, \ x_2, \ y_1, \ y_2 \in \mathbb{R}^D$  and $h$ chosen uniformly from $\mathcal{H}$ satisfies the following condition: if $\text{SIM}(x_1,x_2)\ge Sim(y_2,y_3)$ then ${Pr}_\mathcal{H}(h(x_1) = h(x_2)) \ge {Pr}_\mathcal{H}(h(y_1) = h(y_2)),$ where ${Pr}_\mathcal{H}$ is the probability over the uniform sampling of $h$.
\end{definition}
The above definition is sufficient condition for a family of function to be LSH. While many popular LSH families satisfy the aforementioned property, we only require this condition for the paper. For a complete review of LSH, we refer to \cite{rajaraman_2012}.
% Many popular LSH families satisfy the above property. We only need this condition for this paper.
%More recently, locality sensitive hashing has been utilized as a form of blocking in entity resolution, where
%one tries to achieve scalability and avoid all-to-all record comparisons by partition records into ``partitions" or ``blocks" either using deterministic or probabilistic methods [[CITE Christen, Herzog]]. However, the theoretical understanding concerning the effect of LSH on the record linkage problem is missing to a large extent.

\subsection{Minhashing}
One of the most popular forms of LSH is minhashing~\cite{Proc:Broder}, which has two key properties --- a type of similarity and a type of dimension reduction. The type of similarity the is used is the Jaccard similarity and the type of dimension reduction is known as the minwise hash, which we now define.

Let $\{0,1\}^D$ denote the set of all binary $D$ dimensional vectors, while $\mathbb{R}^D$ refers to the set of all $D$ dimensional vectors (of records). Note that records can be represented as a binary vector (or set) representation via shingling, BoW, or combining these two methods. More specifically, given two record sets (or equivalently binary vectors) $x,y \in \{0,1\}^D,$ the Jaccard similarity between $x, y \in \{0,1\}^D$  is\begin{align*}
\mathcal{J} = \frac{|x \cap y|}{|x \cup y|},
%\footnote{Minwise hashing is the popular locality sensitive hashing (LSH) for Jaccard similarity.}.
\end{align*}
where $|\cdot|$ is the cardinality of the set.

More specifically, the minwise hashing family applies a random permutation $\pi$, on the given set $S$, and stores only the minimum value after the permutation mapping, known as the \emph{minhash}.  Formally, the minhash is defined as $h_{\pi}^{min}(S) = \min(\pi(S))$, where $h(\cdot)$ is a hash function.

Given two sets $S_1$ and $S_2$, it can be shown by an elementary probability argument that
\begin{equation}
\label{eq:MinHash}
Pr_{\pi}({h_{\pi}^{min}(S_1) = h_{\pi}^{min}(S_2)) =  \frac{|S_1 \cap S_2|}{| S_1 \cup S_2|}},
% \mathcal{J}=\frac{a}{f_1+f_2 -a},
\end{equation}
where the probability is over uniform sampling of $\pi$. It follows from Equation~\ref{eq:MinHash} that minhashing is a LSH family for the Jaccard similarity.

\textbf{Remark}: In this paper, we utilize a shingling based approach, and thus, our representation of each record is likely to be very sparse. Moreover, \cite{shrivastava2014defense} showed that minhashing based approaches are superior than the random projection based approaches for very sparse datasets.

\subsubsection{Densified One Permutation Hashing (DOPH)}
For realistic sized entity resolution tasks, sampling based on LSH requires hundreds of hashes (Section~\ref{sec:KLParametrizedLSH}). It is well known that computing several minwise hashes of data is a very costly operation~\citep{li2012gpu}. Fortunately, recent literature on Densified One Permutation Hashing (DOPH) has shown that it is possible to compute several, hundreds of even thousands, hashes of the data vector in one pass with nearly identical statistical properties as minwise hashes \citep{shrivastava2014densifying,shrivastava2014improved,shrivastava2017optimal}. In this paper, we will use the most recent variant of DOPH, which is significantly faster in practice compared to minwise hashing ~\citep{shrivastava2017optimal}. Throughout the paper, our mention of minwise hashing will refer to the DOPH algorithm for computing minhashes, which we have just mentioned. The complete details can be found in the aforementioned papers.

\section{Unique Entity Estimation}
\label{proposal}
In this section, we provide notation used throughout the rest of the paper and provide an illustrative example.
We then propose our estimator, which is unbiased and has provably low variance. In addition, random sampling is a special case of our procedure as explained in section~\ref{sec:randVsAdaptive}.  Finally, we present our unique entity estimation algorithm in section~\ref{analysis}.
%In this section, we provide notation and an illustrative example for unique entity estimation. Next, we give our unbiased and
\subsection{Notation}
\label{form}
The problem of unique entity estimation can be reduced to approximating the number of connected components in a corresponding graph.
%
Given a data set with size $M$, we denote the records as
$$R = \{R_i| 1\leq i \leq M, \; i \in \mathbb{Z}\}.$$
%and
%define $Q(R_i, R_j) = 1$ if $R_i, R_j$ refer to the same entity and 0 otherwise. Now
Next, we define
$$Q(R_i, R_j) =\begin{cases}
1, & \text{if $R_i, R_j$ refer to the same entity }.\\
0, & \text{otherwise}. \end{cases}.$$
Let us represent the data set by a graph $G^* = (E, V),$ with vertices $E,V.$ Let vertex $V_i$ correspond to record $R_i$ and vertex $V_j$ correspond to record $R_j$. Then let edge $E_{ij}$ represent the linkage between records of $R_i$ and $R_j$ (or
vertex $V_i$ and $V_j$). More specifically, we can represent this by the following relationship:
\begin{align*}
V &= \{R_i|1\leq i \leq M, i \in \mathbb{Z}\}, \; \text{and} \;
E = \{(R_i, R_j)| \forall\; 1\leq i, j \leq M, \; Q(R_i, R_j) = 1\}.
\end{align*}

\subsection{Illustrative Example}
\label{sec:toy}

In this section, we provide an illustrative example of how six records are mapped to a graph $G^*$.
%
Consider record 3 (John) and record 5 (Johnathan) which correspond to the same entity (John Schaech). In $G^*$, there is an edge $E_{35}$ that connect these records, denoted by $V_3$ and $V_5.$
%
Now consider records 2, 4, and 6, which all refer to the same entity (Nicholas Cage). In $G^*$, there is are edges $E_{24}, E_{26},$ and $E_{46}$  that connect these records, denoted by $V_2, V_4,$ and $V_6.$
%
Observe that each connected component in $G^*$ is a unique entity and also a clique. Therefore, our task is reduced to estimating the number of connected components in $G^*$.

\begin{figure}[ht]
	\begin{center}
		\includegraphics[width=\linewidth]{fig/map.pdf}
		\caption{A toy example of mapping records to a graph, where vertices represent records and edges refer the relation between records.}
		\label{map}
	\end{center}
	\vskip -0.2in
\end{figure}

\subsection{Proposed Unique Entity Estimator}
\label{analysis}
In this section, we propose our unique entity estimator and provide assumptions that are necessary for our estimation procedure to be practical (scalable).

Since we do not observe the edges of $G^*$ (the linkage), inferring whether there is an edge between two nodes (or whether two records are linked) can be costly, i.e., $O(M^2)$. Hence, one is constrained to probe a small set $\mathcal{S} \subset V \times V$ with $|\mathcal{S}|\ll O(M^2)$ of pairs and query if they have edges. The aim is to use the information about $\mathcal{S}$ to estimate the total number of connected components accurately. More precisely,  given the partial graph $G' = \{V,E'\}$, where $E' = E \cap \mathcal{S}$, one wishes to estimate the connected components $n$ of $G^* = \{V,E\}.$

One key property of our estimation process is that we do not make any modeling assumptions of how duplicate records are generated, and it is not immediately clear how we can obtain unbiased estimation. For sake of simplicity, we first assume the existence of an efficient (sub-quadratic) process that samples a small set (near-linear size) of edges $\mathcal{S}$, such that, every edge in the original graph $G^*$ has (reasonably high) probability $p$ of being in $\mathcal{S}$. Thus, set $\mathcal{S}$, even though small, contains $p$ fraction of the actual edges. For sparse graphs, as in the case of duplicate records, such a sampler will be far more efficient than random sampling. Based on this assumption, we will first describe our estimator and its properties. We then show in why our assumption about existence of adaptive sampler is practical by providing an efficient sampling process based on LSH (Section~\ref{samplep}).

\textbf{Remark}: It is not difficult to see that random sampling is a special case when $p = \frac{|\mathcal{S}|}{O(M^2)}$ which, as we show later, is a very small number for any accurate estimation.

Our proposed estimator and corresponding algorithm obtains the set of vertex pairs (or edges) $\mathcal{S}$  through an efficient (adaptive) sampling process and queries whether there is an edge (linkage) between each pair in $\mathcal{S}$.  Respectively, after the ground truth querying, we observe a sub-sampled graph $G'$, consisting of vertices returned by the sampler. Let $n_i'$ be the number of connected component of size $i$ in the observed graph $G'$, i.e., $n_1'$ is the number of singleton vertices, $n_2'$ is the number of isolated edges, etc. in $G'$. It worth noting that every connected component in $G'$ is a part of some clique (maybe larger) in $G^*$.  Let $n_i^*$ denote the number of connected components (clique) of size $i$ in the original (unobserved) graph $G^*$.

Observe that under the sampling process, any original connected component, say $C_i^*$ (clique), will be sub-sampled and can appear as some possibly smaller connected components in $G'$. For example, a singleton set in $G^*$ will remain the same in $G'$. An isolated edge, on the other hand, can appear as an edge in $G'$ with probability $p$ and as two singleton vertices in $G'$ with probability $1-p$. A triangle can decompose into three possibilities with probability shown in Figure~\ref{p}. Each of these possibilities provides a linear equation connecting $n_i^*$ to $n_i'$. These equations up to cliques of size three are

\begin{align}
\label{n3}
&\mathbb{E}[n_3']= n_3^* \cdot p^2 \cdot (3-2p) \\
&\mathbb{E}[n_2'] = n_2^* \cdot p + n_3^*\cdot (3\cdot(1-p)^2\cdot p) \\
\label{n1}
&\mathbb{E}[n_1'] = n_1^* + n_2^* \cdot(2\cdot(1-p)) + n_3^* \cdot (3\cdot (1-p)^2).
\end{align}

Since we observe $n_i'$,  we can solve for the estimator of each $n_i^*$ and compute the number of connected components by summing up all $n_i^*$.

\begin{figure}[ht]
	\begin{center}
		\includegraphics[width=\linewidth]{fig/p.pdf}
		\caption{A general example illustrating the transformation and probabilities of connected components from $G^*$ to $G'.$}
		\label{p}
	\end{center}
	\vskip -0.2in
\end{figure}

Unfortunately, this process quickly becomes combinatorial, and in fact, is at least $\#P$ hard~\citep{doi:10.1137/0212053} to compute for cliques of larger sizes. A large clique of size $k$ can appear as many separate connected components and the possibilities of smaller size components it can break into are exponential ~\citep{wiki:xxx}.
%
Fortunately, we can safely ignore large connected components without significant loss in estimation with two reasons.
First, in practical entity resolution tasks, when $M$ is large and contains at least one string-valued feature, it is observed that \emph{most} entities are replicated no more than three or four times. Second,  a large clique can only induce large errors if it is broken into many connected components due to undersampling. According to \cite{erdos1960evolution}, it will almost surely stay connected if $p$ is high, which is the case with our sampling method.

{\bf Assumption:} As argued above, we safely assume that the cliques of sizes equal or larger than 4 in the original graph would retain their structures, i.e., $\forall i \ge 4, \ n_i^* = n_i'$. With this assumption, we can write down the formula for estimating $n_1^*$, $n_2^*$, $n_3^*$ by solving Equations~\ref{n3}--\ref{n1} as,
\begin{align}
&n_3^*= \frac{\mathbb{E}[n_3']}{p^2 \cdot (3-2p) }, \quad n_2^* = \frac{\mathbb{E}[n_2'] - n_3^*\cdot (3\cdot(1-p)^2\cdot p)}{p}  \\
&n_1^* = \mathbb{E}[n_1'] - n_2^* \cdot(2\cdot(1-p)) - n_3^* \cdot (3\cdot (1-p)^2)
\end{align}
It directly follows that our estimator, which we call the Locality Sensitive Hashing Estimator (LSHE) for the number of connected components is given by
\begin{align}
\label{main}
\text{LSHE} &= n_1' +  n_2' \cdot \frac{2p-1}{p}  + n_3' \cdot \frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p) }
+ \sum_{i=4}^{M}n_i'.
\end{align}

\subsection{Optimality Properties of LSHE}
%\subsection{Theoretical Analysis}
We now prove two properties of our unique entity estimator, namely, that it is unbiased and that is has provably low variance than random sampling approaches.

\begin{theorem}
	\label{thm:thm1}
	Assuming  $\forall i \ge 4, \ n_i^* = n_i'$, we have
	\begin{align}\label{eq:thm1}
	\mathbb{E}[\text{LSHE}] &=  n   \ \ \ \mbox{unbiased} \\
	\label{var}
	\mathbb{V}[\text{LSHE}] &= n_3^* \cdot \frac{(p-1)^2 \cdot (3p^2-p+1)}{p^2 \cdot (3-2p)} + n_2^*\frac{(1-p)}{p}
	\end{align}
	The above estimator is unbiased and the variance is given by Equation \ref{var}.
\end{theorem}
Theorem \ref{thm:thm2} proves the variance of our estimator is monotonically decreasing with $p$.
\begin{theorem}
	\label{thm:thm2}
	$\mathbb{V}[\text{LSHE}]$ is monotonically decreasing when $p$ increases in  range $(0,1]$.
\end{theorem}

The proof of Theorem \ref{thm:thm2} directly follows from the following Lemma \ref{lemma}.
\begin{lemma}
	\label{lemma}
	First order derivative of $\mathbb{V}[\text{LSHE}]$ is negative when $p \in (0,1]$.
\end{lemma}
Note that when $p=1$, $\mathbb{V}[\text{LSHE}] = 0$ which means the observed graph $G'$ is exactly the same as $G^*$. For detailed proofs of unbiasedness and Lemma \ref{lemma}, see Appendix \ref{sec:app}.

\subsection{Adaptive Sampling versus Random Sampling}
\label{sec:randVsAdaptive}
Before we describe our adaptive sampler, we briefly quantify the advantages of an adaptive sampling over random sampling for the Syrian data set by computing the differences between their variances. Let $p$ be the probability that an edge (correct match) is sampled. On the Syrian data set, our proposed sampler, described in next section, empirically achieves $p = 0.9$, by reporting around a million sampled pairs out of the 45 billion possibilities. Substituting this value of $p$, the corresponding variance can be calculated from Equation~\ref{var} as $$n_3^* \cdot 0.026 + n_2^*\cdot 0.111.$$

Turning to plain random sampling of edges, in order to achieve the same sample size above leads to $p$ as low as $\frac{10^6}{4.5 \time 10^{10}} \simeq 0.00002$. With such minuscule $p$, the resulting variance is $$n_3^* \cdot 833294446 + n_2^*\cdot 49999.$$ Thus, the variance for random sampling is roughly 50,000 times the number of duplicates in the data set and $8.3 \times 10^8$ the number of triplets in the data set.

In section~\ref{experiments}, we illustrate that two other random sampling based algorithms of~\cite{chazelle2005approximating} and ~\cite{1978paper} also have poor accuracy compared to our proposed estimator. The poor performance of random sampling is not surprising from a theoretical perspective, and illustrates a major weakness empirically for the task of unique entity estimation with sparse graphs, where adaptive sampling is significantly advantageous.

\subsection{The Missing Ingredient: (K,L)-LSH Algorithm}
\label{sec:missingVariant}
%\subsection{The Missing Ingredient: Sub-quadratic Adaptive Sampling}
%\textcolor{red}{THIS SECTION IS TOO LONG AND THE SECTION TITLE IS CONFUSING. COME BACK TO THIS.}
Our proposed methodology, for unique entity estimation, assumes that we have an efficient algorithm that adaptively samples a set of record pairs, in sub-quadratic time. In this section, we argue that using a variant of LSH (Section~\ref{sec:lsh}) we can construct such an efficient sampler.

As already noted, we do not make any modeling assumptions on the generation process of the duplicate records. Also, we cannot assume that there is a fixed similarity threshold, because in real datasets duplicates can have arbitrarily large similarity. Instead, we rely on the observation that record pairs with high similarity have a higher chance of being duplicate records.  That is, we assume that when two  entities $R_i$ and $R_j$ are similar in their attributes; it is more likely that they refer to the same entities~\citep{christen_2012}.\footnote{The similarity metric that we use to compare sets of record strings is the Jaccard similarity.} We note that this probabilistic observation is the weakest possible assumption, and almost always true for entity resolution tasks because linking records by a similarity score is one simple way of approaching entity resolution \citep{christen_2012, winkler_2006, fellegi_1969}.

The similarity between entities (records) naturally gives us a notion of adaptiveness. One simple adaptive approach is to sample records pairs with probability proportional to their similarity. However, as a prerequisite for such sampling, we must compute all the pairwise similarities and associated probability values with every edge. Computing such a pairwise similarity score is a quadratic operation ($O(M^2)$) and is intractable for large datasets. Fortunately, recent work has shown that~\cite{spring2017scalable,spring2017new,luo2017Arrays} it is possible to sample pairs adaptively in proportion to the similarity in provably sub-quadratic time using LSH, which we describe in the next section.

%\subsubsection{(K,L)-parameterized Pair Sampling Algorithm}
\subsubsection{(K,L)-LSH Algorithm and Sub-quadratic Adaptive Sampling}
\label{sec:KLParametrizedLSH}

We leverage a very recent observation associated with the traditional $(K, L)$ parameterized LSH algorithm. The $(K, L)$ parameterized LSH algorithm is a popular similarity search algorithm, which given a query $q$, retrieves element $x$ from a preprocessed data set in sub-linear time ($O(KL) \ll M$) with probability $1-(1-\mathcal{J}(q,\ x)^K)^L$.  Here, $J$ denotes the Jaccard similarity between the query and the retrieved data vector $x$. Our proposed method leverages this $(K, L)$-parameterized LSH Algorithm, and we briefly describe the algorithm in this section. For complete details refer to~\cite{Report:E2LSH}.

Before we proceed, we define hash maps and keys. We use hash maps, where every integer (or keys) are associated with a bucket (or a list) of records. In a hash map, searching for the bucket corresponding to a key is a constant time operation. Please refer to algorithms literature~\citep{rajaraman_2012} for details on hashing and its computational complexity. Our algorithm will require several hash maps, $L$ of them, where a record $R_i$ is associated with a unique bucket in every hash map. The key corresponding to this bucket is determined by minwise hashes of the record $R_i$. We encourage readers to refer~\cite{Report:E2LSH} for implementation details.

More precisely, let $h_{ij}$, $i = \{1, \ 2,...,\ L\}$ and $j = \{1, \ 2,...,\ K\}$ be $K \times L$ minwise hash functions (Equation~\ref{eq:MinHash}) with each minwise hash function is formed by independently choosing the underlying permutation $\pi$. Next, we construct $L$ meta-hash functions (or the keys) $H_i = \{h_{i,1}, h_{i,2},...,h_{i,K}\}$, where each of the $H_i$'s are formed by combining $K$ different minwise hash functions. For this variant of the algorithm, we need a total of $K \times L$ functions. With such $L$ meta-hash functions, the algorithm has two main phases, namely the data pre-processing and the sampling pairs phases, which we outline below.

\begin{itemize}
  \item {\bf Data Preprocessing Phase:} We create $L$ different hash maps or popularly referred to as hash tables in computer science, where every hash value maps to a bucket of elements.   For every record $R_i$ in the dataset, we insert $R_j$ in the bucket associated  with the key $H_i(R_j)$, in hash map $i = \{1, \ 2, ..., \ L\}$. To assign $K$-tuples $H_i$ (meta-hash) to a number in a fixed range, we use some universal random mapping function to the desired address range. See~\cite{andoni20052} for details.
  \item {\bf Sample Pair Reporting:} For every $R_j$ in the dataset. From table $i$, get all elements in the bucket associated with key $H_i(R_j)$, where $i = \{1,\ 2,...,\ L\}$. Then take the union of the $L$ buckets obtained from the $L$ hash tables, and denote this (aggregated) set by $A.$ Finally, report pairs of records $(R_i, R_j)$, where $R \in A$.
\end{itemize}

\begin{theorem}
The above algorithm reports a pair $(R_i,\ R_j)$ with probability $1 - (1-\mathcal{J}(R_i,\ R_j)^K)^L$, where $\mathcal{J}(R_i,\ R_j)$ is the Jaccard Similarity between record pairs $(R_i,\ R_j).$
\end{theorem}
{\bf Proof:} Since all the minwise hashes are independent due to an independent sampling of permutations. The probability that both $R_i$ and $R_j$ belongs to same bucket in any of hash table $i$ is exactly $\mathcal{J}(R_i,\ R_j)^K$. Note from equation~\ref{eq:MinHash}, each meta-hash agreement has probability $\mathcal{J}(R_i,\ R_j)$. Therefore, the probability that pair $(R_i, \ R_j)$ is missed by all the $L$ tables is precisely $(1-\mathcal{J}(R_i,\ R_j)^K)^L$, and thus, the required probability of successful retreival is the complement.

The probabilistic expression $1 - (1-\mathcal{J}(R_i,\ R_j)^K)^L$ is a monotonic function of the underlying similarity $Sim(q,y)$ associated with the LSH. In particular, higher similarity pairs have more chance of being retrieved. Thus, LSH provides the required sampling which is adaptive in similarity and is sub-quadratic in running time.

\subsubsection{Computational Complexity}
\label{sec:CC}
The computational compexity for sampling with $M$ records is $O(MKL)$. The procedure requires computing $KL$ minwise hashes for each record. This step is followed by adding every record to $L$ hash tables.  Finally, for each record, we aggregate $L$ buckets to form sample pairs. The result of monotonicity and adaptivity of the samples applies to any value of $K$ and $L$. We choose $O(K \times L) \ll O(M)$ such that we are able to get samples in sub-quadratic time. We further tune $K$ and $L$ using cross-validation to limit the size of our samples. In section~\ref{sec:parameterstudy}, we evaluate the effect of varying $K$ and $L$ in terms of the recall and reduction ratio. (For a review of the recall and reduction ratio, we refer to \cite{christen_2012}.) We address the precision at the very end of our experimental procedure to ensure that the recall, reduction ratio, and precision of our proposed unique entity estimation procedure are all as close to 1 as possible while ensuring that the entire algorithm is computationally efficient. For example, on the Syrian data set, we can generate a million sample in less than $120 \sec$ with an adaptive sampling probability (recall) $p$ as high as $0.9$.  On the other hand, computing all pairwise similarities (45 billion) takes more than 6 days on the same machine with 28 cores capable of running 56 threads in parallel.

Next, we describe how this LSH sampler is related to the adaptive sampler described earlier in  Section~\ref{analysis}.

\subsubsection{Underlying Assumptions and Connections with $p$}
Recall that we can efficiently sample record pairs $R_i, \ R_j$ with probability $1-(1-J(R_i,\ R_j)^K)^L.$
%$1-(1-Sim(R_i,\ R_j)^K)^L$, where $Sim(R_i, \ R_j) = \mathcal{J}(R_i,R_j)$ is the Jaccard Similarity
%(Equation~\ref{eq:jacc}).
Since we are not making any modeling assumptions, we cannot directly link this probability to $p$, the probability of sampling the right duplicated pair (or linked entities) as required by our estimator $\text{LSHE}$. In the absence of any knowledge, we can get the estimate of $p$ using a small set of labeled linked pairs $\mathcal{L}$. Given a small sample set of manually labeled $\mathcal{L}$ record pairs, which is often available for entity resolution datasets (and happens to be available for the Syrian data set), we can estimate the value of $p$ by counting the fraction of matched pairs (true edges) from $\mathcal{L}$ reported by the sampling process.

Note that in practice there is no similarity threshold $\theta$ that guarantees that two record pairs are duplicate records. That is, it is difficult in practice to know a fixed $\theta$ where $\mathcal{J}(R_i,\ R_j) \ge \theta$ ensures that $R_i$ and $R_j$ are the same entities. However, the weakest possible, and a reasonable assumption is that high similarity pairs (textual similarity of records) should have higher chances of being duplicate records than lower similarity pairs.

Formally, this assumption implies that there exists a monotonic function $f$ of similarity $\mathcal{J}(R_i,\ R_j)$ such that the probability of for any $R_i, \ R_j$ being a duplicate records is given by $f(\mathcal{J}(R_i,\ R_j))$.  Since our sampling probability $1-(1-\mathcal{J}(R_i,\ R_j)^K)^L$ is also a monotonic function of $\mathcal{J}(R_i,\ R_j)$, we can also write $$f(\mathcal{J}(R_i,\ R_j)) = g(1-(1-\mathcal{J}(R_i,\ R_j)^K)^L),$$ where $g$ is $f$ composed with $h^{-1}$ which is the inverse of $h(x) = 1-(1-x^K)^L$. Unfortunately, we do not know the form of $f$ or $g$.

Instead of deriving $g$ (or $f$), which requires additional implicit assumptions on the form of the functions, our process estimates $p$ directly. In particular, the estimated value of $p$ is a data dependent mean-field approximation of $g$, or  rather, $$p = \mathbb{E}[g(1-(1-\mathcal{J}(R_i,\ R_j)^K)^L)].$$ Crucially, our estimation procedure does not require any modeling assumptions regarding the generation process of the duplicate records, which is significant for noisy data sets, where such assumptions typically break.

%We stress on the fact that this does not require any modeling assumption about the duplicated entities, which is critical for real and significantly noisy datasets where all assumptions usually break.

\subsubsection{Why LSH?}
\label{sec:WHYLSH}
Although there are several rule-based blocking methodologies, LSH is the only one which is also a random adaptive sampler. In particular, consider a rule-based blocking mechanism, for example on the Syrian data set, which might block on the date of death feature. Such blocking could be a very reasonable strategy for finding candidate pairs. Note that it is still very likely that duplicate records can have different date of deaths because the information could be missing or misrepresented.  In addition, such a blocking method is deterministic, and different independent runs of the blocking algorithm will report the same set of pairs. Even if we find reasonable candidates, we cannot up-sample the linked records to get an unbiased estimate. There will be a systematic bias in the estimates, which does not have any reasonable correction. In fact, random sampling to our knowledge is the only known choice in the existing literature for an unbiased estimation procedure, however, as already mentioned random uninformative sampling is likely to be very inaccurate.

LSH, on the other hand, can also be used as a blocking mechanism~\citep{steorts14comparison}. It is, however, more than just a blocking scheme, it is a provably adaptive sampler.
%a provable adaptive sampler.
Due to randomness in the blocking, different runs of sampler lead to different candidates, unlike deterministic blocking. We can also average over multiple runs to even increase the concentration of our estimates. The adaptive sampling view of LSH has come to light very recently~\citep{spring2017scalable,spring2017new,luo2017Arrays}. With adaptive sampling, we get much sharper unbiased estimators than the random sampling approach. To our knowledge, this is the first study of LSH sampling for unique entity estimation.

\subsection{Putting it all Together: Scalable Unique Entity Estimation}
\label{putit}

We now describe our scalable unique entity estimation algorithm.  As mentioned earlier, assume that we have a data set that contains a text representation of the $M$ records. Suppose that we have a reasonably sized manually labeled training set $\mathcal{T}$. We will denote the set of sampled pairs of records given by our sampling process as $\mathcal{S}$. Note, each element of $\mathcal{S}$ is a pair. Then our scalable entity resolution algorithm consists of three main steps, with the total computational complexity $O(ML+ KL +|\mathcal{S}| +|\mathcal{T}|)$.  In our case, we will always have $|\mathcal{S}| \ll O(M^2)$ and $KL \ll M$ (in fact, $L$ will be a small constant), which ensures that the total cost is strictly sub-quadratic. The complete procedure is summarized in Algorithm~\ref{Est}.

\begin{enumerate}
    \item {\bf Adaptively Sample Record Pairs ($O(ML)$):} We regard each record $R_i$ as a short string and replace it by an ``n-grams" based representation. Then one computes $K \times L$  minwise hashes of each corresponding string. This can be done in a computationally efficient manner using the DOPH algorithm \cite{shrivastava2017optimal}, which is done in data reading time. Next, once these hashes are obtained, one applies the sampling algorithm described in section~\ref{samplep} in order to generate a large enough sample set, which we denote by $\mathcal{S}$. For each record, the sampling step requires exactly $L$, hash table queries which are themselves $O(1)$ memory lookups. Therefore, the computational complexity of this step is $O(ML +KL).$
    \item {\bf Query each Sample Pairs:} Given the set of sampled pairs of records $\mathcal{S}$ from Step 1, for every pair of records in $\mathcal{S}$, we query whether these record pairs are a match or non-match. This step requires, $O(|S|)$, queries for the true labels. Here, one can use manually labeled data if it exists. In the absence of manually labeled data, we can also use a supervised algorithm, such as support vector machines, random forests, that is trained on the manually labeled set $\mathcal{T}$ (Section~\ref{syria}).
    \begin{enumerate}
        \item {\bf Estimate $p$:} Given the sampled set of record pairs $\mathcal{S},$ we need to know the value of $p$, the probability that any given correct pair is sampled. To do so, we use the fraction of true pairs sampled from the labeled training set $\mathcal{T}.$ The sampling probability $p$ can be estimated by computing the fraction of the matched pairs of training set records $\mathcal{T}_{match}$ appearing in $\mathcal{S}$. That is, we estimate $p$ (unbiasedly) by $$p = \frac{|\mathcal{T}_{match} \cap \mathcal{S}|}{|\mathcal{T}_{match}|}.$$ If $T$ is stored in a dictionary, then this step can be done on the fly while generating samples. It only cost $O(\mathcal{T})$ extra work to create the dictionary.

    \item {\bf Count Different Connected Components in $G'$ ($O(M+|\mathcal{S}|)$):}
        The resulting matched sampled pairs, after querying every sample for actual (or inferred) labels, form the edges of $G'$. We now have complete information about our sampled graph $G'$. We can now traverse $G'$ and counts all sizes of connected components in $G'$ to obtain $n_1'$, $n_2'$, $n_3'$ and so on. Traversing the graph has computational complexity $O(M + |\mathcal{S}|)$ time using Breadth First Search (BFS).
    \end{enumerate}
    \item {\bf Estimate the Number of Connected Components in $G^*$ ($O(1)$):} Given the values of $p$, $n_1'$, $n_2'$, and $n_3'$  we use equation~\ref{main} to compute the unique entity estimator $\text{LSHE}$.
\end{enumerate}

\begin{figure}
%\begin{wrapfigure}{R}{0.47\textwidth}
	\begin{minipage}{\textwidth}
		\begin{algorithm}[H]
			\caption{LSH-Based Unique Entity Estimation Algorithm}
			\label{Est}
			\begin{algorithmic}[1]
				\STATE {\bfseries Input:} Records $R$, Labeled Set $\mathcal{T}$, Sample Size $m$
				\STATE {\bfseries Output:} $LSHE$
				\STATE $\mathcal{S}$ = $LSH Sampling$($R$) (Section~\ref{sec:KLParametrizedLSH})
                \STATE Get $\mathcal{T}_{match}$ be the linked pairs (duplicate entities) in $\mathcal{T}$
				\STATE $p$ = $\frac{|\mathcal{T}_{match} \cap \mathcal{S}||}{|\mathcal{T}_{match}|}$
                \STATE Query every pair in $\mathcal{S}$ for match/mismatch (get actual labels). (Graph $G'$)
				\STATE $n_1', n_2', n_3'... n_M'$ = $Traverse(G')$
				\STATE $LSHE$ = $Equation$ \ref{main} ($p$ , $n_1', n_2', n_3'... n_M'$)
			\end{algorithmic}
		\end{algorithm}
	\end{minipage}
	\caption{Overview of our proposed unique entity estimation algorithm.}
\label{samplep}
	\end{figure}
%\end{wrapfigure}

\section{Experiments}
\label{experiments}

We evaluate the effectiveness of our proposed methodology on the Syrian data set and three additional real data sets, where the Syrian data set is only partially labeled while the other three data sets are fully labelled. We first perform evaluations and comparisons on the three fully labeled data sets, and then give an estimate of the documented number of identifiable victims for the Syrian data set.

\looseness=-1
\begin{itemize}
    \item \textbf{Restaurant}: The \textbf{Restaurant} data set contains 864 restaurant records collected from Fodor's and Zagat's restaurant guides.\footnote{Originally provided by Sheila Tejada, downloaded from http://www.cs.utexas.edu/users/ml/riddle/data.html.}  There are a total of 112 duplicate records. Attribute information contains name, address, city, and cuisine.

        \item \textbf{CD}: The \textbf{CD} data set that includes 9,763 CDs randomly extracted from freeDB.\footnote{https://hpi.de/naumann/projects/repeatability/datasets/cd-datasets.html.} There are a total of 299 duplicate records. Attribute information consists of 106 total features such as artist name, title, genre, among others.

    \item \textbf{Voter}: The  \textbf{Voter} data has been scraped and collected by \cite{christen16} beginning in October 2011. We work with a subset of this data set containing 324,074 records. There are a total of 68,627 duplicate records. Attribute information contains personal information on voters from North Carolina including full name, age, gender, race, ethnicity, address, zip code, birth place, and phone number.

    \item \textbf{Syria}: The \textbf{Syria} data set comprises data from the Syrian conflict, which covers the same time period, namely, March 2011 -- November 2013. This data set is not publicly available and was provided by the Human Rights Data Analysis Group (HRDAG). The respective data sets come from the Violation Documentation Centre (VDC), Syrian Center for Statistics and Research (CSR-SY), Syrian Network for Human Rights (SNHR), and Syria Shuhada website (SS). Each database lists a different number of recorded victims killed in the Syrian conflict, along with available identifying information including full Arabic name, date of death, death location, gender, among others.\footnote{These databases include documented identifiable victims and not those who are missing in the conflict. Hence, any estimate reported only refers to the data at hand.}
\end{itemize}

The above datasets cover a wide spectrum of different varieties observed in practice. For each data set, we report summary information in
Table~\ref{table1}.

\begin{table}[t]
	\centering			
	\begin{tabular}{lccccc}
		\toprule
		\textbf{DBname} & \textbf{Domain}& \textbf{Size} & \textbf{\# Matching Pairs} & \textbf{\# Attributes}& \textbf{\# Entities} \\
		\midrule
		Restaurants & Restaurant Guide&864  &112 & 4&752 \\

		CD & Music CDs&9,763  &299 & 106 & 9,508\\

		Voter & Registration Info&324,074  & 70,359 & 6 & 255,447\\

		Syria & Death Records&296,245  & N/A & 7 & N/A\\
		\bottomrule
	\end{tabular}	
	\caption{ We present five important features of the four data sets.
%	(three well-labeled real data set and Syrian data set).
	\textbf{Domain} reflects the variety of the data type we used in the experiments. \textbf{Size} is the number of total records respectively. \textbf{\# Matching Pairs} shows how many pair of records point to the same entity in each data set. \textbf{\# Attributes} represents the dimensionality of individual record. \textbf{\# Entities} is the number of unique records. }
	\label{table1}
\end{table}

\begin{figure}[H]
	%\begin{figure}[!Htb]
	%	\centering
	\mbox{

		\includegraphics[width=\linewidth]{fig/arabic.pdf}
}
	\caption{ \color{green} We show an example of a death record in Syrian dataset.}
	\label{fg1}
\end{figure}

\subsection{Evaluation Settings}
In this section, we outline our evaluation settings. We denote Algorithm~\ref{Est} as the LSH Estimator (LSHE). We make comparisons to the non-adaptive variant of our estimator (PRSE), where we use plain random sampling (instead of adaptive sampling). This baseline uses the same procedure as our proposed LSHE, except that the sampling is done uniformly. A comparison with PRSE quantifies the advantages of the proposed adaptive sampling over random sampling. In addition, we implemented the two other known sampling methods, for connected component estimation, proposed in \cite{1978paper} and \cite{chazelle2005approximating}. For convenience, we denote them as Random Sub-Graph based Estimator (RSGE), and BFS on Random Vertex based Estimator (BFSE) respectively. Since the algorithms are based on sampling (adaptive or random), to ensure fairness, we fix a budget $m$ as the number of pairs of vertices considered by the algorithm. Note that any query for an edge is a part of the budget. If the fixed budget is exhausted, then we stop sampling process and use the corresponding estimate, using all the information available.

We briefly describe the implementation details of the four considered estimators below:
\begin{enumerate}
\item \textbf{LSHE:} In our proposed algorithm, we use the ($K, L$) parameterized LSH algorithm to generate samples of record pairs using Algorithm~\ref{samplep}, where recall $K$ and $L$ control the resulting sample size (section~\ref{sec:parameterstudy}). Given $K, L$ as an input to Algorithm \ref{Est}, we use the sample size as the value of the fixed budget $m$. Table \ref{table2} gives different sample budget sizes (with the corresponding $K$ and $L$) and corresponding values of $p$ for selected samples in three real data sets.
%Parameters $K$ and $L$ controls the size of the sample size. Please refer to ~\ref{sec:parameterstudy} for details. We also report this values of $K$ and $L$ for all the datasets.  Table \ref{table2} gives different sample budget sizes and corresponding values of $p$ for selected samples in three real data sets.
\item \textbf{PRSE:} For a fair comparison, in this algorithm, we randomly sample the same number of record pairs used by LSHE. We then perform the same estimation process as LSHE but instead use $p = \dfrac{2m}{M(M-1)},$ which corresponds to the random sampling probability to get the same number of samples, which is $m$.
\item \textbf{RSGE~\citep{1978paper}:} This algorithm requires performing breadth first search (BFS) on each randomly selected vertices. BFS requires knowing all edges (neighbors) of a node for the next step, which requires $M-1$ edge queries. To ensure the fixed budget $m$, we end the traversal when the number of distinct edge queries reaches the fixed budget $m$.

\item \textbf{BFSE~\citep{chazelle2005approximating}:} This algorithm samples a subgraph and observes it completely. This requires labeling all the pairs of records in the sampled sub-graph. To ensure same budget $m$, the sampled sub-graph has approximately $\sqrt{2m}$ vertices.
\end{enumerate}

\textbf{Remark}: To the best of our knowledge there have been no experimental evaluations of the two algorithms of \cite{1978paper} and \cite{chazelle2005approximating} in the literature. Hence, our results could be of independent interest in itself.

We compute the relative error (RE), calculated as $$\text{RE} = \dfrac{|\text{LSHE} - n|}{n},$$ for each of the estimators, for different values of the budget $m$. We plot the $\text{RE}$ for each of the estimators, over a range of values of $m$, summarizing the results in Figure~\ref{fg1}.

All the estimators require querying pairs of records compared to labeled ground truth data for whether they are a match or a non-match. As already mentioned, in the absence of full labeled ground truth data, we can use a supervised classifiers such as SVMs as a proxy, assuming at least some small amount of labeled data exists. By training an SVM, we can use this as a proxy for labeled data as well. We use such a proxy in the Syrian data set because we are not able to query every pair of records to determine whether they are true duplicates or not.

%manual expert labelers for labeling each query, we use supervised classifiers such as support vector machine (SVM) as a proxy. Note that we have access to a small sample of manually labeled records. We can train an SVM, or other supervised classifiers, which we can be used as a proxy for manual labelers. This proxy is required for the Syrian dataset where we cannot query every pair for match-mismatch.

We start with the three data sets where fully labelled ground truth data exists. For LSHE, we compute both the estimation accuracy using the supervised SVM (Section~\ref{syria}) as well as using the fully labelled ground truth data. The difference in these two numbers quantifies the loss in estimation accuracy due to the use of the proxy SVM prediction instead of using ground truth labeled data. In our use of SVMs, we take less than 0.01$\%$ of total number of the possible record pairs as the training set.

\begin{figure}[H]
%\begin{figure}[!Htb]
	%	\centering
	\mbox{
		\hspace{-0.1in}
		\includegraphics[width=.36\linewidth]{fig/restaurant.eps}\hspace{-0.18in}
		\includegraphics[width=.36\linewidth]{fig/cd.eps}\hspace{-0.16in}
		\includegraphics[width=.36\linewidth]{fig/voter.eps}}
	\caption{ The dashed lines show the $\text{RE}$ of the four estimators on the three real data sets, where the y-axis is on the log-scale. Observe that LSHE outperforms all the other three estimators in one to two order of magnitude. The standard deviation of the $\text{RE}$ for LSHE is also shown in the plots with the red error bars.  In particular, the PRSE performs unreliable estimation on the CD and Voter data sets. The dashed and solid black lines represent RE of LSHE using ground truth labels and a SVM classifier (y-axis is on the log scale). We discuss the LSHE + SVM estimator in section \ref{syria} (solid black line).}
	\label{fg1}
\end{figure}

\subsection{Evaluation Results}
In this section, we summarize our results regarding the aforementioned evaluation metrics by varying the sample size $m$ on the three real data sets (see Figure \ref{fg1}). We notice that for the CD and Voter data sets, we cannot obtain any reliable estimate (for any sample size) using PRSE. Recall that plain random sampling almost always samples pairs of records that correspond to non-matches. Thus, it is not surprising that this method is unreliable because sampling random pairs is unlikely to result in a duplicate pair for entity resolution tasks. Even with repeated trials, there are no edges in the specified sampled pairs of records, leading to an undefined value of $p$. This phenomenon is a common problem in random sampling estimators over sparse graphs. Almost all the sampled nodes are singletons. Subsampling a small sub-graph leads to a graph with most singleton nodes, which leads to a poor accuracy of BFSE. Thus, it is expected that random sampling will perform poorly. Unfortunately, there is no other baseline for unbiased estimation of the number of unique entities.

From Figure \ref{fg1} observe that the $\text{RE}$ for proposed estimator LSHE is approximately one to two orders of magnitude lower than the other considered methods, where the  y-axis is on the log-scale. Undoubtedly, our proposed estimator LSHE consistently leads to significantly lower $\text{RE}$ (lower error rates) than the other three estimators. This is not surprising from the analysis shown in section~\ref{sec:randVsAdaptive}. The variance of random sampling based methodologies will be significantly higher.

%to extreme sparsity in the data (random pairs being duplicate has almost zero probability).

Taking a closer look at LSHE, we notice that we are able to efficiently generate samples with very high values of $p$ (See Table~\ref{table2}).
%{\color{red} We need a line describing the values of $p$ and referring to our precision recall experiments.}
%{\color{green} Here I suggest not relating to the precision recall experiments because they are in next section and we haven't explain those terms.}
In addition, we can clearly see that LSHE achieves high accuracy with very few samples. For example, for the CD data set, with a sample size less than $0.05\%$ of the total possible pairs of records of the entire data set, LSHE achieves $0.0006$ $\text{RE}$. Similarly, for the Voter data set, with a sample size less than $0.012\%$ of the total possible pairs of records of the entire data set, LSHE achieves $0.003$ $\text{RE}$.

Also, note the small values of $K$ and $L$ parameters required to achieve the corresponding sample size. $K$ and $L$ affects the running time, and small values $KL \ll O(M^2)$ indicate significant computational savings as argued in section~\ref{sec:CC}

As mentioned earlier, we also evaluate the effect of using SVM prediction as a proxy for actual labels with our LSHE. The dotted plot show those results. We remark on the results for LSHE + SVM in the next section~\ref{syria}.

\begin{table}[ht]
	\centering
	\setlength\tabcolsep{0.18cm}
		\begin{tabular}{l|cccc|cccc|cccc}
			\toprule
			\multicolumn{4}{c}{\textbf{Restaurant}}&\multicolumn{4}{c}{\textbf{CD}}&\multicolumn{4}{c}{\textbf{Voter}}
			\\
			\midrule
			Size&1.0&2.5&5.0&10&0.005 & 0.01& 0.02 & 0.04&0.002&0.006&0.009&0.013
			\\

			$p$&0.42&0.54&0.65&0.82& 0.72& 0.74& 0.82& 0.92& 0.62&0.72& 0.76& 0.82
			\\

			$K$&1&1&1&1& 1& 1& 1& 1& 4& 4& 4& 4
			\\

			$L$& 4 & 8 &12&20& 5& 6 & 8 & 14 &  25&  32&  35&40
			\\

			\bottomrule			
		\end{tabular}
			\caption{ We illustrate part of the sample sizes (in \% in TOTAL) for different set of samples generated by Min-Wise Hashing and their corresponding $p$ in all three data sets. }
			\label{table2}
\end{table}

\section{Estimation of Casualties in Syrian Conflict}
\label{syria}
In this section, we describe how we estimate the number of documented identifiable deaths for the Syrian data set. As noted before, we do not have ground truth labels for all record pairs, but the data set was partially labeled with 217,788 record pairs (out of 45 billion). We propose an alternative (automatic) method of labeling the sample pairs, which is also needed by our proposed estimation algorithm. More specifically, using the partially labeled pairs, we train an SVM. In fact, other supervised methods could be considered here, such as random forests, Bayesian Adaptive Regression Trees (BART), among others, however, given that SVMs perform very well, we omit such comparisons as we expect the results to be similar if not worse.

To train the SVM, we take every record pair and generate $k$-grams representation for each record. Then we spilt the partially labeled data into training and testing sets, respectively. Each training and testing set contains a pair of records $x_k = [R_i,\ R_j]$. In addition, we can use a binary label indicating whether the record pair is a match or non-match. That is, we can write the data as $\{x_k =[R_i,R_j],y_k\}$ as the set difference of the $k$-grams of the strings of pairs of records $R_i$ and $R_j$, respectively. Observe that $y_k = 1$ if the $R_i$ and $R_j$ is labelled as match and $y_k = -1$ otherwise. Next, we tune the SVM hyper-parameters using 5-fold cross-validation, and we find the accuracy of SVM on the testing set was 99.9\%.  With a precision as high a 0.99, we can reliably query an SVM and now treat this an expert label.

% We can use $R_1$, $R_2$ as well} and a binary classifier indicating whether the record pair is a match or non-match. \textcolor{red}{AS: THIS NEXT PART IS NOT CLEAR AND NEEDS TO BE WRITTEN MORE CLEARLY WITH WORDS AND SYMBOLS. I RECOMMEND WE REMOVE THE NEXT SENTENCE. I THINK IT WILL CONFUSED THE READER.}
%
%
% That is, the training data $\{x_i=[R_a,R_b],y_i\}$ has $x_i$ as the set difference of the $n$-gram features of the pairs of records $R_a$ and $R_b$ respectively.  Observe that $y_i = 1$ if the $R_a$ and $R_b$ is labelled as match and $y_i = -1$ otherwise. Next, we tune the SVM hyper-parameters using 5-fold cross validation, and we find the the SVM of the testing set was 99.9\%. \textcolor{red}{BC: WHAT was the split regarding testing and training? This is something that someone might ask regarding overfitting and actually I want to know more about. We should have a skype call about this and perhaps run more experiments after the submission.} With a precision as high a 0.99, we can reliably query an SVM and now treat this an expert label.

%We tune the SVM hyper parameter using standard 5-fold cross validations. We found that the accuracy of SVM on the unseen test set was over 99.9\%. With this high precision, we can reliably query an SVM and treat that as an expert label.

To understand the effect of using SVM prediction as a proxy to label queries in our proposed unique entity estimation algorithm, we return to observing the behavior in figure \ref{fg1}. Here, we compare to baseline comparison of the LSHE estimator on the other three real datasets which have fully labeled ground truth data. Next, we compare to a new variant of LHSE with an SVM component, where the SVM prediction replaces the querying process. We refer to this variant as LSHE +SVM. Observe in  Figure~\ref{fg1}, that the plot for LSH (solid black line) and LSH+SVM (dotted black line) overlap indicating a negligible loss in performance. This overlap is expected given the high accuracy (high precision) of the SVM classifier.

%We use the LSHE estimator on the other three real datasets which have fully labelled ground truth data.  We then compare the original procedure, which relies on ground truth labeling of samples, with LSHE where the SVM prediction replaces the querying part. We call the method with the SVM proxy as LSHE +SVM. Observe in  Figure~\ref{fg1}, that the plot for LSH (solid black line) and LSH+SVM (dotted black line) overlap indicating a negligible loss in performance. This overlap is expected given the high accuracy (high precision) of SVM classifier.

\subsubsection{Running Time}

We briefly highlight the speed of the sampling process since it could be used for on the fly or online unique entity estimation.
The total running time for producing 917,577 sampled pairs (out of possible 45 billion) used for the LSH sampler (Section~\ref{sec:KLParametrizedLSH}) with $K=25$ and $L=200$ is 119 seconds. On the other hand, it will approximately take 6 days to compute all pairwise similarities across the 300,000 records. Computing all pairwise similarities is the first prerequisite for any known adaptive sampling over pairs based on similarity -- if we are not using the proposed LSH sampler.
Note there are other ways of blocking~\citep{christen_2012}, however as mentioned in Section~\ref{sec:WHYLSH} they are mostly deterministic (or rule-based) and do not provide an estimator of the unique entities.)

\subsection{Unique Number of Documented Identifiable Victims}

In the Syrian dataset, with 300,000 records and possible 45 billion ($4.5 \times 10^{10}$) pairs, our motivating goal was to estimate the unique number of documented identifiable victims in the Syrian data set. Specifically, in our final estimate, we use
917,577 sampled pairs that are given by  LSHE+SVM  ($K = 25$, $L = 200$) which has approximately $p=0.9$ based on the relatively large subset of labeled pairs. The sample size was chosen to balance the computational runtime and the value of $p$.
Specifically, one wants high values of $p$ (for a resulting low variance of our estimate), and to balance running time we limit the sample size to be around 1 million record pairs. (Such settings are determined by the application, but as we have demonstrated they work for a variety of real entity resolution data sets). We chose the SVM as our classifier to label the matches and non-matches. The final unique number of documented identifiable victims in the Syrian data set was estimated to be 190,369$\pm207$, very close to the 190,000 reported casualties by HRDAG in 2014, where their process is described in Appendix~\ref{sec:syrian}.

\subsection{Effects of $L$, $K$, on sample size and $p$}
\label{sec:parameterstudy}
%\textcolor{red}{RS: Smooth this out and think about moving this to the appendix. Maybe call this a sensitivity analysis.}
In this section, we discuss the sensitivity of our proposed method as we vary the choice of $L$, $K$, the sample size $M$, and $p$.

We want both $KL \ll M$ as well as the number of samples to be $\ll M^2$, for the process to be truly sub-quadratic. For accuracy, we want high values of $p$, because the variance is monotonic in $p$ which is also the recall of true labeled pairs. Thus, there is a natural trade-off. If we sample more, we get high $p$ but more computations.

$K$ and $L$ are the basic parameters of our sampler (Section~\ref{sec:KLParametrizedLSH}), which trade of computations with accuracy. Large $K$ makes buckets sparse (exponentially), and so fewer pairs are sampled from each table. Large $L$ increases the repetition of hash tables (linearly) which increases the sample size. The computational cost as argued is $O(MKL)$.

To understand the behavior of $K$, $L$, $p$, and the computational cost, we perform a set of experiments on the Syrian dataset.  We use n-gram of 2---5, we vary L from 100--1000 by steps of 100 and K takes values 15,18,20,23,25,28,30,32,35. For all these combinations, we then plot the recall (also the value of $p$) and the reduction ratio (RR) which is the percentage of computational savings. A 99\% reduction ratio means only 1\% of total sampled pairs. Figure~\ref{syria-takethatAssad} shows the tradeoffs between reduction ratio and recall (or value of $p$). Every dot in the figures is one whole experiment.

Regardless of the n-gram variation from 2--5, the recall and reduction ratio (RR) are close to 1 as illustrated in Figure~\ref{syria-takethatAssad}. We see that an n-gram of 3 overall is most stable in having a recall and RR close to 0.99. We observe that $K=25$ and $L=200$ gives a high recall of around 90\% with less than a million pairs (out of 45 billion possible) to evaluate ($RR \ge 0.9999$).

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{plots/syria_full_recall_rr_allShingles}
\caption{For shingles 2--5, we plot the RR versus the recall. Overall, we see the best behavior for a shingle of 3, where the RR and recall can be reached at 0.98 and 1, respectively. We allow L and K to vary on a grid here. L varies from 100--1000 by steps of 100; and K takes values 15, 18, 20, 23, 25, 28, 30, 32, and 35.}
\label{syria-takethatAssad}
\end{center}
\end{figure}

\section{Discussion}
Motivated by three real entity resolution tasks and the ongoing Syrian conflict, we have proposed a general, scalable algorithm for unique entity estimation. Our proposed method is an adaptive LSH on the edges of a graph, which in turn estimates the connected components in sub-quadratic time. Our estimator is unbiased and has provably low variance, which contrasts other such estimators for unique entity estimation in the literature. In experimental results, it outperforms other estimators in the literature on three real entity resolution data sets. Moreover, we have estimated the number of conflict casualties to be 190,369$\pm207$, which very closely matches the 2014 HRDAG estimate, completed by hand-matching. To our knowledge, we have the first estimate for the number of documented identifiable deaths as well with a standard error associated with such an estimate.
Our methods are scalable, potentially bringing impact to the human rights community, where such estimates could be updated in near real time. It could lead to further impact in public policy and transitional justice, where Syria has currently been at the forefront of both national and international news.
%Add this later.
%
%We should talk about doing some kind of interaactive learning approach with this method that would allow for the training data to be improved. This would be super cool and be great for future work! (Or maybe let's not mention it and work on it).

\clearpage
\newpage

%\begin{supplement}[id=suppA]
%  \sname{Appendix}
%  \stitle{Supplementary Material for ``Regularized Brain Reading with Shrinkage and Smoothing''}
%  \slink[doi]{COMPLETED BY THE TYPESETTER}
%  \sdatatype{.pdf}
%  \sdescription{}
%\end{supplement}
%
%
%\input{appendix_journal_v2}

\clearpage
\newpage

\appendix

\section{Syrian Data Set}
\label{sec:syrian}
In this section, we provide a more detailed description about the Syrian data set. As mentioned in section \ref{syriadata}, via
collaboration with the Human Rights Data Analysis Group (HRDAG), we have access to four databases. They come from the Violation Documentation Centre (VDC), Syrian Center for Statistics and Research (CSR-SY), Syrian Network for Human Rights (SNHR), and Syria Shuhada website (SS). Each database lists each victim killed in the Syrian conflict, along with identifying information about each person (see \cite{price_2013} for further details).

Data collection by these organizations is carried out in a variety of ways. Three of the groups (VDC, CSR-SY, and SNHR) have trusted networks on the ground in Syria.  These networks collect as much information as possible about the victims. For example, information is collected  through direct community contacts. Sometimes information comes from a victim's friends or family members. Other times, information comes from religious leaders, hospital, or morgue records.  These networks also verify information collected via social and traditional media sources.  The fourth source, SS, aggregates records from multiple other sources, including NGOs as well as social and traditional media sources (see \url{http://syrianshuhada.com/} for information about specific sources).

These lists, despite being products of extremely careful, systematic data collection, are not probabilistic samples \citep{Sig2015, IOAS2015, CJLS2015, price2014updated}. Thus, these lists cannot be assumed to represent the underlying population of all victims of conflict violence.  Records collected by each source are subject to biases, stemming from a number of potential causes, including a group's relationship within a community, resource availability, and the current security situation.  Although it is beyond the scope of this paper, final analyses of these sources must appropriately adjust for such biases before drawing conclusions about patterns of violence.

\subsection{Syrian Handmatched Data Set}
\label{app:hand}
We now describe how HRDAG's training data on the Syrian data set was created, which we use in our paper. We would like to note that we only use a small fraction of the training data for two purposes. The first is such that we can see how close our estimate is in practice to their original handmatched estimate, given such a large portion of the data was handmatched. Second, we want to avoid using too much training data to avoid biases and also since such handmatching efforts would not be possible moving forward as the Syrian conflict continues, and our small training data set is meant for one moving forward in practice.
%created the handmatched data set used in our paper.

First, all documented deaths recorded by any of the documentation groups were concatenated together into a single list. From this list, records were broadly grouped according to governorate and year. In other words, all killings recorded in Homs in 2011 were examined as a group, looking for records with similar names and dates.

Next, several experts review these ``blocks", sometimes organized as pairs for comparison and other times organized as entire spreadsheets for review.  These experts determine whether pairs or groups of records refer to the same individual victim or not. Pairs or groups of records determined to refer to the same individual are assigned to the same ``match group." All of the records contributing to a single ``match group" are then combined into a single record. This new single record is then again examined as a pair or group with other records, in an iterative process.

For example, two records with the same name, date, and location may be identified as referring to the same individual, and combined into a single record. In a second review process, it may be found that that record also matches the name and location, but not date, of a third record. The third record may list a date one week later than the two initial records, but still be determined to refer to the same individual. In this second pass, information from this third record will also be included in the single combined record.

When records are combined, the most precise information available from each of the individual records is kept. If some records contain contradictory information (for example, if records A and B record the victim as age 19 and record C records age 20) the most frequently reported information is used (in this case, age 19). If the same number of records report each piece of contradictory information, a value from the contradictory set is randomly selected.

Three of the experts are native Arabic speakers; they review records with the original Arabic content. Two of the experts review records translated into English. These five experts review overlapping sets of records, meaning that some records are evaluated by two, three, four, or all five of the experts. This makes it possible to check the consistency of the reviewers, to ensure that they are each reaching comparable decisions regarding whether two (or more) records refer to the same individual or not.

After an initial round of clustering, subsets of these combined records were then re-examined to identify previously missed groups of records that refer to the same individual, particularly across years (e.g., records with dates of death 2011/12/31 and 2012/01/01 might refer to the same individual) and governorates (e.g., records with neighboring locations of death might refer to the same individual).

\section{Unique Entity Estimation Proofs}
\label{sec:app}
First, we introduce four indicators.
%
First, let $\mathbb{I}_2$ denote every 2-vertex clique in $G^*$ (Recall that $G^*$ is the original graph and $G'$ is the observed one):
	\begin{equation}
	\mathbb{I}_2=\begin{cases}
	1, & \text{if this clique is in } G'.\\
	0, & \text{otherwise}.
	\end{cases}
	\end{equation}
Second, let 	
$\mathbb{I}_{33}$ denote every 3-vertex clique in $G^*$:
	\begin{equation}
	\mathbb{I}_{33}=\begin{cases}
	1, & \text{if this clique mains as a 3-clique in } G'.\\
	0, & \text{otherwise}.
	\end{cases}
	\end{equation}
Third, let
$\mathbb{I}_{32}$ denote every 3-vertex clique in $G^*$:
	\begin{equation}
	\mathbb{I}_{32}=\begin{cases}
	1, & \text{if this clique breaks to a 2-clique in } G'.\\
	0, & \text{otherwise}.
	\end{cases}
	\end{equation}
Finally, let
$\mathbb{I}_{31}$ denote every 3-vertex clique in $G^*$:
	\begin{equation}
	\mathbb{I}_{31}=\begin{cases}
	1, & \text{if this clique breaks to only 1-cliques in } G'.\\
	0, & \text{otherwise}.
	\end{cases}
	\end{equation}

\subsection{Expectation}
We now prove that our estimator is unbiased. Consider
\begin{equation}
\label{C3}
\mathbb{E}[n_3']= \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}]= n_3^* \cdot p^2 \cdot (3-2p),
\end{equation}
\begin{equation}
\label{C2}
\begin{split}
\mathbb{E}[n_2'] &=    \mathbb{E}[\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}] + \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}]     \\
&= n_2^* \cdot p + n_3^*\cdot (3\cdot(1-p)^2\cdot p), \quad \text{and}
\end{split}
\end{equation}
\begin{equation}
\label{C1}
\begin{split}
\mathbb{E}[n_1']&= n_1^*+
\mathbb{E}[\sum_{i=1}^{n_2^*}{(1-\mathbb{I}_{2i})}] + \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}] + \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}}]\\
&= n_1^* + n_2^* \cdot(2\cdot(1-p)) + n_3^* \cdot (3\cdot (1-p)^2\cdot p) \\
&+ n_3^* \cdot (3\cdot(1-p)^3 ).
\end{split}
\end{equation}

Our estimator is unbiased via equations \ref{C1}, \ref{C2}, \ref{C3}:
\begin{equation*}
\begin{split}
\mathbb{E}[LSHE] =&\mathbb{E}[n_1' +  n_2' \cdot \frac{2p-1}{p}  \\
&+ n_3' \cdot \frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p) }
+ \sum_{i=4}^{M} n_i] \\
& = \mathbb{E}[n_1'] +  \frac{2p-1}{p} \cdot \mathbb{E}[n_2']    \\
&+ \frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p)} \cdot \mathbb{E}[n_3']   +  \mathbb{E}[\sum_{i=4}^{M}n_i]\\
& = n_1^* + n_2^* + n_3^* + \sum_{i=4}^{N}n_i^*\\
& = n.
\end{split}
\end{equation*}

%\begin{equation}
%\label{C3}
%\mathbb{E}[\hat{n_3}]= \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}]= n_3^* \cdot p^2 \cdot (3-2p)
%\end{equation}
%\begin{equation}
%\label{C2}
%\begin{split}
%\mathbb{E}[\hat{n_2}] &=    \mathbb{E}[\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}] - \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}]     \\
%&= n_2^* \cdot p - n_3^*\cdot (3\cdot(1-p)^2\cdot p)
%\end{split}
%\end{equation}
%
%\begin{equation}
%\label{C1}
%\begin{split}
%\mathbb{E}[\hat{n_1}]&= n_1^*+
%\mathbb{E}[\sum_{i=1}^{n_2^*}{(1-\mathbb{I}_{2i})}] + \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}] + \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}}]\\
%&= n_1^* + n_2^* \cdot(2\cdot(1-p)) + n_3^* \cdot (3\cdot (1-p)^2\cdot p) \\
%&+ n_3^* \cdot (3\cdot(1-p)^3 )
%\end{split}
%\end{equation}
%
%\begin{equation}
%\begin{split}
%&\mathbb{E}[Est] =\mathbb{E}[\hat{C}_1 +  \hat{C}_2 \cdot \frac{2p-1}{p}  \\
%&+ \hat{C}_3 \cdot \frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p) }
%+ \sum_{n=4}^{N}\hat{C}_n] \\
%& = \mathbb{E}[\hat{C}_1] +  \frac{2p-1}{p} \cdot \mathbb{E}[\hat{C}_2]    \\
%&+ \frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p)} \cdot \mathbb{E}[\hat{C}_3]   +  \mathbb{E}[\sum_{n=4}^{N}\hat{C}_n]\\
%& = n_1 + n_2 + n_3 + \sum_{n=4}^{N}C_n
%\end{split}
%\end{equation}
%{\color{red} I will see what is a better way of explaining how we drop $\sum_{n=4}^{N}\hat{C}_n$}

\subsection{Variance}
We now turn to deriving the variance of our proposed estimator, showing that
$$\mathbb{V}[LSHE] = n_3^* \cdot \frac{(p-1)^2 \cdot (3p^2-p+1)}{p^2 \cdot (3-2p)} + n_2^* \cdot \frac{(1-p)}{p}.$$
Consider
\begin{equation}
\label{v1}
\begin{split}
&\mathbb{V}[LSHE]=
\mathbb{V}[\frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p) } \cdot \sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}\\
&+
\frac{2p-1}{p} \cdot
(\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}
+
\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}})\\
&+
\sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}}
+
\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}
+
\sum_{i=1}^{n_2^*}{(1-\mathbb{I}_{2i})}]\\
&=
\mathbb{V}ar[\frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p) } \cdot \sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}
+
\frac{3p-1}{p} \cdot
\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}\\
&+
3 \cdot
\sum_{i=1}^{n_2^*}{\mathbb{I}_{31i}}
-
\frac{1}{p} \cdot
\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}].
\end{split}
\end{equation}

Next, we replace $1-6 \cdot (1-p)^2 \cdot p$ by $a,$ and by simplifying equation \ref{v1},
we find
\begin{equation}
%\label{v1}
\begin{split}
&=
\frac{a^2}{(p^2 \cdot (3-2p))^2 } \cdot \mathbb{V}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}]
 +
\frac{(3p-1)^2}{p^2} \cdot
\mathbb{V}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}]\\
&+
9 \cdot
\mathbb{V}[\sum_{i=1}^{n_2^*}{\mathbb{I}_{31i}}]
-
\frac{1}{p^2} \cdot
\mathbb{V}[\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}]
+
Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}})\\
&+
Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}})
+
Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}}).
\end{split}
\end{equation}

Note that the covariance of $\sum_{i=1}^{n_3^*}{\mathbb{I}_{2i}}$ with any indicator is zero due to independence.
Furthermore, since the indicators are Bernoulli distributed random variables, the variance is easily found. Consider
\begin{equation}
\label{v4}
\begin{split}
\mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}]
=  \frac{a^2 \cdot (1-p^2 \cdot (3-2p))}{p^2 \cdot (3-2p) } \cdot n_3^*
\end{split}
\end{equation}

\begin{equation}
\label{v5}
\begin{split}
\mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}]
= \frac{3 \cdot (3p-1)^2 \cdot (1-p)^2 \cdot (1-3p \cdot (1-p)^2)}{p} \cdot n_3^*
\end{split}
\end{equation}

\begin{equation}
\label{v6}
\begin{split}
\mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}}]
= 9 \cdot (1-p)^3 \cdot (1-(1-p)^3) \cdot n_3^*
\end{split}
\end{equation}

\begin{equation}
\label{v7}
\begin{split}
\mathbb{V}ar[\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}]
= \frac{(1-p)}{p} \cdot n_2^*
\end{split}
\end{equation}

Using equations \ref{v4} -- \ref{v7}, the covariance simplifies to
\begin{equation*}
\begin{split}
&Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}})\\
&=  \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}} \cdot \sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}) -  	\mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}] \mathbb{E}[\sum_{i=1}^{n_2^*}{\mathbb{I}_{32i}}]\\
&=  \sum_{i=1}^{n_3^*}{\sum_{j=1}^{n_3^*}{
		\mathbb{E}[\mathbb{I}_{33i} \cdot \mathbb{I}_{32j}] -  	
		\mathbb{E}[\mathbb{I}_{33i}]
		\mathbb{E}[\mathbb{I}_{32j}]}}\\
&=-6 \cdot a \cdot n_3^*
\end{split}
\end{equation*}

When $i=j$, since $\mathbb{I}_{33j}$ and $\mathbb{I}_{32j}$ are mutually exclusive, $\mathbb{E}[\mathbb{I}_{33i} \cdot \mathbb{I}_{32j}]=0$. Otherwise when $i \neq j$, $\mathbb{I}_{33j}$ and $\mathbb{I}_{32j}$ are independent and $\mathbb{E}[\mathbb{I}_{33i} \cdot \mathbb{I}_{32j}]=0$. Similarly,
\begin{equation*}
\begin{split}
Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{n_1^*}{\mathbb{I}_{32i}})
= - 6 \cdot a \cdot (1-p)^3 \cdot n_3^*
\end{split}
\end{equation*}
and
\begin{equation*}
\begin{split}
Cov(\sum_{i=1}^{n_2^*}{\mathbb{I}_{32i}}, \sum_{i=1}^{n_1^*}{\mathbb{I}_{31i}})
= - 18 \cdot (1-p)^5 \cdot (3p-1) \cdot n_3^*.
\end{split}
\end{equation*}

It then follows that
\begin{equation*}
\begin{split}
&\mathbb{V}ar[LSHE]=n_3^*\cdot\Big(\frac{3\cdot(3p-1)^2\cdot(1-p)^2\cdot(1-3p\cdot(1-p)^2)}{p}\\
&+\frac{(1-6\cdot(1-p)^2\cdot p)^2\cdot(1-p^2\cdot(3-2p))}{p^2\cdot(3-2p)}\\
&-(6\cdot(1-6\cdot(1-p)^2\cdot p)\cdot(3p-1)\cdot(1-p)^2)\\
&+ 9\cdot(1-p)^6+3\cdot(1-p)^3 \Big)\\
&+ n_2^*\cdot\frac{(1-p)}{p}\\
&= n_3^* \cdot \frac{(p-1)^2 \cdot (3p^2-p+1)}{p^2 \cdot (3-2p)} + n_2^* \cdot \frac{(1-p)}{p}.
\end{split}
\end{equation*}

%\begin{equation}
%\label{v1}
%\begin{split}
%&\mathbb{V}ar[Est]=
%\mathbb{V}ar[\frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p) } \cdot \sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}\\
%&+
%\frac{2p-1}{p} \cdot
%(\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}
%+
%\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}})\\
%&+
%\sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}}
%+
%\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}
%+
%\sum_{i=1}^{n_2^*}{(1-\mathbb{I}_{2i})}]\\
%&=
%\mathbb{V}ar[\frac{1-6 \cdot (1-p)^2 \cdot p}{p^2 \cdot (3-2p) } \cdot \sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}
%+
%\frac{3p-1}{p} \cdot
%\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}\\
%&+
%3 \cdot
%\sum_{i=1}^{n_2^*}{\mathbb{I}_{31i}}
%-
%\frac{1}{p} \cdot
%\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}]\\
%\end{split}
%\end{equation}
%
%We now replace $1-6 \cdot (1-p)^2 \cdot p$ by $a$ to make equation \ref{v1} simpler.
%
% \begin{equation}
% \label{v1}
% \begin{split}
% &=
% \frac{a^2}{(p^2 \cdot (3-2p))^2 } \cdot \mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}]
% +
% \frac{(3p-1)^2}{p^2} \cdot
% \mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}]\\
% &+
% 9 \cdot
% \mathbb{V}ar[\sum_{i=1}^{n_2^*}{\mathbb{I}_{31i}}]
%  -
% \frac{1}{p^2} \cdot
% \mathbb{V}ar[\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}]
% +
% Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}})\\
%&+
% Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}})
%+
% Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}})\\
%\end{split}
%\end{equation}
%
%
%
%The Covariance of $\sum_{i=1}^{n_3^*}{\mathbb{I}_{2i}}$ with any other one is zero because they are independent.
%
%Because indicators are Bernoulli distributed random variables. We can get variance easily by:
%
%
%\begin{equation}
%\begin{split}
%\mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}]
%=  \frac{a^2 \cdot (1-p^2 \cdot (3-2p))}{p^2 \cdot (3-2p) } \cdot n_3^*
%\end{split}
%\end{equation}
%
%\begin{equation}
%\begin{split}
%\mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}]
%= \frac{3 \cdot (3p-1)^2 \cdot (1-p)^2 \cdot (1-3p \cdot (1-p)^2)}{p} \cdot n_3^*
%\end{split}
%\end{equation}
%
%\begin{equation}
%\begin{split}
%\mathbb{V}ar[\sum_{i=1}^{n_3^*}{\mathbb{I}_{31i}}]
%= 9 \cdot (1-p)^3 \cdot (1-(1-p)^3) \cdot n_3^*
%\end{split}
%\end{equation}
%
%\begin{equation}
%\begin{split}
%\mathbb{V}ar[\sum_{i=1}^{n_2^*}{\mathbb{I}_{2i}}]
%= \frac{(1-p)}{p} \cdot n_2^*
%\end{split}
%\end{equation}
%
%
%\begin{equation}
%\begin{split}
% &Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}})\\
%  &=  \mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}} \cdot \sum_{i=1}^{n_3^*}{\mathbb{I}_{32i}}) -  	\mathbb{E}[\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}] \mathbb{E}[\sum_{i=1}^{n_2^*}{\mathbb{I}_{32i}}]\\
%  &=  \sum_{i=1}^{n_3^*}{\sum_{j=1}^{n_3^*}{
%  \mathbb{E}[\mathbb{I}_{33i} \cdot \mathbb{I}_{32j}] -  	
%  \mathbb{E}[\mathbb{I}_{33i}]
%  \mathbb{E}[\mathbb{I}_{32j}]}}\\
%&=-6 \cdot a \cdot n_3^*
%\end{split}
%\end{equation}
%
%When $i=j$, since $\mathbb{I}_{33j}$ and $\mathbb{I}_{32j}$ are mutually excluded, $\mathbb{E}[\mathbb{I}_{33i} \cdot \mathbb{I}_{32j}]=0$. Otherwise when $i \neq j$, $\mathbb{I}_{33j}$ and $\mathbb{I}_{32j}$ are independent $\mathbb{E}[\mathbb{I}_{33i} \cdot \mathbb{I}_{32j}]=0$.
%
%Similarly,
%\begin{equation}
%\begin{split}
% Cov(\sum_{i=1}^{n_3^*}{\mathbb{I}_{33i}}, \sum_{i=1}^{C_1^*}{\mathbb{I}_{32i}})
%  = - 6 \cdot a \cdot (1-p)^3 \cdot n_3^*
%  \end{split}
%\end{equation}
%
%\begin{equation}
%\begin{split}
% Cov(\sum_{i=1}^{n_2^*}{\mathbb{I}_{32i}}, \sum_{i=1}^{C_1^*}{\mathbb{I}_{31i}})
% = - 18 \cdot (1-p)^5 \cdot (3p-1) \cdot n_3^*
%\end{split}
%\end{equation}
%
%\begin{equation}
%\begin{split}
%&\mathbb{V}ar[Est]=n_3^*\cdot\Big(\frac{3\cdot(3p-1)^2\cdot(1-p)^2\cdot(1-3p\cdot(1-p)^2)}{p}\\
%&+\frac{(1-6\cdot(1-p)^2\cdot p)^2\cdot(1-p^2\cdot(3-2p))}{p^2\cdot(3-2p)}\\
%&-(6\cdot(1-6\cdot(1-p)^2\cdot p)\cdot(3p-1)\cdot(1-p)^2)\\
%&+ 9\cdot(1-p)^6+3\cdot(1-p)^3 \Big)\\
%&+ n_2^*\cdot\frac{(1-p)}{p}\\
%&= n_3^* \cdot \frac{(p-1)^2 \cdot (3p^2-p+1)}{p^2*(3-2p)} + n_2^*\frac{(1-p)}{p}\\
%\end{split}
%\end{equation}

\vspace*{4em}
\subsection{Variance Monotonicity}
\label{mo}
We now prove the monotonicity of the variance of our estimator.
\begin{theorem}
	\label{thm}
	$\mathbb{V}ar[LSHE]$ is monotonically decreasing when $p$ increases in  range $(0,1]$.
\end{theorem}

\begin{proof}
	\begin{lemma}
		\label{lemma}
		First order derivative of $\mathbb{V}ar[LSHE]$ is negative when $p \in (0,1]$.
	\end{lemma}
	\begin{proof}
	Consider
	\begin{align*}
	\dfrac{d(\mathbb{V}[LSHE])}{dp} &= \dfrac{3(2-p)(p-1)(p+1)((p-1)^2+p^2)}{p^3(2p-3)^2} \\ &\quad \cdot n_3^*
	-p^2 \cdot n_2^*.
	\end{align*}
	%%		
	When $p\in(0,1]$,  $-p^2 <0$. Because $(2-p), (p+1), (p-1)^2+p^2,p^3(2p-3)^2$ are all positive and $(p-1)$ is the only term that is negative, $\frac{3(2-p)(p-1)(p+1)((p-1)^2+p^2)}{p^3(2p-3)^2} < 0$. Thus, $\frac{d(\mathbb{V}ar[LSHE])}{dp} < 0$.
\end{proof}
	By using Lemma \ref{lemma}, we can consequently prove Theorem \ref{thm} is. We also note that when $p=1$, $\mathbb{V}ar[LSHE] = 0$.
\end{proof}
%
%
%\section{The DOPH Algorithm}
%\label{doph-alg}
%We first generate $L$ different meta-hash functions. Each of these meta-functions is formed by concatenating $K$ different Min-Wise hashes which are denoted as
%\begin{equation}
%\label{eq:bucket}
%B_i(x)  = [h_{i1}(x);h_{i2}(x);...;h_{{iK}}(x)],
%\end{equation}
%where $h_{ij}, i, j \in \{1,2,...,K \}$ are $K \times L$ different Min-Wise hashes.
%%\begin{figure}
%%\vspace{-0.7in}
%%\begin{center}
%%\mbox{
%%\includegraphics[width=5in]{resources/doctopdf/LSH_manyBuckets.pdf}
%%}
%%\end{center}
%%\vspace{-3.2in}
%%\caption{Illustration of hash table for the classical LSH algorithm}\label{fig:LSHtables}
%%%\vspace{-3.5in}
%%\end{figure}
%The algorithm can now be summarized by the following two steps:
%\begin{enumerate}
%\item {\bf Processing Phase ($O(ML)$):} We construct $L$ hash tables from the data by storing records $x \in \mathcal{C}$, at location $B_i(x)$ in hash table $i$.\footnote{We do not store the whole data vectors in the hash tables, or it would be memory consuming. We only store pointers or references to the vector.}
%\item {\bf Sampling Phase ($O(ML)$):} Given a queried record $R_q$, we report the union of all the pointers in the buckets $B_i(R_q)$ $\forall i \in \{1, 2, ..., L\}$ over $L$ hash tables. Note that we only probe $L$ different buckets, one from each hash table, rather than scanning all the elements in $\mathcal{C}$.
%\end{enumerate}
%
%Given the LSH property  $Pr(h(R_q) = h(R_i)) = \text{sim}(R_q,R_i)$, it is not difficult to show that with the above scheme, every pair $(R_i,R_j)$ is reported with probability precisely $1-(1-\text{sim}(R_i,R_j)^K)^L$.
%%With Min-Wise Hashing, the $Sim$ used is the Jaccard Similarity (see equation~\ref{eq:MinHash}).
%

%\begin{proof}
%	\begin{lemma}
%		\label{lemma}
%		First order derivative of $\mathbb{V}ar[Est]$ is negative when $p \in (0,1]$.
%	\end{lemma}
%	\begin{proof}
%		\begin{equation}
%		\begin{split}
%			\frac{d(\mathbb{V}ar[Est])}{dp} &= -\frac{3(2-p)(1-p)(p+1)((p-1)^2+p^2)}{p^3(2p-3)^2}\cdot n_3^*\\
%			&-p^2 \cdot n_2^*
%		\end{split}
%		\end{equation}
%		When $p\in(0,1]$, it is obvious that $-p^2 <0$. Because $(2-p)>(1-p)>0$, $-\frac{3(2-p)(1-p)(p+1)((p-1)^2+p^2)}{p^3(2p-3)^2} < 0$. In conclusion, $\frac{d(\mathbb{V}ar[Est])}{dp} < 0$.
%	\end{proof}
%	By using Lemma \ref{lemma}, we can consequently prove Theorem \ref{thm} is. Also noted that when $p=1$, $\mathbb{V}ar[Est] = 0$.
%\end{proof}

% \clearpage
%\newpage

%%\bibliography{RL}
%\bibliography{}
%\bibliographystyle{icml2017}

%\bibliography{references}
%\bibliographystyle{abbrv}

\clearpage
\newpage

\bibliographystyle{imsart-nameyear}
%\bibliography{chomp,mybib_merged}
\bibliography{mybib_merged,references,chomp}

\end{document}

### 2021-06-09 aoas-template/aoas-sample.tex
bekas stuff

\documentclass[aoas]{imsart}

%% Packages
\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage[authoryear]{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\RequirePackage{graphicx}

\startlocaldefs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Uncomment next line to change            %%
%% the type of equation numbering           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\numberwithin{equation}{section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Axiom, Claim, Corollary, Hypothezis, %%
%% Lemma, Theorem, Proposition              %%
%% use \theoremstyle{plain}                 %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\theoremstyle{plain}
\newtheorem{axiom}{Axiom}
\newtheorem{claim}[axiom]{Claim}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Assumption, Definition, Example,     %%
%% Notation, Property, Remark, Fact         %%
%% use \theoremstyle{remark}                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{remark}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{example}{Example}
\newtheorem*{fact}{Fact}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please put your definitions here:        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\endlocaldefs

\begin{document}

\begin{frontmatter}
\title{A sample article title}
%\title{A sample article title with some additional note\thanksref{t1}}
\runtitle{A sample running head title}
%\thankstext{T1}{A sample additional note to the title.}

\begin{aug}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Only one address is permitted per author. %%
%%Only division, organization and e-mail is %%
%%included in the address.                  %%
%%Additional information can be included in %%
%%the Acknowledgments section if necessary. %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[A]{\fnms{First} \snm{Author}\ead[label=e1]{first@somewhere.com}},
\author[B]{\fnms{Second} \snm{Author}\ead[label=e2,mark]{second@somewhere.com}}
\and
\author[B]{\fnms{Third} \snm{Author}\ead[label=e3,mark]{third@somewhere.com}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Addresses                                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address[A]{Department,
University or Company Name,
\printead{e1}}

\address[B]{Department,
University or Company Name,
\printead{e2,e3}}
\end{aug}

\begin{abstract}
Add in the abstract once the paper is set. 
\begin{enumerate}
\item BK: Follow the outline of Chen et al. (2018) regarding how an AoAS paper should be written given the Sadinle (2017) is written in a pretty methods based way. Sadinle (2014) is much more applied and what AoAS and JASA ACS like to see for an applied paper. 
\end{enumerate}
\end{abstract}

\begin{keyword}
\kwd{First keyword}
\kwd{second keyword}
\end{keyword}

\end{frontmatter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please use \tableofcontents for articles %%
%% with 50 pages and more                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\tableofcontents

\section{Introduction}

Some keys references that you'll want to address:

\cite{sadinle_detecting_2014, sadinle2017bayesian} are both methods of Sadinle. The most seminal method in probabilistic record linkage is that of \cite{fellegi1969theory}.

\subsection{Prior Work on Bipartite Matching for Record Linkage}

You only want to cite prior work on bipartite matching for record linkage in this section, so make sure to avoid my work as it's not comparable. You could mention it briefly given that it's been recently done, but that it would not fall under this framework. You would also want to mention that the first peer reviewed case study to our knowledge was done by Sadinle.

\subsection{The El Salavoran Conflict}

Talk about the data very simply.

\subsection{Our Contributions}

Outline your contributions. You might outline them from now regarding the fact that they can change.

\section{The Fellegi Sunter Method}

Needs to be reviewed and explained as Sadinle did in his 2017 paper. This should set out notation that you use for the rest of the paper.

\section{Bayesian Bipartite Matching for Record Linkage}

Need to review Mauricio's method, which should help outline any shortcoming of his method and notation you want to use for the rest of the paper.

\section{fabl: fast Bayesian Linkage}

Need to introduce your method in generality and explain your contributions. This is where you want things to be very powerful. You want to include proofs and showcase your work here. If you have insights moving forward --- then make sure to forward reference them.

\section{Simulation Studies}
In this section, we provide two simulation studies. First, we provide a simulation study, where we XXX. Second, we provide a simulation study, where we XXX.

\subsection{Title}

\subsection{Title}

\section{Applied to El Salavoran Conflict}

\section{Discussion}

\clearpage
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Example with single Appendix:            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{appendix}
\section*{Title}\label{appn} %% if no title is needed, leave empty \section*{}.
Appendices should be provided in \verb|{appendix}| environment,
before Acknowledgements.

If there is only one appendix,
then please refer to it in text as \ldots\ in the \hyperref[appn]{Appendix}.
\end{appendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Example with multiple Appendixes:        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{appendix}
\section{Title of the first appendix}\label{appA}
If there are more than one appendix, then please refer to it
as \ldots\ in Appendix \ref{appA}, Appendix \ref{appB}, etc.

\section{Title of the second appendix}\label{appB}
\subsection{First subsection of Appendix \protect\ref{appB}}

Use the standard \LaTeX\ commands for headings in \verb|{appendix}|.
Headings and other objects will be numbered automatically.
\begin{equation}
\mathcal{P}=(j_{k,1},j_{k,2},\dots,j_{k,m(k)}). \label{path}
\end{equation}

Sample of cross-reference to the formula (\ref{path}) in Appendix \ref{appB}.
\end{appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Support information, if any,             %%
%% should be provided in the                %%
%% Acknowledgements section.                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acks}[Acknowledgments]
The authors would like to thank the anonymous referees, an Associate
Editor and the Editor for their constructive comments that improved the
quality of this paper.
\end{acks}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Funding information, if any,             %%
%% should be provided in the                %%
%% funding section.                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{funding}
The first author was supported by NSF Grant DMS-??-??????.

The second author was supported in part by NIH Grant ???????????.
\end{funding}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supplementary Material, if any, should   %%
%% be provided in {supplement} environment  %%
%% with title and short description.        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{supplement}
\stitle{Title of Supplement A}
\sdescription{Short description of Supplement A.}
\end{supplement}
\begin{supplement}
\stitle{Title of Supplement B}
\sdescription{Short description of Supplement B.}
\end{supplement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  imsart-nameyear.bst  will be used to                   %%
%%  create a .BBL file for submission.                     %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%  MR numbers will be added by VTeX.                      %%
%%                                                         %%
%%  Use \cite{...} to cite references in text.             %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% if your bibliography is in bibtex format, uncomment commands:
\clearpage
\newpage
\bibliographystyle{imsart-nameyear}
\bibliography{fabl}

\end{document}

### 2021-06-09 template-methods/figures/plate-diagram.tex
bekas stuff

\begin{tikzpicture}
	\tikzset{
		hatch distance/.store in=\hatchdistance,
		hatch distance=8pt,
		hatch thickness/.store in=\hatchthickness,
		hatch thickness=2pt
  }

	\makeatletter
	\pgfdeclarepatternformonly[\hatchdistance,\hatchthickness]{flexible hatch}
	{\pgfqpoint{0pt}{0pt}}
	{\pgfqpoint{\hatchdistance}{\hatchdistance}}
	{\pgfpoint{\hatchdistance-1pt}{\hatchdistance-1pt}}%
	{
		\pgfsetcolor{\tikz@pattern@color}
		\pgfsetlinewidth{\hatchthickness}
		\pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
		\pgfpathlineto{\pgfqpoint{\hatchdistance}{\hatchdistance}}
		\pgfusepath{stroke}
	}
	\makeatother

	\tikzstyle{main}=[circle, minimum size=9mm, thick, draw=black!80, node distance=17mm]
	\tikzstyle{hyparam}=[rectangle, minimum size=5mm, thick, draw=black!80, fill=black!10, node distance=17mm]
	\tikzstyle{param}=[rectangle, minimum size=8mm, thick, draw=black!80, node distance=17mm]
	\tikzstyle{connect}=[-latex, thick]
	%\tikzstyle{selector}=[-latex, -|, snake=snake,segment amplitude=.4mm,segment length=2mm,line after snake=1mm, thick]
	\tikzstyle{shortconnect}=[-latex, thin]
	\tikzstyle{plate}=[rectangle, inner sep=4mm, yshift=-1mm, draw=black, rounded corners=6pt]
	\tikzstyle{highlight}=[draw=blue,text=blue, dash pattern=on 2pt off 2pt]
	\tikzstyle{platelab}=[anchor=south east, xshift=-0.4mm, yshift=0.4mm]
	\tikzstyle{switch}=[circle, minimum size=1mm, fill=black, draw=black]
	\node[param, fill=black!10] (alpha) {$\alpha_{a}$};
	\node[main] (theta) [below right of=alpha, xshift=10mm, yshift=6mm] {$\theta_{ta}$};
	\node[param, fill=black!10] (beta) [below left of=theta, xshift=-10mm, yshift=6mm] {$\beta_{a}$};
	\node[main] (z) [below of=theta] {$z_{tra}$};
	\node[main, pattern=flexible hatch, pattern color=black!10] (x) [below of=z] {$x_{tra}$};
	\node[main] (lambda) [below of=x, yshift=-1mm] {$\lambda_{tr}$};
	\node[main, highlight] (gamma) [below of=lambda] {$\gamma_{tr}$};
	\node[main] (y) [left of=x, xshift=-10mm, yshift=5mm] {$y_{ea}$};
	\node[param, fill=black!10] (phi) [above left of=y, yshift=3mm] {$\phi_{a}$};
	\node[param, highlight] (P) [left of=gamma, xshift=-10mm] {$\partset_{b}$};
	\node[param, fill=black!10, highlight] (eta) [right of=theta] {$\eta_{ta}$};
	\node[main, fill=black!10, highlight] (o) [below of=eta] {$o_{tra}$};
	\path (alpha) edge [connect] (theta)
	(beta) edge [connect] (theta)
	(theta) edge [connect] (z)
	(z) edge [connect] (x)
	(lambda) edge [connect] (x)
	(gamma) edge [connect, highlight, fill=blue] (lambda)
	(eta) edge [connect, highlight, fill=blue] (o)
	(y) edge [connect] (x)
	(y) edge [connect, highlight, fill=blue] (P)
	(P) edge [connect, highlight, fill=blue] (gamma)
  (P) edge [connect, highlight, fill=blue] (lambda)
	(phi) edge [connect] (y);
	\node[plate, fit=(y)] (plate-e) {};
	\node[platelab] at (plate-e.south east) (e) {\footnotesize$E$};
	\node[plate, fit=(z) (gamma) (x) (lambda) (o), inner sep=5.5mm] (plate-r) {};
	\node[platelab] at (plate-r.south east) {\footnotesize$R_t$};
	\node[plate, fit=(theta) (eta) (plate-r), inner sep=4mm] (plate-t) {};
	\node[platelab] at (plate-t.south east) {\footnotesize$T$};
  \node[plate, fit=(P), highlight] (plate-P) {};
  \node[platelab, text=blue] at (plate-P.south east) {\footnotesize$B$};
	\node[plate, fit=(phi) (alpha) (x) (eta) (plate-e), yshift=0mm, inner sep=4mm] (plate-a) {};
	\node[platelab] at (plate-a.south east) {\footnotesize$A$};
\end{tikzpicture}

### 2021-06-15 methods_paper/appendices.tex
nltcs and fabl paper

\documentclass[12pt,letterpaper]{article}

\usepackage{xr-hyper}       % cross-referencing (must be loaded before hyperref)
\usepackage[hypertexnames=false, draft]{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}

% Custom packages
\usepackage{amsthm, amssymb}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{tabularx}
\usepackage{tikz}           % vector graphics
\usepackage{bm}             % bold math symbols
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{import}
\usepackage{titling}
\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
  calc, shadows, plotmarks}

\graphicspath{{./app-figures/}}

% Custom commands
\let\oldvec\vec
\renewcommand\vec{\bm}
\newcommand{\simfn}{\mathtt{sim}} % similarity function
\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
\newcommand{\distfn}{\mathtt{dist}} % distance function
\newcommand{\valset}{\mathcal{V}} % attribute value set
\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
\newcommand{\euler}{\mathrm{e}} % Euler's constant
\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
\def\spacingset#1{\renewcommand{\baselinestretch}%
  {#1}\small\normalsize} \spacingset{1}

\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thealgorithm}{S\arabic{algorithm}}

\newtheorem*{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem*{definition}{Definition}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\externaldocument{unblinded-paper}

\title{\bf Appendices for ``d-blink: Distributed End-to-End Bayesian Entity Resolution''}
\author{Neil G.~Marchant\textsuperscript{a} \and
	Andee Kaplan\textsuperscript{b} \and 
	Daniel N.~Elazar\textsuperscript{c} \and
	Benjamin I.~P.~Rubinstein\textsuperscript{a} \and 
	Rebecca C.~Steorts\textsuperscript{d}}
\date{
	\textsuperscript{a}School of Computing and Information Systems, University 
	of Melbourne\\
	\textsuperscript{b}Department of Statistics, Colorado State University\\
	\textsuperscript{c}Methodology Division, Australian Bureau of Statistics\\
	\textsuperscript{d}Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
	\today}

\begin{document}
\maketitle

\newpage

\appendix
\section{Derivation of the posterior distribution}
\label{app-sec:full-posterior}
Here we sketch the derivation of the joint posterior distribution over the 
unobserved variables conditioned on the observed record attributes 
$\vec{X}^{(o)}$, which is given in 
Equation~\ref{eqn:partitioned-posterior} of the paper.
First we read the factorization off the plate diagram in 
Figure~\ref{fig:plate-diagram}, together with the conditional dependence 
assumptions detailed in Section~\ref{sec:model-specification} of the paper.
We obtain the following expression, up to a normalisation constant:
\begin{equation*}
\begin{split}
& p(\vec{\Gamma}, \vec{\Lambda}, \vec{Y}, \vec{Z}, \vec{\Theta}, \vec{X}^{(m)}|\vec{X}^{(o)}, \vec{O}) \propto 
  \prod_{e,a} p(y_{ea}|\phi_{a}) 
  \times \prod_{t,a} p(\theta_{ta}|\alpha_{a}, \beta_{a}) \\
& \quad {} \times 
  \prod_{t,r} \Big\{ p(\gamma_{tr}|\vec{Y})p(\lambda_{tr}|\gamma_{tr}, \vec{Y}) 
  \prod_{a} p(z_{tra}|\theta_{ta}) \Big\} 
  \times \prod_{\substack{t,r,a\\o_{tra}=1}} p(x_{tra}|z_{tra}, \lambda_{tr}, y_{\lambda_{tr}a}) \\
& \qquad {} \times 
  \prod_{\substack{t,r,a\\o_{tra}=0}} p(x_{tra}|z_{tra}, \lambda_{tr}, y_{\lambda_{tr}a}).
\end{split}
\end{equation*}
Ideally, we'd like to marginalize out all variables except $\vec{\Lambda}$ and 
$\vec{Y}$ (the variables of interest), however this is not tractable 
analytically.
Fortunately, we can marginalize out the missing record attributes 
$\vec{X}^{(m)}$ which yields Equation~\ref{eqn:partitioned-posterior} 
from the paper:
\begin{equation*}
\begin{split}
& p(\vec{\Gamma}, \vec{\Lambda}, \vec{Y}, \vec{Z}, \vec{\Theta}|\vec{X}^{(o)}, \vec{O}) \propto \prod_{e,a} p(y_{ea}|\phi_{a}) 
  \times \prod_{t,a} p(\theta_{ta}|\alpha_{a}, \beta_{a}) \\
& \quad {} \times 
  \prod_{t,r} \Big\{ p(\gamma_{tr}|\vec{Y})p(\lambda_{tr}|\gamma_{tr}, \vec{Y}) 
  \prod_{a} p(z_{tra}|\theta_{ta}) \Big\} 
  \times \prod_{\substack{t,r,a\\o_{tra}=1}} p(x_{tra}|z_{tra}, \lambda_{tr}, y_{\lambda_{tr}a}).
\end{split}
\end{equation*}

We can expand this further by substituting the conditional distributions 
given in Section~\ref{sec:model-specification} of the paper.
This yields:
\begin{equation}
\begin{split}
& p(\vec{\Gamma}, \vec{\Lambda}, \vec{Y}, \vec{Z}, \vec{\Theta}|\vec{X}^{(o)}, \vec{O}) \propto 
  \prod_{e,a} \phi_{a}(y_{ea}) \times \prod_{t,a} \theta_{ta}^{\alpha_{a} - 1} (1 - \theta_{ta})^{\beta_{a} - 1} \\
& \quad {} \times \prod_{t,r} \Big\{ \1{\lambda_{tr} \in 
  \partset_{\gamma_{tr}}(\vec{Y})} \prod_{a} \theta_{ta}^{z_{tra}} (1 - \theta_{ta})^{1-z_{tra}} \Big\} \\
& \mspace{40mu} {} \times  \prod_{\substack{t,r,a\\o_{tra}=1}} 
  \Big\{ (1 - z_{tra}) \1{x_{tra} = y_{\lambda_{tr}a}} 
  + z_{tra} \, \psi_{a}(x_{tra}|y_{\lambda_{tr}a}) \Big\}.
\end{split}
\label{app-eqn:expanded-posterior}
\end{equation}

\section{Equivalence of \texorpdfstring{\dblink}{d-blink} and \texorpdfstring{\blink}{blink}}
In this section, we present proofs of 
Propositions~\ref{thm:sim-dist-equiv} and~\ref{thm:posterior-equiv}, 
which show that the inferences we obtain from \dblink\ are equivalent to those 
we would obtain from \blink\ under certain conditions.

\subsection{Proof of Proposition~\ref{thm:sim-dist-equiv}: equivalence of 
distance\slash similarity representations}
\label{app-sec:proof-sim-dist-equiv}
It is straightforward to show that $\simfn$ as defined in 
Equation~\ref{eqn:simfn-distfn-correspondence} of the paper satisfies 
the requirements of Definition~\ref{def:attribute-sim-measure}.
All that remains is to show that the two parameterizations of the distortion 
distribution $\psi_{a}$ are equivalent.
Beginning with $\psi_{a}$ as parameterized in \blink, we substitute 
Equation~\ref{eqn:simfn-distfn-correspondence} and observe that
\begin{equation*}
\psi_{a}(v|w) \propto \phi_{a}(v) \euler^{-\distfn_{a}(v, w)} 
   = \phi_{a}(v) \euler^{d_\mathrm{max;a} + \simfn_{a}(v, w)} 
   \propto \phi_{a}(v) \euler^{\simfn_{a}(v, w)}.
\end{equation*}
This is identical to our parameterization in 
Equation~\ref{eqn:distortion-dist}. \qed

\subsection{Proof of Proposition~\ref{thm:posterior-equiv}: equivalence of 
\texorpdfstring{\dblink}{d-blink} and \texorpdfstring{\blink}{blink}}
\label{app-sec:proof-posterior-equiv}
Given that 
\begin{itemize}
  \item Proposition~\ref{thm:sim-dist-equiv} holds, 
  \item the distortion hyperparameters are the same for all attributes, and 
  \item all record attributes are observed,
\end{itemize}
the only factor in the posterior that differs from \blink\ is:
\begin{equation}
\prod_{t,r} p(\lambda_{tr}| \gamma_{tr}, \vec{Y}) 
  p(\gamma_{tr}|\vec{Y}).
\label{app-eqn:posterior-part-factor}
\end{equation}
Substituting the density for the conditional distributions 
for a single $t,r$ factor yields:
\begin{equation*}
p(\lambda_{tr}|\gamma_{tr},\vec{Y}) p(\gamma_{tr}|\vec{Y})
= \frac{\1{\lambda_{tr} \in \partset_{\gamma_{tr}}(\vec{Y})}}
{|\partset_{\gamma_{tr}}(\vec{Y})|} 
\times \frac{|\partset_{\gamma_{tr}}(\vec{Y})|}{E} = \frac{1}{E} \1{\lambda_{tr} \in \partset_{\gamma_{tr}}(\vec{Y})}.
\end{equation*}
Putting this in Equation~\ref{app-eqn:posterior-part-factor} and 
marginalizing over $\vec{\Gamma}$ we obtain:
\begin{equation*}
\prod_{t,r} \sum_{\gamma_{tr} = 1}^{B} p(\lambda_{tr}|\gamma_{tr},\vec{Y}) p(\gamma_{tr}|\vec{Y}) 
= \prod_{t,r} \frac{1}{E} \sum_{\gamma_{tr} = 1}^{B} \1{\lambda_{tr} \in \partset_{\gamma_{tr}}(\vec{Y})} 
= \prod_{t,r} \frac{1}{E} \1{\lambda_{tr} \in \{1,\ldots,E\}},
\end{equation*}
which is the factor that appears in the posterior for \blink. \qed

\section{Splitting rules for the \textit{k}-d tree blocking function}
\label{app-sec:splitting-rules}
In Section~\ref{sec:kd} of the paper we outline a blocking function inspired 
by $k$-d trees.
When inserting a node in the tree, we require a splitting rule that partitions 
the input set of values.
In ordinary $k$-d trees, the median is often used for this purpose, 
however it is not appropriate for the discrete input sets that we 
encounter.
As a result, we propose the following alternative splitting rules:
\begin{enumerate}
  \item \emph{Ordered median.}
  This rule is appropriate if the set of input attribute 
  values is large and\slash or has a natural ordering.
  If there is no natural ordering, an artificial ordering 
  must be applied (e.g. lexicographic ordering).
  The splitting rule is determined by sorting the input 
  values and finding the median, accounting for the 
  frequency of each value.
  Attribute values ordered before (after) the median are 
  passed to the left (right) child node.
  \item \emph{Reference set.}
  This rule is appropriate if the set of input attribute 
  values is small with no natural ordering.
  The splitting rule is determined by using a first-fit 
  bin-packing algorithm to split the values into two roughly 
  equal-sized bins, accounting for the frequency of each value.
  One of these bins is then labeled the ``reference set''.
  Attribute values (not) in the reference set are passed to 
  the left (right) child node.
\end{enumerate}

\section{Gibbs update distributions}
\label{app-sec:gibbs}
Here we list the conditional distributions for the Gibbs updates.
These are derived by referring to the posterior distribution in 
Equation~\ref{app-eqn:expanded-posterior}.

\subsection{Update for \texorpdfstring{$\theta_{ta}$}{distortion probabilities}}
\begin{equation}
\theta_{ta}|\vec{Z}, \vec{\Lambda}, \vec{\Gamma}, \vec{Y}, \vec{X}^{(o)}, \vec{O} \sim \operatorname{Beta}[z_{t \cdot a} + \alpha_{a}, R_t - z_{t \cdot a} + \beta_{a}]
\label{app-eqn:theta-update}
\end{equation}
where $z_{t \cdot a} := \sum_{r = 1}^{R_t} z_{tra}$.

\subsection{Update for \texorpdfstring{$z_{tra}$}{distortion indicators}}
\begin{equation}
  \begin{split}
  & z_{tra}|\vec{\Lambda}, \vec{\Gamma}, \vec{Y}, \vec{\Theta}, \vec{X}^{(o)}, \vec{O} \sim 
      (1-o_{tra}) \operatorname{Bernoulli}[\theta_{ta}] +
      o_{tra} \operatorname{Bernoulli}[\zeta_a(\theta_{ta}, x_{tra}, y_{\lambda_{tr}a})]
  \end{split}
  \label{app-eqn:z-update}
\end{equation}
where
$\zeta_a(\theta, x, y) = \begin{cases}
    1, & \text{if } x \neq y, \\
    \frac{\theta \psi_{a}(x|y)}{\theta \psi_{a}(x|y) - \theta + 1}, &\text{otherwise.}
  \end{cases}$

\subsection{Update for \texorpdfstring{$\lambda_{tr}$}{entity assignments}}
\begin{equation}
\begin{split}
& p(\lambda_{tr} | \vec{\Gamma}, \vec{Y}, \vec{\Theta}, \vec{Z}, \vec{X}^{(o)}, \vec{O}) \propto \\
& \qquad \1{\lambda_{tr} \in \partset_{\gamma_{tr}}(\vec{Y})} 
\prod_{\substack{a\\o_{tra}=1}} \Big\{ (1 - z_{tra}) \1{x_{tra} = y_{\lambda_{tr}a}} + z_{tra} \psi_{a}(x_{tra}|y_{\lambda_{tr}a})\Big\}.
\end{split}
\label{app-eqn:lambda-update}
\end{equation}

\section{Perturbation sampling algorithm}
\label{app-sec:pert-sampling}
In Proposition~\ref{thm:pert-sampling} of the paper, we show how to express a 
target pmf $p$ (from which we'd like to draw random variates) as a mixture over 
a base pmf $q$ and a perturbation pmf $v$.
Algorithm~\ref{app-alg:pert-sampling} demonstrates how to efficiently 
draw random variates from the target pmf using this mixture representation.

\begin{algorithm}[H]
  \caption{Perturbation sampling for $p(x|\omega)$}
  \label{app-alg:pert-sampling}
  \begin{algorithmic}[1]
    \Statex \textbf{Input:} 
    map from $x, \omega \in \mathcal{X}^\star \times \Omega \to 
    \epsilon(x|\omega)$; 
    map from $x \in \mathcal{X} \to q(x)$; 
    pre-initialized Alias sampler for $q$.
    \State $v \gets \emptyset$ 
    \Comment{empty map}
    \For {$x \in \mathcal{X}^\star$}
    \State $v(x) \gets q(x) \epsilon(x|\omega)$
    \EndFor
    \State $c \gets 1/\sum_{x \in \mathcal{X}^\star} v(x)$ 
    \Comment{normalization}
    \State $s \sim \mathrm{Bernoulli}\!\left[\frac{c}{1+c}\right]$
    \If {$s = 1$}
    \State \textbf{Return:} $x \sim q(\cdot)$ 
    \Comment{using input Alias sampler}
    \Else 
    \State $v \gets c \cdot v$
    \State \textbf{Return:} $x \sim v(\cdot)$
    \Comment{using new Alias sampler}
    \EndIf
  \end{algorithmic}
\end{algorithm}

\subsection{Proof of Proposition~\ref{thm:complexity-pert-sampling}: 
complexity of perturbation sampling}
Let us analyze the time complexity of Algorithm~\ref{app-alg:pert-sampling}.
Lines 2--6 are $O(|\mathcal{X}^\star|)$.
By properties of the Alias sampler~\citep{vose_linear_1991},
line 8 is $O(1)$ and line 11 is $O(|\mathcal{X}^\star|)$.
Thus the overall complexity is $O(|\mathcal{X}^\star|)$.

\section{Further details on the experimental set-up}
\label{app-sec:experiments-setup}
\subsection{Data sets}
We provide a brief description of each data set below. 
All data sets come with some form of ``ground truth'', which we 
use for evaluation purposes. 
However, the ground truth for \texttt{NCVR} and \texttt{SHIW0810} (two 
of the real data sets) may not be error-free as indicated below.

\begin{itemize}
  \item \texttt{ABSEmployee}. A synthetic data set used 
  internally for linkage experiments at the Australian Bureau of Statistics.
  It simulates an employment census and two supplementary 
  surveys (it is not derived from any real data sources).
  We used four categorical attributes: \texttt{MB}, \texttt{BDAY}, 
  \texttt{BYEAR} and \texttt{SEX}.
  \item \texttt{NCVR}. Two snapshots from the North Carolina 
  Voter Registration database taken two months 
  apart~\citep{christen_preparation_2014}.
  The snapshots are filtered to include only those voters 
  whose details changed over the two-month period.
  We used \texttt{first\_name}, \texttt{middle\_name} and 
  \texttt{last\_name} as string-type attributes; and 
  \texttt{age}, \texttt{gender} and \texttt{zip\_code} as 
  categorical attributes. 
  Unique voter identifiers are provided, however they are known to contain 
  some errors~\citep{christen_preparation_2014}.
  \item \texttt{NLTCS}. A subset of the U.S.\ National Long-Term 
  Care Survey~\citep{manton_nltcs_2010} comprising the 
  1982, 1989 and 1994 waves.
  It was necessary to use a subset, as race was subsampled in the other three 
  years, making it unsuitable for ER.
  We used four categorical attributes: \texttt{SEX}, \texttt{DOB}, 
  \texttt{STATE} and \texttt{REGOFF}.
  Unique identifiers are available which are known to be of high quality.
  \item \texttt{SHIW0810}. A subset from the Bank of Italy's 
  Survey on Household Income and Wealth~\citep{bancaitalia_2010} 
  comprising the 2008 and 2010 waves.
  We used eight categorical attributes: \texttt{IREG}, \texttt{SESSO}, 
  \texttt{ANASC}, \texttt{STUDIO}, \texttt{PAR}, \texttt{STACIV}, 
  \texttt{PERC} and \texttt{CFDIC}. 
  Unique identifiers were inferred using a deterministic algorithm, 
  which may not be error-free. 
  Further information and open-source code is provided at
  \url{http://github.com/ngmarchant/shiw}.
  \item \texttt{RLdata10000}. A synthetic data set provided 
  with the \texttt{RecordLinkage} R 
  package~\citep{sariyar_recordlinkage_2010}.
  We used \texttt{fname\_c1} and \texttt{lname\_c1} as string-type 
  attributes and \texttt{bd}, \texttt{bm}, \texttt{by} as categorical 
  attributes. 
  The \texttt{fname\_c2} and \texttt{lname\_c2} were excluded as they 
  have a high fraction of missing values.
\end{itemize}

\subsection{Implementation and hardware}
Our implementation of \dblink\ is written in Scala and depends on 
Apache Spark 2.3.1 (a distributed computing framework).
Since \dblink\ requires control over the partitioning (entities and 
linked records \emph{must} reside on their assigned partitions), we 
used the RDD API with a custom partitioner.
Our custom-built server ran in local (pseudo-cluster) mode, with
$2 \times$ 28-core Intel Xeon Platinum 8180M CPUs for a total of 112 
threads (with HyperThreading); and 128GB of allocated RAM on the driver.

\subsection{Parameter settings and initialization} 
\label{app-sec:parameter-settings}
We used the following parameter settings for all experiments.
\begin{itemize}
  \item The distortion hyperparameters $\alpha_{a}, \beta_{a}$ were set to 
  encode a prior mean distortion probability of approximately 1\%, with the 
  strength varying in proportion to the total number of records $R$:
  \begin{equation*}
  \alpha_{a} = R \times 10\% \times 1\% \text{ and }
    \beta_{a} = R \times 10\% \text{ for all } a.
  \end{equation*}
  \item The size of the latent entity population $E$ was set to $R$.
  This corresponds to a prior mean number of observed entities of 
  $(1 - \euler^{-1}) R \approx 0.63 R$, as shown by 
  \citet{steorts_bayesian_2016}.
  It is important not to set $E$ too low, as it places an upper bound 
  on the number of entities present in the data. 
  \item The entity attribute distributions $\{\phi_a\}$ were set empirically
  based on the observed record attributes. Specifically, we set 
  \begin{equation*}
    \phi_a(v) = \frac{\sum_{t = 1}^{T} \sum_{r = 1}^{R_t} o_{tra} \1{x_{tra} = v}}
      {\sum_{t = 1}^{T} \sum_{r = 1}^{R_t} o_{tra}} \quad \text{for all } a.
  \end{equation*}
  \item For simplicity, we treated all attributes as either 
  ``categorical-type''  with similarity function $\simfn_{\mathrm{const}}$ or 
  ``string-type'' with similarity function $10.0 \times \simfn_{\mathrm{nEd}}$ 
  (these are defined in Section~\ref{sec:attribute-sim-measure}).
  \item The similarity cut-off for string-type attributes was set to 7.0, 
  following advice in the \texttt{RecordLinkage} R 
  package~\citep{sariyar_recordlinkage_2010}.
  \item We used the $k$-d tree blocking function as defined in 
  Section~\ref{sec:kd}.
  The \emph{reference set} splitting rule was used for input sets with 30 or 
  fewer elements---the \emph{ordered median} splitting rule was used 
  otherwise.
\end{itemize}

To initialize the Markov chain, we linked each record to a unique entity and 
copied the record attributes into the entity attributes, assuming no 
distortion.
Any entity attributes that were missing after this process (due to missing 
record attributes) were filled by drawing an attribute value from the 
empirical distribution.
We set the thinning interval to 10---i.e. we only saved every tenth step along 
the chain.
This increases the effective sample size for a given storage budget.

\section{Results on Amazon EC2}
\label{app-sec:cloud-results}
We repeated two of the experiments described in 
Section~\ref{sec:efficiency-expts} of the main paper on a cluster 
running in the Amazon Elastic Compute Cloud (EC2).
For the worker (executor) nodes, we used varying numbers of \texttt{m5.xlarge} 
instances with 4 vCores, 16 GiB memory and 32 GiB of Elastic Block Store 
(EBS) storage. 
Due to the increased latency and decreased bandwidth between the 
compute nodes, we expected the efficiency to decrease.
This is indeed what we observed.

Figure~\ref{app-fig:speed-up-vs-num-partitions} plots the speed-up as a function 
of the number of blocks $B$ relative to a baseline with no partitioning.
We observe poorer scaling with $B$ compared to the results we obtained 
on our local server (c.f.~Figure~\ref{fig:speed-up-vs-num-partitions} 
in the main paper).
Figure~\ref{app-fig:speed-up-vs-sampler} plots the efficiency as a function of 
the sampling method with $B=16$.
The results are qualitatively similar to the ones we obtained using our local 
server (c.f.~Figure~\ref{fig:speed-up-vs-sampler} in the main paper).
However, the ESS rate was reduced for all samplers as expected due to 
increased communication costs.

\begin{figure}
  \centering
  \includegraphics{plot-partitions-speed-up-aws.pdf}
  \caption{Efficiency of \dblink\ as a function of the
    number of blocks $B$ and summary statistic of interest 
    (larger is better).
    The speed-up measures the ESS rate relative to the ESS rate 
    for $B = 1$ (no partitioning) for the \texttt{NLTCS} data set.}
  \label{app-fig:speed-up-vs-num-partitions}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{plot-sampler-speed-up-aws.pdf}
  \caption{Efficiency of \dblink\ as a function of the 
    sampler and summary statistic of interest (larger is better).
    All measurements are for the \texttt{NLTCS} data set with 
    $B = 16$.
    }
  \label{app-fig:speed-up-vs-sampler}
\end{figure}

\section{Balance of the blocks}
\label{app-sec:balance-partitions}
In Section~\ref{sec:kd}, we proposed a blocking function based on $k$-d 
trees, and argued that it could yield balanced blocks with good entity 
separation.
While running \dblink\ with the $k$-d tree blocking function, we recorded the 
size of the blocks ($|\partset_{b}|$ for all $b$) to assess whether 
they were well-balanced.
Figure~\ref{app-fig:partition-sizes} illustrates the results in terms of the 
relative absolute deviation from the perfectly balanced configuration
(where the entities are divided equally among the blocks). 
We can see that the $k$-d tree partitioner is functioning quite well---the 
deviation from the perfectly balanced configuration is no more than 10\% for 
all data sets.

\begin{figure}
  \centering
  \includegraphics{partition-sizes-plot.pdf}
  \caption{Balance of the blocks for a single run on each
    data set.
    The balance is measured in terms of the relative absolute 
    deviation from the perfectly balanced configuration.
    The number of blocks $B = 64, 64, 16, 2, 8$ for each data set (in the 
    order listed in the legend).
  }
  \label{app-fig:partition-sizes}
\end{figure}

\section{Uncertainty measures}
\label{app-sec:error-num-ents}
\dblink\ allows for measures of uncertainty to be reported, unlike the 
baseline methods, since we have the full posterior distribution.
For example, in Figure~\ref{app-fig:error-num-ents} we compute posterior estimates 
for the number of entities present in each data set, with 95\% 
Bayesian credible intervals.
Note that the posterior estimates are typically quite 
sharp.
This seems to confirm arguments by~\citet{steorts_bayesian_2016} 
regarding the informativeness of the prior for the linkage 
structure in \blink. 
Research on less informative priors is ongoing~\citep{zanella_flexible_2016}.

\begin{figure}
  \centering
  \includegraphics{posterior-bias-plot-dark.pdf}
  \caption{Percentage error in the posterior\slash prior estimates for the 
    number of observed entities for \dblink.
    The posterior estimates are generally sharp and underestimate the true 
    number of observed entities.
  }
  \label{app-fig:error-num-ents}
\end{figure}

%\color{blue}
\section{Sensitivity analysis}
\label{app-sec:sensitivity}
We conducted an empirical sensitivity analysis for \dblink\ 
using the \texttt{RLdata10000} data set.
We selected this data set as it is relatively small, which made it 
quick to run the inference for various hyperparameter combinations.
The parameters tested were:
\begin{itemize}
  \item $\alpha_\ell, \beta_\ell$: the shape parameters for the Beta prior on 
  the distortion probabilities. 
  We used the same values for all attributes ($a$). 
  \item $E$: the size of the latent population.
  \item $s_{\mathrm{max}}$: the scaling factor for the similarity function. 
  This controls the inverse temperature of the softmax distribution for the 
  distorted attribute values. 
\end{itemize}
We varied each of these parameters in turn, while holding all other parameters 
fixed.
For the Beta prior on the distortion probabilities, we first varied the 
strength while fixing the prior mean to $\sim 1\%$, then we varied the mean 
(1\%, 5\% and 10\%) while fixing $\alpha + \beta$ (related to the strength). 
Table~\ref{tbl:sensitivity} presents the evaluation measures for each 
combination of parameters.
The results indicate that the inferred linkage structure is relatively 
sensitive to all of the parameters, however sensitivity is in general 
predictable, following clear and intuitive trends.
Of particular interest is the fact that the model performs best when the 
Beta prior on the distortion probabilities is sharply peaked near zero. 
It seems that the model has a tendency to overestimate the amount of 
distortion, particularly in the absence of ground truth.

\begin{table}
	\centering
  \caption{Sensitivity analysis for various parameters combinations 
  using \texttt{RLdata10000}. The first group of rows tests the effect of 
  varying the \emph{strength} of the Beta prior, the second group tests the 
  effect of varying the \emph{mean} of the Beta prior, the third group 
  tests the effect of varying the population size, and the fourth group 
  tests the effect of varying the scaling factor for the similarity function.}
	\label{tbl:sensitivity}
	\spacingset{1}
  \footnotesize
  \begin{center}
	\begin{tabular}{*{9}{c}}
		\toprule
		\multicolumn{2}{c}{Distortion} & Pop.\ size & Max.\ sim. & \multicolumn{3}{c}{Pairwise measures} & \multicolumn{2}{c}{Cluster measures} \\
		\cmidrule(lr){1-2} \cmidrule(lr){3-3} \cmidrule(lr){4-4} \cmidrule(lr){5-7} \cmidrule(lr){8-9}
		$\alpha$       & $\beta$          & $E$            & $s_{\mathrm{max}}$ & Precision & Recall & F1-score & ARI & Err. \# clust.\ \\
		\midrule 
    \textbf{0.1}   & \textbf{10.0}    & 10000          & 10.0          & 0.5342 & 0.9990 & 0.6962 & 0.6962 & $-17.47\%$ \\
    \textbf{1.0}   & \textbf{100.0}   & 10000          & 10.0          & 0.5435 & 0.9990 & 0.7040 & 0.7040 & $-16.58\%$ \\
    \textbf{10.0}  & \textbf{1000.0}  & 10000          & 10.0          & 0.6334 & 0.9970 & 0.7747 & 0.7747 & $-10.97\%$ \\
    \textbf{100.0} & \textbf{10000.0} & 10000          & 10.0          & 0.9180 & 0.9850 & 0.9503 & 0.9503 & $-1.595\%$ \\
    \midrule
    \textbf{10.0}  & \textbf{1000.0}  & 10000          & 10.0          & 0.6334 & 0.9970 & 0.7747 & 0.7747 & $-10.97\%$ \\
    \textbf{50.5}  & \textbf{959.5}   & 10000          & 10.0          & 0.6132 & 0.9970 & 0.7593 & 0.7593 & $-11.90\%$ \\
    \textbf{101.0} & \textbf{909.0}   & 10000          & 10.0          & 0.5992 & 0.9970 & 0.7485 & 0.7485 & $-12.90\%$ \\
    \midrule
    10.0           & 1000.0           & \textbf{9000}  & 10.0          & 0.5306 & 0.9970 & 0.6926 & 0.6926 & $-15.65\%$ \\
    10.0           & 1000.0           & \textbf{10000} & 10.0          & 0.6334 & 0.9970 & 0.7747 & 0.7747 & $-10.97\%$ \\
    10.0           & 1000.0           & \textbf{11000} & 10.0          & 0.6999 & 0.9960 & 0.8221 & 0.8221 & $-7.365\%$ \\
    \midrule
    10.0           & 1000.0           & 10000          & \textbf{5.0}  & 0.6927 & 0.9940 & 0.8164 & 0.8164 & $-22.12\%$ \\
    10.0           & 1000.0           & 10000          & \textbf{10.0} & 0.6334 & 0.9970 & 0.7747 & 0.7747 & $-10.97\%$ \\
    10.0           & 1000.0           & 10000          & \textbf{50.0} & 0.2112 & 0.3920 & 0.2745 & 0.2745 & $-12.50\%$ \\
		\bottomrule
  \end{tabular}
  \end{center}
\end{table}

\section{Details of inference for the case study to the 2010 Decennial Census}
\label{app-sec:census-mcmc}
We ran inference for 15,000 iterations using the PCG-I sampler.
After removing 5,000 iterations as burn-in and applying thinning with 
an interval of 10, we obtained 1,000 approximate samples from the 
posterior.
Convergence diagnostics are consistent with those reported for 
the other data sets in Appendix~\ref{app-sec:trace-plots}, 
and are complicated to release due to the fact that the data is 
protected under Title~13. 
Releasing each iteration of a Gibbs sampler could potentially say something 
about individuals in the population, and thus, for privacy reasons, these 
diagnostics are omitted.

\pagebreak
\section{Trace plots}
\label{app-sec:trace-plots}
\subsection{Attribute-level distortion}
The following figures relate to the aggregate distortion per attribute 
for each data set.
On the left are the trace plots, which show the aggregate distortion 
for each attribute (stacked vertically) along the Markov chain.
On the right are the corresponding autocorrelation plots.

\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{ABSEmployee_attribute-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{ABSEmployee_attribute-distortions-acf}
  \caption{Attribute-level distortion for \texttt{ABSEmployee}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{NCVR_attribute-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{NCVR_attribute-distortions-acf}
  \caption{Attribute-level distortion for \texttt{NCVR}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{NLTCS_attribute-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{NLTCS_attribute-distortions-acf}
  \caption{Attribute-level distortion for \texttt{NLTCS}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{SHIW_attribute-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{SHIW_attribute-distortions-acf}
  \caption{Attribute-level distortion for \texttt{SHIW0810}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{RLdata10000_attribute-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{RLdata10000_attribute-distortions-acf}
  \caption{Attribute-level distortion for \texttt{RLdata10000}}
\end{figure}

\subsection{Distribution of record distortion}
The following figures relate to the distribution of record distortion 
for each data set.
Specifically, we count the number of records with 0 distorted attributes, 
1 distorted attribute, 2 distorted attributes, etc.
On the left are the trace plots, which show the record counts for each 
distortion level (stacked vertically) along the Markov chain.
On the right are the corresponding autocorrelation plots.

\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{ABSEmployee_record-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{ABSEmployee_record-distortions-acf}
  \caption{Distribution of record distortion for \texttt{ABSEmployee}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{NCVR_record-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{NCVR_record-distortions-acf}
  \caption{Distribution of record distortion for \texttt{NCVR}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{NLTCS_record-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{NLTCS_record-distortions-acf}
  \caption{Distribution of record distortion for \texttt{NLTCS}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{SHIW_record-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{SHIW_record-distortions-acf}
  \caption{Distribution of record distortion for \texttt{SHIW0810}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{RLdata10000_record-distortions} \hfill
  \includegraphics[width=0.48\linewidth]{RLdata10000_record-distortions-acf}
  \caption{Distribution of record distortion for \texttt{RLdata10000}}
\end{figure}

\subsection{Cluster size distribution}
The following figures relate to the distribution of cluster (entity) sizes 
for each data set.
Specifically, we count the number of entities with 0 linked records, 
1 linked record, 2 linked records, etc.
On the left are the trace plots, which show the counts for each cluster size 
(stacked vertically) along the Markov chain.
On the right are the corresponding autocorrelation plots.

\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{ABSEmployee_cluster-size-distribution} \hfill
  \includegraphics[width=0.48\linewidth]{ABSEmployee_cluster-size-distribution-acf}
  \caption{Cluster size distribution for \texttt{ABSEmployee}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{NCVR_cluster-size-distribution} \hfill
  \includegraphics[width=0.48\linewidth]{NCVR_cluster-size-distribution-acf}
  \caption{Cluster size distribution for \texttt{NCVR}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{NLTCS_cluster-size-distribution} \hfill
  \includegraphics[width=0.48\linewidth]{NLTCS_cluster-size-distribution-acf}
  \caption{Cluster size distribution for \texttt{NLTCS}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{SHIW_cluster-size-distribution} \hfill
  \includegraphics[width=0.48\linewidth]{SHIW_cluster-size-distribution-acf}
  \caption{Cluster size distribution for \texttt{SHIW0810}}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=0.48\linewidth]{RLdata10000_cluster-size-distribution} \hfill
  \includegraphics[width=0.48\linewidth]{RLdata10000_cluster-size-distribution-acf}
  \caption{Cluster size distribution for \texttt{RLdata10000}}
\end{figure}

%\let\noopsort\undefined
%\let\printfirst\undefined
%\let\singleletter\undefined
%\let\switchargs\undefined
{
\footnotesize
\bibliographystyle{jasa}
\bibliography{dblink}
}

\end{document}

### 2021-06-15 methods_paper/fabl-paper.tex
nltcs and fabl paper

--- 2021-06-06 paper/parlr_paper_1.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
-%
-\documentclass[
-  12pt,
-]{article}
-\usepackage[]{mathpazo}
-\usepackage{setspace}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
+\documentclass[12pt,letterpaper]{article}
+
+\usepackage[hypertexnames=false]{hyperref}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{enumerate}
+\usepackage{natbib}
+
+% Custom packages
+\usepackage{amsthm, amssymb}
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{algorithm}      % algorithm environment
+\usepackage{algpseudocode}
+\usepackage{multirow}
+\usepackage{sectsty}
+\usepackage{tabularx}
+\usepackage{tikz}           % vector graphics
+\usepackage{bm}             % bold math symbols
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  colorlinks=true,
-  linkcolor=blue,
-  filecolor=Maroon,
-  citecolor=Blue,
-  urlcolor=Blue,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\usepackage{hyperref}
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
+\usepackage{microtype}
+\usepackage{import}
+\usepackage{titling}
+\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
+  calc, shadows, plotmarks}
 
-\author{}
-\date{\vspace{-2.5em}}
+\graphicspath{{./notes/figures/}}
+
+% Custom commands
+\let\oldvec\vec
+\renewcommand\vec{\bm}
+\newcommand{\simfn}{\mathtt{sim}} % similarity function
+\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
+\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
+\newcommand{\distfn}{\mathtt{dist}} % distance function
+\newcommand{\valset}{\mathcal{V}} % attribute value set
+\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
+\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
+\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
+\newcommand{\euler}{\mathrm{e}} % Euler's constant
+\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
+\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+  {#1}\small\normalsize} \spacingset{1}
+
+\newtheorem*{remark}{Remark}
+\newtheorem{proposition}{Proposition}
+\newtheorem*{definition}{Definition}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+\sectionfont{\large\nohang\centering\MakeUppercase}
+
+\title{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}
+\author{Brian Kundinger\textsuperscript{a} \and
+  Jerome Reiter\textsuperscript{a} \and 
+  Rebecca C.~Steorts\textsuperscript{b}}
+\date{
+  \textsuperscript{a}Department of Statistical Science, Duke University\\
+  \textsuperscript{b}Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\
+  \today}
 
 \begin{document}
+\maketitle
 
-\setstretch{1}
-\begin{flushright} 
-    \end{flushright}
-    \begin{center} \textbf{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}
+\bigskip
+\begin{abstract}
+Abstract
+\end{abstract}
 
-    Brian Kundinger, Rebecca Steorts, and Jerome Reiter
+\noindent%
+{\it Keywords:} record linkage, Bayesian methods, paralellized computing, streaming algorithms
 
-    \end{center}
+\newpage
+\spacingset{1.5}
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\section{Introduction}
+\label{sec:introduction}
 
 Entity resolution (or record linkage) is the task of identifying
-duplicate records across multiple data sources. This task is
-straightforward when provided unique identifiers or highly reliable
+duplicate records across multiple data sources.
+In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years.
+In others, we seek to perform some other statistical procedure, such as survival analysis or regression, using linked datasets, and one must  propagate linkage uncertainty into the subsequent analysis.
+
+Record linkage is straightforward when provided unique identifiers or highly reliable
 information, but becomes difficult when data sources are plagued with
-error. Additionally, many entity resolution tasks are large in nature,
-and many existing methods are too computational intensive to be used in
-practice. In this paper, we propose a Bayesian model for record linkage
-that uses hashing techniques and efficient storage mechanisms, and is
-amenable to parallelization, making it scalable to large linkage tasks.
+error.
+This motivates the use of probabilistic record linkage techniques that are able to model the reliability of the information and which are more robust to errors than are deterministic methods.
+Unfortunately, probabilistic record linkage is computationally intense, and many entity resolution tasks are large in nature,
+motivating the search for scalable and efficient record linkage techniques.
 
-The proposed method is inspired by the \emph{Beta Record Linkage}
-(\texttt{BRL}) method proposed in Mauricio Sadinle's 2017 ``Bayesian
-Estimation of Bipartite Matching for Record Linkage''. To recognize this
-lineage, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
-pronounced ``fable''). In what follows, Section 2 describes previous
-work in record linkage that lays the foundation for our method, Section
+In this paper, we propose an extention to the \emph{Beta Record Linkage}
+(\texttt{BRL}) method proposed proposed by \citep{sadinle2017} for this very end.
+We increase the speed of the \texttt{BRL} procedure through the modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
+Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks.
+To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
+pronounced ``fable'')
+
+In what follows, Section 2 describes previous work in record linkage that lays the foundation for our method, Section
 3 provides our model and the Gibbs sampler used for posterior inference,
@@
 interesting properties of the method through a case study of El
-Salvadoran casuality records. Lastly, our discussion highlights open
-questions in the record linkage liturature and motivates future work.
+Salvadoran casualty records. Lastly, our discussion highlights open
+questions in the record linkage literature and motivates future work.
 
-\hypertarget{related-work}{%
-\subsection{Related Work}\label{related-work}}
+\section{Related work}
+\label{sec:related}
 
+\paragraph{Fellegi and Sunter Method}
 Most record linkage techniques are derived from the seminal 1969 paper
@@
 
-One key strength of Sadinle's \texttt{BRL} is that it is garunteed to
-produce a set of bipartite matchings that respect one-to-one
-assumptions--without the need for any post-hoc procedures. Specifically,
-in each iteration of his Gibbs sampler, he considers each record
-\(j\in B\), removes from consideration the records \(i\in A\) that have
+\paragraph{Sadinle's Method for Bipartite Matching} \texttt{BRL} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
+Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have
 already been matched, and then samples a potential link. However,
@@
 
-Outside of the Fellegi-Sunter framework, some researchers have developed
+\paragraph{Deterministic Blocking}
+
+\paragraph{Other Approach (Will probably remove)}Outside of the Fellegi-Sunter framework, some researchers have developed
 methods that model the data directly, not a derived set of comparison
@@
 
-\hypertarget{notation-and-assumptions}{%
-\subsection{Notation and Assumptions}\label{notation-and-assumptions}}
+\section{A scalable model for Bayesian Record Linkage}
+\label{sec:model}
 
@@
 \hypertarget{model-specification}{%
-\subsection{Model Specification}\label{model-specification}}
+	\subsection{Model Specification}
+	\label{model-specification}}
 
@@
 \hypertarget{posterior-sampling}{%
-\subsection{Posterior Sampling}\label{posterior-sampling}}
+	\subsection{Posterior Sampling}
+	\label{posterior-sampling}}
+We work with the following factorization of the joint posterior distribution:
 
-We work with the following factorization of the joint distribution:
-
-\[p(\Gamma, \mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) = p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)\]
+\[p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) \propto p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)\]
 
@@
 \hypertarget{bayes-estimate}{%
-\subsection{Bayes Estimate}\label{bayes-estimate}}
+	\subsection{Bayes Estimate}
+	\label{bayes-estimate}}
 
@@
 
-\hypertarget{efficient-computation}{%
-\section{Efficient Computation}\label{efficient-computation}}
-
+\section{Efficient and Scalable Implementation}
+\label{sec:efficiency}
 Broadly speaking, we increase our computational efficiency by
@@
 
-\hypertarget{gibbs-sampling}{%
-\subsection{Gibbs Sampling}\label{gibbs-sampling}}
+\section{Efficient Posterior Inference}\label{sec:efficient-posterior}
 
@@
 
-\hypertarget{simulation-study-1---accuracy}{%
-\section{Simulation Study 1 -
-Accuracy}\label{simulation-study-1---accuracy}}
+\section{Simulation Studies}
+\label{sec:simulations}
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 
-{\centering \includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot}
+	{\centering
 
+		\includegraphics{../notes/figures/sadinle_sim_plot} 
 	}
@@
 	\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
-For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:unnamed-chunk-1}
+For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:sadinle_simulation}
+
 \end{figure}
-
-\hypertarget{simulation-study-2---speed}{%
-\section{Simulation Study 2 - Speed}\label{simulation-study-2---speed}}
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 
-{\centering \includegraphics[width=17.75in]{../notes/figures/sadinle_speed_plot2}
+	{\centering \includegraphics{../notes/figures/sadinle_speed_plot2}
 
@@
 
-\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:unnamed-chunk-2}
+	\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:speed1}
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[t]
 
-{\centering \includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB}
+	{\centering \includegraphics{../notes/figures/speed_plot_fixed_nB}
 
@@
 
-\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:unnamed-chunk-3}
+	\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:speed2}
 \end{figure}
@@
 
-\hypertarget{data-analysis}{%
-\section{Data Analysis}\label{data-analysis}}
+\section{Case Studies}
+\label{case-studies}
 
-The country of El Salvador was immersed in civil war from 1980 to 1991,
+We now demonstrate the power of our method through two case studies.
+The first is a study of documented identifiable deaths (DID) from the El Salvadoran Civil War.
+Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be acheived by simplifying the construction of the comparison vectors.
+Secondly, the case study reveals the impact of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
+The second case study is the National Long Term Care Survey (NLTCS), a much larger linkage tasks which \texttt{BRL} is unable to complete but which we complete with \texttt{fabl} with relative ease.
+
+\paragraph{El Salvadoran Civil War}The country of El Salvador was immersed in civil war from 1980 to 1991,
 and throughout the time, several organizations attempted to document
@@
 
-\begin{figure}
+\begin{figure}[t]
 
-{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/m_posterior_smallP}
+	{\centering \includegraphics{../notes/figures/el_salvador/m_posterior_smallP}
 
@@
 
-\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:unnamed-chunk-4}
+	\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
 \end{figure}
@@
 \(\mathbf{u}\) paramters under each method are very similar, as shown in
-Figure XX.
+Figure~\ref{fig:m-and-u}.
 
-\hypertarget{violations-of-one-to-one-matching}{%
-\subsection{Violations of One-to-One
-Matching}\label{violations-of-one-to-one-matching}}
+\subsection{Violations of One-to-One Matching}
+\label{violations-of-one-to-one-matching}
 
-Figure XX shows the posterior distribution for \(D\), the number of
+Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
 duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
@@
 
-\begin{figure}
+\begin{figure}[t]
 
-{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+	{\centering \includegraphics{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
 
@@
 	\caption{Posterior distribution of overlap across the two files.
-The solid lines show the Bayes estimate for the amonut of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:unnamed-chunk-5}
+The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:overlap-plot}
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[t]
 
-{\centering \includegraphics[width=6.51in]{../notes/figures/el_salvador/rosa_records}
+	{\centering \includegraphics{../notes/figures/el_salvador/rosa_records}
 
@@
 
-\caption{Example of linkage situation with multiple plausible matches}\label{fig:unnamed-chunk-6}
+	\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[t]
 
-{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/bad_mixing}
+	{\centering \includegraphics{../notes/figures/el_salvador/bad_mixing}
 
@@
 
-\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:unnamed-chunk-7}
+	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
 \end{figure}
 
-The records shown in Figure XX provide one such instance. Note these
+The records shown in Figure~\ref{fig:rosa-maria} provide one such instance. Note these
 records present a near violation of our assumption that there are no
@@
 resulting in a Gibbs process with poor mixing. Additionally,
-\texttt{fabl} allows the modeller to inspect records with mulitple
+\texttt{fabl} allows the modeler to inspect records with multiple
 plausible matches, and if they desire, to then choose the record pairing
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\paragraph{National Long Term Care Survey} The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
+It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989.
+The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office.
+State and regional office have been anomynized, presented as numbers instead of their actual values.
+We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context.
+The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466 files$, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns.
+We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.
+
+With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible.
+Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold.
+Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller datasets.
+We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing.
+We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler.
+
+Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computatonal complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
+However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
+Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conficts are resolved to ensure one-to-one matching.
+
+Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
+With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the seperate blocks.
+This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler.
+However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error.
+In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office.
+We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives.
+Even if not, the modeller may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures.
+However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons.
+Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block.
+Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.
+
+\section{Discussion}
+\label{discussion}
 
@@
 
+\bigskip
+
+\begin{center}
+{\large\bf ACKNOWLEDGEMENTS}
+\end{center}
+
+\bibliographystyle{jasa}
+{
+\footnotesize
+\bibliography{references}
+}
+
 \end{document}

### 2021-06-15 R/CompareBinary.R

@@ -1,0 +1,18 @@
+CompareBinary <- function(df1, df2, cols){
+  nA <- dim(df1)[1]
+  nB <- dim(df2)[1]
+  ids <- expand.grid(1:nA, 1:nB)
+
+  indicators <- lapply(cols, function(x){
+    match <- df1[ids[,1], x] == df2[ids[,2], x]
+    cbind(match, 1 - match)
+  })
+
+  indicators <- do.call(cbind, indicators)
+  cd <- list(comparions = indicators,
+             n1 = nA,
+             n2 = nB,
+             nDisagLevs = rep(2, length(cols)),
+             Ztrue = NULL)
+  cd
+}

### 2021-06-15 R/GetUniquePatterns3.R

@@ -1,43 +1,47 @@
 GetUniquePatterns3 <- function(cd, fast = F, R = NULL){
 
-  gamma <- cd[[1]]
-  N <- dim(gamma)[1]
-  fields <- dim(gamma)[2]
+  indicators <- apply(cd[[1]], 2, as.numeric)
+  N <- dim(indicators)[1]
+  fields <- dim(indicators)[2]
   n1 <- cd[[2]]
   n2 <- cd[[3]]
-  levels <- cd[[4]]
-  possible_patterns <- GetPossiblePatterns(levels)
 
   ids <- expand.grid(1:n1, 1:n2)
   rec1 <- ids[,1]
   rec2 <- ids[,2]
+  levels <- cd[[4]]
+  possible_patterns <- GetPossiblePatternsSad(levels)
+  P <- length(possible_patterns)
 
-  df <- data.frame(gamma, rec1, rec2)
+  df <- data.frame(indicators, rec1, rec2)
 
+  #this step is slow. maybe use data.table?
   pattern_df <- df %>%
-    unite(pattern, 1:fields, sep = "")
+    tidyr::unite(pattern, 1:fields, sep = "")
+
+  thing <- unique(pattern_df$pattern)
 
 
 
+  # hash_id <- vector(length = N)
+  # for(i in seq_along(possible_patterns)){
+  #   hash_id[pattern_df$pattern == possible_patterns[i]] <- i
+  # }
+  # hash_id <- factor(hash_id)
 
-  # unique_hash <- unique(pattern_df$pattern)
-   P <- length(possible_patterns)
-
-  hash_id <- vector(length = n1 * n2)
-  for(i in 1:P){
-    hash_id[pattern_df$pattern == possible_patterns[i]] <- i
-  }
-
-  hash_id <- factor(hash_id)
+  hash_id <- factor(pattern_df$pattern,
+                   levels = possible_patterns,
+                   labels = 1:P)
 
 
-  temp <- data.frame(gamma, rec1, rec2, hash_id)
+  #used to have indicators here too)
+  temp <- data.frame(rec1, rec2, hash_id)
   pattern_counts <- temp %>%
-    group_by(hash_id) %>%
+    group_by(hash_id, .drop = F) %>%
     count() %>%
     pull()
 
-  #unique_patterns <- gamma[!duplicated(hash_id),]
+  #unique_patterns <- indicators[!duplicated(hash_id),]
 
   counts_by_rec <- temp %>%
[…]

### 2021-06-15 R/NLC_hash2.R

@@ -180,7 +180,7 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   Z.SAMPS <- Z.SAMPS[,-(1:burn)]
   L.SAMPS <- L.SAMPS[-(1:burn)]
   M.SAMPS <- M.SAMPS[,-(1:burn)]
-  #U.SAMPS <- U.SAMPS[,-(1:burn)]
+  U.SAMPS <- U.SAMPS[,-(1:burn)]
 
   final_gibbs <- apply(Z.SAMPS, 2, function(z){
     unlist(imap(z, ~sample_with_1(hash_to_rec1[[.y]][[.x]], 1)))
@@ -415,7 +415,7 @@ BKSimple_hash2_big <- function(comparisons, m_prior = 1, u_prior = 1,
     rep(x, comparisons[[1]][[4]][x])
   })))
 
-  ids <- expand.grid(1:n1, 1:n2)
+  #ids <- expand.grid(1:n1, 1:n2)
   #rec2 <- ids[,2]
 
   #hash_id <- patterns[[1]]
@@ -548,9 +548,8 @@ BKSimple_hash2_big <- function(comparisons, m_prior = 1, u_prior = 1,
     # hash$Z.temp <- Z.temp
 
     Z.SAMPS[,s] <- Z
-    #M.SAMPS[,s] <- m
-    #U.SAMPS[,s] <- u
-    L.SAMPS[s] <- L
+    M.SAMPS[,s] <- m
+    U.SAMPS[,s] <- u
 
     if(show_progress){
       if (s %% (S / 100) == 0) {
@@ -563,14 +562,15 @@ BKSimple_hash2_big <- function(comparisons, m_prior = 1, u_prior = 1,
     parallel::stopCluster(cl)
   }
   Z.SAMPS <- Z.SAMPS[,-(1:burn)]
-  #L.SAMPS <- L.SAMPS[-(1:burn)]
-  #M.SAMPS <- M.SAMPS[,-(1:burn)]
-  #U.SAMPS <- U.SAMPS[,-(1:burn)]
+  M.SAMPS <- M.SAMPS[,-(1:burn)]
+  U.SAMPS <- U.SAMPS[,-(1:burn)]
 
   final_gibbs <- apply(Z.SAMPS, 2, function(z){
     unlist(imap(z, ~sample_with_1(hash_to_rec1[[.y]][[.x]], 1)))
   })
 
[…]

### 2021-06-20 methods_paper/fabl-paper.tex
fabl method paper first draft

@@
 \title{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}
-\author{Brian Kundinger\textsuperscript{a} \and
-  Jerome Reiter\textsuperscript{a} \and 
-  Rebecca C.~Steorts\textsuperscript{b}}
+\author{Brian Kundinger \and
+  Jerome Reiter \and 
+  Rebecca C.~Steorts}
 \date{
-  \textsuperscript{a}Department of Statistical Science, Duke University\\
-  \textsuperscript{b}Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\
+  Department of Statistical Science, Duke University\\
   \today}
@@
 
-In this paper, we propose an extention to the \emph{Beta Record Linkage}
-(\texttt{BRL}) method proposed proposed by \citep{sadinle2017} for this very end.
+In this paper, we propose an extension to the \emph{Beta Record Linkage}
+(\texttt{BRL}) method proposed by \citep{sadinle2017} for this very end.
 We increase the speed of the \texttt{BRL} procedure through the modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
@@
 \paragraph{Fellegi and Sunter Method}
-Most record linkage techniques are derived from the seminal 1969 paper
-by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
+Most record linkage techniques are derived from the seminal \citep{fs} paper "A Theory for Record Linkage". The defining
 characteristic of their approach was to transform the sets of records,
@@
 text similarity score, and bin these scores to integers for use in the
-model. Jaro (1989) provided guidelines for appropriate binning for the
+model. \citep{jaro1989} provided guidelines for appropriate binning for the
 comparison vectors, but details of implementation are generally context
@@
 that break expectations of transitivity, which is undesirable. Although
-Jaro devised an procedure to reduce a set of conflicting matches to the
+\citep{jaro1995} devised an procedure to reduce a set of conflicting matches to the
 mostly likely set of one-to-one matchings, the abundance of false
@@
 
-\paragraph{Deterministic Blocking}
-
-\paragraph{Other Approach (Will probably remove)}Outside of the Fellegi-Sunter framework, some researchers have developed
+\paragraph{Other Approaches (Will probably remove)}Outside of the Fellegi-Sunter framework, some researchers have developed
 methods that model the data directly, not a derived set of comparison
-vectors. As an early example, Steorts et al 2016 produced
+vectors. As an early example,  produced
 \texttt{blink}, which had the advantage of a computational complexity
@@
 sampler \emph{inherently serial}. We however use independent priors for
-each \(Z_j\), creating a sampler that is \emph{perfectly
-	parallelizable}, allowing for significant computational gains. More
+each \(Z_j\), creating a sampler that is \emph{perfectly parallelizable}, allowing for significant computational gains. More
 importantly, since only the agreement pattern of \(Z_j\) is used for
@@
 \texttt{BRL}; our sampler does ensure that each record in \(B\) can be
-matched to at most one record in \(A\), but allows for the possiblity
+matched to at most one record in \(A\), but allows for the possibility
 that multiple records in \(B\) match to the same record in \(A\). We
@@
 
-In order to make fair comparisons against the Sadinle 2017 model, we
+In order to make fair comparisons against the \citep{sadinle2017} model, we
 integrate over the posterior of \(\lambda\) and rearrange terms to
@@
 For a more indepth explanation of this function and the induced Bayes
-estimate, see Sadinle 2017.
+estimate, see \citep{sadinle2017}.
 
@@
 patterns rather than the \(n_A \times n_B\) record pairs.
+Some of these techniques were used by \citep{enamorado2019} to produce \texttt{fastlink} a fast a scalable implementation of the Fellegi-Sunter model, and others have been created here for our particular context.
 
@@
 bijective transformation that does not change the meaning of the data,
-but this representation will ease calculations and posterior updates.
+but this representation will ease calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.
 
@@
 
-Laslty, for large data, we can partition the two datasets \(A\) and
+Lastly, for large data, we can partition the two datasets \(A\) and
 \(B\) into smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more
@@
 patterns through the second equation. Sampling uniformly is
-computational simple even for large sets of candidate records, but this
+computationally simple even for large sets of candidate records, but this
 step can also be parallelized when working with large data.
@@
 
+\begin{figure}
 	\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
@@
 	\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
+
+	\caption{Definitions of performance metrics}\label{fig:performance_metrics}
+\end{figure}
 
@@
 
-		\includegraphics{../notes/figures/sadinle_sim_plot} 
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot} 
 	}
@@
 consider only exact matching, so a vector (1, 0) corresponds to
-agreeement and (0, 1) to disagreement. We simulate these data for
+agreement and (0, 1) to disagreement. We simulate these data for
 different values of \(n_A\) and \(n_B\), and compare the run-time of
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 	\centering
@@
 	\end{tabular}
+\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{fig:distributions}
 \end{table}
@@
 
-	{\centering \includegraphics{../notes/figures/sadinle_speed_plot2}
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2}
 
@@
 
-	\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:speed1}
+	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:speed1}
 \end{figure}
@@
 
-	{\centering \includegraphics{../notes/figures/speed_plot_fixed_nB}
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB}
 
@@
 
-	\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:speed2}
+	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:speed2}
 \end{figure}
@@
 The first is a study of documented identifiable deaths (DID) from the El Salvadoran Civil War.
-Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be acheived by simplifying the construction of the comparison vectors.
+Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
 Secondly, the case study reveals the impact of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
@@
 
-\paragraph{El Salvadoran Civil War}The country of El Salvador was immersed in civil war from 1980 to 1991,
+\subsection{El Salvadoran Civil War}
+\label{el_salvador}
+
+The country of El Salvador was immersed in civil war from 1980 to 1991,
 and throughout the time, several organizations attempted to document
 casualties of the conflict. When estimating the total number of
-casualities, one cannot simply sum the numbers recorded by each
+casualties, one cannot simply sum the numbers recorded by each
 organization, as it is likely that the same individuals are recorded in
-multiple casuality lists. To obtain a more accurate estimate of the
+multiple casualty lists. To obtain a more accurate estimate of the
 casualties then, we follow Sadinle's 2017 paper and link records of El
@@
 in the Spanish language, and then compared using a modified Levenstein
-distance to account for the fact that second names are often ommited.
+distance to account for the fact that second names are often omitted.
 Place of birth is recorded by municipality and department within that
@@
 
-	{\centering \includegraphics{../notes/figures/el_salvador/m_posterior_smallP}
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP}
 
@@
 realized in the data. This leads us to believe that such high number of
-aggreement levels creates unnecesary distictions in the data and makes
+agreement levels creates unnecessary distinctions in the data and makes
 the comparison vectors less interpretable. Therefore we re-ran our
@@
 the way that the computational complexity of our method depends on the
-number of unique agreement patterns, and how significant computional
+number of unique agreement patterns, and how significant computational
 gains can be made by simplifying the construction of the comparison
 vectors. We also note that estimates of the \(\mathbf{m}\) and
-\(\mathbf{u}\) paramters under each method are very similar, as shown in
+\(\mathbf{u}\) parameters under each method are very similar, as shown in
 Figure~\ref{fig:m-and-u}.
 
-\subsection{Violations of One-to-One Matching}
-\label{violations-of-one-to-one-matching}
-
-Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
+\paragraph{Violations of One-to-One Matching} Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
 duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
@@
 
-	{\centering \includegraphics{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
 
@@
 an initial Bayes estimate of 195 matches found across files, and after
-resolving matches that violated one-to-one requirements, yeilded 180
+resolving matches that violated one-to-one requirements, yielded 180
 matches across files. This is acceptably close to the 181 matches found
@@
 
-	{\centering \includegraphics{../notes/figures/el_salvador/rosa_records}
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records}
 
@@
 
-	{\centering \includegraphics{../notes/figures/el_salvador/bad_mixing}
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing}
 
@@
 
-Figure XX shows the values that \(Z_{825}\) and \(Z_{826}\) take on
+Figure~\ref{fig:mixing-plot} shows the values that \(Z_{825}\) and \(Z_{826}\) take on
 throughout the Gibbs sampler, and demonstrates how each method handles
@@
 
-\paragraph{National Long Term Care Survey} The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
+\subsection{National Long Term Care Survey}
+\label{nltcs}
+
+The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
 It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989.
 The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office.
-State and regional office have been anomynized, presented as numbers instead of their actual values.
+State and regional office have been anonymized, presented as numbers instead of their actual values.
 We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context.
-The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466 files$, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns.
+The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466$ files, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns.
 We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.
@@
 
-Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computatonal complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
+Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
 However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
-Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conficts are resolved to ensure one-to-one matching.
+Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
 
 Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
-With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the seperate blocks.
+With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks.
 This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler.
@@
 We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives.
-Even if not, the modeller may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures.
+Even if not, the modeler may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures.
 However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons.
@@
 In our case study, we included an exploration of how to conduct record
-linkage when modelling assumptions are not met in practice. We explored
+linkage when modeling assumptions are not met in practice. We explored
 ``one-to-many'' scenarios in which one record in \(A\) has multiple
@@
 there is duplication both across and within datasets. Tuning
-\texttt{fabl} for use in these scenarious is one potential avenue for
+\texttt{fabl} for use in these scenarios is one potential avenue for
 future work.
@@
 \begin{center}
-{\large\bf ACKNOWLEDGEMENTS}
+{\large\bf ACKNOWLEDGMENTS}
+
+Left over from dblink paper, might remove.
 \end{center}

### 2021-09-16 methods_paper/fabl-paper.tex
traceplots and more

@@
 
-\title{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}
-\author{Brian Kundinger \and
-  Jerome Reiter \and 
-  Rebecca C.~Steorts}
+\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+\author{Brian Kundinger\textsuperscript{a} \and
+  Jerome Reiter\textsuperscript{a} \and 
+  Rebecca C.~Steorts \textsuperscript{b}}
 \date{
-  Department of Statistical Science, Duke University\\
+ \textsuperscript{a} Department of Statistical Science, Duke University \\
+ \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
   \today}
@@
 
-Entity resolution (or record linkage) is the task of identifying
+Record linkage is the task of identifying
 duplicate records across multiple data sources.
@@
 
-Record linkage is straightforward when provided unique identifiers or highly reliable
-information, but becomes difficult when data sources are plagued with
-error.
+Record linkage is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error.
 This motivates the use of probabilistic record linkage techniques that are able to model the reliability of the information and which are more robust to errors than are deterministic methods.
@@
 
-In this paper, we propose an extension to the \emph{Beta Record Linkage}
-(\texttt{BRL}) method proposed by \citep{sadinle2017} for this very end.
-We increase the speed of the \texttt{BRL} procedure through the modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
+Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}.
+This method creates comparison vectors for each pair of records in the data, and independently classifies those pairs as a match or a non-match.
+Though this independence is usually not reasonable, it is appealing for its mathematical simplicity and relative computational ease.
+Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}.
+Importantly, \citep{enamorado2019} modified the Fellegi-Sunter method with innovative hashing techniques, greatly expanding the scalability of the independent classification approach.
+
+Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file.
+This setting is known as bipartite matching, and has received much less attention in the literature.
+\citep{sadinle2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing.
+However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.
+
+In this paper, we propose an extension to the \texttt{BRL} method proposed by \citep{sadinle2017} for this very end.
+We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
 Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks.
@@
 
-In what follows, Section 2 describes previous work in record linkage that lays the foundation for our method, Section
+In what follows, Section~\ref{sec:related} describes previous work in record linkage that lays the foundation for our method, Section~\ref{sec:model}
 3 provides our model and the Gibbs sampler used for posterior inference,
-and Section 4 describes strategies used for more efficient and scalable
-implementation. Then, Section 5 demonstrates the accuracy and speed of
-out method through two simulation studies, and Section 6 reveals
+and Section~\ref{sec:efficiency} describes strategies used for more efficient and scalable
+implementation. Then, Section~\ref{sec:simulations} demonstrates the accuracy and speed of
+out method through two simulation studies, and Section~\ref{sec:case-studies} reveals
 interesting properties of the method through a case study of El
-Salvadoran casualty records. Lastly, our discussion highlights open
+Salvadoran casualty records. Lastly, Section~\ref{sec:discussion} is a discussion that highlights open
 questions in the record linkage literature and motivates future work.
@@
 mostly likely set of one-to-one matchings, the abundance of false
-matches can bias the estimation of other parameters of the model.
+matches can bias the estimation of other parameters of the model, and ultimately leads to poor linkage performance.
 
-\paragraph{Sadinle's Method for Bipartite Matching} \texttt{BRL} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
-Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have
-already been matched, and then samples a potential link. However,
+\citep{sadinle2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
+Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link.
+This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach.
+However,
 accounting for these dependencies throughout the linkage process is
@@
 
-\paragraph{Other Approaches (Will probably remove)}Outside of the Fellegi-Sunter framework, some researchers have developed
+Outside of the Fellegi-Sunter framework, some researchers have developed
 methods that model the data directly, not a derived set of comparison
-vectors. As an early example,  produced
-\texttt{blink}, which had the advantage of a computational complexity
+vectors. As an early example,  produced \texttt{blink}, which had the advantage of a computational complexity
 that grew linearly with the data (as opposed to quadratically within
@@
 distributed computing to manage even moderate linkage problems in
-acceptable time. Because these methods fall outside of the
-Fellegi-Sunter framework, we do not consider them further in this paper.
+acceptable time.
+These methods perform entity resolution over arbitrary numbers of datasets with arbitrary amounts of duplication and fall outside the Fellegi-Sunter framework, so we do not discuss them further here.
 
@@
 
-Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
+In this section, we propose a statistical model suitable for efficient computation in large linkage tasks. Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
 respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
@@
 	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
-\end{cases}\] This provides more memory efficient storage for the
+\end{cases}\] This parameter is vector of length $n_B$, providing more memory efficient storage for the
 linkage information than a \(n_A \times n_B\) sparse matrix of
@@
 agreement on another. Though this assumption is often not reasonable
-(for example, first name and gender are two clearly dependent fields),
+(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
 but it is common within the record linkage literature and generally
@@
 matched to at most one record in \(A\), but allows for the possibility
-that multiple records in \(B\) match to the same record in \(A\). We
+that multiple records in \(B\) match to the same record in \(A\). We then resolve these contradictions through a straightfoward post-processing step. We
 explore the ramifications of this distinction with the El Salvador case
@@
 \hypertarget{posterior-sampling}{%
-	\subsection{Posterior Sampling}
-	\label{posterior-sampling}}
+	\subsection{Gibbs Sampler}
+	\label{gibbs_sampling}}
 We work with the following factorization of the joint posterior distribution:
 
-\[p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) \propto p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)\]
-
+\begin{align*}
+p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) \propto p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)
+\end{align*}
 This factorization leads to following Gibbs Sampler:
@@
 
-For a more indepth explanation of this function and the induced Bayes
+For a more in-depth explanation of this function and the induced Bayes
 estimate, see \citep{sadinle2017}.
@@
 resolve these conflicts by accepting only the match with highest
-posterior probability. Such a resolution procedure indeed is equivalent
+posterior probability.
+A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities.
+Such a resolution procedure indeed is equivalent
 to minimizing posterior risk under the restriction of one-to-one
@@
 processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
-and enumerate them \(h_1, \ldots, h_P\). When the \((i,j)\) pair
+and enumerate them \(h_1, \ldots, h_P\), and map record pairs to this hashed value. When the \((i,j)\) pair
 exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
@@
 patterns rather than the \(n_A \times n_B\) record pairs.
-Some of these techniques were used by \citep{enamorado2019} to produce \texttt{fastlink} a fast a scalable implementation of the Fellegi-Sunter model, and others have been created here for our particular context.
+Some of these techniques are similar to those used by \citep{enamorado2019} to produce \texttt{fastlink}, a fast a scalable implementation of the Fellegi-Sunter model, and others have been created here for our particular context.
 
@@
 
+We first hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019} accomplished this efficiently through the hashing function
+$$\tilde{\gamma}_{ij} = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$
+This function maps each agreement pattern to a unique integer, allowing us to store a scalar quanity instead of an entire vector for each record pair.
+For computational ease, we then map these integers to sequential integers from 1 to P.
+
 The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
-comparison vector as vector of length \(F\), with each component
+comparison as a vector of length \(F\), with each component
 \(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). To ease
-computations, we instead represent the comparison as a concatenation of
-\(F\) many binary indicator vectors of lengths \(L_f\). For example, if
+computations, we instead use a one hot encoding of the comparison vector. For example, if
 \(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
@@
 bijective transformation that does not change the meaning of the data,
-but this representation will ease calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.
+but this representation eases calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.
 
@@
 vector for a different \((i,j)\) pair. In contrast, we do not store
-these comparison vectors themselves, but instead only a hashed value
+these comparison vectors themselves, but instead only the hashed value $h_p$
 corresponding to the agreement pattern of the \((i, j)\) pair. We store
@@
 
-To address this, we propose a new strategy called \emph{storage
-	efficient indexing} (SEI). In standard indexing, the modeler decides a
+To address this, we propose a new strategy called \emph{storage	efficient indexing} (SEI). In standard indexing, the modeler decides a
 certain criteria that they expect all true matching pairs to satisfy,
@@
 contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
-vector is simply \(H_p^m \times h_p\). Thus our posterior update for
-\(\alpha\) is simply
+vector is simply \(H_p^m \times h_p\). Thus our posterior update for the
+\(\alpha\) vector is simply
 \(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
@@
 
-In his 2017 paper, Sadinle demonstrated the accuracy of his method by
-running \texttt{BRL} on simulated pairs of files with differing levels
-of errors and overlap. Each file consists of first name, last name,
-gender, postal code, age, and occupation, and following Sadinle's
-implementation, we use only four fields for the linkage process: both
-names, age, and occupation. Similarity scores for the names are binned
-into three levels according the default settings for the
-\texttt{compareRecords} function in the \texttt{BRL} package, and age
-and occupation similarity scores are binary for exact matching. We use
-flat priors for all \(\mathbf{m}\) and \(\mathbf{u}\) parameters, and
-run the Gibbs sampler for 1000 iterations, with the first 100 discarded
-as burn-in.
+We first compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
+We use first name, last name, age, and occupation for this linkage, and create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package.
+Each simulation identifies duplicated individuals between two datasets, each with 500 records.
+We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets.
+We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
+This is a near exact replication of the simulation study provided by \citep{sadinle2017}.
 
-Comparing our method against \texttt{BRL} on these same simulated
-datasets, we find that our method only has weakened performance in the
+In cases when there are only one or two errors in matching records, and in cases with low to moderate duplication across records, we see that \texttt{fabl} provides approximately equivalent accuracy as \texttt{BRL}.
+We find that our method only has weakened performance in the
 most extreme scenario of very high errors and very high overlap across
-files. We use the following measures of accuracy in our analysis:
-although using only recall and precision is common in the literature, we
-find it useful to have a single metric to gauge overall performance, and
-thus also include F measure.
+files.
+In these situations, \texttt{BRL} is removing large numbers of records from consideration throughout the Gibbs Sampler, making its implementation most different from \texttt{fabl}.
+We note however that such extreme linkage tasks, which such high amounts of errors and overlap, are extremely rare in practice.
 
-\begin{figure}
-	\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
-	\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
-	\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
-
-	\caption{Definitions of performance metrics}\label{fig:performance_metrics}
-\end{figure}
+%\begin{figure}
+%	\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
+%	\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
+%	\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
+%	
+%	\caption{Definitions of performance metrics}\label{fig:performance_metrics}
+%\end{figure}
 
@@
 problem. Distributions are meant to emulate the behavior of similarity
-scores across first name, last name, and day, month. For simplicity, we
+scores across first name, last name, and day, month. For example, $u^{\text{month, 1}} = P(\text{Records have same birth-month | Nonmatch}) = \frac{1}{12}$. For simplicity, we
 consider only exact matching, so a vector (1, 0) corresponds to
@@
 		\hline
-		fname & (0.95, 0.05) & (0.01, 0.99) \\ 
-		lname & (0.95, 0.05) & (0.01, 0.99) \\ 
-		day & (0.95, 0.05) & (0.03, 0.97) \\ 
-		month & (0.95, 0.05) & (0.08, 0.92) \\ 
-		year & (0.95, 0.05) & (0.07, 0.93) \\ 
+		fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
+		lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
+		day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
+		month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
+		year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
 		\hline
@@
 
-	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:speed1}
+	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$.
+We see near quadratic growth in runtime for BRL, and near linear growth for fabl.}\label{fig:speed1}
 \end{figure}
@@
 
-	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:speed2}
+	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
+We see linear growth in runtime for BRL, and near constant runtime for fabl.}\label{fig:speed2}
 \end{figure}
@@
 
-\begin{center}
-{\large\bf ACKNOWLEDGMENTS}
+\section{Appendix}
+\label{sec:appendix}
 
-Left over from dblink paper, might remove.
-\end{center}
+\begin{figure}[t]
+
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace}
+
+	}
+
+	\caption{Traceplot for number of matches found across datasets in El Salvador case study}\label{fig:overlap_trace}
+\end{figure}
+
+\begin{figure}[t]
+
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace}
+
+	}
+
+	\caption{Traceplot for m parameter in El Salvador case study}\label{fig:m_trace}
+\end{figure}
+
+\begin{figure}[t]
+
+	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace}
+
+	}
+
+	\caption{Traceplot for u parameter in El Salvador case study}\label{fig:u_trace}
+\end{figure}
+
+\bigskip
 

### 2021-09-16 R/NLC_hash2.R

@@ -1,6 +1,7 @@
 BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
                           alpha = 1, beta = 1, S = 1000, burn = 100,
-                          show_progress = T, fast = F, R = NULL){
+                          show_progress = T, fast = F, R = NULL,
+                          all_patterns = TRUE){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -23,7 +24,7 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   }
 
   ptm <- proc.time()
-  patterns <- GetUniquePatterns2(comparisons, fast, R)
+  patterns <- GetUniquePatterns2(comparisons, fast, R, all_patterns)
   elapsed_hash <- proc.time() - ptm
 
   parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
@@ -177,10 +178,10 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   if (fast){
     parallel::stopCluster(cl)
   }
-  Z.SAMPS <- Z.SAMPS[,-(1:burn)]
-  L.SAMPS <- L.SAMPS[-(1:burn)]
-  M.SAMPS <- M.SAMPS[,-(1:burn)]
-  U.SAMPS <- U.SAMPS[,-(1:burn)]
+  # Z.SAMPS <- Z.SAMPS[,-(1:burn)]
+  # L.SAMPS <- L.SAMPS[-(1:burn)]
+  # M.SAMPS <- M.SAMPS[,-(1:burn)]
+  # U.SAMPS <- U.SAMPS[,-(1:burn)]
 
   final_gibbs <- apply(Z.SAMPS, 2, function(z){
     unlist(imap(z, ~sample_with_1(hash_to_rec1[[.y]][[.x]], 1)))

### 2021-09-18 R/MakeComparisons.R
added install_github command

@@ -1,7 +1,0 @@
-# n <- 50
-# ids <- expand.grid(1:n, 1:n)
-# df1 <- sample(1:5, n, replace = TRUE)
-# df2 <- sample(1:5, n, replace = TRUE)
-# df1[ids[,1]] == df2[ids[,2]]
-
-

### 2021-09-18 R/SynthesizeHashChunks.R

@@ -1,0 +1,24 @@
+SythensizeHash <- function(patterns_hashed){
+  counts_big <- Reduce(`+`, map(patterns_hashed, ~.x[[2]]))
+
+  counts_per_rec_big <- patterns_hashed %>%
+    map(`[[`, 3) %>%
+    flatten()
+
+  hash_list_big <- patterns_hashed %>%
+    map(`[[`, 4) %>%
+    flatten()
+
+  comparisons <- list(indicators = NULL,
+                      n1 = dim(big_file2)[1],
+                      n2 = dim(big_file2)[1],
+                      nDisagLevs = c(4, 4, 2, 2),
+                      Ztrue = NULL)
+
+  patterns <- list(patterns_hashed[[1]][[1]],
+                   counts_big,
+                   counts_per_rec_big,
+                   hash_list_big)
+
+  cd <- list(comparisons, patterns)
+}

### 2021-09-20 methods_paper/fabl-paper.tex
ran large sadinle simulation

@@
 3 provides our model and the Gibbs sampler used for posterior inference,
-and Section~\ref{sec:efficiency} describes strategies used for more efficient and scalable
+and Sections~\ref{sec:efficiency} and~\ref{sec:efficient-posterior} describe strategies used for more efficient and scalable
 implementation. Then, Section~\ref{sec:simulations} demonstrates the accuracy and speed of
@@
 
-Outside of the Fellegi-Sunter framework, some researchers have developed
-methods that model the data directly, not a derived set of comparison
-vectors. As an early example,  produced \texttt{blink}, which had the advantage of a computational complexity
-that grew linearly with the data (as opposed to quadratically within
-Fellegi Sunter), but was plagued with slow mixing times and was unable
-to incorporate text data. Later, Steorts et al 2020 improved on this
-method with \texttt{dblink}, which was able to simulate text data within
-its Gibbs sampler by drawing from an empirical distribution, and
-remedied its slow mixing through a probabilistic blocking approach.
-However, this method remains computationally intensive and requires
-distributed computing to manage even moderate linkage problems in
-acceptable time.
+%Outside of the Fellegi-Sunter framework, some researchers have developed
+%methods that model the data directly, not a derived set of comparison
+%vectors. As an early example,  produced \texttt{blink}, which had the advantage of a computational complexity
+%that grew linearly with the data (as opposed to quadratically within
+%Fellegi Sunter), but was plagued with slow mixing times and was unable
+%to incorporate text data. Later, Steorts et al 2020 improved on this
+%method with \texttt{dblink}, which was able to simulate text data within
+%its Gibbs sampler by drawing from an empirical distribution, and
+%remedied its slow mixing through a probabilistic blocking approach.
+%However, this method remains computationally intensive and requires
+%distributed computing to manage even moderate linkage problems in
+%acceptable time.
 These methods perform entity resolution over arbitrary numbers of datasets with arbitrary amounts of duplication and fall outside the Fellegi-Sunter framework, so we do not discuss them further here.
@@
 
-\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
+%\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
+
+$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$
 
@@
 \begin{align*}
-p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) \propto p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)
+p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda) \\
+&\propto \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right] \\
+&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
+&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_A)\frac{1}{n_A}\lambda + I(Z_j = n_A + 1)(1 - \lambda)\right]
 \end{align*}
@@
 
-\[\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
-\[\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
+\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
+\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
 where
-\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
+\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I(\gamma_{ij}^f = l) I(z_j = i)\)
 and
-\(\beta_{fl}(\mathbf{Z})= \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).
+\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + I(\gamma_{ij}^f = l) I(z_j \neq i)\).
 
@@
 \begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\\
-	&\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} && \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\\
+	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
+	&\propto \prod_{i=1}^{n_A}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
 	&=
@@
 where
-\(w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
+\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
 The interested reader should note that these are precisely the
@@
 
+To aid the reader, we provide summary of the \texttt{fabl} method through pseudocode:
+
+\begin{algorithm}
+	\caption{Summary of fabl algorithm}
+	\begin{algorithmic}[1]
+
+		\Procedure{Hashing and Preprocessing}{}
+		\State Partition files $A$ and $B$ into chunks $\{A_I\}, \{B_J\}$
+		\For{each $I$, $J$}
+		\State Create comparison vectors between $A_I$ and $B_J$
+		\State Hash results and calculate summary statistics
+		\State Use SEI to reduce memory usage
+		\EndFor
+		\State Synthesize results across pairings
+		\EndProcedure
+
+		\Procedure{Gibbs Sampling}{}
+		\State Initialize $m$, $u$, and $Z$ parameters
+
+		\For{$t \in \{1, \ldots, T\}$} 
+		\State Sample $m^{t+1}|Z^{t}, \Gamma$ and  $u^{t+1}|Z^{t}, \Gamma$ 
+		\State Sample $H\left(Z^{t+1}\right)|m^{t+1}, u^{t+1}, \Gamma$  \Comment{Sample agreement pattern, not record}
+		\EndFor
+		\State Sample $Z | H(Z), \Gamma$ \Comment{Fills in record label based on agreement pattern}
+
+		\EndProcedure
+
+	\end{algorithmic}
+\end{algorithm}
+
 \section{Simulation Studies}
@@
 \section{Case Studies}
-\label{case-studies}
+\label{sex:case-studies}
 
@@
 
-\subsection{National Long Term Care Survey}
-\label{nltcs}
-
-The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
+%\subsection{National Long Term Care Survey}
+%\label{nltcs} 
+%
+%The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
 It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989.
@@
 We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.
-
-With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible.
+%
+%With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible.
 Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold.
@@
 We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler.
-
-Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
+%
+%Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
 However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
 Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
-
-Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
+%
+%Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
 With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks.

### 2021-09-20 R/GetUniquePatterns2.R

@@ -6,7 +6,10 @@ hash_field <- function(L_f, k, Lf_vec){
 GetUniquePatterns2 <- function(cd,
                                fast = F,
                                R = NULL,
-                               all_patterns = TRUE){
+                               all_patterns = TRUE,
+                               nA_index = FALSE,
+                               df1_index = NULL,
+                               df2_index = NULL){
 
   indicators <- cd[[1]]
   N <- dim(indicators)[1]
@@ -14,7 +17,17 @@ GetUniquePatterns2 <- function(cd,
   n1 <- cd[[2]]
   n2 <- cd[[3]]
 
-  ids <- expand.grid(1:n1, 1:n2)
+  if(is.null(df1_index)){
+    chunkA_index <- 1:n1
+  }else{
+    chunkA_index <-  df1_index
+  }
+  if(is.null(df2_index)){
+    chunkB_index <- 1:n2
+  }else{
+    chunkB_index <-  df2_index
+  }
+  ids <- expand.grid(chunkA_index, chunkB_index)
   rec1 <- ids[,1]
   rec2 <- ids[,2]
   levels <- cd[[4]]

### 2021-09-20 R/LinkRecordsBK.R

@@ -72,3 +72,235 @@ LinkRecordsBK<- function(Zchain, n1, lFNM=1, lFM1=1, lFM2=2, lR=Inf){
   return(list(Zhat = Zhat,
               post_probs = probMaxProbOption))
 }
+
+CompareRecordsBK <- function(df1,
+                             df2,
+                             flds=NULL,
+                             flds1=NULL,
+                             flds2=NULL,
+                             types=NULL,
+                             breaks=c(0,.25,.5),
+                             df1_index = NULL,
+                             df2_index = NULL){
+
+  warn <- FALSE
+
+  # control the input
+
+  if(!is.data.frame(df1)) stop("'df1' is not a data frame")
+  if(!is.data.frame(df2)) stop("'df2' is not a data frame")
+
+  if(!is.null(flds)){ # user provides 'flds' fields in 'df1' and 'df2' for linkage
+    if( !(class(flds) %in% c("character", "numeric")) )
+      stop("'flds' should be of class character or numeric")
+
+    if(!is.null(flds1)) warning("Argument 'flds' was provided, 'flds1' is ignored")
+    if(!is.null(flds2)) warning("Argument 'flds' was provided, 'flds2' is ignored")
+
+    flds1 <- flds2 <- flds
+
+    if(is.numeric(flds)){
+      fldsCheck1 <- flds %in% seq_len(ncol(df1))
+      if(!all(fldsCheck1)) stop("Some numbers in 'flds' are out of the column range for 'df1'")
+      fldsCheck2 <- flds %in% seq_len(ncol(df2))
+      if(!all(fldsCheck2)) stop("Some numbers in 'flds' are out of the column range for 'df2'")
+      warn <- TRUE
+    }
+    if(is.character(flds)){
+      fldsCheck1 <- flds %in% colnames(df1)
+      if(!all(fldsCheck1)) stop("Some names in 'flds' are not columns of 'df1'")
+      fldsCheck2 <- flds %in% colnames(df2)
+      if(!all(fldsCheck2)) stop("Some names in 'flds' are not columns of 'df2'")
+      flds1 <- match(flds1, colnames(df1))
+      flds2 <- match(flds2, colnames(df2))
+    }
+  }else{
+    if(is.null(flds1) & !is.null(flds2)) stop("Argument 'flds1' also needs to be specified")
[…]

### 2021-09-20 R/SynthesizeHashChunks.R

@@ -1,4 +1,4 @@
-SythensizeHash <- function(patterns_hashed){
+SynthesizeHash <- function(patterns_hashed){
   counts_big <- Reduce(`+`, map(patterns_hashed, ~.x[[2]]))
 
   counts_per_rec_big <- patterns_hashed %>%
@@ -22,3 +22,6 @@ SythensizeHash <- function(patterns_hashed){
 
   cd <- list(comparisons, patterns)
 }
+
+
+

### 2021-09-20 methods_paper/fabl-paper.tex
parallel on cluster

@@
 
+\hypertarget{accuracy}{%
+	\subsection{Accuracy}\label{accuracy}}
+
 We first compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
@@
 \end{figure}
+
+\hypertarget{speed}{%
+	\subsection{Speed}\label{speed}}
 
@@
 status of the record pairs should lead to even more computational gains.
+
+\hypertarget{scale}{%
+	\subsection{Scale}\label{scale}}
+
+In one final simulation, we demonstrate how partitioning and hashing the data through \texttt{fabl} greatly expands the scale of the linkage tasks we can solve.
+Assuming a 16GB personal machine, standard record linkage methods using all-to-all comparisons can only function on tasks where the matrix of comparison vectors $\Gamma$ is less than 8GB.
+In the context of the above simulation, each comparison vector contains four integers,
 

### 2021-09-20 methods_paper/fabl-paper.tex
working on bk paper.

@@
 
-Record linkage is the task of identifying
-duplicate records across multiple data sources.
+%% Reworking introduction for Brian's paper.
+
+Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen2012data}.
+Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+
+\begin{enumerate}
+\item Let's re-visit the case study of Sadinle. Lay this out and the importance of it. 
+\item Next, let's briefly review the landscape of the literature. 
+\item State Brian's contributions. 
+\item Then write a section on related work. 
+\item Then write a section regarding the layout of the paper.
+\end{enumerate}
+
+\newpage
+
 In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years.

### 2021-09-22 Prelim/prelim.Rmd
added prelim slides

---
title: "Prelim"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Record Linkage

# Fellegi and Sunter

# Notation

- File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.

- $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors

- $Z_j$

- $m_{fl} = P\left(\gamma_{ij}^f = l |Z_j = i\right)$

- $u_{fl} = P\left(\gamma_{ij}^f = l |Z_j \neq i\right)$

- $\lambda = P(Z_j \leq n_A)$

# m's and u's

# Previous work

# Model

# Hashing

Demonstration

# Managing large data

# Simulation Results 1

# Simulation Results 2

# Extentions and Directions

- $m_{fl} = $

# Previous work

______________________________________

# El Salvador

# Distributed computation

# Storage Efficient Indexing

### 2021-09-22 methods_paper/fabl-paper.tex
added prelim slides

--- 2021-09-20 methods_paper/fabl-paper.tex
@@
 
+\hypertarget{scale}{%
+	\subsection{Scale}\label{scale}}
+
+% Assuming a 16GB personal machine, standard record linkage methods using all-to-all comparisons can only function on tasks where the matrix of comparison vectors $\Gamma$ is less than 8GB.
+In the context of the above simulation, each comparison vector contains four integers, collectively requiring 16 bytes of memory; assuming files of equal lengths, this means the largest linkage task possible is around $7000 \times 7000$.
+
+ In one final simulation, we demonstrate how partitioning and hashing the data through \texttt{fabl} method allows us undertake significantly larger linkage tasks We concatenate two sets of 40 simulated datasets from the Sadinle simulation to create two larger files, each of 20,000 records.
+Working on one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
+Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting a final comparison matrix about 6.4 GB in size.
+
+ For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
+We then compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
+The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method.
+Excecuted sequentially, these comparisons and the Gibbs sampler take about one hour to run; using distributed computing, it could be completed much faster.
+This simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
+The reader will note that this slightly worse performance than witnessed in the smaller simulation studies; this is expected because it is naturally more difficult to link more files with the same amount of information.
+With more linkage fields \texttt{fabl} maintains the high accuracy seen above.
+
 \section{Case Studies}
@@
 
-We now demonstrate the power of our method through two case studies.
-The first is a study of documented identifiable deaths (DID) from the El Salvadoran Civil War.
+We now demonstrate the power of our method through a case study of documented identifiable deaths (DID) from the El Salvadoran Civil War.
 Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
 Secondly, the case study reveals the impact of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
-The second case study is the National Long Term Care Survey (NLTCS), a much larger linkage tasks which \texttt{BRL} is unable to complete but which we complete with \texttt{fabl} with relative ease.
+
+%The second case study is the National Long Term Care Survey (NLTCS), a much larger linkage tasks which \texttt{BRL} is unable to complete but which we complete with \texttt{fabl} with relative ease.
 

### 2021-09-22 methods_paper/unblinded-paper.tex
added dblink paper

\documentclass[12pt,letterpaper]{article}

\usepackage[hypertexnames=false]{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}

% Custom packages
\usepackage{amsthm, amssymb}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{tabularx}
\usepackage{tikz}           % vector graphics
\usepackage{bm}             % bold math symbols
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{import}
\usepackage{titling}
\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
  calc, shadows, plotmarks}

\graphicspath{{./figures/}}

% Custom commands
\let\oldvec\vec
\renewcommand\vec{\bm}
\newcommand{\simfn}{\mathtt{sim}} % similarity function
\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
\newcommand{\distfn}{\mathtt{dist}} % distance function
\newcommand{\valset}{\mathcal{V}} % attribute value set
\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
\newcommand{\euler}{\mathrm{e}} % Euler's constant
\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
\def\spacingset#1{\renewcommand{\baselinestretch}%
  {#1}\small\normalsize} \spacingset{1}

\newtheorem*{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem*{definition}{Definition}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\sectionfont{\large\nohang\centering\MakeUppercase}

\title{\bf \dblink: Distributed End-to-End Bayesian Entity Resolution}
\author{Neil G.~Marchant\textsuperscript{a} \and
  Andee Kaplan\textsuperscript{b} \and 
  Daniel N.~Elazar\textsuperscript{c} \and
  Benjamin I.~P.~Rubinstein\textsuperscript{a} \and 
  Rebecca C.~Steorts\textsuperscript{d}}
\date{
  \textsuperscript{a}School of Computing and Information Systems, University 
  of Melbourne\\
  \textsuperscript{b}Department of Statistics, Colorado State University\\
  \textsuperscript{c}Methodology Division, Australian Bureau of Statistics\\
  \textsuperscript{d}Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\
  DRB \#: CBDRB-FY20-309\\[2ex]
  \today}

\begin{document}
\maketitle

\bigskip
\begin{abstract}
Entity resolution (ER; also known as record linkage or de-duplication) is 
the process of merging noisy databases, often in the absence of unique 
identifiers. 
A major advancement in ER methodology has been the application of Bayesian 
generative models, which provide a natural framework for inferring latent 
entities with rigorous quantification of uncertainty. 
Despite these advantages, existing models are severely limited in practice, 
as standard inference algorithms scale quadratically in the number of records. 
While scaling can be managed by fitting the model on separate blocks of 
the data, such a na\"{i}ve approach may induce significant error 
in the posterior. 
In this paper, we propose a principled model for scalable Bayesian ER, 
called ``\underline{d}istributed \underline{B}ayesian 
\underline{link}age'' or \dblink, which jointly performs 
blocking and ER without compromising posterior correctness. 
Our approach relies on several key ideas, including: 
(i)~an auxiliary variable representation that induces a partition of 
the entities and records into blocks; 
(ii)~a method for constructing well-balanced blocks based on k-d trees; 
(iii)~a distributed partially-collapsed Gibbs sampler with improved mixing; 
and 
(iv)~fast algorithms for performing Gibbs updates. 
Empirical studies on six data sets---including a case study 
on the 2010 Decennial Census---demonstrate the scalability and 
effectiveness of our approach.
\end{abstract}

\noindent%
{\it Keywords:} auxiliary variable, distributed computing, Markov chain Monte 
Carlo, partially-collapsed Gibbs sampling, record linkage

\newpage
\spacingset{1.5}

\section{Introduction}
\label{sec:introduction}

﻿When information about a statistical population is scattered across 
multiple databases, there may be immense value in combining them. 
A combined database can provide a more accurate and complete view of 
the population by improving coverage, bringing together analytic 
variables, and resolving erroneous and missing values. 
This allows statisticians to draw  richer and more reliable 
conclusions.  
Among the types of questions that can be addressed by combining such 
databases are the following: 
How accurate are census enumerations for minority groups 
\citep{winkler_overview_2006}? 
How many of the elderly are at high risk for sepsis in different parts 
of the country \citep{saria_trillion_2014}? 
How many people were victims of war crimes in recent conflicts in Syria 
\citep{price_2013}?

An important step when combining databases is identifying records that 
refer to the same statistical unit. 
This is challenging in practice because consistent identifiers, such 
as social security numbers, are often not available. 
Identifiers may be omitted due to privacy concerns, they may be 
inconsistent across the databases, or they may have never been 
recorded. 
In such cases, practitioners must rely on  {\em entity resolution} 
(ER) to infer the relationships between records and statistical units 
(entities) using linking variables in the observed data. 
This problem is studied in the statistics, machine learning, database 
and natural language processing communities, and is also known as 
entity disambiguation, merge-purge, record linkage, deduplication 
and co-reference resolution \citep{christen_data_2012, dong_big_2015, 
soon_machine_2001}.

ER is not only a crucial tool for statistical analysis, it is also a 
challenging statistical and computational problem in itself. 
This is because many databases lack reliable linking variables, the 
record comparison space scales quadratically in the number of records, 
and the number of parameters to be estimated grows with the number of 
records \citep{Herzog_2007, lahiri_2005, winkler_state_1999, 
winkler_2000}. 
To meet present and near-future needs, ER methods must be flexible and 
scalable to large databases. 
Furthermore, they must be able to handle uncertainty and be easily 
integrated with post-ER statistical analyses, such as regression. 
All of this must be done while achieving low error rates.

Bayesian models offer a promising framework for ER as they support 
natural uncertainty propagation, flexible modeling assumptions, and 
incorporation of prior information. 
However, existing Bayesian ER models either ignore scalability 
\citep{steorts_entity_2015, zanella_flexible_2016, 
sadinle_bayesian_2017} or manage scalability in an unprincipled 
manner by applying blocking outside the Bayesian framework
\citep{fortini_bayesian_2001, larsen_advances_2005, 
larsen_experiment_2012, tancredi_hierarchical_2011, 
gutman_bayesian_2013, sadinle_detecting_2014, steorts_bayesian_2016}. 
Blocking improves scalability by partitioning records into blocks and 
assuming records in different blocks do not refer to the same entity 
\citep{christen_survey_2012}. 
However, when blocking is performed as a separate deterministic step 
it is not possible to propagate the uncertainty. 
Moreover, since the blocks are fixed, a poor blocking design may compromise 
the accuracy of the entire ER process. 
In other words, one sacrifices uncertainty propagation and accuracy 
for scalability.

In this paper, we propose a principled approach to scaling Bayesian ER 
models, which does not suffer from the limitations of ad-hoc 
deterministic blocking. 
Using the \blink\ ER~model~\citep{steorts_entity_2015} as a 
foundation, we propose a scalable and distributed extension called 
``\underline{d}istributed \blink'' or \dblink\ for short, which 
integrates probabilistic blocking in a fully Bayesian framework. 
To our knowledge, \dblink\ is the first Bayesian ER model which 
supports propagation of uncertainty between the blocking and 
matching\slash linking stages of ER, without compromising the 
correctness of the posterior. 
In addition, \dblink\ supports distributed\slash parallel inference 
at the block level to further improve scalability to large databases.

We make several contributions to the literature. 
First, we propose an auxiliary variable representation of \blink, 
which induces a partitioning of the entities and records into blocks. 
These play a similar role as traditional deterministic blocks, 
however the assignments of latent entities and records to blocks are 
\emph{random} and inferred \emph{jointly} with the other model 
parameters. 
Second, we prove that our auxiliary variable representation 
preserves the marginal posterior distribution over the model 
parameters. 
This is a desirable property, as it means our inferences are 
theoretically independent of the blocking design. 
Third, we propose a method for constructing well-balanced blocks based 
on $k$-d trees. 
Fourth, we design a distributed partially-collapsed Gibbs sampler to perform 
inference, and demonstrate superior mixing times when compared to a standard 
Gibbs sampler. 
Fifth, we propose algorithms for improving computational efficiency 
of the Gibbs updates which leverage indexing data structures and a 
novel perturbation sampling algorithm.

We implement our proposed methodology as an open-source Apache Spark 
package\footnote{Spark package source code available at 
  \url{https://github.com/cleanzr/dblink}.}
and provide an R interface for broad accessibility\footnote{R package 
  source code available at \url{https://github.com/cleanzr/dblinkR}.}. 
We conduct empirical evaluations on two synthetic and three real data sets, 
demonstrating efficiency gains in excess of 300$\times$ compared to 
\blink. 
To illustrate the effectiveness of our approach for realistic ER tasks, 
we present a case study using Census and administrative data from the 
U.S.\ state of Wyoming.

The paper is organized as follows. 
In Section~\ref{sec:related} we review related work in ER methodology 
and approximate inference algorithms. 
We then formulate ER in a Bayesian setting in Section~\ref{sec:model}, and 
present the \dblink\ model with integrated probabilistic blocking.
In Section~\ref{sec:partition-fn} we provide guidelines for selecting 
blocking functions. 
We then discuss inference and propose a distributed partially-collapsed 
Gibbs sampler in Section~\ref{sec:inference}. 
We suggest additional methods for improving computational efficiency of 
inference in Section~\ref{sec:tricks}.
Section~\ref{sec:experiments} provides a comprehensive empirical evaluation, 
and Section~\ref{sec:decennial} presents a case study to U.S.\ Census 
and administrative data. 
We make closing remarks in Section~\ref{sec:conclusions}.

\section{Related work}
\label{sec:related}
We review related work across three main areas---ER methodology, inference 
for Bayesian ER models, and distributed Markov chain Monte Carlo (MCMC).

\paragraph{Entity resolution methodology.}
The first probabilistic approach to ER was due to 
\citet{newcombe_automatic_1959}, who applied matching rules to 
pairs of records. 
This idea was later formalized in a seminal paper by 
\citet{fellegi_theory_1969} within a decision-theoretic framework.
Many variations of the Fellegi-Sunter (FS) approach have been proposed 
(for surveys, see~\citealp{winkler_overview_2006, winkler2014matching}), 
including a generalization to multiple 
databases~\citep{sadinle_generalized_2013}.
Others have addressed scalability of FS-type approaches using 
blocking\slash indexing methods (see 
\citealp{christen_data_2012,steorts_comparison_2014} for surveys) and 
efficient data structures \citep{enamorado_using_2019}.
However, traditional FS approaches do not naturally support propagation of 
ER uncertainty, and existing methods for scaling make approximations that 
sacrifice accuracy.

While the FS approach has been highly influential, it has also been 
criticized due to its lack of support for duplicates within databases; 
misspecified independence assumptions; and its dependence on subjective 
thresholds \citep{tancredi_hierarchical_2011}.
These limitations have prompted development of more sophisticated 
Bayesian models, including models for bipartite 
matching~\citep{fortini_bayesian_2001, larsen_advances_2005, 
  larsen_experiment_2012, tancredi_hierarchical_2011, gutman_bayesian_2013, 
  sadinle_bayesian_2017, mcveigh_scaling_2019}, 
deduplication~\citep{sadinle_detecting_2014, tancredi_unified_2020}
and matching across multiple databases~\citep{steorts_entity_2015, 
  steorts_bayesian_2016}.
Several of these models operate on attribute-level comparisons between 
pairs of records in a similar vein as the FS approach 
\citep{larsen_advances_2005, larsen_experiment_2012, gutman_bayesian_2013, 
  sadinle_detecting_2014, sadinle_bayesian_2017, mcveigh_scaling_2019}.
This contrasts with entity-centric generative models which assume the 
records arise as distortions to some latent entity attributes 
\citep{tancredi_hierarchical_2011, steorts_entity_2015, 
  steorts_bayesian_2016, tancredi_unified_2020}.

In scenarios where training data is scarce or unavailable, Bayesian 
generative models tend to be more robust than discriminative or 
likelihood-based methods, as the priors have a regularizing effect. 
Bayesian generative models are also amenable to theoretical analysis: 
recent work has obtained lower bounds on the probability of misclassifying 
the entity associated with a record \citep{steorts_performance_2017}.
However, a major downside of Bayesian ER models is the computational 
cost of performing inference (see discussion below).

Apart from these advances in Bayesian models for ER (largely undertaken 
in statistics), there have been an abundance of contributions from the 
database and machine learning communities 
\citep[see surveys by][]{getoor_entity_2012, christen_data_2012}.
Their focus has typically been on rule-based 
approaches~\citep{fan_reasoning_2009, singh_generating_2017}, 
supervised learning approaches~\citep{mudgal_deep_2018}, 
hybrid human-machine approaches~\citep{wang_crowder:_2012, 
  gokhale_corleone:_2014}, 
and scalability~\citep{papadakis_comparative_2016}.
Broadly speaking, all of these approaches rely on either humans in-the-loop 
or large amounts of labelled training data, which is not generally the 
case in the Bayesian setting.

\paragraph{Inference for Bayesian ER models.}
Most prior work on Bayesian generative models for ER
\citep[e.g.][]{tancredi_hierarchical_2011, gutman_bayesian_2013,
  steorts_entity_2015} has relied on Gibbs sampling for inference. 
Compared to other Markov chain Monte Carlo (MCMC) algorithms, 
Gibbs sampling is relatively easy to implement, however it may suffer 
from slow convergence and poor mixing owing to its highly local 
moves~\citep{liu_monte_2004}. 
Scalability is also a challenge, as a na\"{i}ve Gibbs update for the 
linkage structure requires all-to-all comparisons between records 
(or between records and entities for entity-centric models). 
This issue is often managed by applying deterministic blocking prior to 
Gibbs sampling, thereby sacrificing accuracy and proper 
treatment of uncertainty \citep{larsen_advances_2005, larsen_experiment_2012, 
tancredi_hierarchical_2011, gutman_bayesian_2013, sadinle_detecting_2014}.

In the broader context of clustering models, the \emph{split-merge 
algorithm}~\citep{jain_split-merge_2004} has been proposed 
as an alternative to Gibbs sampling.
It is a Metropolis-Hastings algorithm, which traverses the space of 
clusterings via proposals that split individual clusters or merge pairs 
of clusters. 
Since multiple cluster items are updated in a single move, it is less 
susceptible to becoming trapped in local modes.
\cite{steorts_bayesian_2016} applied this algorithm, in combination with 
deterministic blocking, to update the linkage structure in an ER model 
similar to \blink. 
A close relative of the split-merge algorithm is the \emph{chaperones 
  algorithm}, which was proposed for inference in microclustering models
\citep{zanella_flexible_2016}. 
The chaperones algorithm is expected to be more efficient, as it 
preferentially focuses on more likely cluster reassignments, 
through a user-specified biased distribution on the product space of 
cluster items. 
However, the biased distribution must be designed so that random 
item pairs can be drawn efficiently, without explicitly constructing 
the product space.

More recently, \citet{zanella_informed_2020} proposed a general 
framework for designing informative proposals in a Metropolis-Hastings 
setting, which is suited for discrete spaces (e.g.\ the space of 
possible linkage structures). 
They show that \emph{locally-balanced proposals} are asymptotically-optimal 
within the class of pointwise informative proposals, and demonstrate 
significant improvements in efficiency when compared to a split-merge-type
algorithm.
However, computing a locally-balanced proposal for the linkage structure 
is computationally challenging due to quadratic scaling. 
This can be mitigated to some extent by running locally-balanced 
updates within randomly-selected sub-blocks of records. 
However to avoid poor mixing, care must be taken to ensure that  
randomly-selected sub-blocks contain likely matching records.

In contrast to much of the literature on Bayesian ER models, 
\citet{mcveigh_scaling_2019} proposed a method that combines deterministic 
blocking and restricted MCMC (based on earlier work by 
\citealp{mcveigh_practical_2017}). 
They balance approximation error by performing coarse-grained 
deterministic blocking\slash indexing as an initial step, followed by 
data-dependent post-hoc blocking.
During inference, the linkage structure is updated using locally-balanced 
proposals, restricted to the post-hoc blocks. 
They demonstrate improved scalability---to data sets with several hundred 
thousand of record---with minimal risk of approximation error. 
However, their approach is not directly compatible with distributed 
inference (see below) and may require modification for use with an 
entity-centric model.

\paragraph{Parallel\slash distributed MCMC.}
Recent literature has focused on using parallel and distributed computing to 
scale up MCMC algorithms, where applications have included Bayesian topic 
models~\citep{newman_distributed_2009, smola_architecture_2010, 
ahn_distributed_2014} and mixture models~\citep{williamson_parallel_2013, 
chang_parallel_2013, lovell_clustercluster:_2013, ge_distributed_2015}.
We review the application to mixture models, as they are 
conceptually similar to ER models.

Existing work has concentrated on Dirichlet process (DP) mixture models and 
hierarchical DP mixture models.
The key to enabling distributed inference for these models is the 
realization that a DP mixture model can be reparameterized as a mixture of 
DPs.
Put simply, the reparameterized model induces a \emph{partitioning} of the 
clusters into blocks, such that clusters assigned to \emph{distinct blocks} 
are conditionally independent. 
As a result, variables within blocks can be updated in parallel.
\cite{williamson_parallel_2013} exploited this idea at the thread level 
to parallelize inference for a DP mixture model.
\cite{chang_parallel_2013} followed a similar approach, but included an 
additional level of parallelization within blocks using a parallelized 
version of the split-merge algorithm.
Others~\citep{lovell_clustercluster:_2013, ge_distributed_2015} have 
developed distributed implementations in the MapReduce framework.

We do not consider DP mixture models in our work, as their behavior 
is ill-suited for ER applications.\footnote{With a DP prior, the 
number of clusters grows logarithmically in the number of records, 
but empirical observations call for near-linear 
growth~\citep{zanella_flexible_2016}.}
However we do borrow the reparameterization idea, albeit with a more flexible 
partition specification which permits similar entities to be co-blocked, 
while facilitating load balancing.
It would be interesting to see whether similar ideas can be applied to 
microclustering models~\citep{zanella_flexible_2016}, 
however preserving the marginal posterior distribution seems 
challenging in this case.

\section{A scalable model for Bayesian ER}
\label{sec:model}

In this section, we present our scalable ER model called \dblink, 
which integrates probabilistic blocking in a fully Bayesian framework.
Our model can be viewed as an extension of the \blink\ model 
\citep{steorts_entity_2015} that incorporates an auxiliary partition 
of the latent entity parameter space into blocks. 
Unlike ad-hoc blocking approaches used previously in the literature 
\citep{larsen_advances_2005, larsen_experiment_2012, 
  tancredi_hierarchical_2011, gutman_bayesian_2013, sadinle_detecting_2014, 
  steorts_bayesian_2016}, 
the blocks in \dblink\ are random, and inferred jointly with the other 
model parameters. 
This enables propagation of uncertainty between the blocking and ER stages.
In addition, \dblink\ extends \blink\ with support for missing values and 
user-defined attribute similarity measures.

We describe notation and assumptions in Section~\ref{sec:problem}, before 
presenting \dblink\ in Section~\ref{sec:model-specification}.
We define attribute similarity measures in 
Section~\ref{sec:attribute-sim-measure}, including an optional truncation 
approximation which can improve scalability.
In Section~\ref{sec:posterior-dist}, we prove that the marginal posterior of  
\dblink\ (integrated over the blocks) reduces to \blink\ under certain 
conditions.
This is a desirable property, as it means our inferences are theoretically 
independent of the blocking design. 
Finally, in Section~\ref{sec:partition-motivation} we explain how the auxiliary 
blocks are beneficial in scaling and distributing inference.

\subsection{Notation and problem formulation}
\label{sec:problem}
In this section, we define notation and formulate ER in a Bayesian 
setting.
Consider a collection of $T$ tables\footnote{We define a \emph{table} 
as an ordered (indexed) collection of records, which may contain
duplicates (records for which all attributes are identical).} 
(databases) indexed by $t$, each with $R_t$ records (rows) indexed by $r$
and $A$ aligned attributes (columns) indexed by $a$.
Associated with the records is a fixed population of entities of size $E$ 
indexed by $e$. 
Each entity $e$ is described by a set of attributes 
$\vec{y}_e = [y_{ea}]_{a = 1 \ldots A}$, which are aligned with the 
record attributes. 
The population of entities is partitioned into $B$ blocks for 
computational convenience, using a blocking function $\blockfn$ that maps an 
entity $e$ to a block based on its attributes $\vec{y}_e$.
We assume each record $(t,r)$ belongs to a block $\gamma_{tr}$ and 
is associated with an entity $\lambda_{tr}$ within that block.
The value of the $a$-th attribute for record $(t,r)$ is 
denoted by $x_{tra}$, and is assumed to be a noisy observation 
of the associated entity's true attribute value $y_{\lambda_{tr}a}$.
We allow for the fact that some attributes $x_{tra}$ may be missing
completely at random through a corresponding indicator variable 
$o_{tra}$~\cite[p.~12]{little_statistical_2002}.

Table~\ref{tbl:notation} summarizes our notation, including 
model-specific parameters which will be introduced shortly.
We adopt the following rules to compactly refer to sets of variables:
\begin{itemize}
  \item A boldface lower-case variable denotes the set of 
  \emph{all attributes}: e.g.~$\vec{x}_{tr} = [x_{tra}]_{a=1\ldots A}$.
  \item A boldface capital variable denotes the set of 
  \emph{all index combinations}: 
  e.g.~$\vec{X} = [x_{tra}]_{t=1\ldots T; r=1 \ldots R_t; a=1 \ldots A}$.
\end{itemize}
We also define notation to separate the record attributes $\mathbf{X}$ 
into an observed part $\mathbf{X}^{(o)}$ (those $x_{tra}$'s for which 
$o_{tra} = 1$) and a missing part $\mathbf{X}^{(m)}$ (those 
$x_{tra}$'s for which $o_{tra}=0$).

After specifying a generative model (see next section), we perform 
ER by inferring the \emph{joint} posterior distribution over:
\begin{itemize}
  \item the block assignments $\vec{\Gamma} = [\gamma_{tr}]_{t = 1 \ldots T; r = 1 \ldots R_t}$,
  \item the linkage structure $\vec{\Lambda} = [\lambda_{tr}]_{t = 1 \ldots T; r = 1 \ldots R_t}$, and 
  \item the true entity attribute values $\vec{Y} = [y_{ea}]_{e = 1 \ldots E; a = 1 \ldots A}$,
\end{itemize}
conditional on the observed record attribute values $\mathbf{X}^{(o)}$. 
Note that we operate in a fully unsupervised setting, since we do 
not condition on ground truth data for the links or entities. 
Inferring $\vec{\Gamma}$ is equivalent to the \emph{blocking} stage 
of ER, where the records are partitioning into blocks to limit the 
comparison space.
Inferring $\vec{\Lambda}$ is equivalent to the \emph{matching\slash linking} 
stage of ER, where records that refer to the same entities are linked together. 
Inferring $\vec{Y}$ is equivalent to the \emph{merging} stage, where 
linked records are combined to produce a single representative record. 
By inferring $\vec{\Gamma}$, $\vec{\Lambda}$ and $\vec{Y}$ jointly, we are 
able to propagate uncertainty between the three stages.

\subsection{Model specification}
\label{sec:model-specification}
We now present our proposed model \dblink\ by describing the generative 
process.
We provide a visual representation of the model in 
Figure~\ref{fig:plate-diagram}, with key differences from \blink\ 
highlighted in a dashed blue line style.

\begin{figure}
  \centering
  \resizebox{0.45\linewidth}{!}{%
    \input{./figures/plate-diagram.tex}
  }
  \caption{Plate diagram for \dblink. 
    Extensions to \blink\ are highlighted in a dashed blue line style.
    Circular nodes represent random variables; square nodes represent 
    deterministic variables; (un)shaded nodes represent (un)observed variables; 
    arrows represent conditional dependence; and plates represent replication 
    over an index.
  }
  \label{fig:plate-diagram}
\end{figure}

\begin{table}
  \caption{Summary of notation.}
  \label{tbl:notation}
  \begin{center}
  \newcolumntype{Z}{>{\arraybackslash}X}
  \spacingset{1}
  \footnotesize
  \begin{tabularx}{\linewidth}{@{}l Z @{} l Z@{}}
  	\toprule
  	Symbol                       & Description & Symbol & Description\\
  	\midrule
    $t \in 1 \ldots T$           & index over tables 
    & $\gamma_{tr}$              & assigned block for record $r$ in table $t$ \\
    $r \in 1 \ldots R_t$         & index over records in table $t$ 
    & $\lambda_{tr}$             & assigned entity for record $r$ in table $t$ \\
    $e \in 1 \ldots E$           & index over entities 
    & $\theta_{ta}$              & prob.\ attribute $a$ in table $t$ is distorted\\
    $b \in 1 \ldots B$           & index over block 
    & $\alpha_a, \beta_a$        & distortion hyperparams.\ for attribute $a$ \\
    $a \in 1 \ldots A$           & index over attributes 
    & $\eta_{ta}$                & prob.\ attribute $a$ in table $t$ is observed \\
    $v \in 1 \ldots |\valset_a| $ & index over domain of attribute $a$ 
    & $\valset_a$                & domain of attribute $a$ \\
    $R = \sum_{t} R_t$           & total number of records 
    & $\phi_{a}(\cdot)$          & distribution over domain of attribute $a$ \\
    $x_{tra}$                    & attribute $a$ for record $r$ in table $t$     
    & $\simfn_a(\cdot, \cdot)$   & similarity measure for attribute $a$ \\
    $z_{tra}$                    & distortion indicator for $x_{tra}$ 
    & $\entset_e$                & set of records assigned to entity $e$\\
    $o_{tra}$                    & observed indicator for $x_{tra}$ 
    & $\partset_b$               & set of entities assigned to block $b$ \\     
    $y_{ea}$                     & attribute $a$ for entity $e$ 
    & $\blockfn(\cdot)$           & block assignment function \\
    \bottomrule
  \end{tabularx}
  \end{center}
\end{table}

\paragraph{Entities.} 
The population of entities is assumed to be of fixed size $E$.
Each entity $e$ is described by a vector of ``true'' attributes 
$\vec{y}_e \in \bigotimes_{a=1}^{A} \valset_{a}$.
The value of the $a$-th attribute $y_{ea}$ is assumed to be drawn 
independently from a distribution 
$\phi_{a}$ over the attribute domain $\valset_{a}$:
\begin{equation}
y_{ea} \overset{\mathrm{ind.}}{\sim}
  \operatorname{Discrete}_{v \in \valset_{a}} [\phi_{a}(v)].
\end{equation}
Following the \blink\ model, we set the population size $E$ and 
the distributions over the attribute domains $\phi_{a}$ empirically.
Recommendations for setting these parameters are provided in 
Appendix~F.3.

\paragraph{Blocks.} 
The parameter space associated with the entities 
$\bigotimes_{a=1}^{A} \valset_{a}$ is partitioned into $B$ blocks. 
The partition is parameterized using a deterministic 
\emph{blocking function}:
\begin{equation}
  \blockfn: \bigotimes_{a} \valset_{a} \to \{1,\ldots, B\},
  \label{eqn:part-fn}
\end{equation}
which is a free parameter and may be selected for inferential 
  convenience. 
We provide recommendations for selecting the blocking function in 
Section~\ref{sec:partition-fn}, including an example based on $k$-d trees.

We shall often need to refer to the entities assigned to a particular block. 
To do this concisely, we introduce the notation 
$\partset_{b}(\vec{Y}) = \{e: \blockfn(\vec{y}_{e}) = b\}$ to denote 
the set of entities assigned to block $b$. 
This is random due to the dependence on $\vec{Y}$, however we shall often 
omit the dependence for brevity.

%We note that the case $B = 1$, where all entities are assigned 
%to a single block, is identical to \blink.

\paragraph{Distortion.} Associated with each table $t$ and 
attribute $a$ is a distortion probability $\theta_{ta}$, 
with assumed prior distribution:
\begin{equation}
\theta_{ta}|\alpha_{a}, \beta_{a} \overset{\mathrm{ind.}}{\sim}
\operatorname{Beta}[\alpha_{a}, \beta_{a}],
\end{equation}
where $\alpha_{a}$ and $\beta_{a}$ are hyperparameters.
We provide recommendations for setting $\alpha_{a}$ and $\beta_{a}$ in 
Appendix~F.
The distortion probabilities feed into the record-generation process below.

\paragraph{Records.} We assume a record is generated 
by selecting an entity uniformly at random and 
copying the entity's attributes subject to distortion. 
The process for generating record $r$ in table $t$ is outlined below. 
Steps (i), (ii), and (v) deviate from \blink.
\begin{enumerate}[(i)]
  \item Choose a block assignment $\gamma_{tr}$ at random 
  in proportion to the block sizes:
  \begin{equation}
  \gamma_{tr}|\vec{Y} \overset{\mathrm{ind.}}{\sim}
    \operatorname{Discrete}_{b \in \{1 \ldots B\}}[|\partset_b|/E].
  \end{equation}
  \item Choose an entity assignment $\lambda_{tr}$ uniformly 
  at random from block $\gamma_{tr}$:
  \begin{equation}
  \lambda_{tr}|\gamma_{tr}, \vec{Y} \overset{\mathrm{ind.}}{\sim}
    \operatorname{DiscreteUniform}[\partset_{\gamma_{tr}}].
  \end{equation}
  \item For each attribute $a$, draw a distortion indicator
  $z_{tra}$:
  \begin{equation}
  z_{tra}|\theta_{ta} \overset{\mathrm{ind.}}{\sim}
    \operatorname{Bernoulli}[\theta_{ta}].
  \end{equation}
  \item For each attribute $a$, draw a record value $x_{tra}$:
  \begin{equation}
  x_{tra}|z_{tra}, y_{\lambda_{tr}a} \overset{\mathrm{ind.}}{\sim}
  (1 - z_{tra}) \delta(y_{\lambda_{tr}a}) + z_{tra} 
  \operatorname{Discrete}_{v \in \valset_{a}}[\psi_{a}(v|y_{\lambda_{tr}a})]
  \end{equation}
  where $\delta(\cdot)$ represents a point mass.
  If $z_{tra} = 0$, $x_{tra}$ is copied directly from the 
  entity.
  Otherwise, $x_{tra}$ is drawn from the domain $\valset_{a}$ 
  according to the distortion distribution $\psi_{a}$.
  In the literature, this is known as a hit-miss 
  model~\citep{copas_record_1990}.
  \item For each attribute $a$, draw an observed indicator $o_{tra}$:
  \begin{equation}
    o_{tra} \overset{\mathrm{ind.}}{\sim} \operatorname{Bernoulli}[\eta_{ta}].
  \end{equation}
  %where $\eta_{ta}$ may be estimated empirically.
  If $o_{tra} = 1$, $x_{tra}$ is observed, otherwise it is 
  missing.
\end{enumerate}

\paragraph{Detail on the distortion distribution.} $\psi_{a}(\cdot|w)$ 
chooses a distorted value for attribute $a$ conditional on 
the true value $w$.
In our parameterization of the model, it is defined as
\begin{equation}
\psi_{a}(v|w) = h_{a}(w) \phi_{a}(v) \euler^{\simfn_{a}(v,w)},
\label{eqn:distortion-dist}
\end{equation}
where $h_{a}(w) = 1 / \sum_{v \in \valset_{a}} \phi_{a}(v) 
\euler^{\simfn_{a}(v,w)}$ is a normalization constant and 
$\simfn_{a}$ is the similarity measure for attribute~$a$
(see Section~\ref{sec:attribute-sim-measure}).
Intuitively, this distribution chooses values in proportion to 
their empirical frequency, while placing more weight on those 
that are ``similar'' to $w$.
This reflects the notion that distorted values are likely 
to be close to the truth, as is the case when modeling
typographical errors.

\paragraph{Posterior distribution.} 
The generative process described above corresponds to a posterior distribution 
over the model parameters, conditioned on the observed records. 
By reading the conditional dependence structure off the plate diagram 
(Figure~\ref{fig:plate-diagram}) and marginalizing over 
the missing record attributes $\vec{X}^{(m)}$, one can show that the 
posterior distribution is of the following form:
\begin{equation}
\begin{split}
p(\vec{\Gamma}, \vec{\Lambda}, \vec{Y}, \vec{Z}, \vec{\Theta}|\vec{X}^{(o)}, \vec{O}) 
& \propto \prod_{e, a} p(y_{ea}|\phi_{a}) \times \prod_{t, a} p(\theta_{ta}|\alpha_{a}, \beta_{a}) \times \prod_{\substack{t, r, a\\o_{tra}=1}} p(x_{tra}|z_{tra}, \lambda_{tr}, y_{\lambda_{tr}a}) \\
& \qquad {} \times \prod_{t, r} \Big\{ p(\gamma_{tr}|\vec{Y}) 
p(\lambda_{tr}|\gamma_{tr}, \vec{Y}) \prod_{a} p(z_{tra}|\theta_{ta}) \Big\}.
\end{split} 
\label{eqn:partitioned-posterior}
\end{equation}
For further detail on the derivation and an expanded form of the 
posterior, we refer the reader to Appendix~A.

\subsection{Attribute similarity measures}
\label{sec:attribute-sim-measure}
We now discuss the attribute similarity measures that appear 
in the distortion distribution of Equation~\ref{eqn:distortion-dist}.
The purpose of these measures is to quantify the 
propensity that some value $v$ in the attribute domain is 
chosen as a distorted alternative to the true value $w$.
\begin{definition}[Attribute similarity measure]
  \label{def:attribute-sim-measure}
  Let $\valset$ be the domain of an attribute.
  An \emph{attribute similarity measure} on $\valset$ is a 
  function $\simfn: \valset \times \valset \to [0, s_\mathrm{max}]$
  that satisfies $0 \leq s_\mathrm{max} < \infty$ and 
  $\simfn(v,w) = \simfn(w,v)$ for all $v, w \in \valset$.
\end{definition}

Note that our parameterization in terms of attribute \emph{similarity} 
measures differs from \blink, which uses \emph{distance} measures.
This allows us to make use of a more efficient sampling method, as described 
in Section~\ref{sec:pert-sampling}. 
The next proposition states that the two parameterizations are equivalent, 
so long as the distance measure is bounded and symmetric (a proof is 
provided in Appendix~B.1).
\begin{proposition}
  \label{thm:sim-dist-equiv}
  Let $\distfn_{a}: \valset \times \valset \to [0, d_{\mathrm{max};a}]$ 
  be the attribute distance measure that appears in \blink, 
  and assume that $0 \leq d_{\mathrm{max};a} < \infty$ and 
  $\distfn_{a}(v, w) = \distfn_{a}(w, v)$ for all $v,w \in \valset$. 
  Define the corresponding attribute similarity measure 
  for \dblink\ as
  \begin{equation}
  \simfn_{a}(v, w) := d_{\mathrm{max};a} - \distfn_{a}(v, w).
  \label{eqn:simfn-distfn-correspondence}
  \end{equation}
  Then the parameterization of %the distortion distribution 
  $\psi_{a}$ used in \dblink\ is equivalent to \blink.
\end{proposition}

In this paper, we restrict our attention to the following 
similarity measures for simplicity:
\begin{itemize}
  \item \emph{Constant similarity measure.} This measure 
  is appropriate for categorical attributes, where there is 
  no reason to believe one value is more likely than any 
  other as a distortion to the true value $w$.
  Without loss of generality, it may be defined as 
  $\simfn_\mathrm{const}(v, w) = s_\mathrm{max}$ for all 
  $v, w \in \valset$.
  \item \emph{Normalized edit similarity measure.} 
  This measure is based on the edit distance metric, and is 
  suitable for modeling distortion in generic string attributes.
  Following~\cite{yujian_normalized_2007}, we define a normalized 
  edit distance metric,
  \begin{equation*}
    \distfn_{\mathrm{nEd}}(v,w) = \frac{2\ \distfn_{\mathrm{Ed}}(v,w)}
    {|v| + |w| + \distfn_{\mathrm{Ed}}(v,w)},
  \end{equation*}
  where $\distfn_{\mathrm{Ed}}$ denotes the regular edit distance 
  and $|v|$ denotes the length of string $v$.
  Note that alternative definitions of the normalized edit distance 
  could be used \citep[see references in][]{yujian_normalized_2007}, 
  however the above definition is unique in that it yields a 
  proper metric.
  Since the normalized edit distance is bounded on the interval 
  $[0,1]$ we can define a corresponding normalized edit similarity 
  measure:
  \begin{equation}
    \simfn_{\mathrm{nEd}}(v, w) = 1 - \distfn_{\mathrm{nEd}}(v, w).
  \label{eqn:norm-edit-sim-fn}
  \end{equation}
\end{itemize}
Ideally, one should select attribute similarity measures based on 
the data at hand.
There are many possibilities to consider, such as  
Jaccard similarity, numeric similarity 
measures~\citep{lesot_similarity_2008} and other 
domain-specific measures~\citep{bilenko_adaptive_2003}.

\subsection{Model equivalence}
\label{sec:posterior-dist}
We have purposely constructed \dblink\ so that it  
reduces to \blink\ under certain conditions. 
Assuming the records are fully observed, the posterior distribution 
of \dblink\ as specified in Equation~\ref{eqn:partitioned-posterior} 
is similar to \blink. 
The difference lies in the factors involving the block assignments 
$\gamma_{tr}$ and the entity assignments $\lambda_{tr}$. 
However, if one marginalizes out the auxiliary block assignments---as is 
done automatically in Markov chain Monte Carlo---the 
posterior distributions are identical.
This statement is made precise below (proof provided in Appendix~B.2):
\begin{proposition}
  \label{thm:posterior-equiv}
  Suppose the conditions of Proposition~\ref{thm:sim-dist-equiv} hold
  and that $\alpha_{a} = \alpha$ and $\beta_{a} = \beta$ for 
  all $a$. 
  Assume furthermore that all record attributes are observed, i.e.\ 
  $o_{tra} = 1$ for all $t, r, a$.
  Then the marginal posterior of $\vec{\Lambda}$, $\vec{Y}$, 
  $\vec{Z}$ and $\vec{\Theta}$ for \dblink\ (i.e. marginalized 
  over $\vec{\Gamma} = [\gamma_{tr}]_{t = 1 \ldots T; r = 1 \ldots R_t}$)
  is identical to the posterior for \blink.
\end{proposition}

This is an important result, as it shows our inferences for the 
meaningful model parameters are the same as we would obtain from \blink. 
Thus we are able to apply blocking to scale the model, without compromising 
the correctness of the posterior distribution.

\subsection{Rationale for introducing block}
\label{sec:partition-motivation}
We now briefly explain the role of the auxiliary block in \dblink.
First, we note that without the block ($B = 1$), the Markov blanket for 
$\lambda_{tr}$ includes the attribute values for \emph{all} of the 
entities $\vec{Y}$.
This presents a major obstacle when it comes to distributing the inference 
on a compute cluster, as the data is not separable.
By incorporating block, we restrict the Markov blanket for $\lambda_{tr}$ 
to include only a subset of the entity attribute values---those in the 
same block as record $(t,r)$.
As a result, it becomes natural to distribute the inference so that each 
compute node is responsible for a single block (see 
Section~\ref{sec:distributed-pcg} for details).
Secondly, we can interpret the block as performing probabilistic blocking in 
the context of MCMC sampling (introduced in Section~\ref{sec:inference}), 
which improves computational efficiency.
In a given iteration, the possible links for a record are restricted to the 
entities residing in the same block.
However, unlike conventional blocking, the block assignments are not 
fixed---between iterations the entities and linked records may move 
between blocks.

\section{Blocking functions}
\label{sec:partition-fn}
In Section~\ref{sec:model-specification} we introduced a generic blocking 
function (Equation~\ref{eqn:part-fn}) that is responsible for assigning 
entities to blocks.
This function may be regarded as a free parameter, 
since it has no bearing on model equivalence according to 
Proposition~\ref{thm:posterior-equiv}.
However, from a practical perspective the blocking function ought to be 
chosen carefully, as it can impact inferential efficiency---both in 
terms of computational and mixing time.
We suggest some guidelines for choosing a blocking function in 
Section~\ref{sec:interpret}, before presenting an example based on $k$-d trees 
in Section~\ref{sec:kd}.

\subsection{Interpretation and guidelines}
\label{sec:interpret}
Recall that the blocking function assigns an entity to 
a block according to its attributes 
$\vec{y}_{e} = [y_{ea}]_{a = 1 \ldots A}$.
Since $\vec{y}_{e}$ is \emph{unobserved}, it must be treated 
as a random variable over the space of possible attributes 
$\valset_\otimes := \bigotimes_{a=1}^{A} \valset_{a}$.
This means the blocking function should not be interpreted 
as partitioning the entities directly.
Rather, it should be interpreted as partitioning the space 
$\valset_\otimes$ in which the entities reside, while taking 
the distribution over $\valset_\otimes$ into account.
With this interpretation in mind, we argue that the 
blocking function should ideally satisfy the following 
properties:
\begin{enumerate}[(i)]
  \item \emph{Balanced weight.} The blocks should have 
  equal weight (probability mass) under the distribution 
  over $\valset_\otimes$, thereby ensuring the entities are 
  distributed evenly (in expectation) among the blocks.
  This is a desirable property, as it ensures proper load balancing 
  for our distributed inference algorithm (see 
  Section~\ref{sec:distributed-pcg}).
  \item \emph{Entity separation.} A pair of entities drawn at 
  random from the same block should have a high degree of 
  similarity, while entities drawn from different blocks 
  should have a low degree of similarity.
  This improves the likelihood that similar records will 
  end up in the same block, and allows them to more 
  readily form likely entities.
\end{enumerate}

These properties need not be satisfied strictly: the extent to which 
they are satisfied is merely expected to improve the efficiency of 
the inference.
For example, satisfying the first property requires 
knowledge of the marginal posterior distribution 
over $\vec{y}_{e}$, which is infeasible to calculate.
We note that there is likely to be tension between the two properties, 
so that a balance must be struck between them.

\subsection{Example: \emph{k}-d tree blocking function}
\label{sec:kd}
We now describe a blocking function based on $k$-d trees, 
which is used in our experiments in Section~\ref{sec:experiments}.

\paragraph{Background.} A $k$-d tree is a binary tree that 
recursively partitions a $k$-dimensional affine 
space~\citep{bentley_multidimensional_1975, 
  friedman_algorithm_1977}.
In the standard set-up, each node of the tree is associated 
with a data point that implicitly splits the input space 
into two half-spaces along a particular dimension.
Owing to its ability to hierarchically group nearby 
points, it is commonly used to speed up nearest-neighbor 
search.
This makes a $k$-d tree a good candidate for a blocking function, 
since it can be balanced while grouping similar points.

\paragraph{Setup.} Our setup differs from a standard $k$-d tree in several 
aspects.
First, we consider a discrete space $\valset_\otimes$ (not an affine space), 
where the ``$k$ dimensions'' are the $A$ attributes. 
Second, we do not store data points in the tree. 
We only require that the tree implicitly stores the boundaries of the blocks, 
so that it can assign an arbitrary $\vec{y} \in \valset_\otimes$ to the 
correct partition (a leaf node).
Finally, since we are working in a discrete space, the input space to a node 
is a countable set. 
The node must split the input set into two parts based on the values of one 
of the attributes.

\paragraph{Fitting the tree.} Since it is infeasible to calculate the 
marginal posterior distribution over $\vec{y}_{e}$ exactly, we use the 
empirical distribution from the tables as an approximation.
As a result, we treat the records (tables) as a sample from the distribution 
over $\vec{y}_{e}$, and fit the tree so that it remains balanced with 
respect to this sample.
The depth of the tree $d$ determines the number of blocks ($2^d$).

\paragraph{Achieving balanced splits.}
When fitting the tree, each node receives an input set of samples and a rule 
must be found that splits the set into two roughly equal (balanced) parts 
based on an attribute.
We consider two types of splitting rules: the \emph{ordered median} and 
the \emph{reference set} (see Appendix~C).
We allow the practitioner to specify an ordered list of attributes to be 
used for splitting.
To ensure balanced splits, we recommend selecting attributes with a 
large domain.
If possible, we recommend preferencing attributes which are known a priori 
to be reliable (low distortion), as this will reduce the shuffling of 
entities\slash records between blocks.
In principle, it is possible to automate the process of fitting a tree: 
one could grow several trees with randomly-selected splits and use the 
one that is most balanced.
We examine balance empirically in Appendix~H.

\section{Inference}
\label{sec:inference}
We now turn to approximating the full joint posterior distribution 
over the unobserved variables $\vec{Z}$, $\vec{Y}$, $\vec{\Theta}$, 
$\vec{\Gamma}$ and $\vec{\Lambda}$, as given in 
Equation~\ref{eqn:partitioned-posterior}.
Since it is infeasible to sample from this distribution directly, 
we design MCMC algorithms based on partially-collapsed Gibbs (PCG) 
sampling~\citep{dyk_partially_2008}.
In addition, we show how to exploit the conditional independence 
induced by the blocks to distribute the PCG sampling across multiple 
threads or machines.

\subsection{Partially-collapsed Gibbs sampling}
\label{sec:pcg-sampling}
Following the \blink\ paper~\citep{steorts_entity_2015}, we initially 
experimented with regular Gibbs sampling.\footnote{
  We define \emph{regular} Gibbs sampling as the most basic variation 
  where variables are updated iteratively one-at-a-time by sampling 
  from their conditional distributions.
} 
However, the resulting Markov chains exhibited slow convergence and poor 
mixing.
This is a known shortcoming of Gibbs sampling which may be remedied by 
collapsing variables and\slash or updating correlated variables in 
groups~\citep{liu_monte_2004}.
These ideas form the basis for a framework called \emph{partially-collapsed 
Gibbs (PCG) sampling}---a generalization of Gibbs sampling that has better 
convergence properties~\citep{dyk_partially_2008}.

Under the PCG framework, variables are updated in groups by sampling 
from their conditional distributions.
These conditional distributions may be taken with respect to the 
joint posterior (like regular Gibbs), or with respect to 
\emph{marginal distributions} of the joint posterior (unique to PCG).
The latter case is called \emph{trimming} and must be handled with 
care so as not to alter the stationary distribution of the Markov 
chain.

In applying PCG sampling to \dblink, we must decide how to apply the 
three tools: \emph{marginalization} (equivalent to grouping), 
\emph{permutation} (changing the order of the updates) and 
\emph{trimming} (removing marginalized variables).
In theory, the convergence rate should improve with more marginalization and 
trimming, however this must be balanced with the following: 
(i)~whether the resulting conditionals can be sampled from 
efficiently, and 
(ii)~whether the resulting dependence structure is 
compatible with our distributed set-up (see 
Section~\ref{sec:distributed-pcg}).
We consider two samplers, PCG-I and PCG-II, described below.
Of the two, we recommend PCG-I as it is more efficient in our empirical 
evaluations (see Section~\ref{sec:efficiency-expts}).
We include the PCG-II sampler, as one would expect the PCG-II sampler 
to perform better than the PCG-I sampler in terms of mixing, 
however when computational efficiency is taken into account the 
performance is worse (see Figure~\ref{fig:speed-up-vs-sampler}).

\begin{figure*}
  \def\svgwidth{\linewidth}
  \import{./figures/}{distributed-transition-operator.pdf_tex}
  \caption{Schematic depicting a single iteration of distributed PCG sampling.
    The entity attributes ($\vec{Y}$---circular nodes), record 
    attributes and their distortion indicators ($\vec{X}$, 
    $\vec{Z}$---square nodes), and links from records to entities 
    ($\vec{\Lambda}$---node connectors) are distributed across the 
    workers (blue rectangular plates) according to their assigned 
    blocks.
    The distortion probabilities ($\vec{\Theta}$) reside on the manager 
    (green rounded-rectangular plate).
  }
  \label{fig:distributed-pcg}
\end{figure*}

\subsubsection{PCG-I sampler} 
The PCG-I sampler uses regular Gibbs updates for $\theta_{ta}$, 
$\lambda_{tr}$ and $z_{tra}$ for all $t$, $r$ and $a$.
The conditional distributions for these updates are listed in 
Appendix~D.
When updating the entity attributes $y_{ea}$ and the block 
assignments $\gamma_{tr}$, marginalization and trimming are 
used.
Specifically, we apply marginalization by jointly updating 
$\vec{y}_{e}$ and $\{\gamma_{tr}, \vec{z}_{tr}\}_{\entset_{e}}$ 
(the set of $\gamma_{tr}$'s and $\vec{z}_{tr}$'s for records $(t,r)$ 
linked to entity $e$).
We then trim (analytically integrate over) 
$\{\vec{z}_{tr}\}_{\entset_{e}}$.

We shall now derive this update. 
Referring to Equation~\ref{eqn:partitioned-posterior}, the joint 
posterior of $\vec{y}_{e}$, $\{\gamma_{tr}, 
\vec{z}_{tr}\}_{\entset_{e}}$ conditioned on the other parameters 
has the form
\begin{equation*}
\begin{split}
& p\!\left(\vec{y}_{e}, \{\gamma_{tr}, \vec{z}_{tr}\}_{\entset_{e}} \middle| 
  \vec{Z}^{\neg \entset_{e}}, \vec{\Gamma}^{\neg \entset_{e}}, \vec{\Theta}, 
    \vec{\Lambda}, \vec{X}^{(o)}, \vec{O}\right) \propto \\
& \quad \prod_{a} \Big\{ p(y_{ea}|\phi_{a}) \times  \prod_{(t,r) \in \entset_{e}} 
  p(\gamma_{tr}|\vec{Y}) p(\lambda_{tr}|\gamma_{tr}, \vec{Y}) p(z_{tra}|\theta_{ta}) 
    \times \prod_{\substack{(t,r) \in \entset_{e}\\o_{tra}=1}} p(x_{tra} | z_{tra}, \lambda_{tr}, y_{ea}) \Big\},
\end{split}
\end{equation*}
where superscript $\neg \entset_{e}$ denotes exclusion of 
any records $(t,r) \in \entset_{e}$ (those currently linked to entity $e$).
Substituting the distributions and trimming 
$\{\vec{z}_{tr}\}_{\entset_{e}}$ yields
\begin{equation}
p\!\left(\vec{y}_{e}, \{\gamma_{tr}\}_{\entset_{e}} \middle| 
  \vec{Z}^{\neg \entset_{e}}, \vec{\Gamma}^{\neg \entset_{e}}, \vec{\Theta}, 
  \vec{\Lambda}, \vec{X}^{(o)}, \vec{O} \right) = p(\{\gamma_{tr}\}_{\entset_{e}}|\entset_{e}, \vec{y}_{e}) 
  \prod_{a} p(y_{ea}|\entset_{e}, \vec{\Theta}, \vec{X}^{(o)}, \vec{O})
\label{eqn:y-gamma-update}
\end{equation}
where
\begin{equation*}
p(y_{ea}|\entset_{e}, \vec{\Theta}, \vec{X}^{(o)}, \vec{O}) 
  \propto \phi_{a}(y_{ea}) \prod_{\substack{(t,r) \in \entset_{e}\\o_{tra}=1}} 
  \left\{(1 - \theta_{ta}) \1{x_{tra} = y_{ea}} 
  + \theta_{ta} \psi_{a}(x_{tra}|y_{ea}) \right\}
\end{equation*}
and
\begin{equation*}
p(\{\gamma_{tr}\}_{\entset_{e}}|\entset_{e}, \vec{y}_{e}) \propto \prod_{(t,r) \in 
\entset_{e}} \1{\gamma_{tr} = \blockfn(\vec{y}_{e})}.
\end{equation*}
Note that the update for $\{\gamma_{tr}\}_{\entset_{e}}$ is 
deterministic, conditional on $\vec{y}_{e}$ and $\entset_e$.

Since we have applied trimming, we must permute the updates so that 
the trimmed variables $\vec{Z}$ are not conditioned on in later 
updates.
This means the updates for $\vec{y}_{e}$ and $\{\gamma_{tr}, 
\vec{z}_{tr}\}_{\entset_{e}}$ must come \emph{after} the updates 
for $\theta_{ta}$ and $\lambda_{tr}$, but \emph{before} the updates 
for $z_{tra}$.

\subsubsection{PCG-II sampler}
The PCG-II sampler is identical to PCG-I, except that it replaces 
the regular Gibbs update for $\lambda_{tr}$ with an update that 
marginalizes and trims $\vec{z}_{tr}$.
To derive the distribution for this update, we first consider the 
joint posterior of $\lambda_{tr}$ and $\vec{z}_{tr}$ conditioned on the 
other parameters:
\begin{equation*}
\begin{split}
& p(\lambda_{tr}, \vec{z}_{tr}|\vec{\Gamma}, \vec{Y}, \vec{\Theta}, \vec{Z}^{\neg (t,r)}, \vec{X}^{(o)}, \vec{O}) \propto \\
& \qquad p(\lambda_{tr}|\gamma_{tr}, \vec{Y}) \times \prod_{a} p(z_{tra}|\theta_{ta}) \times \prod_{\substack{a\\o_{tra}=1}} p(x_{tra}|z_{tra}, \lambda_{tr}, y_{\lambda_{tr}a})
\end{split}
\end{equation*}
where superscript $\neg(t,r)$ denotes exclusion of record $(t,r)$. 
Substituting the distributions and trimming $\vec{z}_{tr}$ yields
\begin{equation}
\begin{split}
& p(\lambda_{tr} | \vec{\Gamma}, \vec{Y}, \vec{\Theta}, \vec{Z}^{\neg (t,r)}, \vec{X}^{(o)}, \vec{O}) \propto \\
& \qquad \1{\lambda_{tr} \in \partset_{\gamma_{tr}}(\vec{Y})} 
  \times \prod_{\substack{a\\o_{tra}=1}} 
  \Big\{(1 - \theta_{ta}) \1{x_{tra} = y_{\lambda_{tr}a}} + \theta_{ta} \psi_{a}(x_{tra}|y_{\lambda_{tr}a})\Big\}.
\end{split}
\label{eqn:lambda-pcg-update}
\end{equation}

\begin{table}
  \spacingset{1}
  \centering
  \caption{Dependencies for the conditional updates used in the PCG-I 
    sampler.}
  \label{tbl:conditional-dependencies}
  \begin{center}
  \begin{tabular}{rl}
    \toprule
    Update variables & Dependencies \\
    \midrule
    $\theta_{ta}$ & $z_{t \cdot a} = \sum_{r} z_{tra}$ \\
    $\lambda_{tr}$   & $\vec{z}_{tr}$, $\vec{x}_{tr}$, $\gamma_{tr}$, 
    $\partset_{\gamma_{tr}}$, 
    $\{\vec{y}_{e}\}_{e \in \partset_{\gamma_{tr}}}$ \\
    $y_{ea}$, $\{\gamma_{tr}, z_{tra}\}_{(t,r) \in \entset_{e}}$ &  
    $\entset_{e}$, $\{x_{tra}\}_{(t,r) \in \entset_{e}}$,
    $\{\theta_{ta}\}_{(t,r) \in \entset_{e}}$ \\
    $z_{tra}$     & $x_{tra}$, $\lambda_{tr}$, $y_{\lambda_{tr}a}$, 
    $\theta_{ta}$ \\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

\subsection{Distributing the sampling}
\label{sec:distributed-pcg}
By examining the conditional distributions derived in the previous 
section and those listed in Appendix~D, one can show that the 
updates for the variables associated with entities and records 
($z_{tra}$, $\lambda_{tr}$, $\gamma_{tr}$ and $y_{ea}$) only depend 
on variables associated with entities and records assigned to the 
\emph{same} block (excluding $\vec{\Theta}$).
These dependencies are summarized in 
Table~\ref{tbl:conditional-dependencies} for the PCG-I sampler.
The distortion probability $\theta_{ta}$ is an exception---it is 
not associated with any block and may depend on $z_{tra}$'s across 
\emph{all} blocks.

This dependence structure---in particular, the conditional 
independence of entities and records across blocks---makes the 
PCG sampling amenable to distributed computing.
As such, we propose a manager-worker architecture where: 
\begin{itemize}
  \item the \emph{manager} is responsible for storing and 
  updating variables \emph{not} associated with any 
  block (i.e.\ $\vec{\Theta}$); and
  \item each \emph{worker} represents a block, and is 
  responsible for storing and updating variables associated 
  with the entities and records assigned to it.
\end{itemize}
The manager\slash workers may be processes running on a single machine 
or on machines in a cluster.
If using a cluster, we recommend that the nodes be tightly 
coupled, as frequent communication between them is required.

Figure~\ref{fig:distributed-pcg} depicts a single 
iteration of PCG sampling using our proposed 
manager-worker architecture. 
Of the four steps depicted, steps~2 and~3---where the links, 
entity attributes and block assignments are updated---are the most 
computationally intensive. 
We therefore expect to achieve a significant speed-up by 
distributing these steps across the workers.

To ensure good load balancing of these steps it is important 
that the blocks are well-balanced (see Section~\ref{sec:interpret}), 
otherwise workers responsible for smaller blocks must wait idly 
for other workers to finish before the next iteration can begin. 
This is because step~1 requires global synchronization of state 
across the workers.
The blocks also have an effect on communication costs, which 
are most significant in step~3, where the entities and linked records 
are shuffled to their newly-assigned blocks.
A well-chosen blocking function can minimize this cost, by ensuring 
similar records\slash entities are co-blocked.

\section{Computational efficiency considerations}\label{sec:tricks}

\subsection{Efficient pruning of candidate links}
\label{sec:pruning-links}
In this section, we describe a trick that is aimed at improving the 
computational efficiency of the Gibbs update for $\lambda_{tr}$ 
(used in the Gibbs and PCG-I samplers).
This particular trick does \emph{not} apply to the joint PCG update 
for $\lambda_{tr}$ and $\vec{z}_{tr}$ (used in the PCG-II sampler).

Consider the conditional distribution for the $\lambda_{tr}$ update in 
Equation~S4 of Appendix~D:
\begin{equation}
\begin{split}
& p(\lambda_{tr} = e | \vec{\Gamma}, \vec{Y}, \vec{Z}, \vec{X}^{(o)}, \vec{O}) \propto \\
& \qquad \1{e \in \partset_{\gamma_{tr}}(\vec{Y})}
  \times \prod_{\substack{a\\o_{tra}=1}} \Big\{(1 - z_{tra}) \1{x_{tra} = y_{ea}} 
  + z_{tra} \psi_{a}(x_{tra}|y_{ea}) \Big\}.
\end{split}
\label{eqn:lambda-km}
\end{equation}
The support of this distribution is the \emph{set of candidate links} 
for record $(t,r)$, which we denote by $\mathcal{L}_{tr}$.
Looking at the first indicator function above, we see that 
$\mathcal{L}_{tr} \subseteq \partset_{\gamma_{tr}}$, i.e.\ the 
candidate links are restricted to the entities in the 
\emph{same block} as record $(t,r)$. 
Thus, a na\"{i}ve sampling approach for this distribution takes 
$O(|\partset_{\gamma_{tr}}|)$ time.

We can improve upon the na\"{i}ve approach by exploiting the fact that 
$\mathcal{L}_{tr}$ is often considerably smaller than 
$\partset_{\gamma_{tr}}$.  
To see why this is the case, note that the second indicator function 
in Equation~\ref{eqn:lambda-km} further restricts $\mathcal{L}_{tr}$ 
if any of the distortion indicators for the observed record attributes 
are zero.
Specifically, if $z_{tra} = 0$ and $o_{tra}=1$, $\mathcal{L}_{tr}$ cannot 
contain any entity  whose $a$-th attribute $y_{ea}$ does not match the 
record's $a$-th attribute $x_{tra}$.
This implies $\mathcal{L}_{tr}$ is likely to be small in the case 
of low distortion.

Putting aside the computation of $\mathcal{L}_{tr}$ for the moment, 
this means we can reduce the time required to update 
$\lambda_{tr}$ to $O(|\mathcal{L}_{tr}|)$.
To compute $\mathcal{L}_{tr}$ efficiently, we propose maintaining 
an inverted index over the entity attributes within each block.
Specifically, the index for the $a$-th attribute in 
block~$b$ should accept a query value $v \in \valset_{a}$ 
and return the set of entities that match on $v$:
\begin{equation}
\mathcal{M}_{pa}(v) = \{n \in \partset_{p}: y_{ea} = v\}.
\end{equation}
Once the index is constructed, we can efficiently retrieve the 
set of candidate links for record $(t,r)$ by computing a 
multiple set intersection:
\begin{equation}
\mathcal{L}_{tr} = \bigcap_{\{a : z_{tra} = 0 \wedge o_{tra} = 1\}}
\mathcal{M}_{\gamma_{tr}a}(x_{tra}).
\label{eqn:set-intersection}
\end{equation}
This assumes at least one of the observed record attributes 
is not distorted. 
Otherwise $\mathcal{L}_{tr} = \partset_{\gamma_{tr}}$.

Since the sizes of the sets $\mathcal{M}_{\gamma_{tr}a}(x_{tra})$ 
are likely to vary significantly, we advise computing 
the intersection iteratively in increasing order of size.
That is, we begin with the smallest set and retain the elements 
that are also in the next largest set, and so on.
With a hash-based set implementation, this scales linearly 
in the size of the first (smallest) set.

\subsection{Caching and truncation of attribute similarities}
\label{sec:trunc-sim}
We have not yet emphasized that the updates for $\vec{\Lambda}$, 
$\vec{Y}$ and $\vec{\Gamma}$ depend on the attribute 
similarities between pairs of values in the attribute domains.
Specifically, for each attribute $a$, we need to access 
the indexed set $\mathcal{S}_{a} = 
  \{\simfn_{a}(v,w): v, w \in \valset_{a} \times \valset_{a} \}$.
These similarities may be expensive to evaluate on-the-fly, 
so we cache the results in memory on the workers.

To manage the quadratic scaling of $\mathcal{S}_{a}$, and 
in anticipation of another trick introduced in 
Section~\ref{sec:pert-sampling}, we transform the similarities so that 
those \emph{below} a cut-off $s_{\mathrm{cut};a}$ are regarded 
as completely disagreeing.
We achieve this by applying the following truncation transformation 
to the raw attribute similarity $\simfn_{a}(v,w)$:
\begin{equation}
\truncsimfn_{a}(v,w) = 
  \max \left(0, \ \frac{\simfn_{a}(v,w) - s_{\mathrm{cut};a}}
    {1 - s_{\mathrm{cut};a}/s_{\mathrm{max};a}} \right).
\end{equation}
as illustrated in Figure~\ref{fig:truncated-sim-fn}.
\begin{figure}
  \centering
  \def\svgwidth{0.45\linewidth}
  \import{./figures/}{truncated-string-function.pdf_tex}
  \caption{Transformation from a raw similarity function 
    ($\simfn$) to a truncated similarity 
    function ($\truncsimfn$).}
  \label{fig:truncated-sim-fn}
\end{figure}
Whenever a raw attribute similarity is called for, we 
replace it with this truncated version.
Only pairs of values with positive truncated similarity are stored 
in the cache---those not stored in the cache have a truncated 
similarity of zero by default.
Note that attributes with a constant similarity function 
$\simfn_{\mathrm{const}}$ are treated specially---there is no 
need to cache the index set of similarities, since they are all 
identical.

It is important to acknowledge that the truncated similarities 
are an approximation to the original model.
We claim that the approximation is reasonable on the following grounds:
\begin{itemize}
  \item \emph{Low loss of information.} 
  Below a certain cut-off, the attribute similarity function is
  unlikely to encode much useful information for modeling the 
  distortion process.
  For example, the fact that
  $\simfn_\mathrm{nEd}(\text{``Smith''},\text{``Chiu''}) = 0.385$ 
  whereas 
  $\simfn_\mathrm{nEd}(\text{``Smith''},\text{``Chen''}) = 0.286$, 
  doesn't necessarily suggest that ``Chiu'' is more likely 
  than ``Chen'' as a distorted alternative to ``Smith''.
  \item \emph{Precedent.}
  In the record linkage literature, value pairs with similarities 
  below a cut-off are regarded as completely 
  disagreeing~\citep{winkler_methods_2002, enamorado_using_2019}.
  \item \emph{Efficiency gains.} 
  As we shall soon see in Section~\ref{sec:pert-sampling}, 
  we can perform the combined $\vec{Y}$, $\vec{\Gamma}$, $\vec{Z}$ 
  update more efficiently by eliminating pairs below the cut-off from 
  consideration. 
\end{itemize}

\subsection{Fast updates of entity attributes using perturbation sampling}
\label{sec:pert-sampling}
We now present a novel sampling algorithm that allows us to efficiently 
perform the PCG update for $y_{ea}$ and $\{\gamma_{tr}, 
z_{tra}\}_{\entset_{e}}$.
The algorithm relies on the observation that the conditional 
distribution for $y_{ea}$ can be expressed as a mixture over two 
components:
\begin{enumerate}[(i)]
  \item a \emph{base distribution} over $\valset_{a}$ which is ideally 
  constant for all entities; and
  \item a \emph{perturbation distribution} which varies for each entity, 
  but has a much smaller support than $\valset_{a}$.
\end{enumerate}
With this representation, we can avoid computing and sampling from 
the full distribution over $\valset_{a}$, which varies for each 
$y_{ea}$ update.
Rather, we only need to compute the perturbation distribution over 
a much smaller support, and then sample from the mixture, which can 
be done efficiently using the Vose-Alias 
method~\citep{vose_linear_1991}.
We refer to this algorithm as \emph{perturbation sampling}.

\subsubsection{Perturbation sampling}
Although we're interested in applying perturbation sampling to a 
specific conditional distribution, we describe the idea in 
generality below.

Consider a target probability mass function (pmf) $p(x|\omega)$ with 
finite support $\mathcal{X}$, which varies as a function of parameters 
$\omega \in \Omega$.
In general, one must recompute the probability tables to draw a new 
variate whenever $\omega$ changes---a computation that takes 
$O(|\mathcal{X}|)$ time.
However, if the dependence on $\omega$ is of a certain restricted 
form, we show that it is possible to achieve better scalability by 
expressing the target as a mixture.
This is made precise in the following result.
\begin{proposition}
  \label{thm:pert-sampling}
  Let $p(x|\omega)$ be a pmf with finite support $\mathcal{X}$, which depends 
  on parameters $\omega \in \Omega$.
  Suppose there exists a ``base'' pmf $q(x)$ over $\mathcal{X}$ which is 
  independent of $\omega$ and a non-negative bounded perturbation term 
  $\epsilon(x|\omega)$, such that $p(x|\omega)$ can be factorized as 
  $p(x|\omega) \propto q(x) (1 + \epsilon(x|\omega))$.
  Then $p(x|\omega)$ can be expressed as a mixture over the base pmf 
  $q(x)$ and a ``perturbation'' pmf $v(x|\omega) := c \, q(x) 
  \epsilon(x|\omega)$ over $\mathcal{X}^\star = \{x \in \mathcal{X}: 
  \epsilon(x|\omega) > 0\}$ as follows:
  \begin{equation}
  p(x|\omega) = \frac{c}{1 + c} q(x) + \frac{1}{1 + c} v(x|\omega)
  \end{equation}
  where $c^{-1} := 
    \sum_{x \in \mathcal{X}^\star} q(x) \epsilon(x|\omega)$.
\end{proposition}
\begin{proof}
  The result is straightforward to verify by substitution.
\end{proof}

Algorithm~S1 (in Appendix~E) shows how to apply this result to draw random 
variates from a target pmf.
Briefly, it consists of three steps:
(i)~the perturbation pmf $v$ and its normalization constant $c$ are computed;
(ii)~a biased coin is tossed to choose between the base pmf $q$ and the 
perturbation pmf $v$; and
(iii)~a random variate is drawn from the selected pmf.
If $q$ is selected, a pre-initialized Alias sampler is used to draw the 
random variate (reused for all $\omega$).
Otherwise if $v$ is selected, a new Alias sampler is instantiated. 
The result below states the time complexity of this algorithm 
(see Appendix~E for a proof).
\begin{proposition}
  \label{thm:complexity-pert-sampling}
  Algorithm~S1 (in Appendix~E) returns a random variate from the 
  target pmf $p(x|\omega)$ for any $\omega \in \Omega$ in 
  $O(|\mathcal{X}^\star|)$ time.
\end{proposition}
This is a promising result, since the size of the perturbation support 
$|\mathcal{X}^\star|$ is typically of order 10 for our application, while the 
size of the full support $|\mathcal{X}|$ may be as large as $10^5$.
Hence, we expect a significant speed-up over the na\"{i}ve approach.

\subsubsection{Application of perturbation sampling}
We now return to our original objective: performing the joint PCG 
update for $y_{ea}$ and $\{\gamma_{tr}, z_{tra}\}_{\entset_{e}}$.
Referring to Equation~\ref{eqn:y-gamma-update}, we can express the 
conditional distribution for $y_{ea}$ (i.e. the target distribution) as
\begin{equation}
p(y_{ea} = v|\entset_{e}, \vec{\Theta}, \vec{X}^{(o)}, \vec{O}) \propto 
  q_{a}(v|\entset_{e}, \mathbf{O}) \left(1 + 
  \epsilon_{a}(v|\entset_{e}, \vec{\Theta}, \vec{X}^{(o)}, \vec{O}) \right).
\label{eqn:y-gamma-perturb}
\end{equation}
The base distribution is given by
\begin{equation}
q_{a}(v|\entset_{e}, \vec{O}) \propto 
\phi_{a}(v) \left( h_{a}(v) \right)^{n_a(\entset_{e}, \vec{O})}
\label{eqn:y-gamma-base}
\end{equation}
where $n_a(\entset_{e}, \vec{O}) = |\{(t,r) \in \entset_{e}: 
o_{tra}=1\}|$ is the number of records linked to entity $e$ with 
observed values for attribute $a$; and the perturbation term is 
given by
\begin{equation}
\epsilon_{a}(v|\entset_{e}, \vec{\Theta}, \{x_{tra}\}_{\entset_{e}}) = 
  \prod_{\substack{(t,r) \in \entset_{e}\\o_{tra}=1}} \Bigg\{ 
  \euler^{\truncsimfn_{a}(x_{tra}, v)} + 
  \frac{(\theta_{ta}^{-1} - 1) \, \1{x_{tra} = v}}{\phi_{a}(x_{tra}) h_{a}(x_{tra})} 
  \Bigg\} - 1.
\label{eqn:y-gamma-pert-term}
\end{equation}

The full support of the target pmf is $\valset_{a}$, 
while the perturbation support is given by
\begin{equation*}
\{x_{tra}: (t,r) \in \entset_{e} \wedge o_{tra} = 1\} \ \cup 
  \{v \in \valset_{a}: \truncsimfn_{a}(v, x_{tra}) > 0 \wedge 
  o_{tra} = 1 \text{ for any } (t,r) \in \entset_{e}\}.
\end{equation*}
In words, this set consists of the observed values for attribute~$a$ 
in the records linked to entity~$e$, plus any sufficiently similar 
values from the attribute domain (for which the truncated similarity 
is non-zero).
The size of the perturbation set will vary depending on the cut-off 
used for the truncation transformation---the higher the cut-off, 
the smaller the set.
This implies that there is a trade-off between efficiency (small 
perturbation set) and accuracy (lower cut-off).

\begin{remark}
  The astute reader may have noticed that the base distribution $q_{a}$ 
  given in Equation~\ref{eqn:y-gamma-base} is not completely 
  independent of the conditioned parameters, as is required by 
  Proposition~\ref{thm:pert-sampling}.
  In particular, $q_{a}$ depends on $n_a(\entset_{e}, 
  \vec{O})$---roughly the size of entity $e$.
  Fortunately, we expect the range of regularly encountered entity sizes to 
  be small, so we sacrifice some memory by instantiating multiple Alias 
  samplers for each $n_a(\entset_{e}, \vec{O})$ in some expected range.
  In the worst case, when a value is encountered outside the expected 
  range and the base distribution is required (unlikely since the 
  weight on the base component is typically small), we instantiate 
  the base distribution on-the-fly (same asymptotic cost as the 
  na\"{i}ve approach).
\end{remark}

\section{Empirical evaluation}
\label{sec:experiments}

\newcolumntype{Y}{>{\centering\arraybackslash}X}
\begin{table*}
  \spacingset{1}
  \centering
  \caption{Summary of data sets. 
    Those marked with a `$\star$' are synthetic.}
  \footnotesize
  \label{tbl:data-sets}
  \begin{center}
  \begin{tabularx}{\textwidth}{l *{3}{c} *{2}{Y}}
    \toprule
    Data set & \# records ($R$) & \# tables ($T$) & \# entities & 
    \multicolumn{2}{c}{\# attributes ($A$)} \\ 
    \cmidrule{5-6}
    & & & & {\footnotesize categorical} & {\footnotesize string } \\
    \midrule
    $\star$ \texttt{ABSEmployee} & 660,000 & 3 & 400,000 & 4 & 0 \\
    \texttt{NCVR}        & 448,134 & 2 & 296,433 & 3 & 3 \\
    \texttt{NLTCS}       &  57,077 & 3 &  34,945 & 6 & 0 \\
    \texttt{SHIW0810}    &  39,743 & 2 &  28,584 & 8 & 0 \\
    $\star$ \texttt{RLdata10000} &  10,000 & 1 &   9,000 & 2 & 3 \\
    \bottomrule
  \end{tabularx}
  \end{center}
\end{table*}

We present an evaluation of \dblink\ using two synthetic and three real 
data sets, as summarized in Table~\ref{tbl:data-sets}. 
The data sets include applications such as ER of employees in administrative 
and survey data (\texttt{ABSEmployee}), ER of voters in registration 
databases (\texttt{NCVR}) and ER of respondents in anonymized survey 
data (\texttt{SHIW0810}).
All results presented here were obtained using a local server in pseudocluster 
mode, however some were replicated on a cluster in the Amazon public cloud 
(see Appendix~G) to test the effect of higher communication costs.
Further details about the data sets, hardware, implementation and parameter 
settings are provided in Appendix~F.

\subsection{Computational and sampling efficiency}
\label{sec:efficiency-expts}
Following \cite{turek_efficient_2016}, we measured the efficiency using the 
rate of effective samples produced per unit time (ESS rate), which 
balances sampling efficiency (related to mixing\slash autocorrelation) and 
computational efficiency.
We used the \texttt{mcmcse} R package~\citep{mcmcse} to compute the 
effective sample size (ESS), which implements a multivariate method 
proposed by \citet{vats_multivariate_2019}.

Since the number of variables in the model is unwieldy (there are at least 
$(E + R + T) A + R$ unobserved variables) we computed the ESS for the 
following summary statistics:
\begin{itemize}
  \item the number of observed entities (scalar);
  \item the aggregate distortion for each attribute (vector); and
  \item the cluster size distribution (vector containing 
  frequency of 0-clusters, 1-clusters, 2-clusters, etc.).
\end{itemize}

\paragraph{\dblink\ versus \blink.}
We compared \dblink\ (using the PCG-I sampler) to our own implementation of 
\blink\ (i.e.\ a Gibbs sampler without any of the tricks described in 
Section~\ref{sec:tricks}).
For a fair comparison, we switched off blocking in \dblink.
We used the relatively small \texttt{RLdata10000} data set, as \blink\ 
cannot cope with larger data sets.
Figure~\ref{fig:convergence-vs-impl} contains trace plots for two 
summary statistics as a function of running time.
It is evident that \blink\ has not converged to the equilibrium 
distribution within the allotted time of 11 hours, while \dblink\ 
converges to equilibrium in 100 seconds.
Looking solely at the time per iteration, \dblink\ is at least 200$\times$ 
faster than \blink.

\begin{figure}[t]
  \centering
  \includegraphics{convergence-num-ent-blink-dblink.pdf} \quad
  \includegraphics{convergence-agg-dist-blink-dblink.pdf}
  \caption{Comparison of convergence rates for \dblink\ and 
  \blink. 
  The summary statistics for \dblink\ (number of observed entities 
  on the left and attribute distortions on the right) rapidly converge 
  to equilibrium, 
  while those for \blink\ fail to converge within 11 hours.}
  \label{fig:convergence-vs-impl}
\end{figure}

\paragraph{Blocking and efficiency.}
We tested the effect of varying the number of blocks $B$ on the efficiency 
of \dblink.
For each value of $B$, we computed the ESS rate 
averaged over 3000 iterations.
We used the \texttt{NLTCS} data set and the PCG-I sampler.
Figure~\ref{fig:speed-up-vs-num-partitions} presents the results in terms of 
the speed-up relative to the ESS rate for $B = 1$. 
We observe a near-linear speed-up in $B$, with the exception of $B=32$.
The speed-up is expected to taper off with increasing numbers of blocks, 
as parallel gains in efficiency are overcome by losses due to communication 
costs and\slash or poorer mixing.
This tipping point seems difficult to predict for a given set up, as it 
depends on complex factors such as the data distribution, the splitting rules 
used, and the hardware characteristics.

\begin{figure}[t]
  \centering
  \includegraphics{plot-partitions-speed-up-muppet.pdf}
  \caption{Efficiency of \dblink\ as a function of the
    number of blocks $B$ and summary statistic of interest 
    (larger is better).
    The speed-up measures the ESS rate relative to the ESS rate 
    for $B = 1$ (no blocking) for the \texttt{NLTCS} data set.}
  \label{fig:speed-up-vs-num-partitions}
\end{figure}

\paragraph{Sampling methods and efficiency.}
We evaluated the efficiency of the three samplers introduced in 
Section~\ref{sec:pcg-sampling} (Gibbs, PCG-I and PCG-II).
As above, we computed the ESS rate as an average over 3000 iterations.
We set $B = 16$ and used the \texttt{NLTCS} data set.
The results, shown in Figure~\ref{fig:speed-up-vs-sampler}, 
indicate that the PCG-I sampler is considerably more efficient (by a factor 
of 1.5--2$\times$) than the baseline Gibbs sampler for this data set.
We also observe that the PCG-II sampler performs quite poorly in comparison: 
between 20--30$\times$ slower than the Gibbs sampler.
This is because the marginalization and trimming for the $\vec{\Lambda}$ 
update for PCG-II prevents us from applying the trick described in 
Section~\ref{sec:pruning-links}.
Thus although PCG-II is expected to be more efficient in terms of 
reducing autocorrelation, it is less efficient overall as each iteration 
is too computationally expensive.

\begin{figure}[t]
  \centering
  \includegraphics{plot-sampler-speed-up-muppet.pdf}
  \caption{Efficiency of \dblink\ as a function of the 
    sampler and summary statistic of interest (larger is better).
    All measurements are for the \text{NLTCS} data set with 
    $B = 16$.
    }
  \label{fig:speed-up-vs-sampler}
\end{figure}

\subsection{Linkage quality}
\label{sec:linkage-quality}
Though not our primary focus, we assessed the performance of \dblink\ 
in terms of its predictions for the linkage structure (the matching step) 
for the data sets in Table~\ref{tbl:data-sets}.
This was not previously possible with \blink, as it could only scale to 
small data sets of around 1000 records.

\paragraph{Point estimate methodology.}
To evaluate the matching performance of \dblink\ with respect to the ground 
truth, we extracted a point estimate of the linkage structure from the 
posterior using the \emph{shared most probable maximal matching sets (sMPMMS)} 
method~\citep{steorts_bayesian_2016}.
This method circumvents the problem of label 
switching~\citep{jasra_markov_2005}---where the identities of the entities 
do not remain constant along the Markov chain.

The sMPMMS method involves two main steps.
In the first step, the most probable entity cluster is computed for each 
record based on the posterior samples.
In general, these entity clusters will conflict with one another---e.g.\ 
the most probable entity cluster for $r_1$ might be $(r_1, r_2)$ while for 
$r_2$ it is $(r_1, r_2, r_3)$.
The second step resolves these conflicts by assigning precedence to links 
between records and their most probable entity clusters.
The result is a globally-consistent estimate of the linkage structure---i.e.\ 
it satisfies transitivity.

We distributed the computation of the sMPMMS method in the Spark framework.
We used $9000$ approximate posterior samples which were derived from a 
Markov chain of length $10^5$ by discarding the first $10^4$ iterations 
as burn-in\footnote{We applied a burn-in of 210k iterations for 
  \texttt{NCVR} as it was slow to converge.} 
and applying a thinning interval of 10.
These parameters were chosen by inspection of trace plots, some of which 
are reported in Appendix~L.
By contrast to the point estimates reported here, we also examined full 
posterior estimation in Appendix~I.

\paragraph{Baseline methods.}
We compared the linkage quality of \dblink\ with three baseline methods as 
described below.
We focused on (scalable) unsupervised methods as we assumed very little to 
no training data was available.
\begin{table}[t]
	\centering
	\caption{Comparison of matching quality.
		``ARI'' stands for adjusted Rand index and ``Err. \# clust.'' 
		is the percentage error in the number of clusters.}
	\label{tbl:linkage-quality}
	\spacingset{1}
  \footnotesize
  \begin{center}
	\begin{tabular}{l l *{5}{c}}
		\toprule
		Data set    & Method & \multicolumn{3}{c}{Pairwise measures} & \multicolumn{2}{c}{Cluster measures} \\
		\cmidrule(lr){3-5} \cmidrule(lr){6-7}
		& & Precision & Recall & F1-score & ARI & Err. \# clust.\ \\
		\midrule
		\multirow{5}{*}{\texttt{ABSEmployee}} 
		& \dblink\              & 0.9763 & 0.8530 & \textbf{0.9105} & \textbf{0.9105} & \textbf{+1.667\%} \\
		& Fellegi-Sunter (10)   & \textbf{0.9963} & 0.8346 & 0.9083 & --- & --- \\
		& Fellegi-Sunter (100)  & \textbf{0.9963} & 0.8346 & 0.9083 & --- & --- \\
		& Near Matching         & 0.0378 & \textbf{0.9930} & 0.0728 & --- & --- \\
		& Exact Matching        & 0.9939 & 0.8346 & 0.9074 & 0.9074 & +9.661\% \\
		\midrule
		\multirow{5}{*}{\texttt{NCVR}}
		& \dblink\              & 0.9146 & \textbf{0.9654} & \textbf{0.9393} & \textbf{0.9392} & \textbf{--3.587\%}\\
		& Fellegi-Sunter (10)   & 0.9868 & 0.7874 & 0.9083  & --- & ---\\
		& Fellegi-Sunter (100)  & 0.9868 & 0.7874 & 0.9083 & --- & --- \\
		& Near Matching         & 0.9899 & 0.7443 & 0.8497 & --- & --- \\
		& Exact Matching        & \textbf{0.9925} & 0.0017 & 0.0034 & 0.0034 & +51.09\% \\
		\midrule
		\multirow{5}{*}{\texttt{NLTCS}}
		& \dblink\              & 0.8319 & 0.9103 & 0.8693 & 0.8693 & --22.09\% \\
		& Fellegi-Sunter (10)   & \textbf{0.9094} & 0.9087 & \textbf{0.9090} & --- & --- \\
		& Fellegi-Sunter (100)  & \textbf{0.9094} & 0.9087 & \textbf{0.9090} & --- & --- \\
		& Near Matching         & 0.0600 & \textbf{0.9563} & 0.1129 & --- & --- \\
		& Exact Matching        & 0.8995 & 0.9087 & 0.9040 & \textbf{0.9040} & \textbf{+2.026\%} \\ 
		\midrule
		\multirow{5}{*}{\texttt{SHIW0810}}
		& \dblink\              & \textbf{0.2514} & 0.5396 & \textbf{0.3430} & \textbf{0.3429} & --37.65\% \\
		& Fellegi-Sunter (10)   & 0.0028 & 0.9050 & 0.0056 & --- & --- \\
		& Fellegi-Sunter (100)  & 0.0025 & \textbf{0.9161} & 0.0050 & --- & --- \\
		& Near Matching         & 0.0043 & 0.9111 & 0.0086 & --- & --- \\
		& Exact Matching        & 0.1263 & 0.7608 & 0.2166 & 0.2166 & \textbf{--37.40\%} \\ 
		\midrule
		\multirow{5}{*}{\texttt{RLdata10000}}
		& \dblink\              & 0.6334 & \textbf{0.9970} & 0.7747 & \textbf{0.7747} & 
		\textbf{--10.97\%} \\
		& Fellegi-Sunter (10)   & 0.9957 & 0.6174 & 0.7622 & --- & --- \\
		& Fellegi-Sunter (100)  & 0.9364 & 0.8734 & 0.9038 & --- & --- \\
		& Near Matching         & 0.9176 & 0.9690 & \textbf{0.9426} & --- & --- \\
		& Exact Matching        & \textbf{1.0000} & 0.0080 & 0.0159 & 0.0159 & +11.02\% \\ 
		\bottomrule
  \end{tabular}
  \end{center}
\end{table}

\begin{itemize}
  \item \emph{Exact Matching.} Links records that match on all $A$ attributes.
  It is unsupervised and ensures transitivity.
  \item \emph{Near Matching.} Links records that match on at least $L - 1$ 
  attributes.
  It is unsupervised, but does not guarantee transitivity.
  \item \emph{Fellegi-Sunter.} Links records according to a pairwise match 
  score that is a weighted sum of attribute-level dis\slash agreements.
  The weights are specified by the Fellegi-Sunter 
  model~\citep{fellegi_theory_1969} and were estimated using the   
  expectation-maximization algorithm, as implemented in the 
  \texttt{RecordLinkage} R package~\citep{sariyar_recordlinkage_2010}.
  We chose the threshold on the match score to optimize 
  the F1-score using a small amount of training data (size 10 and 100). 
  This makes the method semi-supervised.
  Note that the training data was sampled in a biased manner 
  to deal with the imbalance between the matches and non-matches
  (half with match scores above zero and half below).
  The method does not guarantee transitivity.
\end{itemize}

\paragraph{Results.}
Table~\ref{tbl:linkage-quality} presents performance measures categorized 
by data set and method.
The pairwise performance measures (precision, recall and F1-score) are 
provided for all methods, however the cluster performance measures 
(adjusted Rand Index, see~\citealp{vinh_information_2010}, and percentage 
error in the number of clusters) are only valid for methods that guarantee 
transitivity of closure (\dblink\ and Exact Matching). 
Despite being fully unsupervised, \dblink\ achieves competitive performance 
when compared to the semi-supervised Fellegi-Sunter method.
The two simple baselines, Near Matching and Exact Matching, are acceptable 
for data sets with low noise but perform poorly otherwise (e.g.\ \texttt{NCVR} 
and \texttt{RLdata10000}).
We conducted an empirical sensitivity analysis for \dblink\ with respect to 
variations in the hyperparameters.
The results for \texttt{RLdata10000} (included in Appendix~J) show that 
\dblink\ is somewhat sensitive to all of the hyperparameters tested, 
however sensitivity is in general predictable, following clear and 
intuitive trends.
One interesting observation is the fact that \dblink\ tends to overestimate 
the amount of distortion.
This is perhaps not surprising given the absence of ground truth.

\section{Application to the 2010 U.S.\ Decennial Census}
\label{sec:decennial}
National statistics agencies frequently need to link inter- or intra-agency 
data sets, for a number of purposes such as quality control. 
One critical problem in the United States (U.S.) occurs every ten years, when 
the U.S.\ Census Bureau must enumerate the population in each State as 
mandated under the U.S.\ Constitution, Article I, Section 2. 
The enumeration is used to apportion the representation of legislators, 
and to allocate resources for housing, highways, schools, assistance 
programs, and other projects that are vital to the prosperity, welfare, 
and economic growth of the U.S. 
As the country grows and becomes more diverse, it becomes more challenging 
to produce an accurate enumeration. 
Many individuals elect not to fill out census forms, which results in 
them not being counted in the enumeration. 
Other individuals may be counted multiple times due to duplicate responses. 
For example, students attending universities or private schools (living 
in group quarters) are often double counted as they are legally required 
to be counted by their university\slash school, while also being counted 
by their parents\slash guardians as part of a household.

Motivated by these data duplication issues, we apply \dblink\ to conduct 
an enumeration in the state of Wyoming. 
In order to improve coverage, we combine records from the 
2010 Decennial Census with administrative records from the 
Social Security Administration's Numerical Identification System 
(Numident).\footnote{The Numident is the Social Security Administration's 
computer database file of an abstract of the information contained in an 
application for a U.S.\ Social Security number.} 
In total, we consider 1,050,000 records representing the population of 
Wyoming: 
a subset of 494,000 records from the 2010 Decennial Census and 556,000 
records from the Numident.\footnote{These figures 
have been rounded to the nearest thousand as they are protected under 
Title~13.}
Our goal is to recover the unique individuals represented in these records
using unsupervised ER.
We apply \dblink\ using the overlapping attributes from the Census and the 
Numident: first and last name, date of birth, gender, and zip code. 
We treat first and last name as string-type attributes and the remaining 
attributes as categorical. 
To manage scalability, we utilize the $k$-d tree blocking function outlined 
in Section~\ref{sec:kd}, splitting recursively on gender and birth year at 
each level of the tree. 
Inference and MCMC diagnostics are discussed in Appendix~K.

After performing ER using \dblink, we are able to provide a posterior 
estimate of the total number of unique individuals represented in both 
data sets. 
Table~\ref{tbl:census-results} reports a point estimate based on the mean. 
The standard error is quite narrow, which is consistent with 
knowledge of the uniform prior \citep{steorts_bayesian_2016}. 
We find that our estimate is significantly larger than the unadjusted count 
of 563,626 reported by \citet{census_report_2010}. 
The difference may be explained by several factors. 
Firstly, our approach may capture individuals who are not represented 
in the Census, but who are represented in the Numident (assuming they 
have a Social Security number).
Indeed, the participation rate for the Census is known to be lower in 
Wyoming than for other states \citep{census_response}.
Secondly, there may be some double-counting for records that cannot 
be reliably linked---e.g.~due to missing or unreliable attribute values.
Thirdly, there may be minor differences in the Census data---e.g.\ 
whether blank forms are discarded or not.

To assess the reliability of ER, we report pairwise evaluation measures 
(precision, recall and F1-score) in Table~\ref{tbl:census-results}. 
These measures are computed using ground truth identifiers, 
which are available for a limited subset of the records.
To our knowledge, these are the first performance measures that have been 
published for ER of Census and administrative data at the state-level.
However, we note that the measures should be interpreted with caution, 
as the limited ground truth may not be representative of all records 
(hence the need for unsupervised methods).

\begin{table}
  \centering
  \caption{Results for ER of 2010 Census and Numident data in Wyoming. 
  Pairwise evaluation measures are computed using ground truth identifiers 
  available for a subset of the records.}
  \label{tbl:census-results}
  \footnotesize
  \begin{center}
  \begin{tabular}{*{3}{c} *{2}{c}}
    \toprule
    \multicolumn{3}{c}{Pairwise measures} & \multicolumn{2}{c}{Posterior population size} \\
    \cmidrule(lr){1-3} \cmidrule(lr){4-5}
    Precision & Recall & F1-score &    Mean & Std.~error \\
    \midrule
    0.97 &   0.84 &     0.90 & 616,000 & 5,000 \\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

We believe that \dblink\ shows promise in producing enumerations at the 
state-level, while accounting for ER uncertainty.
Moving forward, it would be beneficial to study the accuracy and scalability 
of \dblink\ in other states, to further assess the reliability of our 
methodology for conducting linkage tasks within national statistical 
agencies. 
While it is beyond the scope of this paper, we are also interested in 
incorporating additional sources of administrative data, such as tax records, 
in future work.

\section{Concluding remarks}
\label{sec:conclusions}
In this paper we have proposed \dblink---a method for performing scalable 
ER with integrated blocking in a fully Bayesian framework. 
Our approach leverages an auxiliary variable representation, which 
partitions the latent entities and records into auxiliary blocks. 
Since the auxiliary blocks are not fixed, but inferred during inference, 
we are able to propagate uncertainty between the blocking and ER stages. 
This stands in contrast with the existing literature, where blocking 
and ER are performed in two separate stages without uncertainty 
propagation. 
In addition, we have shown that our approach does not compromise 
the correctness of the marginal posterior over the model parameters. 
In other words, approximate posterior samples produced by \dblink\ 
are independent of the blocking design in the asymptotic limit.

To further improve scalability, we discussed inference for 
\dblink\ in a distributed\slash parallel setting.
We proposed a blocking function based on $k$-d trees, which achieves 
good load balancing at the block level.  
We designed a distributed partially-collapsed Gibbs sampler, 
with superior mixing properties compared to a standard Gibbs sampler. 
We also presented fast algorithms for the Gibbs updates, which 
leverage indexing data structures and perturbation sampling. 
Our empirical evaluation on five data sets demonstrated efficiency 
gains for \dblink\ in excess of 300$\times$ when compared to existing 
methods.
We also demonstrated the potential of \dblink\ in a population 
enumeration case study, using data from the 2010 Decennial 
Census and Social Security Administration. 
The resulting enumeration was output with uncertainty, and achieved 
high precision and recall.

An implementation of \dblink\ is provided as an open-source Apache Spark 
package. 
We also provide an interface for R users, for broad accessibility. 
Our software has been put in place within the United States Census 
Bureau for research purposes.

\bigskip
\begin{center}
{\large\bf SUPPLEMENTAL MATERIALS}
\end{center}

\begin{description}

\item[Appendices:] Includes proofs, further details about the 
experimental setup, and additional results. (PDF file)

\item[Code:] An implementation of \dblink\ in Apache Spark 
and a corresponding R interface. (Zip file)

\item[Data:] An archive containing data sets that we have permission to 
redistribute. (Zip file)

\end{description}

\bigskip
\begin{center}
{\large\bf ACKNOWLEDGEMENTS}
\end{center}
The authors would also like to thank the anonymous reviewers, Associate 
Editor and Editor for their valuable comments and helpful suggestions. 
N.~Marchant acknowledges the support of an Australian Government 
Research Training Program Scholarship and the AMSIIntern program
hosted by the Australian Bureau of Statistics. 
R.~C.~Steorts and A.~Kaplan acknowledge the support of NSF SES-1534412 and 
CAREER-1652431.
B.~Rubinstein acknowledges the support of Australian Research Council grant 
DP150103710.
N.~Marchant and B.~Rubinstein also acknowledge support of Australian Bureau 
of Statistics project ABS2018.363.

\bibliographystyle{jasa}
{
\footnotesize
\bibliography{dblink}
}

\end{document}

### 2021-09-22 methods_paper/fabl-paper.tex
working on edits that i stupidly lost.

@@
 Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+In the literature, bipartite record linkage is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
+
+task consists of merging two disparate datafiles containing information on two overlapping sets of entities.
 

### 2021-09-22 methods_paper/fabl-paper.tex
working on intro.

--- 2021-09-22 methods_paper/fabl-paper.tex
@@
 Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+Many statistical record linkage methods are the results of extensions from the groundbreaking work of \cite{fs, newcombe}.
+Specifically, \cite{newcombe} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match.
+While this independence assumption may not always be reasonable, it is appealing for its mathematical simplicity and relative computational ease.
+Many recent work in the statistical literature has extended the aforementioned work [Brian: add citations here: Winkler and Thibadeau, 1991, Fair, 2004, Wagner and Layne, 2014, Gill and Goldacre, 2003, Singleton, 2013).
+
+In the literature, bipartite record linkage is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle2017}.
+
+Citations to add (Fellegi and Sunter, Jaro, 1989, Winkler 1988, 1993, 1994, Belin and Rubin 1995, Larsen and Rubin 2001, Herzog et al 2007, Tancredi and Liseo, 2011, Gutman et al. 2013, others?)
+
+Talk about the advantages and then what Sadinle has added to the literature. Then talk about our contributions to the Sadinle paper.
+
+%Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}.
+Importantly, \citep{enamorado2019} modified the Fellegi-Sunter method with innovative hashing techniques, greatly expanding the scalability of the independent classification approach.
 
@@
 
-In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years.
+%In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years.
 In others, we seek to perform some other statistical procedure, such as survival analysis or regression, using linked datasets, and one must  propagate linkage uncertainty into the subsequent analysis.
-
-Record linkage is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error.
+%
+%Record linkage is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error.
 This motivates the use of probabilistic record linkage techniques that are able to model the reliability of the information and which are more robust to errors than are deterministic methods.
 Unfortunately, probabilistic record linkage is computationally intense, and many entity resolution tasks are large in nature,
-motivating the search for scalable and efficient record linkage techniques.
+%motivating the search for scalable and efficient record linkage techniques.
 
-Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}.
-This method creates comparison vectors for each pair of records in the data, and independently classifies those pairs as a match or a non-match.
+%Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}.
+
+%This method creates comparison vectors for each pair of records in the data, and independently classifies those pairs as a match or a non-match.
 Though this independence is usually not reasonable, it is appealing for its mathematical simplicity and relative computational ease.
+
 Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}.

### 2021-09-22 methods_paper/fabl-paper.tex
add citations.

@@
 
-Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen2012data}.
+Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
 Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
-Many statistical record linkage methods are the results of extensions from the groundbreaking work of \cite{fs, newcombe}.
-Specifically, \cite{newcombe} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match.
-While this independence assumption may not always be reasonable, it is appealing for its mathematical simplicity and relative computational ease.
-Many recent work in the statistical literature has extended the aforementioned work [Brian: add citations here: Winkler and Thibadeau, 1991, Fair, 2004, Wagner and Layne, 2014, Gill and Goldacre, 2003, Singleton, 2013).
+Many statistical record linkage methods and case studies are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
+Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test (or hypothesis test).
+Recent work in the statistical literature has extended the aforementioned work \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.
 
-In the literature, bipartite record linkage is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle2017}.
+%While this independence assumption may not always be reasonable, it is appealing for its mathematical simplicity and relative computational ease.
+Many recent work in the statistical literature has extended the aforementioned work [Brian: add citations here:
+
+In this paper, we consider bipartite record linkage, which is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle_bayesian_2017}.
+Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, Jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
 
@@
 This setting is known as bipartite matching, and has received much less attention in the literature.
-\citep{sadinle2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing.
+\citep{sadinle_bayesian_2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing.
 However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.
 
-In this paper, we propose an extension to the \texttt{BRL} method proposed by \citep{sadinle2017} for this very end.
+In this paper, we propose an extension to the \texttt{BRL} method proposed by \citep{sadinle_bayesian_2017} for this very end.
 We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
@@
 
-\citep{sadinle2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
+\citep{sadinle_bayesian_2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
 Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link.
@@
 
-In order to make fair comparisons against the \citep{sadinle2017} model, we
+In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
 integrate over the posterior of \(\lambda\) and rearrange terms to
@@
 For a more in-depth explanation of this function and the induced Bayes
-estimate, see \citep{sadinle2017}.
+estimate, see \citep{sadinle_bayesian_2017}.
 
@@
 \footnotesize
-\bibliography{references}
+\bibliography{biblio}
 }

### 2021-09-22 Prelim/prelim.Rmd
worked on prelim slides

@@
 ---
-title: "Prelim"
-output: html_document
+title: "Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)"
+author: "Brian Kundinger"
+output:
+  beamer_presentation:
+    includes:
+      in_header: preamble.tex  
+urlcolor: blue
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
 ```
@@
 
+```{r out.width="90%"}
+knitr::include_graphics("fs_graphic.png")
+```
+
 # Notation
 
-- File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.
+- File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.
 
-- $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors
+- $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors where $\gamma_{ij}^f \in \{1, \ldots, L_f\}$
 
-- $Z_j$
+- $Z_j = \begin{cases} 
+	i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
+	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
+	\end{cases}$
 
@@
 
-# Model
+- Fellegi and Sunter (1969)
+  - independent classification of all $n_A n_B$ record pairs. Transitive closure acheived through postprocessing
+
+- Enamorado et al (2019): \texttt{fastlink} 
+  - provides efficient and scalable FS model
+
+- Sadinle (2017) - Beta Record Linkage (\texttt{BRL}) 
+  - Bayesian method for bipartite matching, strictly enforces one-to-one matching. Does not scale well to large linkage tasks
+
+# Fast Beta Linkage (fabl)
+
+$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$
+
+$$\mathbf{m_{f}} \sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{fL_f})$$
+$$\mathbf{u_{f}} \sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{fL_f})$$
+$$Z_j | \lambda =
+\begin{cases} 
+	\frac{1}{n_A}\lambda  & z_j \leq n_A; \\
+	1-\lambda &  z_j  = n_A + 1 \\
+\end{cases}$$
+$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda})$$
+
+- Model specification allows for **parallel/distributed** computing, **hashing** of comparison vectors, and **storage efficient indexing (SEI)**
 
@@
 
-Demonstration
+- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$
 
-# Managing large data
+- Reduce data to sufficient statistics
+  - $H_p = \sum_{i,j} \mathbf{1}_{(i, j) \in h_p}$
+  - $H_{p_j} = \sum_{i} \mathbf{1}_{(i, j) \in h_p}$
+
+- Gibbs sampler over each record pair has complexity $O(n_A \times n_B \times F)$, but sampler over agreement pattern has complexity $O(P \times n_B \times F)$
+
+- Speeds up posterior calculations and sampling
+
+# Managing Large Data
+
+- **Distributed Computing** - Partition data in to chunks $\{A_I\}$ and $\{B_J\}$. Compare records, hash results, compute summary statistics in parallel, and synthesize results
+
+- **Storage Efficient Indexing (SEI)** - Remove highly unlikely record labels from memory. Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$
 
@@
 
+```{r}
+knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
+```
+
 # Simulation Results 2
+
+```{r}
+knitr::include_graphics("../notes/figures/sadinle_speed_plot2.png")
+```
+
+# Simulation Results 3
+
+```{r}
+knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
+```
 
@@
 
-- $m_{fl} = $
+- Linkage in the face of duplicates within datasets
 
-# Previous work
+- Models that allow reliability of information to differ by subgroup in the data
 
-______________________________________
+- Linkage over blocked data (allows for much larger linkage tasks)
+
+# Questions?
+
+# Additional Slides
 

### 2021-09-22 methods_paper/fabl-paper.tex
worked on prelim slides

--- 2021-09-22 methods_paper/fabl-paper.tex
@@
 
-%% Reworking introduction for Brian's paper.
-
-Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen2012data}.
-Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
-In the literature, bipartite record linkage is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
-
-task consists of merging two disparate datafiles containing information on two overlapping sets of entities.
-
-\begin{enumerate}
-\item Let's re-visit the case study of Sadinle. Lay this out and the importance of it. 
-\item Next, let's briefly review the landscape of the literature. 
-\item State Brian's contributions. 
-\item Then write a section on related work. 
-\item Then write a section regarding the layout of the paper.
-\end{enumerate}
-
-\newpage
-
+Record linkage is the task of identifying
+duplicate records across multiple data sources.
 In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years.
@@
 
-Wherever possible, we reserve superscripts for denoting field and level,
-while reserving subscripts for record indices. For example,
-\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
-distribution governing field \(f\) for matching records, and
-\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
-is product of the relevant of the appropriate \(\mathbf{m}\) parameters
-for record pair \((i,j)\). We hope that these conventions avoid
-overloaded notation in the likelihood and subsequent derivations.
+%Wherever possible, we reserve superscripts for denoting field and level,
+%while reserving subscripts for record indices. For example,
+%\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
+%distribution governing field \(f\) for matching records, and
+%\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
+%is product of the relevant of the appropriate \(\mathbf{m}\) parameters
+%for record pair \((i,j)\). We hope that these conventions avoid
+%overloaded notation in the likelihood and subsequent derivations.
 

### 2021-09-22 methods_paper/fabl-paper.tex
more writing.

--- 2021-09-22 methods_paper/fabl-paper.tex
@@
 
-Citations to add (Fellegi and Sunter, Jaro, 1989, Winkler 1988, 1993, 1994, Belin and Rubin 1995, Larsen and Rubin 2001, Herzog et al 2007, Tancredi and Liseo, 2011, Gutman et al. 2013, others?)
+The Fellegi-Sunter method and its extensions are widely popular mainly due to its simplicity and computational scalability.
+Despite this, like any method, it has limitations, that are well known in the literature.
+Specifically, the assumption of no duplications within a database implies a maximum one-to-one constraint.
+That is, a record from one database ban be linked with only one record in the other database.
+Modern methods of Fellegi-Sunter either ignore this restriction \citep{Winkler1988, belin_1995, larsen_2001}, include this as a post-processing step \citep{Jaro1989}, or include it directly in the modeling framework \citep{sadinle_bayesian_2017}.
+It is important to note that \cite{fellegi_theory_1969} ignored this restriction as well.
+In practice, it seems that one should either directly include it in the modeling framework or correct for this in a post-processing step.
 
-Talk about the advantages and then what Sadinle has added to the literature. Then talk about our contributions to the Sadinle paper.
+To our knowledge, \cite{sadinle_bayesian_2017} was the first to propose correcting the Fellegi and Sunter model for the one-to-one constraint.
+Specifically, the authors propose a Bayesian variant of the original model.
+This satisfies two goals.
+First, it allows one to quantify uncertainty regarding match or non-match status.
+In addition, posterior distributions are readily available to help estimate point estimates of interest (and quantity such estimates).
+Second, under the Bayesian paradigm, the one-to-one matching constraint is easily imbedded.
+The downside to this constraint is an added computational cost.
+Our goal is proposing a model in the spirit of \cite{sadinle_bayesian_2017}, which scales beyond the original proposed model, without sacrificing much in the way of accuracy or precision of the resulting estimates.
+
+Now, mention Brian's contributions.
+
+Then mention related work.
 

### 2021-09-22 methods_paper/fabl-paper.tex

@@
 %Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}. 
+>>>>>>> e17e46b99593a9cae2676267d508dfcd21a5d0ac
 

### 2021-09-22 methods_paper/fabl-paper.tex
working on intro.

@@
 Specifically, the assumption of no duplications within a database implies a maximum one-to-one constraint.
-That is, a record from one database ban be linked with only one record in the other database.
+That is, a record from one database can be linked with only one record in the other database.
 Modern methods of Fellegi-Sunter either ignore this restriction \citep{Winkler1988, belin_1995, larsen_2001}, include this as a post-processing step \citep{Jaro1989}, or include it directly in the modeling framework \citep{sadinle_bayesian_2017}.
@@
 To our knowledge, \cite{sadinle_bayesian_2017} was the first to propose correcting the Fellegi and Sunter model for the one-to-one constraint.
-Specifically, the authors propose a Bayesian variant of the original model.
-This satisfies two goals.
+Specifically, the authors propose a Bayesian variant of the original model that satisfies two goals.
 First, it allows one to quantify uncertainty regarding match or non-match status.
@@
 The downside to this constraint is an added computational cost.
-Our goal is proposing a model in the spirit of \cite{sadinle_bayesian_2017}, which scales beyond the original proposed model, without sacrificing much in the way of accuracy or precision of the resulting estimates.
+The goal of our paper is proposing an extension of \cite{sadinle_bayesian_2017} that scales to large databases, but does not compromise much in terms of accuracy on the original case study considered by the author.
+
+\textbf{Our Contributions}
 
 Now, mention Brian's contributions.
-
-Then mention related work.
 
@@
 
-\begin{enumerate}
-\item Let's re-visit the case study of Sadinle. Lay this out and the importance of it. 
-\item Next, let's briefly review the landscape of the literature. 
-\item State Brian's contributions. 
-\item Then write a section on related work. 
-\item Then write a section regarding the layout of the paper.
-\end{enumerate}
+%\begin{enumerate}
+%\item Let's re-visit the case study of Sadinle. Lay this out and the importance of it. 
+%\item Next, let's briefly review the landscape of the literature. 
+%\item State Brian's contributions. 
+%\item Then write a section on related work. 
+%\item Then write a section regarding the layout of the paper.
+%\end{enumerate}
 

### 2021-09-22 methods_paper/fabl-paper.tex
more writing.

@@
 
-\textbf{Our Contributions}
+\textbf{Our Contribution} 
+In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\text{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations, and to reduce computational complexity.
+Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching.
+Instead, we propose independent priors (breaking the one-to-one matching) in order to create computational speeds and gains.
+Second, we resolve the one-to-one matching issue via a post processing step.
+Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model.
+Fourth, we explore other computational speeds up based upon storage efficient indexing.
+Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices.
+To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
+pronounced ``fable'')
+
+% We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
+Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks.
+To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
+%pronounced ``fable'')
 
@@
 %Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}.
->>>>>>> e17e46b99593a9cae2676267d508dfcd21a5d0ac
 
@@
 
-Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}.
+%Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}.
 Importantly, \citep{enamorado2019} modified the Fellegi-Sunter method with innovative hashing techniques, greatly expanding the scalability of the independent classification approach.
-
-Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file.
+%
+%Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file.
 This setting is known as bipartite matching, and has received much less attention in the literature.
@@
 However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.
-
-In this paper, we propose an extension to the \texttt{BRL} method proposed by \citep{sadinle_bayesian_2017} for this very end.
-We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
-Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks.
-To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
-pronounced ``fable'')
 

### 2021-09-23 methods_paper/fabl-paper.tex
update.

@@
 \begin{abstract}
-Abstract
+In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
+In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity.
+To achieve a scalable model, we propose independent priors in order to create computational speeds and gains, where we relax the one-to-one matching assumption.
+Second, we resolve the one-to-one matching issue via a post processing step.
+Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model.
+Fourth, we explore other computational speeds up based upon storage efficient indexing.
+Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} record linkage, Bayesian methods, paralellized computing, streaming algorithms
+{\it Keywords:} record linkage, bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo
 
@@
 
-Now, mention Brian's contributions.
+\textcolor{red}{Talk about related work and weave in the El Salvadoran study.}
+
+%Now, mention Brian's contributions.
 

### 2021-09-23 methods_paper/fabl-paper.tex
update.

@@
 \begin{abstract}
-In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
+In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
 In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity.
@@
 
+The remainder of the paper is as follows.
+Section~\ref{sec:bipartite} reviews bipartite record linkage; section \ref{sec:fs} reviews the traditional approach of the Fellegi and Sunter framework before introducing our proposed work.
+
+\section{Bipartite Record Linkage and Related Work}
+As already mentioned, bipartite record linkage is the traditional set up studied in record linkage, where we assume there are two databases $\bm{X}_1$ and $\bm{X}_2$ such that there are duplications across the databases but not within each databases.
+We assume that each database has $n_1$ and $n_2$ records, and that $n_1 \leq n_2.$ Denote the number of entities simultaneously recorded by \emph{both databases} by $n_{12}$ such that $0 \leq n_{12} \leq n_2.$ Given this set up, as explained by \cite{steorts_bayesian_2016} and \cite{sadinle_bayesian_2017}, this can be viewed as bipartite record linkage.
+This is important as it allows one to view the matching matrix or linkage structure in the following way:
+\begin{align}
+\Delta_{ij} =
+\begin{cases}
+1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
+0 \quad \text{otherwise}.\\
+\end{cases}
+\end{align}
+Bipartite matching can also be viewed as a matching labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
+\begin{align}
+Z_{j} =
+\begin{cases}
+i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
+n_1 + j \quad n_1 + j \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
+\end{cases}
+\end{align}
+Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.
+
+\textcolor{red}{Now state the related work here in this literature and common approaches.}
+
+ATTN: These are just notes and need to be edited
+
+\begin{enumerate}
+\item The most common approach is FS. 
+\item ``A number of approaches directly model the information contained in the datafiles (Fortini et al., 2002; Matsakis, 2010; Liseo and Tancredi, 2011; Tancredi and Liseo, 2011; Gutman et al., 2013; Steorts et al., 2013), which requires crafting specific models for each type of field in the datafile, and are therefore currently limited to handle nominal categorical
+fields, or continuous variables modeled under normality." [Update this literature] 
+\item Classification. ``A more common way of tackling this problem is to see it as a traditional classification
+problem: we need to classify record pairs into matches and non-matches. If we have access
+to a sample of record pairs for which the true matching statuses are known, we can train a
+classifier on this sample using comparisons between the pairs of records as our predictors,
+and then predict the matching statuses of the remaining record pairs (e.g., Cochinwala et al.,
+2001; Bilenko et al., 2003; Christen, 2008; Ventura et al., 2013)." [Update this literature] 
+\end{enumerate}
+
+\section{The Fellegi Sunter Approach to Record Linkage}
+
+I think this should be reviewed so that the paper is more clear.
+
+%n_1 + j \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1 \\
+
+%describes previous work in record linkage that lays the foundation for our method, Section~\ref{sec:model}
+%3 provides our model and the Gibbs sampler used for posterior inference,
+%and Sections~\ref{sec:efficiency} and~\ref{sec:efficient-posterior} describe strategies used for more efficient and scalable
+%implementation. Then, Section~\ref{sec:simulations} demonstrates the accuracy and speed of
+%out method through two simulation studies, and Section~\ref{sec:case-studies} reveals
+%interesting properties of the method through a case study of El
+%Salvadoran casualty records. Lastly, Section~\ref{sec:discussion} is a discussion that highlights open
+%questions in the record linkage literature and motivates future work.
+
 % We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
@@
 
-\textcolor{red}{Talk about related work and weave in the El Salvadoran study.}
-
-%Now, mention Brian's contributions.
-
-%Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}.
-Importantly, \citep{enamorado2019} modified the Fellegi-Sunter method with innovative hashing techniques, greatly expanding the scalability of the independent classification approach.
-
-%\begin{enumerate}
-%\item Let's re-visit the case study of Sadinle. Lay this out and the importance of it. 
-%\item Next, let's briefly review the landscape of the literature. 
-%\item State Brian's contributions. 
-%\item Then write a section on related work. 
-%\item Then write a section regarding the layout of the paper.
-%\end{enumerate}
-
-\newpage
-
-%In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years.
-In others, we seek to perform some other statistical procedure, such as survival analysis or regression, using linked datasets, and one must  propagate linkage uncertainty into the subsequent analysis.
+%\textcolor{red}{Talk about related work and weave in the El Salvadoran study.}
 %
-%Record linkage is straightforward when provided unique identifiers or highly reliable information, but becomes difficult when data sources are plagued with error.
-This motivates the use of probabilistic record linkage techniques that are able to model the reliability of the information and which are more robust to errors than are deterministic methods.
-Unfortunately, probabilistic record linkage is computationally intense, and many entity resolution tasks are large in nature,
-%motivating the search for scalable and efficient record linkage techniques.
-
-%Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}.
-
-%This method creates comparison vectors for each pair of records in the data, and independently classifies those pairs as a match or a non-match.
-Though this independence is usually not reasonable, it is appealing for its mathematical simplicity and relative computational ease.
-
-%Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}.
-Importantly, \citep{enamorado2019} modified the Fellegi-Sunter method with innovative hashing techniques, greatly expanding the scalability of the independent classification approach.
-%
-%Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file.
-This setting is known as bipartite matching, and has received much less attention in the literature.
-\citep{sadinle_bayesian_2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing.
-However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.
-
-In what follows, Section~\ref{sec:related} describes previous work in record linkage that lays the foundation for our method, Section~\ref{sec:model}
-3 provides our model and the Gibbs sampler used for posterior inference,
-and Sections~\ref{sec:efficiency} and~\ref{sec:efficient-posterior} describe strategies used for more efficient and scalable
-implementation. Then, Section~\ref{sec:simulations} demonstrates the accuracy and speed of
-out method through two simulation studies, and Section~\ref{sec:case-studies} reveals
-interesting properties of the method through a case study of El
-Salvadoran casualty records. Lastly, Section~\ref{sec:discussion} is a discussion that highlights open
-questions in the record linkage literature and motivates future work.
+%\newpage
 
@@
 \label{sec:related}
+In this section, we review work that is most related to our proposed work.
+For surveys of the record linkage literature, see \citep{christen_2012, Herzog_2007, liseo_2013, Winkler2014, Christophides2019, Jurek2019}.
 

### 2021-09-23 Prelim/prelim.Rmd

@@
 
+\begin{columns}
+\begin{column}{0.5\textwidth}
+
+\begin{itemize}
+\item Record linkage is the task of identifying duplicate records over noisy datasets.
+
+\item Easy with unique identifiers, difficult when faced with errors
+
+\item Far ranging applications in business, public health, and human rights
+
+\end{itemize}
+
+\end{column}
+\begin{column}{0.5\textwidth}  %%<--- here
+    \begin{center}
+     picture
+     \end{center}
+\end{column}
+\end{columns}
+
 # Fellegi and Sunter
 
-```{r out.width="90%"}
+```{r out.width="100%"}
 knitr::include_graphics("fs_graphic.png")
@@
 - $Z_j = \begin{cases} 
-	i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
-	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
+	i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
+	n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
 	\end{cases}$
@@
 
-- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$
+- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$.
 
 - Reduce data to sufficient statistics
-  - $H_p = \sum_{i,j} \mathbf{1}_{(i, j) \in h_p}$
-  - $H_{p_j} = \sum_{i} \mathbf{1}_{(i, j) \in h_p}$
+  - $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
+  - $H_{p_j} = ||r_{p_j}||$
+  - $H_p = \sum_{j} H_{p_j}$
 
@@
 
-- **Storage Efficient Indexing (SEI)** - Remove highly unlikely record labels from memory. Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$
+- **Storage Efficient Indexing (SEI)** - Store at most small number $R$ many record labels in each $r_{p_j}$, remove highly unlikely record labels from memory.
+Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$.
+
+- Hashing plus SEI can reduce memory requirements by >99%.
+  - Simulation of $20,000 \times 20,000$ linkage problem with 4 fields. Naive approach requires 6.4GB of storage, hashing and SEI requires 90MB.
 

### 2021-09-23 methods_paper/fabl-paper.tex
update.

@@
 
+There are many approaches to bipartite record linkage, and we review some of the most common approaches below.
+
+\paragraph{Fellegi and Sunter}
+As already mentioned, the most common approach is the original approach of \cite{fellegi_theory_1969}.
+\cite{enamorado2019using} have extended this approach using computational speeds ups to allow for a much more scalable and practical approach, providing open source code for the community.
+
+%These approaches have led to downstream tasks, such as regression approaches in the literature \citep{tang2020}.
+
+\paragraph{Directly Modeling the Data} Instead of relying on comparison vectors, some authors will incorporate the data directly into their bipartite record linkage model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, gutman_bayesian_2013, shan2020bayesian}.
+These approaches have led to extensions regarding downstream tasks in the literature or joint models, such as regression approaches \citep{dalzell2018regression, steorts2018generalized, tancredi2020unified}.
+
+There have been extensions to directly model the data in this literature at combining many databases, however, we do not review this as it's not directly comparable with the situation of bipartite record linkage \citep{steorts_bayesian_2016, steorts_entity_2015, zanella_flexible_2016, marchant_distributed_2019, tancredi_2015_regression, }.
+
+\paragraph{Classification Approaches}
+
 \textcolor{red}{Now state the related work here in this literature and common approaches.}
 
-ATTN: These are just notes and need to be edited
+\textcolor{red}{ATTN: These are just notes and need to be edited}
 

### 2021-09-23 methods_paper/fabl-paper.tex
update.

@@
 These approaches have led to extensions regarding downstream tasks in the literature or joint models, such as regression approaches \citep{dalzell2018regression, steorts2018generalized, tancredi2020unified}.
-
-There have been extensions to directly model the data in this literature at combining many databases, however, we do not review this as it's not directly comparable with the situation of bipartite record linkage \citep{steorts_bayesian_2016, steorts_entity_2015, zanella_flexible_2016, marchant_distributed_2019, tancredi_2015_regression, }.
+There have been extensions to directly model the data in this literature where authors consider merging multiple databases (more than two) \citep{steorts_bayesian_2016, steorts_entity_2015, zanella_flexible_2016, marchant_distributed_2019, tancredi_2015_regression}.
+We do not review this literature given that it is not directly/fairly comparable with the case of bipartite record linkage.
 
 \paragraph{Classification Approaches} 
-
-\textcolor{red}{Now state the related work here in this literature and common approaches.}
-
-\textcolor{red}{ATTN: These are just notes and need to be edited}
-
-\begin{enumerate}
-\item The most common approach is FS. 
-\item ``A number of approaches directly model the information contained in the datafiles (Fortini et al., 2002; Matsakis, 2010; Liseo and Tancredi, 2011; Tancredi and Liseo, 2011; Gutman et al., 2013; Steorts et al., 2013), which requires crafting specific models for each type of field in the datafile, and are therefore currently limited to handle nominal categorical
-fields, or continuous variables modeled under normality." [Update this literature] 
-\item Classification. ``A more common way of tackling this problem is to see it as a traditional classification
-problem: we need to classify record pairs into matches and non-matches. If we have access
-to a sample of record pairs for which the true matching statuses are known, we can train a
-classifier on this sample using comparisons between the pairs of records as our predictors,
-and then predict the matching statuses of the remaining record pairs (e.g., Cochinwala et al.,
-2001; Bilenko et al., 2003; Christen, 2008; Ventura et al., 2013)." [Update this literature] 
-\end{enumerate}
+One common way of approaching the record linkage task, especially in computer science and machine learning is using a classification approach.
+One uses a model (such as random forests) to classify records into matches/non-matches.
+When access to unique identifiers is available, one can create a training data (or reference data) set of record pairs (where the true match status is known).
+Finally, one can train a classifier using the reference data set to predict the match status on any record pairs that were not included in the reference data set \citep{Bilenko2006, Christen2008, ventura2015seeing}.
 

### 2021-09-23 Prelim/prelim.Rmd
updated prelim slides

@@
 
-\begin{columns}
-\begin{column}{0.5\textwidth}
+<!-- \begin{columns} -->
+<!-- \begin{column}{0.5\textwidth} -->
 
-\begin{itemize}
-\item Record linkage is the task of identifying duplicate records over noisy datasets.
+<!-- \begin{itemize} -->
+<!-- \item Record linkage is the task of identifying duplicate records over noisy datasets. -->
 
-\item Easy with unique identifiers, difficult when faced with errors
+<!-- \item Easy with unique identifiers, difficult when faced with errors -->
 
-\item Far ranging applications in business, public health, and human rights
+<!-- \item Far ranging applications in business, public health, and human rights -->
 
-\end{itemize}
+<!-- \end{itemize} -->
 
-\end{column}
-\begin{column}{0.5\textwidth}  %%<--- here
-    \begin{center}
-     picture
-     \end{center}
-\end{column}
-\end{columns}
+<!-- \end{column} -->
+<!-- \begin{column}{0.5\textwidth}  %%<--- here -->
+<!--     \begin{center} -->
+<!--      picture -->
+<!--      \end{center} -->
+<!-- \end{column} -->
+<!-- \end{columns} -->
+
+- Record linkage is the task of identifying duplicate records over noisy datasets.
+
+- Easy with unique identifiers, difficult when faced with errors
+
+- Far ranging applications in business, public health, and human rights
+
+# Examples
+
+```{r}
+knitr::include_graphics("rl_examples.png")
+```
 
@@
 ```
+# Previous work
+
+- Fellegi and Sunter (1969)
+  - independent classification of all $n_A n_B$ record pairs. Transitive closure acheived through postprocessing
+
+- Enamorado et al (2019): \texttt{fastlink} 
+  - provides efficient and scalable FS model
+
+- Sadinle (2017) - Beta Record Linkage (\texttt{BRL}) 
+  - Bayesian method for bipartite matching, strictly enforces one-to-one matching. Does not scale well to large linkage tasks
 
@@
 
-# m's and u's
-
-# Previous work
-
-- Fellegi and Sunter (1969)
-  - independent classification of all $n_A n_B$ record pairs. Transitive closure acheived through postprocessing
-
-- Enamorado et al (2019): \texttt{fastlink} 
-  - provides efficient and scalable FS model
-
-- Sadinle (2017) - Beta Record Linkage (\texttt{BRL}) 
-  - Bayesian method for bipartite matching, strictly enforces one-to-one matching. Does not scale well to large linkage tasks
-
 # Fast Beta Linkage (fabl)
@@
 
-- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$.
+- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $h(i, j)=p.$
+
+<!-- (i, j) \in h_p$. -->
 
@@
 
-- Gibbs sampler over each record pair has complexity $O(n_A \times n_B \times F)$, but sampler over agreement pattern has complexity $O(P \times n_B \times F)$
+- Run Gibbs sampler at level of agreement patterns, not record pairs
+  - Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
+  - Use number of matches for each pattern to update $m$ and $u$
+  - Back fill record labels at the end through $r_{p_j}$
+  - Computational complexity $O(P \times n_B \times F)$ instead of $O(n_A \times n_B \times F)$
 
-- Speeds up posterior calculations and sampling
+<!-- - Gibbs sampler over each record pair has complexity $O(n_A \times n_B \times F)$, but sampler over agreement pattern has complexity $O(P \times n_B \times F)$ -->
 
@@
 - Hashing plus SEI can reduce memory requirements by >99%.
-  - Simulation of $20,000 \times 20,000$ linkage problem with 4 fields. Naive approach requires 6.4GB of storage, hashing and SEI requires 90MB.
+  - Simulation of $20,000 \times 20,000$ linkage task with 4 fields. Naive approach requires 6.4GB of storage for all-to-all comparisons, hashing and SEI requires 90MB.
+
+# Three Simulations
+
+- We compare \texttt{fabl} against \texttt{BRL} in three simulation studies
+  - Measure precision and recall on 100 simulated datasets and varying levels of error and duplication across files
+  - Measure speed when both $n_A$ and $n_B$ are increasing
+  - Measure speed when $n_A$ is increasing and $n_B = 500$ is fixed.
 
@@
 
-# El Salvador
+- Details of intuition behind $m$ and $u$ parameters
 
-# Distributed computation
+- Loss function and Bayes estimate
 
-# Storage Efficient Indexing
+- Details of one-to-one matching; conflict resolution under \texttt{fabl};
+
+- El Salvador case study (time under different values of $P$, violations of one-to-one matching)
+
+- Further details on distributed computation of comparison vectors and storage efficient indexing

### 2021-09-23 methods_paper/fabl-paper.tex
update.

@@
 
-\section{The Fellegi Sunter Approach to Record Linkage}
-
-I think this should be reviewed so that the paper is more clear.
+\section{Review of the Fellegi Sunter Method}
+In this section, we review notation, terminology, and the classical approach of Fellegi and Sunter.
+
+I think this should be reviewed so that the paper is more clear. Consider ordered record pairs $\bm{X}_1 \times \bm{X}_2.$ 
+Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$ Denote the set of non-matches 
+$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ An alternative way of merging two databases and removing 
+duplications can be viewed as identifying the sets of matches/non-matches, or rather $\bm{M}, \bm{U}.$ In the literature, it is well known that record pairs that are estimated as matches are known as links; record pairs that are estimated as non-matches are known as non-links.
+\cite{fellegi_theory_1969} proposed using all-to-all pairwise comparison of records to estimate the matches/non-matches.
+
+\newpage
 

### 2021-09-24 methods_paper/fabl-paper.tex
update.
working on fs.

@@
 \begin{abstract}
-In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
+In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
 In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity.
@@
 \noindent%
-{\it Keywords:} record linkage, bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo
+{\it Keywords:} record linkage, bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
 
@@
 \cite{fellegi_theory_1969} proposed using all-to-all pairwise comparison of records to estimate the matches/non-matches.
+
+\subsection{Comparison Data}
 

### 2021-09-24 methods_paper/fabl-paper.tex
comp data.

@@
 \subsection{Comparison Data}
+In this section, we review the notion of \emph{comparison data}.
+Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar.
+A common way to encode this is using a comparison vector $\bm{\gamma}_{ij}$ which is computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote $F$ as the criteria for compare the records such that
+$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $F$ represents to a single comparison per feature that the databases share, such as gender.
+%
+The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for nominal variables such as gender.
+For more complex measurements, such as textual or numeric feature information, we can take into account partial agreement patterns using common string metrics, such as the Edit, Jaro, or Jaro-Winkler distance functions.
+This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+
+Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$ representing disagreement levels.
+Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
+Thus, we can construct comparison vectors in the following way:
+$$\gamma_{ij} = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$
+
+\subsection{Fellegi-Sunter Model}
 

### 2021-09-24 methods_paper/fabl-paper.tex
more writing.

@@
 \subsection{Fellegi-Sunter Model}
+In this section, we review the Fellegi-Sunter model.
+While we have introduced comparison data, $\bm{\gamma}_{ij}$, this is not sufficient to determine whether a record is a match/non-match given errors that naturally occur in the data.
+This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
+\begin{align}
+w_{ij} = \log \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
+\end{align}
+as a weight to estimate if record pairs were a match/non-match.
 

### 2021-09-24 methods_paper/fabl-paper.tex
done with fs.

@@
 In this paper, we consider bipartite record linkage, which is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle_bayesian_2017}.
-Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, Jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
+Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
 
@@
 That is, a record from one database can be linked with only one record in the other database.
-Modern methods of Fellegi-Sunter either ignore this restriction \citep{Winkler1988, belin_1995, larsen_2001}, include this as a post-processing step \citep{Jaro1989}, or include it directly in the modeling framework \citep{sadinle_bayesian_2017}.
+Modern methods of Fellegi-Sunter either ignore this restriction \citep{Winkler1988, belin_1995, larsen_2001}, include this as a post-processing step \citep{jaro1989}, or include it directly in the modeling framework \citep{sadinle_bayesian_2017}.
 It is important to note that \cite{fellegi_theory_1969} ignored this restriction as well.
@@
 Thus, we can construct comparison vectors in the following way:
-$$\gamma_{ij} = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$
+$$\gamma_{ij} = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$ The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.
 
@@
 \begin{align}
+\label{eqn:wts}
 w_{ij} = \log \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
 \end{align}
-as a weight to estimate if record pairs were a match/non-match.
+as a weight to estimate if record pairs are a match/non-match.
+Assume $\bm{\Gamma}_{ij}$ is a random vector, whose distribution depends on $\Delta_{ij}.$ Then equation \ref{eqn:wts} is a representative value of $\bm{\Gamma}_{ij} \mid \Delta_{ij}.$ The ratio will be large if we favor the pair being a match.
+This is a problem in practice as transitive closures are violated.
+For example, suppose records $i$ and $j$ in $\bm{X}_1$ are extremely similar but are known to be non-matches.
+Suppose that both records $i$ and $j$ in $\bm{X}_1$ are very similar to record $k$ in $\bm{X}_2.$ By transitive closures, this implies that $i$ and $j$ are a match, which cannot be true by assumption.
+
+\cite{jaro1989} proposed addressing this issue via an optimization approach as follows:
+\begin{align}
+\label{eqn:jaro}
+&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
+\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\};\\
+& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
+& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag \\
+\end{align}
+
+Equation \label{eqn:jaro} ensures that each record of $\bm{X}_2$ only matches with at most one record in $\bm{X}_1$ (and the reverse condition).
+This is commonly known as the \emph{maximum-weight bipartite matching problem} or a \emph{linear sum assignment problem}.
+The output of this algorithm is a bipartite matching that maximizes the sum of the weights among matched pairs, where the pairs are not matched are assumed to be non-matches.
+\cite{jaro1989} provided no theoretical justification for using this approach, however, \cite{} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
+
+\subsection{Model Estimation}
+In our review, we have assumed one knows $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0),$ however these are typically not known in practice and must be estimated.
+Assume the comparison vectors are independent given the bipartite matching and we assume that the match status of the record pairs are independent.
+It then follows that we can model
+the comparison data using mixture models in the following way:
+\begin{align}
+&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
+&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \\
+& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \\
+\end{align}
+where $p$ is the proportion of record pairs that match.
+Note that $\bm{m}$ and $\bm{u}$ are vectors of the match/non-match parameters.
+One common way to estimate the parameters of the aforementioned model is using the EM algorithm.
 

### 2021-09-25 methods_paper/fabl-paper.tex
outlining next steps.
working on paper.

@@
 One common way to estimate the parameters of the aforementioned model is using the EM algorithm.
+
+\subsection{Pros and Cons of the Fellegi Sunter Method}
+
+\textcolor{red}{Notes below}
+
+``Even if one implores mixture models with a one-to-one constraint, the Fellegi-Sunter decision rule alone may still lead to many-to-many assignments such as the work by Jaro.
+This led to Sadinle (2017)  incorporating this constraint into the model (as in Fortini et al., 2001, 2002; Liseo and Tancredi, 2011; Tancredi and Liseo, 2011; Larsen, 2002, 2005, 2010, 2012; Gutman et al., 2013) rather than forcing it in a post-processing step."
+
+On the other hand, this greatly adds to the computational complexity as explored by Ted in [CITE]. Explain his paper.
+
+In this paper, we bridge the gap between both methods, in order to propose a method that takes into account the one-to-one matching constraint, while scaling to large databases.
+
+Next, review briefly the two approach and the important technical components we utilize for our paper.
 

### 2021-09-26 methods_paper/fabl-paper.tex
adding notes that will be resolved later.
working on paper.

--- 2021-09-24 methods_paper/fabl-paper.tex
@@
 \subsection{Model Estimation}
-In our review, we have assumed one knows $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0),$ however these are typically not known in practice and must be estimated.
-Assume the comparison vectors are independent given the bipartite matching and we assume that the match status of the record pairs are independent.
+In our review, we have assumed one knows $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0),$ however, these are typically not known in practice and must be estimated.
+Assume the comparison vectors are independent given the bipartite matching and assume that the match status of the record pairs are independent.
 It then follows that we can model
@@
 &\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
-&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \\
-& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \\
+&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
+& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
 \end{align}
-where $p$ is the proportion of record pairs that match.
+where $p$ is the proportion of records that match.
 Note that $\bm{m}$ and $\bm{u}$ are vectors of the match/non-match parameters.
-One common way to estimate the parameters of the aforementioned model is using the EM algorithm.
-
-\newpage
-
-%n_1 + j \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1 \\
-
-%describes previous work in record linkage that lays the foundation for our method, Section~\ref{sec:model}
-%3 provides our model and the Gibbs sampler used for posterior inference,
-%and Sections~\ref{sec:efficiency} and~\ref{sec:efficient-posterior} describe strategies used for more efficient and scalable
-%implementation. Then, Section~\ref{sec:simulations} demonstrates the accuracy and speed of
-%out method through two simulation studies, and Section~\ref{sec:case-studies} reveals
-%interesting properties of the method through a case study of El
-%Salvadoran casualty records. Lastly, Section~\ref{sec:discussion} is a discussion that highlights open
-%questions in the record linkage literature and motivates future work.
-
-% We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity.
-Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks.
-To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
-%pronounced ``fable'')
-
-%\textcolor{red}{Talk about related work and weave in the El Salvadoran study.}
-%
-%\newpage
+One common way to estimate the unknown parameters of the aforementioned model is using the EM algorithm.
+
+One criticism of using mixture models as proposed by \cite{jaro1989} is the fact that the decision rule will lead to ``many-to-many assignments (breaking transitive closures).
+One way of resolving this issue to incorporate the constraint directly into the model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, Larsen05, Larsen12, gutman_bayesian_2013, sadinle_bayesian_2017}, which is appealing as one then obtains uncertainty quantification under the Bayesian paradigm, but often has to sacrifice scaling to very large databases.
+On the other hand, one can turn to post-processing methods in practical situations.
+In this paper, we attempt to bridge the best of both approaches.
+Specifically, we propose a model in the spirit of \cite{sadinle_bayesian_2017} and \cite{enamorado_using_2019}.
+We relax the one-to-one matching constraint proposed by  \cite{sadinle_bayesian_2017} for purely computational reasons, and resolve this in a post-processing step.
+In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}.
+Our framework is then embarassingly parallel, available in the \text{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.
 

### 2021-09-26 methods_paper/fabl-paper.tex
update.

@@
 
-\section{Related work}
-\label{sec:related}
-In this section, we review work that is most related to our proposed work.
+%\section{Related work}
+%\label{sec:related}
+%In this section, we review work that is most related to our proposed work.
 For surveys of the record linkage literature, see \citep{christen_2012, Herzog_2007, liseo_2013, Winkler2014, Christophides2019, Jurek2019}.
-
-\paragraph{Fellegi and Sunter Method}
-Most record linkage techniques are derived from the seminal \citep{fs} paper "A Theory for Record Linkage". The defining
-characteristic of their approach was to transform the sets of records,
-which often contain text data that is difficult to model, into sets of
-comparison vectors governed by parameters that can be more easily
-estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
-records respectively, and if the files share \(F\) fields in common upon
-which to base the linkage, the Fellegi and Sunter approach generates an
-\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
-between each pair of records across datasets. We say \(\gamma_{ij}\) is
-the comparison vector for record \(i \in A\) and record \(j \in B\),
-with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
-score on the \(f^{th}\) field. For ease of modeling and computation, we
-restrict these similarity scores to be discrete, ordinal variables, and
-the construction of these is left to the modeler. We adopt the
-convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
-of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
-is common to use use binary 1-2 variables to indicate exact matching,
-and 1-2-3 variables to provide an option for partial matching. For text
-data, we calculate similarity based on Levenstein distance or some other
-text similarity score, and bin these scores to integers for use in the
-model. \citep{jaro1989} provided guidelines for appropriate binning for the
-comparison vectors, but details of implementation are generally context
-specific.
-
-The likelihood in the original Fellegi Sunter model is a mixture model
-through which each record pair is independently classified as a match or
-nonmatch. This independent classification often leads to sets of matches
-that break expectations of transitivity, which is undesirable. Although
-\citep{jaro1995} devised an procedure to reduce a set of conflicting matches to the
-mostly likely set of one-to-one matchings, the abundance of false
-matches can bias the estimation of other parameters of the model, and ultimately leads to poor linkage performance.
-
-\citep{sadinle_bayesian_2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
-Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link.
-This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach.
-However,
-accounting for these dependencies throughout the linkage process is
-computationally burdensome, leaving \texttt{BRL} only suitable for small
-to moderate linkage problems.
+%
+%\paragraph{Fellegi and Sunter Method}
+%Most record linkage techniques are derived from the seminal \citep{fs} paper "A Theory for Record Linkage". The defining
+%characteristic of their approach was to transform the sets of records,
+%which often contain text data that is difficult to model, into sets of
+%comparison vectors governed by parameters that can be more easily
+%estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
+%records respectively, and if the files share \(F\) fields in common upon
+%which to base the linkage, the Fellegi and Sunter approach generates an
+%\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
+%between each pair of records across datasets. We say \(\gamma_{ij}\) is
+%the comparison vector for record \(i \in A\) and record \(j \in B\),
+%with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
+%score on the \(f^{th}\) field. For ease of modeling and computation, we
+%restrict these similarity scores to be discrete, ordinal variables, and
+%the construction of these is left to the modeler. We adopt the
+%convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
+%of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
+%is common to use use binary 1-2 variables to indicate exact matching,
+%and 1-2-3 variables to provide an option for partial matching. For text
+%data, we calculate similarity based on Levenstein distance or some other
+%text similarity score, and bin these scores to integers for use in the
+%model. \citep{jaro1989} provided guidelines for appropriate binning for the
+%comparison vectors, but details of implementation are generally context
+%specific.
+%
+%The likelihood in the original Fellegi Sunter model is a mixture model
+%through which each record pair is independently classified as a match or
+%nonmatch. This independent classification often leads to sets of matches
+%that break expectations of transitivity, which is undesirable. Although
+%\citep{jaro1995} devised an procedure to reduce a set of conflicting matches to the
+%mostly likely set of one-to-one matchings, the abundance of false
+%matches can bias the estimation of other parameters of the model, and ultimately leads to poor linkage performance.
 
@@
 
-\section{A scalable model for Bayesian Record Linkage}
+\section{Proposed Methodology}
 \label{sec:model}
 
-In this section, we propose a statistical model suitable for efficient computation in large linkage tasks. Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
-respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
-in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
-generality, label the files such that \(n_A \geq n_B\). We assume there
-are no duplicates within files, only across. For each record pair under
-consideration, we generate a comparison vector
-\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
-where \(F\) is the number of fields used in the linkage and each
-\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
-indicating the level agreement between the two records on a specified
-field.
+\textcolor{red}{This section needs to be much more clear regarding what is our own contribution and how it differs from the rest of the literature.
+Right now it's confusing to the reader what comes from Sadinle, FS, etc and why these choices are being made.}
+
+%\citep{sadinle_bayesian_2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
+Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link.
+This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach.
+However,
+%accounting for these dependencies throughout the linkage process is
+%computationally burdensome, leaving \texttt{BRL} only suitable for small
+%to moderate linkage problems.
+
+In this section, we propose the fast beta linkage method for bipartite record linkage, extending the work of \cite{sadinle_bayesian_2017}.
+Specifically, the author used the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
+In each iteration of his Gibbs sampler, the author considered each record \(j\in \bm{X}_2\), and then removed from consideration the records \(i\in \bm{X}_1\) that have already been matched.
+Then a potential link is sampled.
+While this leads to a sampler that is more accurate than the classical Fellegi-Sunter approach, accounting for such dependencies throughout the linkage process is computationally burdensome, and this method is only suitable for small databases.
+
+As such, we propose a framework for efficient computation in large linkage tasks. 
+\textcolor{red}{I think my notation is OK from earlier.}
+
+%Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
+%respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
+%in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
+%generality, label the files such that \(n_A \geq n_B\). We assume there
+%are no duplicates within files, only across. For each record pair under
+%consideration, we generate a comparison vector
+%\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
+%where \(F\) is the number of fields used in the linkage and each
+%\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
+%indicating the level agreement between the two records on a specified
+%field.
 
@@
 indicators.
+
+\textcolor{red}{Brian: This does not match what Sadinle has, so this needs to be justified regarding the change or if this is a typo. Could you please check this?}
 

### 2021-09-27 methods_paper/fabl-paper.tex
minor.

@@
 &\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
-\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\};\\
+\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
 & \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
-& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag \\
+& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
 \end{align}
 
-Equation \label{eqn:jaro} ensures that each record of $\bm{X}_2$ only matches with at most one record in $\bm{X}_1$ (and the reverse condition).
+Equation \ref{eqn:jaro} ensures that each record of $\bm{X}_2$ only matches with at most one record in $\bm{X}_1$ (and the reverse condition).
 This is commonly known as the \emph{maximum-weight bipartite matching problem} or a \emph{linear sum assignment problem}.
 The output of this algorithm is a bipartite matching that maximizes the sum of the weights among matched pairs, where the pairs are not matched are assumed to be non-matches.
-\cite{jaro1989} provided no theoretical justification for using this approach, however, \cite{} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
+\cite{jaro1989} provided no theoretical justification for using this approach, however, \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
 

### 2021-09-27 methods_paper/fabl-paper.tex
update.

@@
 While this leads to a sampler that is more accurate than the classical Fellegi-Sunter approach, accounting for such dependencies throughout the linkage process is computationally burdensome, and this method is only suitable for small databases.
-
 As such, we propose a framework for efficient computation in large linkage tasks.
-\textcolor{red}{I think my notation is OK from earlier.}
+
+\subsection{Notation}
+Before providing our proposed model, we first define notation and assumptions.
+
+%\textcolor{red}{I think my notation is OK from earlier.}
 
@@
 
-Following the Fellegi Sunter framework, we define
-\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
-observing agreement level \(l\) in field \(f\) for records \(i\) and
-\(j\) given that the records are a match, and similarly define
-\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We also
-adopt Fellegi and Sunter's conditionally independent fields assumption
-that the level of agreement on one field is independent of the level of
-agreement on another. Though this assumption is often not reasonable
+\textcolor{blue}{Define \(m^{f\ell}:= P(\Gamma_{ij}^f = l |Z_j = i)\) as  the probability of a match having level $\ell$ of disagreement for feature $f.$ Define \(u^{f\ell}:= P(\gamma_{ij}^f = l |Z_j \neq i)\) as the corresponding probability for non-matches.
+Next, we adopt Fellegi and Sunter's conditionally independent fields assumption
+that the level of agreement on one feature is independent of the level of
+agreement on another.\footnote{Though this assumption is often not reasonable
 (for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
@@
 leads to models that perform well in practice; see discussion for
-further remarks. Lastly, we define \(\lambda\) to be the (marginal)
-probability that some record \(j \in B\) has a match in \(A\).
+further remarks.} Denote \(\lambda\) to be the (marginal)
+probability that some record \(j \in \bm{X}_2\) has a match in \(\bm{X}_1\)}
+
+%Following the Fellegi Sunter framework, we define
+%\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
+%observing agreement level \(l\) in field \(f\) for records \(i\) and
+%\(j\) given that the records are a match, and similarly define
+%\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We also
+%adopt Fellegi and Sunter's conditionally independent fields assumption
+%that the level of agreement on one field is independent of the level of
+%agreement on another. Though this assumption is often not reasonable
+%(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
+%but it is common within the record linkage literature and generally
+%leads to models that perform well in practice; see discussion for
+%further remarks. Lastly, we define \(\lambda\) to be the (marginal)
+%probability that some record \(j \in B\) has a match in \(A\).
 

### 2021-09-27 methods_paper/fabl-paper.tex
updates.

@@
 
+\subsection{Generalized Beta Linkage Model}
+In this section, our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
+
+\begin{align}
+&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
+&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
+& \bm{Z} \sim \mathcal{B}, \notag
+\end{align}
+where $\mathcal{B}$ represents a prior on the space of bipartite matchings. $\mathcal{M}(\bm{m})$ and $\mathcal{M}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
+
 \subsection{Notation}
@@
 further remarks.} Denote \(\lambda\) to be the (marginal)
-probability that some record \(j \in \bm{X}_2\) has a match in \(\bm{X}_1\)}
+probability that some record \(j \in \bm{X}_2\) has a match in \(\bm{X}_1\). In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
+$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$}.
 
@@
 
-%\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
-
 $$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$
+
+\textcolor{red}{It seems to me that this should be $P(\bm{Z}, \Phi \mid \bm{\Gamma})$ above and not what is currently written. }
 
@@
 
-Here the reader should the relationship between our proposed model and
-that of Sadinle 2017. In his model, Sadinle constructs a prior
-distribution on the entire \(\mathbf{Z}\) vector, which induces a Gibbs
+\begin{remark}
+There is a clear relationship between our proposed model and that of \cite{sadinle_bayesian_2017}. 
+\cite{sadinle_bayesian_2017} constructs a prior distribution on the entire \(\mathbf{Z}\) vector, 
+which induces a Gibbs
 sampler that strictly enforces one-to-one matching. In particular, this
@@
 records when sampling \(Z_j\), creating a dependency that makes the
-sampler \emph{inherently serial}. We however use independent priors for
-each \(Z_j\), creating a sampler that is \emph{perfectly parallelizable}, allowing for significant computational gains. More
-importantly, since only the agreement pattern of \(Z_j\) is used for
+sampler \emph{inherently serial}. On the other hand, we propose 
+ independent priors for
+each \(Z_j\), creating a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains. 
+More importantly, since only the agreement pattern of \(Z_j\) is used for
 calculations within the Gibbs sampler, and not the particular record
-label, we can conduct this sampling only at the level of the unique
-agreement patterns, offering even more computational savings. In doing
-so however, we thereby weaken the one-to-one requirement from
-\texttt{BRL}; our sampler does ensure that each record in \(B\) can be
-matched to at most one record in \(A\), but allows for the possibility
-that multiple records in \(B\) match to the same record in \(A\). We then resolve these contradictions through a straightfoward post-processing step. We
-explore the ramifications of this distinction with the El Salvador case
-study.
+label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.
+In doing so however, we thereby weaken the one-to-one requirement from
+\texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be
+matched to at most one record in \(\bm{X}_1\), but allows for the possibility
+that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). We resolve contradictions via a post-processing step \textcolor{red}{state what the post-processing step is Brian}. 
+\end{remark}
 

### 2021-09-27 methods_paper/fabl-paper.tex
update.

@@
 
+\textcolor{red}{Are you handling missing values the way that Sadinle does?
+This isn't clear.
+It would he helpful to follow the Sadinle paper regarding structure as closely as possible given that it's very clear.
+I might recommend that you work on some details here regarding pages 12---13 as some important details are skipped over regarding what's random and observed that may not be obvious to the AE/reviewer.}
+
 \[\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})\]
@@
 
+\newpage
+\textcolor{red}{RCS: I stopped here.}
 \hypertarget{posterior-sampling}{%

### 2021-09-27 methods_paper/fabl-paper.tex
update.

@@
 
+\begin{enumerate}
+\item Let's first write out the likelihood, which should be the same as in Sadinle
+\item Then we should write out the Beta Prior, which should deviate and needs its own section in my opinion so the reader understands why it's a deviation
+\item My understanding, is that section 4.1 and section 4.2 should look as they are, so these just need to be explained to the reader.
+The innovation is in the beta prior and making it more efficient.
+\end{enumerate}
+
 %Following the Fellegi Sunter framework, we define
@@
 %overloaded notation in the likelihood and subsequent derivations.
+
+\hypertarget{model-specification}{%
+	\subsection{New Model Specification}
+	\label{model-specification}}
+
+In this section, we provide the specification for the models of $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u}),$ which follows from assuming that the features are conditionally independent of the level of agreement of another.
+The likelihood can be written as follows:
+
+\begin{align}
+P(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}, 
+\end{align}
+where $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denotes the probability of a match have level $\ell$ of disagreement in feature $f$ and 
+$u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represents the same probability for non-matches. 
+In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
+$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
+
+As pointed out in ``We now need to modify this model to accommodate missing comparison criteria since in practice it is rather common to find records with missing fields of information, which lead in turn to missing comparisons for the corresponding record pairs."
+
+We will assume the the missing comparisons occur at missing at random as done in Little and Rubin (2002) and widely assumed in the record linkage literature [CITE].
+
+The likelihood becomes
+
+\begin{align}
+P(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) = 
+\prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{a_{f\ell}(\bm{Z})}u_{fl}^{a_{f\ell}(\bm{Z})}\right], 
+\end{align}
+where
+
+\begin{align}
+&a_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f I(\gamma_{ij}^f = \ell) I(Z_j = i) \\
+&b_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f I(\gamma_{ij}^f = \ell) I(Z_j \neq i) 
+\end{align}
 

### 2021-09-27 methods_paper/fabl-paper.tex
update.
updating paper after meeting with brian.

@@
 
-I think this should be reviewed so that the paper is more clear. Consider ordered record pairs $\bm{X}_1 \times \bm{X}_2.$ 
+Consider ordered record pairs $\bm{X}_1 \times \bm{X}_2.$ 
 Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$ Denote the set of non-matches 
@@
 
-%\section{Related work}
-%\label{sec:related}
-%In this section, we review work that is most related to our proposed work.
-For surveys of the record linkage literature, see \citep{christen_2012, Herzog_2007, liseo_2013, Winkler2014, Christophides2019, Jurek2019}.
-%
-%\paragraph{Fellegi and Sunter Method}
-%Most record linkage techniques are derived from the seminal \citep{fs} paper "A Theory for Record Linkage". The defining
-%characteristic of their approach was to transform the sets of records,
-%which often contain text data that is difficult to model, into sets of
-%comparison vectors governed by parameters that can be more easily
-%estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
-%records respectively, and if the files share \(F\) fields in common upon
-%which to base the linkage, the Fellegi and Sunter approach generates an
-%\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
-%between each pair of records across datasets. We say \(\gamma_{ij}\) is
-%the comparison vector for record \(i \in A\) and record \(j \in B\),
-%with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
-%score on the \(f^{th}\) field. For ease of modeling and computation, we
-%restrict these similarity scores to be discrete, ordinal variables, and
-%the construction of these is left to the modeler. We adopt the
-%convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
-%of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
-%is common to use use binary 1-2 variables to indicate exact matching,
-%and 1-2-3 variables to provide an option for partial matching. For text
-%data, we calculate similarity based on Levenstein distance or some other
-%text similarity score, and bin these scores to integers for use in the
-%model. \citep{jaro1989} provided guidelines for appropriate binning for the
-%comparison vectors, but details of implementation are generally context
-%specific.
-%
-%The likelihood in the original Fellegi Sunter model is a mixture model
-%through which each record pair is independently classified as a match or
-%nonmatch. This independent classification often leads to sets of matches
-%that break expectations of transitivity, which is undesirable. Although
-%\citep{jaro1995} devised an procedure to reduce a set of conflicting matches to the
-%mostly likely set of one-to-one matchings, the abundance of false
-%matches can bias the estimation of other parameters of the model, and ultimately leads to poor linkage performance.
-
-%Outside of the Fellegi-Sunter framework, some researchers have developed
-%methods that model the data directly, not a derived set of comparison
-%vectors. As an early example,  produced \texttt{blink}, which had the advantage of a computational complexity
-%that grew linearly with the data (as opposed to quadratically within
-%Fellegi Sunter), but was plagued with slow mixing times and was unable
-%to incorporate text data. Later, Steorts et al 2020 improved on this
-%method with \texttt{dblink}, which was able to simulate text data within
-%its Gibbs sampler by drawing from an empirical distribution, and
-%remedied its slow mixing through a probabilistic blocking approach.
-%However, this method remains computationally intensive and requires
-%distributed computing to manage even moderate linkage problems in
-%acceptable time.
-These methods perform entity resolution over arbitrary numbers of datasets with arbitrary amounts of duplication and fall outside the Fellegi-Sunter framework, so we do not discuss them further here.
-
 \section{Proposed Methodology}
@@
 
-\textcolor{red}{This section needs to be much more clear regarding what is our own contribution and how it differs from the rest of the literature.
-Right now it's confusing to the reader what comes from Sadinle, FS, etc and why these choices are being made.}
-
-%\citep{sadinle_bayesian_2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
-Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link.
-This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach.
-However,
-%accounting for these dependencies throughout the linkage process is
-%computationally burdensome, leaving \texttt{BRL} only suitable for small
-%to moderate linkage problems.
-
-In this section, we propose the fast beta linkage method for bipartite record linkage, extending the work of \cite{sadinle_bayesian_2017}.
-Specifically, the author used the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
+In this section, we propose the fast beta linkage method (fabl) for bipartite record linkage, extending the work of \cite{sadinle_bayesian_2017}.
+Specifically, we review the general approach of \cite{sadinle_bayesian_2017} and then provide our extensions for scalability.
+
+%Specifically, the author used the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
 In each iteration of his Gibbs sampler, the author considered each record \(j\in \bm{X}_2\), and then removed from consideration the records \(i\in \bm{X}_1\) that have already been matched.
@@
 While this leads to a sampler that is more accurate than the classical Fellegi-Sunter approach, accounting for such dependencies throughout the linkage process is computationally burdensome, and this method is only suitable for small databases.
-As such, we propose a framework for efficient computation in large linkage tasks.
-
-\subsection{Generalized Beta Linkage Model}
-In this section, our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
-
+%As such, we propose a framework for efficient computation in large linkage tasks.
+
+\subsection{Beta Linkage Model}
+In this section, we review the Beta Linkage Model, which can be written as follows: 
+%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
 \begin{align}
@@
 \end{align}
-where $\mathcal{B}$ represents a prior on the space of bipartite matchings. $\mathcal{M}(\bm{m})$ and $\mathcal{M}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-
-\subsection{Notation}
-Before providing our proposed model, we first define notation and assumptions.
+where $\mathcal{B}$ represents a prior on the space of bipartite matchings and $\mathcal{M}(\bm{m}), \mathcal{M}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
+
+\hypertarget{model-specification}{%
+	\subsection{New Model Specification}
+	\label{model-specification}}
+
+In this section, we provide the specification for the models of $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u}),$ which follows from assuming that the features are conditionally independent of the level of agreement of another.
+The likelihood can be written as follows:
+
+\begin{align}
+P(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}, 
+\end{align}
+where $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denotes the probability of a match have level $\ell$ of disagreement in feature $f$ and 
+$u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represents the same probability for non-matches. 
+In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
+$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
+
+Next, we allow for missing values or rather missing comparisons for corresponding record pairs.
+As commonly done in the record linkage literature, we assume that these occur missing at random [CITE: Little and Rubin (2002), Sadinle (2017), Steorts et al (2016)].
+
+The likelihood becomes 
+\begin{align}
+P(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) = 
+\prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{a_{f\ell}(\bm{Z})}u_{fl}^{a_{f\ell}(\bm{Z})}\right], 
+\end{align}
+where 
+\begin{align}
+&a_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell) I(Z_j = i) \\
+&b_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = \ell) I(Z_j \neq i),
+\end{align}
+where $I_{obs}$ denotes if a value is observed or not. Next, for a particular matching configuration $\bm{Z}$, $a_{f\ell}(\bm{Z})$
+and $b_{f\ell}(\bm{Z})$ denote the number of matches and non-matches with observed disagreement level $\ell$ for comparison feature $f.$ Using this parameterization, we specify independent conjugate priors for all features $f$:
+\begin{align}
+&\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
+&\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).
+\end{align}
+
+\hypertarget{fast-beta-prior}{%
+	\subsection{Fast Beta Prior}
+	\label{fast-beta-prior}}
+
+In this section, we deviate from the approach of \cite{}, and propose a more scalable prior on the matching labelings $\bm{Z}.$ 
+While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.
+
+\textcolor{red}{Lay out Brian's prior here.}
+
+\newpage
 
@@
 %overloaded notation in the likelihood and subsequent derivations.
-
-\hypertarget{model-specification}{%
-	\subsection{New Model Specification}
-	\label{model-specification}}
-
-In this section, we provide the specification for the models of $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u}),$ which follows from assuming that the features are conditionally independent of the level of agreement of another.
-The likelihood can be written as follows:
-
-\begin{align}
-P(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}, 
-\end{align}
-where $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denotes the probability of a match have level $\ell$ of disagreement in feature $f$ and 
-$u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represents the same probability for non-matches. 
-In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
-$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
-
-As pointed out in ``We now need to modify this model to accommodate missing comparison criteria since in practice it is rather common to find records with missing fields of information, which lead in turn to missing comparisons for the corresponding record pairs."
-
-We will assume the the missing comparisons occur at missing at random as done in Little and Rubin (2002) and widely assumed in the record linkage literature [CITE].
-
-The likelihood becomes
-
-\begin{align}
-P(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) = 
-\prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{a_{f\ell}(\bm{Z})}u_{fl}^{a_{f\ell}(\bm{Z})}\right], 
-\end{align}
-where
-
-\begin{align}
-&a_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f I(\gamma_{ij}^f = \ell) I(Z_j = i) \\
-&b_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f I(\gamma_{ij}^f = \ell) I(Z_j \neq i) 
-\end{align}
 

### 2021-09-27 methods_paper/fabl-paper.tex
update and notes for brian.

@@
 i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-n_1 + j \quad n_1 + j \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
+n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
 \end{cases}
@@
 \begin{align}
-P(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}, 
+\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}, 
 \end{align}
@@
 \begin{align}
-P(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) = 
+\mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) = 
 \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{a_{f\ell}(\bm{Z})}u_{fl}^{a_{f\ell}(\bm{Z})}\right], 
@@
 
-In this section, we deviate from the approach of \cite{}, and propose a more scalable prior on the matching labelings $\bm{Z}.$ 
+In this section, we deviate from the approach of \cite{sadinle_bayesian_2017}, and propose a more scalable prior on the matching labelings $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2}),$ where recall $Z_j \in \{1,2, 3, n_1, n_1 + j\}.$
+
 While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.
 
-\textcolor{red}{Lay out Brian's prior here.}
-
-\newpage
-
-%\textcolor{red}{I think my notation is OK from earlier.}
-
-%Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
-%respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
-%in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
-%generality, label the files such that \(n_A \geq n_B\). We assume there
-%are no duplicates within files, only across. For each record pair under
-%consideration, we generate a comparison vector
-%\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
-%where \(F\) is the number of fields used in the linkage and each
-%\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
-%indicating the level agreement between the two records on a specified
-%field.
-
-To indicate matching status, we adopt the \emph{linkage structure
-	parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_B})\) from Sadinle 2017,
-defined as \[Z_j=\begin{cases} 
-	i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
-	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
-\end{cases}\] This parameter is vector of length $n_B$, providing more memory efficient storage for the
-linkage information than a \(n_A \times n_B\) sparse matrix of
-indicators.
-
-\textcolor{red}{Brian: This does not match what Sadinle has, so this needs to be justified regarding the change or if this is a typo. Could you please check this?}
-
-\textcolor{blue}{Define \(m^{f\ell}:= P(\Gamma_{ij}^f = l |Z_j = i)\) as  the probability of a match having level $\ell$ of disagreement for feature $f.$ Define \(u^{f\ell}:= P(\gamma_{ij}^f = l |Z_j \neq i)\) as the corresponding probability for non-matches.
-Next, we adopt Fellegi and Sunter's conditionally independent fields assumption
-that the level of agreement on one feature is independent of the level of
-agreement on another.\footnote{Though this assumption is often not reasonable
-(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
-but it is common within the record linkage literature and generally
-leads to models that perform well in practice; see discussion for
-further remarks.} Denote \(\lambda\) to be the (marginal)
-probability that some record \(j \in \bm{X}_2\) has a match in \(\bm{X}_1\). In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
-$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$}.
-
-\begin{enumerate}
-\item Let's first write out the likelihood, which should be the same as in Sadinle
-\item Then we should write out the Beta Prior, which should deviate and needs its own section in my opinion so the reader understands why it's a deviation
-\item My understanding, is that section 4.1 and section 4.2 should look as they are, so these just need to be explained to the reader.
-The innovation is in the beta prior and making it more efficient.
-\end{enumerate}
-
-%Following the Fellegi Sunter framework, we define
-%\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
-%observing agreement level \(l\) in field \(f\) for records \(i\) and
-%\(j\) given that the records are a match, and similarly define
-%\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We also
-%adopt Fellegi and Sunter's conditionally independent fields assumption
-%that the level of agreement on one field is independent of the level of
-%agreement on another. Though this assumption is often not reasonable
-%(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
-%but it is common within the record linkage literature and generally
-%leads to models that perform well in practice; see discussion for
-%further remarks. Lastly, we define \(\lambda\) to be the (marginal)
-%probability that some record \(j \in B\) has a match in \(A\).
-
-%Wherever possible, we reserve superscripts for denoting field and level,
-%while reserving subscripts for record indices. For example,
-%\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
-%distribution governing field \(f\) for matching records, and
-%\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
-%is product of the relevant of the appropriate \(\mathbf{m}\) parameters
-%for record pair \((i,j)\). We hope that these conventions avoid
-%overloaded notation in the likelihood and subsequent derivations.
-
-\hypertarget{model-specification}{%
-	\subsection{Model Specification}
-	\label{model-specification}}
-
-For fields \(f \in \{1, \ldots, F\}\) and levels
-\(l\in \{1, \ldots, L_f\}\) we adopt the following likelihood and prior
-distributions.
-
-$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$
-
-\textcolor{red}{It seems to me that this should be $P(\bm{Z}, \Phi \mid \bm{\Gamma})$ above and not what is currently written. }
-
-\textcolor{red}{Are you handling missing values the way that Sadinle does?
-This isn't clear.
-It would he helpful to follow the Sadinle paper regarding structure as closely as possible given that it's very clear.
-I might recommend that you work on some details here regarding pages 12---13 as some important details are skipped over regarding what's random and observed that may not be obvious to the AE/reviewer.}
-
-\[\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})\]
-\[\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})\]
+\textcolor{red}{Brian: could you please lay out your proposal here more clear regarding the notation above and add as much details as possible regarding your approach.
+More detail is better than less.
+It will be important to contrast this with the prior that Sadinle uses and note that you're dropping the constraint.
+(Make sure to include any proofs/derivations that might not be obvious to a reader given this is a major contribution to the paper).}
+
 \[Z_j | \lambda =
 \begin{cases} 
-	\frac{1}{n_A}\lambda  & z_j \leq n_A; \\
-	1-\lambda &  z_j  = n_A + 1 \\
+	\frac{1}{n_1}\lambda  & z_j \leq n_1; \\
+	1-\lambda &  z_j  = n_1 + j \\
 \end{cases}\]
@@
 belief that nonmatches and matches are equally likely, and a
-\(\lambda \sim \text{Beta}\left(1, \frac{1}{n_A}\right)\) prior
+\(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\) prior
 corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
@@
 \end{remark}
+
+\newpage
+
+%\textcolor{red}{I think my notation is OK from earlier.}
+
+%Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
+%respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
+%in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
+%generality, label the files such that \(n_A \geq n_B\). We assume there
+%are no duplicates within files, only across. For each record pair under
+%consideration, we generate a comparison vector
+%\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
+%where \(F\) is the number of fields used in the linkage and each
+%\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
+%indicating the level agreement between the two records on a specified
+%field.
+
+%To indicate matching status, we adopt the \emph{linkage structure
+%	parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_B})\) from Sadinle 2017,
+%defined as \[Z_j=\begin{cases} 
+%	i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
+%	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
+%\end{cases}\] This parameter is vector of length $n_B$, providing more memory efficient storage for the
+%linkage information than a \(n_A \times n_B\) sparse matrix of
+%indicators.
+%
+%\textcolor{red}{Brian: This does not match what Sadinle has, so this needs to be justified regarding the change or if this is a typo. Could you please check this?} 
+%
+%
+%\textcolor{blue}{Define \(m^{f\ell}:= P(\Gamma_{ij}^f = l |Z_j = i)\) as  the probability of a match having level $\ell$ of disagreement for feature $f.$ Define \(u^{f\ell}:= P(\gamma_{ij}^f = l |Z_j \neq i)\) as the corresponding probability for non-matches.
+Next, we adopt Fellegi and Sunter's conditionally independent fields assumption
+%that the level of agreement on one feature is independent of the level of
+%agreement on another.\footnote{Though this assumption is often not reasonable
+%(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
+%but it is common within the record linkage literature and generally
+%leads to models that perform well in practice; see discussion for
+%further remarks.} Denote \(\lambda\) to be the (marginal)
+%probability that some record \(j \in \bm{X}_2\) has a match in \(\bm{X}_1\). In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
+%$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$}.
+%
+%\begin{enumerate}
+%\item Let's first write out the likelihood, which should be the same as in Sadinle
+%\item Then we should write out the Beta Prior, which should deviate and needs its own section in my opinion so the reader understands why it's a deviation
+%\item My understanding, is that section 4.1 and section 4.2 should look as they are, so these just need to be explained to the reader.
+The innovation is in the beta prior and making it more efficient.
+%\end{enumerate}
+
+%Following the Fellegi Sunter framework, we define
+%\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
+%observing agreement level \(l\) in field \(f\) for records \(i\) and
+%\(j\) given that the records are a match, and similarly define
+%\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We also
+%adopt Fellegi and Sunter's conditionally independent fields assumption
+%that the level of agreement on one field is independent of the level of
+%agreement on another. Though this assumption is often not reasonable
+%(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
+%but it is common within the record linkage literature and generally
+%leads to models that perform well in practice; see discussion for
+%further remarks. Lastly, we define \(\lambda\) to be the (marginal)
+%probability that some record \(j \in B\) has a match in \(A\).
+
+%Wherever possible, we reserve superscripts for denoting field and level,
+%while reserving subscripts for record indices. For example,
+%\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
+%distribution governing field \(f\) for matching records, and
+%\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
+%is product of the relevant of the appropriate \(\mathbf{m}\) parameters
+%for record pair \((i,j)\). We hope that these conventions avoid
+%overloaded notation in the likelihood and subsequent derivations.
+
+%\hypertarget{model-specification}{%
+%	\subsection{Model Specification}
+%	\label{model-specification}}
+%
+%For fields \(f \in \{1, \ldots, F\}\) and levels
+%\(l\in \{1, \ldots, L_f\}\) we adopt the following likelihood and prior
+%distributions.
+
+%$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$
+
+%\textcolor{red}{It seems to me that this should be $P(\bm{Z}, \Phi \mid \bm{\Gamma})$ above and not what is currently written. }
+%
+%\textcolor{red}{Are you handling missing values the way that Sadinle does?
+This isn't clear.
+It would he helpful to follow the Sadinle paper regarding structure as closely as possible given that it's very clear.
+I might recommend that you work on some details here regarding pages 12---13 as some important details are skipped over regarding what's random and observed that may not be obvious to the AE/reviewer.}
+%
+%\[\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})\]
+%\[\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})\]
 

### 2021-09-27 methods_paper/fabl-paper.tex
update.

@@
 
-Consider ordered record pairs $\bm{X}_1 \times \bm{X}_2.$ 
+Consider ordered record pairs $\bm{X}_1 \times \bm{X}_2,$ where the size of the databases are $n_1$ and $n_2,$ respectively. 
 Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$ Denote the set of non-matches 
@@
 $$\gamma_{ij} = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$ The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.
+
+\subsection{Blocking}
+In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, clusters, or ``blocks" to reduces this computational burden.
+Blocking can be deterministic, such as reducing the dimensionality of the comparison space on a highly reliable feature, such a gender.
+Blocking can be probabilistic, such as utilizing machine learning methods, known as hashing, which seek to use probabilistic functions (with low computational cost) to place similar records into the same bin.
+We propose hashing base methods for Bayesian bipartite record linkage, illustrating their ability to improve upon computational scalability.
+We emphasize that depending on the size of the problem, our proposed methodology can be used with or without hashing.
 
@@
 We work with the following factorization of the joint posterior distribution:
+
+\textcolor{red}{Brian: The Gibbs sampler has some minor typos. It should be written in terms of $\bm{Z}, \Phi, \lambda \mid \gamma_{obs}.$ Also, just make the rest of the notation match please:)}
 

### 2021-09-29 methods_paper/fabl-paper.tex
update paper.

@@
 \label{sec:efficiency}
-Broadly speaking, we increase our computational efficiency by
-recognizing that record pairs contribute to posterior calculations only
-through the agreement pattern of the \(\gamma_{ij}\) vector. Let
+In this section, we propose efficient and scalable proposals using hashing based methods that are similar in spirit to \citep{enamorado2019}, a fast and scalable implementation of the Fellegi-Sunter model.
+ Specifically, one way we can improve our computational efficiency is by recognizing that record pairs contribute to posterior calculations only
+through the agreement pattern of the \(\gamma_{ij}\) vector. To make this more precise, let 
 \(\mathcal{H}\) be the set of unique agreement patterns in the data, let
 \(P = |\mathcal{H}|\) denote the total number of unique agreement
-patterns.  Note that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
-that this bound does not scale with \(n_A\) or \(n_B\). Prior to
+patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
+that this bound \textcolor{red}{does not depend on} \(n_1\) or \(n_2\). Prior to
 processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
-and enumerate them \(h_1, \ldots, h_P\), and map record pairs to this hashed value. When the \((i,j)\) pair
+and enumerate them as follows: \(h_1, \ldots, h_P\), which are called \emph{hashed values.}
+Next, we map record pairs to their corresponding hashed value. 
+More specifically, when the record pair \((i,j)\) 
 exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
-Wherever possible, we conduct calculations over these \(P\) agreement
-patterns rather than the \(n_A \times n_B\) record pairs.
-Some of these techniques are similar to those used by \citep{enamorado2019} to produce \texttt{fastlink}, a fast a scalable implementation of the Fellegi-Sunter model, and others have been created here for our particular context.
+Whenever possible, we conduct calculations over these \(P\) agreement
+patterns, instead of the typical  \(n_1 \times n_2\) record pairs.
+
+%Some of these techniques are similar to those used by \citep{enamorado2019} to produce \texttt{fastlink}, a fast a scalable implementation of the Fellegi-Sunter model, and others have been created here for our particular context.
 
@@
 
-We first hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019} accomplished this efficiently through the hashing function
+In this section, we describe the hashing function we propose.
+
+First, we hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019} accomplished this efficiently through the hashing function
 $$\tilde{\gamma}_{ij} = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$
-This function maps each agreement pattern to a unique integer, allowing us to store a scalar quanity instead of an entire vector for each record pair.
-For computational ease, we then map these integers to sequential integers from 1 to P.
+This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
+For computational ease, we then map these integers to sequential integers from $1, \ldots, P.$
 
@@
 comparison as a vector of length \(F\), with each component
-\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). To ease
-computations, we instead use a one hot encoding of the comparison vector. For example, if
+\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). 
+To ease our computational burden, we instead use a \emph{one hot encoding} of the comparison vector. For example, if
 \(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
@@
 
-In the classic Fellegi Sunter framework, \(\Gamma\) is a
-\(n_A n_B \times F\) matrix, with each row providing the comparison
+In the classic Fellegi Sunter framework, $\bm{\Gamma}$ is a
+\(n_1 n_2 \times F\) matrix, with each row providing the comparison
 vector for a different \((i,j)\) pair. In contrast, we do not store
@@
 corresponding to the agreement pattern of the \((i, j)\) pair. We store
-this information in a nested list \(\tilde{\Gamma}\) where the
+this information in a nested list $\tilde{\bm{\Gamma}}$ where the
 \(p^{th}\) component of the \(j^{th}\) list contains a vector of records
-in \(A\) that share agreement pattern \(p\) with record \(j \in B\). For
+in \(\bm{X}_1\) that share agreement pattern \(p\) with record \(j \in \bm{X}_2\). For
 each \(p\), we also calculate
-\(H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\),
+\(H_p = \sum_{i=1}^{n_1}\sum_{j=1}^{n_2} \mathbf{1}_{(i,j) \in h_p}\),
 the total instances of agreement pattern \(p\) throughout the data, and
 also for each \(j\), we calculate
-\(H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}\) the
+\(H_{p_j} = \sum_{i=1}^{n_1} \mathbf{1}_{{(i,j) \in h_p}}\) the
 instances of agreement pattern \(p\) among the comparison vectors
-between record \(j \in B\) and each of the \(n_A\) records in \(A\).
+between record \(j \in \bm{X}_2\) and each of the \(n_1\) records in \(\bm{X}_1\).
 

### 2021-09-29 methods_paper/fabl-paper.tex
update.

@@
 The hashing procedure described above considerably reduces the memory
-needed to store the comparison information: instead of storing
-\(n_A \times n_B\) comparison vectors, which are relatively long under
+needed to store the comparison information. That is, instead of storing
+\(n_1 \times n_1\) comparison vectors, which are relatively long under
 either the Fellegi Sunter or our modified framework, we only store the
-\(P\) unique vectors, and then \(n_A \times n_B\) scalar quantities
+\(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
 relating record pairs to those vectors. However, even storing these
-\(n_A \times n_B\) scalar labels can become burdensome with large data.
+\(n_1 \times n_2\) scalar labels can become burdensome with large data.
 Worse, the overwhelming majority of these labels relate to record pairs
@@
 
-To address this, we propose a new strategy called \emph{storage	efficient indexing} (SEI). In standard indexing, the modeler decides a
-certain criteria that they expect all true matching pairs to satisfy,
-and a priori label any record pairs that do not meet that criteria as
-non-matching. For example, one might only consider pairs with a certain
+To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI).
+In standard indexing, one decides a certain criteria that they expect all true matching pairs to satisfy.
+In addition, one decides
+a priori to label any record pairs that do not meet that criteria as
+non-matches. For example, one might only consider pairs with a certain
 similarity score on a field deemed to be important (like first name), or
@@
 generally chosen to be be quite loose, establishing these criteria
-requires knowledge of the problem and invites room for human error. We
-propose a method of reducing the comparison space (and reducing storage
-requirements) without these drawbacks. Note that all records of the same
+requires knowledge of the problem and invites room for human error. 
+To improve upon this, we propose to reduce the comparison space and reduce the storage requirements while avoiding the drawbacks of standard indexing.
+
+Observe that all records of the same
 agreement pattern have the same probability when sampling \(Z_j\).
@@
 statistics \(H_p\), and \(H_{p_j}\). Rather than storing
-\(n_A \times n_B\) record labels, SEI allows us to store at most
-\(n_B \times P \times R\) labels, regardless of how large \(n_A\) is.
-
-Lastly, for large data, we can partition the two datasets \(A\) and
-\(B\) into smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more
+\(n_1 \times n_2\) record labels, SEI allows us to store at most
+\(n_2 \times P \times R\) labels, regardless of how large \(n_1\) might be.
+
+Lastly, for large data, we can partition the two datasets \(\bm{X}_1\) and
+\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2m}\}\) for more
 manageable computations. On a single machine, we can read-in data
@@
 chunk of data. With multiple cores or multiple machines, this can be
-done in parallel. Thus the combination of hashing, SEI, and partitioning
+done in parallel. Thus, the combination of hashing, SEI, and partitioning
 allows us to conduct linkage tasks over much larger datasets.
+
+\textcolor{red}{Brian: how large can we go and how fast can we go?}
 

### 2021-09-29 Prelim/prelim.tex
prelim slides on latex

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\begin{document}
	\title{Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger, Jerome Reiter, Rebecca Steorts}
	\institute{Duke University}
	\date{\today}

%	\AtBeginSection[]{
%		\begin{frame}
%			\vfill
%			\centering
%			\begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%				\usebeamerfont{title}\insertsectionhead\par%
%			\end{beamercolorbox}
%			\vfill
%		\end{frame}
%	}

\AtBeginSection[]
{
	\begin{frame}
		\frametitle{Table of Contents}
		\tableofcontents[currentsection]
	\end{frame}
}

	\begin{frame}
		\titlepage
	\end{frame}

%	\begin{frame}
%	\frametitle{Outline}
%	\tableofcontents
%	\end{frame}

\section{Introduction to Record Linkage}

	\begin{frame}{What is Record Linkage?}
	\begin{enumerate}
		\item Record linkage is the task of identifying duplicate records over noisy datasets.

		\item Easy with unique identifiers, difficult when faced with errors

		\item Far ranging applications in business, public health, and human rights
	\end{enumerate}
	\end{frame}

\begin{frame}{Record Linkage in Practice}
	\begin{columns}
		\begin{column}{0.48\textwidth}
			\includegraphics<1->[width = \textwidth, height = 1.2\textwidth ]{ted_article2.png}

		\end{column}
		\begin{column}{0.48\textwidth}
			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{dnc_header.png}
		\end{column}
	\end{columns}
\end{frame}

	\begin{frame}{Linkage for Downstream Analysis}
	\includegraphics<1>[width = \textwidth, height = .7\textwidth ]{datasets.png}
	\includegraphics<2>[width = \textwidth, height = .7\textwidth ]{datasets_linked.png}

\end{frame}

	\begin{frame}{Linkage through Comparison Vectors}
		\includegraphics[width = \textwidth, height = .7\textwidth ]{fs_graphic.png}
	\end{frame}

\begin{frame}{Fellegi and Sunter (1969)}
	\begin{columns}
		\begin{column}{0.6\textwidth}
			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide1.png}
			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide2.png}

		\end{column}
		\begin{column}{0.38\textwidth}
			\begin{itemize}
				\item<3-> scalable to large datasets (\texttt{fastlink}, Enamorado et al 2017)
				\item<4-> no transitive closure, requires post-processing
				\pause 
				\item<5-> overmatches, leading to inaccurate parameter estimation
			\end{itemize}

		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Sadinle (2017) - Beta Record Linkage}
	\begin{columns}
		\begin{column}{0.6\textwidth}
			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide1.png}
			\includegraphics<2>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide3.png}
			\includegraphics<3>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide4.png}
			\includegraphics<4->[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide5.png}

		\end{column}
		\begin{column}{0.38\textwidth}
			\begin{itemize}
				\item<1-> Beta Record Linkage (\texttt{BRL})
				\pause
				\item<5-> strictly enforces one-to-one matching, no post-processing
				\pause
				\item<6-> high accuracy for linkage and other parameters
				\pause 
				\item<7-> inherently serial, not scalable to large linkage tasks
			\end{itemize}

		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Our Contribution - Fast Beta Linkage}
	\begin{columns}
		\begin{column}{0.6\textwidth}
			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide1.png}
			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide6.png}

		\end{column}
		\begin{column}{0.38\textwidth}
			\begin{itemize}
				\item<3-> simple mathematical change, large computational gains
				\item<4-> minimal loss of accuracy for linkage and other parameters, minimal post-processing
			\end{itemize}

		\end{column}
	\end{columns}
\end{frame}

\section{Fast Beta Linkage}
\begin{frame}{Notation}
\begin{itemize}
	\item File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.

	\item $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors where $\gamma_{ij}^f \in \{1, \ldots, L_f\}$

	\item $Z_j = \begin{cases} 
	i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
	n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
	\end{cases}$

	\item $m_{fl} = P\left(\gamma_{ij}^f = l |Z_j = i\right)$

	\item $u_{fl} = P\left(\gamma_{ij}^f = l |Z_j \neq i\right)$

	\item $\lambda = P(Z_j \leq n_A)$
\end{itemize}
\end{frame}

\begin{frame}{Fast Beta Linkage (fabl)}
	$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$

	$$\mathbf{m_{f}} \sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{fL_f})$$
	$$\mathbf{u_{f}} \sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{fL_f})$$
	$$Z_j | \lambda =
	\begin{cases} 
	\frac{1}{n_A}\lambda  & z_j \leq n_A; \\
	1-\lambda &  z_j  = n_A + 1 \\
	\end{cases}$$
	$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda})$$

	Model specification allows for \textbf{parallel/distributed} computing, \textbf{hashing} of comparison vectors, and \textbf{storage efficient indexing (SEI)}
\end{frame}

\begin{frame}{Hashing}
	\begin{itemize}
		\item Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $h(i, j)=p.$

		\item Reduce data to sufficient statistics
		\begin{itemize}
		\item $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
		\item $H_{p_j} = ||r_{p_j}||$
		\item $H_p = \sum_{j} H_{p_j}$
		\end{itemize}

		\item Run Gibbs sampler at level of agreement patterns, not record pairs
		\begin{itemize}
		\item Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
		\item Use number of matches for each pattern to update $m$ and $u$
		\item Back fill record labels at the end through $r_{p_j}$
	\end{itemize}

		\item Reduces computational complexity from $O(n_A \times n_B \times F)$ to $O(P \times n_B \times F)$
	\end{itemize}
\end{frame}

\begin{frame}{Managing Large Data}
	\begin{itemize}
		\item \textbf{Distributed Computing} - Partition data in to chunks $\{A_I\}$ and $\{B_J\}$. Compare records, hash results, compute summary statistics in parallel, and synthesize results.

		\item \textbf{Storage Efficient Indexing (SEI)} - Store at most small number $R$ many record labels in each $r_{p_j}$, remove highly unlikely record labels from memory. Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$.

		\item Hashing plus SEI can reduce memory requirements by >99%
		\begin{itemize}
			\item Simulation of $20,000 \times 20,000$ linkage task with 4 fields. Naive approach requires 6.4GB of storage for all-to-all comparisons, hashing and SEI requires 90MB.
		\end{itemize}
	\end{itemize}
\end{frame}

\section{Results}

\begin{frame}{Three Simulation Studies}
	\begin{itemize}
		\item We compare \texttt{fabl} against \texttt{BRL} in three simulation studies
		\begin{itemize}
			\item Measure precision and recall on 100 simulated datasets and varying levels of error and duplication across files
			\item Measure speed when both $n_A$ and $n_B$ are increasing
			\item Measure speed when $n_A$ is increasing and $n_B = 500$ is fixed. 
		\end{itemize}
	\end{itemize}
\end{frame}

	\begin{frame}{Accuracy Simulation}
	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/sadinle_sim_plot.png}
\end{frame}

	\begin{frame}{Speed Simulation 1}
	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/sadinle_speed_plot2.png}
\end{frame}

	\begin{frame}{Speed Simulation 2}
	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/speed_plot_fixed_nB.png}
\end{frame}

\begin{frame}{Extensions and Directions}
	\begin{itemize}
	\item Linkage in the face of duplicates within datasets

	\item Models that allow reliability of information to differ by subgroup in the data

	\item Linkage over blocked data (allows for much larger linkage tasks)
	\end{itemize}
\end{frame}

\end{document}

### 2021-09-30 Prelim/prelim.tex
new speed graphics for el salvador

@@
 	\begin{frame}{What is Record Linkage?}
-	\begin{enumerate}
+	\begin{itemize}
 		\item Record linkage is the task of identifying duplicate records over noisy datasets.
@@
 		\item Far ranging applications in business, public health, and human rights
-	\end{enumerate}
+
+		\item \textbf{Bipartite matching} is the specific goal of matching one record in one dataset to most one match in another dataset
+	\end{itemize}
 	\end{frame}
@@
 		\begin{column}{0.48\textwidth}
-			\includegraphics<1->[width = \textwidth, height = 1.2\textwidth ]{ted_article2.png}
+			\includegraphics<1->[width = \textwidth, height = .9\textwidth ]{ted_article2.png}
 
@@
 		\begin{column}{0.48\textwidth}
-			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{dnc_header.png}
+			\includegraphics<2->[width = \textwidth, height = .9\textwidth ]{dnc_header.png}
 		\end{column}
@@
 	\begin{frame}{Linkage through Comparison Vectors}
-		\includegraphics[width = \textwidth, height = .7\textwidth ]{fs_graphic.png}
+		\includegraphics[width = \textwidth, height = .6\textwidth ]{fs_graphic.png}
 	\end{frame}
@@
 \begin{itemize}
-	\item File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$.
+	\item File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$, with $n_A \geq n_B$.
 We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.
@@
 
-		\item Hashing plus SEI can reduce memory requirements by >99%
+		\item Hashing plus SEI can reduce memory requirements by $>99$.
 		\begin{itemize}
@@
 
-\section{Results}
+\section{Simulation Studies}
 
@@
 
+\section{El Salvador Homicide Case Study}
+
+\begin{frame}{El Salvador Case Study}
+	\begin{itemize}
+		\item El Salvadoran Civil War (1980-1991)
+		\item Lists of casualties collected by multiple organizations
+		\begin{itemize}
+			\item Salvadoran Human Rights Commissions (CDHES), $n_A = 4420$
+			\item El Rescate - Tutela Regal (ERTL), $n_B = 1323$
+			\item Features include first name, last name, date and place of death
+		\end{itemize}
+		\item We aim to find duplicate records across files
+		\begin{itemize}
+			\item Particularly difficult because families are often killed together, and some children share names with parents
+		\end{itemize}
+
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Run Time}
+		\begin{columns}
+		\begin{column}{0.48\textwidth}
+			\centering
+			\includegraphics<1->[width = .6\textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/time_table_big_P2.png}
+
+		\end{column}
+		\begin{column}{0.48\textwidth}
+			\centering
+			\includegraphics<2>[width = .6\textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/time_table_small_P2.png}	
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Posterior Inference}
+	\includegraphics[width = \textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/m_posterior_smallP.png}
+\end{frame}
+
+\begin{frame}{Posterior Inference}
+	\includegraphics[width = \textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/DID_distribution_smallP_bayes.png}
+\end{frame}
+
+\begin{frame}{Violations of One-to-One Matching}
+	\centering
+	\includegraphics[width = .9\textwidth, height = .4\textwidth ]{../notes/figures/el_salvador/rosa_records.png}
+\end{frame}
+
+\begin{frame}{Violations of One-to-One Matching}
+	\includegraphics[width = \textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/bad_mixing.png}
+\end{frame}
+
+\section{Conclusion}
+
 \begin{frame}{Extensions and Directions}

### 2021-10-01 methods_paper/fabl-paper.tex
update.
BK edits on 4.3, Fast Beta Prior

@@
 
-\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
-prior for \(Z_j\) has equal probability of matching to all records
-\(i\in A\), and non-matching probability governed by \(\lambda\).
-Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds to a prior
-belief that nonmatches and matches are equally likely, and a
-\(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\) prior
-corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
-
-\begin{remark}
+\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \]
+
+{
+	\color{blue}
+	Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
+$\lambda$ itself is given a prior distribution to reflect prior beliefs about the overall rate of matchingness in the data.
+One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
+}
+
+%The prior for \(Z_j\) has equal probability of matching to all records
+%\(i\in A\), and non-matching probability governed by \(\lambda\).
+
+{
+\color{blue}
 There is a clear relationship between our proposed model and that of \cite{sadinle_bayesian_2017}. 
-\cite{sadinle_bayesian_2017} constructs a prior distribution on the entire \(\mathbf{Z}\) vector, 
-which induces a Gibbs
-sampler that strictly enforces one-to-one matching. In particular, this
-sampler removes previously matches records from the set of candidate
-records when sampling \(Z_j\), creating a dependency that makes the
-sampler \emph{inherently serial}. On the other hand, we propose 
- independent priors for
-each \(Z_j\), creating a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains. 
+\cite{sadinle_bayesian_2017} constructs a prior distribution on the entire \(\mathbf{Z}\) vector, called the \emph{beta distribution for bipartite matching}, given by
+$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
+$$
+where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior induces a Gibbs sampler that strictly enforces one-to-one matching,  previously matched records from the set of candidate
+records when sampling \(Z_j\). This creates a dependency that makes the
+sampler \emph{inherently serial}.
+
+On the other hand, we propose independent priors for each \(Z_j\), creating a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains.
 More importantly, since only the agreement pattern of \(Z_j\) is used for
@@
 matched to at most one record in \(\bm{X}_1\), but allows for the possibility
-that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). We resolve contradictions via a post-processing step \textcolor{red}{state what the post-processing step is Brian}. 
-\end{remark}
+that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). \textcolor{blue}{We resolve contradictions when calcuating the Bayes estimate, which is explained in Section 4.5}
+
+NOTE TO BEKA: I think it will be best to do the "post-processing step" along with the Bayes Estimate.
+Its equivalent to minimizing the posterior loss subject to the constraint that $Z_j \neq Z_{j'}, \forall j \neq j'$.
+}
 
@@
 \newpage
-\textcolor{red}{RCS: I stopped here.}
+
 \hypertarget{posterior-sampling}{%
@@
 
+\textcolor{red}{RCS: I stopped here.}
+
 \section{Simulation Studies}

### 2021-10-01 methods_paper/fabl-paper.tex
BK edits on abstract

@@
 Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.
+
+\vspace{1ex}
+{
+\color{blue}
+In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
+Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
+This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}.
+Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
+
+\vspace{1ex}
+QUESTION: Does this abstract necessarily need to read like an outline?
+}
 \end{abstract}

### 2021-10-04 methods_paper/fabl-paper.tex
fix figure and ask bk some questions.

@@
 \label{sec:simulations}
-
-\hypertarget{accuracy}{%
-	\subsection{Accuracy}\label{accuracy}}
+In this section, we consider a simulation study, nearly identical to that in \cite{sadinle2017}.
+\textcolor{red}{Brian: can you explain how it's the same versus different than what it is given in Sadinle as this would be the most helpful.
+Do you utilize Mauricio's exact files and approach and then do something else?
+If so, I would state this for clarity.}
+
+\textcolor{red}{If you look in Sadinle (2017), did you use Table 2 or something else?
+Can you construct something similar to Figure 3 and Figure 4 as it's likely a reviewer will look at this paper and ask why you didn't do this exact comparison.}
 
@@
 
+\hypertarget{accuracy}{%
+	\subsection{Accuracy}\label{accuracy}}
+
 In cases when there are only one or two errors in matching records, and in cases with low to moderate duplication across records, we see that \texttt{fabl} provides approximately equivalent accuracy as \texttt{BRL}.
@@
 
-\begin{figure}[ht]
-
-	{\centering
-
+\begin{figure}[htbp]
+\begin{center}
 \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot} 
-	}
-
 \caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
+For each level of overlap and each level of error, we have 100 paired sets of 500 records.}
+\label{fig:sadinle_simulation}
+\end{center}
+\end{figure}
+
+%
+%\begin{figure}[ht]
+%	
+%	{\centering 
+%		
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot} 
+%	}
+%	
+%	\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
 For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:sadinle_simulation}
-
-\end{figure}
+%	
+%\end{figure}
 

### 2021-10-04 methods_paper/fabl-paper.tex
fix figure and ask bk some questions.
update.

@@
 
-\textcolor{red}{If you look in Sadinle (2017), did you use Table 2 or something else?
+\textcolor{red}{Brian: If you look in Sadinle (2017), did you use Table 2 or something else?
 Can you construct something similar to Figure 3 and Figure 4 as it's likely a reviewer will look at this paper and ask why you didn't do this exact comparison.}
@@
 \hypertarget{accuracy}{%
-	\subsection{Accuracy}\label{accuracy}}
-
+	\subsection{Precision, Recall, and F-measure}\label{accuracy}}
+
+In this section, we compare \texttt{fabl} to 	\texttt{BRL} in terms of precision, recall, and f-measure, which are commonly defined record linkage evaluation metrics that are defined in \citep{christen_2012}.
+%
 In cases when there are only one or two errors in matching records, and in cases with low to moderate duplication across records, we see that \texttt{fabl} provides approximately equivalent accuracy as \texttt{BRL}.
@@
 In these situations, \texttt{BRL} is removing large numbers of records from consideration throughout the Gibbs Sampler, making its implementation most different from \texttt{fabl}.
-We note however that such extreme linkage tasks, which such high amounts of errors and overlap, are extremely rare in practice.
+However,  such extreme linkage tasks, which such high amounts of errors and overlap, are extremely rare in practice.
 
@@
 	\subsection{Speed}\label{speed}}
+
+\textcolor{red}{Brian: I have some general questions here.
+Are you using blocking for BRL to scale?
+Could you use the hashing proposal for BRL.
+It's important that the scale comparison be as fair as possible or there will be some pushback here.
+My suggestion around all this is to look at the computational complexity of both methods as it's hard to attack this.}
 
@@
 \hypertarget{scale}{%
-	\subsection{Scale}\label{scale}}
+	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
 
@@
 
- In one final simulation, we demonstrate how partitioning and hashing the data through \texttt{fabl} method allows us undertake significantly larger linkage tasks We concatenate two sets of 40 simulated datasets from the Sadinle simulation to create two larger files, each of 20,000 records.
+In this section, we investigate scaling to larger data sets using our proposed methodology.
+
+Specifically, we demonstrate how partitioning and hashing the data through \texttt{fabl} method allows us undertake significantly larger linkage tasks.
+We concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{} to create two larger files, each of 20,000 records (40,000 records total).
 Working on one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
@@
 For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
-We then compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
-The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method.
-Excecuted sequentially, these comparisons and the Gibbs sampler take about one hour to run; using distributed computing, it could be completed much faster.
-This simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
+We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
+The resulting data object is now only 90 MB, about 1\% the size of the object required under the standard method.
+Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run; using distributed computing.
+However, this could be sped up.
+
+Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
 The reader will note that this slightly worse performance than witnessed in the smaller simulation studies; this is expected because it is naturally more difficult to link more files with the same amount of information.
 With more linkage fields \texttt{fabl} maintains the high accuracy seen above.
+\textcolor{red}{Brian: you should look to see if you fix these in the post-processing step as this would be interesting/important.}
+
+ \textcolor{red}{How long does it take on the same machine to run BRL. Can you run this? This can close this section out!}
 

### 2021-10-04 methods_paper/fabl-paper.tex
update.

@@
 
-\section{Case Studies}
+\section{Application to Civilian Casualties from the El Salvadoran Civil War}
 \label{sex:case-studies}
 
-We now demonstrate the power of our method through a case study of documented identifiable deaths (DID) from the El Salvadoran Civil War.
+In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of documented identifiable deaths (DID).
 Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-Secondly, the case study reveals the impact of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
+This case study also considers the ramifications of
+independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
 
@@
 organization, as it is likely that the same individuals are recorded in
-multiple casualty lists. To obtain a more accurate estimate of the
-casualties then, we follow Sadinle's 2017 paper and link records of El
-Salvadoran civilian casualties from two sources: El Rescate - Tutela
+multiple casualty lists. Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize 
+two different sources, namely,  El Rescate - Tutela
 Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
@@
 later digitized.
+
+\textcolor{red}{Beka: These data sets need more information regarding how the data sets were collected as this is quite important. See Sadinle (2017) and references for info). }
 
@@
 
-We only utilized records with nonmissing entries for given and last
-name, results in \(n_A = 4420\) files in CHDES and \(n_B = 1323\) files
-in ERTL. The names were standardized to account for common misspellings
-in the Spanish language, and then compared using a modified Levenstein
+In this paper, we utilize records that have non-missing entries for given and last name, which results 
+in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files
+in ERTL. We standardized names to account for common misspellings
+in the Spanish language. The string distance used is a modified Levenstein
 distance to account for the fact that second names are often omitted.
@@
 municipality; however, since department was missing in 95\% of records
-in CHDES and 80\% of records in ERTL, we excluded it from our linkage
-process. Thus we conduct linkage using given name, last name,
+in CHDES and 80\% of records in ERTL, this was excluded from our analyses. 
+Thus, we conduct record linkage using given name, last name,
 municipality, and day, month, and year of death. We again use flat
-priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters.
-
-\begin{figure}[t]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP}
-
-	}
-
-	\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
-\end{figure}
-
-To mirror the original implementation, we constructed the comparison
-vectors using 4 levels of agreement for each field, according to the
-thresholds provided in Figure XX. This took 422 seconds for our proposed
-method and 240 for \texttt{BRL}. However, we observed that posterior
+priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters. To our knowledge, we have followed the same guidelines/settings in \cite{sadinle_bayesian_2017}.
+
+To mirror the original implementation of \cite{sadinle_bayesian_2017}, we construct the comparison
+vectors using four levels of agreement for each field, according to the
+thresholds provided in \textcolor{red}{Figure XX}. This took 422 seconds for our proposed
+method, \texttt{fabl}, and 240 for \texttt{BRL}. However, we observed that posterior
 distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
@@
 agreement levels creates unnecessary distinctions in the data and makes
-the comparison vectors less interpretable. Therefore we re-ran our
-analysis with fewer agreement levels for each field (see Figure XX), and
+the comparison vectors less interpretable.
+
+Therefore we re-ran our analysis with fewer agreement levels for each field \textcolor{red}{(see Figure XX)}, and
 obtained analogous results. With 216 possible agreement patterns, 159
-were realized in the data, and our proposed method became much faster,
+were realized in the data, and our proposed method, \texttt{fabl} became was faster,
 finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
@@
 gains can be made by simplifying the construction of the comparison
-vectors. We also note that estimates of the \(\mathbf{m}\) and
-\(\mathbf{u}\) parameters under each method are very similar, as shown in
+vectors. The estimates of the \(\mathbf{m}\) parameters under each method are very similar, as shown in
 Figure~\ref{fig:m-and-u}.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
+\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
+\label{fig:m-and-u}
+\end{center}
+\end{figure}
 

### 2021-10-04 methods_paper/fabl-paper.tex
update.

@@
 
-\paragraph{Violations of One-to-One Matching} Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
-duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
+\paragraph{Violations of One-to-One Matching}
+
+In this section, we investigate violations of one-to-one matching of our method as compared to \texttt{BRL}.
+
+Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
+duplicates found in both databases for both \texttt{fabl} and \texttt{BRL}. We
 see that \texttt{fabl} consistently overmatches within each Gibbs
@@
 
-\begin{figure}[t]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
-
-	}
-
+\begin{figure}[htbp]
+\begin{center}
+ \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes} 
 \caption{Posterior distribution of overlap across the two files.
+The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}
+\label{fig:overlap-plot}
+\end{center}
+\end{figure}
+
+%\begin{figure}[t]
+%	
+%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes} 
+%		
+%	}
+%	
+%	\caption{Posterior distribution of overlap across the two files.
 The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:overlap-plot}
-\end{figure}
-
-Overall, both methods presented similar results. \texttt{fabl} yielded
-an initial Bayes estimate of 195 matches found across files, and after
-resolving matches that violated one-to-one requirements, yielded 180
-matches across files. This is acceptably close to the 181 matches found
+%\end{figure}
+
+Overall, both methods present similar results. \texttt{fabl} yields
+an initial Bayes estimate of total 195 DID. After
+resolving matches that violated one-to-one requirements,  we estimate 180 DID. 
+\textcolor{red}{Can you report the posterior standard error for each one as we want uncertainty quantification.}
+
+This is very close to the 181 DID found
 by \texttt{BRL}. The main reason for this discrepancy is the difference
-in how each model handles situations in which one file in \(B\) has
-multiple plausible matches in \(A\).
-
-\begin{figure}[t]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records}
-
-	}
-
-	\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
-\end{figure}
-
-\begin{figure}[t]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing}
-
-	}
-
-	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
-\end{figure}
-
-The records shown in Figure~\ref{fig:rosa-maria} provide one such instance. Note these
+in how each model handles situations in which one file in \(\bm{X}_1\) has
+multiple plausible matches in \(\bm{X}_2\).
+
+%
+%\begin{figure}[t]
+%	
+%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records} 
+%		
+%	}
+%	
+%	\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
+%\end{figure}
+
+Figure~\ref{fig:rosa-maria} illustrates an example. Note these
 records present a near violation of our assumption that there are no
@@
 826 in ERTL correspond to different individuals, but their records are
-nearly identical. Additionally, using the modified Levenstein distance
-from Sadinle 2017, the comparison vectors \(\gamma_{2776, 825}\) and
+nearly identical. In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and
 \(\gamma_{2776, 826}\) are exactly identical.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records} 
+\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
+\end{center}
+\end{figure}
 
@@
 throughout the Gibbs sampler, and demonstrates how each method handles
-this situation. Under \texttt{fabl}, both records in \(B\) match to the
-same record in \(A\) throughout the Gibbs process, creating consistent
+this situation. Under \texttt{fabl}, both records in \(\bm{X}_2\) match to the
+same record in \(\bm{X}_1\) throughout the Gibbs process, creating consistent
 violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
-creates one matching configuration stays there for a while. However, if
+creates one matching configuration and stays there for a while. However, if
 one pair ``unmatches,'' then the other record has a chance to latch on.
 Then, the Gibbs process is stuck with that matching status for a while,
-resulting in a Gibbs process with poor mixing. Additionally,
+resulting in a Gibbs process with poor mixing. In addition,
 \texttt{fabl} allows the modeler to inspect records with multiple
@@
 be identified through the Bayes estimate.
+In short, we have pointed out why one may not wish to imbed this constraint in a model, and proposed a simple alternative that performs about the same as \texttt{BRL}, while scaling to much larger data sets.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
+\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
+\end{center}
+\end{figure}
+
+%\begin{figure}[t]
+%	
+%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
+%		
+%	}
+%	
+%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
+%\end{figure}
 

### 2021-10-06 methods_paper/fabl-paper.tex
update.
added demonstration, addressed Bekas points

@@
 \begin{align}
-\mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) = 
+\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}^{obs}) = 
 \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{a_{f\ell}(\bm{Z})}u_{fl}^{a_{f\ell}(\bm{Z})}\right], 
@@
 \begin{align*}
-p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda) \\
-&\propto \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right] \\
+p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
+&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)} \\
 &\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_A)\frac{1}{n_A}\lambda + I(Z_j = n_A + 1)(1 - \lambda)\right]
+&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
 \end{align*}
@@
 determined only by the number of duplicates
-\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
-\(\mathbf{Z}\). Thus we have
+$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
+$\mathbf{Z}$. Thus we have
 
 \[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
-\[\propto \lambda^D (1-\lambda)^{n_B - D} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
-\[ \propto \lambda^{D + \alpha_{\lambda} - 1} (1-\lambda)^{n_B - D + \beta_{\lambda} -1}\]
-\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
+\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
+\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
+\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
 
@@
 \begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
-	&\propto \prod_{i=1}^{n_A}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
+	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
+	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
 	&=
 	\begin{cases} 
-		w_{ij}  & z_j \leq n_A; \\
-		1 &  z_j  = n_A + 1 \\
+		w_{ij}  & z_j \leq n_1; \\
+		1 &  z_j  = n_1 + j \\
 	\end{cases}\\
@@
 \(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
-option to leave the matching undetermined by the model. Specifically, we
-minimize the quantity
-\(L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)\)
-where
+option to leave the matching undetermined by the model. Specifically, we have
 
@@
 which the record has a true match other than the one estimated by the
-model. This loss function leads to closed form decision rules for
-minimizing posterior expected loss. In this paper, we adopt losses
-\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\),
-inducing the intuitive decision rule
+model.
+
+We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$.
+In this paper, we adopt losses
+\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
 
@@
 
-Since our Gibbs procedure does not strictly enforce one-to-one matching,
-it is possible for the final Bayes estimate to link multiple records in
-\(B\) to one record in \(A\). The modeler can either report both such
-matches (with their respective posterior match probabilities), or
-resolve these conflicts by accepting only the match with highest
-posterior probability.
-A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities.
-Such a resolution procedure indeed is equivalent
-to minimizing posterior risk under the restriction of one-to-one
-matching, and as thus theoretically justified within the Bayesian
-framework.
+{
+\color{blue}
+Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable.
+We will explore this issue more deeply through the El Salvador homicide case study.
+
+To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+In the event that two matches.
+In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.
+A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
+}
 
@@
 patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
-that this bound \textcolor{red}{does not depend on} \(n_1\) or \(n_2\). Prior to
+that this bound does not depend on \(n_1\) or \(n_2\). Prior to
 processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
@@
 
-\textcolor{red}{RCS: I stopped here.}
-
 \section{Simulation Studies}
@@
 
+\textcolor{blue}{For Beka: The simulated files contain 6 fields of information, but he only uses 4 in his study.
+One of the fields was giving me problems, so I had been using a different one instead.
+I just figured out how to make the field work, so I can rerun the code.
+At that point, it will be a 100\% true replication.}
+
 \textcolor{red}{Brian: If you look in Sadinle (2017), did you use Table 2 or something else?
 Can you construct something similar to Figure 3 and Figure 4 as it's likely a reviewer will look at this paper and ask why you didn't do this exact comparison.}
+
+{
+\color{blue}For Beka: Yes, I used Table 2 (those are default settings in BRL)
+
+My Figure 1 is equivalent to Sadinle's figure 3.
+He compares against Fellegi Sunter, whereas I compare against BRL.
+I found his style of graphic difficult to read, especcially because I wanted to show F-measure in addition to recall and precision.
+
+Sadinle's Figure 4 just shows the results when you add the option of not declaring matching status (leaving it undecided by the model).
+This didn't seem to be so relevant in my paper.
+I can make the plot by changing just a few lines of code, so I can do it if you think I need it!}
 
@@
 My suggestion around all this is to look at the computational complexity of both methods as it's hard to attack this.}
+
+{
+\color{blue} For Beka: I'm not using any blocking for either method.
+
+Some of the hashing techniques could work for BRL (updating the m and u parameters), and they technically could be used for updating Z, but it would be much less useful because BRL requires knowing which exact record you match to, not just the agreement pattern for that record.
+Its not feasible for me try making those changes since its coded in C.
+
+I'll work out the computational complexity. That can be one of the things we nail down on Thursday!}
 
@@
 
+\textcolor{blue}{This is with post processing already.
+All I meant here was that is that small tasks require less information.
+If you do a $500 \times 500$ linkage task and you only have first and last name (for example), thats likely enough to get all the correct matches.
+However, once you do $20000 \times 20000$, its more likely that more than one person in the second file has the same first and last name, so you need more information to distinguish.
+I mentioned before that the simulated files actually have 6 fields of information, but I only used 4 originally because thats what Sadinle did.
+I can use all 6 though to see what happens?}
+
  \textcolor{red}{How long does it take on the same machine to run BRL. Can you run this? This can close this section out!}
+
+ \textcolor{blue}{When I use the compareRecords from BRL to do the pairwise comparison on the concatenated files, its too  big for R to handle!
+One could do hashing, reduce the data down to summary statistics, and alter the BRL to use that new object, but that would be a whole different project!}
 

### 2021-10-06 R/ResolveConflicts.R

@@ -2,7 +2,7 @@ ResolveConflicts <- function(Zhat){
 
   post_probs <- Zhat[[2]]
   Zhat <- Zhat[[1]]
-  double_matches <- Zhat[duplicated(Zhat)]
+  double_matches <- Zhat[duplicated(Zhat) & Zhat != -1]
   to_resolve <- unlist(lapply(double_matches, function(x){
     dfB_options <- which(Zhat == x)
     dfB_probs <- post_probs[dfB_options]

### 2021-10-07 methods_paper/fabl-paper.tex
added tables

@@
 In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}.
-Our framework is then embarassingly parallel, available in the \text{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.
+Our framework is then embarassingly parallel, available in the \texttt{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.
 
@@
 If so, I would state this for clarity.}
-
-\textcolor{blue}{For Beka: The simulated files contain 6 fields of information, but he only uses 4 in his study.
-One of the fields was giving me problems, so I had been using a different one instead.
-I just figured out how to make the field work, so I can rerun the code.
-At that point, it will be a 100\% true replication.}
-
-\textcolor{red}{Brian: If you look in Sadinle (2017), did you use Table 2 or something else?
-Can you construct something similar to Figure 3 and Figure 4 as it's likely a reviewer will look at this paper and ask why you didn't do this exact comparison.}
-
 {
-\color{blue}For Beka: Yes, I used Table 2 (those are default settings in BRL)
-
-My Figure 1 is equivalent to Sadinle's figure 3.
-He compares against Fellegi Sunter, whereas I compare against BRL.
-I found his style of graphic difficult to read, especcially because I wanted to show F-measure in addition to recall and precision.
-
-Sadinle's Figure 4 just shows the results when you add the option of not declaring matching status (leaving it undecided by the model).
-This didn't seem to be so relevant in my paper.
-I can make the plot by changing just a few lines of code, so I can do it if you think I need it!}
-
-We first compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
+\color{blue}For Beka: I have re-run the simulation exactly as Sadinle implemented it. Here is the rewritten paragraph.
+
+\vspace{1em}
+In this section, we replicate a simulation study from \cite{sadinle2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
 We use first name, last name, age, and occupation for this linkage, and create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package.
+See Table~\ref{Tab:sadinle_simulation_cutoffs} for details on the construction of comparison vectors.
 Each simulation identifies duplicated individuals between two datasets, each with 500 records.
@@
 We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
-This is a near exact replication of the simulation study provided by \citep{sadinle2017}.
+\vspace{1em}
+}
+
+\begin{table}
+	\centering
+	\begin{tabular}[t]{ll|llll}
+		\hline
+		\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
+		\cline{3-6}
+		Fields & Similarity & 1 & 2 & 3 & 4\\
+		\hline
+		First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (5, .1]\\
+		\hline
+		Age and Occupation & Binary & Agree & Disagree &  & \\
+		\hline
+	\end{tabular}
+\caption{Construction of comparison vectors for accuracy study with simulated datasets}
+\label{Tab:sadinle_simulation_cutoffs}
+\end{table}
+
+\textcolor{red}{Brian: If you look in Sadinle (2017), did you use Table 2 or something else?
+Can you construct something similar to Figure 3 and Figure 4 as it's likely a reviewer will look at this paper and ask why you didn't do this exact comparison.}
+
+{
+\color{blue}For Beka: Yes, I used Table 2 (those are default settings in BRL). See above comment
+
+My Figure 1 is equivalent to Sadinle's figure 3.
+He compares against Fellegi Sunter, whereas I compare against BRL.
+I found his style of graphic difficult to read, especially because I wanted to show F-measure in addition to recall and precision.
+
+Sadinle's Figure 4 just shows the results when you add the option of not declaring matching status (leaving it undecided by the model).
+This didn't seem to be so relevant in my paper, so I didn't have it before.
+I just completed the simulation, and I'll include it under a seperate section.
+}
 
@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot} 
+\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
 \caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
@@
 %\end{figure}
+
+\hypertarget{partial}{%
+	\subsection{Accuracy under Partial Estimates}\label{partial}}
+
+{
+\color{blue}
+Lastly, we review the performance of \texttt{fabl} and \texttt{BRL} using loss functions that allow partial estimates of the linkage parameter.
+By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
+In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
+Thus, instead of recall, we use the \emph{negative predictive value} (NPV), defined as the proportion of non-links that are actual non-matches.
+Mathematically, $NPV = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
+We continue to use the precision, which is renamed the \emph{positive predictive value} in this context.
+Lastly, we also report that rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.
+
+We see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
+However, with high amounts of overlap and high amounts of error, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
+This is reasonable because large numbers of errors lead to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
+Since \texttt{BRL} removes previously matched records from the set of available records for a new match, it is able to retain higher confidence in matches.
+In practice, this partial estimate approach is not often used, so we remain confident in \texttt{fabl}'s performance.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
+
+NOTE: Sadinle only shows results for the 10\% overlap case. Should I inlcude all of this?
+NOTE: When I ran this, I didn't realize I needed to calculate Rejection Rate. I'll have to run the simulation again to get this. 
+}
+
+\begin{figure}[htbp]
+\begin{center}
+	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
+	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets}
+	\label{fig:sadinle_simulation_partial}
+\end{center}
+\end{figure}
 
@@
 	\end{tabular}
-\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{fig:distributions}
+\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{Tab:distributions}
 \end{table}
@@
 distance to account for the fact that second names are often omitted.
-Place of birth is recorded by municipality and department within that
+Place of death is recorded by municipality and department within that
 municipality; however, since department was missing in 95\% of records
@@
 priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters. To our knowledge, we have followed the same guidelines/settings in \cite{sadinle_bayesian_2017}.
+\begin{table}
+\begin{tabular}[t]{ll|llll}
+	\hline
+	\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
+	\cline{3-6}
+	Fields & Similarity & 1 & 2 & 3 & 4\\
+	\hline
+	First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (5, 1]\\
+	\hline
+	Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
+	\hline
+	Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
+	\hline
+	Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
+	\hline
+	Municipality & Binary & Agree & Disagree &  & \\
+	\hline
+\end{tabular}
+\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle2018bayesian}.
+This set up leads to 2048 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_1}
+\end{table}
+
+\vspace{1em}
+
+\begin{table}
+	\centering
+	\begin{tabular}[t]{ll|lll}
+		\hline
+		\multicolumn{2}{c|}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
+		\cline{3-5}
+		Fields & Similarity & 1 & 2 & 3\\
+		\hline
+		First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, 1]\\
+		\hline
+		Year of Death & Binary & Agree & Disagree & \\
+		\hline
+		Month of Death & Binary & Agree & Disagree & \\
+		\hline
+		Day of Death & Absolute Difference & 0 & 1 & 2+\\
+		\hline
+		Municipality & Binary & Agree & Disagree & \\
+		\hline
+	\end{tabular}
+\caption{Construction of comparison vectors for El Salvador for increased speed under \texttt{fabl}. This set up leads to 216 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_2}
+\end{table}
 
@@
 vectors using four levels of agreement for each field, according to the
-thresholds provided in \textcolor{red}{Figure XX}. This took 422 seconds for our proposed
+thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}. This took 422 seconds for our proposed
 method, \texttt{fabl}, and 240 for \texttt{BRL}. However, we observed that posterior
@@
 
-Therefore we re-ran our analysis with fewer agreement levels for each field \textcolor{red}{(see Figure XX)}, and
+Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}, and
 obtained analogous results. With 216 possible agreement patterns, 159

### 2021-10-08 methods_paper/fabl-paper.tex
added computational complexity

@@
 \begin{abstract}
-In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
+%In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
 In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity.
@@
 
-\vspace{1ex}
 {
-\color{blue}
+	\color{blue} Thanks Beka for writing that first abstract. I think this one captures the idea better, even if its not exactly in "outline" form.
+	\vspace{1em}
+}
+
 In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
@@
 
-\vspace{1ex}
-QUESTION: Does this abstract necessarily need to read like an outline?
-}
 \end{abstract}
@@
 \textbf{Our Contribution} 
-In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\text{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations, and to reduce computational complexity.
+In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\texttt{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations, and to reduce computational complexity.
 Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching.
@@
 
-In this section, we deviate from the approach of \cite{sadinle_bayesian_2017}, and propose a more scalable prior on the matching labelings $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2}),$ where recall $Z_j \in \{1,2, 3, n_1, n_1 + j\}.$
-
-While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.
-
-\textcolor{red}{Brian: could you please lay out your proposal here more clear regarding the notation above and add as much details as possible regarding your approach.
-More detail is better than less.
-It will be important to contrast this with the prior that Sadinle uses and note that you're dropping the constraint.
-(Make sure to include any proofs/derivations that might not be obvious to a reader given this is a major contribution to the paper).}
+In this section, we deviate from the approach of \cite{sadinle_bayesian_2017}, and propose a more scalable prior on the matching labelings $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2}),$ where recall $Z_j \in \{1,2, 3, n_1, n_1 + j\}.$ While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.
 
@@
 
-{
-	\color{blue}
 Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
@@
 One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
-}
 
@@
 
-{
-\color{blue}
 There is a clear relationship between our proposed model and that of \cite{sadinle_bayesian_2017}. 
@@
 matched to at most one record in \(\bm{X}_1\), but allows for the possibility
-that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). \textcolor{blue}{We resolve contradictions when calcuating the Bayes estimate, which is explained in Section 4.5}
-
-NOTE TO BEKA: I think it will be best to do the "post-processing step" along with the Bayes Estimate.
-Its equivalent to minimizing the posterior loss subject to the constraint that $Z_j \neq Z_{j'}, \forall j \neq j'$.
-}
+that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
 
@@
 
-{
-\color{blue}
 Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
@@
 A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
-}
 
@@
 
-\textcolor{red}{Brian: how large can we go and how fast can we go?}
-
-\section{Efficient Posterior Inference}\label{sec:efficient-posterior}
+\hypertarget{efficient-posterior}{%
+	\subsection{Efficient Posterior Inference }\label{efficient-posterior}}
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[h!]
 	\caption{Summary of fabl algorithm}
@@
 
+\hypertarget{computational-complexity}{%
+	\subsection{Computational Complexity}\label{computational-complexity}}
+
+{
+\color{blue}
+The computational complexity of \texttt{fabl} is determined by its two steps: constructing comparison vectors and Gibbs sampling.
+The computational complexity of all pairwise comparisons across files $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$.
+\textcolor{blue}{TALK TO BEKA.
+I CAN DO THE SAME TRICKS AS TED HERE TO REDUCE COMPLEXITY}.
+Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
+
+Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
+However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the complexity to $O(FP)$.
+
+The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
+With hashing, we split this sampling into two steps; first we sample that agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record that exhibits that pattern with complexity $O(n_2)$.
+Thus we have reduced the complexity of the Gibbs sampler under \texttt{BRL} from $O(F n_1 n_2)$ to $O(n_2 P)$.
+
+Overall, the computational complexity under the complete \texttt{fabl} framework is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
+}
+
 \section{Simulation Studies}
 \label{sec:simulations}
-In this section, we consider a simulation study, nearly identical to that in \cite{sadinle2017}.
-\textcolor{red}{Brian: can you explain how it's the same versus different than what it is given in Sadinle as this would be the most helpful.
-Do you utilize Mauricio's exact files and approach and then do something else?
-If so, I would state this for clarity.}
-{
-\color{blue}For Beka: I have re-run the simulation exactly as Sadinle implemented it. Here is the rewritten paragraph.
-
-\vspace{1em}
+
 In this section, we replicate a simulation study from \cite{sadinle2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
@@
 We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
-\vspace{1em}
-}
-
-\begin{table}
+
+\begin{table}[h!]
 	\centering
@@
 \end{table}
-
-\textcolor{red}{Brian: If you look in Sadinle (2017), did you use Table 2 or something else?
-Can you construct something similar to Figure 3 and Figure 4 as it's likely a reviewer will look at this paper and ask why you didn't do this exact comparison.}
-
-{
-\color{blue}For Beka: Yes, I used Table 2 (those are default settings in BRL). See above comment
-
-My Figure 1 is equivalent to Sadinle's figure 3.
-He compares against Fellegi Sunter, whereas I compare against BRL.
-I found his style of graphic difficult to read, especially because I wanted to show F-measure in addition to recall and precision.
-
-Sadinle's Figure 4 just shows the results when you add the option of not declaring matching status (leaving it undecided by the model).
-This didn't seem to be so relevant in my paper, so I didn't have it before.
-I just completed the simulation, and I'll include it under a seperate section.
-}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[h!]
 \begin{center}
@@
 
-{
-\color{blue}
 Lastly, we review the performance of \texttt{fabl} and \texttt{BRL} using loss functions that allow partial estimates of the linkage parameter.
@@
 
-NOTE: Sadinle only shows results for the 10\% overlap case. Should I inlcude all of this?
-NOTE: When I ran this, I didn't realize I needed to calculate Rejection Rate. I'll have to run the simulation again to get this. 
-}
-
-\begin{figure}[htbp]
+\begin{figure}[h!]
 \begin{center}
@@
 	\subsection{Speed}\label{speed}}
-
-\textcolor{red}{Brian: I have some general questions here.
-Are you using blocking for BRL to scale?
-Could you use the hashing proposal for BRL.
-It's important that the scale comparison be as fair as possible or there will be some pushback here.
-My suggestion around all this is to look at the computational complexity of both methods as it's hard to attack this.}
-
-{
-\color{blue} For Beka: I'm not using any blocking for either method.
-
-Some of the hashing techniques could work for BRL (updating the m and u parameters), and they technically could be used for updating Z, but it would be much less useful because BRL requires knowing which exact record you match to, not just the agreement pattern for that record.
-Its not feasible for me try making those changes since its coded in C.
-
-I'll work out the computational complexity. That can be one of the things we nail down on Thursday!}
 
@@
 
-\begin{table}[t]
+\begin{table}[h!]
 	\centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[h!]
 
@@
 
-\begin{figure}[t]
+\begin{figure}[h!]
 
@@
 
-% Assuming a 16GB personal machine, standard record linkage methods using all-to-all comparisons can only function on tasks where the matrix of comparison vectors $\Gamma$ is less than 8GB.
-In the context of the above simulation, each comparison vector contains four integers, collectively requiring 16 bytes of memory; assuming files of equal lengths, this means the largest linkage task possible is around $7000 \times 7000$.
-
 In this section, we investigate scaling to larger data sets using our proposed methodology.
@@
 We concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{} to create two larger files, each of 20,000 records (40,000 records total).
-Working on one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
 Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting a final comparison matrix about 6.4 GB in size.
+Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task.
+Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
 
@@
 The resulting data object is now only 90 MB, about 1\% the size of the object required under the standard method.
-Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run; using distributed computing.
-However, this could be sped up.
+Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run.
+However, this could be substantially sped up using distributed computing.
 
@@
 With more linkage fields \texttt{fabl} maintains the high accuracy seen above.
-\textcolor{red}{Brian: you should look to see if you fix these in the post-processing step as this would be interesting/important.}
-
-\textcolor{blue}{This is with post processing already.
+
+%\textcolor{blue}{This is with post processing already.
 All I meant here was that is that small tasks require less information.
@@
 
- \textcolor{red}{How long does it take on the same machine to run BRL. Can you run this? This can close this section out!}
-
- \textcolor{blue}{When I use the compareRecords from BRL to do the pairwise comparison on the concatenated files, its too  big for R to handle!
-One could do hashing, reduce the data down to summary statistics, and alter the BRL to use that new object, but that would be a whole different project!}
-
 \section{Application to Civilian Casualties from the El Salvadoran Civil War}
-\label{sex:case-studies}
+\label{sec:case-studies}
 
@@
 \label{el_salvador}
+
+{
+\color{blue} For Beka: You wrote that you wanted more detail here, and I added some of it (with some citations).
+I believe I have all the important informatoin, just more concise.
+If I included Sadinle's level of detail, I worried about plagarism.
+\vspace{1em}
+}
 
@@
 organization, as it is likely that the same individuals are recorded in
-multiple casualty lists. Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize 
+multiple casualty lists.
+This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications} Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize
 two different sources, namely,  El Rescate - Tutela
 Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
-acronym in Spanish). The ERTL dataset consists of digitized reports that
-had been published throughout the conflict. The CDHES dataset consists
+acronym in Spanish). The ERTL dataset consists of digitized denunciations that
+had been published throughout the conflict, and additional investigation was required before the ERTL recorded them as human rights abuses, giving us confidence in its accuracy \cite{howland2008rescate}.
+The CDHES dataset consists
 of casualties that had been reported directly to the organization, and
-later digitized.
-
-\textcolor{red}{Beka: These data sets need more information regarding how the data sets were collected as this is quite important. See Sadinle (2017) and references for info). }
+later digitized; since reports were taken soon after the events occured, this dataset is also thought to be fairly reliable. \cite{ball2000salvadoran}.
 
@@
 \begin{table}
-\begin{tabular}[t]{ll|llll}
+\begin{tabular}[h!]{ll|llll}
 	\hline
@@
 	\centering
-	\begin{tabular}[t]{ll|lll}
+	\begin{tabular}[h!]{ll|lll}
 		\hline
@@
 vectors using four levels of agreement for each field, according to the
-thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}. This took 422 seconds for our proposed
+thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}. Gibbs sampling for these comparison vectors took 422 seconds for our proposed
 method, \texttt{fabl}, and 240 for \texttt{BRL}. However, we observed that posterior
@@
 obtained analogous results. With 216 possible agreement patterns, 159
-were realized in the data, and our proposed method, \texttt{fabl} became was faster,
-finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
+were realized in the data, and our proposed method, \texttt{fabl} became much faster,
+finishing in 61 seconds. Meanwhile \texttt{BRL} took 239 seconds,
 relatively unchanged from the first implementation. This demonstrates
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
-\label{fig:m-and-u}
-\end{center}
-\end{figure}
-
-\paragraph{Violations of One-to-One Matching}
-
-In this section, we investigate violations of one-to-one matching of our method as compared to \texttt{BRL}.
-
-Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
-duplicates found in both databases for both \texttt{fabl} and \texttt{BRL}. We
-see that \texttt{fabl} consistently overmatches within each Gibbs
-iteration when compared to \texttt{BRL}, which is to be expected because
-\texttt{BRL} explicitly prevents matches that violate one-to-one
-matching throughout the entire sampler. Most of such matchings are due
-to the randomness in the sampling procedure, and they occur sporadically
-throughout the sampler in such a way that does not measurably influence
-the eventual Bayes estimate.
-
-\begin{figure}[htbp]
+{
+\color{blue}
+Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets.
+We calcuate posterior samples of size of the overlap across files by summing the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
+The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate is quite conservative.
+This is because there a large number of records in ERTL for which there are multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair a match in the Bayes estimate.
+We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study.
+Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually; this means that after clerical review, the number of individuals replicated across datasets would fall in the interval (136, 311), encapsulating the posterior credible interval.
+More or fewer records could be indentified for clerical review by decreasing or increasing $\theta_R$.
+
+We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
+}
+
+\begin{figure}[h!]
 	\begin{center}
@@
 
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
+\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
+\label{fig:m-and-u}
+\end{center}
+\end{figure}
+
+\paragraph{Violations of One-to-One Matching}
+
+%In this section, we investigate violations of one-to-one matching of our method as compared to \texttt{BRL}.
+{
+\color{blue}
+As noted previously, the application exhibits many situations in which bipartite matching is difficult.
+To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
+These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in \(\bm{X}_1\) has
+multiple plausible matches in \(\bm{X}_2\). 
+}
+
+%Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
+%duplicates found in both databases for both \texttt{fabl} and \texttt{BRL}. We
+%see that \texttt{fabl} consistently overmatches within each Gibbs
+%iteration when compared to \texttt{BRL}, which is to be expected because
+%\texttt{BRL} explicitly prevents matches that violate one-to-one
+%matching throughout the entire sampler. Most of such matchings are due
+%to the randomness in the sampling procedure, and they occur sporadically
+%throughout the sampler in such a way that does not measurably influence
+%the eventual Bayes estimate.
+
 %\begin{figure}[t]
@@
 
-Overall, both methods present similar results. \texttt{fabl} yields
-an initial Bayes estimate of total 195 DID. After
-resolving matches that violated one-to-one requirements,  we estimate 180 DID. 
-\textcolor{red}{Can you report the posterior standard error for each one as we want uncertainty quantification.}
-
-This is very close to the 181 DID found
-by \texttt{BRL}. The main reason for this discrepancy is the difference
-in how each model handles situations in which one file in \(\bm{X}_1\) has
-multiple plausible matches in \(\bm{X}_2\).
+%Overall, both methods present similar results. \texttt{fabl} yields
+%an initial Bayes estimate of total 195 DID. After
+%resolving matches that violated one-to-one requirements,  we estimate 180 DID. 
+%
+%
+%
+%This is very close to the 181 DID found
+%by \texttt{BRL}. The main reason for this discrepancy is the difference
+%in how each model handles situations in which one file in \(\bm{X}_1\) has
+%multiple plausible matches in \(\bm{X}_2\).
 
@@
 
-Figure~\ref{fig:rosa-maria} illustrates an example. Note these
+Figure~\ref{fig:rosa-maria} illustrates an example. Note that these
 records present a near violation of our assumption that there are no

### 2021-10-09 methods_paper/fabl-paper.tex
proposed changes to introduction and first four sections

@@
 
+{
+	\color{blue}
+
+	I have some proposed revisions for the first part of the paper.
+In the introduction, I reordered things and changed some topic sentences to highlight that the main contribution here is speed and scale.
+
+	Also, in the prior version, we didn't get to any of the proposed material until Section 4.3, which did not seem reasonable to me.
+	I know you recommended splitting the review of the literature between the Introduction and Section 2 on bipartite matching, but since we needed the mathemetical overview of FS and Sadinle, I decided to move all of the literature review into the introduction.
+I've seen this done in many papers.
+
+	Almost all of the content is still there, but it's moved around in a way that makes more sense to me. Let me know your thoughts! (But please avoid hard edits until we talk on Monday!)
+
+	I have some additional edits in the later part of the paper about the partial Bayes estimate, computational complexity, and added detail for the case study. All of that is in blue as well.
+
+	---------------------------------------------
+
+	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
+Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+In this paper, we consider bipartite record linkage, which is the merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
+
+	%Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
+
+	Many statistical record linkage methods and case studies are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
+Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test (or hypothesis test).
+Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.
+Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.
+
+	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice.
+in many contexts, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record in the other database.
+Here, additional declared matches are known by assumption to be false.
+Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989, enamorado2019using}, but this model misspecification can still lead to poor results \cite{sadinle_bayesian_2017}.
+
+	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost.
+\cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling from matches that respected one-to-one matching, but such algorithms are slow due to the combinatorial nature of the constrained matching space.
+\cite{fortunato_2010} used simulated annealing to target the space of permitted matches under this constraint, but the method was so computationally intensive it could used to link databases of less than 100 records.
+Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (BRL), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler.
+Additionally, he introduces a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
+Unfortunately, the method becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.
+
+	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends a \texttt{BRL} model for increased efficiency and scalability.
+Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
+This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}.
+Throughout, we argue that even in cases where a bipartite matching is desired, this relaxed model (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure, more information through which to asses model misspecification, and greatly enhanced speed.
+To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
+	pronounced ``fable'')
+
+	In what follows, we first review more thoroughly the Fellegi Sunter and Beta Record Linkage methods.
+Then, we provide the model specification under \texttt{fabl} and derive the Gibbs sampler for posterior inference.
+Next, we introduce the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
+Lastly, we use simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
+
+\section{Review of Prior Work}
+\label{sec:review-of_prior-work}
+
+Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$.
+Without loss of generality, denote files suchs that $n_1 \geq n_2$.
+In the context of bipartite matching, we assume that there are duplications across the databases but not within each database.
+Within this framework, the set of matches across datasets can be viewed in two equivalent ways:
+\begin{align}
+	\Delta_{ij} =
+	\begin{cases}
+		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
+		0 \quad \text{otherwise}.\\
+	\end{cases}
+\end{align}
+This matrix valued representation of the set of matches is of size $n_1 \times n_2$, which can become costly for large linkage tasks or for long posterior samplers.
+More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that
+\begin{align}
+	Z_{j} =
+	\begin{cases}
+		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
+		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
+	\end{cases}
+\end{align}
+Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.
+
+Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$ Denote the set of non-matches 
+$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ Merging two databases and removing 
+duplications can be viewed as identifying the sets of matches/non-matches, or rather $\bm{M}, \bm{U}.$ We refer to record pairs that are estimated as matches as "links", and record pairs that are estimated as non-matches are known as "non-links".
+\cite{fellegi_theory_1969} proposed using all-to-all pairwise comparison of records to estimate the matches and non-matches.
+
+\subsection{Comparison Data}
+
+Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar.
+A common way to encode this is using a comparison vector $\bm{\gamma}_{ij}$ which is computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote $F$ as the criteria for compare the records such that
+$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $F$ represents to a single comparison per feature that the databases share, such as gender.
+
+The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for nominal variables such as gender.
+For more complex measurements, such as textual or numeric feature information, we can take into account partial agreement patterns using common string metrics, such as the Edit, Jaro, or Jaro-Winkler distance functions.
+This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+
+Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$ representing disagreement levels.
+Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
+Thus, we can construct comparison vectors in the following way:
+$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$
+
+The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.
+
+%In what follows we use notation that reflects that these comparison vectors can be modelled in number of ways.
+In practice however, this binning in beneficial because it allows us to model any type of comparison through a multinomial distribution.
+In a Bayesian context, Dirichlet priors for
+
+\subsection{Fellegi-Sunter Model}
+While we have introduced comparison data, $\bm{\gamma}_{ij}$, this is not sufficient to determine whether a record is a match/non-match given errors that naturally occur in the data.
+This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
+\begin{align}
+	\label{eqn:wts}
+	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
+\end{align}
+as a weight to estimate if record pairs are a match/non-match. The ratio will be large if there is strong evidence of the pair being a match, and small otherwise.
+
+In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
+We assume the comparison vectors are independent given the bipartite matching and assume that the match status of the record pairs are independent.
+It then follows that we can model
+the comparison data using mixture models in the following way:
+\begin{align}
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
+	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
+\end{align}
+where $p$ is the proportion of records that match.
+Note that $\bm{m}$ and $\bm{u}$ are vectors of the match/non-match parameters.
+One common way to estimate the unknown parameters of the aforementioned model is using the EM algorithm.
+
+After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ declares as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined).
+This is a problem in practice as transitive closures are violated.
+Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} T_m$, even those such a matching is not allowed by assumption.
+
+\cite{jaro1989} proposed addressing this issue via a post-processing optimization approach as follows:
+\begin{align}
+	\label{eqn:jaro}
+	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
+	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
+	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
+	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
+\end{align}
+
+Equation \ref{eqn:jaro} ensures that each record of $\bm{X}_2$ only matches with at most one record in $\bm{X}_1$ (and the reverse condition).
+This is commonly known as the \emph{maximum-weight bipartite matching problem} or a \emph{linear sum assignment problem}.
+The output of this algorithm is a bipartite matching that maximizes the sum of the weights among matched pairs.
+\cite{jaro1989} provided no theoretical justification for using this approach, however, \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
+
+%One criticism of using mixture models as proposed by \cite{jaro1989} is the fact that the decision rule will lead to ``many-to-many assignments (breaking transitive closures).
+One way of resolving this issue to incorporate the constraint directly into the model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, Larsen05, Larsen12, gutman_bayesian_2013, sadinle_bayesian_2017}, which is appealing as one then obtains uncertainty quantification under the Bayesian paradigm, but often has to sacrifice scaling to very large databases.
+On the other hand, one can turn to post-processing methods in practical situations.
+In this paper, we attempt to bridge the best of both approaches.
+Specifically, we propose a model in the spirit of \cite{sadinle_bayesian_2017} and \cite{enamorado_using_2019}.
+We relax the one-to-one matching constraint proposed by  \cite{sadinle_bayesian_2017} for purely computational reasons, and resolve this in a post-processing step.
+In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}.
+Our framework is then embarassingly parallel, available in the \texttt{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.
+
+\subsection{Beta Linkage Model}
+One criticism of using mixture models is the fact that the decision rule will lead to "many-to-many assignments", breaking transitive closures.
+One way of resolving this issue to incorporate the constraint directly into the model.
+The Beta Linkage Model of \cite{sadinle_bayesian_2017} does through the following model:
+%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
+\begin{align}
+	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
+	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
+	& \bm{Z} \sim \mathcal{B}, \notag
+\end{align}
+where $\mathcal{B}$ represents a prior on the space of bipartite matchings and $\mathcal{M}(\bm{m}), \mathcal{M}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
+
+Here, $\mathcal{B}$ is a prior distribution on the entire \(\mathbf{Z}\) vector, called the \emph{beta distribution for bipartite matching}, given by
+$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
+$$
+where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior induces a Gibbs sampler that strictly enforces one-to-one matching,  previously matched records from the set of candidate
+records when sampling \(Z_j\). This creates a dependency that makes the
+sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.
+
+\subsection{Blocking}
+
+In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or "blocks" to reduces this computational burden.
+In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field.
+The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
+However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.
+
+In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without the need for blocking.
+This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
+In practice, the method the follows can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.
+
+------------------------------------------------------
+
+\textcolor{black}{I think its best just not to mention probabalistic blocking.
+My method doesn't use it, and introducing it is really confusing (Sadinle 2017, Enamorado 2019, and Steorts 2014/2015 don't mention it).}
+
+We note here that the word "hashing" is also used in probabalistic blocking, which uses techniques from machine learning to cluster records together before attempting to link records.
+The use of such probabalistic blocking to provide even greater computational gains will be explored in future work.
+
+-------------------------------------------------------
+
+\section{Fast Beta Linkage}
+\label{sec:fast-beta-linkage}
+
+Recall our compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, 3, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
+In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we propose independent priors for each component $Z_j$.
+We denote the \emph{fast beta prior} as follows:
+
+\[Z_j | \lambda =
+\begin{cases} 
+	\frac{1}{n_1}\lambda  & z_j \leq n_1; \\
+	1-\lambda &  z_j  = n_1 + j \\
+\end{cases}\]
+
+\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \]
+
+Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
+$\lambda$ itself is given a prior distribution to reflect prior beliefs about the overall rate of matchingness in the data.
+One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
+
+Importantly, these independent priors relax the one-to-one requirement from\texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
+This behavior may be desireable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such "many-to-one" matchings are reasonable.
+In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.
+
+%this independent prior for each \(Z_j\), induces a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains over the Gibbs sampler under \texttt{BRL}.
+More importantly, since only the agreement pattern of \(Z_j\) is used forcalculations within the Gibbs sampler, and not the particular record label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.
+
+\hypertarget{fabl-model}{%
+	\subsection{Model Specification}
+	\label{model-specification}}
+
+\textcolor{black}{NOTE: I really want to present the entire model in one place like this.
+I saw this for the first time in Steorts 2014, and I really liked it.
+Part of the reason it took me so long to understand Sadinle 2017 was because the model was presented over several sections}
+
+As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent.
+Additionally, we assume that any missing data is missing at random.
+Let  $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denote the probability of that matching records have level $\ell$ of disagreement in feature $f$, and $u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represent the same probability for non-matches.
+In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, $\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
+The model can then be specified as follows:
+
+$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
+$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
+$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
+$$Z_j | \lambda =
+\begin{cases} 
+	\frac{1}{n_1}\lambda  & z_j \leq n_1; \\
+	1-\lambda &  z_j  = n_1 + j \\
+\end{cases}$$
+$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda})$$
+
+\hypertarget{posterior-sampling}{%
+	\subsection{Gibbs Sampler}
+	\label{gibbs_sampling}}
+We work with the following factorization of the joint posterior distribution:
+
+\begin{align*}
+	p(\mathbf{Z}, \Phi, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
+	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
+	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
+\end{align*}
+This factorization leads to following Gibbs Sampler:
+
+\underline{Sample $\Phi^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
+The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
+standard multinomial-dirichlet mechanics. Thus we have
+
+\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
+\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
+where
+\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
+and
+\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).
+
+\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
+\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
+of successes (when \(z_j < n_A + 1\)) and failures (when
+\(z_j = n_A + 1\)), and therefore
+\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
+determined only by the number of duplicates
+$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
+$\mathbf{Z}$. Thus we have
+
+\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
+\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
+\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
+\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
+
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \Phi^{(s+1)}, \lambda^{(s+1)}$:}
+Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
+only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
+denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
+that as a function of \(Z_j\), the likelihood
+\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
+is a discrete distribution with probabilities proportional to
+
+\begin{align*}
+	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	&=
+	\begin{cases} 
+		w_{ij}  & z_j \leq n_1; \\
+		1 &  z_j  = n_1 + j \\
+	\end{cases}\\
+\end{align*}
+
+where
+\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)I_{obs}(\gamma_{ij}^f)}\).
+The interested reader should note that these are precisely the
+likelihood ratios used in the Fellegi-Sunter model to classify matches
+and non-matches, and we therefore refer to \(w_{ij}\) as the
+\emph{Fellegi Sunter weights}.
+
+With the likelihood in this form, we can derive the full conditional
+\[p(Z_j|\Gamma_{.j}, \Phi, \lambda) \propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\lambda)\]
+
+\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
+\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
+\[ \implies Z_j^{(s+1)} | \Phi, \Gamma, \lambda \propto
+\begin{cases} 
+	\frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
+	1-\lambda &  z_j  = n_A + 1 \\
+\end{cases}\]
+
+In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
+integrate over the posterior of \(\lambda\) and rearrange terms to
+produce the final full conditional:
+
+\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
+\begin{cases} 
+	w_{ij}  & i \leq n_A; \\
+	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
+\end{cases}\]
+
+\hypertarget{bayes-estimate}{%
+	\subsection{Bayes Estimate}
+	\label{bayes-estimate}}
+
+We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
+parameter \(\mathbf{Z}\) by assigning different positive losses to
+different types of errors, and minimizing posterior expected loss. We
+adopt the loss function proposed in Sadinle 2017, in which
+\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
+option to leave the matching undetermined by the model. Specifically, we have
+
+\[L(\hat{Z_j}, Z_j)=\begin{cases} 
+	0  & \text{if } Z_j = \hat{Z_j}; \\
+	\theta_R,  & \text{if } \hat{Z_j} = R; \\
+	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
+	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
+	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
+\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
+on the linkage status, \(\theta_{10}\) is the loss from a false
+non-match, \(\theta_{01}\) is the loss from a false match, and
+\(\theta_{11}\) is the loss from the special case of a false match in
+which the record has a true match other than the one estimated by the
+model.
+
+We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$.
+In this paper, we adopt losses
+\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
+
+\[\hat{Z}_j =\begin{cases} 
+	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
+	0,  & \text{otherwise} ; \\
+\end{cases}\]
+
+For a more in-depth explanation of this function and the induced Bayes
+estimate, see \citep{sadinle_bayesian_2017}.
+
+Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable.
+We will explore this issue more deeply through the El Salvador homicide case study.
+
+To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+In the event that two matches.
+In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.
+A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
+
+}
+
+\newpage
+{
+\color{red}
+Here is the previous version of everything. Note that the numbering of the sections is off at the moment. 
+}
+
 %% Reworking introduction for Brian's paper.
@@
 In addition, posterior distributions are readily available to help estimate point estimates of interest (and quantity such estimates).
-Second, under the Bayesian paradigm, the one-to-one matching constraint is easily imbedded.
+Second, under the Bayesian paradigm, the one-to-one matching constraint is easily embedded.
 The downside to this constraint is an added computational cost.
@@
 \textbf{Our Contribution} 
-In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\texttt{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations, and to reduce computational complexity.
+In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\texttt{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations and to reduce computational complexity.
 Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching.
@@
 One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
-
-%The prior for \(Z_j\) has equal probability of matching to all records
-%\(i\in A\), and non-matching probability governed by \(\lambda\).
 

### 2021-10-10 methods_paper/fabl-paper.tex
additional edits througout paper

@@
 	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
-Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+This is an increasingly important task in "data cleaning," and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
 In this paper, we consider bipartite record linkage, which is the merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
@@
 
-	Many statistical record linkage methods and case studies are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test (or hypothesis test).
+	Many statistical record linkage methods are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
+Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test.
 Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.
@@
 	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice.
-in many contexts, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record in the other database.
+In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record in the other database.
 Here, additional declared matches are known by assumption to be false.
-Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989, enamorado2019using}, but this model misspecification can still lead to poor results \cite{sadinle_bayesian_2017}.
+Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
 
 	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost.
-\cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling from matches that respected one-to-one matching, but such algorithms are slow due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of permitted matches under this constraint, but the method was so computationally intensive it could used to link databases of less than 100 records.
+\cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
+\cite{fortunato_2010} used simulated annealing to target the space of permitted matches under the one-to-one constraint, but the method was so computationally intensive it could used to link databases of less than 100 records.
 Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (BRL), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler.
-Additionally, he introduces a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
-Unfortunately, the method becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.
-
-	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends a \texttt{BRL} model for increased efficiency and scalability.
-Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
+Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
+Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.
+
+	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
+Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a "many-to-one" model for record linkage.
 This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}.
-Throughout, we argue that even in cases where a bipartite matching is desired, this relaxed model (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure, more information through which to asses model misspecification, and greatly enhanced speed.
+Throughout, we argue that even in cases where a bipartite matching is desired, this relaxed model (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
 To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
@@
 Next, we introduce the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
-Lastly, we use simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
+Lastly, we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
 
@@
 Without loss of generality, denote files suchs that $n_1 \geq n_2$.
-In the context of bipartite matching, we assume that there are duplications across the databases but not within each database.
-Within this framework, the set of matches across datasets can be viewed in two equivalent ways:
+In the context of bipartite matching, we assume that there are duplications across, but not within, each database.
+Under this framework, the set of matches across datasets can be represented in two equivalent ways.
+First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
 \begin{align}
@@
 \end{align}
-This matrix valued representation of the set of matches is of size $n_1 \times n_2$, which can become costly for large linkage tasks or for long posterior samplers.
+Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers.
 More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that
@@
 
-Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$ Denote the set of non-matches 
-$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ Merging two databases and removing 
-duplications can be viewed as identifying the sets of matches/non-matches, or rather $\bm{M}, \bm{U}.$ We refer to record pairs that are estimated as matches as "links", and record pairs that are estimated as non-matches are known as "non-links".
-\cite{fellegi_theory_1969} proposed using all-to-all pairwise comparison of records to estimate the matches and non-matches.
+Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$, and the set of non-matches by 
+$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as "links", and record pairs that are estimated as non-matches as "non-links".
 
@@
 Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar.
-A common way to encode this is using a comparison vector $\bm{\gamma}_{ij}$ which is computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote $F$ as the criteria for compare the records such that
-$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $F$ represents to a single comparison per feature that the databases share, such as gender.
-
-The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for nominal variables such as gender.
-For more complex measurements, such as textual or numeric feature information, we can take into account partial agreement patterns using common string metrics, such as the Edit, Jaro, or Jaro-Winkler distance functions.
+\cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that
+$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
+
+The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for categorical variables.
+For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance.
 This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
-
-Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$ representing disagreement levels.
+Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$.
 Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
@@
 \subsection{Fellegi-Sunter Model}
-While we have introduced comparison data, $\bm{\gamma}_{ij}$, this is not sufficient to determine whether a record is a match/non-match given errors that naturally occur in the data.
+The comparison data $\bm{\gamma}_{ij}$ is not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
 This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
@@
 \end{align}
-as a weight to estimate if record pairs are a match/non-match. The ratio will be large if there is strong evidence of the pair being a match, and small otherwise.
+as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large if there is strong evidence of the pair being a match, and small otherwise.
 
 In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
-We assume the comparison vectors are independent given the bipartite matching and assume that the match status of the record pairs are independent.
-It then follows that we can model
+Under the Fellegi Sunter model, we assume that comparison vectors are independent given their matching status, and that match status of the record pairs are independent, allowing us to model
 the comparison data using mixture models in the following way:
@@
 \end{align}
-where $p$ is the proportion of records that match.
-Note that $\bm{m}$ and $\bm{u}$ are vectors of the match/non-match parameters.
-One common way to estimate the unknown parameters of the aforementioned model is using the EM algorithm.
-
-After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ declares as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined).
+where $p$ is the proportion of records that match, and $\bm{m}$ and $\bm{u}$ are vectors of parameters for the matching and nonmatching record pairs.
+These unknown parameters are often estimated using the EM algorithm.
+
+After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined).
 This is a problem in practice as transitive closures are violated.
-Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} T_m$, even those such a matching is not allowed by assumption.
-
-\cite{jaro1989} proposed addressing this issue via a post-processing optimization approach as follows:
+Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
+
+To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
 \begin{align}
@@
 
-Equation \ref{eqn:jaro} ensures that each record of $\bm{X}_2$ only matches with at most one record in $\bm{X}_1$ (and the reverse condition).
-This is commonly known as the \emph{maximum-weight bipartite matching problem} or a \emph{linear sum assignment problem}.
-The output of this algorithm is a bipartite matching that maximizes the sum of the weights among matched pairs.
-\cite{jaro1989} provided no theoretical justification for using this approach, however, \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
+The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
+\cite{jaro1989} provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
 
@@
 \subsection{Beta Linkage Model}
-One criticism of using mixture models is the fact that the decision rule will lead to "many-to-many assignments", breaking transitive closures.
-One way of resolving this issue to incorporate the constraint directly into the model.
-The Beta Linkage Model of \cite{sadinle_bayesian_2017} does through the following model:
+
+One criticism of using mixture models is that the decision rule will lead to "many-to-many assignments," breaking one-to-one matching assumptions.
+One way of resolving this issue to incorporate the constraint directly into the model as follows
+
 %our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
@@
 \end{align}
-where $\mathcal{B}$ represents a prior on the space of bipartite matchings and $\mathcal{M}(\bm{m}), \mathcal{M}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-
-Here, $\mathcal{B}$ is a prior distribution on the entire \(\mathbf{Z}\) vector, called the \emph{beta distribution for bipartite matching}, given by
+where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
+For \texttt{BRL} \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
 $$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
 $$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior induces a Gibbs sampler that strictly enforces one-to-one matching,  previously matched records from the set of candidate
+where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
+This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
 records when sampling \(Z_j\). This creates a dependency that makes the
@@
 
-Recall our compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, 3, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
+Recall our compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
 In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we propose independent priors for each component $Z_j$.
@@
 
-\[Z_j | \lambda =
+$$Z_j | \pi =
 \begin{cases} 
-	\frac{1}{n_1}\lambda  & z_j \leq n_1; \\
-	1-\lambda &  z_j  = n_1 + j \\
-\end{cases}\]
-
-\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \]
-
-Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
-$\lambda$ itself is given a prior distribution to reflect prior beliefs about the overall rate of matchingness in the data.
-One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
-
-Importantly, these independent priors relax the one-to-one requirement from\texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desireable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such "many-to-one" matchings are reasonable.
+	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
+	1-\pi &  z_j  = n_1 + j \\
+\end{cases}$$
+$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
+
+Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
+$\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$..
+One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
+We encourage the reader to integrate out $\pi$ from the $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.
+
+Note that linkage in this setting is conducted at the \emph{record} level, rather than at the \emph{record pair} level as in the Fellegi Sunter model.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimate that proportion of record pairs that have matches.
+We find $\pi$ to be more interpretable parameter than $\lambda$, especially considering that the overwhelming majority of pairs are non matches, meaning that $\lambda$ necessarily tends towards 0 as the size of the linkage task grows.
+Additionally, we emphasize that while the Fellegi Sunter model makes $n_1 \times n_2$ many independent matching decisions and \texttt{BRL} makes $n_2$ many dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$ many independent matching decisions.
+As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows us to achieve many of the accuracy gains from \texttt{BRL} while maintaining the efficiency gains possible by exploiting independence.
+
+Importantly, this independence means that we have relaxed one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
+This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such "many-to-one" matchings are reasonable.
 In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.
@@
 $$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
-$$Z_j | \lambda =
+$$Z_j | \pi =
 \begin{cases} 
-	\frac{1}{n_1}\lambda  & z_j \leq n_1; \\
-	1-\lambda &  z_j  = n_1 + j \\
+	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
+	1-\pi &  z_j  = n_1 + j \\
 \end{cases}$$
-$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda})$$
+$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
 
@@
 \begin{align*}
-	p(\mathbf{Z}, \Phi, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
+	p(\mathbf{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\mathbf{Z} | \pi) p(\Phi) p(\pi) \\
 	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
 	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
+	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
+	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}
 \end{align*}
@@
 
-\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
-\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
+\underline{Sample $\pi^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
+\(\pi\), the linkage structure parameter \(\mathbf{Z}\) is sequence
 of successes (when \(z_j < n_A + 1\)) and failures (when
 \(z_j = n_A + 1\)), and therefore
-\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
+\(p(\mathbf{Z}|\pi) = \mathcal{L}(\pi|\mathbf{Z})\) is
 determined only by the number of duplicates
@@
 
-\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
-\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
-\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
-\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
-
-\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \Phi^{(s+1)}, \lambda^{(s+1)}$:}
+\begin{align*}
+p(\pi | \mathbf{Z}) &\propto p(\mathbf{Z}|\pi)p(\pi) \\
+&\propto \pi^{n_{12}(\mathbf{Z})} (1-\pi)^{n_2 - (n_{12}(\mathbf{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+&\propto \pi^{n_{12}(\mathbf{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\pi} -1} \\
+\implies &\pi^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\pi})
+\end{align*}
+
+\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \Phi^{(s+1)}, \pi^{(s+1)}$:}
 Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
 only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
-denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
-that as a function of \(Z_j\), the likelihood
-\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
+denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\), and note that as a function of \(Z_j\), the likelihood
+\(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\)
 is a discrete distribution with probabilities proportional to
@@
 where
-\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)I_{obs}(\gamma_{ij}^f)}\).
+$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
 The interested reader should note that these are precisely the
@@
 With the likelihood in this form, we can derive the full conditional
-\[p(Z_j|\Gamma_{.j}, \Phi, \lambda) \propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\lambda)\]
-
-\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
-\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
-\[ \implies Z_j^{(s+1)} | \Phi, \Gamma, \lambda \propto
+\[p(Z_j|\Gamma_{.j}, \Phi, \pi) \propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi)\]
+
+\[\propto \left(\sum_{i=1}^{n_1}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\mathbf{1}_{z_j = i} + (1-\pi)\mathbf{1}_{z_j = n_1 + 1}\right)\]
+\[= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\mathbf{1}_{z_j = i} + (1-\pi)\mathbf{1}_{z_j = n_1 + 1} \]
+\[ \implies Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto
 \begin{cases} 
-	\frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
-	1-\lambda &  z_j  = n_A + 1 \\
+	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
+	1-\pi &  z_j  = n_1 + 1 \\
 \end{cases}\]
@@
 In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
-integrate over the posterior of \(\lambda\) and rearrange terms to
+integrate over the posterior of \(\pi\) and rearrange terms to
 produce the final full conditional:
 
-\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
+\[p\left(Z_j^{(s+1)}  = i| \Phi, \mathbf{Z^{(s)}}\right) \propto
 \begin{cases} 
-	w_{ij}  & i \leq n_A; \\
-	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
+	w_{ij}  & i \leq n_1; \\
+	n_1 \frac{n_2 - n_{12}(\mathbf{Z}) + \beta_{\pi}}{n_{12}(\mathbf{Z}) + \alpha_{\pi}} & i  = n_1 + 1 \\
 \end{cases}\]
@@
 To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that two matches.
 In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.

### 2021-10-10 methods_paper/fabl-paper.tex
minor change to abstract. great job bk.

@@
 
-{
-	\color{blue} Thanks Beka for writing that first abstract. I think this one captures the idea better, even if its not exactly in "outline" form.
-	\vspace{1em}
-}
-
-In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
+%{
+%	\color{blue} Thanks Beka for writing that first abstract. I think this one captures the idea better, even if its not exactly in "outline" form.
+%	\vspace{1em}
+%}
+
+In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
 Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
-This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}.
+This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.
+In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}.
 Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
+Finally, we discuss our work and future directions.
 

### 2021-10-10 methods_paper/fabl-paper.tex
comments on section 1.

@@
 \noindent%
-{\it Keywords:} record linkage, bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
+{\it Keywords:} bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
 
@@
 
-	I have some proposed revisions for the first part of the paper.
+%	I have some proposed revisions for the first part of the paper.
 In the introduction, I reordered things and changed some topic sentences to highlight that the main contribution here is speed and scale.
-
-	Also, in the prior version, we didn't get to any of the proposed material until Section 4.3, which did not seem reasonable to me.
-	I know you recommended splitting the review of the literature between the Introduction and Section 2 on bipartite matching, but since we needed the mathemetical overview of FS and Sadinle, I decided to move all of the literature review into the introduction.
+%	
+%	Also, in the prior version, we didn't get to any of the proposed material until Section 4.3, which did not seem reasonable to me.
+%	I know you recommended splitting the review of the literature between the Introduction and Section 2 on bipartite matching, but since we needed the mathemetical overview of FS and Sadinle, I decided to move all of the literature review into the introduction.
 I've seen this done in many papers.
-
-	Almost all of the content is still there, but it's moved around in a way that makes more sense to me. Let me know your thoughts! (But please avoid hard edits until we talk on Monday!)
-
-	I have some additional edits in the later part of the paper about the partial Bayes estimate, computational complexity, and added detail for the case study. All of that is in blue as well.
-
-	---------------------------------------------
+%	
+%	Almost all of the content is still there, but it's moved around in a way that makes more sense to me. Let me know your thoughts! (But please avoid hard edits until we talk on Monday!)
+%	
+%	I have some additional edits in the later part of the paper about the partial Bayes estimate, computational complexity, and added detail for the case study. All of that is in blue as well.
 
 	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
-This is an increasingly important task in "data cleaning," and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
-In this paper, we consider bipartite record linkage, which is the merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
+This is an increasingly important task in ``data cleaning," and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
 
@@
 
-	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a "many-to-one" model for record linkage.
-This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}.
-Throughout, we argue that even in cases where a bipartite matching is desired, this relaxed model (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
-To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
-	pronounced ``fable'')
-
-	In what follows, we first review more thoroughly the Fellegi Sunter and Beta Record Linkage methods.
+	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
+Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one" model for record linkage.
+This allows us to employ hashing techniques that hasten calculations and reduce computational costs.
+Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}.
+Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
+We illustrate our results via simulation studies and a real case study of homicides from the El Salvadoran Civil War.
+
+\textcolor{red}{BK: I would layout all the sections, section by section once they have been set for the reader and I can help you do this.}	
+	% I think this is clear and has already been said. 
+%	To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl}, pronounced ``fable'')
+
+%	In what follows, we first review more thoroughly the Fellegi Sunter and Beta Record Linkage methods.
 Then, we provide the model specification under \texttt{fabl} and derive the Gibbs sampler for posterior inference.
@@
 Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$.
-Without loss of generality, denote files suchs that $n_1 \geq n_2$.
+Without loss of generality, denote files such that $n_1 \geq n_2$.
 In the context of bipartite matching, we assume that there are duplications across, but not within, each database.

### 2021-10-10 methods_paper/fabl-paper.tex
minor edit.

@@
 
-\subsection{Beta Linkage Model}
+\subsection{Beta Record Linkage Model}
 
 One criticism of using mixture models is that the decision rule will lead to "many-to-many assignments," breaking one-to-one matching assumptions.
-One way of resolving this issue to incorporate the constraint directly into the model as follows
-
+One way of resolving this issue to incorporate the constraint directly into the model as follows:
 %our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
@@
 where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-For \texttt{BRL} \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
+For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
 $$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
@@
 
-------------------------------------------------------
-
-\textcolor{black}{I think its best just not to mention probabalistic blocking.
+%------------------------------------------------------
+%
+%\textcolor{black}{I think its best just not to mention probabalistic blocking.
 My method doesn't use it, and introducing it is really confusing (Sadinle 2017, Enamorado 2019, and Steorts 2014/2015 don't mention it).}
-
-We note here that the word "hashing" is also used in probabalistic blocking, which uses techniques from machine learning to cluster records together before attempting to link records.
+%
+%We note here that the word "hashing" is also used in probabalistic blocking, which uses techniques from machine learning to cluster records together before attempting to link records.
 The use of such probabalistic blocking to provide even greater computational gains will be explored in future work.
-
--------------------------------------------------------
+%
+%-------------------------------------------------------
 

### 2021-10-11 methods_paper/fabl-paper.tex
erased auxilliary files

@@
 
-\newpage
-{
-\color{red}
-Here is the previous version of everything. Note that the numbering of the sections is off at the moment. 
-}
-
-%% Reworking introduction for Brian's paper.
-
-Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
-Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
-Many statistical record linkage methods and case studies are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test (or hypothesis test).
-Recent work in the statistical literature has extended the aforementioned work \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.
-
-%While this independence assumption may not always be reasonable, it is appealing for its mathematical simplicity and relative computational ease.
-Many recent work in the statistical literature has extended the aforementioned work [Brian: add citations here:
-
-In this paper, we consider bipartite record linkage, which is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle_bayesian_2017}.
-Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
-
-The Fellegi-Sunter method and its extensions are widely popular mainly due to its simplicity and computational scalability.
-Despite this, like any method, it has limitations, that are well known in the literature.
-Specifically, the assumption of no duplications within a database implies a maximum one-to-one constraint.
-That is, a record from one database can be linked with only one record in the other database.
-Modern methods of Fellegi-Sunter either ignore this restriction \citep{Winkler1988, belin_1995, larsen_2001}, include this as a post-processing step \citep{jaro1989}, or include it directly in the modeling framework \citep{sadinle_bayesian_2017}.
-It is important to note that \cite{fellegi_theory_1969} ignored this restriction as well.
-In practice, it seems that one should either directly include it in the modeling framework or correct for this in a post-processing step.
-
-To our knowledge, \cite{sadinle_bayesian_2017} was the first to propose correcting the Fellegi and Sunter model for the one-to-one constraint.
-Specifically, the authors propose a Bayesian variant of the original model that satisfies two goals.
-First, it allows one to quantify uncertainty regarding match or non-match status.
-In addition, posterior distributions are readily available to help estimate point estimates of interest (and quantity such estimates).
-Second, under the Bayesian paradigm, the one-to-one matching constraint is easily embedded.
-The downside to this constraint is an added computational cost.
-The goal of our paper is proposing an extension of \cite{sadinle_bayesian_2017} that scales to large databases, but does not compromise much in terms of accuracy on the original case study considered by the author.
-
-\textbf{Our Contribution} 
-In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\texttt{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations and to reduce computational complexity.
-Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching.
-Instead, we propose independent priors (breaking the one-to-one matching) in order to create computational speeds and gains.
-Second, we resolve the one-to-one matching issue via a post processing step.
-Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model.
-Fourth, we explore other computational speeds up based upon storage efficient indexing.
-Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices.
-To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
-pronounced ``fable'')
-
-The remainder of the paper is as follows.
-Section~\ref{sec:bipartite} reviews bipartite record linkage; section \ref{sec:fs} reviews the traditional approach of the Fellegi and Sunter framework before introducing our proposed work.
-
-\section{Bipartite Record Linkage and Related Work}
-As already mentioned, bipartite record linkage is the traditional set up studied in record linkage, where we assume there are two databases $\bm{X}_1$ and $\bm{X}_2$ such that there are duplications across the databases but not within each databases.
-We assume that each database has $n_1$ and $n_2$ records, and that $n_1 \leq n_2.$ Denote the number of entities simultaneously recorded by \emph{both databases} by $n_{12}$ such that $0 \leq n_{12} \leq n_2.$ Given this set up, as explained by \cite{steorts_bayesian_2016} and \cite{sadinle_bayesian_2017}, this can be viewed as bipartite record linkage.
-This is important as it allows one to view the matching matrix or linkage structure in the following way:
-\begin{align}
-\Delta_{ij} =
-\begin{cases}
-1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-0 \quad \text{otherwise}.\\
-\end{cases}
-\end{align}
-Bipartite matching can also be viewed as a matching labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
-\begin{align}
-Z_{j} =
-\begin{cases}
-i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
-\end{cases}
-\end{align}
-Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.
-
-There are many approaches to bipartite record linkage, and we review some of the most common approaches below.
-
-\paragraph{Fellegi and Sunter}
-As already mentioned, the most common approach is the original approach of \cite{fellegi_theory_1969}.
-\cite{enamorado2019using} have extended this approach using computational speeds ups to allow for a much more scalable and practical approach, providing open source code for the community.
-
-%These approaches have led to downstream tasks, such as regression approaches in the literature \citep{tang2020}.
-
-\paragraph{Directly Modeling the Data} Instead of relying on comparison vectors, some authors will incorporate the data directly into their bipartite record linkage model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, gutman_bayesian_2013, shan2020bayesian}.
-These approaches have led to extensions regarding downstream tasks in the literature or joint models, such as regression approaches \citep{dalzell2018regression, steorts2018generalized, tancredi2020unified}.
-There have been extensions to directly model the data in this literature where authors consider merging multiple databases (more than two) \citep{steorts_bayesian_2016, steorts_entity_2015, zanella_flexible_2016, marchant_distributed_2019, tancredi_2015_regression}.
-We do not review this literature given that it is not directly/fairly comparable with the case of bipartite record linkage.
-
-\paragraph{Classification Approaches} 
-One common way of approaching the record linkage task, especially in computer science and machine learning is using a classification approach.
-One uses a model (such as random forests) to classify records into matches/non-matches.
-When access to unique identifiers is available, one can create a training data (or reference data) set of record pairs (where the true match status is known).
-Finally, one can train a classifier using the reference data set to predict the match status on any record pairs that were not included in the reference data set \citep{Bilenko2006, Christen2008, ventura2015seeing}.
-
-\section{Review of the Fellegi Sunter Method}
-In this section, we review notation, terminology, and the classical approach of Fellegi and Sunter.
-
-Consider ordered record pairs $\bm{X}_1 \times \bm{X}_2,$ where the size of the databases are $n_1$ and $n_2,$ respectively. 
-Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$ Denote the set of non-matches 
-$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ An alternative way of merging two databases and removing 
-duplications can be viewed as identifying the sets of matches/non-matches, or rather $\bm{M}, \bm{U}.$ In the literature, it is well known that record pairs that are estimated as matches are known as links; record pairs that are estimated as non-matches are known as non-links.
-\cite{fellegi_theory_1969} proposed using all-to-all pairwise comparison of records to estimate the matches/non-matches.
-
-\subsection{Comparison Data}
-In this section, we review the notion of \emph{comparison data}.
-Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar.
-A common way to encode this is using a comparison vector $\bm{\gamma}_{ij}$ which is computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote $F$ as the criteria for compare the records such that
-$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $F$ represents to a single comparison per feature that the databases share, such as gender.
-%
-The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for nominal variables such as gender.
-For more complex measurements, such as textual or numeric feature information, we can take into account partial agreement patterns using common string metrics, such as the Edit, Jaro, or Jaro-Winkler distance functions.
-This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
-
-Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$ representing disagreement levels.
-Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
-Thus, we can construct comparison vectors in the following way:
-$$\gamma_{ij} = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$ The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.
-
-\subsection{Blocking}
-In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, clusters, or ``blocks" to reduces this computational burden.
-Blocking can be deterministic, such as reducing the dimensionality of the comparison space on a highly reliable feature, such a gender.
-Blocking can be probabilistic, such as utilizing machine learning methods, known as hashing, which seek to use probabilistic functions (with low computational cost) to place similar records into the same bin.
-We propose hashing base methods for Bayesian bipartite record linkage, illustrating their ability to improve upon computational scalability.
-We emphasize that depending on the size of the problem, our proposed methodology can be used with or without hashing.
-
-\subsection{Fellegi-Sunter Model}
-In this section, we review the Fellegi-Sunter model.
-While we have introduced comparison data, $\bm{\gamma}_{ij}$, this is not sufficient to determine whether a record is a match/non-match given errors that naturally occur in the data.
-This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
-\begin{align}
-\label{eqn:wts}
-w_{ij} = \log \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
-\end{align}
-as a weight to estimate if record pairs are a match/non-match.
-Assume $\bm{\Gamma}_{ij}$ is a random vector, whose distribution depends on $\Delta_{ij}.$ Then equation \ref{eqn:wts} is a representative value of $\bm{\Gamma}_{ij} \mid \Delta_{ij}.$ The ratio will be large if we favor the pair being a match.
-This is a problem in practice as transitive closures are violated.
-For example, suppose records $i$ and $j$ in $\bm{X}_1$ are extremely similar but are known to be non-matches.
-Suppose that both records $i$ and $j$ in $\bm{X}_1$ are very similar to record $k$ in $\bm{X}_2.$ By transitive closures, this implies that $i$ and $j$ are a match, which cannot be true by assumption.
-
-\cite{jaro1989} proposed addressing this issue via an optimization approach as follows:
-\begin{align}
-\label{eqn:jaro}
-&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
-\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
-& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
-& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
-\end{align}
-
-Equation \ref{eqn:jaro} ensures that each record of $\bm{X}_2$ only matches with at most one record in $\bm{X}_1$ (and the reverse condition).
-This is commonly known as the \emph{maximum-weight bipartite matching problem} or a \emph{linear sum assignment problem}.
-The output of this algorithm is a bipartite matching that maximizes the sum of the weights among matched pairs, where the pairs are not matched are assumed to be non-matches.
-\cite{jaro1989} provided no theoretical justification for using this approach, however, \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
-
-\subsection{Model Estimation}
-In our review, we have assumed one knows $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0),$ however, these are typically not known in practice and must be estimated.
-Assume the comparison vectors are independent given the bipartite matching and assume that the match status of the record pairs are independent.
-It then follows that we can model
-the comparison data using mixture models in the following way:
-\begin{align}
-&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
-&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
-& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
-\end{align}
-where $p$ is the proportion of records that match.
-Note that $\bm{m}$ and $\bm{u}$ are vectors of the match/non-match parameters.
-One common way to estimate the unknown parameters of the aforementioned model is using the EM algorithm.
-
-One criticism of using mixture models as proposed by \cite{jaro1989} is the fact that the decision rule will lead to ``many-to-many assignments (breaking transitive closures).
-One way of resolving this issue to incorporate the constraint directly into the model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, Larsen05, Larsen12, gutman_bayesian_2013, sadinle_bayesian_2017}, which is appealing as one then obtains uncertainty quantification under the Bayesian paradigm, but often has to sacrifice scaling to very large databases.
-On the other hand, one can turn to post-processing methods in practical situations.
-In this paper, we attempt to bridge the best of both approaches.
-Specifically, we propose a model in the spirit of \cite{sadinle_bayesian_2017} and \cite{enamorado_using_2019}.
-We relax the one-to-one matching constraint proposed by  \cite{sadinle_bayesian_2017} for purely computational reasons, and resolve this in a post-processing step.
-In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}.
-Our framework is then embarassingly parallel, available in the \texttt{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.
-
-\section{Proposed Methodology}
-\label{sec:model}
-
-In this section, we propose the fast beta linkage method (fabl) for bipartite record linkage, extending the work of \cite{sadinle_bayesian_2017}.
-Specifically, we review the general approach of \cite{sadinle_bayesian_2017} and then provide our extensions for scalability.
-
-%Specifically, the author used the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures.
-In each iteration of his Gibbs sampler, the author considered each record \(j\in \bm{X}_2\), and then removed from consideration the records \(i\in \bm{X}_1\) that have already been matched.
-Then a potential link is sampled.
-While this leads to a sampler that is more accurate than the classical Fellegi-Sunter approach, accounting for such dependencies throughout the linkage process is computationally burdensome, and this method is only suitable for small databases.
-%As such, we propose a framework for efficient computation in large linkage tasks.
-
-\subsection{Beta Linkage Model}
-In this section, we review the Beta Linkage Model, which can be written as follows: 
-%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
-\begin{align}
-&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
-&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
-& \bm{Z} \sim \mathcal{B}, \notag
-\end{align}
-where $\mathcal{B}$ represents a prior on the space of bipartite matchings and $\mathcal{M}(\bm{m}), \mathcal{M}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-
-\hypertarget{model-specification}{%
-	\subsection{New Model Specification}
-	\label{model-specification}}
-
-In this section, we provide the specification for the models of $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u}),$ which follows from assuming that the features are conditionally independent of the level of agreement of another.
-The likelihood can be written as follows:
-
-\begin{align}
-\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}, 
-\end{align}
-where $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denotes the probability of a match have level $\ell$ of disagreement in feature $f$ and 
-$u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represents the same probability for non-matches. 
-In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
-$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
-
-Next, we allow for missing values or rather missing comparisons for corresponding record pairs.
-As commonly done in the record linkage literature, we assume that these occur missing at random [CITE: Little and Rubin (2002), Sadinle (2017), Steorts et al (2016)].
-
-The likelihood becomes 
-\begin{align}
-\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}^{obs}) = 
-\prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{a_{f\ell}(\bm{Z})}u_{fl}^{a_{f\ell}(\bm{Z})}\right], 
-\end{align}
-where 
-\begin{align}
-&a_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell) I(Z_j = i) \\
-&b_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = \ell) I(Z_j \neq i),
-\end{align}
-where $I_{obs}$ denotes if a value is observed or not. Next, for a particular matching configuration $\bm{Z}$, $a_{f\ell}(\bm{Z})$
-and $b_{f\ell}(\bm{Z})$ denote the number of matches and non-matches with observed disagreement level $\ell$ for comparison feature $f.$ Using this parameterization, we specify independent conjugate priors for all features $f$:
-\begin{align}
-&\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
-&\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).
-\end{align}
-
-\hypertarget{fast-beta-prior}{%
-	\subsection{Fast Beta Prior}
-	\label{fast-beta-prior}}
-
-In this section, we deviate from the approach of \cite{sadinle_bayesian_2017}, and propose a more scalable prior on the matching labelings $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2}),$ where recall $Z_j \in \{1,2, 3, n_1, n_1 + j\}.$ While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.
-
-\[Z_j | \lambda =
-\begin{cases} 
-	\frac{1}{n_1}\lambda  & z_j \leq n_1; \\
-	1-\lambda &  z_j  = n_1 + j \\
-\end{cases}\]
-
-\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \]
-
-Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
-$\lambda$ itself is given a prior distribution to reflect prior beliefs about the overall rate of matchingness in the data.
-One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
-
-There is a clear relationship between our proposed model and that of \cite{sadinle_bayesian_2017}. 
-\cite{sadinle_bayesian_2017} constructs a prior distribution on the entire \(\mathbf{Z}\) vector, called the \emph{beta distribution for bipartite matching}, given by
-$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
-$$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior induces a Gibbs sampler that strictly enforces one-to-one matching,  previously matched records from the set of candidate
-records when sampling \(Z_j\). This creates a dependency that makes the
-sampler \emph{inherently serial}.
-
-On the other hand, we propose independent priors for each \(Z_j\), creating a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains.
-More importantly, since only the agreement pattern of \(Z_j\) is used for
-calculations within the Gibbs sampler, and not the particular record
-label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.
-In doing so however, we thereby weaken the one-to-one requirement from
-\texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be
-matched to at most one record in \(\bm{X}_1\), but allows for the possibility
-that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-
-\newpage
-
-%\textcolor{red}{I think my notation is OK from earlier.}
-
-%Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
-%respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
-%in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
-%generality, label the files such that \(n_A \geq n_B\). We assume there
-%are no duplicates within files, only across. For each record pair under
-%consideration, we generate a comparison vector
-%\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
-%where \(F\) is the number of fields used in the linkage and each
-%\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
-%indicating the level agreement between the two records on a specified
-%field.
-
-%To indicate matching status, we adopt the \emph{linkage structure
-%	parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_B})\) from Sadinle 2017,
-%defined as \[Z_j=\begin{cases} 
-%	i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
-%	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
-%\end{cases}\] This parameter is vector of length $n_B$, providing more memory efficient storage for the
-%linkage information than a \(n_A \times n_B\) sparse matrix of
-%indicators.
-%
-%\textcolor{red}{Brian: This does not match what Sadinle has, so this needs to be justified regarding the change or if this is a typo. Could you please check this?} 
-%
-%
-%\textcolor{blue}{Define \(m^{f\ell}:= P(\Gamma_{ij}^f = l |Z_j = i)\) as  the probability of a match having level $\ell$ of disagreement for feature $f.$ Define \(u^{f\ell}:= P(\gamma_{ij}^f = l |Z_j \neq i)\) as the corresponding probability for non-matches.
-Next, we adopt Fellegi and Sunter's conditionally independent fields assumption
-%that the level of agreement on one feature is independent of the level of
-%agreement on another.\footnote{Though this assumption is often not reasonable
-%(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
-%but it is common within the record linkage literature and generally
-%leads to models that perform well in practice; see discussion for
-%further remarks.} Denote \(\lambda\) to be the (marginal)
-%probability that some record \(j \in \bm{X}_2\) has a match in \(\bm{X}_1\). In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
-%$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$}.
-%
-%\begin{enumerate}
-%\item Let's first write out the likelihood, which should be the same as in Sadinle
-%\item Then we should write out the Beta Prior, which should deviate and needs its own section in my opinion so the reader understands why it's a deviation
-%\item My understanding, is that section 4.1 and section 4.2 should look as they are, so these just need to be explained to the reader.
-The innovation is in the beta prior and making it more efficient.
-%\end{enumerate}
-
-%Following the Fellegi Sunter framework, we define
-%\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
-%observing agreement level \(l\) in field \(f\) for records \(i\) and
-%\(j\) given that the records are a match, and similarly define
-%\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We also
-%adopt Fellegi and Sunter's conditionally independent fields assumption
-%that the level of agreement on one field is independent of the level of
-%agreement on another. Though this assumption is often not reasonable
-%(for example, two individuals of a different genders are more likely to have different names than two individuals of the same gender),
-%but it is common within the record linkage literature and generally
-%leads to models that perform well in practice; see discussion for
-%further remarks. Lastly, we define \(\lambda\) to be the (marginal)
-%probability that some record \(j \in B\) has a match in \(A\).
-
-%Wherever possible, we reserve superscripts for denoting field and level,
-%while reserving subscripts for record indices. For example,
-%\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
-%distribution governing field \(f\) for matching records, and
-%\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
-%is product of the relevant of the appropriate \(\mathbf{m}\) parameters
-%for record pair \((i,j)\). We hope that these conventions avoid
-%overloaded notation in the likelihood and subsequent derivations.
-
-%\hypertarget{model-specification}{%
-%	\subsection{Model Specification}
-%	\label{model-specification}}
-%
-%For fields \(f \in \{1, \ldots, F\}\) and levels
-%\(l\in \{1, \ldots, L_f\}\) we adopt the following likelihood and prior
-%distributions.
-
-%$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$
-
-%\textcolor{red}{It seems to me that this should be $P(\bm{Z}, \Phi \mid \bm{\Gamma})$ above and not what is currently written. }
-%
-%\textcolor{red}{Are you handling missing values the way that Sadinle does?
-This isn't clear.
-It would he helpful to follow the Sadinle paper regarding structure as closely as possible given that it's very clear.
-I might recommend that you work on some details here regarding pages 12---13 as some important details are skipped over regarding what's random and observed that may not be obvious to the AE/reviewer.}
-%
-%\[\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})\]
-%\[\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})\]
-
-\newpage
-
-\hypertarget{posterior-sampling}{%
-	\subsection{Gibbs Sampler}
-	\label{gibbs_sampling}}
-We work with the following factorization of the joint posterior distribution:
-
-\textcolor{red}{Brian: The Gibbs sampler has some minor typos. It should be written in terms of $\bm{Z}, \Phi, \lambda \mid \gamma_{obs}.$ Also, just make the rest of the notation match please:)}
-
-\begin{align*}
-p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
-&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)} \\
-&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
-\end{align*}
-This factorization leads to following Gibbs Sampler:
-
-\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
-The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
-standard multinomial-dirichlet mechanics. Thus we have
-
-\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
-\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
-where
-\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I(\gamma_{ij}^f = l) I(z_j = i)\)
-and
-\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + I(\gamma_{ij}^f = l) I(z_j \neq i)\).
-
-\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
-\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
-of successes (when \(z_j < n_A + 1\)) and failures (when
-\(z_j = n_A + 1\)), and therefore
-\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
-determined only by the number of duplicates
-$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
-$\mathbf{Z}$. Thus we have
-
-\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
-\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
-\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
-\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
-
-\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
-Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
-only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
-denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
-that as a function of \(Z_j\), the likelihood
-\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
-is a discrete distribution with probabilities proportional to
-
-\begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
-	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
-	&=
-	\begin{cases} 
-		w_{ij}  & z_j \leq n_1; \\
-		1 &  z_j  = n_1 + j \\
-	\end{cases}\\
-\end{align*}
-
-where
-\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
-The interested reader should note that these are precisely the
-likelihood ratios used in the Fellegi-Sunter model to classify matches
-and non-matches, and we therefore refer to \(w_{ij}\) as the
-\emph{Fellegi Sunter weights}.
-
-With the likelihood in this form, we can derive the full conditional
-\[p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)\]
-
-\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
-\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
-\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
-\begin{cases} 
-	\frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
-	1-\lambda &  z_j  = n_A + 1 \\
-\end{cases}\]
-
-In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
-integrate over the posterior of \(\lambda\) and rearrange terms to
-produce the final full conditional:
-
-\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
-\begin{cases} 
-	w_{ij}  & i \leq n_A; \\
-	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
-\end{cases}\]
-
-\hypertarget{bayes-estimate}{%
-	\subsection{Bayes Estimate}
-	\label{bayes-estimate}}
-
-We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
-parameter \(\mathbf{Z}\) by assigning different positive losses to
-different types of errors, and minimizing posterior expected loss. We
-adopt the loss function proposed in Sadinle 2017, in which
-\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
-option to leave the matching undetermined by the model. Specifically, we have
-
-\[L(\hat{Z_j}, Z_j)=\begin{cases} 
-	0  & \text{if } Z_j = \hat{Z_j}; \\
-	\theta_R,  & \text{if } \hat{Z_j} = R; \\
-	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
-	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
-	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
-\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
-on the linkage status, \(\theta_{10}\) is the loss from a false
-non-match, \(\theta_{01}\) is the loss from a false match, and
-\(\theta_{11}\) is the loss from the special case of a false match in
-which the record has a true match other than the one estimated by the
-model.
-
-We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$.
+%\newpage
+%
+%\hypertarget{posterior-sampling}{%
+%	\subsection{Gibbs Sampler}
+%	\label{gibbs_sampling}}
+%We work with the following factorization of the joint posterior distribution:
+%
+%\textcolor{red}{Brian: The Gibbs sampler has some minor typos. It should be written in terms of $\bm{Z}, \Phi, \lambda \mid \gamma_{obs}.$ Also, just make the rest of the notation match please:)}
+%
+%\begin{align*}
+%p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
+%&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)} \\
+%&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
+%&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
+%\end{align*}
+%This factorization leads to following Gibbs Sampler:
+%
+%\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
+%The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
+%standard multinomial-dirichlet mechanics. Thus we have
+%
+%\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
+%\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
+%where
+%\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I(\gamma_{ij}^f = l) I(z_j = i)\)
+%and
+%\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + I(\gamma_{ij}^f = l) I(z_j \neq i)\).
+%
+%\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
+%\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
+%of successes (when \(z_j < n_A + 1\)) and failures (when
+%\(z_j = n_A + 1\)), and therefore
+%\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
+%determined only by the number of duplicates
+%$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
+%$\mathbf{Z}$. Thus we have
+%
+%\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
+%\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
+%\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
+%\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
+%
+%\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
+%Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
+%only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
+%denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
+%that as a function of \(Z_j\), the likelihood
+%\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
+%is a discrete distribution with probabilities proportional to
+%
+%\begin{align*}
+%	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
+%	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
+%	&=
+%	\begin{cases} 
+%		w_{ij}  & z_j \leq n_1; \\
+%		1 &  z_j  = n_1 + j \\
+%	\end{cases}\\
+%\end{align*}
+%
+%where
+%\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
+%The interested reader should note that these are precisely the
+%likelihood ratios used in the Fellegi-Sunter model to classify matches
+%and non-matches, and we therefore refer to \(w_{ij}\) as the
+%\emph{Fellegi Sunter weights}.
+%
+%With the likelihood in this form, we can derive the full conditional
+%\[p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)\]
+%
+%\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
+%\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
+%\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
+%\begin{cases} 
+%	\frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
+%	1-\lambda &  z_j  = n_A + 1 \\
+%\end{cases}\]
+%
+%In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
+%integrate over the posterior of \(\lambda\) and rearrange terms to
+%produce the final full conditional:
+%
+%\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
+%\begin{cases} 
+%	w_{ij}  & i \leq n_A; \\
+%	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
+%\end{cases}\]
+%
+%\hypertarget{bayes-estimate}{%
+%	\subsection{Bayes Estimate}
+%	\label{bayes-estimate}}
+%
+%We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
+%parameter \(\mathbf{Z}\) by assigning different positive losses to
+%different types of errors, and minimizing posterior expected loss. We
+%adopt the loss function proposed in Sadinle 2017, in which
+%\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
+%option to leave the matching undetermined by the model. Specifically, we have
+%
+%\[L(\hat{Z_j}, Z_j)=\begin{cases} 
+%	0  & \text{if } Z_j = \hat{Z_j}; \\
+%	\theta_R,  & \text{if } \hat{Z_j} = R; \\
+%	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
+%	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
+%	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
+%\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
+%on the linkage status, \(\theta_{10}\) is the loss from a false
+%non-match, \(\theta_{01}\) is the loss from a false match, and
+%\(\theta_{11}\) is the loss from the special case of a false match in
+%which the record has a true match other than the one estimated by the
+%model. 
+%
+%We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$.
 In this paper, we adopt losses
-\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
-
-\[\hat{Z}_j =\begin{cases} 
-	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
-	0,  & \text{otherwise} ; \\
-\end{cases}\]
-
-For a more in-depth explanation of this function and the induced Bayes
-estimate, see \citep{sadinle_bayesian_2017}.
-
-Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+%\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
+%
+%\[\hat{Z}_j =\begin{cases} 
+%	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
+%	0,  & \text{otherwise} ; \\
+%\end{cases}\]
+%
+%For a more in-depth explanation of this function and the induced Bayes
+%estimate, see \citep{sadinle_bayesian_2017}.
+%
+%Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
 It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable.
 We will explore this issue more deeply through the El Salvador homicide case study.
-
-To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+%
+%To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
 In the event that two matches.

### 2021-10-11 methods_paper/fabl-paper.tex
new edits

@@
 
-%{
-%	\color{blue} Thanks Beka for writing that first abstract. I think this one captures the idea better, even if its not exactly in "outline" form.
-%	\vspace{1em}
-%}
-
 In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
@@
 
-{
-	\color{blue}
-
 %	I have some proposed revisions for the first part of the paper.
@@
 	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
-This is an increasingly important task in ``data cleaning," and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
 In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
@@
 	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one" model for record linkage.
+Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
 This allows us to employ hashing techniques that hasten calculations and reduce computational costs.
@@
 Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
-We illustrate our results via simulation studies and a real case study of homicides from the El Salvadoran Civil War.
-
-\textcolor{red}{BK: I would layout all the sections, section by section once they have been set for the reader and I can help you do this.}	
+
+	In what follows, Section \ref{sec:review-of_prior-work} reviews more thoroughly the Fellegi Sunter and Beta Record Linkage methods.
+Then, Section \ref{sec:fast-beta-linkage} provides the model specification under \texttt{fabl} and derives the Gibbs sampler for posterior inference.
+Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
+Lastly, in Sections \ref{sec:simulations} and \ref{sec:case-studies} we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
+We conclude in Section \ref{discussion} with a discussion of directions for future work.
+
 	% I think this is clear and has already been said. 
@@
 
-%	In what follows, we first review more thoroughly the Fellegi Sunter and Beta Record Linkage methods.
-Then, we provide the model specification under \texttt{fabl} and derive the Gibbs sampler for posterior inference.
-Next, we introduce the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
-Lastly, we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
+%
 
@@
 Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$, and the set of non-matches by 
-$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as "links", and record pairs that are estimated as non-matches as "non-links".
+$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.
 
@@
 \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that
-$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
+$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
 
@@
 This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
-Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$.
+Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$.
 Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
@@
 
-One criticism of using mixture models is that the decision rule will lead to "many-to-many assignments," breaking one-to-one matching assumptions.
+One criticism of using mixture models is that the decision rule will lead to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
 One way of resolving this issue to incorporate the constraint directly into the model as follows:
@@
 
-In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or "blocks" to reduces this computational burden.
+In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or ``blocks'' to reduces this computational burden.
 In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field.
@@
 In practice, the method the follows can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.
-
-%------------------------------------------------------
-%
-%\textcolor{black}{I think its best just not to mention probabalistic blocking.
-My method doesn't use it, and introducing it is really confusing (Sadinle 2017, Enamorado 2019, and Steorts 2014/2015 don't mention it).}
-%
-%We note here that the word "hashing" is also used in probabalistic blocking, which uses techniques from machine learning to cluster records together before attempting to link records.
-The use of such probabalistic blocking to provide even greater computational gains will be explored in future work.
-%
-%-------------------------------------------------------
 
@@
 Importantly, this independence means that we have relaxed one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such "many-to-one" matchings are reasonable.
+This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable.
 In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.
@@
 	\label{model-specification}}
-
-\textcolor{black}{NOTE: I really want to present the entire model in one place like this.
-I saw this for the first time in Steorts 2014, and I really liked it.
-Part of the reason it took me so long to understand Sadinle 2017 was because the model was presented over several sections}
 
@@
 Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable.
+It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable.
 We will explore this issue more deeply through the El Salvador homicide case study.
@@
 A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
-
-}
 
@@
 \label{sec:efficiency}
-In this section, we propose efficient and scalable proposals using hashing based methods that are similar in spirit to \citep{enamorado2019}, a fast and scalable implementation of the Fellegi-Sunter model.
- Specifically, one way we can improve our computational efficiency is by recognizing that record pairs contribute to posterior calculations only
-through the agreement pattern of the \(\gamma_{ij}\) vector. To make this more precise, let 
+
+To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we employ hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
+The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
+To make this more precise, let
 \(\mathcal{H}\) be the set of unique agreement patterns in the data, let
 \(P = |\mathcal{H}|\) denote the total number of unique agreement
-patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
-that this bound does not depend on \(n_1\) or \(n_2\). Prior to
+patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not depend on \(n_1\) or \(n_2\).  Prior to
 processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
-and enumerate them as follows: \(h_1, \ldots, h_P\), which are called \emph{hashed values.}
-Next, we map record pairs to their corresponding hashed value. 
-More specifically, when the record pair \((i,j)\) 
-exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
+and enumerate them \(1, \ldots, P\), which are called \emph{hashed values}.
+Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
 Whenever possible, we conduct calculations over these \(P\) agreement
@@
 
-%Some of these techniques are similar to those used by \citep{enamorado2019} to produce \texttt{fastlink}, a fast a scalable implementation of the Fellegi-Sunter model, and others have been created here for our particular context.
+%In this section, we propose efficient and scalable proposals using hashing based methods that are similar in spirit to \citep{enamorado2019}, a fast and scalable implementation of the Fellegi-Sunter model.
+% Specifically, one way we can improve our computational efficiency is by recognizing that record pairs contribute to posterior calculations only
+%through the agreement pattern of the \(\gamma_{ij}\) vector. To make this more precise, let 
+%\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
+%\(P = |\mathcal{H}|\) denote the total number of unique agreement
+%patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
+%that this bound does not depend on \(n_1\) or \(n_2\). Prior to
+%processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
+%and enumerate them as follows: \(h_1, \ldots, h_P\), which are called \emph{hashed values.}
+%Next, we map record pairs to their corresponding hashed value. 
+%More specifically, when the record pair \((i,j)\) 
+%exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
+%Whenever possible, we conduct calculations over these \(P\) agreement
+%patterns, instead of the typical  \(n_1 \times n_2\) record pairs.
 
@@
 
-In this section, we describe the hashing function we propose.
-
-First, we hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019} accomplished this efficiently through the hashing function
-$$\tilde{\gamma}_{ij} = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$
+First, we hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019using} accomplished this efficiently through the hashing function
+
+$$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$
+
 This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
-For computational ease, we then map these integers to sequential integers from $1, \ldots, P.$
-
-The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
-comparison as a vector of length \(F\), with each component
-\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). 
-To ease our computational burden, we instead use a \emph{one hot encoding} of the comparison vector. For example, if
-\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
-under the classical framework becomes
-\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
-bijective transformation that does not change the meaning of the data,
-but this representation eases calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.
-
-In the classic Fellegi Sunter framework, $\bm{\Gamma}$ is a
-\(n_1 n_2 \times F\) matrix, with each row providing the comparison
-vector for a different \((i,j)\) pair. In contrast, we do not store
-these comparison vectors themselves, but instead only the hashed value $h_p$
-corresponding to the agreement pattern of the \((i, j)\) pair. We store
-this information in a nested list $\tilde{\bm{\Gamma}}$ where the
-\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
-in \(\bm{X}_1\) that share agreement pattern \(p\) with record \(j \in \bm{X}_2\). For
-each \(p\), we also calculate
-\(H_p = \sum_{i=1}^{n_1}\sum_{j=1}^{n_2} \mathbf{1}_{(i,j) \in h_p}\),
-the total instances of agreement pattern \(p\) throughout the data, and
-also for each \(j\), we calculate
-\(H_{p_j} = \sum_{i=1}^{n_1} \mathbf{1}_{{(i,j) \in h_p}}\) the
-instances of agreement pattern \(p\) among the comparison vectors
-between record \(j \in \bm{X}_2\) and each of the \(n_1\) records in \(\bm{X}_1\).
+For computational ease, we then map these integers to sequential integers from $1, \ldots, P$ corresponding to the enumerated patterns $\mathcal{H}$.
+When the $(i, j)$ pair exhbits the $p^{th}$ pattern, we say $h(i,j) = p$.
+
+With all record pairs converted to hashed values, we can create a more compact nested list 
+$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
+where $r_{j_p} = \{i \in \bm{X_1} | h(i,j) = p\}$.
+Note that the pair $(\mathcal{H}, \mathcal{R})$ fully characterizes the comparison matrix $\Gamma$ with no loss of information.
+This representation is useful because it allows us to easily compute sufficient statistics
+$$H_{j_p} = \sum_{i} I(h(i,j) = p) = ||r_{j_p}||$$
+and
+$$H_p =  \sum_{i, j} I(h(i,j) = \sum_{j} H_{j_p}$$
+As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
+
+\hypertarget{large-data}{%
+	\subsection{Working with Large Data}\label{large-data}}
 
@@
 needed to store the comparison information. That is, instead of storing
-\(n_1 \times n_1\) comparison vectors, which are relatively long under
-either the Fellegi Sunter or our modified framework, we only store the
+\(n_1 \times n_2\) comparison vectors, we only store the
 \(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
@@
 
+%The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
+%comparison as a vector of length \(F\), with each component
+%\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). 
+%To ease our computational burden, we instead use a \emph{one hot encoding} of the comparison vector. For example, if
+%\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
+%under the classical framework becomes
+%\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
+%bijective transformation that does not change the meaning of the data,
+%but this representation eases calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.
+
+%In the classic Fellegi Sunter framework, $\bm{\Gamma}$ is a
+%\(n_1 n_2 \times F\) matrix, with each row providing the comparison
+%vector for a different \((i,j)\) pair. In contrast, we do not store
+%these comparison vectors themselves, but instead only the hashed value $h_p$
+%corresponding to the agreement pattern of the \((i, j)\) pair. We store
+%this information in a nested list $\tilde{\bm{\Gamma}}$ where the
+%\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
+%in \(\bm{X}_1\) that share agreement pattern \(p\) with record \(j \in \bm{X}_2\). For
+%each \(p\), we also calculate
+%\(H_p = \sum_{i=1}^{n_1}\sum_{j=1}^{n_2} \mathbf{1}_{(i,j) \in h_p}\),
+%the total instances of agreement pattern \(p\) throughout the data, and
+%also for each \(j\), we calculate
+%\(H_{p_j} = \sum_{i=1}^{n_1} \mathbf{1}_{{(i,j) \in h_p}}\) the
+%instances of agreement pattern \(p\) among the comparison vectors
+%between record \(j \in \bm{X}_2\) and each of the \(n_1\) records in \(\bm{X}_1\).
+
 To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI).
-In standard indexing, one decides a certain criteria that they expect all true matching pairs to satisfy.
-In addition, one decides
-a priori to label any record pairs that do not meet that criteria as
+In standard indexing, one decides a priori certain criteria that they expect all true matching pairs to satisfy,
+and labels any record pairs that do not meet that criteria as
 non-matches. For example, one might only consider pairs with a certain
@@
 only pairs with exact matching on a specified number of fields. While
-generally chosen to be be quite loose, establishing these criteria
+generally chosen to be quite loose, establishing these criteria
 requires knowledge of the problem and invites room for human error.
-To improve upon this, we propose to reduce the comparison space and reduce the storage requirements while avoiding the drawbacks of standard indexing.
-
-Observe that all records of the same
-agreement pattern have the same probability when sampling \(Z_j\).
-Therefore we know that records belonging to an \(h_p\) such that
-\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
-to be deemed a match through the Bayes estimate, even without
-considering the form of the agreement pattern itself.
-
-In SEI, rather than store all of these unlikely record labels, we choose
-to store only a small number \(R\) of them. Posterior calculations still
-attribute the appropriate weight to all records through the summary
-statistics \(H_p\), and \(H_{p_j}\). Rather than storing
+
+With SEI however, we can reduce comparison space, and its associated storage costs, while avoiding these drawbacks.
+Observe that all records of the same agreement pattern have the same probability when sampling \(Z_j\).
+Therefore we know that records $\i in r_{j_p}$  such that
+\(H_{j_p}\) is large are very unlikely to be sampled consistently enough
+to be deemed a match through the Bayes estimate.
+We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
+Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$.
+Posterior calculations still attribute the appropriate weight to all records through the summary statistics \(H_p\), and \(H_{j_p}\).
+Rather than storing
 \(n_1 \times n_2\) record labels, SEI allows us to store at most
-\(n_2 \times P \times R\) labels, regardless of how large \(n_1\) might be.
+\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
 
 Lastly, for large data, we can partition the two datasets \(\bm{X}_1\) and
-\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2m}\}\) for more
+\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more
 manageable computations. On a single machine, we can read-in data
@@
 chunk of data. With multiple cores or multiple machines, this can be
-done in parallel. Thus, the combination of hashing, SEI, and partitioning
+done in parallel.
+Summary statistics from each pairwise chunk comparison can then be easily synthesized to recover sufficient statistics for the larger linkage task.
+Thus, the combination of hashing, SEI, and partitioning
 allows us to conduct linkage tasks over much larger datasets.
@@
 
-\underline{Updating $\mathbf{m}$ and $\mathbf{u}$:} After receiving
-matching statuses from \(\mathbf{Z}\), the Sadinle method calculates
-\(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each
-field and level. This constitutes \(2 \times \sum L_f\) many summations
-over \(n_A \times n_B\) quantities, and becomes computationally
-burdensome with large data. In contrast, we recognize that each unique
-agreement pattern contributes to the posterior \(\alpha(\mathbf{Z})\)
-and \(\beta(\mathbf{Z})\) vectors in the same way. In fact, if we denote
-\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
-number of matching record pairs with agreement pattern \(p\), then the
-contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
-vector is simply \(H_p^m \times h_p\). Thus our posterior update for the
-\(\alpha\) vector is simply
-\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
-we can easily calculate \(H_p^u\), the number of nonmatching record
-pairs of agreement pattern \(p\), by subtracting the number of matching
-pairs from the total present in the data; that is
-\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
-parameter through
-\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
-that these constitute \(P\) many summations over \(n_B\) quantities, and
-thus avoid the \(n_A \times n_B\) summation from the original method.
-
-\underline{Updating $\mathbf{Z}$:} Although sampling \(Z_j\) from a the
+\underline{Sample $\Phi^{(s+1)}|\mathcal{H}, \mathcal{R}, \mathbf{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
+These are simple calculations, but become computationally burdensome when working on large linkage tasks.
+
+Instead, we use one-hot encodings of the agreement patterns $\mathcal{H}$ for more efficient calculations. Denote
+$H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the
+number of matching record pairs with agreement pattern $p$.
+We can also easily calculate the number of nonmatching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\).
+Then, if $\mathbf{\alpha_0}$ and $\mathbf{\beta_0}$ are vectors of prior parameters for the $\mathbf{m}$ and $\mathbf{u}$ distributions respectively, the posterior update becomes simply
+
+$$\alpha(\mathbf{Z}) = \mathbf{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
+$$\beta(\mathbf{Z}) = \mathbf{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$
+
+Note that these constitute \(P\) many summations over \(n_B\) quantities, and thus avoid the \(n_A \times n_B\) summation from the original method.
+
+%After receiving
+%matching statuses from \(\mathbf{Z}\), the Sadinle method calculates
+%\(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each
+%field and level. This constitutes \(2 \times \sum L_f\) many summations
+%over \(n_A \times n_B\) quantities, and becomes computationally
+%burdensome with large data. In contrast, we recognize that each unique
+%agreement pattern contributes to the posterior \(\alpha(\mathbf{Z})\)
+%and \(\beta(\mathbf{Z})\) vectors in the same way. In fact, if we denote
+%\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
+%number of matching record pairs with agreement pattern \(p\), then the
+%contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
+%vector is simply \(H_p^m \times h_p\). Thus our posterior update for the
+%\(\alpha\) vector is simply
+%\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
+%we can easily calculate \(H_p^u\), the number of nonmatching record
+%pairs of agreement pattern \(p\), by subtracting the number of matching
+%pairs from the total present in the data; that is
+%\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
+%parameter through
+%\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
+%that these constitute \(P\) many summations over \(n_B\) quantities, and
+%thus avoid the \(n_A \times n_B\) summation from the original method.
+
+\underline{Sample $\mathbf{Z}^{(s+1)}|\mathcal{H}, \mathcal{R}, \Phi^{(s+1)}$:} Although sampling \(Z_j\) from a the
 full conditional provided earlier is conceptually straightforward, it

### 2021-10-11 methods_paper/fabl-paper.tex
update.

@@
 
-\begin{figure}[h!]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace}
-
-	}
-
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
 \caption{Traceplot for number of matches found across datasets in El Salvador case study}\label{fig:overlap_trace}
+\end{center}
 \end{figure}
 
-\begin{figure}[h!]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace}
-
-	}
-
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
 \caption{Traceplot for m parameter in El Salvador case study}\label{fig:m_trace}
+\end{center}
 \end{figure}
 
-\begin{figure}[h!]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace}
-
-	}
-
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
 \caption{Traceplot for u parameter in El Salvador case study}\label{fig:u_trace}
+\end{center}
 \end{figure}
 
+\newpage
 \bigskip

### 2021-10-11 methods_paper/fabl-paper.tex
update.

@@
 different types of errors, and minimizing posterior expected loss. We
-adopt the loss function proposed in Sadinle 2017, in which
+adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
 \(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
@@
 
-In this section, we replicate a simulation study from \cite{sadinle2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
+In this section, we replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
 We use first name, last name, age, and occupation for this linkage, and create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package.
@@
 \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
-\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
+\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
 For each level of overlap and each level of error, we have 100 paired sets of 500 records.}

### 2021-10-11 methods_paper/fabl-paper.tex
edited Efficiency section

@@
 The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-To make this more precise, let
-\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
-\(P = |\mathcal{H}|\) denote the total number of unique agreement
-patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not depend on \(n_1\) or \(n_2\).  Prior to
-processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
-and enumerate them \(1, \ldots, P\), which are called \emph{hashed values}.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
+To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+Here,
+$P = |\mathcal{P}|$ is the total number of unique agreement
+patterns; observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not depend on \(n_1\) or \(n_2\).
+In this context, the integers $1, \ldots, P$ serve as \emph{hashed values} that encode the same information as the orignial vectors themselves.
 Whenever possible, we conduct calculations over these \(P\) agreement
 patterns, instead of the typical \(n_1 \times n_2\) record pairs.
+
+Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
 
@@
 This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
-For computational ease, we then map these integers to sequential integers from $1, \ldots, P$ corresponding to the enumerated patterns $\mathcal{H}$.
-When the $(i, j)$ pair exhbits the $p^{th}$ pattern, we say $h(i,j) = p$.
+For computational ease, we then map these integers to sequential integers from $1, \ldots, P$ corresponding to the enumerated patterns in $\mathcal{P}$.
+When the $(i, j)$ pair exhibits the $p^{th}$ pattern, we say $(i, j) \in h_p$.
 
@@
 $$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
-where $r_{j_p} = \{i \in \bm{X_1} | h(i,j) = p\}$.
-Note that the pair $(\mathcal{H}, \mathcal{R})$ fully characterizes the comparison matrix $\Gamma$ with no loss of information.
-This representation is useful because it allows us to easily compute sufficient statistics
-$$H_{j_p} = \sum_{i} I(h(i,j) = p) = ||r_{j_p}||$$
+where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
+$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||$$
 and
-$$H_p =  \sum_{i, j} I(h(i,j) = \sum_{j} H_{j_p}$$
+$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
 As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
+Note that $\mathcal{P}$ and $\mathcal{R}$ fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}\}$ for our efficient posterior inference.
 
@@
 
-\underline{Sample $\Phi^{(s+1)}|\mathcal{H}, \mathcal{R}, \mathbf{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
+\underline{Sample $\Phi^{(s+1)}|\tilde{\Gamma}, \mathbf{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
 These are simple calculations, but become computationally burdensome when working on large linkage tasks.
 
-Instead, we use one-hot encodings of the agreement patterns $\mathcal{H}$ for more efficient calculations. Denote
+Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote
 $H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the
@@
 
-$$\alpha(\mathbf{Z}) = \mathbf{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
-$$\beta(\mathbf{Z}) = \mathbf{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$
+$$\alpha(\mathbf{Z}) = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
+$$\beta(\mathbf{Z}) = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$
 
@@
 
-\underline{Sample $\mathbf{Z}^{(s+1)}|\mathcal{H}, \mathcal{R}, \Phi^{(s+1)}$:} Although sampling \(Z_j\) from a the
+\underline{Sample $\mathbf{Z}^{(s+1)}|\tilde{\Gamma}, \Phi^{(s+1)}$:} Although sampling \(Z_j\) from a the
 full conditional provided earlier is conceptually straightforward, it
-becomes computational burdensome when \(n_A\) is larger. The reader can
-confirm that sampling a value from a large set of unequal probabilities
-becomes difficult in most programming languages. To speed up
+becomes computational burdensome when $n_1$ is large.
+This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
+To speed up
 computation, we break this sampling step into two simpler steps. First,
@@
 
-\[P\left(h\left(Z_j^{(s+1)}\right) = p | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
+$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \mathbf{Z^{(s)}}\right) \propto
 \begin{cases} 
-	w_{p}\times H_{p_j}  & p \leq P; \\
+	w_{p}\times H_{j_p}  & p \leq P; \\
 	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
-\end{cases}\]
-\[P\left(Z_j^{(s+1)} = i\bigg|\;\; h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
-	\frac{1}{H_{p_j}} & (i, j) \in h_p \\
+\end{cases}$$
+$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) \begin{cases} 
+	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
 	0 & \text{otherwise} \\
-\end{cases}\]
+\end{cases}$$
 
@@
 level of the \(P\) agreement patterns with the first equation above.
-After, we can back-fill the records corresponding to the agreement
-patterns through the second equation. Sampling uniformly is
-computationally simple even for large sets of candidate records, but this
-step can also be parallelized when working with large data.
+After all Gibbs iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.
+
+We emphasize the computational gains of this split sampler: the first step is a sample from $P$ many options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
+These changes in themselves drastically improves the speed of the sampler, and each can be parallelized if desired for additional computational gains.
 
@@
 The computational complexity of all pairwise comparisons across files $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$.
-\textcolor{blue}{TALK TO BEKA.
-I CAN DO THE SAME TRICKS AS TED HERE TO REDUCE COMPLEXITY}.
 Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
@@
 
-In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of documented identifiable deaths (DID).
-Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-This case study also considers the ramifications of
-independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
-
-%The second case study is the National Long Term Care Survey (NLTCS), a much larger linkage tasks which \texttt{BRL} is unable to complete but which we complete with \texttt{fabl} with relative ease.
+In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in each dataset.
+Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
+This case study also considers the ramifications of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
 
@@
 \label{el_salvador}
-
-{
-\color{blue} For Beka: You wrote that you wanted more detail here, and I added some of it (with some citations).
-I believe I have all the important informatoin, just more concise.
-If I included Sadinle's level of detail, I worried about plagarism.
-\vspace{1em}
-}
 
@@
 
-{
-\color{blue}
 Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets.
@@
 We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
-}
 
@@
 		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes} 
-		\caption{Posterior distribution of overlap across the two files.
-The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}
+		\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
 		\label{fig:overlap-plot}
@@
 \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
+\caption{Posterior estimates of m parameters with 95\% credible intervals. We note they are quite similar across the two methods. }\label{fig:m-and-u}
 \label{fig:m-and-u}
@@
 
-%In this section, we investigate violations of one-to-one matching of our method as compared to \texttt{BRL}.
-{
-\color{blue}
 As noted previously, the application exhibits many situations in which bipartite matching is difficult.
@@
 multiple plausible matches in \(\bm{X}_2\).
-}
 
@@
 
-%\begin{figure}[t]
-%	
-%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes} 
-%		
-%	}
-%	
-%	\caption{Posterior distribution of overlap across the two files.
-The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:overlap-plot}
-%\end{figure}
-
-%Overall, both methods present similar results. \texttt{fabl} yields
-%an initial Bayes estimate of total 195 DID. After
-%resolving matches that violated one-to-one requirements,  we estimate 180 DID. 
-%
-%
-%
-%This is very close to the 181 DID found
-%by \texttt{BRL}. The main reason for this discrepancy is the difference
-%in how each model handles situations in which one file in \(\bm{X}_1\) has
-%multiple plausible matches in \(\bm{X}_2\).
-
-%
-%\begin{figure}[t]
-%	
-%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records} 
-%		
-%	}
-%	
-%	\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
-%\end{figure}
-
-Figure~\ref{fig:rosa-maria} illustrates an example. Note that these
+Figure~\ref{Tab:rosa-maria} illustrates an example. Note that these
 records present a near violation of our assumption that there are no
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records} 
-\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
-\end{center}
-\end{figure}
+\begin{table}
+	\centering
+	\begin{tabular}[h!]{llllllll}
+		\hline
+		Record Number & dataset & lastname & firstname & day & month & year & dept \\
+		\hline
+		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
+		\hline
+		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
+		\hline
+		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & NA\\
+		\hline
+	\end{tabular}
+	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.} \label{Tab:rosa_maria}
+\end{table}
 
@@
 \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
+\caption{Gibbs sampling in situation with multiple plausible matches.
+Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability.
+Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
 \end{center}
@@
 
-\bigskip
+\clearpage
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[h]
 \begin{center}
@@
 
-\newpage
+\clearpage
+
 \bigskip

### 2021-10-12 methods_paper/fabl-paper.tex
added traceplots from one simulation

@@
 Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
+Open source software to use \texttt{fabl} in \texttt{R} is available through Github.
 
@@
 $$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
-As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
-Note that $\mathcal{P}$ and $\mathcal{R}$ fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}\}$ for our efficient posterior inference.
+For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{j_p}\}\right\}$As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
+Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for our efficient posterior inference.
 
@@
 Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$.
-Posterior calculations still attribute the appropriate weight to all records through the summary statistics \(H_p\), and \(H_{j_p}\).
 Rather than storing
@@
 \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
+Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
+For simplicty, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
 
@@
 	\begin{center}
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
+		\caption{Representative traceplot of overlap between files from simulation study in \ref{accuracy}}\label{fig:sim_overlap_trace}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
+		\caption{Representative traceplot of m parameter from simulation study in \ref{accuracy}}\label{fig:sim_m_trace}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
+		\caption{Representative traceplot of u parameters from simulation study in \ref{accuracy}}\label{fig:sim_u_trace}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
 \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 

### 2021-10-12 methods_paper/fabl-paper.tex
moved bk computational complexity to proof.

@@
 \newtheorem*{definition}{Definition}
+\newtheorem*{lemma}{Lemma}
 
@@
 	\subsection{Computational Complexity}\label{computational-complexity}}
-
-{
-\color{blue}
-The computational complexity of \texttt{fabl} is determined by its two steps: constructing comparison vectors and Gibbs sampling.
+In this section, we provide the computational complexity of \texttt{fabl} in Lemma \ref{lemma:fabl}.
+
+\newpage
+\begin{lemma}
+The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
+\end{lemma}
+
+\begin{proof}
+To prove the computational complexity, we consider two steps --- constructing the comparison vectors and the Gibbs sampler. 
+%
+The computational complexity of all pairwise comparisons across the two databases $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$.
+Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.\footnote{Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
+However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the overall complexity to $O(FP)$, where $F$ is the total number of features and $P$ is the total number of agreement patterns.}
+The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
+With hashing, we split this sampling into two steps.
+First we sample that agreement pattern of the match with complexity $O(n_2 P).$ Next,  we sample the record that exhibits that pattern with complexity $O(n_2)$.
+Thus we have reduced the complexity of the Gibbs sampler (under \texttt{BRL})from $O(F n_1 n_2)$ to $O(n_2 P)$.
+In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
+\end{proof}
+
+%{
+%\color{blue}
+%The computational complexity of \texttt{fabl} is determined by its two steps: constructing comparison vectors and Gibbs sampling.
 The computational complexity of all pairwise comparisons across files $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$.
 Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
-
-Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
+%
+%Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
 However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the complexity to $O(FP)$.
-
-The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
+%
+%The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
 With hashing, we split this sampling into two steps; first we sample that agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record that exhibits that pattern with complexity $O(n_2)$.
 Thus we have reduced the complexity of the Gibbs sampler under \texttt{BRL} from $O(F n_1 n_2)$ to $O(n_2 P)$.
-
-Overall, the computational complexity under the complete \texttt{fabl} framework is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
-}
+%
+%Overall, the computational complexity under the complete \texttt{fabl} framework is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
+%}
 

### 2021-10-12 methods_paper/fabl-paper.tex
formatting appendix

--- 2021-10-12 methods_paper/fabl-paper.tex
@@
 
-To aid the reader, we provide summary of the \texttt{fabl} method through pseudocode:
-
-\begin{algorithm}[h!]
-	\caption{Summary of fabl algorithm}
-	\begin{algorithmic}[1]
-
-		\Procedure{Hashing and Preprocessing}{}
-		\State Partition files $A$ and $B$ into chunks $\{A_I\}, \{B_J\}$
-		\For{each $I$, $J$}
-		\State Create comparison vectors between $A_I$ and $B_J$
-		\State Hash results and calculate summary statistics
-		\State Use SEI to reduce memory usage
-		\EndFor
-		\State Synthesize results across pairings
-		\EndProcedure
-
-		\Procedure{Gibbs Sampling}{}
-		\State Initialize $m$, $u$, and $Z$ parameters
-
-		\For{$t \in \{1, \ldots, T\}$} 
-		\State Sample $m^{t+1}|Z^{t}, \Gamma$ and  $u^{t+1}|Z^{t}, \Gamma$ 
-		\State Sample $H\left(Z^{t+1}\right)|m^{t+1}, u^{t+1}, \Gamma$  \Comment{Sample agreement pattern, not record}
-		\EndFor
-		\State Sample $Z | H(Z), \Gamma$ \Comment{Fills in record label based on agreement pattern}
-
-		\EndProcedure
-
-	\end{algorithmic}
-\end{algorithm}
+To aid the reader, we provide summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
 
@@
 We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
+Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}.
 
@@
 vectors. The estimates of the \(\mathbf{m}\) parameters under each method are very similar, as shown in
-Figure~\ref{fig:m-and-u}.
+Figure~\ref{fig:m-and-u}. Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}
 
@@
 
+\hypertarget{app:fabl-psuedocode}{%
+	\subsection{Summary of Fast Beta Linkage Method}\label{app:fabl-psuedocode}}
+
+\begin{algorithm}[h!]
+	\begin{algorithmic}[1]
+
+		\Procedure{Hashing and Preprocessing}{}
+		\State Partition files $A$ and $B$ into chunks $\{A_I\}, \{B_J\}$
+		\For{each $I$, $J$}
+		\State Create comparison vectors between $A_I$ and $B_J$
+		\State Hash results and calculate summary statistics
+		\State Use SEI to reduce memory usage
+		\EndFor
+		\State Synthesize results across pairings
+		\EndProcedure
+
+		\Procedure{Gibbs Sampling}{}
+		\State Initialize $m$, $u$, and $Z$ parameters
+
+		\For{$t \in \{1, \ldots, T\}$} 
+		\State Sample $m^{t+1}|Z^{t}, \Gamma$ and  $u^{t+1}|Z^{t}, \Gamma$ 
+		\State Sample $H\left(Z^{t+1}\right)|m^{t+1}, u^{t+1}, \Gamma$  \Comment{Sample agreement pattern, not record}
+		\EndFor
+		\State Sample $Z | H(Z), \Gamma$ \Comment{Fills in record label based on agreement pattern}
+
+		\EndProcedure
+
+	\end{algorithmic}
+\end{algorithm}
+
+\hypertarget{appendix-sim}{%
+	\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
+
 \begin{figure}[h]
@@
 
+\hypertarget{appendix-es}{%
+	\subsection{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
+
 \begin{figure}[h]

### 2021-10-12 methods_paper/fabl-paper.tex
several edits

@@
 In this section, we provide the computational complexity of \texttt{fabl} in Lemma \ref{lemma:fabl}.
-
-\newpage
+Let $L^{*} = \sum_{f=1}^F L_f$ denote the total number of levels used across all features in the linkage task.
+
 \begin{lemma}
@@
 The computational complexity of all pairwise comparisons across the two databases $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$.
-Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.\footnote{Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the overall complexity to $O(FP)$, where $F$ is the total number of features and $P$ is the total number of agreement patterns.}
-The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
+The hashing procedure for all pairwise comparisons is also of complexity $O(F n_1 n_2)$.
+Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
+There are then minimal computational costs associated with synthesizing summary statistics across these partitions.
+
+Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
+However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the overall complexity to $O(L^{*}P)$, where $L^{*} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task and $P$ is the total number of agreement patterns.The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
 With hashing, we split this sampling into two steps.
 First we sample that agreement pattern of the match with complexity $O(n_2 P).$ Next, we sample the record that exhibits that pattern with complexity $O(n_2)$.
-Thus we have reduced the complexity of the Gibbs sampler (under \texttt{BRL})from $O(F n_1 n_2)$ to $O(n_2 P)$.
+Since $L^{*} << n_1$ in all reasonable applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
 In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
@@
 
-In this section, we replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
-We use first name, last name, age, and occupation for this linkage, and create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package.
-See Table~\ref{Tab:sadinle_simulation_cutoffs} for details on the construction of comparison vectors.
+We demonstrate the accuracy, speed, and scale of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
+
+%Sections \ref{accuracy} and \ref{partial} utilize simulated datasets from \cite{sadinle_bayesian_2017} made available by the author to show the similarity of both methods across several accuracy metrics.
+These same datasets are used in Section \ref{scale} to create a larger linkage test to demonstrate the scale of problems \texttt{fabl} can undertake.
+Finally, Section \ref{speed} demonstrates the computational gains.
+
+\hypertarget{accuracy}{%
+	\subsection{Accuracy under Full Estimates}\label{accuracy}}
+
+We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
+We use first name, last name, age, and occupation for this linkage, where age and occupation take on 9 and 10 unique values respectively.
+We creage comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table~\ref{Tab:sadinle_simulation_cutoffs}.
 Each simulation identifies duplicated individuals between two datasets, each with 500 records.
@@
 
-\hypertarget{accuracy}{%
-	\subsection{Precision, Recall, and F-measure}\label{accuracy}}
-
-In this section, we compare \texttt{fabl} to 	\texttt{BRL} in terms of precision, recall, and f-measure, which are commonly defined record linkage evaluation metrics that are defined in \citep{christen_2012}.
-%
-In cases when there are only one or two errors in matching records, and in cases with low to moderate duplication across records, we see that \texttt{fabl} provides approximately equivalent accuracy as \texttt{BRL}.
-We find that our method only has weakened performance in the
-most extreme scenario of very high errors and very high overlap across
-files.
+We compare \texttt{fabl} to \texttt{BRL} in terms of precision, recall, and f-measure, which are commonly defined record linkage evaluation metrics that are defined in \citep{christen_2012}.
+At all levels of error and overlap, we see that the two methods have comparable performance.
+In the specific case of high error and high overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, since both methods exhibit high variance in performance in this setting, we remain confident in \texttt{fabl}'s overal accuracy.
+
+%In cases when there are only one or two errors in matching records, and in cases with low to moderate duplication across records, we see that \texttt{fabl} provides approximately equivalent accuracy as \texttt{BRL}.
+%We find that our method only has weakened performance in the
+%most extreme scenario of very high errors and very high overlap across
+%files.
 In these situations, \texttt{BRL} is removing large numbers of records from consideration throughout the Gibbs Sampler, making its implementation most different from \texttt{fabl}.
@@
 
-Lastly, we review the performance of \texttt{fabl} and \texttt{BRL} using loss functions that allow partial estimates of the linkage parameter.
 By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
@@
 municipality, and day, month, and year of death. We again use flat
-priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters. To our knowledge, we have followed the same guidelines/settings in \cite{sadinle_bayesian_2017}.
+priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters.
+
 \begin{table}
@@
 	First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (5, 1]\\
-	\hline
 	Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
-	\hline
 	Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
-	\hline
 	Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
-	\hline
 	Municipality & Binary & Agree & Disagree &  & \\
@@
 
-\vspace{1em}
-
 \begin{table}
@@
 		First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, 1]\\
-		\hline
 		Year of Death & Binary & Agree & Disagree & \\
-		\hline
 		Month of Death & Binary & Agree & Disagree & \\
-		\hline
 		Day of Death & Absolute Difference & 0 & 1 & 2+\\
-		\hline
 		Municipality & Binary & Agree & Disagree & \\
@@
 
-To mirror the original implementation of \cite{sadinle_bayesian_2017}, we construct the comparison
+To mirror the original implementation of \cite{sadinle_bayesian_2017}, we first construct the comparison
 vectors using four levels of agreement for each field, according to the
@@
 
-\paragraph{Violations of One-to-One Matching}
+\hypertarget{violations}{%
+	\subsection{Violations of One-to-One Matching}\label{violations}}
 
@@
 
-Figure~\ref{Tab:rosa-maria} illustrates an example. Note that these
+Table ~\ref{Tab:rosa-maria} illustrates an example. Note that these
 records present a near violation of our assumption that there are no
@@
 		\hline
-		Record Number & dataset & lastname & firstname & day & month & year & dept \\
+		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
 		\hline
 		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
-		\hline
 		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
-		\hline
-		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & NA\\
+		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
 		\hline
 	\end{tabular}
-	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.} \label{Tab:rosa_maria}
+	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
 \end{table}
@@
 		\Procedure{Hashing and Preprocessing}{}
-		\State Partition files $A$ and $B$ into chunks $\{A_I\}, \{B_J\}$
-		\For{each $I$, $J$}
-		\State Create comparison vectors between $A_I$ and $B_J$
-		\State Hash results and calculate summary statistics
-		\State Use SEI to reduce memory usage
+		\State Construct and enumerate set of unique pattern $\mathcal{P}$ from $F$ and $\{L_f\}$
+		\State Partition files $\bm{X_1}$ and $\bm{X_1}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$
+		\For{each $n$, $m$}
+		\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$
+		\State Hash records to $\mathcal{R}_{nm}$ and calculate summary statistics $\mathcal{H}_{nm}$
+		\State Use SEI to reduce memory usage; $\mathcal{R}_{nm} \to \mathcal{R}_{nm}^{\text{SEI}}$
 		\EndFor
-		\State Synthesize results across pairings
+		\State Synthesize results across pairings to get $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$
 		\EndProcedure
@@
 		\For{$t \in \{1, \ldots, T\}$} 
-		\State Sample $m^{t+1}|Z^{t}, \Gamma$ and  $u^{t+1}|Z^{t}, \Gamma$ 
-		\State Sample $H\left(Z^{t+1}\right)|m^{t+1}, u^{t+1}, \Gamma$  \Comment{Sample agreement pattern, not record}
+		\State Sample $\Phi^{t+1}|Z^{t}, \tilde{\Gamma}$
+		\State Sample $h\left(Z^{t+1}\right)|\Phi^{t+1}, \tilde{\Gamma}$  \Comment{Sample agreement pattern, not record}
 		\EndFor
-		\State Sample $Z | H(Z), \Gamma$ \Comment{Fills in record label based on agreement pattern}
+		\State Sample $Z | h(Z), \Gamma$ \Comment{Fills in record label based on agreement pattern}
 
@@
 \end{algorithm}
+\clearpage
 
@@
 	\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
-
-\begin{figure}[h]
+Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}. It is set up with one error across the linkage fields and 50 duplicates across files.
+
+\begin{figure}[!h]
 	\begin{center}
 		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-		\caption{Representative traceplot of overlap between files from simulation study in \ref{accuracy}}\label{fig:sim_overlap_trace}
+		\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}}\label{fig:sim_overlap_trace}
 	\end{center}
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
 	\begin{center}
 		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-		\caption{Representative traceplot of m parameter from simulation study in \ref{accuracy}}\label{fig:sim_m_trace}
+		\caption{Representative traceplot of m parameter from simulation study in Section \ref{accuracy}}\label{fig:sim_m_trace}
 	\end{center}
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
 	\begin{center}
 		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-		\caption{Representative traceplot of u parameters from simulation study in \ref{accuracy}}\label{fig:sim_u_trace}
+		\caption{Representative traceplot of u parameters from simulation study in Section \ref{accuracy}}\label{fig:sim_u_trace}
 	\end{center}
@@
 
+\clearpage
+
 \hypertarget{appendix-es}{%
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
 \begin{center}

### 2021-10-13 methods_paper/fabl-paper.tex
fabl paper ready for review

@@
 
-	Many statistical record linkage methods are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test.
+	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
+Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test.
 Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.
@@
 	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice.
-In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record in the other database.
+In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record from the other.
 Here, additional declared matches are known by assumption to be false.
@@
 \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of permitted matches under the one-to-one constraint, but the method was so computationally intensive it could used to link databases of less than 100 records.
-Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (BRL), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler.
+\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records.
+Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler.
 Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
@@
 
-	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
-This allows us to employ hashing techniques that hasten calculations and reduce computational costs.
-Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}.
-Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
+	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the the Beta Record Linkage model for increased efficiency and scalability.
+Specifically, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
+This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs; (2) compute the pairwise record comparisons over large datasets via parallel computing; and (3) reduce memory costs through our proposed method of \emph{storage efficient indexing}.
+Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
 Open source software to use \texttt{fabl} in \texttt{R} is available through Github.
@@
 Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
-Lastly, in Sections \ref{sec:simulations} and \ref{sec:case-studies} we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
-We conclude in Section \ref{discussion} with a discussion of directions for future work.
+Lastly, Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War.
+Section \ref{discussion} concludes our paper with a discussion of directions for future work.
 
@@
 Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$.
-Without loss of generality, denote files such that $n_1 \geq n_2$.
+Without loss of generality, denote the files such that $n_1 \geq n_2$.
 In the context of bipartite matching, we assume that there are duplications across, but not within, each database.
@@
 \subsection{Comparison Data}
-
-Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar.
+\label{comparison-data}
+
+Intuitively, co-referent records (those that refer to the same entity) should be similar; records that are not co-referent should not be similar.
 \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that
@@
 
-The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for categorical variables.
+The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables.
 For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance.
-This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+We can then propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
 Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$.
@@
 \subsection{Fellegi-Sunter Model}
+\label{fellegi-sunter}
 The comparison data $\bm{\gamma}_{ij}$ is not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
@@
 In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
-Under the Fellegi Sunter model, we assume that comparison vectors are independent given their matching status, and that match status of the record pairs are independent, allowing us to model
+Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent.
+This allows us to model
 the comparison data using mixture models in the following way:
@@
 
-After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined).
+After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined).
 This is a problem in practice as transitive closures are violated.
@@
 The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
-\cite{jaro1989} provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.
-
-%One criticism of using mixture models as proposed by \cite{jaro1989} is the fact that the decision rule will lead to ``many-to-many assignments (breaking transitive closures).
-One way of resolving this issue to incorporate the constraint directly into the model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, Larsen05, Larsen12, gutman_bayesian_2013, sadinle_bayesian_2017}, which is appealing as one then obtains uncertainty quantification under the Bayesian paradigm, but often has to sacrifice scaling to very large databases.
-On the other hand, one can turn to post-processing methods in practical situations.
-In this paper, we attempt to bridge the best of both approaches.
-Specifically, we propose a model in the spirit of \cite{sadinle_bayesian_2017} and \cite{enamorado_using_2019}.
-We relax the one-to-one matching constraint proposed by  \cite{sadinle_bayesian_2017} for purely computational reasons, and resolve this in a post-processing step.
-In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}.
-Our framework is then embarassingly parallel, available in the \texttt{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.
+Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate for bipartite matching.
 
 \subsection{Beta Record Linkage Model}
-
-One criticism of using mixture models is that the decision rule will lead to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
+\label{BRL}
+
+One criticism of using mixture models is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
 One way of resolving this issue to incorporate the constraint directly into the model as follows:
-%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
 \begin{align}
@@
 \end{align}
-where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
+where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
 For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
-$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
+$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
 $$
@@
 
-In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or ``blocks'' to reduces this computational burden.
+In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden.
 In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field.
@@
 This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
-In practice, the method the follows can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.
+In practice, the following method can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.
 
@@
 Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
-$\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$..
-One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
-We encourage the reader to integrate out $\pi$ from the $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.
+$\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
+One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
+We encourage the reader to integrate out $\pi$ from $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.
 
 Note that linkage in this setting is conducted at the \emph{record} level, rather than at the \emph{record pair} level as in the Fellegi Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimate that proportion of record pairs that have matches.
-We find $\pi$ to be more interpretable parameter than $\lambda$, especially considering that the overwhelming majority of pairs are non matches, meaning that $\lambda$ necessarily tends towards 0 as the size of the linkage task grows.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of \emph{records} in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of \emph{record pairs} that are matches.
+We find $\pi$ to be more interpretable parameter than $\lambda$; especcially in the bipartite case, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
+necessarily tends towards 0 as the size of the linkage task grows.
 Additionally, we emphasize that while the Fellegi Sunter model makes $n_1 \times n_2$ many independent matching decisions and \texttt{BRL} makes $n_2$ many dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$ many independent matching decisions.
@@
 
-Importantly, this independence means that we have relaxed one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable.
+Importantly, this independence means that we have relaxed the one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
+This behavior may be desirable in its own right; \cite{Newcombe1962b} were interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable.
 In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.
@@
 \begin{align*}
-	p(\mathbf{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\mathbf{Z} | \pi) p(\Phi) p(\pi) \\
+	p(\bm{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\bm{Z} | \pi) p(\Phi) p(\pi) \\
 	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
@@
 
-\underline{Sample $\Phi^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
-The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
-standard multinomial-dirichlet mechanics. Thus we have
-
-\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
-\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
+\underline{Sample $\Phi^{(s+1)}|\Gamma, \bm{Z}^{(s)}$:}
+
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have
+
+$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
+$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z}))$$
 where
-\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
+\(\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
 and
-\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).
-
-\underline{Sample $\pi^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
-\(\pi\), the linkage structure parameter \(\mathbf{Z}\) is sequence
+\(\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).
+
+\underline{Sample $\pi^{(s+1)}|\bm{Z}^{(s)}$:} As a function of
+\(\pi\), the linkage structure parameter \(\bm{Z}\) is sequence
 of successes (when \(z_j < n_A + 1\)) and failures (when
 \(z_j = n_A + 1\)), and therefore
-\(p(\mathbf{Z}|\pi) = \mathcal{L}(\pi|\mathbf{Z})\) is
+\(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is
 determined only by the number of duplicates
-$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
-$\mathbf{Z}$. Thus we have
+$n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
+$\bm{Z}$. Thus we have
 
 \begin{align*}
-p(\pi | \mathbf{Z}) &\propto p(\mathbf{Z}|\pi)p(\pi) \\
-&\propto \pi^{n_{12}(\mathbf{Z})} (1-\pi)^{n_2 - (n_{12}(\mathbf{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-&\propto \pi^{n_{12}(\mathbf{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\pi} -1} \\
-\implies &\pi^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\pi})
+p(\pi | \bm{Z}) &\propto p(\bm{Z}|\pi)p(\pi) \\
+&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1} \\
+\implies &\pi^{(s+1)}|\bm{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})
 \end{align*}
 
-\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \Phi^{(s+1)}, \pi^{(s+1)}$:}
+\underline{Sample $\bm{Z}^{(s+1)}|\Gamma, \Phi^{(s+1)}, \pi^{(s+1)}$:}
 Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
@@
 
-\[\propto \left(\sum_{i=1}^{n_1}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\mathbf{1}_{z_j = i} + (1-\pi)\mathbf{1}_{z_j = n_1 + 1}\right)\]
-\[= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\mathbf{1}_{z_j = i} + (1-\pi)\mathbf{1}_{z_j = n_1 + 1} \]
+\[\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right)\]
+\[= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1} \]
 \[ \implies Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto
@@
 
-\[p\left(Z_j^{(s+1)}  = i| \Phi, \mathbf{Z^{(s)}}\right) \propto
+\[p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
 	w_{ij}  & i \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\mathbf{Z}) + \beta_{\pi}}{n_{12}(\mathbf{Z}) + \alpha_{\pi}} & i  = n_1 + 1 \\
+	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + 1 \\
 \end{cases}\]
@@
 
-We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
-parameter \(\mathbf{Z}\) by assigning different positive losses to
+We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage
+parameter \(\bm{Z}\) by assigning different positive losses to
 different types of errors, and minimizing posterior expected loss. We
@@
 
-We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$.
-In this paper, we adopt losses
-\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
+We seek to minimize the posterior expected loss, given by 
+$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs})$$. 
+In this paper, we adopt losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule
 
@@
 
-%\newpage
-%
-%\hypertarget{posterior-sampling}{%
-%	\subsection{Gibbs Sampler}
-%	\label{gibbs_sampling}}
-%We work with the following factorization of the joint posterior distribution:
-%
-%\textcolor{red}{Brian: The Gibbs sampler has some minor typos. It should be written in terms of $\bm{Z}, \Phi, \lambda \mid \gamma_{obs}.$ Also, just make the rest of the notation match please:)}
-%
-%\begin{align*}
-%p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
-%&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)} \\
-%&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-%&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
-%\end{align*}
-%This factorization leads to following Gibbs Sampler:
-%
-%\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
-%The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
-%standard multinomial-dirichlet mechanics. Thus we have
-%
-%\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
-%\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
-%where
-%\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I(\gamma_{ij}^f = l) I(z_j = i)\)
-%and
-%\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + I(\gamma_{ij}^f = l) I(z_j \neq i)\).
-%
-%\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
-%\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
-%of successes (when \(z_j < n_A + 1\)) and failures (when
-%\(z_j = n_A + 1\)), and therefore
-%\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
-%determined only by the number of duplicates
-%$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
-%$\mathbf{Z}$. Thus we have
-%
-%\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
-%\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
-%\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
-%\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
-%
-%\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
-%Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
-%only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
-%denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
-%that as a function of \(Z_j\), the likelihood
-%\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
-%is a discrete distribution with probabilities proportional to
-%
-%\begin{align*}
-%	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
-%	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
-%	&=
-%	\begin{cases} 
-%		w_{ij}  & z_j \leq n_1; \\
-%		1 &  z_j  = n_1 + j \\
-%	\end{cases}\\
-%\end{align*}
-%
-%where
-%\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
-%The interested reader should note that these are precisely the
-%likelihood ratios used in the Fellegi-Sunter model to classify matches
-%and non-matches, and we therefore refer to \(w_{ij}\) as the
-%\emph{Fellegi Sunter weights}.
-%
-%With the likelihood in this form, we can derive the full conditional
-%\[p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)\]
-%
-%\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
-%\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
-%\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
-%\begin{cases} 
-%	\frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
-%	1-\lambda &  z_j  = n_A + 1 \\
-%\end{cases}\]
-%
-%In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
-%integrate over the posterior of \(\lambda\) and rearrange terms to
-%produce the final full conditional:
-%
-%\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
-%\begin{cases} 
-%	w_{ij}  & i \leq n_A; \\
-%	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
-%\end{cases}\]
-%
-%\hypertarget{bayes-estimate}{%
-%	\subsection{Bayes Estimate}
-%	\label{bayes-estimate}}
-%
-%We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
-%parameter \(\mathbf{Z}\) by assigning different positive losses to
-%different types of errors, and minimizing posterior expected loss. We
-%adopt the loss function proposed in Sadinle 2017, in which
-%\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
-%option to leave the matching undetermined by the model. Specifically, we have
-%
-%\[L(\hat{Z_j}, Z_j)=\begin{cases} 
-%	0  & \text{if } Z_j = \hat{Z_j}; \\
-%	\theta_R,  & \text{if } \hat{Z_j} = R; \\
-%	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
-%	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
-%	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
-%\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
-%on the linkage status, \(\theta_{10}\) is the loss from a false
-%non-match, \(\theta_{01}\) is the loss from a false match, and
-%\(\theta_{11}\) is the loss from the special case of a false match in
-%which the record has a true match other than the one estimated by the
-%model. 
-%
-%We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$.
-In this paper, we adopt losses
-%\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
-%
-%\[\hat{Z}_j =\begin{cases} 
-%	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
-%	0,  & \text{otherwise} ; \\
-%\end{cases}\]
-%
-%For a more in-depth explanation of this function and the induced Bayes
-%estimate, see \citep{sadinle_bayesian_2017}.
-%
-%Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable.
-We will explore this issue more deeply through the El Salvador homicide case study.
-%
-%To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that two matches.
-In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.
-A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
-
 \section{Efficient and Scalable Implementation}
@@
 The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
 Here,
@@
 patterns; observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not depend on \(n_1\) or \(n_2\).
-In this context, the integers $1, \ldots, P$ serve as \emph{hashed values} that encode the same information as the orignial vectors themselves.
+In this context, the integers $\{1, \ldots, P\}$ serve as \emph{hashed values} that encode the same information as the orignial vectors themselves.
 Whenever possible, we conduct calculations over these \(P\) agreement
@@
 Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
-
-%In this section, we propose efficient and scalable proposals using hashing based methods that are similar in spirit to \citep{enamorado2019}, a fast and scalable implementation of the Fellegi-Sunter model.
-% Specifically, one way we can improve our computational efficiency is by recognizing that record pairs contribute to posterior calculations only
-%through the agreement pattern of the \(\gamma_{ij}\) vector. To make this more precise, let 
-%\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
-%\(P = |\mathcal{H}|\) denote the total number of unique agreement
-%patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
-%that this bound does not depend on \(n_1\) or \(n_2\). Prior to
-%processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
-%and enumerate them as follows: \(h_1, \ldots, h_P\), which are called \emph{hashed values.}
-%Next, we map record pairs to their corresponding hashed value. 
-%More specifically, when the record pair \((i,j)\) 
-%exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
-%Whenever possible, we conduct calculations over these \(P\) agreement
-%patterns, instead of the typical  \(n_1 \times n_2\) record pairs.
 
@@
 
-First, we hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019using} accomplished this efficiently through the hashing function
+First, we hash record pairs of the same agreement pattern to unique integer values. \cite{enamorado2019using} accomplished this efficiently through the hashing function
 
@@
 This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
-For computational ease, we then map these integers to sequential integers from $1, \ldots, P$ corresponding to the enumerated patterns in $\mathcal{P}$.
+For computational ease, we then map these integers to sequential integers from $\{1, \ldots, P\}$ corresponding to the enumerated patterns in $\mathcal{P}$.
 When the $(i, j)$ pair exhibits the $p^{th}$ pattern, we say $(i, j) \in h_p$.
@@
 $$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
-For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{j_p}\}\right\}$As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
+For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}$As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
 Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for our efficient posterior inference.
 
-\hypertarget{large-data}{%
-	\subsection{Working with Large Data}\label{large-data}}
+\hypertarget{SEI}{%
+	\subsection{Storage Efficient Indexing}\label{SEI}}
 
@@
 
-%The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
-%comparison as a vector of length \(F\), with each component
-%\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). 
-%To ease our computational burden, we instead use a \emph{one hot encoding} of the comparison vector. For example, if
-%\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
-%under the classical framework becomes
-%\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
-%bijective transformation that does not change the meaning of the data,
-%but this representation eases calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.
-
-%In the classic Fellegi Sunter framework, $\bm{\Gamma}$ is a
-%\(n_1 n_2 \times F\) matrix, with each row providing the comparison
-%vector for a different \((i,j)\) pair. In contrast, we do not store
-%these comparison vectors themselves, but instead only the hashed value $h_p$
-%corresponding to the agreement pattern of the \((i, j)\) pair. We store
-%this information in a nested list $\tilde{\bm{\Gamma}}$ where the
-%\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
-%in \(\bm{X}_1\) that share agreement pattern \(p\) with record \(j \in \bm{X}_2\). For
-%each \(p\), we also calculate
-%\(H_p = \sum_{i=1}^{n_1}\sum_{j=1}^{n_2} \mathbf{1}_{(i,j) \in h_p}\),
-%the total instances of agreement pattern \(p\) throughout the data, and
-%also for each \(j\), we calculate
-%\(H_{p_j} = \sum_{i=1}^{n_1} \mathbf{1}_{{(i,j) \in h_p}}\) the
-%instances of agreement pattern \(p\) among the comparison vectors
-%between record \(j \in \bm{X}_2\) and each of the \(n_1\) records in \(\bm{X}_1\).
-
 To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI).
@@
 With SEI however, we can reduce comparison space, and its associated storage costs, while avoiding these drawbacks.
-Observe that all records of the same agreement pattern have the same probability when sampling \(Z_j\).
-Therefore we know that records $\i in r_{j_p}$  such that
-\(H_{j_p}\) is large are very unlikely to be sampled consistently enough
-to be deemed a match through the Bayes estimate.
+Observe that all record $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{x}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
+Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
 We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
 Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$.
-Rather than storing
-\(n_1 \times n_2\) record labels, SEI allows us to store at most
+Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most
 \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
@@
 
-Lastly, for large data, we can partition the two datasets \(\bm{X}_1\) and
+\hypertarget{distributed}{%
+	\subsection{Distributed Computation for Pairwise Comparison}\label{distributed}}
+
+For large data, we can partition the two datasets \(\bm{X}_1\) and
 \(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more
@@
 
-\underline{Sample $\Phi^{(s+1)}|\tilde{\Gamma}, \mathbf{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
+\underline{Sample $\Phi^{(s+1)}|\tilde{\Gamma}, \bm{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
 These are simple calculations, but become computationally burdensome when working on large linkage tasks.
@@
 Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote
-$H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the
+$H_p^m = \sum_{j=1}^{n_B} \bm{1}_{(Z_j, j) \in h_p}$ to be the
 number of matching record pairs with agreement pattern $p$.
 We can also easily calculate the number of nonmatching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\).
-Then, if $\mathbf{\alpha_0}$ and $\mathbf{\beta_0}$ are vectors of prior parameters for the $\mathbf{m}$ and $\mathbf{u}$ distributions respectively, the posterior update becomes simply
-
-$$\alpha(\mathbf{Z}) = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
-$$\beta(\mathbf{Z}) = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$
+Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes simply
+
+$$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
+$$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$
 
@@
 
-%After receiving
-%matching statuses from \(\mathbf{Z}\), the Sadinle method calculates
-%\(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each
-%field and level. This constitutes \(2 \times \sum L_f\) many summations
-%over \(n_A \times n_B\) quantities, and becomes computationally
-%burdensome with large data. In contrast, we recognize that each unique
-%agreement pattern contributes to the posterior \(\alpha(\mathbf{Z})\)
-%and \(\beta(\mathbf{Z})\) vectors in the same way. In fact, if we denote
-%\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
-%number of matching record pairs with agreement pattern \(p\), then the
-%contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
-%vector is simply \(H_p^m \times h_p\). Thus our posterior update for the
-%\(\alpha\) vector is simply
-%\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
-%we can easily calculate \(H_p^u\), the number of nonmatching record
-%pairs of agreement pattern \(p\), by subtracting the number of matching
-%pairs from the total present in the data; that is
-%\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
-%parameter through
-%\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
-%that these constitute \(P\) many summations over \(n_B\) quantities, and
-%thus avoid the \(n_A \times n_B\) summation from the original method.
-
-\underline{Sample $\mathbf{Z}^{(s+1)}|\tilde{\Gamma}, \Phi^{(s+1)}$:} Although sampling \(Z_j\) from a the
+\underline{Sample $\bm{Z}^{(s+1)}|\tilde{\Gamma}, \Phi^{(s+1)}$:} Although sampling \(Z_j\) from a the
 full conditional provided earlier is conceptually straightforward, it
@@
 pattern between \(j\) and its potential match, and say
-\(h(Z_j) = h_{P+1}\) when \(Z_j = n_A + 1\). Then,
-
-$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \mathbf{Z^{(s)}}\right) \propto
+\(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + 1\). Then,
+
+$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
 	w_{p}\times H_{j_p}  & p \leq P; \\
-	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
+	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
 \end{cases}$$
@@
 
-Lastly, we recognize that all posterior updates are governed by the
-agreement patterns of the record pairs rather than the record labels
-themselves. Thus we complete the entire Gibbs procedure first at the
-level of the \(P\) agreement patterns with the first equation above.
+Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
+Thus we complete the entire Gibbs procedure first at the level of the \(P\) agreement patterns with the first equation above.
 After all Gibbs iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.
@@
 To prove the computational complexity, we consider two steps --- constructing the comparison vectors and the Gibbs sampler.
-%
-The computational complexity of all pairwise comparisons across the two databases $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$.
+
+The computational complexity of all pairwise comparisons across the two databases $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
 The hashing procedure for all pairwise comparisons is also of complexity $O(F n_1 n_2)$.
 Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
-There are then minimal computational costs associated with synthesizing summary statistics across these partitions.
-
-Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the overall complexity to $O(L^{*}P)$, where $L^{*} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task and $P$ is the total number of agreement patterns.The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
+There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
+
+Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_1 n_2)$.
+However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the overall complexity to $O(\tilde{L}^{*}P)$, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.
+The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
 With hashing, we split this sampling into two steps.
-First we sample that agreement pattern of the match with complexity $O(n_2 P).$ Next, we sample the record that exhibits that pattern with complexity $O(n_2)$.
-Since $L^{*} << n_1$ in all reasonable applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
+First we sample that agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
+Thus the complexity of sampling $\bm{Z}$ is $O(n_2 P)$.
+Since $\tilde{L} << n_1$ in all reasonable applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
 In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
@@
 
-%{
-%\color{blue}
-%The computational complexity of \texttt{fabl} is determined by its two steps: constructing comparison vectors and Gibbs sampling.
-The computational complexity of all pairwise comparisons across files $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$.
-Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
-%
-%Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the complexity to $O(FP)$.
-%
-%The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
-With hashing, we split this sampling into two steps; first we sample that agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record that exhibits that pattern with complexity $O(n_2)$.
-Thus we have reduced the complexity of the Gibbs sampler under \texttt{BRL} from $O(F n_1 n_2)$ to $O(n_2 P)$.
-%
-%Overall, the computational complexity under the complete \texttt{fabl} framework is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
-%}
-
 \section{Simulation Studies}
@@
 
-%Sections \ref{accuracy} and \ref{partial} utilize simulated datasets from \cite{sadinle_bayesian_2017} made available by the author to show the similarity of both methods across several accuracy metrics.
-These same datasets are used in Section \ref{scale} to create a larger linkage test to demonstrate the scale of problems \texttt{fabl} can undertake.
-Finally, Section \ref{speed} demonstrates the computational gains.
-
 \hypertarget{accuracy}{%
@@
 
-We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap.
+We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap.
 We use first name, last name, age, and occupation for this linkage, where age and occupation take on 9 and 10 unique values respectively.
-We creage comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table~\ref{Tab:sadinle_simulation_cutoffs}.
+We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table~\ref{Tab:sadinle_simulation_cutoffs}.
 Each simulation identifies duplicated individuals between two datasets, each with 500 records.
@@
 We compare \texttt{fabl} to \texttt{BRL} in terms of precision, recall, and f-measure, which are commonly defined record linkage evaluation metrics that are defined in \citep{christen_2012}.
-At all levels of error and overlap, we see that the two methods have comparable performance.
-In the specific case of high error and high overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, since both methods exhibit high variance in performance in this setting, we remain confident in \texttt{fabl}'s overal accuracy.
-
-%In cases when there are only one or two errors in matching records, and in cases with low to moderate duplication across records, we see that \texttt{fabl} provides approximately equivalent accuracy as \texttt{BRL}.
-%We find that our method only has weakened performance in the
-%most extreme scenario of very high errors and very high overlap across
-%files.
-In these situations, \texttt{BRL} is removing large numbers of records from consideration throughout the Gibbs Sampler, making its implementation most different from \texttt{fabl}.
-However,  such extreme linkage tasks, which such high amounts of errors and overlap, are extremely rare in practice.
-
-%\begin{figure}
-%	\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
-%	\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
-%	\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
-%	
-%	\caption{Definitions of performance metrics}\label{fig:performance_metrics}
-%\end{figure}
+In Figure~\ref{fig:sadinle_simulation}, we see that we see that the two methods have comparable performance at all levels of error and overlap, .
+In the specific case of high error and high overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, since both methods exhibit high variance in performance in this setting, we remain confident in \texttt{fabl}'s overall accuracy.
 
@@
 \caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
-For each level of overlap and each level of error, we have 100 paired sets of 500 records.}
+For each level of overlap and each level of error, we have 100 paired sets of 500 records.
+We see comparable performance for all levels of error and overlap.}
 \label{fig:sadinle_simulation}
@@
 \end{figure}
-
-%
-%\begin{figure}[ht]
-%	
-%	{\centering 
-%		
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot} 
-%	}
-%	
-%	\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017.
-For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:sadinle_simulation}
-%	
-%\end{figure}
 
@@
 Thus, instead of recall, we use the \emph{negative predictive value} (NPV), defined as the proportion of non-links that are actual non-matches.
-Mathematically, $NPV = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
-We continue to use the precision, which is renamed the \emph{positive predictive value} in this context.
-Lastly, we also report that rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.
-
-We see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
-However, with high amounts of overlap and high amounts of error, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
-This is reasonable because large numbers of errors lead to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
-Since \texttt{BRL} removes previously matched records from the set of available records for a new match, it is able to retain higher confidence in matches.
-In practice, this partial estimate approach is not often used, so we remain confident in \texttt{fabl}'s performance.
+Mathematically, $\text{NPV} = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
+We continue to use the precision, which is renamed the \emph{positive predictive value} (PPV) in this context.
+Lastly, we also report the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.
+
+In Figure~\ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
+However, with high amounts of overlap, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
+Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probability of matches can at times be more split across different records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
+In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
 If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
+
+%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
 
@@
 	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
-	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets}
+	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
 	\label{fig:sadinle_simulation_partial}
@@
 distributions so that we can easily increase the size of the linkage
-problem. Distributions are meant to emulate the behavior of similarity
-scores across first name, last name, and day, month. For example, $u^{\text{month, 1}} = P(\text{Records have same birth-month | Nonmatch}) = \frac{1}{12}$. For simplicity, we
-consider only exact matching, so a vector (1, 0) corresponds to
+problem.
+The distributions used (see Table~\ref{Tab:distributions}) are meant to emulate the rates of agreement across first name, last name, day of birth, and month of birth.
+For example, $u^{\text{month, agree}} = P(\text{Records have same birth-month} \mid \text{Nonmatch}) = \frac{1}{12}$.
+The parameters for year would vary largely by context, so we chose them here to adjust the desired difficulty of the linkage task.For simplicity, we consider only exact matching, so a vector (1, 0) corresponds to
 agreement and (0, 1) to disagreement. We simulate these data for
-different values of \(n_A\) and \(n_B\), and compare the run-time of
-\texttt{fabl} against \texttt{BRL}. Note that the number of unique
-patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is
+different values of \(n_1\) and \(n_2\), and compare the run-time of
+\texttt{fabl} against \texttt{BRL}. Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is
 consistently attained in the larger simulations.
-
-We see that at low data size, \texttt{BRL} outperforms, but that
-\texttt{fabl} is significantly faster at handling larger data. In
-particular, run-time for \texttt{BRL} seems to grow quadratically (or
-linearly with the size of both \(A\) and \(B\)) while run-time for
-\texttt{fabl} seems to grow linearly (in the size of \(B\)).
 
@@
 
+In Figure~\ref{fig:speed1}, we see that at low data size, \texttt{BRL} outperforms, but that
+\texttt{fabl} is significantly faster at handling larger data. In
+particular, run-time for \texttt{BRL} seems to grow quadratically (or
+linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for
+\texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
+
 \begin{figure}[h!]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2}
-
-	}
-
+	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
 	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$.
 We see near quadratic growth in runtime for BRL, and near linear growth for fabl.}\label{fig:speed1}
-\end{figure}
-
-The above discussion suggests that for fixed \(n_B\), computation time
-should remain mostly constant with growing \(n_A\). Simulation study
-suggests that this is true. In the plot below, fixing \(n_B = 500\), we
-see linear growth for the run-time under \texttt{BRL} as \(n_A\)
+	\end{center}
+\end{figure}
+
+The above discussion suggests that for fixed \(n_2\), computation time
+should remain mostly constant with growing \(n_1\). Simulation study
+suggests that this is true. In Figure~\ref{fig:speed2} fixing \(n_2 = 500\), we
+see linear growth for the run-time under \texttt{BRL} as \(n_1\)
 increases, with much more static run-time under \texttt{fabl}. The
@@
 \begin{figure}[h!]
-
-	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB}
-
-	}
-
+	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB} 
 	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
 We see linear growth in runtime for BRL, and near constant runtime for fabl.}\label{fig:speed2}
+	\end{center}
 \end{figure}
@@
 
-In this section, we investigate scaling to larger data sets using our proposed methodology.
-
-Specifically, we demonstrate how partitioning and hashing the data through \texttt{fabl} method allows us undertake significantly larger linkage tasks.
-We concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{} to create two larger files, each of 20,000 records (40,000 records total).
-Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting a final comparison matrix about 6.4 GB in size.
+To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
+Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
 Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task.
@@
 We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
-The resulting data object is now only 90 MB, about 1\% the size of the object required under the standard method.
+The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method.
 Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run.
@@
 The reader will note that this slightly worse performance than witnessed in the smaller simulation studies; this is expected because it is naturally more difficult to link more files with the same amount of information.
-With more linkage fields \texttt{fabl} maintains the high accuracy seen above.
-
-%\textcolor{blue}{This is with post processing already.
-All I meant here was that is that small tasks require less information.
-If you do a $500 \times 500$ linkage task and you only have first and last name (for example), thats likely enough to get all the correct matches.
-However, once you do $20000 \times 20000$, its more likely that more than one person in the second file has the same first and last name, so you need more information to distinguish.
-I mentioned before that the simulated files actually have 6 fields of information, but I only used 4 originally because thats what Sadinle did.
-I can use all 6 though to see what happens?}
+With more linkage fields, \texttt{fabl} maintains the high accuracy in Section~\ref{accuracy}.
 
@@
 
-In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in each dataset.
+In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in both datasets.
 Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
@@
 multiple casualty lists.
-This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications} Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize
-two different sources, namely,  El Rescate - Tutela
-Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
-acronym in Spanish). The ERTL dataset consists of digitized denunciations that
-had been published throughout the conflict, and additional investigation was required before the ERTL recorded them as human rights abuses, giving us confidence in its accuracy \cite{howland2008rescate}.
-The CDHES dataset consists
-of casualties that had been reported directly to the organization, and
-later digitized; since reports were taken soon after the events occured, this dataset is also thought to be fairly reliable. \cite{ball2000salvadoran}.
-
-There are several challenges with working with such data. Firstly, both
+This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications}.
+We utilize lists of casualties from the war, one collected by El Rescate - Tutela
+Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its
+acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that
+had been published throughout the conflict, and the CDHES dataset consists
+of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
+The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both datasets are thought to be fairly reliable.
+
+There are several challenges with working with this data. Firstly, both
 datasets have been automatically digitized, which inherently leads to
-some degree of typographical error. Secondly, the CDHES records are all
-second hand accounts reported by individuals, which can result in
-additional errors. Lastly, the only fields recorded are given name, last
-name, date of death, and place of death; it is relatively common for a
+some degree of typographical error. Additionally, the only fields recorded are given name, last name, date of death, and place of death; it is relatively common for a
 parent and child to share the same given name, resulting in
@@
 
-In this paper, we utilize records that have non-missing entries for given and last name, which results 
-in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files
-in ERTL. We standardized names to account for common misspellings
-in the Spanish language. The string distance used is a modified Levenstein
-distance to account for the fact that second names are often omitted.
-Place of death is recorded by municipality and department within that
-municipality; however, since department was missing in 95\% of records
-in CHDES and 80\% of records in ERTL, this was excluded from our analyses. 
-Thus, we conduct record linkage using given name, last name,
-municipality, and day, month, and year of death. We again use flat
-priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters.
+Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files in ERTL.
+We standardized names to account for common misspellings and used a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
+Place of death is recorded by municipality and department within that municipality; however, since department was missing in 95\% of records in CHDES and 80\% of records in ERTL, this was excluded from our analyses.
+Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
+We again use flat priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
+
+To mirror the original implementation of \cite{sadinle_bayesian_2017}, we first construct the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
+Gibbs sampling for these comparison vectors took 422 seconds for our proposed method, \texttt{fabl}, and 240 for \texttt{BRL}.
+However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, and that of the \(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 were realized in the data.
+This leads us to believe that such high number of agreement levels creates unnecessary distinctions in the data and makes the comparison vectors less interpretable.
 
@@
 		\hline
-	First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (5, 1]\\
+		First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (.5, 1]\\
 		Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
@@
 	\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle2018bayesian}.
-This set up leads to 2048 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_1}
+This set up leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
 \end{table}
+
+Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}, and obtained analogous inference.
+With 216 possible agreement patterns, 159 were realized in the data, and our proposed method, \texttt{fabl} became much faster, finishing in 61 seconds.
+Meanwhile \texttt{BRL} took 239 seconds, relatively unchanged from the first implementation.
+This demonstrates the way that the computational complexity of our method depends on the number of unique agreement patterns, and how significant computational gains can be made by simplifying the construction of the comparison vectors.
+The estimates of the \(\bm{m}\) parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
+Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}
 
@@
 	\end{tabular}
-\caption{Construction of comparison vectors for El Salvador for increased speed under \texttt{fabl}. This set up leads to 216 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_2}
+	\caption{Construction of comparison vectors for El Salvador for increased speed under \texttt{fabl}. This set up leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
 \end{table}
 
-To mirror the original implementation of \cite{sadinle_bayesian_2017}, we first construct the comparison
-vectors using four levels of agreement for each field, according to the
-thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}. Gibbs sampling for these comparison vectors took 422 seconds for our proposed
-method, \texttt{fabl}, and 240 for \texttt{BRL}. However, we observed that posterior
-distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
-parameters were nearly identical, and that of the
-\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
-realized in the data. This leads us to believe that such high number of
-agreement levels creates unnecessary distinctions in the data and makes
-the comparison vectors less interpretable.
-
-Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}, and
-obtained analogous results. With 216 possible agreement patterns, 159
-were realized in the data, and our proposed method, \texttt{fabl} became much faster,
-finishing in 61 seconds. Meanwhile \texttt{BRL} took 239 seconds,
-relatively unchanged from the first implementation. This demonstrates
-the way that the computational complexity of our method depends on the
-number of unique agreement patterns, and how significant computational
-gains can be made by simplifying the construction of the comparison
-vectors. The estimates of the \(\mathbf{m}\) parameters under each method are very similar, as shown in
-Figure~\ref{fig:m-and-u}. Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}
-
 Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets.
-We calcuate posterior samples of size of the overlap across files by summing the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
+We calcuate posterior samples of the size of the overlap across files by finding the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
 The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate is quite conservative.
-This is because there a large number of records in ERTL for which there are multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair a match in the Bayes estimate.
+This is because there a large number of records in ERTL for which there are multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
 We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study.
@@
 	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-\caption{Posterior estimates of m parameters with 95\% credible intervals. We note they are quite similar across the two methods. }\label{fig:m-and-u}
+\caption{Posterior estimates of m parameters with 95\% credible intervals. We note they are quite similar across the two methods.}\label{fig:m-and-u}
 \label{fig:m-and-u}
@@
 To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
-These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in \(\bm{X}_1\) has
-multiple plausible matches in \(\bm{X}_2\).
+These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $\bm{X}_1$ has
+multiple plausible matches in $\bm{X}_2$.
 
@@
 
-Table ~\ref{Tab:rosa-maria} illustrates an example. Note that these
-records present a near violation of our assumption that there are no
-duplications within files; we continue to assume that records 825 and
-826 in ERTL correspond to different individuals, but their records are
-nearly identical. In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and
-\(\gamma_{2776, 826}\) are exactly identical.
+Table~\ref{Tab:rosa-maria} illustrates an example.
+Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
+In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and \(\gamma_{2776, 826}\) are exactly identical.
 
@@
 
-Figure~\ref{fig:mixing-plot} shows the values that \(Z_{825}\) and \(Z_{826}\) take on
-throughout the Gibbs sampler, and demonstrates how each method handles
-this situation. Under \texttt{fabl}, both records in \(\bm{X}_2\) match to the
-same record in \(\bm{X}_1\) throughout the Gibbs process, creating consistent
-violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
-creates one matching configuration and stays there for a while. However, if
-one pair ``unmatches,'' then the other record has a chance to latch on.
-Then, the Gibbs process is stuck with that matching status for a while,
-resulting in a Gibbs process with poor mixing. In addition,
-\texttt{fabl} allows the modeler to inspect records with multiple
-plausible matches, and if they desire, to then choose the record pairing
-with the highest posterior probability. \texttt{BRL} in contrast, in
-strictly enforcing one-to-one matching throughout the sampler, can lead
-to situations where none of the plausible matches reach the threshold to
-be identified through the Bayes estimate.
-In short, we have pointed out why one may not wish to imbed this constraint in a model, and proposed a simple alternative that performs about the same as \texttt{BRL}, while scaling to much larger data sets.
+Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
+Under \texttt{fabl}, both records in $\bm{X}_2$ match to the same record in $\bm{X}_1$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
+Under \texttt{BRL}, the Gibbs process creates one matching configuration and stays there for a while.
+However, if one pair ``unmatches,'' then the other record has a chance to latch on.
+Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
+
+Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
+\texttt{BRL} in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
+Thus relaxing the one-to-one constraint has served as a useful tool for identifying model misspecification.
 
@@
 \end{figure}
+
+\section{Discussion}
+\label{discussion}
+
+We have presented a method for Bayesian record linkage that is feasible for large datasets.
+In particular, our hashing procedure and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making this method particularly powerful in linking records when one datafile is
+substantially smaller than the other.
+
+In our case study, we included an exploration of how to conduct record linkage when modeling assumptions are not met in practice.
+We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
+Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple
+plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
+Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
+future work.
 
@@
 
-\section{Discussion}
-\label{discussion}
-
-We have presented a method for Bayesian record linkage that is feasible
-for large datasets. In particular, our hashing procedure and model
-assumptions allow for a linkage procedure whose computational complexity
-does not scale with the size of the larger dataset, making this method
-particularly powerful in linking records when one datafile is
-substantially smaller than the other.
-
-In our case study, we included an exploration of how to conduct record
-linkage when modeling assumptions are not met in practice. We explored
-``one-to-many'' scenarios in which one record in \(A\) has multiple
-plausible matches in \(B\), and showed how both \texttt{fabl} and
-\texttt{BRL} demonstrated undesirable qualities. Other issues arise
-under ``many-to-one'' scenarios, where one record in \(B\) has multiple
-plausible matches in \(A\), and ``many-to-many'' scenarios in which
-there is duplication both across and within datasets. Tuning
-\texttt{fabl} for use in these scenarios is one potential avenue for
-future work.
-
 \clearpage
@@
 		\Procedure{Hashing and Preprocessing}{}
-		\State Construct and enumerate set of unique pattern $\mathcal{P}$ from $F$ and $\{L_f\}$
+		\State Construct and enumerate set of unique patterns $\mathcal{P}$ from $F$ and $\{L_f\}$
 		\State Partition files $\bm{X_1}$ and $\bm{X_1}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$
@@
 		\EndFor
-		\State Sample $Z | h(Z), \Gamma$ \Comment{Fills in record label based on agreement pattern}
+		\State Sample $Z | h(Z), \tilde{\Gamma}$ \Comment{Fills in record label based on agreement pattern}
 

### 2021-10-13 methods_paper/fabl-paper.tex
suggestions to abstract.

@@
 Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
-This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.
+\textcolor{red}{We derive the associated conditional distributions and provide a Gibbs sampling algorithm for our proposed framework.}
+\textcolor{red}{Our modified prior} allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.
 In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}.
-Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
+\textcolor{red}{We derive the associated conditional distributions and provide a Gibbs sampling algorithm for our proposed framework.}
+\textcolor{red}{Moreover, we provide the computational complexity of \texttt{fabl}.} Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
 Finally, we discuss our work and future directions.
-
 \end{abstract}

### 2021-10-13 methods_paper/fabl-paper.tex
minor updates.

@@
 This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
-In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.
-
+\textcolor{red}{In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.}
 	%Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
-
 	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
@@
 \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records.
-Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler.
+Most recently, \cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler.
 Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
@@
 Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
-Open source software to use \texttt{fabl} in \texttt{R} is available through Github.
-
-	In what follows, Section \ref{sec:review-of_prior-work} reviews more thoroughly the Fellegi Sunter and Beta Record Linkage methods.
-Then, Section \ref{sec:fast-beta-linkage} provides the model specification under \texttt{fabl} and derives the Gibbs sampler for posterior inference.
-Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
-Lastly, Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War.
-Section \ref{discussion} concludes our paper with a discussion of directions for future work.
+Open source software to use \texttt{fabl} in \texttt{R} is available through Github at \textcolor{red}{BK: Add the URL even if it's private at the moment: \url{}}.
+
+	\textcolor{red}{The remainder of the paper is as follows.} Section~\ref{sec:review-of_prior-work} reviews \textcolor{red}{the prior work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.} In section~\ref{sec:fast-beta-linkage}, we propose our the \texttt{fabl} model specification.
+\textcolor{red}{Also, we derive the corresponding conditional distributions and provide a Gibbs sampler for posterior inference.}  Section \ref{bayes-estimate} \textcolor{red}{Brian: add in what you do in this section.}
+	Section~\ref{sec:efficiency} \textcolor{red}{proposes our approach} for hashing and  and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
+In addition, we prove the computational complexity of \texttt{fabl}.
+Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War.
+Section~\ref{discussion} concludes our paper with a discussion of directions for future work.
 
@@
 Thus, we can construct comparison vectors in the following way:
-$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$
-
+$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$
 The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.
@@
 This behavior may be desirable in its own right; \cite{Newcombe1962b} were interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable.
-In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.
+In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in  section~\ref{bayes-estimate} to resolve illegal matchings.
 
@@
 We seek to minimize the posterior expected loss, given by 
-$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs})$$. 
+$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
 In this paper, we adopt losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule
@@
 In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.
-A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
+A similar approach can be see in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities, and is justified within this Bayesian framework.
 

### 2021-10-13 methods_paper/fabl-paper.tex
fixed number of lemma.

@@
 
-\newtheorem*{remark}{Remark}
+\newtheorem{remark}{Remark}
 \newtheorem{proposition}{Proposition}
-\newtheorem*{definition}{Definition}
-\newtheorem*{lemma}{Lemma}
+\newtheorem{definition}{Definition}
+\newtheorem{lemma}{Lemma}
 
@@
 
-To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we employ hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
+To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we \textcolor{red}{propose} hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
 The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
@@
 $$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
-For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}$As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
+For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
 Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for our efficient posterior inference.
@@
 The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
+\label{lemma:fabl}
 \end{lemma}
@@
 To prove the computational complexity, we consider two steps --- constructing the comparison vectors and the Gibbs sampler. 
-
 The computational complexity of all pairwise comparisons across the two databases $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.

### 2021-10-13 methods_paper/fabl-paper.tex
update.

@@
 \textcolor{red}{We derive the associated conditional distributions and provide a Gibbs sampling algorithm for our proposed framework.}
-\textcolor{red}{Moreover, we provide the computational complexity of \texttt{fabl}.} Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
+\textcolor{red}{Moreover, we prove the computational complexity of \texttt{fabl}.} Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
 Finally, we discuss our work and future directions.
@@
 	\subsection{Computational Complexity}\label{computational-complexity}}
-In this section, we provide the computational complexity of \texttt{fabl} in Lemma \ref{lemma:fabl}.
+In this section, we prove the computational complexity of \texttt{fabl} in Lemma \ref{lemma:fabl}.
 Let $L^{*} = \sum_{f=1}^F L_f$ denote the total number of levels used across all features in the linkage task.
@@
 We compare \texttt{fabl} to \texttt{BRL} in terms of precision, recall, and f-measure, which are commonly defined record linkage evaluation metrics that are defined in \citep{christen_2012}.
-In Figure~\ref{fig:sadinle_simulation}, we see that we see that the two methods have comparable performance at all levels of error and overlap, .
+In Figure~\ref{fig:sadinle_simulation}, we see that we see that the two methods have comparable performance at all levels of error and overlap.
 In the specific case of high error and high overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, since both methods exhibit high variance in performance in this setting, we remain confident in \texttt{fabl}'s overall accuracy.
@@
 
-We have presented a method for Bayesian record linkage that is feasible for large datasets.
-In particular, our hashing procedure and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making this method particularly powerful in linking records when one datafile is
-substantially smaller than the other.
-
-In our case study, we included an exploration of how to conduct record linkage when modeling assumptions are not met in practice.
+\textcolor{red}{In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. Specifically, our proposed hashing method}
+%We have presented a method for Bayesian record linkage that is feasible for large datasets. In particular, our hashing procedure 
+and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making our approach particularly powerful in linking records when one datafile is substantially smaller than the other.
+\textcolor{red}{In our paper, we have provided details regarding derivations of conditional distributions and our proposed Gibbs sampler.
+In addition, we prove the computational complexity of our algorithm.
+In addition, we provided both simulation studies and revisited the same case study of the main methodology that we build upon.}  In our case study, we included an exploration of how to conduct record linkage when modeling assumptions are not met in practice.
 We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
-Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple
-plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
-Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
-future work.
+Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
+Extending \texttt{fabl} in such scenarios is currently investigation by the authors regarding future work.
+
+\textcolor{red}{Brian: could you talk about a few other directions that might be useful to take your work as this will come up in your prelim.
+For example, how might you extend beyond bipartite and still scale?
+Or what ideas do you think are the most exciting to take your work in next.
+This is your opportunity to talk about anything here and give a comment to the community about what you think is most important!}
+
+%Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
+%future work.
 
@@
 
-\section{Appendix}
+\section*{Appendix}
 \label{sec:appendix}

### 2021-10-14 methods_paper/fabl-paper.tex
dont make biblio small.
references come before appendix.
addressed bekas comments

--- 2021-10-13 methods_paper/fabl-paper.tex
@@
 Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
-\textcolor{red}{We derive the associated conditional distributions and provide a Gibbs sampling algorithm for our proposed framework.}
-\textcolor{red}{Our modified prior} allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.
+This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.
 In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}.
-\textcolor{red}{We derive the associated conditional distributions and provide a Gibbs sampling algorithm for our proposed framework.}
-\textcolor{red}{Moreover, we provide the computational complexity of \texttt{fabl}.} Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
-Finally, we discuss our work and future directions.
+Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
 \end{abstract}
@@
 
-%	I have some proposed revisions for the first part of the paper.
-In the introduction, I reordered things and changed some topic sentences to highlight that the main contribution here is speed and scale.
-%	
-%	Also, in the prior version, we didn't get to any of the proposed material until Section 4.3, which did not seem reasonable to me.
-%	I know you recommended splitting the review of the literature between the Introduction and Section 2 on bipartite matching, but since we needed the mathemetical overview of FS and Sadinle, I decided to move all of the literature review into the introduction.
-I've seen this done in many papers.
-%	
-%	Almost all of the content is still there, but it's moved around in a way that makes more sense to me. Let me know your thoughts! (But please avoid hard edits until we talk on Monday!)
-%	
-%	I have some additional edits in the later part of the paper about the partial Bayes estimate, computational complexity, and added detail for the case study. All of that is in blue as well.
-
 	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
 This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
-\textcolor{red}{In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.}
-	%Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
+In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.
+
 	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
@@
 Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.
+	% Possible citation for bipartite matching. 
+	%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
 
@@
 
-	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost.
+	Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost.
 \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
@@
 Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
-Open source software to use \texttt{fabl} in \texttt{R} is available through Github at \textcolor{red}{BK: Add the URL even if it's private at the moment: \url{}}.
-
-	\textcolor{red}{The remainder of the paper is as follows.} Section~\ref{sec:review-of_prior-work} reviews \textcolor{red}{the prior work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.} In section~\ref{sec:fast-beta-linkage}, we propose our the \texttt{fabl} model specification.
-\textcolor{red}{Also, we derive the corresponding conditional distributions and provide a Gibbs sampler for posterior inference.}  Section \ref{bayes-estimate} \textcolor{red}{Brian: add in what you do in this section.}
-	Section~\ref{sec:efficiency} \textcolor{red}{proposes our approach} for hashing and  and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
+Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
+
+	In what follows, Section~\ref{sec:review-of_prior-work} reviews the prior work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
+Section~\ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching.
+	Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
 In addition, we prove the computational complexity of \texttt{fabl}.
@@
 Section~\ref{discussion} concludes our paper with a discussion of directions for future work.
-
-	% I think this is clear and has already been said. 
-%	To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl}, pronounced ``fable'')
-
-%
 
@@
 where $p$ is the proportion of records that match, and $\bm{m}$ and $\bm{u}$ are vectors of parameters for the matching and nonmatching record pairs.
-These unknown parameters are often estimated using the EM algorithm.
+These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.
 
@@
 Importantly, this independence means that we have relaxed the one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desirable in its own right; \cite{Newcombe1962b} were interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable.
+This behavior may be desirable in its own right; \cite{Newcombe1963} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable.
 In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in  section~\ref{bayes-estimate} to resolve illegal matchings.
@@
 With the likelihood in this form, we can derive the full conditional
-\[p(Z_j|\Gamma_{.j}, \Phi, \pi) \propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi)\]
-
-\[\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right)\]
-\[= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1} \]
-\[ \implies Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto
-\begin{cases} 
+\begin{align*}
+	p(Z_j|\Gamma_{.j}, \Phi, \pi) &\propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi) \\
+	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
+	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1} \\
+	&\implies Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
 	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
 	1-\pi &  z_j  = n_1 + 1 \\
-\end{cases}\]
+\end{cases}
+\end{align*}
 
@@
 
-To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we \textcolor{red}{propose} hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
+To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
 The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
@@
 
-We have presented a method for Bayesian record linkage that is feasible for large datasets.
-In particular, our hashing procedure and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making this method particularly powerful in linking records when one datafile is
-substantially smaller than the other.
-
-In our case study, we included an exploration of how to conduct record linkage when modeling assumptions are not met in practice.
+In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. Specifically, our proposed hashing method
+and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making our approach particularly powerful in linking records when one datafile is substantially smaller than the other.
+
+Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
 We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
-Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple
-plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
-Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
-future work.
+Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
+Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is currently under investigation by the authors for future work.
+
+To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
+This motivates extentions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
+We see these types of heirarchical record linkage models as potential contributions to the growing algorithmic fairness literature, and useful to practitioners working with or studying marginalized populations.
+
+%Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
+%future work.
 
@@
 
-\section{Appendix}
+\bigskip
+
+\bibliographystyle{jasa}
+\bibliography{biblio}
+
+\clearpage
+
+\section*{Appendix}
 \label{sec:appendix}
@@
 
-\clearpage
-
-\bigskip
-
-\bibliographystyle{jasa}
-{
-\footnotesize
-\bibliography{biblio}
-}
-
 \end{document}

### 2021-10-14 bayesian-analysis/kundinger-ba-2021.tex
all there but names.

--- 2021-10-14 methods_paper/fabl-paper.tex
@@
-\documentclass[12pt,letterpaper]{article}
+% \documentclass[ba,preprint]{imsart}% use this for supplement article
+\documentclass[ba]{imsart}
+%
+\pubyear{2021}
+\volume{TBA}
+\issue{TBA}
+% \doi{0000}
+%\arxiv{}
+\firstpage{1}
+\lastpage{1}
 
@@
 
+\usepackage{amsthm}
+\usepackage{amsmath}
+\usepackage{natbib}
+%\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
+%\usepackage{graphicx}
+
 % Custom commands
@@
 
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
-
-\sectionfont{\large\nohang\centering\MakeUppercase}
-
+\startlocaldefs
+% ** Local definitions **
+\endlocaldefs
+
+\begin{document}
+
+%% *** Frontmatter ***
+
+\begin{frontmatter}
 \title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
-\author{Brian Kundinger\textsuperscript{a} \and
-  Jerome Reiter\textsuperscript{a} \and 
-  Rebecca C.~Steorts \textsuperscript{b}}
-\date{
- \textsuperscript{a} Department of Statistical Science, Duke University \\
- \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
-  \today}
-
-\begin{document}
-\maketitle
-
-\bigskip
+%\title{\support{}}
+\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+
+\begin{aug}
+%\author{\fnms{} \snm{}}
+%\author{\fnms{<firstname>} \snm{<surname>}\thanksref{}\ead[label=e1]{}}
+%\and
+%\author{\fnms{} \snm{}}
+
+\runauthor{}
+
+\address[addr1]{}
+
+%\thankstext{<id>}{<text>}
+
+\end{aug}
+
 \begin{abstract}
-%In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
-In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity.
-To achieve a scalable model, we propose independent priors in order to create computational speeds and gains, where we relax the one-to-one matching assumption.
-Second, we resolve the one-to-one matching issue via a post processing step.
-Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model.
-Fourth, we explore other computational speeds up based upon storage efficient indexing.
-Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.
-
 In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
@@
 
-\noindent%
-{\it Keywords:} bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
-
-\newpage
-\spacingset{1.5}
+%% ** Keywords **
+\begin{keyword}%[class=MSC]
+bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
+%\kwd{}
+%\kwd[]{}
+\end{keyword}
+
+\end{frontmatter}
 
@@
 
-\bibliographystyle{jasa}
+\bibliographystyle{ba}
 \bibliography{biblio}

### 2021-10-14 bayesian-analysis/kundinger-ba-2021.tex
update.

@@
 \begin{frontmatter}
-\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
-%\title{\support{}}
-\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+\title{Bayesian Graphical Entity Resolution using Exchangeable Random Partition Priors}
+%\title{A Sample Document\thanksref{T1}}
+%\thankstext{T1}{<thanks text>}
+\runtitle{Entity Resolution using Exchangeable Random Partition Priors}
 
 \begin{aug}
-%\author{\fnms{} \snm{}}
-%\author{\fnms{<firstname>} \snm{<surname>}\thanksref{}\ead[label=e1]{}}
-%\and
-%\author{\fnms{} \snm{}}
-
-\runauthor{}
-
-\address[addr1]{}
+\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
+\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jerry@stat.duke.edu}}
+\and
+\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
+
+\runauthor{Brian Kundinger et al.}
+
+\address[addr1]{Department of Statistical Science,
+ Duke University,
+  P.O.\ Box 90251,
+  Durham, NC 27708, USA
+  \printead*{e1}, % print email address of "e1"
+  \printead*{e2}
+   \printead*{e3}
+}
+
+%\address[addr2]{Departments of Statistical Science and Computer Science,
+%  Duke University,
+%  P.O.\ Box 90251,
+%  Durham, NC 27708, USA
+%  \printead{e3}
+%}
 

### 2021-10-14 bayesian-analysis/kundinger-ba-2021.tex
update.

@@
 \begin{frontmatter}
-\title{Bayesian Graphical Entity Resolution using Exchangeable Random Partition Priors}
+\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
 %\title{A Sample Document\thanksref{T1}}
 %\thankstext{T1}{<thanks text>}
-\runtitle{Entity Resolution using Exchangeable Random Partition Priors}
+\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
 

### 2021-11-15 Prelim/prelim.tex
NLTCS case study

--- 2021-09-29 Prelim/prelim.tex
@@
 \begin{document}
-	\title{Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)}
+	\title{Efficient and Scalable Bayesian Bipartite Matching through Fast Beta Linkage (\texttt{fabl})}
 	%\subtitle{Using Beamer}
@@
 
-		\item Far ranging applications in business, public health, and human rights
-	\end{enumerate}
+		\item \textbf{Bipartite matching} is the specific goal of matching one record in one dataset to most one match in another dataset
+	\end{itemize}
+	\end{frame}
+
+%\begin{frame}{Record Linkage in Practice}
+%	\begin{columns}
+%		\begin{column}{0.48\textwidth}
+%			\includegraphics<1->[width = \textwidth, height = .9\textwidth ]{ted_article2.png}
+%			
+%		\end{column}
+%		\begin{column}{0.48\textwidth}
+%			\includegraphics<2->[width = \textwidth, height = .9\textwidth ]{dnc_header.png}
+%		\end{column}
+%	\end{columns}
+%\end{frame}
+
+\begin{frame}{Record Linkage in Practice}
+	\centering
+	\includegraphics<1->[width = .8\textwidth, height = .6\textwidth ]{graphics/syria_article_big.png}
 \end{frame}
@@
 \begin{frame}{Record Linkage in Practice}
-	\begin{columns}
-		\begin{column}{0.48\textwidth}
-			\includegraphics<1->[width = \textwidth, height = 1.2\textwidth ]{ted_article2.png}
-
-		\end{column}
-		\begin{column}{0.48\textwidth}
-			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{dnc_header.png}
-		\end{column}
-	\end{columns}
+	\includegraphics<1->[width = \textwidth, height = .6\textwidth ]{graphics/dnc_big.png}
 \end{frame}
@@
 \begin{frame}{Linkage for Downstream Analysis}
-	\includegraphics<1>[width = \textwidth, height = .7\textwidth ]{datasets.png}
-	\includegraphics<2>[width = \textwidth, height = .7\textwidth ]{datasets_linked.png}
-
+	\includegraphics<1>[width = \textwidth, height = .7\textwidth ]{graphics/Slide1.png}
+	\includegraphics<2>[width = \textwidth, height = .7\textwidth ]{graphics/Slide2.png}
 \end{frame}
@@
 \begin{frame}{Linkage through Comparison Vectors}
-		\includegraphics[width = \textwidth, height = .7\textwidth ]{fs_graphic.png}
+	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide3.png}
+	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide4.png}
+	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide5.png}
+\end{frame}
+
+\begin{frame}{Linkage through Comparison Vectors}
+	Represent linkage structure through vector $\mathbf{Z} = \{Z_1, \ldots, Z_{n_B}\}$, where
+\begin{center}
+	$Z_j = \begin{cases} 
+		i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
+		n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
+	\end{cases}$
+\end{center}
+\end{frame}
+
+\begin{frame}{Linkage through Comparison Vectors}
+	\includegraphics[width = \textwidth, height = .6\textwidth ]{graphics/Slide6.png}
 \end{frame}
@@
 		\begin{column}{0.6\textwidth}
-			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide1.png}
-			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide2.png}
-
+			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide1.png}
+			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide2.png}
 		\end{column}
@@
 			\begin{itemize}
-				\item<3-> scalable to large datasets (\texttt{fastlink}, Enamorado et al 2017)
-				\item<4-> no transitive closure, requires post-processing
+				\item<3-> scalable to large datasets (\texttt{fastlink}, Enamorado et al 2019)
+				\item<4-> not bipartite, requires post-processing
 				\pause 
@@
 
-\begin{frame}{Sadinle (2017) - Beta Record Linkage}
+\begin{frame}{Sadinle (2017)}
 	\begin{columns}
 		\begin{column}{0.6\textwidth}
-			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide1.png}
-			\includegraphics<2>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide3.png}
-			\includegraphics<3>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide4.png}
-			\includegraphics<4->[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide5.png}
-
+			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide1.png}
+			\includegraphics<2>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide3.png}
+			\includegraphics<3>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide4.png}
+			\includegraphics<4->[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide5.png}
 		\end{column}
@@
 		\begin{column}{0.6\textwidth}
-			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide1.png}
-			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{rl_graphics_all/Slide6.png}
-
+			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide1.png}
+			\includegraphics<2-4>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide6.png}
+			\includegraphics<5>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide7.png}
+			\includegraphics<6>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide8.png}
 		\end{column}
@@
 			\begin{itemize}
-				\item<3-> simple mathematical change, large computational gains
-				\item<4-> minimal loss of accuracy for linkage and other parameters, minimal post-processing
+				\item<3-> relaxation proposed by Heck Wortman (2019)
+				\item<4-> minimal loss of accuracy, large computational gains
+				\item<5-> allows for "one to many" matchings
+				\item<6-> simple postprocessing to obtain bipartite matching
 			\end{itemize}
@@
 \section{Fast Beta Linkage}
-\begin{frame}{Notation}
-\begin{itemize}
-	\item File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$.
+%\begin{frame}{Notation}
+%\begin{itemize}
+%	\item File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$, with $n_A \geq n_B$.
 We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.
+%	
+%	\item $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors where $\gamma_{ij}^f \in \{1, \ldots, L_f\}$
+%	
+%	\item $Z_j = \begin{cases} 
+%	i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
+%	n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
+%	\end{cases}$
+%	
+%	\item $m_{fl} = P\left(\gamma_{ij}^f = l |Z_j = i\right)$
+%	
+%	\item $u_{fl} = P\left(\gamma_{ij}^f = l |Z_j \neq i\right)$
+%	
+%	\item $\lambda = P(Z_j \leq n_A)$
+%\end{itemize}
+%\end{frame}
 
-	\item $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors where $\gamma_{ij}^f \in \{1, \ldots, L_f\}$
-
-	\item $Z_j = \begin{cases} 
-	i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
-	n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
-	\end{cases}$
-
-	\item $m_{fl} = P\left(\gamma_{ij}^f = l |Z_j = i\right)$
-
-	\item $u_{fl} = P\left(\gamma_{ij}^f = l |Z_j \neq i\right)$
-
-	\item $\lambda = P(Z_j \leq n_A)$
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Fast Beta Linkage (fabl)}
+\begin{frame}{Fast Beta Linkage (\texttt{fabl})}
 	$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$
@@
 	$$\mathbf{u_{f}} \sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{fL_f})$$
-	$$Z_j | \lambda =
-	\begin{cases} 
-	\frac{1}{n_A}\lambda  & z_j \leq n_A; \\
-	1-\lambda &  z_j  = n_A + 1 \\
+	$$Z_j | \pi	\begin{cases} 
+	\frac{\pi}{n_A}  & z_j \leq n_A; \\
+	1-\pi &  z_j  = n_A + 1 \\
 	\end{cases}$$
-	$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda})$$
+	$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
 
-	Model specification allows for \textbf{parallel/distributed} computing, \textbf{hashing} of comparison vectors, and \textbf{storage efficient indexing (SEI)}
+	Model specification allows for \textbf{hashing} of comparison vectors, \textbf{distributed} computing, and \textbf{storage efficient indexing (SEI)}
 \end{frame}
@@
 \begin{frame}{Hashing}
+		\begin{columns}
+		\begin{column}{0.6\textwidth}
 			\begin{itemize}
-		\item Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $h(i, j)=p.$
-
-		\item Reduce data to sufficient statistics
+				\item<1-> Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records (Enamorado et al 2019).
 				\begin{itemize}
-		\item $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
-		\item $H_{p_j} = ||r_{p_j}||$
-		\item $H_p = \sum_{j} H_{p_j}$
+					\item<2-> $L = \{3, 3, 2, 2\}$ implies 36 unique patterns 
 				\end{itemize} 
-
-		\item Run Gibbs sampler at level of agreement patterns, not record pairs
-		\begin{itemize}
-		\item Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
-		\item Use number of matches for each pattern to update $m$ and $u$
-		\item Back fill record labels at the end through $r_{p_j}$
+				\item<3-> When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p.$
+				\item<4-> Allows us to compute sufficient statistics and reduce computational complexity from $O(n_A \times n_B)$ to $O(P \times n_B)$
 			\end{itemize}
-
-		\item Reduces computational complexity from $O(n_A \times n_B \times F)$ to $O(P \times n_B \times F)$
-	\end{itemize}
+		\end{column}
+		\begin{column}{0.38\textwidth}
+			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{graphics/unique_patterns.png}
+		\end{column}
+	\end{columns}
 \end{frame}
 
-\begin{frame}{Managing Large Data}
-	\begin{itemize}
-		\item \textbf{Distributed Computing} - Partition data in to chunks $\{A_I\}$ and $\{B_J\}$. Compare records, hash results, compute summary statistics in parallel, and synthesize results.
-
-		\item \textbf{Storage Efficient Indexing (SEI)} - Store at most small number $R$ many record labels in each $r_{p_j}$, remove highly unlikely record labels from memory.
-Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$.
-
-		\item Hashing plus SEI can reduce memory requirements by >99%
-		\begin{itemize}
-			\item Simulation of $20,000 \times 20,000$ linkage task with 4 fields. Naive approach requires 6.4GB of storage for all-to-all comparisons, hashing and SEI requires 90MB.
-		\end{itemize}
-	\end{itemize}
+\begin{frame}{Hashing}
+	\centering
+	\includegraphics<1>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide17.png}
+	\includegraphics<2>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide18.png}
+	\includegraphics<3>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide19.png}
+	\includegraphics<4>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide20.png}
+	\includegraphics<5>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide21.png}	
 \end{frame}
 
-\section{Results}
+\begin{frame}{Hashing to Scale to Large Linkage Tasks}
+	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide23.png}
+	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide24.png}
+	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide25.png}
+	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide26.png}
+	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide27.png}
+\end{frame}
 
-\begin{frame}{Three Simulation Studies}
+\begin{frame}{Hashing for Speed (Efficient Gibbs Sampling)}
+	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide37.png}
+	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide38.png}
+	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide39.png}
+	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide40.png}
+	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide41.png}
+	\includegraphics<6>[width = \textwidth, height = .6\textwidth ]{graphics/Slide42.png}
+\end{frame}
+
+%\begin{frame}{Managing Large Data}
+%	\begin{itemize}
+%		\item \textbf{Distributed Computing} - Partition data in to chunks $\{A_I\}$ and $\{B_J\}$. Compare records, hash results, compute summary statistics in parallel, and synthesize results.
+%		
+%		\item \textbf{Storage Efficient Indexing (SEI)} - Store at most small number $R$ many record labels in each $r_{p_j}$, remove highly unlikely record labels from memory.
+Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$.
+%		
+%		\item Hashing plus SEI can reduce memory requirements by $>99$.
+%		\begin{itemize}
+%			\item Simulation of $20,000 \times 20,000$ linkage task with 4 fields. Naive approach requires 6.4GB of storage for all-to-all comparisons, hashing and SEI requires 90MB.
+%		\end{itemize}
+%	\end{itemize}
+%\end{frame}
+
+\section{Simulation Studies}
+
+%\begin{frame}{Three Simulation Studies}
+%	\begin{itemize}
+%		\item We compare \texttt{fabl} against \texttt{BRL} in three simulation studies
+%		\begin{itemize}
+%			\item Measure precision and recall on 100 simulated datasets and varying levels of error and duplication across files
+%			\item Measure speed when both $n_A$ and $n_B$ are increasing
+%			\item Measure speed when $n_A$ is increasing and $n_B = 500$ is fixed. 
+%		\end{itemize}
+%	\end{itemize}
+%\end{frame}
+
+\begin{frame}{Speed Simulation 1}
+	\begin{columns}
+		\begin{column}{0.38\textwidth}
 			\begin{itemize}
-		\item We compare \texttt{fabl} against \texttt{BRL} in three simulation studies
+				\item $F$ = 5 comparison fields
+				\item L = \{2, 2, 2, 2, 2\}, all binary comparisons
+				\item 32 possible patterns
+				\item Increase both $n_A$ and $n_B$
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.58\textwidth}
+			\includegraphics[width = \textwidth, height = \textwidth ]{../notes/figures/sadinle_speed_plot_slides.png}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Speed Simulation 2}
+	\begin{columns}
+		\begin{column}{0.38\textwidth}
 			\begin{itemize}
-			\item Measure precision and recall on 100 simulated datasets and varying levels of error and duplication across files
-			\item Measure speed when both $n_A$ and $n_B$ are increasing
-			\item Measure speed when $n_A$ is increasing and $n_B = 500$ is fixed. 
+				\item $F$ = 5 comparison fields
+				\item L = \{2, 2, 2, 2, 2\}, all binary comparisons
+				\item 32 possible patterns
+				\item Fix $n_B = 500$, increase $n_A$
 			\end{itemize}
+		\end{column}
+		\begin{column}{0.58\textwidth}
+			\includegraphics[width = \textwidth, height = \textwidth ]{../notes/figures/speed_plot_fixed_nB_slides.png}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Accuracy Simulation}
+	\begin{itemize}
+		\item Sadinle (2017) used 900 simulated linkage tasks to show accuracy of \texttt{BRL}
+		\item Find matches across two datasets, each with 500 records and 4 fields in common. 
+		\item One, two or three errors across matching records
+		\item 10\% matching, 50\% matching, or 90\% matching
+		\item Calculate recall, precision, and F-measure 
 	\end{itemize}
@@
 \begin{frame}{Accuracy Simulation}
-	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/sadinle_sim_plot.png}
+	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/sadinle_sim_plot2.png}
 \end{frame}
 
-	\begin{frame}{Speed Simulation 1}
-	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/sadinle_speed_plot2.png}
+%\begin{frame}{Speed Simulation 1}
+%	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/sadinle_speed_plot2.png}
+%\end{frame}
+%
+%	\begin{frame}{Speed Simulation 2}
+%	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/speed_plot_fixed_nB.png}
+%\end{frame}
+
+%\section{El Salvador Homicide Case Study}
+%
+%\begin{frame}{El Salvador Case Study}
+%	\begin{itemize}
+%		\item El Salvadoran Civil War (1980-1991)
+%		\item Lists of casualties collected by multiple organizations
+%		\begin{itemize}
+%			\item Salvadoran Human Rights Commissions (CDHES), $n_A = 4420$
+%			\item El Rescate - Tutela Regal (ERTL), $n_B = 1323$
+%			\item Features include first name, last name, date and place of death
+%		\end{itemize}
+%		\item We aim to find duplicate records across files
+%		\begin{itemize}
+%			\item Particularly difficult because families are often killed together, and some children share names with parents
+%		\end{itemize}
+%
+%	\end{itemize}
+%\end{frame}
+%
+%\begin{frame}{Run Time}
+%		\begin{columns}
+%		\begin{column}{0.48\textwidth}
+%			\centering
+%			\includegraphics<1->[width = .6\textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/time_table_big_P2.png}
+%			
+%		\end{column}
+%		\begin{column}{0.48\textwidth}
+%			\centering
+%			\includegraphics<2>[width = .6\textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/time_table_small_P2.png}	
+%		\end{column}
+%	\end{columns}
+%\end{frame}
+%
+%\begin{frame}{Posterior Inference}
+%	\includegraphics[width = \textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/m_posterior_smallP.png}
+%\end{frame}
+%
+%\begin{frame}{Posterior Inference}
+%	\includegraphics[width = \textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/DID_distribution_smallP_bayes.png}
+%\end{frame}
+%
+%\begin{frame}{Violations of One-to-One Matching}
+%	\centering
+%	\includegraphics[width = .9\textwidth, height = .4\textwidth ]{../notes/figures/el_salvador/rosa_records.png}
+%\end{frame}
+%
+%\begin{frame}{Violations of One-to-One Matching}
+%	\includegraphics[width = \textwidth, height = .6\textwidth ]{../notes/figures/el_salvador/bad_mixing.png}
+%\end{frame}
+%
+
+\section{Conclusion}
+
+\begin{frame}{Benefits of \texttt{fabl}}
+	\begin{itemize}
+		\item Faster computation for larger linkage tasks
+
+		\item Accurate estimation of linkage structure $\mathbf{Z}$, and additional parameters $\mathbf{m}$ and $\mathbf{u}$
+
+		\item Bayesian model with natural uncertainty quantification
+	\end{itemize}
 \end{frame}
 
-	\begin{frame}{Speed Simulation 2}
-	\includegraphics[width = \textwidth, height = .7\textwidth ]{../notes/figures/speed_plot_fixed_nB.png}
-\end{frame}
+\begin{frame}{Extensions and Future Directions}
+	\begin{itemize}
+		\item Linkage when reliability of information and rates of matching differs by subgroup in the data
 
-\begin{frame}{Extensions and Directions}
-	\begin{itemize}
-	\item Linkage in the face of duplicates within datasets
-
-	\item Models that allow reliability of information to differ by subgroup in the data
+		\item Linkage when there are duplicates within datasets
 
@@
 
+\section{Appendix}
+
+\begin{frame}{Storage Efficient Indexing}
+	\begin{itemize}
+		\item Storing $n_A \times n_B$ hashed values burdensome for large tasks
+		\item Overwhelming majority of record pairs show nonagreement
+		\item<3> Correct counts for calculations stored in $\mathcal{H}$
+	\end{itemize}
+	\includegraphics<1>[width = \textwidth]{graphics/before_SEI.png}
+	\includegraphics<2->[width = \textwidth]{graphics/after_SEI2.png}
+\end{frame}
+
+%		\item Reduce data to sufficient statistics
+%		\begin{itemize}
+%		\item $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
+%		\item $H_{p_j} = ||r_{p_j}||$
+%		\item $H_p = \sum_{j} H_{p_j}$
+%		\end{itemize}
+%		
+%		\item Run Gibbs sampler at level of agreement patterns, not record pairs
+%		\begin{itemize}
+%		\item Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
+%		\item Use number of matches for each pattern to update $m$ and $u$
+%		\item Back fill record labels at the end through $r_{p_j}$
+%	\end{itemize}
+%
+%		\item Reduces computational complexity from $O(n_A \times n_B \times F)$ to $O(P \times n_B \times F)$
+
 \end{document}

### 2021-11-15 methods_paper/fabl-paper.tex
NLTCS case study

@@
 
-\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+\title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
 \author{Brian Kundinger\textsuperscript{a} \and
@@
 \begin{abstract}
-%In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing.
-In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity.
-To achieve a scalable model, we propose independent priors in order to create computational speeds and gains, where we relax the one-to-one matching assumption.
-Second, we resolve the one-to-one matching issue via a post processing step.
-Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model.
-Fourth, we explore other computational speeds up based upon storage efficient indexing.
-Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.
-
-In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
+Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage.
+These have the crucial advantage of quantifying uncertainty from imperfect linkages.
+However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
+We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.
+Specifically, in \texttt{fabl} we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
+This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
+Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
+
+%In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
 Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
@@
 \noindent%
-{\it Keywords:} bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
+{\it Keywords:} record linkage, fusion, hashing, distributed computing, Fellegi Sunter
 
@@
 
-	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}.
-This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others.
+	In many data analysis tasks, analysts seek to identify duplicate records across two databases.
+This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, dalzell2018regression, tang2020}.
 In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.
@@
 Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test.
-Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.
-Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.
+Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english, enamorado2019using}.
 	% Possible citation for bipartite matching. 
@@
 \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records.
-Most recently, \cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler.
+\cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout his Gibbs sampler.
 Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
-Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.
-
-	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the the Beta Record Linkage model for increased efficiency and scalability.
-Specifically, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
-This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs; (2) compute the pairwise record comparisons over large datasets via parallel computing; and (3) reduce memory costs through our proposed method of \emph{storage efficient indexing}.
-Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
+Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets.
+
+	In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
+As suggested by \cite{heck2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
+This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We argue that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
 Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
 
-	In what follows, Section~\ref{sec:review-of_prior-work} reviews the prior work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
+	In what follows, Section~\ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
 Section~\ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching.
-	Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake.
-In addition, we prove the computational complexity of \texttt{fabl}.
+Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}.
 Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War.
-Section~\ref{discussion} concludes our paper with a discussion of directions for future work.
+Section~\ref{discussion} concludes with a discussion of directions for future work.
 
@@
 In the context of bipartite matching, we assume that there are duplications across, but not within, each database.
-Under this framework, the set of matches across datasets can be represented in two equivalent ways.
+Under this framework, the set of matches across databases can be represented in two equivalent ways.
 First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
@@
 \end{align}
-Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers.
-More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that
+Though intuitive, this sparse matrix representation can become costly for large linkage tasks.
+More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that
 \begin{align}
@@
 \end{align}
-Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.
+Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.
 
@@
 
-\subsection{Comparison Data}
-\label{comparison-data}
+%\subsection{Comparison Vectors}
+%\label{comparison-vectors}
 
 Intuitively, co-referent records (those that refer to the same entity) should be similar; records that are not co-referent should not be similar.
-\cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that
-$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
+\cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
 
 The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables.
-For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance.
-We can then propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
+We then can propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
 Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$.
@@
 
-%In what follows we use notation that reflects that these comparison vectors can be modelled in number of ways.
-In practice however, this binning in beneficial because it allows us to model any type of comparison through a multinomial distribution.
-In a Bayesian context, Dirichlet priors for
+In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
+In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field.
+The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
+However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations.
+
+In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
+This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
+In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.
 
@@
 \label{fellegi-sunter}
-The comparison data $\bm{\gamma}_{ij}$ is not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
-This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
+The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data.
+This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
 \begin{align}
@@
 \end{align}
-as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large if there is strong evidence of the pair being a match, and small otherwise.
+as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
 
@@
 \end{align}
-where $p$ is the proportion of records that match, and $\bm{m}$ and $\bm{u}$ are vectors of parameters for the matching and nonmatching record pairs.
+where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters $m_{fl} = P(\gamma_{ij}^f = l|Z_j =i)$ for the matching record pairs, and $\bm{u}$ is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
 These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.
 
-After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined).
-This is a problem in practice as transitive closures are violated.
-Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
+After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+This is a problem in practice as transitive closures can be violated.
+Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
 
@@
 The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
-Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate for bipartite matching.
+Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
 
@@
 
-One criticism of using mixture models is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
+One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
 One way of resolving this issue to incorporate the constraint directly into the model as follows:
@@
 where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
+For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
 $$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
@@
 where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
-records when sampling \(Z_j\). This creates a dependency that makes the
-sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.
-
-\subsection{Blocking}
-
-In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden.
-In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field.
-The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.
-
-In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without the need for blocking.
-This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
-In practice, the following method can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.
+This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
+This creates a dependency that makes the sampler inherently serial, which can slow the sampler when working on large linkage tasks.
 
@@
 
-Recall our compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
-In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we propose independent priors for each component $Z_j$.
-We denote the \emph{fast beta prior} as follows:
+Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
+In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{heck2019} and use independent priors for each component $Z_j$.
+However, unlike Heck Wortman who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
+We denote the fast beta prior as follows:
 
@@
 $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-We encourage the reader to integrate out $\pi$ from $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.
-
-Note that linkage in this setting is conducted at the \emph{record} level, rather than at the \emph{record pair} level as in the Fellegi Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of \emph{records} in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of \emph{record pairs} that are matches.
-We find $\pi$ to be more interpretable parameter than $\lambda$; especcially in the bipartite case, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
-necessarily tends towards 0 as the size of the linkage task grows.
-Additionally, we emphasize that while the Fellegi Sunter model makes $n_1 \times n_2$ many independent matching decisions and \texttt{BRL} makes $n_2$ many dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$ many independent matching decisions.
-As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows us to achieve many of the accuracy gains from \texttt{BRL} while maintaining the efficiency gains possible by exploiting independence.
-
-Importantly, this independence means that we have relaxed the one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desirable in its own right; \cite{Newcombe1963} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable.
-In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in  section~\ref{bayes-estimate} to resolve illegal matchings.
+One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
+We note here that prior from \cite{heck2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
+
+Note that linkage in this setting is conducted at the record level, rather than at the record pair level as in the Fellegi Sunter model.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of record pairs that are matches.
+We find $\pi$ to be more interpretable parameter than $\lambda$ in the bipartite case.
+There are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
+Additionally, while the Fellegi Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions.
+As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to achieve the benefits from \texttt{BRL} while making efficiency gains possible by exploiting independence.
+
+Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
+This behavior may be desirable; for example, we may be interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable \citep{Newcombe1963} .
+In situations when a bipartite estimate is desired, we can use the post processing procedure shown in section~\ref{bayes-estimate} to resolve many-to-one matchings.
 
@@
 
-\hypertarget{fabl-model}{%
-	\subsection{Model Specification}
-	\label{model-specification}}
-
-As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent.
+%\hypertarget{fabl-model}{%
+%	\subsection{Model Specification}
+%	\label{model-specification}} 
+%
+%As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent.
 Additionally, we assume that any missing data is missing at random.
@@
 In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, $\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
-The model can then be specified as follows:
+
+%The model can then be specified as follows:
+With the \texttt{fabl} prior defined, we present the full model as follows:
 
@@
 \end{align*}
-This factorization leads to following Gibbs Sampler:
-
-\underline{Sample $\Phi^{(s+1)}|\Gamma, \bm{Z}^{(s)}$:}
-
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have
+
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. Thus we have
 
@@
 
-\underline{Sample $\pi^{(s+1)}|\bm{Z}^{(s)}$:} As a function of
-\(\pi\), the linkage structure parameter \(\bm{Z}\) is sequence
-of successes (when \(z_j < n_A + 1\)) and failures (when
-\(z_j = n_A + 1\)), and therefore
-\(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is
-determined only by the number of duplicates
-$n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
-$\bm{Z}$. Thus we have
+As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
+Thus we have
 
@@
 &\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1} \\
-\implies &\pi^{(s+1)}|\bm{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})
 \end{align*}
-
-\underline{Sample $\bm{Z}^{(s+1)}|\Gamma, \Phi^{(s+1)}, \pi^{(s+1)}$:}
+Thus the posterior $\pi^{(s+1)}|\bm{Z}^{(s+1)}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
+
 Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
 only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
-denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\), and note that as a function of \(Z_j\), the likelihood
+denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).  As a function of \(Z_j\), 
 \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\)
@@
 where
-$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
-The interested reader should note that these are precisely the
-likelihood ratios used in the Fellegi-Sunter model to classify matches
-and non-matches, and we therefore refer to \(w_{ij}\) as the
-\emph{Fellegi Sunter weights}.
+%$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
+$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$.
+
+These \(w_{ij}\) are precisely the likelihood ratios used in the Fellegi-Sunter model to classify matches and non-matches, and we therefore call them the Fellegi Sunter weights.
 
@@
 	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1} \\
-	&\implies Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
+	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}
+\end{align*}
+Thus we have
+$$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
 	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
 	1-\pi &  z_j  = n_1 + 1 \\
-\end{cases}
-\end{align*}
-
-In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
-integrate over the posterior of \(\pi\) and rearrange terms to
-produce the final full conditional:
+\end{cases}$$
+Finally, we integrate over the posterior of \(\pi\) and rearrange terms to produce the final full conditional:
 
@@
 $$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
-In this paper, we adopt losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule
+Here, we set $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule
 
@@
 	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
-	0,  & \text{otherwise} ; \\
+	0,  & \text{otherwise}. \\
 \end{cases}\]
@@
 For a more in-depth explanation of this function and the induced Bayes
-estimate, see \citep{sadinle_bayesian_2017}.
-
-Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+estimate, see \cite{sadinle_bayesian_2017}.
+
+Since the Gibbs sampler does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
 It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable.
-We will explore this issue more deeply through the El Salvador homicide case study.
+We will explore this issue through the El Salvador homicide case study.
 
 To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.
-A similar approach can be see in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities, and is justified within this Bayesian framework.
+In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the highest posterior probability, and declare the other to have no match.
+A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
 
@@
 
-To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
+To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
 The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
@@
 $P = |\mathcal{P}|$ is the total number of unique agreement
-patterns; observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not depend on \(n_1\) or \(n_2\).
-In this context, the integers $\{1, \ldots, P\}$ serve as \emph{hashed values} that encode the same information as the orignial vectors themselves.
-Whenever possible, we conduct calculations over these \(P\) agreement
-patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-
+patterns.
+\(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
+In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the orignial vectors themselves.
+Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
 Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
@@
 
-$$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$
+$$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}.$$
 
@@
 $$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
-For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.
-Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for our efficient posterior inference.
-
-\hypertarget{SEI}{%
-	\subsection{Storage Efficient Indexing}\label{SEI}}
+For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will shown in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
+Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.
+
+\hypertarget{scaling}{%
+	\subsection{Scaling to Large Linkage Tasks}\label{scaling}}
 
@@
 
-To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI).
-In standard indexing, one decides a priori certain criteria that they expect all true matching pairs to satisfy,
+To address this issue, we propose a new method called storage efficient indexing (SEI). In standard indexing, one decides a priori certain criteria that all true matching pairs must satisfy, 
 and labels any record pairs that do not meet that criteria as
@@
 similarity score on a field deemed to be important (like first name), or
-only pairs with exact matching on a specified number of fields. While
+only pairs that exactly match on a specified number of fields. While
 generally chosen to be quite loose, establishing these criteria
@@
 
-With SEI however, we can reduce comparison space, and its associated storage costs, while avoiding these drawbacks.
+With SEI however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks.
 Observe that all record $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{x}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
@@
 
-\hypertarget{distributed}{%
-	\subsection{Distributed Computation for Pairwise Comparison}\label{distributed}}
+%\hypertarget{distributed}{%
+%	\subsection{Distributed Computation for Pairwise Comparison}\label{distributed}}
 
@@
 
-\underline{Sample $\Phi^{(s+1)}|\tilde{\Gamma}, \bm{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
+Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
 These are simple calculations, but become computationally burdensome when working on large linkage tasks.
@@
 number of matching record pairs with agreement pattern $p$.
-We can also easily calculate the number of nonmatching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\).
-Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes simply
+We also can easily calculate the number of non-matching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\).
+Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
 
 $$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
-$$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$
-
-Note that these constitute \(P\) many summations over \(n_B\) quantities, and thus avoid the \(n_A \times n_B\) summation from the original method.
-
-\underline{Sample $\bm{Z}^{(s+1)}|\tilde{\Gamma}, \Phi^{(s+1)}$:} Although sampling \(Z_j\) from a the
-full conditional provided earlier is conceptually straightforward, it
-becomes computational burdensome when $n_1$ is large.
+$$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
+
+These constitute $\sum L_f$ many summations over $P$ quantities.
+
+Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
 This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
-To speed up
-computation, we break this sampling step into two simpler steps. First,
-we calculate the Fellegi Sunter weight \(w_{p}\) associated with each
-unique pattern and sample the agreement pattern between \(j\) and its
-potential match. Second, we sample the record label uniformly among
-records associated with that agreement pattern for that particular
-\(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement
-pattern between \(j\) and its potential match, and say
-\(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + 1\). Then,
+To speed up computation, we break this sampling step into two.
+First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link.
+Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\).
+More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + 1\).
+Then,
 
@@
 \end{cases}$$
-$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) \begin{cases} 
+$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
 	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
-	0 & \text{otherwise} \\
+	0 & \text{otherwise.} \\
 \end{cases}$$
@@
 Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
-Thus we complete the entire Gibbs procedure first at the level of the \(P\) agreement patterns with the first equation above.
-After all Gibbs iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.
-
-We emphasize the computational gains of this split sampler: the first step is a sample from $P$ many options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
-These changes in themselves drastically improves the speed of the sampler, and each can be parallelized if desired for additional computational gains.
-
-To aid the reader, we provide summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
+Thus we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with the first equation above.
+After all iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.
+
+We emphasize the computational gains of this split sampler: the first step is a sample from $P$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
+These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
+We provide summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
 
@@
 	\subsection{Computational Complexity}\label{computational-complexity}}
-In this section, we prove the computational complexity of \texttt{fabl} in Lemma \ref{lemma:fabl}.
-Let $L^{*} = \sum_{f=1}^F L_f$ denote the total number of levels used across all features in the linkage task.
+In this section, we prove the computational complexity of \texttt{fabl}.
 
@@
 Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the overall complexity to $O(\tilde{L}^{*}P)$, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.
+However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(\tilde{L}^{*}P)$, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.
 The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
 With hashing, we split this sampling into two steps.
-First we sample that agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
+First we sample the agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
 Thus the complexity of sampling $\bm{Z}$ is $O(n_2 P)$.
@@
 
-We demonstrate the accuracy, speed, and scale of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
-
-\hypertarget{accuracy}{%
-	\subsection{Accuracy under Full Estimates}\label{accuracy}}
-
-We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap.
-We use first name, last name, age, and occupation for this linkage, where age and occupation take on 9 and 10 unique values respectively.
-We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table~\ref{Tab:sadinle_simulation_cutoffs}.
-Each simulation identifies duplicated individuals between two datasets, each with 500 records.
-We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets.
-We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
-Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}.
-
-\begin{table}[h!]
-	\centering
-	\begin{tabular}[t]{ll|llll}
-		\hline
-		\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
-		\cline{3-6}
-		Fields & Similarity & 1 & 2 & 3 & 4\\
-		\hline
-		First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (5, .1]\\
-		\hline
-		Age and Occupation & Binary & Agree & Disagree &  & \\
-		\hline
-	\end{tabular}
-\caption{Construction of comparison vectors for accuracy study with simulated datasets}
-\label{Tab:sadinle_simulation_cutoffs}
-\end{table}
-
-We compare \texttt{fabl} to \texttt{BRL} in terms of precision, recall, and f-measure, which are commonly defined record linkage evaluation metrics that are defined in \citep{christen_2012}.
-In Figure~\ref{fig:sadinle_simulation}, we see that we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and high overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, since both methods exhibit high variance in performance in this setting, we remain confident in \texttt{fabl}'s overall accuracy.
-
-\begin{figure}[h!]
-	\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
-\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
-For each level of overlap and each level of error, we have 100 paired sets of 500 records.
-We see comparable performance for all levels of error and overlap.}
-\label{fig:sadinle_simulation}
-	\end{center}
-\end{figure}
-
-\hypertarget{partial}{%
-	\subsection{Accuracy under Partial Estimates}\label{partial}}
-
-By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
-In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
-Thus, instead of recall, we use the \emph{negative predictive value} (NPV), defined as the proportion of non-links that are actual non-matches.
-Mathematically, $\text{NPV} = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
-We continue to use the precision, which is renamed the \emph{positive predictive value} (PPV) in this context.
-Lastly, we also report the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.
-
-In Figure~\ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
-However, with high amounts of overlap, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
-Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probability of matches can at times be more split across different records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
-In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
-
-\begin{figure}[h!]
-\begin{center}
-	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
-	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
-	\label{fig:sadinle_simulation_partial}
-\end{center}
-\end{figure}
+We demonstrate the speed, and accuracy of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
 
@@
 
-To demonstrate speed, we generate comparison vectors from pre-specified
-distributions so that we can easily increase the size of the linkage
-problem.
+In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
 The distributions used (see Table~\ref{Tab:distributions}) are meant to emulate the rates of agreement across first name, last name, day of birth, and month of birth.
 For example, $u^{\text{month, agree}} = P(\text{Records have same birth-month} \mid \text{Nonmatch}) = \frac{1}{12}$.
-The parameters for year would vary largely by context, so we chose them here to adjust the desired difficulty of the linkage task.For simplicity, we consider only exact matching, so a vector (1, 0) corresponds to
-agreement and (0, 1) to disagreement. We simulate these data for
-different values of \(n_1\) and \(n_2\), and compare the run-time of
-\texttt{fabl} against \texttt{BRL}. Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is
-consistently attained in the larger simulations.
-
-\begin{table}[h!]
+The parameters for year would vary largely by context, so we choose them here to adjust the desired difficulty of the linkage task.
+For simplicity, we consider only exact matching, so a vector (1, 0) corresponds to agreement and (0, 1) to disagreement.
+We simulate these data for different values of \(n_1\) and \(n_2\), and compare the run-time of \texttt{fabl} against \texttt{BRL}.
+Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in the larger simulations.
+
+\begin{table}[t]
 	\centering
@@
 
-In Figure~\ref{fig:speed1}, we see that at low data size, \texttt{BRL} outperforms, but that
-\texttt{fabl} is significantly faster at handling larger data. In
-particular, run-time for \texttt{BRL} seems to grow quadratically (or
-linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for
-\texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
-
-\begin{figure}[h!]
+We note here that \texttt{BRL} is coded in C, which makes for unfair
+comparison against \texttt{fabl}, currently built in R.
+However, The computational speed gains for \texttt{fabl} persist, especially for the larger simulations.
+Additionally, although \texttt{fabl} is amenable to parallelization, this simulation was run on a single core.
+Running \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
+
+In Figure~\ref{fig:speed1}, we see that at low data size, \texttt{BRL} outperforms, but that \texttt{fabl} is significantly faster at handling larger data.
+In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
+
+\begin{figure}[t]
 	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
@@
 \begin{figure}[h!]
-	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB} 
+	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
 		\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
@@
 
-We note here that \texttt{BRL} is coded in C, which makes for unfair
-comparison against \texttt{fabl}, currently only built in R.
-Additionally, although \texttt{fabl} is amenable to parallelization,
-this simulation was run on a single core. Running \texttt{fabl} in C++
-with paralellization for the hashing step and sampling the matching
-status of the record pairs should lead to even more computational gains.
-
-\hypertarget{scale}{%
-	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
-
-To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
+\hypertarget{accuracy}{%
+	\subsection{Accuracy under Full Estimates}\label{accuracy}}
+
+We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap.
+To generate these data, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
+Fields are then chosen for distortion uniformly at random.
+Names are subject to string insertions, deletions and substitutions, as well as common keyboard, phonetic, and optical recognition errors.
+Age and occupation are distorted through keyboard errors and missingness.
+For further details on the generation of these data, see \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}.
+
+We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table~\ref{Tab:sadinle_simulation_cutoffs}.
+Each simulation identifies matched individuals between two datasets, each with 500 records.
+We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets.
+We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
+We calculate Bayes estimates of the linkage structure though the losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$.
+Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}.
+
+\begin{table}[t]
+	\centering
+	\begin{tabular}[t]{llllll}
+
+		\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
+		\cline{3-6}
+		Fields & Similarity & 1 & 2 & 3 & 4\\
+		\hline
+		First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (.5, .1]\\
+		Age and Occupation & Binary & Agree & Disagree &  & \\
+		\hline
+	\end{tabular}
+\caption{Construction of comparison vectors for accuracy study with simulated datasets}
+\label{Tab:sadinle_simulation_cutoffs}
+\end{table}
+
+We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and f-measure, as defined in \cite{christen_2012}.
+Recall is the proportion of true matches found by the model, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$, and precision is proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
+F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
+In Figure~\ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
+In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, we remain confident in \texttt{fabl}'s overall accuracy.
+
+%Each metric is provided explicitly below:
+%$$\text{Recall } = \frac{\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1)}{\sum_{j=1}^{n_2} I(Z_j \leq n_1)}$$
+%$$\text{Precision } = \frac{\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1)}{ \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)}$$
+%$$\text{F-measure } = 2 \frac{\text{Recall } + \text{Precision}}{\text{Recall } \times \text{Precision}}$$
+
+\begin{figure}[t]
+	\begin{center}
+\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
+\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
+For each level of overlap and each level of error, we have 100 paired sets of 500 records.
+We see comparable performance for all levels of error and overlap.}
+\label{fig:sadinle_simulation}
+	\end{center}
+\end{figure}
+
+\hypertarget{partial}{%
+	\subsection{Accuracy under Partial Estimates}\label{partial}}
+
+By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
+In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
+Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual non-matches.
+Mathematically, $\text{NPV} = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
+We continue to use the precision, which is renamed the positive predictive value (PPV) in this context.
+Lastly, we also report the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.
+
+In Figure~\ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
+However, with high amounts of overlap, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
+Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probability of matches can at times be more split across different records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
+
+%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
+
+%This is reasonable because large numbers of errors blur the distinction between matched and non-matched pairs, leading to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
+
+\begin{figure}[t]
+\begin{center}
+	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
+	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
+	\label{fig:sadinle_simulation_partial}
+\end{center}
+\end{figure}
+
+%\hypertarget{scale}{%
+%	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
+%
+%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
 Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
@@
 Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
-
-For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
+% 
+%For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
 We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
@@
 However, this could be substantially sped up using distributed computing.
-
-Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
-The reader will note that this slightly worse performance than witnessed in the smaller simulation studies; this is expected because it is naturally more difficult to link more files with the same amount of information.
+%
+%Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
+This is slightly worse performance than  in the smaller simulation studies; this is expected because it is naturally more difficult to link more records with the same amount of information.
 With more linkage fields, \texttt{fabl} maintains the high accuracy in Section~\ref{accuracy}.
 
-\section{Application to Civilian Casualties from the El Salvadoran Civil War}
+\section{Case Studies}
 \label{sec:case-studies}
 
-In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in both datasets.
+In our first case study, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in both datasets.
 Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-This case study also considers the ramifications of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.
-
-\subsection{El Salvadoran Civil War}
+In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study (NLTCS), a larger linkage task that is not possible under \texttt{BRL}.
+
+\subsection{Civilian Casualties from the El Salvadoran Civil War}
 \label{el_salvador}
 
-The country of El Salvador was immersed in civil war from 1980 to 1991,
-and throughout the time, several organizations attempted to document
-casualties of the conflict. When estimating the total number of
-casualties, one cannot simply sum the numbers recorded by each
-organization, as it is likely that the same individuals are recorded in
-multiple casualty lists.
+The country of El Salvador was immersed in civil war from 1980 to 1991, and throughout the time, several organizations attempted to document casualties of the conflict.
+When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
 This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications}.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela
-Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its
-acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that
-had been published throughout the conflict, and the CDHES dataset consists
-of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
+We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that had been published throughout the conflict, and the CDHES dataset consists of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
 The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both datasets are thought to be fairly reliable.
 
-There are several challenges with working with this data. Firstly, both
-datasets have been automatically digitized, which inherently leads to
-some degree of typographical error. Additionally, the only fields recorded are given name, last name, date of death, and place of death; it is relatively common for a
-parent and child to share the same given name, resulting in
-indistinguishable records for two different individuals. This last point
-nearly breaks the earlier mentioned assumption that there are no
-duplicates within files, and reveals a key difference between
-\texttt{BRL} and our proposed method.
+There are several challenges with working with this data.
+First, both datasets have been automatically digitized, which inherently leads to some degree of typographical error.
+Second, the only fields recorded are given name, last name, date of death, and place of death.
+It is relatively common for aparent and child to share the same given name, resulting in indistinguishable records for two different individuals.
+This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between
+\texttt{BRL} and \texttt{fabl}.
 
@@
 To mirror the original implementation of \cite{sadinle_bayesian_2017}, we first construct the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
-Gibbs sampling for these comparison vectors took 422 seconds for our proposed method, \texttt{fabl}, and 240 for \texttt{BRL}.
-However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, and that of the \(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 were realized in the data.
-This leads us to believe that such high number of agreement levels creates unnecessary distinctions in the data and makes the comparison vectors less interpretable.
+This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
+Gibbs sampling for these comparison vectors took 265 seconds for \texttt{fabl} and 240 for \texttt{BRL}; here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
+However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
 
 \begin{table}
-	\begin{tabular}[h!]{ll|llll}
-		\hline
-		\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
+	\begin{tabular}[t]{ll|llll}
+		\hline
+		\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
 		\cline{3-6}
@@
 The estimates of the \(\bm{m}\) parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
-Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}
+Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
 
@@
 	\centering
-	\begin{tabular}[h!]{ll|lll}
-		\hline
-		\multicolumn{2}{c|}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
+	\begin{tabular}[t]{lllll}
+		\hline
+		\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
 		\cline{3-5}
@@
 Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets.
-We calcuate posterior samples of the size of the overlap across files by finding the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
+We calculate posterior samples of the size of the overlap across files by finding the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
 The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate is quite conservative.
@@
 We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study.
-Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually; this means that after clerical review, the number of individuals replicated across datasets would fall in the interval (136, 311), encapsulating the posterior credible interval.
-More or fewer records could be indentified for clerical review by decreasing or increasing $\theta_R$.
+Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
+This means that after clerical review, the number of individuals replicated across datasets would fall in the interval (136, 311), encapsulating the posterior credible interval.
+More or fewer records could be identified for clerical review by decreasing or increasing $\theta_R$.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
 	\begin{center}
@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
 \begin{center}
 	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-\caption{Posterior estimates of m parameters with 95\% credible intervals. We note they are quite similar across the two methods.}\label{fig:m-and-u}
+\caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. We note they are quite similar across the two methods.}\label{fig:m-and-u}
 \label{fig:m-and-u}
@@
 
-\hypertarget{violations}{%
-	\subsection{Violations of One-to-One Matching}\label{violations}}
-
-As noted previously, the application exhibits many situations in which bipartite matching is difficult.
+\subsection{National Long Term Care Study}
+\label{nltcs}
+
+The National Long Term Care Study (NLTCS) is a longitidunal study tracking the health outcomes of Medicare recipients.
+The initial began in 1982, with follow-up surveys taken approximately every 5 years.
+For our purposes, we are interested in finding matching records over the $n_A = 20485$ records from 1982 to the $n_B = 17466$ records from 1989.
+
+We conduct linkage using sex, date of birth, and location using the thresholds shown in Table~\ref{Tab:nltcs-comparisons}.
+Note that storing 5 comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
+Standard settings on 16BG personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task.
+However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and compress information through storage efficient indexing; the resulting data object is just 10 MB, approximately 0.1\% of what is required for the raw comparisons.
+Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.
+
+We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
+Our Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+Since we access to unique identifiers for each record, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
+
+\begin{table}
+	\centering
+	\begin{tabular}[t]{lllll}
+		\hline
+		\multicolumn{2}{c|}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
+		\cline{3-5}
+		Fields & Similarity & 1 & 2 & 3\\
+		\hline
+		Sex & Binary & Agree & Disagree & \\
+		Year of Birth & Binary & Agree & Disagree & \\
+		Month of Birth & Binary & Agree & Disagree & \\
+		Day of Birth & Binary & Agree & Disagree & \\
+		Location & Custom & Same State and Office & Same State & Otherwise \\
+		\hline
+	\end{tabular}
+	\caption{Construction of comparison vectors for NTLCS data}\label{Tab:nltcs-comparisons}
+\end{table}
+
+\begin{figure}[t]
+	\begin{center}
+		\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
+		\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
+		\label{fig:overlap-plot}
+	\end{center}
+\end{figure}
+
+%\hypertarget{violations}{%
+%	\subsection{Violations of One-to-One Matching}\label{violations}}
+%
+%As noted previously, the application exhibits many situations in which bipartite matching is difficult.
 To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
 These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $\bm{X}_1$ has
-multiple plausible matches in $\bm{X}_2$.
-
-%Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
-%duplicates found in both databases for both \texttt{fabl} and \texttt{BRL}. We
-%see that \texttt{fabl} consistently overmatches within each Gibbs
-%iteration when compared to \texttt{BRL}, which is to be expected because
-%\texttt{BRL} explicitly prevents matches that violate one-to-one
-%matching throughout the entire sampler. Most of such matchings are due
-%to the randomness in the sampling procedure, and they occur sporadically
-%throughout the sampler in such a way that does not measurably influence
-%the eventual Bayes estimate.
-
-Table~\ref{Tab:rosa-maria} illustrates an example.
+%multiple plausible matches in $\bm{X}_2$. 
+%
+%Table~\ref{Tab:rosa-maria} illustrates an example.
 Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
 In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and \(\gamma_{2776, 826}\) are exactly identical.
-
-\begin{table}
-	\centering
-	\begin{tabular}[h!]{llllllll}
-		\hline
-		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
-		\hline
-		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
-		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
-		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
-		\hline
-	\end{tabular}
-	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
-\end{table}
-
-Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
+%
+%\begin{table}
+%	\centering
+%	\begin{tabular}[h!]{llllllll}
+%		\hline
+%		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
+%		\hline
+%		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
+%		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
+%		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
+%		\hline
+%	\end{tabular}
+%	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
+%\end{table}
+%
+%Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
 Under \texttt{fabl}, both records in $\bm{X}_2$ match to the same record in $\bm{X}_1$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
@@
 Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
-
-Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
+%
+%Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
 \texttt{BRL} in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
 Thus relaxing the one-to-one constraint has served as a useful tool for identifying model misspecification.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-\caption{Gibbs sampling in situation with multiple plausible matches.
+%
+%\begin{figure}[h!]
+%\begin{center}
+%\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
+%\caption{Gibbs sampling in situation with multiple plausible matches.
 Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability.
 Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
-\end{center}
-\end{figure}
+%\end{center}
+%\end{figure}
 
@@
 Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
-Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is currently under investigation by the authors for future work.
+Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
 
 To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
-This motivates extentions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
+This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
+Similar work has been been proposed by \cite{heck2019} and \cite{larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
 We see these types of heirarchical record linkage models as potential contributions to the growing algorithmic fairness literature, and useful to practitioners working with or studying marginalized populations.
@@
 		\Procedure{Hashing and Preprocessing}{}
-		\State Construct and enumerate set of unique patterns $\mathcal{P}$ from $F$ and $\{L_f\}$
-		\State Partition files $\bm{X_1}$ and $\bm{X_1}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$
+		\State Construct and enumerate set of unique patterns $\mathcal{P}$ from $F$ and $\{L_f\}$.
+		\State Partition files $\bm{X_1}$ and $\bm{X_1}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$.
 		\For{each $n$, $m$}
-		\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$
-		\State Hash records to $\mathcal{R}_{nm}$ and calculate summary statistics $\mathcal{H}_{nm}$
-		\State Use SEI to reduce memory usage; $\mathcal{R}_{nm} \to \mathcal{R}_{nm}^{\text{SEI}}$
+		\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$.
+		\State Hash records to $\mathcal{R}_{nm}$ and calculate summary statistics $\mathcal{H}_{nm}$.
+		\State Use SEI to reduce memory usage; $\mathcal{R}_{nm} \to \mathcal{R}_{nm}^{\text{SEI}}$.
 		\EndFor
-		\State Synthesize results across pairings to get $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$
+		\State Synthesize results across pairings to get $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$.
 		\EndProcedure
@@
 		\Procedure{Gibbs Sampling}{}
-		\State Initialize $m$, $u$, and $Z$ parameters
+		\State Initialize $m$, $u$, and $Z$ parameters.
 
 		\For{$t \in \{1, \ldots, T\}$} 
-		\State Sample $\Phi^{t+1}|Z^{t}, \tilde{\Gamma}$
-		\State Sample $h\left(Z^{t+1}\right)|\Phi^{t+1}, \tilde{\Gamma}$  \Comment{Sample agreement pattern, not record}
+		\State Sample $\Phi^{t+1}|Z^{t}, \tilde{\Gamma}$.
+		\State Sample $h\left(Z^{t+1}\right)|\Phi^{t+1}, \tilde{\Gamma}$.  \Comment{Sample agreement pattern, not record}
 		\EndFor
-		\State Sample $Z | h(Z), \tilde{\Gamma}$ \Comment{Fills in record label based on agreement pattern}
+		\State Sample $Z | h(Z), \tilde{\Gamma}$. \Comment{Fills in record label based on agreement pattern}
 
@@
 	\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
-Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}. It is set up with one error across the linkage fields and 50 duplicates across files.
+Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+It is set up with one error across the linkage fields and 50 duplicates across files.
+Traceplots across other settings exhibit similar behavior.
+Note that traceplots for $\bf{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
 

### 2021-11-15 R/CompareBinary.R

@@ -16,3 +16,88 @@ CompareBinary <- function(df1, df2, cols){
              Ztrue = NULL)
   cd
 }
+
+CompareBinary2 <- function(df1, df2){
+  nA <- length(df1)
+  nB <- length(df2)[1]
+  ids <- expand.grid(1:nA, 1:nB)
+  match <- df1[ids[, 1]]== df2[ids[, 2]]
+  comparison <- cbind(match, 1 - match) %>%
+    unname()
+  comparison
+}
+
+CompareDOB <- function(vec1, vec2){
+  nA <- length(vec1)
+  nB <- length(vec2)
+  ids <- expand.grid(1:nA, 1:nB)
+  matched_components <- matrix(NA, nrow = nrow(ids), ncol = 3)
+  for(i in 1:3){
+    s <- 2*i - 1
+    match <- str_sub(vec1[ids[, 1]], s, s+1) == str_sub(vec2[ids[, 2]], s, s+1)
+    matched_components[, i] <- match
+  }
+  score <- rowSums(matched_components)
+  score[score == 0] <- 1
+
+  comparison <- matrix(0, nrow = nrow(ids), ncol = 3)
+  comparison[score == 3, 1] <- 1
+  comparison[score == 2, 2] <- 1
+  comparison[score == 1, 3] <- 1
+  comparison
+}
+
+CompareLocation <- function(vec1, vec2){
+  nA <- length(vec1)
+  nB <- length(vec2)
+  ids <- expand.grid(1:nA, 1:nB)
+  matched_components <- matrix(NA, nrow = nrow(ids), ncol = 2)
+  for(i in 1:2){
+    s <- 2*i - 1
+    match <- str_sub(vec1[ids[, 1]], s, s+1) == str_sub(vec2[ids[, 2]], s, s+1)
+    matched_components[, i] <- match
+  }
+  score <- rep(1, nrow(ids))
+  score[matched_components[, 2] == 1] <- 3
+  score[matched_components[, 2] == 1 & matched_components[, 1] == 0] <- 2
+
+  comparison <- matrix(0, nrow = nrow(ids), ncol = 3)
+  comparison[score == 3, 1] <- 1
+  comparison[score == 2, 2] <- 1
+  comparison[score == 1, 3] <- 1
+  comparison
+}
+
+CompareNLTCS <- function(df1, df2){
+  nA <- nrow(df1)
[…]

### 2021-11-15 R/NLC_hash2.R

@@ -1,7 +1,7 @@
 BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
                           alpha = 1, beta = 1, S = 1000, burn = 100,
                           show_progress = T, fast = F, R = NULL,
-                          all_patterns = TRUE){
+                          all_patterns = TRUE, empirical = FALSE){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -32,11 +32,7 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   })))
 
   ids <- expand.grid(1:n1, 1:n2)
-  #rec2 <- ids[,2]
 
-  #hash_id <- patterns[[1]]
-  #hash <- data.frame(hash_id)
-  #hash_split <- split(hash_id, ids[,2])
   unique_patterns <- patterns[[1]]
   pattern_counts <- patterns[[2]]
   P <- dim(unique_patterns)[1]
@@ -51,27 +47,26 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   Z.temp <- rep(0, n1*n2)
   Z <- rep(n1+1, n2)
   L <- 0
-  #AZ <- BZ <- rep(0, length(parameter_split))
+
   m <- u <- rep(0, length(parameter_split))
   matches <- rep(0,P)
 
-  #indicators <- data.frame(indicators_raw, Z.temp)
+  #Empirical Bayes
+  alpha_emp <- pattern_counts[1]
+  beta_emp <- n2 - alpha_emp
 
-
+  if(empirical == TRUE){
+    alpha <- 170
+    if(alpha == 0){
+      alpha <- 1
+    }
+    beta <- beta_emp
+  }
 
 
 
   # Gibbs
   for(s in 1:S){
-
-    #map2_df(hash_split, Z, ~ .x[.y])
-
-    # hash$Z.temp <- Z.temp
-    # matches <- hash %>%
-    #   group_by(hash_id) %>%
-    #   summarize(matches = sum(Z.temp)) %>%
-    #   select(matches) %>%
-    #   pull()
 
     AZ <- sweep(unique_patterns, MARGIN = 1, STAT = matches, FUN = "*") %>%
       colSums() %>%
@@ -121,47 +116,6 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
       count() %>%
[…]

### 2021-11-15 R/fellegi_sunter.R

@@ -1,0 +1,135 @@
+FellegiSunter <- function(comparisons, S = 1000, show_progress = T, R = NULL,
+                           all_patterns = TRUE, threshold = .75){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRecords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
+  # alpha = first parameter of prior for linkage probability
+  # beta = second parameter of prior for linkage probability
+  # S = number of Gibbs iterations
+  # burn = number of iterations to be discarded as burn-in
+  # show_progress = set to false to show simulation progress
+
+  fields <- length(comparisons[[4]])
+  levels <- comparisons[[4]]
+  n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
+  indicators_raw <-comparisons[[1]]
+
+  ptm <- proc.time()
+  patterns <- GetUniquePatterns2(comparisons, fast, R, all_patterns)
+  elapsed_hash <- proc.time() - ptm
+
+  parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
+    rep(x, comparisons[[4]][x])
+  })))
+
+  ids <- expand.grid(1:n1, 1:n2)
+  #rec2 <- ids[,2]
+
+  unique_patterns <- patterns[[1]]
+  pattern_counts <- patterns[[2]]
+  P <- dim(unique_patterns)[1]
+  counts_by_rec <- patterns[[3]]
+  hash_to_rec1 <- patterns[[4]]
+  # candidates <- 1:n1
+  # candidates_P <- 1:(P+1)
+  Z.SAMPS <- matrix(NA, nrow = nrow(indicators_raw), ncol = S)
+  M.SAMPS <- matrix(NA, nrow = length(parameter_split), ncol = S)
+  U.SAMPS <- matrix(NA, nrow = length(parameter_split), ncol = S)
+  LAMBDA.SAMPS <- vector(length = S)
+
+  #Initialize
+  lambda <-  1 / n1
+  m <- sapply(levels, function(x){
+    ((x:1)^2)/sum((x:1)^2)
+  }) %>%
+    unlist() %>%
+    as.vector()
+
+  u <- sapply(levels, function(x){
[…]

### 2021-12-06 methods_paper/fabl-paper.tex
many additions

@@
 Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}.
-Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War.
-Section~\ref{discussion} concludes with a discussion of directions for future work.
+Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
+Section~\ref{discussion} concludes with a discussion of future work.
 
@@
 $\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$
-The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.
+The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case studies.
 
@@
 This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
-This creates a dependency that makes the sampler inherently serial, which can slow the sampler when working on large linkage tasks.
+This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.
 
@@
 In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{heck2019} and use independent priors for each component $Z_j$.
-However, unlike Heck Wortman who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
+However, unlike \cite{heck2019} who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
 We denote the fast beta prior as follows:
@@
 Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desirable; for example, we may be interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable \citep{Newcombe1963} .
+This behavior may be desirable; for example, we may be interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable \citep{Newcombe1963}.
 In situations when a bipartite estimate is desired, we can use the post processing procedure shown in section~\ref{bayes-estimate} to resolve many-to-one matchings.
@@
 $$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
-
-These constitute $\sum L_f$ many summations over $P$ quantities.
+These constitute $\tilde{L}$ summations over $P$ quantities, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.
 
@@
 These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
-We provide summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
+We provide a summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
+{\color{blue}We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
+\texttt{BRL} would see considerable speed up through the procedure, but this step would keep the complexity of the sampler at $O(n_A \times n_B)$.}
 
@@
 Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(\tilde{L}^{*}P)$, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.
+However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(\tilde{L}P)$.
 The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
@@
 comparison against \texttt{fabl}, currently built in R.
-However, The computational speed gains for \texttt{fabl} persist, especially for the larger simulations.
+However, the computational speed gains for \texttt{fabl} persist, especially for the larger simulations.
 Additionally, although \texttt{fabl} is amenable to parallelization, this simulation was run on a single core.
-Running \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
+Implementing \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
 
@@
 \begin{table}
-	\begin{tabular}[t]{ll|llll}
+	\begin{tabular}[t]{llllll}
 		\hline
@@
 		\hline
-		\multicolumn{2}{c|}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
+		\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
 		\cline{3-5}

### 2021-12-06 R/NLC_hash2.R

@@ -1,7 +1,7 @@
 BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
                           alpha = 1, beta = 1, S = 1000, burn = 100,
                           show_progress = T, fast = F, R = NULL,
-                          all_patterns = TRUE, empirical = FALSE){
+                          all_patterns = TRUE){
   # Implements bipartite record linkage with BK Sampling Mechanism
   #
   # Arguments
@@ -50,20 +50,6 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
 
   m <- u <- rep(0, length(parameter_split))
   matches <- rep(0,P)
-
-  #Empirical Bayes
-  alpha_emp <- pattern_counts[1]
-  beta_emp <- n2 - alpha_emp
-
-  if(empirical == TRUE){
-    alpha <- 170
-    if(alpha == 0){
-      alpha <- 1
-    }
-    beta <- beta_emp
-  }
-
-
 
   # Gibbs
   for(s in 1:S){
@@ -142,7 +128,8 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   list(Z = final_gibbs,
        m = M.SAMPS,
        u = U.SAMPS,
-       hash_time = elapsed_hash[3])
+       hash_time = elapsed_hash[3],
+       overlap = L.SAMPS)
 
 }
 
@@ -661,5 +648,183 @@ BKSimple_hash2_big <- function(comparisons, m_prior = 1, u_prior = 1,
   list(Z = final_gibbs,
        m = M.SAMPS,
        u = U.SAMPS)
+
+}
+
+BKSimple_hash_discount <- function(comparisons, m_prior = 1, u_prior = 1,
+                           alpha = 1, beta = 1, S = 1000, burn = 100,
+                           show_progress = T, fast = F, R = NULL,
+                           all_patterns = TRUE, lag = 10){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRecords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
[…]

### 2021-12-06 methods_paper/cover-letter/ba-cover-letter.tex

\documentclass[11pt]{letter}
%\usepackage{fullpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[top=3in,  % FIRST, CHANGE THIS NUMBER TO GET THE SECOND PAGE 
                      % LOOKING THE WAY YOU WANT IT
	bottom=1in, left=1in, right=1in]{geometry}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eso-pic}
\usepackage{graphicx}

\newcommand\BackgroundPic{
\put(0,0){
\parbox[b][\paperheight]{\paperwidth}{
\vfill
\centering
\includegraphics[width=\paperwidth,height=\paperheight,
keepaspectratio]{beka_lh_duke.pdf}%
\vfill
}}}

%\setlength\headheight{25em}
%\setlength\headsep{0em}
%\date{June 9, 2010}
\date{\raggedright November 29, 2021}

\begin{document}

\AddToShipoutPicture*{\BackgroundPic}
\signature{\vspace*{-4em}\includegraphics[scale=1]{sign.pdf}} 
\thispagestyle{empty}

\begin{letter}{
%Dear Graduate Admissions Committee
%https://www.franklin.uga.edu/jobs/statistics
%National Science Foundation Graduate Research Fellowship
%Carnegie Mellon University\\Department of Statistics\\
%Baker/Porter Hall\\
%Pittsburgh, PA 15213}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-.2in}  % AFTER YOU HAVE THE SECOND PAGE THE WAY YOU WANT IT,
                % CHANGE THIS NUMBER TO GET THE FIRST PAGE HOW YOU WANT IT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{left}
%This goes left\\
%BxLah\\
%\end{left}

\opening{Dear Editor Guindani:}

I am a PhD student at Duke University in the Department of Statistical Science. I am writing to submit my paper with my coauthors
Jerome Reiter and Rebecca C. Steorts entitled ``Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)" to \emph{Bayesian Analysis}.

Very often information about social entities is scattered across multiple databases.  Combining that information into one database can result in enormous benefits for analysis, resulting in richer and more reliable conclusions. The application areas touched by record linkage include voter registration, official statistics, human rights violations, patent inventions, health applications, among many others. Disciplines that utilize both applications and methods include statistics, biostatistics, computer science, machine learning, political science, and others. The goal of our review article is to review nearly almost all of entity resolution in such a way that it can be read and understood by as many communities as possible given that the current state of the literature is mathematical and overly technical, making the entry point difficult for those in the computational social sciences.

Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or fabl. Specifically, in fabl we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that fabl has markedly increased speed with minimal loss of accuracy.

To help the editorial office, we have listed a number of potential reviewers that do not have any conflicts with the authors:

Mauricio Sadinle, msadinle@uw.edu, University of Washington; Serge Aleshin-Guendel, aleshing@uw.edu, University of Washington; Yves Thibaudeau, Yves.Thibaudeau@census.gov, United States Census Bureau; Megan Price, meganp@hrdag.org, Human Rights Data Analysis Group; Ted Enamorado, ted@wustl.edu, University of Washington St. Louis; Nick Monath, nmonath@cs.umass.edu; Dinusha Vatsalan, Dinusha.Vatsalan@data61.csiro.au, Ben Rubinstein; benjamin.rubinstein@unimelb.edu.au; Neil Marchant; nmarchant@unimelb.edu.au; Andy Toole; Andrew.Toole@uspto.gov;  jph34@stat.duke.edu; Giacomo Zanella, zanella.gcm@gmail.com, Brunero Liseo, brunero.liseo@uniroma1.it, Andrea Tancredi, andrea.tancredi@uniroma1.it;
Peter Christen, peter.christen@anu.edu.au, Partha Lahiri, plahiri@umd.edu, Rebecca Nugent, rnugent@stat.cmu.edu, Kayla Frisoli, kaylafrisoli@gmail.com.

We appreciate consideration of our paper, and please do not hesitate to reach out if you have any questions.

 \closing{Sincerely,}

 Rebecca C. Steorts

\end{letter}

\end{document}

### 2021-12-06 bayesian-analysis/kundinger-ba-2021.tex
working on compiling ba paper.

@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
-%
-\pubyear{2021}
-\volume{TBA}
-\issue{TBA}
-% \doi{0000}
-%\arxiv{}
-\firstpage{1}
-\lastpage{1}
+\documentclass[ba,preprint]{imsart}% use this for supplement article
+%\documentclass[ba]{imsart}
+%%
+%\pubyear{2021}
+%\volume{TBA}
+%\issue{TBA}
+%% \doi{0000}
+%%\arxiv{}
+%\firstpage{1}
+%\lastpage{1}
+%\input{subfiles/preamble.tex}
 
@@
 
-\startlocaldefs
-% ** Local definitions **
-\endlocaldefs
+%\startlocaldefs
+%% ** Local definitions **
+%\endlocaldefs
 
 \begin{document}
+
+\begin{frontmatter}
 
@@
 \runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+%
+%\begin{aug}
+%\author{\fnms{Neil G.} \snm{Marchant}\thanksref{addr1}\ead[label=e1]{nmarchant@unimelb.edu.au}},
+%\author{\fnms{Benjamin I. P.} \snm{Rubinstein}\thanksref{addr1}\ead[label=e2]{brubinstein@unimelb.edu.au}}
+%\and
+%\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
+%
+%\runauthor{N. G. Marchant et al.}
+%
+%\address[addr1]{School of Computing and Information Systems,
+%  University of Melbourne,
+%  Parkville, VIC 3010, Australia,
+%  \printead{e1}, % print email address of "e1"
+%  \printead*{e2}
+%}
+%
+%\address[addr2]{Departments of Statistical Science and Computer Science,
+%  Duke University,
+%  P.O.\ Box 90251,
+%  Durham, NC 27708, USA
+%  \printead{e3}
+%}
+%
+%%\thankstext{<id>}{<text>}
+%
+%\end{aug}
 
@@
 
-%\address[addr2]{Departments of Statistical Science and Computer Science,
-%  Duke University,
-%  P.O.\ Box 90251,
-%  Durham, NC 27708, USA
-%  \printead{e3}
-%}
+\address[addr2]{Departments of Statistical Science and Computer Science,
+  Duke University,
+  P.O.\ Box 90251,
+  Durham, NC 27708, USA
+  \printead{e3}
+}
 
@@
 \begin{keyword}%[class=MSC]
-bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
-%\kwd{}
+\kwd{entity resolution}
+\kwd{record linkage}
+\kwd{deduplication}
+\kwd{Bayesian methods}
+\kwd{Markov chain Monte Carlo}
 %\kwd[]{}
 \end{keyword}
+
+\end{frontmatter}
+
+%%% ** Keywords **
+%\begin{keyword}%[class=MSC]
+%bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing
+%%\kwd{}
+%%\kwd[]{}
+%\end{keyword}
 

### 2021-12-06 bayesian-analysis/kundinger-submit.tex
working on compiling ba paper.

--- 2021-12-06 methods_paper/fabl-paper.tex
@@
-\documentclass[12pt,letterpaper]{article}
+\documentclass[ba]{imsart}
+%
+\pubyear{0000}
+\volume{00}
+\issue{0}
+\doi{0000}
+%\arxiv{}
+\firstpage{1}
+\lastpage{1}
 
-\usepackage[hypertexnames=false]{hyperref}
+\usepackage{amsthm}
 \usepackage{amsmath}
+\usepackage{natbib}
+\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
 \usepackage{graphicx}
-\usepackage{enumerate}
-\usepackage{natbib}
 
-% Custom packages
-\usepackage{amsthm, amssymb}
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{algorithm}      % algorithm environment
-\usepackage{algpseudocode}
-\usepackage{multirow}
-\usepackage{sectsty}
-\usepackage{tabularx}
-\usepackage{tikz}           % vector graphics
-\usepackage{bm}             % bold math symbols
-\usepackage{xcolor}
-\usepackage{microtype}
-\usepackage{import}
-\usepackage{titling}
-\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
-  calc, shadows, plotmarks}
-
-\graphicspath{{./notes/figures/}}
-
-% Custom commands
-\let\oldvec\vec
-\renewcommand\vec{\bm}
-\newcommand{\simfn}{\mathtt{sim}} % similarity function
-\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
-\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
-\newcommand{\distfn}{\mathtt{dist}} % distance function
-\newcommand{\valset}{\mathcal{V}} % attribute value set
-\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
-\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
-\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
-\newcommand{\euler}{\mathrm{e}} % Euler's constant
-\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
-\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-  {#1}\small\normalsize} \spacingset{1}
-
-\newtheorem{remark}{Remark}
-\newtheorem{proposition}{Proposition}
-\newtheorem{definition}{Definition}
-\newtheorem{lemma}{Lemma}
-
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
-
-\sectionfont{\large\nohang\centering\MakeUppercase}
-
-\title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
-\author{Brian Kundinger\textsuperscript{a} \and
-  Jerome Reiter\textsuperscript{a} \and 
-  Rebecca C.~Steorts \textsuperscript{b}}
-\date{
- \textsuperscript{a} Department of Statistical Science, Duke University \\
- \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
-  \today}
+\startlocaldefs
+\input{subfiles/preamble.tex}
+\endlocaldefs
 
 \begin{document}
-\maketitle
 
-\bigskip
+%% *** Frontmatter ***
+
+\begin{frontmatter}
+\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+%\title{A Sample Document\thanksref{T1}}
+%\thankstext{T1}{<thanks text>}
+\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+
+\begin{aug}
+\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
+\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
+\and
+\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
+
+\runauthor{N. G. Marchant et al.}
+
+\address[addr1]{Department of Statistical Science,
+  Duke University,
+  P.O.\ Box 90251,
+  Durham, NC 27708, USA
+  \printead{e1}, % print email address of "e1"
+  \printead*{e2}
+}
+
+\address[addr2]{Departments of Statistical Science and Computer Science,
+  Duke University,
+  P.O.\ Box 90251,
+  Durham, NC 27708, USA
+  \printead{e3}
+}
+
+%\thankstext{<id>}{<text>}
+\end{aug}
+
 \begin{abstract}
@@
 Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
-
-%In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
-Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
-This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.
-In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}.
-Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
 \end{abstract}
 
-\noindent%
-{\it Keywords:} record linkage, fusion, hashing, distributed computing, Fellegi Sunter
+%% ** Keywords **
+\begin{keyword}%[class=MSC]
+\kwd{bipartite record linkage}
+\kwd{Bayesian methods}
+\kwd{hashing techniques}
+\kwd{parallel/distributed computing}
+\kwd{Markov chain Monte Carlo}
+\end{keyword}
 
-\newpage
-\spacingset{1.5}
+\end{frontmatter}
+
+%% ** Mainmatter **
 
@@
 
-\subsection{Fellegi-Sunter Model}
-\label{fellegi-sunter}
-The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data.
-This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
-\begin{align}
-	\label{eqn:wts}
-	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
-\end{align}
-as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
+%\section{}\label{}
 
-In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
-Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent.
-This allows us to model
-the comparison data using mixture models in the following way:
-\begin{align}
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
-	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
-\end{align}
-where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters $m_{fl} = P(\gamma_{ij}^f = l|Z_j =i)$ for the matching record pairs, and $\bm{u}$ is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
-These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.
+% \begin{figure} 
+% \includegraphics{<eps-file>}% place <eps-file> in ./img  subfolder
+% \caption{}
+% \label{}
+% \end{figure}
 
-After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
-This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
+% \begin{table} 
+% *****************
+% \begin{tabular}{lll}
+% \end{tabular}
+% *****************
+% \caption{}
+% \label{}
+% \end{figure}
 
-To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
-\begin{align}
-	\label{eqn:jaro}
-	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
-	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
-	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
-	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
-\end{align}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Supplementary Material, if any, should   %%
+%% be provided in {supplement} environment  %%
+%% with title and short description.        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{supplement}
+% \stitle{???}
+% \sdescription{???.}
+% \end{supplement}
 
-The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
-Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
-
-\subsection{Beta Record Linkage Model}
-\label{BRL}
-
-One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
-One way of resolving this issue to incorporate the constraint directly into the model as follows:
-\begin{align}
-	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
-	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
-	& \bm{Z} \sim \mathcal{B}, \notag
-\end{align}
-where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
-$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
-$$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
-This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.
-
-\section{Fast Beta Linkage}
-\label{sec:fast-beta-linkage}
-
-Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
-In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{heck2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{heck2019} who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
-We denote the fast beta prior as follows:
-
-$$Z_j | \pi =
-\begin{cases} 
-	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + j \\
-\end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
-
-Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
-$\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-We note here that prior from \cite{heck2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
-
-Note that linkage in this setting is conducted at the record level, rather than at the record pair level as in the Fellegi Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of record pairs that are matches.
-We find $\pi$ to be more interpretable parameter than $\lambda$ in the bipartite case.
-There are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions.
-As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to achieve the benefits from \texttt{BRL} while making efficiency gains possible by exploiting independence.
-
-Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desirable; for example, we may be interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable \citep{Newcombe1963}.
-In situations when a bipartite estimate is desired, we can use the post processing procedure shown in section~\ref{bayes-estimate} to resolve many-to-one matchings.
-
-%this independent prior for each \(Z_j\), induces a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains over the Gibbs sampler under \texttt{BRL}.
-More importantly, since only the agreement pattern of \(Z_j\) is used forcalculations within the Gibbs sampler, and not the particular record label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.
-
-%\hypertarget{fabl-model}{%
-%	\subsection{Model Specification}
-%	\label{model-specification}} 
-%
-%As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent.
-Additionally, we assume that any missing data is missing at random.
-Let  $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denote the probability of that matching records have level $\ell$ of disagreement in feature $f$, and $u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represent the same probability for non-matches.
-In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, $\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
-
-%The model can then be specified as follows:
-With the \texttt{fabl} prior defined, we present the full model as follows:
-
-$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
-$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
-$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
-$$Z_j | \pi =
-\begin{cases} 
-	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + j \\
-\end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
-
-\hypertarget{posterior-sampling}{%
-	\subsection{Gibbs Sampler}
-	\label{gibbs_sampling}}
-We work with the following factorization of the joint posterior distribution:
-
-\begin{align*}
-	p(\bm{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\bm{Z} | \pi) p(\Phi) p(\pi) \\
-	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
-	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}
-\end{align*}
-
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. Thus we have
-
-$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
-$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z}))$$
-where
-\(\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
-and
-\(\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).
-
-As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
-Thus we have
-
-\begin{align*}
-p(\pi | \bm{Z}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1} \\
-\end{align*}
-Thus the posterior $\pi^{(s+1)}|\bm{Z}^{(s+1)}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
-only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
-denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).  As a function of \(Z_j\), 
-\(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\)
-is a discrete distribution with probabilities proportional to
-
-\begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
-	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&=
-	\begin{cases} 
-		w_{ij}  & z_j \leq n_1; \\
-		1 &  z_j  = n_1 + j \\
-	\end{cases}\\
-\end{align*}
-
-where
-%$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
-$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$.
-
-These \(w_{ij}\) are precisely the likelihood ratios used in the Fellegi-Sunter model to classify matches and non-matches, and we therefore call them the Fellegi Sunter weights.
-
-With the likelihood in this form, we can derive the full conditional
-\begin{align*}
-	p(Z_j|\Gamma_{.j}, \Phi, \pi) &\propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi) \\
-	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}
-\end{align*}
-Thus we have
-$$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
-	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + 1 \\
-\end{cases}$$
-Finally, we integrate over the posterior of \(\pi\) and rearrange terms to produce the final full conditional:
-
-\[p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
-\begin{cases} 
-	w_{ij}  & i \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + 1 \\
-\end{cases}\]
-
-\hypertarget{bayes-estimate}{%
-	\subsection{Bayes Estimate}
-	\label{bayes-estimate}}
-
-We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage
-parameter \(\bm{Z}\) by assigning different positive losses to
-different types of errors, and minimizing posterior expected loss. We
-adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
-\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
-option to leave the matching undetermined by the model. Specifically, we have
-
-\[L(\hat{Z_j}, Z_j)=\begin{cases} 
-	0  & \text{if } Z_j = \hat{Z_j}; \\
-	\theta_R,  & \text{if } \hat{Z_j} = R; \\
-	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
-	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
-	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
-\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
-on the linkage status, \(\theta_{10}\) is the loss from a false
-non-match, \(\theta_{01}\) is the loss from a false match, and
-\(\theta_{11}\) is the loss from the special case of a false match in
-which the record has a true match other than the one estimated by the
-model.
-
-We seek to minimize the posterior expected loss, given by 
-$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
-Here, we set $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule
-
-\[\hat{Z}_j =\begin{cases} 
-	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
-	0,  & \text{otherwise}. \\
-\end{cases}\]
-
-For a more in-depth explanation of this function and the induced Bayes
-estimate, see \cite{sadinle_bayesian_2017}.
-
-Since the Gibbs sampler does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable.
-We will explore this issue through the El Salvador homicide case study.
-
-To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the highest posterior probability, and declare the other to have no match.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
-
-\section{Efficient and Scalable Implementation}
-\label{sec:efficiency}
-
-To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
-The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-Here,
-$P = |\mathcal{P}|$ is the total number of unique agreement
-patterns.
-\(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
-In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the orignial vectors themselves.
-Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
-
-\hypertarget{data-representation-hashing-and-storage}{%
-	\subsection{Data Representation, Hashing, and
-		Storage}\label{data-representation-hashing-and-storage}}
-
-First, we hash record pairs of the same agreement pattern to unique integer values. \cite{enamorado2019using} accomplished this efficiently through the hashing function
-
-$$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}.$$
-
-This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
-For computational ease, we then map these integers to sequential integers from $\{1, \ldots, P\}$ corresponding to the enumerated patterns in $\mathcal{P}$.
-When the $(i, j)$ pair exhibits the $p^{th}$ pattern, we say $(i, j) \in h_p$.
-
-With all record pairs converted to hashed values, we can create a more compact nested list 
-$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
-where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
-$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||$$
-and
-$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
-For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will shown in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
-Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.
-
-\hypertarget{scaling}{%
-	\subsection{Scaling to Large Linkage Tasks}\label{scaling}}
-
-The hashing procedure described above considerably reduces the memory
-needed to store the comparison information. That is, instead of storing
-\(n_1 \times n_2\) comparison vectors, we only store the
-\(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
-relating record pairs to those vectors. However, even storing these
-\(n_1 \times n_2\) scalar labels can become burdensome with large data.
-Worse, the overwhelming majority of these labels relate to record pairs
-that are clear non-matches.
-
-To address this issue, we propose a new method called storage efficient indexing (SEI). In standard indexing, one decides a priori certain criteria that all true matching pairs must satisfy, 
-and labels any record pairs that do not meet that criteria as
-non-matches. For example, one might only consider pairs with a certain
-similarity score on a field deemed to be important (like first name), or
-only pairs that exactly match on a specified number of fields. While
-generally chosen to be quite loose, establishing these criteria
-requires knowledge of the problem and invites room for human error.
-
-With SEI however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks.
-Observe that all record $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{x}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
-Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
-We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
-Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$.
-Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most
-\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
-Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
-For simplicty, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
-
-%\hypertarget{distributed}{%
-%	\subsection{Distributed Computation for Pairwise Comparison}\label{distributed}}
-
-For large data, we can partition the two datasets \(\bm{X}_1\) and
-\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more
-manageable computations. On a single machine, we can read-in data
-sequentially, conduct hashing, compress information through SEI, and
-delete the original data from memory before continuing with the next
-chunk of data. With multiple cores or multiple machines, this can be
-done in parallel.
-Summary statistics from each pairwise chunk comparison can then be easily synthesized to recover sufficient statistics for the larger linkage task.
-Thus, the combination of hashing, SEI, and partitioning
-allows us to conduct linkage tasks over much larger datasets.
-
-\hypertarget{efficient-posterior}{%
-	\subsection{Efficient Posterior Inference }\label{efficient-posterior}}
-
-Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
-These are simple calculations, but become computationally burdensome when working on large linkage tasks.
-
-Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote
-$H_p^m = \sum_{j=1}^{n_B} \bm{1}_{(Z_j, j) \in h_p}$ to be the
-number of matching record pairs with agreement pattern $p$.
-We also can easily calculate the number of non-matching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\).
-Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
-
-$$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
-$$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
-These constitute $\tilde{L}$ summations over $P$ quantities, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.
-
-Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
-This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
-To speed up computation, we break this sampling step into two.
-First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link.
-Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\).
-More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + 1\).
-Then,
-
-$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
-\begin{cases} 
-	w_{p}\times H_{j_p}  & p \leq P; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
-\end{cases}$$
-$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
-	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
-	0 & \text{otherwise.} \\
-\end{cases}$$
-
-Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
-Thus we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with the first equation above.
-After all iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.
-
-We emphasize the computational gains of this split sampler: the first step is a sample from $P$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
-These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
-We provide a summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
-{\color{blue}We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
-\texttt{BRL} would see considerable speed up through the procedure, but this step would keep the complexity of the sampler at $O(n_A \times n_B)$.}
-
-\hypertarget{computational-complexity}{%
-	\subsection{Computational Complexity}\label{computational-complexity}}
-In this section, we prove the computational complexity of \texttt{fabl}.
-
-\begin{lemma}
-The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
-\label{lemma:fabl}
-\end{lemma}
-
-\begin{proof}
-To prove the computational complexity, we consider two steps --- constructing the comparison vectors and the Gibbs sampler. 
-The computational complexity of all pairwise comparisons across the two databases $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
-The hashing procedure for all pairwise comparisons is also of complexity $O(F n_1 n_2)$.
-Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
-There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
-
-Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(\tilde{L}P)$.
-The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
-With hashing, we split this sampling into two steps.
-First we sample the agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
-Thus the complexity of sampling $\bm{Z}$ is $O(n_2 P)$.
-Since $\tilde{L} << n_1$ in all reasonable applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
-In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
-\end{proof}
-
-\section{Simulation Studies}
-\label{sec:simulations}
-
-We demonstrate the speed, and accuracy of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
-
-\hypertarget{speed}{%
-	\subsection{Speed}\label{speed}}
-
-In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-The distributions used (see Table~\ref{Tab:distributions}) are meant to emulate the rates of agreement across first name, last name, day of birth, and month of birth.
-For example, $u^{\text{month, agree}} = P(\text{Records have same birth-month} \mid \text{Nonmatch}) = \frac{1}{12}$.
-The parameters for year would vary largely by context, so we choose them here to adjust the desired difficulty of the linkage task.
-For simplicity, we consider only exact matching, so a vector (1, 0) corresponds to agreement and (0, 1) to disagreement.
-We simulate these data for different values of \(n_1\) and \(n_2\), and compare the run-time of \texttt{fabl} against \texttt{BRL}.
-Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in the larger simulations.
-
-\begin{table}[t]
-	\centering
-	\begin{tabular}{rrr}
-		\hline
-		& m & u \\ 
-		\hline
-		fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
-		lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
-		day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
-		month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
-		year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
-		\hline
-	\end{tabular}
-	\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{Tab:distributions}
-\end{table}
-
-We note here that \texttt{BRL} is coded in C, which makes for unfair
-comparison against \texttt{fabl}, currently built in R.
-However, the computational speed gains for \texttt{fabl} persist, especially for the larger simulations.
-Additionally, although \texttt{fabl} is amenable to parallelization, this simulation was run on a single core.
-Implementing \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
-
-In Figure~\ref{fig:speed1}, we see that at low data size, \texttt{BRL} outperforms, but that \texttt{fabl} is significantly faster at handling larger data.
-In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
-
-\begin{figure}[t]
-	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
-		\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$.
-We see near quadratic growth in runtime for BRL, and near linear growth for fabl.}\label{fig:speed1}
-	\end{center}
-\end{figure}
-
-The above discussion suggests that for fixed \(n_2\), computation time
-should remain mostly constant with growing \(n_1\). Simulation study
-suggests that this is true. In Figure~\ref{fig:speed2} fixing \(n_2 = 500\), we
-see linear growth for the run-time under \texttt{BRL} as \(n_1\)
-increases, with much more static run-time under \texttt{fabl}. The
-slight increases in run-time that we do see are due primarily to the
-hashing step, which again can be run in parallel for large data.
-
-\begin{figure}[h!]
-	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-		\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
-We see linear growth in runtime for BRL, and near constant runtime for fabl.}\label{fig:speed2}
-	\end{center}
-\end{figure}
-
-\hypertarget{accuracy}{%
-	\subsection{Accuracy under Full Estimates}\label{accuracy}}
-
-We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap.
-To generate these data, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
-Fields are then chosen for distortion uniformly at random.
-Names are subject to string insertions, deletions and substitutions, as well as common keyboard, phonetic, and optical recognition errors.
-Age and occupation are distorted through keyboard errors and missingness.
-For further details on the generation of these data, see \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}.
-
-We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table~\ref{Tab:sadinle_simulation_cutoffs}.
-Each simulation identifies matched individuals between two datasets, each with 500 records.
-We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets.
-We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
-We calculate Bayes estimates of the linkage structure though the losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$.
-Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}.
-
-\begin{table}[t]
-	\centering
-	\begin{tabular}[t]{llllll}
-
-		\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
-		\cline{3-6}
-		Fields & Similarity & 1 & 2 & 3 & 4\\
-		\hline
-		First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (.5, .1]\\
-		Age and Occupation & Binary & Agree & Disagree &  & \\
-		\hline
-	\end{tabular}
-\caption{Construction of comparison vectors for accuracy study with simulated datasets}
-\label{Tab:sadinle_simulation_cutoffs}
-\end{table}
-
-We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and f-measure, as defined in \cite{christen_2012}.
-Recall is the proportion of true matches found by the model, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$, and precision is proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
-F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
-In Figure~\ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, we remain confident in \texttt{fabl}'s overall accuracy.
-
-%Each metric is provided explicitly below:
-%$$\text{Recall } = \frac{\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1)}{\sum_{j=1}^{n_2} I(Z_j \leq n_1)}$$
-%$$\text{Precision } = \frac{\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1)}{ \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)}$$
-%$$\text{F-measure } = 2 \frac{\text{Recall } + \text{Precision}}{\text{Recall } \times \text{Precision}}$$
-
-\begin{figure}[t]
-	\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
-\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
-For each level of overlap and each level of error, we have 100 paired sets of 500 records.
-We see comparable performance for all levels of error and overlap.}
-\label{fig:sadinle_simulation}
-	\end{center}
-\end{figure}
-
-\hypertarget{partial}{%
-	\subsection{Accuracy under Partial Estimates}\label{partial}}
-
-By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
-In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
-Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual non-matches.
-Mathematically, $\text{NPV} = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
-We continue to use the precision, which is renamed the positive predictive value (PPV) in this context.
-Lastly, we also report the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.
-
-In Figure~\ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
-However, with high amounts of overlap, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
-Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probability of matches can at times be more split across different records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
-
-%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-%This is reasonable because large numbers of errors blur the distinction between matched and non-matched pairs, leading to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
-
-\begin{figure}[t]
-\begin{center}
-	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
-	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
-	\label{fig:sadinle_simulation_partial}
-\end{center}
-\end{figure}
-
-%\hypertarget{scale}{%
-%	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
-%
-%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
-Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
-Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task.
-Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
-% 
-%For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
-We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
-The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method.
-Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run.
-However, this could be substantially sped up using distributed computing.
-%
-%Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
-This is slightly worse performance than  in the smaller simulation studies; this is expected because it is naturally more difficult to link more records with the same amount of information.
-With more linkage fields, \texttt{fabl} maintains the high accuracy in Section~\ref{accuracy}.
-
-\section{Case Studies}
-\label{sec:case-studies}
-
-In our first case study, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in both datasets.
-Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study (NLTCS), a larger linkage task that is not possible under \texttt{BRL}.
-
-\subsection{Civilian Casualties from the El Salvadoran Civil War}
-\label{el_salvador}
-
-The country of El Salvador was immersed in civil war from 1980 to 1991, and throughout the time, several organizations attempted to document casualties of the conflict.
-When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
-This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications}.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that had been published throughout the conflict, and the CDHES dataset consists of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
-The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both datasets are thought to be fairly reliable.
-
-There are several challenges with working with this data.
-First, both datasets have been automatically digitized, which inherently leads to some degree of typographical error.
-Second, the only fields recorded are given name, last name, date of death, and place of death.
-It is relatively common for aparent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between
-\texttt{BRL} and \texttt{fabl}.
-
-Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files in ERTL.
-We standardized names to account for common misspellings and used a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
-Place of death is recorded by municipality and department within that municipality; however, since department was missing in 95\% of records in CHDES and 80\% of records in ERTL, this was excluded from our analyses.
-Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We again use flat priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
-
-To mirror the original implementation of \cite{sadinle_bayesian_2017}, we first construct the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
-This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
-Gibbs sampling for these comparison vectors took 265 seconds for \texttt{fabl} and 240 for \texttt{BRL}; here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
-However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
-
-\begin{table}
-	\begin{tabular}[t]{llllll}
-		\hline
-		\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
-		\cline{3-6}
-		Fields & Similarity & 1 & 2 & 3 & 4\\
-		\hline
-		First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (.5, 1]\\
-		Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
-		Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
-		Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
-		Municipality & Binary & Agree & Disagree &  & \\
-		\hline
-	\end{tabular}
-	\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle2018bayesian}.
-This set up leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
-\end{table}
-
-Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}, and obtained analogous inference.
-With 216 possible agreement patterns, 159 were realized in the data, and our proposed method, \texttt{fabl} became much faster, finishing in 61 seconds.
-Meanwhile \texttt{BRL} took 239 seconds, relatively unchanged from the first implementation.
-This demonstrates the way that the computational complexity of our method depends on the number of unique agreement patterns, and how significant computational gains can be made by simplifying the construction of the comparison vectors.
-The estimates of the \(\bm{m}\) parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
-Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
-
-\begin{table}
-	\centering
-	\begin{tabular}[t]{lllll}
-		\hline
-		\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
-		\cline{3-5}
-		Fields & Similarity & 1 & 2 & 3\\
-		\hline
-		First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, 1]\\
-		Year of Death & Binary & Agree & Disagree & \\
-		Month of Death & Binary & Agree & Disagree & \\
-		Day of Death & Absolute Difference & 0 & 1 & 2+\\
-		Municipality & Binary & Agree & Disagree & \\
-		\hline
-	\end{tabular}
-	\caption{Construction of comparison vectors for El Salvador for increased speed under \texttt{fabl}. This set up leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
-\end{table}
-
-Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets.
-We calculate posterior samples of the size of the overlap across files by finding the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
-The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate is quite conservative.
-This is because there a large number of records in ERTL for which there are multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
-We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study.
-Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
-This means that after clerical review, the number of individuals replicated across datasets would fall in the interval (136, 311), encapsulating the posterior credible interval.
-More or fewer records could be identified for clerical review by decreasing or increasing $\theta_R$.
-
-We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
-
-\begin{figure}[t]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
-		\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
-		\label{fig:overlap-plot}
-	\end{center}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-\caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. We note they are quite similar across the two methods.}\label{fig:m-and-u}
-\label{fig:m-and-u}
-\end{center}
-\end{figure}
-
-\subsection{National Long Term Care Study}
-\label{nltcs}
-
-The National Long Term Care Study (NLTCS) is a longitidunal study tracking the health outcomes of Medicare recipients.
-The initial began in 1982, with follow-up surveys taken approximately every 5 years.
-For our purposes, we are interested in finding matching records over the $n_A = 20485$ records from 1982 to the $n_B = 17466$ records from 1989.
-
-We conduct linkage using sex, date of birth, and location using the thresholds shown in Table~\ref{Tab:nltcs-comparisons}.
-Note that storing 5 comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
-Standard settings on 16BG personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task.
-However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and compress information through storage efficient indexing; the resulting data object is just 10 MB, approximately 0.1\% of what is required for the raw comparisons.
-Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.
-
-We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-Our Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
-Since we access to unique identifiers for each record, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
-
-\begin{table}
-	\centering
-	\begin{tabular}[t]{lllll}
-		\hline
-		\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
-		\cline{3-5}
-		Fields & Similarity & 1 & 2 & 3\\
-		\hline
-		Sex & Binary & Agree & Disagree & \\
-		Year of Birth & Binary & Agree & Disagree & \\
-		Month of Birth & Binary & Agree & Disagree & \\
-		Day of Birth & Binary & Agree & Disagree & \\
-		Location & Custom & Same State and Office & Same State & Otherwise \\
-		\hline
-	\end{tabular}
-	\caption{Construction of comparison vectors for NTLCS data}\label{Tab:nltcs-comparisons}
-\end{table}
-
-\begin{figure}[t]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
-		\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
-		\label{fig:overlap-plot}
-	\end{center}
-\end{figure}
-
-%\hypertarget{violations}{%
-%	\subsection{Violations of One-to-One Matching}\label{violations}}
-%
-%As noted previously, the application exhibits many situations in which bipartite matching is difficult.
-To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
-These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $\bm{X}_1$ has
-%multiple plausible matches in $\bm{X}_2$. 
-%
-%Table~\ref{Tab:rosa-maria} illustrates an example.
-Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
-In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and \(\gamma_{2776, 826}\) are exactly identical.
-%
-%\begin{table}
-%	\centering
-%	\begin{tabular}[h!]{llllllll}
-%		\hline
-%		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
-%		\hline
-%		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
-%		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
-%		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
-%		\hline
-%	\end{tabular}
-%	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
-%\end{table}
-%
-%Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
-Under \texttt{fabl}, both records in $\bm{X}_2$ match to the same record in $\bm{X}_1$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
-Under \texttt{BRL}, the Gibbs process creates one matching configuration and stays there for a while.
-However, if one pair ``unmatches,'' then the other record has a chance to latch on.
-Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
-%
-%Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
-\texttt{BRL} in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
-Thus relaxing the one-to-one constraint has served as a useful tool for identifying model misspecification.
-%
-%\begin{figure}[h!]
-%\begin{center}
-%\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-%\caption{Gibbs sampling in situation with multiple plausible matches.
-Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability.
-Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
-%\end{center}
-%\end{figure}
-
-\section{Discussion}
-\label{discussion}
-
-In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. Specifically, our proposed hashing method
-and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making our approach particularly powerful in linking records when one datafile is substantially smaller than the other.
-
-Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
-We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
-Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
-Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
-
-To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
-This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
-Similar work has been been proposed by \cite{heck2019} and \cite{larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
-We see these types of heirarchical record linkage models as potential contributions to the growing algorithmic fairness literature, and useful to practitioners working with or studying marginalized populations.
-
-%Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
-%future work.
-
-%\begin{figure}[t]
-%	
-%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-%		
-%	}
-%	
-%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
-%\end{figure}
-
-%\subsection{National Long Term Care Survey}
-%\label{nltcs} 
-%
-%The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
-It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989.
-The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office.
-State and regional office have been anonymized, presented as numbers instead of their actual values.
-We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context.
-The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466$ files, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns.
-We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.
-%
-%With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible.
-Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold.
-Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller datasets.
-We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing.
-We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler.
-%
-%Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
-However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
-Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
-%
-%Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
-With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks.
-This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler.
-However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error.
-In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office.
-We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives.
-Even if not, the modeler may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures.
-However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons.
-Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block.
-Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.
-
-\clearpage
-
-\bigskip
-
-\bibliographystyle{jasa}
+\bibliographystyle{ba}
 \bibliography{biblio}
 
-\clearpage
+% ** Acknowledgements **
+% \begin{acknowledgement}
+% \end{acknowledgement}
 
-\section*{Appendix}
-\label{sec:appendix}
-
-\hypertarget{app:fabl-psuedocode}{%
-	\subsection{Summary of Fast Beta Linkage Method}\label{app:fabl-psuedocode}}
-
-\begin{algorithm}[h!]
-	\begin{algorithmic}[1]
-
-		\Procedure{Hashing and Preprocessing}{}
-		\State Construct and enumerate set of unique patterns $\mathcal{P}$ from $F$ and $\{L_f\}$.
-		\State Partition files $\bm{X_1}$ and $\bm{X_1}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$.
-		\For{each $n$, $m$}
-		\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$.
-		\State Hash records to $\mathcal{R}_{nm}$ and calculate summary statistics $\mathcal{H}_{nm}$.
-		\State Use SEI to reduce memory usage; $\mathcal{R}_{nm} \to \mathcal{R}_{nm}^{\text{SEI}}$.
-		\EndFor
-		\State Synthesize results across pairings to get $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$.
-		\EndProcedure
-
-		\Procedure{Gibbs Sampling}{}
-		\State Initialize $m$, $u$, and $Z$ parameters.
-
-		\For{$t \in \{1, \ldots, T\}$} 
-		\State Sample $\Phi^{t+1}|Z^{t}, \tilde{\Gamma}$.
-		\State Sample $h\left(Z^{t+1}\right)|\Phi^{t+1}, \tilde{\Gamma}$.  \Comment{Sample agreement pattern, not record}
-		\EndFor
-		\State Sample $Z | h(Z), \tilde{\Gamma}$. \Comment{Fills in record label based on agreement pattern}
-
-		\EndProcedure
-
-	\end{algorithmic}
-\end{algorithm}
-\clearpage
-
-\hypertarget{appendix-sim}{%
-	\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
-Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
-It is set up with one error across the linkage fields and 50 duplicates across files.
-Traceplots across other settings exhibit similar behavior.
-Note that traceplots for $\bf{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
-
-\begin{figure}[!h]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-		\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}}\label{fig:sim_overlap_trace}
-	\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-		\caption{Representative traceplot of m parameter from simulation study in Section \ref{accuracy}}\label{fig:sim_m_trace}
-	\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-		\caption{Representative traceplot of u parameters from simulation study in Section \ref{accuracy}}\label{fig:sim_u_trace}
-	\end{center}
-\end{figure}
-
-\clearpage
-
-\hypertarget{appendix-es}{%
-	\subsection{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-\caption{Traceplot for number of matches found across datasets in El Salvador case study}\label{fig:overlap_trace}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-\caption{Traceplot for m parameter in El Salvador case study}\label{fig:m_trace}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-\caption{Traceplot for u parameter in El Salvador case study}\label{fig:u_trace}
-\end{center}
-\end{figure}
+% ATTN: Should not include appendices in the main manuscript
+%\input{subfiles/appendix.tex}
 

### 2021-12-06 bayesian-analysis/kundinger-submit.tex
update.

--- 2021-12-06 methods_paper/fabl-paper.tex
@@
-\documentclass[12pt,letterpaper]{article}
-
-\usepackage[hypertexnames=false]{hyperref}
+\documentclass[ba]{imsart}
+%
+\pubyear{0000}
+\volume{00}
+\issue{0}
+\doi{0000}
+%\arxiv{}
+\firstpage{1}
+\lastpage{1}
+
+\usepackage{amsthm}
 \usepackage{amsmath}
+\usepackage{natbib}
+\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
 \usepackage{graphicx}
-\usepackage{enumerate}
-\usepackage{natbib}
-
-% Custom packages
-\usepackage{amsthm, amssymb}
+\newtheorem{lemma}{Lemma}
 \usepackage{booktabs}       % professional-quality tables
@@
 \usepackage{algpseudocode}
-\usepackage{multirow}
-\usepackage{sectsty}
-\usepackage{tabularx}
-\usepackage{tikz}           % vector graphics
-\usepackage{bm}             % bold math symbols
-\usepackage{xcolor}
-\usepackage{microtype}
-\usepackage{import}
-\usepackage{titling}
-\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
-  calc, shadows, plotmarks}
-
-\graphicspath{{./notes/figures/}}
-
-% Custom commands
-\let\oldvec\vec
-\renewcommand\vec{\bm}
-\newcommand{\simfn}{\mathtt{sim}} % similarity function
-\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
-\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
-\newcommand{\distfn}{\mathtt{dist}} % distance function
-\newcommand{\valset}{\mathcal{V}} % attribute value set
-\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
-\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
-\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
-\newcommand{\euler}{\mathrm{e}} % Euler's constant
-\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
-\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-  {#1}\small\normalsize} \spacingset{1}
-
-\newtheorem{remark}{Remark}
-\newtheorem{proposition}{Proposition}
-\newtheorem{definition}{Definition}
-\newtheorem{lemma}{Lemma}
-
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
-
-\sectionfont{\large\nohang\centering\MakeUppercase}
-
-\title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
-\author{Brian Kundinger\textsuperscript{a} \and
-  Jerome Reiter\textsuperscript{a} \and 
-  Rebecca C.~Steorts \textsuperscript{b}}
-\date{
- \textsuperscript{a} Department of Statistical Science, Duke University \\
- \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
-  \today}
+
+\startlocaldefs
+\input{subfiles/preamble.tex}
+\endlocaldefs
 
 \begin{document}
-\maketitle
-
-\bigskip
+
+%% *** Frontmatter ***
+
+\begin{frontmatter}
+\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+%\title{A Sample Document\thanksref{T1}}
+%\thankstext{T1}{<thanks text>}
+\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+
+\begin{aug}
+\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
+\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
+\and
+\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
+
+\runauthor{Kundinger et al.}
+
+\address[addr1]{Department of Statistical Science,
+  Duke University,
+  P.O.\ Box 90251,
+  Durham, NC 27708, USA
+  \printead{e1}, % print email address of "e1"
+  \printead*{e2}
+}
+
+\address[addr2]{Departments of Statistical Science and Computer Science,
+  Duke University,
+  P.O.\ Box 90251,
+  Durham, NC 27708, USA
+  \printead{e3}
+}
+
+%\thankstext{<id>}{<text>}
+\end{aug}
+
 \begin{abstract}
@@
 Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
-
-%In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability.
-Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space.
-This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.
-In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}.
-Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
 \end{abstract}
 
-\noindent%
-{\it Keywords:} record linkage, fusion, hashing, distributed computing, Fellegi Sunter
-
-\newpage
-\spacingset{1.5}
+%% ** Keywords **
+\begin{keyword}%[class=MSC]
+\kwd{bipartite record linkage}
+\kwd{Bayesian methods}
+\kwd{hashing techniques}
+\kwd{parallel/distributed computing}
+\kwd{Markov chain Monte Carlo}
+\end{keyword}
+
+\end{frontmatter}
+
+%% ** Mainmatter **
 
@@
 
-%\hypertarget{scale}{%
-%	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
-%
-%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
-Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
-Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task.
-Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
-% 
-%For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
-We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
-The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method.
-Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run.
-However, this could be substantially sped up using distributed computing.
-%
-%Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
-This is slightly worse performance than  in the smaller simulation studies; this is expected because it is naturally more difficult to link more records with the same amount of information.
-With more linkage fields, \texttt{fabl} maintains the high accuracy in Section~\ref{accuracy}.
-
 \section{Case Studies}
@@
 
-%\hypertarget{violations}{%
-%	\subsection{Violations of One-to-One Matching}\label{violations}}
-%
-%As noted previously, the application exhibits many situations in which bipartite matching is difficult.
-To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
-These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $\bm{X}_1$ has
-%multiple plausible matches in $\bm{X}_2$. 
-%
-%Table~\ref{Tab:rosa-maria} illustrates an example.
-Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
-In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and \(\gamma_{2776, 826}\) are exactly identical.
-%
-%\begin{table}
-%	\centering
-%	\begin{tabular}[h!]{llllllll}
-%		\hline
-%		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
-%		\hline
-%		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
-%		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
-%		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
-%		\hline
-%	\end{tabular}
-%	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
-%\end{table}
-%
-%Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
-Under \texttt{fabl}, both records in $\bm{X}_2$ match to the same record in $\bm{X}_1$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
-Under \texttt{BRL}, the Gibbs process creates one matching configuration and stays there for a while.
-However, if one pair ``unmatches,'' then the other record has a chance to latch on.
-Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
-%
-%Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
-\texttt{BRL} in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
-Thus relaxing the one-to-one constraint has served as a useful tool for identifying model misspecification.
-%
-%\begin{figure}[h!]
-%\begin{center}
-%\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-%\caption{Gibbs sampling in situation with multiple plausible matches.
-Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability.
-Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
-%\end{center}
-%\end{figure}
-
 \section{Discussion}
@@
 
-%Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
-%future work.
-
-%\begin{figure}[t]
-%	
-%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-%		
-%	}
-%	
-%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
-%\end{figure}
-
-%\subsection{National Long Term Care Survey}
-%\label{nltcs} 
-%
-%The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
-It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989.
-The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office.
-State and regional office have been anonymized, presented as numbers instead of their actual values.
-We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context.
-The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466$ files, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns.
-We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.
-%
-%With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible.
-Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold.
-Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller datasets.
-We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing.
-We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler.
-%
-%Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
-However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
-Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
-%
-%Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
-With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks.
-This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler.
-However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error.
-In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office.
-We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives.
-Even if not, the modeler may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures.
-However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons.
-Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block.
-Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.
-
-\clearpage
-
-\bigskip
-
-\bibliographystyle{jasa}
+%\section{}\label{}
+
+% \begin{figure} 
+% \includegraphics{<eps-file>}% place <eps-file> in ./img  subfolder
+% \caption{}
+% \label{}
+% \end{figure}
+
+% \begin{table} 
+% *****************
+% \begin{tabular}{lll}
+% \end{tabular}
+% *****************
+% \caption{}
+% \label{}
+% \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Supplementary Material, if any, should   %%
+%% be provided in {supplement} environment  %%
+%% with title and short description.        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{supplement}
+% \stitle{???}
+% \sdescription{???.}
+% \end{supplement}
+
+\bibliographystyle{ba}
 \bibliography{biblio}
+
+% ** Acknowledgements **
+% \begin{acknowledgement}
+% \end{acknowledgement}
+
+% ATTN: Should not include appendices in the main manuscript
+%\input{subfiles/appendix.tex}
 

### 2021-12-07 bayesian-analysis/kundinger-submit.tex
update.
adding punctutation.

@@
 \end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
+$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$
 
@@
 	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
-	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}
+	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}.
 \end{align*}
 
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. Thus we have
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. Thus, we have
 
 $$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
-$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z}))$$
+$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
 where
@@
 
-As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
-Thus we have
-
+As a function of \(\pi\), \(\bm{Z}\) is a sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
+It follows that
 \begin{align*}
@@
 &\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1} \\
+&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
 \end{align*}
-Thus the posterior $\pi^{(s+1)}|\bm{Z}^{(s+1)}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
+Hence, the posterior $\pi^{(s+1)}|\bm{Z}^{(s+1)}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
 
@@
 is a discrete distribution with probabilities proportional to
-
 \begin{align*}
@@
 		w_{ij}  & z_j \leq n_1; \\
-		1 &  z_j  = n_1 + j \\
+		1 &  z_j  = n_1 + j,\\
 	\end{cases}\\
 \end{align*}
-
 where
 %$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
-$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$.
+$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
 
@@
 	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}
+	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}.
 \end{align*}
-Thus we have
+It follows that
 $$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
 	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + 1 \\
+	1-\pi &  z_j  = n_1 + 1. \\
 \end{cases}$$
@@
 	w_{ij}  & i \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + 1 \\
+	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + 1. \\
 \end{cases}\]
@@
 We provide a summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
-{\color{blue}We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
+{We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
 \texttt{BRL} would see considerable speed up through the procedure, but this step would keep the complexity of the sampler at $O(n_A \times n_B)$.}
@@
 
-\clearpage
+%\clearpage
 

### 2021-12-07 bayesian-analysis/kundinger-submit.tex
fixing typos.

@@
 With all record pairs converted to hashed values, we can create a more compact nested list 
-$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
+$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
 where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
@@
 Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
-
 $$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
@@
 
-We demonstrate the speed, and accuracy of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
+We demonstrate the speed and accuracy of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
 

### 2021-12-07 bayesian-analysis/kundinger-submit.tex
paper at 28 pages.

@@
 \begin{figure}[t]
-	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
+	\begin{center} \includegraphics[width=0.45\textwidth]{../notes/figures/sadinle_speed_plot2} 
 		\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$.
@@
 \begin{figure}[h!]
-	\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
+	\begin{center} \includegraphics[width=0.45\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
 		\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
@@
 	\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
+\includegraphics[width=0.5\textwidth]{../notes/figures/sadinle_sim_plot2} 
 \caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
@@
 \begin{center}
-	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
+	\includegraphics[width=0.5\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
 	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
@@
 	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+		\includegraphics[width=0.45\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
 		\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
@@
 \begin{center}
-	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
+	\includegraphics[width=0.45\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
 \caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. We note they are quite similar across the two methods.}\label{fig:m-and-u}
@@
 	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
+		\includegraphics[width=0.45\textwidth]{../notes/figures/nltcs/overlap_posterior4}
 		\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
@@
 
+\hypertarget{appendix-sim}{%
+	\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
+Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+It is set up with one error across the linkage fields and 50 duplicates across files.
+Traceplots across other settings exhibit similar behavior.
+Note that traceplots for $\bf{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{../notes/figures/sim_overlap_trace} 
+		\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}}\label{fig:sim_overlap_trace}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{../notes/figures/sim_m_trace} 
+		\caption{Representative traceplot of m parameter from simulation study in Section \ref{accuracy}}\label{fig:sim_m_trace}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.45\textwidth]{../notes/figures/sim_u_trace} 
+		\caption{Representative traceplot of u parameters from simulation study in Section \ref{accuracy}}\label{fig:sim_u_trace}
+	\end{center}
+\end{figure}
+
+\clearpage
+
+\hypertarget{appendix-es}{%
+	\subsection{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.45\textwidth]{../notes/figures/el_salvador/overlap_trace} 
+\caption{Traceplot for number of matches found across datasets in El Salvador case study}\label{fig:overlap_trace}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.45\textwidth]{../notes/figures/el_salvador/m_trace} 
+\caption{Traceplot for m parameter in El Salvador case study}\label{fig:m_trace}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.45\textwidth]{../notes/figures/el_salvador/u_trace} 
+\caption{Traceplot for u parameter in El Salvador case study}\label{fig:u_trace}
+\end{center}
+\end{figure}
+
 \hypertarget{app:fabl-psuedocode}{%
@@
 \end{algorithm}
-\clearpage
-
-\hypertarget{appendix-sim}{%
-	\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
-Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
-It is set up with one error across the linkage fields and 50 duplicates across files.
-Traceplots across other settings exhibit similar behavior.
-Note that traceplots for $\bf{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
-
-\begin{figure}[!h]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-		\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}}\label{fig:sim_overlap_trace}
-	\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-		\caption{Representative traceplot of m parameter from simulation study in Section \ref{accuracy}}\label{fig:sim_m_trace}
-	\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-		\caption{Representative traceplot of u parameters from simulation study in Section \ref{accuracy}}\label{fig:sim_u_trace}
-	\end{center}
-\end{figure}
-
-\clearpage
-
-\hypertarget{appendix-es}{%
-	\subsection{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-\caption{Traceplot for number of matches found across datasets in El Salvador case study}\label{fig:overlap_trace}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-\caption{Traceplot for m parameter in El Salvador case study}\label{fig:m_trace}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-\caption{Traceplot for u parameter in El Salvador case study}\label{fig:u_trace}
-\end{center}
-\end{figure}
 

### 2021-12-07 bayesian-analysis/kundinger-submit-blind.tex
blinded version for sbss.

--- 2021-12-07 bayesian-analysis/kundinger-submit.tex
@@
 \begin{aug}
-\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
-\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
-\and
-\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
+\author{\fnms{Anonymous} \snm{Authors}\thanksref{addr1}\ead[label=e1]{}
+%\author{\fnms{} \snm{}\thanksref{addr1}\ead[label=e2]{}}
+%\and
+%\author{\fnms{} \snm{}\thanksref{addr2}\ead[label=e3]{}
+}
 
@@
 
-\address[addr1]{Department of Statistical Science,
-  Duke University,
-  P.O.\ Box 90251,
-  Durham, NC 27708, USA
+\address[addr1]{Anonymous Address
   \printead{e1}, % print email address of "e1"
-  \printead*{e2}
+%  \printead*{e2}
 }
 
-\address[addr2]{Departments of Statistical Science and Computer Science,
-  Duke University,
-  P.O.\ Box 90251,
-  Durham, NC 27708, USA
-  \printead{e3}
-}
+%\address[addr2]{Departments of Statistical Science and Computer Science,
+%  Duke University,
+%  P.O.\ Box 90251,
+%  Durham, NC 27708, USA
+%  \printead{e3}
+%}
 
@@
 \end{aug}
+
+%\begin{aug}
+%\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
+%\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
+%\and
+%\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
+%
+%\runauthor{Kundinger et al.}
+%
+%\address[addr1]{Department of Statistical Science,
+%  Duke University,
+%  P.O.\ Box 90251,
+%  Durham, NC 27708, USA
+%  \printead{e1}, % print email address of "e1"
+%  \printead*{e2}
+%}
+%
+%\address[addr2]{Departments of Statistical Science and Computer Science,
+%  Duke University,
+%  P.O.\ Box 90251,
+%  Durham, NC 27708, USA
+%  \printead{e3}
+%}
+%
+%%\thankstext{<id>}{<text>}
+%\end{aug}
 

### 2021-12-20 bayesian-analysis/kundinger-submit-final.tex
updated BA submission

--- 2021-12-06 methods_paper/fabl-paper.tex
@@
 In many data analysis tasks, analysts seek to identify duplicate records across two databases.
-This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, dalzell2018regression, tang2020}.
-In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.
+This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
+In this article, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.
 
 Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test.
-Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english, enamorado2019using}.
+Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified each pair as a match or a non-match using a likelihood ratio test.
+Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
+Additionally, some methods model records directly \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.
+
 % Possible citation for bipartite matching. 
@@
 
-	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice.
-In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record from the other.
-Here, additional declared matches are known by assumption to be false.
+The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice.
+In many situations, we know that there are no duplications within a database, meaning that each record should be linked with at most one other record.
+Thus, when the procedure results in many to one matches, some of these links must be false.
 Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
@@
 Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost.
-\cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records.
-\cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout his Gibbs sampler.
-Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
-Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets.
+\cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
+\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is quite computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records.
+\cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
+Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
+\texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
 
 In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-As suggested by \cite{heck2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
-This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-We argue that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
+Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
+This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We demonstrate that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
 Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
@@
 Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
-Section~\ref{discussion} concludes with a discussion of future work.
 
@@
 
-Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$.
+Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with $n_1$ and $n_2$ records respectively.
 Without loss of generality, denote the files such that $n_1 \geq n_2$.
@@
 \end{align}
-Though intuitive, this sparse matrix representation can become costly for large linkage tasks.
+Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks.
 More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that
@@
 \end{align}
-Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.
+We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.
 
@@
 The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables.
-For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
-We then can propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
+We then can set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
 Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$.
@@
 The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations.
-
-In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
+However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
+
+In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
 This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
@@
 
-In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
+In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
 Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent.
@@
 \end{align}
-where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters $m_{fl} = P(\gamma_{ij}^f = l|Z_j =i)$ for the matching record pairs, and $\bm{u}$ is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
+where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters for the matching record pairs $m_{fl} = P(\gamma_{ij}^f = l|Z_j =i)$ for all fields $f$ and agreement levels $\ell_f$, and $\bm{u}$ is defined similarly for non-matching pairs.
 These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.
+
+%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
 
@@
 The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
-Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
+Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
 
@@
 $$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
+where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
 This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
@@
 Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
-In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{heck2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{heck2019} who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
+In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
 We denote the fast beta prior as follows:
@@
 $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-We note here that prior from \cite{heck2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
+One choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
+We note here that prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
 
 Note that linkage in this setting is conducted at the record level, rather than at the record pair level as in the Fellegi Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of record pairs that are matches.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in the Fellegi Sunter model estimates the proportion of record pairs that are matches.
 We find $\pi$ to be more interpretable parameter than $\lambda$ in the bipartite case.
@@
 
-$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
-$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
-$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
-$$Z_j | \pi =
+%$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
+%$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
+%$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
+%$$Z_j | \pi =
+%\begin{cases} 
+%	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
+%	1-\pi &  z_j  = n_1 + j \\
+%\end{cases}$$
+%$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
+
+\begin{align*}
+	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}) \\
+	Z_j | \pi &=
 	\begin{cases} 
@@
 		1-\pi &  z_j  = n_1 + j \\
-\end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
+	\end{cases} \\
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).
+\end{align*}
 
@@
 	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
-	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}
+	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}.
 \end{align*}
@@
 As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
-Thus we have
+We have
 
 \begin{align*}
-p(\pi | \bm{Z}) &\propto p(\bm{Z}|\pi)p(\pi) \\
+	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
 	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1} \\
+	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
 \end{align*}
-Thus the posterior $\pi^{(s+1)}|\bm{Z}^{(s+1)}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
-only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
-denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).  As a function of \(Z_j\), 
-\(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\)
-is a discrete distribution with probabilities proportional to
+Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
+
+Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\).
+Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).
+As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with
 
@@
 %$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
-$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$.
+$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
 
@@
 	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}
+	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}.
 \end{align*}
-Thus we have
+Thus, we have
 $$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
 	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + 1 \\
+	1-\pi &  z_j  = n_1 + 1. \\
 \end{cases}$$
-Finally, we integrate over the posterior of \(\pi\) and rearrange terms to produce the final full conditional:
+We integrate over \(\pi\) and rearrange terms to produce the final full conditional:
 
@@
 	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
-	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
+	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
 \end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
@@
 
+%In this article, we set false matches and false non-matches to be equally costly, and thus set $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$; when we want a full estimate of the linkage parameter, we set $\theta_R = \infty$, and when we want to leave some records unmatched by the algorithm, we set $\theta_R = 0.1$.
+
 We seek to minimize the posterior expected loss, given by 
 $$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
-Here, we set $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule
+
+Here, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the intuitive decision rule
+
+%$\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$ ,
 
@@
 Since the Gibbs sampler does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable.
-We will explore this issue through the El Salvador homicide case study.
-
 To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
@@
 A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
+
+%It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable.
+We will explore this issue through the El Salvador homicide case study.
 
@@
 Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
+See Appendix ~\ref{app:ohe} for details.
 
@@
 and
-$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
-For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will shown in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
+$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}.$$
+For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
 Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.
@@
 
-The hashing procedure described above considerably reduces the memory
-needed to store the comparison information. That is, instead of storing
-\(n_1 \times n_2\) comparison vectors, we only store the
-\(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
-relating record pairs to those vectors. However, even storing these
-\(n_1 \times n_2\) scalar labels can become burdensome with large data.
-Worse, the overwhelming majority of these labels relate to record pairs
-that are clear non-matches.
-
-To address this issue, we propose a new method called storage efficient indexing (SEI). In standard indexing, one decides a priori certain criteria that all true matching pairs must satisfy, 
-and labels any record pairs that do not meet that criteria as
-non-matches. For example, one might only consider pairs with a certain
-similarity score on a field deemed to be important (like first name), or
-only pairs that exactly match on a specified number of fields. While
-generally chosen to be quite loose, establishing these criteria
-requires knowledge of the problem and invites room for human error.
-
-With SEI however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks.
-Observe that all record $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{x}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
+The hashing procedure described above considerably reduces the memory needed to store the comparison information.
+That is, instead of storing \(n_1 \times n_2\) comparison vectors, we only store the \(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities relating record pairs to those vectors.
+However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large databases.
+Additionally, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
+
+To address this issue, we propose storage efficient indexing (SEI).
+In standard indexing, one decides a priori certain criteria that all true matching pairs must satisfy, and labels any record pairs that do not meet that criteria as non-matches.
+For example, one might only consider pairs with a certain similarity score on a field deemed to be important (like first name), or only pairs that exactly match on a specified number of fields.
+While generally chosen to be quite loose, establishing these criteria requires knowledge of the problem and invites room for human error.
+
+With SEI, however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks.
+Observe that all records $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{x}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
 Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
@@
 
-Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote
-$H_p^m = \sum_{j=1}^{n_B} \bm{1}_{(Z_j, j) \in h_p}$ to be the
-number of matching record pairs with agreement pattern $p$.
-We also can easily calculate the number of non-matching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\).
+Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations.
+Denote $H_p^m = \sum_{j=1}^{n_B} \bm{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$.
+The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
 Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
@@
 
-$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
+%$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
+%\begin{cases} 
+%	w_{p}\times H_{j_p}  & p \leq P; \\
+%	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
+%\end{cases}$$
+%$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
+	%	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
+	%	0 & \text{otherwise.} \\
+	%\end{cases}$$
+	%
+	\begin{align}
+		\label{eqn:gibbs1}
+		P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
 		\begin{cases} 
@@
 			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
-\end{cases}$$
-$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
+		\end{cases}
+	\end{align}
+	\begin{align}
+		\label{eqn:gibbs2}
+		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
 			\frac{1}{H_{j_p}} & (i, j) \in h_p \\
 			0 & \text{otherwise.} \\
-\end{cases}$$
+		\end{cases}
+	\end{align}
 
 	Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
-Thus we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with the first equation above.
-After all iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.
+Thus we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with (\ref{eqn:gibbs1}) above.
+After all iterations are complete, we can back-fill the records corresponding to the agreement patterns through (\ref{eqn:gibbs2}).
 
@@
 We provide a summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
-{\color{blue}We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
-\texttt{BRL} would see considerable speed up through the procedure, but this step would keep the complexity of the sampler at $O(n_A \times n_B)$.}
+We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
+\texttt{BRL} would see considerable speed up through the procedure, but computation time would still scale with the size of the larger dataset.
 
@@
 The hashing procedure for all pairwise comparisons is also of complexity $O(F n_1 n_2)$.
-Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
+With $B$ processors available, we can split these computations across $B$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
 There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
@@
 We simulate these data for different values of \(n_1\) and \(n_2\), and compare the run-time of \texttt{fabl} against \texttt{BRL}.
-Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in the larger simulations.
+Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
 
@@
 
-We note here that \texttt{BRL} is coded in C, which makes for unfair
-comparison against \texttt{fabl}, currently built in R.
-However, the computational speed gains for \texttt{fabl} persist, especially for the larger simulations.
+	We note here that the implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
+In contrast, we use non-optimized code written in R for \texttt{fabl}.
+While this complicates comparisons, and indeed disfavors fabl, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
 Additionally, although \texttt{fabl} is amenable to parallelization, this simulation was run on a single core.
@@
 
-In Figure~\ref{fig:speed1}, we see that at low data size, \texttt{BRL} outperforms, but that \texttt{fabl} is significantly faster at handling larger data.
+	In Figure~\ref{fig:speed1}, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data.
 In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
@@
 			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$.
-We see near quadratic growth in runtime for BRL, and near linear growth for fabl.}\label{fig:speed1}
-	\end{center}
-\end{figure}
-
-The above discussion suggests that for fixed \(n_2\), computation time
-should remain mostly constant with growing \(n_1\). Simulation study
-suggests that this is true. In Figure~\ref{fig:speed2} fixing \(n_2 = 500\), we
-see linear growth for the run-time under \texttt{BRL} as \(n_1\)
-increases, with much more static run-time under \texttt{fabl}. The
-slight increases in run-time that we do see are due primarily to the
-hashing step, which again can be run in parallel for large data.
+We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:speed1}
+		\end{center}
+	\end{figure}
+
+	The above discussion suggests that for fixed \(n_2\), computation time for \texttt{fabl} should remain mostly constant with growing \(n_1\).
+In the simulation results displayed in Figure~\ref{fig:speed2} fixing \(n_2 = 500\), we see linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, with much more static run-time under \texttt{fabl}.
+The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
 
@@
 			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
-We see linear growth in runtime for BRL, and near constant runtime for fabl.}\label{fig:speed2}
+We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:speed2}
 		\end{center}
@@
 
-We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap.
-To generate these data, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
+	We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap (the number of records in common across files).
+Following methods proposed by \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
 Fields are then chosen for distortion uniformly at random.
@@
 Age and occupation are distorted through keyboard errors and missingness.
-For further details on the generation of these data, see \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}.
 
@@
 We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
-We calculate Bayes estimates of the linkage structure though the losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$.
-Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}.
+We calculate Bayes estimates of the linkage structure using the losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$.
+Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
 
@@
 		\end{tabular}
-\caption{Construction of comparison vectors for accuracy study with simulated datasets}
+		\caption{Construction of comparison vectors for accuracy study with simulated datasets.}
 		\label{Tab:sadinle_simulation_cutoffs}
@@
 
-We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and f-measure, as defined in \cite{christen_2012}.
-Recall is the proportion of true matches found by the model, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$, and precision is proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
-F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
+	We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}.
+Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$.
+Precision is proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
+The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
 In Figure~\ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, we remain confident in \texttt{fabl}'s overall accuracy.
+In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, the overall accuracy level remains high.
 
@@
 	In Figure~\ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
-However, with high amounts of overlap, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
-Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probability of matches can at times be more split across different records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
+However, with high amounts of error, and thus fewer accurate and discerning fields of information, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
+Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probabilities of matches at times can be split across more records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand than would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
 
@@
 
-In our first case study, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in both datasets.
-Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study (NLTCS), a larger linkage task that is not possible under \texttt{BRL}.
+	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017}.
+Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
+In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study (NLTCS), a larger linkage task that is not feasible in reasonable time under \texttt{BRL} with typical computing setups.
 
@@
 
-The country of El Salvador was immersed in civil war from 1980 to 1991, and throughout the time, several organizations attempted to document casualties of the conflict.
+	The country of El Salvador was immersed in civil war from 1980 to 1991.
+Throughout the time, several organizations attempted to document casualties of the conflict.
 When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
@@
 
-There are several challenges with working with this data.
+	There are several challenges with these data.
 First, both datasets have been automatically digitized, which inherently leads to some degree of typographical error.
 Second, the only fields recorded are given name, last name, date of death, and place of death.
-It is relatively common for aparent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between
-\texttt{BRL} and \texttt{fabl}.
-
-Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files in ERTL.
-We standardized names to account for common misspellings and used a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
-Place of death is recorded by municipality and department within that municipality; however, since department was missing in 95\% of records in CHDES and 80\% of records in ERTL, this was excluded from our analyses.
+It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
+
+	%This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between \texttt{BRL} and \texttt{fabl}.
+
+	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in \(n_1 = 4420\) records in CHDES and \(n_2 = 1323\) records in ERTL.
+We standardize names to account for common misspellings and use a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
+Place of death is recorded by municipality and department within that municipality; however, since department is missing in 95\% of records in CHDES and 80\% of records in ERTL, we exclude department from our analysis.
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
@@
 
-To mirror the original implementation of \cite{sadinle_bayesian_2017}, we first construct the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
+	\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
 This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
-Gibbs sampling for these comparison vectors took 265 seconds for \texttt{fabl} and 240 for \texttt{BRL}; here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
-However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
+We initially used these comparisons to estimate fabl and BRL.
+However, after investigating the posterior distributions of the $\bm{m}$ and $\bm{u}$ parameters, we noticed that posterior distributions of several levels of the m and u parameters are nearly identical, suggesting that these levels are unnecessary for meaningfully modeling the distribution of the comparison vectors.
+
+	%However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
+
+	%here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
 
@@
 
-Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}, and obtained analogous inference.
-With 216 possible agreement patterns, 159 were realized in the data, and our proposed method, \texttt{fabl} became much faster, finishing in 61 seconds.
-Meanwhile \texttt{BRL} took 239 seconds, relatively unchanged from the first implementation.
-This demonstrates the way that the computational complexity of our method depends on the number of unique agreement patterns, and how significant computational gains can be made by simplifying the construction of the comparison vectors.
-The estimates of the \(\bm{m}\) parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
-Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+	Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}.
+Among the 216 possible agreement patterns, 159 are realized in the data.
+With this revised comparison specification, fabl ran in 61 seconds, approximately 4 times faster than BRL 239 second run time.
+The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
+Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+
+	For completeness, we note that the more detailed comparison vector produced a run time of 240 seconds for BRL, and 261 seconds for fabl.
+Apparently, the number of patterns was sufficiently many that the computational savings from fabl does not overcome the inherent speed differences of C as opposed to R.
 
@@
 We calculate posterior samples of the size of the overlap across files by finding the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
-The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate is quite conservative.
-This is because there a large number of records in ERTL for which there are multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
-We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study.
+The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate identifies fewer matches than the Gibbs sampler identifies on average.
+This is because a large number of records in ERTL have multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
+We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Section ~\ref{accuracy}.
 Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
@@
 	We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
+We note that Bayes estimates falling outside of posterior credible intervals has occurred previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
+
+	We note that reconciling Bayes estimates that fall outside of posterior credible intervals is a common challenge in record linkage and an avenue of future work.
 
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-\caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. We note they are quite similar across the two methods.}\label{fig:m-and-u}
+			\caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
 			\label{fig:m-and-u}
@@
 
-The National Long Term Care Study (NLTCS) is a longitidunal study tracking the health outcomes of Medicare recipients.
-The initial began in 1982, with follow-up surveys taken approximately every 5 years.
-For our purposes, we are interested in finding matching records over the $n_A = 20485$ records from 1982 to the $n_B = 17466$ records from 1989.
-
-We conduct linkage using sex, date of birth, and location using the thresholds shown in Table~\ref{Tab:nltcs-comparisons}.
-Note that storing 5 comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
-Standard settings on 16BG personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task.
+	The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients.
+The initial began in 1982, with follow-up surveys taken approximately every five years.
+For this study, we seek to link records over the $n_A = 20485$ records from 1982 to the $n_B = 17466$ records from 1989.
+The NLTCS data have longitudinal links, so that in reality one does not need to do a record linkage.
+However, following the strategy in \cite{guha:reiter:BA} we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone files.
+
+	We link records using sex, date of birth, and location using the thresholds shown in Table~\ref{Tab:nltcs-comparisons}.
+Storing five comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
+Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
 However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and compress information through storage efficient indexing; the resulting data object is just 10 MB, approximately 0.1\% of what is required for the raw comparisons.
@@
 	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-Our Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
-Since we access to unique identifiers for each record, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
+The Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
+Traceplots indicate quick convergence, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
 
@@
 
-\section{Discussion}
+	\section{Conclusion}
 	\label{discussion}
 
-In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. Specifically, our proposed hashing method
-and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making our approach particularly powerful in linking records when one datafile is substantially smaller than the other.
-
-Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
+	In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. We have proven that the proposed hashing method
+	and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset.
+This makes \texttt{fabl} computationally advantageous in many linkage scenarios, particularly when one datafile is substantially smaller than the other.
+We have demonstrated the speed and accuracy of \texttt{fabl} by replicating a simulation study and a case study in \cite{sadinle_bayesian_2017}, and through an additional case study that is computationally infeasible under \texttt{BRL}.
+Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
+
+	%Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
 We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
@@
 Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
-
-To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
+	%
+	%To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
 This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
-Similar work has been been proposed by \cite{heck2019} and \cite{larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
+Similar work has been been proposed by \cite{wortman2019} and \cite{Larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
 We see these types of heirarchical record linkage models as potential contributions to the growing algorithmic fairness literature, and useful to practitioners working with or studying marginalized populations.
@@
 
+	\hypertarget{app:ohe}{%
+		\subsection{One Hot Encoding Transformation}\label{app:ohe}}
+
+	\texttt{fabl} makes use of one-hot encodings to aid in vectorized computations.
+For $\gamma_{ij}^f$ with $L_f$ levels, define $e_{ij}^f$ to be an $L_f \times 1$ vector.
+When $\gamma_{ij}^f = \ell$, we set the $\ell^{th}$ element of $e_{ij}^f$ to be 1, and set the other $L_f - 1$ elements of $e_{ij}^f$ to be 0.
+We then concatenate the $e_{ij}^f$ for all $f \in \{1, \ldots, F\}$, resulting in the one-hot encoded comparison vector $e_{ij}$ of length $\sum_{f=1}^F L_f$.
+
+	For example, consider comparing the toy records shown in Table \ref{tab:ohe} with $L = (3, 3, 2, 2)$ levels of agreement for last name, first name, DOB, and city respectively.
+Since the first name differs by only one letter, a reasonable comparison vector for this pair would be $\gamma_{ij} = (1, 2, 1, 2)$.
+The one hot encoding representation of this vector is $e_{ij} = (1, 0, 0, 0, 1, 0, 1, 0, 0, 1)$.
+	\begin{table}[h!]
+		\centering
+		\begin{tabular}[h!]{llll}
+			\hline
+			Last Name & First Name & DOB & City \\
+			\hline
+			Smith & Taylor & 01/01/2000 & Durham\\
+			Smith & Tayler & 01/01/2000 & Raleigh\\
+			\hline
+		\end{tabular}
+		\caption{Example records for one hot encoding.}\label{tab:ohe}
+	\end{table}
+
 	\hypertarget{appendix-sim}{%
 		\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
-Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+	Below are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
 It is set up with one error across the linkage fields and 50 duplicates across files.
@@
 	\end{figure}
-
 \end{document}

### 2021-12-20 methods_paper/fabl-paper-blind.tex
updated BA submission

--- 2021-12-06 methods_paper/fabl-paper.tex
@@
 \title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
-\author{Brian Kundinger\textsuperscript{a} \and
-  Jerome Reiter\textsuperscript{a} \and 
-  Rebecca C.~Steorts \textsuperscript{b}}
-\date{
- \textsuperscript{a} Department of Statistical Science, Duke University \\
- \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
-  \today}
+%\author{Brian Kundinger\textsuperscript{a} \and
+%  Jerome Reiter\textsuperscript{a} \and 
+%  Rebecca C.~Steorts \textsuperscript{b}}
+%\date{
+% \textsuperscript{a} Department of Statistical Science, Duke University \\
+% \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
+%  \today}
 

### 2021-12-20 methods_paper/fabl-paper.tex
updated BA submission

@@
 \author{Brian Kundinger\textsuperscript{a} \and
-  Jerome Reiter\textsuperscript{a} \and 
+  Jerome P. Reiter\textsuperscript{a} \and 
   Rebecca C.~Steorts \textsuperscript{b}}
@@
 However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.
-Specifically, in \texttt{fabl} we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
+We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.
+Specifically, in \texttt{fabl} we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
 This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
-Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
+Through simulations and two case studies, we show that \texttt{fabl} has increased speed with minimal loss of accuracy.
 
@@
 	In many data analysis tasks, analysts seek to identify duplicate records across two databases.
-This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, dalzell2018regression, tang2020}.
-In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.
+This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
+In this article, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.
 
 	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test.
-Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english, enamorado2019using}.
+Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified each pair as a match or a non-match using a likelihood ratio test.
+Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
+Additionally, some methods model records directly \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.
+
 	% Possible citation for bipartite matching. 
@@
 
-	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice.
-In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record from the other.
-Here, additional declared matches are known by assumption to be false.
+	The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice.
+In many situations, we know that there are no duplications within a database, meaning that each record should be linked with at most one other record.
+Thus, when the procedure results in many to one matches, some of these links must be false.
 Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
@@
 	Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost.
-\cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records.
-\cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout his Gibbs sampler.
-Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review.
-Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets.
+\cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
+\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is quite computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records.
+\cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
+Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
+\texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
 
 	In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-As suggested by \cite{heck2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
-This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-We argue that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
+Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
+This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We demonstrate that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
 Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
@@
 
-$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
-$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
-$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
-$$Z_j | \pi =
+%$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
+%$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
+%$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
+%$$Z_j | \pi =
+%\begin{cases} 
+%	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
+%	1-\pi &  z_j  = n_1 + j \\
+%\end{cases}$$
+%$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
+
+\begin{align*}
+\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
+\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}) \\
+Z_j | \pi &=
 \begin{cases} 
@@
 	1-\pi &  z_j  = n_1 + j \\
-\end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
+\end{cases} \\
+\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).
+\end{align*}
 
@@
 As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
-Thus we have
+We have
 
 \begin{align*}
-p(\pi | \bm{Z}) &\propto p(\bm{Z}|\pi)p(\pi) \\
+p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
 &\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1} \\
+&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
 \end{align*}
-Thus the posterior $\pi^{(s+1)}|\bm{Z}^{(s+1)}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
-only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
-denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).  As a function of \(Z_j\), 
-\(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\)
-is a discrete distribution with probabilities proportional to
+Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
+
+Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\).
+Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).
+As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with
 
@@
 
-$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
+%$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
+%\begin{cases} 
+%	w_{p}\times H_{j_p}  & p \leq P; \\
+%	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
+%\end{cases}$$
+%$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
+%	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
+%	0 & \text{otherwise.} \\
+%\end{cases}$$
+%
+\begin{align}
+	\label{eqn:gibbs1}
+	P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
 	\begin{cases} 
@@
 		n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
-\end{cases}$$
-$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
+	\end{cases}
+\end{align}
+\begin{align}
+	\label{eqn:gibbs2}
+	P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
 	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
 	0 & \text{otherwise.} \\
-\end{cases}$$
+\end{cases}
+\end{align}
 
 Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
-Thus we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with the first equation above.
-After all iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.
+Thus we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with (\ref{eqn:gibbs1}) above.
+After all iterations are complete, we can back-fill the records corresponding to the agreement patterns through (\ref{eqn:gibbs2}).
 
@@
 
-We note here that \texttt{BRL} is coded in C, which makes for unfair
-comparison against \texttt{fabl}, currently built in R.
-However, the computational speed gains for \texttt{fabl} persist, especially for the larger simulations.
+We note here that the implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
+In contrast, we use non-optimized code written in R for \texttt{fabl}.
+While this complicates comparisons, and indeed disfavors fabl, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
 Additionally, although \texttt{fabl} is amenable to parallelization, this simulation was run on a single core.
@@
 
-In Figure~\ref{fig:speed1}, we see that at low data size, \texttt{BRL} outperforms, but that \texttt{fabl} is significantly faster at handling larger data.
+In Figure~\ref{fig:speed1}, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data.
 In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
@@
 		\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$.
-We see near quadratic growth in runtime for BRL, and near linear growth for fabl.}\label{fig:speed1}
-	\end{center}
-\end{figure}
-
-The above discussion suggests that for fixed \(n_2\), computation time
-should remain mostly constant with growing \(n_1\). Simulation study
-suggests that this is true. In Figure~\ref{fig:speed2} fixing \(n_2 = 500\), we
-see linear growth for the run-time under \texttt{BRL} as \(n_1\)
-increases, with much more static run-time under \texttt{fabl}. The
-slight increases in run-time that we do see are due primarily to the
-hashing step, which again can be run in parallel for large data.
+We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:speed1}
+	\end{center}
+\end{figure}
+
+The above discussion suggests that for fixed \(n_2\), computation time for \texttt{fabl} should remain mostly constant with growing \(n_1\).
+In the simulation results displayed in Figure~\ref{fig:speed2} fixing \(n_2 = 500\), we see linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, with much more static run-time under \texttt{fabl}.
+The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
 
@@
 		\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
-We see linear growth in runtime for BRL, and near constant runtime for fabl.}\label{fig:speed2}
+We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:speed2}
 	\end{center}
@@
 
-We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap.
-To generate these data, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
+We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap (the number of records in common across files).
+Following methods proposed by \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
 Fields are then chosen for distortion uniformly at random.
@@
 	\end{tabular}
-\caption{Construction of comparison vectors for accuracy study with simulated datasets}
+\caption{Construction of comparison vectors for accuracy study with simulated datasets.}
 \label{Tab:sadinle_simulation_cutoffs}
@@
 
-We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and f-measure, as defined in \cite{christen_2012}.
-Recall is the proportion of true matches found by the model, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$, and precision is proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
-F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
+We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}.
+Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$.
+Precision is proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
+The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
 In Figure~\ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, we remain confident in \texttt{fabl}'s overall accuracy.
+In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, the overall accuracy level remains high.
 
@@
 In Figure~\ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
-However, with high amounts of overlap, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
-Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probability of matches can at times be more split across different records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
+However, with high amounts of error, and thus fewer accurate and discerning fields of information, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
+Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probabilities of matches at times can be split across more records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand than would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
 
@@
 
-In our first case study, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in both datasets.
-Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study (NLTCS), a larger linkage task that is not possible under \texttt{BRL}.
+In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017}.
+Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
+In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study (NLTCS), a larger linkage task that is not feasible in reasonable time under \texttt{BRL} with typical computing setups.
 
@@
 
-The country of El Salvador was immersed in civil war from 1980 to 1991, and throughout the time, several organizations attempted to document casualties of the conflict.
+The country of El Salvador was immersed in civil war from 1980 to 1991.
+Throughout the time, several organizations attempted to document casualties of the conflict.
 When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
@@
 
-There are several challenges with working with this data.
+There are several challenges with these data.
 First, both datasets have been automatically digitized, which inherently leads to some degree of typographical error.
 Second, the only fields recorded are given name, last name, date of death, and place of death.
-It is relatively common for aparent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between
-\texttt{BRL} and \texttt{fabl}.
-
-Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files in ERTL.
-We standardized names to account for common misspellings and used a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
-Place of death is recorded by municipality and department within that municipality; however, since department was missing in 95\% of records in CHDES and 80\% of records in ERTL, this was excluded from our analyses.
+It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
+
+%This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between \texttt{BRL} and \texttt{fabl}.
+
+Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in \(n_1 = 4420\) records in CHDES and \(n_2 = 1323\) records in ERTL.
+We standardize names to account for common misspellings and use a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
+Place of death is recorded by municipality and department within that municipality; however, since department is missing in 95\% of records in CHDES and 80\% of records in ERTL, we exclude department from our analysis.
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
@@
 
-To mirror the original implementation of \cite{sadinle_bayesian_2017}, we first construct the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
+\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
 This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
-Gibbs sampling for these comparison vectors took 265 seconds for \texttt{fabl} and 240 for \texttt{BRL}; here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
-However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
+We initially used these comparisons to estimate fabl and BRL.
+However, after investigating the posterior distributions of the $\bm{m}$ and $\bm{u}$ parameters, we noticed that posterior distributions of several levels of the m and u parameters are nearly identical, suggesting that these levels are unnecessary for meaningfully modeling the distribution of the comparison vectors.
+
+%However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
+
+%here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
 
@@
 
-Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}, and obtained analogous inference.
-With 216 possible agreement patterns, 159 were realized in the data, and our proposed method, \texttt{fabl} became much faster, finishing in 61 seconds.
-Meanwhile \texttt{BRL} took 239 seconds, relatively unchanged from the first implementation.
-This demonstrates the way that the computational complexity of our method depends on the number of unique agreement patterns, and how significant computational gains can be made by simplifying the construction of the comparison vectors.
-The estimates of the \(\bm{m}\) parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
-Additionally, traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}.
+Among the 216 possible agreement patterns, 159 are realized in the data.
+With this revised comparison specification, fabl ran in 61 seconds, approximately 4 times faster than BRL 239 second run time.
+The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
+Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+
+For completeness, we note that the more detailed comparison vector produced a run time of 240 seconds for BRL, and 261 seconds for fabl.
+Apparently, the number of patterns was sufficiently many that the computational savings from fabl does not overcome the inherent speed differences of C as opposed to R.
 
@@
 We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
+We note that Bayes estimates falling outside of posterior credible intervals has occurred previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
+
+We note that reconciling Bayes estimates that fall outside of posterior credible intervals is a common challenge in record linkage and an avenue of future work.
 
@@
 	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-\caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. We note they are quite similar across the two methods.}\label{fig:m-and-u}
+\caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
 \label{fig:m-and-u}
@@
 
-The National Long Term Care Study (NLTCS) is a longitidunal study tracking the health outcomes of Medicare recipients.
-The initial began in 1982, with follow-up surveys taken approximately every 5 years.
-For our purposes, we are interested in finding matching records over the $n_A = 20485$ records from 1982 to the $n_B = 17466$ records from 1989.
-
-We conduct linkage using sex, date of birth, and location using the thresholds shown in Table~\ref{Tab:nltcs-comparisons}.
-Note that storing 5 comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
-Standard settings on 16BG personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task.
+The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients.
+The initial began in 1982, with follow-up surveys taken approximately every five years.
+For this study, we seek to link records over the $n_A = 20485$ records from 1982 to the $n_B = 17466$ records from 1989.
+The NLTCS data have longitudinal links, so that in reality one does not need to do a record linkage.
+However, following the strategy in \cite{guha:reiter:BA} we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone files.
+
+We link records using sex, date of birth, and location using the thresholds shown in Table~\ref{Tab:nltcs-comparisons}.
+Storing five comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
+Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
 However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and compress information through storage efficient indexing; the resulting data object is just 10 MB, approximately 0.1\% of what is required for the raw comparisons.
@@
 We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-Our Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
-Since we access to unique identifiers for each record, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
+The Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
+Traceplots indicate quick convergence, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
 
@@
 
-\section{Discussion}
+\section{Conclusion}
 \label{discussion}
 
-In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. Specifically, our proposed hashing method
-and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset, making our approach particularly powerful in linking records when one datafile is substantially smaller than the other.
-
-Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
+In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. We have proven that the proposed hashing method
+and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset.
+This makes \texttt{fabl} computationally advantageous in many linkage scenarios, particularly when one datafile is substantially smaller than the other.
+We have demonstrated the speed and accuracy of \texttt{fabl} by replicating a simulation study and a case study in \cite{sadinle_bayesian_2017}, and through an additional case study that is computationally infeasible under \texttt{BRL}.
+Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
+
+%Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
 We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
@@
 Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
-
-To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
+%
+%To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
 This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
-Similar work has been been proposed by \cite{heck2019} and \cite{larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
+Similar work has been been proposed by \cite{wortman2019} and \cite{Larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
 We see these types of heirarchical record linkage models as potential contributions to the growing algorithmic fairness literature, and useful to practitioners working with or studying marginalized populations.
@@
 
+\hypertarget{app:ohe}{%
+	\subsection{One Hot Encoding Transformation}\label{app:ohe}}
+
+\texttt{fabl} makes use of one-hot encodings to aid in vectorized computations.
+For $\gamma_{ij}^f$ with $L_f$ levels, define $e_{ij}^f$ to be an $L_f \times 1$ vector.
+When $\gamma_{ij}^f = \ell$, we set the $\ell^{th}$ element of $e_{ij}^f$ to be 1, and set the other $L_f - 1$ elements of $e_{ij}^f$ to be 0.
+We then concatenate the $e_{ij}^f$ for all $f \in \{1, \ldots, F\}$, resulting in the one-hot encoded comparison vector $e_{ij}$ of length $\sum_{f=1}^F L_f$.
+
+For example, consider comparing the toy records shown in Table \ref{tab:ohe} with $L = (3, 3, 2, 2)$ levels of agreement for last name, first name, DOB, and city respectively.
+Since the first name differs by only one letter, a reasonable comparison vector for this pair would be $\gamma_{ij} = (1, 2, 1, 2)$.
+The one hot encoding representation of this vector is $e_{ij} = (1, 0, 0, 0, 1, 0, 1, 0, 0, 1)$.
+\begin{table}[h!]
+	\centering
+	\begin{tabular}[h!]{llll}
+		\hline
+		Last Name & First Name & DOB & City \\
+		\hline
+		Smith & Taylor & 01/01/2000 & Durham\\
+		Smith & Tayler & 01/01/2000 & Raleigh\\
+		\hline
+	\end{tabular}
+	\caption{Example records for one hot encoding.}\label{tab:ohe}
+\end{table}
+
 \hypertarget{appendix-sim}{%
 	\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
-Below are traceplots for one 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+Below are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
 It is set up with one error across the linkage fields and 50 duplicates across files.

### 2022-07-14 Prelim/google_research_overview.tex
uploaded editor comments

--- 2021-11-15 Prelim/prelim.tex
@@
 	\title{Efficient and Scalable Bayesian Bipartite Matching through Fast Beta Linkage (\texttt{fabl})}
-	%\subtitle{Using Beamer}
+	\subtitle{Google Version}
 	\author{Brian Kundinger, Jerome Reiter, Rebecca Steorts}
@@
 
-\begin{frame}{Linkage through Comparison Vectors}
-	Represent linkage structure through vector $\mathbf{Z} = \{Z_1, \ldots, Z_{n_B}\}$, where
-\begin{center}
-	$Z_j = \begin{cases} 
-		i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
-		n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
-	\end{cases}$
-\end{center}
-\end{frame}
-
-\begin{frame}{Linkage through Comparison Vectors}
-	\includegraphics[width = \textwidth, height = .6\textwidth ]{graphics/Slide6.png}
-\end{frame}
+%\begin{frame}{Linkage through Comparison Vectors}
+%	Represent linkage structure through vector $\mathbf{Z} = \{Z_1, \ldots, Z_{n_B}\}$, where
+%\begin{center}
+%	$Z_j = \begin{cases} 
+%		i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
+%		n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
+%	\end{cases}$
+%\end{center}
+%\end{frame}
+%
+%\begin{frame}{Linkage through Comparison Vectors}
+%	\includegraphics[width = \textwidth, height = .6\textwidth ]{graphics/Slide6.png}
+%\end{frame}
 
@@
 
-	Model specification allows for \textbf{hashing} of comparison vectors, \textbf{distributed} computing, and \textbf{storage efficient indexing (SEI)}
+	Note: the first three lines are common to many record linkage models, and bear similarities to a larger family of \textit{latent class models.} The last two lines are the newly developed prior distribution for the set of matches.
 \end{frame}
 
-\begin{frame}{Hashing}
-		\begin{columns}
-		\begin{column}{0.6\textwidth}
-			\begin{itemize}
-				\item<1-> Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records (Enamorado et al 2019).
-				\begin{itemize}
-					\item<2-> $L = \{3, 3, 2, 2\}$ implies 36 unique patterns 
-				\end{itemize} 
-				\item<3-> When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p.$
-				\item<4-> Allows us to compute sufficient statistics and reduce computational complexity from $O(n_A \times n_B)$ to $O(P \times n_B)$
-			\end{itemize}
-		\end{column}
-		\begin{column}{0.38\textwidth}
-			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{graphics/unique_patterns.png}
-		\end{column}
-	\end{columns}
-\end{frame}
-
-\begin{frame}{Hashing}
-	\centering
-	\includegraphics<1>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide17.png}
-	\includegraphics<2>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide18.png}
-	\includegraphics<3>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide19.png}
-	\includegraphics<4>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide20.png}
-	\includegraphics<5>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide21.png}	
-\end{frame}
-
-\begin{frame}{Hashing to Scale to Large Linkage Tasks}
-	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide23.png}
-	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide24.png}
-	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide25.png}
-	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide26.png}
-	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide27.png}
-\end{frame}
-
-\begin{frame}{Hashing for Speed (Efficient Gibbs Sampling)}
-	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide37.png}
-	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide38.png}
-	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide39.png}
-	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide40.png}
-	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide41.png}
-	\includegraphics<6>[width = \textwidth, height = .6\textwidth ]{graphics/Slide42.png}
-\end{frame}
+%\begin{frame}{Hashing}
+%		\begin{columns}
+%		\begin{column}{0.6\textwidth}
+%			\begin{itemize}
+%				\item<1-> Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records (Enamorado et al 2019).
+%				\begin{itemize}
+%					\item<2-> $L = \{3, 3, 2, 2\}$ implies 36 unique patterns 
+%				\end{itemize} 
+%				\item<3-> When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p.$
+%				\item<4-> Allows us to compute sufficient statistics and reduce computational complexity from $O(n_A \times n_B)$ to $O(P \times n_B)$
+%			\end{itemize}
+%		\end{column}
+%		\begin{column}{0.38\textwidth}
+%			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{graphics/unique_patterns.png}
+%		\end{column}
+%	\end{columns}
+%\end{frame}
+%
+%\begin{frame}{Hashing}
+%	\centering
+%	\includegraphics<1>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide17.png}
+%	\includegraphics<2>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide18.png}
+%	\includegraphics<3>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide19.png}
+%	\includegraphics<4>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide20.png}
+%	\includegraphics<5>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide21.png}	
+%\end{frame}
+%
+%\begin{frame}{Hashing to Scale to Large Linkage Tasks}
+%	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide23.png}
+%	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide24.png}
+%	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide25.png}
+%	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide26.png}
+%	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide27.png}
+%\end{frame}
+%
+%\begin{frame}{Hashing for Speed (Efficient Gibbs Sampling)}
+%	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide37.png}
+%	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide38.png}
+%	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide39.png}
+%	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide40.png}
+%	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide41.png}
+%	\includegraphics<6>[width = \textwidth, height = .6\textwidth ]{graphics/Slide42.png}
+%\end{frame}
 
@@
 
-\begin{frame}{Extensions and Future Directions}
+\begin{frame}{Glaring Questions}
 	\begin{itemize}
-		\item Linkage when reliability of information and rates of matching differs by subgroup in the data
+		\item Can you do record linkage with arbitrary amounts of duplicated files?
+		\begin{itemize}
+			\item<2-> Yes! See Aleshin-Guendel and Sadinle (2021)
+		\end{itemize}
 
-		\item Linkage when there are duplicates within datasets
+		\item Isn't $n_A \times n_B$ record pairs computationally infeasible?
+		\begin{itemize}
+			\item<3-> Yes! My work with the DNC is forthcoming, Kundinger, Wortman, Reiter, and Steorts (2022)
+		\end{itemize}
 
-		\item Linkage over blocked data (allows for much larger linkage tasks)
+		\item Can you account situations when the reliability of information differs throughout the data?
+		\begin{itemize}
+			\item<4-> Yes! I have a heirarchical model in the works. Hopefully Kundinger et al (2023)
+		\end{itemize}
+
+		\item Can you do record linkage on the records themselves, rather than transforming to comparison vectors?
+		\begin{itemize}
+			\item<5-> Yes! See \texttt{d-blink} package from Marchant et al (2021)
+		\end{itemize}
+
 	\end{itemize}
@@
 
-\section{Appendix}
+\section{Possible Relevance to Google}
+\begin{frame}{Latent Class Analysis}
+	The record linkage model presented is a special case of a "latent class model". In my work, we attempt to classify record pairs into two distinct classes: matching, and non-matching pairs.
+	\linebreak
 
-\begin{frame}{Storage Efficient Indexing}
-	\begin{itemize}
-		\item Storing $n_A \times n_B$ hashed values burdensome for large tasks
-		\item Overwhelming majority of record pairs show nonagreement
-		\item<3> Correct counts for calculations stored in $\mathcal{H}$
-	\end{itemize}
-	\includegraphics<1>[width = \textwidth]{graphics/before_SEI.png}
-	\includegraphics<2->[width = \textwidth]{graphics/after_SEI2.png}
+	One can imagine however, that we can look at articles instead of records, look at blocks of text instead of record pairs, and look at attributes of those blocks instead of comparison vectors.
+This may provide a strategy for \textbf{entirely unsupervised} byline detection.
+	\linebreak
+
+	Additionally, if model parameters are accurate, we can train the model on a small set of articles, and then apply those parameters out onto a larger set of articles, for nearly instantaneous latent class analysis.
 \end{frame}
 
-%		\item Reduce data to sufficient statistics
-%		\begin{itemize}
-%		\item $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
-%		\item $H_{p_j} = ||r_{p_j}||$
-%		\item $H_p = \sum_{j} H_{p_j}$
-%		\end{itemize}
-%		
-%		\item Run Gibbs sampler at level of agreement patterns, not record pairs
-%		\begin{itemize}
-%		\item Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
-%		\item Use number of matches for each pattern to update $m$ and $u$
-%		\item Back fill record labels at the end through $r_{p_j}$
-%	\end{itemize}
-%
-%		\item Reduces computational complexity from $O(n_A \times n_B \times F)$ to $O(P \times n_B \times F)$
-
 \end{document}

### 2022-07-14 methods_paper/cover-letter/ba-cover-letter.tex
uploaded editor comments

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[top=3in,  % FIRST, CHANGE THIS NUMBER TO GET THE SECOND PAGE 
+\usepackage[top=1in,  % FIRST, CHANGE THIS NUMBER TO GET THE SECOND PAGE 
                       % LOOKING THE WAY YOU WANT IT
@@
 
-\usepackage{eso-pic}
-\usepackage{graphicx}
+%\usepackage{eso-pic}
+%\usepackage{graphicx}
+\usepackage{parskip}
 
@@
 %\date{June 9, 2010}
-\date{\raggedright November 29, 2021}
+%\date{\raggedright November 29, 2021}
 
 \begin{document}
-
-\AddToShipoutPicture*{\BackgroundPic}
-\signature{\vspace*{-4em}\includegraphics[scale=1]{sign.pdf}} 
-\thispagestyle{empty}
+\date{\raggedright December 21, 2021}
+%\AddToShipoutPicture*{\BackgroundPic}
+%\signature{\vspace*{-4em}\includegraphics[scale=1]{sign.pdf}} 
+%\thispagestyle{empty}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace*{-.2in}  % AFTER YOU HAVE THE SECOND PAGE THE WAY YOU WANT IT,
+%\vspace*{-.2in}  % AFTER YOU HAVE THE SECOND PAGE THE WAY YOU WANT IT,
                 % CHANGE THIS NUMBER TO GET THE FIRST PAGE HOW YOU WANT IT
@@
 I am a PhD student at Duke University in the Department of Statistical Science. I am writing to submit my paper with my coauthors
-Jerome Reiter and Rebecca C. Steorts entitled ``Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)" to \emph{Bayesian Analysis}.
+Jerome P. Reiter and Rebecca C. Steorts entitled ``Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)" to \emph{Bayesian Analysis}.
 
@@
 Disciplines that utilize both applications and methods include statistics, biostatistics, computer science, machine learning, political science, and others.
-The goal of our review article is to review nearly almost all of entity resolution in such a way that it can be read and understood by as many communities as possible given that the current state of the literature is mathematical and overly technical, making the entry point difficult for those in the computational social sciences.
 
@@
 
-To help the editorial office, we have listed a number of potential reviewers that do not have any conflicts with the authors:
+To help the editorial office, we have listed a number of potential reviewers:
 
-Mauricio Sadinle, msadinle@uw.edu, University of Washington; Serge Aleshin-Guendel, aleshing@uw.edu, University of Washington; Yves Thibaudeau, Yves.Thibaudeau@census.gov, United States Census Bureau; Megan Price, meganp@hrdag.org, Human Rights Data Analysis Group; Ted Enamorado, ted@wustl.edu, University of Washington St.
-Louis; Nick Monath, nmonath@cs.umass.edu; Dinusha Vatsalan, Dinusha.Vatsalan@data61.csiro.au, Ben Rubinstein; benjamin.rubinstein@unimelb.edu.au; Neil Marchant; nmarchant@unimelb.edu.au; Andy Toole; Andrew.Toole@uspto.gov;  jph34@stat.duke.edu; Giacomo Zanella, zanella.gcm@gmail.com, Brunero Liseo, brunero.liseo@uniroma1.it, Andrea Tancredi, andrea.tancredi@uniroma1.it;
-Peter Christen, peter.christen@anu.edu.au, Partha Lahiri, plahiri@umd.edu, Rebecca Nugent, rnugent@stat.cmu.edu, Kayla Frisoli, kaylafrisoli@gmail.com.
+\setlength{\parindent}{0mm}
+\hangindent=10mm
+\hangafter=0
+Mauricio Sadinle, msadinle@uw.edu \newline
+Serge Aleshin-Guendel, aleshing@uw.edu \newline
+Yves Thibaudeau, Yves.Thibaudeau@census.gov \newline
+Megan Price, meganp@hrdag.org \newline
+Ted Enamorado, ted@wustl.edu \newline
+Nick Monath, nmonath@cs.umass.edu \newline
+Dinusha Vatsalan, Dinusha.Vatsalan@data61.csiro.au \newline
+Ben Rubinstein, benjamin.rubinstein@unimelb.edu.au \newline
+Neil Marchant, nmarchant@unimelb.edu.au \newline
+Andy Toole, Andrew.Toole@uspto.gov \newline
+Giacomo Zanella, zanella.gcm@gmail.com \newline
+Brunero Liseo, brunero.liseo@uniroma1.it \newline
+Andrea Tancredi, andrea.tancredi@uniroma1.it \newline
+Peter Christen, peter.christen@anu.edu.au \newline
+Partha Lahiri, plahiri@umd.edu \newline
+Rebecca Nugent, rnugent@stat.cmu.edu \newline
+Kayla Frisoli, kaylafrisoli@gmail.com \newline
 
@@
  \closing{Sincerely,}
-
- Rebecca C. Steorts
+ Brian Kundinger
 

### 2022-07-14 notes/beta_linkage_blocked_data_slides.Rmd
uploaded editor comments

---
title: "Beta Linkage on Blocked Data"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
library(RecordLinkage)
library(tidyverse)
```

# Overview

- Big Picture

- Review of Base Model

- Modified Model for Blocking

- Challenges From Blocking

- Informative Priors

- "Train and Test" Method for Scalability

# Big Picture

The biggest drawback to the Fellegi-Sunter approach of record linkage is the extremely large number of comparison vectors that must be created. This can be heavily reduced through blocking, but this blocking is not well addressed in the literature. Generally, one performs blocking on a reliable (usually categorical) field, and then applies a record linkage algorithm independently on each block of data.

# Intensive Blocking

However, if you block on more discerning fields, like first and last name you end up with many more, and much smaller, blocks. It is often mathematically impossible to run independent record linkage over such small blocks.

(If a block consists of one Brian Kundinger in file A and one Brian Kundinger in file B, how can I separate pairs into matches and nonmatches? How could I estimate the $2\sum_{f=1}^F L_f + 1$ parameters of the model on only one pair?)

Previous approaches run the record linkage algorithm unmodified over this filtered data all at once (Wortman 2021, Sadinle 2014). However, we have seen poor estimation of $u$ parameters, and thus poor overall performance.

# Fast Beta Linkage

By now, we are all familiar with the model used in Fast Beta Linkage. For fields $f \in \{1, \ldots, F\}$ and levels $l\in \{1, \ldots, L_f\}$ we adopt the following likelihood and prior distributions.

$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \pi) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
$$\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})$$
$$\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})$$
$$Z_j | \pi =
\begin{cases} 
    \frac{1}{n_A}\pi  & z_j \leq n_A; \\
     1-\pi &  z_j  = n_A + 1 \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) $$

# Model Specification Under Blocking

Under blocking, the likelihood becomes

$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \pi) =\prod_{j=1}^{n_B}  \prod_{i \in N_j}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
where $N_j$ is the set of candidate records in $A$ for record $j \in B$ after blocking. For the new likelihood, we need to modify the prior:

$$Z_j | \pi =
\begin{cases} 
    \frac{1}{N_j}\pi  & z_j \leq n_A; \\
     1-\pi &  z_j  = n_A + 1 \\
\end{cases}$$

# Gibbs Sampling

Under this new likelihood and prior, the full conditional for the components of the linkage parameter $mathbf{Z}$ becomes

$$P\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
    w_{ij}  & i \leq n_A; \\
     N_j \frac{N_B - D + \beta_{\pi}}{D + \alpha_{\pi}} & i  = n_A + 1 \\
\end{cases}$$

where $N_B$ is the number of records in the smaller dataset. Note here that $N_j$ differs across the records in $B$. If you are blocking on first and last name and come across a rare name (Brian Kundinger), you'll have a small $N_{\text{BrianK}}$. However, for a more common name like Tom Smith, you'll have a large $N_{\text{TomS}}$.

Thus, records that have been more fine distinguished through blocking require less information in the other features in order to be deemed a match!

# U-Estimation

We have found that blocking led to poor estimation of the $u$ parameters in the model. Intutively, this is because blocking removes a large number of non-matching record pairs from the data, so there are less pairs available to estimate these parameters.

A little more rigorously, linking $n_A$ records to $n_B$ records results in $n_A n_B$ total record pairs. Under the assumption of one-to-one matching, at most $n_B$ of these pairs are matches, so at least $n_A (n_B - 1)$ are non matches. Note that this number is **huge**.

Under blocking however, we are not guaranteed any nonmatching pairs at all! If we have unique values for blocking criteria, then we can end up in a situation with 0 nonmatching record pairs. In this setting, the model cannot separate pairs into matches and nonmatches (because there no nonmatches!)

# Data Driven Priors for U-Parameter

Instead of needing $n_A (n_B - 1)$ nonmatching pairs that would result from all-to-all comparions, or allowing for the possibility of 0 nonmatching pairs, we seek a middle ground. For each feature $f$, we take $R$ random records in $B$ and compute agreement levels with $S$ random records $i \notin N_j \subset A$. This is a form of *bootstrapping* that simulates the behavior of nonmatching pairs. We use these agreement level counts as a *data driven prior* on the $u$ parameters.

Intuitively, this procedure is meant to recover the mean of the $u$ parameters that would have resulted from all-to-all comparisons, but with higher variance (because there are less counts). However, we have seen that the $u$ parameters are generally *absurdly precise*, this is a negligible sacrifice.

Over the summer, we found that this step greatly improved model performance.

# Data Driven Prior on Rate of Matching

Note that we can write $N = N_{\text{in}} + N_{\text{out}}$, where $N_{\text{in}}$ is the number of records in $B$ that pass through blocking, and $N_{\text{out}}$ is the number of records that don't.

The Beta linkage method uses a parameter $\pi$ for the overall rate of matching across two datasets. If we use blocking, and remove the $N_{\text{out}}$ records in $B$ that have no plausible match in $A$, this parameter becomes biased upwards.

However, we can maintain the relevant information from these nonmatching records through an informative prior on the rate of matching! If $\pi \sim \text{Beta} (\alpha_{\pi}, \beta_{\pi})$, simply set $\beta_{\pi} = N_{\text{out}}$. Then, calculations for the rate of matching parameter are essentially equivalent to if they done without blocking!

# Elements of Paper

- Plot showing how increasing "prior counts" of nonmatching pairs through $R$ and $S$ leads to parameters that converge to those found under all-to-all comparisons.

- Simulation study showing
  - Results without blocking
  - Results with blocking, but no informative priors (should be poor)
  - Results with blocking and informative priors (should nearly recover results without blocking)

- Memory comparisons (less pairs to compute) and time comparisons

- Data
  - Same data Jody is using for other paper
  - Data from [North Carolina State Board of Elections](https://dl.ncsbe.gov/?prefix=)

# Bonus piece

In fastLink, one of the ways the increase scalability is by learning model parameters on a subset of record pairs, and then applying those parameters to the full set of record pairs in order to determine matching status (see paragraph before S4 in the appendix [here](https://imai.fas.harvard.edu/research/files/linkage-app.pdf#page=4)).

One issue I see with doing this under fastLink that it would bias the estimate of the rate of matching parameter. In order for a matching record pair to make it into the algorithm is for the relevant record to be sampled from *both* datasets. Thus I suspect that this rate of matching parameter would be biased downward.

With blocking however, we (mostly) ensure that a record and its match are both in the analysis we conduct. We can then run the model on a subsample of the blocks, and apply those parameters to the rest of the blocks after. This was pretty successful when we tested it out over the summer, and provides *massive* gains in scalability.

# Notes about other methods

This "train and test" method wouldn't be possible under the Sadinle (2017) model because its not possible to simply apply parameters in a way that respects one-to-one matching.

Also, I don't think it would work under `d-blink` either, because there are no parameters that you can simply apply to records to get posterior match probabilities.

So I think there would be really interesting comparisons between `fastlink`, `BRL`, and `d-blink`!

### 2022-07-14 ba_resubmit/response_to_reviewers.tex
response to reviewers

\documentclass[letterpaper, parskip]{scrartcl}

\usepackage{amsmath,amssymb,bm}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=1.4in]{geometry}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{tcolorbox}            % provides shaded box for reviewer comments
\usepackage{enumitem}             % use more compact lists
%\usepackage[parfill]{parskip}     % use skips between paragraphs and no indents
\usepackage{bibentry}             % provides \nobibliography command
\usepackage{xspace}

% Make itemize and enumerate environments more compact
\setlist[itemize]{noitemsep, topsep=0pt, parsep=0pt}
\setlist[enumerate]{noitemsep, topsep=0pt, parsep=0pt}

% -------- Define point raised command --------- %

\definecolor{darkpurple}{HTML}{500050}
\tcbuselibrary{breakable}       % allow tcolorbox to break across pages
\tcbuselibrary{skins}
\tcbset{%
	enhanced,%
	frame hidden,%
	sharp corners,%
	breakable,%
	parbox=false,%
	colback=lightgray!30,%
	enlarge left by=6pt,%
	width=\linewidth-6pt,%
	colback=gray!0,
	left=4pt,%
	right=0pt,%
	top=0pt,%
	bottom=0pt,%
	borderline west={2pt}{0pt}{lightgray},%
	%coltext=darkpurple,%
	before skip=14pt plus 2pt,%
	after skip=14pt plus 2pt}

\newcommand{\pointRaised}[1]{%
	\begin{tcolorbox}
		\itshape #1
	\end{tcolorbox}
}

\newcounter{responsectr}[section]     % counter for response resets at each section
\newcommand{\reply}[2]{%
	\refstepcounter{responsectr}%
	\textbf{#1.\theresponsectr:} #2
}

\newcommand{\todo}{\textcolor{red}{[TODO]}\xspace}

\newcommand{\ncomment}[1]{{\color{teal} NM: #1}}

\begin{document}

	% ****************** TITLE ****************************************

	\title{Response to Reviewer Comments}

	\maketitle
	We thank the Editor and reviewers for their encouraging assessment and constructive 
	feedback.

	\section*{Editor}

	Dear Dr Kundinger,

	Thank you for submitting your paper "Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)" for possible publication in Bayesian Analysis.
	It has now been carefully reviewed and my decision is: Major revision required.

	These are the comments I received from the Editor who has handled your paper:

	The paper has been reviewed by a referee and an associate editor and I have read the manuscript independently, before looking at the reports.  When I confront my notes and overall impression with the reviewers’ comments, I find myself in substantial agreement with the reviewers.

	As indicated by both reviewers, the work expands on previous work by Sadinle, introducing computational advances that are achieved by relaxing some requirements in the original formulation of the problem.  The close connections with the previous work and the extent of the novel contributions are not adequately explained.  This makes it hard for the reader to understand what is new and why the new contributions are valuable.  A careful editing of the presentation is needed to address this shortcoming.

	Regarding the presentation, the AE points out a number of problems with notation, typos, and various inconsistencies.  Independently, I found many similar problems.  The intersection between what I found and what the AE found is non-empty and so is the symmetric difference.  Also, the AE found some issues that I did not find and vice versa.  So, I am almost certain that an additional review would uncover more problems.  Now, the results presented in the paper seem plausible as do the broad strokes of the derivations, but I must admit that I was not able to follow all the details. This, in large part, was because of the issues that I just mentioned, and I believe that most readers would find it difficult to follow the developments, as these issues are exceedingly distracting.

	There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.  For example, I_obs() is never defined, n_12(Z) first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.  Throughout the manuscript, n_1 and n_2 are used interchangeably with n_A and n_B, sometimes in the same section, as it happens, for example, in Section 3.1.  The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein (\theta_11, if Z_j, \hat{Z_j}, …).  In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)  Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.  For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.

	Confusion can also arise from lack of clarity in the exposition.  Take comment 4 from the AE, for example.  Whose marriage certificates and whose birth records are involved?  What records are in X_1 and what records are in X_2?  The reader should not be expected to go back to Newcombe’s paper to clarify the issue.

	This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.  Many more specific points are raised in the AE’s report and here are a couple of other minor issues I noticed:

	p. 12, second paragraph:  insert “is” after Section 3.1

	p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.”

	As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies.

	The AE finds the simulation studies to be incomplete and gives detailed suggestions on how those should be improved.  The AE also gives important suggestions about other aspects of the manuscript that must be carefully considered and addressed.  All of the AE’s comments are right on the mark.

	To summarize, the paper contains an interesting algorithmic contribution that can speed up calculations at the expense of relaxing some of the modeling conditions, without much adverse impact on the resulting inferences.  This aspect should be emphasized in a revised presentation.  The presentation should also make clearer the close connections to the relevant work by Sadinle, and all the issues that I mentioned above should be ironed out.

	The authors should prepare a careful and substantive revision that remains within the editorial limit of 25 pages and that answers the various comments satisfactorily.  The authors should upload the revised manuscript together with a document detailing how they addressed the reviewers’ comments.

%	\reply{E}{%
%		Thank you for the opportunity to submit a revision. 
%		We have provided detailed responses to the reviewers' comments below, 
%		which prompted significant changes in our paper.
%	}

	\clearpage
	\newpage

	\section*{Reviewer 2}
	\setcounter{responsectr}{0}
	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast
	computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases
	without resorting to ad hoc blocking. It is built upon the Felegi-Sunter model of record linkages, where
	linkages are made within strata of agreement between records across the two databases, and relies
	heavily on advances made by Sadinle (2017).
	Overall, the paper presents a practical implementation of a modified Beta Linkage approach to
	matching. It demonstrates that even though the new method suffers from the limitations of not
	seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of
	features on which to match. Though the novelty of the presented methodology is primarily in the
	computational approach, I believe it to be a real advance for practical implementation of record linkage
	in modern applications.
	I have no major concerns about the content or accuracy of the presentation. However, substantial
	improvements are warranted to confirm my impression. In addition to the recommendation made by
	Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the
	explanation of the method should be written more clearly, the simulation comparisons should be
	repeated/expanded, and various features of the simplification should be fleshed out a bit more. Finally,
	more practical advice would be welcome. All of these points are more specifically described in the
	itemized list below

		\pointRaised{%
Overall, the notation is, somewhat by necessity, quite extensive. However, I’m not convinced
that it’s entirely consistent. A glossary in the appendix may help the reader navigate the paper
more easily.
	}

		\pointRaised{%
Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as implied
by equations (1) and (2).
}
		\pointRaised{%
Section 2.2, beta distribution: the n_{12}(.) notation needs to be defined.
}

		\pointRaised{%
Section 3, 4th full paragraph: My understanding of the example of matching birth records and
marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as
presented, as I would think that one certificate (element in X1) would match multiple birth
records (elements in X2) which violates the size ordering. Can the algorithm be easily adjusted
for this reverse sizing?
}
		\pointRaised{%
5. Section 3, full model: I_obs(.) seems to be taken from a different paper dealing with missing
fields, which is not at all discussed in this manuscript. This consideration of missing fields and
the accompanying discussion of missing data mechanism should be included if such data is to be
considered for this manuscript. This comment also applies to the Gamma^{obs} notation in the
following subsection.
}
		\pointRaised{%
Section 3, full model: The Phi notation needs to be defined.
}
		\pointRaised{%
Section 3, full model: The square brackets seem to be misplaced, as the exponent includes
indices of the sum contained in the brackets.
}
		\pointRaised{%
Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in
that they seem to mix the posterior behavior (e.g., parameters are updated though standard
multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…).
This subsection might be easier to understand if its purpose was more bluntly described in an
opening sentence or paragraph and streamlined to focus only on the full conditional
distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in
intermediate conclusions. In particular, the derivation of the probability of Gamma_{.j} seems
to be missing a factor of u_{fl}^I(gamma_{ij}^f=l). This is not important to the eventual full
conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
}
		\pointRaised{%
Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the
conditional distribution of Z was included using the summation-style notation rather than the
binomial/multinomial-style multiplicative notation that naturally motivates the full conditional
presented on the first line of page 8.
}
		\pointRaised{%
Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
}
		\pointRaised{%
Section 3.1, pmf for Gamma_{.j} and full conditional for Z: The notation in the final equation
may be more clear if you replace i with z_j, as in w_{z_j j}. I believe that “n_1+1” should be “n_1
+ j” in several places.
}
		\pointRaised{%
Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing
these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It
would be helpful to provide a brief justification of integrating out pi.
}
		\pointRaised{%
Section 4.2, SEI procedure. I’m not sure I follow the description of the SEI procedure. My
understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to
(should) be identified in the posterior distribution. Thus, you take a small subsample of the
records of size S < H_j_p, and store these in R^{SEI} rather than the complete index of pairs in
each R. Can you quantify the computational savings due to this method? Do you have advice for
how small S can be? In your simulations, how does the use of this method or choice of S affect
the results?
}
		\pointRaised{%
Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
}
		\pointRaised{%
Section 5.1 simulations – there is a detailed comparison for one set of simulation settings,
including a binary definition of a “match”. Do these results differ as these settings change?
}
		\pointRaised{%
Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results
hold over multiple simulated datasets, Or was this just a chance result for each setting?
}
		\pointRaised{%
Section 5.3 simulation – RR results are not presented.
}
		\pointRaised{%
	Section 6.1 last sentences. I believe the last sentence should be removed, as it is redundant.
}
		\pointRaised{%
Section 6.2, second sentence. I believe the word “survey” is missing, as in “the initial SURVEY
began…”
}
		\pointRaised{%
Section 6.2. In the NLTCS, are participants added in later cohorts, or may we assume that every
member of the smaller dataset should also be included in the earlier one?
}
		\pointRaised{%
	Figures 5 and 7 are not referenced in the text.
}

	\clearpage
	\newpage

	\section*{Reviewer 3}
	\setcounter{responsectr}{0}

	The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation
	results are presented and the variant is applied on a new, significantly larger data set.
	A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in
	section 4. It is simple but consequential. The simulations and applications show how the variant
	accelerates the computation relative to the original method and reveals the trade-offs.
	The value of the paper resides in that computational complexity continues to be a difficult obstacle in
	the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s
	approach to an extent not possible before.
	It is important to present the research for what it is: An extension of the method of Sadinle motivated
	by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting
	point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should
	be made clearer, beginning in the abstract and reiterated in the conclusion.

\end{document}

### 2022-07-20 ba_resubmit/response_to_reviewers.tex
Corrected some of the easy typos

@@
 }
+		\reply{R1}{%
+			We thank the reviewers for identifying this oversight, and have defined the notation. 
+		}
 
@@
 }
+		\reply{R1}{%
+	We thank the reviewers for identifying this oversight, and have defined the notation. 
+}
+
 		\pointRaised{%
@@
 indices of the sum contained in the brackets.
+}
+
+		\reply{R1}{%
+	We thank the reviewers for identifying this oversight, and have corrected the expression. 
 }
@@
 Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
+}
+		\reply{R1}{%
+	We thank the reviewers for identifying this oversight, and have corrected the expression. 
 }

### 2022-09-01 ba_resubmit/response_markdown.Rmd
Markdown file for addressing reviewer comments

--- 2022-07-20 ba_resubmit/response_to_reviewers.tex
@@
-\documentclass[letterpaper, parskip]{scrartcl}
+---
+title: "Response Markdown"
+output: pdf_document
+date: '2022-08-31'
+---
 
-\usepackage{amsmath,amssymb,bm}
-\usepackage{url}
-\usepackage[hidelinks]{hyperref}
-\usepackage[margin=1.4in]{geometry}
-\usepackage{natbib}
-\usepackage{xcolor}
-\usepackage{tcolorbox}            % provides shaded box for reviewer comments
-\usepackage{enumitem}             % use more compact lists
-%\usepackage[parfill]{parskip}     % use skips between paragraphs and no indents
-\usepackage{bibentry}             % provides \nobibliography command
-\usepackage{xspace}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
 
-% Make itemize and enumerate environments more compact
-\setlist[itemize]{noitemsep, topsep=0pt, parsep=0pt}
-\setlist[enumerate]{noitemsep, topsep=0pt, parsep=0pt}
-
-% -------- Define point raised command --------- %
-
-\definecolor{darkpurple}{HTML}{500050}
-\tcbuselibrary{breakable}       % allow tcolorbox to break across pages
-\tcbuselibrary{skins}
-\tcbset{%
-	enhanced,%
-	frame hidden,%
-	sharp corners,%
-	breakable,%
-	parbox=false,%
-	colback=lightgray!30,%
-	enlarge left by=6pt,%
-	width=\linewidth-6pt,%
-	colback=gray!0,
-	left=4pt,%
-	right=0pt,%
-	top=0pt,%
-	bottom=0pt,%
-	borderline west={2pt}{0pt}{lightgray},%
-	%coltext=darkpurple,%
-	before skip=14pt plus 2pt,%
-	after skip=14pt plus 2pt}
-
-\newcommand{\pointRaised}[1]{%
-	\begin{tcolorbox}
-		\itshape #1
-	\end{tcolorbox}
-}
-
-\newcounter{responsectr}[section]     % counter for response resets at each section
-\newcommand{\reply}[2]{%
-	\refstepcounter{responsectr}%
-	\textbf{#1.\theresponsectr:} #2
-}
-
-\newcommand{\todo}{\textcolor{red}{[TODO]}\xspace}
-
-\newcommand{\ncomment}[1]{{\color{teal} NM: #1}}
-
-\begin{document}
-
-	% ****************** TITLE ****************************************
-
-	\title{Response to Reviewer Comments}
-
-	\maketitle
-	We thank the Editor and reviewers for their encouraging assessment and constructive 
-	feedback.
-
-	\section*{Editor}
+# Editor
 
@@
 There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.
-For example, I_obs() is never defined, n_12(Z) first appears on p.
-5, but is not defined (in passing) until the bottom of p.
-7, and it seems to become D on p.
-8.
-Throughout the manuscript, n_1 and n_2 are used interchangeably with n_A and n_B, sometimes in the same section, as it happens, for example, in Section 3.1.
-The loss function on p.
-9 is out of Sadinle’s paper and it inherits the typo therein (\theta_11, if Z_j, \hat{Z_j}, …).
-In the statement of Lemma 1 on p.13, B comes out of nowhere.
-(Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)  Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.
-For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
 
-	Confusion can also arise from lack of clarity in the exposition.
+**For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.**
+\begin{quote}
+\textcolor{red}{How and when should I define $I_{obs}$? I probably need to lay out certain assumptions about missingness?}
+
+Thank you for identifying inconsistencies with the notation $n_{12}(Z)$. We have made these corrections.
+\end{quote}
+
+**Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.**
+\begin{quote}
+Thank you for addressing this oversight. We have made these corrections throughout the text.
+\end{quote}
+
+**The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_{11},$ $if Z_j, \hat{Z_j}, …)$.**
+\begin{quote}
+Thank you for addressing this oversight. We have corrected the loss function.
+\end{quote}
+
+**In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)**
+\begin{quote}
+We have updated the Lemma to include the definition of $B$.
+\end{quote}
+
+**Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.
+For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.**
+\begin{quote}
+\textcolor{red}{Is this true?}
+\end{quote}
+
+**Confusion can also arise from lack of clarity in the exposition.
 Take comment 4 from the AE, for example.
 Whose marriage certificates and whose birth records are involved?
-What records are in X_1 and what records are in X_2?
-The reader should not be expected to go back to Newcombe’s paper to clarify the issue.
+What records are in $X_1$ and what records are in $X_2$?
+The reader should not be expected to go back to Newcombe’s paper to clarify the issue.**
+\begin{quote}
+\textcolor{red}{Please see comment in Reviewer 2 responses.}
+\end{quote}
 
@@
 
-	p. 12, second paragraph:  insert “is” after Section 3.1
+**p. 12, second paragraph:  insert “is” after Section 3.1**
+\begin{quote}
+Thank you for identifying this. We have made the correction.
+\end{quote}
 
-	p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.”
+**p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.”**
+\begin{quote}
+\textcolor{red}{I can't find this typo}
+\end{quote}
 
@@
 
-%	\reply{E}{%
-%		Thank you for the opportunity to submit a revision. 
-%		We have provided detailed responses to the reviewers' comments below, 
-%		which prompted significant changes in our paper.
-%	}
+# Reviewer 2
+
+This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases without resorting to ad hoc blocking.
+It is built upon the Felegi-Sunter model of record linkages, where linkages are made within strata of agreement between records across the two databases, and relies heavily on advances made by Sadinle (2017).
+
+Overall, the paper presents a practical implementation of a modified Beta Linkage approach to matching.
+It demonstrates that even though the new method suffers from the limitations of not seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of features on which to match.
+Though the novelty of the presented methodology is primarily in the computational approach, I believe it to be a real advance for practical implementation of record linkage in modern applications.
+
+I have no major concerns about the content or accuracy of the presentation.
+However, substantial improvements are warranted to confirm my impression.
+In addition to the recommendation made by Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the explanation of the method should be written more clearly, the simulation comparisons should be repeated/expanded, and various features of the simplification should be fleshed out a bit more.
+Finally, more practical advice would be welcome.
+All of these points are more specifically described in the itemized list below
+
+**Overall, the notation is, somewhat by necessity, quite extensive.
+However, I’m not convinced that it’s entirely consistent.
+A glossary in the appendix may help the reader navigate the paper more easily.**
+
+\begin{quote}
+\textcolor{red}{I don't think we require an appendix. It's not much more notation than Sadinle or Fellegi Sunter.}
+\end{quote}
+
+**Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as impliedby equations (1) and (2).**
+\begin{quote}
+\textcolor{red}{I'm not sure I agree with this point}
+\end{quote}
+
+**Section 2.2, beta distribution: the $n_{12}(.)$ notation needs to be defined.**
+
+\begin{quote}
+We thank the reviewers for identifying this oversight, and have defined the notation.
+\end{quote}
+
+**Section 3, 4th full paragraph: My understanding of the example of matching birth records and marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as presented, as I would think that one certificate (element in X1) would match multiple birth records (elements in X2) which violates the size ordering.
+Can the algorithm be easily adjusted for this reverse sizing?**
+
+\begin{quote}
+\textcolor{red}{PLEASE READ: Part of me thinks we should just remove this element of the model from the paper.
+This topic also begs the question of whether linkages stay the same depending on which dataset you label X1 and X2.
+I have no good way of answering that.}
+
+\textcolor{red}{Essentially though,the answer is yes, the algorithm can easily be adapted for the reverse sizing. The issue is mostly notational.}
+
+\textcolor{red}{It seems as though Sadinle chose to label the datasets $X_1$ and $X_2$ such that $n_1 > n_2$ for mostly notational reasons.
+He uses the notation $Z_j = n_1 + j$ when record $j \in X_1$ has no match in $X_2$, which works because when $n_1 > n_2$, $n_1 + j$ never refers to any record in $X_2$.
+Jody uses the simplified notation $Z_j = n_1 + 1$ to refer to unmatched records, and this works for the same reason.
+}
+
+\textcolor{red}{However, when we relax the one-to-one constraint, or even encourage a modeller to allow multiple linkages in one dataset, we shouldn't denote the datasets according to simple notational convenience.
+This might require clever edits of the "Review of Prior Work" section, and other areas of the paper.}
+
+\textcolor{red}{Notationally, we can use  $Z_j = 0$ or $Z_j = -j$ as a way to unambiguously denote a record as nonmatching.}
+
+\textcolor{red}{But again, it may be easier to just strike the paragraph about the advantages of many-to-one matchings.}
+\end{quote}
+
+**Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing fields, which is not at all discussed in this manuscript.
+This consideration of missing fields and the accompanying discussion of missing data mechanism should be included if such data is to be considered for this manuscript.
+This comment also applies to the $Gamma^{obs}$ notation in the following subsection.**
+
+\begin{quote}
+\textcolor{red}{I can add these definitions, no problem. But is it worth including mention of missing data?}
+\end{quote}
+
+**Section 3, full model: The Phi notation needs to be defined.**
+
+\begin{quote}
+We thank the reviewers for identifying this oversight, and have defined the notation.
+\end{quote}
+
+**Section 3, full model: The square brackets seem to be misplaced, as the exponent includes indices of the sum contained in the brackets.**
+
+\begin{quote}
+We thank the reviewers for identifying this oversight, and have corrected the expression. 
+\end{quote}
+
+**Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in that they seem to mix the posterior behavior (e.g., parameters are updated though standard multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…).
+This subsection might be easier to understand if its purpose was more bluntly described in an opening sentence or paragraph and streamlined to focus only on the full conditional distributions for a Gibbs Sampler.
+In addition, the mixed motivation may have led to errors in intermediate conclusions.
+In particular, the derivation of the probability of $Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$.
+This is not important to the eventual full conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.**
+
+\begin{quote}
+\textcolor{red}{I'm not entirely sure how to address the first part of the response. Should I just eliminate a lot of the exposition? Also, I see no issue with $p(\Gamma_{.j})$}
+\end{quote}
+
+**Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the conditional distribution of Z was included using the summation-style notation rather than the binomial/multinomial-style multiplicative notation that naturally motivates the full conditional presented on the first line of page 8.**
+
+\begin{quote}
+The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
+The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$.
+We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$.
+We have updated our explanation to make this more clear.
+\end{quote}
+
+**Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).**
+
+\begin{quote}
+We thank the reviewers for identifying this oversight, and have corrected the expression.
+\end{quote}
+
+**Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation may be more clear if you replace i with $z_j$, as in $w_{z_j,j}$.
+I believe that “$n_1+1$” should be “$n_1+j$” in several places.**
+
+\begin{quote}
+\textcolor{red}{Ask Beka if $i$ or $z_j$ is more clear.}
+
+We agree that using $z_j$ instead of $i$ in that context is more clear. We have also changed $n_1 + 1$ to $n_1 + j$ in several instances. Thank you for bringing this to our attention. 
+\end{quote}
+
+**Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing these equations as full conditional distributions vs.
+steps in the Gibbs Sampler algorithm.
+It would be helpful to provide a brief justification of integrating out pi.**
+
+\begin{quote}
+\textcolor{red}{Should I just remove this step?
+I'm not sure how to justify this step, but also this step isn't actually necessary.
+We could just sample $\pi$ and use it, as in a standard Gibbs sampler.
+The only reason to integrate it out is to see the similarity to Sadinle's full conditional, but its not necessary at all.}
+\end{quote}
+
+**Section 4.2, SEI procedure.
+I’m not sure I follow the description of the SEI procedure.
+My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution.
+Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R.
+Can you quantify the computational savings due to this method?
+Do you have advice for how small S can be?
+In your simulations, how does the use of this method or choice of S affect the results?**
+
+\begin{quote}
+\textcolor{red}{All computational savings occur from conducting the Gibbs sampler over the $P$ patterns instead of the $n_2$ records.
+SEI is for memory savings.
+I think I need help articulating this clearly.}
+\end{quote}
+
+**Section 4.4, Assumptions and definitions should be included in the statement of this lemma.**
+
+\begin{quote}
+\textcolor{red}{Do I really need this to be a Lemma? It seems overly clunky to me, but if its convention to provide computational complexity as a lemma, thats OK with me.}
+\end{quote}
+
+**Section 5.1 simulations – there is a detailed comparison for one set of simulation settings, including a binary definition of a “match”. Do these results differ as these settings change?**
+
+\begin{quote}\textcolor{red}{No, the time savings would be the same.
+In all cases, the fabl method replaces dependency on $n_2$ for dependency on $P$, the number of patterns.
+To me, this is clear, I'm not sure how to clarify this in the paper}
+\end{quote}
+
+**Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results hold over multiple simulated datasets, Or was this just a chance result for each setting?**
+
+\begin{quote}
+In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error. We have added this description in the body of the paper as well for clarity.
+\end{quote}
+
+**Section 5.3 simulation – RR results are not presented.**
+
+\begin{quote}
+\textcolor{red}{I will redo simulations in order to get this}
+\end{quote}
+
+**Section 6.1 last sentences. I believe the last sentence should be removed, as it is redundant.**
+
+\begin{quote}
+We have deleted this last sentence. Thank you for addressing this oversight.
+\end{quote}
+
+**Section 6.2, second sentence. I believe the word “survey” is missing, as in “the initial SURVEY began…”**
+
+\begin{quote}
+Thank you for addressing this oversight.
+\end{quote}
+
+**Section 6.2. In the NLTCS, are participants added in later cohorts, or may we assume that every member of the smaller dataset should also be included in the earlier one?**
+
+\begin{quote}
+\textcolor{red}{Beka, can you help answer this?}
+\end{quote}
+
+**Figures 5 and 7 are not referenced in the text.**
+
+\begin{quote}
+Thank you for addressing this oversight. We have included pointers to these figures in the main text.
+\end{quote}
 
@@
 
-	\section*{Reviewer 2}
-	\setcounter{responsectr}{0}
-	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast
-	computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases
-	without resorting to ad hoc blocking. It is built upon the Felegi-Sunter model of record linkages, where
-	linkages are made within strata of agreement between records across the two databases, and relies
-	heavily on advances made by Sadinle (2017).
-	Overall, the paper presents a practical implementation of a modified Beta Linkage approach to
-	matching. It demonstrates that even though the new method suffers from the limitations of not
-	seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of
-	features on which to match. Though the novelty of the presented methodology is primarily in the
-	computational approach, I believe it to be a real advance for practical implementation of record linkage
-	in modern applications.
-	I have no major concerns about the content or accuracy of the presentation. However, substantial
-	improvements are warranted to confirm my impression. In addition to the recommendation made by
-	Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the
-	explanation of the method should be written more clearly, the simulation comparisons should be
-	repeated/expanded, and various features of the simplification should be fleshed out a bit more. Finally,
-	more practical advice would be welcome. All of these points are more specifically described in the
-	itemized list below
+# Reviewer 3
 
-		\pointRaised{%
-Overall, the notation is, somewhat by necessity, quite extensive. However, I’m not convinced
-that it’s entirely consistent. A glossary in the appendix may help the reader navigate the paper
-more easily.
-	}
+The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation results are presented and the variant is applied on a new, significantly larger data set.
 
-		\pointRaised{%
-Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as implied
-by equations (1) and (2).
-}
-		\pointRaised{%
-Section 2.2, beta distribution: the n_{12}(.) notation needs to be defined.
-}
-		\reply{R1}{%
-			We thank the reviewers for identifying this oversight, and have defined the notation. 
-		}
+A significant part of the paper is a reproduction of the paper of Sadinle.
+The variant is introduced in section 4.
+It is simple but consequential.
+The simulations and applications show how the variant accelerates the computation relative to the original method and reveals the trade-offs.
+The value of the paper resides in that computational complexity continues to be a difficult obstacle in the application of Bayesian statistics.
+The proposals of the paper enable the application of Sadinle’s approach to an extent not possible before.
 
-		\pointRaised{%
-Section 3, 4th full paragraph: My understanding of the example of matching birth records and
-marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as
-presented, as I would think that one certificate (element in X1) would match multiple birth
-records (elements in X2) which violates the size ordering. Can the algorithm be easily adjusted
-for this reverse sizing?
-}
-		\pointRaised{%
-5. Section 3, full model: I_obs(.) seems to be taken from a different paper dealing with missing
-fields, which is not at all discussed in this manuscript. This consideration of missing fields and
-the accompanying discussion of missing data mechanism should be included if such data is to be
-considered for this manuscript. This comment also applies to the Gamma^{obs} notation in the
-following subsection.
-}
-		\pointRaised{%
-Section 3, full model: The Phi notation needs to be defined.
-}
-		\reply{R1}{%
-	We thank the reviewers for identifying this oversight, and have defined the notation. 
-}
-
-		\pointRaised{%
-Section 3, full model: The square brackets seem to be misplaced, as the exponent includes
-indices of the sum contained in the brackets.
-}
-
-		\reply{R1}{%
-	We thank the reviewers for identifying this oversight, and have corrected the expression. 
-}
-		\pointRaised{%
-Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in
-that they seem to mix the posterior behavior (e.g., parameters are updated though standard
-multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…).
-This subsection might be easier to understand if its purpose was more bluntly described in an
-opening sentence or paragraph and streamlined to focus only on the full conditional
-distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in
-intermediate conclusions. In particular, the derivation of the probability of Gamma_{.j} seems
-to be missing a factor of u_{fl}^I(gamma_{ij}^f=l). This is not important to the eventual full
-conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
-}
-		\pointRaised{%
-Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the
-conditional distribution of Z was included using the summation-style notation rather than the
-binomial/multinomial-style multiplicative notation that naturally motivates the full conditional
-presented on the first line of page 8.
-}
-		\pointRaised{%
-Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
-}
-		\reply{R1}{%
-	We thank the reviewers for identifying this oversight, and have corrected the expression. 
-}
-		\pointRaised{%
-Section 3.1, pmf for Gamma_{.j} and full conditional for Z: The notation in the final equation
-may be more clear if you replace i with z_j, as in w_{z_j j}. I believe that “n_1+1” should be “n_1
-+ j” in several places.
-}
-		\pointRaised{%
-Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing
-these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It
-would be helpful to provide a brief justification of integrating out pi.
-}
-		\pointRaised{%
-Section 4.2, SEI procedure. I’m not sure I follow the description of the SEI procedure. My
-understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to
-(should) be identified in the posterior distribution. Thus, you take a small subsample of the
-records of size S < H_j_p, and store these in R^{SEI} rather than the complete index of pairs in
-each R. Can you quantify the computational savings due to this method? Do you have advice for
-how small S can be? In your simulations, how does the use of this method or choice of S affect
-the results?
-}
-		\pointRaised{%
-Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
-}
-		\pointRaised{%
-Section 5.1 simulations – there is a detailed comparison for one set of simulation settings,
-including a binary definition of a “match”. Do these results differ as these settings change?
-}
-		\pointRaised{%
-Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results
-hold over multiple simulated datasets, Or was this just a chance result for each setting?
-}
-		\pointRaised{%
-Section 5.3 simulation – RR results are not presented.
-}
-		\pointRaised{%
-	Section 6.1 last sentences. I believe the last sentence should be removed, as it is redundant.
-}
-		\pointRaised{%
-Section 6.2, second sentence. I believe the word “survey” is missing, as in “the initial SURVEY
-began…”
-}
-		\pointRaised{%
-Section 6.2. In the NLTCS, are participants added in later cohorts, or may we assume that every
-member of the smaller dataset should also be included in the earlier one?
-}
-		\pointRaised{%
-	Figures 5 and 7 are not referenced in the text.
-}
-
-	\clearpage
-	\newpage
-
-	\section*{Reviewer 3}
-	\setcounter{responsectr}{0}
-
-	The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation
-	results are presented and the variant is applied on a new, significantly larger data set.
-	A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in
-	section 4. It is simple but consequential. The simulations and applications show how the variant
-	accelerates the computation relative to the original method and reveals the trade-offs.
-	The value of the paper resides in that computational complexity continues to be a difficult obstacle in
-	the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s
-	approach to an extent not possible before.
-	It is important to present the research for what it is: An extension of the method of Sadinle motivated
-	by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting
-	point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should
-	be made clearer, beginning in the abstract and reiterated in the conclusion.
-
-\end{document}
+It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations.
+Here Sadinle supersedes Fellegi-Sunter in the sense while the starting point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work.
+This should be made clearer, beginning in the abstract and reiterated in the conclusion.

### 2022-09-01 ba_resubmit/response_to_reviewers.tex
Markdown file for addressing reviewer comments

@@
 	There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.
-For example, I_obs() is never defined, n_12(Z) first appears on p.
+For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p.
 5, but is not defined (in passing) until the bottom of p.
@@
 8.
-Throughout the manuscript, n_1 and n_2 are used interchangeably with n_A and n_B, sometimes in the same section, as it happens, for example, in Section 3.1.
+Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.
 The loss function on p.
-9 is out of Sadinle’s paper and it inherits the typo therein (\theta_11, if Z_j, \hat{Z_j}, …).
+9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_11, if Z_j, \hat{Z_j}, …)$.
 In the statement of Lemma 1 on p.13, B comes out of nowhere.
@@
 Whose marriage certificates and whose birth records are involved?
-What records are in X_1 and what records are in X_2?
+What records are in $X_1$ and what records are in $X_2$?
 The reader should not be expected to go back to Newcombe’s paper to clarify the issue.
@@
 		\pointRaised{%
-Overall, the notation is, somewhat by necessity, quite extensive. However, I’m not convinced
-that it’s entirely consistent. A glossary in the appendix may help the reader navigate the paper
-more easily.
+Overall, the notation is, somewhat by necessity, quite extensive.
+However, I’m not convinced that it’s entirely consistent.
+A glossary in the appendix may help the reader navigate the paper more easily.
 	}
@@
 }
+\textcolor{red}{I'm not sure I agree with this point}
+
 		\pointRaised{%
-Section 2.2, beta distribution: the n_{12}(.) notation needs to be defined.
+Section 2.2, beta distribution: the $n_{12}(.)$ notation needs to be defined.
 }
@@
 }
+
+\color{red}{I need some help addressing this in the paper. The answer is yes, the algorithm can easily be adapted for the reverse sizing. The issue is mostly notational.
+
+It seems as though Sadinle chose to label the datasets $X_1$ and $X_2$ such that $n_1 > n_2$ for mostly notational reasons.
+He uses the notation $Z_j = n_1 + j$ when record $j \in X_1$ has no match in $X_2$, which works because when $n_1 > n_2$, $n_1 + j$ never refers to any record in $X_2$.
+Jody uses the simplified notation $Z_j = n_1 + 1$ to refer to unmatched records, and this works for the same reason.
+
+However, when we relax the one-to-one constraint, or even encourage a modeller to allow multiple linkages in one dataset, we shouldn't denote the datasets according to simple notational convenience.
+This might require clever edits of the "Review of Prior Work" section, and other areas of the paper.
+
+Notationally, we can use  $Z_j = 0$ or $Z_j = -j$ as a way to unambiguously denote a record as nonmatching.
+
+Additionally, we could just strike the paragraph about the advantages of many-to-one matchings, and keep this strictly a paper about bipartite matching}
+
 		\pointRaised{%
-5. Section 3, full model: I_obs(.) seems to be taken from a different paper dealing with missing
+5. Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing
 fields, which is not at all discussed in this manuscript. This consideration of missing fields and
 the accompanying discussion of missing data mechanism should be included if such data is to be
-considered for this manuscript. This comment also applies to the Gamma^{obs} notation in the
+considered for this manuscript. This comment also applies to the $Gamma^{obs}$ notation in the
 following subsection.
 }
+\textcolor{red}{I can add these definitions, no problem. But is it worth including mention of missing data?}
 		\pointRaised{%
@@
 distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in
-intermediate conclusions. In particular, the derivation of the probability of Gamma_{.j} seems
-to be missing a factor of u_{fl}^I(gamma_{ij}^f=l). This is not important to the eventual full
+intermediate conclusions. In particular, the derivation of the probability of $Gamma_{.j}$ seems
+to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$. This is not important to the eventual full
 conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
@@
 		\pointRaised{%
-Section 3.1, pmf for Gamma_{.j} and full conditional for Z: The notation in the final equation
-may be more clear if you replace i with z_j, as in w_{z_j j}. I believe that “n_1+1” should be “n_1
-+ j” in several places.
+Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation
+may be more clear if you replace i with $z_j$, as in $w_{z_j j}$. I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
 }
@@
 }
+\textcolor{red}{I'm not sure how to justify this step, but also this step isn't actually necessary.
+We could just sample $\pi$ and use it, as in a standard Gibbs sampler.
+Should I just remove this step?}
+
 		\pointRaised{%
-Section 4.2, SEI procedure. I’m not sure I follow the description of the SEI procedure. My
-understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to
-(should) be identified in the posterior distribution. Thus, you take a small subsample of the
-records of size S < H_j_p, and store these in R^{SEI} rather than the complete index of pairs in
-each R. Can you quantify the computational savings due to this method? Do you have advice for
-how small S can be? In your simulations, how does the use of this method or choice of S affect
-the results?
-}
+Section 4.2, SEI procedure.
+I’m not sure I follow the description of the SEI procedure.
+My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution.
+Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in
+each R.
+Can you quantify the computational savings due to this method?
+Do you have advice for how small S can be?
+In your simulations, how does the use of this method or choice of S affect the results?}
+
+\textcolor{red}{All computational savings occur from conducting the Gibbs sampler over the $P$ patterns instead of the $n_2$ records.
+SEI is for memory savings.
+I think I need help articulating this clearly.}
 		\pointRaised{%
@@
 }
+\textcolor{red}{Do I really need this to be a Lemma? It seems overly clunky to me, but if its convention to provide computational complexity as a lemma, thats OK with me.}
 		\pointRaised{%
@@
 }
+\textcolor{red}{Yes, results would be the same. To me, this is clear, I'm not sure how to clarify this in the paper}
 		\pointRaised{%
@@
 }
+		\reply{R1}{%
+	In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error. We have added this description in the body of the paper as well for clarity. 
+}
 		\pointRaised{%
@@
 }
+\textcolor{red}{I will redo simulations in order to get this}
+
 		\pointRaised{%
 	Section 6.1 last sentences. I believe the last sentence should be removed, as it is redundant.
+}
+		\reply{R1}{%
+			We have deleted this last sentence. Thank you for addressing this oversight. 
 }
@@
 }
+		\reply{R1}{%
+	Thank you for addressing this oversight. 
+}
 		\pointRaised{%
@@
 }
+\textcolor{red}{Beka, can you help answer this?}
+
 		\pointRaised{%
 	Figures 5 and 7 are not referenced in the text.
+}
+		\reply{R1}{%
+	Thank you for addressing this oversight. We have included pointers to these figures in the main text. 
 }

### 2022-09-01 ba_resubmit/kundinger_resubmit.tex
Markdown file for addressing reviewer comments

--- 2021-12-20 bayesian-analysis/kundinger-submit-final.tex
@@
 \begin{keyword}%[class=MSC]
-\kwd{bipartite record linkage}
-\kwd{Bayesian methods}
-\kwd{hashing techniques}
-\kwd{parallel/distributed computing}
-\kwd{Markov chain Monte Carlo}
+\kwd{entity resolution}
+\kwd{data fusion}
+\kwd{hashing}
+\kwd{record linkage}
+\kwd{Bayesian}
+\kwd{distributed computing}
 \end{keyword}
@@
 For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
-$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
+$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
 $$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
+where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12} = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
 This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
@@
 \begin{align*}
-	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
 	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
@@
 	\end{cases} \\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
 \end{align*}
+where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.
 
@@
 
-As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
+Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$ (and hyperparameters).
+Thus we use the abbreviated $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have
+
+%As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_1 + j\)) and failures (when \(z_j = n_1 + j\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
 We have
@@
 	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
 	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
 \end{align*}
-Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
+Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
 
@@
 	\begin{cases} 
-		w_{ij}  & z_j \leq n_1; \\
+		w_{z_j, j}  & z_j \leq n_1; \\
 		1 &  z_j  = n_1 + j \\
@@
 $$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
-	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
+	\frac{\pi}{n_1}w_{z_{j},j}   & z_j \leq n_1; \\
 	1-\pi &  z_j  = n_1 + 1. \\
@@
 	w_{ij}  & i \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + 1. \\
+	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + j. \\
 \end{cases}\]
@@
 adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
-\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
+\(\hat{Z}_j \in \{1, \ldots, n_1 + j, R\}\), with \(R\) representing the
 option to leave the matching undetermined by the model. Specifically, we have
@@
 	\theta_R,  & \text{if } \hat{Z_j} = R; \\
-	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
-	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
-	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
+	\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
+	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
+	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
 \end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
@@
 
-Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities.
+Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_1 \times n_2\) quantities.
 These are simple calculations, but become computationally burdensome when working on large linkage tasks.
@@
 Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations.
-Denote $H_p^m = \sum_{j=1}^{n_B} \bm{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$.
+Denote $H_p^m = \sum_{j=1}^{n_2} \bm{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$.
 The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
@@
 Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\).
-More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + 1\).
+More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\).
 Then,
@@
 	\begin{lemma}
-		The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
+		The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$, where $B$ is the number of processors available for parallelization.
 		\label{lemma:fabl}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$.
+			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_1$ and $n_2$.
 We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:speed1}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500.
+			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_1$, and $n_2$ fixed at 500.
 We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:speed2}
@@
 We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets.
+Under each of these settings, we have 100 pairs of simulated datasets in order to obtain uncertainty quantification on our performance metrics.
 We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
@@
 	We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
+See `\ref{fig:overlap-plot} for a visual comparison of the Bayes estimate and posterior credible intervals for the two methods.
 We note that Bayes estimates falling outside of posterior credible intervals has occurred previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
-
-	We note that reconciling Bayes estimates that fall outside of posterior credible intervals is a common challenge in record linkage and an avenue of future work.
 
@@
 	The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients.
-The initial began in 1982, with follow-up surveys taken approximately every five years.
-For this study, we seek to link records over the $n_A = 20485$ records from 1982 to the $n_B = 17466$ records from 1989.
+The initial survey began in 1982, with follow-up surveys taken approximately every five years.
+For this study, we seek to link records over the $n_1 = 20485$ records from 1982 to the $n_2 = 17466$ records from 1989.
 The NLTCS data have longitudinal links, so that in reality one does not need to do a record linkage.
@@
 	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-The Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+As shown in ~\ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
 Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
@@
 
-	\begin{figure}[t]
+	\begin{figure}[h]
 		\begin{center}
@@
 			\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
-			\label{fig:overlap-plot}
+			\label{fig:nltcs-overlap-plot}
 		\end{center}

### 2022-09-01 ba_resubmit/kundinger_resubmit.tex
Paper edits

@@
 $$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12} = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
+where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
 This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
@@
 	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
-	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
+	\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
 \end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
@@
 
-Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
+Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
 This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.

### 2022-09-01 ba_resubmit/response_to_reviewers.tex
working on fabl paper.

@@
 
-	Dear Dr Kundinger,
-
-	Thank you for submitting your paper "Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)" for possible publication in Bayesian Analysis.
-	It has now been carefully reviewed and my decision is: Major revision required.
-
-	These are the comments I received from the Editor who has handled your paper:
-
+	\pointRaised{%
 	The paper has been reviewed by a referee and an associate editor and I have read the manuscript independently, before looking at the reports.
@@
 This makes it hard for the reader to understand what is new and why the new contributions are valuable.
-A careful editing of the presentation is needed to address this shortcoming.
+A careful editing of the presentation is needed to address this shortcoming.}
 
+			\reply{E}{%
+	\textbf{We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
+In summary, we have revised our paper in the abstract and introduction as to make our contributions more clear to that of prior work.
+Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
+	}
+
+	\newpage
+
+	\pointRaised{%
 	Regarding the presentation, the AE points out a number of problems with notation, typos, and various inconsistencies.
@@
 
-	There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.
-For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p.
-5, but is not defined (in passing) until the bottom of p.
-7, and it seems to become D on p.
-8.
-Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.
-The loss function on p.
-9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_11, if Z_j, \hat{Z_j}, …)$.
-In the statement of Lemma 1 on p.13, B comes out of nowhere.
-(Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)  Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.
-For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
+	\begin{enumerate}
+	\item There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.  
+	\item For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.  
+	\item Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.  
+	\item The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_11, if Z_j, \hat{Z_j}, …)$. 
+	\item  In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)  
+	\item Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.  
+	\item For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
+	\end{enumerate}
+	}
 
+		\reply{E}{%
+	\textbf{We thank the Editor, the AE, and the reviewer for their feedback.
+We have made an effort to make it more clear our contributions in this work as opposed to previous work.
+Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
+	}
+
+		\pointRaised{%
 	Confusion can also arise from lack of clarity in the exposition.
@@
 The reader should not be expected to go back to Newcombe’s paper to clarify the issue.
+	}
+	\reply{E}{%
+	Go fix this! \todo
+	}
 
-	This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.
+		\pointRaised{%
+
+		Other comments that were raised are as follows:
+
+	\begin{enumerate}
+
+	\item This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.
 Many more specific points are raised in the AE’s report and here are a couple of other minor issues I noticed:
 
-	p. 12, second paragraph:  insert “is” after Section 3.1
+	\item p. 12, second paragraph:  insert “is” after Section 3.1
 
-	p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.”
+	\item p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.”
 
-	As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies.
+	\item As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies.
 
-	The AE finds the simulation studies to be incomplete and gives detailed suggestions on how those should be improved.
+	\item The AE finds the simulation studies to be incomplete and gives detailed suggestions on how those should be improved.
 The AE also gives important suggestions about other aspects of the manuscript that must be carefully considered and addressed.
 All of the AE’s comments are right on the mark.
+
+	\end{enumerate}
+	}
+
+	\reply{E}{%
+	\textbf{Go fix these!} \todo
+	}
 
@@
 
-%	\reply{E}{%
-%		Thank you for the opportunity to submit a revision. 
-%		We have provided detailed responses to the reviewers' comments below, 
-%		which prompted significant changes in our paper.
-%	}
+	\reply{E}{%
+		\textbf{Thank you for the opportunity to submit a revision and improve the quality of our paper. 
+		We have provided detailed responses to the reviewers' comments below, 
+		which prompted significant changes in our paper.}
+	}
+
+\textcolor{blue}{Suggestions/Tips for Brian to Discuss: Let's work on the easy fixes first.
+Go fix all the notation, check the grammar, and fix the point by point responses of the Editor as these can be done hopefully quickly.
+Let's use this more organized format so we can more quickly talk about what comment the Editor is making.
+It also makes it easier for the Editor/AE to review, so they like this (just a pointer).
+Suggest putting lines in the paper so we can also know where we are and it's often helpful to put changes in color so we can check every detail before resubmission.
+These are just some tips that have worked in the past that I hope that will help.
+I typically like to handle the big changes last as they take a lot of time and more attention to detail.
+Antidote is a nice tool for checking grammar or asking someone who is really great at proofing to find these issues and mark it up for you.
+Olivier is pretty good at spotting such things or Ted, so they might be good choices.
+Finally, given that we have some sloppy stuff going on, I'd recommend having this really polished before sending it to HRDAG or Patrick Ball will think it's a mess.
+I personally don't like sending things that aren't polished out the door, and this my fault for not catching this prior to submission.}
 
@@
 	\section*{Reviewer 2}
+
+\textcolor{blue}{Subtle comment to Brian: The Editor told us that this review is the AE, so we want to satisfy all these comments if possible since the Editor thinks the comments are "spot on"}
+
 	\setcounter{responsectr}{0}
+
+	\pointRaised{%
 	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast
@@
 	itemized list below
+	}
+
+	\reply{R1}{%
+	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017). 
+	Our review paper clarifies our novel contributions, improves upon the exposition of the writing, provides a practical guidance for users, and expands the simulations.
+We expand on these points further below.} \todo
+	}
 
@@
 A glossary in the appendix may help the reader navigate the paper more easily.
+            \reply{R1}{%
+            \textcolor{blue}{I think this is an excellent idea and it could make sure that we don't miss anything.
+It would really help the reader.
+You could make a table of the notation, and Neil does this quite well in the Marchant 2021 paper as an example.}
+            }
+
 	}
@@
 \textcolor{red}{I'm not sure I agree with this point}
+\textcolor{blue}{The reviewer is stating that equations 1 and 2 are vectorized, so everything is else needs to be vectorized as well. In short, all notation needs to be consistent.}
 
@@
 		\reply{R1}{%
-			We thank the reviewers for identifying this oversight, and have defined the notation. 
+			\textbf{We thank the reviewers for identifying this oversight, and have defined the notation.}
 		}

### 2022-09-01 ba_resubmit/response_to_reviewers.tex
update.

@@
 
-\color{red}{I need some help addressing this in the paper. The answer is yes, the algorithm can easily be adapted for the reverse sizing. The issue is mostly notational.
-
+\textcolor{red}{I need some help addressing this in the paper.
+The answer is yes, the algorithm can easily be adapted for the reverse sizing.
+The issue is mostly notational.
 It seems as though Sadinle chose to label the datasets $X_1$ and $X_2$ such that $n_1 > n_2$ for mostly notational reasons.
@@
 Jody uses the simplified notation $Z_j = n_1 + 1$ to refer to unmatched records, and this works for the same reason.
-
 However, when we relax the one-to-one constraint, or even encourage a modeller to allow multiple linkages in one dataset, we shouldn't denote the datasets according to simple notational convenience.
 This might require clever edits of the "Review of Prior Work" section, and other areas of the paper.
+Notationally, we can use  $Z_j = 0$ or $Z_j = -j$ as a way to unambiguously denote a record as nonmatching.
+Additionally, we could just strike the paragraph about the advantages of many-to-one matchings, and keep this strictly a paper about bipartite matching}
 
-Notationally, we can use  $Z_j = 0$ or $Z_j = -j$ as a way to unambiguously denote a record as nonmatching.
-
-Additionally, we could just strike the paragraph about the advantages of many-to-one matchings, and keep this strictly a paper about bipartite matching}
+\textcolor{blue}{This is a good point.
+I wonder if we could perhaps remove a lot of the notation by Sadinle/FS and redo the notation such that the ordering wouldn't matter.
+Also, it would make more sense to follow with the examples that we use in the paper -- NLTCS and EL Salvador so that we don't confuse the reader.
+Let's talk through this point as this would be a big change.
+If the algorithm and notation can be easily modified to reverse things, this would be a nice addition to the paper and improvement over Sadinle.}
 
@@
 }
+
+		\reply{R1}{%
 \textcolor{red}{I can add these definitions, no problem. But is it worth including mention of missing data?}		
+\textcolor{blue}{Given that we have missing data (or could have missing data), this needs to be added into the paper.
+Otherwise, the authors are going to complain that we don't have it.
+If you don't have this in your code, then remove it.
+Make sense?}
+}
+
 		\pointRaised{%
@@
 		\reply{R1}{%
-	We thank the reviewers for identifying this oversight, and have defined the notation. 
+\textbf{We thank the reviewers for identifying this oversight, and have defined the notation. \todo}
 }
@@
 		\reply{R1}{%
-	We thank the reviewers for identifying this oversight, and have corrected the expression. 
+\textbf{We thank the reviewers for identifying this oversight, and have corrected the expression. \todo}
 }
@@
 }
+
+\reply{R1}{%
+\textcolor{blue}{Let's talk through why the reviewer find this to be confusing.
+I think that the AE is asking for the joint posterior to be written in such as way that we can easily decompose the full conditionals more naturally.
+So, we can start with what you have and then simplify the joint posterior such the full conditionals fall out more easily.
+The point below makes me think this is his/her suggestion.}
+}
+
 		\pointRaised{%
@@
 }
+
+\reply{R1}{%
+\textcolor{blue}{This comment goes with the one above.}
+}
+
 		\pointRaised{%
@@
 		\reply{R1}{%
-	We thank the reviewers for identifying this oversight, and have corrected the expression. 
+	\textbf{We thank the reviewers for identifying this oversight, and have corrected the expression.} \todo
 }
@@
 }
+
+\reply{R1}{%
+\textbf{I would recommend fixing this to make it more clear since this is the AE.}
+}
+
 		\pointRaised{%
@@
 Should I just remove this step?}
+\textcolor{blue}{The AE makes a good point.
+We have given the full conditionals, but have not written down the Gibbs algorithm.
+The AE is expecting something like this for the algorithm: https://arxiv.org/pdf/1403.0211.pdf}
 
@@
 In your simulations, how does the use of this method or choice of S affect the results?}
+\textcolor{blue}{Do you think you could write the computational complexity of this out Brian as this would help?
+In addition, how does the SEI come into play in the Gibbs sampler?
+If it comes into play, it will need to be in the algorithm.}
 
@@
 Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
+
+            	\reply{R1}{%
+	\textcolor{red}{Do I really need this to be a Lemma?
+It seems overly clunky to me, but if its convention to provide computational complexity as a lemma, thats OK with me.} \textcolor{blue}{Yes, this is convention.}
 	        }
-\textcolor{red}{Do I really need this to be a Lemma? It seems overly clunky to me, but if its convention to provide computational complexity as a lemma, thats OK with me.}
+}
+
 		\pointRaised{%
@@
 }
-\textcolor{red}{Yes, results would be the same. To me, this is clear, I'm not sure how to clarify this in the paper}
+
+	\reply{R1}{%
+	\textcolor{red}{Yes, results would be the same. To me, this is clear, I'm not sure how to clarify this in the paper.} 
+\textcolor{blue}{Did you run them on other settings? What were they? Don't include this in the paper, but explain this here and we'll work on this together.}
+	}
+
 		\pointRaised{%
@@
 }
+
 		\reply{R1}{%
-	In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error. We have added this description in the body of the paper as well for clarity. 
+	\textbf{In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error.
+We have added this description in the body of the paper as well for clarity.
+\textcolor{blue}{We need to more clearly answer the question.
+Was this one data set?
+This isn't clear to me from the answer.}}
 }
@@
 }
-\textcolor{red}{I will redo simulations in order to get this}
+\textbf{\textcolor{red}{I will redo simulations in order to get this}}
 
@@
 		\reply{R1}{%
-			We have deleted this last sentence. Thank you for addressing this oversight. 
+			\textbf{This has been removed.}
 }
@@
 		\reply{R1}{%
-	Thank you for addressing this oversight. 
+	\textbf{This has been fixed.}
 }
@@
 }
-\textcolor{red}{Beka, can you help answer this?}
+\textbf{Participants can be added in later cohorts. This has been clarified in the paper. \todo}
 
@@
 		\reply{R1}{%
-	Thank you for addressing this oversight. We have included pointers to these figures in the main text. 
+	\textbf{This has been fixed.}
 	}
@@
 
+	\textbf{We thank the reviewer for their comments, and we have revised the clarity of our paper. \todo}
+
 \end{document}

### 2022-09-01 ba_resubmit/response_to_reviewers.tex

@@
 \textcolor{red}{I'm not sure I agree with this point}
-\textcolor{blue}{The reviewer is stating that equations 1 and 2 are vectorized, so everything is else needs to be vectorized as well. In short, all notation needs to be consistent.}
+\textcolor{blue}{I think the comment is being made as otherwise it's not so clear to the reader. I would agree with this comment as not everyone reading this will know record linkage.}
 
@@
 
-\textcolor{red}{I need some help addressing this in the paper.
+\textbf{Thank you for the interesting comment/question.
+In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper.
+We have removed the confusing text mentioned by the reviewer.
+We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue, however, this is beyond the scope of this paper.}
+
+%\textcolor{red}{I need some help addressing this in the paper.
 The answer is yes, the algorithm can easily be adapted for the reverse sizing.
@@
 Jody uses the simplified notation $Z_j = n_1 + 1$ to refer to unmatched records, and this works for the same reason.
-However, when we relax the one-to-one constraint, or even encourage a modeller to allow multiple linkages in one dataset, we shouldn't denote the datasets according to simple notational convenience.
+%However, when we relax the one-to-one constraint, or even encourage a modeller to allow multiple linkages in one dataset, we shouldn't denote the datasets according to simple notational convenience.
 This might require clever edits of the "Review of Prior Work" section, and other areas of the paper.
-Notationally, we can use  $Z_j = 0$ or $Z_j = -j$ as a way to unambiguously denote a record as nonmatching.
-Additionally, we could just strike the paragraph about the advantages of many-to-one matchings, and keep this strictly a paper about bipartite matching}
+%Notationally, we can use  $Z_j = 0$ or $Z_j = -j$ as a way to unambiguously denote a record as nonmatching.
+%Additionally, we could just strike the paragraph about the advantages of many-to-one matchings, and keep this strictly a paper about bipartite matching}
 
-\textcolor{blue}{This is a good point.
+%\textcolor{blue}{This is a good point.
 I wonder if we could perhaps remove a lot of the notation by Sadinle/FS and redo the notation such that the ordering wouldn't matter.
@@
 \textcolor{red}{I can add these definitions, no problem. But is it worth including mention of missing data?}		
-\textcolor{blue}{Given that we have missing data (or could have missing data), this needs to be added into the paper.
-Otherwise, the authors are going to complain that we don't have it.
-If you don't have this in your code, then remove it.
-Make sense?}
+\textcolor{blue}{It's my understanding that there is missing data in SV, so we need to include this.
+(If we don't include it, the reviewers will complain that we're not including a very easy component of missing data), so this is something that really should be included.}
 }

### 2022-09-02 ba_resubmit/response_to_reviewers.tex
update.
Changed formatting of response_to_reviewer docu

@@
 
-\newcommand{\pointRaised}[1]{%
+%\newcommand{\pointRaised}[1]{%
+%	\begin{tcolorbox}
+%		\itshape #1
+%	\end{tcolorbox}
+%}
+%
+%\newcounter{responsectr}[section]     % counter for response resets at each section
+%\newcommand{\reply}[2]{%
+%	\refstepcounter{responsectr}%
+%	\textbf{#1.\theresponsectr:} #2
+%}
+
+\newcommand{\pointRaised}[2]{%
+	\textbf{#1.\theresponsectr:} #2
+}
+
+\newcounter{responsectr}[section]     % counter for response resets at each section
+\newcommand{\reply}[1]{%
+	\refstepcounter{responsectr}%
 		\begin{tcolorbox}
@@
 		\end{tcolorbox}
-}
-
-\newcounter{responsectr}[section]     % counter for response resets at each section
-\newcommand{\reply}[2]{%
-	\refstepcounter{responsectr}%
-	\textbf{#1.\theresponsectr:} #2
 }
@@
 
-	\pointRaised{%
+	\pointRaised{E}{%
 	The paper has been reviewed by a referee and an associate editor and I have read the manuscript independently, before looking at the reports.
@@
 
-			\reply{E}{%
-	\textbf{We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
+			\reply{%
+	We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
 In summary, we have revised our paper in the abstract and introduction as to make our contributions more clear to that of prior work.
-Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
+Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.
+\todo
 	}
@@
 
-	\pointRaised{%
+	\pointRaised{E}{%
 	Regarding the presentation, the AE points out a number of problems with notation, typos, and various inconsistencies.
@@
 	\item Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.  
-	\item The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_11, if Z_j, \hat{Z_j}, …)$. 
+	\item The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_11, if Z_j, \hat{Z_j}, \ldots)$. 
 	\item  In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)  
@@
 
-		\reply{E}{%
-	\textbf{We thank the Editor, the AE, and the reviewer for their feedback.
+\reply{%
+	We thank the Editor, the AE, and the reviewer for their feedback.
 We have made an effort to make it more clear our contributions in this work as opposed to previous work.
-Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
+Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.\todo
 	}
 
-		\pointRaised{%
+	\pointRaised{E}{%
 	Confusion can also arise from lack of clarity in the exposition.
@@
 	}
-	\reply{E}{%
+	\reply{%
 	Go fix this! \todo
@@
 
-		\pointRaised{%
+	\pointRaised{E}{%
 
@@
 
-	\reply{E}{%
-	\textbf{Go fix these!} \todo
+	\reply{%
+	Go fix these! \todo
 	}
@@
 
-	\reply{E}{%
-		\textbf{Thank you for the opportunity to submit a revision and improve the quality of our paper. 
+	\reply{%
+		Thank you for the opportunity to submit a revision and improve the quality of our paper. 
 		We have provided detailed responses to the reviewers' comments below, 
-		which prompted significant changes in our paper.}
+		which prompted significant changes in our paper.
 	}
@@
 
-	\section*{Reviewer 2}
+	\section*{Reviewer 1}
 
@@
 
-	\pointRaised{%
+	\pointRaised{R1}{%
 	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast
@@
 
-	\reply{R1}{%
-	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017). 
+	\reply{%
+	We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017).
 Our review paper clarifies our novel contributions, improves upon the exposition of the writing, provides a practical guidance for users, and expands the simulations.
-We expand on these points further below.} \todo
+We expand on these points further below.
+\todo
 	}
 
-		\pointRaised{%
+	\pointRaised{R1}{%
 Overall, the notation is, somewhat by necessity, quite extensive.
@@
 A glossary in the appendix may help the reader navigate the paper more easily.
-            \reply{R1}{%
+            \reply{%
             \textcolor{blue}{I think this is an excellent idea and it could make sure that we don't miss anything.
@@
 
-		\pointRaised{%
+	\pointRaised{R1}{%
 Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as implied
@@
 
-		\pointRaised{%
+	\pointRaised{R1}{%
 Section 2.2, beta distribution: the $n_{12}(.)$ notation needs to be defined.
 }
-		\reply{R1}{%
-			\textbf{We thank the reviewers for identifying this oversight, and have defined the notation.}
+		\reply{%
+			We thank the reviewers for identifying this oversight, and have defined the notation.
 		}
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 3, 4th full paragraph: My understanding of the example of matching birth records and
@@
 
-\textbf{Thank you for the interesting comment/question.
+\reply{Thank you for the interesting comment/question.
 In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper.
@@
 
-		\pointRaised{%
+\pointRaised{R1}{%
 5. Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing
@@
 
-		\reply{R1}{%
+\reply{%
 \textcolor{red}{I can add these definitions, no problem. But is it worth including mention of missing data?}		
@@
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 3, full model: The Phi notation needs to be defined.
 }
-		\reply{R1}{%
-\textbf{We thank the reviewers for identifying this oversight, and have defined the notation. \todo}
+\reply{%
+We thank the reviewers for identifying this oversight, and have defined the notation. \todo
 }
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 3, full model: The square brackets seem to be misplaced, as the exponent includes
@@
 
-		\reply{R1}{%
-\textbf{We thank the reviewers for identifying this oversight, and have corrected the expression. \todo}
+		\reply{%
+We thank the reviewers for identifying this oversight, and have corrected the expression. \todo
 }
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in
@@
 
-\reply{R1}{%
+\reply{%
 \textcolor{blue}{Let's talk through why the reviewer find this to be confusing.
@@
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the
@@
 
-\reply{R1}{%
+\reply{%
 \textcolor{blue}{This comment goes with the one above.}
@@
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
 }
-		\reply{R1}{%
-	\textbf{We thank the reviewers for identifying this oversight, and have corrected the expression.} \todo
+	\reply{%
+	We thank the reviewers for identifying this oversight, and have corrected the expression. \todo
 	}
-		\pointRaised{%
+
+\pointRaised{R1}{%
 Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation
@@
 
-\reply{R1}{%
-\textbf{I would recommend fixing this to make it more clear since this is the AE.}
+	\reply{%
+	I would recommend fixing this to make it more clear since this is the AE.
 	}
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing
@@
 }
+\reply{
 \textcolor{red}{I'm not sure how to justify this step, but also this step isn't actually necessary.
@@
 The AE is expecting something like this for the algorithm: https://arxiv.org/pdf/1403.0211.pdf}
+}
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 4.2, SEI procedure.
@@
 
+\reply{%
 	\textcolor{red}{All computational savings occur from conducting the Gibbs sampler over the $P$ patterns instead of the $n_2$ records.
@@
 I think I need help articulating this clearly.}
-		\pointRaised{%
+}
+\pointRaised{R1}{%
 Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
 
-            	\reply{R1}{%
+\reply{%
 	\textcolor{red}{Do I really need this to be a Lemma?
@@
 
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 5.1 simulations – there is a detailed comparison for one set of simulation settings,
@@
 
-	\reply{R1}{%
+\reply{%
 	\textcolor{red}{Yes, results would be the same. To me, this is clear, I'm not sure how to clarify this in the paper.} 
@@
 
-		\pointRaised{%
+\pointRaised{R1}{%
 	Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results
@@
 
-		\reply{R1}{%
-	\textbf{In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error.
+\reply{%
+	In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error.
 We have added this description in the body of the paper as well for clarity.
@@
 Was this one data set?
-This isn't clear to me from the answer.}}
+This isn't clear to me from the answer.}
 }
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 5.3 simulation – RR results are not presented.
 }
-\textbf{\textcolor{red}{I will redo simulations in order to get this}}
+\reply{%
+	\textcolor{red}{I will redo simulations in order to get this}
+}
 
-		\pointRaised{%
+\pointRaised{R1}{%
 	Section 6.1 last sentences. I believe the last sentence should be removed, as it is redundant.
 }
-		\reply{R1}{%
-			\textbf{This has been removed.}
+\reply{%
+	This has been removed.
 }
-		\pointRaised{%
+\pointRaised{R1}{%
 Section 6.2, second sentence. I believe the word “survey” is missing, as in “the initial SURVEY
@@
 }
-		\reply{R1}{%
-	\textbf{This has been fixed.}
+\reply{%
+	This has been fixed.
 }
-		\pointRaised{%
+
+\pointRaised{R1}{%
 Section 6.2. In the NLTCS, are participants added in later cohorts, or may we assume that every
@@
 }
-\textbf{Participants can be added in later cohorts. This has been clarified in the paper. \todo}
+\reply{%
+	Participants can be added in later cohorts. This has been clarified in the paper. \todo
+}
 
-		\pointRaised{%
+\pointRaised{R1}{%
 	Figures 5 and 7 are not referenced in the text.
 }
-		\reply{R1}{%
-	\textbf{This has been fixed.}
+\reply{%
+	This has been fixed.
 }

### 2022-09-02 ba_resubmit/response_to_reviewers.tex
formatting changes

@@
 	\item There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.
-	\item For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.  
+For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p.
+5, but is not defined (in passing) until the bottom of p.
+7, and it seems to become D on p.
+8.
 	\item Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.
@@
 	\item Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.
-	\item For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
+For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
 	\end{enumerate}
@@
 We have made an effort to make it more clear our contributions in this work as opposed to previous work.
-Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.\todo
+Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.
+
+		\begin{enumerate}
+		\item We have included an explanation of our missingness assumptions along with the defintion of $I_{obs}.$ Additionally, we have defined $n_{12}(Z)$ at its first occurance.
+		\item We have fixed the inconsistency, using $n_1$ and $n_2$ throughout the paper.
+		\item The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1$ We agree that the latter expression is more clear, and have made that change.
+		\item  We have updated the Lemma to include the definition of $B$, in addition to all other variables included in the Lemma. 
+		\item \textcolor{red}{BEKA/JERRY: This exact reference seems purely stylistic. We need to decide how to respond to this.}
+	\end{enumerate}
 }
@@
 	}
+
 	\reply{%
-	Go fix this! \todo
+We agree with this comment, and we have removed this example as it does not add to the paper.
+Furthermore, we have gone through the paper to improve the overall exposition and clarity of the writing.\todo
 }
@@
 	\pointRaised{E}{%
-
-		Other comments that were raised are as follows:
-
+		Other comments that were raised are as follows: This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.
+Many more specific points are raised in the AE’s report and here are a couple of other minor issues I noticed:
 	\begin{enumerate}
-
-	\item This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.
-Many more specific points are raised in the AE’s report and here are a couple of other minor issues I noticed:
-
 	\item p. 12, second paragraph:  insert “is” after Section 3.1
@@
 	\reply{%
-	Go fix these! \todo
+		\begin{enumerate}
+		\item This change has been made
+
+		\item \textcolor{red}{I could not find this error}
+
+		\item We have read the paper for typos and inconsistencies, and thank the reviewers for bringing these to our attention.
+
+		\item We have responded to the AE's comments about the simulation studies
+
+	\end{enumerate}
 }
 
-	To summarize, the paper contains an interesting algorithmic contribution that can speed up calculations at the expense of relaxing some of the modeling conditions, without much adverse impact on the resulting inferences.
+\pointRaised{E}{To summarize, the paper contains an interesting algorithmic contribution that can speed up calculations at the expense of relaxing some of the modeling conditions, without much adverse impact on the resulting inferences.
 This aspect should be emphasized in a revised presentation.
@@
 The authors should prepare a careful and substantive revision that remains within the editorial limit of 25 pages and that answers the various comments satisfactorily.
-The authors should upload the revised manuscript together with a document detailing how they addressed the reviewers’ comments.
+The authors should upload the revised manuscript together with a document detailing how they addressed the reviewers’ comments.}
 
@@
 	Thank you for the opportunity to submit a revision and improve the quality of our paper.
-		We have provided detailed responses to the reviewers' comments below, 
-		which prompted significant changes in our paper.
+We have provided detailed responses to the reviewers' comments below, which prompted significant changes in our paper.
 }
@@
 However, I’m not convinced that it’s entirely consistent.
-A glossary in the appendix may help the reader navigate the paper more easily.
+A glossary in the appendix may help the reader navigate the paper more easily.}
 \reply{%
@@
 
+\pointRaised{R1}{%
+	Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as implied by equations (1) and (2).
 }
-
-	\pointRaised{R1}{%
-Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as implied
-by equations (1) and (2).
-}
+\reply{
 	\textcolor{red}{I'm not sure I agree with this point}
 	\textcolor{blue}{I think the comment is being made as otherwise it's not so clear to the reader. I would agree with this comment as not everyone reading this will know record linkage.}
+}
 
@@
 \pointRaised{R1}{%
-5. Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing
-fields, which is not at all discussed in this manuscript. This consideration of missing fields and
-the accompanying discussion of missing data mechanism should be included if such data is to be
-considered for this manuscript. This comment also applies to the $Gamma^{obs}$ notation in the
-following subsection.
+5.
+Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing fields, which is not at all discussed in this manuscript.
+This consideration of missing fields and the accompanying discussion of missing data mechanism should be included if such data is to be considered for this manuscript.
+This comment also applies to the $Gamma^{obs}$ notation in the following subsection.
 }
@@
 \reply{%
-We thank the reviewers for identifying this oversight, and have defined the notation. \todo
+	We thank the reviewers for identifying this oversight, and have defined the notation.
 }
@@
 \pointRaised{R1}{%
-Section 3, full model: The square brackets seem to be misplaced, as the exponent includes
-indices of the sum contained in the brackets.
+	Section 3, full model: The square brackets seem to be misplaced, as the exponent includes indices of the sum contained in the brackets.
 }
@@
 \reply{%
-We thank the reviewers for identifying this oversight, and have corrected the expression. \todo
+	We thank the reviewers for identifying this oversight, and have corrected the expression.
 }
+
 \pointRaised{R1}{%
-Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in
-that they seem to mix the posterior behavior (e.g., parameters are updated though standard
-multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…).
-This subsection might be easier to understand if its purpose was more bluntly described in an
-opening sentence or paragraph and streamlined to focus only on the full conditional
-distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in
-intermediate conclusions. In particular, the derivation of the probability of $Gamma_{.j}$ seems
-to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$. This is not important to the eventual full
+Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in that they seem to mix the posterior behavior (e.g., parameters are updated though standard multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…).
+This subsection might be easier to understand if its purpose was more bluntly described in an opening sentence or paragraph and streamlined to focus only on the full conditional distributions for a Gibbs Sampler.
+In addition, the mixed motivation may have led to errors in intermediate conclusions.
+In particular, the derivation of the probability of $Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$.
+This is not important to the eventual full
 conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
@@
 \pointRaised{R1}{%
-Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the
-conditional distribution of Z was included using the summation-style notation rather than the
-binomial/multinomial-style multiplicative notation that naturally motivates the full conditional
+Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the conditional distribution of Z was included using the summation-style notation rather than the binomial/multinomial-style multiplicative notation that naturally motivates the full conditional
 presented on the first line of page 8.
@@
 \reply{%
-\textcolor{blue}{This comment goes with the one above.}
+The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
+The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$.
+We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$.
+We have updated our explanation to make this more clear.
 }
@@
 \reply{%
-	We thank the reviewers for identifying this oversight, and have corrected the expression. \todo
+	We thank the reviewers for identifying this oversight, and have corrected the expression.
 }
@@
 \pointRaised{R1}{%
-Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation
-may be more clear if you replace i with $z_j$, as in $w_{z_j j}$. I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
+Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation may be more clear if you replace i with $z_j$, as in $w_{z_j j}$.
+I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
 }
@@
 \reply{%
-	I would recommend fixing this to make it more clear since this is the AE.
+	\todo
 }
@@
 \pointRaised{R1}{%
-Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing
-these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It
-would be helpful to provide a brief justification of integrating out pi.
+Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing these equations as full conditional distributions vs.
+steps in the Gibbs Sampler algorithm.
+It would be helpful to provide a brief justification of integrating out pi.
 }
-\reply{
-\textcolor{red}{I'm not sure how to justify this step, but also this step isn't actually necessary.
-We could just sample $\pi$ and use it, as in a standard Gibbs sampler.
-Should I just remove this step?}
-\textcolor{blue}{The AE makes a good point.
-We have given the full conditionals, but have not written down the Gibbs algorithm.
-The AE is expecting something like this for the algorithm: https://arxiv.org/pdf/1403.0211.pdf}
+\reply{%
+	We have included the details for integrating out $\pi$ in the appendix. \todo
 }
@@
 My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution.
-Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in
-each R.
+Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R.
 Can you quantify the computational savings due to this method?
@@
 In your simulations, how does the use of this method or choice of S affect the results?}
-\textcolor{blue}{Do you think you could write the computational complexity of this out Brian as this would help?
-In addition, how does the SEI come into play in the Gibbs sampler?
-If it comes into play, it will need to be in the algorithm.}
 
 \reply{%
-	\textcolor{red}{All computational savings occur from conducting the Gibbs sampler over the $P$ patterns instead of the $n_2$ records.
-SEI is for memory savings.
-I think I need help articulating this clearly.}
+	\todo
 }
@@
 \pointRaised{R1}{%
-Section 5.1 simulations – there is a detailed comparison for one set of simulation settings,
-including a binary definition of a “match”. Do these results differ as these settings change?
+Section 5.1 simulations – there is a detailed comparison for one set of simulation settings, including a binary definition of a “match”. Do these results differ as these settings change?
 }
@@
 \pointRaised{R1}{%
-	Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results
-	hold over multiple simulated datasets, Or was this just a chance result for each setting?
+	Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results hold over multiple simulated datasets, Or was this just a chance result for each setting?
 }
@@
 }
+
 \pointRaised{R1}{%
@@
 }
+
 \reply{%

### 2022-09-02 ba_resubmit/response_to_reviewers.tex
Worked on responses to reviewers

@@
 		\begin{enumerate}
-		\item We have included an explanation of our missingness assumptions along with the defintion of $I_{obs}.$ Additionally, we have defined $n_{12}(Z)$ at its first occurance.
+		\item \textcolor{red}{I'm unsure where I should put our missing at random (MAR) assumption, and therefore also unsure where to put the definition of $I_{obs}$}
+
+		Additionally, we have defined $n_{12}(Z)$ at its first occurance.
 		\item We have fixed the inconsistency, using $n_1$ and $n_2$ throughout the paper.
@@
 	\reply{%
-We agree with this comment, and we have removed this example as it does not add to the paper.
-Furthermore, we have gone through the paper to improve the overall exposition and clarity of the writing.\todo
+We agree with this comment, and we have removed this example as it does not add to the paper. Furthermore, we have gone through the paper to improve the overall exposition and clarity of the writing.
 }
@@
 		\item We have responded to the AE's comments about the simulation studies
-
 	\end{enumerate}
@@
 	\pointRaised{R1}{%
-	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast
-	computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases
-	without resorting to ad hoc blocking. It is built upon the Felegi-Sunter model of record linkages, where
-	linkages are made within strata of agreement between records across the two databases, and relies
-	heavily on advances made by Sadinle (2017).
-	Overall, the paper presents a practical implementation of a modified Beta Linkage approach to
-	matching. It demonstrates that even though the new method suffers from the limitations of not
-	seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of
-	features on which to match. Though the novelty of the presented methodology is primarily in the
-	computational approach, I believe it to be a real advance for practical implementation of record linkage
-	in modern applications.
-	I have no major concerns about the content or accuracy of the presentation. However, substantial
-	improvements are warranted to confirm my impression. In addition to the recommendation made by
-	Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the
-	explanation of the method should be written more clearly, the simulation comparisons should be
-	repeated/expanded, and various features of the simplification should be fleshed out a bit more. Finally,
-	more practical advice would be welcome. All of these points are more specifically described in the
-	itemized list below
+	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases without resorting to ad hoc blocking.
+It is built upon the Felegi-Sunter model of record linkages, where linkages are made within strata of agreement between records across the two databases, and relies heavily on advances made by Sadinle (2017).
+Overall, the paper presents a practical implementation of a modified Beta Linkage approach to	matching.
+It demonstrates that even though the new method suffers from the limitations of not seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of features on which to match.
+Though the novelty of the presented methodology is primarily in the computational approach, I believe it to be a real advance for practical implementation of record linkage in modern applications.
+
+	I have no major concerns about the content or accuracy of the presentation.
+However, substantial improvements are warranted to confirm my impression.
+In addition to the recommendation made by Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the explanation of the method should be written more clearly, the simulation comparisons should be repeated/expanded, and various features of the simplification should be fleshed out a bit more.
+Finally, more practical advice would be welcome.
+All of these points are more specifically described in the itemized list below
 	}
@@
 \reply{
-	\textcolor{red}{I'm not sure I agree with this point}
-	\textcolor{blue}{I think the comment is being made as otherwise it's not so clear to the reader. I would agree with this comment as not everyone reading this will know record linkage.}
+	\textcolor{red}{I'm not sure the best way to approach this. Here are two options:}
+
+	\textcolor{red}{OPTION 1: Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+Without loss of generality, denote files such that $n_1 \geq n_2$.
+We follow the convention established by Sadinle (2017) and say "record i $\in X_1$ rather than the more compact $x_{1i}$ in order to avoid double subscripts.}
+
+	\textcolor{red}{OPTION 2: We agree that the notation provided conflates the records themselves with the record indices.
+However, this notation was established in Sadinle 2017 and has been replicated in further work (Wortman 2019, Alishen-Guendel, one of Jerry's students), and we have decided to use it as well for consistency within the literature.}
 }
@@
 \reply{%
-\textcolor{red}{I can add these definitions, no problem. But is it worth including mention of missing data?}
-\textcolor{blue}{It's my understanding that there is missing data in SV, so we need to include this.
-(If we don't include it, the reviewers will complain that we're not including a very easy component of missing data), so this is something that really should be included.}
+	\textcolor{red}{I'm unsure where I should put our missing at random (MAR) assumption, and therefore also unsure where to put the definition of $I_{obs}$}
 }
@@
 In addition, the mixed motivation may have led to errors in intermediate conclusions.
-In particular, the derivation of the probability of $Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$.
-This is not important to the eventual full
-conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
+In particular, the derivation of the probability of $\Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$.
+This is not important to the eventual full conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
 }
@@
 \reply{%
-\textcolor{blue}{Let's talk through why the reviewer find this to be confusing.
-I think that the AE is asking for the joint posterior to be written in such as way that we can easily decompose the full conditionals more naturally.
-So, we can start with what you have and then simplify the joint posterior such the full conditionals fall out more easily.
-The point below makes me think this is his/her suggestion.}
+\textcolor{red}{We should meet specifically to talk about how to best present the Gibbs Sampler}
+
+We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
+\textcolor{red}{I want you guys to look at it though.
+Its pretty ugly.}
 }
@@
 \reply{%
-	\todo
+	We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper. 
 }
@@
 \reply{%
-	We have included the details for integrating out $\pi$ in the appendix. \todo
+	\textcolor{red}{I don't fully understand the first part of the comment}
+	We have included the details for integrating out $\pi$.
 }
@@
 \reply{%
-	\todo
+	In line 276, we have quantified the memory savings: "Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.
+
+	I included the following text to answer the second question.
+	\textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches.
+Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results.}
 }
@@
 Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
+}
 
 \reply{%
-	\textcolor{red}{Do I really need this to be a Lemma?
-It seems overly clunky to me, but if its convention to provide computational complexity as a lemma, thats OK with me.} \textcolor{blue}{Yes, this is convention.}
-	        }
+	We have revised the lemma to include definitions for all variables used within the statement. 
 }
@@
 \reply{%
-	\textcolor{red}{Yes, results would be the same. To me, this is clear, I'm not sure how to clarify this in the paper.}
-	\textcolor{blue}{Did you run them on other settings? What were they? Don't include this in the paper, but explain this here and we'll work on this together.}
+	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl.
+The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity.
+
+	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} 
 }
@@
 	In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error.
+Thus the table represents results over 900 pairs of datasets.
 We have added this description in the body of the paper as well for clarity.
-\textcolor{blue}{We need to more clearly answer the question.
-Was this one data set?
-This isn't clear to me from the answer.}
 }
@@
 \pointRaised{R1}{%
-Section 6.2, second sentence. I believe the word “survey” is missing, as in “the initial SURVEY
-began…”
+	Section 6.2, second sentence. I believe the word “survey” is missing, as in “the initial SURVEY began…”
 }
@@
 \pointRaised{R1}{%
-Section 6.2. In the NLTCS, are participants added in later cohorts, or may we assume that every
-member of the smaller dataset should also be included in the earlier one?
+	Section 6.2. In the NLTCS, are participants added in later cohorts, or may we assume that every member of the smaller dataset should also be included in the earlier one?
 }
 \reply{%
-	Participants can be added in later cohorts. This has been clarified in the paper. \todo
+	In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death. This has been clarified in the paper. 
 }

### 2022-09-02 ba_resubmit/kundinger_resubmit.tex
Worked on responses to reviewers

@@
 
+\usepackage{lineno}
 \usepackage{amsthm}
@@
 %% *** Frontmatter *** 
-
+\linenumbers
 \begin{frontmatter}
@@
 Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent.
-This allows us to model
-the comparison data using mixture models in the following way:
+This allows us to model the comparison data using mixture models in the following way:
 \begin{align}
@@
 Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
-In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
+In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
 However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
@@
 Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-This behavior may be desirable; for example, we may be interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable \citep{Newcombe1963}.
-In situations when a bipartite estimate is desired, we can use the post processing procedure shown in section~\ref{bayes-estimate} to resolve many-to-one matchings.
+To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in section~\ref{bayes-estimate}.
+
+Lastly, we note that it is common to encounter missing information in datasets used for record linkage.
+Whether information is missing in record $i \in \bf{X}_1$ or $j \in \bf{X}_2$, the comparison vector $\bf{\gamma}_{ij}$ will have missing values.
+We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002})
 
@@
 
-Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$ (and hyperparameters).
+Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
 Thus we use the abbreviated $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have
 
-%As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_1 + j\)) and failures (when \(z_j = n_1 + j\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$.
+%As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_1 + j\)) and failures (when \(z_j = n_1 + j\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$.
 We have
@@
 
+\textcolor{red}{I added the second line, showing the division by all the u's. Its pretty ugly! What do you guys think?}
+
 \begin{align*}
 	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
 	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
@@
 where
-%$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
 $$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
@@
 	p(Z_j|\Gamma_{.j}, \Phi, \pi) &\propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi) \\
-	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}.
+	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
+	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_1 + j).
 \end{align*}
@@
 $$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
-	\frac{\pi}{n_1}w_{z_{j},j}   & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + 1. \\
+	\frac{\pi}{n_1}w_{z_{j}, j}   & z_j \leq n_1; \\
+	1-\pi &  z_j  = n_1 + j. \\
 \end{cases}$$
@@
 
+\textcolor{red}{This is the derivation for integrating out $\pi$. Should it go here, or in the appendix?}
+\begin{align*}
+	\int_{\pi}p(Z_j | \Phi, \Gamma, \pi) d\pi &= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right]d\pi \\
+	&= \frac{\int_{\pi} \pi d\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + \left(1-\int_{\pi} \pi d\pi \right) I(z_j = n_1 + j) \\
+	&= \frac{n_{12}(\bf{Z}) + \alpha_{\pi}}{n_1 (n_2 + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_1) + \frac{n_2 - n_{12}(\bf{Z}) + \beta_{\pi}}{n_2 + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_1 + j) \\
+	&\propto w_{z_j, j}I(z_j\leq n_1) + n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_1 + j) \\
+	&\propto 
+	\begin{cases} 
+		w_{z_j, j}  & z_j \leq n_1; \\
+		n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
+	\end{cases} \\
+\end{align*}
+
 \[p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
-	w_{ij}  & i \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + j. \\
+	w_{z_j, j}  & z_j \leq n_1; \\
+	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
 \end{cases}\]
@@
 With SEI, however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks.
-Observe that all records $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{x}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
+Observe that all records $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
 Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
@@
 \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
-Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
-For simplicty, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
+\textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches.
+Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results.}Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
+For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
 
@@
 	\begin{lemma}
-		The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$, where $B$ is the number of processors available for parallelization.
+		Recall that $n_1$ and $n_2$ are the number of records in $\bf{X}_1$ and $\bf{X}_2$ respectively, $F$ is the number of fields used for comparisons across records, and $P$ is the number of patterns that comparison vectors can exhibit.
+We assume $B$ processors available for parallelization.
+Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
 		\label{lemma:fabl}
@@
 For each level of overlap and each level of error, we have 100 paired sets of 500 records.
+Thus this table represents results for 900 datasets.
 We see comparable performance for all levels of error and overlap.}
@@
 For this study, we seek to link records over the $n_1 = 20485$ records from 1982 to the $n_2 = 17466$ records from 1989.
-The NLTCS data have longitudinal links, so that in reality one does not need to do a record linkage.
+\textcolor{red}{In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death.} The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
 However, following the strategy in \cite{guha:reiter:BA} we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone files.

### 2022-09-06 ba_resubmit/response_to_reviewers.tex
Added RR to plot, many edits

@@
 		\begin{enumerate}
-		\item \textcolor{red}{I'm unsure where I should put our missing at random (MAR) assumption, and therefore also unsure where to put the definition of $I_{obs}$}
+		\item \textcolor{red}{We have included a discussion of our missingness assumptions and the definition of $I_{obs}()$ near the end of Section 3.}
 
@@
 Without loss of generality, denote files such that $n_1 \geq n_2$.
-We follow the convention established by Sadinle (2017) and say "record i $\in X_1$ rather than the more compact $x_{1i}$ in order to avoid double subscripts.}
+We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.}
 
@@
 \reply{%
-	\textcolor{red}{I'm unsure where I should put our missing at random (MAR) assumption, and therefore also unsure where to put the definition of $I_{obs}$}
+	\textcolor{red}{We have included a discussion of our missingness assumptions and the definition of $I_{obs}()$ near the end of Section 3.}
 }
@@
 \reply{%
-\textcolor{red}{We should meet specifically to talk about how to best present the Gibbs Sampler}
+\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
+I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.}
 
@@
 \reply{%
-	\textcolor{red}{I will redo simulations in order to get this}
+	\textcolor{red}{We have updated the image with these results.
+Since its is presented along side NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results.}
 }
@@
 
-	The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation
-	results are presented and the variant is applied on a new, significantly larger data set.
-	A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in
-	section 4. It is simple but consequential. The simulations and applications show how the variant
-	accelerates the computation relative to the original method and reveals the trade-offs.
-	The value of the paper resides in that computational complexity continues to be a difficult obstacle in
-	the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s
-	approach to an extent not possible before.
-	It is important to present the research for what it is: An extension of the method of Sadinle motivated
-	by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting
-	point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should
-	be made clearer, beginning in the abstract and reiterated in the conclusion.
+	The paper presents a computational variant of the method introduced by Sadinle (2017).
+Simulation results are presented and the variant is applied on a new, significantly larger data set.
+A significant part of the paper is a reproduction of the paper of Sadinle.
+The variant is introduced in section 4.
+It is simple but consequential.
+The simulations and applications show how the variant accelerates the computation relative to the original method and reveals the trade-offs.
+
+	The value of the paper resides in that computational complexity continues to be a difficult obstacle in the application of Bayesian statistics.
+The proposals of the paper enable the application of Sadinle’s approach to an extent not possible before.
+It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations.
+Here Sadinle supersedes Fellegi-Sunter in the sense while the starting point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work.
+This should be made clearer, beginning in the abstract and reiterated in the conclusion.
 

### 2022-09-06 ba_resubmit/kundinger_resubmit.tex
Added RR to plot, many edits

@@
 
-Lastly, we note that it is common to encounter missing information in datasets used for record linkage.
-Whether information is missing in record $i \in \bf{X}_1$ or $j \in \bf{X}_2$, the comparison vector $\bf{\gamma}_{ij}$ will have missing values.
-We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002})
+\textcolor{red}{Lastly, we note that it is common to encounter missing information in datasets used for record linkage.
+Whether information is missing in record $i \in \bf{X}_1$ or $j \in \bf{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
+With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}
 
@@
 %The model can then be specified as follows:
-With the \texttt{fabl} prior defined, we present the full model as follows:
+With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model as follows:
 
@@
 In contrast, we use non-optimized code written in R for \texttt{fabl}.
-While this complicates comparisons, and indeed disfavors fabl, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
+While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
 Additionally, although \texttt{fabl} is amenable to parallelization, this simulation was run on a single core.
@@
 We continue to use the precision, which is renamed the positive predictive value (PPV) in this context.
-Lastly, we also report the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.
+Lastly, we also examine the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)/n_2$.
+\textcolor{red}{To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.}
+
+	\textcolor{red}{Should I skip defining RR? And just define decision rate DR = $\sum_{j=1}^{n_2} I(\hat{Z}_j \neq R)/n_2$?}
 
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
-			\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
+			\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on simulated datasets.
+We see poorer performance for \texttt{fabl} only in situations with high overlap.}
 			\label{fig:sadinle_simulation_partial}
@@
 			\Procedure{Gibbs Sampling}{}
-			\State Initialize $m$, $u$, and $Z$ parameters.
+			\State Initialize $\Phi$, and $Z$ parameters.
 

### 2022-09-08 ba_resubmit/response_to_reviewers.tex
working on reviewer comments.

@@
 \reply{
-	\textcolor{red}{I'm not sure the best way to approach this. Here are two options:}
 
-	\textcolor{red}{OPTION 1: Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+	\textcolor{red}{We thank the reviewer for the suggestion, which has been incorporated into the paper in the following way: ``Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
 Without loss of generality, denote files such that $n_1 \geq n_2$.
-We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.}
+We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts."} \todo
 
-	\textcolor{red}{OPTION 2: We agree that the notation provided conflates the records themselves with the record indices.
+%	\textcolor{red}{OPTION 2: We agree that the notation provided conflates the records themselves with the record indices.
 However, this notation was established in Sadinle 2017 and has been replicated in further work (Wortman 2019, Alishen-Guendel, one of Jerry's students), and we have decided to use it as well for consistency within the literature.}
+%	\textcolor{blue}{We appreciate the suggestion, however, we believe it would be confusing to the existing prior work by Fellegi and Sunter (1969) that has been established as well as others that have extended the work, such as Sadinle (2014+), Wortman (2019), and Asheshin-Gundel (2020+).
+Given that it's widely used in this literature, we continue to utilize it.}
 }
@@
 \reply{%
-	We thank the reviewers for identifying this oversight, and have defined the notation.
+	This has been defined.
 }
@@
 \reply{%
-	We thank the reviewers for identifying this oversight, and have defined the notation.
+	This has been defined. 
 }
@@
 \reply{%
-	We thank the reviewers for identifying this oversight, and have corrected the expression.
+	The expression is now corrected.
 }

### 2022-09-08 ba_resubmit/response_to_reviewers.tex
update.

@@
 \textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
-I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.}
+I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
 
@@
 \reply{%
-	\textcolor{red}{I don't fully understand the first part of the comment}
+	\textcolor{red}{I don't fully understand the first part of the comment} \textcolor{blue}{I believe the person is asking you to make a connection to your full conditionals to how you make Gibbs draws.
+To do this, state that now you have all the full conditionals to run a Gibbs sampler, which can be found in Appendix XX.}
 	We have included the details for integrating out $\pi$.
@@
 
-	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} 
+	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question.
+You show a comparison for one set of simulation settings, but if you change the settings what would occur?
+I think the reviewer is asking for more clarify regarding how the lines would change.
+You may wish to give a few different scenarios to illustrate how things change?}
 }

### 2022-09-08 ba_resubmit/response_to_reviewers.tex
more additions.

@@
 	\item There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.
-For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p.
+For example, $I_{obs()}$ is never defined, $n_{12(Z)}$ first appears on p.
 5, but is not defined (in passing) until the bottom of p.
@@
 	\item Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.
-	\item The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_11, if Z_j, \hat{Z_j}, \ldots)$.
+	\item The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_{11}, \text{if} \; Z_j, \;\hat{Z_j}, \ldots)$.
 	\item  In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)
@@
 	\reply{%
-We agree with this comment, and we have removed this example as it does not add to the paper. Furthermore, we have gone through the paper to improve the overall exposition and clarity of the writing.
+\textbf{We agree with this comment, and we have removed this example as it does not add to the paper.
+Furthermore, we have gone through the paper to improve the overall exposition and clarity of the writing.}
 }
@@
 \reply{%
-	Thank you for the opportunity to submit a revision and improve the quality of our paper.
-We have provided detailed responses to the reviewers' comments below, which prompted significant changes in our paper.
+	\textbf{Thank you for the opportunity to submit a revision and improve the quality of our paper.
+We have provided detailed responses to the reviewers' comments below, which prompted significant changes in our paper.} \todo
 }
 
-\textcolor{blue}{Suggestions/Tips for Brian to Discuss: Let's work on the easy fixes first.
+%\textcolor{blue}{Suggestions/Tips for Brian to Discuss: Let's work on the easy fixes first.
 Go fix all the notation, check the grammar, and fix the point by point responses of the Editor as these can be done hopefully quickly.
@@
 
-\textcolor{blue}{Subtle comment to Brian: The Editor told us that this review is the AE, so we want to satisfy all these comments if possible since the Editor thinks the comments are "spot on"}
+%\textcolor{blue}{Subtle comment to Brian: The Editor told us that this review is the AE, so we want to satisfy all these comments if possible since the Editor thinks the comments are "spot on"}
 
@@
 It would really help the reader.
-You could make a table of the notation, and Neil does this quite well in the Marchant 2021 paper as an example.}
+You could make a table of the notation, and Neil does this quite well in the Marchant 2021 paper as an example.} \todo
 }
@@
 \reply{%
-	This has been defined.
+	\textbf{This has been defined.}
 }
@@
 
-\reply{Thank you for the interesting comment/question.
+\reply{\textbf{Thank you for the interesting comment/question.
 In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper.
 We have removed the confusing text mentioned by the reviewer.
-We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue, however, this is beyond the scope of this paper.}
+We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue, however, this is beyond the scope of this paper.}}
 
@@
 \reply{%
-	\textcolor{red}{We have included a discussion of our missingness assumptions and the definition of $I_{obs}()$ near the end of Section 3.}
+	\textcolor{red}{\textbf{Our missingness assumptions and the definition of $I_{obs}()$ are near the end of Section 3.}}
 }
@@
 \reply{%
-	This has been defined. 
+	\textbf{This has been defined.}
 }
@@
 \reply{%
-	The expression is now corrected.
+	\textbf{The expression is now corrected.}
 }
@@
 \reply{%
-	We thank the reviewers for identifying this oversight, and have corrected the expression.
+	\textbf{This has been corrected.}
 }
@@
 \reply{%
-	We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper. 
+	\textbf{We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$.
+We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.}
 }
@@
 \reply{%
-	We have revised the lemma to include definitions for all variables used within the statement. 
+	\textbf{The lemma now includes assumptions/definitions for all variables used within the statement.}
 }
@@
 \reply{%
-	This has been removed.
+	\textbf{This has been removed.}
 }
@@
 \reply{%
-	This has been fixed.
+	\textbf{This has been fixed.}
 }

### 2022-09-08 ba_resubmit/response_to_reviewers.tex
update.

@@
 This, in large part, was because of the issues that I just mentioned, and I believe that most readers would find it difficult to follow the developments, as these issues are exceedingly distracting.
+	}
 
-	\begin{enumerate}
-	\item There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.
+	\reply{
+	\textbf{We appreciate these comments immensely, and have done our utmost to address the ones listed by yourself and the AE.
+Furthermore, we have carefully gone through the paper ourselves, where we have corrected other issues with notation, typos, and various inconsistencies.} \todo
+	}
+
+	\pointRaised{E}{%
+	There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.
 For example, $I_{obs()}$ is never defined, $n_{12(Z)}$ first appears on p.
@@
 8.
-	\item Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.
-	\item The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_{11}, \text{if} \; Z_j, \;\hat{Z_j}, \ldots)$.
-	\item  In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)
-	\item Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.
-For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
-	\end{enumerate}
 	}
 
-\reply{%
-	We thank the Editor, the AE, and the reviewer for their feedback.
-We have made an effort to make it more clear our contributions in this work as opposed to previous work.
-Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.
+	\reply{
+	\textbf{The definition of $I_{obs}()$ now defined.
+We have defined $n_{12(Z)}$ in the correct place (and removed the error with $D$).
+We have checked for other notational inconsistencies.
+Additionally, we have defined $n_{12}(Z)$ when it first appears.}
+	}
 
-		\begin{enumerate}
-		\item \textcolor{red}{We have included a discussion of our missingness assumptions and the definition of $I_{obs}()$ near the end of Section 3.}
+	\pointRaised{E}{%
+	Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1. 
+	}
 
-		Additionally, we have defined $n_{12}(Z)$ at its first occurance.
-		\item We have fixed the inconsistency, using $n_1$ and $n_2$ throughout the paper.
-		\item The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1$ We agree that the latter expression is more clear, and have made that change.
-		\item  We have updated the Lemma to include the definition of $B$, in addition to all other variables included in the Lemma. 
-		\item \textcolor{red}{BEKA/JERRY: This exact reference seems purely stylistic. We need to decide how to respond to this.}
-	\end{enumerate}
+	\reply{
+	\textbf{We have fixed the inconsistency, using $n_1$ and $n_2$ throughout the paper.}
+
 	}
+
+		\pointRaised{E}{%
+	The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_{11}, \text{if} \; Z_j, \;\hat{Z_j}, \ldots)$.
+	}
+
+	\reply{
+	\textbf{The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1.$ We agree that the latter expression is more clear.
+This has been updated in our revision.}
+	}
+
+		\pointRaised{E}{%
+	In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)
+	}
+
+	\reply{
+	\textbf{We have updated the Lemma to include the definition of $B$, in addition to all other variables included in the Lemma.}
+	}
+
+        	\pointRaised{E}{%
+	Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.
+For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.
+	}
+
+	\reply{
+	\textcolor{red}{BEKA/JERRY: This exact reference seems purely stylistic.
+We need to decide how to respond to this.} \textcolor{blue}{The Editor is unhappy as the displayed equations are not consistent regarding punctuation.
+They must all be checked and be the same.
+This could be due to the fact that Brian/myself edited the paper separately.
+I would follow the notation that Mauricio uses (as one example) or that of Marchant et al.
+(2021).
+The idea is to be very consistent in every single equation.
+Yes, this is very nitpicky, but we need to satisfy the Editor.
+}
+	}
+
+.
 
@@
 \reply{%
-	In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death. This has been clarified in the paper. 
+	In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death. This has been clarified in the paper. \todo
 }

### 2022-09-09 ba_resubmit/response_to_reviewers.tex
many edits

--- 2022-09-08 ba_resubmit/response_to_reviewers.tex
@@
 \reply{
-
-	\textcolor{red}{We thank the reviewer for the suggestion, which has been incorporated into the paper in the following way: ``Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+	We thank the reviewer for the suggestion, which has been incorporated into the paper in the following way: ``Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
 Without loss of generality, denote files such that $n_1 \geq n_2$.
-We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts."} \todo
-
+We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts."
+}
 %	\textcolor{red}{OPTION 2: We agree that the notation provided conflates the records themselves with the record indices.
@@
 Given that it's widely used in this literature, we continue to utilize it.}
-}
 
@@
 \pointRaised{R1}{%
-Section 3, 4th full paragraph: My understanding of the example of matching birth records and
-marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as
-presented, as I would think that one certificate (element in X1) would match multiple birth
-records (elements in X2) which violates the size ordering. Can the algorithm be easily adjusted
-for this reverse sizing?
+Section 3, 4th full paragraph: My understanding of the example of matching birth records and marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as presented, as I would think that one certificate (element in X1) would match multiple birth records (elements in X2) which violates the size ordering.
+Can the algorithm be easily adjusted for this reverse sizing?
 }
 
-\reply{Thank you for the interesting comment/question.
+\reply{
+	Thank you for the interesting comment/question.
 In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper.
@@
 \pointRaised{R1}{%
-5.
 	Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing fields, which is not at all discussed in this manuscript.
@@
 \reply{%
-\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
-I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
-
-We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
-\textcolor{red}{I want you guys to look at it though.
-Its pretty ugly.}
+	\textcolor{red}{We have revised our presentation of the Gibbs sampler with this feedback in mind.
+The main body of the text now presents the final full conditionals, with revised derivations moved now in the appendix.
+In particular, we have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.}
 }
@@
 \reply{%
-	\textcolor{red}{I don't fully understand the first part of the comment} \textcolor{blue}{I believe the person is asking you to make a connection to your full conditionals to how you make Gibbs draws.
-To do this, state that now you have all the full conditionals to run a Gibbs sampler, which can be found in Appendix XX.}
-	We have included the details for integrating out $\pi$.
+	\textcolor{red}{We have revised the presentation of the Gibbs sampler based on this feedback.
+The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the appendix.}
 }
@@
 The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity.
-
-	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question.
-You show a comparison for one set of simulation settings, but if you change the settings what would occur?
-I think the reviewer is asking for more clarify regarding how the lines would change.
-You may wish to give a few different scenarios to illustrate how things change?}
+\textcolor{red}{To illustrate this, we have included in the appendix an additional simulation study under different comparison vector settings.}
 }

### 2022-09-09 ba_resubmit/response_to_reviewers_rcs.tex
update.

--- 2022-09-08 ba_resubmit/response_to_reviewers.tex
@@
 	\maketitle
-	We thank the Editor and reviewers for their encouraging assessment and constructive
-	feedback.
+	\textbf{We thank the Editor and reviewers for their encouraging assessment and constructive
+	feedback.}
 
@@
 			\reply{%
-	We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
+\textbf{We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
 In summary, we have revised our paper in the abstract and introduction as to make our contributions more clear to that of prior work.
-Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.
-\todo
+Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
 	}
@@
 	\reply{
-	\textbf{We have updated the Lemma to include the definition of $B$, in addition to all other variables included in the Lemma.}
+	\textbf{We have updated the Lemma to include the definition of $B$ as well as other variables in the Lemma.}
 	}
@@
 Yes, this is very nitpicky, but we need to satisfy the Editor.
-}
+Brian: I'm happy to work with you if this is still unclear or you need some additional help with it.}
 	}
@@
 	Other comments that were raised are as follows: This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.
-Many more specific points are raised in the AE’s report and here are a couple of other minor issues I noticed:
+We summarize these below.
+}
+
+\reply{	
 	\begin{enumerate}
-	\item p. 12, second paragraph:  insert “is” after Section 3.1
-
-	\item p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.”
-
+	\item p. 12, second paragraph:  insert “is” after Section 3.1.  \textbf{This has been fixed.}
+	\item p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.” \textbf{We are not able to find this error. Go check this.}
 	\item As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies.
-
+\textbf{This has been done by all authors.} \todo
 	\item The AE finds the simulation studies to be incomplete and gives detailed suggestions on how those should be improved.
@@
 All of the AE’s comments are right on the mark.
-
+\textbf{We have provided detailed comments to the AE.}
 	\end{enumerate}
@@
 
-	\reply{%
-	\begin{enumerate}
-		\item This change has been made
-
-		\item \textcolor{red}{I could not find this error}
-
-		\item We have read the paper for typos and inconsistencies, and thank the reviewers for bringing these to our attention.
-
-		\item We have responded to the AE's comments about the simulation studies
-	\end{enumerate}
-}
+%	\reply{%
+%	\begin{enumerate}
+%		\item This change has been made
+%		
+%		\item \textcolor{red}{I could not find this error}
+%		
+%		\item We have read the paper for typos and inconsistencies, and thank the reviewers for bringing these to our attention. 
+%		
+%		\item We have responded to the AE's comments about the simulation studies
+%	\end{enumerate}
+%}
 
@@
 	\reply{%
-	We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017).
+	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017).
 Our review paper clarifies our novel contributions, improves upon the exposition of the writing, provides a practical guidance for users, and expands the simulations.
-We expand on these points further below.
-\todo
+We expand on these points further below.} \todo
 	}
@@
 \reply{%
-The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
+\textbf{The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
 The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$.
-We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$.
-We have updated our explanation to make this more clear.
+We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi \mid Z)$.
+We have tried to improve our explanation of this in our revision.}
 }

### 2022-09-09 ba_resubmit/kundinger_resubmit.tex

@@
 
-Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with $n_1$ and $n_2$ records respectively.
-Without loss of generality, denote the files such that $n_1 \geq n_2$.
-In the context of bipartite matching, we assume that there are duplications across, but not within, each database.
+\textcolor{red}{Consider two databases $\bf{X}_1$ and $\bf{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+Without loss of generality, denote files such that $n_1 \geq n_2$.
+We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.} In the context of bipartite matching, we assume that there are duplications across, but not within, each database.
 Under this framework, the set of matches across databases can be represented in two equivalent ways.
@@
 	\label{gibbs_sampling}}
-We work with the following factorization of the joint posterior distribution:
-
-\begin{align*}
-	p(\bm{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\bm{Z} | \pi) p(\Phi) p(\pi) \\
-	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
-	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}.
-\end{align*}
-
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. Thus we have
+
+We initialize $\Phi$ from random draws from the prior distributions of $\bm{m}$ and $\bm{u}$, and initialize $\bm{Z}$ to reflect no matches across datasets, that is $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
+We then iterativly sample $\Phi$ and $\bm{Z}$ from their respective full conditionals:
 
@@
 $$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z}))$$
-where
-\(\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
-and
-\(\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).
-
-Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
-Thus we use the abbreviated $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have
-
-%As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_1 + j\)) and failures (when \(z_j = n_1 + j\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$.
-We have
-
-\begin{align*}
-	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
-\end{align*}
-Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\).
-Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).
-As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with
-
-\textcolor{red}{I added the second line, showing the division by all the u's. Its pretty ugly! What do you guys think?}
-
-\begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
-	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
-	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&=
-	\begin{cases} 
-		w_{z_j, j}  & z_j \leq n_1; \\
-		1 &  z_j  = n_1 + j \\
-	\end{cases}\\
-\end{align*}
-
-where
-$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
-
-These \(w_{ij}\) are precisely the likelihood ratios used in the Fellegi-Sunter model to classify matches and non-matches, and we therefore call them the Fellegi Sunter weights.
-
-With the likelihood in this form, we can derive the full conditional
-\begin{align*}
-	p(Z_j|\Gamma_{.j}, \Phi, \pi) &\propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi) \\
-	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_1 + j).
-\end{align*}
-Thus, we have
-$$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
-	\frac{\pi}{n_1}w_{z_{j}, j}   & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + j. \\
-\end{cases}$$
-We integrate over \(\pi\) and rearrange terms to produce the final full conditional:
-
-\textcolor{red}{This is the derivation for integrating out $\pi$. Should it go here, or in the appendix?}
-\begin{align*}
-	\int_{\pi}p(Z_j | \Phi, \Gamma, \pi) d\pi &= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right]d\pi \\
-	&= \frac{\int_{\pi} \pi d\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + \left(1-\int_{\pi} \pi d\pi \right) I(z_j = n_1 + j) \\
-	&= \frac{n_{12}(\bf{Z}) + \alpha_{\pi}}{n_1 (n_2 + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_1) + \frac{n_2 - n_{12}(\bf{Z}) + \beta_{\pi}}{n_2 + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_1 + j) \\
-	&\propto w_{z_j, j}I(z_j\leq n_1) + n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_1 + j) \\
-	&\propto 
+$$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
@@
 	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
-	\end{cases} \\
-\end{align*}
-
-\[p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
-\begin{cases} 
-	w_{z_j, j}  & z_j \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
-\end{cases}\]
+\end{cases},$$
+where 
+$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
+Derivations for these full conditionals can be found in Section ~\ref{app:derivations}.
+
+%We work with the following factorization of the joint posterior distribution:
+%
+%\begin{align*}
+%	p(\bm{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\bm{Z} | \pi) p(\Phi) p(\pi) \\
+%	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+%	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
+%	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
+%	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}.
+%\end{align*}
 
@@
 	\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
-\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
-on the linkage status, \(\theta_{10}\) is the loss from a false
-non-match, \(\theta_{01}\) is the loss from a false match, and
-\(\theta_{11}\) is the loss from the special case of a false match in
-which the record has a true match other than the one estimated by the
-model.
+\end{cases}\] 
+Here, \(\theta_R\) is the loss from not making a decision on the linkage status, \(\theta_{10}\) is the loss from a false non-match, \(\theta_{01}\) is the loss from a false match, and \(\theta_{11}\) is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
 
@@
 To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-Here,
-$P = |\mathcal{P}|$ is the total number of unique agreement
-patterns.
+Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns.
 \(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
@@
 These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
-We provide a summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}.
+We provide a summary of the \texttt{fabl} method through psuedocode in Appendix \ref{app:fabl-psuedocode}.
 We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
@@
 The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
+\textcolor{red}{To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in }
 
@@
 
+	\hypertarget{app:derivations}{%
+	\subsection{Derivations of Full Conditionals}\label{app:derivations}}
+
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{f\ell}$, we have
+
+\begin{align*}
+	p(m_{fl}| \Gamma, \bf{Z}) &\propto \prod_{i=1}^{n_1} \prod_{j=1}^{n_2} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
+	&= m_{fl}^{\alpha_{fl}(\bm{Z}) - 1}
+\end{align*}
+
+where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$.
+Analogous procedures lead to the posterior distribution $p(u_{fl}| \Gamma, \bf{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
+Thus for the vectors of parameters $\bf{m}_f$ and $\bf{u}_f$, we have
+
+$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
+$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})).$$
+
+Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
+Thus we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have
+
+%As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_1 + j\)) and failures (when \(z_j = n_1 + j\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$.
+We have
+
+\begin{align*}
+	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
+	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
+\end{align*}
+Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
+
+Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\).
+Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).
+We have
+
+%As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with
+
+\begin{align*}
+	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
+	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	&=
+	\begin{cases} 
+		w_{z_j, j}  & z_j \leq n_1; \\
+		1 &  z_j  = n_1 + j \\
+	\end{cases},\\
+\end{align*}
+where
+$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
+
+With the likelihood in this form, we derive an expression for the posterior distribution of $Z_j$:
+\begin{align*}
+	p\left(Z_j^{(s+1)}|\Gamma_{.j}, \Phi, \bm{Z^{(s)}}, \pi\right) &\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \Phi) P(Z_j^{(s+1)}|\pi) \\
+	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
+	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_1 + j). \\
+	&= \begin{cases} 
+		\frac{\pi}{n_1}w_{z_{j}, j}   & z_j \leq n_1; \\
+		1-\pi &  z_j  = n_1 + j. \\
+	\end{cases}
+\end{align*}
+
+For more direct comparability with the method from \cite{sadinle2018bayesian}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
+
+\begin{align*}
+	p\left(Z_j^{(s+1)}  = i| \Gamma, \Phi, \bm{Z^{(s)}}\right) &= \int_{\pi}p(Z_j^{(s+1)} | \Gamma, \Phi, \bm{Z^{(s)}}, \pi) d\pi \\
+	&= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right]d\pi \\
+	&= \frac{\int_{\pi} \pi d\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + \left(1-\int_{\pi} \pi d\pi \right) I(z_j = n_1 + j) \\
+	&= \frac{n_{12}(\bm{Z}) + \alpha_{\pi}}{n_1 (n_2 + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_1) + \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_2 + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_1 + j) \\
+	&\propto w_{z_j, j}I(z_j\leq n_1) + n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_1 + j) \\
+	&\propto 
+	\begin{cases} 
+		w_{z_j, j}  & z_j \leq n_1; \\
+		n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
+	\end{cases} \\
+\end{align*}
+
+	\clearpage
+
 	\hypertarget{app:fabl-psuedocode}{%
@@
 
+	\hypertarget{appendix-speed}{%
+	\subsection{Additional Speed Simulation Study}\label{app:appendix-speed}}
+	To illustrate that different constructions of the comparison vectors lead to similar speed gains, we replicate the speed study of Section ~\ref{speed} under different settings.
+Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.
+
+	\begin{figure}[h!]
+		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
+			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_1$ and $n_2$.
+We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:app-speed1}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[h!]
+		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
+			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_1$, and $n_2$ fixed at 500.
+We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:app-speed2}
+		\end{center}
+	\end{figure}
+
+\clearpage
+
 	\hypertarget{appendix-es}{%

### 2022-09-09 ba_resubmit/response_to_reviewers_rcs.tex
update.

@@
 \reply{%
-\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
+	\textcolor{red}{We have revised our presentation of the Gibbs sampler with this feedback in mind.
+The main body of the text now presents the final full conditionals, with revised derivations moved now in the appendix.
+In particular, we have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.}
+}
+
+%\reply{%
+%\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
 I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
-
-We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
+%
+%We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
 \textcolor{red}{I want you guys to look at it though.
 Its pretty ugly.}
-}
+%}
 
@@
 The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$.
-We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi \mid Z)$.
-We have tried to improve our explanation of this in our revision.}
+We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$.
+We have updated our explanation to make this more clear.}
 }
@@
 I’m not sure I follow the description of the SEI procedure.
-My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution.
-Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R.
-Can you quantify the computational savings due to this method?
-Do you have advice for how small S can be?
-In your simulations, how does the use of this method or choice of S affect the results?}
+(1) My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution.
+(2) Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R.
+(3) Can you quantify the computational savings due to this method?
+(4) Do you have advice for how small S can be?
+(5) In your simulations, how does the use of this method or choice of S affect the results?
+}
 
 \reply{%
-	In line 276, we have quantified the memory savings: "Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.
+\textcolor{blue}{Brian: There are five questions here that should be answered.
+The first two should be yes, the understanding is correct or no, it's not correct.
+(If the reviewer doesn't understand, then provide an explanation of your approach.
+For (3), we quantify the SEI procedure in Section 4.2, where we state:
+"Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.
+Turning to (4), in practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches.
+Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results.
+I don't see an answer to the question (5).}
+}
 
-	I included the following text to answer the second question.
-	\textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches.
-Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results.}
-}
 \pointRaised{R1}{%
@@
 \reply{%
-	\textcolor{red}{We have updated the image with these results.
+	\textcolor{red}{We have updated the \textcolor{blue}{Figure XXXX} with these results.
 Since its is presented along side NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results.}

### 2022-09-09 ba_resubmit/response_to_reviewers_rcs.tex
update.

@@
 \reply{%
-	\textcolor{red}{I don't fully understand the first part of the comment} \textcolor{blue}{I believe the person is asking you to make a connection to your full conditionals to how you make Gibbs draws.
-To do this, state that now you have all the full conditionals to run a Gibbs sampler, which can be found in Appendix XX.}
-	We have included the details for integrating out $\pi$.
+ \textbf{We have revised the presentation of the Gibbs sampler based on this feedback.
+The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the appendix.}
 }

### 2022-09-09 ba_resubmit/response_to_reviewers_rcs.tex
update.

@@
 The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity.
+\textcolor{red}{To illustrate this, we have included in the appendix an additional simulation study under different comparison vector settings.}
+}
 
-	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question.
+%\reply{%
+%	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl.
+The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity.
+%	
+%	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question.
 You show a comparison for one set of simulation settings, but if you change the settings what would occur?
@@
 You may wish to give a few different scenarios to illustrate how things change?}
-}
+%}
 

### 2022-09-09 ba_resubmit/deprecated/response_markdown.Rmd
clean up.

--- 2022-09-01 ba_resubmit/response_markdown.Rmd
@@
 \begin{quote}
-Thank you for addressing this oversight. We have corrected the loss function.
+Thank you for addressing this oversight.
+The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1$ We agree that the latter expression is more clear, and have made that change.
 \end{quote}
@@
 \begin{quote}
-\textcolor{red}{Is this true?}
+\textcolor{red}{Is this true? I don't mind making the change, I've just never seen commas in that way.}
 \end{quote}
@@
 
-It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations.
+**It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations.
 Here Sadinle supersedes Fellegi-Sunter in the sense while the starting point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work.
-This should be made clearer, beginning in the abstract and reiterated in the conclusion.
+This should be made clearer, beginning in the abstract and reiterated in the conclusion.**
+\begin{quote}
+\textcolor{red}{I'm not sure how to address this.}
+\end{quote}

### 2022-09-09 ba_resubmit/deprecated/kundinger-submit-final.tex
clean up.

--- 2021-12-20 bayesian-analysis/kundinger-submit-final.tex
@@
 $$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
+where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12} = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
 This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
@@
 	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
 	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\

### 2022-09-09 ba_resubmit/kundinger_resubmit.tex
cleaning up

@@
 	\begin{cases}
-		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-		0 \quad \text{otherwise}.\\
+		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
+		0, & \text{otherwise}.\\
 	\end{cases}
@@
 	\begin{cases}
-		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
+		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
+		n_1 + j, & \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
 	\end{cases}
@@
 \begin{align}
-	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
-	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
+	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
+	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
 	& \bm{Z} \sim \mathcal{B}, \notag
@@
 For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
-$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
+$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
 $$
@@
 \begin{cases} 
-	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
-	1-\pi &  z_j  = n_1 + j \\
+	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
+	1-\pi, &  z_j  = n_1 + j; \\
 \end{cases}$$
@@
 Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in section~\ref{bayes-estimate}.
+To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in Section~\ref{bayes-estimate}.
 
@@
 \begin{align*}
-	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}) \\
+	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \\
 	Z_j | \pi &=
 	\begin{cases} 
-		\frac{1}{n_1}\pi  & z_j \leq n_1; \\
-		1-\pi &  z_j  = n_1 + j \\
+		\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
+		1-\pi, &  z_j  = n_1 + j; \\
 	\end{cases} \\
@@
 
-$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
-$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z}))$$
+$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
+$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
 $$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
-	w_{z_j, j}  & z_j \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
+	w_{z_j, j},  & z_j \leq n_1; \\
+	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
 \end{cases},$$
@@
 $$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
-Derivations for these full conditionals can be found in Section ~\ref{app:derivations}.
+Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.
 
@@
 \[L(\hat{Z_j}, Z_j)=\begin{cases} 
-	0  & \text{if } Z_j = \hat{Z_j}; \\
+	0,  & \text{if } Z_j = \hat{Z_j}; \\
 	\theta_R,  & \text{if } \hat{Z_j} = R; \\
@@
 $$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
-
-Here, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the intuitive decision rule
-
-%$\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$ ,
-
-\[\hat{Z}_j =\begin{cases} 
+In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the intuitive decision rule
+$$\hat{Z}_j =\begin{cases} 
 	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
 	0,  & \text{otherwise}. \\
-\end{cases}\]
-
+\end{cases}$$
 For a more in-depth explanation of this function and the induced Bayes
@@
 With all record pairs converted to hashed values, we can create a more compact nested list 
-$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
+$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
 where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
-$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||$$
+$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||,$$
 and
@@
 For large data, we can partition the two datasets \(\bm{X}_1\) and
-\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more
-manageable computations. On a single machine, we can read-in data
+\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more manageable computations. On a single machine, we can read-in data
 sequentially, conduct hashing, compress information through SEI, and
@@
 Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations.
-Denote $H_p^m = \sum_{j=1}^{n_2} \bm{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$.
+Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p\right)$ to be the number of matching record pairs with agreement pattern $p$.
 The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
@@
 
-Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
+In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihooh relevant to a particular record $j \in \bm{X}_2$.
+Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
 Thus we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have
@@
 
-Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\).
-Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\).
-We have
-
+Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\). We have
 %As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with
-
 \begin{align*}
@@
 
-With the likelihood in this form, we derive an expression for the posterior distribution of $Z_j$:
+With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$.
+Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\):
 \begin{align*}
@@
 
-For more direct comparability with the method from \cite{sadinle2018bayesian}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
+For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
 

### 2022-09-09 ba_resubmit/response_to_reviewers.tex
updated speed simulations

--- 2022-09-08 ba_resubmit/response_to_reviewers.tex
@@
 	\reply{
-	\textcolor{red}{BEKA/JERRY: This exact reference seems purely stylistic.
-We need to decide how to respond to this.} \textcolor{blue}{The Editor is unhappy as the displayed equations are not consistent regarding punctuation.
-They must all be checked and be the same.
-This could be due to the fact that Brian/myself edited the paper separately.
-I would follow the notation that Mauricio uses (as one example) or that of Marchant et al.
-(2021).
-The idea is to be very consistent in every single equation.
-Yes, this is very nitpicky, but we need to satisfy the Editor.
+		Punctuation for all math has been checked.
 	}
-	}
-
-.
 
@@
 \reply{
-
-	\textcolor{red}{We thank the reviewer for the suggestion, which has been incorporated into the paper in the following way: ``Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+	We thank the reviewer for the suggestion, which has been incorporated into the paper in the following way: ``Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
 Without loss of generality, denote files such that $n_1 \geq n_2$.
-We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts."} \todo
+We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts."
 
@@
 
-\reply{\textbf{Thank you for the interesting comment/question.
+\reply{
+	\textbf{Thank you for the interesting comment/question.
 In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper.
 We have removed the confusing text mentioned by the reviewer.
-We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue, however, this is beyond the scope of this paper.}}
+We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue, however, this is beyond the scope of this paper.}
+}
 
@@
 \reply{%
-	\textcolor{red}{\textbf{Our missingness assumptions and the definition of $I_{obs}()$ are near the end of Section 3.}}
+	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ are near the end of Section 3.}
 }
@@
 \reply{%
-\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
+	We have revised our presentation of the Gibbs sampler with this feedback in mind.
+The main body of the text now presents the final full conditionals, with revised derivations moved now in the appendix.
+In particular, we have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
+}
+
+%\reply{%
+%\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
 I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
-
-We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
+%
+%We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
 \textcolor{red}{I want you guys to look at it though.
 Its pretty ugly.}
-}
+%}
 
 \pointRaised{R1}{%
-Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the conditional distribution of Z was included using the summation-style notation rather than the binomial/multinomial-style multiplicative notation that naturally motivates the full conditional
-presented on the first line of page 8.
+Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the conditional distribution of Z was included using the summation-style notation rather than the binomial/multinomial-style multiplicative notation that naturally motivates the full conditional presented on the first line of page 8.
 }
@@
 \reply{%
-The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
+	\textbf{The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
 The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$.
 We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$.
-We have updated our explanation to make this more clear.
+We have updated our explanation to make this more clear.}
 }
@@
 \reply{%
-	\textcolor{red}{I don't fully understand the first part of the comment} \textcolor{blue}{I believe the person is asking you to make a connection to your full conditionals to how you make Gibbs draws.
-To do this, state that now you have all the full conditionals to run a Gibbs sampler, which can be found in Appendix XX.}
-	We have included the details for integrating out $\pi$.
+	\textbf{We have revised the presentation of the Gibbs sampler based on this feedback.
+The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the appendix.}
 }
@@
 I’m not sure I follow the description of the SEI procedure.
-My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution.
-Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R.
-Can you quantify the computational savings due to this method?
-Do you have advice for how small S can be?
-In your simulations, how does the use of this method or choice of S affect the results?}
+(1) My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution.
+(2) Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R.
+(3) Can you quantify the computational savings due to this method?
+(4) Do you have advice for how small S can be?
+(5) In your simulations, how does the use of this method or choice of S affect the results?
+}
 
 \reply{%
-	In line 276, we have quantified the memory savings: "Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.
+	We have numbered each sentence in the comment above to clearly address each point.
+(1) and (2) are correct understandings of the SEI procedure.
+For (3), we quantify computational savings of SEI in Section 4.2, where we state:
+	"Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.
 
-	I included the following text to answer the second question.
-	\textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches.
-Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results.}
+	For (4), we have included the following sentence in the paper ``In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches."
+
+	For (5), we have included the following sentence in the paper ``Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results." 
 }
+
 \pointRaised{R1}{%
@@
 The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity.
+To illustrate this, we have included in the appendix an additional simulation study under different comparison vector settings.
+}
 
-	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question.
+%\reply{%
+%	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl.
+The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity.
+%	
+%	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question.
 You show a comparison for one set of simulation settings, but if you change the settings what would occur?
@@
 \reply{%
-	In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error.
-Thus the table represents results over 900 pairs of datasets.
+	In the caption under Figure 3, we explain that we have 100 pairs datasets for each level of overlap and error.
+Thus the table represents results from 900 pairs of datasets.
 We have added this description in the body of the paper as well for clarity.
@@
 \reply{%
-	\textcolor{red}{We have updated the image with these results.
-Since its is presented along side NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results.}
+	We have updated the Figure 4 with these results. Since its is presented along side NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results.
 }
@@
 \reply{%
-	In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death. This has been clarified in the paper. \todo
+	In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death. This has been clarified in the paper.
 }
@@
 
-	\textbf{We thank the reviewer for their comments, and we have revised the clarity of our paper. \todo}
+	\reply{%
+		\textcolor{red}{BEKA/JERRY: I think this reviewer wants us to explicitly mention Sadinle in the abstract.
+I think I had that originally, but we removed it.
+Please advise on the best way to address this comment.}
+	}
 

### 2022-09-09 ba_resubmit/kundinger_resubmit.tex
updated speed simulations

@@
 However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.
-Specifically, in \texttt{fabl} we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
-This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
+We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.
+Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
+This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
 Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
@@
 
-In what follows, Section~\ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
-Section~\ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching.
-Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}.
-Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
+In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
+Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching.
+Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}.
+Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
 
@@
 Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in Section~\ref{bayes-estimate}.
+To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in Section \ref{bayes-estimate}.
 
@@
 We initialize $\Phi$ from random draws from the prior distributions of $\bm{m}$ and $\bm{u}$, and initialize $\bm{Z}$ to reflect no matches across datasets, that is $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
-We then iterativly sample $\Phi$ and $\bm{Z}$ from their respective full conditionals:
+We then sample $\Phi$ from its full conditional:
 
@@
 $$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
+where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$, and $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
+
+Next, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
 $$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
@@
 Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
-See Appendix ~\ref{app:ohe} for details.
+See Appendix \ref{app:ohe} for details.
 
@@
 Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$.
-Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most
-\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
+Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
 \textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches.
-Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results.}Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
+Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results.}Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
 For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
@@
 Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
-
-$$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
+$$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p,$$
 $$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
@@
 
-Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
+Although sampling \(Z_j\) from a the full conditional provided in Section \ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
 This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
@@
 		\begin{cases} 
-			w_{p}\times H_{j_p}  & p \leq P; \\
-			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
+			w_{p}\times H_{j_p},  & p \leq P; \\
+			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}}, &   p = P + 1; \\
 		\end{cases}
@@
 		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
-			\frac{1}{H_{j_p}} & (i, j) \in h_p \\
-			0 & \text{otherwise.} \\
+			\frac{1}{H_{j_p}}, & (i, j) \in h_p; \\
+			0, & \text{otherwise.} \\
 		\end{cases}
@@
 	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-The distributions used (see Table~\ref{Tab:distributions}) are meant to emulate the rates of agreement across first name, last name, day of birth, and month of birth.
+The distributions used (see Table \ref{Tab:distributions}) are meant to emulate the rates of agreement across first name, last name, day of birth, and month of birth.
 For example, $u^{\text{month, agree}} = P(\text{Records have same birth-month} \mid \text{Nonmatch}) = \frac{1}{12}$.
@@
 
-	In Figure~\ref{fig:speed1}, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data.
+	In Figure \ref{fig:speed1}, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data.
 In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
@@
 	The above discussion suggests that for fixed \(n_2\), computation time for \texttt{fabl} should remain mostly constant with growing \(n_1\).
-In the simulation results displayed in Figure~\ref{fig:speed2} fixing \(n_2 = 500\), we see linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, with much more static run-time under \texttt{fabl}.
+In the simulation results displayed in Figure \ref{fig:speed2} fixing \(n_2 = 500\), we see linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, with much more static run-time under \texttt{fabl}.
 The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-\textcolor{red}{To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in }
+\textcolor{red}{To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.}
 
@@
 
-	We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table~\ref{Tab:sadinle_simulation_cutoffs}.
+	We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table \ref{Tab:sadinle_simulation_cutoffs}.
 Each simulation identifies matched individuals between two datasets, each with 500 records.
@@
 The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
-In Figure~\ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
+In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
 In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, the overall accuracy level remains high.
@@
 
-	In Figure~\ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
+	In Figure \ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
 However, with high amounts of error, and thus fewer accurate and discerning fields of information, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
@@
 This is slightly worse performance than  in the smaller simulation studies; this is expected because it is naturally more difficult to link more records with the same amount of information.
-With more linkage fields, \texttt{fabl} maintains the high accuracy in Section~\ref{accuracy}.
+With more linkage fields, \texttt{fabl} maintains the high accuracy in Section \ref{accuracy}.
 
@@
 
-	\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}.
+	\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
 This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
@@
 
-	Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}.
+	Therefore we re-ran our analysis with fewer agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
 Among the 216 possible agreement patterns, 159 are realized in the data.
 With this revised comparison specification, fabl ran in 61 seconds, approximately 4 times faster than BRL 239 second run time.
-The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure~\ref{fig:m-and-u}.
+The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure \ref{fig:m-and-u}.
 Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
@@
 This is because a large number of records in ERTL have multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
-We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Section ~\ref{accuracy}.
+We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Section \ref{accuracy}.
 Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
@@
 
-	We link records using sex, date of birth, and location using the thresholds shown in Table~\ref{Tab:nltcs-comparisons}.
+	We link records using sex, date of birth, and location using the thresholds shown in Table \ref{Tab:nltcs-comparisons}.
 Storing five comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
@@
 	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-As shown in ~\ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+As shown in  \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
 Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
@@
 	%
-	%Table~\ref{Tab:rosa-maria} illustrates an example.
+	%Table \ref{Tab:rosa-maria} illustrates an example.
 Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
@@
 The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{f\ell}$, we have
-
 \begin{align*}
@@
 \end{align*}
-
 where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$.
@@
 Thus we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have
-
-%As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_1 + j\)) and failures (when \(z_j = n_1 + j\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$.
-We have
-
 \begin{align*}
@@
 	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
+	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}.
 \end{align*}
@@
 	\subsection{Additional Speed Simulation Study}\label{app:appendix-speed}}
-	To illustrate that different constructions of the comparison vectors lead to similar speed gains, we replicate the speed study of Section ~\ref{speed} under different settings.
+	To illustrate that different constructions of the comparison vectors lead to similar speed gains, we replicate the speed study of Section \ref{speed} under different settings.
 Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.

### 2022-09-11 ba_resubmit/response_to_reviewers.tex
edits

@@
 	\reply{
-\textbf{We appreciate these comments immensely, and have done our utmost to address the ones listed by yourself and the AE.
+\textbf{We appreciate these comments, and have done our utmost to address the ones listed by yourself and the AE.
 Furthermore, we have carefully gone through the paper ourselves, where we have corrected other issues with notation, typos, and various inconsistencies.} \todo
@@
 \textbf{The definition of $I_{obs}()$ now defined.
-We have defined $n_{12(Z)}$ in the correct place (and removed the error with $D$).
+We have defined $n_{12}(Z)$ in the correct place (and removed the error with $D$).
 We have checked for other notational inconsistencies.
@@
 \reply{%
-	\textcolor{blue}{I think this is an excellent idea and it could make sure that we don't miss anything.
-It would really help the reader.
-You could make a table of the notation, and Neil does this quite well in the Marchant 2021 paper as an example.} \todo
+	We have provided a summary of notation in Appendix 7.1. \textcolor{red}{I'm not sure where I should reference this in the text. Do I need to?}
 }
@@
 In addition, the mixed motivation may have led to errors in intermediate conclusions.
-In particular, the derivation of the probability of $\Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$.
+In particular, the derivation of the probability of $\Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(\gamma_{ij}^f=l)$.
 This is not important to the eventual full conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
@@
 \pointRaised{R1}{%
-Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
+Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-$\pi$).
 }

### 2022-09-11 ba_resubmit/kundinger_resubmit.tex
edits

@@
 $$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
-where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
+where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ \textcolor{red}{is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$}.
+This representation is useful because it allows us to easily compute sufficient statistics
 $$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||,$$
-and
-$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}.$$
+which is the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, and 
+$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p},$$
+which is the number of comparison vectors that exhibit pattern $p$.
 For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
@@
 		\end{tabular}
-		\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{Tab:distributions}
+		\caption{Distributions used for $m$ and $u$ probabilities in simulation studies}\label{Tab:distributions}
 	\end{table}
@@
 	\label{sec:appendix}
+
+	\hypertarget{app:glossary}{%
+	\subsection{Summary of Notation}\label{app:glossary}}
+
+		\begin{table}[h!]
+		\centering
+		\begin{tabular}[h!]{ll}
+			\hline
+			Symbol & Description \\
+			\hline
+			$\bm{X}_1, \bm{X}_2$ & datasets\\
+			$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
+			$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
+			$f \in 1, \ldots F$ & index over fields used for comparisons \\
+			$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
+			$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
+			$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
+			$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
+			$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
+			$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
+			$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
+			$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
+			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+			$h_p$ & one hot encoding of agreement pattern $p$ \\
+			$H_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+			$H_p$ & number of comparison vectors that exhibit pattern $p$ \\
+			\hline
+		\end{tabular}
+	\end{table}
 

### 2022-09-11 ba_resubmit/response_to_reviewers.tex
response to reviewer edits

@@
 \textbf{We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
-In summary, we have revised our paper in the abstract and introduction as to make our contributions more clear to that of prior work.
+We have revised our paper in the abstract and introduction as to make our contributions more clear in relation to prior work.
 Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
@@
 	\reply{
-\textbf{The definition of $I_{obs}()$ now defined.
-We have defined $n_{12}(Z)$ in the correct place (and removed the error with $D$).
-We have checked for other notational inconsistencies.
-Additionally, we have defined $n_{12}(Z)$ when it first appears.}
+\textbf{$I_{obs}()$ is now defined. We have defined $n_{12}(Z)$ at its first occurrence (and removed the error with $D$). We have checked for other notational inconsistencies.
+	}
 }
@@
 	\reply{%
-\textbf{We agree with this comment, and we have removed this example as it does not add to the paper.
-Furthermore, we have gone through the paper to improve the overall exposition and clarity of the writing.}
+\textbf{We agree with this comment, and we have removed this example as it does not add to the paper.}
 }
@@
 	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017).
-Our review paper clarifies our novel contributions, improves upon the exposition of the writing, provides a practical guidance for users, and expands the simulations.
+Our revised paper clarifies our novel contributions, improves upon the exposition of the writing, provides practical guidance for users, and expands the simulations.
 We expand on these points further below.} \todo
@@
 We have removed the confusing text mentioned by the reviewer.
-We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue, however, this is beyond the scope of this paper.}
+We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue; however, this is beyond the scope of this paper.}
 }
@@
 \reply{%
-	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ are near the end of Section 3.}
+	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.} \textcolor{red}{Think about where best to put this.}
 }
@@
 	We have revised our presentation of the Gibbs sampler with this feedback in mind.
-The main body of the text now presents the final full conditionals, with revised derivations moved now in the appendix.
+The main body of the text now presents the final full conditionals, with revised derivations now in the Appendix.
 In particular, we have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
@@
 	\textbf{We have revised the presentation of the Gibbs sampler based on this feedback.
-The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the appendix.}
+The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the Appendix.}
 }
@@
 For (3), we quantify computational savings of SEI in Section 4.2, where we state:
-	"Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.
+	''Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.
 
-	For (4), we have included the following sentence in the paper ``In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches."
+	For (4), we have included the following sentence in the paper ``In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for plausible matches."
 
@@
 \reply{%
-	\textbf{The lemma now includes assumptions/definitions for all variables used within the statement.}
+	\textbf{The lemma now includes assumptions and definitions for all variables used within the statement.}
 }
@@
 \reply{%
-	We have updated the Figure 4 with these results. Since its is presented along side NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results.
+	We have updated the Figure 4 with these results.
+Since it is presented alongside NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results, instead of rejection rate (RR) results.
 }

### 2022-09-11 ba_resubmit/kundinger_resubmit.tex
revisions to paper

@@
 Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
+
+\textcolor{red}{BEKA/JERRY: Reviewer 2 wanted a specific reference to Sadinle in the abstract. I remember we actively decided not to do that. What do you think?}
 \end{abstract}
@@
 This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
-In this article, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.
+In this article, we consider bipartite record linkage, which merges two databases that contain duplications across but not within the respective databases.
 
@@
 The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice.
-In many situations, we know that there are no duplications within a database, meaning that each record should be linked with at most one other record.
-Thus, when the procedure results in many to one matches, some of these links must be false.
+In many situations, we know that there are no duplications within a database, meaning that each record in one file should be linked with at most one other record in the other file.
+Thus, when the procedure results in many-to-one matches, some of these links must be false.
 Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
@@
 \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is quite computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records.
-\cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
-Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
+\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records.
+\cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
+Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
 \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
@@
 In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage.
+Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space.
 This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-We demonstrate that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.
+We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed.
 Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
@@
 In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
-Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching.
-Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}.
+Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching.
+Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}.
 Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
@@
 
-\textcolor{red}{Consider two databases $\bf{X}_1$ and $\bf{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+\textcolor{red}{Consider two databases $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
 Without loss of generality, denote files such that $n_1 \geq n_2$.
@@
 
-Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$, and the set of non-matches by 
+Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
 $\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.
@@
 
-Intuitively, co-referent records (those that refer to the same entity) should be similar; records that are not co-referent should not be similar.
+Intuitively, matching records (those that refer to the same entity) should be similar; records that are not non-matching should not be similar.
 \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
+We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.
 
@@
 Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
-Thus, we can construct comparison vectors in the following way:
-$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$
-The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case studies.
-
-In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
-In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field.
+Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choice of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.
+
+\textcolor{red}{(I don't love this right here, but I don't really know where else to put it.
+But if it won't cause a problem with the reviewers, I'll just leave it.)}In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
+In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
 The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
@@
 \label{fellegi-sunter}
-The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data.
-This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
+
+Under the Fellegi Sunter model, we assume first, that comparison vectors are independent given their matching status, and second, that the match status of the record pairs are independent.
+This allows us to model the comparison data using mixture models in the following way:
+\begin{align}
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \text{Multinomial}(\bm{m}), \\
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \text{Multinomial}(\bm{u}), \notag\\
+	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
+\end{align}
+where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters for the matching record pairs $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$, and $\bm{u}$ is defined similarly for non-matching pairs.
+These $\bm{m}$ and $\bm{u}$ parameters are then used to calculate the likelihood ratio
 \begin{align}
 	\label{eqn:wts}
-	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
+	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
 \end{align}
-as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
-
-In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
-Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent.
-This allows us to model the comparison data using mixture models in the following way:
-\begin{align}
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
-	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
-\end{align}
-where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters for the matching record pairs $m_{fl} = P(\gamma_{ij}^f = l|Z_j =i)$ for all fields $f$ and agreement levels $\ell_f$, and $\bm{u}$ is defined similarly for non-matching pairs.
-These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.
-
-%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
-
-After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
+
+The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
 This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
+Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.
 
@@
 	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
-	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
+	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
 \end{align}
@@
 
+%The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data.
+This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
+%\begin{align}
+%	\label{eqn:wts}
+%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
+%\end{align}
+%as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
+%
+%In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
+%
+%%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
+%
+%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+This is a problem in practice as transitive closures can be violated.
+Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
+
 \subsection{Beta Record Linkage Model}
@@
 
-One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
+To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model.
+He proposed the  beta distribution for bipartite matching, given by
+$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+$$
+where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
+This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
+This creates a dependency that makes the sampler inherently serial, which can be slow when working on large linkage tasks.
+
+%One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
 One way of resolving this issue to incorporate the constraint directly into the model as follows:
-\begin{align}
-	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
-	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
-	& \bm{Z} \sim \mathcal{B}, \notag
-\end{align}
-where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
+%\begin{align}
+%	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
+%	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
+%	& \bm{Z} \sim \mathcal{B}, \notag
+%\end{align}
+%where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
 For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
-$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
-$$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
+%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+%$$
+%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
 This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
@@
 
-Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink.
+Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a match and $Z_j = n_1 + j$ indicates a non-match.
 In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we propose a proper and flexible prior.
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper and flexible prior.
 We denote the fast beta prior as follows:
-
 $$Z_j | \pi =
@@
 $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-One choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-We note here that prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
+One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
+We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
 
@@
 We find $\pi$ to be more interpretable parameter than $\lambda$ in the bipartite case.
-There are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
+In this setting, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
 Additionally, while the Fellegi Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions.
@@
 Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in Section \ref{bayes-estimate}.
+To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure explained in Section \ref{bayes-estimate}.
 
@@
 
+\textcolor{red}{(I don't love this right here, but I don't really know where else to put it.)}
+
 %this independent prior for each \(Z_j\), induces a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains over the Gibbs sampler under \texttt{BRL}.
@@
 With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model as follows: 
-
-%$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
-%$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
-%$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
-%$$Z_j | \pi =
-%\begin{cases} 
-%	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
-%	1-\pi &  z_j  = n_1 + j \\
-%\end{cases}$$
-%$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$
-
 \begin{align*}
@@
 
+\textcolor{red}{I actually don't like the $\Phi$ notation. It seems unnecessary. The only reason it's here is because Sadinle used it. What do you guys think?}
+
 \hypertarget{posterior-sampling}{%
@@
 
-$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
-$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
+$$\bm{m}_f|\Gamma, \bm{Z} \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
+$$\bm{u}_f|\Gamma, \bm{Z} \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
 where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$, and $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
@@
 Next, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
-$$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
+$$p\left(Z_j^{(s+1)}  = i|\Gamma, \Phi, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
@@
 adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
-\(\hat{Z}_j \in \{1, \ldots, n_1 + j, R\}\), with \(R\) representing the
+\(\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\}\), with \(R\) representing the
 option to leave the matching undetermined by the model. Specifically, we have
@@
 
-We seek to minimize the posterior expected loss, given by 
-$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
+%We seek to minimize the posterior expected loss, given by 
+%$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
 In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the intuitive decision rule
@@
 $$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p},$$
-which is the number of comparison vectors that exhibit pattern $p$.
+which is the number of comparison vectors in $\Gamma$ that exhibit pattern $p$.
 For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
@@
 However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large databases.
-Additionally, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
+Worse, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
 
 To address this issue, we propose storage efficient indexing (SEI).
-In standard indexing, one decides a priori certain criteria that all true matching pairs must satisfy, and labels any record pairs that do not meet that criteria as non-matches.
+In standard indexing, one decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important (like first name), or only pairs that exactly match on a specified number of fields.
@@
 With SEI, however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks.
-Observe that all records $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
+Observe that all records $i \in \bm{X}_1$ that share agreement pattern $p$ with $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
 Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
@@
 Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$.
-Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
-\textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches.
-Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results.}Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
+instead of storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
+\textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for plausible matches.
+Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results.} Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
 For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
@@
 
-For large data, we can partition the two datasets \(\bm{X}_1\) and
-\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more manageable computations. On a single machine, we can read-in data
-sequentially, conduct hashing, compress information through SEI, and
-delete the original data from memory before continuing with the next
-chunk of data. With multiple cores or multiple machines, this can be
-done in parallel.
+For large data, we can partition the two datasets \(\bm{X}_1\) and \(\bm{X}_2\) into smaller chunks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more manageable computations.
+On a single machine, we can read-in data sequentially, conduct hashing, compress information through SEI, and delete the original data from memory before continuing with the next chunk of data.
+With multiple cores or multiple machines, this can be done in parallel.
 Summary statistics from each pairwise chunk comparison can then be easily synthesized to recover sufficient statistics for the larger linkage task.
@@
 
-Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_1 \times n_2\) quantities.
+Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum_f L_f\) many summations over \(n_1 \times n_2\) quantities.
 These are simple calculations, but become computationally burdensome when working on large linkage tasks.
@@
 Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p\right)$ to be the number of matching record pairs with agreement pattern $p$.
-The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
-Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
+It follows that the number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
+Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors  of length $\sum_f L_f$ of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
 $$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p,$$
 $$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
-These constitute $\tilde{L}$ summations over $P$ quantities, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.
-
-Although sampling \(Z_j\) from a the full conditional provided in Section \ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
+
+Although sampling \(Z_j\) from the full conditional provided in Section \ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
 This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
@@
 First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link.
-Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\).
+Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in \bm{X}_2\).
 More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\).
 Then,
-
-%$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
-%\begin{cases} 
-%	w_{p}\times H_{j_p}  & p \leq P; \\
-%	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
-%\end{cases}$$
-%$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
-	%	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
-	%	0 & \text{otherwise.} \\
-	%\end{cases}$$
-	%
 	\begin{align}
 		\label{eqn:gibbs1}
-		P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
+		P\left(h\left(Z_j^{(s+1)}\right) = p \mid\tilde{\Gamma} \Phi, \bm{Z^{(s)}}\right) \propto
 		\begin{cases} 
@@
 	\end{align}
+	Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with (\ref{eqn:gibbs1}) above.
+After all iterations are complete, we can back-fill the records uniformally among records of the sampled agreeement pattern through (\ref{eqn:gibbs1}) below:
 	\begin{align}
@@
 
-	Lastly, we recognize that all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves.
-Thus we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with (\ref{eqn:gibbs1}) above.
-After all iterations are complete, we can back-fill the records corresponding to the agreement patterns through (\ref{eqn:gibbs2}).
-
 	We emphasize the computational gains of this split sampler: the first step is a sample from $P$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
@@
 
-	\hypertarget{computational-complexity}{%
-		\subsection{Computational Complexity}\label{computational-complexity}}
-	In this section, we prove the computational complexity of \texttt{fabl}.
+%	\hypertarget{computational-complexity}{%
+%		\subsection{Computational Complexity}\label{computational-complexity}}
+	We close this section by proving the computational complexity of \texttt{fabl}.
 
 	\begin{lemma}
-		Recall that $n_1$ and $n_2$ are the number of records in $\bf{X}_1$ and $\bf{X}_2$ respectively, $F$ is the number of fields used for comparisons across records, and $P$ is the number of patterns that comparison vectors can exhibit.
+		Recall that $n_1$ and $n_2$ are the number of records in $\bm{X}_1$ and $\bm{X}_2$ respectively, $F$ is the number of fields used for comparisons across records, and $P$ is the number of patterns that comparison vectors can exhibit.
 We assume $B$ processors available for parallelization.
@@
 		Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(\tilde{L}P)$.
+However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
 The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
@@
 Thus the complexity of sampling $\bm{Z}$ is $O(n_2 P)$.
-Since $\tilde{L} << n_1$ in all reasonable applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
+Since $P << n_1$ in all reasonable applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
 		In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
@@
 
-	We demonstrate the speed, and accuracy of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
+	We demonstrate the speed and accuracy of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
 
@@
 	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-The distributions used (see Table \ref{Tab:distributions}) are meant to emulate the rates of agreement across first name, last name, day of birth, and month of birth.
-For example, $u^{\text{month, agree}} = P(\text{Records have same birth-month} \mid \text{Nonmatch}) = \frac{1}{12}$.
+We use five binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}).
+$\bm{u}$ parameters are chosen to emulate what is seen in practice for first name, last name, day of birth, and month of birth.
 The parameters for year would vary largely by context, so we choose them here to adjust the desired difficulty of the linkage task.
-For simplicity, we consider only exact matching, so a vector (1, 0) corresponds to agreement and (0, 1) to disagreement.
-We simulate these data for different values of \(n_1\) and \(n_2\), and compare the run-time of \texttt{fabl} against \texttt{BRL}.
+We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across datasets, and compare the run-time of \texttt{fabl} against \texttt{BRL}.
 Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
@@
 		\centering
-		\begin{tabular}{rrr}
-			\hline
-			& m & u \\ 
-			\hline
-			fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
-			lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
-			day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
-			month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
-			year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
+		\begin{tabular}{lll|ll}
+			\multicolumn{1}{c}{ } & \multicolumn{2}{c}{m} & \multicolumn{2}{c}{u} \\
+			\cline{2-3} \cline{4-5}
+			& Agree & Disagree & Agree & Disagree \\
+			\hline
+			First Name & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{100}$ &  $\frac{99}{100}$ \\ 
+			Last Name & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{100}$ &  $\frac{99}{100}$ \\ 
+			Day & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{30}$ &  $\frac{29}{30}$ \\ 
+			Month & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{12}$ &  $\frac{11}{12}$ \\ 
+			Year & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{12}$ &  $\frac{11}{12}$ \\  
 			\hline
 		\end{tabular}
-		\caption{Distributions used for $m$ and $u$ probabilities in simulation studies}\label{Tab:distributions}
+		\caption{Probabilities used for $m$ and $u$ distributions in simulation studies}\label{Tab:distributions}
 	\end{table}
@@
 
-	In Figure \ref{fig:speed1}, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data.
-In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
+	In Figure \ref{fig:speed1}, where we increase both $n_1$ and $n_2$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data.
+In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of only $\bm{X}_2$).
 
@@
 Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$.
-Precision is proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
+Precision is proportion the of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
 The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
@@
 
-	\textcolor{red}{Should I skip defining RR? And just define decision rate DR = $\sum_{j=1}^{n_2} I(\hat{Z}_j \neq R)/n_2$?}
-
 	In Figure \ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
@@
 When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
-This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications}.
+Instead, record linkage techniques must be used to merge datasets before analyzing the data \citep{lum2013applications}.
 We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that had been published throughout the conflict, and the CDHES dataset consists of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
@@
 
-	\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
+	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
+This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
+However, after investigating the posterior distributions of the $\bm{m}$ and $\bm{u}$ parameters, we noticed that posterior distributions of several levels of the m and u parameters are nearly identical, suggesting that these levels are unnecessary.
+
+%	\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
 This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
@@
 		\begin{tabular}[t]{llllll}
-			\hline
 			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
@@
 Among the 216 possible agreement patterns, 159 are realized in the data.
-With this revised comparison specification, fabl ran in 61 seconds, approximately 4 times faster than BRL 239 second run time.
+With this revised comparison specification, fabl ran in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
 The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure \ref{fig:m-and-u}.
@@
 		\begin{tabular}[t]{lllll}
-			\hline
 			\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
@@
 	We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
-See `\ref{fig:overlap-plot} for a visual comparison of the Bayes estimate and posterior credible intervals for the two methods.
+See Figure \ref{fig:overlap-plot} for a visual comparison of the Bayes estimate and posterior credible intervals for the two methods.
 We note that Bayes estimates falling outside of posterior credible intervals has occurred previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-			\caption{Posterior estimates of m parameters with 95\% credible intervals for El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
+			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
 			\label{fig:m-and-u}
@@
 	We link records using sex, date of birth, and location using the thresholds shown in Table \ref{Tab:nltcs-comparisons}.
-Storing five comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
+Storing three comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
 Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
@@
 \begin{align*}
-	p\left(Z_j^{(s+1)}  = i| \Gamma, \Phi, \bm{Z^{(s)}}\right) &= \int_{\pi}p(Z_j^{(s+1)} | \Gamma, \Phi, \bm{Z^{(s)}}, \pi) d\pi \\
-	&= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right]d\pi \\
-	&= \frac{\int_{\pi} \pi d\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + \left(1-\int_{\pi} \pi d\pi \right) I(z_j = n_1 + j) \\
+	p\left(Z_j^{(s+1)}  = i| \Gamma, \Phi, \bm{Z^{(s)}}\right) &= \int_{\pi}p\left(Z_j^{(s+1)} | \Gamma, \Phi, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+	&= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(z_j = n_1 + j) \\
 	&= \frac{n_{12}(\bm{Z}) + \alpha_{\pi}}{n_1 (n_2 + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_1) + \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_2 + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_1 + j) \\

### 2022-09-11 ba_resubmit/kundinger_resubmit.tex
BK edits

@@
 \begin{align}
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \text{Multinomial}(\bm{m}), \\
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \text{Multinomial}(\bm{u}), \notag\\
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
 	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
 \end{align}
-where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters for the matching record pairs $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$, and $\bm{u}$ is defined similarly for non-matching pairs.
+where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match.
 These $\bm{m}$ and $\bm{u}$ parameters are then used to calculate the likelihood ratio
@@
 %The model can then be specified as follows:
-With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model as follows: 
+With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model.
+Let $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$.
+Let $\bm{u}$ parameters be defined analogously, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$.
+Then, using the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$, our model is:
 \begin{align*}
@@
 \end{align*}
-where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.
+%where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.
 

### 2022-09-19 ba_resubmit/kundinger_resubmit.tex
update.

@@
 \end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$
-
+$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),$$
+$\alpha_{\pi}$ and $\beta_{\pi}$ are fixed hyperparameters that encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}.$ Note the prior from \cite{wortman2019} can be viewed as a special case of the fabl prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
+
+\begin{remark}
 Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
+One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
+\end{remark}
+
+\textcolor{blue}{Furthermore, note that it is common to encounter missing information in datasets used for record linkage.
+Whether information is missing in record $i \in \bf{X}_1$ or $j \in \bf{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
+With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}
+\textcolor{red}{(I don't love this right here, but I don't really know where else to put it.)}
+
+%Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
 $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
@@
 To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure explained in Section \ref{bayes-estimate}.
-
-\textcolor{red}{Lastly, we note that it is common to encounter missing information in datasets used for record linkage.
-Whether information is missing in record $i \in \bf{X}_1$ or $j \in \bf{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
-With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
-To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}
-
-\textcolor{red}{(I don't love this right here, but I don't really know where else to put it.)}
 

### 2022-09-20 ba_resubmit/response_to_reviewers.tex
update.

@@
 
+\newpage
+
+\section{Additional Points Addressed}
+
+We highlight some additional points addressed in our revision. \todo
+\textcolor{blue}{Brian: These are just some suggestions that I have, so we can chat through these in case you don't agree!
+I have these marked up on the paper itself to help you more easily make the changes.}
+
+\begin{enumerate}
+\item We have added introductory sentences to each section/subsection to provide more guidance to the reader. 
+\item The word databases is used throughout instead of files or lists. 
+\item We have checked for the proper use of commas throughout the paper. 
+\item We have re-written our proposed work on page 2 to make it more clear our contributions to the literature as it was missing our improvements over existing work in terms of speed with minimal loss in accuracy as well as providing two case studies.
+One case study is from the original paper of our extension, while another case study illustrates our ability to scale to larger databases.
+We provide a more comprehensive discussion regarding our work.
+\item We have broken up sentences that are too long.
+\item We have removed repetitive words, such as intuitive, additionally.
+\textcolor{blue}{Brian: you can use something like antidote to check for these or go through yourself.
+Be careful with the word intuitive.
+While it might be obvious to you, it may not be obvious to the Editor and this word can be quite jarring).
+As an example, in my job talk, Mike West did not find the linkage structure to be intuitive and found it to be quite weird and confusing!}
+\item We have checked that we define terms before they are introduced. 
+\item We have defined the parameters $m$ and $u$ on page 4. 
+\item One page 5, section 2.2, we have made it more clear why our extension is warranted and needed in the literature. 
+\item We have removed confusing language, such as "compactful representation of" on page 5. 
+\item We have checked that fast beta linkage is represented by fabl throughout. 
+\item \textcolor{blue}{See page 10 for a long list of comments.}
+\item We do not start sentences with a Greek letter or a parameter.
+\item We have fixed the sentence ``sampling a value from...." that has a missing verb (see page 11).
+\item We have fixed a typo in equation 6, which is a missing (). 
+\end{enumerate}
 \end{document}

### 2022-09-20 ba_resubmit/kundinger_resubmit.tex
proposed change on p11.

@@
 We provide a summary of the \texttt{fabl} method through psuedocode in Appendix \ref{app:fabl-psuedocode}.
+\textcolor{blue}{The computational complexity of \texttt{fabl} is given in Lemma \ref{lemma:fabl}.}
+
+%	We close this section by proving the computational complexity of \texttt{fabl}.
+
+	\begin{remark}
 	We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
 \texttt{BRL} would see considerable speed up through the procedure, but computation time would still scale with the size of the larger dataset.
+	\end{remark}
 
@@
 %		\subsection{Computational Complexity}\label{computational-complexity}}
-	We close this section by proving the computational complexity of \texttt{fabl}.
 

### 2022-09-20 ba_resubmit/response_to_reviewers.tex
update.

@@
 \textcolor{blue}{Brian: These are just some suggestions that I have, so we can chat through these in case you don't agree!
-I have these marked up on the paper itself to help you more easily make the changes.}
+I have these marked up on a printed copy of the paper itself to help you more easily make the changes.
+Please note I have not made these changes, but I'm marking them on a marked up paper so you can go back through and make changes.
+It's likely that there may be more errors, so please go over things very carefully for all errors as the content is all there, we just have to weed out these tedious details.}
 
@@
 \item We have fixed a typo in equation 6, which is a missing (). 
+\item We have corrected many typos in section 5.1 page 12 (first paragraph). These include a comma that should not be present, a period instead of a comma, and a Greek letter that starts a sentence.
+\item Break up the first long sentence in Lemma 1 as it's hard to read. 
 \end{enumerate}

### 2022-09-20 ba_resubmit/kundinger_resubmit.tex
update nltcs.

@@
 For this study, we seek to link records over the $n_1 = 20485$ records from 1982 to the $n_2 = 17466$ records from 1989.
-\textcolor{red}{In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death.} The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
-However, following the strategy in \cite{guha:reiter:BA} we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone files.
+\textcolor{blue}{Within each database there are no duplications, however, there can be duplications across the databases.
+In addition, patients can either drop out of the study, pass away, or enter as new patients.
+Hence, the assumptions of our model hold for this study.}
+%	\textcolor{red}{In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death.} 
+	The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
+However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone databases.
 

### 2022-09-20 ba_resubmit/response_to_reviewers.tex
done with pas of full paper.

@@
 \item Break up the first long sentence in Lemma 1 as it's hard to read. 
+\item "In the simulation results displayed..." This sentence is awkward and should be rephrased. Page 13. 
+\item Appendices should be done as A, B, C or the convention in BA (please check). 
+\item There is a comma present that needs to be removed in Figure 2. 
+\item Page 15, provide a reference to the Appendix.
+\item proportion should be proportional on page 15
+\item datasets should be databases on page 17
+\item bold m and u on page 18
+\item Therefore needs a comma after it on page 18
+\item equation 7 is not consistent with other equations on page 20
+\item Convergence is stated on page 20. We have shown no issue of lack of convergence. Please address. 
+\item Conclusion needs to be longer most likely or they will ask us to address this. 
 \end{enumerate}

### 2022-09-20 ba_resubmit/response_to_reviewers.tex
update.

@@
 \item Conclusion needs to be longer most likely or they will ask us to address this. 
+\item Reference summary of notation in main paper and make sure it is complete. 
+\item Refer the reader back to the main body of appendix with a reference so it's easier to find the section. 
+\item Provide an introduction to each subsection so it's more clear what you're presenting. Example: Section 7.2: In this section, we derive the full conditional distributions of section \ref{}. 
+\item Similar minor details in appendix that are marked up regarding punctuation/typos.
 \end{enumerate}

### 2022-09-20 ba_resubmit/response_to_reviewers.tex
answered q of brian.

@@
 	We have provided a summary of notation in Appendix 7.1. \textcolor{red}{I'm not sure where I should reference this in the text. Do I need to?}
+	\textcolor{blue}{I would recommend a table for notation of the model and a table for notation of your hashing method.
+Why?
+It will be strange to reference it all together.
+I would try and put these in the main body if possible as it will really help the reader.
+You could also write a graphical model with notation of your model to help the reader.
+These are just some thoughts to think about.
+(I'm going to leave this up to you Brian if you want to abandon the able as it's quite annoying).}
 }
@@
 \reply{%
-	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.} \textcolor{red}{Think about where best to put this.}
+	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.} \textcolor{red}{Think about where best to put this.} \textcolor{blue}{I put this somewhere; feel free to move it if you don't like it!}
 }
@@
 	\reply{%
-		\textcolor{red}{BEKA/JERRY: I think this reviewer wants us to explicitly mention Sadinle in the abstract.
-I think I had that originally, but we removed it.
-Please advise on the best way to address this comment.}
+	Thank you for your positive feedback regarding our paper.
+We have clarified that we are motivation by the work of Sadinle and computational considerations.
+We have tried to make this more clear in both the abstract, introduction, and conclusion.}
+		\textcolor{red}{BEKA/JERRY: I think this reviewer wants us to explicitly mention Sadinle in the abstract. I think I had that originally, but we removed it. 
 	}

### 2022-09-20 ba_resubmit/kundinger_resubmit.tex
answered q of brian.

@@
 However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.
+\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.
+Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}
 Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
@@
 
-\textcolor{red}{BEKA/JERRY: Reviewer 2 wanted a specific reference to Sadinle in the abstract. I remember we actively decided not to do that. What do you think?}
+\textcolor{red}{BEKA/JERRY: Reviewer 2 wanted a specific reference to Sadinle in the abstract.
+I remember we actively decided not to do that.
+What do you think?} \textcolor{blue}{I added this back in.
+He only has one concern, so let's make sure to satisfy the person.}
 \end{abstract}

### 2022-09-20 ba_resubmit/kundinger_resubmit.tex
update.

@@
 
+\textcolor{blue}{In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
+In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
+The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
+However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.}
+
 In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
 Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space.
-This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed.
+This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs and (2) compute the pairwise record comparisons over large datasets via parallel computing.
+\textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
+ \textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking.
+This can be useful when there is no reliable blocking field available or one desires estimates of model parameters for the entire sample in question.
+In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.
+We illustrate our methodology on two case studies and two simulation scenarios, where Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.}
+
+%We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed.
 Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
@@
 
-\textcolor{red}{(I don't love this right here, but I don't really know where else to put it.
-But if it won't cause a problem with the reviewers, I'll just leave it.)}In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
+\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}
+\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}
+
+%In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
 In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
@@
 However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
-
-In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
+%
+%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
 This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
@@
 \textcolor{red}{I actually don't like the $\Phi$ notation. It seems unnecessary. The only reason it's here is because Sadinle used it. What do you guys think?}
+\textcolor{blue}{I would leave it since they didn't complain unless you have a better idea.}
 
@@
 
+\textcolor{blue}{This needs to address the concern of R3, where we didn't cite Sadinle/FS. I would also suggest that we take the time to elaborate on some future work.}
+
 	%Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.

### 2022-09-20 ba_resubmit/kundinger_resubmit.tex
update.

@@
 
-\textcolor{blue}{In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
+\textcolor{blue}{In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2),$ where $n_1$ and $n_2$ are the sizes of each database.
+The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
 In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
@@
 
-\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}
-\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}
+%\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}
+%\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}
 

### 2022-09-20 ba_resubmit/response_to_reviewers.tex
update.

@@
 	\reply{%
-	Thank you for your positive feedback regarding our paper.
+	\textbf{Thank you for your positive feedback regarding our paper.
 We have clarified that we are motivation by the work of Sadinle and computational considerations.
-We have tried to make this more clear in both the abstract, introduction, and conclusion.}
-		\textcolor{red}{BEKA/JERRY: I think this reviewer wants us to explicitly mention Sadinle in the abstract. I think I had that originally, but we removed it. 
-	}
+We have tried to make this more clear in both the abstract, introduction, and conclusion.}} \todo
 

### 2022-09-21 ba_resubmit/response_to_reviewers.tex
response to comments

--- 2022-09-20 ba_resubmit/response_to_reviewers.tex
@@
 \textcolor{blue}{Brian: These are just some suggestions that I have, so we can chat through these in case you don't agree!
-I have these marked up on a printed copy of the paper itself to help you more easily make the changes.
-Please note I have not made these changes, but I'm marking them on a marked up paper so you can go back through and make changes.
-It's likely that there may be more errors, so please go over things very carefully for all errors as the content is all there, we just have to weed out these tedious details.}
+I have these marked up on the paper itself to help you more easily make the changes.}
 
@@
 \item We have added introductory sentences to each section/subsection to provide more guidance to the reader. 
+\begin{itemize}
+	\item \textcolor{red}{I originally had sentences like this, and Jerry removed them. I also like it better without these introductory sentences, and don't find them necessary.}
+\end{itemize}
 \item The word databases is used throughout instead of files or lists. 
+\begin{itemize}
+	\item \todo
+\end{itemize}
 \item We have checked for the proper use of commas throughout the paper. 
+\begin{itemize}
+	\item \textcolor{red}{Done}
+\end{itemize}
 \item We have re-written our proposed work on page 2 to make it more clear our contributions to the literature as it was missing our improvements over existing work in terms of speed with minimal loss in accuracy as well as providing two case studies.
@@
 We provide a more comprehensive discussion regarding our work.
+\begin{itemize}
+	\item \textcolor{red}{I don't really understand what you want me to change there.}
+\end{itemize}
 \item We have broken up sentences that are too long.
+\begin{itemize}
+	\item \textcolor{red}{Done.}
+\end{itemize}
 \item We have removed repetitive words, such as intuitive, additionally.
@@
 \item We have checked that we define terms before they are introduced.
+\begin{itemize}
+	\item \textcolor{red}{Done}
+\end{itemize} 
 \item We have defined the parameters $m$ and $u$ on page 4. 
+\begin{itemize}
+	\item \textcolor{red}{I originally had the precise definition of $m$ and $u$ at this spot, giving the $m_{fl} = p(\gamma_{ij}^f = l | Z_{j} = i)$ definitions.
+However, technically, FS do not specify that comparison vectors be thresholded to discrete values, so this definition technically doesn't make sense here.
+I checked, and Sadinle defines the $m$ and $u$ parameters during the FS review section in the same way.}
+
+	\textcolor{red}{I agree that reviewers might take note of this, so I could add the following sentence.}
+
+	\textcolor{red}{When using comparison vectors with discrete agreement levels as described above, the $\bm{m}$ parameter takes the form $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$.}
+\end{itemize}
 \item One page 5, section 2.2, we have made it more clear why our extension is warranted and needed in the literature. 
+\begin{itemize}
+	\item \textcolor{red}{I don't really understand what you want me to change there.
+Its not really possible to say the Sadinle doesn't work for X many records in any precise way, because it depends on the number of comparison fields and the number of records in each dataset.
+Also, I don't know how to say exactly when something is "too slow" in a precise way.}
+\end{itemize}
 \item We have removed confusing language, such as "compactful representation of" on page 5.
+\begin{itemize}
+	\item \textcolor{red}{Done}
+\end{itemize}
 \item We have checked that fast beta linkage is represented by fabl throughout.
+\begin{itemize}
+	\item \textcolor{red}{Done. However, I think "fast beta prior" is fine, and I have left that as is.}
+\end{itemize} 
 \item \textcolor{blue}{See page 10 for a long list of comments.}
+\begin{itemize}
+	\item \textcolor{red}{I don't understand what this is.}
+\end{itemize}
 \item We do not start sentences with a Greek letter or a parameter.
+\begin{itemize}
+	\item \textcolor{red}{Done.}
+\end{itemize}
 \item We have fixed the sentence ``sampling a value from...." that has a missing verb (see page 11).
+\begin{itemize}
+	\item \textcolor{red}{Done.}
+\end{itemize}
 \item We have fixed a typo in equation 6, which is a missing (). 
-\item We have corrected many typos in section 5.1 page 12 (first paragraph). These include a comma that should not be present, a period instead of a comma, and a Greek letter that starts a sentence.
-\item Break up the first long sentence in Lemma 1 as it's hard to read. 
-\item "In the simulation results displayed..." This sentence is awkward and should be rephrased. Page 13. 
-\item Appendices should be done as A, B, C or the convention in BA (please check). 
-\item There is a comma present that needs to be removed in Figure 2. 
-\item Page 15, provide a reference to the Appendix.
-\item proportion should be proportional on page 15
-\item datasets should be databases on page 17
-\item bold m and u on page 18
-\item Therefore needs a comma after it on page 18
-\item equation 7 is not consistent with other equations on page 20
-\item Convergence is stated on page 20. We have shown no issue of lack of convergence. Please address. 
-\item Conclusion needs to be longer most likely or they will ask us to address this. 
-\item Reference summary of notation in main paper and make sure it is complete. 
-\item Refer the reader back to the main body of appendix with a reference so it's easier to find the section. 
-\item Provide an introduction to each subsection so it's more clear what you're presenting. Example: Section 7.2: In this section, we derive the full conditional distributions of section \ref{}. 
-\item Similar minor details in appendix that are marked up regarding punctuation/typos.
+\begin{itemize}
+	\item \textcolor{red}{Done. But this was a new typo introduced in my editing, not in the original submission. So I'll remove this one from the list.}
+\end{itemize}
 \end{enumerate}

### 2022-10-01 ba_resubmit/response_to_reviewers.tex

--- 2022-09-20 ba_resubmit/response_to_reviewers.tex
@@
 	\item p. 12, second paragraph:  insert “is” after Section 3.1.  \textbf{This has been fixed.}
-	\item p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.” \textbf{We are not able to find this error. Go check this.}
+	\item p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.” \textbf{This has been fixed.}
 	\item As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies.
@@
 \reply{%
-	We have provided a summary of notation in Appendix 7.1. \textcolor{red}{I'm not sure where I should reference this in the text. Do I need to?}
+	We have provided a summary of notation in Appendix 7.1.
+
 	\textcolor{blue}{I would recommend a table for notation of the model and a table for notation of your hashing method.
@@
 (I'm going to leave this up to you Brian if you want to abandon the able as it's quite annoying).}
+
+	\textcolor{red}{Done}
 }
@@
 \reply{%
-	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.} \textcolor{red}{Think about where best to put this.} \textcolor{blue}{I put this somewhere; feel free to move it if you don't like it!}
+	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.} 
 }
@@
 }
+
 \reply{%
-	\textbf{This has been defined.}
+	\textbf{We have removed the $\Phi$ notation, using the parameters $\bm{m}$ and $\bm{u}$ throughout the paper.}
 }
@@
 }
-
-%\reply{%
-%\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler.
-I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
-%
-%We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
-\textcolor{red}{I want you guys to look at it though.
-Its pretty ugly.}
-%}
 
@@
 
-%\reply{%
-%	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl.
-The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity.
-%	
-%	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question.
-You show a comparison for one set of simulation settings, but if you change the settings what would occur?
-I think the reviewer is asking for more clarify regarding how the lines would change.
-You may wish to give a few different scenarios to illustrate how things change?}
-%}
-
 \pointRaised{R1}{%
@@
 We have clarified that we are motivation by the work of Sadinle and computational considerations.
-We have tried to make this more clear in both the abstract, introduction, and conclusion.}} \todo
+We have tried to make this more clear in both the abstract, introduction, and conclusion.}}
 
@@
 We highlight some additional points addressed in our revision. \todo
+
 \textcolor{blue}{Brian: These are just some suggestions that I have, so we can chat through these in case you don't agree!
@@
 
+\textcolor{Red}{For the purpose of our revisions, I commented out the comments about minor typos. The comments that remain are things I want you to read so we can discuss.}
+
 \begin{enumerate}
 \item We have added introductory sentences to each section/subsection to provide more guidance to the reader. 
+\begin{itemize}
+	\item \textcolor{red}{I originally had sentences like this, and Jerry removed them. I also like it better without these introductory sentences, and don't find them necessary.}
+\end{itemize}
+
 \item The word databases is used throughout instead of files or lists. 
-\item We have checked for the proper use of commas throughout the paper. 
-\item We have re-written our proposed work on page 2 to make it more clear our contributions to the literature as it was missing our improvements over existing work in terms of speed with minimal loss in accuracy as well as providing two case studies.
+\begin{itemize}
+	\item \textcolor{red}{Sadinle uses the words datafile and file. I'll remove database, and do that instead.}
+\end{itemize}
+
+%\item We have checked for the proper use of commas throughout the paper. 
+%\begin{itemize}
+%	\item \textcolor{red}{Done}
+%\end{itemize}
+
+%\item We have re-written our proposed work on page 2 to make it more clear our contributions to the literature as it was missing our improvements over existing work in terms of speed with minimal loss in accuracy as well as providing two case studies.
 One case study is from the original paper of our extension, while another case study illustrates our ability to scale to larger databases.
 We provide a more comprehensive discussion regarding our work.
+%\begin{itemize}
+%	\item \textcolor{red}{I don't think that the proposed changes made the contribution more clear. I did other edits, adding the specifics }
+%\end{itemize}
+
 \item We have broken up sentences that are too long.
-\item We have removed repetitive words, such as intuitive, additionally.
+\begin{itemize}
+	\item \textcolor{red}{We disagree on some of these. For all of the longer sentences that remain, I believe they are clearer together than broken up.}
+\end{itemize}
+
+%\item We have removed repetitive words, such as intuitive, additionally.
 \textcolor{blue}{Brian: you can use something like antidote to check for these or go through yourself.
@@
 As an example, in my job talk, Mike West did not find the linkage structure to be intuitive and found it to be quite weird and confusing!}
-\item We have checked that we define terms before they are introduced. 
+%
+%\item We have checked that we define terms before they are introduced.
+%\begin{itemize}
+%	\item \textcolor{red}{Done}
+%\end{itemize}
+
 \item We have defined the parameters $m$ and $u$ on page 4. 
+\begin{itemize}
+	\item \textcolor{red}{I originally had the precise definition of $m$ and $u$ at this spot, giving the $m_{fl} = p(\gamma_{ij}^f = l | Z_{j} = i)$ definitions.
+However, technically, FS do not specify that comparison vectors be thresholded to discrete values, so this definition technically doesn't make sense here.
+I checked, and Sadinle defines the $m$ and $u$ parameters during the FS review section the same way that I do.}
+
+	\textcolor{red}{I agree that reviewers might take note of this, so I could add the following sentence.}
+
+	\textcolor{red}{"When using comparison vectors with discrete agreement levels as described above, the $\bm{m}$ parameter takes the form $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$."}
+\end{itemize}
+
 \item One page 5, section 2.2, we have made it more clear why our extension is warranted and needed in the literature. 
+\begin{itemize}
+	\item \textcolor{red}{Done.}
+\end{itemize}
+
 \item We have removed confusing language, such as "compactful representation of" on page 5.
+\begin{itemize}
+	\item \textcolor{red}{Done}
+\end{itemize}
+
 \item We have checked that fast beta linkage is represented by fabl throughout.
+\begin{itemize}
+	\item \textcolor{red}{Done. However, I think "fast beta prior" is fine, and I have left that as is.}
+\end{itemize}
+
 \item \textcolor{blue}{See page 10 for a long list of comments.}
-\item We do not start sentences with a Greek letter or a parameter.
-\item We have fixed the sentence ``sampling a value from...." that has a missing verb (see page 11).
-\item We have fixed a typo in equation 6, which is a missing (). 
-\item We have corrected many typos in section 5.1 page 12 (first paragraph). These include a comma that should not be present, a period instead of a comma, and a Greek letter that starts a sentence.
-\item Break up the first long sentence in Lemma 1 as it's hard to read. 
-\item "In the simulation results displayed..." This sentence is awkward and should be rephrased. Page 13. 
+\begin{itemize}
+	\item \textcolor{red}{I don't understand what this is.}
+\end{itemize}
+
+%\item We do not start sentences with a Greek letter or a parameter.
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+
+%\item We have fixed the sentence ``sampling a value from...." that has a missing verb (see page 11).
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item We have fixed a typo in equation 6, which is a missing (). 
+%\begin{itemize}
+%	\item \textcolor{red}{Done. But this was a new typo introduced in my editing, not in the original submission. So I'll remove this one from the list.}
+%\end{itemize}
+%\item We have corrected many typos in section 5.1 page 12 (first paragraph). These include a comma that should not be present, a period instead of a comma, and a Greek letter that starts a sentence.
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item Break up the first long sentence in Lemma 1 as it's hard to read. 
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item "In the simulation results displayed..." This sentence is awkward and should be rephrased. Page 13. 
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
 \item Appendices should be done as A, B, C or the convention in BA (please check). 
-\item There is a comma present that needs to be removed in Figure 2. 
-\item Page 15, provide a reference to the Appendix.
-\item proportion should be proportional on page 15
-\item datasets should be databases on page 17
-\item bold m and u on page 18
-\item Therefore needs a comma after it on page 18
-\item equation 7 is not consistent with other equations on page 20
-\item Convergence is stated on page 20. We have shown no issue of lack of convergence. Please address. 
+\begin{itemize}
+	\item \textcolor{red}{None of the articles I saw on the BA homepage had appendices. Some had "Supplementary Material," and those had numbered headings. So I'm leaving it as is for now. }
+\end{itemize}
+
+%\item There is a comma present that needs to be removed in Figure 2. 
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%
+%\item Page 15, provide a reference to the Appendix.
+%\begin{itemize}
+%	\item \textcolor{red}{Already there.}
+%\end{itemize}
+%
+%\item proportion should be proportional on page 15
+%\begin{itemize}
+%	\item \textcolor{red}{Fixed.}
+%\end{itemize}
+%\item datasets should be databases on page 17
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item bold m and u on page 18
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item Therefore needs a comma after it on page 18
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item equation 7 is not consistent with other equations on page 20
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item Convergence is stated on page 20. We have shown no issue of lack of convergence. Please address.
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
 \item Conclusion needs to be longer most likely or they will ask us to address this.
+\begin{itemize}
+	\item \textcolor{red}{I added a paragraph (in red), check it out!}
+\end{itemize}
+
 \item Reference summary of notation in main paper and make sure it is complete. 
-\item Refer the reader back to the main body of appendix with a reference so it's easier to find the section. 
-\item Provide an introduction to each subsection so it's more clear what you're presenting. Example: Section 7.2: In this section, we derive the full conditional distributions of section \ref{}. 
-\item Similar minor details in appendix that are marked up regarding punctuation/typos.
+\begin{itemize}
+	\item \textcolor{red}{I included the summary of notation (one for the general Gibbs sampler, one for the hashing notation) in the main body.}
+\end{itemize}
+
+%\item Refer the reader back to the main body of appendix with a reference so it's easier to find the section. 
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+%\item Provide an introduction to each subsection so it's more clear what you're presenting. Example: Section 7.2: In this section, we derive the full conditional distributions of section \ref{}. 
+%\begin{itemize}
+%	\item \textcolor{red}{See earlier response.}
+%\end{itemize}
+%\item Similar minor details in appendix that are marked up regarding punctuation/typos.
+%\begin{itemize}
+%	\item \textcolor{red}{Done.}
+%\end{itemize}
+
+\item \textcolor{red}{I moved the discussion of blocking and indexing the beginning of Section 4. It makes the most sense there, because that's when I introduce the computational speed ups.}
+
+\item \textcolor{red}{Do I need to provide open source software? I don't really want to be responsible for having a perfectly functioning R package at this stage.}
+
+\item \textcolor{red}{I got rid of both remarks. I figure if the information isn't relevant enough to be in the main flow of the paper, its probably not relevant enough to include.}
+
+\item \textcolor{red}{I have for the moment removed the psuedo-code for fabl from the appendix. It didn't seem necessary, and the notation was difficult in that format.}
+
 \end{enumerate}

### 2022-10-01 ba_resubmit/kundinger_resubmit.tex

@@
 However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.
+\textcolor{red}{To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.} Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
+This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
+Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy \textcolor{red}{when compared to \texttt{BRL}.}
+%
+%\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.
 Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}
-Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
-This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
-Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
-
-\textcolor{red}{BEKA/JERRY: Reviewer 2 wanted a specific reference to Sadinle in the abstract.
-I remember we actively decided not to do that.
-What do you think?} \textcolor{blue}{I added this back in.
-He only has one concern, so let's make sure to satisfy the person.}
+
 \end{abstract}
@@
 
-In many data analysis tasks, analysts seek to identify duplicate records across two databases.
+In many data analysis tasks, analysts seek to identify duplicate records across two datafiles.
 This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
-In this article, we consider bipartite record linkage, which merges two databases that contain duplications across but not within the respective databases.
+In this article, we consider bipartite record linkage, which merges two datafiles that contain duplications across, but not within, the respective datafiles.
 
@@
 The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice.
-In many situations, we know that there are no duplications within a database, meaning that each record in one file should be linked with at most one other record in the other file.
+In many situations, we know that there are no duplications within a datafile, meaning that each record in one file should be linked with at most one other record in the other file.
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
@@
 \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on databases with more than 100 records.
+\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on datafiles with more than 100 records.
 \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
-Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weight the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
+Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
 \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
-
-\textcolor{blue}{In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2),$ where $n_1$ and $n_2$ are the sizes of each database.
-The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
-In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
-The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.}
 
@@
 Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space.
-This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs and (2) compute the pairwise record comparisons over large datasets via parallel computing.
-\textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
- \textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking.
+This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+\textcolor{red}{These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy.
+In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.}
+
+%\textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
+%\textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking.
 This can be useful when there is no reliable blocking field available or one desires estimates of model parameters for the entire sample in question.
@@
 In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
-Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching.
+Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching.
 Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}.
@@
 
-\textcolor{red}{Consider two databases $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+Consider two datafiles $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
 Without loss of generality, denote files such that $n_1 \geq n_2$.
-We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.} In the context of bipartite matching, we assume that there are duplications across, but not within, each database.
-Under this framework, the set of matches across databases can be represented in two equivalent ways.
+We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.
+In the context of bipartite matching, we assume that there are duplications across, but not within, each datafile.
+Under this framework, the set of matches across datafiles can be represented in two equivalent ways.
 First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
@@
 Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks.
-More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that
+More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in datafile $\bm{X}_2$ such that
 \begin{align}
@@
 		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-		n_1 + j, & \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
+		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in datafile}\; \bm{X}_1. \\
 	\end{cases}
@@
 Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
-$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.
+$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
 
@@
 The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables.
-For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
-We then can set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+For numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
+We can then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
 Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$.
 Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
-Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choice of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.
+Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.
+
+%For more complex measurements, we can take into account partial agreement to more richly characterize the comparison;
 
@@
 %\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}
-
 %In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
@@
 
-Under the Fellegi Sunter model, we assume first, that comparison vectors are independent given their matching status, and second, that the match status of the record pairs are independent.
+The Fellegi Sunter model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
 This allows us to model the comparison data using mixture models in the following way:
@@
 where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match.
-These $\bm{m}$ and $\bm{u}$ parameters are then used to calculate the likelihood ratio
+The $\bm{m}$ and $\bm{u}$ parameters are used to calculate the likelihood ratio
 \begin{align}
@@
 This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
+
+%These unknown parameters are often estimated using the EM algorithm \cite{winkler_state_1999}.
 
@@
 
-To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model.
+To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
 He proposed the  beta distribution for bipartite matching, given by
@@
 This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
-This creates a dependency that makes the sampler inherently serial, which can be slow when working on large linkage tasks.
+This creates a dependency that makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
 
@@
 
-Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a match and $Z_j = n_1 + j$ indicates a non-match.
 In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper and flexible prior.
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior.
 We denote the fast beta prior as follows:
@@
 \end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),$$
-$\alpha_{\pi}$ and $\beta_{\pi}$ are fixed hyperparameters that encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}.$ Note the prior from \cite{wortman2019} can be viewed as a special case of the fabl prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
-
-\begin{remark}
-Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
-One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-\end{remark}
-
-\textcolor{blue}{Furthermore, note that it is common to encounter missing information in datasets used for record linkage.
-Whether information is missing in record $i \in \bf{X}_1$ or $j \in \bf{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
-With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
-To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}
-\textcolor{red}{(I don't love this right here, but I don't really know where else to put it.)}
+$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$
+This set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
+The parameter $\pi$ is given hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ that encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}.$
+
+% I don't really like using remarks. If its not relevante enough for the main body, we can just remove. 
+%\begin{remark}
+%One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
+Note the prior from \cite{wortman2019} can be viewed as a special case of the fabl prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
+%\end{remark}
 
@@
 
+Lastly, we note that it is common to encounter missing information in datasets used for record linkage.
+Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
+With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.
+
 %this independent prior for each \(Z_j\), induces a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains over the Gibbs sampler under \texttt{BRL}.
@@
 %The model can then be specified as follows:
+
+\begin{table}[t!]
+	\centering
+	\begin{tabular}[t!]{ll}
+		Symbol & Description \\
+		\hline
+		$\bm{X}_1, \bm{X}_2$ & datasets\\
+		$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
+		$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
+		$f \in 1, \ldots F$ & index over fields used for comparisons \\
+		$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
+		$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
+		$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
+		$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
+		$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
+		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
+		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
+		$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
+		\hline
+	\end{tabular}\caption{Summary of model notation}\label{table_notation_1}
+\end{table}
+
 With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model.
@@
 Let $\bm{u}$ parameters be defined analogously, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$.
-Then, using the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$, our model is:
+Then, our model is:
 \begin{align*}
-	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
 	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \\
@@
 	\end{cases} \\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).
 \end{align*}
+For convenience, a summary of notation is provided in Table \ref{table_notation_1}.
 %where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.
-
-\textcolor{red}{I actually don't like the $\Phi$ notation. It seems unnecessary. The only reason it's here is because Sadinle used it. What do you guys think?}
-\textcolor{blue}{I would leave it since they didn't complain unless you have a better idea.}
 
@@
 
-We initialize $\Phi$ from random draws from the prior distributions of $\bm{m}$ and $\bm{u}$, and initialize $\bm{Z}$ to reflect no matches across datasets, that is $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
-We then sample $\Phi$ from its full conditional:
-
+We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across datasets; that is, $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
+We then sample $\bm{m}$ and $\bm{u}$ from their full conditionals:
 $$\bm{m}_f|\Gamma, \bm{Z} \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
 $$\bm{u}_f|\Gamma, \bm{Z} \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
-where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$, and $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
+where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j = i)$, and $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j \neq i)$.
 
 Next, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
-$$p\left(Z_j^{(s+1)}  = i|\Gamma, \Phi, \bm{Z^{(s)}}\right) \propto
+$$p\left(Z_j^{(s+1)}  = i|\Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
@@
 	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
-\end{cases},$$
+\end{cases}$$
 where 
@@
 different types of errors, and minimizing posterior expected loss. We
-adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
+adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which
 \(\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\}\), with \(R\) representing the
@@
 %$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
-In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the intuitive decision rule
+In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
 $$\hat{Z}_j =\begin{cases} 
@@
 
-To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
+The scale of linkage tasks possible within the Fellegi Sunter framework is limited by (1) the memory costs of storing $n_1 \times n_2$ comparison vectors for every pair of records across the two datafiles, and (2) the speed of the linkage algorithm over those comparison vectors.
+One common solution to these challenges is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{christen2019data}.
+In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
+The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
+However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
+
+Another technique is indexing, in which the modeller decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
+For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like first name, or only pairs that exactly match on a specified number of fields.
+However, this kind of indexing requires knowledge  of the application and invites room for human error \citep{murray2016probabilistic}.
+
+With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage without the drawbacks of blocking and indexing.
+First, we propose hashing methods that allow us to compute sufficient statistics which reduce the computational complexity of the Gibbs sampler.
+Second, we introduce storage efficient indexing (SEI), which reduces the memory costs associated with unlikely matches.
+
+%To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
 The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
@@
 
+Following insight from \cite{enamorado2019using}, we recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
+To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns.
+\(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
+In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the original vectors themselves.
+Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
+Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
+See Appendix \ref{app:ohe} for details.
+
 First, we hash record pairs of the same agreement pattern to unique integer values. \cite{enamorado2019using} accomplished this efficiently through the hashing function
-
 $$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}.$$
-
 This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
@@
 $$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
-where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ \textcolor{red}{is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$}.
-This representation is useful because it allows us to easily compute sufficient statistics
-$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||,$$
-which is the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, and 
-$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p},$$
-which is the number of comparison vectors in $\Gamma$ that exhibit pattern $p$.
-For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
-Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.
+where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$.
+This representation is useful because it allows us to easily compute two sets of sufficient statistics.
+First, we compute the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, given by
+$$N_{j_p} = \sum_{i=1}^{n_1} I((i, j) \in h_p) = |r_{j_p}|.$$
+Second, we compute the number of comparison vectors in $\Gamma$ that exhibit pattern $p$, given by
+$$H_p =  \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} I((i, j) \in h_p) = \sum_{j = 1}^{n_2} N_{j_p}.$$
+We collect these sufficient statistics in the sets $\mathcal{N} = \left\{\{N_{j_p}\}_{j=1}^{n_2}\right\}_{p=1}^P$ and $\mathcal{H} = \{H_{p}\}_{p=1}^P.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
+Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}, \mathcal{H}\}$ for posterior inference.
 
@@
 
-The hashing procedure described above considerably reduces the memory needed to store the comparison information.
-That is, instead of storing \(n_1 \times n_2\) comparison vectors, we only store the \(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities relating record pairs to those vectors.
-However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large databases.
+%In practice, it is not feasible to make all-to-all record comparisons because creating and storing $n_1 \times n_2$ comparison vectors is computationally infeasible for large datafiles.
+The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
+In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
+The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
+However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
+%
+%The hashing procedure described above considerably reduces the memory needed to store the comparison information.
+That is, instead of storing \(n_1 \times n_2\) comparison vectors, we only store the \(P\) unique vectors, and \(n_1 \times n_2\) scalar quantities relating record pairs to those vectors.
+However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large datafiles.
 Worse, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
-
-To address this issue, we propose storage efficient indexing (SEI).
+%
+%To address this issue, we propose storage efficient indexing (SEI).
 In standard indexing, one decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
-For example, one might only consider pairs with a certain similarity score on a field deemed to be important (like first name), or only pairs that exactly match on a specified number of fields.
-While generally chosen to be quite loose, establishing these criteria requires knowledge of the problem and invites room for human error.
-
-With SEI, however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks.
-Observe that all records $i \in \bm{X}_1$ that share agreement pattern $p$ with $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\).
-Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
+For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like first name, or only pairs that exactly match on a specified number of fields.
+this kind of indexing requires knowledge of the application of the application and invites room for human error \cite{murray2016probabilistic}.
+
+With storage efficient indexing (SEI), we can calculate the above sufficient statistics for all $n_1 \times n_2$ record pairs, but greatly reduce the memory costs associated with unlikely matches, allowing us to conduct all-to-all comparisons for substantially larger linkage tasks.
+Observe that all records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$.
+Therefore, these records have the same probability to be identified as a link when sampling \(Z_j\).
+Thus, we know that records $i \in r_{j_p}$  such that $N_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
 We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
-Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$.
-instead of storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
-\textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for plausible matches.
-Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results.} Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$.
-For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
+Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested list $\mathcal{R}^{\text{SEI}}$.
+Instead of storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
+In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for plausible matches.
+Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results.
+Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}, \mathcal{H}\}$.
+
+%For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
 
@@
 Summary statistics from each pairwise chunk comparison can then be easily synthesized to recover sufficient statistics for the larger linkage task.
-Thus, the combination of hashing, SEI, and partitioning
-allows us to conduct linkage tasks over much larger datasets.
+Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger datasets.
 
@@
 Although sampling \(Z_j\) from the full conditional provided in Section \ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
-This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
+This is because sampling a value from $n_1$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
 To speed up computation, we break this sampling step into two.
-First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link.
-Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in \bm{X}_2\).
-More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\).
-Then,
+First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link, according to (\ref{eqn:gibbs1}).
 \begin{align}
 	\label{eqn:gibbs1}
-		P\left(h\left(Z_j^{(s+1)}\right) = p \mid\tilde{\Gamma} \Phi, \bm{Z^{(s)}}\right) \propto
+	P\left(\left(Z_j^{(s+1)}, j\right) \in h_p \mid \tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
 	\begin{cases} 
-			w_{p}\times H_{j_p},  & p \leq P; \\
-			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}}, &   p = P + 1; \\
+		w_{p}\times N_{j_p},  & p \leq P; \\
+		n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, &   p = P + 1. \\
 	\end{cases}
 \end{align}
-	Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns with (\ref{eqn:gibbs1}) above.
-After all iterations are complete, we can back-fill the records uniformally among records of the sampled agreeement pattern through (\ref{eqn:gibbs1}) below:
+%More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,
+%	\begin{align}
+%		\label{eqn:gibbs1}
+%		P\left(h\left(Z_j^{(s+1)}\right) = h_p \mid\tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+%		\begin{cases} 
+%			w_{p}\times H_{j_p},  & p \leq P; \\
+%			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, &   p = P + 1; \\
+%		\end{cases}
+%	\end{align}
+Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns.
+After all iterations are complete, we can back-fill the records uniformally among records of the sampled agreement pattern through (\ref{eqn:gibbs2}).
 \begin{align}
 		\label{eqn:gibbs2}
-		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
-			\frac{1}{H_{j_p}}, & (i, j) \in h_p; \\
+		P\left(Z_j^{(s+1)} = i \mid \left(Z_j^{(s+1)}, j\right) \in h_p \right) = \begin{cases} 
+			\frac{1}{N_{j_p}}, & (z_j, j) \in h_p; \\
 			0, & \text{otherwise.} \\
@@
 	\end{align} 
+%	\begin{align}
+%		\label{eqn:gibbs2}
+%		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
+%			\frac{1}{H_{j_p}}, & (i, j) \in h_p; \\
+%			0, & \text{otherwise.} \\
+%		\end{cases}
+%	\end{align}
 
@@
 These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
-We provide a summary of the \texttt{fabl} method through psuedocode in Appendix \ref{app:fabl-psuedocode}.
-\textcolor{blue}{The computational complexity of \texttt{fabl} is given in Lemma \ref{lemma:fabl}.}
+We provide a summary of the notation used for the hashing procedure in Table \ref{table_notation_2}.
+The computational complexity of \texttt{fabl} is given in Lemma \ref{lemma:fabl}.
+%, and a summary of the \texttt{fabl} method through psuedocode in Appendix \ref{app:fabl-psuedocode}
+
+	\begin{table}[t!]
+		\centering
+		\begin{tabular}[t!]{ll}
+			Symbol & Description \\
+			\hline
+			$h_p$ & one hot encoding of agreement pattern $p$ \\
+			$(i,j) \in h_p$ & comparison vector between records $i \in \bm{X}_1$ and $i \in \bm{X}_2$ exhibits pattern $p$ \\
+			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+			$N_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+			$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+			$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+			$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
+			\hline
+		\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_2}
+	\end{table}
 
@@
 
-	\begin{remark}
-	We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
+%	\begin{remark}
+%	We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
 \texttt{BRL} would see considerable speed up through the procedure, but computation time would still scale with the size of the larger dataset.
-	\end{remark}
+%	\end{remark}
 
@@
 	\begin{lemma}
-		Recall that $n_1$ and $n_2$ are the number of records in $\bm{X}_1$ and $\bm{X}_2$ respectively, $F$ is the number of fields used for comparisons across records, and $P$ is the number of patterns that comparison vectors can exhibit.
+		Recall that $n_1$ and $n_2$ are the number of records in $\bm{X}_1$ and $\bm{X}_2$, respectively.
+Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors can exhibit.
 We assume $B$ processors available for parallelization.
@@
 	\begin{proof}
-		To prove the computational complexity, we consider two steps --- constructing the comparison vectors and the Gibbs sampler. 
-		The computational complexity of all pairwise comparisons across the two databases $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
-The hashing procedure for all pairwise comparisons is also of complexity $O(F n_1 n_2)$.
+		To prove the computational complexity, we consider two steps: constructing the comparison vectors and the Gibbs sampler.
+The computational complexity of all pairwise comparisons across the two datafiles $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
+The hashing procedure for all pairwise comparisons is also $O(F n_1 n_2)$.
 With $B$ processors available, we can split these computations across $B$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
@@
 However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
-The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$.
+The complexity of updating $\bm{Z}$ sequentially at the record level is $O(n_1 n_2)$.
 With hashing, we split this sampling into two steps.
-First we sample the agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
-Thus the complexity of sampling $\bm{Z}$ is $O(n_2 P)$.
-Since $P << n_1$ in all reasonable applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
+First, we sample the agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
+Thus, the complexity of sampling $\bm{Z}$ is $O(n_2 P)$.
+Since $P << n_1$ in most applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
 		In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
@@
 	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-We use five binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}).
-$\bm{u}$ parameters are chosen to emulate what is seen in practice for first name, last name, day of birth, and month of birth.
+We use five binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
+The $\bm{u}$ parameters are chosen to emulate what is seen in practice for first name, last name, day of birth, and month of birth.
 The parameters for year would vary largely by context, so we choose them here to adjust the desired difficulty of the linkage task.
@@
 	We note here that the implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
-In contrast, we use non-optimized code written in R for \texttt{fabl}.
+In contrast, we use non-optimized code written in \texttt{R} for \texttt{fabl}.
 While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
@@
 
-	In Figure \ref{fig:speed1}, where we increase both $n_1$ and $n_2$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data.
-In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of only $\bm{X}_2$).
+	In Figure \ref{fig:speed1}, where we increase both $n_1$ and $n_2$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger data.
+In particular, run-time for \texttt{BRL} grows quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} grows linearly (in the size of only $\bm{X}_2$).
 
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_1$ and $n_2$.
-We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:speed1}
-		\end{center}
-	\end{figure}
-
-	The above discussion suggests that for fixed \(n_2\), computation time for \texttt{fabl} should remain mostly constant with growing \(n_1\).
-In the simulation results displayed in Figure \ref{fig:speed2} fixing \(n_2 = 500\), we see linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, with much more static run-time under \texttt{fabl}.
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$.
+We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
+		\end{center}
+	\end{figure}
+
+	%The above discussion suggests that for fixed \(n_2\), computation time for \texttt{fabl} should remain mostly constant with growing \(n_1\).
+
+	In Figure \ref{fig:speed2}, where we fix \(n_2 = 500\), we see linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, and much more static run-time under \texttt{fabl}.
 The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-\textcolor{red}{To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.}
+To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.
 
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_1$, and $n_2$ fixed at 500.
-We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:speed2}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with increasing $n_1$ and $n_2$ fixed at 500.
+We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
 		\end{center}
@@
 Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$.
-Precision is proportion the of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
+Precision is the proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
 The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
@@
 Lastly, we also examine the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)/n_2$.
-\textcolor{red}{To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.}
+To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.
 
@@
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We again use flat priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
+We again flat priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
 
@@
 This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
-However, after investigating the posterior distributions of the $\bm{m}$ and $\bm{u}$ parameters, we noticed that posterior distributions of several levels of the m and u parameters are nearly identical, suggesting that these levels are unnecessary.
+However, conducting record linkage, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical, suggesting that these levels were unnecessary.
 
@@
 
-	Therefore we re-ran our analysis with fewer agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
+	Therefore, we re-ran our analysis with fewer agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
 Among the 216 possible agreement patterns, 159 are realized in the data.
-With this revised comparison specification, fabl ran in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
+With this revised comparison specification, \texttt{fabl} ran in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
 The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure \ref{fig:m-and-u}.
-Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
-
-	For completeness, we note that the more detailed comparison vector produced a run time of 240 seconds for BRL, and 261 seconds for fabl.
-Apparently, the number of patterns was sufficiently many that the computational savings from fabl does not overcome the inherent speed differences of C as opposed to R.
+\textcolor{red}{Estimates of $\bm{u}$ are indistinguishable, and thus omitted.} Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+
+	For completeness, we note that linkage with the more detailed comparison vectors required240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
+Apparently, the number of patterns was sufficiently many that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
 
@@
 The initial survey began in 1982, with follow-up surveys taken approximately every five years.
-For this study, we seek to link records over the $n_1 = 20485$ records from 1982 to the $n_2 = 17466$ records from 1989.
-\textcolor{blue}{Within each database there are no duplications, however, there can be duplications across the databases.
+As such, patients are surveyed at most once in a given year, and many patients are surveyed across multiple years.
 In addition, patients can either drop out of the study, pass away, or enter as new patients.
-Hence, the assumptions of our model hold for this study.}
-%	\textcolor{red}{In each iteration of the study, some participants are added to the lists, and others are lost due to loss of contact or death.} 
+Hence, the assumptions of our model hold for this study.
+We seek to link records over the $n_1 = 20485$ records from 1982 to the $n_2 = 17466$ records from 1989.
 The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
-However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone databases.
+However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone datafiles.
 
@@
 Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
-However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and compress information through storage efficient indexing; the resulting data object is just 10 MB, approximately 0.1\% of what is required for the raw comparisons.
+However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and compress information through storage efficient indexing.
+The resulting data object is just 10 MB, approximately 0.1\% of what is required for the raw comparisons.
 Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.
@@
 	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-As shown in  \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
 Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
-Traceplots indicate quick convergence, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
+Traceplots do not suggest convergence issues, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
 
@@
 
-	In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that scales to large data sets. We have proven that the proposed hashing method
-	and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset.
+	In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that extends the work of \cite{sadinle_bayesian_2017} to scale to large data sets.
+We have proven that the proposed hashing method and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset.
 This makes \texttt{fabl} computationally advantageous in many linkage scenarios, particularly when one datafile is substantially smaller than the other.
+We have also shown that storage efficient indexing, in tandem with hashing, greatly reduces the memory costs required for all-to-all comparisons, giving practictioners an option for larger record linkage tasks without the use of blocking or indexing.
 We have demonstrated the speed and accuracy of \texttt{fabl} by replicating a simulation study and a case study in \cite{sadinle_bayesian_2017}, and through an additional case study that is computationally infeasible under \texttt{BRL}.
-Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
+
+	\textcolor{red}{Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, the computing all $n_1 \times n_2$ record pairs can still be infeasible for larger linkage tasks.
+Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
+Due to the quadratic nature of the comparison space, this computation time would grow quickly with the size of the linkage task, and would be infeasibly slow when dealing with millions of records.
+Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task.
+Thus in future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to arbitrarily large linkage tasks.}
+
+	%Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
 
@@
 
-	\hypertarget{app:glossary}{%
-	\subsection{Summary of Notation}\label{app:glossary}}
-
-		\begin{table}[h!]
-		\centering
-		\begin{tabular}[h!]{ll}
-			\hline
-			Symbol & Description \\
-			\hline
-			$\bm{X}_1, \bm{X}_2$ & datasets\\
-			$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
-			$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
-			$f \in 1, \ldots F$ & index over fields used for comparisons \\
-			$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
-			$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
-			$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
-			$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
-			$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
-			$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
-			$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-			$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
-			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-			$h_p$ & one hot encoding of agreement pattern $p$ \\
-			$H_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-			$H_p$ & number of comparison vectors that exhibit pattern $p$ \\
-			\hline
-		\end{tabular}
-	\end{table}
+%	\hypertarget{app:glossary}{%
+%	\subsection{Summary of Notation}\label{app:glossary}}
+%	
+%		\begin{table}[h!]
+%		\centering
+%		\begin{tabular}[h!]{ll}
+%			\hline
+%			Symbol & Description \\
+%			\hline
+%			$\bm{X}_1, \bm{X}_2$ & datasets\\
+%			$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
+%			$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
+%			$f \in 1, \ldots F$ & index over fields used for comparisons \\
+%			$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
+%			$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
+%			$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
+%			$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
+%			$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
+%			$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
+%			$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
+%			$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
+%			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+%			$h_p$ & one hot encoding of agreement pattern $p$ \\
+%			$H_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+%			$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+%			$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+%			$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
+%			\hline
+%		\end{tabular}
+%	\end{table}
 
@@
 
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{f\ell}$, we have
+We provide detailed derivations of the full-conditionals provided in Section \ref{gibbs_sampling}.
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
+For a particular $m_{fl}$, we have
 \begin{align*}
 	p(m_{fl}| \Gamma, \bf{Z}) &\propto \prod_{i=1}^{n_1} \prod_{j=1}^{n_2} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
-	&= m_{fl}^{\alpha_{fl}(\bm{Z}) - 1}
+	&= m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
 \end{align*}
@@
 Analogous procedures lead to the posterior distribution $p(u_{fl}| \Gamma, \bf{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
-Thus for the vectors of parameters $\bf{m}_f$ and $\bf{u}_f$, we have
-
-$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
+Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
+
+$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
 $$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})).$$
 
-In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihooh relevant to a particular record $j \in \bm{X}_2$.
-Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
-Thus we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have
+In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihood relevant to a particular record $j \in \bm{X}_2$.
+Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
+Thus, we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
 \begin{align*}
@@
 Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\). We have
-%As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with
 \begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+	p(\Gamma_{.j}|Z_j = z_j, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
 	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
@@
 		w_{z_j, j}  & z_j \leq n_1; \\
-		1 &  z_j  = n_1 + j \\
-	\end{cases},\\
+		1 &  z_j  = n_1 + j, \\
+	\end{cases}\\
 \end{align*}
@@
 \begin{align*}
-	p\left(Z_j^{(s+1)}|\Gamma_{.j}, \Phi, \bm{Z^{(s)}}, \pi\right) &\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \Phi) P(Z_j^{(s+1)}|\pi) \\
+	p&\left(Z_j^{(s+1)}|\Gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
+	&\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
 	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
@@
 \begin{align*}
-	p\left(Z_j^{(s+1)}  = i| \Gamma, \Phi, \bm{Z^{(s)}}\right) &= \int_{\pi}p\left(Z_j^{(s+1)} | \Gamma, \Phi, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+	p&\left(Z_j^{(s+1)}  = i| \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
+	&= \int_{\pi}p\left(Z_j^{(s+1)} | \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
 	&= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
@@
 	&\propto w_{z_j, j}I(z_j\leq n_1) + n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_1 + j) \\
-	&\propto 
+	&=
 	\begin{cases} 
@@
 
-	\hypertarget{app:fabl-psuedocode}{%
-		\subsection{Summary of Fast Beta Linkage Method}\label{app:fabl-psuedocode}}
-
-	\begin{algorithm}[h!]
-		\begin{algorithmic}[1]
-
-			\Procedure{Hashing and Preprocessing}{}
-			\State Construct and enumerate set of unique patterns $\mathcal{P}$ from $F$ and $\{L_f\}$.
-			\State Partition files $\bm{X_1}$ and $\bm{X_1}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$.
-			\For{each $n$, $m$}
-			\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$.
-			\State Hash records to $\mathcal{R}_{nm}$ and calculate summary statistics $\mathcal{H}_{nm}$.
-			\State Use SEI to reduce memory usage; $\mathcal{R}_{nm} \to \mathcal{R}_{nm}^{\text{SEI}}$.
-			\EndFor
-			\State Synthesize results across pairings to get $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$.
-			\EndProcedure
-
-			\Procedure{Gibbs Sampling}{}
-			\State Initialize $\Phi$, and $Z$ parameters.
-
-			\For{$t \in \{1, \ldots, T\}$} 
-			\State Sample $\Phi^{t+1}|Z^{t}, \tilde{\Gamma}$.
-			\State Sample $h\left(Z^{t+1}\right)|\Phi^{t+1}, \tilde{\Gamma}$.  \Comment{Sample agreement pattern, not record}
-			\EndFor
-			\State Sample $Z | h(Z), \tilde{\Gamma}$. \Comment{Fills in record label based on agreement pattern}
-
-			\EndProcedure
-
-		\end{algorithmic}
-	\end{algorithm}
-	\clearpage
+%	\hypertarget{app:fabl-psuedocode}{%
+%		\subsection{Summary of Fast Beta Linkage Method}\label{app:fabl-psuedocode}}
+%	
+%	\textcolor{red}{I am getting rid of the notation $h(Z_j)$ for the agreement pattern between $j$ and its potential match.
+Instead, I am keeping notation that already exists, saying $(Z_j, j) \in h_p$.
+With this change, I don't see a good way to write lines 15 and 17.
+For that reason, I think I should remove this pseudo-code; it doesn't seem necessary, and invites criticism.}
+%	
+%	\begin{algorithm}[h!]
+%		\begin{algorithmic}[1]
+%			
+%			\Procedure{Hashing and Preprocessing}{}
+%			\State Construct and enumerate set of unique patterns $\mathcal{P}$ from $F$ and $\{L_f\}$.
+%			\State Partition files $\bm{X_1}$ and $\bm{X_2}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$.
+%			\For{each $n$, $m$}
+%			\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$.
+%			\State Hash records to $\mathcal{R}_{nm}$ and calculate summary statistics $\mathcal{H}_{nm}$.
+%			\State Use SEI to reduce memory usage; $\mathcal{R}_{nm} \to \mathcal{R}_{nm}^{\text{SEI}}$.
+%			\EndFor
+%			\State Synthesize results across pairings to get $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$.
+%			\EndProcedure
+%			
+%			\Procedure{Gibbs Sampling}{}
+%			\State Initialize $\bm{m}$, $\bm{u}$, and $Z$ parameters.
+%			
+%			\For{$t \in \{1, \ldots, T\}$} 
+%			\State Sample $\bm{m}^{t+1}|Z^{t}, \tilde{\Gamma}$ and  $\bm{u}^{t+1}|\bm{Z}^{t}, \tilde{\Gamma}$
+%			\State Sample $h\left(\bm{Z}^{t+1}\right)|\bm{m}^{t+1}, \bm{u}^{t+1}, \tilde{\Gamma}$.  \Comment{Sample agreement pattern, not record}
+%			\EndFor
+%			\State Sample $\bm{Z} | h(\bm{Z}), \tilde{\Gamma}$. \Comment{Fills in record label based on agreement pattern}
+%			
+%			\EndProcedure
+%			
+%		\end{algorithmic}
+%	\end{algorithm}
+%	\clearpage
 
@@
 
-	\texttt{fabl} makes use of one-hot encodings to aid in vectorized computations.
+	As described in Section \ref{sec:efficiency}, \texttt{fabl} makes use of one-hot encodings to aid in vectorized computations.
 For $\gamma_{ij}^f$ with $L_f$ levels, define $e_{ij}^f$ to be an $L_f \times 1$ vector.
-When $\gamma_{ij}^f = \ell$, we set the $\ell^{th}$ element of $e_{ij}^f$ to be 1, and set the other $L_f - 1$ elements of $e_{ij}^f$ to be 0.
+When $\gamma_{ij}^f = l$, we set the $l^{th}$ element of $e_{ij}^f$ to be 1, and set the other $L_f - 1$ elements of $e_{ij}^f$ to be 0.
 We then concatenate the $e_{ij}^f$ for all $f \in \{1, \ldots, F\}$, resulting in the one-hot encoded comparison vector $e_{ij}$ of length $\sum_{f=1}^F L_f$.
@@
 		\begin{tabular}[h!]{llll}
-			\hline
 			Last Name & First Name & DOB & City \\
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
-			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_1$ and $n_2$.
-We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:app-speed1}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$.
+We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
 		\end{center}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
-			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_1$, and $n_2$ fixed at 500.
-We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:app-speed2}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with increasing $n_1$ and $n_2$ fixed at 500.
+We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-			\caption{Traceplot for number of matches found across datasets in El Salvador case study}\label{fig:overlap_trace}
+			\caption{Traceplot for number of matches found across datasets in El Salvador case study.} \label{fig:overlap_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-			\caption{Traceplot for m parameter in El Salvador case study}\label{fig:m_trace}
+			\caption{Traceplot for m parameter in El Salvador case study.} \label{fig:m_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-			\caption{Traceplot for u parameter in El Salvador case study}\label{fig:u_trace}
+			\caption{Traceplot for u parameter in El Salvador case study.} \label{fig:u_trace}
 		\end{center}

### 2022-10-04 ba_resubmit/response_to_reviewers.tex
added red comment
final edits before Jerry

@@
 	\maketitle
-	\textbf{We thank the Editor and reviewers for their encouraging assessment and constructive
-	feedback.}
+	We thank the Editor and reviewers for their encouraging assessment and constructive
+	feedback.
 
@@
 	\reply{%
-\textbf{We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
+We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper.
 We have revised our paper in the abstract and introduction as to make our contributions more clear in relation to prior work.
-Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
+Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.
 	}
@@
 	\reply{
-\textbf{We appreciate these comments, and have done our utmost to address the ones listed by yourself and the AE.
-Furthermore, we have carefully gone through the paper ourselves, where we have corrected other issues with notation, typos, and various inconsistencies.} \todo
+We appreciate these comments, and have done our utmost to address the ones listed by yourself and the AE.
+Furthermore, we have carefully gone through the paper ourselves, where we have corrected other issues with notation, typos, and various inconsistencies.
 	}
@@
 	\reply{
-\textbf{$I_{obs}()$ is now defined. We have defined $n_{12}(Z)$ at its first occurrence (and removed the error with $D$). We have checked for other notational inconsistencies.
-	}
+$I_{obs}()$ is now defined. We have defined $n_{12}(Z)$ at its first occurrence (and removed the error with $D$). We have checked for other notational inconsistencies.
 	}
@@
 	\reply{
-	\textbf{We have fixed the inconsistency, using $n_1$ and $n_2$ throughout the paper.}
+	We have fixed the inconsistency, using $n_1$ and $n_2$ throughout the paper.
 
@@
 	\reply{
-	\textbf{The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1.$ We agree that the latter expression is more clear.
-This has been updated in our revision.}
+	The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1.$ We agree that the latter expression is more clear.
+This has been updated in our revision.
 	}
@@
 	\reply{
-	\textbf{We have updated the Lemma to include the definition of $B$ as well as other variables in the Lemma.}
+	We have updated the Lemma to include the definition of $B$ as well as other variables in the Lemma.
 	}
@@
 	\reply{%
-\textbf{We agree with this comment, and we have removed this example as it does not add to the paper.}
+We agree with this comment, and we have removed this example as it does not add to the paper.
 }
@@
 \reply{%
-	\textbf{Thank you for the opportunity to submit a revision and improve the quality of our paper.
-We have provided detailed responses to the reviewers' comments below, which prompted significant changes in our paper.} \todo
+	Thank you for the opportunity to submit a revision and improve the quality of our paper.
+We have provided detailed responses to the reviewers' comments below, which prompted significant changes in our paper.
 }
@@
 	\reply{%
-	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017).
+	We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017).
 Our revised paper clarifies our novel contributions, improves upon the exposition of the writing, provides practical guidance for users, and expands the simulations.
-We expand on these points further below.} \todo
+We expand on these points further below.
 	}
@@
 \reply{%
-	We have provided a summary of notation in Appendix 7.1.
+	We have provided summaries of notation in Tables 1 and 2.
 
-	\textcolor{blue}{I would recommend a table for notation of the model and a table for notation of your hashing method.
-Why?
-It will be strange to reference it all together.
-I would try and put these in the main body if possible as it will really help the reader.
-You could also write a graphical model with notation of your model to help the reader.
-These are just some thoughts to think about.
-(I'm going to leave this up to you Brian if you want to abandon the able as it's quite annoying).}
-
-	\textcolor{red}{Done}
 }
@@
 \reply{%
-	\textbf{This has been defined.}
+	This has been defined.
 }
@@
 \reply{
-	\textbf{Thank you for the interesting comment/question.
+	Thank you for the interesting comment/question.
 In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper.
 We have removed the confusing text mentioned by the reviewer.
-We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue; however, this is beyond the scope of this paper.}
+We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue; however, this is beyond the scope of this paper.
 }
@@
 \reply{%
-	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.} 
+	Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.
 }
@@
 \reply{%
-	\textbf{We have removed the $\Phi$ notation, using the parameters $\bm{m}$ and $\bm{u}$ throughout the paper.}
+	We have removed the $\Phi$ notation, using the parameters $\bm{m}$ and $\bm{u}$ throughout the paper.
 }
@@
 \reply{%
-	\textbf{The expression is now corrected.}
+	The expression is now corrected.
 }
@@
 \reply{%
-	\textbf{The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
+	The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior.
 The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$.
 We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$.
-We have updated our explanation to make this more clear.}
+We have updated our explanation to make this more clear.
 }
@@
 \reply{%
-	\textbf{This has been corrected.}
+	This has been corrected.
 }
@@
 \reply{%
-	\textbf{We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$.
-We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.}
+	We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.
 }
@@
 \reply{%
-	\textbf{We have revised the presentation of the Gibbs sampler based on this feedback.
-The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the Appendix.}
+	We have revised the presentation of the Gibbs sampler based on this feedback.
+The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the Appendix.
 }
@@
 \reply{%
-	\textbf{The lemma now includes assumptions and definitions for all variables used within the statement.}
+	The lemma now includes assumptions and definitions for all variables used within the statement.
 }
@@
 \reply{%
-	\textbf{This has been removed.}
+	This has been removed.
 }
@@
 \reply{%
-	\textbf{This has been fixed.}
+	This has been fixed.
 }
@@
 	\reply{%
-	\textbf{Thank you for your positive feedback regarding our paper.
+	Thank you for your positive feedback regarding our paper.
 We have clarified that we are motivation by the work of Sadinle and computational considerations.
-We have tried to make this more clear in both the abstract, introduction, and conclusion.}}
-
-\newpage
-
-\section{Additional Points Addressed}
-
-We highlight some additional points addressed in our revision. \todo
-
-\textcolor{blue}{Brian: These are just some suggestions that I have, so we can chat through these in case you don't agree!
-I have these marked up on a printed copy of the paper itself to help you more easily make the changes.
-Please note I have not made these changes, but I'm marking them on a marked up paper so you can go back through and make changes.
-It's likely that there may be more errors, so please go over things very carefully for all errors as the content is all there, we just have to weed out these tedious details.}
-
-\textcolor{Red}{For the purpose of our revisions, I commented out the comments about minor typos. The comments that remain are things I want you to read so we can discuss.}
-
-\begin{enumerate}
-\item We have added introductory sentences to each section/subsection to provide more guidance to the reader. 
-\begin{itemize}
-	\item \textcolor{red}{I originally had sentences like this, and Jerry removed them. I also like it better without these introductory sentences, and don't find them necessary.}
-\end{itemize}
-
-\item The word databases is used throughout instead of files or lists. 
-\begin{itemize}
-	\item \textcolor{red}{Sadinle uses the words datafile and file. I'll remove database, and do that instead.}
-\end{itemize}
-
-%\item We have checked for the proper use of commas throughout the paper. 
-%\begin{itemize}
-%	\item \textcolor{red}{Done}
-%\end{itemize}
-
-%\item We have re-written our proposed work on page 2 to make it more clear our contributions to the literature as it was missing our improvements over existing work in terms of speed with minimal loss in accuracy as well as providing two case studies.
-One case study is from the original paper of our extension, while another case study illustrates our ability to scale to larger databases.
-We provide a more comprehensive discussion regarding our work.
-%\begin{itemize}
-%	\item \textcolor{red}{I don't think that the proposed changes made the contribution more clear. I did other edits, adding the specifics }
-%\end{itemize}
-
-\item We have broken up sentences that are too long.
-\begin{itemize}
-	\item \textcolor{red}{We disagree on some of these. For all of the longer sentences that remain, I believe they are clearer together than broken up.}
-\end{itemize}
+We have tried to make this more clear in both the abstract, introduction, and conclusion.}
 
@@
 %\end{itemize}
-
-\item We have defined the parameters $m$ and $u$ on page 4. 
-\begin{itemize}
-	\item \textcolor{red}{I originally had the precise definition of $m$ and $u$ at this spot, giving the $m_{fl} = p(\gamma_{ij}^f = l | Z_{j} = i)$ definitions.
-However, technically, FS do not specify that comparison vectors be thresholded to discrete values, so this definition technically doesn't make sense here.
-I checked, and Sadinle defines the $m$ and $u$ parameters during the FS review section the same way that I do.}
-
-	\textcolor{red}{I agree that reviewers might take note of this, so I could add the following sentence.}
-
-	\textcolor{red}{"When using comparison vectors with discrete agreement levels as described above, the $\bm{m}$ parameter takes the form $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$.
-The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$."}
-\end{itemize}
-
-\item One page 5, section 2.2, we have made it more clear why our extension is warranted and needed in the literature. 
-\begin{itemize}
-	\item \textcolor{red}{Done.}
-\end{itemize}
-
-\item We have removed confusing language, such as "compactful representation of" on page 5.
-\begin{itemize}
-	\item \textcolor{red}{Done}
-\end{itemize}
-
-\item We have checked that fast beta linkage is represented by fabl throughout.
-\begin{itemize}
-	\item \textcolor{red}{Done. However, I think "fast beta prior" is fine, and I have left that as is.}
-\end{itemize}
-
-\item \textcolor{blue}{See page 10 for a long list of comments.}
-\begin{itemize}
-	\item \textcolor{red}{I don't understand what this is.}
-\end{itemize}
 
@@
 %\end{itemize}
-\item Appendices should be done as A, B, C or the convention in BA (please check). 
-\begin{itemize}
-	\item \textcolor{red}{None of the articles I saw on the BA homepage had appendices. Some had "Supplementary Material," and those had numbered headings. So I'm leaving it as is for now. }
-\end{itemize}
 
-%\item There is a comma present that needs to be removed in Figure 2. 
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-%
-%\item Page 15, provide a reference to the Appendix.
-%\begin{itemize}
-%	\item \textcolor{red}{Already there.}
-%\end{itemize}
-%
-%\item proportion should be proportional on page 15
-%\begin{itemize}
-%	\item \textcolor{red}{Fixed.}
-%\end{itemize}
-%\item datasets should be databases on page 17
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-%\item bold m and u on page 18
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-%\item Therefore needs a comma after it on page 18
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-%\item equation 7 is not consistent with other equations on page 20
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-%\item Convergence is stated on page 20. We have shown no issue of lack of convergence. Please address.
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-\item Conclusion needs to be longer most likely or they will ask us to address this.
-\begin{itemize}
-	\item \textcolor{red}{I added a paragraph (in red), check it out!}
-\end{itemize}
-
-\item Reference summary of notation in main paper and make sure it is complete. 
-\begin{itemize}
-	\item \textcolor{red}{I included the summary of notation (one for the general Gibbs sampler, one for the hashing notation) in the main body.}
-\end{itemize}
-
-%\item Refer the reader back to the main body of appendix with a reference so it's easier to find the section. 
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-%\item Provide an introduction to each subsection so it's more clear what you're presenting. Example: Section 7.2: In this section, we derive the full conditional distributions of section \ref{}. 
-%\begin{itemize}
-%	\item \textcolor{red}{See earlier response.}
-%\end{itemize}
-%\item Similar minor details in appendix that are marked up regarding punctuation/typos.
-%\begin{itemize}
-%	\item \textcolor{red}{Done.}
-%\end{itemize}
-
-\item \textcolor{red}{I moved the discussion of blocking and indexing the beginning of Section 4. It makes the most sense there, because that's when I introduce the computational speed ups.}
-
-\item \textcolor{red}{Do I need to provide open source software? I don't really want to be responsible for having a perfectly functioning R package at this stage.}
-
-\item \textcolor{red}{I got rid of both remarks. I figure if the information isn't relevant enough to be in the main flow of the paper, its probably not relevant enough to include.}
-
-\item \textcolor{red}{I have for the moment removed the psuedo-code for fabl from the appendix. It didn't seem necessary, and the notation was difficult in that format.}
-
-\end{enumerate}
 \end{document}

### 2022-10-04 ba_resubmit/kundinger_resubmit.tex
final edits before Jerry

@@
 However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-\textcolor{red}{To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.} Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
+To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
+Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
 This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
-Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy \textcolor{red}{when compared to \texttt{BRL}.}
+Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
 %
@@
 This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-\textcolor{red}{These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy.
-In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.}
+These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy.
+In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.
 
@@
 
-%\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}
-%\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}
-%In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
-In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
-The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
 %
@@
 where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match.
+When using comparison vectors with discrete agreement levels as described above, the $\bm{m}$ parameter takes the form $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$.
 The $\bm{m}$ and $\bm{u}$ parameters are used to calculate the likelihood ratio
@@
 
-With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model.
-Let $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$.
-Let $\bm{u}$ parameters be defined analogously, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$.
-Then, our model is:
+With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model:
 \begin{align*}
@@
 The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure \ref{fig:m-and-u}.
-\textcolor{red}{Estimates of $\bm{u}$ are indistinguishable, and thus omitted.} Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+Estimates of $\bm{u}$ are indistinguishable, and thus omitted.
+Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
 
@@
 
-	\textcolor{red}{Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, the computing all $n_1 \times n_2$ record pairs can still be infeasible for larger linkage tasks.
+	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, the computing all $n_1 \times n_2$ record pairs can still be infeasible for larger linkage tasks.
 Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
@@
 Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task.
-Thus in future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to arbitrarily large linkage tasks.}
+Thus in future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to arbitrarily large linkage tasks.
 
 	%Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
-
-\textcolor{blue}{This needs to address the concern of R3, where we didn't cite Sadinle/FS. I would also suggest that we take the time to elaborate on some future work.}
 

### 2022-10-31 Prelim/DANE_presentation.tex
JErrys edits

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\begin{document}
	\title{Introduction to Probabilistic Record Linkage}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger, Jerome Reiter, Rebecca Steorts}
	\institute{Duke University}
	\date{\today}

%	\AtBeginSection[]{
%		\begin{frame}
%			\vfill
%			\centering
%			\begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%				\usebeamerfont{title}\insertsectionhead\par%
%			\end{beamercolorbox}
%			\vfill
%		\end{frame}
%	}

\AtBeginSection[]
{
	\begin{frame}
		\frametitle{Table of Contents}
		\tableofcontents[currentsection]
	\end{frame}
}

	\begin{frame}
		\titlepage
	\end{frame}

%	\begin{frame}
%	\frametitle{Outline}
%	\tableofcontents
%	\end{frame}

\section{Introduction to Record Linkage}

	\begin{frame}{What is Record Linkage?}
	\begin{itemize}
		\item Record linkage is the task of identifying duplicate records over noisy datasets.

		\item Easy with unique identifiers, difficult when faced with errors

		\item \textbf{Bipartite matching} is the specific goal of matching one record in one dataset to most one match in another dataset
	\end{itemize}
	\end{frame}

%\begin{frame}{Record Linkage in Practice}
%	\begin{columns}
%		\begin{column}{0.48\textwidth}
%			\includegraphics<1->[width = \textwidth, height = .9\textwidth ]{ted_article2.png}
%			
%		\end{column}
%		\begin{column}{0.48\textwidth}
%			\includegraphics<2->[width = \textwidth, height = .9\textwidth ]{dnc_header.png}
%		\end{column}
%	\end{columns}
%\end{frame}

\begin{frame}{Record Linkage in Practice}
	\centering
	\includegraphics<1->[width = .8\textwidth, height = .6\textwidth ]{graphics/syria_article_big.png}
\end{frame}

\begin{frame}{Linkage for Downstream Analysis}
	\includegraphics<1>[width = \textwidth, height = .7\textwidth ]{graphics/Slide1.png}
	\includegraphics<2>[width = \textwidth, height = .7\textwidth ]{graphics/Slide2.png}
\end{frame}

\begin{frame}{Linkage through Comparison Vectors}
	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide3.png}
	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide4.png}
	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide5.png}
\end{frame}

\begin{frame}{Linkage through Comparison Vectors}
	Represent linkage structure through vector $\mathbf{Z} = \{Z_1, \ldots, Z_{n_B}\}$, where
\begin{center}
	$Z_j = \begin{cases} 
		i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
		n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
	\end{cases}$
\end{center}
\end{frame}

\begin{frame}{Linkage through Comparison Vectors}
	\includegraphics[width = \textwidth, height = .6\textwidth ]{graphics/Slide6.png}
\end{frame}

\begin{frame}{Fellegi and Sunter (1969)}
	\begin{columns}
		\begin{column}{0.6\textwidth}
			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide1.png}
			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide2.png}
		\end{column}
		\begin{column}{0.38\textwidth}
			\begin{itemize}
				\item<3-> scalable to large datasets (\texttt{fastlink}, Enamorado et al 2019)
				\item<4-> not bipartite, requires post-processing
				\pause 
				\item<5-> overmatches, leading to inaccurate parameter estimation
			\end{itemize}

		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Fellegi and Sunter (1969)}
	\begin{itemize}
		\item The Fellegi and Sunter model is essentially a \emph{mixture model}
		\item We posit that comparison vectors come from two different distributions, one for matching, and another for nonmatching record pairs
		\item The model learns the parameters for each distribution and independently classifies each record pair as matching or nonmatching.
	\end{itemize}
\end{frame}

\begin{frame}{Sadinle (2017)}
	\begin{columns}
		\begin{column}{0.6\textwidth}
			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide1.png}
			\includegraphics<2>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide3.png}
			\includegraphics<3>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide4.png}
			\includegraphics<4->[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide5.png}
		\end{column}
		\begin{column}{0.38\textwidth}
			\begin{itemize}
				\item<1-> Beta Record Linkage (\texttt{BRL})
				\pause
				\item<5-> strictly enforces one-to-one matching, no post-processing
				\pause
				\item<6-> high accuracy for linkage and other parameters
				\pause 
				\item<7-> inherently serial, not scalable to large linkage tasks
			\end{itemize}

		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Our Contribution - Fast Beta Linkage}
	\begin{columns}
		\begin{column}{0.6\textwidth}
			\includegraphics<1>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide1.png}
			\includegraphics<2-4>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide6.png}
			\includegraphics<5>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide7.png}
			\includegraphics<6>[width = \textwidth, height = 1.2\textwidth ]{graphics_square/Slide8.png}
		\end{column}
		\begin{column}{0.38\textwidth}
			\begin{itemize}
				\item<3-> relaxation proposed by Heck Wortman (2019)
				\item<4-> minimal loss of accuracy, large computational gains
				\item<5-> allows for "one to many" matchings
				\item<6-> simple postprocessing to obtain bipartite matching
			\end{itemize}

		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Model Parameters}
	\begin{itemize}
		\item Both models utilize $\mathbf{m}$ and $\mathbf{u}$ parameters to model the data.
		\item Define $m_{f\ell} = P(\gamma_{ij}^f = \ell | \text{Records Match})$ and $u_{f\ell} = P(\gamma_{ij}^f = \ell | \text{Records Do Not Match})$
		\item We can think of $\mathbf{m}$ as reliability parameters, that show how likely it is for a feature to remain unchanged across files
		\item We can think of $\mathbf{u}$ as discernment parameters, that show how unique the values of a feature can be
	\end{itemize}
\end{frame}

\begin{frame}{Fellegi Sunter Weights}
	\begin{itemize}
		\item We multiply these parameters for all $f$ and the relevant $\ell$ to get $\mathbf{m}_{ij} = P(\gamma_{ij}| \text{Records Match})$ and $\mathbf{u}_{ij} = P(\gamma_{ij}| \text{Records Do Not Match})$
		\item These probabilities form the likelihood ratio 
		$$w_{ij} = \frac{P(\gamma_{ij}| \text{Records Match})}{P(\gamma_{ij}| \text{Records Do Not Match})}$$
		\item This ratio is large when $\mathbf{m}_{ij}$ is large and $\mathbf{u}_{ij}$ is small
	\end{itemize}
\end{frame}

\section{Practical Considerations}

\begin{frame}{Constructing Comparison Vectors}
	\begin{itemize}
		\item We measure the distance between two fields using whatever metric we like, and then threshold this distance into discrete values
		\item We can use binary agreement, or use multiple agreement levels to capture partial agreement
		\item Existing packages perform binary agreement, numerical distance, and string distance
	\end{itemize}
\end{frame}

\begin{frame}{String Distance}
	\begin{itemize}
		\item Many packages compare strings through the normalized Levenstein distance. This is the minimum number of edits required to transform one string to the other, divided by the length of the larger string.
		\item Additional metrics could be language specific Soundex codes
		\item Care should be taken when transforming metric into discreet agreement level. For example, should a Soundex match be regarded as a full match or a partial match?
		\item Code from existing packages can be edited to accomodate additional string distances
	\end{itemize}
\end{frame}

\begin{frame}{Conditionally Independent Features}
	\begin{itemize}
		\item The record linkage model assumes that agreement levels are \emph{conditionally independent}, given the matching status of the records. This is satisfied for most fields
		\item For example, knowing that two records agree on birth month does not tell you anything about whether they agree on birth year. These fields are independent
		\item In contrast, if two records disagree on department, we know they probably disagree on city. These fields are not independent
		\item Using dependent features is essentially using the same information multiple times, and can harm results
	\end{itemize}
\end{frame}

\begin{frame}{Conditionally Independent Features}
	\begin{itemize}
		\item One solution to dependent features is to create a more nuanced comparison incorporating multiple features
		\item For example, you can define full agreement for "location" to be matching on city AND department, partial agreement to be matching on only department, and no agreement not be matching on neither.
		\item Current packages do not execute this type of comparison, but this can be done with some work
		\item Depending on the type of dependent fields, I may recommend keeping both or just omitting one. 
	\end{itemize}
\end{frame}

\begin{frame}{Conditionally Independent Features}
	\begin{itemize}
		\item One solution to dependent features is to create a more nuanced comparison incorporating multiple features
		\item For example, you can define full agreement for "location" to be matching on city AND department, partial agreement to be matching on only department, and no agreement not be matching on neither.
		\item Current packages do not execute this type of comparison, but this can be done with some work
		\item Depending on the type of dependent fields, I may recommend keeping both or just omitting one. 
	\end{itemize}
\end{frame}

\begin{frame}{Random Sampling and Generalization}
	\begin{itemize}
		\item To avoid calculating $n_A n_B$ comparison vectors, we can use \emph{blocking}
		\item We only create comparison vectors when records agree on a particular field
		\item It may be tempting to block on a unique identifier, but this can create blocks that are too small for record linkage to run effectively
	\end{itemize}
\end{frame}

\begin{frame}{Thoughts for DANE}
	\begin{itemize}
		\item With \texttt{fastLink}, you may be able to do the entire linkage task (60,000 $\times$ 1,000,000) records at one time. If not blocking by gender or by department should make it feasible
		\item You should not need to use the generalization technique
		\item \texttt{fastLink} currently does not do Soundex comparisons, but using allowing for partial agreement on strings should be sufficient. \texttt{fastLink} code can be modified to incorporate Soundex comparison, but may take some work
	\end{itemize}
\end{frame}

\section{Evaluation}

\begin{frame}{Performance Metrics with Known Labels}
	\begin{itemize}
		\item When true matching status is known, we use two primary metrics:
		\begin{itemize}
		\item $\text{Recall} = \frac{\text{True Matches Declared}}{\text{Total True Matches}}$
		\item $\text{Precision} = \frac{\text{Matches Declared}}{\text{Total Matches Declared}}$
		\end{itemize}
		\item To balance these two metrics, we sometimes also consider:
		\begin{itemize}
		\item  $\text{F-Measure} = \frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}$
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Performance Metrics without Known Labels}
	\begin{itemize}
		\item Without known labels, evaluating model performance is more difficult, but there are general checks you can make
		\begin{itemize}
			\item The histogram of match probabilities should be clearly separated, with most record pairs having probabilities near 0 or near 1, and minimal number of pairs in between
			\item  Fellegi Sunter weights for impactful matching variables should be high. Weights for partial agreement should be lower than those for full agreement
			\item Visually inspect a subset of matches and nonmatches. In particular, examine matches near the decision threshold.
		\end{itemize}
	\item These approaches are demonstrated in the accompanying Rmarkdown file
	\end{itemize}
\end{frame}

\section{Software}
\begin{frame}{Software}
	\begin{itemize}
		\item \texttt{fastLink} can be found on \href{https://github.com/kosukeimai/fastLink}{GitHub}
		\item \texttt{BRL} can also be found on \href{https://github.com/cran/BRL}{GitHub}
		\item \texttt{fabl} is not yet publicly available
		\item Due to the scale of DANE's record linkage task, I recommend you to use \texttt{fastLink}
	\end{itemize}
\end{frame}

\section{References}
\begin{frame}{References}
	\begin{itemize}
		\item Fellegi I. and Sunter A., A theory for record linkage, J. Am. Stat. Assoc. 64 (1969), pp. 1183–1210
		\item Enamorado T., Fifield B. and Imai K., Using a probabilistic model to assist merging of large-scale administrative records, Am. Political Sci. Rev. 113 (2019), pp. 353–371
		\item Sadinle M., Bayesian estimation of bipartite matchings for record linkage, J. Am. Stat. Assoc. 112 (2017), pp. 600–612
	\end{itemize}
\end{frame}

\end{document}

### 2022-10-31 ba_resubmit/kundinger_resubmit.tex
JErrys edits

@@
 \author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
-\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
+\author{\fnms{Jerome P.} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
 \and
@@
 \begin{abstract}
-Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage.
-These have the crucial advantage of quantifying uncertainty from imperfect linkages.
-However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
+%Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
+\textcolor{red}{Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.} However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
 To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
-Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
+Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
 This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
@@
 \begin{keyword}%[class=MSC]
+\kwd{data fusion}
+\kwd{distributed computing}
 \kwd{entity resolution}
-\kwd{data fusion}
 \kwd{hashing}
 \kwd{record linkage}
-\kwd{Bayesian}
-\kwd{distributed computing}
 \end{keyword}
@@
 In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space.
+Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space.
 This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
@@
 Without loss of generality, denote files such that $n_1 \geq n_2$.
-We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.
-In the context of bipartite matching, we assume that there are duplications across, but not within, each datafile.
-Under this framework, the set of matches across datafiles can be represented in two equivalent ways.
+We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in X_1$" rather than the more compact $x_{1i}$ \textcolor{red}{for ease of readability}.
+Under bipartite matching, the set of matches across datafiles can be represented in two equivalent ways.
 First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
@@
 Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks.
-More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in datafile $\bm{X}_2$ such that
+More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that
 \begin{align}
@@
 		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in datafile}\; \bm{X}_1. \\
+		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
 	\end{cases}
@@
 Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
-$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
+$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
 
@@
 
-Intuitively, matching records (those that refer to the same entity) should be similar; records that are not non-matching should not be similar.
-\cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
+Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
 We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.
 
-The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables.
-For numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
-We can then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features.
+For numerical features, we can use absolute difference between the two feature values.
+For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
+We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+
 Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$.
 Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
-Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.
+Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
+
+\textcolor{red}{In the construction of comparison vectors, we note it is common to encounter missing information in datasets used for record linkage.
+Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We assume that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
+With the MAR, we can marginalize over the missing data, and do all computation simply using the observed data.
+To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}
 
@@
 
-\subsection{Fellegi-Sunter Model}
+\subsection{Fellegi-Sunter Models}
 \label{fellegi-sunter}
 
-The Fellegi Sunter model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
-This allows us to model the comparison data using mixture models in the following way:
+The seminal \cite{fellegi_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
+These allow us to model the comparison data using a mixture model, where
 \begin{align}
+	\label{eqn:fs_model}
 	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
 	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
-	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
+	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \notag
 \end{align}
-where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match.
-When using comparison vectors with discrete agreement levels as described above, the $\bm{m}$ parameter takes the form $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$.
-The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$.
-The $\bm{m}$ and $\bm{u}$ parameters are used to calculate the likelihood ratio
+Here $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
+When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
+Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
+
+\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
 \begin{align}
@@
 \end{align}
-This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
+This ratio is large when there is strong evidence of the pair being a match, and small otherwise.
+All record pairs with weights above a specified threshold are considered a match.
+Such independent classifications often leads to matchings that violate one-on-one matching assumptions, which prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi Sunter weights among matched pairs.
 
@@
 
-The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+%The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
 This is a problem in practice as transitive closures can be violated.
@@
 
-To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
-\begin{align}
-	\label{eqn:jaro}
-	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
-	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
-	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
-	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
-\end{align}
-
-The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
+%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
+%\begin{align}
+%	\label{eqn:jaro}
+%	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
+%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
+%	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
+%	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
+%\end{align}
+%
+%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
 Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
@@
 
-\subsection{Beta Record Linkage Model}
-\label{BRL}
-
-To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
-He proposed the  beta distribution for bipartite matching, given by
-$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
-$$
-where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
-This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
-This creates a dependency that makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
-
+%\subsection{Beta Record Linkage Model}
+%\label{BRL}
+
+%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
+
+\textcolor{red}{To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a Bayesian framework.
+In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposed the "beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
+To construct this prior, he assigned a prior to the probability of the indicator that a record in $\bm{X}_2$ had a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
+It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$.
+Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 !
+/ (n_1 - n_{12}(\bm{Z})!$ bipartite matchings are taken to be equally likely.} Thus the beta distribution for bipartite matching, is given by
+\begin{align}
+\label{eqn:sadinle_prior}
+P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+\end{align}
+where $\text{B}(\cdot, \cdot)$ represents the Beta function.
+This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling \(Z_j\).
+This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
+
+%hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X}_2$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
 %One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
@@
 However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior.
-We denote the fast beta prior as follows:
-$$Z_j | \pi =
-\begin{cases} 
+We denote the fast Beta prior as follows:
+\begin{align}
+	\label{eqn:fast_beta_prior}
+	Z_j | \pi &= \begin{cases} 
 	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
 	1-\pi, &  z_j  = n_1 + j; \\
-\end{cases}$$
-$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$
-This set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
-The parameter $\pi$ is given hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ that encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}.$
-
-% I don't really like using remarks. If its not relevante enough for the main body, we can just remove. 
-%\begin{remark}
-%One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-Note the prior from \cite{wortman2019} can be viewed as a special case of the fabl prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
-%\end{remark}
+\end{cases} \\
+\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
+\end{align}
+We can interpret equation \ref{eqn:fast_beta_prior} as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match.
+The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$
+
+\textcolor{red}{Previously, \cite{wortman2019} proposed a flat prior over the labeling of $\bm{Z}$, meaning that each value $\{1, \ldots, n_1, n_1 +j\}$ was considered equally likely for $Z_j$.
+This however amounts to a prior probability of $\frac{n_1}{n_1 + 1}$ that record $j$ has a match in $\bm{X}_1$, and thus tended to over match, leading to poor precision when used in practice.
+Indeed we find that the flat prior is equivalent to a special case of the fast Beta prior where $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.}
 
@@
 
-Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
+%Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
 To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure explained in Section \ref{bayes-estimate}.
 
-Lastly, we note that it is common to encounter missing information in datasets used for record linkage.
+\textcolor{red}{To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}, reproduced in Appendix .
+Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+To achieve a Bayes estimate that fulfills one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the highest posterior probability, and declare the other to have no match.
+A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.}
+
+%It is common to encounter missing information in datasets used for record linkage.
 Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
-With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+We assume that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
+With the MAR, we can marginalize over the missing data, and do all computation simply using the observed data.
 To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.
@@
 		$f \in 1, \ldots F$ & index over fields used for comparisons \\
-		$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
+		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
 		$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
@@
 
-With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model:
-\begin{align*}
+For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_1}.
+\begin{align}
+	\label{eqn:full_model}
 	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \notag \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \notag\\
 	Z_j | \pi &=
@@
 		1-\pi, &  z_j  = n_1 + j; \\
-	\end{cases} \\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).
-\end{align*}
-For convenience, a summary of notation is provided in Table \ref{table_notation_1}.
+	\end{cases} \notag\\
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\notag.
+\end{align}
+
 %where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.
@@
 Next, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
-$$p\left(Z_j^{(s+1)}  = i|\Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+\begin{align}
+\label{eqn:z_full_conditional}
+p\left(Z_j^{(s+1)}  = i|\Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
@@
 	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
-\end{cases}$$
+\end{cases}
+\end{align}
 where 
@@
 
-\hypertarget{bayes-estimate}{%
-	\subsection{Bayes Estimate}
-	\label{bayes-estimate}}
-
-We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage
-parameter \(\bm{Z}\) by assigning different positive losses to
-different types of errors, and minimizing posterior expected loss. We
-adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which
-\(\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\}\), with \(R\) representing the
-option to leave the matching undetermined by the model. Specifically, we have
-
-\[L(\hat{Z_j}, Z_j)=\begin{cases} 
-	0,  & \text{if } Z_j = \hat{Z_j}; \\
-	\theta_R,  & \text{if } \hat{Z_j} = R; \\
-	\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
-	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
-	\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
-\end{cases}\] 
-Here, \(\theta_R\) is the loss from not making a decision on the linkage status, \(\theta_{10}\) is the loss from a false non-match, \(\theta_{01}\) is the loss from a false match, and \(\theta_{11}\) is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
-
-%In this article, we set false matches and false non-matches to be equally costly, and thus set $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$; when we want a full estimate of the linkage parameter, we set $\theta_R = \infty$, and when we want to leave some records unmatched by the algorithm, we set $\theta_R = 0.1$.
-
-%We seek to minimize the posterior expected loss, given by 
-%$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
-In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
-$$\hat{Z}_j =\begin{cases} 
-	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
-	0,  & \text{otherwise}. \\
-\end{cases}$$
-For a more in-depth explanation of this function and the induced Bayes
-estimate, see \cite{sadinle_bayesian_2017}.
-
-Since the Gibbs sampler does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+%Since the Gibbs sampler does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
 To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
@@
 
-%It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable.
-We will explore this issue through the El Salvador homicide case study.
-
 \section{Efficient and Scalable Implementation}
@@
 
-The scale of linkage tasks possible within the Fellegi Sunter framework is limited by (1) the memory costs of storing $n_1 \times n_2$ comparison vectors for every pair of records across the two datafiles, and (2) the speed of the linkage algorithm over those comparison vectors.
+The scale of linkage tasks possible through \texttt{BRL} is limited by (1) the memory costs of storing $n_1 \times n_2$ comparison vectors for every pair of records across the two datafiles, and (2) the speed of the linkage algorithm over those comparison vectors.
 One common solution to these challenges is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{christen2019data}.
@@
 The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
+\textcolor{blue}{DELETE?
+However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}}.
 
@@
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like first name, or only pairs that exactly match on a specified number of fields.
-However, this kind of indexing requires knowledge  of the application and invites room for human error \citep{murray2016probabilistic}.
-
-With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage without the drawbacks of blocking and indexing.
+\textcolor{blue}{DELETE?
+However, this kind of indexing requires knowledge  of the application and invites room for human error \citep{murray2016probabilistic}.}
+
+With \texttt{fabl}, we introduce two techniques \textcolor{red}{further expand} the scalability of probabilistic record linkage \textcolor{red}{in addition to} blocking and indexing.
 First, we propose hashing methods that allow us to compute sufficient statistics which reduce the computational complexity of the Gibbs sampler.
-Second, we introduce storage efficient indexing (SEI), which reduces the memory costs associated with unlikely matches.
+Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
 
@@
 
-Following insight from \cite{enamorado2019using}, we recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
+Following insight from \cite{enamorado2019using}, we recognize that record pairs contribute to posterior calculations only through the agreement pattern $\gamma_{ij}$.
 To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
@@
 Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
+Additionally, we store ``one-hot encodings'' of these patterns rather than the original $\gamma_{ij}$ to aid in vectorized computations.
 See Appendix \ref{app:ohe} for details.
@@
 When the $(i, j)$ pair exhibits the $p^{th}$ pattern, we say $(i, j) \in h_p$.
-
-With all record pairs converted to hashed values, we can create a more compact nested list 
-$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
-where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$.
-This representation is useful because it allows us to easily compute two sets of sufficient statistics.
+\textcolor{red}{When $j$ has no match in $\bm{X}_1$, we write $(n_1 + j, j) \in h_{P+1}$.}
+
+\textcolor{red}{With all record pairs converted to hashed values, we identify the records in $\bm{X}_1$ with comparison vectors corresponding to each pattern $p$ for each record $j \in \bm{X}_2$.
+We denote this set $r_{j_p} = \{i \in \bm{X}_1 | (i, j) \in h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{j_p} | j \in \bm{X}_1, p \in 1, \ldots, P\}$.}This representation is useful because it allows us to easily compute two sets of sufficient statistics.
 First, we compute the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, given by
-$$N_{j_p} = \sum_{i=1}^{n_1} I((i, j) \in h_p) = |r_{j_p}|.$$
+$$N_{j_p} = |r_{j_p}| = \sum_{i=1}^{n_1} I((i, j) \in h_p).$$
 Second, we compute the number of comparison vectors in $\Gamma$ that exhibit pattern $p$, given by
-$$H_p =  \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} I((i, j) \in h_p) = \sum_{j = 1}^{n_2} N_{j_p}.$$
-We collect these sufficient statistics in the sets $\mathcal{N} = \left\{\{N_{j_p}\}_{j=1}^{n_2}\right\}_{p=1}^P$ and $\mathcal{H} = \{H_{p}\}_{p=1}^P.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
-Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}, \mathcal{H}\}$ for posterior inference.
+$$H_p = \sum_{j = 1}^{n_2} N_{j_p}.$$
+\textcolor{red}{We collect these sufficient statistics in the sets $\mathcal{N} = \{N_{j_p} |j \in \bm{X}_1, p \in 1, \ldots, P \}$ and $\mathcal{H} = \{H_{p} | p \in 1, \ldots, P \}$.} As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
+Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we can use $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}, \mathcal{H}\}$ for posterior inference.
+
+%With all record pairs converted to hashed values, we can create a more compact nested list 
+%$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
+%where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$.
+
+%We collect these sufficient statistics in the sets $\mathcal{N} = \left\{\{N_{j_p}\}_{j=1}^{n_2}\right\}_{p=1}^P$ and $\mathcal{H} = \{H_{p}\}_{p=1}^P.$
 
@@
 
-With storage efficient indexing (SEI), we can calculate the above sufficient statistics for all $n_1 \times n_2$ record pairs, but greatly reduce the memory costs associated with unlikely matches, allowing us to conduct all-to-all comparisons for substantially larger linkage tasks.
-Observe that all records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$.
+With storage efficient indexing (SEI), we can calculate $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs, but greatly reduce the memory costs associated with unlikely matches, allowing us to conduct all-to-all comparisons for substantially larger linkage tasks.
+Observe that all records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same $w_{ij}$.
 Therefore, these records have the same probability to be identified as a link when sampling \(Z_j\).
@@
 
-For large data, we can partition the two datasets \(\bm{X}_1\) and \(\bm{X}_2\) into smaller chunks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more manageable computations.
-On a single machine, we can read-in data sequentially, conduct hashing, compress information through SEI, and delete the original data from memory before continuing with the next chunk of data.
-With multiple cores or multiple machines, this can be done in parallel.
-Summary statistics from each pairwise chunk comparison can then be easily synthesized to recover sufficient statistics for the larger linkage task.
+For datafiles with large amounts of records, we can partition the two datasets \(\bm{X}_1\) and \(\bm{X}_2\) into smaller chunks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more manageable computations.
+\textcolor{red}{On a single machine, we can read-in data and construct comparison vectors sequentially.
+Then we can conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.} With multiple cores or multiple machines, this can be done in parallel.
+Summary statistics from each pairwise chunk comparison can  be easily synthesized to recover sufficient statistics for the $\Gamma$.
 Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger datasets.
@@
 
-Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum_f L_f\) many summations over \(n_1 \times n_2\) quantities.
+Updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum_f L_f\) many summations over \(n_1 \times n_2\) quantities.
 These are simple calculations, but become computationally burdensome when working on large linkage tasks.
@@
 Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations.
-Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p\right)$ to be the number of matching record pairs with agreement pattern $p$.
-It follows that the number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
+\textcolor{red}{Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p, Z_j \leq n_1 \right)$ to be the number of matching record pairs with agreement pattern $p$.} It follows that the number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
 Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors  of length $\sum_f L_f$ of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
@@
 
-Although sampling \(Z_j\) from the full conditional provided in Section \ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
+Although sampling \(Z_j\) from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
 This is because sampling a value from $n_1$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
@@
 Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns.
-After all iterations are complete, we can back-fill the records uniformally among records of the sampled agreement pattern through (\ref{eqn:gibbs2}).
+\textcolor{red}{Since all records in $\bm{X}_1$ sharing the same agreement pattern with $j \in \bm{X}_2$ are equally likely, we can sample among candidate records uniformally through (\ref{eqn:gibbs2}).}
+
+%After all iterations are complete, we can back-fill the records uniformally among records of the sampled agreement pattern through (\ref{eqn:gibbs2}).
 \begin{align}
@@
 
-We emphasize the computational gains of this split sampler: the first step is a sample from $P$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
-These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
+We emphasize the computational gains of this split sampler: the first step is a sample from $P + 1$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
+These changes greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
 We provide a summary of the notation used for the hashing procedure in Table \ref{table_notation_2}.
@@
 	\begin{proof}
-		To prove the computational complexity, we consider two steps: constructing the comparison vectors and the Gibbs sampler.
-The computational complexity of all pairwise comparisons across the two datafiles $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
+		 We consider two steps: constructing the comparison vectors and the Gibbs sampler.
+The computational complexity of all pairwise comparisons across $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
 The hashing procedure for all pairwise comparisons is also $O(F n_1 n_2)$.
@@
 	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-We use five binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
+We use $F = 5$ binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
+\textcolor{red}{For each record in $\bm{X}_2$, we simulate $n_1$ comparison vectors, resulting in a comparison matrix $\Gamma \in \mathbb{R}^{n_1 n_2 \times F}$.
+For $n_2/2$ of these records, there is no match in $\bm{X}_1$, so we simulate $n_1$ comparison vectors from the $\bm{u}$ distribution.
+For the other $n_2/2$ of these records, there is one match in $\bm{X}_1$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_1 - 1$ comparison vectors from the $\bm{u}$ distribution.
+We compare the run-time of \texttt{fabl} against \texttt{BRL} as we increase $n_1$ and $n_2$.
+Note that since we have five binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.}
+
+%	We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across datasets.
+and compare the run-time of \texttt{fabl} against \texttt{BRL}.
+Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
+
+	 %\textcolor{blue}{DELETE?
 The $\bm{u}$ parameters are chosen to emulate what is seen in practice for first name, last name, day of birth, and month of birth.
-The parameters for year would vary largely by context, so we choose them here to adjust the desired difficulty of the linkage task.
-We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across datasets, and compare the run-time of \texttt{fabl} against \texttt{BRL}.
-Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
+The parameters for year would vary largely by context, so we choose them here to adjust the desired difficulty of the linkage task.}
 
@@
 		\end{tabular}
-		\caption{Probabilities used for $m$ and $u$ distributions in simulation studies}\label{Tab:distributions}
+		\caption{Probabilities used for $m$ and $u$ distributions in simulation study in Section \ref{speed}.}\label{Tab:distributions}
 	\end{table}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$, \textcolor{red}{as described in Section \ref{speed}.} We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
 		\end{center}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with increasing $n_1$ and $n_2$ fixed at 500.
-We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_2$ fixed at 500, \textcolor{red}{as described in Section \ref{speed}.} We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
 		\end{center}
@@
 
-	We replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several synthetic datasets with varying amounts of error and overlap (the number of records in common across files).
+	We replicate a simulation study from \cite{sadinle_bayesian_2017} used to illustrate the effectiveness of \texttt{BRL}.
+We adapt it to compare the accuracy of \texttt{fabl} to \texttt{BRL}.
+The simulations employ a collection of synthetic synthetic datasets with varying amounts of error and overlap (the number of records in common across files).
 Following methods proposed by \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
@@
 Under each of these settings, we have 100 pairs of simulated datasets in order to obtain uncertainty quantification on our performance metrics.
-We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
+We use flat priors for all $\bm{m}$ and $\bm{u}$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
 We calculate Bayes estimates of the linkage structure using the losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$.
 Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
+We replicate this simulation under partial estimates, allowing the \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review, in Appendix \ref{partial}.
 
@@
 		\end{tabular}
-		\caption{Construction of comparison vectors for accuracy study with simulated datasets.}
+		\caption{Construction of comparison vectors for accuracy study with simulated datasets of Section \ref{accuracy}.}
 		\label{Tab:sadinle_simulation_cutoffs}
@@
 Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$.
-Precision is the proportion of links found by the model that are true matches, that is $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
+Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
 The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
@@
 
-	\hypertarget{partial}{%
-		\subsection{Accuracy under Partial Estimates}\label{partial}}
-
-	By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
-In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
-Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual non-matches.
-Mathematically, $\text{NPV} = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
-We continue to use the precision, which is renamed the positive predictive value (PPV) in this context.
-Lastly, we also examine the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)/n_2$.
-To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.
-
-	In Figure \ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
-However, with high amounts of error, and thus fewer accurate and discerning fields of information, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
-Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probabilities of matches at times can be split across more records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand than would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-	%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-	%This is reasonable because large numbers of errors blur the distinction between matched and non-matched pairs, leading to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
-
-	\begin{figure}[t]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
-			\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on simulated datasets.
-We see poorer performance for \texttt{fabl} only in situations with high overlap.}
-			\label{fig:sadinle_simulation_partial}
-		\end{center}
-	\end{figure}
-
 	%\hypertarget{scale}{%
@@
 
-	The country of El Salvador was immersed in civil war from 1980 to 1991.
-Throughout the time, several organizations attempted to document casualties of the conflict.
+	The country of El Salvador was immersed in civil war from 1980 to 1991, and we are interested in estimating the total number of casualties from the war.
+We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that had been published throughout the conflict, and the CDHES dataset consists of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
+The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both datasets are thought to be fairly reliable.
 When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
 Instead, record linkage techniques must be used to merge datasets before analyzing the data \citep{lum2013applications}.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that had been published throughout the conflict, and the CDHES dataset consists of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
-The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both datasets are thought to be fairly reliable.
 
@@
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We again flat priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
-
-	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
+We flat priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
+
+	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
 This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
-However, conducting record linkage, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical, suggesting that these levels were unnecessary.
+However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
 
@@
 		\end{tabular}
-		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle2018bayesian}.
+		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle_bayesian_2017}.
 This set up leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
@@
 
-	Therefore, we re-ran our analysis with fewer agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
+	Therefore, we perform our analysis with the agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
 Among the 216 possible agreement patterns, 159 are realized in the data.
 With this revised comparison specification, \texttt{fabl} ran in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
-The estimates of the $\bm{m}$ parameters under each method are very similar, as shown in Figure \ref{fig:m-and-u}.
+The estimates of the $\bm{m}$ parameters under each method are similar, as shown in Figure \ref{fig:m-and-u}.
 Estimates of $\bm{u}$ are indistinguishable, and thus omitted.
@@
 
-	For completeness, we note that linkage with the more detailed comparison vectors required240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
+	For completeness, we note that linkage with the more detailed comparison vectors required 240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
 Apparently, the number of patterns was sufficiently many that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
@@
 This is because a large number of records in ERTL have multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
-We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Section \ref{accuracy}.
+We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Appendix \ref{partial}.
 Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
@@
 See Figure \ref{fig:overlap-plot} for a visual comparison of the Bayes estimate and posterior credible intervals for the two methods.
-We note that Bayes estimates falling outside of posterior credible intervals has occurred previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
+We note that Bayes estimates falling outside of posterior credible intervals has been observed previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
 
@@
 As such, patients are surveyed at most once in a given year, and many patients are surveyed across multiple years.
-In addition, patients can either drop out of the study, pass away, or enter as new patients.
+In addition, patients can either drop out of the study, pass away, or enter as new patients \citep{steorts_bayesian_2016}.
 Hence, the assumptions of our model hold for this study.
@@
 
-	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, the computing all $n_1 \times n_2$ record pairs can still be infeasible for larger linkage tasks.
+	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, the computing all $n_1 \times n_2$ record pairs still can be infeasible for larger linkage tasks.
 Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
@@
 	\label{sec:appendix}
+
+	\hypertarget{bayes-estimate}{%
+		\subsection{Bayes Estimate}
+		\label{bayes-estimate}}
+
+	We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage
+	parameter \(\bm{Z}\) by assigning different positive losses to
+	different types of errors, and minimizing posterior expected loss. We
+	adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which
+	\(\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\}\), with \(R\) representing the
+	option to leave the matching undetermined by the model. Specifically, we have
+
+	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
+		0,  & \text{if } Z_j = \hat{Z_j}; \\
+		\theta_R,  & \text{if } \hat{Z_j} = R; \\
+		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
+		\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
+		\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
+	\end{cases}\] 
+	Here, \(\theta_R\) is the loss from not making a decision on the linkage status, \(\theta_{10}\) is the loss from a false non-match, \(\theta_{01}\) is the loss from a false match, and \(\theta_{11}\) is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
+
+	In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
+	$$\hat{Z}_j =\begin{cases} 
+		i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
+		0,  & \text{otherwise}. \\
+	\end{cases}$$
+When we seek a partial estimate of the linkage structure, leaving a portion of record pairs to be classified manually in clerical review, we adopt losses $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, .1)$.
+For a more in-depth explanation of this function and the induced Bayes
+	estimate, see \cite{sadinle_bayesian_2017}.
 
@@
 
+	\hypertarget{partial}{%
+	\subsection{Accuracy under Partial Estimates}\label{partial}}
+
+By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
+In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
+Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual non-matches.
+Mathematically, $\text{NPV} = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
+We continue to use the precision, which is renamed the positive predictive value (PPV) in this context.
+Lastly, we also examine the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)/n_2$.
+To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.
+
+In Figure \ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
+However, with high amounts of error, and thus fewer accurate and discerning fields of information, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
+Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probabilities of matches at times can be split across more records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand than would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
+
+%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
+
+%This is reasonable because large numbers of errors blur the distinction between matched and non-matched pairs, leading to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
+
+\begin{figure}[t]
+	\begin{center}
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
+		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on simulated datasets.
+We see poorer performance for \texttt{fabl} only in situations with high overlap.}
+		\label{fig:sadinle_simulation_partial}
+	\end{center}
+\end{figure}
+
 	\hypertarget{appendix-sim}{%
@@
 Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.
+The $\bm{m}$ and $\bm{u}$ parameters for this simulation are shown Table \ref{Tab:distributions_2}.
+
+		\begin{table}[t]
+		\centering
+		\begin{tabular}{llll|lll}
+			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{m} & \multicolumn{3}{c}{u} \\
+			\cline{2-4} \cline{5-7}
+			& Agree & Partial & Disagree & Agree & Partial & Disagree \\
+			\hline
+			Feature 1 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+			Feature 2 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+			Feature 3 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+			Feature 4 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+			\hline
+		\end{tabular}
+		\caption{Probabilities used for $m$ and $u$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
+	\end{table}
 

### 2022-10-31 R/GetUniquePatterns2.R

@@ -1,5 +1,5 @@
 hash_field <- function(L_f, k, Lf_vec){
-  level_seq <- seq_len(L_f) - 1
+  level_seq <- seq_len(L_f)
   as.numeric(level_seq > 0) * 2 ^ ((level_seq) + (as.numeric(k > 1)  * Lf_vec[k]))
 }
 

### 2022-10-31 R/NLC_hash2.R

@@ -24,7 +24,7 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   }
 
   ptm <- proc.time()
-  patterns <- GetUniquePatterns2(comparisons, fast, R, all_patterns)
+  patterns <- GetUniquePatterns2(comparisons)
   elapsed_hash <- proc.time() - ptm
 
   parameter_split <- as.vector(unlist(sapply(1:fields, function(x){

### 2022-11-19 ba_resubmit/response_to_reviewers.tex
final revisions

@@
 	\item As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies.
-\textbf{This has been done by all authors.} \todo
+\textbf{This has been done by all authors.}
 	\item The AE finds the simulation studies to be incomplete and gives detailed suggestions on how those should be improved.
@@
 \reply{%
-	Our missingness assumptions and the definition of $I_{obs}()$ have been added near end of Section 3.
+	Our missingness assumptions and the definition of $I_{obs}()$ have been added right before Section 2.1.
 }
@@
 \reply{%
-	We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.
+	We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$.
+We have updated additional parts of the paper with this revision.
+We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.
 }
@@
 \reply{%
-	We have updated the Figure 4 with these results.
+	We have updated the Figure with these results.
 Since it is presented alongside NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results, instead of rejection rate (RR) results.
+Additionally, this analysis was moved to the appendix so the main body more clearly focuses on our contributions.
 }

### 2022-11-19 ba_resubmit/kundinger_resubmit.tex
final revisions

@@
 \begin{abstract}
-%Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
-\textcolor{red}{Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.} However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
+%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
+Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.
+However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
 To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
@@
 %
-%\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.
+%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.
 Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}
@@
 
-In many data analysis tasks, analysts seek to identify duplicate records across two datafiles.
+In many data analysis tasks, analysts seek to identify duplicate records across two data files.
 This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
-In this article, we consider bipartite record linkage, which merges two datafiles that contain duplications across, but not within, the respective datafiles.
+In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.
 
@@
 Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
-Additionally, some methods model records directly \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.
+Additionally, some methods directly model the linkage variables \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.
 
@@
 The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice.
-In many situations, we know that there are no duplications within a datafile, meaning that each record in one file should be linked with at most one other record in the other file.
+In many situations, we know that there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
@@
 \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on datafiles with more than 100 records.
+\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records.
 \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
@@
 This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy.
-In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.
+These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with minimal loss of accuracy.
+In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
 
@@
 
-Consider two datafiles $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
+Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
 Without loss of generality, denote files such that $n_1 \geq n_2$.
-We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in X_1$" rather than the more compact $x_{1i}$ \textcolor{red}{for ease of readability}.
-Under bipartite matching, the set of matches across datafiles can be represented in two equivalent ways.
+For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$.
+Under bipartite matching, the set of matches across data files can be represented in two equivalent ways.
 First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
@@
 Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
-We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.
+
+%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
 
 The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features.
+For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for record $i$ and $j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
 For numerical features, we can use absolute difference between the two feature values.
+For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for record $i$ and $j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for record $i$ and $j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
 For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
@@
 
-Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$.
-Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
+More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
+Here, $I_{f1}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
 Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
 
-\textcolor{red}{In the construction of comparison vectors, we note it is common to encounter missing information in datasets used for record linkage.
-Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We assume that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
-With the MAR, we can marginalize over the missing data, and do all computation simply using the observed data.
-To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}
+In the construction of comparison vectors, it is common to encounter missing information in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$.
+As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We assume that this missingness occurs at random (MAR, per \cite{LittleRubin2002}).
+With the MAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
 
@@
 
-The seminal \cite{fellegi_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
-These allow us to model the comparison data using a mixture model, where
+The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
+Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
 \begin{align}
@@
 \end{align}
-Here $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
+Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
 When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
@@
 
-\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
-\begin{align}
-	\label{eqn:wts}
-	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
-\end{align}
-This ratio is large when there is strong evidence of the pair being a match, and small otherwise.
-All record pairs with weights above a specified threshold are considered a match.
-Such independent classifications often leads to matchings that violate one-on-one matching assumptions, which prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi Sunter weights among matched pairs.
+%\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
+%\begin{align}
+%	\label{eqn:wts}
+%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
+%\end{align}
+%This ratio is large when there is strong evidence of the pair being a match, and small otherwise. All record pairs with weights above a specified threshold are considered a match.
+
+Within this framework, record pairs are independently classified as matches and non-matches based on the estimated parameters.
+However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results.
+To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching.
+\cite{sadinle_bayesian_2017} later showed that this method could be thought of as a maximum likelihood estimate.
+
+%This prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi-Sunter weights among matched pairs.
 
@@
 
-%The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+%The Fellegi-Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
 This is a problem in practice as transitive closures can be violated.
@@
 %
-%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs.
+%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs.
 Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
@@
 %
-%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
 This is a problem in practice as transitive closures can be violated.
@@
 
-\textcolor{red}{To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a Bayesian framework.
-In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposed the "beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
-To construct this prior, he assigned a prior to the probability of the indicator that a record in $\bm{X}_2$ had a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
+To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a Bayesian framework.
+In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposed the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
+He assigns a prior distribution for the probability of the indicator that a record in $\bm{X}_2$ has a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
 It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$.
 Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 !
-/ (n_1 - n_{12}(\bm{Z})!$ bipartite matchings are taken to be equally likely.} Thus the beta distribution for bipartite matching, is given by
+/ (n_1 - n_{12}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
+Thus, the beta distribution for bipartite matching, is given by
 \begin{align}
@@
 where $\text{B}(\cdot, \cdot)$ represents the Beta function.
-This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling \(Z_j\).
+This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each \(Z_j\).
 This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
@@
 However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior.
-We denote the fast Beta prior as follows:
+We denote the fast Beta prior as follows.
+For each $Z_j$, we have
 \begin{align}
 	\label{eqn:fast_beta_prior}
-	Z_j | \pi &= \begin{cases} 
+	p(Z_j = z_j| \pi) &= \begin{cases} 
 	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
@@
 \end{align}
-We can interpret equation \ref{eqn:fast_beta_prior} as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match.
+We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match.
 The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$
 
-\textcolor{red}{Previously, \cite{wortman2019} proposed a flat prior over the labeling of $\bm{Z}$, meaning that each value $\{1, \ldots, n_1, n_1 +j\}$ was considered equally likely for $Z_j$.
-This however amounts to a prior probability of $\frac{n_1}{n_1 + 1}$ that record $j$ has a match in $\bm{X}_1$, and thus tended to over match, leading to poor precision when used in practice.
-Indeed we find that the flat prior is equivalent to a special case of the fast Beta prior where $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.}
+In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_1, n_1 +j\}$ is a priori equally likely for $Z_j$.
+This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$.
+In our simulation studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}).
+We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.
 
@@
 
-Note that linkage in this setting is conducted at the record level, rather than at the record pair level as in the Fellegi Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in the Fellegi Sunter model estimates the proportion of record pairs that are matches.
-We find $\pi$ to be more interpretable parameter than $\lambda$ in the bipartite case.
+Note that linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
+We find $\pi$ to be more a interpretable parameter than $\lambda$ in the bipartite case.
 In this setting, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions.
-As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to achieve the benefits from \texttt{BRL} while making efficiency gains possible by exploiting independence.
-
-%Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).
-To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure explained in Section \ref{bayes-estimate}.
-
-\textcolor{red}{To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}, reproduced in Appendix .
+Additionally, while the Fellegi-Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions.
+As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
+
+%To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}, reproduced in Appendix \ref{bayes-estimate}.
 Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-To achieve a Bayes estimate that fulfills one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the highest posterior probability, and declare the other to have no match.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.}
-
-%It is common to encounter missing information in datasets used for record linkage.
-Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We assume that this missingness occurs at random (MAR, by \cite{LittleRubin2002}).
-With the MAR, we can marginalize over the missing data, and do all computation simply using the observed data.
-To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.
-
-%this independent prior for each \(Z_j\), induces a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains over the Gibbs sampler under \texttt{BRL}.
-More importantly, since only the agreement pattern of \(Z_j\) is used forcalculations within the Gibbs sampler, and not the particular record label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.
+To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
+A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
+
+To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
+Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
 
@@
 For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_1}.
-\begin{align}
-	\label{eqn:full_model}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \notag \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \notag\\
-	Z_j | \pi &=
+\begin{subequations}
+\begin{align}
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \label{eqn:m} \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}), \label{eqn:u}\\
+	p(Z_j = z_j| \pi)  &=
 	\begin{cases} 
@@
 		1-\pi, &  z_j  = n_1 + j; \\
-	\end{cases} \notag\\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\notag.
-\end{align}
+	\end{cases} \label{eqn:z}\\
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
+\end{align}
+\end{subequations}
 
@@
 We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across datasets; that is, $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
-We then sample $\bm{m}$ and $\bm{u}$ from their full conditionals:
-$$\bm{m}_f|\Gamma, \bm{Z} \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
-$$\bm{u}_f|\Gamma, \bm{Z} \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
-where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j = i)$, and $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j \neq i)$.
+To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^s$, we use the full conditionals,
+\begin{subequations}
+\begin{align}
+	\bm{m}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
+	\bm{u}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update}
+\end{align}
+\end{subequations}
+where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} \neq i)$.
 
@@
 \label{eqn:z_full_conditional}
-p\left(Z_j^{(s+1)}  = i|\Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+p\left(Z_j^{(s+1)}  = z_j|\Gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
 \begin{cases} 
 	w_{z_j, j},  & z_j \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
+	n_1 \frac{n_2 - n_{12}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{12}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
 \end{cases}
 \end{align}
-where 
-$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
+where, for all $i \in \bm{X}_1$, 
+\begin{align}
+	\label{eqn:fs_weight}
+	w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+\end{align}
 Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.
@@
 
-The scale of linkage tasks possible through \texttt{BRL} is limited by (1) the memory costs of storing $n_1 \times n_2$ comparison vectors for every pair of records across the two datafiles, and (2) the speed of the linkage algorithm over those comparison vectors.
-One common solution to these challenges is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{christen2019data}.
-In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
-The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-\textcolor{blue}{DELETE?
-However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}}.
-
-Another technique is indexing, in which the modeller decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
+The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_1 \times n_2$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors.
+One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}.
+In deterministic blocking, the modeler chooses fields thought to be highly reliable, and only compares records that agree on those fields.
+The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains.
+Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
+
+After computing all comparison vectors within a block, the modeller can further reduce the number of comparison vectors used in the linkage algorithm through indexing.
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like first name, or only pairs that exactly match on a specified number of fields.
-\textcolor{blue}{DELETE?
-However, this kind of indexing requires knowledge  of the application and invites room for human error \citep{murray2016probabilistic}.}
-
-With \texttt{fabl}, we introduce two techniques \textcolor{red}{further expand} the scalability of probabilistic record linkage \textcolor{red}{in addition to} blocking and indexing.
-First, we propose hashing methods that allow us to compute sufficient statistics which reduce the computational complexity of the Gibbs sampler.
+However, the impact of indexing on model parameters is not well understood; \citep{murray2016probabilistic} reviewed this issue in the context of the frequentist Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
+%Another technique is indexing, in which the modeller decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
+
+With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
+First, we propose hashing methods that allow us to compute sufficient statistics that reduce the computational complexity of the Gibbs sampler.
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
-
-%To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method.
+For convenience, Table \ref{table_notation_2} summarizes the notation introduced throughout this section.
+
+\begin{table}[t!]
+	\centering
+	\begin{tabular}[t!]{ll}
+		Symbol & Description \\
+		\hline
+		$h_p$ & one hot encoding of agreement pattern $p$ \\
+		$(i,j) \in h_p$ & comparison vector between records $i \in \bm{X}_1$ and $i \in \bm{X}_2$ exhibits pattern $p$ \\
+		$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+		$N_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+		$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+		$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+		$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
+		\hline
+	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_2}
+\end{table}
+
+%To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi-Sunter method.
 The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
@@
 Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.
+Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi-Sunter comparison vectors, to aid in vectorized computations.
 See Appendix \ref{app:ohe} for details.
@@
 
-Following insight from \cite{enamorado2019using}, we recognize that record pairs contribute to posterior calculations only through the agreement pattern $\gamma_{ij}$.
-To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+Following an insight from \cite{enamorado2019using}, we recognize that record pairs contribute to posterior calculations only through their agreement pattern as represented by $\gamma_{ij}$.
+To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns observed in $\bm{X}_1 \times \bm{X}_2$, and let $\mathcal{P} = \{h_1, \ldots, h_P\}$.
 Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns.
@@
 First, we hash record pairs of the same agreement pattern to unique integer values. \cite{enamorado2019using} accomplished this efficiently through the hashing function
-$$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}.$$
+\begin{align}
+\label{eqn:hashing}
+h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}.
+\end{align}
 This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
-For computational ease, we then map these integers to sequential integers from $\{1, \ldots, P\}$ corresponding to the enumerated patterns in $\mathcal{P}$.
-When the $(i, j)$ pair exhibits the $p^{th}$ pattern, we say $(i, j) \in h_p$.
-\textcolor{red}{When $j$ has no match in $\bm{X}_1$, we write $(n_1 + j, j) \in h_{P+1}$.}
-
-\textcolor{red}{With all record pairs converted to hashed values, we identify the records in $\bm{X}_1$ with comparison vectors corresponding to each pattern $p$ for each record $j \in \bm{X}_2$.
-We denote this set $r_{j_p} = \{i \in \bm{X}_1 | (i, j) \in h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{j_p} | j \in \bm{X}_1, p \in 1, \ldots, P\}$.}This representation is useful because it allows us to easily compute two sets of sufficient statistics.
+For computational ease, we map these integers to sequential integers from $\{1, \ldots, P\}$ corresponding to the enumerated patterns in $\mathcal{P}$.
+When the $(i, j)$ pair exhibits pattern $p$, we say $(i, j) \in h_p$.
+When $j$ has no match in $\bm{X}_1$, we write $(n_1 + j, j) \in h_{P+1}$.
+
+With all record pairs converted to hashed values, we identify the records in $\bm{X}_1$ with comparison vectors corresponding to each pattern $p$ for each record $j \in \bm{X}_2$.
+We denote this set $r_{j_p} = \{i \in \bm{X}_1 | (i, j) \in h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{j_p} | j \in \bm{X}_2, p \in 1, \ldots, P\}$.
+This representation is useful because it allows us to easily compute two sets of sufficient statistics.
 First, we compute the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, given by
-$$N_{j_p} = |r_{j_p}| = \sum_{i=1}^{n_1} I((i, j) \in h_p).$$
+\begin{align}
+\label{eqn:N}
+N_{j_p} = |r_{j_p}| = \sum_{i=1}^{n_1} I((i, j) \in h_p).
+\end{align}
 Second, we compute the number of comparison vectors in $\Gamma$ that exhibit pattern $p$, given by
-$$H_p = \sum_{j = 1}^{n_2} N_{j_p}.$$
-\textcolor{red}{We collect these sufficient statistics in the sets $\mathcal{N} = \{N_{j_p} |j \in \bm{X}_1, p \in 1, \ldots, P \}$ and $\mathcal{H} = \{H_{p} | p \in 1, \ldots, P \}$.} As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
+\begin{align}
+	\label{eqn:H}
+	H_p = \sum_{j = 1}^{n_2} N_{j_p}.
+\end{align}
+We collect these sufficient statistics in the sets $\mathcal{N} = \{N_{j_p} |j \in \bm{X}_2, p \in 1, \ldots, P \}$ and $\mathcal{H} = \{H_{p} | p \in 1, \ldots, P \}$.
+As we show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
 Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we can use $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}, \mathcal{H}\}$ for posterior inference.
@@
 
-%In practice, it is not feasible to make all-to-all record comparisons because creating and storing $n_1 \times n_2$ comparison vectors is computationally infeasible for large datafiles.
+%In practice, it is not feasible to make all-to-all record comparisons because creating and storing $n_1 \times n_2$ comparison vectors is computationally infeasible for large data files.
 The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
@@
 That is, instead of storing \(n_1 \times n_2\) comparison vectors, we only store the \(P\) unique vectors, and \(n_1 \times n_2\) scalar quantities relating record pairs to those vectors.
-However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large datafiles.
+However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large data files.
 Worse, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
@@
 
-With storage efficient indexing (SEI), we can calculate $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs, but greatly reduce the memory costs associated with unlikely matches, allowing us to conduct all-to-all comparisons for substantially larger linkage tasks.
-Observe that all records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same $w_{ij}$.
-Therefore, these records have the same probability to be identified as a link when sampling \(Z_j\).
+We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs, but greatly reduce the memory costs associated with unlikely matches.
+This allows all-to-all comparisons for substantially larger linkage tasks.
+All records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same $w_{ij}$.
+Therefore, these records have the same probability to be identified as the link for record $j$.
 Thus, we know that records $i \in r_{j_p}$ such that $N_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
 We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
-Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested list $\mathcal{R}^{\text{SEI}}$.
-Instead of storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
-In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for plausible matches.
-Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results.
+Therefore, rather than store all of these record labels, we store only a small number \(S\) of them in a new nested list $\mathcal{R}^{\text{SEI}}$.
+Instead of storing \(n_1 \times n_2\) record labels, with SEI we store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
+In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but is not likely to eliminate any of the indices for plausible matches.
+Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on few matches and distort linkage results.
 Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}, \mathcal{H}\}$.
 
-%For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.
-
-%\hypertarget{distributed}{%
-	%	\subsection{Distributed Computation for Pairwise Comparison}\label{distributed}}
-
-For datafiles with large amounts of records, we can partition the two datasets \(\bm{X}_1\) and \(\bm{X}_2\) into smaller chunks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more manageable computations.
-\textcolor{red}{On a single machine, we can read-in data and construct comparison vectors sequentially.
-Then we can conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.} With multiple cores or multiple machines, this can be done in parallel.
-Summary statistics from each pairwise chunk comparison can  be easily synthesized to recover sufficient statistics for the $\Gamma$.
+%For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations.
+For each chunked data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each chunked data file $\bm{X}_{2d}$ to construct the chunked comparison matrix $\Gamma^{cd}$.
+We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.
+These all-to-all comparisons can be computed serially or in parallel.
+Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
 Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger datasets.
+
+For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller disjoint chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations.
+For each data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each $\bm{X}_{2d}$ to construct the comparison matrix $\Gamma^{cd}$.
+We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.
+In detail, we calculate
+\begin{subequations}
+	\begin{align}
+		r_{j_p}^{cd} &= \{i \in \bm{X}_{1c} | (i, j) \in h_p, j \in \bm{X}_{2d}\}, \label{r_{j_p}^{cd}} \\
+		N_{j_p}^{cd} &= |r_{j_p}^{cd}|, \label{N_{j_p}^{cd}} \\
+		H_p^{cd} &= \sum_{j \in \bm{X}_{2d}} N_{j_p}.\label{H_p^{cd}}
+	\end{align}
+\end{subequations}
+These can computed serially or in parallel.
+We reduce memory costs through SEI on each $r_{j_p}$, while retaining counts of each agreement pattern through $N_{j_p}^{cd}$ and $H_p^{cd}$.
+Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
+Suppressing SEI notation, we combine information through
+\begin{subequations}
+	\begin{align}
+		r_{j_p} &= (r_{j_p}^{11}, \ldots, r_{j_p}^{cd}) \text{ for } c = 1, \ldots, C \text{ and } d = 1, \ldots, D,\label{r_{j_p}} \\
+		N_{j_p} &= \sum_{c = 1}^C \sum_{d = 1}^D N_{j_p}^{cd},\label{N_{j_p}} \\
+		H_p &= \sum_{c = 1}^C \sum_{d = 1}^D H_p^{cd}.\label{H_p}
+	\end{align}
+\end{subequations}
 
@@
 Updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum_f L_f\) many summations over \(n_1 \times n_2\) quantities.
-These are simple calculations, but become computationally burdensome when working on large linkage tasks.
-
-Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations.
-\textcolor{red}{Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p, Z_j \leq n_1 \right)$ to be the number of matching record pairs with agreement pattern $p$.} It follows that the number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
-Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors  of length $\sum_f L_f$ of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
-$$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p,$$
-$$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
-
-Although sampling \(Z_j\) from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large.
+These are simple calculations, but they become computationally burdensome when working on large linkage tasks.
+
+Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$.
+Let $H_p^m = \\ \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p, Z_j \leq n_1 \right)$ be the number of matching record pairs with agreement pattern $p$.
+It follows that the number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
+Let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_f})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define  $\bm{\beta_0} = (\beta_{11}, \ldots, \beta_{F L_f})$ similarly for the $\bm{u}$ distributions.
+Then, the terms needed for the posterior updates in (\ref{eqn:m_update}) and (\ref{eqn:u_update}) are contained in the vectors
+\begin{subequations}
+\begin{align}
+\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p, \label{efficient_alpha} \\
+\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p. \label{efficient_alpha}
+\end{align}
+\end{subequations}
+
+Although sampling \(Z_j\) from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_1$ is large.
 This is because sampling a value from $n_1$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
 To speed up computation, we break this sampling step into two.
-First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link, according to (\ref{eqn:gibbs1}).
+First, we calculate the quantity expressed in (\ref{eqn:fs_weight}) associated with each unique pattern, and denote this quantity as $w_{p}$.
+We then sample the agreement pattern between \(j\) and its potential link, according to
 \begin{align}
@@
 		w_{p}\times N_{j_p},  & p \leq P; \\
-		n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, &   p = P + 1. \\
+		n_1 \frac{n_2 - n_{12}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{12}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   p = P + 1. \\
 	\end{cases}
@@
 Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns.
-\textcolor{red}{Since all records in $\bm{X}_1$ sharing the same agreement pattern with $j \in \bm{X}_2$ are equally likely, we can sample among candidate records uniformally through (\ref{eqn:gibbs2}).}
+Since all records in $\bm{X}_1$ sharing the same agreement pattern with $j \in \bm{X}_2$ are equally likely, we can sample among candidate records uniformally using
 
@@
 		\label{eqn:gibbs2}
-		P\left(Z_j^{(s+1)} = i \mid \left(Z_j^{(s+1)}, j\right) \in h_p \right) = \begin{cases} 
+		P\left(Z_j^{(s+1)} = z_j \mid \left(Z_j^{(s+1)}, j\right) \in h_p \right) = \begin{cases} 
 			\frac{1}{N_{j_p}}, & (z_j, j) \in h_p; \\
@@
 
-We emphasize the computational gains of this split sampler: the first step is a sample from $P + 1$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
-These changes greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational gains.
-We provide a summary of the notation used for the hashing procedure in Table \ref{table_notation_2}.
+We emphasize the computational gains of this split sampler: the first step is a sample from $P + 1$ options, where $P$ does not scale with the size of the linkage task; the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
+These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
+
 The computational complexity of \texttt{fabl} is given in Lemma \ref{lemma:fabl}.
 %, and a summary of the \texttt{fabl} method through psuedocode in Appendix \ref{app:fabl-psuedocode}
-
-	\begin{table}[t!]
-		\centering
-		\begin{tabular}[t!]{ll}
-			Symbol & Description \\
-			\hline
-			$h_p$ & one hot encoding of agreement pattern $p$ \\
-			$(i,j) \in h_p$ & comparison vector between records $i \in \bm{X}_1$ and $i \in \bm{X}_2$ exhibits pattern $p$ \\
-			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-			$N_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-			$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
-			$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
-			$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
-			\hline
-		\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_2}
-	\end{table}
 
@@
 		Recall that $n_1$ and $n_2$ are the number of records in $\bm{X}_1$ and $\bm{X}_2$, respectively.
-Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors can exhibit.
-We assume $B$ processors available for parallelization.
-Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
+Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $\bm{X}_1 \times \bm{X}_2$.
+We assume $B$ processors available for parallelization and a Gibbs sampler with $T$ iterations.
+Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(\frac{T}{B}n_2 P)$.
 		\label{lemma:fabl}
@@
 First, we sample the agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
-Thus, the complexity of sampling $\bm{Z}$ is $O(n_2 P)$.
-Since $P << n_1$ in most applications, we have reduced the complexity of the Gibbs sampler from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
-		In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
+Thus, the complexity of sampling $\bm{Z}$ in a single iteration is $O(n_2 P)$.
+Since $P << n_1$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
+With parallelization, this complexity is further reduced to $O(\frac{1}{B}n_2 P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{B}n_2 P)$
+		In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(\frac{T}{B}n_2 P).$
 	\end{proof}
@@
 We use $F = 5$ binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
-\textcolor{red}{For each record in $\bm{X}_2$, we simulate $n_1$ comparison vectors, resulting in a comparison matrix $\Gamma \in \mathbb{R}^{n_1 n_2 \times F}$.
+For each record in $\bm{X}_2$, we simulate $n_1$ comparison vectors, resulting in a comparison matrix $\Gamma \in \mathbb{R}^{n_1 n_2 \times F}$.
 For $n_2/2$ of these records, there is no match in $\bm{X}_1$, so we simulate $n_1$ comparison vectors from the $\bm{u}$ distribution.
@@
 We compare the run-time of \texttt{fabl} against \texttt{BRL} as we increase $n_1$ and $n_2$.
-Note that since we have five binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.}
+Since we have five binary comparison fields, the number of unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
 
@@
 		\begin{tabular}{lll|ll}
-			\multicolumn{1}{c}{ } & \multicolumn{2}{c}{m} & \multicolumn{2}{c}{u} \\
+			\multicolumn{1}{c}{ } & \multicolumn{2}{c}{$\bm{m}$} & \multicolumn{2}{c}{$\bm{u}$} \\
 			\cline{2-3} \cline{4-5}
@@
 		\end{tabular}
-		\caption{Probabilities used for $m$ and $u$ distributions in simulation study in Section \ref{speed}.}\label{Tab:distributions}
+		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Section \ref{speed}.}\label{Tab:distributions}
 	\end{table}
 
-	We note here that the implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
+	The implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
 In contrast, we use non-optimized code written in \texttt{R} for \texttt{fabl}.
 While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
-Additionally, although \texttt{fabl} is amenable to parallelization, this simulation was run on a single core.
+Additionally, although \texttt{fabl} is amenable to parallelization, this simulation is run on a single core.
 Implementing \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
 
-	In Figure \ref{fig:speed1}, where we increase both $n_1$ and $n_2$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger data.
+	In Figure \ref{fig:speed1}, where we increase both $n_1$ and $n_2$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger sample sizes.
 In particular, run-time for \texttt{BRL} grows quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} grows linearly (in the size of only $\bm{X}_2$).
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$, \textcolor{red}{as described in Section \ref{speed}.} We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$, as described in Section \ref{speed}.
+We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
 		\end{center}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_2$ fixed at 500, \textcolor{red}{as described in Section \ref{speed}.} We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_2$ fixed at 500, as described in Section \ref{speed}.
+We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
 		\end{center}
@@
 	\hypertarget{accuracy}{%
-		\subsection{Accuracy under Full Estimates}\label{accuracy}}
-
-	We replicate a simulation study from \cite{sadinle_bayesian_2017} used to illustrate the effectiveness of \texttt{BRL}.
-We adapt it to compare the accuracy of \texttt{fabl} to \texttt{BRL}.
-The simulations employ a collection of synthetic synthetic datasets with varying amounts of error and overlap (the number of records in common across files).
+		\subsection{Accuracy}\label{accuracy}}
+
+	Computational speed-ups are only worthwhile if not accompanied by a notable loss of record linkage accuracy.
+Therefore, we examine the accuracy of \texttt{fabl} relative to \texttt{BRL} by replicating a simulation study from \cite{sadinle_bayesian_2017}.
+The simulations employ a collection of synthetic datasets with varying amounts of error and overlap (the number of records in common across files).
 Following methods proposed by \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
@@
 Age and occupation are distorted through keyboard errors and missingness.
+These synthetic datasets are available in the supplement to \cite{sadinle_bayesian_2017}.
 
@@
 Under each of these settings, we have 100 pairs of simulated datasets in order to obtain uncertainty quantification on our performance metrics.
-We use flat priors for all $\bm{m}$ and $\bm{u}$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
-We calculate Bayes estimates of the linkage structure using the losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$.
+We use uniform priors for all $\bm{m}$ and $\bm{u}$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
+We run the Gibbs sampler for 1000 iterations, and discard the first 100 as burn-in.
+We calculate Bayes estimates $\hat{\bm{Z}}$ of the linkage structure using the  loss function and post-processing procedure described in Appendix \ref{bayes-estimate}.
 Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
-We replicate this simulation under partial estimates, allowing the \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review, in Appendix \ref{partial}.
+We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; these results are in Appendix \ref{partial}.
 
@@
 In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse on average; however, the overall accuracy level remains high.
+In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.
 
@@
 For each level of overlap and each level of error, we have 100 paired sets of 500 records.
-Thus this table represents results for 900 datasets.
+Thus this table summarizes results for 900 datasets.
 We see comparable performance for all levels of error and overlap.}
@@
 	%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
-Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
+Under standard Fellegi-Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
 Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task.
@@
 Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study (NLTCS), a larger linkage task that is not feasible in reasonable time under \texttt{BRL} with typical computing setups.
+In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study, a larger linkage task that is not feasible in reasonable time under \texttt{BRL} with typical computing setups.
 
@@
 	The country of El Salvador was immersed in civil war from 1980 to 1991, and we are interested in estimating the total number of casualties from the war.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to this data.} The ERTL dataset consists of digitized denunciations that had been published throughout the conflict, and the CDHES dataset consists of casualties that had been reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
+We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
 The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both datasets are thought to be fairly reliable.
@@
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We flat priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
+We use uniform priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
 
@@
 
-	\begin{table}
+	\begin{table}[t]
 		\begin{tabular}[t]{llllll}
@@
 		\end{tabular}
-		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle_bayesian_2017}.
+		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \cite{sadinle_bayesian_2017}.
 This set up leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
@@
 
-	Therefore, we perform our analysis with the agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
-Among the 216 possible agreement patterns, 159 are realized in the data.
-With this revised comparison specification, \texttt{fabl} ran in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
-The estimates of the $\bm{m}$ parameters under each method are similar, as shown in Figure \ref{fig:m-and-u}.
-Estimates of $\bm{u}$ are indistinguishable, and thus omitted.
-Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
-
-	For completeness, we note that linkage with the more detailed comparison vectors required 240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
-Apparently, the number of patterns was sufficiently many that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
-
-	\begin{table}
+	\begin{table}[t]
 	\centering
@@
 	\end{tabular}
-		\caption{Construction of comparison vectors for El Salvador for increased speed under \texttt{fabl}. This set up leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
+	\caption{Construction of comparison vectors for El Salvador data for increased speed under \texttt{fabl}.
+This set up leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
 \end{table}
 
+	Therefore, we perform our analysis with the agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
+Among the 216 possible agreement patterns, 159 are realized in the data.
+With this revised comparison specification, \texttt{fabl} runs in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
+The estimates of the $\bm{m}$ parameters under each method are similar, as shown in Figure \ref{fig:m-and-u}.
+Estimates of $\bm{u}$ are indistinguishable, and thus omitted.
+Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+
+	For completeness, we note that linkage with the more detailed comparison vectors requires 240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
+Apparently, the number of patterns is sufficiently many that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
+
 	Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets.
-We calculate posterior samples of the size of the overlap across files by finding the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
+We calculate posterior samples of the size of the overlap across files by finding the number of links in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
 The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate identifies fewer matches than the Gibbs sampler identifies on average.
 This is because a large number of records in ERTL have multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
-We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Appendix \ref{partial}.
+We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both data files, and a posterior 95\% credible interval of (211, 244).
+See Figure \ref{fig:overlap-plot} for a visual comparison of the Bayes estimates and posterior credible intervals for the two methods.
+We note that Bayes estimates falling outside of posterior credible intervals has been observed previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
+
+	%We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Appendix \ref{partial}.
 Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
@@
 
-	We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.
-See Figure \ref{fig:overlap-plot} for a visual comparison of the Bayes estimate and posterior credible intervals for the two methods.
-We note that Bayes estimates falling outside of posterior credible intervals has been observed previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
+	%and a range of (140, 294) after the partial estimate and clerical review.
 
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
+			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
 			\label{fig:m-and-u}
@@
 	\label{nltcs}
+
+	\begin{table}[t]
+		\centering
+		\begin{tabular}[t]{lllll}
+			\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
+			\cline{3-5}
+			Fields & Similarity & 1 & 2 & 3\\
+			\hline
+			Sex & Binary & Agree & Disagree & \\
+			Year of Birth & Binary & Agree & Disagree & \\
+			Month of Birth & Binary & Agree & Disagree & \\
+			Day of Birth & Binary & Agree & Disagree & \\
+			Location & Custom & Same State and Office & Same State & Otherwise \\
+			\hline
+		\end{tabular}
+		\caption{Construction of comparison vectors for NTLCS data}\label{Tab:nltcs-comparisons}
+	\end{table}
 
@@
 Hence, the assumptions of our model hold for this study.
-We seek to link records over the $n_1 = 20485$ records from 1982 to the $n_2 = 17466$ records from 1989.
+We seek to link records over the $n_1 = 20485$ individuals from 1982 to the $n_2 = 17466$ individuals from 1989.
 The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
-However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone datafiles.
+However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone data files.
 
@@
 Traceplots do not suggest convergence issues, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
-
-	\begin{table}
-		\centering
-		\begin{tabular}[t]{lllll}
-			\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
-			\cline{3-5}
-			Fields & Similarity & 1 & 2 & 3\\
-			\hline
-			Sex & Binary & Agree & Disagree & \\
-			Year of Birth & Binary & Agree & Disagree & \\
-			Month of Birth & Binary & Agree & Disagree & \\
-			Day of Birth & Binary & Agree & Disagree & \\
-			Location & Custom & Same State and Office & Same State & Otherwise \\
-			\hline
-		\end{tabular}
-		\caption{Construction of comparison vectors for NTLCS data}\label{Tab:nltcs-comparisons}
-	\end{table}
 
@@
 	In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that extends the work of \cite{sadinle_bayesian_2017} to scale to large data sets.
-We have proven that the proposed hashing method and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger dataset.
-This makes \texttt{fabl} computationally advantageous in many linkage scenarios, particularly when one datafile is substantially smaller than the other.
-We have also shown that storage efficient indexing, in tandem with hashing, greatly reduces the memory costs required for all-to-all comparisons, giving practictioners an option for larger record linkage tasks without the use of blocking or indexing.
-We have demonstrated the speed and accuracy of \texttt{fabl} by replicating a simulation study and a case study in \cite{sadinle_bayesian_2017}, and through an additional case study that is computationally infeasible under \texttt{BRL}.
-
-	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, the computing all $n_1 \times n_2$ record pairs still can be infeasible for larger linkage tasks.
+We have proven that the proposed hashing method and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger data file.
+This makes \texttt{fabl} computationally advantageous in many linkage scenarios, particularly when one data file is substantially smaller than the other.
+We have also shown that storage efficient indexing, in tandem with hashing, greatly reduces the memory costs required for all-to-all comparisons, giving practitioners an option for larger record linkage tasks potentially even without the use of blocking or indexing.
+We have demonstrated the speed and accuracy of \texttt{fabl} by replicating a simulation study and a case study in \cite{sadinle_bayesian_2017}, and through an additional case study that is computationally impractical under \texttt{BRL}.
+
+	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing all $n_1 \times n_2$ record pairs still can be prohibitive for larger linkage tasks.
 Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
@@
 Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task.
-Thus in future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to arbitrarily large linkage tasks.
+In future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to arbitrarily large linkage tasks.
 
@@
 	%
-	%To our knowledge, \texttt{fabl} is the first method within the Fellegi Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
+	%To our knowledge, \texttt{fabl} is the first method within the Fellegi-Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
 This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
@@
 
-	\section*{Appendix}
+	\section{Appendix}
 	\label{sec:appendix}
@@
 
-	We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage
-	parameter \(\bm{Z}\) by assigning different positive losses to
-	different types of errors, and minimizing posterior expected loss. We
-	adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which
-	\(\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\}\), with \(R\) representing the
-	option to leave the matching undetermined by the model. Specifically, we have
-
+We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage parameter \(\bm{Z}\) by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
+We adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which \(\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\}\), with \(R\) representing the option to leave the matching undetermined by the model.
+Specifically, we have
 	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
@@
 	\end{cases}$$
+Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
+Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
+
 When we seek a partial estimate of the linkage structure, leaving a portion of record pairs to be classified manually in clerical review, we adopt losses $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, .1)$.
-For a more in-depth explanation of this function and the induced Bayes
-	estimate, see \cite{sadinle_bayesian_2017}.
+For a more in-depth explanation of this function and the induced Bayes estimate, see \cite{sadinle_bayesian_2017}.
 
@@
 Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\):
+\newpage
 \begin{align*}
-	p&\left(Z_j^{(s+1)}|\Gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
+	p&\left(Z_j^{(s+1)} = z_j|\Gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
 	&\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
@@
 For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
-
 \begin{align*}
-	p&\left(Z_j^{(s+1)}  = i| \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
+	p&\left(Z_j^{(s+1)}  = z_j| \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
 	&= \int_{\pi}p\left(Z_j^{(s+1)} | \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
@@
 
-By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
+In this section, we repeat the simulation study in Section \ref{accuracy} of the main text, allowing for clerical review rather than forcing all records to have or not have links.
+Specifically, by leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
 In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
@@
 		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
-		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on simulated datasets.
+		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on datasets in the simulation in Appendix \ref{partial}.
 We see poorer performance for \texttt{fabl} only in situations with high overlap.}
@@
 Traceplots across other settings exhibit similar behavior.
-Note that traceplots for $\bf{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
+Note that traceplots for $\bm{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
 
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of m parameter from simulation study in Section \ref{accuracy}}\label{fig:sim_m_trace}
+			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}}\label{fig:sim_m_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of u parameters from simulation study in Section \ref{accuracy}}\label{fig:sim_u_trace}
+			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}}\label{fig:sim_u_trace}
 		\end{center}
@@
 		\begin{tabular}{llll|lll}
-			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{m} & \multicolumn{3}{c}{u} \\
+			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{$\bm{m}$} & \multicolumn{3}{c}{$\bm{u}$} \\
 			\cline{2-4} \cline{5-7}
@@
 		\end{tabular}
-		\caption{Probabilities used for $m$ and $u$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
+		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
 	\end{table}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$.
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$.
 We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with increasing $n_1$ and $n_2$ fixed at 500.
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_1$ and $n_2$ fixed at 500.
 We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-			\caption{Traceplot for m parameter in El Salvador case study.} \label{fig:m_trace}
+			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} \label{fig:m_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-			\caption{Traceplot for u parameter in El Salvador case study.} \label{fig:u_trace}
+			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} \label{fig:u_trace}
 		\end{center}

### 2022-11-23 ba_resubmit/response_to_reviewers.tex
submitted revision

@@
 	Thank you for your positive feedback regarding our paper.
-We have clarified that we are motivation by the work of Sadinle and computational considerations.
-We have tried to make this more clear in both the abstract, introduction, and conclusion.}
+We have clarified that our work is an extension of Sadinle (2017) for enhanced speed and scalability.
+We have tried to make this more clear in both the abstract, introduction, and conclusion.
+We have also condensed the discussion of Fellegi and Sunter and moved the explanation of the Bayes estimate to the Appendix; we hope that these changes allow the reader to more clearly see our contributions.
+}
 

### 2022-11-23 ba_resubmit/kundinger_resubmit.tex
submitted revision

@@
 Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
-This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
+This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
 Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
@@
 \begin{keyword}%[class=MSC]
-\kwd{data fusion}
+\kwd{Bayesian methods}
 \kwd{distributed computing}
@@
 In many data analysis tasks, analysts seek to identify duplicate records across two data files.
-This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
+This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
 In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.
@@
 Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified each pair as a match or a non-match using a likelihood ratio test.
+Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data files and independently classified each pair as a match or a non-match using a likelihood ratio test.
 Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
@@
 Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space.
-This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with minimal loss of accuracy.
@@
 Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
+Finally, Section \ref{discussion} summarizes our contributions and highlights areas for further research.
 
@@
 \end{align}
-Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks.
+This sparse matrix representation can become cumbersome for large linkage tasks.
 More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that
@@
 
-%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.
+%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two data files.
 
@@
 Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
-\begin{align}
-	\label{eqn:fs_model}
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
-	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
-	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \notag
-\end{align}
+\begin{subequations}
+\begin{align}
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
+	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
+	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
+\end{align}
+\end{subequations}
 Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
@@
 To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching.
-\cite{sadinle_bayesian_2017} later showed that this method could be thought of as a maximum likelihood estimate.
+\cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.
 
@@
 
-To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a Bayesian framework.
-In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposed the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
+To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework.
+In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
 He assigns a prior distribution for the probability of the indicator that a record in $\bm{X}_2$ has a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
@@
 This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$.
-In our simulation studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}).
+In our preliminary studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}).
 We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.
@@
 		\hline
-		$\bm{X}_1, \bm{X}_2$ & datasets\\
+		$\bm{X}_1, \bm{X}_2$ & data files\\
 		$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
@@
 		\hline
-	\end{tabular}\caption{Summary of model notation}\label{table_notation_1}
+	\end{tabular}\caption{Summary of model notation.}\label{table_notation_1}
 \end{table}
@@
 
-We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across datasets; that is, $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
+We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
 To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^s$, we use the full conditionals,
@@
 
-We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs, but greatly reduce the memory costs associated with unlikely matches.
+We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs but greatly reduce the memory costs associated with unlikely matches.
 This allows all-to-all comparisons for substantially larger linkage tasks.
@@
 Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
-Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger datasets.
+Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger data files.
 
@@
 
-%	We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across datasets.
+%	We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across data files.
 and compare the run-time of \texttt{fabl} against \texttt{BRL}.
@@
 
-	The implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
-In contrast, we use non-optimized code written in \texttt{R} for \texttt{fabl}.
+	The Gibbs sampler in the implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
+In contrast, we use non-optimized code written only in \texttt{R} for \texttt{fabl}.
 While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
@@
 
-	In Figure \ref{fig:speed2}, where we fix \(n_2 = 500\), we see linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, and much more static run-time under \texttt{fabl}.
+	In Figure \ref{fig:speed2}, where we fix \(n_2 = 500\), we see near linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, and much more static run-time under \texttt{fabl}.
 The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
@@
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_2$ fixed at 500, as described in Section \ref{speed}.
-We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
+We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
 		\end{center}
@@
 Therefore, we examine the accuracy of \texttt{fabl} relative to \texttt{BRL} by replicating a simulation study from \cite{sadinle_bayesian_2017}.
-The simulations employ a collection of synthetic datasets with varying amounts of error and overlap (the number of records in common across files).
+The simulations employ a collection of synthetic data files with varying amounts of error and overlap (the number of records in common across files).
 Following methods proposed by \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
@@
 Age and occupation are distorted through keyboard errors and missingness.
-These synthetic datasets are available in the supplement to \cite{sadinle_bayesian_2017}.
+These synthetic data files are available in the supplement to \cite{sadinle_bayesian_2017}.
 
 	We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table \ref{Tab:sadinle_simulation_cutoffs}.
-Each simulation identifies matched individuals between two datasets, each with 500 records.
-We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets.
-Under each of these settings, we have 100 pairs of simulated datasets in order to obtain uncertainty quantification on our performance metrics.
+Each simulation identifies matched individuals between two data files, each with 500 records.
+We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across data files.
+Under each of these settings, we have 100 pairs of simulated data files in order to obtain uncertainty quantification on our performance metrics.
 We use uniform priors for all $\bm{m}$ and $\bm{u}$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
@@
 		\end{tabular}
-		\caption{Construction of comparison vectors for accuracy study with simulated datasets of Section \ref{accuracy}.}
+		\caption{Construction of comparison vectors for accuracy study with simulated data files of Section \ref{accuracy}.}
 		\label{Tab:sadinle_simulation_cutoffs}
@@
 For each level of overlap and each level of error, we have 100 paired sets of 500 records.
-Thus this table summarizes results for 900 datasets.
+Thus this table summarizes results for 900 data files.
 We see comparable performance for all levels of error and overlap.}
@@
 	%
-	%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
+	%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated data files from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
 Under standard Fellegi-Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
@@
 We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
-The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both datasets are thought to be fairly reliable.
+The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
 When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
-Instead, record linkage techniques must be used to merge datasets before analyzing the data \citep{lum2013applications}.
+Instead, record linkage techniques must be used to merge data files before analyzing the data \citep{lum2013applications}.
 
 	There are several challenges with these data.
-First, both datasets have been automatically digitized, which inherently leads to some degree of typographical error.
+First, both data files have been automatically digitized, which inherently leads to some degree of typographical error.
 Second, the only fields recorded are given name, last name, date of death, and place of death.
@@
 
-	Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets.
+	Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both data files.
 We calculate posterior samples of the size of the overlap across files by finding the number of links in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
@@
 Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
-This means that after clerical review, the number of individuals replicated across datasets would fall in the interval (136, 311), encapsulating the posterior credible interval.
+This means that after clerical review, the number of individuals replicated across data files would fall in the interval (136, 311), encapsulating the posterior credible interval.
 More or fewer records could be identified for clerical review by decreasing or increasing $\theta_R$.
@@
 		\end{tabular}
-		\caption{Construction of comparison vectors for NTLCS data}\label{Tab:nltcs-comparisons}
+		\caption{Construction of comparison vectors for NTLCS data.}\label{Tab:nltcs-comparisons}
 	\end{table}
 
-	The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients.
+	The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients \citep{steorts_bayesian_2016}.
 The initial survey began in 1982, with follow-up surveys taken approximately every five years.
 As such, patients are surveyed at most once in a given year, and many patients are surveyed across multiple years.
-In addition, patients can either drop out of the study, pass away, or enter as new patients \citep{steorts_bayesian_2016}.
+In addition, patients can either drop out of the study, pass away, or enter as new patients.
 Hence, the assumptions of our model hold for this study.
@@
 We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
-Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within datasets.
+Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within data files.
 Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
 	%
-	%To our knowledge, \texttt{fabl} is the first method within the Fellegi-Sunter comparison vector approaches with an interpretable parameter for the rate of matching across datasets.
+	%To our knowledge, \texttt{fabl} is the first method within the Fellegi-Sunter comparison vector approaches with an interpretable parameter for the rate of matching across data files.
 This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
@@
 Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold.
-Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller datasets.
+Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller data files.
 We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing.
@@
 However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
-Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
+Posterior analysis yields an initial estimate of 9991 individuals linked across the the two data files, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
 	%
@@
 %			\hline
-%			$\bm{X}_1, \bm{X}_2$ & datasets\\
+%			$\bm{X}_1, \bm{X}_2$ & data files\\
 %			$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
@@
 In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihood relevant to a particular record $j \in \bm{X}_2$.
-Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
+Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
 Thus, we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
@@
 		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
-		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on datasets in the simulation in Appendix \ref{partial}.
+		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
 We see poorer performance for \texttt{fabl} only in situations with high overlap.}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}}\label{fig:sim_overlap_trace}
+			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}}\label{fig:sim_m_trace}
+			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}}\label{fig:sim_u_trace}
+			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-			\caption{Traceplot for number of matches found across datasets in El Salvador case study.} \label{fig:overlap_trace}
+			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
 		\end{center}

### 2023-03-28 Prelim/latent_class_model.tex
stuff

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\begin{document}
	\title{Latent Class Models}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger}
	\date{\today}

%	\AtBeginSection[]{
%		\begin{frame}
%			\vfill
%			\centering
%			\begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%				\usebeamerfont{title}\insertsectionhead\par%
%			\end{beamercolorbox}
%			\vfill
%		\end{frame}
%	}

%	\begin{frame}
%	\frametitle{Outline}
%	\tableofcontents
%	\end{frame}

\section{Latent Class Models}

	\begin{frame}{Latent Class Models}
	\begin{itemize}
		\item In Dunson's STA 841 class "Categorical Data Analysis," we learned about a \emph{latent class models}
		\item In class, we used to them to classify DNA sequences, but I think they may be effective in classifying individuals based on their travel path in a network
		\item DNA sequences have complicated dependencies. The idea here is to model the sequences such that by conditioning on latent classes, the expressions in each position are independent.

		\item DISCLAIMER: I see many shortcomings of the model, but just wanted to share it as a potential idea
	\end{itemize}
	\end{frame}

\begin{frame}{Data}
	\begin{itemize}
		\item Individuals $i \in \{1, \ldots, I \}$
		\item Nucleotide position $p \in \{1, \ldots, P \}$
		\item Nucleotide expression  $y_{ip} = d \in \{A, C, G, T\}$
		\item Expert classification $e_i \in \{IE, EI, N\}$

		<2->\item We attempt to recover the expert classification through latent class $z_i = k \in \{1, \ldots, K\}$. Number of classes $K$ can be fixed or estimated nonparametrically
		<2->\item Vector of expression probabilities $\theta_{kp} = (\theta_{kpA}, \ldots, \theta_{kpT})$ for each class and position
	\end{itemize}
\end{frame}

\begin{frame}{Data}
	\begin{itemize}
		\includegraphics[width = \textwidth, height = .7\textwidth ]{data_snapshot.png}
	\end{itemize}
\end{frame}

\begin{frame}{Model Formulation}
	\begin{itemize}
		\item $y_{ip} \sim \text{Multinomial} (\theta_{z_i, p})$
		\item $z_i \sim \text{Dirichlet}(\alpha_1, \ldots, \alpha_K)$
		\item $\theta_{z_i, p} \sim \text{Dirichlet}(\beta_1, \ldots, \beta_D)$
		<2->\item $z_i$ can also be modeled through a Dirichlet process stick-breaking prior so that the number of clusters is inferred through the data
		<3-> Parameters are estimated through straightforward Gibbs sampler
	\end{itemize}
\end{frame}

\begin{frame}{Result}
	\begin{itemize}
		\includegraphics[width = \textwidth, height = .7\textwidth ]{latent_variable_plot.png}
	\end{itemize}
\end{frame}

\end{document}

### 2023-03-28 cluster_jobs/combine_results.R

@@ -1,0 +1,64 @@
+
+nltcs <- read.csv("../data/proc_nltcs.csv")
+true_matches <- 8866
+# df82 <- nltcs %>%
+#   filter(FILE == 82) %>%
+#   select(-FILE, -SEQ) %>%
+#   mutate(unique_ID = str_sub(REC, start = 3L),
+#          DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+#          DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+#          DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+#          STATE = str_pad(STATE, 2, pad = "0"),
+#          REGOFF = str_pad(REGOFF, 2, pad = "0"))
+#
+#
+# df89 <- nltcs %>%
+#   filter(FILE == 89) %>%
+#   select(-FILE, -SEQ) %>%
+#   mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+#   mutate(unique_ID = str_sub(REC, start = 3L),
+#          DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+#          DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+#          DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+#          STATE = str_pad(STATE, 2, pad = "0"),
+#          REGOFF = str_pad(REGOFF, 2, pad = "0"))
+#
+# df82 <- df82[-c(1:2000), ]
+# df89 <- df89[-c(1:2000), ]
+# sum(df82$unique_ID %in% df89$unique_ID)
+
+
+# nA <- nrow(df82)
+# nB <- nrow(df89)
+# Ztrue <- sapply(df89$unique_ID, function(x){
+#   match <- which(df82$unique_ID == x)
+#   if (length(match) == 0){
+#     match <- nA + 1
+#   }
+#   match
+# }) %>%
+#   unname()
+
+#sum(Ztrue < nA +1)
+
+evals <- list.files("eval", full.names = T)
+eval_df <- lapply(evals, readRDS) %>%
+  do.call(rbind, .)
+
+eval_df
+
+eval_df %>%
+  group_by(method) %>%
+  summarize(total_declared = sum(total_declared),
+            true_declared = sum(true_declared),
+            false_declared = sum(false_declared)) %>%
+  mutate(recall = true_declared / true_matches,
+         precision = true_declared / total_declared,
[…]

### 2023-03-28 cluster_jobs/combine_results_italy.R

@@ -1,0 +1,22 @@
+library(italy)
+
+train_set <- 1:1250
+id1 <- italy10$id[-train_set]
+id2 <- italy08$id[-train_set]
+
+true_matches <- sum(id2 %in% id1)
+
+
+evals <- list.files("eval_italy", full.names = T)
+eval_df <- lapply(evals, readRDS) %>%
+  do.call(rbind, .)
+
+eval_df %>%
+  group_by(method) %>%
+  summarize(total_declared = sum(total_declared),
+            true_declared = sum(true_declared),
+            false_declared = sum(false_declared)) %>%
+  mutate(recall = true_declared / true_matches,
+         precision = true_declared / total_declared,
+         fmeasure = 2 * (recall  * precision) / (recall + precision)) %>%
+  select(recall, precision, fmeasure)

### 2023-03-28 cluster_jobs/comp_data.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --array=1-78
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --job-name=comp_data
+#SBATCH --time=24:00:00
+#SBATCH --mem=128GB
+
+#
+# modules
+#
+module load R
+Rscript nltcs_comp_data.R

### 2023-03-28 cluster_jobs/comp_data_italy.R

@@ -1,0 +1,28 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(italy)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+train_set <- 1:1250
+id1 <- italy10$id[-train_set]
+id2 <- italy08$id[-train_set]
+
+df1 <- italy10 %>%
+  select(-id, -PARENT)
+
+df2 <- italy08 %>%
+  select(-id, -PARENT)
+
+df1 <- df1[-train_set, ]
+df2 <- df2[-train_set, ]
+
+job_marker <- rep(1:63, each =200)
+job_marker <- job_marker[1:nrow(df2)]
+chunk_marker <- job_marker == taskID
+
+chunk2 <- df2[chunk_marker, ]
+
+cd <- compare.linkage(df1, chunk2,
+                identity1 = id1, identity2 = id2[chunk_marker])
+saveRDS(cd, file = paste0("comp_vectors/", "cd_", taskID))

### 2023-03-28 cluster_jobs/comp_hash_nltcs.R

@@ -1,0 +1,46 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+nltcs <- read.csv("data/proc_nltcs.csv")
+df82 <- nltcs %>%
+  filter(FILE == 82) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+
+df89 <- nltcs %>%
+  filter(FILE == 89) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+# df82 <- df82[-c(1:2000), ]
+# df89 <- df89[-c(1:2000), ]
+
+job_marker <- rep(1:88, each =200)
+job_marker <- job_marker[1:nrow(df89)]
+chunk_marker <- job_marker == taskID
+
+chunk89 <- df89[chunk_marker, ]
+
+cd <- (BRL::compareRecords(df82[, -c(7, 8)], chunk89[, -c(7, 8)],
+                                     c(1, 2, 3, 4, 5, 6),
+                                     types = c("bi", "bi", "bi", "bi", "bi", "bi")))
+cd[[1]] <- apply(cd[[1]], 2, as.numeric)
+hash <- GetUniquePatterns2(cd, R = 10)
+saveRDS(hash, file = paste0("hash/", "hash_", taskID))

### 2023-03-28 cluster_jobs/nltcs_comp_data.R

@@ -1,0 +1,43 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+nltcs <- read.csv("data/proc_nltcs.csv")
+df82 <- nltcs %>%
+  filter(FILE == 82) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+  #unite(dob, 2:4, sep = "") %>%
+  #unite(location, c(STATE, REGOFF), sep = "")
+
+
+df89 <- nltcs %>%
+  filter(FILE == 89) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+df82 <- df82[-c(1:2000), ]
+df89 <- df89[-c(1:2000), ]
+
+job_marker <- rep(1:78, each =200)
+job_marker <- job_marker[1:nrow(df89)]
+chunk_marker <- job_marker == taskID
+
+chunk89 <- df89[chunk_marker, ]
+
+cd <- compare.linkage(df82[, -c(7, 8)], chunk89[, -c(7, 8)],
+                identity1 = df82$unique_ID, identity2 = chunk89$unique_ID)
+saveRDS(cd, file = paste0("comp_vectors/", "cd_", taskID))

### 2023-03-28 cluster_jobs/nltcs_fastlink.R

@@ -1,0 +1,38 @@
+
+nltcs <- read.csv("../data/proc_nltcs.csv")
+true_matches <- 8866 #depends on training set
+train_set <- 2000
+df82 <- nltcs %>%
+  filter(FILE == 82) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+
+df89 <- nltcs %>%
+  filter(FILE == 89) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+df82 <- df82[-c(1:2000), ]
+df89 <- df89[-c(1:2000), ]
+
+fl_full <- fastLink(df82[, -c(7, 8)], df89[, -c(7, 8)],
+                   varnames = names(df82[, -c(7, 8)]),
+                   dedupe.matches = F)
+
+fl_dedup <- fastLink(df82[, -c(7, 8)], df89[, -c(7, 8)],
+                    varnames = names(df82[, -c(7, 8)]),
+                    dedupe.matches = T)
+
+sum(df82$unique_ID[fl_out$matches$inds.a] == df89$unique_ID[fl_out$matches$inds.b])

### 2023-03-28 cluster_jobs/test.R

@@ -1,0 +1,30 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+
+#setwd("cluster_jobs")
+ApplyModel <- function(method){
+  model <- readRDS(paste0("trained_models/", method))
+  out <- classifySupv(model, cd)
+  out$pairs %>%
+    as.data.frame() %>%
+    mutate(prediction = out$prediction[]) %>%
+    summarize(total_declared = sum(prediction == "L"),
+              true_declared = sum(is_match == 1 & prediction == "L"),
+              false_declared = sum(is_match == 0 & prediction == "L")) %>%
+    mutate(method = method,
+           array_job = taskID)
+}
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+#taskID <- 3
+# model_paths <- list.files("trained_models", full.names = T)
+# models <- lapply(model_paths, readRDS)
+methods <- list.files("trained_models", full.names = F)
+cd_paths <- list.files("comp_vectors", full.names = T)
+cd_path <- cd_paths[taskID]
+cd <- readRDS(cd_path)
+
+results <- lapply(methods, ApplyModel)
+results_df <- do.call(rbind, results)
+saveRDS(results_df, file = paste0("eval/eval_", taskID))

### 2023-03-28 cluster_jobs/test.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --array=1-78
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --job-name=test
+#SBATCH --time=24:00:00
+#SBATCH --mem=128G
+
+#
+# modules
+#
+module load R
+Rscript test.R

### 2023-03-28 cluster_jobs/train.R

@@ -1,0 +1,42 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+methods <- c("svm", "rpart")
+method <- methods[taskID]
+nltcs <- read.csv("data/proc_nltcs.csv")
+df82 <- nltcs %>%
+  filter(FILE == 82) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+#unite(dob, 2:4, sep = "") %>%
+#unite(location, c(STATE, REGOFF), sep = "")
+
+
+df89 <- nltcs %>%
+  filter(FILE == 89) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+train82 <- df82[1:2000, ]
+train89 <- df89[1:2000, ]
+
+
+cd <- compare.linkage(train82[, -c(7, 8)], train89[, -c(7, 8)],
+                      identity1 = train82$unique_ID, identity2 = train89$unique_ID)
+
+model <- trainSupv(cd, method = paste(method))
+saveRDS(model, file = paste0("trained_models/", method))

### 2023-03-28 cluster_jobs/train_italy.R

@@ -1,0 +1,28 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(italy)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+methods <- c("svm", "rpart")
+method <- methods[taskID]
+train_set <- 1:1250
+id1 <- italy10$id[train_set]
+id2 <- italy08$id[train_set]
+
+df1 <- italy10 %>%
+  select(-id, -PARENT)
+
+df2 <- italy08 %>%
+  select(-id, -PARENT)
+
+df1 <- df1[train_set, ]
+df2 <- df2[train_set, ]
+
+
+cd <- compare.linkage(df1, df2,
+                      identity1 = id1, identity2 = id2)
+
+model <- trainSupv(cd, method = paste(method))
+saveRDS(model, file = paste0("trained_models/", method))

### 2023-04-03 R/MakeZhat_from_fastlink.R
new stuff

@@ -1,0 +1,5 @@
+MakeZhat_from_fastlink <- function(idA, idB){
+  Zhat <- rep(nA + 1, nB)
+  Zhat[idB] <- idA
+  Zhat
+}

### 2023-04-03 cluster_jobs/combine_results.R

@@ -1,64 +1,0 @@
-
-nltcs <- read.csv("../data/proc_nltcs.csv")
-true_matches <- 8866
-# df82 <- nltcs %>%
-#   filter(FILE == 82) %>%
-#   select(-FILE, -SEQ) %>%
-#   mutate(unique_ID = str_sub(REC, start = 3L),
-#          DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
-#          DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
-#          DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
-#          STATE = str_pad(STATE, 2, pad = "0"),
-#          REGOFF = str_pad(REGOFF, 2, pad = "0"))
-#
-#
-# df89 <- nltcs %>%
-#   filter(FILE == 89) %>%
-#   select(-FILE, -SEQ) %>%
-#   mutate(unique_ID = str_sub(REC, start = 3L)) %>%
-#   mutate(unique_ID = str_sub(REC, start = 3L),
-#          DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
-#          DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
-#          DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
-#          STATE = str_pad(STATE, 2, pad = "0"),
-#          REGOFF = str_pad(REGOFF, 2, pad = "0"))
-#
-# df82 <- df82[-c(1:2000), ]
-# df89 <- df89[-c(1:2000), ]
-# sum(df82$unique_ID %in% df89$unique_ID)
-
-
-# nA <- nrow(df82)
-# nB <- nrow(df89)
-# Ztrue <- sapply(df89$unique_ID, function(x){
-#   match <- which(df82$unique_ID == x)
-#   if (length(match) == 0){
-#     match <- nA + 1
-#   }
-#   match
-# }) %>%
-#   unname()
-
-#sum(Ztrue < nA +1)
-
-evals <- list.files("eval", full.names = T)
-eval_df <- lapply(evals, readRDS) %>%
-  do.call(rbind, .)
-
-eval_df
-
-eval_df %>%
-  group_by(method) %>%
-  summarize(total_declared = sum(total_declared),
-            true_declared = sum(true_declared),
-            false_declared = sum(false_declared)) %>%
-  mutate(recall = true_declared / true_matches,
-         precision = true_declared / total_declared,
[…]

### 2023-04-03 cluster_jobs/comp_hash_nltcs_dep.R

@@ -1,0 +1,55 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+library(tidyverse)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+HashRecords <- function(df1, df2){
+  cd <- CompareNLTCS(df1, df2)
+  patterns <- GetUniquePatterns2(cd, R = 10)
+  patterns
+}
+
+nltcs <- read.csv("data/proc_nltcs.csv")
+df82 <- nltcs %>%
+  filter(FILE == 82) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0")) %>%
+  unite(dob, 2:4, sep = "") %>%
+  unite(location, c(STATE, REGOFF), sep = "")
+
+
+df89 <- nltcs %>%
+  filter(FILE == 89) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0")) %>%
+  unite(dob, 2:4, sep = "") %>%
+  unite(location, c(STATE, REGOFF), sep = "")
+
+
+chunks <- 30
+chunk_size <- ceiling(dim(df89)[1]/chunks)
+chunk_id <- rep(1:chunks, each = chunk_size)[1:dim(df89)[1]]
+
+#job_marker <- rep(1:88, each =200)
+#job_marker <- job_marker[1:nrow(df89)]
+chunk_marker <- chunk_id == taskID
[…]

### 2023-04-03 cluster_jobs/run_fabl_dep.R

@@ -1,0 +1,53 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+library(tidyverse)
+
+setwd("cluster_jobs")
+#nltcs <- read.csv("data/proc_nltcs.csv")
+df82 <- nltcs %>%
+  filter(FILE == 82) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+df89 <- nltcs %>%
+  filter(FILE == 89) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+nA <- nrow(df82)
+nB <- nrow(df89)
+
+Ztrue <- sapply(df89$unique_ID, function(x){
+  match <- which(df82$unique_ID == x)
+  if (length(match) == 0){
+    match <- nA + 1
+  }
+  match
+}) %>%
+  unname()
+
+comparisons <- readRDS("full_hash/dependent/full_hash")
+chain <- BKSimple_hash2_big(comparisons)
+Zchain <- chain$Z
+Zhat <- LinkRecordsBK(Zchain, nA, 1, 1, 2, Inf)
+eval <- GetEvaluations(Zhat[[1]], Ztrue, nA)
+Zhat_resolved <- ResolveConflicts(Zhat)
+eval_resolved <- GetEvaluations(Zhat_resolved, Ztrue, nA)
+results <- rbind(eval, eval_resolved) %>%
+  as.data.frame()
[…]

### 2023-04-03 cluster_jobs/run_fabl_ind.R

@@ -1,0 +1,52 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+library(tidyverse)
+
+nltcs <- read.csv("data/proc_nltcs.csv")
+df82 <- nltcs %>%
+  filter(FILE == 82) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+df89 <- nltcs %>%
+  filter(FILE == 89) %>%
+  select(-FILE, -SEQ) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+  mutate(unique_ID = str_sub(REC, start = 3L),
+         DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+         DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+         DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+         STATE = str_pad(STATE, 2, pad = "0"),
+         REGOFF = str_pad(REGOFF, 2, pad = "0"))
+
+nA <- nrow(df82)
+nB <- nrow(df89)
+
+Ztrue <- sapply(df89$unique_ID, function(x){
+  match <- which(df82$unique_ID == x)
+  if (length(match) == 0){
+    match <- nA + 1
+  }
+  match
+}) %>%
+  unname()
+
+comparisons <- readRDS("full_hash/independent/full_hash")
+chain <- BKSimple_hash2_big(comparisons)
+Zchain <- chain$Z
+Zhat <- LinkRecordsBK(Zchain, nA, 1, 1, 2, Inf)
+eval <- GetEvaluations(Zhat[[1]], Ztrue, nA)
+Zhat_resolved <- ResolveConflicts(Zhat)
+eval_resolved <- GetEvaluations(Zhat_resolved, Ztrue, nA)
+results <- rbind(eval, eval_resolved) %>%
+  as.data.frame()
+row.names(results) <- NULL
[…]

### 2023-04-03 cluster_jobs/train.R

@@ -40,3 +40,4 @@ cd <- compare.linkage(train82[, -c(7, 8)], train89[, -c(7, 8)],
 
 model <- trainSupv(cd, method = paste(method))
 saveRDS(model, file = paste0("trained_models/", method))
+

### 2023-04-03 ncvr_raleigh/combine_hash.R

@@ -1,0 +1,37 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+
+dfA <- readRDS("data/raleigh_A")
+dfB <- readRDS("data/raleigh_B")
+nA <- nrow(dfA)
+nB <- nrow(dfB)
+hash_paths <- list.files("hash/", full.names = T)
+patterns_hashed <- lapply(hash_paths, readRDS)
+
+counts_big <- Reduce(`+`, map(patterns_hashed, ~.x[[2]]))
+
+counts_per_rec_big <- patterns_hashed %>%
+  map(`[[`, 3) %>%
+  flatten()
+
+hash_list_big <- patterns_hashed %>%
+  map(`[[`, 4) %>%
+  flatten()
+
+cd <- list(indicators = NULL,
+                    n1 = nA,
+                    n2 = nB,
+                    nDisagLevs =rep(2, 6),
+                    Ztrue = NULL)
+
+patterns <- list(patterns_hashed[[1]][[1]],
+                 counts_big,
+                 counts_per_rec_big,
+                 hash_list_big)
+
+comparisons <- list(cd, patterns)
+saveRDS(comparisons, file = paste0("full_hash/", "full_hash"))
+

### 2023-04-03 ncvr_raleigh/combine_hash.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+#SBATCH --job-name=combine_hash
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --time=24:00:00
+#SBATCH --mem=128G
+
+#
+# modules
+#
+module load R
+Rscript %x.R

### 2023-04-03 ncvr_raleigh/combine_results_supervised.R

@@ -1,0 +1,72 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+
+dfA <- readRDS("data/raleigh_A")
+dfB <- readRDS("data/raleigh_B")
+
+testA <- dfA[-c(1:1500), ]
+testB <- dfB[-c(1:1500), ]
+nA <- nrow(testA)
+nB <- nrow(testB)
+
+Ztrue <- MakeZtrue(testA$voter_id, testB$voter_id,
+                   nA, nB)
+
+true_matches <- sum(Ztrue < nA + 1)
+# df82 <- nltcs %>%
+#   filter(FILE == 82) %>%
+#   select(-FILE, -SEQ) %>%
+#   mutate(unique_ID = str_sub(REC, start = 3L),
+#          DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+#          DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+#          DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+#          STATE = str_pad(STATE, 2, pad = "0"),
+#          REGOFF = str_pad(REGOFF, 2, pad = "0"))
+#
+#
+# df89 <- nltcs %>%
+#   filter(FILE == 89) %>%
+#   select(-FILE, -SEQ) %>%
+#   mutate(unique_ID = str_sub(REC, start = 3L)) %>%
+#   mutate(unique_ID = str_sub(REC, start = 3L),
+#          DOB_DAY = str_pad(DOB_DAY, 2, pad = "0"),
+#          DOB_MONTH = str_pad(DOB_MONTH, 2, pad = "0"),
+#          DOB_YEAR = str_pad(DOB_YEAR, 2, pad = "0"),
+#          STATE = str_pad(STATE, 2, pad = "0"),
+#          REGOFF = str_pad(REGOFF, 2, pad = "0"))
+#
+# df82 <- df82[-c(1:2000), ]
+# df89 <- df89[-c(1:2000), ]
+# sum(df82$unique_ID %in% df89$unique_ID)
+
+
+# nA <- nrow(df82)
+# nB <- nrow(df89)
+# Ztrue <- sapply(df89$unique_ID, function(x){
+#   match <- which(df82$unique_ID == x)
+#   if (length(match) == 0){
+#     match <- nA + 1
+#   }
+#   match
+# }) %>%
+#   unname()
+
[…]

### 2023-04-03 ncvr_raleigh/comp_data.R

@@ -1,0 +1,27 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+
+#setwd("ncvr_raleigh")
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+train_set <- 1500
+dfA <- readRDS("data/raleigh_A")
+dfB <- readRDS("data/raleigh_B")
+nA <- nrow(dfA)
+nB <- nrow(dfB)
+
+testA <- dfA[-c(1:train_set), ]
+testB <- dfB[-c(1:train_set), ]
+
+chunks <- 50
+chunk_size <- ceiling(dim(testB)[1]/chunks)
+chunk_id <- rep(1:chunks, each = chunk_size)[1:dim(testB)[1]]
+
+chunk_marker <- chunk_id == taskID
+chunkB <- testB[chunk_marker, ]
+linkage_vars <- c(4, 5, 6, 7, 9, 13)
+
+cd <- compare.linkage(testA[, linkage_vars], chunkB[, linkage_vars],
+                      identity1 = testA$voter_id, identity2 = chunkB$voter_id)
+saveRDS(cd, file = paste0("comp_vectors/", "cd_", taskID))
+

### 2023-04-03 ncvr_raleigh/comp_data.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --array=1-50
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --job-name=comp_data
+#SBATCH --time=24:00:00
+#SBATCH --mem=128GB
+
+#
+# modules
+#
+module load R
+Rscript comp_data.R

### 2023-04-03 ncvr_raleigh/comp_hash.R

@@ -1,0 +1,28 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+#setwd("ncvr_raleigh")
+dfA <- readRDS("data/raleigh_A")
+dfB <- readRDS("data/raleigh_B")
+nA <- nrow(dfA)
+nB <- nrow(dfB)
+
+chunks <- 50
+chunk_size <- ceiling(dim(dfB)[1]/chunks)
+chunk_id <- rep(1:chunks, each = chunk_size)[1:dim(dfB)[1]]
+
+chunk_marker <- chunk_id == taskID
+chunkB <- dfB[chunk_marker, ]
+
+linkage_vars <- c(4, 5, 6, 7, 9, 13)
+cd <- CompareRecords_fabl(dfA[, linkage_vars], chunkB[, linkage_vars],
+                          as.numeric(seq_along(linkage_vars)),
+                          types = rep("bi", length(linkage_vars)))
+cd[[1]] <- apply(cd[[1]], 2, as.numeric)
+hash <- GetUniquePatterns2(cd, R = 10)
+saveRDS(hash, file = paste0("hash/", "hash_",
+                            str_pad(taskID, 2, pad = "0")))

### 2023-04-03 ncvr_raleigh/comp_hash.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --array=1-50
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --job-name=comp_hash
+#SBATCH --time=24:00:00
+#SBATCH --mem=128GB
+
+#
+# modules
+#
+module load R
+Rscript comp_hash.R

### 2023-04-03 ncvr_raleigh/make_data.R

@@ -1,0 +1,33 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+
+ncvr <- read.csv("data/ncvr.csv")
+# ncvr_clean <- ncvr %>%
+#   apply(., 2, function(x){
+#     x[x == ""] <- NA
+#     x
+#   }) %>%
+#   as.data.frame()
+
+
+ncvr$full_phone_num[ncvr$full_phone_num == ""] <- NA
+ncvr$middle_name[ncvr$middle_name == ""] <- NA
+
+ncvr_clean <- ncvr
+saveRDS(ncvr_clean, "data/ncvr_clean")
+
+dfA <- ncvr_clean %>%
+  filter(file_id == "a") %>%
+  filter(city == "raleigh")
+
+dfB <- ncvr_clean %>%
+  filter(file_id == "b") %>%
+  filter(city == "raleigh")
+
+dfA <- dfA[!duplicated(dfA$voter_id), ]
+dfB <- dfB[!duplicated(dfB$voter_id), ]
+
+saveRDS(dfA, "data/raleigh_A")
+saveRDS(dfB, "data/raleigh_B")
+

### 2023-04-03 ncvr_raleigh/test.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --array=1-50
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --job-name=test
+#SBATCH --time=24:00:00
+#SBATCH --mem=128G
+
+#
+# modules
+#
+module load R
+Rscript test.R

### 2023-04-03 ncvr_raleigh/train.R

@@ -1,0 +1,23 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+train_set <- 1500
+methods <- c("svm", "rpart")
+method <- methods[taskID]
+dfA <- readRDS("data/raleigh_A")
+dfB <- readRDS("data/raleigh_B")
+# nA <- nrow(dfA)
+# nB <- nrow(dfB)
+
+trainA <- dfA[1:train_set, ]
+trainB <- dfB[1:train_set, ]
+
+linkage_vars <- c(4, 5, 6, 7, 9, 13)
+cd <- compare.linkage(trainA[, linkage_vars], trainB[, linkage_vars],
+                      identity1 = trainA$voter_id, identity2 = trainB$voter_id)
+
+model <- trainSupv(cd, method = paste(method))
+saveRDS(model, file = paste0("trained_models/", method))
+

### 2023-04-03 ncvr_raleigh/train.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+#SBATCH --array=1-2
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --job-name=training
+#SBATCH --time=24:00:00
+#SBATCH --mem-per-cpu=128G
+#
+# modules
+#
+module load R
+Rscript train.R

### 2023-05-02 ba_resubmit/response_to_reviewers2.tex
response to reviewers

\documentclass[letterpaper, parskip]{scrartcl}

\usepackage{amsmath,amssymb,bm}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=1.4in]{geometry}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{tcolorbox}            % provides shaded box for reviewer comments
\usepackage{enumitem}             % use more compact lists
%\usepackage[parfill]{parskip}     % use skips between paragraphs and no indents
\usepackage{bibentry}             % provides \nobibliography command
\usepackage{xspace}

% Make itemize and enumerate environments more compact
\setlist[itemize]{noitemsep, topsep=0pt, parsep=0pt}
\setlist[enumerate]{noitemsep, topsep=0pt, parsep=0pt}

% -------- Define point raised command --------- %

\definecolor{darkpurple}{HTML}{500050}
\tcbuselibrary{breakable}       % allow tcolorbox to break across pages
\tcbuselibrary{skins}
\tcbset{%
	enhanced,%
	frame hidden,%
	sharp corners,%
	breakable,%
	parbox=false,%
	colback=lightgray!30,%
	enlarge left by=6pt,%
	width=\linewidth-6pt,%
	colback=gray!0,
	left=4pt,%
	right=0pt,%
	top=0pt,%
	bottom=0pt,%
	borderline west={2pt}{0pt}{lightgray},%
	%coltext=darkpurple,%
	before skip=14pt plus 2pt,%
	after skip=14pt plus 2pt}

%\newcommand{\pointRaised}[1]{%
%	\begin{tcolorbox}
%		\itshape #1
%	\end{tcolorbox}
%}
%
%\newcounter{responsectr}[section]     % counter for response resets at each section
%\newcommand{\reply}[2]{%
%	\refstepcounter{responsectr}%
%	\textbf{#1.\theresponsectr:} #2
%}

\newcommand{\pointRaised}[2]{%
	\textbf{#1.\theresponsectr:} #2
}

\newcounter{responsectr}[section]     % counter for response resets at each section
\newcommand{\reply}[1]{%
	\refstepcounter{responsectr}%
		\begin{tcolorbox}
			\itshape #1
		\end{tcolorbox}
}

\newcommand{\todo}{\textcolor{red}{[TODO]}\xspace}

\newcommand{\ncomment}[1]{{\color{teal} NM: #1}}

\begin{document}

	% ****************** TITLE ****************************************

	\title{Response to Reviewer Comments}
	\maketitle

Thank you for the thoughtful feedback. We agree that much of the notation iscumbersome, and we in particular thank the reviewers for seeing the merit in our methods despite those difficulties.  We have made several broad changes to address the Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{AE Comments}

\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}

\reply{We agree that our notation is at times cumbersome. We have made several efforts to simplify notation in this revision.

\begin{itemize}
	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. We find this generally more readable, and particularly simplifies the notation in Section 4.2. 
	\item At the beginning of Section 4.3, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics discussed in Section 4.1. This is the text added:

''To ground the discussion of posterior inference, we first re-express the likelihood function from (6a) in terms of the agreement patterns and summary statistics in Table 2. Let $m(p) =  p(\gamma_{ij} = h_p|Z_j = i)$ be probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match. Similarly define $u(p)$. These probabilities are simply products of relevant $\bm{m}$ and $\bm{u}$ parameters. Viewed through the perspective of patterns, the likelihood is
\begin{align}
	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\bm{\Gamma}}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}. \label{likelihood_efficient}"
\end{align}

	This likelihood is then used throughout the discussion of efficient posterior inference. We hope this expression clarifies the meaning of "preservation of weight (Comment 5), and strengthens the argument for $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ being sufficient statistics for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ (Comment 4).

	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the clunky $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$, which we feel is more direct and intuitive.

	\item Section 4.1 introduces the sufficient statistic $N_{j_p}$. After careful review, we realized that by changing the order of subscripts and instead using $N_{p_j}$, we can replace the sum of counts across all records (previously denoted $H_p$), with the more intuitive $N_p$. Thus, we have removed all of the ''$H$" quantities throughout Section 4. We believe this substantially improves readability.  
\end{itemize}
}

\pointRaised{AE}{Line 80 on p. 3 says: “For ease of readability, we follow the convention established by Sadinle (2017) and say “record $i \in X_1$” rather than the more compact $x_{1i}$.” Taking this at face value implies that, in any sentence, one could swap “record $i \in X_1$” for “$x_{1i}$,” which clearly is not true.  So, as it stands, the sentence does little to improve ease of readability.  In fact, if the point is that “$x_{1i}$” will never be used again, why is the notation introduced in the first place?}
\reply{We added this sentence in an attempt to address a previous reviewer comment pointing out a conflation between the record itself and the index directing towards that record. We did in fact intend for “record $i \in X1$” to be mean precisely $x_1i$. However, we see now that introducing $x_{1i}$ has introduced new confusion. With the simplified reference to the two datasets as $A$ and $B$, that material now reads:

	``Consider a data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$".
}

\setcounter{responsectr}{0}
\section{Reviewer Comments}

	\pointRaised{R}{%
	For the missing data treatment now included just before Section 2.1, is the assumption truly
	missing at random, or missing completely at random? Or, does this distinction not matter
	because of the independence assumed across elements of the comparison vector?}

	\reply{%
\textcolor{red}{I used MAR because that is what Sadinle wrote in his paper. Please review, and confirm whether its MCAR or not.}

The assumption is missing completely at random. You are correct that this is a result of the independence assumed across elements of the comparison vector. This has been clarified.
	}

	\pointRaised{R}{%
	Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and
	think it was a good choice to move the full derivation to an appendix. However, I do not follow
	the re-expression of the pmf for $\Gamma_{.j}$. First, the opening square brackets are still misplaced
	throughout. In the second line of this derivation, the authors divide the expression by a product
	of the element-wise conditional probabilities of a match in the comparison vector (u), raised by
	an indicator that the element of the comparison vector equals a particular value. Perhaps I
	misunderstand, but I believe this product is not constant in $\Gamma_{.j}$ or $u$, and so the total expression
	is not proportional to the line above, as interpreted as a pmf. I think that the end effect is that
	the last line of the newly expressed pmf is missing a factor of $\prod_i \prod_f \prod_l u_{fl}^{I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}$ regardless of the value of $z_j$.}

\reply{%
	The square brackets have been corrected, and we apologize for the oversight.

	We have provided an alternative derivation of the full conditional, separately handling the cases for $Z_j \leq n_A$ and for $Z_j = n_A + j$. We provide it here for your reference:

	We consider the likelihood in two cases. When $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
	\end{align}
	When $Z_j = i$ for some $i \leq n_A$, we have
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = i) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \times \prod_{i' != i}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{i'j}^f = l)I_{obs}(\gamma_{i'j}^f)}.
	\end{align}
	We multiply and divide by the $u$ parameters for the matching record pair to obtain
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = i) &=\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \times \prod_{i' = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{i'j}^f = l)I_{obs}(\gamma_{i'j}^f)} \\
		&= w_{ij} \times c_j.
	\end{align}
	Combining these results, we have 
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = z_j) = \begin{cases} 
		w_{z_j, j}  & z_j \leq n_A; \\
		1 &  z_j  = n_A + j, \\
	\end{cases}
	\end{align}
}

%	{\color{red}{I don't really understand his confusion. Why wouldn't the product of $u$ parameters be constant in $u$ and the data? Perhaps if I use $\mathcal{L}(Z_j| m, y, \Gamma_{.j})$, it might be clearer?}}}

%	In the second line of the derivation, we divide by the $u$ parameters for all of the records in $A$ (note the product outside of the fraction). This quantity does not depend on $Z_j$ and is constant in $u$ and $\Gamma_{.j}$.  This is easiest to see through the likelihood expressed at the level of the agreement patterns. We are also notating this likelihood as $\mathcal{L}(Z_j | m, u, \Gamma_{.j})$ to make it clear what is considered constant.
%	
%	\begin{align*}
%	 \mathcal{L}(Z_j | m, u, \Gamma_{.j})  &= \prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}\\
%	 &= \prod_{p=1}^P u(p) ^ {N_{p_j}} \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{- I(Z_j = i)}\\
%	 &= \prod_{p=1}^P u(p) ^ {N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}\\
%	\end{align*} 
%where $w_p = \frac{m(p)}{u(p)}$. The term outside of the inner product does not depend on $Z_j$. Thus we have
%\begin{align*}
%	\mathcal{L}(Z_j | m, u, \Gamma_{.j})   \propto
%	\begin{cases} 
%		w_p,  & z_j \leq n_1; \\
%		1, &  z_j  = n_1 + j; \\
%	\end{cases} \label{eqn:z}\\
%\end{align*}
%
%REVISIT

	\pointRaised{R}{%
Appendix 8.2: Thank you for including the details on integrating out pi from the full
conditionals. They surprised me. I had assumed that the authors had integrated out pi in the
prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite
matchings”. Does the alternate approach presented in this paper provide a different algorithm
than directly integrating the prior distribution? Is the presented alternative approach justified?}

\reply{%
 {\color{red}{I really don't know how to answer this. I don't know how I would ''directly integrate" $\pi$, and I don't know if what I do is ''justified." Its kind of an intellectual relic to how I was originally thinking about the problem, and it reveals a quantity similar to what Sadinle was working with.

 It is theoretically more straightforward to just not integrate out $\pi$, and use samples from $\pi$ in Gibbs sampler like I do with all other parameters. Would that be too large a change at this stage?

 Without integrating out $\pi$, the full conditional is given on page 26 of the Appendix.}}
}

\pointRaised{R}{%
	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are
	based on observed comparison vectors without consideration of Z (matches). In the
	second paragraph of Section 4.3, the H notation includes the matches (Z), in notation
	and definition that seems to conflict with the statement just below Equation (10). This
	also comes into play in Equation (16)}

	\reply{%
	You are correct that the the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$. This notation was created purely to be able to denote when a record was left unlinked during the Gibbs Sampler as in Equation 16 in the submitted draft.

	I can see why this is confusing. Therefore, I have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
	}

\pointRaised{R}{%
	Third paragraph of Section 4.1: The authors claim they are computing “sufficient
	statistics”. What exactly are these statistics sufficient for?
	}

	\reply{%

We show in the revised Section 4.3 that $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ is a set of statistics that are sufficient for $m$, $u$, and $Z$.

{\color{red}{I drafted a whole bunch of material to argue more rigoursly that $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ is  set of sufficient statistics. If this argument seems weak, or is too much new material, I can also just keep what I had, and use less rigorous term "summary statistics" instead.}} \todo

}

\pointRaised{R}{%
First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute
the appropriate weight to all records through the summary statistics…” What is meant
by the term “weight”? Which records are appropriately weighted – those in X2?}

\reply{%
	For each $j \in \{1, \ldots, n_B\}$, I use SEI to eliminate from memory the indices of  many records in $A$ that are unlikely matches for $B_j$ However, since I know the number of such records that exhibit each pattern $p$, I can still include their contribution to the likelihood through $N_{p_j}$.

	We have changed that sentence in the text to read ''All original record pairs still contribute to the likelihood through $\mathcal{N}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}\}$."

	This is clearer when looking at the conditional likelihoods in Equations 15 and 17 using $\tilde{\Gamma}$ in the revised Section 4.3. 
}

\pointRaised{R}{%
Second paragraph of Section 4.2: “and delete those comparison vectors”. Which are
“those” vectors?}

\reply{%
	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
}

\pointRaised{R}{%
Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations
(13) and (14)? I recommend that the authors either refrain from suppressing the SEI
notation or further explain how the SEI algorithm has changed the quantities in these
equations.}

\reply{%
	The written expression in (13a) is correct. We then compress $r_{p_j}^{cd}$, and store in memory $r_{p_j}^{SEI, cd}$ Equation (14a) is conducted through $r_{p_j}^{SEI, cd}$. Equations (13b-c) and (14b-c) are exact, unchanged by SEI. We have clarified this.
}

\pointRaised{R}{%
Section 4.2: I appreciate the practical advice about choosing S for the SEI method. However,
this choice seems arbitrary in the absence of further discussion/evidence. Given that the
primary novelty of the manuscript is in methods to speed and otherwise improve computation, I
am surprised that this aspect of computational innovation is presented with virtually no
theoretical or empirical exploration. Presumably the SEI method has some sort of accuracy
trade-off, as the authors warn that linkage results may be “distorted” if S is low. However, this
trade-off is not quantified or even discussed in practical terms beyond the terse
recommendation to choose S=10.
}

\reply{%

{\color{red}{I'm not sure the best way to address this. Here is the where the situation arises:

Imagine file $B$ has one record $B_1$ for Brian Kundinger, and file $A$ has 20 nearly matching records for Brian Kundinger. Perhaps Brian A Kundinger, Brian B Kundinger, Brian C Kundinger, $\ldots$, ect. All of these records share the same agreement pattern (say $p = 2$) with $B_1$.

Without SEI, posterior probability would be divided evenly among all 20 near matches, such that the posterior probability would be at most $\frac{1}{20}$. If I chose $S=2$ however and I only retained 2 representatives in $r_{2_1}$, I could likely have one of those records having posterior probability above .5, and I would erroneously declare it a match.

So issues only arise in very degenerate situations. How do you think I should address this? Should I do a simulation showing degeneracy when S=1 or 2?
}
}
}
\pointRaised{R}{%
Page 5, 2 sentences before equation (4): I believe the sum should be of $I(Z_j \leq n_1)$, not $I(Z_j \leq n_1 +1)$.}

\reply{%
	That is correct. This has been corrected, and we apologize for the oversight. 
}

\pointRaised{R}{%
Equation (6a): The indices do not match the subscripts in the indicator function in each
summand, or their standard meaning in table 1.}

\reply{%
	{\color{red}{I do not understand this objection.}}
}

\pointRaised{R}{%
Generally, the authors seem to arbitrarily use upper and lower case z interchangeably in
function definitions.}

\reply{%
	{\color{red}{When writing, I attempted to use $Z$ when I was discussing a random quantity, and $z$ to reference a realized value. I haven't found anything that I see wrong. I think I'm going to need help with this.}}
}

\pointRaised{R}{%
Equations (8) and (9): Should the weights have superscript (s) (as the Zs do)?}

\reply{%
	When writing, I believed it was sufficient to include the superscripts on the $m$ and $u$ in the definition of $w$. However, you are correct, it is more clear to explicitly include it on the $w$. This has been updated. 
}

\pointRaised{R}{%
Gamma is in some places described as a set and in others as a matrix (particularly in
sections 4.2 and 5.1).
}

\reply{%
	$\Gamma$ is a matrix comprised by comparison vectors. It is certainly not a set in the mathematical sense. All references to $\Gamma$ as a set have been removed.

	{\color{red}{I found one example of me saying that $\Gamma$ is ''comprised of the set of comparison vectors.' I'll triple check before resubmission.}}
}

\pointRaised{R}{%
Section 4.3, second paragraph, the definitions of the concatenated vectors $\alpha_0$ and $\beta_0$
should have final elements subscripted by $L_f$, where the sub-subscript is capitalized.}

\reply{%
	You are correct. This has been revised. 
}

\end{document}

### 2023-05-02 ba_resubmit/kundinger_resubmit2.tex
response to reviewers

--- 2022-11-23 ba_resubmit/kundinger_resubmit.tex
@@
 % Possible citation for bipartite matching. 
-%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
+%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_A995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
 
@@
 
-Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
-Without loss of generality, denote files such that $n_1 \geq n_2$.
-For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$.
+%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.
+
+Consider two data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$.
+Without loss of generality, denote files such that $n_A \geq n_B$.
 Under bipartite matching, the set of matches across data files can be represented in two equivalent ways.
-First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
+First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
 \begin{align}
@@
 	\begin{cases}
-		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
+		1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
 		0, & \text{otherwise}.\\
@@
 This sparse matrix representation can become cumbersome for large linkage tasks.
-More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that
+More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
 \begin{align}
@@
 	\begin{cases}
-		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
-		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
+		i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
+		n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
 	\end{cases}
@@
 
-Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
-$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
+Denote the set of matches by $\bm{M} = \{(i,j): \Delta_{ij} = 1\}$, and the set of non-matches by 
+$\bm{U} =  \{(i,j): \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
 
@@
 Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ We collect the vectors $\bm{\gamma}_{ij}$ as the comparison matrix $\Gamma \in \mathbb{R}^{n_A n_B \times F}$.
 
@@
 The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features.
-For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for record $i$ and $j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
+For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
 For numerical features, we can use absolute difference between the two feature values.
-For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for record $i$ and $j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for record $i$ and $j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
+For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
 For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
@@
 
-More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
-Here, $I_{f1}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
+More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
+Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
 Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
 
-In the construction of comparison vectors, it is common to encounter missing information in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$.
+{\color{red}{
+In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
 As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We assume that this missingness occurs at random (MAR, per \cite{LittleRubin2002}).
-With the MAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
-To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.}
+}
 
@@
 	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
-	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
+	&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
 \end{align}
@@
 
-%\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
-%\begin{align}
-%	\label{eqn:wts}
-%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
-%\end{align}
-%This ratio is large when there is strong evidence of the pair being a match, and small otherwise. All record pairs with weights above a specified threshold are considered a match.
-
 Within this framework, record pairs are independently classified as matches and non-matches based on the estimated parameters.
@@
 %	\label{eqn:jaro}
-%	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
+%	&\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij} 
 %	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
-%	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
-%	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
+%	& \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
+%	& \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
 %\end{align}
@@
 In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
-He assigns a prior distribution for the probability of the indicator that a record in $\bm{X}_2$ has a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
-It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$.
-Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 !
-/ (n_1 - n_{12}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
+He assigns a prior distribution for the probability of the indicator that a record in $B$ has a match in $A$, so that $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
+It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
+Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A !
+/ (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
 Thus, the beta distribution for bipartite matching, is given by
@@
 \label{eqn:sadinle_prior}
-P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
 \end{align}
 where $\text{B}(\cdot, \cdot)$ represents the Beta function.
-This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each \(Z_j\).
+This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
 This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
 
-%hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X}_2$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
+%hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $B$ that have matches in $\bm{X_1}$, and $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A + j)$ is the number of records in $B$ that have a match in $A$.
 %One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
@@
 For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
-%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
 %$$
-%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
-This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
+%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A + j)$ is the records in $B$ that have a match in $A$.
+This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling $Z_j$.
 This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.
@@
 	p(Z_j = z_j| \pi) &= \begin{cases} 
-	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
-	1-\pi, &  z_j  = n_1 + j; \\
+	\frac{1}{n_A}\pi,  & z_j \leq n_A; \\
+	1-\pi, &  z_j  = n_A + j; \\
 \end{cases} \\
@@
 \end{align}
-We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match.
-The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$
-
-In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_1, n_1 +j\}$ is a priori equally likely for $Z_j$.
-This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$.
+We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match.
+The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$
+
+In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
+This however amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
 In our preliminary studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}).
-We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.
-
-%Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
+We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
+
+%Intuitively, this set of priors says that record $B_j$ has some match in $A$ with probability $\pi$, and that each record $A_i$ is equally likely to be that match.
 $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
+One non-informative choice might be $\pi \sim \text{Beta}(1, 1)$, which corresponds to a prior belief that non-matches and matches are equally likely, and another might be $\pi \sim \text{Beta}\left(1, \frac{1}{n_A}\right)$, which corresponds to a uniform prior on the labeling of $\bm{Z}$.
+We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_A}\right)$ random variable.
 
 Note that linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
 We find $\pi$ to be more a interpretable parameter than $\lambda$ in the bipartite case.
-In this setting, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions.
+In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows.
+Additionally, while the Fellegi-Sunter model makes $n_A \times n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
 As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
 
-%To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}, reproduced in Appendix \ref{bayes-estimate}.
-Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
-
 To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
 To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
 Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
-
-%\hypertarget{fabl-model}{%
-	%	\subsection{Model Specification}
-	%	\label{model-specification}} 
-%
-%As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent.
-Additionally, we assume that any missing data is missing at random.
-Let  $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denote the probability of that matching records have level $\ell$ of disagreement in feature $f$, and $u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represent the same probability for non-matches.
-In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, $\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
-
-%The model can then be specified as follows:
 
@@
 		\hline
-		$\bm{X}_1, \bm{X}_2$ & data files\\
-		$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
-		$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
+		$A, B$ & data files\\
+		$i \in 1, \ldots, n_A $ & index over records in $A$\\
+		$j \in 1, \ldots, n_B $ & index over records in $B$\\
 		$f \in 1, \ldots F$ & index over fields used for comparisons \\
 		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
-		$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
-		$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
-		$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
-		$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
+		$n_{AB}$ & number of entities in common between $A$ and $B$\\
+		$\bm{\gamma}_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+		$Z_j = i$ & records $A_i$ and $B_j$ match \\
+		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
 		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
 		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-		$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
-		\hline
-	\end{tabular}\caption{Summary of model notation.}\label{table_notation_1}
+		$\pi$ & probability that a record in $B$ has a match in $A$ \\
+		\hline
+	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
 \end{table}
 
-For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_1}.
+For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_A}.
 \begin{subequations}
 \begin{align}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
 	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \label{eqn:m} \\
@@
 	\begin{cases} 
-		\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
-		1-\pi, &  z_j  = n_1 + j; \\
+		\frac{1}{n_A}\pi,  & z_j \leq n_A; \\
+		1-\pi, &  z_j  = n_A + j; \\
 	\end{cases} \label{eqn:z}\\
@@
 \end{subequations}
-
-%where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.
 
@@
 
-We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$.
+We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
 To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^s$, we use the full conditionals,
@@
 \begin{cases} 
-	w_{z_j, j},  & z_j \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{12}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
+	w_{z_j, j}^{(s)},  & z_j \leq n_A; \\
+	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
 \end{cases}
 \end{align}
-where, for all $i \in \bm{X}_1$, 
+where, for all $i \in \{1, \ldots, n_A\}$, 
 \begin{align}
 	\label{eqn:fs_weight}
-	w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+	w_{ij}^{(s)} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
 \end{align}
 Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.
-
-%We work with the following factorization of the joint posterior distribution:
-%
-%\begin{align*}
-%	p(\bm{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\bm{Z} | \pi) p(\Phi) p(\pi) \\
-%	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-%	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-%	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
-%	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}.
-%\end{align*}
-
-%Since the Gibbs sampler does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the highest posterior probability, and declare the other to have no match.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
 
@@
 
-The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_1 \times n_2$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors.
+The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A \times n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors.
 One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}.
@@
 However, the impact of indexing on model parameters is not well understood; \citep{murray2016probabilistic} reviewed this issue in the context of the frequentist Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
-%Another technique is indexing, in which the modeller decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
 
@@
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
-For convenience, Table \ref{table_notation_2} summarizes the notation introduced throughout this section.
+For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.
 
@@
 		\hline
-		$h_p$ & one hot encoding of agreement pattern $p$ \\
-		$(i,j) \in h_p$ & comparison vector between records $i \in \bm{X}_1$ and $i \in \bm{X}_2$ exhibits pattern $p$ \\
-		$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-		$N_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-		$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
-		$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
-		$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
-		\hline
-	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_2}
+		$P$ & number of agreement patterns exhibited in $\Gamma$ \\
+		$h_p$ & agreement pattern $p$ \\
+		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $i \in B$ exhibits pattern $p$ \\
+		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+		%$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+		%$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+		%$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
+		\hline
+	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_B}
 \end{table}
-
-%To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi-Sunter method.
-The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns.
-\(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
-In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the orignial vectors themselves.
-Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi-Sunter comparison vectors, to aid in vectorized computations.
-See Appendix \ref{app:ohe} for details.
 
@@
 
-Following an insight from \cite{enamorado2019using}, we recognize that record pairs contribute to posterior calculations only through their agreement pattern as represented by $\gamma_{ij}$.
-To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns observed in $\bm{X}_1 \times \bm{X}_2$, and let $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns.
-\(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
-In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the original vectors themselves.
-Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original $\gamma_{ij}$ to aid in vectorized computations.
-See Appendix \ref{app:ohe} for details.
-
-First, we hash record pairs of the same agreement pattern to unique integer values. \cite{enamorado2019using} accomplished this efficiently through the hashing function
+Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
+Let $P$ be the number of patterns realized in $\Gamma$ and note that it is bounded above by $P^{*} = F + \prod_{f=1}^F L_f$, where we have noted the possibility of missing values.
+Note that this quantity is controlled by the modeller in their choices for $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
+To obtain a memory efficient representation of the data, we map the agreement pattern of each record pair to a unique integer.
+\cite{enamorado2019using} accomplished this through the hashing function
 \begin{align}
 	\label{eqn:hashing}
-h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}.
-\end{align}
-This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
-For computational ease, we map these integers to sequential integers from $\{1, \ldots, P\}$ corresponding to the enumerated patterns in $\mathcal{P}$.
-When the $(i, j)$ pair exhibits pattern $p$, we say $(i, j) \in h_p$.
-When $j$ has no match in $\bm{X}_1$, we write $(n_1 + j, j) \in h_{P+1}$.
-
-With all record pairs converted to hashed values, we identify the records in $\bm{X}_1$ with comparison vectors corresponding to each pattern $p$ for each record $j \in \bm{X}_2$.
-We denote this set $r_{j_p} = \{i \in \bm{X}_1 | (i, j) \in h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{j_p} | j \in \bm{X}_2, p \in 1, \ldots, P\}$.
-This representation is useful because it allows us to easily compute two sets of sufficient statistics.
-First, we compute the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, given by
+	h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(f>1)\sum_{d=1}^{f-1}(L_d -1)}.
+\end{align}
+We then map these integers to sequential integers $\{1, \ldots, P\}$.
+We refer to each unique agreement pattern as $h_p$, and to the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
+
+With all record pairs converted to hashed values, we identify the records in $A$ with comparison vectors corresponding to each pattern $p$ for each record $B_j$.
+We denote this set $r_{p_j} = \{i | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
+We then compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$, given by
 \begin{align}
 \label{eqn:N}
-N_{j_p} = |r_{j_p}| = \sum_{i=1}^{n_1} I((i, j) \in h_p).
-\end{align}
-Second, we compute the number of comparison vectors in $\Gamma$ that exhibit pattern $p$, given by
-\begin{align}
-	\label{eqn:H}
-	H_p = \sum_{j = 1}^{n_2} N_{j_p}.
-\end{align}
-We collect these sufficient statistics in the sets $\mathcal{N} = \{N_{j_p} |j \in \bm{X}_2, p \in 1, \ldots, P \}$ and $\mathcal{H} = \{H_{p} | p \in 1, \ldots, P \}$.
-As we show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
-Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we can use $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}, \mathcal{H}\}$ for posterior inference.
+N_{p_j} = |r_{p_j}| = \sum_{i=1}^{n_A} I(\gamma_{ij} = h_p).
+\end{align}
+%Second, we compute the number of comparison vectors in $\Gamma$ that exhibit pattern $p$, given by
+%\begin{align}
+%	\label{eqn:H}
+%	N_p = \sum_{j = 1}^{n_B} N_{p_j}.
+%\end{align}
+We collect these counts in $\mathcal{N} = \{N_{p_j} |p \in 1, \ldots, P, j \in 1, \ldots, n_B \}$.
+As we show in Section \ref{efficient-posterior}, $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$  serves as a set of sufficient statistics and fully characterizes the comparison matrix $\Gamma$ with no loss of information.
 
 %With all record pairs converted to hashed values, we can create a more compact nested list 
-%$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
-%where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$.
-
-%We collect these sufficient statistics in the sets $\mathcal{N} = \left\{\{N_{j_p}\}_{j=1}^{n_2}\right\}_{p=1}^P$ and $\mathcal{H} = \{H_{p}\}_{p=1}^P.$
+%$$\mathcal{R} = \left\{\{r_{p_j} \}_{p=1}^{P} \right\}_{j = 1}^{n_B},$$
+%where $r_{p_j} = \{A_i | \gamma_{ij} = h_p\}$ is a list of records in $A$ that share agreement pattern $p$ with record $B_j$.
+
+%We collect these sufficient statistics in the sets $\mathcal{N} = \left\{\{N_{p_j}\}_{j=1}^{n_B}\right\}_{p=1}^P$ and $\mathcal{H} = \{H_{p}\}_{p=1}^P.$
 
@@
 
-%In practice, it is not feasible to make all-to-all record comparisons because creating and storing $n_1 \times n_2$ comparison vectors is computationally infeasible for large data files.
-The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
-In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
-The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
-%
-%The hashing procedure described above considerably reduces the memory needed to store the comparison information.
-That is, instead of storing \(n_1 \times n_2\) comparison vectors, we only store the \(P\) unique vectors, and \(n_1 \times n_2\) scalar quantities relating record pairs to those vectors.
-However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large data files.
-Worse, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
-%
-%To address this issue, we propose storage efficient indexing (SEI).
-In standard indexing, one decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
-For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like first name, or only pairs that exactly match on a specified number of fields.
-this kind of indexing requires knowledge of the application of the application and invites room for human error \cite{murray2016probabilistic}.
-
-We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs but greatly reduce the memory costs associated with unlikely matches.
+We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ for all $n_A \times n_B$ record pairs but greatly reduce the memory costs associated with unlikely matches.
 This allows all-to-all comparisons for substantially larger linkage tasks.
-All records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same $w_{ij}$.
-Therefore, these records have the same probability to be identified as the link for record $j$.
-Thus, we know that records $i \in r_{j_p}$ such that $N_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
+All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{ij}$.
+Therefore, these records have the same probability to be identified as the link for record $B_j$.
+Thus, we know that records $i \in r_{p_j}$ such that $N_{p_j}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
 We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
-Therefore, rather than store all of these record labels, we store only a small number \(S\) of them in a new nested list $\mathcal{R}^{\text{SEI}}$.
-Instead of storing \(n_1 \times n_2\) record labels, with SEI we store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
+Therefore, rather than store all of these record labels, we store only a small number $S$ of them in a new nested list $\mathcal{R}^{\text{SEI}}$.
+Instead of storing $n_A \times n_B$ record labels, with SEI we store at most $n_B \times P \times S$ labels, regardless of how large $n_A$ might be.
+
 In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but is not likely to eliminate any of the indices for plausible matches.
 Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on few matches and distort linkage results.
-Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}, \mathcal{H}\}$.
-
-%For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations.
-For each chunked data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each chunked data file $\bm{X}_{2d}$ to construct the chunked comparison matrix $\Gamma^{cd}$.
-We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.
-These all-to-all comparisons can be computed serially or in parallel.
-Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
-Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger data files.
-
-For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller disjoint chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations.
-For each data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each $\bm{X}_{2d}$ to construct the comparison matrix $\Gamma^{cd}$.
-We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.
+All original record pairs still contribute to the likelihood through $\mathcal{N}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}\}$.
+
+For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $A^{*}$ and $B^{*}$ smaller disjoint chunks $\{A^a | a = 1, \ldots, A^{*} \}$ and $\{B^b | b = 1, \ldots, B^{*} \}$ for more manageable computations.
+For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\Gamma^{ab}$.
+We then conduct hashing, obtain the compressed $\tilde{\Gamma}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data.
 In detail, we calculate
@@
 	\begin{align}
-		r_{j_p}^{cd} &= \{i \in \bm{X}_{1c} | (i, j) \in h_p, j \in \bm{X}_{2d}\}, \label{r_{j_p}^{cd}} \\
-		N_{j_p}^{cd} &= |r_{j_p}^{cd}|, \label{N_{j_p}^{cd}} \\
-		H_p^{cd} &= \sum_{j \in \bm{X}_{2d}} N_{j_p}.\label{H_p^{cd}}
+		r_{p_j}^{ab} &= \{i \in  | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+		N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
 	\end{align}
@@
 These can computed serially or in parallel.
-We reduce memory costs through SEI on each $r_{j_p}$, while retaining counts of each agreement pattern through $N_{j_p}^{cd}$ and $H_p^{cd}$.
+We reduce memory costs through SEI on each $r_{p_j}$, while retaining counts of each agreement pattern through $N_{p_j}^{ab}$.
 Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
-Suppressing SEI notation, we combine information through
+Specifically, we combine information through
 \begin{subequations}
 	\begin{align}
-		r_{j_p} &= (r_{j_p}^{11}, \ldots, r_{j_p}^{cd}) \text{ for } c = 1, \ldots, C \text{ and } d = 1, \ldots, D,\label{r_{j_p}} \\
-		N_{j_p} &= \sum_{c = 1}^C \sum_{d = 1}^D N_{j_p}^{cd},\label{N_{j_p}} \\
-		H_p &= \sum_{c = 1}^C \sum_{d = 1}^D H_p^{cd}.\label{H_p}
-	\end{align}
+		r_{p_j}^{SEI} &= (r_{p_j}^{SEI, 11}, \ldots, r_{p_j}^{SEI, A^{*}B^{*}}) \text{ for } a = 1, \ldots, A^{*} \text{ and } b = 1, \ldots, B^{*},\label{r_{p_j}} \\
+		N_{p_j} &= \sum_{a = 1}^{A^{*}} \sum_{b = 1}^{B^{*}} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
 \end{subequations}
@@
 
-Updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum_f L_f\) many summations over \(n_1 \times n_2\) quantities.
+%prod_{f = 1}^F \prod_{l = 1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = 1)}
+
+To ground the discussion of posterior inference, we first re-express the likelihood function from (\ref{eqn:likelihood}) in terms of the agreement patterns and summary statistics in \label{table_notation_B}.
+Let $m(p) =  p(\gamma_{ij} = h_p|Z_j = i)$ be probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match.
+Similarly define $u(p)$.
+These probabilities are simply products of relevant $\bm{m}$ and $\bm{u}$ parameters.
+Viewed through the perspective of patterns, the likelihood is
+\begin{align}
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\bm{\Gamma}}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}. \label{likelihood_efficient}
+\end{align}
+{\color{red}{Should this first paragraph go at the end of Section 4.1? And should I be writing $r^{SEI}$ each time it comes up?}}
+
+In this form, we can see that the statistics collected in $\tilde{\Gamma}$ are sufficient for each of the model parameters in their respective conditional likelihoods and provide straightforward posterior updates.
+Conditioning on $\bm{Z}$, we can express the conditional likelihood for $\bm{m}$ and $\bm{u}$ as
+\begin{align}
+	\mathcal{L}(\bm{m}, \bm{u} \mid \bm{\Gamma}, \bm{Z})  &=  \prod_{p=1}^P m(p)^{N_p^m}u(p)^{N_p - N_p^m}  \label{likelihood_efficient_m_u}
+\end{align}
+where $N_p^m = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ is the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ is the total occurrence of pattern $p$ in the data across all record pairs.
+Note that the $\bm{m}$ and $\bm{u}$ parameters depend on the data only through the quantities calculated from $\mathcal{N}$ and $\mathcal{P}$.
+
+Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define $\bm{\beta_0}$ similarly for the $\bm{u}$ distributions.
+Let $e(h_p)$ be a one hot encoding of agreement pattern $p$, as described in Appendix \ref{app:ohe}.
+Then the terms needed for the posterior updates for level $l$ of field $f$ of the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
+\begin{subequations}
+	\begin{align}
+		\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P N_p^m \times e(h_p), \label{efficient_alpha} \\
+		\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P (N_p - N_p^m) \times e(h_p). \label{efficient_beta}
+	\end{align}
+\end{subequations}
+These quantities are identical those presented in (\ref{eqn:m_update}),and (\ref{eqn:u_update}), but are calculated through $P$ multiplications of scalar quantities over the vectors $h_p$, and then vectorized summations.
+This is markedly more efficient than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+Additionally, we can express the conditional likelihood for $Z_{j}$ as
+%\begin{align}
+%	\mathcal{L}(Z_j\mid  \bm{m}, \bm{u}, \tilde{\bm{\Gamma}}) = \prod_{p=1}^P N_{p_j} \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{- I(Z_j = i)} \label{likelihood_efficient_z}.
+%\end{align}
+\begin{align}
+	\mathcal{L}(Z_j\mid  \bm{m}, \bm{u}, \tilde{\bm{\Gamma}}) = \prod_{p=1}^P u(p)^{N_{p_j}} \prod_{i \in r_{p_j}} w(p)^{I(Z_j = i)} \propto \prod_{p=1}^P \prod_{i \in r_{p_j}} w(p)^{I(Z_j = i)} \label{likelihood_efficient_z}
+\end{align}
+{\color{red}{Should I include the proportionality step?}}
+where $w(p) = \frac{m(p)}{u(p)}$.
+Note that $Z_j$ interacts with the data only through $\mathcal{N}$ and $\mathcal{R}$.
+This likelihood also lends itself to markedly faster posterior computations.
+Although sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_A$ is large.
+This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
+To speed up computation, we break this sampling step into two.
+We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
+\begin{align}
+	\label{eqn:gibbs1}
+	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+	\begin{cases} 
+		 N_{p_j} \times w_{p},  & \gamma_{z_j, j} = h_p; \\
+		n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   \text{otherwise}. \\
+	\end{cases}
+\end{align}
+Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the $P$ agreement patterns.
+Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformally using
+\begin{align}
+	\label{eqn:gibbs2}
+	P\left(Z_j^{(s+1)} = z_j \mid \gamma_{z_j^{(s+1)}, j} = h_p \right) = \begin{cases} 
+		\frac{1}{N_{p_j}}, & z_j \in r_{p_j}; \\
+		0, & \text{otherwise.} \\
+	\end{cases}
+\end{align}
+
+These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
+We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
+
+%We emphasize the computational gains of this split sampler: the first step is a sample from $P + 1$ options, where $P$ does not scale with the size of the linkage task; the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
+\begin{lemma}
+	Recall that $n_A$ and $n_B$ are the number of records in $A$ and $B$, respectively.
+Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $A \times B$.
+We assume $C$ cores available for parallelization and a Gibbs sampler with $T$ iterations.
+Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P)$.
+	\label{lemma:fabl}
+\end{lemma}
+\begin{proof}
+	We consider two steps: constructing the comparison vectors and the Gibbs sampler.
+The computational complexity of all pairwise comparisons across $A$ and $B$ is $O(F n_A n_B)$.
+The hashing procedure for all pairwise comparisons is also $O(F n_A n_B)$.
+With $B$ processors available, we can split these computations across $B$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{B} n_A n_B)$.
+There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
+
+	Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_A n_B)$.
+However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
+The complexity of updating $\bm{Z}$ sequentially at the record level is $O(n_A n_B)$.
+With hashing, we split this sampling into two steps.
+First, we sample the agreement pattern of the match with complexity $O(n_B P)$, and then we sample the record exhibiting that pattern with complexity $O(n_B)$.
+Thus, the complexity of sampling $\bm{Z}$ in a single iteration is $O(n_B P)$.
+Since $P << n_A$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_A n_B)$ under \texttt{BRL} to $O(n_B P)$ under \texttt{fabl}.
+With parallelization, this complexity is further reduced to $O(\frac{1}{C}n_B P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{C}n_B P)$
+	In summary, the total computational complexity is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P).$
+\end{proof} 
+%\begin{align}
+%	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\bm{\Gamma}}) &= \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+%&=  \prod_{j=1}^{n_B}\prod_{i=1}^{n_A}\prod_{p=1}^P \left[ m_{fl}^{I(Z_j = i)}u_{fl}^{1 - I(Z_j = i)}\right]^{I(e(i,j) = h_p)}\\
+%&=  \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}\\
+%\end{align}
+
+%In this form, the likelihood reveals straightforward posterior updates for the $\bm{m}$ and $\bm{u}$ parameters.
+Let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define $\bm{\beta_0}$ similarly for the $\bm{u}$ distributions.
+Then, conditional on $\bm{Z}$, we have
+%\begin{align}
+%	\mathcal{L}(\bm{m}, \bm{u} \mid \bm{\Gamma}, \bm{Z})  &=  \prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{N_p^m}u(p)^{N_p - N_p^m}\\
+%\end{align}
+%where $N_p^m = \sum_{j=1}^{n_B} I\left(e_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ is total occurrence of pattern $p$ in the data across all record pairs.
+In this form, we see through factorization that $\mathcal{N}$, $\mathcal{P}$, and $\mathcal{R}$ are sufficient for $\bm{m}$ and $\bm{u}$.
+%
+%Viewing the likelihood at the level of patterns also allows us for efficient  We also see that the terms needed for the posterior updates needed for level $l$ of field $f$ are given by the appropriate component of the vectors
+%\begin{subequations}
+%	\begin{align}
+%		\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P N_p^m \times h_p, \label{efficient_alpha} \\
+%		\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P (N_p - N_p^m) \times h_p. \label{efficient_beta}
+%	\end{align}
+%\end{subequations}
+%}
+
+%Updating  $\alpha_{fl}(\bm{Z})$ and $\beta_{fl}(\bm{Z})$ for each field and level in the linkage task constitutes $2 \times \sum_f L_f$ many summations over $n_A \times n_B$ quantities.
 These are simple calculations, but they become computationally burdensome when working on large linkage tasks.
 
-Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$.
-Let $H_p^m = \\ \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p, Z_j \leq n_1 \right)$ be the number of matching record pairs with agreement pattern $p$.
-It follows that the number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
+%Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$.
+Let $H_p^m = \\ \sum_{j=1}^{n_B} I\left((Z_j, j) \in h_p, Z_j \leq n_A \right)$ be the number of matching record pairs with agreement pattern $p$.
+It follows that the number of non-matching record pairs with agreement pattern $p$ is $H_p^u = H_p - H_p^m$.
 Let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_f})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define  $\bm{\beta_0} = (\beta_{11}, \ldots, \beta_{F L_f})$ similarly for the $\bm{u}$ distributions.
 Then, the terms needed for the posterior updates in (\ref{eqn:m_update}) and (\ref{eqn:u_update}) are contained in the vectors
-\begin{subequations}
-\begin{align}
-\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p, \label{efficient_alpha} \\
-\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p. \label{efficient_alpha}
-\end{align}
-\end{subequations}
-
-Although sampling \(Z_j\) from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_1$ is large.
-This is because sampling a value from $n_1$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
+
+%Although sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_A$ is large.
+This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
 To speed up computation, we break this sampling step into two.
 First, we calculate the quantity expressed in (\ref{eqn:fs_weight}) associated with each unique pattern, and denote this quantity as $w_{p}$.
-We then sample the agreement pattern between \(j\) and its potential link, according to
-\begin{align}
-	\label{eqn:gibbs1}
-	P\left(\left(Z_j^{(s+1)}, j\right) \in h_p \mid \tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
-	\begin{cases} 
-		w_{p}\times N_{j_p},  & p \leq P; \\
-		n_1 \frac{n_2 - n_{12}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{12}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   p = P + 1. \\
-	\end{cases}
-\end{align}
-%More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,
-%	\begin{align}
+We then sample the agreement pattern between $B_j$ and its potential link, according to
+%\begin{align}
 %	\label{eqn:gibbs1}
-%		P\left(h\left(Z_j^{(s+1)}\right) = h_p \mid\tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+%	P\left(\left(Z_j^{(s+1)}, j\right) \in h_p \mid \tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
 %	\begin{cases} 
-%			w_{p}\times H_{j_p},  & p \leq P; \\
-%			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, &   p = P + 1; \\
+%		w_{p}\times N_{p_j},  & p \leq P; \\
+%		n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   \text{otherwise}. \\
 %	\end{cases}
-%	\end{align}
-Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns.
-Since all records in $\bm{X}_1$ sharing the same agreement pattern with $j \in \bm{X}_2$ are equally likely, we can sample among candidate records uniformally using
-
-%After all iterations are complete, we can back-fill the records uniformally among records of the sampled agreement pattern through (\ref{eqn:gibbs2}).
-\begin{align}
-		\label{eqn:gibbs2}
-		P\left(Z_j^{(s+1)} = z_j \mid \left(Z_j^{(s+1)}, j\right) \in h_p \right) = \begin{cases} 
-			\frac{1}{N_{j_p}}, & (z_j, j) \in h_p; \\
-			0, & \text{otherwise.} \\
-		\end{cases}
-	\end{align} 
-%	\begin{align}
+%\end{align}
+%
+%Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the $P$ agreement patterns.
+Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we can sample among candidate records uniformally using
+%\begin{align}
 %		\label{eqn:gibbs2}
-%		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
-%			\frac{1}{H_{j_p}}, & (i, j) \in h_p; \\
+%		P\left(Z_j^{(s+1)} = z_j \mid e_{z_j^{(s+1)}, j} = h_p \right) = \begin{cases} 
+%			\frac{1}{N_{p_j}}, & e_{z_j, j} = h_p; \\
 %			0, & \text{otherwise.} \\
@@
 %	\end{align} 
-
-We emphasize the computational gains of this split sampler: the first step is a sample from $P + 1$ options, where $P$ does not scale with the size of the linkage task; the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
-These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
-
-The computational complexity of \texttt{fabl} is given in Lemma \ref{lemma:fabl}.
-%, and a summary of the \texttt{fabl} method through psuedocode in Appendix \ref{app:fabl-psuedocode}
-
-%	We close this section by proving the computational complexity of \texttt{fabl}.
-
-%	\begin{remark}
-%	We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
-\texttt{BRL} would see considerable speed up through the procedure, but computation time would still scale with the size of the larger dataset.
-%	\end{remark}
-
-%	\hypertarget{computational-complexity}{%
-%		\subsection{Computational Complexity}\label{computational-complexity}}
-
-	\begin{lemma}
-		Recall that $n_1$ and $n_2$ are the number of records in $\bm{X}_1$ and $\bm{X}_2$, respectively.
-Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $\bm{X}_1 \times \bm{X}_2$.
-We assume $B$ processors available for parallelization and a Gibbs sampler with $T$ iterations.
-Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(\frac{T}{B}n_2 P)$.
-		\label{lemma:fabl}
-	\end{lemma}
-
-	\begin{proof}
-		 We consider two steps: constructing the comparison vectors and the Gibbs sampler.
-The computational complexity of all pairwise comparisons across $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
-The hashing procedure for all pairwise comparisons is also $O(F n_1 n_2)$.
-With $B$ processors available, we can split these computations across $B$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
-There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
-
-		Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
-The complexity of updating $\bm{Z}$ sequentially at the record level is $O(n_1 n_2)$.
-With hashing, we split this sampling into two steps.
-First, we sample the agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
-Thus, the complexity of sampling $\bm{Z}$ in a single iteration is $O(n_2 P)$.
-Since $P << n_1$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
-With parallelization, this complexity is further reduced to $O(\frac{1}{B}n_2 P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{B}n_2 P)$
-		In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(\frac{T}{B}n_2 P).$
-	\end{proof}
-
 	\section{Simulation Studies}
@@
 We use $F = 5$ binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
-For each record in $\bm{X}_2$, we simulate $n_1$ comparison vectors, resulting in a comparison matrix $\Gamma \in \mathbb{R}^{n_1 n_2 \times F}$.
-For $n_2/2$ of these records, there is no match in $\bm{X}_1$, so we simulate $n_1$ comparison vectors from the $\bm{u}$ distribution.
-For the other $n_2/2$ of these records, there is one match in $\bm{X}_1$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_1 - 1$ comparison vectors from the $\bm{u}$ distribution.
-We compare the run-time of \texttt{fabl} against \texttt{BRL} as we increase $n_1$ and $n_2$.
-Since we have five binary comparison fields, the number of unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
-
-%	We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across data files.
+For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\Gamma \in \mathbb{R}^{n_A n_B \times F}$.
+For $n_B/2$ of these records, there is no match in $A$, so we simulate $n_A$ comparison vectors from the $\bm{u}$ distribution.
+For the other $n_B/2$ of these records, there is one match in $A$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_A - 1$ comparison vectors from the $\bm{u}$ distribution.
+We compare the run-time of \texttt{fabl} against \texttt{BRL} as we increase $n_A$ and $n_B$.
+Since we have five binary comparison fields with no missingness, the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
+
+%	We simulate these data for different values of $n_A$ and $n_B$, always with $n_B/2$ records in common across data files.
 and compare the run-time of \texttt{fabl} against \texttt{BRL}.
-Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
+Since we have 5 binary comparison fields, the number is unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
 
@@
 Additionally, although \texttt{fabl} is amenable to parallelization, this simulation is run on a single core.
-Implementing \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
-
-	In Figure \ref{fig:speed1}, where we increase both $n_1$ and $n_2$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger sample sizes.
-In particular, run-time for \texttt{BRL} grows quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} grows linearly (in the size of only $\bm{X}_2$).
+Implementing \texttt{fabl} in C++ with parallelization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
+
+	In Figure \ref{fig:speed1}, where we increase both $n_A$ and $n_B$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger sample sizes.
+In particular, run-time for \texttt{BRL} grows quadratically (or linearly with the size of both $A$ and $B$) while run-time for \texttt{fabl} grows linearly (in the size of only $B$).
 
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$, as described in Section \ref{speed}.
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
 We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
@@
 
-	%The above discussion suggests that for fixed \(n_2\), computation time for \texttt{fabl} should remain mostly constant with growing \(n_1\).
-
-	In Figure \ref{fig:speed2}, where we fix \(n_2 = 500\), we see near linear growth for the run-time under \texttt{BRL} as \(n_1\) increases, and much more static run-time under \texttt{fabl}.
+	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}.
 The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_2$ fixed at 500, as described in Section \ref{speed}.
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
 We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
@@
 	We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}.
-Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(Z_j \leq n_1)$.
-Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1) / \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)$.
+Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(Z_j \leq n_A)$.
+Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(\hat{Z}_j \leq n_A)$.
 The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
@@
 In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.
-
-	%Each metric is provided explicitly below:
-	%$$\text{Recall } = \frac{\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1)}{\sum_{j=1}^{n_2} I(Z_j \leq n_1)}$$
-	%$$\text{Precision } = \frac{\sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j, Z_j \leq n_1)}{ \sum_{j=1}^{n_2} I(\hat{Z}_j \leq n_1)}$$
-	%$$\text{F-measure } = 2 \frac{\text{Recall } + \text{Precision}}{\text{Recall } \times \text{Precision}}$$
 
@@
 
-	%\hypertarget{scale}{%
-		%	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
-	%
-	%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated data files from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
-Under standard Fellegi-Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
-Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task.
-Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
-	% 
-	%For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
-We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
-The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method.
-Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run.
-However, this could be substantially sped up using distributed computing.
-	%
-	%Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
-This is slightly worse performance than  in the smaller simulation studies; this is expected because it is naturally more difficult to link more records with the same amount of information.
-With more linkage fields, \texttt{fabl} maintains the high accuracy in Section \ref{accuracy}.
-
 	\section{Case Studies}
@@
 
-	%This last point nearly breaks the earlier mentioned assumption that there are no duplicates within files, and reveals a key difference between \texttt{BRL} and \texttt{fabl}.
-
-	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in \(n_1 = 4420\) records in CHDES and \(n_2 = 1323\) records in ERTL.
+	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in $n_A = 4420$ records in CHDES and $n_B = 1323$ records in ERTL.
 We standardize names to account for common misspellings and use a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
@@
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We use uniform priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
+We use uniform priors for the $\bm{m}$ and $\bm{u}$ parameters.
 
 	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
+This results in $4^5 \times 2 = 2048$ possible agreement patterns, with 1173 patterns realized in the data.
 However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
-
-%	\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
-We initially used these comparisons to estimate fabl and BRL.
-However, after investigating the posterior distributions of the $\bm{m}$ and $\bm{u}$ parameters, we noticed that posterior distributions of several levels of the m and u parameters are nearly identical, suggesting that these levels are unnecessary for meaningfully modeling the distribution of the comparison vectors.
-
-	%However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
-
-	%here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
 
@@
 		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \cite{sadinle_bayesian_2017}.
-This set up leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
+This setup leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
 	\end{table}
@@
 	\caption{Construction of comparison vectors for El Salvador data for increased speed under \texttt{fabl}.
-This set up leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
+This setup leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
 \end{table}
@@
 Hence, the assumptions of our model hold for this study.
-We seek to link records over the $n_1 = 20485$ individuals from 1982 to the $n_2 = 17466$ individuals from 1989.
+We seek to link records over the $n_A = 20485$ individuals from 1982 to the $n_B = 17466$ individuals from 1989.
 The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
@@
 To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
-These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $\bm{X}_1$ has
-	%multiple plausible matches in $\bm{X}_2$. 
+These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $A$ has
+	%multiple plausible matches in $B$. 
 	%
@@
 Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
-In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and \(\gamma_{2776, 826}\) are exactly identical.
+In addition, using the modified Levenstein distance, the comparison vectors $\gamma_{2776, 825}$ and $\gamma_{2776, 826}$ are exactly identical.
 	%
@@
 	%Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
-Under \texttt{fabl}, both records in $\bm{X}_2$ match to the same record in $\bm{X}_1$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
+Under \texttt{fabl}, both records in $B$ match to the same record in $A$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
 Under \texttt{BRL}, the Gibbs process creates one matching configuration and stays there for a while.
@@
 
-	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing all $n_1 \times n_2$ record pairs still can be prohibitive for larger linkage tasks.
+	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing all $n_A \times n_B$ record pairs still can be prohibitive for larger linkage tasks.
 Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
@@
 	%Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
-We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
-Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within data files.
+We explored ``one-to-many'' scenarios in which one record in $B$ has multiple plausible matches in $A$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
+Other issues arise under ``many-to-one'' scenarios, where one record in $B$ has multiple plausible matches in $A$, and ``many-to-many'' scenarios in which there is duplication both across and within data files.
 Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
@@
 However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons.
-Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block.
-Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.
+Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $\bm{m}$ and $\bm{u}$ parameters that are interpretale for the entire study, rather than separate parameters for each block.
+Although these $\bm{m}$ and $\bm{u}$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.
 
@@
 
-We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage parameter \(\bm{Z}\) by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
-We adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which \(\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\}\), with \(R\) representing the option to leave the matching undetermined by the model.
+We calculate a Bayes estimate $\hat{\bm{Z}}$ for the linkage parameter $\bm{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
+We adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which $\hat{Z}_j \in \{1, \ldots, n_A, n_A + j, R\}$, with $R$ representing the option to leave the matching undetermined by the model.
 Specifically, we have
@@
 		\theta_R,  & \text{if } \hat{Z_j} = R; \\
-		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
-		\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
-		\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
+		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_A + j ; \\
+		\theta_{01},  & \text{if } Z_j = n_A + j,\hat{Z_j} \leq n_A ; \\
+		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
 	\end{cases}\] 
-Here, \(\theta_R\) is the loss from not making a decision on the linkage status, \(\theta_{10}\) is the loss from a false non-match, \(\theta_{01}\) is the loss from a false match, and \(\theta_{11}\) is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
+Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false non-match, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
 
@@
 Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
+In the event that we have two records $B_j$ and $B_{j'}$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
 Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
@@
 %			\hline
-%			$\bm{X}_1, \bm{X}_2$ & data files\\
-%			$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
-%			$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
+%			$A, B$ & data files\\
+%			$i \in 1, \ldots, n_A $ & index over records in $A$\\
+%			$j \in 1, \ldots, n_B $ & index over records in $B$\\
 %			$f \in 1, \ldots F$ & index over fields used for comparisons \\
 %			$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
-%			$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
-%			$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
-%			$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
-%			$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
+%			$n_{AB}$ & number of entities in common between $A$ and $B$\\
+%			$\bm{\gamma}_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+%			$Z_j = i$ & records $A_i$ and $B_j$ match \\
+%			$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
 %			$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
 %			$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-%			$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
-%			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+%			$\pi$ & probability that a record in $B$ has a match in $A$ \\
+%			$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
 %			$h_p$ & one hot encoding of agreement pattern $p$ \\
-%			$H_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+%			$H_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
 %			$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
@@
 \begin{align*}
-	p(m_{fl}| \Gamma, \bf{Z}) &\propto \prod_{i=1}^{n_1} \prod_{j=1}^{n_2} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
+	p(m_{fl}| \Gamma, \bm{Z}) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
 	&= m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
@@
 where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$.
-Analogous procedures lead to the posterior distribution $p(u_{fl}| \Gamma, \bf{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
+Analogous procedures lead to the posterior distribution $p(u_{fl}| \Gamma, \bm{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
 Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
@@
 
-In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihood relevant to a particular record $j \in \bm{X}_2$.
-Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
-Thus, we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
+In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihood relevant to a particular record $B_j$.
+Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(z_j < n_A + j)$ encoded by $\bm{Z}$ (and hyperparameters).
+Thus, we use $p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
 \begin{align*}
 	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}.
+	&\propto \pi^{n_{AB}(\bm{Z})} (1-\pi)^{n_B - n_{AB}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+	&\propto \pi^{n_{AB}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_A - n_{AB}(\bm{Z}) + \beta_{\pi} -1}.
 \end{align*}
-Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\). We have
+Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})$ distribution.
+
+Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $B_j$. We have
 \begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
-	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
-	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	p(\Gamma_{.j}|Z_j = z_j, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_A}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+	&\propto \prod_{i=1}^{n_A}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
+	&\propto \prod_{i=1}^{n_A}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
 	&=
 	\begin{cases} 
-		w_{z_j, j}  & z_j \leq n_1; \\
-		1 &  z_j  = n_1 + j, \\
+		w_{z_j, j}  & z_j \leq n_A; \\
+		1 &  z_j  = n_A + j, \\
 	\end{cases}\\
@@
 With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$.
-Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\):
+Because we sample each $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$:
 \newpage
@@
 	&\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
-	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_1 + j). \\
+	&\propto \left(\sum_{i=1}^{n_A}w_{z_j, j}I(z_j = i) + I(z_j = n_A + j) \right) \left (\pi\sum_{i=1}^{n_A}\frac{1}{n_A}I(z_j = i) + (1-\pi)I(z_j = n_A + j) \right) \\	
+	&= \frac{\pi}{n_A}\sum_{i=1}^{n_A}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_A + j). \\
 	&= \begin{cases} 
-		\frac{\pi}{n_1}w_{z_{j}, j}   & z_j \leq n_1; \\
-		1-\pi &  z_j  = n_1 + j. \\
+		\frac{\pi}{n_A}w_{z_{j}, j}   & z_j \leq n_A; \\
+		1-\pi &  z_j  = n_A + j. \\
 	\end{cases}
@@
 	&= \int_{\pi}p\left(Z_j^{(s+1)} | \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-	&= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(z_j = n_1 + j) \\
-	&= \frac{n_{12}(\bm{Z}) + \alpha_{\pi}}{n_1 (n_2 + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_1) + \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_2 + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_1 + j) \\
-	&\propto w_{z_j, j}I(z_j\leq n_1) + n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_1 + j) \\
+	&= \int_{\pi}\left[\frac{\pi}{n_A}w_{z_{j}, j} I(z_j\leq n_A) + (1-\pi)I(z_j = n_A + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_A}w_{z_{j}, j} I(z_j\leq n_A) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(z_j = n_A + j) \\
+	&= \frac{n_{AB}(\bm{Z}) + \alpha_{\pi}}{n_A (n_B + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_A) + \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_B + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_A + j) \\
+	&\propto w_{z_j, j}I(z_j\leq n_A) + n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_A + j) \\
 	&=
 	\begin{cases} 
-		w_{z_j, j}  & z_j \leq n_1; \\
-		n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
+		w_{z_j, j}  & z_j \leq n_A; \\
+		n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_A + j. \\
 	\end{cases} \\
@@
 %	
-%	\textcolor{red}{I am getting rid of the notation $h(Z_j)$ for the agreement pattern between $j$ and its potential match.
+%	\textcolor{red}{I am getting rid of the notation $h(Z_j)$ for the agreement pattern between $B_j$ and its potential match.
 Instead, I am keeping notation that already exists, saying $(Z_j, j) \in h_p$.
@@
 %			\State Partition files $\bm{X_1}$ and $\bm{X_2}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$.
-%			\For{each $n$, $m$}
+%			\For{each $n$, $\bm{m}$}
 %			\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$.
@@
 Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual non-matches.
-Mathematically, $\text{NPV} = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$.
+Mathematically, $\text{NPV} = \sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j = n_A + j)$/$\sum_{j=1}^{n_B} I(\hat{Z}_j = n_A + j)$.
 We continue to use the precision, which is renamed the positive predictive value (PPV) in this context.
-Lastly, we also examine the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)/n_2$.
+Lastly, we also examine the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_B} I(\hat{Z}_j = R)/n_B$.
 To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.
@@
 
-%This is reasonable because large numbers of errors blur the distinction between matched and non-matched pairs, leading to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records.
+%This is reasonable because large numbers of errors blur the distinction between matched and non-matched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
 
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_1$ and $n_2$.
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
 We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
@@
 		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_1$ and $n_2$ fixed at 500.
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
 We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}

### 2023-05-02 ncvr_raleigh/fastlink.R

@@ -1,0 +1,57 @@
+library(dplyr)
+library(fastLink)
+library(parlrdev)
+library(purrr)
+
+
+dfA <- readRDS("data/raleigh_A")
+dfB <- readRDS("data/raleigh_B")
+nA <- nrow(dfA)
+nB <- nrow(dfB)
+
+Ztrue <- MakeZtrue(dfA$voter_id, dfB$voter_id,
+                   nA, nB)
+true_matches <- sum(Ztrue < nA + 1)
+
+linkage_vars <- c(4, 5, 6, 7, 9, 13)
+
+fl_dedup <- fastLink(dfA[, linkage_vars], dfB[, linkage_vars],
+                    varnames = names(dfB[, linkage_vars]),
+                    dedupe.matches = T,
+                    threshold.match = .5)
+
+Zhat <- MakeZhat_from_fastlink(fl_dedup$matches$inds.a,
+                               fl_dedup$matches$inds.b)
+
+declared_correct <- sum(Zhat == Ztrue & Zhat < (nA + 1))
+eval <- c(GetEvaluations(Zhat, Ztrue, nA), 'fastlink_resolved')
+names(eval)[4] <- "method"
+
+# Full
+fl_full <- fastLink(dfA[, linkage_vars], dfB[, linkage_vars],
+                     varnames = names(dfB[, linkage_vars]),
+                     dedupe.matches = F,
+                     threshold.match = .5)
+
+declared_correct <- imap(Ztrue, ~sum((fl_full$matches$inds.a == .x) &
+                           (fl_full$matches$inds.b == .y))) %>%
+  unlist() %>%
+  sum()
+
+recall <- declared_correct/true_matches
+
+total_declared <- length(fl_full$matches$inds.a)
+precision <- declared_correct/total_declared
+fmeasure <- 2 * (recall * precision) / (recall + precision)
+
+eval_full <- c(recall, precision, fmeasure, "fastlink")
+names(eval_full) <- c("Recall", "Precision", "Fmeasure", "method")
+
+result <- rbind(eval_full, eval)
+rownames(result) <- NULL
+
+saveRDS(result, "eval/fastlink/result")
+# fl_full <- fastLink(df82[, -c(7, 8)], df89[, -c(7, 8)],
+#                     varnames = names(df82[, -c(7, 8)]),
+#                     dedupe.matches = F)
+# sum(df82$unique_ID[fl_out$matches$inds.a] == df89$unique_ID[fl_out$matches$inds.b])

### 2023-05-02 ncvr_raleigh/gather_results.R

@@ -2,11 +2,14 @@ library(RecordLinkage)
 library(dplyr)
 library(stringr)
 
-result_paths <- c("eval/fabl/result",
+result_paths <- c("eval/fabl/independent",
   "eval/supervised/aggregate/result",
   "eval/fastlink/result")
 
 results <- lapply(result_paths, readRDS) %>%
   do.call(rbind, .)
 
+# results <- results[-2, ]
+results[1, 4] <- "fabl"
+results[1, 4] <- "fabl_resolved"
 saveRDS(results, "eval/all_results")

### 2023-05-09 ba_resubmit/third-submission/suggestions/third-revision-review-comments.tex
update rcs notes.

\documentclass{article}
\usepackage[numbers]{natbib}
%\usepackage{hyperref}
\usepackage{lineno}
\usepackage{amsthm}
\usepackage{amsmath, color}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
%\usepackage{graphicx}
%\newtheorem{lemma}{Lemma}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}

\newcommand{\bX}{\boldsymbol X}
\newcommand{\bZ}{ \boldsymbol Z}
\newcommand{\bh}{\boldsymbol h}
\newcommand{\bc}{\boldsymbol c}
\newcommand{\btheta}{\boldsymbol \theta}
\newcommand{\LL}{\mathcal{L}} 
\newcommand{\C}{\mathcal{C}} 
\newcommand{\R}{\mathbb{R}} 
\newcommand{\CC}{\mathfrak{C}} 
\newcommand{\PP}{\mathbb P} 
\newcommand{\pairP}{\mathcal{P}}

\begin{document}
\title{AE Comments}
\author{}
\date{}
\maketitle

%\begin{enumerate}
%\item I would revert back to the original notation as I find the A, B to be confusing and it's not easy to extend to multiple files. (See Aleshin-Guendel (2022) for reference. I'm not sure what Jerry thinks about this though. Typically changes should be agreed upon by all authors or some standard.
%\item Perhaps to make things easy on you, I can propose some changes in the document and help and iterate and then we can ask Jerry regarding his opinions regarding this. 
%\item Let us also make a plan for outlining your thesis and a schedule so it can move ahead. Would you like suggestions or to propose something yourself? 
%\end{enumerate}

\section{AE Comments with responses}

The comments to the AE need to be extremely precise and accurate. I have attempted to write one in painful detail to provide guidance and an example.

\begin{enumerate}
\item Line 111. Is this missing at random or something else? This needs to be clarified (AE Comment 1). This needs to be addressed with extreme care regarding the precise wording. \\
\textcolor{blue}{This distinction between truly missing at random and missing completely at random does not matter due to two assumptions. Following Sadinle (2017), we assume that the fields in the comparison vectors are conditionally independent given the coreference matrix and we assume that the missing comparisons are ignorable (Sadinle (2017). Therefore, our likelihood is the same as in Sadinle (2017) (see equations (3) and (4) of his paper). As stated by Sadinle (2017), ``we work under the assumption of ignorability of the missingness mechanism for the comparisons so that we can base our inferences on the marginal distribution of the observed comparisons (Little and Rubin, p. 90)."  We refer to Sadinle (2017), Section 4.2, page 604 for further details.}
\item Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and
	think it was a good choice to move the full derivation to an appendix. However, I do not follow
	the re-expression of the pmf for $\Gamma_{.j}$. First, the opening square brackets are still misplaced
	throughout. In the second line of this derivation, the authors divide the expression by a product
	of the element-wise conditional probabilities of a match in the comparison vector (u), raised by
	an indicator that the element of the comparison vector equals a particular value. Perhaps I
	misunderstand, but I believe this product is not constant in $\Gamma_{.j}$ or $u$, and so the total expression
	is not proportional to the line above, as interpreted as a pmf. I think that the end effect is that
	the last line of the newly expressed pmf is missing a factor of \\
%	$\prod_i \prod_f \prod_l u_{fl}^{I(\gamma_{ij}^f = l) I_{obs}$\\
	\textcolor{blue}{Is the new proof, Brian, missing an indicator over the cluster assignment? I believe that it should match those 
	of Sadinle (2017), Aleshin and Guendel (2022), and Wortman (2019). Where is the indicator regarding the cluster assignment that you condition on? Specifically, the update for each gamma will depend upon your cluster assignment, so I believe there is a typo for this reason. (To help with notation,  Jody calls this $S_k$, Sadinle calls it $Z_{-j}$, and Serge calls it $\C_{-j}$). Could you carefully check this please? I think this question is related to the next point/comment by the AE.}
    \item Appendix 8.2: Thank you for including the details on integrating out pi from the full
conditionals. They surprised me. I had assumed that the authors had integrated out pi in the
prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite
matchings”. Does the alternate approach presented in this paper provide a different algorithm
than directly integrating the prior distribution? Is the presented alternative approach justified? \\
\textcolor{blue}{
Thank you for noting your surprise regarding our proposal. We do not extend the approach of Sadinle (2017), but relax the one-to-one assumption for computational reasons. This motivates the choice of our prior (instead of the beta prior for bipartite matchings). 
Note: I do not believe that we should be integrating out the prior distribution. I think we did this at the confusion from the first revision of the  AR. We should instead just update the conditionals to form a Gibbs update, which will make the algorithm justified. If you believe that the algorithm is correct, we will need to convince the AE why it is a valid Markov chain or rather Gibbs update of the partition. 
}
\textcolor{blue}{It will not be justified unless the full conditionals of the Gibbs sampler contain a sequential update or one proves that this is a justified Gibbs sampler. This was the point that Serge mentioned in his email. Currently, it is not a valid Gibbs sampler. I attempt to explain this more clearly below.}

It would need to be something like this, which comes from Aleshin-Guendel and Steorts (2023). The partition is equivalent to the co-reference matrix or linkage structure. The partition approach is more widely understood by the Bayesian non-parametrics community (or statistics community), which is why we are using this terminology. (They will not easily understand the linkage structure or coreference matrix without connections being made for them).

\begin{quote}
			`` The most common way to update the partition, $\C$, in Bayesian entity resolution is the Gibbs update \citep{Sadinle_2014, Steorts_2015, Sadinle_2017, Tancredi_2020, Marchant_2021, Aleshin-Guendel_2022, Marchant_2023}, which originated in the Bayesian clustering literature \citep[see e.g.][]{Neal_2000}.
			 We emphasize that this is only referring how the \textit{partition} is being updated.

The Gibbs update works by sampling the cluster assignment for each record $i \in [R]$ sequentially. To describe the update we will represent a partition $\C$ by a  vector of labels $\bZ=(Z_1,\dots, Z_R)$, where $Z_i\in\{1,\dots, R\}$ such that $Z_i=Z_j$ if records $i$ and $j$ are in the same cluster in $\C$. Suppose that we have already sampled the cluster assignment for the first $i-1$ records. Sampling the cluster assignment for record $i$ proceeds as follows:
			\begin{enumerate}
				\item Let $\bZ^{[t]}_{-i}=(Z_1^{[t]},\dots,Z_{i-1}^{[t]},Z_{i+1}^{[t-1]},\dots, Z_R^{[t-1]})$ denote the current partition of $[R]$ after removing record $i$. 
				\item Let $\bc$ denote the union of the unique labels occupied in $\bZ^{[t]}_{-i}$ and one unoccupied label. Each of these labels correspond to a cluster to which record $i$ could be assigned, with the unoccupied cluster representing the creation of new cluster with only record $i$.
				\item Sample record $i$'s cluster assignment, by sampling the cluster $c\in \bc$ with probability proportional to $p(Z_{i}^{[t]}=c\mid f(\bX), \bZ^{[t]}_{-i}, \btheta^{[t]})$.
			\end{enumerate}
			"

\end{quote}
\item Sections 4.1 and 4.2 (I have gone through this, however, I suggest that we all go through this many times and write out responses very clearly). I am doing my best to get through this as quickly as possible and make recommendations.  
\end{enumerate}

\end{document}

### 2023-05-09 ba_resubmit/third-submission/suggestions/third-revision-review-comments.tex
updated comments from ae.

@@
 Note: I do not believe that we should be integrating out the prior distribution.
-I think we did this at the confusion from the first revision of the  AR.
+I think we did this at the confusion from the first revision of the  AE.
 We should instead just update the conditionals to form a Gibbs update, which will make the algorithm justified.
@@
 Currently, it is not a valid Gibbs sampler.
-I attempt to explain this more clearly below.}
+I attempt to explain this more clearly below.
+Looking at Pitman, Pruenster, and the BNP literature, I cannot find a reference regarding why the Markov chain would be valid in this case.
+If you were to prove this, it would be a new result (to my knowledge).}
 
-It would need to be something like this, which comes from Aleshin-Guendel and Steorts (2023).
+It would need to be something like this, which comes from Aleshin-Guendel and Steorts (2023), which extends the general clustering literature.
 The partition is equivalent to the co-reference matrix or linkage structure.

### 2023-06-07 ba_resubmit/response_to_reviewers2.tex
corrected SEI

@@
 \reply{%
- {\color{red}{I really don't know how to answer this.
+	The reviewer raises important questions.
+The main difference between the full conditionals in Sadinle (2017) and ours is that for him, $Z$ is a partition, while for us, it is not.
+This is what required him to integrate out $\pi$ in the prior, and then sequentially update each component of the $Z$ vector.
+
+	The approach we have is justified, as indicated in the proof provided.
+However, under closer review, we have determined that it is unnecessary.
+Without integrating out $\pi$, all paramaters are updated through a standard Gibbs sampler.
+Furthermore, in other work we are developing, we actually need samples of $\pi$ in different applications.
+
+	We reran all case studies and simulation using samples of $\pi$ in the full conditional for $Z$, and found no difference in results.
+This is unsuprsing, because with high numbers of observations, the posterior of the beta distrubution is highly concentrated around its mean.
+
+	Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler. We leave the remark about marginalization in the Appendix to highlight the similarity to Sadinle's method.
+
+%	
+% {\color{red}{I really don't know how to answer this.
 I don't know how I would ''directly integrate" $\pi$, and I don't know if what I do is ''justified." Its kind of an intellectual relic to how I was originally thinking about the problem, and it reveals a quantity similar to what Sadinle was working with.
-
- It is theoretically more straightforward to just not integrate out $\pi$, and use samples from $\pi$ in Gibbs sampler like I do with all other parameters.
+% 		
+% It is theoretically more straightforward to just not integrate out $\pi$, and use samples from $\pi$ in Gibbs sampler like I do with all other parameters.
 Would that be too large a change at this stage?
-
- Without integrating out $\pi$, the full conditional is given on page 26 of the Appendix.}}
+%
+% Without integrating out $\pi$, the full conditional is given on page 26 of the Appendix.}}
 }
@@
 If this argument seems weak, or is too much new material, I can also just keep what I had, and use less rigorous term "summary statistics" instead.}} \todo
-
 }

### 2023-06-07 Prelim/bcm_proposal.tex
corrected SEI

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }

\begin{document}
	\title{Bayesian Causal Matching with Fellegi Sunter}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger}
	\institute{Duke University}
	\date{\today}

%	\AtBeginSection[]{
%		\begin{frame}
%			\vfill
%			\centering
%			\begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%				\usebeamerfont{title}\insertsectionhead\par%
%			\end{beamercolorbox}
%			\vfill
%		\end{frame}
%	}

\AtBeginSection[]
{
	\begin{frame}
		\frametitle{Table of Contents}
		\tableofcontents[currentsection]
	\end{frame}
}

	\begin{frame}
		\titlepage
	\end{frame}

\section{Matching in Causal Inference}
\begin{frame}{Overview of Approaches}
	\begin{itemize}
		\item Exact Matching
		\item Mahalonobis Distance
		\item Propensity Score
		\item Almost Exact Matching
	\end{itemize}
\end{frame}

\begin{frame}{Mahalonobis Distance}
	\begin{itemize}
	\item One method of matching has been to calculate distance in the covariate space through
		\begin{align}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1} |X_1 - X_2|}
		\end{align}
	\item An algorithm then matches each unit in control to closest unit in treated
	\end{itemize}
\end{frame}

\begin{frame}{Propensity Score}
	\begin{itemize}
	\item Another method is to estimate propensity scores $p(Z_n = 1 |X_n)$ for each unit in the data, and match units based on this score.
	\end{itemize}
\end{frame}

\begin{frame}{Pitfalls of Both Approaches}
	\begin{itemize}
		\item The pros and cons of each method are summarized very well on this \href{https://stats.stackexchange.com/questions/511294/what-are-the-pros-and-cons-of-using-mahalanobis-distance-instead-of-propensity-s}{Stack Exchange} page

		\item Essentially, propensity scores can end up matching units with that have similar probability of treatment, but have very different covariate values
		\item Conversely, Mahalonobis can end up matching units that are similar on features that are irrelevant to the outcome of interest
	\end{itemize}
\end{frame}

\begin{frame}{Blending the Two Approaches}
	\begin{itemize}
		\item There is also a good summary on page 5 of \href{https://www.jstatsoft.org/article/view/v042i07}{Sekhot (2011)} recommends including the propensity score into the Malahalonobis distance. This was the genesis for the idea to come
	\end{itemize}
\end{frame}

\section{Proposed Idea}

\begin{frame}{My Idea}
	\begin{itemize}
		\item The machinery of Fellegi Sunter record linkage is very good at identifying duplicate records. It should also be good at identifying separate individuals that are similar.
		\item FS does matching purely on the reliability and uniqueness ($m$ and $u$ parameters) of fields of information. It does not take into account any relationship with an outcome variable. Conceptually, this is very similar to the Mahalonobis distance, but allows for uncertainty quantification.
		\item If we include propensity scores into FS, the model will place high weight on units with ``similar" scores. Then, the FS machinery for the other fields should find the units that are most similar. 
	\end{itemize}
\end{frame}

\begin{frame}{A Similar Idea}
	\begin{itemize}
		\item\href{https://arxiv.org/pdf/2105.02362.pdf}{Alvarez and Levin (2021)} (Section 2, pages 5 - 7) use a Bayesian model to create draws of the propensity scores. For set of scores, they use a deterministic algorithm to match records. 
		\item I imagine that each particular draw of matched pairs may have the issues presented earlier. Propensities are matched, but units may actually look quite different.
		\item Information does not seem to propagate correctly through the Gibbs Sampler
	\end{itemize}
\end{frame}

\begin{frame}{My Idea}
	\begin{itemize}
		\item Data is $\{y_n, z_n, x_n\}_{n = 1}^N$ where $y_n$ is the outcome, $z_n$ is the treatment indicator, and $x_n$ is a vector of $F$ fields of covariates. 
		\item Split data into a datasets $T$ with records $i \in \{1, \ldots, n_T\}$ for treated units and $C$ with records $j \in \{1, \ldots, n_C\}$. 
		\item At each step of the Gibbs Sampler, calculate propensity score $PS(i) = P(Z_i = 1 | X)$ for each data point (logistic or probit regression)
	\end{itemize}
\end{frame}

\begin{frame}{My Idea}
	\begin{itemize}
		\item Define comparison field 
		\begin{align}
			\Gamma_{0} = \begin{cases}
				1 & $PS(i)$ and $PS(j)$ \text{ are sufficently close} \\
				0 & \text{otherwise} \\
			\end{cases}
		\end{align} 
		\item ''Sufficiently close" could mean all units with scores within a certain caliper, or it could mean the $K$ nearest scores.
		\item Conduct Fellegi Sunter Matching on new full set of comparison vectors $\Gamma^{*} = (\Gamma_0, \gamma)$.
		Note $\gamma$ is fixed data, but the $\Gamma$ are random. 
	\end{itemize}
\end{frame}

\begin{frame}
	\begin{itemize}
		\item Each iteration of the Gibbs sampler will provide a matched dataset, from which you can calculate an MCMC sample of the treatment effect (TE) or other quantities of interest
		\item We obtain measures of match quality as usual, through the posterior samples of the linkage structure
		\item I believe this to be first causal matching method to produce matchings probabilistically
		\item The proposed method is fully model based, with no deterministic algorithms. It propogates uncertainty between the matching procedure and the treatment effect estimation. 
	\end{itemize}
\end{frame}

\begin{frame}{Model (Basic)}
	\begin{align}
		Z_n &= x_n' \beta + \epsilon \\
		\beta &\sim N(0, \tau) \\
		\epsilon &\sim N(0, \sigma^2) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l) \\
		m, u \sim \text{Dirichlet}
		\Delta \sim \text{Desired Record Linkage Prior}
	\end{align}
\end{frame}

\begin{frame}{Model (With Variable Selection)}
	\begin{itemize}
		\item In causal matching, it is especially important not to match on irrelevant features. Therefore, we can use a spike and slab prior on $\beta$, and only use fields where $\beta_f \neq 0$ in the linkage.
	\end{itemize}
\begin{align}
	\beta_j &\sim \pi_0 \delta_0 + \(1 - \pi\)N(0, \tau_j) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\left[\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l)\right]^{\beta_j \neq 0} \\
\end{align}
\end{frame}

\begin{frame}{Full Conditional for $\beta$}
	\begin{itemize}
		\item The full condition for $\beta_j$ includes a measure of the quality of matches along field $f$
	\end{itemize}
\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}
	\begin{itemize}
		\item Mathematically, any record linkage prior could work. Sadinle, fabl, or original FS (one-to-one, one-to-many, many-to-many) may all be useful , depending on the analysis
		\item If this works, would be a very strong final thesis chapter
	\end{itemize}
\end{frame}

\begin{frame}{Caveats}
	\begin{itemize}
		\item Using a comparison vector made from the propensity score may technically violate the conditional independence assumption of FS
		\item I can't imagine this idea totally failing, but I'm sure that it won't be cutting edge. Sekhot further devised a method that estimates
		\begin{align}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1/2} W S^{-1/2} |X_1 - X_2|}
		\end{align}
		where the entries in $W$ are fine tuned through optimization methods that encourage covariate balance across datasets, based on t-tests and all sorts of other measures. This is almost certainly outperform what I have presented, because it is able to use more information.
		\item Propensity and Mahalonobis matching both have theoretic garuntees. I don't know right now if this method would have them
		\item I know very little about causal inference! This would be a large undertaking
	\end{itemize}
\end{frame}

\begin{frame}{Questions}
	\begin{itemize}
		\item What do you all think of this idea?
		\item Jerry, how well versed are you in matching for causal inference? To what extent to should I seek guidance/collaboration from Fan Li or Alex Volfovsky?
	\end{itemize}
\end{frame}

\begin{frame}{Possible Theoretical Question}
	\begin{itemize}
		\item I'm really struck by the conceptual similarity between Mahalanobis distance and the FS weights. If you make some assumptions (categorical data, one-to-one matching), it might be possible to prove a correspondence between the two. 
	\end{itemize}
\end{frame}

\end{document}

### 2023-06-07 ba_resubmit/kundinger_resubmit2.tex
corrected SEI

@@
 Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
-Let $P$ be the number of patterns realized in $\Gamma$ and note that it is bounded above by $P^{*} = F + \prod_{f=1}^F L_f$, where we have noted the possibility of missing values.
+Let $P$ be the number of patterns realized in $\Gamma$ and note that it is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where we have noted the possibility of missing values.
 Note that this quantity is controlled by the modeller in their choices for $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.

### 2023-06-07 R/NLC_hash2.R

@@ -24,7 +24,7 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
   }
 
   ptm <- proc.time()
-  patterns <- GetUniquePatterns2(comparisons)
+  patterns <- GetUniquePatterns2(comparisons, R = R)
   elapsed_hash <- proc.time() - ptm
 
   parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
@@ -50,6 +50,7 @@ BKSimple_hash2 <- function(comparisons, m_prior = 1, u_prior = 1,
 
   m <- u <- rep(0, length(parameter_split))
   matches <- rep(0,P)
+  #set.seed(1)
 
   # Gibbs
   for(s in 1:S){

### 2023-06-07 SEI_sensitvity/SEI_analysis.R

@@ -1,0 +1,20 @@
+
+
+Z_paths <- list.files(path = "SEI_sensitvity/Zs", full.names = T)
+
+thing <- readRDS("SEI_sensitvity/Zs/Zs_01")
+Zs_list <- lapply(Z_paths, readRDS)
+
+distinct_Z <- lapply(Zs_list, function(x){
+  dim(unique(x, MARGIN = 2))[2]
+})
+
+eval_paths <- list.files(path = "SEI_sensitvity/eval", full.names = T)
+eval_list <- lapply(eval_paths, readRDS)
+recall_var <- lapply(eval_list, function(x){
+  var(x[1, ])
+})
+
+precision_var <- lapply(eval_list, function(x){
+  var(x[2, ])
+})

### 2023-06-07 cluster_jobs/SEI_sensitivity.R

@@ -1,0 +1,81 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+library(readr)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+files <- list.files(path = "data/SimulationDataFiles", full.names = T)
+all_patterns = TRUE
+
+m_prior = 1
+u_prior = 1
+alpha = 1
+beta = 1
+S = 1000
+iterations <- 20
+burn = 100
+show_progress = T
+fast = F
+all_patterns = TRUE
+
+
+overlap_vals <- c(50, 250, 450)
+
+j <-  2
+i <-  297
+
+overlap <- overlap_vals[j]
+
+records <- read_csv(files[i], col_types = cols())
+records$file <- rep(2:1, length.out = dim(records)[1])
+
+records <- records %>%
+  janitor::clean_names() %>%
+  mutate(rec_id = as.numeric(str_extract(rec_id, "\\d{3}")) + 1)
+
+n1 <- 500
+n2 <- 500
+#overlap <- n2/2
+
+Ztrue <- n1 + 1:n2
+Ztrue[1:overlap] <- 1:overlap
+
+file1 <- records %>%
+  filter(file ==1,
+         rec_id <= n1) %>%
+  select(-rec_id) %>%
+  as.matrix(.) %>%
+  data.frame(.) %>%
+  mutate(occup = as.numeric(occup))
+
+file2 <- records %>%
+  filter(file == 2,
+         rec_id %in% c(1:overlap, (n1 +1):(1000 - overlap))) %>%
+  select(-rec_id) %>%
+  as.matrix() %>%
+  data.frame(.) %>%
+  mutate(occup = as.numeric(occup))
+
+
+cd <- suppressMessages(BRL::compareRecords(file1, file2, c(2, 3, 5, 6),
+                                           types = c("lv", "lv", "bi", "bi")))
+cd[[1]] <- apply(cd[[1]], 2, as.numeric)
+R_vals <- c(1, 2, 5, 10, 20)
+R <- R_vals[taskID]
+
+Zs <- matrix(NA, nrow = n2, ncol = iterations)
+evals <- matrix(NA, nrow = 3, ncol = iterations)
+
+for(iter in 1:iterations){
[…]

### 2023-06-07 cluster_jobs/SEI_sensitivity.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --array=1-5
+#SBATCH --mail-type=end
+#SBATCH --mail-user=bak47@duke.edu
+#SBATCH --output=logs/%x_%a.out
+#SBATCH --error=logs/%x_%a.err
+#SBATCH --job-name=SEI
+#SBATCH --time=24:00:00
+#SBATCH --mem=128G
+
+#
+# modules
+#
+module load R
+Rscript code/SEI_sensitivity.R

### 2023-06-07 notes-steorts/third-revision-review-comments-suggestions.tex

\documentclass{article}
\usepackage[numbers]{natbib}
%\usepackage{hyperref}
\usepackage{lineno}
\usepackage{amsthm}
\usepackage{amsmath, color}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
%\usepackage{graphicx}
%\newtheorem{lemma}{Lemma}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bm}
\usepackage{natbib}

\newcommand{\bam}{\boldsymbol{m}}    
\newcommand{\bu}{\boldsymbol{u}}    
\newcommand{\bM}{\boldsymbol{M}}  
\newcommand{\bU}{\boldsymbol{U}}    
\newcommand{\bZ}{\boldsymbol{Z}}  
\newcommand{\bX}{\boldsymbol{X}}
\newcommand{\bDelta}{\boldsymbol{\Delta}}
\newcommand{\bGamma}{\boldsymbol{\Gamma}}
\newcommand{\bgamma}{\boldsymbol{\gamma}}

%\newcommand{\bX}{\boldsymbol X}
%\newcommand{\bZ}{ \boldsymbol Z}
\newcommand{\bh}{\boldsymbol h}
\newcommand{\bc}{\boldsymbol c}
\newcommand{\btheta}{\boldsymbol \theta}
\newcommand{\LL}{\mathcal{L}} 
\newcommand{\C}{\mathcal{C}} 
\newcommand{\R}{\mathbb{R}} 
\newcommand{\CC}{\mathfrak{C}} 
\newcommand{\PP}{\mathbb P} 
\newcommand{\pairP}{\mathcal{P}}

\begin{document}
\title{AE Comments}
\author{}
\date{}
\maketitle

\section{AE Comments}

\textcolor{red}{Find some proposed responses to the reviewer for some of the comments based upon the May 11 in person meeting.}

\begin{enumerate}
\item Line 111. Is this missing at random or something else? This needs to be clarified (AE Comment 1). This needs to be addressed with extreme care regarding the precise wording. \\
\textcolor{blue}{This distinction between truly missing at random and missing completely at random does not matter due to two assumptions. Following Sadinle (2017), we assume that the fields in the comparison vectors are conditionally independent given the coreference matrix and we assume that the missing comparisons are ignorable (Sadinle (2017). Therefore, our likelihood is the same as in Sadinle (2017) (see equations (3) and (4) of his paper). As stated by Sadinle (2017), ``we work under the assumption of ignorability of the missingness mechanism for the comparisons so that we can base our inferences on the marginal distribution of the observed comparisons (Little and Rubin, p. 90)."  We refer to Sadinle (2017), Section 4.2, page 604 for further details.}
\item Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and
	think it was a good choice to move the full derivation to an appendix. However, I do not follow
	the re-expression of the pmf for $\Gamma_{.j}$. First, the opening square brackets are still misplaced
	throughout. In the second line of this derivation, the authors divide the expression by a product
	of the element-wise conditional probabilities of a match in the comparison vector (u), raised by
	an indicator that the element of the comparison vector equals a particular value. Perhaps I
	misunderstand, but I believe this product is not constant in $\Gamma_{.j}$ or $u$, and so the total expression
	is not proportional to the line above, as interpreted as a pmf. I think that the end effect is that
	the last line of the newly expressed pmf is missing a factor of \\
%	$\prod_i \prod_f \prod_l u_{fl}^{I(\gamma_{ij}^f = l) I_{obs}$\\
	\textcolor{blue}{Is the new proof, Brian, missing an indicator over the cluster assignment? I believe that it should match those 
	of Sadinle (2017), Aleshin and Guendel (2022), and Wortman (2019). Where is the indicator regarding the cluster assignment that you condition on? Specifically, the update for each gamma will depend upon your cluster assignment, so I believe there is a typo for this reason. (To help with notation,  Jody calls this $S_k$, Sadinle calls it $Z_{-j}$, and Serge calls it $\C_{-j}$). Could you carefully check this please? I think this question is related to the next point/comment by the AE.} \\

	\textcolor{blue}{Brian: please re-derive these very carefully as I have found typos in the appendix after going back through them.}
    \item Appendix 8.2: Thank you for including the details on integrating out pi from the full
conditionals. They surprised me. I had assumed that the authors had integrated out pi in the
prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite
matchings”. Does the alternate approach presented in this paper provide a different algorithm
than directly integrating the prior distribution? Is the presented alternative approach justified? \\
\textcolor{blue}{
Thank you for noting your surprise regarding our proposal. We do not extend the approach of Sadinle (2017), but relax the one-to-one assumption for computational reasons. This motivates the choice of our prior (instead of the beta prior for bipartite matchings). 
\begin{enumerate}
\item Both approaches that were previously provided in the revision are justified, however, did contain typos. The first one is providing a standard update of the partition or coreference matrix. The second provides the update to the partition or coreference matrix, where $\pi$ is integrated out for a computational benefit. Furthermore, when integrating out $\pi$ from the full conditional of the partition, there are connections to the original derivation the full conditional of Sadinle (2017). Specifically, each update to the full conditional is similar in nature to the original derivation of Sadinle (2017), where the main difference is that we no longer enforce the bipartite matching constraint. [TODO: Make this connection more clear in the revision].
\item Yes, the presented approach is justified as it is simply a relaxation of Sadline (2017). We utilize a Gibbs update in the same spirit of Sadinle (2017) and the rest of the record linkage literature. We provide details regarding both Gibbs samplers below.
\end{enumerate}
}

\newpage
\textbf{Gibbs sampler}

We present a Gibbs sampler to explore the joint posterior distribution of $\bZ$, $\Phi$, and $\pi$ given the observed comparison data $\bgamma_{\text{obs}},$ for the likelihood and priors presented earlier. 
We start the sampler with an empty bipartite matching, such that, $Z_{j}^{[0]} = n_1 + j$ for all $j \in \{1, \ldots, n_2 \}.$ For a current value of the matching labeling $\bZ^{[t]},$ we obtain the next values of the Gibbs sampler iteration 
$\bam_f^{[t+1]} = (m_{f0}^{[t+1]}, \ldots, m_{f L_f}^{[t+1]})$, 
$\bu_f^{[t+1]} =  (u_{f0}^{[t+1]}, \ldots, u_{f L_f}^{[t+1]}),$ for $f=1, \ldots, F,$ 
$\pi^{[t+1]}$, and $\bZ^{(t+1)} = (Z_1^{(t+1)}, \ldots, Z_{n2}^{(t+1)})$ as follows:

\begin{enumerate}
\item For $f=1, \ldots, F$ sample
$$\bam_f^{[t+1]} \mid \bgamma_{\text{obs}}, \bZ^{[t]} \sim
\text{Dirichlet}(
a_{f0}(\bZ^{(t)})  + \alpha_{f0}, \ldots, a_{f L_f}(\bZ^{(t)} ) + \alpha_{f L_f})
),
$$
and
$$\bu_f^{[t+1]} \mid \bgamma_{\text{obs}}, \bZ^{[t]} \sim
\text{Dirichlet}(
b_{f0}(\bZ^{(t)})  + \beta_{f0}, \ldots, b_{f L_f}(\bZ^{(t)} ) + \beta_{f L_f})
). 
$$
Collect these updated parameters into $\Phi^{[t+1]},$ where the functions $a_{f \ell}(\cdot)$ and $b_{f \ell}(\cdot)$ were defined in equation \ref{eqn:ab}. 
\item Sample $$\pi^{[t+1]} \
\sim \text{Beta}(
n_{12}(\bZ^{(t)} )+ \alpha_{\pi}, n_2 - n_{12}(\bZ^{(t)}) + \beta_{\pi} - 1)
)
$$
\item Sample the entries of $\bZ^{[t+1]}$ sequentially. As we have sampled the first $j-1$ entries of $\bZ^{[t+1]},$ we define 
$\bZ_{-j}^{(t + (j-1)/n_2)} = (
Z_{1}^{[t+1]}, \ldots, Z_{j-1}^{[t+1]}, Z_{j+1}^{[t+1]}, \ldots, Z_{n_2}^{[t+1]}
).
$
We sample a new label $Z_{j}^{[t+1]}$ with probability of \textcolor{blue}{selecting label} $q \in \{1, \ldots, n_1, n_1 + j \}$ given by

\begin{align}
P(\bZ_{-j} \mid \Phi, \bgamma_{obs}) \propto 
\begin{cases}
\frac{\pi}{n_1}\exp(w_{qj}(\Phi) )  & \text{if} \; q \leq n_1 \\
1- \pi
%  \frac{n_1}\frac{n_2}{n_{12}(\bZ_{-j}) + \beta_{\pi}}{n_{12}(\bZ_{-j}) + \alpha_{\pi}}
   &  \text{if} \; q = n_1 + j.
\end{cases}
\end{align}
\end{enumerate}

Observe that the update for the partition does not depend upon any sequential updates, which makes this sampler extremely fast to update compared with the collapsed sampler below.

We present a collapsed Gibbs sampler to explore the joint posterior distribution of $\bZ$, $\Phi$ given the observed comparison data 
$\bgamma_{\text{obs}},$ for the likelihood and priors presented earlier, where we collapse (or integrate out) $\pi$ in order to illustrate similarities with the original proposal of Sadinle (2017).

\textbf{Collapsed Gibb sampler}
We start the sampler with an empty bipartite matching, such that, $Z_{j}^{[0]} = n_1 + j$ for all $j \in \{1, \ldots, n_2 \}.$ For a current value of the matching labeling $\bZ^{[t]},$ we obtain the next values of the Gibbs sampler iteration 
$\bam_f^{[t+1]} = (m_{f0}^{[t+1]}, \ldots, m_{f L_f}^{[t+1]})$, 
$\bu_f^{[t+1]} =  (u_{f0}^{[t+1]}, \ldots, u_{f L_f}^{[t+1]}),$ for $f=1, \ldots, F,$ and $\bZ^{(t+1)} = (Z_1^{(t+1)}, \ldots, Z_{n2}^{(t+1)})$ as follows:

\begin{enumerate}
\item For $f=1, \ldots, F$ sample
$$\bam_f^{[t+1]} \mid \bgamma_{\text{obs}}, \bZ^{[t]} \sim
\text{Dirichlet}(
a_{f0}(\bZ^{(t)})  + \alpha_{f0}, \ldots, a_{f L_f}(\bZ^{(t)} ) + \alpha_{f L_f})
),
$$
and
$$\bu_f^{[t+1]} \mid \bgamma_{\text{obs}}, \bZ^{[t]} \sim
\text{Dirichlet}(
b_{f0}(\bZ^{(t)})  + \beta_{f0}, \ldots, b_{f L_f}(\bZ^{(t)} ) + \beta_{f L_f})
). 
$$
Collect these updated parameters into $\Phi^{[t+1]},$ where the functions $a_{f \ell}(\cdot)$ and $b_{f \ell}(\cdot)$ were defined in equation \ref{eqn:ab}.

\item Sample the entries of $\bZ^{[t+1]}$ sequentially. As we have sampled the first $j-1$ entries of $\bZ^{[t+1]},$ we define 
$\bZ_{-j}^{(t + (j-1)/n_2)} = (
Z_{1}^{[t+1]}, \ldots, Z_{j-1}^{[t+1]}, Z_{j+1}^{[t+1]}, \ldots, Z_{n_2}^{[t+1]}
).
$
We sample a new label $Z_{j}^{[t+1]}$ with probability of \textcolor{blue}{selecting label} $q \in \{1, \ldots, n_1, n_1 + j \}$ given by

\begin{align}
P(\bZ_{-j} \mid \Phi, \bgamma_{obs}) \propto 
\begin{cases}
\exp(w_{qj}(\Phi) )  & \text{if} \; q \leq n_1 \\
n_1 \frac{n_2- n_{12}(\bZ_{-j}) + \beta_{\pi}}
{n_{12}(\bZ_{-j}) + \alpha_{\pi}
}
%  \frac{n_1}\frac{n_2}{n_{12}(\bZ_{-j}) + \beta_{\pi}}{n_{12}(\bZ_{-j}) + \alpha_{\pi}}
   &  \text{if} \; q = n_1 + j.
\end{cases}
\end{align}
\end{enumerate}

The Gibbs sampler above is more directly comparable with that of \cite{} (see the supplement), where in our collapsed sampler we have relaxed the one-to-one assumption (or rather the bipartite matching) regarding the update to update to the partition. For computational reasons, we correct for this in our post-processing stage. Observe that our weights $\exp(w_{qj}(\Phi))$ are the same as in the work of Sadinle (2017).

\end{enumerate}

\end{document}

### 2023-07-07 ba_resubmit/response_to_reviewers2.tex
updated draft

@@
 	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. We find this generally more readable, and particularly simplifies the notation in Section 4.2. 
-	\item At the beginning of Section 4.3, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics discussed in Section 4.1. This is the text added:
+	\item At the end of Section 4.1, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics. This is the text added:
 
-''To ground the discussion of posterior inference, we first re-express the likelihood function from (6a) in terms of the agreement patterns and summary statistics in Table 2.
-Let $m(p) =  p(\gamma_{ij} = h_p|Z_j = i)$ be probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match.
-Similarly define $u(p)$.
-These probabilities are simply products of relevant $\bm{m}$ and $\bm{u}$ parameters.
-Viewed through the perspective of patterns, the likelihood is
-\begin{align}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\bm{\Gamma}}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}. \label{likelihood_efficient}"
-\end{align}
-
-	This likelihood is then used throughout the discussion of efficient posterior inference.
-We hope this expression clarifies the meaning of "preservation of weight (Comment 5), and strengthens the argument for $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ being sufficient statistics for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ (Comment 4).
+	This likelihood is then used throughout Section 4.2 in the discussion of efficient posterior inference.
+We hope this expression clarifies the meaning of "preservation of weight (Comment 5), and strengthens the argument for $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ being summary statistics for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ (Comment 4).
 
@@
 
-	\item Section 4.1 introduces the sufficient statistic $N_{j_p}$.
+	\item In Section 4.1, we have provided a specific example of a comparison vector and how the hashing procedure works.
+
+	\item Section 4.1 introduces the summary statistic $N_{j_p}$.
 After careful review, we realized that by changing the order of subscripts and instead using $N_{p_j}$, we can replace the sum of counts across all records (previously denoted $H_p$), with the more intuitive $N_p$.
@@
 In fact, if the point is that “$x_{1i}$” will never be used again, why is the notation introduced in the first place?}
+
 \reply{We added this sentence in an attempt to address a previous reviewer comment pointing out a conflation between the record itself and the index directing towards that record.
@@
 	\pointRaised{R}{%
-	For the missing data treatment now included just before Section 2.1, is the assumption truly
-	missing at random, or missing completely at random? Or, does this distinction not matter
-	because of the independence assumed across elements of the comparison vector?}
+	For the missing data treatment now included just before Section 2.1, is the assumption truly missing at random, or missing completely at random?
+Or, does this distinction not matter because of the independence assumed across elements of the comparison vector?}
 
 	\reply{% 
-\textcolor{red}{I used MAR because that is what Sadinle wrote in his paper. Please review, and confirm whether its MCAR or not.}
-
 		The assumption is missing completely at random. You are correct that this is a result of the independence assumed across elements of the comparison vector. This has been clarified.
@@
 	\pointRaised{R}{%
-	Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and
-	think it was a good choice to move the full derivation to an appendix. However, I do not follow
-	the re-expression of the pmf for $\Gamma_{.j}$. First, the opening square brackets are still misplaced
-	throughout. In the second line of this derivation, the authors divide the expression by a product
-	of the element-wise conditional probabilities of a match in the comparison vector (u), raised by
-	an indicator that the element of the comparison vector equals a particular value. Perhaps I
-	misunderstand, but I believe this product is not constant in $\Gamma_{.j}$ or $u$, and so the total expression
-	is not proportional to the line above, as interpreted as a pmf. I think that the end effect is that
-	the last line of the newly expressed pmf is missing a factor of $\prod_i \prod_f \prod_l u_{fl}^{I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}$ regardless of the value of $z_j$.}
+	Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and think it was a good choice to move the full derivation to an appendix.
+However, I do not follow the re-expression of the pmf for $\Gamma_{.j}$.
+First, the opening square brackets are still misplaced
+	throughout.
+In the second line of this derivation, the authors divide the expression by a product of the element-wise conditional probabilities of a match in the comparison vector (u), raised by an indicator that the element of the comparison vector equals a particular value.
+Perhaps I misunderstand, but I believe this product is not constant in $\Gamma_{.j}$ or $u$, and so the total expression is not proportional to the line above, as interpreted as a pmf.
+I think that the end effect is that the last line of the newly expressed pmf is missing a factor of $\prod_i \prod_f \prod_l u_{fl}^{I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}$ regardless of the value of $z_j$.}
 
@@
 	The square brackets have been corrected, and we apologize for the oversight.
-
-	We have provided an alternative derivation of the full conditional, separately handling the cases for $Z_j \leq n_A$ and for $Z_j = n_A + j$. We provide it here for your reference:
-
-	We consider the likelihood in two cases. When $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
-	\begin{align}
-		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
-	\end{align}
-	When $Z_j = i$ for some $i \leq n_A$, we have
-	\begin{align}
-		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = i) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \times \prod_{i' != i}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{i'j}^f = l)I_{obs}(\gamma_{i'j}^f)}.
-	\end{align}
-	We multiply and divide by the $u$ parameters for the matching record pair to obtain
-	\begin{align}
-		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = i) &=\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \times \prod_{i' = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{i'j}^f = l)I_{obs}(\gamma_{i'j}^f)} \\
-		&= w_{ij} \times c_j.
-	\end{align}
-	Combining these results, we have 
-	\begin{align}
-		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = z_j) = \begin{cases} 
-		w_{z_j, j}  & z_j \leq n_A; \\
-		1 &  z_j  = n_A + j, \\
-	\end{cases}
-	\end{align}
+We have also prepared an alternative derivation of the likelihood for $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
+We believe this is more clear.
 }
 
-%	{\color{red}{I don't really understand his confusion.
-Why wouldn't the product of $u$ parameters be constant in $u$ and the data?
-Perhaps if I use $\mathcal{L}(Z_j| m, y, \Gamma_{.j})$, it might be clearer?}}}
-
-%	In the second line of the derivation, we divide by the $u$ parameters for all of the records in $A$ (note the product outside of the fraction).
-This quantity does not depend on $Z_j$ and is constant in $u$ and $\Gamma_{.j}$.
-This is easiest to see through the likelihood expressed at the level of the agreement patterns.
-We are also notating this likelihood as $\mathcal{L}(Z_j | m, u, \Gamma_{.j})$ to make it clear what is considered constant.
-%	
-%	\begin{align*}
-%	 \mathcal{L}(Z_j | m, u, \Gamma_{.j})  &= \prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}\\
-%	 &= \prod_{p=1}^P u(p) ^ {N_{p_j}} \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{- I(Z_j = i)}\\
-%	 &= \prod_{p=1}^P u(p) ^ {N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}\\
-%	\end{align*} 
-%where $w_p = \frac{m(p)}{u(p)}$. The term outside of the inner product does not depend on $Z_j$. Thus we have
-%\begin{align*}
-%	\mathcal{L}(Z_j | m, u, \Gamma_{.j})   \propto
-%	\begin{cases} 
-%		w_p,  & z_j \leq n_1; \\
-%		1, &  z_j  = n_1 + j; \\
-%	\end{cases} \label{eqn:z}\\
-%\end{align*}
-%
-%REVISIT
-
 	\pointRaised{R}{%
-Appendix 8.2: Thank you for including the details on integrating out pi from the full
-conditionals. They surprised me. I had assumed that the authors had integrated out pi in the
-prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite
-matchings”. Does the alternate approach presented in this paper provide a different algorithm
-than directly integrating the prior distribution? Is the presented alternative approach justified?}
+Appendix 8.2: Thank you for including the details on integrating out pi from the full conditionals.
+They surprised me.
+I had assumed that the authors had integrated out pi in the prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite matchings”.
+Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution?
+Is the presented alternative approach justified?}
 
@@
 	We reran all case studies and simulation using samples of $\pi$ in the full conditional for $Z$, and found no difference in results.
-This is unsuprsing, because with high numbers of observations, the posterior of the beta distrubution is highly concentrated around its mean.
+This is unsurprising, because with high numbers of observations, the posterior of the beta distribution is highly concentrated around its mean.
 
@@
 
-%	
-% {\color{red}{I really don't know how to answer this.
-I don't know how I would ''directly integrate" $\pi$, and I don't know if what I do is ''justified." Its kind of an intellectual relic to how I was originally thinking about the problem, and it reveals a quantity similar to what Sadinle was working with.
-% 		
-% It is theoretically more straightforward to just not integrate out $\pi$, and use samples from $\pi$ in Gibbs sampler like I do with all other parameters.
-Would that be too large a change at this stage?
-%
-% Without integrating out $\pi$, the full conditional is given on page 26 of the Appendix.}}
 }
@@
 \pointRaised{R}{%
-	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are
-	based on observed comparison vectors without consideration of Z (matches). In the
-	second paragraph of Section 4.3, the H notation includes the matches (Z), in notation
-	and definition that seems to conflict with the statement just below Equation (10). This
-	also comes into play in Equation (16)}
+	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are based on observed comparison vectors without consideration of Z (matches).
+In the second paragraph of Section 4.3, the H notation includes the matches (Z), in notation and definition that seems to conflict with the statement just below Equation (10).
+This also comes into play in Equation (16)}
 
 	\reply{%
-	You are correct that the the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$.
+	You are correct that the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$.
 This notation was created purely to be able to denote when a record was left unlinked during the Gibbs Sampler as in Equation 16 in the submitted draft.
@@
 \pointRaised{R}{%
-	Third paragraph of Section 4.1: The authors claim they are computing “sufficient
-	statistics”. What exactly are these statistics sufficient for?
+	Third paragraph of Section 4.1: The authors claim they are computing “sufficient statistics”. What exactly are these statistics sufficient for?
 	}
@@
 	\reply{%
+		We show in the revised Section 4.2 that the perspective of patterns allows us to write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ entirely throught the statistics in $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
+However, we cannot express the full likelihood in (14) through these statistics.
+Therefore, we have revised to text to say ``summary statistics" instead of the more technical ``sufficient statistics."}
 
-We show in the revised Section 4.3 that $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ is a set of statistics that are sufficient for $m$, $u$, and $Z$.
+\pointRaised{R}{%
+	First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute the appropriate weight to all records through the summary statistics…” What is meant by the term “weight”?
+Which records are appropriately weighted – those in X2?}
 
-{\color{red}{I drafted a whole bunch of material to argue more rigoursly that $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ is  set of sufficient statistics.
-If this argument seems weak, or is too much new material, I can also just keep what I had, and use less rigorous term "summary statistics" instead.}} \todo
+\reply{%
+
+	We have switched the order of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI.
+With this new ordering, and with the newlyu provided equations (15 -19) it is clearer that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
 }
@@
 \pointRaised{R}{%
-First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute
-the appropriate weight to all records through the summary statistics…” What is meant
-by the term “weight”? Which records are appropriately weighted – those in X2?}
-
-\reply{%
-	For each $j \in \{1, \ldots, n_B\}$, I use SEI to eliminate from memory the indices of  many records in $A$ that are unlikely matches for $B_j$ However, since I know the number of such records that exhibit each pattern $p$, I can still include their contribution to the likelihood through $N_{p_j}$.
-
-	We have changed that sentence in the text to read ''All original record pairs still contribute to the likelihood through $\mathcal{N}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}\}$."
-
-	This is clearer when looking at the conditional likelihoods in Equations 15 and 17 using $\tilde{\Gamma}$ in the revised Section 4.3. 
-}
-
-\pointRaised{R}{%
-Second paragraph of Section 4.2: “and delete those comparison vectors”. Which are
-“those” vectors?}
+	Second paragraph of Section 4.2: “and delete those comparison vectors”. Which are “those” vectors?}
 
@@
 \pointRaised{R}{%
-Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations
-(13) and (14)? I recommend that the authors either refrain from suppressing the SEI
-notation or further explain how the SEI algorithm has changed the quantities in these
-equations.}
+	Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations (13) and (14)?
+I recommend that the authors either refrain from suppressing the SEI notation or further explain how the SEI algorithm has changed the quantities in these equations.}
 
 \reply{%
-	The written expression in (13a) is correct.
-We then compress $r_{p_j}^{cd}$, and store in memory $r_{p_j}^{SEI, cd}$ Equation (14a) is conducted through $r_{p_j}^{SEI, cd}$.
-Equations (13b-c) and (14b-c) are exact, unchanged by SEI.
-We have clarified this.
+	We have reordered Section 4 so that all posterior inference is presented before SEI.
+This makes it clear that the posterior updates in (16a), (16b), and (18) depend only on $\mathcal{N}$.
+SEI only affects the step shown in (19).
+We have provided an additional simulation study to show robustness to the SEI procedure.
 }
@@
 \pointRaised{R}{%
-Section 4.2: I appreciate the practical advice about choosing S for the SEI method. However,
-this choice seems arbitrary in the absence of further discussion/evidence. Given that the
-primary novelty of the manuscript is in methods to speed and otherwise improve computation, I
-am surprised that this aspect of computational innovation is presented with virtually no
-theoretical or empirical exploration. Presumably the SEI method has some sort of accuracy
-trade-off, as the authors warn that linkage results may be “distorted” if S is low. However, this
-trade-off is not quantified or even discussed in practical terms beyond the terse
-recommendation to choose S=10.
+	Section 4.2: I appreciate the practical advice about choosing S for the SEI method.
+However, this choice seems arbitrary in the absence of further discussion/evidence.
+Given that the primary novelty of the manuscript is in methods to speed and otherwise improve computation, I am surprised that this aspect of computational innovation is presented with virtually no theoretical or empirical exploration.
+Presumably the SEI method has some sort of accuracy trade-off, as the authors warn that linkage results may be “distorted” if S is low.
+However, this trade-off is not quantified or even discussed in practical terms beyond the terse recommendation to choose S=10.
 }
@@
 \reply{% 
-
-{\color{red}{I'm not sure the best way to address this. Here is the where the situation arises:
-
-Imagine file $B$ has one record $B_1$ for Brian Kundinger, and file $A$ has 20 nearly matching records for Brian Kundinger.
-Perhaps Brian A Kundinger, Brian B Kundinger, Brian C Kundinger, $\ldots$, ect.
-All of these records share the same agreement pattern (say $p = 2$) with $B_1$.
-
-Without SEI, posterior probability would be divided evenly among all 20 near matches, such that the posterior probability would be at most $\frac{1}{20}$.
-If I chose $S=2$ however and I only retained 2 representatives in $r_{2_1}$, I could likely have one of those records having posterior probability above .5, and I would erroneously declare it a match.
-
-So issues only arise in very degenerate situations. How do you think I should address this? Should I do a simulation showing degeneracy when S=1 or 2?
-}
-}
+	We have provided an additional simulation study in Section 5.3 to show robustness to different choices of $S$.
 }
@@
 \pointRaised{R}{%
-Equation (6a): The indices do not match the subscripts in the indicator function in each
-summand, or their standard meaning in table 1.}
+	Equation (6a): The indices do not match the subscripts in the indicator function in each summand, or their standard meaning in table 1.}
 
 \reply{%
-	{\color{red}{I do not understand this objection.}}
+	This has been fixed.
 }
@@
 \pointRaised{R}{%
-Generally, the authors seem to arbitrarily use upper and lower case z interchangeably in
-function definitions.}
+	Generally, the authors seem to arbitrarily use upper and lower case z interchangeably in function definitions.}
 
 \reply{%
-	{\color{red}{When writing, I attempted to use $Z$ when I was discussing a random quantity, and $z$ to reference a realized value.
-I haven't found anything that I see wrong.
-I think I'm going to need help with this.}}
+	We use $Z$ when discussing a random quantity, and $z$ to reference a realized value. We have revised, and have found all instances of inconsistency. 
 }
@@
 \reply{%
-	When writing, I believed it was sufficient to include the superscripts on the $m$ and $u$ in the definition of $w$.
-However, you are correct, it is more clear to explicitly include it on the $w$.
-This has been updated.
+	This has been fixed. 
 }
@@
 \pointRaised{R}{%
-Gamma is in some places described as a set and in others as a matrix (particularly in
-sections 4.2 and 5.1).
+	Gamma is in some places described as a set and in others as a matrix (particularly in sections 4.2 and 5.1).
 }
@@
 \reply{%
-	$\Gamma$ is a matrix comprised by comparison vectors. It is certainly not a set in the mathematical sense. All references to $\Gamma$ as a set have been removed.
-
-	{\color{red}{I found one example of me saying that $\Gamma$ is ''comprised of the set of comparison vectors.' I'll triple check before resubmission.}}
+	$\gamma$ is a matrix comprised by comparison vectors. It is certainly not a set in the mathematical sense. All references to $\gamma$ as a set have been removed.
 }
@@
 \pointRaised{R}{% 
-Section 4.3, second paragraph, the definitions of the concatenated vectors $\alpha_0$ and $\beta_0$
-should have final elements subscripted by $L_f$, where the sub-subscript is capitalized.}
+	Section 4.3, second paragraph, the definitions of the concatenated vectors $\alpha_0$ and $\beta_0$ should have final elements subscripted by $L_f$, where the sub-subscript is capitalized.}
 

### 2023-07-07 ba_resubmit/kundinger_resubmit2.tex
updated draft

--- 2023-05-02 ba_resubmit/kundinger_resubmit2.tex
@@
 \usepackage{algpseudocode}
+\usepackage{subcaption}
 
@@
 
-In many data analysis tasks, analysts seek to identify duplicate records across two data files.
+Before conducting data analysis, it is often necessary to identify duplicate records across two data files.
 This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
@@
 
-Consider two data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$.
+Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
+Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
 Without loss of generality, denote files such that $n_A \geq n_B$.
@@
 
-Denote the set of matches by $\bm{M} = \{(i,j): \Delta_{ij} = 1\}$, and the set of non-matches by 
-$\bm{U} =  \{(i,j): \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
+{\color{red}{The following paragraph is pretty unnecessary, I never use this notation}}
+Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of non-matches by 
+$\bm{U} =  \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
 
@@
 
+{\color{red}{Beka suggested eliminating the following two paragraphs. Thoughts? I think the first one is good, the second seems unneccesary.}}
+
 Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ We collect the vectors $\bm{\gamma}_{ij}$ as the comparison matrix $\Gamma \in \mathbb{R}^{n_A n_B \times F}$.
-
-%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two data files.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ {\color{red}{We use $\Gamma_{ij}$ when referring to the vector as a random variable, and $\gamma_{ij}$ when referring to observed data.
+We collect the vectors $\gamma_{ij}$ as the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$}}.
+
+%Frequently, $\gamma_{ij}$ consists of one comparison for each feature shared between the two data files.
 
@@
 We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
 With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
-To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.}
+}
 }
@@
 The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
-Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
+Using these assumptions, \cite{winkler_state_1999}, \cite{jaro1989}, \cite{larsen_2001}, \cite{enamorado2019using} and others model the comparison data through mixture models of the form
 \begin{subequations}
@@
 
-%The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data.
+%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data.
 This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
@@
 To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework.
-In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
+In addition to {\color{red}{uniform}} Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
 He assigns a prior distribution for the probability of the indicator that a record in $B$ has a match in $A$, so that $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
@@
 / (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
-Thus, the beta distribution for bipartite matching, is given by
+Thus, the prior is given by
 \begin{align}
@@
 
-%hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $B$ that have matches in $\bm{X_1}$, and $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A + j)$ is the number of records in $B$ that have a match in $A$.
-%One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
-One way of resolving this issue to incorporate the constraint directly into the model as follows:
-%\begin{align}
-%	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
-%	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
-%	& \bm{Z} \sim \mathcal{B}, \notag
-%\end{align}
-%where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
-%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
-%$$
-%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A + j)$ is the records in $B$ that have a match in $A$.
-This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling $Z_j$.
-This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.
-
 \section{Fast Beta Linkage}
@@
 In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior.
-We denote the fast Beta prior as follows.
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta prior as follows.
 For each $Z_j$, we have
@@
 	\label{eqn:fast_beta_prior}
-	p(Z_j = z_j| \pi) &= \begin{cases} 
-	\frac{1}{n_A}\pi,  & z_j \leq n_A; \\
-	1-\pi, &  z_j  = n_A + j; \\
+	p(Z_j = q| \pi) &= \begin{cases} 
+	\frac{1}{n_A}\pi,  & q \leq n_A; \\
+	1-\pi, &  q  = n_A + j; \\
 \end{cases} \\
@@
 To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+However, there is a crucial difference: since \ref{eqn:fast_beta_prior} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
 To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
@@
 		$n_{AB}$ & number of entities in common between $A$ and $B$\\
-		$\bm{\gamma}_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
 		$Z_j = i$ & records $A_i$ and $B_j$ match \\
@@
 \begin{align}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \label{eqn:m} \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}), \label{eqn:u}\\
-	p(Z_j = z_j| \pi)  &=
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
+	p(Z_j = q| \pi)  &=
 	\begin{cases} 
-		\frac{1}{n_A}\pi,  & z_j \leq n_A; \\
-		1-\pi, &  z_j  = n_A + j; \\
+		\frac{1}{n_A}\pi,  & q \leq n_A; \\
+		1-\pi, &  q  = n_A + j; \\
 	\end{cases} \label{eqn:z}\\
@@
 \begin{align}
-	\bm{m}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
-	\bm{u}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update}
+	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
+	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
+	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
+	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}
+	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
 \end{align}
 \end{subequations}
-where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} \neq i)$.
-
-Next, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
+Next, the full conditional for $\pi$ is given by
+\begin{align}
+	\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi}).
+\end{align}
+
+Lastly, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
 \begin{align}
 	\label{eqn:z_full_conditional}
-p\left(Z_j^{(s+1)}  = z_j|\Gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
+	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
 	\begin{cases} 
-	w_{z_j, j}^{(s)},  & z_j \leq n_A; \\
-	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
+		\frac{\pi^{(s+1)}}{n_A} w_{q, j}^{(s+1)},  & q \leq n_A; \\
+		1 - \pi^{(s+1)}, & q  = n_A + j, \\
 	\end{cases}
 \end{align}
+%\begin{align}
+%\label{eqn:z_full_conditional}
+%p\left(Z_j^{(s+1)}  = z_j|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
+%\begin{cases} 
+%	w_{z_j, j}^{(s+1)},  & z_j \leq n_A; \\
+%	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
+%\end{cases}
+%\end{align}
 where, for all $i \in \{1, \ldots, n_A\}$, 
@@
 With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
-First, we propose hashing methods that allow us to compute sufficient statistics that reduce the computational complexity of the Gibbs sampler.
+First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler.
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
@@
 		\hline
-		$P$ & number of agreement patterns exhibited in $\Gamma$ \\
+		$P$ & number of agreement patterns exhibited in $\gamma$ \\
 		$h_p$ & agreement pattern $p$ \\
-		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $i \in B$ exhibits pattern $p$ \\
+		$e(h_p)$ & one-hot encoding of agreement pattern $p$ \\
+		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $B_j$ exhibits pattern $p$ \\
 		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
 		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
-		%$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
-		%$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
-		%$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
+		$N_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+		$N_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+
 		\hline
@@
 
+	{\color{red}{NOTE: This section used to be organized 
+		\begin{itemize}
+			\item Hashing
+			\item SEI
+			\item Inference
+		\end{itemize}
+	but now it is organized
+		\begin{itemize}
+		\item Hashing
+		\item Inference
+		\item Scaling (chunking method)
+		\item SEI
+	\end{itemize}
+
+I think inference is much more natural and after hashing.
+And having all the equations and updates written out in the inference part makes it clearer that SEI does not affect posterior updates at all.
+}}
+
 Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
-Let $P$ be the number of patterns realized in $\Gamma$ and note that it is bounded above by $P^{*} = F + \prod_{f=1}^F L_f$, where we have noted the possibility of missing values.
-Note that this quantity is controlled by the modeller in their choices for $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
+Let $P$ be the number of patterns realized in $\gamma$ and note that it is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the additional count notes the possibility of missing values.
+This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
 To obtain a memory efficient representation of the data, we map the agreement pattern of each record pair to a unique integer.
-\cite{enamorado2019using} accomplished this through the hashing function
+\cite{enamorado2019using} accomplished this through a hashing function, which we modify to explicitly handle missing values:
 \begin{align}
 	\label{eqn:hashing}
-	h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(f>1)\sum_{d=1}^{f-1}(L_d -1)}.
+	h^{*}(\gamma_{ij}) = \sum_{f = 1}^F I_{obs}(\gamma_{ij}^f)2^{\gamma_{ij}^f + I(f>1)\sum_{d=1}^{f-1}(L_d)}.
 \end{align}
@@
 
-With all record pairs converted to hashed values, we identify the records in $A$ with comparison vectors corresponding to each pattern $p$ for each record $B_j$.
+{\color{red}{In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
+This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
+A more complete example of the one-hot encoding is provided in Appendix \ref{app:ohe}.}}
+
+{\color{red}{
+For example, with five fields with binary agreements, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
+Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
+Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
+We map the unique values of $h^{*}(\gamma)$ to sequential integers (for in example in \texttt{R}, by converting the integers into factors, and then back to integers).
+If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
+}}
+
+We then identify the records in $A$ with comparison vectors corresponding to each pattern $p$ for each record $B_j$.
 We denote this set $r_{p_j} = \{i | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
 We then compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$, given by
-\begin{align}
-\label{eqn:N}
+\begin{align}\label{eqn:N}
 N_{p_j} = |r_{p_j}| = \sum_{i=1}^{n_A} I(\gamma_{ij} = h_p).
 \end{align}
-%Second, we compute the number of comparison vectors in $\Gamma$ that exhibit pattern $p$, given by
-%\begin{align}
-%	\label{eqn:H}
-%	N_p = \sum_{j = 1}^{n_B} N_{p_j}.
-%\end{align}
 We collect these counts in $\mathcal{N} = \{N_{p_j} |p \in 1, \ldots, P, j \in 1, \ldots, n_B \}$.
-As we show in Section \ref{efficient-posterior}, $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$  serves as a set of sufficient statistics and fully characterizes the comparison matrix $\Gamma$ with no loss of information.
-
-%With all record pairs converted to hashed values, we can create a more compact nested list 
-%$$\mathcal{R} = \left\{\{r_{p_j} \}_{p=1}^{P} \right\}_{j = 1}^{n_B},$$
-%where $r_{p_j} = \{A_i | \gamma_{ij} = h_p\}$ is a list of records in $A$ that share agreement pattern $p$ with record $B_j$.
-
-%We collect these sufficient statistics in the sets $\mathcal{N} = \left\{\{N_{p_j}\}_{j=1}^{n_B}\right\}_{p=1}^P$ and $\mathcal{H} = \{H_{p}\}_{p=1}^P.$
-
-\hypertarget{scaling}{%
-	\subsection{Scaling to Large Linkage Tasks}\label{scaling}}
-
-We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ for all $n_A \times n_B$ record pairs but greatly reduce the memory costs associated with unlikely matches.
-This allows all-to-all comparisons for substantially larger linkage tasks.
-All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{ij}$.
-Therefore, these records have the same probability to be identified as the link for record $B_j$.
-Thus, we know that records $i \in r_{p_j}$ such that $N_{p_j}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
-We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
-Therefore, rather than store all of these record labels, we store only a small number $S$ of them in a new nested list $\mathcal{R}^{\text{SEI}}$.
-Instead of storing $n_A \times n_B$ record labels, with SEI we store at most $n_B \times P \times S$ labels, regardless of how large $n_A$ might be.
-
-In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but is not likely to eliminate any of the indices for plausible matches.
-Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on few matches and distort linkage results.
-All original record pairs still contribute to the likelihood through $\mathcal{N}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}\}$.
-
-For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $A^{*}$ and $B^{*}$ smaller disjoint chunks $\{A^a | a = 1, \ldots, A^{*} \}$ and $\{B^b | b = 1, \ldots, B^{*} \}$ for more manageable computations.
-For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\Gamma^{ab}$.
-We then conduct hashing, obtain the compressed $\tilde{\Gamma}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data.
-In detail, we calculate
-\begin{subequations}
-	\begin{align}
-		r_{p_j}^{ab} &= \{i \in  | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
-		N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
-	\end{align}
-\end{subequations}
-These can computed serially or in parallel.
-We reduce memory costs through SEI on each $r_{p_j}$, while retaining counts of each agreement pattern through $N_{p_j}^{ab}$.
-Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
-Specifically, we combine information through
-\begin{subequations}
-	\begin{align}
-		r_{p_j}^{SEI} &= (r_{p_j}^{SEI, 11}, \ldots, r_{p_j}^{SEI, A^{*}B^{*}}) \text{ for } a = 1, \ldots, A^{*} \text{ and } b = 1, \ldots, B^{*},\label{r_{p_j}} \\
-		N_{p_j} &= \sum_{a = 1}^{A^{*}} \sum_{b = 1}^{B^{*}} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
-\end{subequations}
+
+{
+\color{red}{
+Together, the set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ with no loss of information.
+To see this, we employ the condensed notation
+\begin{align}
+	m_p =  p(\gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}
+\end{align}
+to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match.
+We similarly define $u_p$.
+Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) is equivalent to
+\begin{align}\label{likelihood_efficient}
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\gamma}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m_p^{I(Z_j = i)}u_p^{1 - I(Z_j = i)}. 
+\end{align}
+}}
 
@@
 
-%prod_{f = 1}^F \prod_{l = 1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = 1)}
-
-To ground the discussion of posterior inference, we first re-express the likelihood function from (\ref{eqn:likelihood}) in terms of the agreement patterns and summary statistics in \label{table_notation_B}.
-Let $m(p) =  p(\gamma_{ij} = h_p|Z_j = i)$ be probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match.
-Similarly define $u(p)$.
-These probabilities are simply products of relevant $\bm{m}$ and $\bm{u}$ parameters.
-Viewed through the perspective of patterns, the likelihood is
-\begin{align}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\bm{\Gamma}}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}. \label{likelihood_efficient}
-\end{align}
-{\color{red}{Should this first paragraph go at the end of Section 4.1? And should I be writing $r^{SEI}$ each time it comes up?}}
-
-In this form, we can see that the statistics collected in $\tilde{\Gamma}$ are sufficient for each of the model parameters in their respective conditional likelihoods and provide straightforward posterior updates.
+Using the likelihood in (\ref{likelihood_efficient}), we can see that the statistics collected in $\tilde{\gamma}$ allow for more efficient posterior inference for all model parameters.
 Conditioning on $\bm{Z}$, we can express the conditional likelihood for $\bm{m}$ and $\bm{u}$ as
 \begin{align}
-	\mathcal{L}(\bm{m}, \bm{u} \mid \bm{\Gamma}, \bm{Z})  &=  \prod_{p=1}^P m(p)^{N_p^m}u(p)^{N_p - N_p^m}  \label{likelihood_efficient_m_u}
+	\mathcal{L}(\bm{m}, \bm{u} \mid \gamma, \bm{Z})  &=  \prod_{p=1}^P m_p^{N_p^m}u_p^{N_p - N_p^m}  \label{likelihood_efficient_m_u}
 \end{align}
@@
 Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define $\bm{\beta_0}$ similarly for the $\bm{u}$ distributions.
-Let $e(h_p)$ be a one hot encoding of agreement pattern $p$, as described in Appendix \ref{app:ohe}.
-Then the terms needed for the posterior updates for level $l$ of field $f$ of the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
+{\color{red}{Using the one-hot encodings described in Section \ref{data-representation-hashing-and-storage},} the terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
 \begin{subequations}
@@
 \end{subequations}
-These quantities are identical those presented in (\ref{eqn:m_update}),and (\ref{eqn:u_update}), but are calculated through $P$ multiplications of scalar quantities over the vectors $h_p$, and then vectorized summations.
-This is markedly more efficient than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-Additionally, we can express the conditional likelihood for $Z_{j}$ as
-%\begin{align}
-%	\mathcal{L}(Z_j\mid  \bm{m}, \bm{u}, \tilde{\bm{\Gamma}}) = \prod_{p=1}^P N_{p_j} \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{- I(Z_j = i)} \label{likelihood_efficient_z}.
-%\end{align}
-\begin{align}
-	\mathcal{L}(Z_j\mid  \bm{m}, \bm{u}, \tilde{\bm{\Gamma}}) = \prod_{p=1}^P u(p)^{N_{p_j}} \prod_{i \in r_{p_j}} w(p)^{I(Z_j = i)} \propto \prod_{p=1}^P \prod_{i \in r_{p_j}} w(p)^{I(Z_j = i)} \label{likelihood_efficient_z}
-\end{align}
-{\color{red}{Should I include the proportionality step?}}
-where $w(p) = \frac{m(p)}{u(p)}$.
+Specifically, the $l + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}),and (\ref{eqn:beta_update}).
+In this new form however, they are calculated through $P$ multiplications of scalar quantities over the vectors $h_p$, and then vectorized summations.} This is markedly more efficient than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+We can also use (\ref{likelihood_efficient}) to express the conditional likelihood for $Z_{j}$. Using $w_p = \frac{m_p}{u_p}$, we have
+\begin{align}
+	\mathcal{L}(Z_j\mid  \bm{m}, \bm{u}, \tilde{\gamma}) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}\label{likelihood_efficient_z}
+\end{align}
 Note that $Z_j$ interacts with the data only through $\mathcal{N}$ and $\mathcal{R}$.
@@
 	\label{eqn:gibbs1}
-	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
 	\begin{cases} 
-		 N_{p_j} \times w_{p},  & \gamma_{z_j, j} = h_p; \\
+		N_{p_j} \times w_{p},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
 		n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   \text{otherwise}. \\
@@
 	\label{eqn:gibbs2}
-	P\left(Z_j^{(s+1)} = z_j \mid \gamma_{z_j^{(s+1)}, j} = h_p \right) = \begin{cases} 
-		\frac{1}{N_{p_j}}, & z_j \in r_{p_j}; \\
+	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p \right) = \begin{cases} 
+		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
 		0, & \text{otherwise.} \\
@@
 
-%We emphasize the computational gains of this split sampler: the first step is a sample from $P + 1$ options, where $P$ does not scale with the size of the linkage task; the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
 \begin{lemma}
@@
 \end{proof}
-%\begin{align}
-%	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\bm{\Gamma}}) &= \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
-%&=  \prod_{j=1}^{n_B}\prod_{i=1}^{n_A}\prod_{p=1}^P \left[ m_{fl}^{I(Z_j = i)}u_{fl}^{1 - I(Z_j = i)}\right]^{I(e(i,j) = h_p)}\\
-%&=  \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}\\
-%\end{align}
-
-%In this form, the likelihood reveals straightforward posterior updates for the $\bm{m}$ and $\bm{u}$ parameters.
-Let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define $\bm{\beta_0}$ similarly for the $\bm{u}$ distributions.
-Then, conditional on $\bm{Z}$, we have
-%\begin{align}
-%	\mathcal{L}(\bm{m}, \bm{u} \mid \bm{\Gamma}, \bm{Z})  &=  \prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{N_p^m}u(p)^{N_p - N_p^m}\\
-%\end{align}
-%where $N_p^m = \sum_{j=1}^{n_B} I\left(e_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ is total occurrence of pattern $p$ in the data across all record pairs.
-In this form, we see through factorization that $\mathcal{N}$, $\mathcal{P}$, and $\mathcal{R}$ are sufficient for $\bm{m}$ and $\bm{u}$.
-%
-%Viewing the likelihood at the level of patterns also allows us for efficient  We also see that the terms needed for the posterior updates needed for level $l$ of field $f$ are given by the appropriate component of the vectors
-%\begin{subequations}
-%	\begin{align}
-%		\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P N_p^m \times h_p, \label{efficient_alpha} \\
-%		\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P (N_p - N_p^m) \times h_p. \label{efficient_beta}
-%	\end{align}
-%\end{subequations}
-%}
-
-%Updating  $\alpha_{fl}(\bm{Z})$ and $\beta_{fl}(\bm{Z})$ for each field and level in the linkage task constitutes $2 \times \sum_f L_f$ many summations over $n_A \times n_B$ quantities.
-These are simple calculations, but they become computationally burdensome when working on large linkage tasks.
-
-%Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$.
-Let $H_p^m = \\ \sum_{j=1}^{n_B} I\left((Z_j, j) \in h_p, Z_j \leq n_A \right)$ be the number of matching record pairs with agreement pattern $p$.
-It follows that the number of non-matching record pairs with agreement pattern $p$ is $H_p^u = H_p - H_p^m$.
-Let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_f})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define  $\bm{\beta_0} = (\beta_{11}, \ldots, \beta_{F L_f})$ similarly for the $\bm{u}$ distributions.
-Then, the terms needed for the posterior updates in (\ref{eqn:m_update}) and (\ref{eqn:u_update}) are contained in the vectors
-
-%Although sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_A$ is large.
-This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-To speed up computation, we break this sampling step into two.
-First, we calculate the quantity expressed in (\ref{eqn:fs_weight}) associated with each unique pattern, and denote this quantity as $w_{p}$.
-We then sample the agreement pattern between $B_j$ and its potential link, according to
-%\begin{align}
-%	\label{eqn:gibbs1}
-%	P\left(\left(Z_j^{(s+1)}, j\right) \in h_p \mid \tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
-%	\begin{cases} 
-%		w_{p}\times N_{p_j},  & p \leq P; \\
-%		n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   \text{otherwise}. \\
-%	\end{cases}
-%\end{align}
-%
-%Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the $P$ agreement patterns.
-Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we can sample among candidate records uniformally using
-%\begin{align}
-%		\label{eqn:gibbs2}
-%		P\left(Z_j^{(s+1)} = z_j \mid e_{z_j^{(s+1)}, j} = h_p \right) = \begin{cases} 
-%			\frac{1}{N_{p_j}}, & e_{z_j, j} = h_p; \\
-%			0, & \text{otherwise.} \\
-%		\end{cases}
-%	\end{align} 
+
+\hypertarget{scaling}{%
+	\subsection{Scaling to Large Linkage Tasks}\label{scaling}}
+
+		For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks $\{A^a | a = 1, \ldots, t_A \}$ and $\{B^b | b = 1, \ldots, t_B\}$ for more manageable computations.
+{\color{red}{For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each chunk $A^a$ and $B^b$ would contain 500 records.}} For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
+We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$ for later calculations, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
+In detail, we calculate
+\begin{subequations}
+	\begin{align}
+		r_{p_j}^{ab} &= \{i | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+		N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
+	\end{align}
+\end{subequations}
+These can computed serially or in parallel.
+Summary statistics from each pairwise chunk comparison can be easily synthesized to recover summary statistics for the full comparison matrix $\gamma$.
+Specifically, we combine information and obtain the statistics comprising $\tilde{\gamma}$ through
+\begin{subequations}
+	\begin{align}
+		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
+		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
+\end{subequations}
+
+\hypertarget{SEI}{%
+	\subsection{Storage Efficient Indexing}\label{SEI}}
+
+As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses dramatically less memory than storing the full comparison matrix $\gamma$.
+However, the memory requirements of each are still quadratic in nature.
+For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
+We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ for all $n_A \times n_B$ record pairs, while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
+This allows all-to-all comparisons for substantially larger linkage tasks.
+
+{\color{red}{All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$.
+Therefore, these records have the same probability to be identified as the link for record $B_j$.
+Thus, we know that records $i \in r_{p_j}$ such that $N_{p_j}$ is large are unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
+We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
+Therefore, rather than store all of these record labels, we store only a small number $S$.
+For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $SEI(r_{p_j}^{ab})$.
+We collect these memory reduced lists to form $SEI(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $SEI(\mathcal{R})$.
+
+Let $n_{A^a}$ and $n_{B^b}$ be the number of records in chunks $A^a$ and $B^b$ respectively.
+Instead of storing $n_A \times n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a} \times n_{B^b} \times P \times S$ labels.
+As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $SEI(\tilde{\gamma}) = \{\mathcal{P}, SEI(\mathcal{R}), \mathcal{N}\}$.
+We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.
+
+NOTE: I switched the order of the chunking and the SEI because SEI only makes sense to use when you split the comparisons into chunks.
+Its also clearer mathematically.
+I think it also avoids the issue of the overloaded notation that used to be in equations \ref{r_{p_j}}.
+The reviewer asked to keep the SEI in the notation to reflect how SEI changed the likelihood and the quantities of interest, but I think the sentence "As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$" makes things very clear.
+Thoughts?}
+}
+
+%{\color{red}{For example, if file $A$ with $n_A = 500$ contains record $A_1$ that is an exact match for record $B_5$ and 499 records with complete disagreement on all fields with record $B_5$, we would only store $S$ indices to represent the records with full disagreement.I added that last example.
+I don't love it though.
+Thoughts?}}
+
+%{\color{red}{When the linkage task is large enough that all $n_A \times n_B$ comparison vectors cannot be stored in memory at the same time, we propose a modified approach where we construct comparison vectors and calculate $\mathcal{R}$ and $\mathcal{N}$ for smaller subsets of the data, and combine this information before conducting the Gibbs sampler.
+We use this approach in the case study in Section \ref{nltcs}, and provide details in Appendix \ref{app:chunks}}}.
+
 	\section{Simulation Studies}
@@
 We use $F = 5$ binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
-For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\Gamma \in \mathbb{R}^{n_A n_B \times F}$.
+For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$.
 For $n_B/2$ of these records, there is no match in $A$, so we simulate $n_A$ comparison vectors from the $\bm{u}$ distribution.
@@
 
+	\hypertarget{SEI-sensitivity}{%
+	\subsection{SEI Sensitivity}\label{SEI-sensitivity}}
+
+{
+\color{red}{Finally, our last simulation demonstrates our method's robustness to different values of $S$ for the SEI memory reduction procedure.
+We perform record linkage on one set of synthetic datafile described in Section \ref{accuracy} with 500 records in each datafile, 250 entities in common across datafiles, and 3 errors present across matching records.
+To achieve more drastic results, we perform SEI without chunking the data, with $t_A = t_B = 1$.
+In practice, if it is possible to create and store the comparison matrix for all record pairs at one time, there is no need to reduce the memory of the hashed matrix through SEI.
+For illustration however, it is easier to see degenerate behavior in this setting.
+
+We preform linkage using SEI with $S = (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler. 
+Note that any particular SEI memory reduction may lead to better or worse linkage performance; if the SEI procedure happens to only remove pairs that are not match by random chance, recall and precision will improve.
+Therefore, we perform linkage under each setting 100 times, recording the linkage estimate $\hat{\bm{Z}}$, and recall and precision.
+
+In Figure \ref{fig:SEI_Z},  see that the largest number of distinct linkage estimates occurs when $S = 1$.
+This makes sense, because the SEI procedure arbitrarily removes large numbers of record labels from consideration, resulting in a noisier estimate of the linkage structure.
+The number of distinct linkage estimates decreases as $S$ increases, with larger values of $S$ providing results more similar to the linkage without SEI.
+In Figure \ref{fig:SEI_eval}, we see similar patterns in precision.
+Setting $S=1$ can arbitrarily remove the index of a true match, leading the Gibbs sampler to concentrate probability on a false match, while larger values of $S$ produce results mirroring implementation with no SEI.
+We note however that even with $S=1$, the loss in precision is very small.
+
+Although the figures provided suggest that $S=2$ is adequate for maintaining linkage performance, we suggest a more conservative value like $S=10$.
+When evaluating the performance of a record linkage algorithm, researchers often examine posterior probabilities.
+By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce suspiciously high posterior probability for certain record pairs, providing a warped perception of model performance.
+We want to be sure that the set of records pairs exhibiting agreement patterns that have considerable probability of matching do not get reduced through SEI.}
+}
+
+%\begin{figure}[t]
+%	\begin{center}
+%		\includegraphics[width=0.6\textwidth]{../SEI_sensitivity/Z_plot}
+%		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
+%		\label{fig:SEI_Z}
+%	\end{center}
+%\end{figure}
+%
+%\begin{figure}[t]
+%	\begin{center}
+%		\includegraphics[width=0.6\textwidth]{../SEI_sensitivity/eval_plot} 
+%		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation}
+%		\label{fig:SEI_eval}
+%	\end{center}
+%\end{figure}
+%
+%  \begin{figure}[hbt]
+%	\centering
+%	\begin{subfigure}{0.4\textwidth}
+%		\includegraphics[width=\textwidth]{../SEI_sensitivity/Z_plot} 
+%		\caption{}\label{fig:SEI-Z}
+%	\end{subfigure}\hspace{\columnsep}%
+%	\begin{subfigure}{0.4\textwidth}
+%		\includegraphics[width=\textwidth]{../SEI_sensitivity/eval_plot} 
+%		\caption{}\label{fig:SEI-eval}
+%	\end{subfigure}
+%\caption{Simulation results Section \ref{SEI-sensitivity}. In \ref{fig:SEI-Z}, 
+%\end{figure}
+%Figure \ref{fig:here}
+
+\begin{figure}
+	\begin{minipage}[c]{0.4\linewidth}
+		\begin{center}
+		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
+		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
+		\label{fig:SEI_Z}
+		\end{center}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[c]{0.5\linewidth}
+		\begin{center}
+		\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
+		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation}
+		\label{fig:SEI_eval}
+	\end{center}
+	\end{minipage}%
+\end{figure}
 	\section{Case Studies}
@@
 	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in $4^5 \times 2 = 2048$ possible agreement patterns, with 1173 patterns realized in the data.
+This results in $5^5 \times 3 = 6025$ possible agreement patterns, with 1173 patterns realized in the data.
 However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
@@
 Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
-However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and compress information through storage efficient indexing.
-The resulting data object is just 10 MB, approximately 0.1\% of what is required for the raw comparisons.
+However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and combine results before conducting linkage.
+Without SEI, the resulting data object is about 2.2 GB, and with SEI using $S=10$, the resulting data object is about 760 MB.
 Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.
@@
 
-	%Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
-
-	%Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
-We explored ``one-to-many'' scenarios in which one record in $B$ has multiple plausible matches in $A$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
-Other issues arise under ``many-to-one'' scenarios, where one record in $B$ has multiple plausible matches in $A$, and ``many-to-many'' scenarios in which there is duplication both across and within data files.
-Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
-	%
-	%To our knowledge, \texttt{fabl} is the first method within the Fellegi-Sunter comparison vector approaches with an interpretable parameter for the rate of matching across data files.
-This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
-Similar work has been been proposed by \cite{wortman2019} and \cite{Larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
-We see these types of heirarchical record linkage models as potential contributions to the growing algorithmic fairness literature, and useful to practitioners working with or studying marginalized populations.
-
-	%Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
-	%future work.
-
-	%\begin{figure}[t]
-	%	
-	%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-		%		
-		%	}
-	%	
-	%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
-	%\end{figure}
-
-	%\subsection{National Long Term Care Survey}
-	%\label{nltcs} 
-	%
-	%The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
-It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989.
-The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office.
-State and regional office have been anonymized, presented as numbers instead of their actual values.
-We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context.
-The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466$ files, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns.
-We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.
-	%
-	%With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible.
-Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold.
-Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller data files.
-We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing.
-We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler.
-	%
-	%Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
-However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
-Posterior analysis yields an initial estimate of 9991 individuals linked across the the two data files, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
-	%
-	%Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
-With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks.
-This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler.
-However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error.
-In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office.
-We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives.
-Even if not, the modeler may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures.
-However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons.
-Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $\bm{m}$ and $\bm{u}$ parameters that are interpretale for the entire study, rather than separate parameters for each block.
-Although these $\bm{m}$ and $\bm{u}$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.
-
 	\clearpage
@@
 	$$\hat{Z}_j =\begin{cases} 
-		i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
+		i,  & \text{if } P(Z_j = i |\gamma) > \frac{1}{2}; \\
 		0,  & \text{otherwise}. \\
@@
 Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-In the event that we have two records $B_j$ and $B_{j'}$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
+In the event that we have two records $B_j$ and $B_{j'}$ such that both $P(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
 Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
@@
 %			$n_{AB}$ & number of entities in common between $A$ and $B$\\
-%			$\bm{\gamma}_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+%			$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
 %			$Z_j = i$ & records $A_i$ and $B_j$ match \\
@@
 \begin{align*}
-	p(m_{fl}| \Gamma, \bm{Z}) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
+	p(m_{fl}| \gamma, \bm{Z}) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
 	&= m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
 \end{align*}
-where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$.
-Analogous procedures lead to the posterior distribution $p(u_{fl}| \Gamma, \bm{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
+where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
+Analogous procedures lead to the posterior distribution $p(u_{fl}| \gamma, \bm{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
 Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
 
-$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
-$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})).$$
-
-In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihood relevant to a particular record $B_j$.
-Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(z_j < n_A + j)$ encoded by $\bm{Z}$ (and hyperparameters).
+$$\bm{m}_f|\bm{Z}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
+$$\bm{u}_f|\bm{Z}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})).$$
+
+Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ (and hyperparameters).
 Thus, we use $p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
@@
 \end{align*}
-Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $B_j$. We have
-\begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_A}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
-	&\propto \prod_{i=1}^{n_A}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
-	&\propto \prod_{i=1}^{n_A}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&=
+Thus $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})$ distribution.
+
+To express the full conditional for $\bm{Z}$, we consider the likelihood in two cases.
+Because we sample each $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
+Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+\end{align}
+When $Z_j = q$ for some $q\leq n_A$, we have
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)} \times \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+\end{align}
+We multiply and divide by the $u$ parameters for the matching record pair to obtain
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)} \times \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	&= w_{qj} \times c_j.
+\end{align}
+We can divide the result of each case by $c_j$ to get
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j) = \begin{cases} 
+		w_{qj}, & q \leq n_A; \\
+		1, &  q  = n_A + j. \
+	\end{cases}
+\end{align}
+Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain
+\begin{align}
+	\label{eqn:z_full_conditional2}
+	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
 	\begin{cases} 
-		w_{z_j, j}  & z_j \leq n_A; \\
-		1 &  z_j  = n_A + j, \\
-	\end{cases}\\
-\end{align*}
-where
-$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
-
-With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$.
+		\frac{\pi^{(s+1)}}{n_A} w_{q, j}^{(s+1)},  & q \leq n_A; \\
+		1 - \pi^{(s+1)}, & q  = n_A + j. \
+	\end{cases}
+\end{align}
+
+%Let $\gamma_{.j}$ denote the set of $n_A$ comparison vectors with $B_j$. We have
+%\begin{align*}
+%	p(\gamma_{.j}|Z_j = q, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_A}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+%	&\propto \prod_{i=1}^{n_A}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
+%	&\propto \prod_{i=1}^{n_A}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(q = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+%	&=
+%	\begin{cases} 
+%		w_{q, j}  & q \leq n_A; \\
+%		1 &  q  = n_A + j. \
+%	\end{cases}\\
+%\end{align*}
+%where
+%$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
+
+%With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$.
 Because we sample each $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$:
-\newpage
-\begin{align*}
-	p&\left(Z_j^{(s+1)} = z_j|\Gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
-	&\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
-	&\propto \left(\sum_{i=1}^{n_A}w_{z_j, j}I(z_j = i) + I(z_j = n_A + j) \right) \left (\pi\sum_{i=1}^{n_A}\frac{1}{n_A}I(z_j = i) + (1-\pi)I(z_j = n_A + j) \right) \\	
-	&= \frac{\pi}{n_A}\sum_{i=1}^{n_A}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_A + j). \\
-	&= \begin{cases} 
-		\frac{\pi}{n_A}w_{z_{j}, j}   & z_j \leq n_A; \\
-		1-\pi &  z_j  = n_A + j. \\
-	\end{cases}
-\end{align*}
-
-For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
-\begin{align*}
-	p&\left(Z_j^{(s+1)}  = z_j| \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
-	&= \int_{\pi}p\left(Z_j^{(s+1)} | \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-	&= \int_{\pi}\left[\frac{\pi}{n_A}w_{z_{j}, j} I(z_j\leq n_A) + (1-\pi)I(z_j = n_A + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_A}w_{z_{j}, j} I(z_j\leq n_A) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(z_j = n_A + j) \\
-	&= \frac{n_{AB}(\bm{Z}) + \alpha_{\pi}}{n_A (n_B + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_A) + \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_B + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_A + j) \\
-	&\propto w_{z_j, j}I(z_j\leq n_A) + n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_A + j) \\
-	&=
-	\begin{cases} 
-		w_{z_j, j}  & z_j \leq n_A; \\
-		n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_A + j. \\
-	\end{cases} \\
-\end{align*}
+%\newpage
+%\begin{align*}
+%	p&\left(Z_j^{(s+1)} = q|\gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
+%	&\propto p(\gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
+%	&\propto \left(\sum_{i=1}^{n_A}w_{q, j}I(q = i) + I(q = n_A + j) \right) \left (\pi\sum_{i=1}^{n_A}\frac{1}{n_A}I(q = i) + (1-\pi)I(q = n_A + j) \right) \\	
+%	&= \frac{\pi}{n_A}\sum_{i=1}^{n_A}w_{q, j}I(q = i) + (1-\pi)I(q = n_A + j). \\
+%	&= \begin{cases} 
+%		\frac{\pi}{n_A}w_{q, j}   & q \leq n_A; \\
+%		1-\pi &  q  = n_A + j. \\
+%	\end{cases}
+%\end{align*}
+%
+%For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
+%\begin{align*}
+%	p&\left(Z_j^{(s+1)}  = q| \gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
+%	&= \int_{\pi}p\left(Z_j^{(s+1)} | \gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+%	&= \int_{\pi}\left[\frac{\pi}{n_A}w_{z_{j}, j} I(q\leq n_A) + (1-\pi)I(q = n_A + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+%	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_A}w_{z_{j}, j} I(q\leq n_A) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(q = n_A + j) \\
+%	&= \frac{n_{AB}(\bm{Z}) + \alpha_{\pi}}{n_A (n_B + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(q\leq n_A) + \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_B + \alpha_{\pi} + \beta_{\pi}}I(q = n_A + j) \\
+%	&\propto w_{q, j}I(q\leq n_A) + n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}}I(q = n_A + j) \\
+%	&=
+%	\begin{cases} 
+%		w_{q, j}  & q \leq n_A; \\
+%		n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}} & q   = n_A + j. \\
+%	\end{cases} \\
+%\end{align*}
 
@@
 
+%	\hypertarget{app:chunks}{%
+%	\subsection{SEI for Large Linkage Tasks}\label{app:chunks}}
+
+%	For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks $\{A^a | a = 1, \ldots, t_A \}$ and $\{B^b | b = 1, \ldots, t_B\}$ for more manageable computations.
+{\color{red}{For example, if $n_A = n_B = 500$ and $t_A = t_B = 10$, each chunk $A^a$ and $B^b$ would contain 50 records.
+(Jerry asked for more precision here.
+Does this example work?) }}For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
+We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$ for later calculations, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
+In detail, we calculate
+%	\begin{subequations}
+%		\begin{align}
+%			r_{p_j}^{ab} &= \{i \in  | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+%			N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
+%		\end{align}
+%	\end{subequations}
+%	These can computed serially or in parallel.
+We reduce memory costs through SEI on each $r_{p_j}$, while retaining counts of each agreement pattern through $N_{p_j}^{ab}$.
+Summary statistics from each pairwise chunk comparison can be easily synthesized to recover summary statistics for the full comparison matrix $\gamma$.
+Specifically, we combine information through
+%	\begin{subequations}
+%		\begin{align}
+%			r_{p_j}^{SEI} &= (r_{p_j}^{SEI, 11}, \ldots, r_{p_j}^{SEI, A^{*}B^{*}}) \text{ for } a = 1, \ldots, A^{*} \text{ and } b = 1, \ldots, B^{*},\label{r_{p_j}} \\
+%			N_{p_j} &= \sum_{a = 1}^{A^{*}} \sum_{b = 1}^{B^{*}} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
+%	\end{subequations}
+%	
+%	{\color{red}{Note that with this approach, each $r_{p_j}^{SEI}$ can contain up to $S \times t_A \times t_B$ indices, so we store at most  $n_B \times P \times S \times t_A \times t_B$ indices in total.}}
+
+%		For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks $\{A^a | a = 1, \ldots, t_A \}$ and $\{B^b | b = 1, \ldots, t_B\}$ for more manageable computations.
+{\color{red}{For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each chunk $A^a$ and $B^b$ would contain 500 records.
+(Jerry asked for more precision here.
+Does this example work?) }}For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
+We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$ for later calculations, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
+In detail, we calculate
+%	\begin{subequations}
+%		\begin{align}
+%			r_{p_j}^{ab} &= \{i \in  | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+%			N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
+%		\end{align}
+%	\end{subequations}
+%	These can computed serially or in parallel.
+Summary statistics from each pairwise chunk comparison can be easily synthesized to recover summary statistics for the full comparison matrix $\gamma$.
+Specifically, we combine information through
+%	\begin{subequations}
+%		\begin{align}
+%			r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B,\label{r_{p_j}} \\
+%			N_{p_j} &= \sum_{a = 1}^{t-A} \sum_{b = 1}^{t_B} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
+%	\end{subequations}
+%	
+%	
 %	\hypertarget{app:fabl-psuedocode}{%
@@
 %			\EndFor
-%			\State Synthesize results across pairings to get $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$.
+%			\State Synthesize results across pairings to get $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$.
 %			\EndProcedure
@@
 %			\For{$t \in \{1, \ldots, T\}$} 
-%			\State Sample $\bm{m}^{t+1}|Z^{t}, \tilde{\Gamma}$ and  $\bm{u}^{t+1}|\bm{Z}^{t}, \tilde{\Gamma}$
-%			\State Sample $h\left(\bm{Z}^{t+1}\right)|\bm{m}^{t+1}, \bm{u}^{t+1}, \tilde{\Gamma}$.  \Comment{Sample agreement pattern, not record}
+%			\State Sample $\bm{m}^{t+1}|Z^{t}, \tilde{\gamma}$ and  $\bm{u}^{t+1}|\bm{Z}^{t}, \tilde{\gamma}$
+%			\State Sample $h\left(\bm{Z}^{t+1}\right)|\bm{m}^{t+1}, \bm{u}^{t+1}, \tilde{\gamma}$.  \Comment{Sample agreement pattern, not record}
 %			\EndFor
-%			\State Sample $\bm{Z} | h(\bm{Z}), \tilde{\Gamma}$. \Comment{Fills in record label based on agreement pattern}
+%			\State Sample $\bm{Z} | h(\bm{Z}), \tilde{\gamma}$. \Comment{Fills in record label based on agreement pattern}
 %			

### 2023-07-07 notes-steorts/kundinger-2023-june-reworking.tex
updated draft

\documentclass[ba]{imsart}
%
\pubyear{0000}
\volume{00}
\issue{0}
\doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

\usepackage{amsthm, amssymb}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}
\newtheorem{lemma}{Lemma}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}

\newcommand{\bem}{\boldsymbol{m}}    
\newcommand{\bu}{\boldsymbol{u}}    
\newcommand{\bM}{\boldsymbol{M}}  
\newcommand{\bU}{\boldsymbol{U}}    
\newcommand{\bZ}{\boldsymbol{Z}}  
\newcommand{\bX}{\boldsymbol{X}}
\newcommand{\bDelta}{\boldsymbol{\Delta}}
\newcommand{\bGamma}{\boldsymbol{\Gamma}}
\newcommand{\bgamma}{\boldsymbol{\gamma}}

\startlocaldefs
%\input{subfiles/preamble.tex}
\endlocaldefs

\begin{document}

%% *** Frontmatter ***

\begin{frontmatter}
\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl) \\ DO NOT SHARE}
%\title{A Sample Document\thanksref{T1}}
%\thankstext{T1}{<thanks text>}
\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}

\begin{aug}
\author{\fnms{Anonymous} \snm{Authors}\thanksref{addr1}\ead[label=e1]{}
%\author{\fnms{} \snm{}\thanksref{addr1}\ead[label=e2]{}}
%\and
%\author{\fnms{} \snm{}\thanksref{addr2}\ead[label=e3]{}
}

%\runauthor{Kundinger et al.}

\address[addr1]{Anonymous Address
  \printead{e1}, % print email address of "e1"
%  \printead*{e2}
}

%\address[addr2]{Departments of Statistical Science and Computer Science,
%  Duke University,
%  P.O.\ Box 90251,
%  Durham, NC 27708, USA
%  \printead{e3}
%}

%\thankstext{<id>}{<text>}
\end{aug}

%\begin{aug}
%\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
%\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
%\and
%\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
%
%\runauthor{Kundinger et al.}
%
%\address[addr1]{Department of Statistical Science,
%  Duke University,
%  P.O.\ Box 90251,
%  Durham, NC 27708, USA
%  \printead{e1}, % print email address of "e1"
%  \printead*{e2}
%}
%
%\address[addr2]{Departments of Statistical Science and Computer Science,
%  Duke University,
%  P.O.\ Box 90251,
%  Durham, NC 27708, USA
%  \printead{e3}
%}
%
%%\thankstext{<id>}{<text>}
%\end{aug}

\begin{abstract}

\end{abstract}

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{bipartite record linkage}
\kwd{Bayesian methods}
\kwd{hashing techniques}
\kwd{parallel/distributed computing}
\kwd{Markov chain Monte Carlo}
\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{sec:intro}

\section{The Bipartite Record Linkage Problem}
\label{sec:problem}
Assume two databases (data sources or data files) $\bX_1$ and $\bX_2$ contain $n_1$ and $n_2$ records, respectively. Without loss of generality, assume $n_1 \geq n_2.$ In addition, assume duplicate records occur across databases but not within a single database, which is known as \textit{bipartite record linkage}. The goal of this problem is to ascertain which records refer to the same entity across the two databases.

\subsection{The Bipartite Record Linkage Matching}
\label{sec:bipartite}

This problem can be formulated by a \textit{bipartite matching} between two sets of records from the two databases. Within the bipartite record linkage framework, let us consider records from $\bX_1$ and $\bX_2$ as disjoint set of nodes (containing all the feature or attribute information). An edge between two records represents them being a \textit{match} or \textit{coreferent}.

One way of representing a bipartite matching is through a \textit{coreference matrix} (or linkage structure) \citep{Fortinietal01, Larsen02, Larsen05,  liseo_2011, sadinle_detecting_2014, steorts_bayesian_2016}. The coreference matrix $\bDelta$ is a $n_1 \times n_2$ matrix whose $(i,j)$th entry is
%%
\begin{align}
\Delta_{ij} &=
\begin{cases}
1  & \text{if records} \; i \in \bX_1 \; \text{and}\; j \in \bX_2 \; \text{refer to the same entity;} \\
0 &  \text{otherwise.}
\end{cases}
\end{align}
While this construction has been used in the literature, it is not computationally convenient. The \textit{matching label} 
$\bZ = (Z_1, \ldots, Z_{n_2})$ for the records in database $n_2$ was introduced by \cite{sadinle_bayesian_2017} such that

\begin{align}
Z_j &=
\begin{cases}
i  & \text{if records} \; i \in \bX_1 \; \text{and}\; j \in \bX_2 \; \text{refer to the same entity;} \\
n_1 + j &  \text{if record} \; j \in \bX_2 \; \text{does not have a match in database} \; \bX_1. 
\end{cases}
\end{align}
The coreference matrix and the matching label are interchangeable because $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.

\subsection{Approaches to Bipartite Record Linkage}
\label{sec:approaches}

There are two main approaches to Bayesian bipartite record linkage --- comparison and direct based frameworks. Direct-based approaches generatively model the field (feature or attribute) information in the databases \citep{gutman_bayesian_2013, steorts_smered_2014, steorts_entity_2015, zanella_flexible_2016, steorts_performance_2017, kaplan_posterior_2018}. On the other hand, comparison-based approaches model the field comparisons between record pairs \citep{sadinle_detecting_2014, sadinle_bayesian_2017, mcveigh2019scaling}.

In this paper, we propose a scalable extension of \cite{sadinle_bayesian_2017}, which relaxes the bipartite extension implicitly in the prior and ``corrects" for this in a post-processing step of \cite{jaro_1989}. As such, we provide the first independent updates to the matching label $\bZ,$ contrasting prior work of sequential updates. Furthermore, we propose additional computational speeds up (see section \ref{}).

\section{The Fellegi-Sunter Approach}
\label{sec:fellegi}

Based upon the original paper of \cite{fellegi_theory_1969}, we can think of the set of ordered record pairs $\bX_1 \times \bX_2$ as the union of the set of matches and non-matches
\begin{align}
&\bM = \{
(i,j); \;  i \in \bX_1, \; j \in \bX_2,\;  \Delta_{ij} = 1
\};\\
&\bU = \{
(i,j); \;  i \in \bX_1, \; j \in \bX_2,\;  \Delta_{ij} = 0
\}.
\end{align}
Record pairs that are estimated as matches are called \textit{links}; pairs estimated as non-matches are called \textit{non-links}. The FS approach uses pairwise comparisons of the records to estimate their match status.

\subsection{Comparison Data}
\label{sec:comparison}

In record linkage applications, records that refer to the same entity should be similar. Records that refer to different entities should be dissimilar. This motivates a \textit{comparison vector} 
\begin{align}
\bgamma_{ij} &= (\gamma_{ij}^1, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F),
\end{align}
where $F$ denotes the number of criteria used to compare the records. Often the $F$ criteria correspond to one comparison per field that the databases have in common. For instance, binary agreement is the commonly used due to its simplicity; others have used partial agreement patterns using similarity metrics or distance functions \citep{sadinle_bayesian_2017}.

%\subsection{Fellegi-Sunter Model}
%\textcolor{red}{talk about FS Model here.}

%\subsection{Fellegi-Sunter Decision Rule}
%\textcolor{red}{remove.}
%The comparison vector $\bgamma_{ij}$ is not enough to determine the match status of record pairs due to typographical, transcription, or other errors in data. \cite{fellegi_theory_1969} proposed a likelihood ratio based method, where they computed
%\begin{align}
%\label{eqn:fellegi}
%w_{ij} &= \log \bigg[
%\mathbb{P}
%\frac{\bgamma_{ij} \mid \Delta_{ij} = 1}
%{\bgamma_{ij} \mid \Delta_{ij} = 0}
%\bigg]
%\end{align}
%as weights to estimated matched record pairs. Equation \ref{eqn:fellegi} assumes the realization  $\bgamma_{ij}$ of the random variable $\bGamma_{ij}$ depends on the coreference matrix (or match status) $\Delta_{ij}$ of record pair $(i,j).$ When this ratio is large, equation~\ref{eqn:fellegi} favors a record pair being a match; otherwise it favors a nonmatch. The authors formalized their proposal (the Fellegi and Sunter (FS) decision rule) and showed that it was optimal \cite{fellegi_theory_1969}. Note that the original FS decision rule does not enforce the bipartite record linkage assumption (or rather the one-to-one assignment restriction).

\subsection{Fellegi-Sunter Model}
\label{sec:fellegi-sunter-model}
In this section, we review the Fellegi and Sunter (FS) model of \cite{fellegi_theory_1969}.
%Unfortunately, the comparison vector $\bgamma_{ij}$ is insufficient to estimate links and nonlinks because observed data is likely to have noise and or missing observations. 
Note that $\bgamma_{ij}$ is a realization of a random variable $\bGamma_{ij}$ whose distribution depends on the coreference matrix $\Delta_{ij}$ of a record pair $(i,j).$ Let $\mathcal{M}$ and $\mathcal{U}$ denote a product of individual models for each of the comparison components under a conditional independence assumption \citep{Winkler1988, jaro_1989} or a complex log-linear model \citep{larsen_2001}. 
Let the comparison vector $\bgamma_{ij}$ be a realization of the random vector $\bGamma_{ij}$ that can take distribution 
$\mathcal{M}(\bem)$ or $\mathcal{U}(\bu),$ depending on whether the observed record pair is a match or non-match with 
$\bem$ and $\bu$ representing vectors of the parameters. Let $p$ denote the proportion of matches. 
Following \cite{Winkler1988, jaro_1989, larsen_2001} consider

\begin{align}
\label{eqn:model-estimation}
\bGamma_{ij} \mid \Delta_{ij} = 1 &\stackrel{iid}{\sim} \mathcal{M}(\bem), \notag \\
\bGamma_{ij} \mid \Delta_{ij} = 0 &\stackrel{iid}{\sim} \mathcal{U}(\bu), \quad \Delta_{ij} \stackrel{iid}{\sim} \text{Bernoulli}(p),
\end{align}
where $p$ denotes the proportion of matches.

The mixture model of equation \ref{eqn:model-estimation} relies on two assumptions: the comparison vectors are independent given the bipartite matching and the matching status of the record pairs are independent of another. Both assumptions imply the comparison vectors are marginally independent. Estimation has been typically accomplished via the expectation-maximization (EM) algorithm \citep{dempster1977maximum}. 
%It is important to note that the FS method does not satisfy the one-to-one assignment restriction of bipartite record linkage without some post-processing.

\subsection{Beta Record Linkage Model}
\label{sec:brl}
As such, this motivated the mixture model of \cite{sadinle_bayesian_2017}, who considered the matching status of the record pairs via a bipartite matching
%%
\begin{align}
\bGamma_{ij} \mid Z_i = i &\stackrel{iid}{\sim} \mathcal{M}(\bem), \notag \\
\bGamma_{ij} \mid Z_i \neq i &\stackrel{iid}{\sim} \mathcal{U}(\bu), \quad \bZ \sim \mathcal{B},
\end{align}
where $\mathcal{M}(\bem)$ and $\mathcal{U}(\bu)$ are models for the comparison vectors for matches and non-matches and $\mathcal{B}$ is a prior of the space of bipartite matches such as the beta prior of \cite{sadinle_bayesian_2017}. For complete details, we refer to the aforementioned paper, which inherently provides the one-to-one constraint of the bipartite matching problem. While theoretically desirable, unfortunately, the Gibbs sampling updates to the matching label $\bZ$ are sequential.

\section{The Fast Beta Linkage (fabl) Model}
\label{sec:fabl}
In this section, we propose the fast beta linkage method (fabl), which relaxes the bipartite assumption of \cite{sadinle_bayesian_2017} to induce a Gibbs sampler that is independent in its updates for the matching label $\bZ.$ This allows for much scalable computation.

\subsection{Likelihood for Comparison Data}
\label{sec:like}

Following \cite{sadinle_bayesian_2017}, the comparison fields are conditionally independent given the matching status of the record pairs, which provides the likelihood of the comparison data 
\begin{align}
\label{eqn:likelihood}
\mathcal{L}(\bZ, \Phi \mid \bgamma) &= \prod_{i=1}^{n_1} 
\prod_{j=1}^{n_2}
\prod_{f=1}^{F}
\prod_{\ell=1}^{L_f}
\bigg[
m_{f\ell}^{
I(Z_j = i)
}
u_{f\ell}^{
I(Z_j \neq i)
}
\bigg]^{
I(\gamma^f_{ij} = \ell)
},
\end{align}
where
$
m_{f\ell}
= \mathbb{P}(\Gamma_{ij}^f = \ell \mid Z_j = i )
$
denotes the probability of a match having a level $\ell$ of disagreement in field $f$ and 
$
u_{f\ell}
= \mathbb{P}(\Gamma_{ij}^f = \ell \mid Z_j \neq i )
$
denotes the probability of a non-match having a level $\ell$ of disagreement in field $f.$
%%
Denote $\bem_f = (m_{f1}, \ldots, m_{fL_f}),$  $\bu_f = (u_{f1}, \ldots, u_{fL_f}),$ $\bem = (\bem_1, \ldots, \bem_F),$
$\bu = (\bu_1, \ldots, \bu_F),$ and $\Phi = (\bem, \bu).$

Following \cite{sadinle_bayesian_2017}, we modify the model to take into account data that is missing completely at random. That is, using equation \ref{eqn:likelihood} and the assumption of ignorability, we marginalize over the missing comparisons to find
%%%
\begin{align}
\label{eqn:sadinle-likelihood}
\mathcal{L}(\bZ, \Phi \mid \bgamma^{\text{obs}}) 
&= 
\prod_{f=1}^{F}
\prod_{\ell=1}^{L_f}
\bigg[
m_{f\ell}^{
\sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j = i)
}
\times
u_{f\ell}^{
\sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j \neq i)
}
\bigg]\\
& = 
\prod_{f=1}^{F}
\prod_{\ell=1}^{L_f}
\bigg[
m_{f\ell}^{a_{f\ell}(\bZ)}
u_{f\ell}^{b_{f\ell}(\bZ)}
\bigg],
\end{align}
%%%
where 
\begin{align}
\label{eqn:ab}
a_{f\ell}(\bZ) &= \sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j = i) \quad \text{and} \\
b_{f\ell}(\bZ) &= \sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j \neq i), \notag
\end{align}

which for a given matching label $\bZ$ represents the number of matches and nonmatches with observed disagreement level $\ell$ in comparison $f$ and $I_{\text{obs}}(\cdot)$ is the indicator of whether its argument is observed.

%Note how $a_{f\ell}(\bZ)$ and $b_{f\ell}(\bZ)$ see the combination of the conditional independence and ignorabililty assumptions that allow us to ignore comparisons that are not observed using a simple, tractable solution.

Under equation \ref{eqn:sadinle-likelihood}, it is convenient to use independent, conditionally conjugate priors
\begin{align}
\label{eqn:fabl-priors}
&\bem_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{fL_f}) \\
&\bu_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{fL_f}),
\end{align}
for $f=1, \ldots, F.$

\subsection{Fast Beta Linkage Prior}
\label{sec:fabl-prior}

While the bipartite prior of \cite{sadinle_bayesian_2017} may be theoretically desirable, we propose a \textit{non-bipartite prior} for purely computational reasons. That is, for each $Z_j,$ $j \in \{1, \ldots, n_1, n_1 + j \}$ assume independent draws from

\begin{align}
\mathbb{P}(Z_j = i \mid \pi) 
&=
\begin{cases}
 \dfrac{\pi}{n_1}  &  \text{if}  \; z_j \leq n_1\\
  1- \pi & z_j = n_1 + j.
\end{cases}
\end{align}

Also, assume

\begin{align}
  \pi &\stackrel{ind}{\sim} \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
\end{align}
where $\alpha_{\pi}$ and $\beta_{\pi}$ are fixed and known. As already mentioned, the independence assumption above leads to a Gibbs sampler that is independent and does not require sequential updates of the matching label $\bZ.$

%\subsection{Appendix A of fabl -- describe here regarding the loss function}
%We represent a Bayes estimate as $\hat{\bZ} = (\hat{Z}_1, \ldots, \hat{Z}_{n2}),$ where $\hat{Z}_j \in \{1, \ldots, n_1, n_1 + j, R\},$ where 
%$R$ represents a rejection option. We use the loss of \cite{} 
%$L(\bZ, \hat{\bZ}) = \sum_{j=1}^{n_2} L(Z_j, \hat{Z}_j),$ where
%\begin{align}
%L(Z_j, \hat{Z}_j) 
%&=
%\begin{cases}
%0, \quad  &  \text{if} \quad Z_j = \hat{Z}_j; \\
%\lambda_{R}, \quad  &  \text{if} \quad \hat{Z}_j = R; \\
%\lambda_{\text{FNM}}, \quad  &  \text{if} \quad Z_j \leq n_1, \hat{Z}_j = n_1 + j; \\
%\lambda_{\text{FM1}}, \quad  &  \text{if} \quad Z_j = n_1, \hat{Z}_j \leq n_1; \\
%\lambda_{\text{FM2}}, \quad  &  \text{if} \quad Z_j, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z}_j; \\
%\end{cases}
%\end{align}
%where $\lambda_{R}$ is the loss from not taking a decision, $\lambda_{\text{FNM}}$ is the loss from a false non-match decision (FNM), $\lambda_{\text{FM1}}$
%is the loss from a false match decision when the record does not actually match any other record, and $\lambda_{\text{FM2}}$ is the loss from a false match decision when the record truly matches a different record than the one assigned to it. The posterior expected loss is given by 
%$E[L(\bZ, \hat{Z} ) \mid \bgamma_{obs}] = \sum_{j=1}^{n_2} \epsilon_j (\hat{Z}_j), $ where 
%$\epsilon_j (\hat{Z}_j) = E[L(Z_j, \hat{Z}_j ) \mid \bgamma_{obs}].$ 
%
%\textcolor{red}{todo, formalize this idea and proof more cleaning and clearly.}

\subsection{Gibbs Sampler}
\label{sec:gibbs}
In this section, we provide the Gibbs sampler of fabl for completeness that explores the joint posterior distribution of $\bZ$, $\Phi$, and $\pi$ given the observed comparison data $\bgamma_{\text{obs}},$ for the likelihood and priors presented earlier.

We start the sampler with an empty matching, such that, $Z_{j}^{[0]} = n_1 + j$ for all $j \in \{1, \ldots, n_2 \}.$ For a current value of the matching labeling $\bZ^{[t]},$ we obtain the next values of the Gibbs sampler iteration 
$\bem_f^{[t+1]} = (m_{f0}^{[t+1]}, \ldots, m_{f L_f}^{[t+1]})$, 
$\bu_f^{[t+1]} =  (u_{f0}^{[t+1]}, \ldots, u_{f L_f}^{[t+1]}),$ for $f=1, \ldots, F,$ 
$\pi^{[t+1]}$, and $\bZ^{(t+1)} = (Z_1^{(t+1)}, \ldots, Z_{n2}^{(t+1)})$ as follows:

\begin{enumerate}
\item For $f=1, \ldots, F$ sample
$$\bem_f^{[t+1]} \mid \bgamma_{\text{obs}}, \bZ^{[t]} \sim
\text{Dirichlet}(
a_{f0}(\bZ^{(t)})  + \alpha_{f0}, \ldots, a_{f L_f}(\bZ^{(t)} ) + \alpha_{f L_f})
),
$$
and
$$\bu_f^{[t+1]} \mid \bgamma_{\text{obs}}, \bZ^{[t]} \sim
\text{Dirichlet}(
b_{f0}(\bZ^{(t)})  + \beta_{f0}, \ldots, b_{f L_f}(\bZ^{(t)} ) + \beta_{f L_f})
). 
$$
Collect these updated parameters into $\Phi^{[t+1]},$ where the functions $a_{f \ell}(\cdot)$ and $b_{f \ell}(\cdot)$ were defined in equation \ref{eqn:ab}. 
\item Sample $$\pi^{[t+1]} \mid  \bgamma_{\text{obs}}, \bZ^{(t)}
\sim \text{Beta}(
n_{12}(\bZ^{(t)} )+ \alpha_{\pi}, n_2 - n_{12}(\bZ^{(t)}) + \beta_{\pi} - 1)
)
$$
\item Sample the entries of $\bZ^{[t+1]}$ as follows:

%As we have sampled the first $j-1$ entries of $\bZ^{[t+1]},$ we define 
%$\bZ_{-j}^{(t + (j-1)/n_2)} = (
%Z_{1}^{[t+1]}, \ldots, Z_{j-1}^{[t+1]}, Z_{j+1}^{[t+1]}, \ldots, Z_{n_2}^{[t+1]}
%).
%$
%We sample a new label $Z_{j}^{[t+1]}$ with probability of \textcolor{blue}{selecting label} $q \in \{1, \ldots, n_1, n_1 + j \}$ given by

\begin{align}
\label{eqn:partition-independent}
P(Z_{j}^{[t+1]} \mid \Phi, \bgamma_{obs}, \pi) \propto 
\begin{cases}
\frac{\pi}{n_1}\exp(w_{qj}(\Phi) )  & \text{if} \; q \leq n_1 \\
1- \pi
%  \frac{n_1}\frac{n_2}{n_{12}(\bZ_{-j}) + \beta_{\pi}}{n_{12}(\bZ_{-j}) + \alpha_{\pi}}
   &  \text{if} \; q = n_1 + j.
\end{cases}
\end{align}
\end{enumerate}

\begin{align}
P(\bZ^{[t+1]}\mid \Phi, \bgamma_{obs}, \pi) \propto \prod_j  \prod_q
P(Z_{j}^{[t+1]} = q \mid \Phi, \bgamma_{obs}, \pi),
\end{align}
which does not depend on the update of $Z_{j}^{[t+1]}.$ Therefore, we can update the partition independently and in parallel.

See Appendix \ref{app:full-conditional} for derivations of the full conditional distributions.

\newpage

\bibliographystyle{ba}
\bibliography{vabl}

\appendix

\section{Full Conditionals}
\label{app:full-conditional}

In this section, we provide the full conditional distributions for our Gibbs sampler in section \ref{sec:gibbs}.

First, we consider the full conditional distributions for $\bem_{f\ell}$ and $\bu_{f\ell}.$
It follows from equations \ref{eqn:sadinle-likelihood} and \ref{eqn:fabl-priors} that this is a standard Multinomial-Dirichlet:
%
\begin{align}
\bem_{f\ell} \mid \bZ, \bgamma_{\text{obs}} &\sim \text{Dirichlet}(a_{f\ell}(\bZ) + \alpha_{f\ell})\\
&=  \text{Dirichlet}(\sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j = i) + \alpha_{f\ell})
\end{align}
and 
\begin{align}
\bu_{f\ell} \mid \bZ, \bgamma_{\text{obs}} &\sim \text{Dirichlet}(b_{f\ell}(\bZ) + \beta_{f\ell})\\
&=  \text{Dirichlet}(\sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f \neq \ell)I(Z_j = i) + \beta_{f\ell}).
\end{align}

Second, we consider the full conditional distribution for $\pi.$ Recall that $\pi$ provides the proportion of matches across the two databases subject to the restriction that the total number of links across both databases is $n_{12} (\bZ).$ Recall that $n_2$ is the larger of the two databases. 
This implies that the conditional distribution is
\begin{align}
p(\pi \mid \bZ)& \propto p(\bZ \mid \pi) p(\pi) \\
& \propto \pi^{n_{12}(\bZ)} (1-\pi)^{n_2(\bZ) - n_{12}(\bZ)} \times \pi^{\alpha_\pi - 1} (1 - \pi)^{\beta_{\pi} - 1} \\
& \propto \pi^{n_{12}(\bZ) + \alpha_\pi - 1} (1-\pi)^{n_2(\bZ) - n_{12}(\bZ) + \beta_{\pi} - 1} 
\end{align}
This implies $$\pi \mid \bZ \sim \text{Beta}(n_{12}(\bZ) + \alpha_\pi , n_2(\bZ) - n_{12}(\bZ) + \beta_{\pi}).$$

\newpage
Recall the full likelihood from equation \ref{eqn:sadinle-likelihood}

\begin{align}
\mathcal{L}(\bZ, \Phi \mid \bgamma^{\text{obs}}) 
&= 
\prod_{f=1}^{F}
\prod_{\ell=1}^{L_f}
\bigg[
m_{f\ell}^{
\sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j = i)
}
\times
u_{f\ell}^{
\sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j \neq i)
}
\bigg]\\
& = 
\prod_{f=1}^{F}
\prod_{\ell=1}^{L_f}
\bigg[
m_{f\ell}^{a_{f\ell}(\bZ)}
u_{f\ell}^{b_{f\ell}(\bZ)}
\bigg],
\end{align}
%%%
where 
\begin{align}
\label{eqn:ab}
a_{f\ell}(\bZ) &= \sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j = i) \quad \text{and} \\
b_{f\ell}(\bZ) &= \sum_{i,j} I_{\text{obs}}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell)I(Z_j \neq i). \notag
\end{align}

Now, let us consider when we have a match. This corresponds when record $i \leq n_1$ and leads to the following expression:
\begin{align}
\mathcal{L}_{ij}(\bZ, \Phi \mid \bgamma^{\text{obs}}) 
&= \prod_{f=1}^F 
\bigg[
\prod_{\ell=1}^{L_f} 
\bigg(
\frac{m_{f \ell}}
{u_{f \ell}
}
\bigg) ^{
I(\gamma_{ij}^f = \ell)
}
\bigg]^{I_{\text{obs}}(\gamma_{ij}^f)} \\
&= \exp  
\bigg \{
\sum_{f=1}^F
I_{\text{obs}}(\gamma_{ij}^f)
\sum_{\ell =1}^{L_f}
\log 
\bigg(
\frac{m_{f \ell}}
{u_{f \ell}
}
\bigg)
I(\gamma_{ij}^f = \ell)
\bigg\} \\
&=: \exp \{ w_{ij}(\Phi) \}.
\end{align}
In the case of a non-match, $i = n_1 + j,$ and the likelihood is proportional to a constant. This result immediately follows from Sadinle (2017) and Wortman (2019), where the later provides more explicit derivational details for each of the two cases.

It follows that 
\begin{align}
\mathcal{L}_{ij}(\bZ, \Phi \mid \bgamma^{\text{obs}}) 
&\propto
\begin{cases}
  \exp \{ w_{ij}(\Phi) \}  & \text{if} \quad  i \leq n_1 \\
  1 & \text{if} \quad i = n_1 + j
\end{cases}
\end{align}

It follows that we can update $Z_j$ using the following expression:
\begin{align}
\label{eqn:partition}
p_{ij} (Z_{j} \mid \Phi, \bgamma^{\text{obs}}, \pi) 
\propto P(Z_j \mid \pi) \times \mathcal{L}_{ij}(\bZ, \Phi \mid \bgamma^{\text{obs}}) 
&\propto
\begin{cases}
 \frac{\pi}{n_1} \exp \{ w_{ij}(\Phi) \}  & \text{if} \quad  i \leq n_1 \\
  1 - \pi & \text{if} \quad i = n_1 + j
\end{cases}\\
\end{align}

We sample the conditional distribution for each $Z_j$ by normalizing the probabilities $\frac{\pi}{n_1} \exp \{ w_{ij}(\Phi) \}$ in equation \ref{eqn:partition} for each $i \leq n_1$ and equal to $1-\pi$ for $i = n_1 + j,$ where $j$ is any new label.

Using these probabilities, we randomly sample $Z_j$ conditional on $\Phi, \bgamma^{\text{obs}}, \pi.$ Observe that for any $i, j$ $(i \neq j)$ the conditional update does not depend on the matching label providing a posteriori independence. It follows that we can update the conditional distribution of the entire matching label independently:
\begin{align}
p(\bZ \mid \Phi, \bgamma^{\text{obs}}, \pi) \propto \prod_j \left[ \prod_i p_{ij} (Z_{j} \mid \Phi, \bgamma^{\text{obs}}, \pi) \right].
\end{align}

\end{document}

### 2023-07-07 R/fabl_gibbs.R

@@ -1,0 +1,135 @@
+fabl_gibbs <- function(comparisons, m_prior = 1, u_prior = 1,
+                           alpha = 1, beta = 1, S = 1000, burn = 100,
+                           show_progress = T, fast = F, R = NULL,
+                           all_patterns = TRUE){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRecords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
+  # alpha = first parameter of prior for linkage probability
+  # beta = second parameter of prior for linkage probability
+  # S = number of Gibbs iterations
+  # burn = number of iterations to be discarded as burn-in
+  # show_progress = set to false to show simulation progress
+
+  fields <- length(comparisons[[4]])
+  n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
+  indicators_raw <-comparisons[[1]]
+
+  if (fast) {
+    numCores <- parallel::detectCores()
+    cl <- parallel::makeCluster(numCores)
+  }
+
+  ptm <- proc.time()
+  patterns <- GetUniquePatterns2(comparisons, R = R)
+  elapsed_hash <- proc.time() - ptm
+
+  parameter_split <- as.vector(unlist(sapply(1:fields, function(x){
+    rep(x, comparisons[[4]][x])
+  })))
+
+  ids <- expand.grid(1:n1, 1:n2)
+
+  unique_patterns <- patterns[[1]]
+  pattern_counts <- patterns[[2]]
+  P <- dim(unique_patterns)[1]
+  counts_by_rec <- patterns[[3]]
+  hash_to_rec1 <- patterns[[4]]
+  candidates <- 1:n1
+  candidates_P <- 1:(P+1)
+  Z.SAMPS <- matrix(NA, nrow = n2, ncol = S)
+  M.SAMPS <- matrix(NA, nrow = length(parameter_split), ncol = S)
+  U.SAMPS <- matrix(NA, nrow = length(parameter_split), ncol = S)
+  L.SAMPS <- vector(length = S)
+  Z.temp <- rep(0, n1*n2)
+  Z <- rep(n1+1, n2)
+  L <- 0
+
[…]

### 2023-07-07 SEI_sensitivity/SEI_analysis.R

@@ -1,0 +1,140 @@
+
+R_vals <- c(1, 2, 5, 10, 20, "No SEI") %>%
+  factor(., levels = .)
+Z_paths <- list.files(path = "SEI_sensitivity/out/Zs", full.names = T)
+
+Zs_list <- lapply(Z_paths, readRDS)
+names(Zs_list) <- R_vals
+
+Zs <- lapply(Zs_list, function(x){
+  dim(unique(x, MARGIN = 2))[2]
+})
+
+Zs <- Zs %>%
+  unlist() %>%
+  data.frame(R_vals, .) %>%
+  mutate(metric = "Zs",
+         stat = "distinct")
+
+
+
+names(Zs) <- c("R", "value", "metric", "stat")
+rownames(Zs) <- NULL
+
+Z_plot <-  Zs %>%
+  ggplot() +
+  aes(x = R, y = value, group = 1) +
+  geom_line() +
+  labs(y = "Distinct Z Estimates", x = "S") +
+  theme_bw(base_size = 10)
+
+Z_plot
+
+#ggsave(plot = Z_plot, filename = "SEI_sensitvity/Z_plot.png")
+
+eval_paths <- list.files(path = "SEI_sensitivity/out/eval", full.names = T)
+eval_list <- lapply(eval_paths, readRDS)
+
+thing <- map2(eval_list, R_vals, ~data.frame(t(.x), .y)) %>%
+  do.call(rbind, .)
+names(thing) <- c("recall", "precision", "fmeasure", "R")
+
+eval_summary <- thing %>%
+  pivot_longer(cols = 1:3, names_to = "metric") %>%
+  group_by(metric, R) %>%
+  summarize(avg = mean(value),
+            lower = quantile(value, .025),
+            upper = quantile(value, .975),
+            .groups = "drop")
+
+eval_plot <- eval_summary %>%
+  filter(metric != "fmeasure") %>%
+  ggplot(aes(x = R, y = avg,
+             ymin  = lower, ymax = upper)) +
+  geom_pointrange(position = position_dodge2(width = .5)) +
+  facet_wrap(~metric, scales = "free") +
+  labs(y = "Value", x = "S")
+  theme_bw(base_size = 10)
+
+eval_plot
+
+ggsave(plot = eval_plot, filename = "SEI_sensitivity/eval_plot.png")
+
+
+
+
+
+
+# recall_mean <- lapply(eval_list, function(x){
+#   mean(x[1, ])
[…]

### 2023-07-07 SEI_sensitivity/SEI_sensitivity.R

@@ -1,0 +1,81 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+library(readr)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+set.seed(41)
+files <- list.files(path = "data/SimulationDataFiles", full.names = T)
+all_patterns = TRUE
+
+m_prior = 1
+u_prior = 1
+alpha = 1
+beta = 1
+S = 500
+iterations <- 100
+burn = 100
+show_progress = T
+fast = F
+all_patterns = TRUE
+
+
+overlap_vals <- c(50, 250, 450)
+
+j <-  2
+i <-  296
+
+overlap <- overlap_vals[j]
+
+records <- read_csv(files[i], col_types = cols())
+records$file <- rep(2:1, length.out = dim(records)[1])
+
+records <- records %>%
+  janitor::clean_names() %>%
+  mutate(rec_id = as.numeric(str_extract(rec_id, "\\d{3}")) + 1)
+
+n1 <- 500
+n2 <- 500
+#overlap <- n2/2
+
+Ztrue <- n1 + 1:n2
+Ztrue[1:overlap] <- 1:overlap
+
+file1 <- records %>%
+  filter(file ==1,
+         rec_id <= n1) %>%
+  select(-rec_id) %>%
+  as.matrix(.) %>%
+  data.frame(.) %>%
+  mutate(occup = as.numeric(occup))
+
+file2 <- records %>%
+  filter(file == 2,
+         rec_id %in% c(1:overlap, (n1 +1):(1000 - overlap))) %>%
+  select(-rec_id) %>%
+  as.matrix() %>%
+  data.frame(.) %>%
+  mutate(occup = as.numeric(occup))
+
+
+cd <- suppressMessages(BRL::compareRecords(file1, file2, c(2, 3, 5, 6),
+                                           types = c("lv", "lv", "bi", "bi")))
+cd[[1]] <- apply(cd[[1]], 2, as.numeric)
+R_vals <- c(1, 2, 5, 10, 20)
+R <- R_vals[taskID]
+
+Zs <- matrix(NA, nrow = n2, ncol = iterations)
+evals <- matrix(NA, nrow = 3, ncol = iterations)
+
+for(iter in 1:iterations){
[…]

### 2023-07-07 SEI_sensitivity/SEI_sensitivity_null.R

@@ -1,0 +1,80 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+library(readr)
+
+set.seed(41)
+files <- list.files(path = "data/SimulationDataFiles", full.names = T)
+all_patterns = TRUE
+
+m_prior = 1
+u_prior = 1
+alpha = 1
+beta = 1
+S = 500
+iterations <- 100
+burn = 100
+show_progress = T
+fast = F
+all_patterns = TRUE
+
+
+overlap_vals <- c(50, 250, 450)
+
+j <-  2
+i <-  296
+
+overlap <- overlap_vals[j]
+
+records <- read_csv(files[i], col_types = cols())
+records$file <- rep(2:1, length.out = dim(records)[1])
+
+records <- records %>%
+  janitor::clean_names() %>%
+  mutate(rec_id = as.numeric(str_extract(rec_id, "\\d{3}")) + 1)
+
+n1 <- 500
+n2 <- 500
+#overlap <- n2/2
+
+Ztrue <- n1 + 1:n2
+Ztrue[1:overlap] <- 1:overlap
+
+file1 <- records %>%
+  filter(file ==1,
+         rec_id <= n1) %>%
+  select(-rec_id) %>%
+  as.matrix(.) %>%
+  data.frame(.) %>%
+  mutate(occup = as.numeric(occup))
+
+file2 <- records %>%
+  filter(file == 2,
+         rec_id %in% c(1:overlap, (n1 +1):(1000 - overlap))) %>%
+  select(-rec_id) %>%
+  as.matrix() %>%
+  data.frame(.) %>%
+  mutate(occup = as.numeric(occup))
+
+
+cd <- suppressMessages(BRL::compareRecords(file1, file2, c(2, 3, 5, 6),
+                                           types = c("lv", "lv", "bi", "bi")))
+cd[[1]] <- apply(cd[[1]], 2, as.numeric)
+R_vals <- c(1, 2, 5, 10, 20)
+R <- NULL
+
+Zs <- matrix(NA, nrow = n2, ncol = iterations)
+evals <- matrix(NA, nrow = 3, ncol = iterations)
+
+for(iter in 1:iterations){
+  Zchain <- BKSimple_hash2(cd, R = R, S = S)
+  Zhat <- LinkRecordsBK(Zchain[[1]], n1, 1, 1, 2, Inf)
[…]

### 2023-07-07 SEI_sensitvity/SEI_analysis.R

@@ -1,20 +1,0 @@
-
-
-Z_paths <- list.files(path = "SEI_sensitvity/Zs", full.names = T)
-
-thing <- readRDS("SEI_sensitvity/Zs/Zs_01")
-Zs_list <- lapply(Z_paths, readRDS)
-
-distinct_Z <- lapply(Zs_list, function(x){
-  dim(unique(x, MARGIN = 2))[2]
-})
-
-eval_paths <- list.files(path = "SEI_sensitvity/eval", full.names = T)
-eval_list <- lapply(eval_paths, readRDS)
-recall_var <- lapply(eval_list, function(x){
-  var(x[1, ])
-})
-
-precision_var <- lapply(eval_list, function(x){
-  var(x[2, ])
-})

### 2023-07-07 cluster_jobs/run_fabl_ind.R

@@ -38,7 +38,7 @@ Ztrue <- sapply(df89$unique_ID, function(x){
 }) %>%
   unname()
 
-comparisons <- readRDS("full_hash/independent/full_hash")
+comparisons <- readRDS("cluster_jobs/full_hash/independent/full_hash")
 chain <- BKSimple_hash2_big(comparisons)
 Zchain <- chain$Z
 Zhat <- LinkRecordsBK(Zchain, nA, 1, 1, 2, Inf)
@@ -50,3 +50,8 @@ results <- rbind(eval, eval_resolved) %>%
 row.names(results) <- NULL
 results$method <- c("fabl_ind_flds", "fabl_ind_flds_resolved")
 saveRDS(results, "eval/fabl/independent")
+
+# thing <- comparisons[[2]][[4]] %>%
+#   sapply(., function(x){sapply(x, length)})
+
+sort(thing, decreasing = T)

### 2023-07-07 ba_resubmit/kundinger_resubmit2.tex
udpated draft

@@
 
-{\color{red}{The following paragraph is pretty unnecessary, I never use this notation}}
+%{\color{red}{The following paragraph is pretty unnecessary, I never use this notation}}
 Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of non-matches by 
@@
 
-{\color{red}{Beka suggested eliminating the following two paragraphs. Thoughts? I think the first one is good, the second seems unneccesary.}}
+%{\color{red}{Beka suggested eliminating the following two paragraphs. Thoughts? I think the first one is good, the second seems unneccesary.}}
 

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex
beka and jerry edits

@@
 
-Thank you for the thoughtful feedback.
-We agree that much of the notation iscumbersome, and we in particular thank the reviewers for seeing the merit in our methods despite those difficulties.
-We have made several broad changes to address the Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.
+Thank you for the thoughtful feedback. We have made several broad changes to address the Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.
 
@@
 
-\reply{We agree that our notation is at times cumbersome. We have made several efforts to simplify notation in this revision.
+\reply{We have made several efforts to simplify notation in this revision.
 
 \begin{itemize}
-	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. We find this generally more readable, and particularly simplifies the notation in Section 4.2. 
-	\item At the end of Section 4.1, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics. This is the text added:
+	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. This particularly simplifies the notation in Section 4.2.
 
+	\item At the end of Section 4.1, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics.
 	This likelihood is then used throughout Section 4.2 in the discussion of efficient posterior inference.
@@
 
-	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the clunky $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$, which we feel is more direct and intuitive.
+	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$.
 
@@
 Thus, we have removed all of the ''$H$" quantities throughout Section 4.
-We believe this substantially improves readability.
 \end{itemize}
@@
 \reply{We added this sentence in an attempt to address a previous reviewer comment pointing out a conflation between the record itself and the index directing towards that record.
-We did in fact intend for “record $i \in X1$” to be mean precisely $x_1i$.
-However, we see now that introducing $x_{1i}$ has introduced new confusion.
 With the simplified reference to the two datasets as $A$ and $B$, that material now reads:
@@
 	The square brackets have been corrected, and we apologize for the oversight.
-We have also prepared an alternative derivation of the likelihood for $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-We believe this is more clear.
+We have clarified derivation of the likelihood for $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
 }
@@
 \reply{%
-	The reviewer raises important questions.
-The main difference between the full conditionals in Sadinle (2017) and ours is that for him, $Z$ is a partition, while for us, it is not.
-This is what required him to integrate out $\pi$ in the prior, and then sequentially update each component of the $Z$ vector.
+	We realize that the integration of pi out of the full conditionals is not necessary for our sampler.
+For clarity, we decided to redo the simulations using a Gibbs sampler that samples $\pi$ from its full conditional, and then samples $Z$ from its full conditionals.
+We obtained equivalent results, which is unsurprising, because with high numbers of observations, the posterior of the beta distribution is highly concentrated around its mean.
 
-	The approach we have is justified, as indicated in the proof provided.
-However, under closer review, we have determined that it is unnecessary.
-Without integrating out $\pi$, all paramaters are updated through a standard Gibbs sampler.
-Furthermore, in other work we are developing, we actually need samples of $\pi$ in different applications.
-
-	We reran all case studies and simulation using samples of $\pi$ in the full conditional for $Z$, and found no difference in results.
-This is unsurprising, because with high numbers of observations, the posterior of the beta distribution is highly concentrated around its mean.
-
-	Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler. We leave the remark about marginalization in the Appendix to highlight the similarity to Sadinle's method.
+	Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler. We leave the remark about marginalization in the Appendix to facilitate comparisons to Sadinle's method.
 
@@
 
-	I can see why this is confusing. Therefore, I have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
+	I can see why this is confusing. Therefore, we have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
 	}
@@
 	\reply{%
-		We show in the revised Section 4.2 that the perspective of patterns allows us to write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ entirely throught the statistics in $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
+		We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
 However, we cannot express the full likelihood in (14) through these statistics.
-Therefore, we have revised to text to say ``summary statistics" instead of the more technical ``sufficient statistics."}
+Therefore, we have revised to text to say ``summary statistics" instead of  ``sufficient statistics."}
 
@@
 	We have switched the order of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI.
-With this new ordering, and with the newlyu provided equations (15 -19) it is clearer that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
+With this new ordering, and with the newly provided equations (15 -19) it is clearer that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
 }
@@
 SEI only affects the step shown in (19).
-We have provided an additional simulation study to show robustness to the SEI procedure.
 }
@@
 \reply{% 
-	We have provided an additional simulation study in Section 5.3 to show robustness to different choices of $S$.
+	We have provided an additional simulation study in Section 5.3 to show the sensitivity to different choices of $S$.
 }
@@
 \reply{%
-	That is correct. This has been corrected, and we apologize for the oversight. 
+	This has been corrected, thank you.
 }
@@
 \reply{%
-	We use $Z$ when discussing a random quantity, and $z$ to reference a realized value. We have revised, and have found all instances of inconsistency. 
+	We use $Z$ when discussing a random quantity, and $z$ to reference a realized value. We have revised all instances of inconsistency. 
 }
@@
 \reply{%
-	$\gamma$ is a matrix comprised by comparison vectors. It is certainly not a set in the mathematical sense. All references to $\gamma$ as a set have been removed.
+	$\gamma$ is a matrix comprised by comparison vectors. All references to $\gamma$ as a set have been removed.
 }
@@
 \reply{%
-	You are correct. This has been revised. 
+	This has been revised, thank you
 }

### 2023-07-17 Prelim/cpm_proposal2.tex
beka and jerry edits

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	filecolor=magenta,      
	urlcolor=cyan,
	pdftitle={Overleaf Example},
	pdfpagemode=FullScreen,
}

\begin{document}
	\title{Coarse Probabilistic Matching for Causal Inference through Bayesian Fellegi Sunter}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger}
	\institute{Duke University}
	\date{\today}

%	\AtBeginSection[]{
%		\begin{frame}
%			\vfill
%			\centering
%			\begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%				\usebeamerfont{title}\insertsectionhead\par%
%			\end{beamercolorbox}
%			\vfill
%		\end{frame}
%	}

\AtBeginSection[]
{
	\begin{frame}
		\frametitle{Table of Contents}
		\tableofcontents[currentsection]
	\end{frame}
}

	\begin{frame}
		\titlepage
	\end{frame}

\section{Record Linkage with Fellegi Sunter}

\begin{frame}{Linkage through Comparison Vectors}
	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide3.png}
	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide4.png}
	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide5.png}
\end{frame}

\begin{frame}{Review of Fellegi Sunter}
	Let $\Delta_{ij} \sim \text{Bernoulli}(\lambda)$ be the indicator that  $(A_i, B_j)$ is a match. Let $m_{fl} = p(\gamma_{ij}^f = l | \Delta_{ij} = 1)$, and $u_{fl} = p(\gamma_{ij}^f = l | \Delta_{ij} = 0)$. \linebreak

	The Fellegi Sunter Model is:
	\begin{align*}
		p(\Gamma| m, u, \Delta) = \prod_{i}^{n_A} \prod_j^{n_B} \lambda\left[\prod_{f=1}^F \prod_{l = 1}^{L_f}m_{fl}^{\gamma_{ij}^f = l}\right]  + (1 - \lambda) \left[\prod_{f=1}^F \prod_{l = 1}^{L_f}u_{fl}^{\gamma_{ij}^f = l}\right] 
	\end{align*}
\end{frame}

\begin{frame}{Bayesian Fellegi Sunter}
	\begin{itemize}
		\item Different Bayesian priors on $\Delta$ can enforce many-to-many, many-to-one, or one-to-one matchings.
		\item Bayesian methods average results over sets of plausible matchings, and provide uncertainty quantification.
		\item Extensions for enhanced speed and scalability
	\end{itemize}
\end{frame}

\section{Matching in Causal Inference}

\begin{frame}{Matching in Causal Inference}
	\begin{itemize}
		\item When estimating causal effects using observational data, it is desirable to replicate a randomized experiment as closely as possible by obtaining treated and control groups with similar covariate distributions.
		\item This goal can often be achieved by choosing well-matched samples of the original treated and control groups, thereby reducing bias due to the covariates
		<2->\item If Fellegi Sunter works for record linkage, why not for causal inference?
	\end{itemize}
\end{frame}

\begin{frame}{Mahalonobis Distance and Propensity Matching}
	\begin{itemize}
		\item One approach to matching is to use the Mahalonobis distance, which measures distance between units purely in the covariate space, without regard to which features are related to the outcome of interest. It is defined by
		\begin{align*}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1} |X_1 - X_2|}
		\end{align*}
		\item Another method is to estimate propensity scores $p(Z_n = 1 |X_n)$ for each unit in the data, and match units based on this score.
		\item Both methods are Equal Percent Bias Reducing (EPBR), meaning that
			$$E(\bar{X}_{m_T} - \bar{X}_{m_C})  \gamma E(\bar{X}_{n_T} - \bar{X}_{n_C})$$
		for $0 \leq \gamma < 1,$ where $m_T, m_C$ are the matched treated and control data, and $n_T, n_C$ are the full data
	\end{itemize}
\end{frame}

\begin{frame}{Critique of Propensity Score Matching}
	\begin{itemize}
		\item David King published a strong critique of EPBR methods in general, and PS matching in particular
		\item The theoretical garauntees for EPBR occur in expectation across many experiments (expectation on the RHS)
		\item Only one parameter $\gamma$ for all fields, no way to match more precisely for specific fields
		\item PS matching reduces all of the information in $X$ to single score, often matches units that do not actually appear similar
		\item PS attempts to recreate a \emph{completely} randomized experiment, but it is better to recreate a \emph{blocked} randomized experiment
	\end{itemize}
\end{frame}

\begin{frame}{King's Coarse Exact Matching (CEM)}
	\begin{itemize}
		\item Coarsen each variable to integer values based on reasonable knowledge (or using automated recommendations)
		\item Each data unit has effectively been transformed to a set of integer values. 
		\item Call treated and control units a match if they match at the coarsened level for \emph{all} variables
		\item This coarsening is controversial, but it allows CEM to be Monotonic Imbalance Bounding (MIB). Shown to have better covariate balance, at expense of sample size.
		\item<2> THOUGHT: This coarsening sounds a lot like Fellegi-Sunter...
	\end{itemize}
\end{frame}

\section{Fellegi Sunter as Expected Imbalance Bounding}

\begin{frame}{Monotonic Imbalance Bounding}
	As defined by \href{https://gking.harvard.edu/files/gking/files/cem_jasa.pdf}{Iacus, King, Porro (2012)}, a matching method is \emph{Monotonic Imbalance Bounding} on a function $f$ with respect to a distance $D(\cdot, \cdot)$, if for a monotonically increasing function $\gamma_{f, D}(\cdot)$ and any $\pi \in \mathbb{R}_{+}^K$, we have
	\begin{align*}
		D(f(\mathcal{X}_{m_T(\pi)}), f(\mathcal{X}_{m_C(\pi)})) \leq \gamma_{f, D}(\pi)
	\end{align*}

	Here, $\pi$ is a vector of scalar parameters relating the coarsening of each feature. Also ${m_T(\pi)}$ and ${m_C(\pi)}$ are the sizes of the matched sets, importantly determined by $\pi$, and not determined by the modeller. Lastly, $\mathcal{X}_{m_T}$ and $\mathcal{X}_{m_C}$ are the matched subsets of $T$ and $C$.
\end{frame}

\begin{frame}{Expected Imbalance Bounding}
	King's CEM is deterministic, and therefore is able to satisfy stronger criteria. No probabilistic method can fulfill this. However, I propose we think about  a class of Expected Imbalance Bounding (EIB) matching methods, as somewhat of a middle ground between MIB and EPBR. Such methods would satisfy:
	\begin{align*}
		E\left[D(f(\mathcal{X}_{m_T(\pi)}), f(\mathcal{X}_{m_C(\pi)}))\right] \leq \gamma_{f, D}(\pi)
	\end{align*}

\end{frame}

\begin{frame}{Review of Fellegi Sunter}
	Let $\Delta_{ij} \sim \text{Bernoulli}(\lambda)$ be the indicator that  $(A_i, B_j)$ is a match. Let $m_{fl} = p(\gamma_{ij}^f = l | \Delta_{ij} = 1)$, and $u_{fl} = p(\gamma_{ij}^f = l | \Delta_{ij} = 0)$. \linebreak

	The Fellegi Sunter Model is:
	\begin{align*}
		p(\Gamma| m, u, \Delta) = \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} \lambda\left[\prod_{f=1}^F \prod_{l = 1}^{L_f}m_{fl}^{\gamma_{ij}^f = l}\right]  + (1 - \lambda) \left[\prod_{f=1}^F \prod_{l = 1}^{L_f}u_{fl}^{\gamma_{ij}^f = l}\right] 
	\end{align*}
\end{frame}

\begin{frame}{MIB for Binary Covariates}
	Let $X_{m_T}$ and $X_{m_C}$ be vectors of binary covariates for units matched through Fellegi-Sunter (and with matchings encoded in $\Delta$)

	\begin{align*}
		E[|\bar{X}_{m_T} - \bar{X}_{m_C}|] &= \frac{1}{n}E\left[|\sum{X}_{m_T, i} - \sum{X}_{m_C, j}|| \Delta\right] \\
		&\leq \frac{1}{n} E\left[\sum |X_i - X_j| | \Delta_{ij} = 1\right] \\
		&= 0 \cdot m_{1} + 1 \cdot m_{2} \\
		&=m_{2}
	\end{align*}
\end{frame}

\begin{frame}{MIB for Continuous Covariates}
	I can prove this for continuous covariates as well. Suppose $X$ has range $R = X_{\max} - X_{\min}$, and bin the data according to $\epsilon = \frac{R}{K}$, where $\pi_j = K$ is our coarsening parameter. Then the upper bound becomes $\epsilon \sum_{k=1}^K k m_k$. Note that this is increasing in $K$. \linebreak

	Note that we can directly compute this bound, and check that it holds on actual matched samples. Therefore, we can use desired bounds on covariate imbalance as a way to set priors for the $m$ parameters!
\end{frame}

\section{Fellegi Sunter as Course Probabilistic Matching}

\begin{frame}{Course Probabilistic Matching}
	In this context, I'd like to propose Course Probabilistic Matching (CPM). As in Fellegi-Sunter, we create rules to characterize the level of agreement between different values. For example, we could have $\gamma_{ij}^f = |C(X_{i, f}) - C(X_{j, f})|$, where $C(\cdot)$ is a coarsening function mapping bins in the covariate space to integer values. \linebreak

	We calculate probabilities exactly as we would under FS:
	\begin{align*}
		P(\gamma_{ij}| \Delta_{ij} = 1) &= \prod_{f=1}^F p(\gamma_{ij}^f|\Delta_{ij}= 1) \\
		P(\gamma_{ij}| \Delta_{ij} = 0) &= \prod_{f=1}^F p(\gamma_{ij}^f|\Delta_{ij}= 0)
	\end{align*}
These are just products of relevant $m$ and $u$ probabilities.
\end{frame}

\begin{frame}{Course Exact Matching}
	Then, CEM can be viewed as a special case of CPM, where comparison vectors are constructed through $\gamma_{ij}^f = I(C(X_{i, f}) = C(X_{j, f}))$, and probabilities are assigned deterministically through \linebreak
\begin{align*}
	P(\Delta_{ij} = 1 | C(X_i) = C(X_j)) &= P(\Delta_{ij} = 1 |\gamma_{ij}^f = 1 , \forall f) = 1 \\
	P(\Delta_{ij} = 1 | C(X_i) \neq C(X_j)) &= 0 \\
\end{align*}
Additionally, if the $m$ parameter gives all its probability mass to exact agreement, then the MIB bound for FS,  $\epsilon \sum_{k=1}^K k m_k$,  reduces down to just $\epsilon$, as provided in the paper.
\end{frame}

\begin{frame}{One Difficulty}
	Ordinarily, we use $m^f \sim \text{Dirichlet}(\alpha_1, \ldots, \alpha_{L_f})$ as a prior for each field. However, I sense that this will be too unstructured, and the sampler would be too flexible in choosing matches. \linebreak

	However, if we decide on acceptable bounds for covariate imbalance, we can use these bounds to construct these priors.
\end{frame}

\begin{frame}{Stacked Truncated Beta Priors}
	Sadinle 2014 used a series of beta distributions as follows. Define $m^{*}_{f0} = P(\Gamma_{ij}^f = 0)$ and $m^{*}_{fl} = P(\Gamma_{ij}^f = l|\Gamma_{ij}^f > l-1)$. \linebreak

	These are related to the standard Fellegi Sunter parameters through:
	\begin{align}
		m_{fl} = P(\Gamma_{ij}^f = l | \Delta_{ij} = 1) = \prod_{l=1}^{L_f - 1} m_{fl}^{*I(\gamma_{ij} = l)}(1 - m_{fl})^{*I(\gamma_{ij} > l)}
	\end{align}

\end{frame}

\begin{frame}{Prior Specification}
	This parameterization is more useful because we can more directly control the kinds of matches we want to make through setting the support of the prior. We would use $m^{*}_{f0} \sim \text{TBeta}(\alpha_{fl}, \beta_{fl}, \lambda_{fl}, 1)$ prior distribution, where $\lambda_{fl}$ is a lower bound for the support. \linebreak

	This may seem overly informative, but these $\lambda$ effectively set bounds for the allowable covariate imbalance, and it is common for this to be chosen by the matcher.
\end{frame}

\section{Conclusion}
\begin{frame}{Concluding Thoughts}
	\begin{itemize}
		\item I suspect the Coarse Probabilistic Matching would outperform Course Exact Matching pretty easily. It would find all of the coarsened "exact" matches, plus additional near matches, resulting in mild increases in covariate imbalance with substantial increases in sample size after matching
		\item Since this matching method is fully Bayesian and probabilistic, we can create joint models for the matching and treatment effect estimation, propagating uncertainty between the steps, and averaging over many plausible matchings.
	\end{itemize}
\end{frame}

\begin{frame}{Tons of Ways Forward}
	\begin{itemize}
		\item Establish theory for CPM in the context of other matching methods, compare out of box performance. Particularly interested in comparisons with King's Matching Frontier and Morucci's Matching Bounds.
		\item Incorporate propensity scores in the CPM method (construct a comparison feature based on that score)
		\item Variable selection prior for propensity score, feeding into a record linkage model on the selected covariates
		\item Many other ideas!
		%\item Empirical Process for deriving bounds on covariate imbalance
	\end{itemize}
\end{frame}

\end{document}

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
beka and jerry edits

--- 2023-07-07 ba_resubmit/kundinger_resubmit2.tex
@@
 This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
-Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
+Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
 %
@@
 \begin{keyword}%[class=MSC]
-\kwd{Bayesian methods}
+\kwd{data fusion}
 \kwd{distributed computing}
@@
 
-Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data files and independently classified each pair as a match or a non-match using a likelihood ratio test.
-Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
-Additionally, some methods directly model the linkage variables \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.
+Many probabilistic record linkage methods are rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
+In their approach, the data analyst first creates comparison vectors for each pair of records in the data files.
+These vectors indicate how similar the records are on a set of variables measured in both files, known as the linkage variables.
+Using these comparison vectors, the analyst classifies each pair as a match or nonmatch using a likelihood ratio test.
+Crucially, these decisions are made independently for each pair.
+The \cite{fellegi_theory_1969} approach has been extended for a wide variety of applications \citep[e.g.,][]{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
+An alternative paradigm is to model the linkage variables directly \citep[e.g.,][]{tancredi2011hierarchical, steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}.
+In this article, we build on the contributions to the comparison vector approach.
 
@@
 
-The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice.
-In many situations, we know that there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
+The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity.
+However, In many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
-Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
-
-Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost.
+To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
+
+Alternatively, analysts can embed one-to-one matching requirements into the model specification, at an additional computational cost.
 \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
@@
 Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
-\texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
-
-In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space.
-This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with minimal loss of accuracy.
+\texttt{BRL} has been shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
+
+In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
+{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
 In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
 
-%\textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
-%\textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking.
-This can be useful when there is no reliable blocking field available or one desires estimates of model parameters for the entire sample in question.
-In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.
-We illustrate our methodology on two case studies and two simulation scenarios, where Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.}
-
-%We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed.
-Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
+%Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space.
 
@@
 
-\section{Review of Prior Work}
+\section{Review of Fellegi-Sunter Approaches for Record Linkage}
 \label{sec:review-of_prior-work}
 
-%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.
-
-Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
+Consider two data files $A$ and $B$ comprising $n_A$ and $n_B$ records, respectively, and including $F$ linkage variables measured in both files.
+For $i=1, \dots, n_A$, let record $i$ be given by $A_i = (A_{i1}, \dots, A_{iF})$, so that $A = (A_i : i = 1, \dots, n_A)$.
+Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.
+Without loss of generality, denote files such that $n_A \geq n_B$.
+
+%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
 Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
-Without loss of generality, denote files such that $n_A \geq n_B$.
-Under bipartite matching, the set of matches across data files can be represented in two equivalent ways.
-First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
+
+Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
+For notational convenience, we define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
+
+When linking variable $f$ is categorical, a common way to define $\gamma_{ij}^f$ is an indicator for exact agreement.
+For example, if zip code is linking variable $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
+For numerical linking variables, we can use the absolute difference of the two values.
+For example, if age is linking variable $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
+For text variables like names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
+We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+
+More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for linking variable $f$ of records $A_i$ and $B_j.$ The range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
+Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
+Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
+
+In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
+As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+
+%\subsection{Fellegi-Sunter Models}
+%\label{fellegi-sunter}
+
+Having defined comparison vectors, we now turn to the \cite{fellegi_theory_1969} model for record linkage.
+We begin with notation for defining linkages.
+Under bipartite matching, the set of matches across $A$ and $B$ can be represented in two equivalent ways.
+First, we may use a matrix $\Delta \in \{0, 1\}^{n_An_B}$, where
 \begin{align}
@@
 We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.
-
-{\color{red}{The following paragraph is pretty unnecessary, I never use this notation}}
-Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of non-matches by 
-$\bm{U} =  \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
-
-%\subsection{Comparison Vectors}
-%\label{comparison-vectors}
-
-{\color{red}{Beka suggested eliminating the following two paragraphs. Thoughts? I think the first one is good, the second seems unneccesary.}}
-
-Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ {\color{red}{We use $\Gamma_{ij}$ when referring to the vector as a random variable, and $\gamma_{ij}$ when referring to observed data.
-We collect the vectors $\gamma_{ij}$ as the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$}}.
-
-%Frequently, $\gamma_{ij}$ consists of one comparison for each feature shared between the two data files.
-
-The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features.
-For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
-For numerical features, we can use absolute difference between the two feature values.
-For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
-For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
-We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
-
-More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
-Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
-Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
-
-{\color{red}{
-In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
-As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
-To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
-}
-}
-
-%For more complex measurements, we can take into account partial agreement to more richly characterize the comparison;
-
-%
-%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
-This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
-In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.
-
-\subsection{Fellegi-Sunter Models}
-\label{fellegi-sunter}
-
-The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
-Using these assumptions, \cite{winkler_state_1999}, \cite{jaro1989}, \cite{larsen_2001}, \cite{enamorado2019using} and others model the comparison data through mixture models of the form
+When presenting the models, we use both representations for convenience.
+
+Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.
+Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
+For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
+Using these independence assumptions,  one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.
+We have
 \begin{subequations}
@@
 &\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
-	&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
+&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:lambda_dist}
 \end{align}
 \end{subequations}
-Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
+Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
 When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
-Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
-The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
-
-Within this framework, record pairs are independently classified as matches and non-matches based on the estimated parameters.
+Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
+
+%Within this framework, record pairs are independently classified as matches and nonmatches based on the estimated parameters.
 However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results.
@@
 \cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.
-
-%This prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi-Sunter weights among matched pairs.
-
-%These unknown parameters are often estimated using the EM algorithm \cite{winkler_state_1999}.
 
@@
 
-%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data.
+%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
 This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
@@
 %     \label{eqn:wts}
-%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
+%     w_{ij} = \frac{p(\gamma_{ij} \mid \Delta_{ij} = 1)}{p(\gamma_{ij} \mid \Delta_{ij} = 0)}
 %\end{align}
-%as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
-%
-%In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
-%
-%%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
+%as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
+%
+%In practice, $p(\cdot \mid \Delta_{ij} = 1)$ and $p(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
+%
+%%is the set of parameters $u_{fl} = p(\gamma_{ij}^f = l|Z_j \neq i)$ for the nonmatching pairs.
 %
@@
 
-To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework.
+%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a
+
+\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.
+He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.
+As the prior distriburion for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
+Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
+It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
+Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A !
+/ (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
+Thus, the prior distribution used by \cite{sadinle_bayesian_2017} is given by
+\begin{align}
+\label{eqn:sadinle_prior}
+p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+\end{align}
+where $\text{B}(\cdot, \cdot)$ represents the Beta function.
+This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
+This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
+
+%\section{Review of Prior Work}
+%\label{sec:review-of_prior-work}
+%
+%%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.
+%
+%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
+Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
+Without loss of generality, denote files such that $n_A \geq n_B$.
+Under bipartite matching, the set of matches across data files can be represented in two equivalent ways.
+First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
+%\begin{align}
+%	\Delta_{ij} =
+%	\begin{cases}
+%		1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
+%		0, & \text{otherwise}.\\
+%	\end{cases}
+%\end{align}
+%This sparse matrix representation can become cumbersome for large linkage tasks.
+More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
+%\begin{align}
+%	Z_{j} =
+%	\begin{cases}
+%		i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
+%		n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
+%	\end{cases}
+%\end{align}
+%We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise. 
+%
+
+%Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of nonmatches by 
+%$\bm{U} =  \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as nonmatches as ``non-links.''
+%
+%%\subsection{Comparison Vectors}
+%%\label{comparison-vectors}
+%
+
+%
+%Intuitively, matching records (those that refer to the same entity) should be similar; records that are nonmatching should be dissimilar.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ {\color{red}{We use $\Gamma_{ij}$ when referring to the vector as a random variable, and $\gamma_{ij}$ when referring to observed data.
+We collect the vectors $\gamma_{ij}$ as the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$}}.
+%
+%%Frequently, $\gamma_{ij}$ consists of one comparison for each feature shared between the two data files. 
+%
+%The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features.
+For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
+For numerical features, we can use absolute difference between the two feature values.
+For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
+For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
+We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
+%
+%More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
+Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
+Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
+%
+%{\color{red}{
+%In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
+As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+}
+%}
+%
+%%For more complex measurements, we can take into account partial agreement to more richly characterize the comparison;  
+%
+%%
+%%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
+This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
+In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.
+%
+%\subsection{Fellegi-Sunter Models}
+%\label{fellegi-sunter}
+%
+%The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
+Using these assumptions, \cite{winkler_state_1999}, \cite{jaro1989}, \cite{larsen_2001}, \cite{enamorado2019using} and others model the comparison data through mixture models of the form
+%\begin{subequations}
+%\begin{align}
+%	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
+%	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
+%	&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
+%\end{align}
+%\end{subequations}
+%Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
+When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
+Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
+%
+%Within this framework, record pairs are independently classified as matches and nonmatches based on the estimated parameters.
+However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results.
+To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching.
+\cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.
+%
+%%This prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi-Sunter weights among matched pairs.
+%
+%%These unknown parameters are often estimated using the EM algorithm \cite{winkler_state_1999}. 
+%
+%%The Fellegi-Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+This is a problem in practice as transitive closures can be violated.
+Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.
+%
+%%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
+%%\begin{align}
+%%	\label{eqn:jaro}
+%%	&\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij} 
+%%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
+%%	& \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
+%%	& \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
+%%\end{align}
+%%
+%%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs.
+Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
+%
+%%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
+This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
+%%\begin{align}
+%%	\label{eqn:wts}
+%%	w_{ij} = \frac{p(\gamma_{ij} \mid \Delta_{ij} = 1)}{p(\gamma_{ij} \mid \Delta_{ij} = 0)}
+%%\end{align}
+%%as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
+%%
+%%In practice, $p(\cdot \mid \Delta_{ij} = 1)$ and $p(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
+%%
+%%%is the set of parameters $u_{fl} = p(\gamma_{ij}^f = l|Z_j \neq i)$ for the nonmatching pairs.
+%%
+%%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+This is a problem in practice as transitive closures can be violated.
+Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
+%
+%%\subsection{Beta Record Linkage Model}
+%%\label{BRL}
+%
+%%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
+%
+%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework.
 In addition to {\color{red}{uniform}} Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
@@
 Thus, the prior is given by
-\begin{align}
-\label{eqn:sadinle_prior}
-P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
-\end{align}
-where $\text{B}(\cdot, \cdot)$ represents the Beta function.
+%\begin{align}
+%\label{eqn:sadinle_prior}
+%p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+%\end{align}
+%where $\text{B}(\cdot, \cdot)$ represents the Beta function.
 This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
@@
 
-In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
+Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
 However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta prior as follows.
 For each $Z_j$, we have
-\begin{align}
-	\label{eqn:fast_beta_prior}
+\begin{subequations}
+\begin{align}
 	p(Z_j = q| \pi) &= \begin{cases} 
-	\frac{1}{n_A}\pi,  & q \leq n_A; \\
+	\frac{1}{n_A}\pi,  & q \leq n_A;\\
 	1-\pi, &  q  = n_A + j;  \\
-\end{cases} \\
-\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
-\end{align}
+\end{cases} \label{eqn:fast_beta_prior}  \\
+\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \label{eqn:fast_beta_prior2}.
+\end{align}
+\end{subequations}
 We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match.
@@
 In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
-This however amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
-In our preliminary studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}).
+This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
+In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates.
+Hence, we prefer (\ref{eqn:fast_beta_prior}).
 We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
 
-%Intuitively, this set of priors says that record $B_j$ has some match in $A$ with probability $\pi$, and that each record $A_i$ is equally likely to be that match.
-$\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-One non-informative choice might be $\pi \sim \text{Beta}(1, 1)$, which corresponds to a prior belief that non-matches and matches are equally likely, and another might be $\pi \sim \text{Beta}\left(1, \frac{1}{n_A}\right)$, which corresponds to a uniform prior on the labeling of $\bm{Z}$.
-We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_A}\right)$ random variable.
-
-Note that linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
+Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
 That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
-We find $\pi$ to be more a interpretable parameter than $\lambda$ in the bipartite case.
+In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
 In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_A \times n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
+Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
 As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
@@
 To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-However, there is a crucial difference: since \ref{eqn:fast_beta_prior} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+However, since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+To obtain a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
 Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
 
-\begin{table}[t!]
-	\centering
-	\begin{tabular}[t!]{ll}
-		Symbol & Description \\
-		\hline
-		$A, B$ & data files\\
-		$i \in 1, \ldots, n_A $ & index over records in $A$\\
-		$j \in 1, \ldots, n_B $ & index over records in $B$\\
-		$f \in 1, \ldots F$ & index over fields used for comparisons \\
-		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
-		$n_{AB}$ & number of entities in common between $A$ and $B$\\
-		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
-		$Z_j = i$ & records $A_i$ and $B_j$ match \\
-		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
-		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
-		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-		$\pi$ & probability that a record in $B$ has a match in $A$ \\
-		\hline
-	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
-\end{table}
-
-For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_A}.
+%\begin{table}[t!]
+%	\centering
+%	\begin{tabular}[t!]{ll}
+%		Symbol & Description \\
+%		\hline
+%		$A, B$ & data files\\
+%		$i \in 1, \ldots, n_A $ & index over records in $A$\\
+%		$j \in 1, \ldots, n_B $ & index over records in $B$\\
+%		$f \in 1, \ldots F$ & index over fields used for comparisons \\
+%		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
+%		$n_{AB}$ & number of matches between $A$ and $B$\\
+%		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+%		$Z_j = i$ & records $A_i$ and $B_j$ match \\
+%		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
+%		$m_{fl}$ & $p(\gamma_{ij}^f = l | Z_j = i)$ \\
+%		$u_{fl}$ & $p(\gamma_{ij}^f = l | Z_j \neq i)$ \\
+%		$\pi$ & probability that a record in $B$ has a match in $A$ \\
+%		\hline
+%	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
+%\end{table}
+%
+%For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_A}.
 \begin{subequations}
@@
 
-\hypertarget{posterior-sampling}{%
-	\subsection{Gibbs Sampler}
-	\label{gibbs_sampling}}
-
+%\hypertarget{posterior-sampling}{%
+%	\subsection{Gibbs Sampler}
+%	\label{gibbs_sampling}}
+
+We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler.
 We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^s$, we use the full conditionals,
+To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
 \begin{subequations}
@@
 	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
-	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}
+	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
+	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
 	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
@@
 \begin{align}
-	\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi}).
-\end{align}
-
-Lastly, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
+	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
+\end{align}
+where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.
+
+Lastly, we sample $\bm{Z}$ componentwise from the full conditional for each $Z_j$:
 \begin{align}
 	\label{eqn:z_full_conditional}
-	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
+	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)},  \pi^{(s+1)}\right) \propto
 	\begin{cases} 
-		\frac{\pi^{(s+1)}}{n_A} w_{q, j}^{(s+1)},  & q \leq n_A; \\
+		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
 		1 - \pi^{(s+1)}, & q  = n_A + j, \\
@@
 %\end{align}
-where, for all $i \in \{1, \ldots, n_A\}$, 
+where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
 \begin{align}
@@
 
-The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A \times n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors.
+The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A  n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors.
 One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}.
-In deterministic blocking, the modeler chooses fields thought to be highly reliable, and only compares records that agree on those fields.
+In deterministic blocking, the modeler chooses fields thought to be highly reliable and only compares records that agree on those fields.
 The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains.
@@
 
-After computing all comparison vectors within a block, the modeller can further reduce the number of comparison vectors used in the linkage algorithm through indexing.
-For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like first name, or only pairs that exactly match on a specified number of fields.
+After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing.
+For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields.
 However, the impact of indexing on model parameters is not well understood; \citep{murray2016probabilistic} reviewed this issue in the context of the frequentist Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
@@
 With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
-First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler.
+First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors.
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
-For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.
-
-\begin{table}[t!]
-	\centering
-	\begin{tabular}[t!]{ll}
-		Symbol & Description \\
-		\hline
-		$P$ & number of agreement patterns exhibited in $\gamma$ \\
-		$h_p$ & agreement pattern $p$ \\
-		$e(h_p)$ & one-hot encoding of agreement pattern $p$ \\
-		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $B_j$ exhibits pattern $p$ \\
-		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
-		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
-		$N_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
-		$N_p$ & number of total comparison vectors that exhibit pattern $p$ \\
-
-		\hline
-	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_B}
-\end{table}
+%For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.
+
+%\begin{table}[t!]
+%	\centering
+%	\begin{tabular}[t!]{ll}
+%		Symbol & Description \\
+%		\hline
+%		$P$ & number of agreement patterns exhibited in $\gamma$ \\
+%		$h_p$ & agreement pattern $p$ \\
+%		$e(h_p)$ & one-hot encoding of agreement pattern $p$ \\
+%		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $B_j$ exhibits pattern $p$ \\
+%		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+%		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+%		$N_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+%		$N_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+%
+%		\hline
+%	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_B}
+%\end{table}
 
@@
 
-	{\color{red}{NOTE: This section used to be organized 
-		\begin{itemize}
-			\item Hashing
-			\item SEI
-			\item Inference
-		\end{itemize}
-	but now it is organized
-		\begin{itemize}
-		\item Hashing
-		\item Inference
-		\item Scaling (chunking method)
-		\item SEI
-	\end{itemize}
-
-I think inference is much more natural and after hashing.
-And having all the equations and updates written out in the inference part makes it clearer that SEI does not affect posterior updates at all.
-}}
-
 Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
-Let $P$ be the number of patterns realized in $\gamma$ and note that it is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the additional count notes the possibility of missing values.
+Let $P$ be the number of patterns realized in $\gamma$.
+It is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.
 This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
-To obtain a memory efficient representation of the data, we map the agreement pattern of each record pair to a unique integer.
+To obtain a memory efficient representation, we map the agreement pattern of each record pair to a unique integer.
 \cite{enamorado2019using} accomplished this through a hashing function, which we modify to explicitly handle missing values:
@@
 \end{align}
+
 We then map these integers to sequential integers $\{1, \ldots, P\}$.
@@
 When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
-
-{\color{red}{In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
+In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
 This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
-A more complete example of the one-hot encoding is provided in Appendix \ref{app:ohe}.}}
-
-{\color{red}{
-For example, with five fields with binary agreements, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
+%A more complete example of the one-hot encoding is provided in Appendix \ref{app:ohe}.
+
+For example, with five fields with binary agreements and possible missingness, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
 Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
 Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
-We map the unique values of $h^{*}(\gamma)$ to sequential integers (for in example in \texttt{R}, by converting the integers into factors, and then back to integers).
+We map the unique values of $h^{*}(\gamma)$ to sequential integers.
 If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
-}}
 
 We then identify the records in $A$ with comparison vectors corresponding to each pattern $p$ for each record $B_j$.
-We denote this set $r_{p_j} = \{i | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
-We then compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$, given by
+We denote this set $r_{p_j} = \{i \in 1, \dots, n_A | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
+We compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$, given by
 \begin{align}\label{eqn:N}
@@
 
-{
-\color{red}{
-Together, the set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ with no loss of information.
+Together, the set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for purposes of writing the likelihood function for \texttt{fabl}.
 To see this, we employ the condensed notation
-\begin{align}
-	m_p =  p(\gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}
-\end{align}
-to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match.
-We similarly define $u_p$.
+\begin{subequations}
+	\begin{align}
+		m_p =  p(\gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:m_p} \\
+		u_p =  p(\gamma_{ij} = h_p|Z_j \neq i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:u_p}
+	\end{align}
+\end{subequations}
+to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match, in (\ref{eqn:m_p}), and not a match, in (\ref{eqn:u_p}).
 Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) is equivalent to
+
 \begin{align}\label{likelihood_efficient}
@@
 \end{align}
-}}
+The likelihood in (\ref{likelihood_efficient}) allows for more
+efficient posterior inference for $\bm{m}$, $\bm{u}$, and $\bm{Z}$, as we now describe.
 
@@
 
-Using the likelihood in (\ref{likelihood_efficient}), we can see that the statistics collected in $\tilde{\gamma}$ allow for more efficient posterior inference for all model parameters.
-Conditioning on $\bm{Z}$, we can express the conditional likelihood for $\bm{m}$ and $\bm{u}$ as
-\begin{align}
-	\mathcal{L}(\bm{m}, \bm{u} \mid \gamma, \bm{Z})  &=  \prod_{p=1}^P m_p^{N_p^m}u_p^{N_p - N_p^m}  \label{likelihood_efficient_m_u}
-\end{align}
-where $N_p^m = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ is the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ is the total occurrence of pattern $p$ in the data across all record pairs.
-Note that the $\bm{m}$ and $\bm{u}$ parameters depend on the data only through the quantities calculated from $\mathcal{N}$ and $\mathcal{P}$.
-
-Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define $\bm{\beta_0}$ similarly for the $\bm{u}$ distributions.
-{\color{red}{Using the one-hot encodings described in Section \ref{data-representation-hashing-and-storage},} the terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
+Turning first to the posterior updates for $\bm{m}$ and $\bm{u}$, we first note that these parameters depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$.
+Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
+Then, conditional on $Z$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
+\begin{align}
+	\mathcal{L}(\bm{m}, \bm{u} \mid \tilde{\gamma}, \bm{Z}, \pi)  &=  \prod_{p=1}^P m_p^{n_p(\bm{Z})}u_p^{N_p - n_p(\bm{Z})}  \label{likelihood_efficient_m_u}.
+\end{align}
+Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be a concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
+The terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
 \begin{subequations}
 	\begin{align}
-		\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P N_p^m \times e(h_p), \label{efficient_alpha} \\
-		\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P (N_p - N_p^m) \times e(h_p). \label{efficient_beta}
+		\bm{\alpha(Z}^{(s)}\bm{)} &= \bm{\alpha_0} + \sum_{p=1}^P n_p\left(\bm{Z}^{(s)}\right) \times e(h_p), \label{efficient_alpha} \\
+		\bm{\beta(Z}^{(s)}\bm{)} &= \bm{\beta_0} + \sum_{p=1}^P \left(N_p - n_p\left(\bm{Z}^{(s)}\right)\right) \times e(h_p). \label{efficient_beta}
 	\end{align}
@@
 Specifically, the $l + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}),and (\ref{eqn:beta_update}).
-In this new form however, they are calculated through $P$ multiplications of scalar quantities over the vectors $h_p$, and then vectorized summations.} This is markedly more efficient than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-We can also use (\ref{likelihood_efficient}) to express the conditional likelihood for $Z_{j}$. Using $w_p = \frac{m_p}{u_p}$, we have
-\begin{align}
-	\mathcal{L}(Z_j\mid  \bm{m}, \bm{u}, \tilde{\gamma}) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}\label{likelihood_efficient_z}
-\end{align}
-Note that $Z_j$ interacts with the data only through $\mathcal{N}$ and $\mathcal{R}$.
-This likelihood also lends itself to markedly faster posterior computations.
-Although sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_A$ is large.
+However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{eqn:beta_update}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+We also use (\ref{likelihood_efficient}) to sample Z efficiently.
+For each pattern $p$, let $w_p =\frac{m_p}{u_p}$.
+The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
+\begin{align}
+	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
+\end{align}
+This likelihood lends itself to markedly faster posterior computations.
+Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large.
 This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
 To speed up computation, we break this sampling step into two.
+
 We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
@@
 	\label{eqn:gibbs1}
-	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
 	\begin{cases} 
-		N_{p_j} \times w_{p},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
-		n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   \text{otherwise}. \\
+		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
+		1- \pi^{(s+1)} , &   \text{otherwise}. \\
 	\end{cases}
 \end{align}
-Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the $P$ agreement patterns.
-Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformally using
+Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
 \begin{align}
 	\label{eqn:gibbs2}
-	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p \right) = \begin{cases} 
+	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
 		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
@@
 \begin{lemma}
-	Recall that $n_A$ and $n_B$ are the number of records in $A$ and $B$, respectively.
+	Let $n_A$ and $n_B$ be the number of records in files $A$ and $B$, respectively.
 Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $A \times B$.
 We assume $C$ cores available for parallelization and a Gibbs sampler with $T$ iterations.
-Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P)$.
+Then, the overall computational complexity of \texttt{fabl} with hashing is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P)$.
 	\label{lemma:fabl}
@@
 The hashing procedure for all pairwise comparisons is also $O(F n_A n_B)$.
-With $B$ processors available, we can split these computations across $B$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{B} n_A n_B)$.
+With $C$ processors available, we can split these computations across $C$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{C} n_A n_B)$.
 There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
+We note that this contribution to computational complexity applies for all versions of \cite{fellegi_theory_1969} algorithms, unless they use blocking to reduce the comparison space.
 
@@
 However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
-The complexity of updating $\bm{Z}$ sequentially at the record level is $O(n_A n_B)$.
-With hashing, we split this sampling into two steps.
-First, we sample the agreement pattern of the match with complexity $O(n_B P)$, and then we sample the record exhibiting that pattern with complexity $O(n_B)$.
+The complexity of updating $\bm{Z}$ sequentially at the record level as in \cite{sadinle_bayesian_2017} is $O(n_A n_B)$.
+With hashing, we first sample the agreement pattern of the match with complexity $O(n_B P)$, and then we sample the record exhibiting that pattern with complexity $O(n_B)$.
 Thus, the complexity of sampling $\bm{Z}$ in a single iteration is $O(n_B P)$.
@@
 With parallelization, this complexity is further reduced to $O(\frac{1}{C}n_B P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{C}n_B P)$
-	In summary, the total computational complexity is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P).$
+	In summary, the total computational complexity for \texttt{fabl} is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P).$
 \end{proof}
@@
 
-		For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks $\{A^a | a = 1, \ldots, t_A \}$ and $\{B^b | b = 1, \ldots, t_B\}$ for more manageable computations.
-{\color{red}{For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each chunk $A^a$ and $B^b$ would contain 500 records.}} For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
-We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$ for later calculations, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
+		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations.
+Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.
+Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B and B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
+For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
+For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
+We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
 In detail, we calculate
@@
 	\begin{align}
-		r_{p_j}^{ab} &= \{i | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
-		N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
+		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
 	\end{align}
 \end{subequations}
-These can computed serially or in parallel.
-Summary statistics from each pairwise chunk comparison can be easily synthesized to recover summary statistics for the full comparison matrix $\gamma$.
-Specifically, we combine information and obtain the statistics comprising $\tilde{\gamma}$ through
+These can computed serially or in parallel. Summary statistics from each pairwise chunk comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
 \begin{subequations}
@@
 		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
-		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
+		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab}.\label{N_{p_j}}	\end{align}
 \end{subequations}
@@
 
-As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses dramatically less memory than storing the full comparison matrix $\gamma$.
+As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses less memory than storing the full comparison matrix $\gamma$.
 However, the memory requirements of each are still quadratic in nature.
 For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
-We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ for all $n_A \times n_B$ record pairs, while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
+We next introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
 This allows all-to-all comparisons for substantially larger linkage tasks.
 
-{\color{red}{All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$.
-Therefore, these records have the same probability to be identified as the link for record $B_j$.
-Thus, we know that records $i \in r_{p_j}$ such that $N_{p_j}$ is large are unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
-We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
-Therefore, rather than store all of these record labels, we store only a small number $S$.
+All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$.
+These records have the same probability to be identified as the link for record $B_j$.
+Thus, we know that records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate.
+Rather than store all of these record labels, we store only a small number $S$.
 For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $SEI(r_{p_j}^{ab})$.
@@
 Let $n_{A^a}$ and $n_{B^b}$ be the number of records in chunks $A^a$ and $B^b$ respectively.
-Instead of storing $n_A \times n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a} \times n_{B^b} \times P \times S$ labels.
+Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels.
 As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $SEI(\tilde{\gamma}) = \{\mathcal{P}, SEI(\mathcal{R}), \mathcal{N}\}$.
@@
 
-NOTE: I switched the order of the chunking and the SEI because SEI only makes sense to use when you split the comparisons into chunks.
-Its also clearer mathematically.
-I think it also avoids the issue of the overloaded notation that used to be in equations \ref{r_{p_j}}.
-The reviewer asked to keep the SEI in the notation to reflect how SEI changed the likelihood and the quantities of interest, but I think the sentence "As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$" makes things very clear.
-Thoughts?}
-}
-
-%{\color{red}{For example, if file $A$ with $n_A = 500$ contains record $A_1$ that is an exact match for record $B_5$ and 499 records with complete disagreement on all fields with record $B_5$, we would only store $S$ indices to represent the records with full disagreement.I added that last example.
-I don't love it though.
-Thoughts?}}
-
-%{\color{red}{When the linkage task is large enough that all $n_A \times n_B$ comparison vectors cannot be stored in memory at the same time, we propose a modified approach where we construct comparison vectors and calculate $\mathcal{R}$ and $\mathcal{N}$ for smaller subsets of the data, and combine this information before conducting the Gibbs sampler.
-We use this approach in the case study in Section \ref{nltcs}, and provide details in Appendix \ref{app:chunks}}}.
-
 	\section{Simulation Studies}
@@
 	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-We use $F = 5$ binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
+We use $F = 5$ binary comparisons with probabilities for matching and nonmatching pairs shown in Table \ref{Tab:distributions}.
 For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$.
@@
 For the other $n_B/2$ of these records, there is one match in $A$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_A - 1$ comparison vectors from the $\bm{u}$ distribution.
-We compare the run-time of \texttt{fabl} against \texttt{BRL} as we increase $n_A$ and $n_B$.
+We compare the run-time of \texttt{fabl} (with no SEI) against \texttt{BRL} as we increase $n_A$ and $n_B$.
 Since we have five binary comparison fields with no missingness, the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
@@
 
-	The Gibbs sampler in the implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.
-In contrast, we use non-optimized code written only in \texttt{R} for \texttt{fabl}.
+	In Figure \ref{fig:speed1}, where we increase both $n_A$ and $n_B$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger sample sizes.
+In particular, run-time for \texttt{BRL} grows quadratically (or linearly with the size of both $A$ and $B$) while run-time for \texttt{fabl} grows linearly (in the size of only $B$).
+
+	\begin{figure}[t]
+		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
+We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
+		\end{center}
+	\end{figure}
+
+	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}.
+The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
+To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.
+
+	Importantly, \texttt{BRL} implements a Gibbs sampler that is coded C \citep{sadinle_bayesian_2017}, while \texttt{fabl} currently uses non-optimized code written only in \texttt{R}.
 While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
@@
 
-	In Figure \ref{fig:speed1}, where we increase both $n_A$ and $n_B$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger sample sizes.
-In particular, run-time for \texttt{BRL} grows quadratically (or linearly with the size of both $A$ and $B$) while run-time for \texttt{fabl} grows linearly (in the size of only $B$).
-
-	\begin{figure}[t]
-		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
-		\end{center}
-	\end{figure}
-
-	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}.
-The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.
-
 	\begin{figure}[h!]
-		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
+		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
@@
 We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across data files.
-Under each of these settings, we have 100 pairs of simulated data files in order to obtain uncertainty quantification on our performance metrics.
+Under each of these settings, we use 100 pairs of simulated data files in order to obtain uncertainty quantification on our performance metrics.
 We use uniform priors for all $\bm{m}$ and $\bm{u}$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
@@
 Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
-We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; these results are in Appendix \ref{partial}.
+We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; those results are in Appendix \ref{partial}.
 
@@
 Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(\hat{Z}_j \leq n_A)$.
-The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2 \times (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
+The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2  (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
 In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
+			\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
 			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
@@
 
-{
-\color{red}{Finally, our last simulation demonstrates our method's robustness to different values of $S$ for the SEI memory reduction procedure.
+Finally, our last simulation demonstrates our method's robustness to different values of $S$ for the SEI memory reduction procedure.
 We perform record linkage on one set of synthetic datafile described in Section \ref{accuracy} with 500 records in each datafile, 250 entities in common across datafiles, and 3 errors present across matching records.
-To achieve more drastic results, we perform SEI without chunking the data, with $t_A = t_B = 1$.
+To achieve more drastic results, we perform SEI without chunking the data, that is, $t_A = t_B = 1$.
 In practice, if it is possible to create and store the comparison matrix for all record pairs at one time, there is no need to reduce the memory of the hashed matrix through SEI.
-For illustration however, it is easier to see degenerate behavior in this setting.
+For illustration however, it is easier to illustrate the effects of choices of $S$ in this setting.
 
 We preform linkage using SEI with $S = (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler.
-Note that any particular SEI memory reduction may lead to better or worse linkage performance; if the SEI procedure happens to only remove pairs that are not match by random chance, recall and precision will improve.
+As any particular SEI implementation may improve or worsen linkage performance; if the SEI procedure happens to only remove pairs that are not matches, recall and precision will improve.
 Therefore, we perform linkage under each setting 100 times, recording the linkage estimate $\hat{\bm{Z}}$, and recall and precision.
 
-In Figure \ref{fig:SEI_Z},  see that the largest number of distinct linkage estimates occurs when $S = 1$.
+In Figure \ref{fig:SEI_Z}, the largest number of distinct linkage estimates occurs when $S = 1$.
 This makes sense, because the SEI procedure arbitrarily removes large numbers of record labels from consideration, resulting in a noisier estimate of the linkage structure.
@@
 Setting $S=1$ can arbitrarily remove the index of a true match, leading the Gibbs sampler to concentrate probability on a false match, while larger values of $S$ produce results mirroring implementation with no SEI.
-We note however that even with $S=1$, the loss in precision is very small.
-
-Although the figures provided suggest that $S=2$ is adequate for maintaining linkage performance, we suggest a more conservative value like $S=10$.
+We note however that even with $S=1$, the loss in precision is small.
+
+Although the figures suggest that $S=2$ is adequate for maintaining linkage performance, we suggest a more conservative value like $S=10$.
 When evaluating the performance of a record linkage algorithm, researchers often examine posterior probabilities.
-By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce suspiciously high posterior probability for certain record pairs, providing a warped perception of model performance.
-We want to be sure that the set of records pairs exhibiting agreement patterns that have considerable probability of matching do not get reduced through SEI.}
-}
+By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce suspiciously high posterior probability for certain record pairs, providing a misleading perception of model performance.
+%We want to be sure that the set of records pairs exhibiting agreement patterns that have considerable probability of matching do not get reduced through SEI.
 
@@
 		\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
-		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation}
+		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation.}
 		\label{fig:SEI_eval}
@@
 
-	The country of El Salvador was immersed in civil war from 1980 to 1991, and we are interested in estimating the total number of casualties from the war.
+	The country of El Salvador was immersed in civil war from 1980 to 1991.
+We are interested in estimating the total number of casualties from the war.
 We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
@@
 	We link records using sex, date of birth, and location using the thresholds shown in Table \ref{Tab:nltcs-comparisons}.
-Storing three comparison scores for each of $20485 \times 17466 \approx 400,000,000$ record pairs would require approximately 8GB of memory.
+Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx 400,000,000$ record pairs would require approximately 8GB of memory.
 Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
-However, through the \texttt{fabl} framework, we compute comparisons over 30 smaller comparison tasks, hash results, and combine results before conducting linkage.
-Without SEI, the resulting data object is about 2.2 GB, and with SEI using $S=10$, the resulting data object is about 760 MB.
-Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.
+{\color{red}{However, through the method described in Section \ref{scaling},  we perform 30 smaller comparison tasks, using $t_A = 1$ and $t_B = 30$.
+We conduct linkage with all record indices recorded and also with SEI using $S=10$, and obtain identical results.
+The hashed $\tilde{\gamma}$ without SEI is about 2.2 GB, and with SEI, it is about 760 MB.
+Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.}}
 
@@
 
-	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing all $n_A \times n_B$ record pairs still can be prohibitive for larger linkage tasks.
+	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing all $n_A  n_B$ record pairs still can be prohibitive for larger linkage tasks.
 Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
@@
 Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task.
-In future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to arbitrarily large linkage tasks.
+In future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to even larger linkage tasks.
 
@@
 	\end{cases}\] 
-Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false non-match, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
+Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
 
@@
 	$$\hat{Z}_j =\begin{cases} 
-		i,  & \text{if } P(Z_j = i |\gamma) > \frac{1}{2}; \\
+		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
 		0,  & \text{otherwise}. \\
 	\end{cases}$$
-Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-In the event that we have two records $B_j$ and $B_{j'}$ such that both $P(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
+Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+In the event that we have two records $B_j$ and $B_{j'}$ such that both $p(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ p(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
 Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
@@
 For a more in-depth explanation of this function and the induced Bayes estimate, see \cite{sadinle_bayesian_2017}.
-
-%	\hypertarget{app:glossary}{%
-%	\subsection{Summary of Notation}\label{app:glossary}}
-%	
-%		\begin{table}[h!]
-%		\centering
-%		\begin{tabular}[h!]{ll}
-%			\hline
-%			Symbol & Description \\
-%			\hline
-%			$A, B$ & data files\\
-%			$i \in 1, \ldots, n_A $ & index over records in $A$\\
-%			$j \in 1, \ldots, n_B $ & index over records in $B$\\
-%			$f \in 1, \ldots F$ & index over fields used for comparisons \\
-%			$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
-%			$n_{AB}$ & number of entities in common between $A$ and $B$\\
-%			$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
-%			$Z_j = i$ & records $A_i$ and $B_j$ match \\
-%			$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
-%			$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
-%			$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-%			$\pi$ & probability that a record in $B$ has a match in $A$ \\
-%			$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
-%			$h_p$ & one hot encoding of agreement pattern $p$ \\
-%			$H_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
-%			$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
-%			$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
-%			$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
-%			\hline
-%		\end{tabular}
-%	\end{table}
 
@@
 \begin{align*}
-	p(m_{fl}| \gamma, \bm{Z}) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
-	&= m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
+	\mathcal{L}(m_{fl}|\gamma, \alpha_{fl}, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
 \end{align*}
 where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
-Analogous procedures lead to the posterior distribution $p(u_{fl}| \gamma, \bm{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
+Analogous procedures lead to the posterior distribution $\mathcal{L}(u_{fl}| \gamma, \beta_{fl}, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
 Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
 
-$$\bm{m}_f|\bm{Z}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
-$$\bm{u}_f|\bm{Z}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})).$$
-
-Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ (and hyperparameters).
-Thus, we use $p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
+$$\bm{m}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})),$$
+$$\bm{u}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).$$
+
+Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ and hyperparameters.
+We have the full conditional
 \begin{align*}
@@
 \end{align*}
-Thus $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-To express the full conditional for $\bm{Z}$, we consider the likelihood in two cases.
-Because we sample each $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
-Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+Thus, $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
+
+To express the full conditional for $\bm{Z}$, we consider the cases where $Z_j$ does not have a link in $A$ and where $Z_j$ does have a link in $A$.
+Because we sample each $Z_j$ independently of all other $Z_{j'}$ (for $j \neq j'$), we need only the full conditional for each individual $Z_j$.
+{\color{red}{Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.}} Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
 \end{align}
@@
 \begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)} \times \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
 \end{align}
@@
 \begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)} \times \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&= w_{qj} \times c_j.
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	&= w_{qj}  c_j.
 \end{align}
@@
 \begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j) = \begin{cases} 
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
 		w_{qj}, & q \leq n_A; \\
@@
 \end{align}
-Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain
+Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
 \begin{align}
 	\label{eqn:z_full_conditional2}
-	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
+	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
 	\begin{cases} 
-		\frac{\pi^{(s+1)}}{n_A} w_{q, j}^{(s+1)},  & q \leq n_A; \\
+		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
 		1 - \pi^{(s+1)}, & q  = n_A + j. \
@@
 %	p&\left(Z_j^{(s+1)} = q|\gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
-%	&\propto p(\gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
+%	&\propto p(\gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) p(Z_j^{(s+1)}|\pi) \\
 %	&\propto \left(\sum_{i=1}^{n_A}w_{q, j}I(q = i) + I(q = n_A + j) \right) \left (\pi\sum_{i=1}^{n_A}\frac{1}{n_A}I(q = i) + (1-\pi)I(q = n_A + j) \right) \\	
@@
 %	
-%	{\color{red}{Note that with this approach, each $r_{p_j}^{SEI}$ can contain up to $S \times t_A \times t_B$ indices, so we store at most  $n_B \times P \times S \times t_A \times t_B$ indices in total.}}
+%	{\color{red}{Note that with this approach, each $r_{p_j}^{SEI}$ can contain up to $S  t_A  t_B$ indices, so we store at most  $n_B  P  S  t_A  t_B$ indices in total.}}
 
@@
 
-	\hypertarget{app:ohe}{%
-		\subsection{One Hot Encoding Transformation}\label{app:ohe}}
-
-	As described in Section \ref{sec:efficiency}, \texttt{fabl} makes use of one-hot encodings to aid in vectorized computations.
+%	\hypertarget{app:ohe}{%
+%		\subsection{One Hot Encoding Transformation}\label{app:ohe}}
+%	
+%	As described in Section \ref{sec:efficiency}, \texttt{fabl} makes use of one-hot encodings to aid in vectorized computations.
 For $\gamma_{ij}^f$ with $L_f$ levels, define $e_{ij}^f$ to be an $L_f \times 1$ vector.
@@
 We then concatenate the $e_{ij}^f$ for all $f \in \{1, \ldots, F\}$, resulting in the one-hot encoded comparison vector $e_{ij}$ of length $\sum_{f=1}^F L_f$.
-
-	For example, consider comparing the toy records shown in Table \ref{tab:ohe} with $L = (3, 3, 2, 2)$ levels of agreement for last name, first name, DOB, and city respectively.
+%	
+%	For example, consider comparing the toy records shown in Table \ref{tab:ohe} with $L = (3, 3, 2, 2)$ levels of agreement for last name, first name, DOB, and city respectively.
 Since the first name differs by only one letter, a reasonable comparison vector for this pair would be $\gamma_{ij} = (1, 2, 1, 2)$.
 The one hot encoding representation of this vector is $e_{ij} = (1, 0, 0, 0, 1, 0, 1, 0, 0, 1)$.
-	\begin{table}[h!]
-		\centering
-		\begin{tabular}[h!]{llll}
-			Last Name & First Name & DOB & City \\
-			\hline
-			Smith & Taylor & 01/01/2000 & Durham\\
-			Smith & Tayler & 01/01/2000 & Raleigh\\
-			\hline
-		\end{tabular}
-		\caption{Example records for one hot encoding.}\label{tab:ohe}
-	\end{table}
+%	\begin{table}[h!]
+%		\centering
+%		\begin{tabular}[h!]{llll}
+%			Last Name & First Name & DOB & City \\
+%			\hline
+%			Smith & Taylor & 01/01/2000 & Durham\\
+%			Smith & Tayler & 01/01/2000 & Raleigh\\
+%			\hline
+%		\end{tabular}
+%		\caption{Example records for one hot encoding.}\label{tab:ohe}
+%	\end{table}
 
@@
 In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
-Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual non-matches.
+Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual nonmatches.
 Mathematically, $\text{NPV} = \sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j = n_A + j)$/$\sum_{j=1}^{n_B} I(\hat{Z}_j = n_A + j)$.
@@
 Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probabilities of matches at times can be split across more records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand than would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand than would be left under \texttt{BRL}, but the decisions made by each method should have nearly equal accuracy.
 
 %In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-%This is reasonable because large numbers of errors blur the distinction between matched and non-matched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
+If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
+
+%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
 
@@
 		\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
-	Below are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
 It is set up with one error across the linkage fields and 50 duplicates across files.
@@
 	\subsection{Additional Speed Simulation Study}\label{app:appendix-speed}}
-	To illustrate that different constructions of the comparison vectors lead to similar speed gains, we replicate the speed study of Section \ref{speed} under different settings.
+	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
+We replicate the speed study of Section \ref{speed} under different settings.
 Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex
updates to response to reviewers for brian to look at.

@@
 
-\newcommand{\ncomment}[1]{{\color{teal} NM: #1}}
+\newcommand{\beka}[1]{{\color{teal}}}
 
@@
 \pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}
+
+\textcolor{teal}{Suggestion: Write a strong paragraph to the Editor after the paper is finalized.
+Come back to this.
+Explain that we have tried to simplify the notation when possible, however, in some places, the heavy notation is not possible.
+Thus, we have worked on improving more precise wording and removing sections or wording that is distracting or not needed.
+Thank the Editor and Associate Editor for helping us improve upon what we believe to be an important paper and breakthrough in the literature.}
+
+\textcolor{teal}{I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.}x
 
@@
 
-\reply{We added this sentence in an attempt to address a previous reviewer comment pointing out a conflation between the record itself and the index directing towards that record.
-With the simplified reference to the two datasets as $A$ and $B$, that material now reads:
+\textcolor{teal}{Thank the Editor, explain the change. I have removed repetition and shortened this bit.}
+
+\reply{
+%We added this sentence in an attempt to address a previous reviewer comment pointing out a conflation between the record itself and the index directing towards that record.
+\textcolor{teal}{We thank the Editor for bringing up this point. In order to simplify notation, we refer to the two data files as} 
+ $A$ and $B$, where the updated material reads:
 
@@
 
+\textcolor{teal}{Don't apologize!}
+
+\reply{%
+	\textcolor{teal}{The square brackets have been corrected.
+We have clarified the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
+}
+}
+
 \reply{%
@@
 I had assumed that the authors had integrated out pi in the prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite matchings”.
-Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution?
-Is the presented alternative approach justified?}
+(1) Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution?
+(2) Is the presented alternative approach justified?}
 
-\reply{%
-	We realize that the integration of pi out of the full conditionals is not necessary for our sampler.
+\textcolor{teal}{Brian: I think it's very important that we answer this carefully and in a strong manner.
+See below.
+Note: The edits in Appendix 8.2 regarding the justification that I wrote have been removed.}
+
+\textcolor{teal}{No, we do not integrate out $\pi$ as done in Sadinle as this would result in a sequential sampler as opposed to the parallel one that we propose.
+This is one of our contributions to our paper.
+(1) Yes, our alternative approach is different than directly integrating over the prior distribution.
+We have attempted to make this more clear in our revised version.
+In short, our prior distribution over $Z$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of Sadinle (2018).
+Of course, we could integrate out $\pi$ in our approach, however, this would lead to sequential updates, which would be slow in practice.
+(2) Yes, our alternative approach is justified.
+We provide a justification in Appendix 8.2 regarding why the updates are independent and not sequential.
+Thank you for the excellent questions, which have greatly improved our paper.
+}
+
+%
+%\reply{%
+%	We realize that the integration of pi out of the full conditionals is not necessary for our sampler.
 For clarity, we decided to redo the simulations using a Gibbs sampler that samples $\pi$ from its full conditional, and then samples $Z$ from its full conditionals.
 We obtained equivalent results, which is unsurprising, because with high numbers of observations, the posterior of the beta distribution is highly concentrated around its mean.
-
-	Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler. We leave the remark about marginalization in the Appendix to facilitate comparisons to Sadinle's method.
-
-}
+%	
+%	Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler. We leave the remark about marginalization in the Appendix to facilitate comparisons to Sadinle's method.
+%
+%}
 

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
add in clarification to AE regarding parallel versus sequential updates in appendix 8.2 for brian to look at.

@@
 
+\textcolor{blue}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
+\begin{align}
+p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi) \propto \prod_q \left(\prod_j
+p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi \right)
+\right).
+\end{align}
+Therefore, we update the full conditional for $\bm{Z}^{(s+1)}$ in parallel.
+This contrasts the Gibbs sampler of  \cite{sadinle_bayesian_2017}, where the update to the matching matrix is sequential and accomplished using a partially collapsed Gibbs sampler, where one integrates out $\pi.$ We do not opt for this approach as it would lead to similar sequential updates, which are computationally slow.}
+
 %Let $\gamma_{.j}$ denote the set of $n_A$ comparison vectors with $B_j$. We have

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
formatted appendix in A,B,C, etc fashion.

@@
 
-	\section{Appendix}
-	\label{sec:appendix}
+%	\section{Appendix}
+%	\label{sec:appendix}
+
+\appendix
 
 	\hypertarget{bayes-estimate}{%
-		\subsection{Bayes Estimate}
+		\section{Bayes Estimate}
 		\label{bayes-estimate}}
@@
 	\hypertarget{app:derivations}{%
-	\subsection{Derivations of Full Conditionals}\label{app:derivations}}
+	\section{Derivations of Full Conditionals}\label{app:derivations}}
 
@@
 	\hypertarget{partial}{%
-	\subsection{Accuracy under Partial Estimates}\label{partial}}
+	\section{Accuracy under Partial Estimates}\label{partial}}
 
@@
 	\hypertarget{appendix-sim}{%
-		\subsection{Traceplots for Simulation Study}\label{app:appendix-sim}}
+		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
 	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
@@
 	\hypertarget{appendix-speed}{%
-	\subsection{Additional Speed Simulation Study}\label{app:appendix-speed}}
+	\section{Additional Speed Simulation Study}\label{app:appendix-speed}}
 	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
@@
 	\hypertarget{appendix-es}{%
-		\subsection{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
+		\section{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
 

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex
working on response to reviewers.

@@
 
-\textcolor{teal}{I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.}x
+\textcolor{teal}{I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.}
 
-\reply{We have made several efforts to simplify notation in this revision.
+\reply{
+We thank the Editor for the opportunity to revise our paper and for the suggestions.
+We have improved notation, when possible, however, as noted by the Editor in certain places, such as in the computational speeds ups of Section 4, this is difficult to avoid.
+We have done our utmost to simplify notation (when possible) and improve the parsing of sentences.
+Finally, we have attempted to make it more clear the general contributions of our paper, which we highlight for the Editor below.
 
-\begin{itemize}
-	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. This particularly simplifies the notation in Section 4.2.
+We propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
+{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
+We employ the technique of \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
+This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
+In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
+}
 
-	\item At the end of Section 4.1, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics.
-	This likelihood is then used throughout Section 4.2 in the discussion of efficient posterior inference.
+%\reply{We have made several efforts to simplify notation in this revision.
+%
+%\begin{itemize}
+%	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. This particularly simplifies the notation in Section 4.2.
+%	
+%	\item At the end of Section 4.1, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics.
+%	This likelihood is then used throughout Section 4.2 in the discussion of efficient posterior inference.
 We hope this expression clarifies the meaning of "preservation of weight (Comment 5), and strengthens the argument for $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ being summary statistics for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ (Comment 4).
-
-	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$.
-
-	\item In Section 4.1, we have provided a specific example of a comparison vector and how the hashing procedure works.
-
-	\item Section 4.1 introduces the summary statistic $N_{j_p}$.
+%	
+%	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$.
+%	
+%	\item In Section 4.1, we have provided a specific example of a comparison vector and how the hashing procedure works. 
+%	
+%	\item Section 4.1 introduces the summary statistic $N_{j_p}$.
 After careful review, we realized that by changing the order of subscripts and instead using $N_{p_j}$, we can replace the sum of counts across all records (previously denoted $H_p$), with the more intuitive $N_p$.
 Thus, we have removed all of the ''$H$" quantities throughout Section 4.
-\end{itemize}
-}
+%\end{itemize}
+%}
 
@@
 
+	\clearpage
+
+	\bigskip
+
+	\bibliographystyle{jasa}
+	\bibliography{biblio}
+
 \end{document}

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
working on response to reviewers.

@@
 \begin{keyword}%[class=MSC]
-\kwd{data fusion}
-\kwd{distributed computing}
+\kwd{bipartite record linkage}
 \kwd{entity resolution}
+\kwd{data cleaning}
+\kwd{parallelization}
 \kwd{hashing}
-\kwd{record linkage}
 \end{keyword}
@@
 The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity.
-However, In many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
+However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
-To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
+To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}.
+For example, the decision theoretic work of \cite{steorts_bayesian_2016}, ensures that the resulting records are bipartite (or rather transitive).}
+
+%but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
 
@@
 In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
+We employ the technique of \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
+This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
attempting to clarify some of brian's questions.

@@
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
-To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}.
-For example, the decision theoretic work of \cite{steorts_bayesian_2016}, ensures that the resulting records are bipartite (or rather transitive).}
+To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}
 
@@
 {\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the technique of \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
+We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
 This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
@@
 This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
-This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
-
-%\section{Review of Prior Work}
-%\label{sec:review-of_prior-work}
-%
-%%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.
-%
-%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
-Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
-Without loss of generality, denote files such that $n_A \geq n_B$.
-Under bipartite matching, the set of matches across data files can be represented in two equivalent ways.
-First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
-%\begin{align}
-%	\Delta_{ij} =
-%	\begin{cases}
-%		1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
-%		0, & \text{otherwise}.\\
-%	\end{cases}
-%\end{align}
-%This sparse matrix representation can become cumbersome for large linkage tasks.
-More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
-%\begin{align}
-%	Z_{j} =
-%	\begin{cases}
-%		i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
-%		n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
-%	\end{cases}
-%\end{align}
-%We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise. 
-%
-
-%Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of nonmatches by 
-%$\bm{U} =  \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as nonmatches as ``non-links.''
-%
-%%\subsection{Comparison Vectors}
-%%\label{comparison-vectors}
-%
-
-%
-%Intuitively, matching records (those that refer to the same entity) should be similar; records that are nonmatching should be dissimilar.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ {\color{red}{We use $\Gamma_{ij}$ when referring to the vector as a random variable, and $\gamma_{ij}$ when referring to observed data.
-We collect the vectors $\gamma_{ij}$ as the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$}}.
-%
-%%Frequently, $\gamma_{ij}$ consists of one comparison for each feature shared between the two data files. 
-%
-%The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features.
-For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
-For numerical features, we can use absolute difference between the two feature values.
-For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
-For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
-We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
-%
-%More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
-Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
-Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
-%
-%{\color{red}{
-%In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
-As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
-To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
-}
-%}
-%
-%%For more complex measurements, we can take into account partial agreement to more richly characterize the comparison;  
-%
-%%
-%%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking.
-This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question.
-In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.
-%
-%\subsection{Fellegi-Sunter Models}
-%\label{fellegi-sunter}
-%
-%The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
-Using these assumptions, \cite{winkler_state_1999}, \cite{jaro1989}, \cite{larsen_2001}, \cite{enamorado2019using} and others model the comparison data through mixture models of the form
-%\begin{subequations}
-%\begin{align}
-%	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
-%	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
-%	&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
-%\end{align}
-%\end{subequations}
-%Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
-When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
-Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
-The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
-%
-%Within this framework, record pairs are independently classified as matches and nonmatches based on the estimated parameters.
-However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results.
-To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching.
-\cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.
-%
-%%This prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi-Sunter weights among matched pairs.
-%
-%%These unknown parameters are often estimated using the EM algorithm \cite{winkler_state_1999}. 
-%
-%%The Fellegi-Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
-This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.
-%
-%%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
-%%\begin{align}
-%%	\label{eqn:jaro}
-%%	&\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij} 
-%%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
-%%	& \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
-%%	& \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
-%%\end{align}
-%%
-%%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs.
-Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
-%
-%%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
-This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
-%%\begin{align}
-%%	\label{eqn:wts}
-%%	w_{ij} = \frac{p(\gamma_{ij} \mid \Delta_{ij} = 1)}{p(\gamma_{ij} \mid \Delta_{ij} = 0)}
-%%\end{align}
-%%as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
-%%
-%%In practice, $p(\cdot \mid \Delta_{ij} = 1)$ and $p(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
-%%
-%%%is the set of parameters $u_{fl} = p(\gamma_{ij}^f = l|Z_j \neq i)$ for the nonmatching pairs.
-%%
-%%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
-This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
-%
-%%\subsection{Beta Record Linkage Model}
-%%\label{BRL}
-%
-%%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
-%
-%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework.
-In addition to {\color{red}{uniform}} Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
-He assigns a prior distribution for the probability of the indicator that a record in $B$ has a match in $A$, so that $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
-It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
-Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A !
-/ (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
-Thus, the prior is given by
-%\begin{align}
-%\label{eqn:sadinle_prior}
-%p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
-%\end{align}
-%where $\text{B}(\cdot, \cdot)$ represents the Beta function.
-This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
-This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
+This makes the sampler inherently \textcolor{teal}{sequential}, which can be slow when working on linkage tasks with more than a few thousand records.
 

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex
working on comments for brian.

@@
 
+	\textcolor{teal}{See two comments below regarding MAR versus MCAR.}
+
+	\textcolor{teal}{Answer for MAR: Yes, the assumption is truly missing at random.
+This distinction between truly missing at random and missing completely at random does not matter due to two assumptions.
+Following Sadinle (2017), we assume that the fields in the comparison vectors are conditionally independent given the coreference matrix and we assume that the missing comparisons are ignorable (Sadinle (2017).
+Therefore, our likelihood is the same as in Sadinle (2017) (see equations (3) and (4) of his paper).
+As stated by Sadinle (2017), ``we work under the assumption of ignorability of the missingness mechanism for the comparisons so that we can base our inferences on the marginal distribution of the observed comparisons (Little and Rubin, p.
+90)."  We refer to Sadinle (2017), Section 4.2, page 604 for further details.}
+
+	\textcolor{teal}{Answer for MCAR: The assumption is missing completely at random due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix.
+The assumption is not missing at random as we do not assume that the missing comparisons are ignorable.}
+
 	\reply{% 
@@
 
+	\textcolor{teal}{Comments: Tell the AE they are correct and mention where you correct the confusion.
+Then move on.
+(No need to state what you intended).
+Mention what equations numbers have been updated to help the AE now given the major revision.
+\textcolor{red}{Please redo this one to answer the question more directly.}}
+
 	\reply{%
@@
 
+	\textcolor{teal}{Comment: Make it more brief and no need to provide notation when it's not needed.}
+
+	\textcolor{teal}{Thank for pointing out our typo. We have revised the text to state ``summary statistics" instead of ``sufficient statistics."}
+
 	\reply{%
@@
 Which records are appropriately weighted – those in X2?}
+
+	\textcolor{teal}{Comment: Brian, you have not answered the reviewer's question regarding what is meant by weight and what records are appropriately weighted.
+\textcolor{red}{Please revise to directly answer the question in the revised text with equation numbers.}}
+
+	\textcolor{teal}{
+	We have revised the ordering of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI.
+With this new ordering, and with the newly provided equations 15 -19, we hope that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$ is now more clear.
+	}
 
@@
 
-\reply{%
-	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
+	\reply{
+
+	\textcolor{teal}{``Those comparison vectors" referred to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$ }
+
+	The revised text states: ``\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison vectors} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove} the \textcolor{teal}{memory-intensive} \textcolor{teal}{comparison vectors} $\gamma^{ab}$ before continuing with the next chunk of data.
+See Section 4.3, page 10.
 	}
+
+%\reply{%
+%	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
+%}
 

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
working on comments for brian.

@@
 \begin{abstract} 
-%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
 Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.
@@
 Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
-%
-%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.
-Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}
-
 \end{abstract}
@@
 %% ** Keywords **
-\begin{keyword}%[class=MSC]
+\begin{keyword}
 \kwd{bipartite record linkage}
@@
 
-% Possible citation for bipartite matching. 
-%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_A995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
-
 The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity.
@@
 To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}
-
-%but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
 
@@
 
-%Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space.
-
 In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
@@
 Without loss of generality, denote files such that $n_A \geq n_B$.
-
-%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
-Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
 
@@
 
-%\subsection{Fellegi-Sunter Models}
-%\label{fellegi-sunter}
-
 Having defined comparison vectors, we now turn to the \cite{fellegi_theory_1969} model for record linkage.
@@
 The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
-
-%Within this framework, record pairs are independently classified as matches and nonmatches based on the estimated parameters.
-However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results.
-To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching.
-\cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.
-
-%The Fellegi-Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
-This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.
-
-%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
-%\begin{align}
-%     \label{eqn:jaro}
-%     &\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij}
-%     \quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
-%     & \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
-%     & \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
-%\end{align}
-%
-%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs.
-Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
-
-%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
-This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
-%\begin{align}
-%     \label{eqn:wts}
-%     w_{ij} = \frac{p(\gamma_{ij} \mid \Delta_{ij} = 1)}{p(\gamma_{ij} \mid \Delta_{ij} = 0)}
-%\end{align}
-%as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
-%
-%In practice, $p(\cdot \mid \Delta_{ij} = 1)$ and $p(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
-%
-%%is the set of parameters $u_{fl} = p(\gamma_{ij}^f = l|Z_j \neq i)$ for the nonmatching pairs.
-%
-%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
-This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
-
-%\subsection{Beta Record Linkage Model}
-%\label{BRL}
-
-%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
-
-%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a
 
@@
 
-%\begin{table}[t!]
-%	\centering
-%	\begin{tabular}[t!]{ll}
-%		Symbol & Description \\
-%		\hline
-%		$A, B$ & data files\\
-%		$i \in 1, \ldots, n_A $ & index over records in $A$\\
-%		$j \in 1, \ldots, n_B $ & index over records in $B$\\
-%		$f \in 1, \ldots F$ & index over fields used for comparisons \\
-%		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
-%		$n_{AB}$ & number of matches between $A$ and $B$\\
-%		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
-%		$Z_j = i$ & records $A_i$ and $B_j$ match \\
-%		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
-%		$m_{fl}$ & $p(\gamma_{ij}^f = l | Z_j = i)$ \\
-%		$u_{fl}$ & $p(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-%		$\pi$ & probability that a record in $B$ has a match in $A$ \\
-%		\hline
-%	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
-%\end{table}
-%
-%For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_A}.
 \begin{subequations}
@@
 
-%\hypertarget{posterior-sampling}{%
-%	\subsection{Gibbs Sampler}
-%	\label{gibbs_sampling}}
-
 We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler.
@@
 	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
 \end{align}
@@
 \end{align}
-%\begin{align}
-%\label{eqn:z_full_conditional}
-%p\left(Z_j^{(s+1)}  = z_j|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
-%\begin{cases} 
-%	w_{z_j, j}^{(s+1)},  & z_j \leq n_A; \\
-%	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
-%\end{cases}
-%\end{align}
+
 where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
@@
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
-%For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.
-
-%\begin{table}[t!]
-%	\centering
-%	\begin{tabular}[t!]{ll}
-%		Symbol & Description \\
-%		\hline
-%		$P$ & number of agreement patterns exhibited in $\gamma$ \\
-%		$h_p$ & agreement pattern $p$ \\
-%		$e(h_p)$ & one-hot encoding of agreement pattern $p$ \\
-%		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $B_j$ exhibits pattern $p$ \\
-%		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
-%		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
-%		$N_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
-%		$N_p$ & number of total comparison vectors that exhibit pattern $p$ \\
-%
-%		\hline
-%	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_B}
-%\end{table}
 
@@
 This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
-%A more complete example of the one-hot encoding is provided in Appendix \ref{app:ohe}.
 
@@
 For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
-We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
+
+		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison vectors} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
+		the 
+\textcolor{teal}{memory-intensive}
+		\textcolor{teal}{comparison vectors} $\gamma^{ab}$ 
+		before continuing with the next chunk of data. 
 		In detail, we calculate
@@
 Since we have five binary comparison fields with no missingness, the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
-
-%	We simulate these data for different values of $n_A$ and $n_B$, always with $n_B/2$ records in common across data files.
-and compare the run-time of \texttt{fabl} against \texttt{BRL}.
-Since we have 5 binary comparison fields, the number is unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
-
-	 %\textcolor{blue}{DELETE?
-The $\bm{u}$ parameters are chosen to emulate what is seen in practice for first name, last name, day of birth, and month of birth.
-The parameters for year would vary largely by context, so we choose them here to adjust the desired difficulty of the linkage task.}
 
@@
 By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce suspiciously high posterior probability for certain record pairs, providing a misleading perception of model performance.
-%We want to be sure that the set of records pairs exhibiting agreement patterns that have considerable probability of matching do not get reduced through SEI.
-
-%\begin{figure}[t]
-%	\begin{center}
-%		\includegraphics[width=0.6\textwidth]{../SEI_sensitivity/Z_plot}
-%		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
-%		\label{fig:SEI_Z}
-%	\end{center}
-%\end{figure}
-%
-%\begin{figure}[t]
-%	\begin{center}
-%		\includegraphics[width=0.6\textwidth]{../SEI_sensitivity/eval_plot} 
-%		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation}
-%		\label{fig:SEI_eval}
-%	\end{center}
-%\end{figure}
-%
-%  \begin{figure}[hbt]
-%	\centering
-%	\begin{subfigure}{0.4\textwidth}
-%		\includegraphics[width=\textwidth]{../SEI_sensitivity/Z_plot} 
-%		\caption{}\label{fig:SEI-Z}
-%	\end{subfigure}\hspace{\columnsep}%
-%	\begin{subfigure}{0.4\textwidth}
-%		\includegraphics[width=\textwidth]{../SEI_sensitivity/eval_plot} 
-%		\caption{}\label{fig:SEI-eval}
-%	\end{subfigure}
-%\caption{Simulation results Section \ref{SEI-sensitivity}. In \ref{fig:SEI-Z}, 
-%\end{figure}
-%Figure \ref{fig:here}
 
@@
 
-	%We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Appendix \ref{partial}.
-Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
-This means that after clerical review, the number of individuals replicated across data files would fall in the interval (136, 311), encapsulating the posterior credible interval.
-More or fewer records could be identified for clerical review by decreasing or increasing $\theta_R$.
-
-	%and a range of (140, 294) after the partial estimate and clerical review.
-
 	\begin{figure}[t]
@@
 
-	%\hypertarget{violations}{%
-		%	\subsection{Violations of One-to-One Matching}\label{violations}}
-	%
-	%As noted previously, the application exhibits many situations in which bipartite matching is difficult.
-To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
-These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $A$ has
-	%multiple plausible matches in $B$. 
-	%
-	%Table \ref{Tab:rosa-maria} illustrates an example.
-Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
-In addition, using the modified Levenstein distance, the comparison vectors $\gamma_{2776, 825}$ and $\gamma_{2776, 826}$ are exactly identical.
-	%
-	%\begin{table}
-	%	\centering
-	%	\begin{tabular}[h!]{llllllll}
-		%		\hline
-		%		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
-		%		\hline
-		%		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
-		%		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
-		%		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
-		%		\hline
-		%	\end{tabular}
-	%	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
-	%\end{table}
-	%
-	%Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
-Under \texttt{fabl}, both records in $B$ match to the same record in $A$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
-Under \texttt{BRL}, the Gibbs process creates one matching configuration and stays there for a while.
-However, if one pair ``unmatches,'' then the other record has a chance to latch on.
-Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
-	%
-	%Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
-\texttt{BRL} in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
-Thus relaxing the one-to-one constraint has served as a useful tool for identifying model misspecification.
-	%
-	%\begin{figure}[h!]
-	%\begin{center}
-	%\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-	%\caption{Gibbs sampling in situation with multiple plausible matches.
-Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability.
-Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
-	%\end{center}
-	%\end{figure}
-
 	\section{Conclusion}
@@
 	\clearpage
-
-%	\section{Appendix}
-%	\label{sec:appendix}
 

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex
more suggested comments that go with brian's revisions.

--- 2023-07-17 ba_resubmit/response_to_reviewers2.tex
@@
 
-Thank you for the thoughtful feedback. We have made several broad changes to address the Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.
+\textcolor{blue}{See suggestions below in teal color.}
+
+Thank you for the thoughtful feedback.
+We have made several broad changes to address \textcolor{teal}{the Editor's and Associate Editor's} overall critiques, and have addressed each specific comment from the reviewer.
 
@@
 
-\textcolor{teal}{Suggestion: Write a strong paragraph to the Editor after the paper is finalized.
+%% RCS: Suggestion: Write a strong paragraph to the Editor after the paper is finalized.
 Come back to this.
@@
 Thus, we have worked on improving more precise wording and removing sections or wording that is distracting or not needed.
-Thank the Editor and Associate Editor for helping us improve upon what we believe to be an important paper and breakthrough in the literature.}
+Thank the Editor and Associate Editor for helping us improve upon what we believe to be an important paper and breakthrough in the literature.
 
-\textcolor{teal}{I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.}
+%% I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.
 
@@
 We have done our utmost to simplify notation (when possible) and improve the parsing of sentences.
-Finally, we have attempted to make it more clear the general contributions of our paper, which we highlight for the Editor below.
+Finally, we have attempted to clarify our overall contributions to the literature, which we highlight below.
 
 We propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the technique of \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
-This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.
+This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
+We employ the decision theoretic  technique of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.
+This discovery allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
 In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
+Finally, we provide two real illustrations of our methodology and simulation studies with reproducible software.
 }
@@
 \pointRaised{AE}{Line 80 on p.
-3 says: “For ease of readability, we follow the convention established by Sadinle (2017) and say “record $i \in X_1$” rather than the more compact $x_{1i}$.” Taking this at face value implies that, in any sentence, one could swap “record $i \in X_1$” for “$x_{1i}$,” which clearly is not true.
+3 says: “For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say “record $i \in X_1$” rather than the more compact $x_{1i}$.” Taking this at face value implies that, in any sentence, one could swap “record $i \in X_1$” for “$x_{1i}$,” which clearly is not true.
 So, as it stands, the sentence does little to improve ease of readability.
@@
 
-\textcolor{teal}{Thank the Editor, explain the change. I have removed repetition and shortened this bit.}
+%%% RCS: Thank the Editor, explain the change. I have removed repetition and shortened this bit.
 
 \reply{
-%We added this sentence in an attempt to address a previous reviewer comment pointing out a conflation between the record itself and the index directing towards that record.
-\textcolor{teal}{We thank the Editor for bringing up this point. In order to simplify notation, we refer to the two data files as} 
+We thank the Editor for bringing up this point. In order to simplify notation, we refer to the two data files as
  $A$ and $B$, where the updated material reads:
 
-	``Consider a data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$".
+	``Consider a data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$". See Section 2, page 3. 
 }
@@
 
-	\reply{% 
-		The assumption is missing completely at random. You are correct that this is a result of the independence assumed across elements of the comparison vector. This has been clarified.
-	}
+	%% RCS: See two comments below regarding MAR versus MCAR. I have an alternative response for the one regarding miss at random.
+
+%	\textcolor{teal}{Answer for MAR: Yes, the assumption is truly missing at random.
+This distinction between truly missing at random and missing completely at random does not matter due to two assumptions.
+Following Sadinle (2017), we assume that the fields in the comparison vectors are conditionally independent given the coreference matrix and we assume that the missing comparisons are ignorable (Sadinle (2017).
+Therefore, our likelihood is the same as in Sadinle (2017) (see equations (3) and (4) of his paper).
+As stated by Sadinle (2017), ``we work under the assumption of ignorability of the missingness mechanism for the comparisons so that we can base our inferences on the marginal distribution of the observed comparisons (Little and Rubin, p.
+90)."  We refer to Sadinle (2017), Section 4.2, page 604 for further details.}
+
+	\reply{The assumption is missing completely at random (MCAR) due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix (or matching label) .
+The assumption is not missing at random (MAR) as we do not assume that the missing comparisons are also ignorable.
+See Section 2, page 2 regarding clarification of this point \citep{LittleRubin2002}.}
+
+%	\reply{% 
+%		The assumption is missing completely at random. You are correct that this is a result of the independence assumed across elements of the comparison vector. This has been clarified. 
+%	}
 
@@
 
-\textcolor{teal}{Don't apologize!}
+%% Don't apologize. Just answer the questions very completely.
+
+%\textcolor{teal}{Don't apologize!}
 
 \reply{%
-	\textcolor{teal}{The square brackets have been corrected.
+	The square brackets have been corrected.
 We have clarified the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-}
+Refer to Appendix B, page 26 -- 27 regarding this extended derivation.
 	}
 
-\reply{%
-	The square brackets have been corrected, and we apologize for the oversight.
+%\reply{%
+%	The square brackets have been corrected, and we apologize for the oversight.
 We have clarified derivation of the likelihood for $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-}
+%}
 
@@
 
-\textcolor{teal}{Brian: I think it's very important that we answer this carefully and in a strong manner.
-See below.
-Note: The edits in Appendix 8.2 regarding the justification that I wrote have been removed.}
+\reply{No, we do not integrate out $\pi$ as done in \cite{sadinle_bayesian_2017} as this would result in a sequential sampler as opposed to the parallel one that we propose.
+This is one of our contributions to our paper.
 
-\textcolor{teal}{No, we do not integrate out $\pi$ as done in Sadinle as this would result in a sequential sampler as opposed to the parallel one that we propose.
-This is one of our contributions to our paper.
-(1) Yes, our alternative approach is different than directly integrating over the prior distribution.
-We have attempted to make this more clear in our revised version.
+(1) Yes, our alternative approach is different than directly integrating over the prior distribution $\pi.$ We have attempted to make this more clear in our revised version.
 In short, our prior distribution over $Z$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of Sadinle (2018).
 Of course, we could integrate out $\pi$ in our approach, however, this would lead to sequential updates, which would be slow in practice.
+See Appendix B, page 27.
+
 (2) Yes, our alternative approach is justified.
-We provide a justification in Appendix 8.2 regarding why the updates are independent and not sequential.
+We provide a justification in Appendix B, page 27 regarding why the updates are independent and not sequential.
 Thank you for the excellent questions, which have greatly improved our paper.
+
 }
@@
 
-	\reply{%
-		We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
+	%%Comment: Make it more brief and no need to provide notation when it's not needed.
+
+	\reply{
+	Thank for pointing out our typo. We have revised the text to state ``summary statistics" instead of ``sufficient statistics."
+	}
+
+%	\reply{%
+%		We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
 However, we cannot express the full likelihood in (14) through these statistics.
@@
 Which records are appropriately weighted – those in X2?}
+
+	\textcolor{teal}{Comment: Brian, you have not answered the reviewer's question regarding what is meant by weight and what records are appropriately weighted.
+\textcolor{red}{Please revise to directly answer the question in the revised text with equation numbers.}}
+
+%	\textcolor{teal}{
+%	We have revised the ordering of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI.
+With this new ordering, and with the newly provided equations 15 -19, we hope that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$ is now more clear.
+%	}
 
@@
 
-\reply{%
-	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
+	\reply{
+
+	``Those comparison vectors" referr to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$
+
+	The revised text states: ``Then, we conduct hashing, obtain the compressed comparison vectors $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison vectors $\gamma^{ab}$ before continuing with the next chunk of data.
+See Section 4.3, page 10.
 	}
+
+%\reply{%
+%	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
+%}
 
@@
 I recommend that the authors either refrain from suppressing the SEI notation or further explain how the SEI algorithm has changed the quantities in these equations.}
+
+	\textcolor{red}{I'm checking over this. This probably needs to be explained in more detail. Go check very carefully.}
 
@@
 \reply{% 
-	We have provided an additional simulation study in Section 5.3 to show the sensitivity to different choices of $S$.
+	Refer to Section 5.3, which explores the trade-offs regarding different choices of $S$ and discusses this in practical terms. 
 }
@@
 \reply{%
-	This has been corrected, thank you.
+	This has been corrected; thank you.
 }
@@
 \reply{%
-	This has been revised, thank you
+	This has been revised; thank you.
 }

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex
still working on comments.

@@
 
-\textcolor{blue}{See suggestions below in teal color.}
-
-Thank you for the thoughtful feedback.
-We have made several broad changes to address \textcolor{teal}{the Editor's and Associate Editor's} overall critiques, and have addressed each specific comment from the reviewer.
+Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.
 
@@
 
+\newpage
 \pointRaised{R}{%
@@
 
-	\textcolor{teal}{Comments: Tell the AE they are correct and mention where you correct the confusion.
-Then move on.
-(No need to state what you intended).
-Mention what equations numbers have been updated to help the AE now given the major revision.
-\textcolor{red}{Please redo this one to answer the question more directly.}}
+	%% RCS: Tried to answer this from the original paper and the updated paper. Tried to make it shorter and close it out very quickly.
 
-	\reply{%
-	You are correct that the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$.
+	\reply{Yes, in the original manuscript, the comparison vectors are created and therefore, patterns are assigned) without regard to $Z$.
+Given the confusion caused to the Associate Editor, we removed the confusing sentence.	}
+
+%	\reply{%
+%	You are correct that the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$.
 This notation was created purely to be able to denote when a record was left unlinked during the Gibbs Sampler as in Equation 16 in the submitted draft.
-
-	I can see why this is confusing. Therefore, we have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
-	}
+%	
+%	I can see why this is confusing. Therefore, we have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
+%	}
 
@@
 
-	\textcolor{teal}{Comment: Brian, you have not answered the reviewer's question regarding what is meant by weight and what records are appropriately weighted.
-\textcolor{red}{Please revise to directly answer the question in the revised text with equation numbers.}}
+		%% Comment: Brian, you have not answered the reviewer's question regarding what is meant by weight and what records are appropriately weighted.
+\textcolor{red}{Please revise to directly answer the question in the revised text with equation numbers.
+To my understanding, the weights have been removed and section 4 has been rewritten }}
+	\reply{
+	\textcolor{teal}{In the original manuscript, we consider all records in the first file ($X_{1}$) that share an agreement pattern with the second data file ($X_{2}$).
+In this situation, all these records have the same Fellegi-Sunter likelihood ratio weight, $w_{ij}$.
+Given this result, we denoted the weight $w_p$ since these records map to the same agreement pattern.}
+
+	\textcolor{teal}{In the revised manuscript, we define $m_p$ and $u_p,$ which are the 
+	 probabilities that records $A_i$ and $B_j$ form agreement pattern $p$ given that
+ they are a match and non-match, respectively.
+For each pattern $p,$ we define $w_p = m_p/u_p.$ We hope that this will be more clear than our previous presentation of the material.
+See Section 4.1, pages 8--9.}
+ }
 
@@
 
-\reply{%
-
-	We have switched the order of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI.
+%\reply{%
+%	
+%	We have switched the order of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI.
 With this new ordering, and with the newly provided equations (15 -19) it is clearer that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
-}
+%}
 
@@
 
-	``Those comparison vectors" referr to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$
+	``Those comparison vectors" refer to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$
 

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex
typo.

@@
 	\reply{Yes, in the original manuscript, the comparison vectors are created and therefore, patterns are assigned) without regard to $Z$.
-Given the confusion caused to the Associate Editor, we removed the confusing sentence.	}
+Given the confusion caused to the Associate Editor, we removed the sentence.	}
 
@@
 
-	\textcolor{red}{I'm checking over this. This probably needs to be explained in more detail. Go check very carefully.}
+\reply{
+	\textcolor{teal}{In order to attempt to make the SEI algorithm more clear, we have revised the ordering of Sections 4.2 and 4.3, so that 1) hashing, 2) posterior inference, 3) chunkwise computation of the comparison matrix, and 4) SEI.
+With this new ordering, and with the newly provided equations 15 -19, we hope that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$ is now more clear.
+In addition, we hope that the SEI algorithm updates will now be more clear that the updates in (16a), (16b), and (18) only depend on $\mathcal{N}$.
+The SEI algorithm affects the step shown in (19).
+	}
+	}
 
-\reply{%
-	We have reordered Section 4 so that all posterior inference is presented before SEI.
+%\reply{%
+%	We have reordered Section 4 so that all posterior inference is presented before SEI.
 This makes it clear that the posterior updates in (16a), (16b), and (18) depend only on $\mathcal{N}$.
 SEI only affects the step shown in (19).
-}
+%}
 

### 2023-07-18 ba_resubmit/kundinger_resubmit2.tex
proposed edits for jerry/brian to look at once finished.

@@
 To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
-Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
+Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
+\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
+\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
+Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
+}
+%This allows
+%us to use hashing techniques that reduce computational overhead.
 This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
@@
 
-In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
-As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. 
 We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
@@
 Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta prior as follows.
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$,
+\textcolor{teal}{we use propose an alternative, called the fast beta linkage prior (fabl).}
+%the fast Beta prior as follows. 
 For each $Z_j$, we have
@@
 
-In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
-This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
-In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates.
-Hence, we prefer (\ref{eqn:fast_beta_prior}).
-We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
-
-Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
-In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
-In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
-As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
-
-To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-However, since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-To obtain a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
+\textcolor{teal}{We summarize our proposed approach as follows:}
 
@@
 
-We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler.
+\begin{remark}
+In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
+This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
+In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates.
+Hence, we prefer (\ref{eqn:fast_beta_prior}).
+We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
+\end{remark}
+
+\begin{remark}
+Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
+In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
+In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows.
+Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
+As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
+\end{remark}
+
+We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
 We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
@@
 Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.
+
+\textcolor{teal}{To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
+However, since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+To obtain a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.}
 
@@
 \appendix
-
-	\hypertarget{bayes-estimate}{%
-		\section{Bayes Estimate}
-		\label{bayes-estimate}}
-
-We calculate a Bayes estimate $\hat{\bm{Z}}$ for the linkage parameter $\bm{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
-We adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which $\hat{Z}_j \in \{1, \ldots, n_A, n_A + j, R\}$, with $R$ representing the option to leave the matching undetermined by the model.
-Specifically, we have
-	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
-		0,  & \text{if } Z_j = \hat{Z_j}; \\
-		\theta_R,  & \text{if } \hat{Z_j} = R; \\
-		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_A + j ; \\
-		\theta_{01},  & \text{if } Z_j = n_A + j,\hat{Z_j} \leq n_A ; \\
-		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
-	\end{cases}\] 
-Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
-
-In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
-	$$\hat{Z}_j =\begin{cases} 
-		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
-		0,  & \text{otherwise}. \\
-	\end{cases}$$
-Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-In the event that we have two records $B_j$ and $B_{j'}$ such that both $p(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ p(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
-Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
-
-When we seek a partial estimate of the linkage structure, leaving a portion of record pairs to be classified manually in clerical review, we adopt losses $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, .1)$.
-For a more in-depth explanation of this function and the induced Bayes estimate, see \cite{sadinle_bayesian_2017}.
 
@@
 
+\hypertarget{bayes-estimate}{%
+		\section{Bayes Estimate}
+		\label{bayes-estimate}}
+
+We calculate a Bayes estimate $\hat{\bm{Z}}$ for the linkage parameter $\bm{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
+We adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which $\hat{Z}_j \in \{1, \ldots, n_A, n_A + j, R\}$, with $R$ representing the option to leave the matching undetermined by the model.
+Specifically, we have
+	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
+		0,  & \text{if } Z_j = \hat{Z_j}; \\
+		\theta_R,  & \text{if } \hat{Z_j} = R; \\
+		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_A + j ; \\
+		\theta_{01},  & \text{if } Z_j = n_A + j,\hat{Z_j} \leq n_A ; \\
+		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
+	\end{cases}\] 
+Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
+
+In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
+	$$\hat{Z}_j =\begin{cases} 
+		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
+		0,  & \text{otherwise}. \\
+	\end{cases}$$
+Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+In the event that we have two records $B_j$ and $B_{j'}$ such that both $p(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ p(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
+Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
+
+When we seek a partial estimate of the linkage structure, leaving a portion of record pairs to be classified manually in clerical review, we adopt losses $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, .1)$.
+For a more in-depth explanation of this function and the induced Bayes estimate, see \cite{sadinle_bayesian_2017}.
+
 	\hypertarget{partial}{%

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex
working on comments.

@@
 
-In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. 
-We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
+As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
 To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}
 
@@
 The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$
+
+%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
+As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}
+
+\textcolor{teal}{We now explain how we accomodate missing fields, which leads to missing comparison vectors.
+To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing completely at random (MCAR, per \cite{LittleRubin2002}).
+With the MCAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.}
+
+%are ignorable (or random) \citep{LittleRubin2002}.
+The likelihood becomes that in (\ref{eqn:likelihood}), following directly from \cite{sadinle_bayesian_2017}.
+Note that both assumptions used here are for computational convenience.
+Missing completely at random (MCAR) applies when the probability that a value of a variable is missing does not depend on the observed or unobserved variables.
+Missing at random (MAR) is present when the probability of missingness depends only on (the values of other) observed variables.}
+%\textcolor{red}{In Sadinle (2017), he assumes MAR and in the JASA paper he assumes ignorability.}
 

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex
working on edits/comments.

@@
 As the prior distriburion for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
-Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
+Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
+
 It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
@@
 	\end{cases} \label{eqn:z}\\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi},
 \end{align}
 \end{subequations}
+\textcolor{teal}{where $\alpha_{\pi}, \beta_{\pi}$ are fixed hyper-parameters.}
 
@@
 
+%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
+%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
+To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
+
 We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
@@
 To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
-\begin{subequations}
-\begin{align}
-	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
-	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
-	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-\end{align}
-\end{subequations}
+
+\textcolor{teal}{We present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m}, \bm{u}$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi}).
+While it is easy to marginalize over $\pi$ and derive a collapsed Gibbs sampler, the sampler is sequential in nature as is that of \cite{}.
+Thus, we proceed with a traditional Gibbs sampler, where we show in Appendix \ref{}, the updates are parallel, making our approach scalable.}
+
+We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
+\textcolor{red}{Need to initialize all random variables, such as $\pi$.} \textcolor{teal}{Denote the current draw of the sampler by $s$ and the updated draw by $s+1$.} \textcolor{teal}{For all $f=1, \ldots, F$ sample}
+
+%\begin{subequations}
+%\begin{align}
+%	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  \\
+%	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
+%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
+%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
+%\end{align}
+%\end{subequations}
+
+%\begin{subequations}
+%
+%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
+%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
+%\end{align}
+%\end{subequations}
+
+\begin{align}
+	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  
+\end{align}
+and	
+\begin{align}
+	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} 
+\end{align}
+
+where
+
+\begin{align}
+\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \quad \text{and}\\
+\beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
+\end{align}
+
 Next, the full conditional for $\pi$ is given by
@@
 \end{align}
-Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.
-
-\textcolor{teal}{To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
+
+\textcolor{teal}{
+Observe that the full conditional in (\ref{eqn:z_full_conditional}) does not depend on $\bm{Z}^{(s)},$ implying that 
+\begin{align}
+p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}) \propto \prod_q \left(\prod_j
+p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right)
+\right).
+\end{align}
+}
+
+\textcolor{teal}{This implies that we can make parallel updates to $\bm{Z}$ instead of the traditional sequential updates as in \cite{sadinle_bayesian_2017}.
+In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), contrasting the Bayesian record linkage literature \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}.
+See Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
+
+%Derivations for the full conditional distributions can be found in Appendix~\ref{app:derivations}.
+
+\textcolor{teal}{Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
 However, since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-To obtain a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+To obtain a Bayes estimate that fulfills the one-to-one matching (or bipartite) requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
 Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.}
@@
 \end{subequations}
-These can computed serially or in parallel. Summary statistics from each pairwise chunk comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
+These can computed \textcolor{teal}{sequentially} or in parallel.
+Summary statistics from each pairwise chunk comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
 \begin{subequations}
@@
 	\section{Derivations of Full Conditionals}\label{app:derivations}}
-
-We provide detailed derivations of the full-conditionals provided in Section \ref{gibbs_sampling}.
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
-For a particular $m_{fl}$, we have
+\textcolor{red}{Note: Please use s, s+1 consistently throughout.
+Or supress the notation and tell the reader.
+It looks weird to go back and forth.
+I've tried to provide an intermediate solution.
+Perhaps Jerry can state if he would think this is okay as an Editor.}\\
+
+\textcolor{teal}{We provide detailed derivations of the full-conditionals provided in Section~\ref{sec:fast-beta-linkage}.
+In our derivations below, we assume that the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the final update of the full conditional that appears in Section~\ref{sec:fast-beta-linkage}.}
+
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{fl}$, we have
 \begin{align*}
@@
 
-\textcolor{blue}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
+\textcolor{teal}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
 \begin{align}
@@
 This contrasts the Gibbs sampler of  \cite{sadinle_bayesian_2017}, where the update to the matching matrix is sequential and accomplished using a partially collapsed Gibbs sampler, where one integrates out $\pi.$ We do not opt for this approach as it would lead to similar sequential updates, which are computationally slow.}
-
-%Let $\gamma_{.j}$ denote the set of $n_A$ comparison vectors with $B_j$. We have
-%\begin{align*}
-%	p(\gamma_{.j}|Z_j = q, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_A}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
-%	&\propto \prod_{i=1}^{n_A}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
-%	&\propto \prod_{i=1}^{n_A}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(q = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-%	&=
-%	\begin{cases} 
-%		w_{q, j}  & q \leq n_A; \\
-%		1 &  q  = n_A + j. \
-%	\end{cases}\\
-%\end{align*}
-%where
-%$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
-
-%With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$.
-Because we sample each $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$:
-%\newpage
-%\begin{align*}
-%	p&\left(Z_j^{(s+1)} = q|\gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
-%	&\propto p(\gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) p(Z_j^{(s+1)}|\pi) \\
-%	&\propto \left(\sum_{i=1}^{n_A}w_{q, j}I(q = i) + I(q = n_A + j) \right) \left (\pi\sum_{i=1}^{n_A}\frac{1}{n_A}I(q = i) + (1-\pi)I(q = n_A + j) \right) \\	
-%	&= \frac{\pi}{n_A}\sum_{i=1}^{n_A}w_{q, j}I(q = i) + (1-\pi)I(q = n_A + j). \\
-%	&= \begin{cases} 
-%		\frac{\pi}{n_A}w_{q, j}   & q \leq n_A; \\
-%		1-\pi &  q  = n_A + j. \\
-%	\end{cases}
-%\end{align*}
-%
-%For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
-%\begin{align*}
-%	p&\left(Z_j^{(s+1)}  = q| \gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
-%	&= \int_{\pi}p\left(Z_j^{(s+1)} | \gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-%	&= \int_{\pi}\left[\frac{\pi}{n_A}w_{z_{j}, j} I(q\leq n_A) + (1-\pi)I(q = n_A + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-%	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_A}w_{z_{j}, j} I(q\leq n_A) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(q = n_A + j) \\
-%	&= \frac{n_{AB}(\bm{Z}) + \alpha_{\pi}}{n_A (n_B + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(q\leq n_A) + \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_B + \alpha_{\pi} + \beta_{\pi}}I(q = n_A + j) \\
-%	&\propto w_{q, j}I(q\leq n_A) + n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}}I(q = n_A + j) \\
-%	&=
-%	\begin{cases} 
-%		w_{q, j}  & q \leq n_A; \\
-%		n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}} & q   = n_A + j. \\
-%	\end{cases} \\
-%\end{align*}
-
-	\clearpage
-
-%	\hypertarget{app:chunks}{%
-%	\subsection{SEI for Large Linkage Tasks}\label{app:chunks}}
-
-%	For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks $\{A^a | a = 1, \ldots, t_A \}$ and $\{B^b | b = 1, \ldots, t_B\}$ for more manageable computations.
-{\color{red}{For example, if $n_A = n_B = 500$ and $t_A = t_B = 10$, each chunk $A^a$ and $B^b$ would contain 50 records.
-(Jerry asked for more precision here.
-Does this example work?) }}For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
-We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$ for later calculations, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
-In detail, we calculate
-%	\begin{subequations}
-%		\begin{align}
-%			r_{p_j}^{ab} &= \{i \in  | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
-%			N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
-%		\end{align}
-%	\end{subequations}
-%	These can computed serially or in parallel.
-We reduce memory costs through SEI on each $r_{p_j}$, while retaining counts of each agreement pattern through $N_{p_j}^{ab}$.
-Summary statistics from each pairwise chunk comparison can be easily synthesized to recover summary statistics for the full comparison matrix $\gamma$.
-Specifically, we combine information through
-%	\begin{subequations}
-%		\begin{align}
-%			r_{p_j}^{SEI} &= (r_{p_j}^{SEI, 11}, \ldots, r_{p_j}^{SEI, A^{*}B^{*}}) \text{ for } a = 1, \ldots, A^{*} \text{ and } b = 1, \ldots, B^{*},\label{r_{p_j}} \\
-%			N_{p_j} &= \sum_{a = 1}^{A^{*}} \sum_{b = 1}^{B^{*}} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
-%	\end{subequations}
-%	
-%	{\color{red}{Note that with this approach, each $r_{p_j}^{SEI}$ can contain up to $S  t_A  t_B$ indices, so we store at most  $n_B  P  S  t_A  t_B$ indices in total.}}
-
-%		For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks $\{A^a | a = 1, \ldots, t_A \}$ and $\{B^b | b = 1, \ldots, t_B\}$ for more manageable computations.
-{\color{red}{For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each chunk $A^a$ and $B^b$ would contain 500 records.
-(Jerry asked for more precision here.
-Does this example work?) }}For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
-We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$ for later calculations, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
-In detail, we calculate
-%	\begin{subequations}
-%		\begin{align}
-%			r_{p_j}^{ab} &= \{i \in  | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
-%			N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
-%		\end{align}
-%	\end{subequations}
-%	These can computed serially or in parallel.
-Summary statistics from each pairwise chunk comparison can be easily synthesized to recover summary statistics for the full comparison matrix $\gamma$.
-Specifically, we combine information through
-%	\begin{subequations}
-%		\begin{align}
-%			r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B,\label{r_{p_j}} \\
-%			N_{p_j} &= \sum_{a = 1}^{t-A} \sum_{b = 1}^{t_B} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
-%	\end{subequations}
-%	
-%	
-%	\hypertarget{app:fabl-psuedocode}{%
-%		\subsection{Summary of Fast Beta Linkage Method}\label{app:fabl-psuedocode}}
-%	
-%	\textcolor{red}{I am getting rid of the notation $h(Z_j)$ for the agreement pattern between $B_j$ and its potential match.
-Instead, I am keeping notation that already exists, saying $(Z_j, j) \in h_p$.
-With this change, I don't see a good way to write lines 15 and 17.
-For that reason, I think I should remove this pseudo-code; it doesn't seem necessary, and invites criticism.}
-%	
-%	\begin{algorithm}[h!]
-%		\begin{algorithmic}[1]
-%			
-%			\Procedure{Hashing and Preprocessing}{}
-%			\State Construct and enumerate set of unique patterns $\mathcal{P}$ from $F$ and $\{L_f\}$.
-%			\State Partition files $\bm{X_1}$ and $\bm{X_2}$ into chunks $\{\bm{X_{1n}}\}$, $\{\bm{X_{2m}}\}$.
-%			\For{each $n$, $\bm{m}$}
-%			\State Create comparison vectors between $\bm{X_{1n}}$ and $\bm{X_{2m}}$.
-%			\State Hash records to $\mathcal{R}_{nm}$ and calculate summary statistics $\mathcal{H}_{nm}$.
-%			\State Use SEI to reduce memory usage; $\mathcal{R}_{nm} \to \mathcal{R}_{nm}^{\text{SEI}}$.
-%			\EndFor
-%			\State Synthesize results across pairings to get $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H} \}$.
-%			\EndProcedure
-%			
-%			\Procedure{Gibbs Sampling}{}
-%			\State Initialize $\bm{m}$, $\bm{u}$, and $Z$ parameters.
-%			
-%			\For{$t \in \{1, \ldots, T\}$} 
-%			\State Sample $\bm{m}^{t+1}|Z^{t}, \tilde{\gamma}$ and  $\bm{u}^{t+1}|\bm{Z}^{t}, \tilde{\gamma}$
-%			\State Sample $h\left(\bm{Z}^{t+1}\right)|\bm{m}^{t+1}, \bm{u}^{t+1}, \tilde{\gamma}$.  \Comment{Sample agreement pattern, not record}
-%			\EndFor
-%			\State Sample $\bm{Z} | h(\bm{Z}), \tilde{\gamma}$. \Comment{Fills in record label based on agreement pattern}
-%			
-%			\EndProcedure
-%			
-%		\end{algorithmic}
-%	\end{algorithm}
-%	\clearpage
-
-%	\hypertarget{app:ohe}{%
-%		\subsection{One Hot Encoding Transformation}\label{app:ohe}}
-%	
-%	As described in Section \ref{sec:efficiency}, \texttt{fabl} makes use of one-hot encodings to aid in vectorized computations.
-For $\gamma_{ij}^f$ with $L_f$ levels, define $e_{ij}^f$ to be an $L_f \times 1$ vector.
-When $\gamma_{ij}^f = l$, we set the $l^{th}$ element of $e_{ij}^f$ to be 1, and set the other $L_f - 1$ elements of $e_{ij}^f$ to be 0.
-We then concatenate the $e_{ij}^f$ for all $f \in \{1, \ldots, F\}$, resulting in the one-hot encoded comparison vector $e_{ij}$ of length $\sum_{f=1}^F L_f$.
-%	
-%	For example, consider comparing the toy records shown in Table \ref{tab:ohe} with $L = (3, 3, 2, 2)$ levels of agreement for last name, first name, DOB, and city respectively.
-Since the first name differs by only one letter, a reasonable comparison vector for this pair would be $\gamma_{ij} = (1, 2, 1, 2)$.
-The one hot encoding representation of this vector is $e_{ij} = (1, 0, 0, 0, 1, 0, 1, 0, 0, 1)$.
-%	\begin{table}[h!]
-%		\centering
-%		\begin{tabular}[h!]{llll}
-%			Last Name & First Name & DOB & City \\
-%			\hline
-%			Smith & Taylor & 01/01/2000 & Durham\\
-%			Smith & Tayler & 01/01/2000 & Raleigh\\
-%			\hline
-%		\end{tabular}
-%		\caption{Example records for one hot encoding.}\label{tab:ohe}
-%	\end{table}
 

### 2023-07-19 SEI_sensitivity/Sadinle_accuracy_cluster.Rmd
pi in gibbs sampler

---
title: "Sadinle_accuracy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sadinle_accuracy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(devtools)
load_all()
library(parlrdev)
library(tidyverse)
rm(list = c("hash_field"))
```

```{r}
taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
i = taskID
set.seed(41)
files <- list.files(path = "../data/SimulationDataFiles", full.names = T)

m_prior = 1
u_prior = 1
alpha = 1
beta = 1
S = 1000
burn = 100
show_progress = T
fast = F
R = NULL
all_patterns = TRUE

overlap_vals <- c(50, 250, 450)

fabl_acc_samps <- matrix(NA, nrow = 5, ncol = 3)
fabl_acc_samps_resolved <- matrix(NA, nrow = 5, ncol = 3)
sad_acc_samps <- matrix(NA, nrow = 5, ncol = 3)

fabl_acc_samps2 <- matrix(NA, nrow = 6, ncol = 3)
fabl_acc_samps_resolved2 <- matrix(NA, nrow = 6, ncol = 3)
sad_acc_samps2 <- matrix(NA, nrow = 6, ncol = 3)

for(j in seq_along(overlap_vals)){

  overlap <- overlap_vals[j]

#for(i in seq_along(files)){

records <- read_csv(files[i], col_types = cols())
records$file <- rep(2:1, length.out = dim(records)[1])

records <- records %>% 
  janitor::clean_names() %>% 
  mutate(rec_id = as.numeric(str_extract(rec_id, "\\d{3}")) + 1)

n1 <- 500
n2 <- 500
#overlap <- n2/2

Ztrue <- n1 + 1:n2
Ztrue[1:overlap] <- 1:overlap

file1 <- records %>% 
  filter(file ==1,
         rec_id <= n1) %>% 
  select(-rec_id) %>% 
  as.matrix(.) %>% 
  data.frame(.) %>% 
  mutate(occup = as.numeric(occup))

file2 <- records %>% 
  filter(file == 2,
         rec_id %in% c(1:overlap, (n1 +1):(1000 - overlap))) %>% 
  select(-rec_id) %>% 
  as.matrix() %>% 
  data.frame(.) %>% 
  mutate(occup = as.numeric(occup))

cd <- suppressMessages(BRL::compareRecords(file1, file2, c(2, 3, 5, 6), 
                          types = c("lv", "lv", "bi", "bi")))
cd[[1]] <- apply(cd[[1]], 2, as.numeric)

  #parlr method
  ptm <- proc.time()
  Zchain <- BKSimple_hash2(cd, R = R)
  Zchain <- fabl_gibbs(cd, R = R)
  elapsed <- proc.time() - ptm
  Zhat <- LinkRecordsBK(Zchain[[1]], n1, 1, 1, 2, Inf)
  eval <- GetEvaluations(Zhat[[1]], Ztrue, n1)
  eval

  # Resolution Step
  Zhat_resolved <- ResolveConflicts(Zhat)
  eval_resolved <- GetEvaluations(Zhat_resolved, Ztrue, n1)
  fabl_acc_samps_resolved[, j] <- c(eval_resolved, elapsed[3], overlap)
  fabl_acc_samps[, j] <- c(eval, elapsed[3], overlap)

  #Partial Estimate
  Zhat <- LinkRecordsBK(Zchain[[1]], n1, 1, 1, 2, .1)
  Zhat <- ResolveConflicts(Zhat)
  eval <- GetEvaluations(Zhat, Ztrue, n1)
  RR <- sum(Zhat == -1)/n2
  eval[1] <- sum(Zhat == Ztrue & Ztrue > n1) / (sum(Zhat > n1))
  fabl_acc_samps2[, j] <- c(eval, RR, elapsed[3], overlap)

  #Sadinle 2017 Method
  ptm <- proc.time()
  Zchain <- BRL::bipartiteGibbs(cd)[[1]]
  Zchain <- Zchain[,101:1000]
  elapsed <- proc.time() - ptm
  Zhat <- BRL::linkRecords(Zchain, n1, 1, 1, 2, Inf)
  eval <- GetEvaluations(Zhat, Ztrue, n1)
  sad_acc_samps[, j] <- c(eval, elapsed[3], overlap)

  Zhat <- BRL::linkRecords(Zchain, n1, 1, 1, 2, .1)
  RR <- sum(Zhat == -1)/n2
  eval <- GetEvaluations(Zhat, Ztrue, n1)
  eval[1] <- sum(Zhat == Ztrue & Ztrue > n1) / (sum(Zhat > n1))
  sad_acc_samps2[, j] <- c(eval, RR, elapsed[3], overlap)

#  print(i)
#}
}

saveRDS(Zs, file = paste0("out/fabl_acc/fabl_acc_", 
                          str_pad(fabl_acc_samps, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/fabl_acc_res/fabl_acc_res_", 
                          str_pad(fabl_acc_samps_resolved, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/sad_acc/sad_acc_", 
                          str_pad(sad_acc_samps, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/fabl_acc2/fabl_acc2_", 
                          str_pad(fabl_acc_samps2, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/sad_acc2/sad_acc2_", 
                          str_pad(sad_acc_samps2, 3, pad = "0")))
# list(BK = parlr_acc_samps,
#      Sadinle = sad_acc_samps)

```

### 2023-07-19 ba_resubmit/response_to_reviewers2.tex

@@
 
-	\reply{The assumption is missing completely at random (MCAR) due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix (or matching label) .
+	\reply{The assumption is missing completely at random (MCAR) due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix (or matching label).
 The assumption is not missing at random (MAR) as we do not assume that the missing comparisons are also ignorable.
-See Section 2, page 2 regarding clarification of this point \citep{LittleRubin2002}.}
+See Section 3, page 5 regarding clarification of this point \citep{LittleRubin2002}.}
+	\textcolor{red}{Still checking on this and looking at Little and Rubin.}
 

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex

@@
 The \cite{fellegi_theory_1969} approach has been extended for a wide variety of applications \citep[e.g.,][]{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
-An alternative paradigm is to model the linkage variables directly \citep[e.g.,][]{tancredi2011hierarchical, steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}.
+An alternative paradigm is to model the linkage variables directly \citep[e.g.,][]{tancredi2011hierarchical, steorts_bayesian_2016, marchant_distributed_2019, betancourt2022prior}.
 In this article, we build on the contributions to the comparison vector approach.
@@
 {\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
+We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
 This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
@@
 However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$,
-\textcolor{teal}{we use propose an alternative, called the fast beta linkage prior (fabl).}
+\textcolor{teal}{we propose an alternative, called the fast beta linkage (fabl) prior.}
 %the fast Beta prior as follows. 
@@
 	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}),\quad  \forall  \; f = 1, \ldots, F, \label{eqn:m} \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}), \quad \forall \; f = 1, \ldots, F,  \label{eqn:u}\\
 	p(Z_j = q| \pi)  &=
@@
 \end{subequations}
-\textcolor{teal}{where $\alpha_{\pi}, \beta_{\pi}$ are fixed hyper-parameters.}
+\textcolor{teal}{where recall that $\alpha_{\pi}, \beta_{\pi}$ are fixed hyper-parameters.}
 
@@
 
-We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
-We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
+%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
+%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
 To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
 
-\textcolor{teal}{We present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m}, \bm{u}$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi}).
-While it is easy to marginalize over $\pi$ and derive a collapsed Gibbs sampler, the sampler is sequential in nature as is that of \cite{}.
-Thus, we proceed with a traditional Gibbs sampler, where we show in Appendix \ref{}, the updates are parallel, making our approach scalable.}
+\textcolor{teal}{We present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m},$ and  $\bm{u}$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi}).
+While we could marginalize over $\pi$ and derive a collapsed Gibbs sampler, in the spirit of \cite{sadinle_bayesian_2017}, this leads to sequential updates, which are computationally slow.
+Instead, we proceed with a traditional Gibbs sampler, where the updates to the matching label (or coreference matrix) are parallel, allowing for more efficient updates.
+Refer to Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
 
@@
 In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), contrasting the Bayesian record linkage literature \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}.
-See Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
+}
+%See Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
 
@@
 To obtain a Bayes estimate that fulfills the one-to-one matching (or bipartite) requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.}
+See Appendix \ref{bayes-estimate} for details regarding the initial Bayes estimate and the post-processing procedure.}
+
+%Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
 

### 2023-07-19 ba_resubmit/response_to_reviewers2.tex
more comments on 4.2

@@
 
+	\reply{The distinction does not matter due to assuming that the comparison vectors are conditionally independent given the coreference matrix (or matching label).
+We now state the following in Section 3, page 5:
+
+	``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
+With either the ignorability or MAR assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data.
+For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
+Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p.
+12]{LittleRubin2002}).
+To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation."
+	}
+
 	%% RCS: See two comments below regarding MAR versus MCAR. I have an alternative response for the one regarding miss at random.
@@
 
-	\reply{The assumption is missing completely at random (MCAR) due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix (or matching label).
+%	\reply{The assumption is missing completely at random (MCAR) due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix (or matching label).
 The assumption is not missing at random (MAR) as we do not assume that the missing comparisons are also ignorable.
@@
 
-\newpage
 \pointRaised{R}{%
@@
 	}
+
+	\newpage
 

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex
more comments on 4.2

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{example}{Example}
+\newtheorem{rem}{Remark}
 \usepackage{booktabs}       % professional-quality tables
@@
 He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.
-As the prior distriburion for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
+As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
 Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
@@
 
-\textcolor{teal}{We now explain how we accomodate missing fields, which leads to missing comparison vectors.
+\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
 To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing completely at random (MCAR, per \cite{LittleRubin2002}).
-With the MCAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.}
+Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
+With either the ignorability or MAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.
+For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
+Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p.
+12]{LittleRubin2002}).
+To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation.}
 
@@
 
-\begin{remark}
+\begin{rem}
 In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
@@
 We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
-\end{remark}
-
-\begin{remark}
+\end{rem}
+
+\begin{rem}
 Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
@@
 As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
-\end{remark}
+\end{rem}
 
@@
 The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains.
-Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
+Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking \textcolor{teal}{often}
+ undesirable %in some situations 
+\citep{steorts_comparison_2014}.
 
@@
 
-With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
+With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage.
 First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors.
@@
 
-Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
-Let $P$ be the number of patterns realized in $\gamma$.
-It is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.
-This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
+Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. 
+%It is
+\textcolor{teal}{The comparison vector} is
+ bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1).$
+% This comment below seems obvious and not needed.  
+% where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values. 
+\textcolor{teal}{The quantity $P^{*}$ clearly depends only on $F$ and $L_f.$ Observe it does not depend (or scale) with $n_A$ or $n_B.$}
+%is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
+
 To obtain a memory efficient representation, we map the agreement pattern of each record pair to a unique integer.
@@
 
-We then map these integers to sequential integers $\{1, \ldots, P\}$.
-We refer to each unique agreement pattern as $h_p$, and to the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
+%\textcolor{red}{Do you mean what I have below or something else? This still is a bit hard to parse.}
+We then map \textcolor{teal}{these integers in (\ref{eqn:hashing})} to sequential integers $\{1, \ldots, P\}$.
+\textcolor{teal}{Denote} each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+When record pair, \textcolor{teal}{$(i,j)$,} exhibits agreement pattern $p$, \textcolor{teal}{let} $\gamma_{ij} = h_p$.
+
 In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
@@
 
-For example, with five fields with binary agreements and possible missingness, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
-Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
+\begin{example}
+\label{ex:one-hot}
+For example, with five fields with binary agreements and possible missingness (NA), the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
+Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, \text{NA}, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
 Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
 We map the unique values of $h^{*}(\gamma)$ to sequential integers.
-If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
+\textcolor{teal}{Suppose that 1066 maps to 42.
+This implies $\gamma_{5,7} = h_{42}.$}
+%If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ 
+This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
+\end{example}
 
@@
 to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match, in (\ref{eqn:m_p}), and not a match, in (\ref{eqn:u_p}).
-Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) is equivalent to
+Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) \textcolor{teal}{can be expressed as}
+%is equivalent to
 
@@
 The likelihood in (\ref{likelihood_efficient}) allows for more
-efficient posterior inference for $\bm{m}$, $\bm{u}$, and $\bm{Z}$, as we now describe.
+efficient \textcolor{teal}{approximate posterior inference of the Gibbs sampler in Section~\ref{sec:fast-beta-linkage},}  as we \textcolor{teal}{describe in the next section}.
+
+%% These are not the full conditional distributions.
+%for $\bm{m}$, $\bm{u}$, and $\bm{Z}$
 
@@
 
-Turning first to the posterior updates for $\bm{m}$ and $\bm{u}$, we first note that these parameters depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$.
+\textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.}
+
+Turning first to the \textcolor{teal}{full conditional updates}
+%posterior updates 
+for $\bm{m}$ and $\bm{u}$, we observe that these parameters only depend on the \textcolor{teal}{observed} data  through quantities that are calculated through $\mathcal{N}$ and $\mathcal{P}$.
 Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
-Then, conditional on $Z$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
+Then, conditional on $Z$, we can express the contribution to the likelihood in the full conditional distribution for $\bm{m}$ and $\bm{u}$ as
 \begin{align}
@@
 Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be a concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
-The terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
+The terms needed for the \textcolor{teal}{full conditional} updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
 \begin{subequations}
@@
 \end{subequations}
-Specifically, the $l + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}),and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{eqn:beta_update}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-We also use (\ref{likelihood_efficient}) to sample Z efficiently.
+%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
+However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
+However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+\textcolor{teal}{Observe that we compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{efficient_alpha}) and (\ref{efficient_beta}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
+
+We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
 For each pattern $p$, let $w_p =\frac{m_p}{u_p}$.
@@
 This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-To speed up computation, we break this sampling step into two.
+\textcolor{teal}{We consider two sampling steps.}
+%To speed up computation, we break this sampling step into two.
 

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex
small comment.

@@
 
+	\textcolor{red}{Suggestion: Chunk is award. Perhaps partition/block/batch/segment. I might like segment since partition/block have meanings in record linkage already. }
+
 		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations.
 Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.
-Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B and B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
+
+		Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
 For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
@@
 
-		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison vectors} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
+		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison matrix} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
 		the 
 \textcolor{teal}{memory-intensive}
-		\textcolor{teal}{comparison vectors} $\gamma^{ab}$ 
+		\textcolor{teal}{comparison matrix} $\gamma^{ab}$ 
 		before continuing with the next chunk of data. 
@@
 For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
-We next introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
+We introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
 This allows all-to-all comparisons for substantially larger linkage tasks.
@@
 Rather than store all of these record labels, we store only a small number $S$.
-For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $SEI(r_{p_j}^{ab})$.
-We collect these memory reduced lists to form $SEI(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $SEI(\mathcal{R})$.
+For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $\text{SEI}(r_{p_j}^{ab})$.
+We collect these memory reduced lists to form $\text{SEI}(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $\text{SEI}(\mathcal{R})$.
 
@@
 Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels.
-As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $SEI(\tilde{\gamma}) = \{\mathcal{P}, SEI(\mathcal{R}), \mathcal{N}\}$.
-We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.
+As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $\text{SEI}(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
+We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
 

### 2023-07-19 ba_resubmit/response_to_reviewers2.tex
working on response to reviewers.

@@
 
-	\newpage
-
 %	\reply{%
@@
 See Section 4.1, pages 8--9.}
+
  }
@@
 
-	``Those comparison vectors" refer to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$
+%	In the revised paper, ``vector" is replaced with ``matrix" when referring to $\gamma.$ (See R.13 regarding making the paper consistent below).
 
-	The revised text states: ``Then, we conduct hashing, obtain the compressed comparison vectors $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison vectors $\gamma^{ab}$ before continuing with the next chunk of data.
+	\textcolor{teal}{In the original manuscript, ``those comparison vectors" refer to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$ }
+
+	\textcolor{teal}{The revised text states: ``Then, we conduct hashing, obtain the compressed comparison matrix, $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison matrix, $\gamma^{ab}$, before continuing with the next chunk of data.
 See Section 4.3, page 10.
+}
+
+%	 \textcolor{red}{Question for Brian: is lowercase $\gamma$ always a matrix?
+Let's make sure we're not swapping between vectors and matrices.
+In the previous version this was a vector and now its listed as a matrix.
+I would argue that a matrix should use capital notation, such as $\Gamma.$}
 	}
@@
 \reply{%
-	This has been fixed.
+	\textcolor{teal}{The typo has been fixed.}
 }
@@
 \reply{%
-	$\gamma$ is a matrix comprised by comparison vectors. All references to $\gamma$ as a set have been removed.
+	\textcolor{teal}{Note that $\gamma$ is a comparison matrix comprised by comparison vectors.
+All references to $\gamma$ as a set have been removed.
+See Section 2, page 3, where we define both the comparison vector and comparison matrix.}
 }

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex
working on response to reviewers.

@@
 Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
-For notational convenience, we define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define a \emph{comparison vector} $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
+For notational convenience, we define the \emph{comparison matrix} $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
 

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex
more edits.

@@
 
-\textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.}
+\textcolor{teal}{In our derivations below, we assume the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the updates to each full conditional distribution based upon our computational proposals using hashing.}
+
+\begin{enumerate}	
+\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.}
+\item \textcolor{red}{There appears to be no update for $\pi$ in this section.}
+\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1.
+Example, it's missing on the update for $m$ and $u$ so it's not immediately clear how one would draw this directly.
+Wouldn't it just be a vectorized Dirichlet, where the computational savings come from vectorizing the hyperparameters of the Dirichlet?
+\end{enumerate}
 
@@
 
-\textcolor{teal}{Observe that we compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{efficient_alpha}) and (\ref{efficient_beta}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
+\textcolor{teal}{Observe that we compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
+
+\textcolor{red}{Some statement regarding $\pi$ needs to go here even if it's that you draw from the same distribution as in the original sampler.}
 

### 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
making more comments.
working on revisions.

@@
 \begin{align}
+\label{eqn:pi}
 	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
@@
 \begin{enumerate}	
-\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.}
-\item \textcolor{red}{There appears to be no update for $\pi$ in this section.}
+\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.
+Also, do you more formally mean something like as follows: Consider $\gamma_{q,q} = h_p \mid Z_j = q$.
+It seems that the conditioning may actually be important here.
+I mention this for two reasons, it appears a bit strange in equation (22).
+When I think through it, it's okay.
+I just wanted to bring it up as a conversation before submission.}
+\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
 \item Make sure when you give the update for the new efficient full conditional, you give the s, s+1.
@@
 
-\textcolor{red}{Some statement regarding $\pi$ needs to go here even if it's that you draw from the same distribution as in the original sampler.}
+\textcolor{teal}{Turning to sampling for $\pi,$ it remains the same as in (\ref{eqn:pi}). }
 
@@
 \textcolor{red}{Note: Please use s, s+1 consistently throughout.
-Or supress the notation and tell the reader.
+Or surpress the notation and tell the reader.
 It looks weird to go back and forth.

### 2023-07-20 ba_resubmit/response_to_reviewers2.tex
address fix that patrick ball asked for over email.

@@
 \reply{
-We thank the Editor for the opportunity to revise our paper and for the suggestions.
+\textcolor{teal}{We thank the Editor for the opportunity to revise our paper and for the suggestions.
 We have improved notation, when possible, however, as noted by the Editor in certain places, such as in the computational speeds ups of Section 4, this is difficult to avoid.
@@
 Finally, we have attempted to clarify our overall contributions to the literature, which we highlight below.
+In addition, we provide responses to all the Editor's and Associate Editor's comments below.}
 
@@
 This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the decision theoretic  technique of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.
-This discovery allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We employ the decision theoretic  technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.
+This discovery allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs via a novel method, storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
@@
 Finally, we provide two real illustrations of our methodology and simulation studies with reproducible software.
+
 }
@@
 
-	\reply{The distinction does not matter due to assuming that the comparison vectors are conditionally independent given the coreference matrix (or matching label).
-We now state the following in Section 3, page 5:
+	\reply{\textcolor{teal}{The distinction does not matter due to assuming that the comparison vectors are conditionally independent given the coreference matrix (or matching label).
+We now state the following in Section 3, page 5:}
 
-	``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
+	\textcolor{teal}{``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
+Due the the conditional independence assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data.
+For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
+
+%	Furthermore, both MAR and MCAR lead to assuming the missing data is ignorable in this situation, which implies that the marginalized likelihood under all three assumptions is the same."
+
+%	``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
+With either the ignorability or MAR assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data.
+For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
+Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p.
+12]{LittleRubin2002}).
+To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation."
+%
+
+%	(MAR, per \cite{LittleRubin2002}).
 With either the ignorability or MAR assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data.
@@
 See Section 3, page 5 regarding clarification of this point \citep{LittleRubin2002}.}
-	\textcolor{red}{Still checking on this and looking at Little and Rubin.}
+%	\textcolor{red}{Still checking on this and looking at Little and Rubin.}
 

### 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
address fix that patrick ball asked for over email.

@@
 To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
+Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
+Due the the conditional independence assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}) and do all computation simply using the observed data.
+For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
+
+%Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
 With either the ignorability or MAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.
@@
 
-	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017}.
+	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017} and others \citep{lum2013applications, jewell2018accounting}.
 Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
@@
 We are interested in estimating the total number of casualties from the war.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
+We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL)\citep{howland2008rescate} and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish)\citep{ball2000salvadoran}.\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization.
 The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
@@
 Second, the only fields recorded are given name, last name, date of death, and place of death.
-It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
+\textcolor{teal}{While it may be common in other data sets for a parent and child to share the same given name, in this data set, it is more common for cousins to share this information, especially in villages, where there may be only four to five extended families.
+In Latin America, an individual receives a surname from their father and their mother, leading to a limited number of surnames in small villages.}
+	%% I believe that Patrick said that this is not common in this data set
+%	It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
 

### 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
reviewer comments.

@@
 Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
-Due the the conditional independence assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}) and do all computation simply using the observed data.
+Due the the conditional independence assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data.
 For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}

### 2023-07-20 ba_resubmit/deprecated/kundinger_resubmit3.tex
cleaning up repo.

--- 2022-11-23 ba_resubmit/kundinger_resubmit.tex
@@
 
-Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively.
-Without loss of generality, denote files such that $n_1 \geq n_2$.
-For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$.
+%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.
+
+Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
+Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
+Without loss of generality, denote files such that $n_A \geq n_B$.
 Under bipartite matching, the set of matches across data files can be represented in two equivalent ways.
-First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
+First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
 \begin{align}
@@
 
-Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
-$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
+%{\color{red}{The following paragraph is pretty unnecessary, I never use this notation}}
+Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of non-matches by 
+$\bm{U} =  \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''
 
@@
 
+%{\color{red}{Beka suggested eliminating the following two paragraphs. Thoughts? I think the first one is good, the second seems unneccesary.}}
+
 Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.
-
-%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two data files.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ {\color{red}{We use $\Gamma_{ij}$ when referring to the vector as a random variable, and $\gamma_{ij}$ when referring to observed data.
+We collect the vectors $\gamma_{ij}$ as the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$}}.
+
+%Frequently, $\gamma_{ij}$ consists of one comparison for each feature shared between the two data files.
 
 The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features.
-For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for record $i$ and $j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
+For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
 For numerical features, we can use absolute difference between the two feature values.
-For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for record $i$ and $j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for record $i$ and $j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
+For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
 For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
@@
 
-More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
-Here, $I_{f1}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
+More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
+Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
 Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
 
-In the construction of comparison vectors, it is common to encounter missing information in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$.
+{\color{red}{
+In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
 As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-We assume that this missingness occurs at random (MAR, per \cite{LittleRubin2002}).
-With the MAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
 To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+}
+}
 
@@
 To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework.
-In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
-He assigns a prior distribution for the probability of the indicator that a record in $\bm{X}_2$ has a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
-It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$.
-Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 !
-/ (n_1 - n_{12}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
-Thus, the beta distribution for bipartite matching, is given by
+In addition to {\color{red}{uniform}} Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$.
+He assigns a prior distribution for the probability of the indicator that a record in $B$ has a match in $A$, so that $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$.
+It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
+Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A !
+/ (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
+Thus, the prior is given by
 \begin{align}
 \label{eqn:sadinle_prior}
-P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
+P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
 \end{align}
 where $\text{B}(\cdot, \cdot)$ represents the Beta function.
-This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each \(Z_j\).
+This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
 This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.
-
-%hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X}_2$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
-%One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions.
-One way of resolving this issue to incorporate the constraint directly into the model as follows:
-%\begin{align}
-%	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
-%	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
-%	& \bm{Z} \sim \mathcal{B}, \notag
-%\end{align}
-%where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches.
-For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
-%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
-%$$
-%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$.
-This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\).
-This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.
 
@@
 	\label{eqn:fast_beta_prior}
-	p(Z_j = z_j| \pi) &= \begin{cases} 
-	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
-	1-\pi, &  z_j  = n_1 + j; \\
+	p(Z_j = q| \pi) &= \begin{cases} 
+	\frac{1}{n_A}\pi,  & q \leq n_A; \\
+	1-\pi, &  q  = n_A + j; \\
 \end{cases} \\
@@
 \end{align}
-We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match.
-The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$
-
-In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_1, n_1 +j\}$ is a priori equally likely for $Z_j$.
-This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$.
+We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match.
+The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$
+
+In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
+This however amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
 In our preliminary studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}).
-We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.
-
-%Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match.
+We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
+
+%Intuitively, this set of priors says that record $B_j$ has some match in $A$ with probability $\pi$, and that each record $A_i$ is equally likely to be that match.
 $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.
-One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\).
-We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.
+One non-informative choice might be $\pi \sim \text{Beta}(1, 1)$, which corresponds to a prior belief that non-matches and matches are equally likely, and another might be $\pi \sim \text{Beta}\left(1, \frac{1}{n_A}\right)$, which corresponds to a uniform prior on the labeling of $\bm{Z}$.
+We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_A}\right)$ random variable.
 
 Note that linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
 We find $\pi$ to be more a interpretable parameter than $\lambda$ in the bipartite case.
-In this setting, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions.
+In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows.
+Additionally, while the Fellegi-Sunter model makes $n_A \times n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
 As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
 
-%To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}, reproduced in Appendix \ref{bayes-estimate}.
-Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
-
 To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
+However, there is a crucial difference: since \ref{eqn:fast_beta_prior} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
 To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
 Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
-
-%\hypertarget{fabl-model}{%
-	%	\subsection{Model Specification}
-	%	\label{model-specification}} 
-%
-%As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent.
-Additionally, we assume that any missing data is missing at random.
-Let  $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denote the probability of that matching records have level $\ell$ of disagreement in feature $f$, and $u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represent the same probability for non-matches.
-In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, $\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
-
-%The model can then be specified as follows:
 
@@
 		\hline
-		$\bm{X}_1, \bm{X}_2$ & data files\\
-		$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
-		$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
+		$A, B$ & data files\\
+		$i \in 1, \ldots, n_A $ & index over records in $A$\\
+		$j \in 1, \ldots, n_B $ & index over records in $B$\\
 		$f \in 1, \ldots F$ & index over fields used for comparisons \\
 		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
-		$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
-		$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
-		$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
-		$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
+		$n_{AB}$ & number of entities in common between $A$ and $B$\\
+		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+		$Z_j = i$ & records $A_i$ and $B_j$ match \\
+		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
 		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
 		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-		$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
-		\hline
-	\end{tabular}\caption{Summary of model notation.}\label{table_notation_1}
+		$\pi$ & probability that a record in $B$ has a match in $A$ \\
+		\hline
+	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
 \end{table}
 
-For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_1}.
+For clarity, we present our full model below. A summary of notation is provided in Table \ref{table_notation_A}.
 \begin{subequations}
 \begin{align}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \label{eqn:m} \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}), \label{eqn:u}\\
-	p(Z_j = z_j| \pi)  &=
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
+	p(Z_j = q| \pi)  &=
 	\begin{cases} 
-		\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
-		1-\pi, &  z_j  = n_1 + j; \\
+		\frac{1}{n_A}\pi,  & q \leq n_A; \\
+		1-\pi, &  q  = n_A + j; \\
 	\end{cases} \label{eqn:z}\\
@@
 \begin{align}
-	\bm{m}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
-	\bm{u}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update}
+	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
+	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
+	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
+	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}
+	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
 \end{align}
 \end{subequations}
-where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} \neq i)$.
-
-Next, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
+Next, the full conditional for $\pi$ is given by
+\begin{align}
+	\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi}).
+\end{align}
+
+Lastly, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
 \begin{align}
 	\label{eqn:z_full_conditional}
-p\left(Z_j^{(s+1)}  = z_j|\Gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
+	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
 	\begin{cases} 
-	w_{z_j, j},  & z_j \leq n_1; \\
-	n_1 \frac{n_2 - n_{12}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{12}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
+		\frac{\pi^{(s+1)}}{n_A} w_{q, j}^{(s+1)},  & q \leq n_A; \\
+		1 - \pi^{(s+1)}, & q  = n_A + j, \\
 	\end{cases}
 \end{align}
-where, for all $i \in \bm{X}_1$, 
+%\begin{align}
+%\label{eqn:z_full_conditional}
+%p\left(Z_j^{(s+1)}  = z_j|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
+%\begin{cases} 
+%	w_{z_j, j}^{(s+1)},  & z_j \leq n_A; \\
+%	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
+%\end{cases}
+%\end{align}
+where, for all $i \in \{1, \ldots, n_A\}$, 
 \begin{align}
 	\label{eqn:fs_weight}
-	w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+	w_{ij}^{(s)} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
 \end{align}
 Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.
-
-%We work with the following factorization of the joint posterior distribution:
-%
-%\begin{align*}
-%	p(\bm{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\bm{Z} | \pi) p(\Phi) p(\pi) \\
-%	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-%	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
-%	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
-%	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}.
-%\end{align*}
-
-%Since the Gibbs sampler does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the highest posterior probability, and declare the other to have no match.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
 
@@
 However, the impact of indexing on model parameters is not well understood; \citep{murray2016probabilistic} reviewed this issue in the context of the frequentist Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
-%Another technique is indexing, in which the modeller decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
 
 With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
-First, we propose hashing methods that allow us to compute sufficient statistics that reduce the computational complexity of the Gibbs sampler.
+First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler.
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
-For convenience, Table \ref{table_notation_2} summarizes the notation introduced throughout this section.
+For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.
 
@@
 		\hline
-		$h_p$ & one hot encoding of agreement pattern $p$ \\
-		$(i,j) \in h_p$ & comparison vector between records $i \in \bm{X}_1$ and $i \in \bm{X}_2$ exhibits pattern $p$ \\
-		$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-		$N_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
-		$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
-		$H_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
-		$H_p^u$ & number of non-matching comparison vectors that exhibit pattern $p$ \\
-		\hline
-	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_2}
+		$P$ & number of agreement patterns exhibited in $\gamma$ \\
+		$h_p$ & agreement pattern $p$ \\
+		$e(h_p)$ & one-hot encoding of agreement pattern $p$ \\
+		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $B_j$ exhibits pattern $p$ \\
+		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+		$N_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+		$N_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+
+		\hline
+	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_B}
 \end{table}
-
-%To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi-Sunter method.
-The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector.
-To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns.
-\(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
-In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the orignial vectors themselves.
-Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi-Sunter comparison vectors, to aid in vectorized computations.
-See Appendix \ref{app:ohe} for details.
 
@@
 
-Following an insight from \cite{enamorado2019using}, we recognize that record pairs contribute to posterior calculations only through their agreement pattern as represented by $\gamma_{ij}$.
-To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns observed in $\bm{X}_1 \times \bm{X}_2$, and let $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns.
-\(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\).
-In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the original vectors themselves.
-Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs.
-Additionally, we store ``one-hot encodings'' of these patterns rather than the original $\gamma_{ij}$ to aid in vectorized computations.
-See Appendix \ref{app:ohe} for details.
-
-First, we hash record pairs of the same agreement pattern to unique integer values. \cite{enamorado2019using} accomplished this efficiently through the hashing function
+	{\color{red}{NOTE: This section used to be organized 
+		\begin{itemize}
+			\item Hashing
+			\item SEI
+			\item Inference
+		\end{itemize}
+	but now it is organized
+		\begin{itemize}
+		\item Hashing
+		\item Inference
+		\item Scaling (chunking method)
+		\item SEI
+	\end{itemize}
+
+I think inference is much more natural and after hashing.
+And having all the equations and updates written out in the inference part makes it clearer that SEI does not affect posterior updates at all.
+}}
+
+Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
+Let $P$ be the number of patterns realized in $\gamma$ and note that it is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the additional count notes the possibility of missing values.
+This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
+To obtain a memory efficient representation of the data, we map the agreement pattern of each record pair to a unique integer.
+\cite{enamorado2019using} accomplished this through a hashing function, which we modify to explicitly handle missing values:
 \begin{align}
 	\label{eqn:hashing}
-h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}.
-\end{align}
-This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair.
-For computational ease, we map these integers to sequential integers from $\{1, \ldots, P\}$ corresponding to the enumerated patterns in $\mathcal{P}$.
-When the $(i, j)$ pair exhibits pattern $p$, we say $(i, j) \in h_p$.
-When $j$ has no match in $\bm{X}_1$, we write $(n_1 + j, j) \in h_{P+1}$.
-
-With all record pairs converted to hashed values, we identify the records in $\bm{X}_1$ with comparison vectors corresponding to each pattern $p$ for each record $j \in \bm{X}_2$.
-We denote this set $r_{j_p} = \{i \in \bm{X}_1 | (i, j) \in h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{j_p} | j \in \bm{X}_2, p \in 1, \ldots, P\}$.
-This representation is useful because it allows us to easily compute two sets of sufficient statistics.
-First, we compute the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, given by
-\begin{align}
-\label{eqn:N}
-N_{j_p} = |r_{j_p}| = \sum_{i=1}^{n_1} I((i, j) \in h_p).
-\end{align}
-Second, we compute the number of comparison vectors in $\Gamma$ that exhibit pattern $p$, given by
-\begin{align}
-	\label{eqn:H}
-	H_p = \sum_{j = 1}^{n_2} N_{j_p}.
-\end{align}
-We collect these sufficient statistics in the sets $\mathcal{N} = \{N_{j_p} |j \in \bm{X}_2, p \in 1, \ldots, P \}$ and $\mathcal{H} = \{H_{p} | p \in 1, \ldots, P \}$.
-As we show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics.
-Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we can use $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}, \mathcal{H}\}$ for posterior inference.
-
-%With all record pairs converted to hashed values, we can create a more compact nested list 
-%$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
-%where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$.
-
-%We collect these sufficient statistics in the sets $\mathcal{N} = \left\{\{N_{j_p}\}_{j=1}^{n_2}\right\}_{p=1}^P$ and $\mathcal{H} = \{H_{p}\}_{p=1}^P.$
+	h^{*}(\gamma_{ij}) = \sum_{f = 1}^F I_{obs}(\gamma_{ij}^f)2^{\gamma_{ij}^f + I(f>1)\sum_{d=1}^{f-1}(L_d)}.
+\end{align}
+We then map these integers to sequential integers $\{1, \ldots, P\}$.
+We refer to each unique agreement pattern as $h_p$, and to the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
+
+{\color{red}{In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
+This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
+A more complete example of the one-hot encoding is provided in Appendix \ref{app:ohe}.}}
+
+{\color{red}{
+For example, with five fields with binary agreements, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
+Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
+Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
+We map the unique values of $h^{*}(\gamma)$ to sequential integers (for in example in \texttt{R}, by converting the integers into factors, and then back to integers).
+If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
+}}
+
+We then identify the records in $A$ with comparison vectors corresponding to each pattern $p$ for each record $B_j$.
+We denote this set $r_{p_j} = \{i | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
+We then compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$, given by
+\begin{align}\label{eqn:N}
+N_{p_j} = |r_{p_j}| = \sum_{i=1}^{n_A} I(\gamma_{ij} = h_p).
+\end{align}
+We collect these counts in $\mathcal{N} = \{N_{p_j} |p \in 1, \ldots, P, j \in 1, \ldots, n_B \}$.
+
+{
+\color{red}{
+Together, the set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ with no loss of information.
+To see this, we employ the condensed notation
+\begin{align}
+	m_p =  p(\gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}
+\end{align}
+to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match.
+We similarly define $u_p$.
+Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) is equivalent to
+\begin{align}\label{likelihood_efficient}
+	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\gamma}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m_p^{I(Z_j = i)}u_p^{1 - I(Z_j = i)}. 
+\end{align}
+}}
+
+\hypertarget{efficient-posterior}{%
+	\subsection{Efficient Posterior Inference }\label{efficient-posterior}}
+
+Using the likelihood in (\ref{likelihood_efficient}), we can see that the statistics collected in $\tilde{\gamma}$ allow for more efficient posterior inference for all model parameters.
+Conditioning on $\bm{Z}$, we can express the conditional likelihood for $\bm{m}$ and $\bm{u}$ as
+\begin{align}
+	\mathcal{L}(\bm{m}, \bm{u} \mid \gamma, \bm{Z})  &=  \prod_{p=1}^P m_p^{N_p^m}u_p^{N_p - N_p^m}  \label{likelihood_efficient_m_u}
+\end{align}
+where $N_p^m = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ is the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ is the total occurrence of pattern $p$ in the data across all record pairs.
+Note that the $\bm{m}$ and $\bm{u}$ parameters depend on the data only through the quantities calculated from $\mathcal{N}$ and $\mathcal{P}$.
+
+Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define $\bm{\beta_0}$ similarly for the $\bm{u}$ distributions.
+{\color{red}{Using the one-hot encodings described in Section \ref{data-representation-hashing-and-storage},} the terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
+\begin{subequations}
+	\begin{align}
+		\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P N_p^m \times e(h_p), \label{efficient_alpha} \\
+		\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P (N_p - N_p^m) \times e(h_p). \label{efficient_beta}
+	\end{align}
+\end{subequations}
+Specifically, the $l + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}),and (\ref{eqn:beta_update}).
+In this new form however, they are calculated through $P$ multiplications of scalar quantities over the vectors $h_p$, and then vectorized summations.} This is markedly more efficient than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+We can also use (\ref{likelihood_efficient}) to express the conditional likelihood for $Z_{j}$. Using $w_p = \frac{m_p}{u_p}$, we have
+\begin{align}
+	\mathcal{L}(Z_j\mid  \bm{m}, \bm{u}, \tilde{\gamma}) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}\label{likelihood_efficient_z}
+\end{align}
+Note that $Z_j$ interacts with the data only through $\mathcal{N}$ and $\mathcal{R}$.
+This likelihood also lends itself to markedly faster posterior computations.
+Although sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_A$ is large.
+This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
+To speed up computation, we break this sampling step into two.
+We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
+\begin{align}
+	\label{eqn:gibbs1}
+	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
+	\begin{cases} 
+		N_{p_j} \times w_{p},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
+		n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   \text{otherwise}. \\
+	\end{cases}
+\end{align}
+Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the $P$ agreement patterns.
+Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformally using
+\begin{align}
+	\label{eqn:gibbs2}
+	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p \right) = \begin{cases} 
+		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
+		0, & \text{otherwise.} \\
+	\end{cases}
+\end{align}
+
+These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
+We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
+
+\begin{lemma}
+	Recall that $n_A$ and $n_B$ are the number of records in $A$ and $B$, respectively.
+Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $A \times B$.
+We assume $C$ cores available for parallelization and a Gibbs sampler with $T$ iterations.
+Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P)$.
+	\label{lemma:fabl}
+\end{lemma}
+\begin{proof}
+	We consider two steps: constructing the comparison vectors and the Gibbs sampler.
+The computational complexity of all pairwise comparisons across $A$ and $B$ is $O(F n_A n_B)$.
+The hashing procedure for all pairwise comparisons is also $O(F n_A n_B)$.
+With $B$ processors available, we can split these computations across $B$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{B} n_A n_B)$.
+There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
+
+	Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_A n_B)$.
+However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
+The complexity of updating $\bm{Z}$ sequentially at the record level is $O(n_A n_B)$.
+With hashing, we split this sampling into two steps.
+First, we sample the agreement pattern of the match with complexity $O(n_B P)$, and then we sample the record exhibiting that pattern with complexity $O(n_B)$.
+Thus, the complexity of sampling $\bm{Z}$ in a single iteration is $O(n_B P)$.
+Since $P << n_A$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_A n_B)$ under \texttt{BRL} to $O(n_B P)$ under \texttt{fabl}.
+With parallelization, this complexity is further reduced to $O(\frac{1}{C}n_B P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{C}n_B P)$
+	In summary, the total computational complexity is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P).$
+\end{proof}
 
@@
 
-%In practice, it is not feasible to make all-to-all record comparisons because creating and storing $n_1 \times n_2$ comparison vectors is computationally infeasible for large data files.
-The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}.
-In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field.
-The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains.
-However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
-%
-%The hashing procedure described above considerably reduces the memory needed to store the comparison information.
-That is, instead of storing \(n_1 \times n_2\) comparison vectors, we only store the \(P\) unique vectors, and \(n_1 \times n_2\) scalar quantities relating record pairs to those vectors.
-However, even storing these \(n_1 \times n_2\) scalar labels can become burdensome with large data files.
-Worse, the overwhelming majority of these labels relate to record pairs that are clear non-matches.
-%
-%To address this issue, we propose storage efficient indexing (SEI).
-In standard indexing, one decides a priori certain criteria that all linked pairs must satisfy, and labels any record pairs that do not meet that criteria as non-links.
-For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like first name, or only pairs that exactly match on a specified number of fields.
-this kind of indexing requires knowledge of the application of the application and invites room for human error \cite{murray2016probabilistic}.
-
-We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs but greatly reduce the memory costs associated with unlikely matches.
-This allows all-to-all comparisons for substantially larger linkage tasks.
-All records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same $w_{ij}$.
-Therefore, these records have the same probability to be identified as the link for record $j$.
-Thus, we know that records $i \in r_{j_p}$ such that $N_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
-We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
-Therefore, rather than store all of these record labels, we store only a small number \(S\) of them in a new nested list $\mathcal{R}^{\text{SEI}}$.
-Instead of storing \(n_1 \times n_2\) record labels, with SEI we store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.
-In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but is not likely to eliminate any of the indices for plausible matches.
-Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on few matches and distort linkage results.
-Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}, \mathcal{H}\}$.
-
-%For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations.
-For each chunked data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each chunked data file $\bm{X}_{2d}$ to construct the chunked comparison matrix $\Gamma^{cd}$.
-We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.
-These all-to-all comparisons can be computed serially or in parallel.
-Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
-Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger data files.
-
-For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller disjoint chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations.
-For each data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each $\bm{X}_{2d}$ to construct the comparison matrix $\Gamma^{cd}$.
-We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data.
+		For linkage tasks with large amounts of records, we can partition the two data files $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks $\{A^a | a = 1, \ldots, t_A \}$ and $\{B^b | b = 1, \ldots, t_B\}$ for more manageable computations.
+{\color{red}{For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each chunk $A^a$ and $B^b$ would contain 500 records.}} For each data file $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
+We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$ for later calculations, and delete the larger $\gamma^{ab}$ from memory before continuing with the next chunk of data.
 In detail, we calculate
@@
 	\begin{align}
-		r_{j_p}^{cd} &= \{i \in \bm{X}_{1c} | (i, j) \in h_p, j \in \bm{X}_{2d}\}, \label{r_{j_p}^{cd}} \\
-		N_{j_p}^{cd} &= |r_{j_p}^{cd}|, \label{N_{j_p}^{cd}} \\
-		H_p^{cd} &= \sum_{j \in \bm{X}_{2d}} N_{j_p}.\label{H_p^{cd}}
+		r_{p_j}^{ab} &= \{i | \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+		N_{p_j}^{ab} &= |r_{p_j}^{ab}|, \label{N_{p_j}^{ab}}
 	\end{align}
@@
 These can computed serially or in parallel.
-We reduce memory costs through SEI on each $r_{j_p}$, while retaining counts of each agreement pattern through $N_{j_p}^{cd}$ and $H_p^{cd}$.
-Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$.
-Suppressing SEI notation, we combine information through
+Summary statistics from each pairwise chunk comparison can be easily synthesized to recover summary statistics for the full comparison matrix $\gamma$.
+Specifically, we combine information and obtain the statistics comprising $\tilde{\gamma}$ through
 \begin{subequations}
 	\begin{align}
-		r_{j_p} &= (r_{j_p}^{11}, \ldots, r_{j_p}^{cd}) \text{ for } c = 1, \ldots, C \text{ and } d = 1, \ldots, D,\label{r_{j_p}} \\
-		N_{j_p} &= \sum_{c = 1}^C \sum_{d = 1}^D N_{j_p}^{cd},\label{N_{j_p}} \\
-		H_p &= \sum_{c = 1}^C \sum_{d = 1}^D H_p^{cd}.\label{H_p}
-	\end{align}
+		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
+		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab},\label{N_{p_j}}	\end{align}
 \end{subequations}
 
-\hypertarget{efficient-posterior}{%
-	\subsection{Efficient Posterior Inference }\label{efficient-posterior}}
-
-Updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum_f L_f\) many summations over \(n_1 \times n_2\) quantities.
-These are simple calculations, but they become computationally burdensome when working on large linkage tasks.
-
-Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$.
-Let $H_p^m = \\ \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p, Z_j \leq n_1 \right)$ be the number of matching record pairs with agreement pattern $p$.
-It follows that the number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\).
-Let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_f})$ be a concatenated vector of prior parameters for the $\bm{m}$ distributions, and define  $\bm{\beta_0} = (\beta_{11}, \ldots, \beta_{F L_f})$ similarly for the $\bm{u}$ distributions.
-Then, the terms needed for the posterior updates in (\ref{eqn:m_update}) and (\ref{eqn:u_update}) are contained in the vectors
-\begin{subequations}
-\begin{align}
-\bm{\alpha(Z)} &= \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p, \label{efficient_alpha} \\
-\bm{\beta(Z)} &= \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p. \label{efficient_alpha}
-\end{align}
-\end{subequations}
-
-Although sampling \(Z_j\) from the full conditional provided in (\ref{eqn:z_full_conditional}) is conceptually straightforward, it can become computationally expensive when $n_1$ is large.
-This is because sampling a value from $n_1$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$.
-To speed up computation, we break this sampling step into two.
-First, we calculate the quantity expressed in (\ref{eqn:fs_weight}) associated with each unique pattern, and denote this quantity as $w_{p}$.
-We then sample the agreement pattern between \(j\) and its potential link, according to
-\begin{align}
-	\label{eqn:gibbs1}
-	P\left(\left(Z_j^{(s+1)}, j\right) \in h_p \mid \tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
-	\begin{cases} 
-		w_{p}\times N_{j_p},  & p \leq P; \\
-		n_1 \frac{n_2 - n_{12}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{12}(\bm{Z}^{(s)}) + \alpha_{\pi}}, &   p = P + 1. \\
-	\end{cases}
-\end{align}
-%More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,
-%	\begin{align}
-%		\label{eqn:gibbs1}
-%		P\left(h\left(Z_j^{(s+1)}\right) = h_p \mid\tilde{\Gamma}, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right) \propto
-%		\begin{cases} 
-%			w_{p}\times H_{j_p},  & p \leq P; \\
-%			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, &   p = P + 1; \\
-%		\end{cases}
-%	\end{align}
-Since all posterior updates are governed by the agreement patterns of the record pairs rather than the record labels themselves, we complete the entire Gibbs sampler first at the level of the \(P\) agreement patterns.
-Since all records in $\bm{X}_1$ sharing the same agreement pattern with $j \in \bm{X}_2$ are equally likely, we can sample among candidate records uniformally using
-
-%After all iterations are complete, we can back-fill the records uniformally among records of the sampled agreement pattern through (\ref{eqn:gibbs2}).
-\begin{align}
-		\label{eqn:gibbs2}
-		P\left(Z_j^{(s+1)} = z_j \mid \left(Z_j^{(s+1)}, j\right) \in h_p \right) = \begin{cases} 
-			\frac{1}{N_{j_p}}, & (z_j, j) \in h_p; \\
-			0, & \text{otherwise.} \\
-		\end{cases}
-	\end{align} 
-%	\begin{align}
-%		\label{eqn:gibbs2}
-%		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
-%			\frac{1}{H_{j_p}}, & (i, j) \in h_p; \\
-%			0, & \text{otherwise.} \\
-%		\end{cases}
-%	\end{align}
-
-We emphasize the computational gains of this split sampler: the first step is a sample from $P + 1$ options, where $P$ does not scale with the size of the linkage task; the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records.
-These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
-
-The computational complexity of \texttt{fabl} is given in Lemma \ref{lemma:fabl}.
-%, and a summary of the \texttt{fabl} method through psuedocode in Appendix \ref{app:fabl-psuedocode}
-
-%	We close this section by proving the computational complexity of \texttt{fabl}.
-
-%	\begin{remark}
-%	We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record.
-\texttt{BRL} would see considerable speed up through the procedure, but computation time would still scale with the size of the larger dataset.
-%	\end{remark}
-
-%	\hypertarget{computational-complexity}{%
-%		\subsection{Computational Complexity}\label{computational-complexity}}
-
-	\begin{lemma}
-		Recall that $n_1$ and $n_2$ are the number of records in $\bm{X}_1$ and $\bm{X}_2$, respectively.
-Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $\bm{X}_1 \times \bm{X}_2$.
-We assume $B$ processors available for parallelization and a Gibbs sampler with $T$ iterations.
-Then, the overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(\frac{T}{B}n_2 P)$.
-		\label{lemma:fabl}
-	\end{lemma}
-
-	\begin{proof}
-		 We consider two steps: constructing the comparison vectors and the Gibbs sampler.
-The computational complexity of all pairwise comparisons across $\bm{X}_1$ and $\bm{X}_2$ is $O(F n_1 n_2)$.
-The hashing procedure for all pairwise comparisons is also $O(F n_1 n_2)$.
-With $B$ processors available, we can split these computations across $B$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.
-There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
-
-		Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_1 n_2)$.
-However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
-The complexity of updating $\bm{Z}$ sequentially at the record level is $O(n_1 n_2)$.
-With hashing, we split this sampling into two steps.
-First, we sample the agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record exhibiting that pattern with complexity $O(n_2)$.
-Thus, the complexity of sampling $\bm{Z}$ in a single iteration is $O(n_2 P)$.
-Since $P << n_1$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_1 n_2)$ under \texttt{BRL} to $O(n_2 P)$ under \texttt{fabl}.
-With parallelization, this complexity is further reduced to $O(\frac{1}{B}n_2 P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{B}n_2 P)$
-		In summary, the total computational complexity is $O(\frac{F}{B} n_1 n_2) + O(\frac{T}{B}n_2 P).$
-	\end{proof}
+\hypertarget{SEI}{%
+	\subsection{Storage Efficient Indexing}\label{SEI}}
+
+As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses dramatically less memory than storing the full comparison matrix $\gamma$.
+However, the memory requirements of each are still quadratic in nature.
+For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
+We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ for all $n_A \times n_B$ record pairs, while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
+This allows all-to-all comparisons for substantially larger linkage tasks.
+
+{\color{red}{All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$.
+Therefore, these records have the same probability to be identified as the link for record $B_j$.
+Thus, we know that records $i \in r_{p_j}$ such that $N_{p_j}$ is large are unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate.
+We know this regardless of the form of the agreement pattern itself, or its associated probabilities.
+Therefore, rather than store all of these record labels, we store only a small number $S$.
+For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $SEI(r_{p_j}^{ab})$.
+We collect these memory reduced lists to form $SEI(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $SEI(\mathcal{R})$.
+
+Let $n_{A^a}$ and $n_{B^b}$ be the number of records in chunks $A^a$ and $B^b$ respectively.
+Instead of storing $n_A \times n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a} \times n_{B^b} \times P \times S$ labels.
+As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $SEI(\tilde{\gamma}) = \{\mathcal{P}, SEI(\mathcal{R}), \mathcal{N}\}$.
+We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.
+
+NOTE: I switched the order of the chunking and the SEI because SEI only makes sense to use when you split the comparisons into chunks.
+Its also clearer mathematically.
+I think it also avoids the issue of the overloaded notation that used to be in equations \ref{r_{p_j}}.
+The reviewer asked to keep the SEI in the notation to reflect how SEI changed the likelihood and the quantities of interest, but I think the sentence "As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$" makes things very clear.
+Thoughts?}
+}
+
+%{\color{red}{For example, if file $A$ with $n_A = 500$ contains record $A_1$ that is an exact match for record $B_5$ and 499 records with complete disagreement on all fields with record $B_5$, we would only store $S$ indices to represent the records with full disagreement.I added that last example.
+I don't love it though.
+Thoughts?}}
+
+%{\color{red}{When the linkage task is large enough that all $n_A \times n_B$ comparison vectors cannot be stored in memory at the same time, we propose a modified approach where we construct comparison vectors and calculate $\mathcal{R}$ and $\mathcal{N}$ for smaller subsets of the data, and combine this information before conducting the Gibbs sampler.
+We use this approach in the case study in Section \ref{nltcs}, and provide details in Appendix \ref{app:chunks}}}.
 
@@
 We use $F = 5$ binary comparisons with probabilities for matching and non-matching pairs shown in Table \ref{Tab:distributions}.
-For each record in $\bm{X}_2$, we simulate $n_1$ comparison vectors, resulting in a comparison matrix $\Gamma \in \mathbb{R}^{n_1 n_2 \times F}$.
-For $n_2/2$ of these records, there is no match in $\bm{X}_1$, so we simulate $n_1$ comparison vectors from the $\bm{u}$ distribution.
-For the other $n_2/2$ of these records, there is one match in $\bm{X}_1$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_1 - 1$ comparison vectors from the $\bm{u}$ distribution.
-We compare the run-time of \texttt{fabl} against \texttt{BRL} as we increase $n_1$ and $n_2$.
-Since we have five binary comparison fields, the number of unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
-
-%	We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across data files.
+For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$.
+For $n_B/2$ of these records, there is no match in $A$, so we simulate $n_A$ comparison vectors from the $\bm{u}$ distribution.
+For the other $n_B/2$ of these records, there is one match in $A$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_A - 1$ comparison vectors from the $\bm{u}$ distribution.
+We compare the run-time of \texttt{fabl} against \texttt{BRL} as we increase $n_A$ and $n_B$.
+Since we have five binary comparison fields with no missingness, the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
+
+%	We simulate these data for different values of $n_A$ and $n_B$, always with $n_B/2$ records in common across data files.
 and compare the run-time of \texttt{fabl} against \texttt{BRL}.
-Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.
+Since we have 5 binary comparison fields, the number is unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
 
@@
 
-	%\hypertarget{scale}{%
-		%	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
-	%
-	%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated data files from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total).
-Under standard Fellegi-Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size.
-Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task.
-Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.
-	% 
-	%For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact.
-We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons.
-The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method.
-Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run.
-However, this could be substantially sped up using distributed computing.
-	%
-	%Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure.
-This is slightly worse performance than  in the smaller simulation studies; this is expected because it is naturally more difficult to link more records with the same amount of information.
-With more linkage fields, \texttt{fabl} maintains the high accuracy in Section \ref{accuracy}.
-
+	\hypertarget{SEI-sensitivity}{%
+	\subsection{SEI Sensitivity}\label{SEI-sensitivity}}
+
+{
+\color{red}{Finally, our last simulation demonstrates our method's robustness to different values of $S$ for the SEI memory reduction procedure.
+We perform record linkage on one set of synthetic datafile described in Section \ref{accuracy} with 500 records in each datafile, 250 entities in common across datafiles, and 3 errors present across matching records.
+To achieve more drastic results, we perform SEI without chunking the data, with $t_A = t_B = 1$.
+In practice, if it is possible to create and store the comparison matrix for all record pairs at one time, there is no need to reduce the memory of the hashed matrix through SEI.
+For illustration however, it is easier to see degenerate behavior in this setting.
+
+We preform linkage using SEI with $S = (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler. 
+Note that any particular SEI memory reduction may lead to better or worse linkage performance; if the SEI procedure happens to only remove pairs that are not match by random chance, recall and precision will improve.
+Therefore, we perform linkage under each setting 100 times, recording the linkage estimate $\hat{\bm{Z}}$, and recall and precision.
+
+In Figure \ref{fig:SEI_Z},  see that the largest number of distinct linkage estimates occurs when $S = 1$.
+This makes sense, because the SEI procedure arbitrarily removes large numbers of record labels from consideration, resulting in a noisier estimate of the linkage structure.
+The number of distinct linkage estimates decreases as $S$ increases, with larger values of $S$ providing results more similar to the linkage without SEI.
+In Figure \ref{fig:SEI_eval}, we see similar patterns in precision.
+Setting $S=1$ can arbitrarily remove the index of a true match, leading the Gibbs sampler to concentrate probability on a false match, while larger values of $S$ produce results mirroring implementation with no SEI.
+We note however that even with $S=1$, the loss in precision is very small.
+
+Although the figures provided suggest that $S=2$ is adequate for maintaining linkage performance, we suggest a more conservative value like $S=10$.
+When evaluating the performance of a record linkage algorithm, researchers often examine posterior probabilities.
+By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce suspiciously high posterior probability for certain record pairs, providing a warped perception of model performance.
+We want to be sure that the set of records pairs exhibiting agreement patterns that have considerable probability of matching do not get reduced through SEI.}
+}
+
+%\begin{figure}[t]
+%	\begin{center}
+%		\includegraphics[width=0.6\textwidth]{../SEI_sensitivity/Z_plot}
+%		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
+%		\label{fig:SEI_Z}
+%	\end{center}
+%\end{figure}
+%
+%\begin{figure}[t]
+%	\begin{center}
+%		\includegraphics[width=0.6\textwidth]{../SEI_sensitivity/eval_plot} 
+%		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation}
+%		\label{fig:SEI_eval}
+%	\end{center}
+%\end{figure}
+%
+%  \begin{figure}[hbt]
+%	\centering
+%	\begin{subfigure}{0.4\textwidth}
+%		\includegraphics[width=\textwidth]{../SEI_sensitivity/Z_plot} 
+%		\caption{}\label{fig:SEI-Z}
+%	\end{subfigure}\hspace{\columnsep}%
+%	\begin{subfigure}{0.4\textwidth}
+%		\includegraphics[width=\textwidth]{../SEI_sensitivity/eval_plot} 
+%		\caption{}\label{fig:SEI-eval}
+%	\end{subfigure}
+%\caption{Simulation results Section \ref{SEI-sensitivity}. In \ref{fig:SEI-Z}, 
+%\end{figure}
+%Figure \ref{fig:here}
+
+\begin{figure}
+	\begin{minipage}[c]{0.4\linewidth}
+		\begin{center}
+		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
+		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
+		\label{fig:SEI_Z}
+		\end{center}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[c]{0.5\linewidth}
+		\begin{center}
+		\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
+		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation}
+		\label{fig:SEI_eval}
+	\end{center}
+	\end{minipage}%
+\end{figure}
 	\section{Case Studies}
@@
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We use uniform priors for the \(\bm{m}\) and \(\bm{u}\) parameters.
+We use uniform priors for the $\bm{m}$ and $\bm{u}$ parameters.
 
 	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
+This results in $5^5 \times 3 = 6025$ possible agreement patterns, with 1173 patterns realized in the data.
 However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
-
-%	\cite{sadinle_bayesian_2017} constructs the comparison vectors using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in \(4^5 \times 2 = 2048\) possible agreement patterns, with 1173 patterns realized in the data.
-We initially used these comparisons to estimate fabl and BRL.
-However, after investigating the posterior distributions of the $\bm{m}$ and $\bm{u}$ parameters, we noticed that posterior distributions of several levels of the m and u parameters are nearly identical, suggesting that these levels are unnecessary for meaningfully modeling the distribution of the comparison vectors.
-
-	%However, we observed that posterior distributions of several levels of the \(\bm{m}\) and \(\bm{u}\) parameters were nearly identical, suggesting that the number of levels used are unnecessary for meaningfully representing the data.
-
-	%here, the number of patterns was sufficiently high enough that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
 
@@
 
-	%Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.
-
-	%Throughout our paper, we have discussed independent pairwise matching assumptions, one-to-one matching assumptions, and when such assumptions are not met in practice.
-We explored ``one-to-many'' scenarios in which one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and showed how both \texttt{fabl} and \texttt{BRL} demonstrated undesirable qualities.
-Other issues arise under ``many-to-one'' scenarios, where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, and ``many-to-many'' scenarios in which there is duplication both across and within data files.
-Extending \texttt{fabl} in such scenarios with a more robust linkage structure parameter $\bm{Z}$ is a topic for future work.
-	%
-	%To our knowledge, \texttt{fabl} is the first method within the Fellegi-Sunter comparison vector approaches with an interpretable parameter for the rate of matching across data files.
-This motivates extensions to \texttt{fabl} with additional parameters that allow the quality of information (the $\bm{m}$ and $\bm{u}$ parameters) and the rate of matching (the $\lambda$ parameter) to differ by subgroups in the linkage task.
-Similar work has been been proposed by \cite{wortman2019} and \cite{Larsen2005}, and would need to be adapted for our prior distribution and model assumptions.
-We see these types of heirarchical record linkage models as potential contributions to the growing algorithmic fairness literature, and useful to practitioners working with or studying marginalized populations.
-
-	%Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
-	%future work.
-
-	%\begin{figure}[t]
-	%	
-	%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
-		%		
-		%	}
-	%	
-	%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
-	%\end{figure}
-
-	%\subsection{National Long Term Care Survey}
-	%\label{nltcs} 
-	%
-	%The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients.
-It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989.
-The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office.
-State and regional office have been anonymized, presented as numbers instead of their actual values.
-We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context.
-The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466$ files, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns.
-We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.
-	%
-	%With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible.
-Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold.
-Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller data files.
-We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing.
-We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler.
-	%
-	%Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$.
-However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds.
-Posterior analysis yields an initial estimate of 9991 individuals linked across the the two data files, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching.
-	%
-	%Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking.
-With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks.
-This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler.
-However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error.
-In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office.
-We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives.
-Even if not, the modeler may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures.
-However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons.
-Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block.
-Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.
-
 	\clearpage
@@
 Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$.
-In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
+In the event that we have two records $B_j$ and $B_{j'}$ such that both $P(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
 Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
@@
 %			\hline
-%			$\bm{X}_1, \bm{X}_2$ & data files\\
-%			$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
-%			$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
+%			$A, B$ & data files\\
+%			$i \in 1, \ldots, n_A $ & index over records in $A$\\
+%			$j \in 1, \ldots, n_B $ & index over records in $B$\\
 %			$f \in 1, \ldots F$ & index over fields used for comparisons \\
 %			$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
-%			$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
-%			$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
-%			$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
-%			$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
+%			$n_{AB}$ & number of entities in common between $A$ and $B$\\
+%			$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+%			$Z_j = i$ & records $A_i$ and $B_j$ match \\
+%			$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
 %			$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
 %			$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-%			$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
-%			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+%			$\pi$ & probability that a record in $B$ has a match in $A$ \\
+%			$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
 %			$h_p$ & one hot encoding of agreement pattern $p$ \\
-%			$H_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
+%			$H_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
 %			$H_p$ & number of total comparison vectors that exhibit pattern $p$ \\
@@
 \begin{align*}
-	p(m_{fl}| \Gamma, \bf{Z}) &\propto \prod_{i=1}^{n_1} \prod_{j=1}^{n_2} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
+	p(m_{fl}| \gamma, \bm{Z}) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)} \times  m_{fl}^{\alpha_{fl} - 1} \\
 	&= m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
 \end{align*}
-where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$.
-Analogous procedures lead to the posterior distribution $p(u_{fl}| \Gamma, \bf{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.
+where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
+Analogous procedures lead to the posterior distribution $p(u_{fl}| \gamma, \bm{Z})$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
 Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
 
-$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
-$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})).$$
-
-In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihood relevant to a particular record $j \in \bm{X}_2$.
-Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters).
-Thus, we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
+$$\bm{m}_f|\bm{Z}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
+$$\bm{u}_f|\bm{Z}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})).$$
+
+Since $\pi$ encodes the rate of matching across the two data files, the posterior distribution $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ (and hyperparameters).
+Thus, we use $p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi})$ and have
 \begin{align*}
 	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}.
+	&\propto \pi^{n_{AB}(\bm{Z})} (1-\pi)^{n_B - n_{AB}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+	&\propto \pi^{n_{AB}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_A - n_{AB}(\bm{Z}) + \beta_{\pi} -1}.
 \end{align*}
-Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
-
-Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\). We have
-\begin{align*}
-	p(\Gamma_{.j}|Z_j = z_j, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
-	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
-	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&=
+Thus $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})$ distribution.
+
+To express the full conditional for $\bm{Z}$, we consider the likelihood in two cases.
+Because we sample each $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$.
+Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+\end{align}
+When $Z_j = q$ for some $q\leq n_A$, we have
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)} \times \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+\end{align}
+We multiply and divide by the $u$ parameters for the matching record pair to obtain
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)} \times \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	&= w_{qj} \times c_j.
+\end{align}
+We can divide the result of each case by $c_j$ to get
+\begin{align}
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j) = \begin{cases} 
+		w_{qj}, & q \leq n_A; \\
+		1, &  q  = n_A + j. \
+	\end{cases}
+\end{align}
+Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain
+\begin{align}
+	\label{eqn:z_full_conditional2}
+	p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
 	\begin{cases} 
-		w_{z_j, j}  & z_j \leq n_1; \\
-		1 &  z_j  = n_1 + j, \\
-	\end{cases}\\
-\end{align*}
-where
-$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
-
-With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$.
-Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\):
-\newpage
-\begin{align*}
-	p&\left(Z_j^{(s+1)} = z_j|\Gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
-	&\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
-	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
-	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_1 + j). \\
-	&= \begin{cases} 
-		\frac{\pi}{n_1}w_{z_{j}, j}   & z_j \leq n_1; \\
-		1-\pi &  z_j  = n_1 + j. \\
+		\frac{\pi^{(s+1)}}{n_A} w_{q, j}^{(s+1)},  & q \leq n_A; \\
+		1 - \pi^{(s+1)}, & q  = n_A + j. \
 	\end{cases}
-\end{align*}
-
-For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
-\begin{align*}
-	p&\left(Z_j^{(s+1)}  = z_j| \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
-	&= \int_{\pi}p\left(Z_j^{(s+1)} | \Gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-	&= \int_{\pi}\left[\frac{\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + (1-\pi)I(z_j = n_1 + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
-	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_1}w_{z_{j}, j} I(z_j\leq n_1) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(z_j = n_1 + j) \\
-	&= \frac{n_{12}(\bm{Z}) + \alpha_{\pi}}{n_1 (n_2 + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(z_j\leq n_1) + \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_2 + \alpha_{\pi} + \beta_{\pi}}I(z_j = n_1 + j) \\
-	&\propto w_{z_j, j}I(z_j\leq n_1) + n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}I(z_j = n_1 + j) \\
-	&=
-	\begin{cases} 
-		w_{z_j, j}  & z_j \leq n_1; \\
-		n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
-	\end{cases} \\
-\end{align*}
+\end{align}
+
+\textcolor{blue}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
+\begin{align}
+p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi) \propto \prod_q \left(\prod_j
+p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi \right)
+\right).
+\end{align}
+Therefore, we update the full conditional for $\bm{Z}^{(s+1)}$ in parallel.
+This contrasts the Gibbs sampler of  \cite{sadinle_bayesian_2017}, where the update to the matching matrix is sequential and accomplished using a partially collapsed Gibbs sampler, where one integrates out $\pi.$ We do not opt for this approach as it would lead to similar sequential updates, which are computationally slow.}
+
+%Let $\gamma_{.j}$ denote the set of $n_A$ comparison vectors with $B_j$. We have
+%\begin{align*}
+%	p(\gamma_{.j}|Z_j = q, \bm{m}, \bm{u}) &\propto \prod_{i=1}^{n_A}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
+%	&\propto \prod_{i=1}^{n_A}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
+%	&\propto \prod_{i=1}^{n_A}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(q = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+%	&=
+%	\begin{cases} 
+%		w_{q, j}  & q \leq n_A; \\
+%		1 &  q  = n_A + j. \
+%	\end{cases}\\
+%\end{align*}
+%where
+%$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
+
+%With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$.
+Because we sample each $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$:
+%\newpage
+%\begin{align*}
+%	p&\left(Z_j^{(s+1)} = q|\gamma_{.j}, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) \\
+%	&\propto p(\gamma_{.j}| Z_j^{(s+1)}, \bm{m}, \bm{u}) P(Z_j^{(s+1)}|\pi) \\
+%	&\propto \left(\sum_{i=1}^{n_A}w_{q, j}I(q = i) + I(q = n_A + j) \right) \left (\pi\sum_{i=1}^{n_A}\frac{1}{n_A}I(q = i) + (1-\pi)I(q = n_A + j) \right) \\	
+%	&= \frac{\pi}{n_A}\sum_{i=1}^{n_A}w_{q, j}I(q = i) + (1-\pi)I(q = n_A + j). \\
+%	&= \begin{cases} 
+%		\frac{\pi}{n_A}w_{q, j}   & q \leq n_A; \\
+%		1-\pi &  q  = n_A + j. \\
+%	\end{cases}
+%\end{align*}
+%
+%For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
+%\begin{align*}
+%	p&\left(Z_j^{(s+1)}  = q| \gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}\right)\\
+%	&= \int_{\pi}p\left(Z_j^{(s+1)} | \gamma, \bm{m}, \bm{u}, \bm{Z^{(s)}}, \pi\right) p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+%	&= \int_{\pi}\left[\frac{\pi}{n_A}w_{z_{j}, j} I(q\leq n_A) + (1-\pi)I(q = n_A + j)\right] p\left(\pi| \bm{Z^{(s)}}\right) d\pi \\
+%	&= \frac{\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi}{n_A}w_{z_{j}, j} I(q\leq n_A) + \left(1-\int_{\pi} \pi p\left(\pi| \bm{Z^{(s)}}\right) d\pi \right) I(q = n_A + j) \\
+%	&= \frac{n_{AB}(\bm{Z}) + \alpha_{\pi}}{n_A (n_B + \alpha_{\pi} + \beta_{\pi})}w_{z_{j}, j} I(q\leq n_A) + \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_B + \alpha_{\pi} + \beta_{\pi}}I(q = n_A + j) \\
+%	&\propto w_{q, j}I(q\leq n_A) + n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}}I(q = n_A + j) \\
+%	&=
+%	\begin{cases} 
+%		w_{q, j}  & q \leq n_A; \\
+%		n_A \frac{n_B - n_{AB}(\bm{Z}) + \beta_{\pi}}{n_{AB}(\bm{Z}) + \alpha_{\pi}} & q   = n_A + j. \\
+%	\end{cases} \\
+%\end{align*}
 

### 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
cleaning up repo for submission and in case jerry needs to make edits to tex.

@@
 	\begin{figure}[t]
-		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
+		\begin{center}
+		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_speed_plot2} 
+	% \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
@@
 	\begin{figure}[h!]
-		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
+		\begin{center} 
+		\includegraphics[width=0.7\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides} 
+	%	\includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
@@
 		\begin{center}
-			\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
+		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_sim_plot2} 
+			%\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
 			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
@@
 		\begin{center}
-		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
+		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/Z_plot}
+%		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
 		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
@@
 		\begin{center}
-		\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
+		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/eval_plot} 
+%	\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
 		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation.}
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_distribution_smallP_bayes}
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
 			\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
+		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_posterior_smallP} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
 			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
+		        \includegraphics[width=0.6\textwidth]{finalFigures/nltcs/overlap_posterior4}
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
 			\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
@@
 	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
+	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
+		%\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
 		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
+		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
 			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
+		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
 			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
+		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
 			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
@@
 	\begin{figure}[h!]
-		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
+		\begin{center} 
+		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
+		%\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
@@
 	\begin{figure}[h!]
-		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
+		\begin{center}
+		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
+		% \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
+		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
 			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} \label{fig:m_trace}
+		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
+			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} 
+			\label{fig:m_trace}
 		\end{center}
@@
 		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} \label{fig:u_trace}
+		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
+			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} 
+			\label{fig:u_trace}
 		\end{center}

### 2023-07-20 presentations/bcm_proposal.tex
cleaning up repo for submission and in case jerry needs to make edits to tex.

--- 2023-06-07 Prelim/bcm_proposal.tex
@@
 	\insertframenumber / \inserttotalframenumber }
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	linkcolor=blue,
+	filecolor=magenta,      
+	urlcolor=cyan,
+	pdftitle={Overleaf Example},
+	pdfpagemode=FullScreen,
+}
 
 \begin{document}
-	\title{Bayesian Causal Matching with Fellegi Sunter}
+	\title{Joint Matching and Treatment Effect Estimation in Causal Inference through Bayesian Fellegi-Sunter Methods}
 	%\subtitle{Using Beamer}
@@
 
-\begin{frame}{Mahalonobis Distance}
+\begin{frame}{Mahalonobis Distance and Propensity Matching}
 	\begin{itemize}
-	\item One method of matching has been to calculate distance in the covariate space through
+	\item One approach to matching is to use the Mahalonobis distance, which measures distance between units purely in the covariate space, without regard to which features are related to the outcome of interest.
+It is defined by
 		\begin{align}
@@
 		\end{align}
-	\item An algorithm then matches each unit in control to closest unit in treated
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Propensity Score}
-	\begin{itemize}
 	\item Another method is to estimate propensity scores $p(Z_n = 1 |X_n)$ for each unit in the data, and match units based on this score.
+	\item Most matching methods in causal inference then use deterministic algorithms to match treated units to their nearest control unit based on the chosen metric
 	\end{itemize}
@@
 		\item The pros and cons of each method are summarized very well on this \href{https://stats.stackexchange.com/questions/511294/what-are-the-pros-and-cons-of-using-mahalanobis-distance-instead-of-propensity-s}{Stack Exchange} page
-
 		\item Essentially, propensity scores can end up matching units with that have similar probability of treatment, but have very different covariate values
@@
 
-\begin{frame}{Blending the Two Approaches}
+\begin{frame}{Almost Exact Matching (AME)}
 	\begin{itemize}
-		\item There is also a good summary on page 5 of \href{https://www.jstatsoft.org/article/view/v042i07}{Sekhot (2011)} recommends including the propensity score into the Malahalonobis distance.
-This was the genesis for the idea to come
+		\item I am new to these readings, but it seems like most of them
+		\begin{itemize}
+			\item Learn a scoring function (or equivalently, weights on individuals covariates)
+			\item Apply scoring function to all units, and find units 
+		\end{itemize}
 	\end{itemize}
@@
 
-\section{Proposed Idea}
+\begin{frame}{Additional Thoughts}
+	\begin{itemize}
+		\item  \href{https://www.jstatsoft.org/article/view/v042i07}{Sekhot (2011)} recommends including the propensity score into the Malahalonobis distance.
+This is just practical advice, but it forms the foundation of thoughts to come
+		\item \href{https://www.lexjansen.com/pharmasug/2006/PublicHealthResearch/PR05.pdf}{This paper} estimates PS, identifies all units in the treatment have PS within a specified caliper, and then chooses the unit with the smallest Mahalanobis distance.
+No uncertainty quantification
+		\item \href{https://arxiv.org/pdf/2105.02362.pdf}{Alvarez and Levin (2021)} (Section 2, pages 5 - 7) use a Bayesian model to create draws of the propensity scores.
+For set of scores, they use a deterministic algorithm to match records.
+They do not use any other covariates
+	\end{itemize}
+\end{frame}
 
-\begin{frame}{My Idea}
+\section{New Research Idea}
+
+\begin{frame}{Intuition}
 	\begin{itemize}
@@
 
-\begin{frame}{A Similar Idea}
+\begin{frame}{Nuts and Bolts}
 	\begin{itemize}
-		\item\href{https://arxiv.org/pdf/2105.02362.pdf}{Alvarez and Levin (2021)} (Section 2, pages 5 - 7) use a Bayesian model to create draws of the propensity scores.
-For set of scores, they use a deterministic algorithm to match records.
-		\item I imagine that each particular draw of matched pairs may have the issues presented earlier. Propensities are matched, but units may actually look quite different.
-		\item Information does not seem to propagate correctly through the Gibbs Sampler
+		\item Data is $\{y_n, z_n, x_n\}_{n = 1}^N$ where $y_n$ is the outcome, $z_n$ is the treatment indicator, and $x_n$ is a vector of $F$ fields of covariates. 
+		\item Split data into a datasets $T$ with records $i \in \{1, \ldots, n_T\}$ for treated units and $C$ with records $j \in \{1, \ldots, n_C\}$. 
+		\item At each step of the Gibbs Sampler, calculate propensity score $PS(n) = P(Z_n = 1 | X)$ for each data point (logistic or probit regression)
 	\end{itemize}
@@
 
-\begin{frame}{My Idea}
-	\begin{itemize}
-		\item Data is $\{y_n, z_n, x_n\}_{n = 1}^N$ where $y_n$ is the outcome, $z_n$ is the treatment indicator, and $x_n$ is a vector of $F$ fields of covariates. 
-		\item Split data into a datasets $T$ with records $i \in \{1, \ldots, n_T\}$ for treated units and $C$ with records $j \in \{1, \ldots, n_C\}$. 
-		\item At each step of the Gibbs Sampler, calculate propensity score $PS(i) = P(Z_i = 1 | X)$ for each data point (logistic or probit regression)
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{My Idea}
+\begin{frame}{Nuts and Bolts}
 	\begin{itemize}
 		\item Define comparison field 
-		\begin{align}
+		\begin{align*}
 			\Gamma_{0} = \begin{cases}
-				1 & $PS(i)$ and $PS(j)$ \text{ are sufficently close} \\
+				1 & $PS(i)$ \text{ and } $PS(j)$ \text{ are within specified caliper} \\
 				0 & \text{otherwise} \\
 			\end{cases}
-		\end{align} 
-		\item ''Sufficiently close" could mean all units with scores within a certain caliper, or it could mean the $K$ nearest scores.
+		\end{align*} 
 		\item Conduct Fellegi Sunter Matching on new full set of comparison vectors $\Gamma^{*} = (\Gamma_0, \gamma)$.
@@
 
-\begin{frame}
+\begin{frame}{Big Picture}
 	\begin{itemize}
@@
 
+\section{Proposed Models}
+
 \begin{frame}{Model (Basic)}
-	\begin{align}
+	\begin{align*}
 		Z_n &= x_n' \beta + \epsilon \\
@@
 		\epsilon &\sim N(0, \sigma^2) \\
-	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l) \\
-		m, u \sim \text{Dirichlet}
-		\Delta \sim \text{Desired Record Linkage Prior}
-	\end{align}
+	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l)} \\
+		m, u &\sim \text{Dirichlet} \\
+		\Delta &\sim \text{Desired Record Linkage Prior}
+	\end{align*}
+Different record linkage priors can induce one-to-one, many-to-one, and many-to-many matches. Any one of them can work within this framework. 
 \end{frame}
@@
 	\end{itemize}
-\begin{align}
-	\beta_j &\sim \pi_0 \delta_0 + \(1 - \pi\)N(0, \tau_j) \\
-	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\left[\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l)\right]^{\beta_j \neq 0} \\
-\end{align}
+\begin{align*}
+	\beta_j & \sim \pi_0 \delta_0 + (1 - \pi_0)N(0, \tau_j) \\
+	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= 
+	\prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\left[\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{1 - \Delta_{ij} = 1}\right]^{I(\gamma_{ij}^f = l)}\right]^{\beta_f \neq 0} \\
+\end{align*}
 \end{frame}
@@
 \begin{frame}{Full Conditional for $\beta$}
-	\begin{itemize}
-		\item The full condition for $\beta_j$ includes a measure of the quality of matches along field $f$
-	\end{itemize}
+	Reexpress the likelihood as
+		\begin{align*}
+			\mathcal{L}(\Delta, m, u, \beta \mid \Gamma^{*}) &= 
+			\prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F} c_f \prod_{l=1}^{L_f}\left(  \frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I(\Delta_{ij} = 1)I(\beta_f \neq 0)} \\
+			&= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F} c_f \prod_{l=1}^{L_f}w_{fl}^{I(\gamma_{ij}^f = l)I(\Delta_{ij} = 1)I(\beta_f \neq 0)}
+		\end{align*}
+where $w_{fl} =  \frac{m_{fl}}{u_{fl}}$. Let $M_{fl} = \sum_{i, j} I(\Delta_{ij} = 1)I(\gamma_{ij}^f = l)$.
+
+Then, conditional on $\Delta$, we have 
+	\begin{align*}
+	\mathcal{L}(m, u, \beta \mid \Gamma^{*}, \Delta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F} \left[c_f \prod_{l=1}^{L_f}w_{fl}^{M_{fl}}\right]^{I(\beta_f \neq 0)}
+\end{align*}
+\end{frame}
+
+\begin{frame}{Full Conditional for $\beta$}
+	The quantity $\prod_{l=1}^{L_f} w_{fl}^{M_{fl}}$ is measure of how strong field $f$ is in the linking procedure.
+And it shows up in the full conditional for $\beta_f$, showing how the regression and matching procedure are linked.
 \end{frame}
@@
 
-\begin{frame}{Conclusion}
-	\begin{itemize}
-		\item Mathematically, any record linkage prior could work. Sadinle, fabl, or original FS (one-to-one, one-to-many, many-to-many) may all be useful , depending on the analysis
-		\item If this works, would be a very strong final thesis chapter
-	\end{itemize}
-\end{frame}
-
 \begin{frame}{Caveats}
@@
 		\item Using a comparison vector made from the propensity score may technically violate the conditional independence assumption of FS
-		\item I can't imagine this idea totally failing, but I'm sure that it won't be cutting edge. Sekhot further devised a method that estimates
-		\begin{align}
-			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1/2} W S^{-1/2} |X_1 - X_2|}
-		\end{align}
-		where the entries in $W$ are fine tuned through optimization methods that encourage covariate balance across datasets, based on t-tests and all sorts of other measures.
+%		\item I can't imagine this idea totally failing, but I'm sure that it won't be cutting edge. Sekhot further devised a method that estimates
+%		\begin{align}
+%			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1/2} W S^{-1/2} |X_1 - X_2|}
+%		\end{align}
+%		where the entries in $W$ are fine tuned through optimization methods that encourage covariate balance across datasets, based on t-tests and all sorts of other measures.
 This is almost certainly outperform what I have presented, because it is able to use more information.
@@
 		\item I'm really struck by the conceptual similarity between Mahalanobis distance and the FS weights.
-If you make some assumptions (categorical data, one-to-one matching), it might be possible to prove a correspondence between the two.
+If you make some assumptions (categorical data, one-to-one matching, known parameters), it might be possible to prove a correspondence between the two.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Possibly Better Approach}
+	\begin{itemize}
+		\item The FS model assumes that agreement levels are independent conditional on the matching status of the ``record" pair. This is often violated, with minimal issues.
+
+		However, the propensity score is very much not independent from other agreement levels.
+It may therefore to incorporate proximity in PS not as a feature, but rather as a \emph{latent class} for the agreement vector.
+
+		See \href{https://projecteuclid.org/journals/annals-of-applied-statistics/volume-13/issue-3/Incorporating-conditional-dependence-in-latent-class-models-for-probabilistic-record/10.1214/19-AOAS1256.full}{Xu et.
+al (2019)} for more on this.
+
+		I think the first method would work well, and presents enough innovation to stand on its own. I can do this latent class stuff in an extension. 
 	\end{itemize}

### 2023-07-20 ba_resubmit/rebuttal/response.tex
organize response folder.

--- 2023-07-19 ba_resubmit/response_to_reviewers2.tex
@@
 
-Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.
+\textcolor{teal}{We thank the Editor and reviewers for their assessment and constructive feedback.}
+
+%Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.
 
@@
 
-%% RCS: Suggestion: Write a strong paragraph to the Editor after the paper is finalized.
-Come back to this.
-Explain that we have tried to simplify the notation when possible, however, in some places, the heavy notation is not possible.
-Thus, we have worked on improving more precise wording and removing sections or wording that is distracting or not needed.
-Thank the Editor and Associate Editor for helping us improve upon what we believe to be an important paper and breakthrough in the literature.
+\reply{
+\textcolor{teal}{Thank you for the opportunity to submit a revision. 
+We have provided detailed responses to the reviewers' comments below, 
+which prompted significant changes in our article. Specifically, we have attempted to improve notation and wording choices, at the reviewers' suggestion. 
+}
+}
 
-%% I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.
-
-\reply{
-We thank the Editor for the opportunity to revise our paper and for the suggestions.
-We have improved notation, when possible, however, as noted by the Editor in certain places, such as in the computational speeds ups of Section 4, this is difficult to avoid.
-We have done our utmost to simplify notation (when possible) and improve the parsing of sentences.
-Finally, we have attempted to clarify our overall contributions to the literature, which we highlight below.
-
-We propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
+%We propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
 We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.
 This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the decision theoretic  technique of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.
-This discovery allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We employ the decision theoretic  technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.
+This discovery allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs via a novel method, storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
@@
 Finally, we provide two real illustrations of our methodology and simulation studies with reproducible software.
-}
+
+%}
 
@@
 \reply{
-We thank the Editor for bringing up this point. In order to simplify notation, we refer to the two data files as
+Thank you for the suggestion. We now refer to the two data files as
  $A$ and $B$, where the updated material reads:
@@
 
-	\reply{The distinction does not matter due to assuming that the comparison vectors are conditionally independent given the coreference matrix (or matching label).
-We now state the following in Section 3, page 5:
+	\reply{\textcolor{teal}{The distinction does not matter due to assuming that the comparison vectors are conditionally independent given the coreference matrix (or matching label).
+We now state the following in Section 3, page 5:}
 
-	``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
-With either the ignorability or MAR assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
-Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p.
-12]{LittleRubin2002}).
-To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation."
+	\textcolor{teal}{``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
+Due the the conditional independence assumption, we can marginalize over the missing data in (6a) and perform all computation using the observed data.
+For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
+
 	}
-
-	%% RCS: See two comments below regarding MAR versus MCAR. I have an alternative response for the one regarding miss at random.
-
-%	\textcolor{teal}{Answer for MAR: Yes, the assumption is truly missing at random.
-This distinction between truly missing at random and missing completely at random does not matter due to two assumptions.
-Following Sadinle (2017), we assume that the fields in the comparison vectors are conditionally independent given the coreference matrix and we assume that the missing comparisons are ignorable (Sadinle (2017).
-Therefore, our likelihood is the same as in Sadinle (2017) (see equations (3) and (4) of his paper).
-As stated by Sadinle (2017), ``we work under the assumption of ignorability of the missingness mechanism for the comparisons so that we can base our inferences on the marginal distribution of the observed comparisons (Little and Rubin, p.
-90)."  We refer to Sadinle (2017), Section 4.2, page 604 for further details.}
-
-%	\reply{The assumption is missing completely at random (MCAR) due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix (or matching label).
-The assumption is not missing at random (MAR) as we do not assume that the missing comparisons are also ignorable.
-See Section 3, page 5 regarding clarification of this point \citep{LittleRubin2002}.}
-	\textcolor{red}{Still checking on this and looking at Little and Rubin.}
-
-%	\reply{% 
-%		The assumption is missing completely at random. You are correct that this is a result of the independence assumed across elements of the comparison vector. This has been clarified. 
-%	}
 
@@
 	The square brackets have been corrected.
-We have clarified the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-Refer to Appendix B, page 26 -- 27 regarding this extended derivation.
+We have revised the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
+Refer to Appendix A, pages 26 -- 27 regarding this extended derivation.
 	}
@@
 
-\reply{No, we do not integrate out $\pi$ as done in \cite{sadinle_bayesian_2017} as this would result in a sequential sampler as opposed to the parallel one that we propose.
-This is one of our contributions to our paper.
+\reply{No, in \text{fabl}, we do not integrate out $\pi$ as done in \cite{sadinle_bayesian_2017} because this leads to a sequential sampler.
 
-(1) Yes, our alternative approach is different than directly integrating over the prior distribution $\pi.$ We have attempted to make this more clear in our revised version.
-In short, our prior distribution over $Z$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of Sadinle (2018).
-Of course, we could integrate out $\pi$ in our approach, however, this would lead to sequential updates, which would be slow in practice.
-See Appendix B, page 27.
+(1) Yes, our alternative approach is different than directly integrating over the prior distribution $\pi.$ 
+%We have attempted to make this more clear in our revised version. 
+In short, our prior distribution over $Z$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of \cite{sadinle_bayesian_2017}.
+Of course, we could integrate out $\pi$ in our approach, however, this leads to sequential updates, which would be slow in practice.
+See Appendix A, page 27.
 
 (2) Yes, our alternative approach is justified.
-We provide a justification in Appendix B, page 27 regarding why the updates are independent and not sequential.
+We provide a justification in Appendix A, page 27 regarding why the updates are independent and not sequential.
 Thank you for the excellent questions, which have greatly improved our paper.
@@
 
-	\reply{Yes, in the original manuscript, the comparison vectors are created and therefore, patterns are assigned) without regard to $Z$.
-Given the confusion caused to the Associate Editor, we removed the sentence.	}
+	\reply{\textcolor{teal}{Yes, in the original manuscript, the comparison vectors are created and therefore, patterns are assigned) without regard to $Z$.
+Given the confusion, we have removed the sentence.	}}
 
@@
 \reply{%
-	We use $Z$ when discussing a random quantity, and $z$ to reference a realized value. We have revised all instances of inconsistency. 
+	We use $Z$ when discussing a random quantity, and $z$ to reference a realized value. We have revised all instances of the inconsistency. 
 }
@@
 \reply{%
-	This has been fixed. 
+	This has been fixed. \textcolor{teal}{Refer to equations (12) and (13), page 7.}
 }
@@
 \reply{%
-	This has been revised; thank you.
+	This has been revised; thank you. \textcolor{teal}{Refer to Section 4.2, page 10.}
 }

### 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
update.

@@
 Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
-Due the the conditional independence assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data.
+It is important to note that for likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
+Therefore, we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data.
+For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
+}
+
+%Due the the conditional independence assumption, MAR or MCAR lead to the same treatment of the missing data.
+That is, for likelihood based inference both MAR and MCAR lead to ignorability, or rather, both lead to being able to consider just the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
+
+%we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data.
 For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}

### 2023-07-20 ba_resubmit/rebuttal/response.tex
revised wording.

@@
 
-	\textcolor{teal}{``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
-Due the the conditional independence assumption, we can marginalize over the missing data in (6a) and perform all computation using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
-
+	\textcolor{teal}{``\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
+To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
+It is important to note that for likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
+Therefore, we can marginalize over the missing data and perform all computation using the observed data, leading to (6a).
+For further details on these connections, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}}."}
 	}

### 2023-07-20 ba_resubmit/kundinger_final.tex
done with pass. sent to jerry and brian.

--- 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
@@
 
+\textcolor{red}{Jerry: could you read through this carefully please regarding the missing data part? Note: we only give the marginalized likelihood and not the likelihood without the missing data.}
+
 \textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
@@
 It is important to note that for likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
-Therefore, we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
+Therefore, we can marginalize over the missing data and perform all computation using the observed data, leading to (\ref{eqn:likelihood}).
+For further details on these connections, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
 }

### 2023-07-20 ba_resubmit/kundinger_final.tex
putting jerry's proposed changes in red text.

@@
 To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
-Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
-\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
-\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
+Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, \textcolor{red}{unlike} the sequential updates of \texttt{BRL}.}
+\textcolor{teal}{We ensure \textcolor{red}{\text{fabl} results in a bipartite matching} using decision theoretic post-processing techniques from the existing literature.}
+\textcolor{red}{To reduce computational overhead and memory costs, we use hashing and a new technique that we call storage efficient indexing. }
+%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
 Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
-}
+%}
 %This allows
@@
 \begin{keyword}
-\kwd{bipartite record linkage}
+\kwd{record linkage}
 \kwd{entity resolution}
@@
 In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. 
-{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
+\textcolor{red}{We use independent prior distributions for the matching status of each record, an idea suggested in \cite{wortman2019}.}
+%% Jerry thought the wording below was too vague and made a suggestion that is more direct and clear. 
+%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}}
+%
+\textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, \textcolor{red}{unlike the} sequential updates of \cite{sadinle_bayesian_2017}.
 We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
-This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+This \textcolor{red}{approach} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
@@
 He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.
-As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
+\textcolor{red}{However, he replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\mathbf{Z}$ that he calls the ``beta distributions for bipartite matching."}
+%
+%As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  
+This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
 Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}

### 2023-07-20 ba_resubmit/kundinger_final.tex
done with edits through page 5.

@@
 
-Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$,
+\textcolor{red}{For \text{fabl}, we also use (\ref{eqn:fs_model}) and (\ref{eqn:m_dist}), along with Dirichlet prior distributions for the $\mathbf{m}$ and $\mathbf{u}$ parameters.
+However, instead of the prior over $\bm{Z}$}
+%Instead of the prior over $\bm{Z}$ 
+from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, 
 \textcolor{teal}{we propose an alternative, called the fast beta linkage (fabl) prior.}
@@
 
-%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
-As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
-To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}
-
-\textcolor{red}{Jerry: could you read through this carefully please regarding the missing data part? Note: we only give the marginalized likelihood and not the likelihood without the missing data.}
-
-\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
+\textcolor{red}{In many record linkage tasks, the records in $A$ or $B$ have missing fields, so that some $\gamma_{ij}^f$ could be missing.
+To handle this, we follow the approach described in \cite{sadinle_bayesian_2017}.
+Let $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+We assume that any missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
+For likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}, so that we can marginalize over the missing data and perform all computation using the observed data \citep{sadinle_detecting_2014, sadinle_bayesian_2017}.}
+
+%\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
 To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
@@
 For further details on these connections, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
-}
-
-%Due the the conditional independence assumption, MAR or MCAR lead to the same treatment of the missing data.
-That is, for likelihood based inference both MAR and MCAR lead to ignorability, or rather, both lead to being able to consider just the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
-
-%we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
-
-%Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
-With either the ignorability or MAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
-Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p.
-12]{LittleRubin2002}).
-To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation.}
-
-%are ignorable (or random) \citep{LittleRubin2002}.
-The likelihood becomes that in (\ref{eqn:likelihood}), following directly from \cite{sadinle_bayesian_2017}.
-Note that both assumptions used here are for computational convenience.
-Missing completely at random (MCAR) applies when the probability that a value of a variable is missing does not depend on the observed or unobserved variables.
-Missing at random (MAR) is present when the probability of missingness depends only on (the values of other) observed variables.}
-%\textcolor{red}{In Sadinle (2017), he assumes MAR and in the JASA paper he assumes ignorability.}
-
-\textcolor{teal}{We summarize our proposed approach as follows:}
+%}
+
+\textcolor{teal}{We summarize \textcolor{red}{the model for \text{fabl}} as follows:}
 

### 2023-07-20 ba_resubmit/kundinger_final.tex
suggested edits to sec 3.

@@
 \usepackage{subcaption}
+\usepackage{soul}
 
@@
 For each $Z_j$, we have
+
 \begin{subequations}
@@
 The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$
+
+\begin{rem}
+\textcolor{red}{
+In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
+This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
+In our preliminary studies, this highly informative prior weighting on matching can result in overly optimistic match rates.
+Hence, \textcolor{red}{we prefer (\ref{eqn:fast_beta_prior}).} We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
+}
+\end{rem}
+
+\begin{rem}
+Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
+That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in \textcolor{red}{the \cite{fellegi_theory_1969} model} estimates the proportion of record pairs that are matches.
+In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
+In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by \textcolor{red}{$1/n_A$} and tends towards 0 as the size of the linkage task grows.
+Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
+As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
+\end{rem}
 
@@
 	\end{cases} \label{eqn:z}\\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi},
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi-prior},
 \end{align}
 \end{subequations}
-\textcolor{teal}{where recall that $\alpha_{\pi}, \beta_{\pi}$ are fixed hyper-parameters.}
-
-\begin{rem}
-In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
-This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
-In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates.
-Hence, we prefer (\ref{eqn:fast_beta_prior}).
-We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
-\end{rem}
-
-\begin{rem}
-Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
-In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
-In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
-As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
-\end{rem}
+\textcolor{teal}{where recall that \textcolor{red}{$(\alpha_{\pi}, \beta_{\pi})$} are fixed hyper-parameters.}
 
@@
 
-\textcolor{teal}{We present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m},$ and  $\bm{u}$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi}).
-While we could marginalize over $\pi$ and derive a collapsed Gibbs sampler, in the spirit of \cite{sadinle_bayesian_2017}, this leads to sequential updates, which are computationally slow.
+\textcolor{teal}{We \textcolor{red}{now} present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m},$ $\bm{u},$ and $\pi$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi-prior}).
+%While we could marginalize over $\pi$ and derive a collapsed Gibbs sampler, in the spirit of \cite{sadinle_bayesian_2017}, this leads to sequential updates, which are computationally slow.
 Instead, we proceed with a traditional Gibbs sampler, where the updates to the matching label (or coreference matrix) are parallel, allowing for more efficient updates.
 Refer to Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
-
+%%
 We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-\textcolor{red}{Need to initialize all random variables, such as $\pi$.} \textcolor{teal}{Denote the current draw of the sampler by $s$ and the updated draw by $s+1$.} \textcolor{teal}{For all $f=1, \ldots, F$ sample}
+\textcolor{blue}{Brian: state how you initialize $\pi$.} \textcolor{teal}{Denote the current draw of the sampler by $s$ and the updated draw by $s+1$.} \textcolor{red}{To update $\mathbf{m}$ and $\mathbf{u}$, for all, $f=1, \ldots, F,$ sample}
 
@@
 
-Next, the full conditional for $\pi$ is given by
+\textcolor{red}{Next, we sample $\pi$ from its full conditional, given by}
+%Next, the full conditional for $\pi$ is given by
 \begin{align}
@@
 \textcolor{teal}{
-Observe that the full conditional in (\ref{eqn:z_full_conditional}) does not depend on $\bm{Z}^{(s)},$ implying that 
+The full conditional in (\ref{eqn:z_full_conditional}) does not depend on $\bm{Z}^{(s)},$ implying that 
 \begin{align}
@@
 
-\textcolor{teal}{This implies that we can make parallel updates to $\bm{Z}$ instead of the traditional sequential updates as in \cite{sadinle_bayesian_2017}.
-In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), contrasting the Bayesian record linkage literature \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}.
+\textcolor{teal}{Thus, we can make parallel updates to $\bm{Z}$ instead of sequential updates as in \cite{sadinle_bayesian_2017}.
+In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), \textcolor{red}{in contrast to} \textcolor{red}{other} Bayesian record linkage \textcolor{red}{methods} \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}.
 }
@@
 \textcolor{teal}{Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-However, since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-To obtain a Bayes estimate that fulfills the one-to-one matching (or bipartite) requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-See Appendix \ref{bayes-estimate} for details regarding the initial Bayes estimate and the post-processing procedure.}
+Since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+To obtain a Bayes estimate that fulfills the \textcolor{red}{bipartite} requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+See Appendix \ref{bayes-estimate} for details regarding the initial Bayes estimate and \textcolor{red}{this} post-processing procedure.}
 

### 2023-07-20 ba_resubmit/kundinger_final.tex
done with comments in 4.2

@@
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields.
-However, the impact of indexing on model parameters is not well understood; \citep{murray2016probabilistic} reviewed this issue in the context of the frequentist Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
+However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the \textcolor{red}{classical} Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
 
@@
 Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. 
-%It is
-\textcolor{teal}{The comparison vector} is
- bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1).$
+\textcolor{red}{The quantity $P$ is
+ bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1),$ which clearly depends only on $F$ and $L_f.$ It does not depend (or scale with) with $n_A$ or $n_B.$}
 % This comment below seems obvious and not needed.  
 % where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.
-\textcolor{teal}{The quantity $P^{*}$ clearly depends only on $F$ and $L_f.$ Observe it does not depend (or scale) with $n_A$ or $n_B.$}
+
 %is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
@@
 %\textcolor{red}{Do you mean what I have below or something else? This still is a bit hard to parse.}
-We then map \textcolor{teal}{these integers in (\ref{eqn:hashing})} to sequential integers $\{1, \ldots, P\}$.
+We then map \textcolor{teal}{the integers in (\ref{eqn:hashing})} to sequential integers $\{1, \ldots, P\}$.
 \textcolor{teal}{Denote} each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
@@
 \label{ex:one-hot}
-For example, with five fields with binary agreements and possible missingness (NA), the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
-Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, \text{NA}, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
+%% Jerry propose with (want to avoid using with twice in the same sentence :)
+\textcolor{teal}{Consider} five fields with binary agreements and possible missingness (NA).
+The number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
+\textcolor{red}{Suppose that} records $A_5$ and $B_7$ exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, \text{NA}, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
 Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
-We map the unique values of $h^{*}(\gamma)$ to sequential integers.
-\textcolor{teal}{Suppose that 1066 maps to 42.
-This implies $\gamma_{5,7} = h_{42}.$}
+\textcolor{red}{All records with a hashed value of 1066 map to the integer 42. Then, $\gamma_{5,7} = h_{42}.$}
+%We map the unique values of $h^{*}(\gamma)$ to sequential integers. \textcolor{teal}{Suppose that 1066 maps to 42. This implies $\gamma_{5,7} = h_{42}.$}
 %If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ 
@@
 We denote this set $r_{p_j} = \{i \in 1, \dots, n_A | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
-We compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$, given by
+We compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$,
 \begin{align}\label{eqn:N}
@@
 
-Together, the set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for purposes of writing the likelihood function for \texttt{fabl}.
+The set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for purposes of writing the likelihood function for \texttt{fabl}.
 To see this, we employ the condensed notation
@@
 	\begin{align}
-		m_p =  p(\gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:m_p} \\
-		u_p =  p(\gamma_{ij} = h_p|Z_j \neq i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:u_p}
+		m_p =  \textcolor{red}{p(\Gamma_{ij} = h_p|Z_j = i)} = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:m_p} \\
+		u_p =  \textcolor{red}{p(\Gamma_{ij} = h_p|Z_j \neq i)} = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:u_p}
 	\end{align}
@@
 Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
-Then, conditional on $Z$, we can express the contribution to the likelihood in the full conditional distribution for $\bm{m}$ and $\bm{u}$ as
+Conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional distribution for $\bm{m}$ and $\bm{u}$ as
 \begin{align}
@@
 
-\textcolor{teal}{Observe that we compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
-
-\textcolor{teal}{Turning to sampling for $\pi,$ it remains the same as in (\ref{eqn:pi}). }
+\textcolor{teal}{\textcolor{red}{We can} compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
+
+\textcolor{teal}{Turning to sampling for $\pi,$ it remains the same as in (\ref{eqn:pi}), \textcolor{red}{so we do not repeat it.} }
 
 We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
-For each pattern $p$, let $w_p =\frac{m_p}{u_p}$.
+For each pattern $p$, let $w_p = m_p/u_p$.
 The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
@@
 This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-\textcolor{teal}{We consider two sampling steps.}
+\textcolor{red}{We use two steps in sampling each $Z_j$.}
 %To speed up computation, we break this sampling step into two.
@@
 We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
+
 \begin{align}
 	\label{eqn:gibbs1}
-	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+	p\left( \Gamma_{Z_j^{(s+1)},j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+%	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
 	\begin{cases} 
-		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
+%		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
+                 \frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
 		1- \pi^{(s+1)} , &   \text{otherwise}. \\
@@
 \end{align}
+
 Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
@@
 	\label{eqn:gibbs2}
-	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
+%	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) 
+P\left(Z_j^{(s+1)} = q \mid \Gamma_j=h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) 
+	= 
+	\begin{cases} 
 		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\

### 2023-07-20 ba_resubmit/kundinger_final.tex
update sec 4.3

@@
 
-		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations.
+		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint \textcolor{green}{chunks} for more manageable computations.
 Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.
-
 Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
 For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
-For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
-
-		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison matrix} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
+For \textcolor{red}{all $a$ and $b$}, we conduct all-to-all comparisons \textcolor{red}{of the pairs of records in each $(A^a, B^b)$} to construct the comparison matrix $\gamma^{ab}$.
+
+		We conduct hashing, obtain the compressed \textcolor{teal}{comparison matrix} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
 		the 
@@
 
-Let $n_{A^a}$ and $n_{B^b}$ be the number of records in chunks $A^a$ and $B^b$ respectively.
+Let $n_{A^a}$ and $n_{B^b}$ be the number of records in \textcolor{green}{chunks} $A^a$ and $B^b$ respectively.
 Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels.

### 2023-07-20 ba_resubmit/kundinger_final.tex
done with most edits from jerry.
update.
cleaning up comments.

@@
 Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
-
+%
 It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
@@
 
-\begin{enumerate}	
-\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.
+%\begin{enumerate}	
+%\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.
 Also, do you more formally mean something like as follows: Consider $\gamma_{q,q} = h_p \mid Z_j = q$.
@@
 I just wanted to bring it up as a conversation before submission.}
-\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
-\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1.
+%\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
+%\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1.
 Example, it's missing on the update for $m$ and $u$ so it's not immediately clear how one would draw this directly.
 Wouldn't it just be a vectorized Dirichlet, where the computational savings come from vectorizing the hyperparameters of the Dirichlet?
-\end{enumerate}
+%\end{enumerate}
 
@@
 These records have the same probability to be identified as the link for record $B_j$.
-Thus, we know that records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate.
+Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate.
 Rather than store all of these record labels, we store only a small number $S$.

### 2023-07-20 ba_resubmit/rebuttal/response.tex
update R.0.

--- 2023-07-20 ba_resubmit/rebuttal/response.tex
@@
 
-	\textcolor{teal}{``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
-Due the the conditional independence assumption, we can marginalize over the missing data in (6a) and perform all computation using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
+	\textcolor{teal}{In many record linkage tasks, the records in $A$ or $B$ have missing fields, so that some $\gamma_{ij}^f$ could be missing.
+To handle this, we follow the approach described in \cite{sadinle_bayesian_2017}.
+Let $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+We assume that any missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
+For likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}, so that we can marginalize over the missing data and perform all computation using the observed data \citep{sadinle_detecting_2014, sadinle_bayesian_2017}.}
 
+%	\textcolor{teal}{``\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
+To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
+It is important to note that for likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
+Therefore, we can marginalize over the missing data and perform all computation using the observed data, leading to (6a).
+For further details on these connections, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}}."}
 	}

### 2023-07-20 ba_resubmit/kundinger_final.tex
update.
update R.0.

@@
 	\subsection{Scaling to Large Linkage Tasks}\label{scaling}}
-
-	\textcolor{red}{Suggestion: Chunk is award. Perhaps partition/block/batch/segment. I might like segment since partition/block have meanings in record linkage already. }
 
@@
 		\begin{center}
-		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_speed_plot2} 
-	% \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
+%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_speed_plot2}  
+	 \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
@@
 		\begin{center} 
-		\includegraphics[width=0.7\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides} 
-	%	\includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
+%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides} 
+		\includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
@@
 		\begin{center}
-		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_sim_plot2} 
-			%\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
+%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_sim_plot2} 
+			\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
 			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
@@
 		\begin{center}
-		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/Z_plot}
-%		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
+%		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/Z_plot}
+		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
 		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
@@
 		\begin{center}
-		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_distribution_smallP_bayes}
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_distribution_smallP_bayes}
+	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
 			\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
@@
 		\begin{center}
-		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_posterior_smallP} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
+%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_posterior_smallP} 
+                  	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
 			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
@@
 		\begin{center}
-		        \includegraphics[width=0.6\textwidth]{finalFigures/nltcs/overlap_posterior4}
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
+%		        \includegraphics[width=0.6\textwidth]{finalFigures/nltcs/overlap_posterior4}
+ 		\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
 			\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
@@
 	\begin{center}
-	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
-		%\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
+%	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
 		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
@@
 		\begin{center}
-		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
+%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
 			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
@@
 		\begin{center}
-		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
+%		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
 			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
@@
 		\begin{center}
-		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
+%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
 			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
@@
 		\begin{center} 
-		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
-		%\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
+%		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
+		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
@@
 		\begin{center}
-		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
-		% \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
+%		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
+		 \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
 			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
@@
 		\begin{center}
-		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
+%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
+		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
 			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
@@
 		\begin{center}
-		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
+%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
+			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
 			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} 
@@
 		\begin{center}
-		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
+%		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
+		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
 			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} 

### 2023-07-29 ba_resubmit/response_to_reviewers3.tex
lots and lots of edits

--- 2023-07-20 ba_resubmit/rebuttal/response.tex
@@
 
-\reply{
-\textcolor{teal}{Thank you for the opportunity to submit a revision. 
-We have provided detailed responses to the reviewers' comments below, 
-which prompted significant changes in our article. Specifically, we have attempted to improve notation and wording choices, at the reviewers' suggestion. 
+\reply{We have made several revisions to make the paper easier to read, as described in our responses below. 
 }
-}
-
-%We propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.
-This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the decision theoretic  technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.
-This discovery allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs via a novel method, storage efficient indexing.
-These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
-In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
-Finally, we provide two real illustrations of our methodology and simulation studies with reproducible software.
 
@@
 \reply{
-Thank you for the suggestion. We now refer to the two data files as
- $A$ and $B$, where the updated material reads:
+To lighten the notation and allow for greater precision, we now refer to the two data files as $A$ and $B$. The updated material now reads:
 
-	``Consider a data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$". See Section 2, page 3. 
+``Consider two data files $A$ and $B$ comprising $n_A$ and $n_B$ records, respectively, and including $F$ linkage variables measured in both files.
+For $i=1, \dots, n_A$, let record $i$ be given by $A_i = (A_{i1}, \dots, A_{iF})$, so that $A = (A_i : i = 1, \dots, n_A)$.
+Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.
+Without loss of generality, denote files such that $n_A \geq n_B$.".
+See Section 2.
 }
@@
 
-	\reply{\textcolor{teal}{The distinction does not matter due to assuming that the comparison vectors are conditionally independent given the coreference matrix (or matching label).
-We now state the following in Section 3, page 5:}
+	\reply{Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
+For simplicity, we have chosen to use MCAR in the article.
+We have included the following explanation of missingness in Section 2.
 
-	\textcolor{teal}{``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
-Due the the conditional independence assumption, we can marginalize over the missing data in (6a) and perform all computation using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
-
+	``In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
+As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
+We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data."
 	}
@@
 However, I do not follow the re-expression of the pmf for $\Gamma_{.j}$.
-First, the opening square brackets are still misplaced
-	throughout.
+First, the opening square brackets are still misplaced throughout.
 In the second line of this derivation, the authors divide the expression by a product of the element-wise conditional probabilities of a match in the comparison vector (u), raised by an indicator that the element of the comparison vector equals a particular value.
@@
 
-%% Don't apologize. Just answer the questions very completely.
-
-%\textcolor{teal}{Don't apologize!}
-
 \reply{%
 	The square brackets have been corrected.
+
 	We have revised the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-Refer to Appendix A, pages 26 -- 27 regarding this extended derivation.
+Refer to Appendix A for this extended derivation.
 	}
-
-%\reply{%
-%	The square brackets have been corrected, and we apologize for the oversight.
-We have clarified derivation of the likelihood for $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-%}
 
@@
 
-\reply{No, in \text{fabl}, we do not integrate out $\pi$ as done in \cite{sadinle_bayesian_2017} because this leads to a sequential sampler.
+\reply{%
+		We realize that the integration of $\pi$ out of the full conditionals is not necessary for our sampler. Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler.
 
-(1) Yes, our alternative approach is different than directly integrating over the prior distribution $\pi.$ 
-%We have attempted to make this more clear in our revised version. 
-In short, our prior distribution over $Z$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of \cite{sadinle_bayesian_2017}.
+		We have redone all simulations and case studies using a Gibbs sampler that samples $\pi$ from its full conditional, and then samples $\bm{Z}$ from its full conditionals.
+We obtained equivalent results.
+	}
+
+%\reply{No, in \text{fabl}, we do not integrate out $\pi$ as done in \cite{sadinle_bayesian_2017} because this leads to a sequential sampler. 
+%
+%(1) Yes, our alternative approach is different than directly integrating over the prior distribution $\pi.$ 
+%%We have attempted to make this more clear in our revised version. 
+%In short, our prior distribution over $\bm{Z}$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of \cite{sadinle_bayesian_2017}.
 Of course, we could integrate out $\pi$ in our approach, however, this leads to sequential updates, which would be slow in practice.
 See Appendix A, page 27.
-
-(2) Yes, our alternative approach is justified.
+%
+%(2) Yes, our alternative approach is justified.
 We provide a justification in Appendix A, page 27 regarding why the updates are independent and not sequential.
 Thank you for the excellent questions, which have greatly improved our paper.
-
-}
-
-%
-%\reply{%
-%	We realize that the integration of pi out of the full conditionals is not necessary for our sampler.
-For clarity, we decided to redo the simulations using a Gibbs sampler that samples $\pi$ from its full conditional, and then samples $Z$ from its full conditionals.
-We obtained equivalent results, which is unsurprising, because with high numbers of observations, the posterior of the beta distribution is highly concentrated around its mean.
-%	
-%	Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler. We leave the remark about marginalization in the Appendix to facilitate comparisons to Sadinle's method.
 %
@@
 
-	\reply{\textcolor{teal}{Yes, in the original manuscript, the comparison vectors are created and therefore, patterns are assigned) without regard to $Z$.
-Given the confusion, we have removed the sentence.	}}
-
-%	\reply{%
-%	You are correct that the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$.
-This notation was created purely to be able to denote when a record was left unlinked during the Gibbs Sampler as in Equation 16 in the submitted draft.
-%	
-%	I can see why this is confusing. Therefore, we have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
-%	}
+	\reply{Yes, the comparison vectors are created (and therefore, patterns are assigned) without regard to $\bm{Z}$.
+We have removed that sentence, and updated the equations in the Section 4.3 to avoid confusion}
 
@@
 
-	%%Comment: Make it more brief and no need to provide notation when it's not needed.
+	\reply{We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
+However, we cannot express the full likelihood in (14) through these statistics.
+Therefore, we have revised to text to say ``summary statistics" instead of  ``sufficient statistics."
 
-	\reply{
-	Thank for pointing out our typo. We have revised the text to state ``summary statistics" instead of ``sufficient statistics."
+	%Thank for pointing out our typo. We have revised the text to state ``summary statistics" instead of ``sufficient statistics."
 	}
-
-%	\reply{%
-%		We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
-However, we cannot express the full likelihood in (14) through these statistics.
-Therefore, we have revised to text to say ``summary statistics" instead of  ``sufficient statistics."}
 
@@
 To my understanding, the weights have been removed and section 4 has been rewritten }}
-	\reply{
-	\textcolor{teal}{In the original manuscript, we consider all records in the first file ($X_{1}$) that share an agreement pattern with the second data file ($X_{2}$).
+%	\reply{
+%	\textcolor{teal}{In the original manuscript, we consider all records in the first file ($X_{1}$) that share an agreement pattern with the second data file ($X_{2}$).
 In this situation, all these records have the same Fellegi-Sunter likelihood ratio weight, $w_{ij}$.
 Given this result, we denoted the weight $w_p$ since these records map to the same agreement pattern.}
-
-	\textcolor{teal}{In the revised manuscript, we define $m_p$ and $u_p,$ which are the 
-	 probabilities that records $A_i$ and $B_j$ form agreement pattern $p$ given that
- they are a match and non-match, respectively.
+%	
+%	\textcolor{teal}{In the revised manuscript, we define $m_p$ and $u_p,$ which are the 
+%	 probabilities that records $A_i$ and $B_j$ form agreement pattern $p$ given that
+% they are a match and non-match, respectively.
 For each pattern $p,$ we define $w_p = m_p/u_p.$ We hope that this will be more clear than our previous presentation of the material.
 See Section 4.1, pages 8--9.}
+% }
+% 
+	\reply{Our intended meaning was ``All record pairs contribute to the likelihood through the summary statistics."
 
- }
-
-%	\textcolor{teal}{
-%	We have revised the ordering of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI.
-With this new ordering, and with the newly provided equations 15 -19, we hope that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$ is now more clear.
-%	}
+	In our revised submission, we have reexpressed the likelihood from (6a) through the perspective of agreement patterns in (14).
+This form of the likelihood is then used to express contributions to the likelihood for the $\bm{m}$ and $\bm{u}$ parameters in (15) and for $\bm{Z}$ in (17), both of which use the summary statistics.
+We hope that this revision makes this point more clear.}
 
@@
 
-	\textcolor{teal}{In the original manuscript, ``those comparison vectors" refer to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$ }
+	In the original manuscript, ``those comparison vectors" refer to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$
 
-	\textcolor{teal}{The revised text states: ``Then, we conduct hashing, obtain the compressed comparison matrix, $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison matrix, $\gamma^{ab}$, before continuing with the next chunk of data.
-See Section 4.3, page 10.
-}
+	The revised text states: ``Then, we conduct hashing, obtain the compressed comparison matrix, $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison matrix, $\gamma^{ab}$, before continuing with the next chunk of data.
+See Section 4.3}
 
@@
 I would argue that a matrix should use capital notation, such as $\Gamma.$}
-	}
 
@@
 
-\reply{
-	\textcolor{teal}{In order to attempt to make the SEI algorithm more clear, we have revised the ordering of Sections 4.2 and 4.3, so that 1) hashing, 2) posterior inference, 3) chunkwise computation of the comparison matrix, and 4) SEI.
-With this new ordering, and with the newly provided equations 15 -19, we hope that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$ is now more clear.
-In addition, we hope that the SEI algorithm updates will now be more clear that the updates in (16a), (16b), and (18) only depend on $\mathcal{N}$.
-The SEI algorithm affects the step shown in (19).
-	}
+\reply{To make the SEI algorithm more clear, we have revised the ordering of Sections 4, so that we have 1) hashing, 2) posterior inference, 3) chunkwise computation of the comparison matrix, and 4) SEI.
+With this new ordering, and with the newly provided equations (15) through (19), we hope it clearer that that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
+In particular, (16a), (16b), and (18) depend only $\mathcal{N}$.
+The SEI algorithm only affects the step shown in (19).
 	}
@@
 \reply{% 
-	Refer to Section 5.3, which explores the trade-offs regarding different choices of $S$ and discusses this in practical terms. 
+	We have included an additional simulation in Section 5.3, which explores the trade-offs regarding different choices of $S$. 
 }
@@
 \reply{%
-	This has been corrected; thank you.
+	This has been corrected.
 }
@@
 \reply{%
-	\textcolor{teal}{The typo has been fixed.}
+	The typo has been fixed.
 }
@@
 \reply{%
-	We use $Z$ when discussing a random quantity, and $z$ to reference a realized value. We have revised all instances of the inconsistency. 
+	We use $\bm{Z}$ when discussing a random quantity.
+In all instances where we discuss $Z_j$ taking on a particular value, we now use $q$ to avoid confusion and to follow notation used by \cite{sadinle_bayesian_2017}.
 }
@@
 \reply{%
-	This has been fixed. \textcolor{teal}{Refer to equations (12) and (13), page 7.}
+	This has been fixed. See Equations (9) and (10).
 }
@@
 \reply{%
-	\textcolor{teal}{Note that $\gamma$ is a comparison matrix comprised by comparison vectors.
+	$\gamma$ is a comparison matrix comprised by comparison vectors.
 All references to $\gamma$ as a set have been removed.
-See Section 2, page 3, where we define both the comparison vector and comparison matrix.}
+See Section 2, where we define both the comparison vector and comparison matrix.
 }
@@
 \reply{%
-	This has been revised; thank you. \textcolor{teal}{Refer to Section 4.2, page 10.}
+	This has been revised; thank you. See Section 4.2
 }

### 2023-07-29 ba_resubmit/kundinger_resubmit2.tex
lots and lots of edits

--- 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
@@
 \newtheorem{lemma}{Lemma}
-\newtheorem{example}{Example}
-\newtheorem{rem}{Remark}
 \usepackage{booktabs}       % professional-quality tables
@@
 To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
-Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
-\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
-\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
-Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
-}
-%This allows
-%us to use hashing techniques that reduce computational overhead.
+Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
 This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
@@
 \end{abstract}
+
+%\textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
+%\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
+%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
+Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
+%}
 
@@
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
-To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}
+To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989}.
+%\textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}
 
@@
 Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define a \emph{comparison vector} $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
-For notational convenience, we define the \emph{comparison matrix} $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
+For notational convenience, we define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
 
@@
 He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.
-As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
+As the prior distriburion for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
 Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
@@
 
-\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
+\textcolor{teal}{We now explain how we accomodate missing fields, which leads to missing comparison vectors.
 To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR).
-Due the the conditional independence assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}
-
-%Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}).
-With either the ignorability or MAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.
-For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
-Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p.
-12]{LittleRubin2002}).
-To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation.}
+Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing completely at random (MCAR, per \cite{LittleRubin2002}).
+With the MCAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.}
 
@@
 
-\begin{rem}
+\begin{remark}
 In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
@@
 We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
-\end{rem}
-
-\begin{rem}
+\end{remark}
+
+\begin{remark}
 Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
@@
 As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
-\end{rem}
+\end{remark}
 
@@
 \begin{align}
-\label{eqn:pi}
 	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
@@
 The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains.
-Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking \textcolor{teal}{often}
- undesirable %in some situations 
-\citep{steorts_comparison_2014}.
+Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
 
@@
 
-With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage.
+With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
 First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors.
@@
 
-Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. 
-%It is
-\textcolor{teal}{The comparison vector} is
- bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1).$
-% This comment below seems obvious and not needed.  
-% where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values. 
-\textcolor{teal}{The quantity $P^{*}$ clearly depends only on $F$ and $L_f.$ Observe it does not depend (or scale) with $n_A$ or $n_B.$}
-%is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
-
+Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
+Let $P$ be the number of patterns realized in $\gamma$.
+It is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.
+This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
 To obtain a memory efficient representation, we map the agreement pattern of each record pair to a unique integer.
@@
 
-%\textcolor{red}{Do you mean what I have below or something else? This still is a bit hard to parse.}
-We then map \textcolor{teal}{these integers in (\ref{eqn:hashing})} to sequential integers $\{1, \ldots, P\}$.
-\textcolor{teal}{Denote} each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-When record pair, \textcolor{teal}{$(i,j)$,} exhibits agreement pattern $p$, \textcolor{teal}{let} $\gamma_{ij} = h_p$.
-
+We then map these integers to sequential integers $\{1, \ldots, P\}$.
+We refer to each unique agreement pattern as $h_p$, and to the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
 In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
@@
 
-\begin{example}
-\label{ex:one-hot}
-For example, with five fields with binary agreements and possible missingness (NA), the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
-Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, \text{NA}, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
+For example, with five fields with binary agreements and possible missingness, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
+Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
 Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
 We map the unique values of $h^{*}(\gamma)$ to sequential integers.
-\textcolor{teal}{Suppose that 1066 maps to 42.
-This implies $\gamma_{5,7} = h_{42}.$}
-%If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ 
-This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
-\end{example}
+If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
 
@@
 to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match, in (\ref{eqn:m_p}), and not a match, in (\ref{eqn:u_p}).
-Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) \textcolor{teal}{can be expressed as}
-%is equivalent to
+Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) is equivalent to
 
@@
 The likelihood in (\ref{likelihood_efficient}) allows for more
-efficient \textcolor{teal}{approximate posterior inference of the Gibbs sampler in Section~\ref{sec:fast-beta-linkage},}  as we \textcolor{teal}{describe in the next section}.
-
-%% These are not the full conditional distributions.
-%for $\bm{m}$, $\bm{u}$, and $\bm{Z}$
+efficient posterior inference for $\bm{m}$, $\bm{u}$, and $\bm{Z}$, as we now describe.
 
@@
 
-\textcolor{teal}{In our derivations below, we assume the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the updates to each full conditional distribution based upon our computational proposals using hashing.}
-
-\begin{enumerate}	
-\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.
-Also, do you more formally mean something like as follows: Consider $\gamma_{q,q} = h_p \mid Z_j = q$.
-It seems that the conditioning may actually be important here.
-I mention this for two reasons, it appears a bit strange in equation (22).
-When I think through it, it's okay.
-I just wanted to bring it up as a conversation before submission.}
-\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
-\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1.
-Example, it's missing on the update for $m$ and $u$ so it's not immediately clear how one would draw this directly.
-Wouldn't it just be a vectorized Dirichlet, where the computational savings come from vectorizing the hyperparameters of the Dirichlet?
-\end{enumerate}
-
-Turning first to the \textcolor{teal}{full conditional updates}
-%posterior updates 
-for $\bm{m}$ and $\bm{u}$, we observe that these parameters only depend on the \textcolor{teal}{observed} data  through quantities that are calculated through $\mathcal{N}$ and $\mathcal{P}$.
+Turning first to the posterior updates for $\bm{m}$ and $\bm{u}$, we first note that these parameters depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$.
 Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
-Then, conditional on $Z$, we can express the contribution to the likelihood in the full conditional distribution for $\bm{m}$ and $\bm{u}$ as
+Then, conditional on $Z$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
 \begin{align}
@@
 Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be a concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
-The terms needed for the \textcolor{teal}{full conditional} updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
+The terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
 \begin{subequations}
@@
 \end{subequations}
-%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-\textcolor{teal}{Observe that we compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
-
-\textcolor{teal}{Turning to sampling for $\pi,$ it remains the same as in (\ref{eqn:pi}). }
-
-We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
+Specifically, the $l + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}),and (\ref{eqn:beta_update}).
+However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{eqn:beta_update}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+We also use (\ref{likelihood_efficient}) to sample Z efficiently.
 For each pattern $p$, let $w_p =\frac{m_p}{u_p}$.
@@
 This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-\textcolor{teal}{We consider two sampling steps.}
-%To speed up computation, we break this sampling step into two.
+To speed up computation, we break this sampling step into two.
 
@@
 
-	\textcolor{red}{Suggestion: Chunk is award. Perhaps partition/block/batch/segment. I might like segment since partition/block have meanings in record linkage already. }
-
 		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations.
 Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.
-
-		Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
+Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B and B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
 For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
@@
 
-		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison matrix} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
+		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison vectors} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
 		the 
 \textcolor{teal}{memory-intensive}
-		\textcolor{teal}{comparison matrix} $\gamma^{ab}$ 
+		\textcolor{teal}{comparison vectors} $\gamma^{ab}$ 
 		before continuing with the next chunk of data. 
@@
 For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
-We introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
+We next introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
 This allows all-to-all comparisons for substantially larger linkage tasks.
@@
 Rather than store all of these record labels, we store only a small number $S$.
-For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $\text{SEI}(r_{p_j}^{ab})$.
-We collect these memory reduced lists to form $\text{SEI}(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $\text{SEI}(\mathcal{R})$.
+For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $SEI(r_{p_j}^{ab})$.
+We collect these memory reduced lists to form $SEI(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $SEI(\mathcal{R})$.
 
@@
 Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels.
-As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $\text{SEI}(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
-We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
+As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $SEI(\tilde{\gamma}) = \{\mathcal{P}, SEI(\mathcal{R}), \mathcal{N}\}$.
+We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.
 
@@
 
-	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017} and others \citep{lum2013applications, jewell2018accounting}.
+	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017}.
 Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
@@
 We are interested in estimating the total number of casualties from the war.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL)\citep{howland2008rescate} and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish)\citep{ball2000salvadoran}.\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization.
+We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
 The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
@@
 Second, the only fields recorded are given name, last name, date of death, and place of death.
-\textcolor{teal}{While it may be common in other data sets for a parent and child to share the same given name, in this data set, it is more common for cousins to share this information, especially in villages, where there may be only four to five extended families.
-In Latin America, an individual receives a surname from their father and their mother, leading to a limited number of surnames in small villages.}
-	%% I believe that Patrick said that this is not common in this data set
-%	It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
+It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
 
@@
 \textcolor{red}{Note: Please use s, s+1 consistently throughout.
-Or surpress the notation and tell the reader.
+Or supress the notation and tell the reader.
 It looks weird to go back and forth.

### 2023-07-29 ba_resubmit/kundinger_resubmit3.tex
lots and lots of edits

--- 2023-07-29 ba_resubmit/kundinger_resubmit2.tex
@@
 
+\usepackage{afterpage}
 \usepackage{lineno}
@@
 \begin{abstract}
+%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
 Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.
 However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
+To address these computational difficulties, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
 Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
@@
 Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
+%
+%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.
+Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}
+
 \end{abstract}
 
-%\textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
-%\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
-%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
-Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
-%}
-
 %% ** Keywords **
-\begin{keyword}
-\kwd{bipartite record linkage}
+\begin{keyword}%[class=MSC]
+\kwd{data fusion}
+\kwd{distributed computing}
 \kwd{entity resolution}
-\kwd{data cleaning}
-\kwd{parallelization}
 \kwd{hashing}
+\kwd{record linkage}
 \end{keyword}
@@
 
-Many probabilistic record linkage methods are rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
+Many probabilistic record linkage methods rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
 In their approach, the data analyst first creates comparison vectors for each pair of records in the data files.
@@
 
-The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity.
+% Possible citation for bipartite matching. 
+%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_A995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
+
+The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular for its mathematical simplicity.
 However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
-To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989}.
-%\textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}
+Analysts typically use an additional post-processing step to turn the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
 
 Alternatively, analysts can embed one-to-one matching requirements into the model specification, at an additional computational cost.
-\cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
+\cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respect one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
 \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records.
 \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
-Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
 \texttt{BRL} has been shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
 
+%Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
+
 In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
-This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We use independent prior distributions for the matching statuses of each record, and modify the decision theoretic technique of \cite{sadinle_bayesian_2017} to ensure our linkage estimates are bipartite.
+This approach allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
@@
 
+%
+
+%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}}
+
 In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
-Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching.
+Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and describes the loss function used to calculate the Bayes estimate for the bipartite matching.
 Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}.
@@
 
+%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
+Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
+
 Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values.
 \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
-For notational convenience, we define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
+We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs $\gamma_{ij}$.
+Additionally, we let $\Gamma_{ij}$ represent a random variable for the comparison vector for $A_i$ and $B_j$, such that $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
 
@@
 
-%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
+In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
 As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
 To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+
+%\subsection{Fellegi-Sunter Models}
+%\label{fellegi-sunter}
 
@@
 
-Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.
-Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
+%Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.  Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
+
 For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
@@
 
+%Within this framework, record pairs are independently classified as matches and nonmatches based on the estimated parameters.
+However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results.
+To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching.
+\cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.
+
+%The Fellegi-Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+This is a problem in practice as transitive closures can be violated.
+Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.
+
+%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
+%\begin{align}
+%     \label{eqn:jaro}
+%     &\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij}
+%     \quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
+%     & \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
+%     & \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
+%\end{align}
+%
+%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs.
+Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
+
+%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
+This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
+%\begin{align}
+%     \label{eqn:wts}
+%     w_{ij} = \frac{p(\gamma_{ij} \mid \Delta_{ij} = 1)}{p(\gamma_{ij} \mid \Delta_{ij} = 0)}
+%\end{align}
+%as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
+%
+%In practice, $p(\cdot \mid \Delta_{ij} = 1)$ and $p(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
+%
+%%is the set of parameters $u_{fl} = p(\gamma_{ij}^f = l|Z_j \neq i)$ for the nonmatching pairs.
+%
+%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
+This is a problem in practice as transitive closures can be violated.
+Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
+
+%\subsection{Beta Record Linkage Model}
+%\label{BRL}
+
+%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
+
+%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a
+
 \cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.
-He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.
-As the prior distriburion for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
-Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
-
+He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$, and replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\bm{Z}$ that he calls the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
+Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, where $\alpha_{\pi}$ and $\beta_{\pi}$ are known hyperparameters.
 It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
@@
 This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
-This makes the sampler inherently \textcolor{teal}{sequential}, which can be slow when working on linkage tasks with more than a few thousand records.
+This makes the sampler inherently sequential, which can be slow when working on linkage tasks with more than a few thousand records.
 
@@
 Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$,
-\textcolor{teal}{we propose an alternative, called the fast beta linkage (fabl) prior.}
-%the fast Beta prior as follows. 
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta linkage (\texttt{fabl}) prior below.
 For each $Z_j$, we have
-\begin{subequations}
+%\begin{subequations}
 \begin{align}
@@
 \end{cases} \label{eqn:fast_beta_prior}  \\
-\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \label{eqn:fast_beta_prior2}.
-\end{align}
-\end{subequations}
+\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \nonumber
+%\label{eqn:fast_beta_prior2}.
+\end{align}
+%\end{subequations}
 We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match.
@@
 
-%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
-As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
-To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}
-
-\textcolor{teal}{We now explain how we accomodate missing fields, which leads to missing comparison vectors.
-To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing completely at random (MCAR, per \cite{LittleRubin2002}).
-With the MCAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.}
-
-%are ignorable (or random) \citep{LittleRubin2002}.
-The likelihood becomes that in (\ref{eqn:likelihood}), following directly from \cite{sadinle_bayesian_2017}.
-Note that both assumptions used here are for computational convenience.
-Missing completely at random (MCAR) applies when the probability that a value of a variable is missing does not depend on the observed or unobserved variables.
-Missing at random (MAR) is present when the probability of missingness depends only on (the values of other) observed variables.}
-%\textcolor{red}{In Sadinle (2017), he assumes MAR and in the JASA paper he assumes ignorability.}
-
-\textcolor{teal}{We summarize our proposed approach as follows:}
-
+In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
+This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
+In our preliminary studies, this highly informative prior weighting on matching can result in overly high match rates.
+Hence, we prefer (\ref{eqn:fast_beta_prior}).
+We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
+
+Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the \cite{fellegi_theory_1969} model.
+That is, $\pi$  in (\ref{eqn:fast_beta_prior}) under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in (\ref{eqn:lambda_dist}) in the \cite{fellegi_theory_1969} model estimates the proportion of record pairs that are matches.
+In the bipartite case, we conjecture that some analysts will find $\pi$ to be more a interpretable parameter than $\lambda$.
+In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $1/n_A$ and tends towards 0 as the size of the linkage task grows.
+Additionally, while the \cite{fellegi_theory_1969} model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
+As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
+
+%\begin{table}[t!]
+%	\centering
+%	\begin{tabular}[t!]{ll}
+%		Symbol & Description \\
+%		\hline
+%		$A, B$ & data files\\
+%		$i \in 1, \ldots, n_A $ & index over records in $A$\\
+%		$j \in 1, \ldots, n_B $ & index over records in $B$\\
+%		$f \in 1, \ldots F$ & index over fields used for comparisons \\
+%		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
+%		$n_{AB}$ & number of matches between $A$ and $B$\\
+%		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
+%		$Z_j = i$ & records $A_i$ and $B_j$ match \\
+%		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
+%		$m_{fl}$ & $p(\gamma_{ij}^f = l | Z_j = i)$ \\
+%		$u_{fl}$ & $p(\gamma_{ij}^f = l | Z_j \neq i)$ \\
+%		$\pi$ & probability that a record in $B$ has a match in $A$ \\
+%		\hline
+%	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
+%\end{table}
+%
+For clarity, we present our full model for \texttt{fabl} below:
 \begin{subequations}
@@
 	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}),\quad  \forall  \; f = 1, \ldots, F, \label{eqn:m} \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}), \quad \forall \; f = 1, \ldots, F,  \label{eqn:u}\\
+	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
+	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
 	p(Z_j = q| \pi)  &=
@@
 	\end{cases} \label{eqn:z}\\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi},
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
 \end{align}
 \end{subequations}
-\textcolor{teal}{where recall that $\alpha_{\pi}, \beta_{\pi}$ are fixed hyper-parameters.}
-
-\begin{remark}
-In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
-This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
-In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates.
-Hence, we prefer (\ref{eqn:fast_beta_prior}).
-We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
-\end{remark}
-
-\begin{remark}
-Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches.
-In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
-In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
-As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
-\end{remark}
-
-%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
-%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
-
-%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
-%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
-
-\textcolor{teal}{We present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m},$ and  $\bm{u}$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi}).
-While we could marginalize over $\pi$ and derive a collapsed Gibbs sampler, in the spirit of \cite{sadinle_bayesian_2017}, this leads to sequential updates, which are computationally slow.
-Instead, we proceed with a traditional Gibbs sampler, where the updates to the matching label (or coreference matrix) are parallel, allowing for more efficient updates.
-Refer to Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
-
-We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-\textcolor{red}{Need to initialize all random variables, such as $\pi$.} \textcolor{teal}{Denote the current draw of the sampler by $s$ and the updated draw by $s+1$.} \textcolor{teal}{For all $f=1, \ldots, F$ sample}
-
-%\begin{subequations}
-%\begin{align}
-%	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  \\
-%	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
-%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-%\end{align}
-%\end{subequations}
-
-%\begin{subequations}
-%
-%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-%\end{align}
-%\end{subequations}
-
-\begin{align}
-	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  
-\end{align}
-and	
-\begin{align}
-	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} 
-\end{align}
-
-where
-
-\begin{align}
-\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \quad \text{and}\\
-\beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-\end{align}
-
-Next, the full conditional for $\pi$ is given by
-\begin{align}
-	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
+
+%\hypertarget{posterior-sampling}{%
+%	\subsection{Gibbs Sampler}
+%	\label{gibbs_sampling}}
+
+We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler.
+Appendix \ref{app:derivations} presents derivations for the full conditional distributions.
+We initialize $\bm{m}$, $\bm{u}$ and $\pi$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
+Denote the current draw of the sampler by $(s)$ and the updated draw by $(s+1)$.
+To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, sample
+\begin{subequations}
+\begin{align}
+	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
+	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
+	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
+	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
+	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
+\end{align}
+\end{subequations}
+Next, we sample $\pi$ from its full conditional, given by
+\begin{align}
+	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
 \end{align}
@@
 \end{align}
-
+%\begin{align}
+%\label{eqn:z_full_conditional}
+%p\left(Z_j^{(s+1)}  = z_j|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
+%\begin{cases} 
+%	w_{z_j, j}^{(s+1)},  & z_j \leq n_A; \\
+%	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
+%\end{cases}
+%\end{align}
 where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
@@
 
-\textcolor{teal}{
-Observe that the full conditional in (\ref{eqn:z_full_conditional}) does not depend on $\bm{Z}^{(s)},$ implying that 
-\begin{align}
-p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}) \propto \prod_q \left(\prod_j
-p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right)
-\right).
-\end{align}
-}
-
-\textcolor{teal}{This implies that we can make parallel updates to $\bm{Z}$ instead of the traditional sequential updates as in \cite{sadinle_bayesian_2017}.
-In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), contrasting the Bayesian record linkage literature \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}.
-}
-%See Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
-
-%Derivations for the full conditional distributions can be found in Appendix~\ref{app:derivations}.
-
-\textcolor{teal}{Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-However, since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-To obtain a Bayes estimate that fulfills the one-to-one matching (or bipartite) requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-See Appendix \ref{bayes-estimate} for details regarding the initial Bayes estimate and the post-processing procedure.}
-
-%Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
+Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
+Since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+To obtain a Bayes estimate that fulfills the bipartite requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+See Appendix \ref{bayes-estimate} for details redarding the initial Bayes estimate and this post-processing procedure.
 
@@
 The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains.
-Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
+Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking often undesirable \citep{steorts_comparison_2014}.
 
@@
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields.
-However, the impact of indexing on model parameters is not well understood; \citep{murray2016probabilistic} reviewed this issue in the context of the frequentist Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
+However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
 
@@
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
+%For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.
+
+%\begin{table}[t!]
+%	\centering
+%	\begin{tabular}[t!]{ll}
+%		Symbol & Description \\
+%		\hline
+%		$P$ & number of agreement patterns exhibited in $\gamma$ \\
+%		$h_p$ & agreement pattern $p$ \\
+%		$e(h_p)$ & one-hot encoding of agreement pattern $p$ \\
+%		$\gamma_{ij} = h_p$ & comparison vector between records $A_i$ and $B_j$ exhibits pattern $p$ \\
+%		$r_{p_j}$ & list of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+%		$N_{p_j}$ & number of records in $A$ that share agreement pattern $p$ with record $B_j$ \\
+%		$N_p^m$ & number of matching comparison vectors that exhibit pattern $p$ \\
+%		$N_p$ & number of total comparison vectors that exhibit pattern $p$ \\
+%
+%		\hline
+%	\end{tabular}\caption{Summary of hashing notation.}\label{table_notation_B}
+%\end{table}
 
@@
 This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
+
 To obtain a memory efficient representation, we map the agreement pattern of each record pair to a unique integer.
@@
 
-We then map these integers to sequential integers $\{1, \ldots, P\}$.
-We refer to each unique agreement pattern as $h_p$, and to the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+We then map the integers in \eqref{eqn:hashing} to sequential integers $\{1, \ldots, P\}$.
+Denote each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
 When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
@@
 This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
-
-For example, with five fields with binary agreements and possible missingness, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
-Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
-Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
-We map the unique values of $h^{*}(\gamma)$ to sequential integers.
-If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
+%A more complete example of the one-hot encoding is provided in Appendix \ref{app:ohe}.
+
+For example, consider five fields with binary agreements and possible missingness, denoted $NA$.
+The number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
+Suppose that records $A_5$ and $B_7$ exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
+The expression in  (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
+All records with a hashed value of $1066$ map to the integer $42$.
+Thus, $\gamma_{5,7} = h_{42},$ and this agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
 
@@
 
-Together, the set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for purposes of writing the likelihood function for \texttt{fabl}.
+The set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for writing the likelihood function for \texttt{fabl}.
 To see this, we employ the condensed notation
@@
 	\begin{align}
-		m_p =  p(\gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:m_p} \\
-		u_p =  p(\gamma_{ij} = h_p|Z_j \neq i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:u_p}
+		m_p =  p(\Gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(h_p^f = l)I_{obs}(\Gamma_{ij}^f)}\label{eqn:m_p} \\
+		u_p =  p(\Gamma_{ij} = h_p|Z_j \neq i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(h_p^f = l)I_{obs}(\Gamma_{ij}^f)}\label{eqn:u_p}
 	\end{align}
 \end{subequations}
-to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match, in (\ref{eqn:m_p}), and not a match, in (\ref{eqn:u_p}).
+to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match in (\ref{eqn:m_p}), and not a match in (\ref{eqn:u_p}).
 Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) is equivalent to
@@
 
-Turning first to the posterior updates for $\bm{m}$ and $\bm{u}$, we first note that these parameters depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$.
+The posterior updates for $\bm{m}$ and $\bm{u}$ depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$.
 Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
-Then, conditional on $Z$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
+Then, conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
 \begin{align}
@@
 \end{subequations}
-Specifically, the $l + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}),and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{eqn:beta_update}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-We also use (\ref{likelihood_efficient}) to sample Z efficiently.
-For each pattern $p$, let $w_p =\frac{m_p}{u_p}$.
+Specifically, the $l + \sum_{k=1}^{f-1} L_k$ components of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provide the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
+However, the vectorized summations can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
+
+Similarly, the posterior update for $\bm{Z}$ depend on the the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
+For each pattern $p$, let $w_p =m_p/u_p$.
 The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
@@
 \end{align}
-This likelihood lends itself to markedly faster posterior computations.
 Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large.
 This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-To speed up computation, we break this sampling step into two.
+To speed up computation, we use \eqref{likelihood_efficient_z} to break this sampling step into two.
 
@@
 	\label{eqn:gibbs1}
-	P\left( \gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
 	\begin{cases} 
-		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
+		%\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
+		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
 		1- \pi^{(s+1)} , &   \text{otherwise}. \\
@@
 	\label{eqn:gibbs2}
-	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
+	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
 		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
@@
 We assume $C$ cores available for parallelization and a Gibbs sampler with $T$ iterations.
-Then, the overall computational complexity of \texttt{fabl} with hashing is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P)$.
+Then, the overall computational complexity of \texttt{fabl} with hashing is $O(F n_A n_B /C) + O(T n_B P / C)$.
 	\label{lemma:fabl}
@@
 The hashing procedure for all pairwise comparisons is also $O(F n_A n_B)$.
-With $C$ processors available, we can split these computations across $C$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{C} n_A n_B)$.
+With $C$ processors available, we can split these computations across $C$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(F n_A n_B /C)$.
 There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
@@
 Since $P << n_A$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_A n_B)$ under \texttt{BRL} to $O(n_B P)$ under \texttt{fabl}.
-With parallelization, this complexity is further reduced to $O(\frac{1}{C}n_B P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{C}n_B P)$
-	In summary, the total computational complexity for \texttt{fabl} is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P).$
+With parallelization, this complexity is further reduced to $O(n_B P /C)$, and so the entire Gibbs sampler has complexity $O(T n_B P / C)$
+	In summary, the total computational complexity for \texttt{fabl} is $O(F n_A n_B / C) + O(T n_B P / C).$
 \end{proof}
@@
 
-		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations.
+		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint batches for more manageable computations.
 Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.
-Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B and B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
-For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
-For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
-
-		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison vectors} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
-		the 
-\textcolor{teal}{memory-intensive}
-		\textcolor{teal}{comparison vectors} $\gamma^{ab}$ 
-		before continuing with the next chunk of data. 
-		In detail, we calculate
+Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
+For each $a$ and $b$, we compute comparison vectors for all records in $A^a \times B^b$ to construct the comparison matrix $\gamma^{ab}$.
+
+		 %For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
+
+		We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$, and delete the memory intensive matrix $\gamma^{ab}$ before continuing with the next batch of data. In detail, we calculate
 \begin{subequations}
 	\begin{align}
-		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, B_j \in B^b\}, \label{r_{p_j}^{ab}} \\
 		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
@@
 \end{subequations}
-These can computed \textcolor{teal}{sequentially} or in parallel.
-Summary statistics from each pairwise chunk comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
+These can computed sequentially or in parallel. Summary statistics from each pairwise batch comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
 \begin{subequations}
@@
 As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses less memory than storing the full comparison matrix $\gamma$.
-However, the memory requirements of each are still quadratic in nature.
-For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
+However, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome for very large linkage tasks.
 We next introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
@@
 These records have the same probability to be identified as the link for record $B_j$.
-Thus, we know that records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate.
+Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate.
 Rather than store all of these record labels, we store only a small number $S$.
-For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $SEI(r_{p_j}^{ab})$.
-We collect these memory reduced lists to form $SEI(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $SEI(\mathcal{R})$.
-
-Let $n_{A^a}$ and $n_{B^b}$ be the number of records in chunks $A^a$ and $B^b$ respectively.
-Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels.
-As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $SEI(\tilde{\gamma}) = \{\mathcal{P}, SEI(\mathcal{R}), \mathcal{N}\}$.
+For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form SEI$(r_{p_j}^{ab})$.
+We collect these memory reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.
+
+Let $n_a$ and $n_b$ be the number of records in batches $A^a$ and $B^b$ respectively.
+Instead of storing $n_A n_B$ record labels, with SEI we store at most $t_A \sum_{b = 1}^{t_B} n_b  P  S$ labels.
+As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
 We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.
@@
 
-	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-We use $F = 5$ binary comparisons with probabilities for matching and nonmatching pairs shown in Table \ref{Tab:distributions}.
-For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$.
-For $n_B/2$ of these records, there is no match in $A$, so we simulate $n_A$ comparison vectors from the $\bm{u}$ distribution.
-For the other $n_B/2$ of these records, there is one match in $A$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_A - 1$ comparison vectors from the $\bm{u}$ distribution.
-We compare the run-time of \texttt{fabl} (with no SEI) against \texttt{BRL} as we increase $n_A$ and $n_B$.
-Since we have five binary comparison fields with no missingness, the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
-
+		\afterpage{
 		\begin{table}[t]
@@
 			\end{tabular}
-		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Section \ref{speed}.}\label{Tab:distributions}
+			\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ in simulation study in Section \ref{speed}.}\label{Tab:distributions}
 		\end{table}
+
+		\begin{figure}[t]
+			\centering
+			\begin{subfigure}{.5\textwidth}
+				\centering
+				\includegraphics[width=0.95\textwidth]{../notes/figures/sadinle_speed_plot_new}
+				\caption{}
+				\label{fig:speed1}
+			\end{subfigure}%
+			\begin{subfigure}{.5\textwidth}
+				\centering
+				\includegraphics[width=0.95\textwidth]{../notes/figures/speed_plot_fixed_nB_new} 
+				\caption{}
+				\label{fig:speed2}
+			\end{subfigure}
+			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulations described in Section \ref{speed}.
+In (\ref{fig:speed1}), both $n_A$ and $n_B$ are increasing.
+We see quadratic growth in \texttt{BRL} and linear growth in \texttt{fabl}.
+In (\ref{fig:speed2}), only $n_A$ only is increasing.
+We see linear growth in \texttt{BRL} and approximately constant run-time in \texttt{fabl}.}
+			\label{fig:speed_sims}
+		\end{figure}
+	}
+
+	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
+We use $F = 5$ binary comparisons with probabilities for matching and nonmatching pairs shown in Table \ref{Tab:distributions}.
+For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$.
+For $n_B/2$ of these records, there is no match in $A$, so we simulate $n_A$ comparison vectors from the $\bm{u}$ probabilities.
+For the other $n_B/2$ of these records, there is one match in $A$, so we simulate $1$ comparison vector from the $\bm{m}$ probabilities, and $n_A - 1$ comparison vectors from the $\bm{u}$ probabilities.
+We compare the run-time of \texttt{fabl} (with no SEI) against \texttt{BRL} as we increase $n_A$ and $n_B$.
+Since we have five binary comparison fields with no missingness, the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
 
@@
 
-	\begin{figure}[t]
-		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
+%	\begin{figure}[t]
+%		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot_new} 
+%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
 We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
-		\end{center}
-	\end{figure}
+%		\end{center}
+%	\end{figure}
+%
+%	\begin{figure}[h!]
+%	\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_new} 
+%		\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
+We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
+%	\end{center}
+%\end{figure}
 
@@
 
-	\begin{figure}[h!]
-		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
-We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
-		\end{center}
-	\end{figure}
+%		\begin{figure}
+%		\begin{minipage}[c]{0.45\linewidth}
+%			\begin{center}
+%				\includegraphics[width=0.8\textwidth]{../notes/figures/sadinle_speed_plot_new}
+%				\caption{Simulations with increasing values of both $n_A$ and $n_B$. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
+%			\end{center}
+%		\end{minipage}
+%		\hfill
+%		\begin{minipage}[c]{0.45\linewidth}
+%			\begin{center}
+%				\includegraphics[width=0.8\textwidth]{../notes/figures/speed_plot_fixed_nB_new} 
+%				\caption{Simulations with $n_B$ = 500, We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
+%			\end{center}
+%		\end{minipage}%
+%	\end{figure}
 
@@
 Under each of these settings, we use 100 pairs of simulated data files in order to obtain uncertainty quantification on our performance metrics.
-We use uniform priors for all $\bm{m}$ and $\bm{u}$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
+We use uniform priors for the $\bm{m}$, $\bm{u}$, and $\pi$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
 We run the Gibbs sampler for 1000 iterations, and discard the first 100 as burn-in.
@@
 
-	\begin{table}[t]
-		\centering
-		\begin{tabular}[t]{llllll}
+	We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}.
+Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(Z_j \leq n_A)$.
+Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(\hat{Z}_j \leq n_A)$.
+The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2  (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
+In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
+	%In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.
+
+		\begin{table}
+		\centering
+		\begin{tabular}{llllll}
 
@@
 
-	We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}.
-Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(Z_j \leq n_A)$.
-Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(\hat{Z}_j \leq n_A)$.
-The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2  (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
-In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.
-
-	\begin{figure}[t]
+	\begin{figure}
 		\begin{center}
-			\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
+			\includegraphics[width=0.99\textwidth]{../notes/figures/sadinle_sim_plot3} 
 			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
@@
 
-Finally, our last simulation demonstrates our method's robustness to different values of $S$ for the SEI memory reduction procedure.
-We perform record linkage on one set of synthetic datafile described in Section \ref{accuracy} with 500 records in each datafile, 250 entities in common across datafiles, and 3 errors present across matching records.
-To achieve more drastic results, we perform SEI without chunking the data, that is, $t_A = t_B = 1$.
-In practice, if it is possible to create and store the comparison matrix for all record pairs at one time, there is no need to reduce the memory of the hashed matrix through SEI.
-For illustration however, it is easier to illustrate the effects of choices of $S$ in this setting.
-
-We preform linkage using SEI with $S = (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler.
-As any particular SEI implementation may improve or worsen linkage performance; if the SEI procedure happens to only remove pairs that are not matches, recall and precision will improve.
+Finally, our last simulation demonstrates the robustness of \texttt{fabl} to different values of $S$ for the SEI memory reduction procedure.
+We perform record linkage on one set of synthetic data file described in Section \ref{accuracy} with 500 records in each data file, 250 entities in common across data files, and 3 errors present across matching records.
+To achieve more drastic results, we perform SEI without batching the data, that is, $t_A = t_B = 1$.
+In practice, when it is computationally feasible to create and store $\gamma$ without batching, there is no need to reduce the memory of the hashed matrix through SEI.
+For illustration, however, it is easier to examine the effects of choices of $S$ in this setting.
+
+We perform linkage using SEI with $S \in (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler.
+Any particular SEI implementation may improve or worsen linkage performance; if the SEI procedure happens to only remove pairs that are not matches, recall and precision will improve.
 Therefore, we perform linkage under each setting 100 times, recording the linkage estimate $\hat{\bm{Z}}$, and recall and precision.
@@
 In Figure \ref{fig:SEI_Z}, the largest number of distinct linkage estimates occurs when $S = 1$.
-This makes sense, because the SEI procedure arbitrarily removes large numbers of record labels from consideration, resulting in a noisier estimate of the linkage structure.
+In this case, the SEI procedure arbitrarily removes large numbers of record labels from consideration, resulting in a noisier estimate of the linkage structure.
 The number of distinct linkage estimates decreases as $S$ increases, with larger values of $S$ providing results more similar to the linkage without SEI.
@@
 Setting $S=1$ can arbitrarily remove the index of a true match, leading the Gibbs sampler to concentrate probability on a false match, while larger values of $S$ produce results mirroring implementation with no SEI.
-We note however that even with $S=1$, the loss in precision is small.
+We note, however, that even with $S=1$, the loss in precision is small.
 
@@
 When evaluating the performance of a record linkage algorithm, researchers often examine posterior probabilities.
-By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce suspiciously high posterior probability for certain record pairs, providing a misleading perception of model performance.
+By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce artificially high posterior probability for certain record pairs, providing a misleading perception of model performance.
 
@@
 \end{figure}
+
 	\section{Case Studies}
@@
 	The country of El Salvador was immersed in civil war from 1980 to 1991.
-We are interested in estimating the total number of casualties from the war.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}.
-The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
-When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
+We are interested in estimating the total number of individuals killed in the war.
+We utilize lists of documented deaths from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises killings reported directly to the organization \citep{howland2008rescate, ball2000salvadoran, green2019civilian}.
+The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CDHES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
+When estimating the total number of individuals killed, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
 Instead, record linkage techniques must be used to merge data files before analyzing the data \citep{lum2013applications}.
@@
 There are several challenges with these data.
-First, both data files have been automatically digitized, which inherently leads to some degree of typographical error.
-Second, the only fields recorded are given name, last name, date of death, and place of death.
-It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-
-	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in $n_A = 4420$ records in CHDES and $n_B = 1323$ records in ERTL.
+First, the ERTL data file was automatically digitized, which inherently leads to some degree of typographical error.
+Second, it is common for villages in El Salvador to consist of only four to five extended families.
+This means there are a small number of last names in use, so many individuals have the same first and last name.
+Since the only fields recorded are given name, last name, date of death, and place of death, this leads to several instances in which distinct individuals have identical records.
+These individuals may be distant cousins, or are perhaps entirely unrelated, but would pose challenges for any record linkage method.
+
+	%It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
+
+	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in $n_A = 4420$ records in CDHES and $n_B = 1323$ records in ERTL.
 We standardize names to account for common misspellings and use a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
-Place of death is recorded by municipality and department within that municipality; however, since department is missing in 95\% of records in CHDES and 80\% of records in ERTL, we exclude department from our analysis.
+Place of death is recorded by municipality and department within that municipality; however, since department is missing in 95\% of records in CDHES and 80\% of records in ERTL, we exclude department from our analysis.
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We use uniform priors for the $\bm{m}$ and $\bm{u}$ parameters.
-
-	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in $5^5 \times 3 = 6025$ possible agreement patterns, with 1173 patterns realized in the data.
-However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
-
-	\begin{table}[t]
+We use uniform priors for the $\bm{m}$, $\bm{u}$, and $\pi$ parameters.
+
+	\begin{table}[!t]
 		\begin{tabular}[t]{llllll}
@@
 		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \cite{sadinle_bayesian_2017}.
-This setup leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
-	\end{table}
-
-	\begin{table}[t]
+This setup leads to 1875 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
+		%	\end{table}
+
+	%	\begin{table}[t]
 		\centering
@@
 		\caption{Construction of comparison vectors for El Salvador data for increased speed under \texttt{fabl}.
-This setup leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
+This setup leads to 432 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
 	\end{table}
 
+	\begin{figure}[p]
+	\begin{center}
+		\includegraphics[width=0.99\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+		\caption{Posterior distribution and Bayes estimate of overlap across the two files in the El Salvador case study.. We note they are quite similar under both methods.}
+		\label{fig:overlap-plot}
+	\end{center}
+	%	\end{figure}
+
+%	\begin{figure}[t]
+	\begin{center}
+		\includegraphics[width=0.99\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
+		\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
+		\label{fig:m-and-u}
+	\end{center}
+\end{figure}
+
+	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
+This results in 1875 possible agreement patterns, with 1173 patterns realized in the data.
+However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
+
 	Therefore, we perform our analysis with the agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
-Among the 216 possible agreement patterns, 159 are realized in the data.
-With this revised comparison specification, \texttt{fabl} runs in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
+Among the 432 possible agreement patterns, 159 are realized in the data.
+With this revised comparison specification, \texttt{fabl} runs in 109 seconds, approximately 3 times faster than the \texttt{BRL} run time of 313 seconds.
 The estimates of the $\bm{m}$ parameters under each method are similar, as shown in Figure \ref{fig:m-and-u}.
-Estimates of $\bm{u}$ are indistinguishable, and thus omitted.
+Estimates of the $\bm{u}$ parameters are indistinguishable, and thus omitted.
 Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
 
-	For completeness, we note that linkage with the more detailed comparison vectors requires 240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
+	For completeness, we note that linkage with the more detailed comparison vectors requires 945 seconds for \texttt{BRL}, and 1093 seconds for \texttt{fabl}.
 Apparently, the number of patterns is sufficiently many that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
 
-	Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both data files.
+	Through \texttt{fabl}, we arrive at a Bayes estimate of 178 individuals recorded in both data files.
 We calculate posterior samples of the size of the overlap across files by finding the number of links in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
-The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate identifies fewer matches than the Gibbs sampler identifies on average.
+The posterior 95\% credible interval for the overlap across files is (205, 236), indicating that the Bayes estimate identifies fewer matches than the Gibbs sampler identifies on average.
 This is because a large number of records in ERTL have multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
@@
 
-	\begin{figure}[t]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
-			\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
-			\label{fig:overlap-plot}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[t]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
-			\label{fig:m-and-u}
-		\end{center}
-	\end{figure}
+	%We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Appendix \ref{partial}.
+Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
+This means that after clerical review, the number of individuals replicated across data files would fall in the interval (136, 311), encapsulating the posterior credible interval.
+More or fewer records could be identified for clerical review by decreasing or increasing $\theta_R$.
+
+	%and a range of (140, 294) after the partial estimate and clerical review.
 
@@
 
-	\begin{table}[t]
+	The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients \citep{steorts_bayesian_2016}.
+The initial survey began in 1982, with follow-up surveys taken approximately every five years.
+As such, patients are surveyed at most once in a given year, and many patients are surveyed across multiple years.
+In addition, patients can either drop out of the study, pass away, or enter as new patients.
+Hence, the assumptions of our model hold for this study.
+We seek to link records over the $n_A = 20485$ individuals from 1982 to the $n_B = 17466$ individuals from 1989.
+The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
+However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone data files.
+
+	\begin{table}
 		\centering
@@
 
-	The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients \citep{steorts_bayesian_2016}.
-The initial survey began in 1982, with follow-up surveys taken approximately every five years.
-As such, patients are surveyed at most once in a given year, and many patients are surveyed across multiple years.
-In addition, patients can either drop out of the study, pass away, or enter as new patients.
-Hence, the assumptions of our model hold for this study.
-We seek to link records over the $n_A = 20485$ individuals from 1982 to the $n_B = 17466$ individuals from 1989.
-The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
-However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone data files.
-
-	We link records using sex, date of birth, and location using the thresholds shown in Table \ref{Tab:nltcs-comparisons}.
+	We link records using sex, location, and day, month, and year of birth using the criteria shown in Table \ref{Tab:nltcs-comparisons}.
 Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx 400,000,000$ record pairs would require approximately 8GB of memory.
 Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
-{\color{red}{However, through the method described in Section \ref{scaling},  we perform 30 smaller comparison tasks, using $t_A = 1$ and $t_B = 30$.
+However, through the method described in Section \ref{scaling},  we perform 30 smaller comparison tasks, using $t_A = 1$ and $t_B = 30$.
 We conduct linkage with all record indices recorded and also with SEI using $S=10$, and obtain identical results.
 The hashed $\tilde{\gamma}$ without SEI is about 2.2 GB, and with SEI, it is about 760 MB.
-Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.}}
+Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.
 
@@
 
-	\begin{figure}[h]
+	\begin{figure}
 		\begin{center}
@@
 
+	%\hypertarget{violations}{%
+		%	\subsection{Violations of One-to-One Matching}\label{violations}}
+	%
+	%As noted previously, the application exhibits many situations in which bipartite matching is difficult.
+To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation.
+These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $A$ has
+	%multiple plausible matches in $B$. 
+	%
+	%Table \ref{Tab:rosa-maria} illustrates an example.
+Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical.
+In addition, using the modified Levenstein distance, the comparison vectors $\gamma_{2776, 825}$ and $\gamma_{2776, 826}$ are exactly identical.
+	%
+	%\begin{table}
+	%	\centering
+	%	\begin{tabular}[h!]{llllllll}
+		%		\hline
+		%		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
+		%		\hline
+		%		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
+		%		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
+		%		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
+		%		\hline
+		%	\end{tabular}
+	%	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
+	%\end{table}
+	%
+	%Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation.
+Under \texttt{fabl}, both records in $B$ match to the same record in $A$ throughout the Gibbs process, creating consistent violations of one-to-one matching.
+Under \texttt{BRL}, the Gibbs process creates one matching configuration and stays there for a while.
+However, if one pair ``unmatches,'' then the other record has a chance to latch on.
+Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing.
+	%
+	%Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability.
+\texttt{BRL} in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate.
+Thus relaxing the one-to-one constraint has served as a useful tool for identifying model misspecification.
+	%
+	%\begin{figure}[h!]
+	%\begin{center}
+	%\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
+	%\caption{Gibbs sampling in situation with multiple plausible matches.
+Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability.
+Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
+	%\end{center}
+	%\end{figure}
+
 	\section{Conclusion}
@@
 
-	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing all $n_A  n_B$ record pairs still can be prohibitive for larger linkage tasks.
+	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing the comparisons for all $n_A  n_B$ record pairs still can be prohibitive for larger linkage tasks.
 Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
@@
 		\section{Derivations of Full Conditionals}\label{app:derivations}}
-\textcolor{red}{Note: Please use s, s+1 consistently throughout.
-Or supress the notation and tell the reader.
-It looks weird to go back and forth.
-I've tried to provide an intermediate solution.
-Perhaps Jerry can state if he would think this is okay as an Editor.}\\
-
-\textcolor{teal}{We provide detailed derivations of the full-conditionals provided in Section~\ref{sec:fast-beta-linkage}.
-In our derivations below, we assume that the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the final update of the full conditional that appears in Section~\ref{sec:fast-beta-linkage}.}
-
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{fl}$, we have
-\begin{align*}
-	\mathcal{L}(m_{fl}|\gamma, \alpha_{fl}, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
-\end{align*}
-where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
-Analogous procedures lead to the posterior distribution $\mathcal{L}(u_{fl}| \gamma, \beta_{fl}, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
+
+	We provide detailed derivations of the full-conditionals provided in Section \ref{sec:fast-beta-linkage}.
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
+For a particular $m_{fl}$, we have
+	\begin{align}
+		\mathcal{L}(m_{fl}|\gamma, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
+	\end{align}
+	where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
+Analogous procedures lead to the posterior distribution $\mathcal{L}(u_{fl}| \gamma, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
 Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
-
-$$\bm{m}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})),$$
-$$\bm{u}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).$$
-
-Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ and hyperparameters.
+	\begin{align}
+	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \\
+	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).
+	\end{align}
+
+	Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j \leq n_A)$ encoded by $\bm{Z}$ and hyperparameters.
 We have the full conditional
-\begin{align*}
-	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
+	\begin{align}
+		p(\pi |\gamma, \bm{Z}, \bm{m}, \bm{u}) &\propto p(\bm{Z}|\pi)p(\pi) \\
 		&\propto \pi^{n_{AB}(\bm{Z})} (1-\pi)^{n_B - n_{AB}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
 		&\propto \pi^{n_{AB}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_A - n_{AB}(\bm{Z}) + \beta_{\pi} -1}.
-\end{align*}
-Thus, $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
-
-To express the full conditional for $\bm{Z}$, we consider the cases where $Z_j$ does not have a link in $A$ and where $Z_j$ does have a link in $A$.
-Because we sample each $Z_j$ independently of all other $Z_{j'}$ (for $j \neq j'$), we need only the full conditional for each individual $Z_j$.
-{\color{red}{Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.}} Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+	\end{align}
+	Thus, $\pi^{(s+1)}|\gamma, \bm{Z}^{(s+1)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
+
+	Due to the independence in the fast beta prior in \eqref{eqn:fast_beta_prior}, we can obtain the full conditional for $\bm{Z}$ through the full conditionals for each individual $Z_j$.
+Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.
+We have
+	\begin{align}
+		p(\bm{Z} | \gamma, \bm{m}, \bm{u}, \pi) &= \prod_{j=1}^{n_B} p(Z_j | \gamma_{.j}, \bm{m}, \bm{u}, \pi).
+		%&\propto p(\Gamma| \bm{Z}, \bm{m}, \bm{u}) p(\bm{Z} | \pi) \\
+		%&=\prod_{j=1}^{n_B} p(\Gamma_{.j}| Z_j, \bm{m}, \bm{u}) p(Z_j | \pi) \\
+	\end{align}
+
+Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
 \begin{align}
@@
 \begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
 	&= w_{qj}  c_j.
@@
 		w_{qj}, & q \leq n_A; \\
-		1, &  q  = n_A + j. \
+		1, &  q  = n_A + j.
 	\end{cases}
@@
 		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
-		1 - \pi^{(s+1)}, & q  = n_A + j. \
+		1 - \pi^{(s+1)}, & q  = n_A + j.
 	\end{cases}
@@
 
-\textcolor{teal}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
-\begin{align}
-p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi) \propto \prod_q \left(\prod_j
-p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi \right)
-\right).
-\end{align}
-Therefore, we update the full conditional for $\bm{Z}^{(s+1)}$ in parallel.
-This contrasts the Gibbs sampler of  \cite{sadinle_bayesian_2017}, where the update to the matching matrix is sequential and accomplished using a partially collapsed Gibbs sampler, where one integrates out $\pi.$ We do not opt for this approach as it would lead to similar sequential updates, which are computationally slow.}
+%	To express the full conditional for $\bm{Z}$, we consider the cases where $Z_j$ does not have a link in $A$ and where $Z_j$ does have a link in $A$.
+Because we sample each $Z_j$ independently of all other $Z_{j'}$ (for $j \neq j'$), we need only the full conditional for each individual $Z_j$.
+Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.
+Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+%	\begin{align}
+%		p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+%	\end{align}
+%	When $Z_j = q$ for some $q\leq n_A$, we have
+%	\begin{align}
+%		p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+%	\end{align}
+%	We multiply and divide by the $u$ parameters for the matching record pair to obtain
+%	\begin{align}
+%		p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+%		&= w_{qj}  c_j.
+%	\end{align}
+%	We can divide the result of each case by $c_j$ to get
+%	\begin{align}
+%		p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
+%			w_{qj}, & q \leq n_A; \\
+%			1, &  q  = n_A + j. \
+%		\end{cases}
+%	\end{align}
+%	Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
+%	\begin{align}
+%		\label{eqn:z_full_conditional2}
+%		p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+%		\begin{cases} 
+%			\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
+%			1 - \pi^{(s+1)}, & q  = n_A + j. \
+%		\end{cases}
+%	\end{align}
 
@@
 Specifically, we have
-	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
+\begin{align}
+	L(\hat{Z_j}, Z_j)=\begin{cases} 
 		0,  & \text{if } Z_j = \hat{Z_j}; \\
@@
 		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
-	\end{cases}\] 
+	\end{cases}
+\end{align}
 Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
@@
 In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
-	$$\hat{Z}_j =\begin{cases} 
+	\begin{align}
+		\hat{Z}_j =\begin{cases} 
 		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
 		0,  & \text{otherwise}. \\
-	\end{cases}$$
+	\end{cases}
+\end{align}
 Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
@@
 For a more in-depth explanation of this function and the induced Bayes estimate, see \cite{sadinle_bayesian_2017}.
+
+	\hypertarget{appendix-sim}{%
+		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
+	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+It is set up with one error across the linkage fields and 50 duplicates across files.
+Traceplots across other settings exhibit similar behavior.
+Note that traceplots for $\bm{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
+
+	\begin{figure}[h]
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
+			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
+			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
+		\end{center}
+%	\end{figure}
+
+%	\begin{figure}[!h]
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
+			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
+		\end{center}
+	\end{figure}
+
+\clearpage
 
@@
 
-\begin{figure}[t]
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
+		\includegraphics[width=0.99\textwidth]{../notes/figures/sadinle_sim_plot_partial2} 
 		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
@@
 		\label{fig:sadinle_simulation_partial}
-	\end{center}
-\end{figure}
-
-	\hypertarget{appendix-sim}{%
-		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
-	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
-It is set up with one error across the linkage fields and 50 duplicates across files.
-Traceplots across other settings exhibit similar behavior.
-Note that traceplots for $\bm{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
-
-	\begin{figure}[!h]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
 	\end{center}
@@
 
-		\begin{table}[t]
+		\begin{table}[h]
 		\centering
@@
 
-	\begin{figure}[h!]
-		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
+%	\begin{figure}[h!]
+%		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2_new} 
+%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
 We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
+%		\end{center}
+%	\end{figure}
+%	
+%	\begin{figure}[h!]
+%		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB2_new} 
+%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
 We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
-		\end{center}
+%		\end{center}
+%	\end{figure}
+
+	\begin{figure}[h]
+	\centering
+	\begin{subfigure}{.5\textwidth}
+		\centering
+		\includegraphics[width=0.95\textwidth]{../notes/figures/sadinle_speed_plot2_new}
+		\caption{}
+		\label{fig:app-speed1}
+	\end{subfigure}%
+	\begin{subfigure}{.5\textwidth}
+		\centering
+		\includegraphics[width=0.95\textwidth]{../notes/figures/speed_plot_fixed_nB2_new} 
+		\caption{}
+		\label{fig:app-speed2}
+	\end{subfigure}
+	\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulations described in Appendix \ref{app:appendix-speed}.
+In (\ref{fig:app-speed1}), both $n_A$ and $n_B$ are increasing.
+We see quadratic growth in \texttt{BRL} and linear growth in \texttt{fabl}.
+In (\ref{fig:app-speed2}), only $n_A$ only is increasing.
+We see linear growth in \texttt{BRL} and approximately constant run-time in \texttt{fabl}.}
+	\label{fig:app-speed_sims}
 \end{figure}
@@
 
-	\begin{figure}[!h]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} \label{fig:m_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} \label{fig:u_trace}
-		\end{center}
+%	\begin{figure}[!h]
+%		\begin{center}
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
+%			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
+%		\end{center}
+%	\end{figure}
+%
+%	\begin{figure}[!h]
+%	\begin{center}
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/pi_trace} 
+%		\caption{Traceplot for $\pi$ parameter in El Salvador case study.} \label{fig:pi_trace}
+%	\end{center}
+%\end{figure}
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
+%			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} \label{fig:m_trace}
+%		\end{center}
+%	\end{figure}
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
+%			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} \label{fig:u_trace}
+%		\end{center}
+%	\end{figure}
+
+    \begin{figure}[h!]
+	\centering
+	\begin{subfigure}[b]{0.475\textwidth}
+		\centering
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/overlap_trace} 
+		\caption{Traceplot for overlap between files.}    
+		\label{fig:overlap_trace}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}  
+		\centering 
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/pi_trace} 
+		\caption{Traceplot for $\pi$.}  
+		\label{fig:pi_trace}
+	\end{subfigure}
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/m_trace} 
+		\caption{Traceplot for $\bm{m}$ parameters.}   
+		\label{fig:m_trace}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/u_trace} 
+		\caption{Traceplot for $\bm{u}$ parameters.}
+		\label{fig:u_trace}
+	\end{subfigure}
+	\caption{Traceplots for parameters of interest in El Salvador case study in Section \ref{el_salvador}}.
+	\label{fig:mean and std of nets}
 \end{figure}
 \end{document}
+
+\end{document}

### 2023-07-29 SEI_sensitivity/sadine_accuracy_plots.R

@@ -1,0 +1,95 @@
+library(tidyverse)
+
+result_files <- list.files(path = "SEI_sensitivity/out/sim_acc", full.names = T)
+results <- lapply(result_files, readRDS) %>%
+  do.call(rbind, .) %>%
+  as.data.frame()
+
+results_df <- results %>%
+  select(-RR, -Time_Elapsed) %>%
+  filter(Method %in% c("fabl_resolved", "BRL")) %>%
+  pivot_longer(cols = 1:3, names_to = "Metric") %>%
+  group_by(Method, Metric, Error, Overlap) %>%
+  summarize(avg = mean(value),
+            lower = quantile(value, .025),
+            upper = quantile(value, .975),
+            .groups = "drop") %>%
+  mutate(Metric = factor(Metric, c("Recall", "Precision", "F-Measure")),
+         Error = factor(Error, c("One Error", "Two Errors", "Three Errors")),
+         Overlap = case_when(
+           Overlap == 50 ~ "10% Overlap",
+           Overlap == 250 ~ "50% Overlap",
+           Overlap == 450 ~ "90% Overlap"
+         ),
+         Method = case_when(
+           Method == "BRL" ~ "BRL",
+           Method == "fabl_resolved" ~ "fabl"
+         ))
+
+acc_plot <- results_df %>%
+  ggplot(aes(x = Metric, y = avg,
+             ymin  = lower, ymax = upper,
+             color = Method)) +
+  geom_pointrange(position = position_dodge2(width = .5),
+                  size = .3) +
+  facet_grid(Overlap ~ Error) +
+  labs(x = NULL,
+       y = NULL,
+       color = "Method") +
+  theme_bw(base_size = 7) +
+  theme(plot.title = element_text(hjust = .5),
+        plot.subtitle = element_text(hjust = .5)) +
+  scale_color_brewer(palette="Set1")
+
+acc_plot
+
+ggsave(filename = "notes/figures/sadinle_sim_plot3.png",
+       plot = acc_plot)
+
+results_partial_df <- results %>%
+  mutate(DR = 1 - RR) %>%
+  select(-RR, -Time_Elapsed) %>%
+  filter(Method %in% c("fabl_partial", "BRL_partial")) %>%
+  pivot_longer(cols = c(1, 2, 3, 7), names_to = "Metric") %>%
+  group_by(Method, Metric, Error, Overlap) %>%
+  summarize(avg = mean(value),
+            lower = quantile(value, .025),
[…]

### 2023-07-30 ba_resubmit/kundinger_supplement_final.tex
add supplement.

--- 2023-07-20 ba_resubmit/kundinger_final.tex
@@
 
+%%% only for appendix
+\renewcommand{\appendixname}{Supplement}
+\renewcommand{\thefigure}{SI\arabic{figure}} \setcounter{figure}{0}
+\renewcommand{\thetable}{SI\arabic{table}} \setcounter{table}{0}
+\renewcommand{\theequation}{SI\arabic{table}} \setcounter{equation}{0}
+%%%
+
 \begin{document}
@@
 \begin{frontmatter}
-\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
+\title{Supplementary Material for ``Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)"}
 %\title{A Sample Document\thanksref{T1}}
@@
 
-\begin{abstract} 
-Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.
-However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
-Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, \textcolor{red}{unlike} the sequential updates of \texttt{BRL}.}
-\textcolor{teal}{We ensure \textcolor{red}{\text{fabl} results in a bipartite matching} using decision theoretic post-processing techniques from the existing literature.}
-\textcolor{red}{To reduce computational overhead and memory costs, we use hashing and a new technique that we call storage efficient indexing. }
-%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
-Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
-%}
-%This allows
-%us to use hashing techniques that reduce computational overhead.
-This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
-Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
-\end{abstract}
-
-%% ** Keywords **
-\begin{keyword}
-\kwd{record linkage}
-\kwd{entity resolution}
-\kwd{data cleaning}
-\kwd{parallelization}
-\kwd{hashing}
-\end{keyword}
-
 \end{frontmatter}
-
-%% ** Mainmatter **
-
-\section{Introduction}
-\label{sec:introduction}
-
-Before conducting data analysis, it is often necessary to identify duplicate records across two data files.
-This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
-In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.
-
-Many probabilistic record linkage methods are rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
-In their approach, the data analyst first creates comparison vectors for each pair of records in the data files.
-These vectors indicate how similar the records are on a set of variables measured in both files, known as the linkage variables.
-Using these comparison vectors, the analyst classifies each pair as a match or nonmatch using a likelihood ratio test.
-Crucially, these decisions are made independently for each pair.
-The \cite{fellegi_theory_1969} approach has been extended for a wide variety of applications \citep[e.g.,][]{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}.
-An alternative paradigm is to model the linkage variables directly \citep[e.g.,][]{tancredi2011hierarchical, steorts_bayesian_2016, marchant_distributed_2019, betancourt2022prior}.
-In this article, we build on the contributions to the comparison vector approach.
-
-The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity.
-However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
-Thus, when the procedure results in many-to-one matches, some of these links must be false.
-To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}
-
-Alternatively, analysts can embed one-to-one matching requirements into the model specification, at an additional computational cost.
-\cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
-\cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records.
-\cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
-Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
-\texttt{BRL} has been shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
-
-In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. 
-\textcolor{red}{We use independent prior distributions for the matching status of each record, an idea suggested in \cite{wortman2019}.}
-%% Jerry thought the wording below was too vague and made a suggestion that is more direct and clear. 
-%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}}
-%
-\textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, \textcolor{red}{unlike the} sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
-This \textcolor{red}{approach} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
-In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
-
-In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
-Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching.
-Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}.
-Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
-Finally, Section \ref{discussion} summarizes our contributions and highlights areas for further research.
-
-\section{Review of Fellegi-Sunter Approaches for Record Linkage}
-\label{sec:review-of_prior-work}
-
-Consider two data files $A$ and $B$ comprising $n_A$ and $n_B$ records, respectively, and including $F$ linkage variables measured in both files.
-For $i=1, \dots, n_A$, let record $i$ be given by $A_i = (A_{i1}, \dots, A_{iF})$, so that $A = (A_i : i = 1, \dots, n_A)$.
-Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.
-Without loss of generality, denote files such that $n_A \geq n_B$.
-
-Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define a \emph{comparison vector} $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
-For notational convenience, we define the \emph{comparison matrix} $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
-
-When linking variable $f$ is categorical, a common way to define $\gamma_{ij}^f$ is an indicator for exact agreement.
-For example, if zip code is linking variable $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not.
-For numerical linking variables, we can use the absolute difference of the two values.
-For example, if age is linking variable $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart.
-For text variables like names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}.
-We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.
-
-More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for linking variable $f$ of records $A_i$ and $B_j.$ The range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$.
-Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement).
-Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.
-
-%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
-As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
-To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}
-
-Having defined comparison vectors, we now turn to the \cite{fellegi_theory_1969} model for record linkage.
-We begin with notation for defining linkages.
-Under bipartite matching, the set of matches across $A$ and $B$ can be represented in two equivalent ways.
-First, we may use a matrix $\Delta \in \{0, 1\}^{n_An_B}$, where
-\begin{align}
-\Delta_{ij} =
-\begin{cases}
-1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
-0, & \text{otherwise}.\\
-\end{cases}
-\end{align}
-This sparse matrix representation can become cumbersome for large linkage tasks.
-More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
-\begin{align}
-Z_{j} =
-\begin{cases}
-i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
-n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
-\end{cases}
-\end{align}
-We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.
-When presenting the models, we use both representations for convenience.
-
-Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.
-Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
-For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
-Using these independence assumptions,  one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.
-We have
-\begin{subequations}
-\begin{align}
-&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
-&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
-&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:lambda_dist}
-\end{align}
-\end{subequations}
-Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match.
-When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
-Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
-The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
-
-\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.
-He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.
-\textcolor{red}{However, he replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\mathbf{Z}$ that he calls the ``beta distributions for bipartite matching."}
-%
-%As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  
-This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
-Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
-%
-It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
-Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A !
-/ (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely.
-Thus, the prior distribution used by \cite{sadinle_bayesian_2017} is given by
-\begin{align}
-\label{eqn:sadinle_prior}
-p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
-\end{align}
-where $\text{B}(\cdot, \cdot)$ represents the Beta function.
-This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$.
-This makes the sampler inherently \textcolor{teal}{sequential}, which can be slow when working on linkage tasks with more than a few thousand records.
-
-\section{Fast Beta Linkage}
-\label{sec:fast-beta-linkage}
-
-\textcolor{red}{For \text{fabl}, we also use (\ref{eqn:fs_model}) and (\ref{eqn:m_dist}), along with Dirichlet prior distributions for the $\mathbf{m}$ and $\mathbf{u}$ parameters.
-However, instead of the prior over $\bm{Z}$}
-%Instead of the prior over $\bm{Z}$ 
-from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, 
-\textcolor{teal}{we propose an alternative, called the fast beta linkage (fabl) prior.}
-%the fast Beta prior as follows. 
-For each $Z_j$, we have
-
-\begin{subequations}
-\begin{align}
-	p(Z_j = q| \pi) &= \begin{cases} 
-	\frac{1}{n_A}\pi,  & q \leq n_A;\\
-	1-\pi, &  q  = n_A + j;  \\
-\end{cases} \label{eqn:fast_beta_prior}  \\
-\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \label{eqn:fast_beta_prior2}.
-\end{align}
-\end{subequations}
-We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match.
-The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$
-
-\begin{rem}
-\textcolor{red}{
-In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
-This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
-In our preliminary studies, this highly informative prior weighting on matching can result in overly optimistic match rates.
-Hence, \textcolor{red}{we prefer (\ref{eqn:fast_beta_prior}).} We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
-}
-\end{rem}
-
-\begin{rem}
-Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in \textcolor{red}{the \cite{fellegi_theory_1969} model} estimates the proportion of record pairs that are matches.
-In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
-In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by \textcolor{red}{$1/n_A$} and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
-As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
-\end{rem}
-
-\textcolor{red}{In many record linkage tasks, the records in $A$ or $B$ have missing fields, so that some $\gamma_{ij}^f$ could be missing.
-To handle this, we follow the approach described in \cite{sadinle_bayesian_2017}.
-Let $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-We assume that any missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
-For likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}, so that we can marginalize over the missing data and perform all computation using the observed data \citep{sadinle_detecting_2014, sadinle_bayesian_2017}.}
-
-%\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
-To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
-It is important to note that for likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
-Therefore, we can marginalize over the missing data and perform all computation using the observed data, leading to (\ref{eqn:likelihood}).
-For further details on these connections, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
-%}
-
-\textcolor{teal}{We summarize \textcolor{red}{the model for \text{fabl}} as follows:}
-
-\begin{subequations}
-\begin{align}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
-	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}),\quad  \forall  \; f = 1, \ldots, F, \label{eqn:m} \\
-	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}), \quad \forall \; f = 1, \ldots, F,  \label{eqn:u}\\
-	p(Z_j = q| \pi)  &=
-	\begin{cases} 
-		\frac{1}{n_A}\pi,  & q \leq n_A; \\
-		1-\pi, &  q  = n_A + j; \\
-	\end{cases} \label{eqn:z}\\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi-prior},
-\end{align}
-\end{subequations}
-\textcolor{teal}{where recall that \textcolor{red}{$(\alpha_{\pi}, \beta_{\pi})$} are fixed hyper-parameters.}
-
-%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
-%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
-
-%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
-%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
-
-\textcolor{teal}{We \textcolor{red}{now} present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m},$ $\bm{u},$ and $\pi$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi-prior}).
-%While we could marginalize over $\pi$ and derive a collapsed Gibbs sampler, in the spirit of \cite{sadinle_bayesian_2017}, this leads to sequential updates, which are computationally slow.
-Instead, we proceed with a traditional Gibbs sampler, where the updates to the matching label (or coreference matrix) are parallel, allowing for more efficient updates.
-Refer to Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
-%%
-We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-\textcolor{blue}{Brian: state how you initialize $\pi$.} \textcolor{teal}{Denote the current draw of the sampler by $s$ and the updated draw by $s+1$.} \textcolor{red}{To update $\mathbf{m}$ and $\mathbf{u}$, for all, $f=1, \ldots, F,$ sample}
-
-%\begin{subequations}
-%\begin{align}
-%	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  \\
-%	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
-%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-%\end{align}
-%\end{subequations}
-
-%\begin{subequations}
-%
-%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-%\end{align}
-%\end{subequations}
-
-\begin{align}
-	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  
-\end{align}
-and	
-\begin{align}
-	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} 
-\end{align}
-
-where
-
-\begin{align}
-\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \quad \text{and}\\
-\beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-\end{align}
-
-\textcolor{red}{Next, we sample $\pi$ from its full conditional, given by}
-%Next, the full conditional for $\pi$ is given by
-\begin{align}
-\label{eqn:pi}
-	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
-\end{align}
-where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.
-
-Lastly, we sample $\bm{Z}$ componentwise from the full conditional for each $Z_j$:
-\begin{align}
-	\label{eqn:z_full_conditional}
-	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)},  \pi^{(s+1)}\right) \propto
-	\begin{cases} 
-		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
-		1 - \pi^{(s+1)}, & q  = n_A + j, \\
-	\end{cases}
-\end{align}
-
-where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
-\begin{align}
-	\label{eqn:fs_weight}
-	w_{ij}^{(s)} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
-\end{align}
-
-\textcolor{teal}{
-The full conditional in (\ref{eqn:z_full_conditional}) does not depend on $\bm{Z}^{(s)},$ implying that 
-\begin{align}
-p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}) \propto \prod_q \left(\prod_j
-p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right)
-\right).
-\end{align}
-}
-
-\textcolor{teal}{Thus, we can make parallel updates to $\bm{Z}$ instead of sequential updates as in \cite{sadinle_bayesian_2017}.
-In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), \textcolor{red}{in contrast to} \textcolor{red}{other} Bayesian record linkage \textcolor{red}{methods} \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}.
-}
-%See Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
-
-%Derivations for the full conditional distributions can be found in Appendix~\ref{app:derivations}.
-
-\textcolor{teal}{Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
-Since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-To obtain a Bayes estimate that fulfills the \textcolor{red}{bipartite} requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-See Appendix \ref{bayes-estimate} for details regarding the initial Bayes estimate and \textcolor{red}{this} post-processing procedure.}
-
-%Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
-
-\section{Efficient and Scalable Implementation}
-\label{sec:efficiency}
-
-The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A  n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors.
-One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}.
-In deterministic blocking, the modeler chooses fields thought to be highly reliable and only compares records that agree on those fields.
-The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains.
-Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking \textcolor{teal}{often}
- undesirable %in some situations 
-\citep{steorts_comparison_2014}.
-
-After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing.
-For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields.
-However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the \textcolor{red}{classical} Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
-
-With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage.
-First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors.
-Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
-
-\hypertarget{data-representation-hashing-and-storage}{%
-	\subsection{Data Representation, Hashing, and
-		Storage}\label{data-representation-hashing-and-storage}}
-
-Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. 
-\textcolor{red}{The quantity $P$ is
- bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1),$ which clearly depends only on $F$ and $L_f.$ It does not depend (or scale with) with $n_A$ or $n_B.$}
-% This comment below seems obvious and not needed.  
-% where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.
-
-%is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
-
-To obtain a memory efficient representation, we map the agreement pattern of each record pair to a unique integer.
-\cite{enamorado2019using} accomplished this through a hashing function, which we modify to explicitly handle missing values:
-\begin{align}
-	\label{eqn:hashing}
-	h^{*}(\gamma_{ij}) = \sum_{f = 1}^F I_{obs}(\gamma_{ij}^f)2^{\gamma_{ij}^f + I(f>1)\sum_{d=1}^{f-1}(L_d)}.
-\end{align}
-
-%\textcolor{red}{Do you mean what I have below or something else? This still is a bit hard to parse.}
-We then map \textcolor{teal}{the integers in (\ref{eqn:hashing})} to sequential integers $\{1, \ldots, P\}$.
-\textcolor{teal}{Denote} each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-When record pair, \textcolor{teal}{$(i,j)$,} exhibits agreement pattern $p$, \textcolor{teal}{let} $\gamma_{ij} = h_p$.
-
-In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
-This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
-
-\begin{example}
-\label{ex:one-hot}
-%% Jerry propose with (want to avoid using with twice in the same sentence :)
-\textcolor{teal}{Consider} five fields with binary agreements and possible missingness (NA).
-The number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
-\textcolor{red}{Suppose that} records $A_5$ and $B_7$ exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, \text{NA}, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
-Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
-\textcolor{red}{All records with a hashed value of 1066 map to the integer 42. Then, $\gamma_{5,7} = h_{42}.$}
-%We map the unique values of $h^{*}(\gamma)$ to sequential integers. \textcolor{teal}{Suppose that 1066 maps to 42. This implies $\gamma_{5,7} = h_{42}.$}
-%If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ 
-This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
-\end{example}
-
-We then identify the records in $A$ with comparison vectors corresponding to each pattern $p$ for each record $B_j$.
-We denote this set $r_{p_j} = \{i \in 1, \dots, n_A | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
-We compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$,
-\begin{align}\label{eqn:N}
-N_{p_j} = |r_{p_j}| = \sum_{i=1}^{n_A} I(\gamma_{ij} = h_p).
-\end{align}
-We collect these counts in $\mathcal{N} = \{N_{p_j} |p \in 1, \ldots, P, j \in 1, \ldots, n_B \}$.
-
-The set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for purposes of writing the likelihood function for \texttt{fabl}.
-To see this, we employ the condensed notation
-\begin{subequations}
-	\begin{align}
-		m_p =  \textcolor{red}{p(\Gamma_{ij} = h_p|Z_j = i)} = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:m_p} \\
-		u_p =  \textcolor{red}{p(\Gamma_{ij} = h_p|Z_j \neq i)} = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:u_p}
-	\end{align}
-\end{subequations}
-to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match, in (\ref{eqn:m_p}), and not a match, in (\ref{eqn:u_p}).
-Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) \textcolor{teal}{can be expressed as}
-%is equivalent to
-
-\begin{align}\label{likelihood_efficient}
-	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\gamma}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m_p^{I(Z_j = i)}u_p^{1 - I(Z_j = i)}. 
-\end{align}
-The likelihood in (\ref{likelihood_efficient}) allows for more
-efficient \textcolor{teal}{approximate posterior inference of the Gibbs sampler in Section~\ref{sec:fast-beta-linkage},}  as we \textcolor{teal}{describe in the next section}.
-
-%% These are not the full conditional distributions.
-%for $\bm{m}$, $\bm{u}$, and $\bm{Z}$
-
-\hypertarget{efficient-posterior}{%
-	\subsection{Efficient Posterior Inference }\label{efficient-posterior}}
-
-\textcolor{teal}{In our derivations below, we assume the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the updates to each full conditional distribution based upon our computational proposals using hashing.}
-
-%\begin{enumerate}	
-%\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.
-Also, do you more formally mean something like as follows: Consider $\gamma_{q,q} = h_p \mid Z_j = q$.
-It seems that the conditioning may actually be important here.
-I mention this for two reasons, it appears a bit strange in equation (22).
-When I think through it, it's okay.
-I just wanted to bring it up as a conversation before submission.}
-%\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
-%\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1.
-Example, it's missing on the update for $m$ and $u$ so it's not immediately clear how one would draw this directly.
-Wouldn't it just be a vectorized Dirichlet, where the computational savings come from vectorizing the hyperparameters of the Dirichlet?
-%\end{enumerate}
-
-Turning first to the \textcolor{teal}{full conditional updates}
-%posterior updates 
-for $\bm{m}$ and $\bm{u}$, we observe that these parameters only depend on the \textcolor{teal}{observed} data  through quantities that are calculated through $\mathcal{N}$ and $\mathcal{P}$.
-Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
-Conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional distribution for $\bm{m}$ and $\bm{u}$ as
-\begin{align}
-	\mathcal{L}(\bm{m}, \bm{u} \mid \tilde{\gamma}, \bm{Z}, \pi)  &=  \prod_{p=1}^P m_p^{n_p(\bm{Z})}u_p^{N_p - n_p(\bm{Z})}  \label{likelihood_efficient_m_u}.
-\end{align}
-Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be a concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
-The terms needed for the \textcolor{teal}{full conditional} updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
-\begin{subequations}
-	\begin{align}
-		\bm{\alpha(Z}^{(s)}\bm{)} &= \bm{\alpha_0} + \sum_{p=1}^P n_p\left(\bm{Z}^{(s)}\right) \times e(h_p), \label{efficient_alpha} \\
-		\bm{\beta(Z}^{(s)}\bm{)} &= \bm{\beta_0} + \sum_{p=1}^P \left(N_p - n_p\left(\bm{Z}^{(s)}\right)\right) \times e(h_p). \label{efficient_beta}
-	\end{align}
-\end{subequations}
-%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-\textcolor{teal}{\textcolor{red}{We can} compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
-
-\textcolor{teal}{Turning to sampling for $\pi,$ it remains the same as in (\ref{eqn:pi}), \textcolor{red}{so we do not repeat it.} }
-
-We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
-For each pattern $p$, let $w_p = m_p/u_p$.
-The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
-\begin{align}
-	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
-\end{align}
-This likelihood lends itself to markedly faster posterior computations.
-Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large.
-This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-\textcolor{red}{We use two steps in sampling each $Z_j$.}
-%To speed up computation, we break this sampling step into two.
-
-We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
-
-\begin{align}
-	\label{eqn:gibbs1}
-	p\left( \Gamma_{Z_j^{(s+1)},j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
-%	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
-	\begin{cases} 
-%		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
-                 \frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
-		1- \pi^{(s+1)} , &   \text{otherwise}. \\
-	\end{cases}
-\end{align}
-
-Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
-\begin{align}
-	\label{eqn:gibbs2}
-%	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) 
-P\left(Z_j^{(s+1)} = q \mid \Gamma_j=h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) 
-	= 
-	\begin{cases} 
-		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
-		0, & \text{otherwise.} \\
-	\end{cases}
-\end{align}
-
-These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
-We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
-
-\begin{lemma}
-	Let $n_A$ and $n_B$ be the number of records in files $A$ and $B$, respectively.
-Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $A \times B$.
-We assume $C$ cores available for parallelization and a Gibbs sampler with $T$ iterations.
-Then, the overall computational complexity of \texttt{fabl} with hashing is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P)$.
-	\label{lemma:fabl}
-\end{lemma}
-\begin{proof}
-	We consider two steps: constructing the comparison vectors and the Gibbs sampler.
-The computational complexity of all pairwise comparisons across $A$ and $B$ is $O(F n_A n_B)$.
-The hashing procedure for all pairwise comparisons is also $O(F n_A n_B)$.
-With $C$ processors available, we can split these computations across $C$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(\frac{F}{C} n_A n_B)$.
-There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
-We note that this contribution to computational complexity applies for all versions of \cite{fellegi_theory_1969} algorithms, unless they use blocking to reduce the comparison space.
-
-	Without hashing, the computational complexity of updating the $\bm{m}$ and $\bm{u}$ parameters is $O(F n_A n_B)$.
-However, by doing calculations over the agreement patterns rather than the individual records, hashing reduces the overall complexity to $O(P)$.
-The complexity of updating $\bm{Z}$ sequentially at the record level as in \cite{sadinle_bayesian_2017} is $O(n_A n_B)$.
-With hashing, we first sample the agreement pattern of the match with complexity $O(n_B P)$, and then we sample the record exhibiting that pattern with complexity $O(n_B)$.
-Thus, the complexity of sampling $\bm{Z}$ in a single iteration is $O(n_B P)$.
-Since $P << n_A$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_A n_B)$ under \texttt{BRL} to $O(n_B P)$ under \texttt{fabl}.
-With parallelization, this complexity is further reduced to $O(\frac{1}{C}n_B P)$, and so the entire Gibbs sampler has complexity $O(\frac{T}{C}n_B P)$
-	In summary, the total computational complexity for \texttt{fabl} is $O(\frac{F}{C} n_A n_B) + O(\frac{T}{C}n_B P).$
-\end{proof}
-
-\hypertarget{scaling}{%
-	\subsection{Scaling to Large Linkage Tasks}\label{scaling}}
-
-		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint \textcolor{green}{chunks} for more manageable computations.
-Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.
-Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
-For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
-For \textcolor{red}{all $a$ and $b$}, we conduct all-to-all comparisons \textcolor{red}{of the pairs of records in each $(A^a, B^b)$} to construct the comparison matrix $\gamma^{ab}$.
-
-		We conduct hashing, obtain the compressed \textcolor{teal}{comparison matrix} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
-		the 
-\textcolor{teal}{memory-intensive}
-		\textcolor{teal}{comparison matrix} $\gamma^{ab}$ 
-		before continuing with the next chunk of data. 
-		In detail, we calculate
-\begin{subequations}
-	\begin{align}
-		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
-		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
-	\end{align}
-\end{subequations}
-These can computed \textcolor{teal}{sequentially} or in parallel.
-Summary statistics from each pairwise chunk comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
-\begin{subequations}
-	\begin{align}
-		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
-		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab}.\label{N_{p_j}}	\end{align}
-\end{subequations}
-
-\hypertarget{SEI}{%
-	\subsection{Storage Efficient Indexing}\label{SEI}}
-
-As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses less memory than storing the full comparison matrix $\gamma$.
-However, the memory requirements of each are still quadratic in nature.
-For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
-We introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
-This allows all-to-all comparisons for substantially larger linkage tasks.
-
-All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$.
-These records have the same probability to be identified as the link for record $B_j$.
-Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate.
-Rather than store all of these record labels, we store only a small number $S$.
-For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $\text{SEI}(r_{p_j}^{ab})$.
-We collect these memory reduced lists to form $\text{SEI}(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $\text{SEI}(\mathcal{R})$.
-
-Let $n_{A^a}$ and $n_{B^b}$ be the number of records in \textcolor{green}{chunks} $A^a$ and $B^b$ respectively.
-Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels.
-As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $\text{SEI}(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
-We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
-
-	\section{Simulation Studies}
-	\label{sec:simulations}
-
-	We demonstrate the speed and accuracy of \texttt{fabl} as compared to \texttt{BRL} through several simulation studies.
-
-	\hypertarget{speed}{%
-		\subsection{Speed}\label{speed}}
-
-	In our first simulation, we generate comparison vectors from pre-specified distributions so that we can easily increase the size of the linkage problem.
-We use $F = 5$ binary comparisons with probabilities for matching and nonmatching pairs shown in Table \ref{Tab:distributions}.
-For each record in $B$, we simulate $n_A$ comparison vectors, resulting in a comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$.
-For $n_B/2$ of these records, there is no match in $A$, so we simulate $n_A$ comparison vectors from the $\bm{u}$ distribution.
-For the other $n_B/2$ of these records, there is one match in $A$, so we simulate $1$ comparison vector from the $\bm{m}$ distribution, and $n_A - 1$ comparison vectors from the $\bm{u}$ distribution.
-We compare the run-time of \texttt{fabl} (with no SEI) against \texttt{BRL} as we increase $n_A$ and $n_B$.
-Since we have five binary comparison fields with no missingness, the number of unique patterns $P$ is bounded above by $2^5 = 32$, a bound which is consistently attained in simulations with more records.
-
-	\begin{table}[t]
-		\centering
-		\begin{tabular}{lll|ll}
-			\multicolumn{1}{c}{ } & \multicolumn{2}{c}{$\bm{m}$} & \multicolumn{2}{c}{$\bm{u}$} \\
-			\cline{2-3} \cline{4-5}
-			& Agree & Disagree & Agree & Disagree \\
-			\hline
-			First Name & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{100}$ &  $\frac{99}{100}$ \\ 
-			Last Name & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{100}$ &  $\frac{99}{100}$ \\ 
-			Day & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{30}$ &  $\frac{29}{30}$ \\ 
-			Month & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{12}$ &  $\frac{11}{12}$ \\ 
-			Year & $\frac{19}{20}$ & $\frac{1}{20}$ & $\frac{1}{12}$ &  $\frac{11}{12}$ \\  
-			\hline
-		\end{tabular}
-		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Section \ref{speed}.}\label{Tab:distributions}
-	\end{table}
-
-	In Figure \ref{fig:speed1}, where we increase both $n_A$ and $n_B$, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but \texttt{fabl} is significantly faster at handling larger sample sizes.
-In particular, run-time for \texttt{BRL} grows quadratically (or linearly with the size of both $A$ and $B$) while run-time for \texttt{fabl} grows linearly (in the size of only $B$).
-
-	\begin{figure}[t]
-		\begin{center}
-%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_speed_plot2}  
-	 \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
-		\end{center}
-	\end{figure}
-
-	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}.
-The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.
-
-	Importantly, \texttt{BRL} implements a Gibbs sampler that is coded C \citep{sadinle_bayesian_2017}, while \texttt{fabl} currently uses non-optimized code written only in \texttt{R}.
-While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.
-Additionally, although \texttt{fabl} is amenable to parallelization, this simulation is run on a single core.
-Implementing \texttt{fabl} in C++ with parallelization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
-
-	\begin{figure}[h!]
-		\begin{center} 
-%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides} 
-		\includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
-We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
-		\end{center}
-	\end{figure}
-
-	\hypertarget{accuracy}{%
-		\subsection{Accuracy}\label{accuracy}}
-
-	Computational speed-ups are only worthwhile if not accompanied by a notable loss of record linkage accuracy.
-Therefore, we examine the accuracy of \texttt{fabl} relative to \texttt{BRL} by replicating a simulation study from \cite{sadinle_bayesian_2017}.
-The simulations employ a collection of synthetic data files with varying amounts of error and overlap (the number of records in common across files).
-Following methods proposed by \cite{christen_pudjijono2009} and \cite{christen_vatsalan2013}, clean records are first simulated from frequency tables for first name, last name, age, and occupation in Australia.
-Fields are then chosen for distortion uniformly at random.
-Names are subject to string insertions, deletions and substitutions, as well as common keyboard, phonetic, and optical recognition errors.
-Age and occupation are distorted through keyboard errors and missingness.
-These synthetic data files are available in the supplement to \cite{sadinle_bayesian_2017}.
-
-	We create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package, shown in Table \ref{Tab:sadinle_simulation_cutoffs}.
-Each simulation identifies matched individuals between two data files, each with 500 records.
-We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across data files.
-Under each of these settings, we use 100 pairs of simulated data files in order to obtain uncertainty quantification on our performance metrics.
-We use uniform priors for all $\bm{m}$ and $\bm{u}$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
-We run the Gibbs sampler for 1000 iterations, and discard the first 100 as burn-in.
-We calculate Bayes estimates $\hat{\bm{Z}}$ of the linkage structure using the  loss function and post-processing procedure described in Appendix \ref{bayes-estimate}.
-Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
-We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; those results are in Appendix \ref{partial}.
-
-	\begin{table}[t]
-		\centering
-		\begin{tabular}[t]{llllll}
-
-			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
-			\cline{3-6}
-			Fields & Similarity & 1 & 2 & 3 & 4\\
-			\hline
-			First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (.5, .1]\\
-			Age and Occupation & Binary & Agree & Disagree &  & \\
-			\hline
-		\end{tabular}
-		\caption{Construction of comparison vectors for accuracy study with simulated data files of Section \ref{accuracy}.}
-		\label{Tab:sadinle_simulation_cutoffs}
-	\end{table}
-
-	We compare \texttt{fabl} to \texttt{BRL} in terms of recall, precision and F-measure, as defined in \cite{christen_2012}.
-Recall is the proportion of true matches found by the model, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(Z_j \leq n_A)$.
-Precision is the proportion of links found by the model that are true matches, that is, $\sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j, Z_j \leq n_A) / \sum_{j=1}^{n_B} I(\hat{Z}_j \leq n_A)$.
-The F-measure balances the two metrics to provide an overall measure of accuracy, and is defined as $2  (\text{Recall } + \text{ Precision}) / (\text{Recall } \times \text{ Precision})$.
-In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.
-
-	\begin{figure}[t]
-		\begin{center}
-%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_sim_plot2} 
-			\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
-			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
-For each level of overlap and each level of error, we have 100 paired sets of 500 records.
-Thus this table summarizes results for 900 data files.
-We see comparable performance for all levels of error and overlap.}
-			\label{fig:sadinle_simulation}
-		\end{center}
-	\end{figure}
-
-	\hypertarget{SEI-sensitivity}{%
-	\subsection{SEI Sensitivity}\label{SEI-sensitivity}}
-
-Finally, our last simulation demonstrates our method's robustness to different values of $S$ for the SEI memory reduction procedure.
-We perform record linkage on one set of synthetic datafile described in Section \ref{accuracy} with 500 records in each datafile, 250 entities in common across datafiles, and 3 errors present across matching records.
-To achieve more drastic results, we perform SEI without chunking the data, that is, $t_A = t_B = 1$.
-In practice, if it is possible to create and store the comparison matrix for all record pairs at one time, there is no need to reduce the memory of the hashed matrix through SEI.
-For illustration however, it is easier to illustrate the effects of choices of $S$ in this setting.
-
-We preform linkage using SEI with $S = (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler.
-As any particular SEI implementation may improve or worsen linkage performance; if the SEI procedure happens to only remove pairs that are not matches, recall and precision will improve.
-Therefore, we perform linkage under each setting 100 times, recording the linkage estimate $\hat{\bm{Z}}$, and recall and precision.
-
-In Figure \ref{fig:SEI_Z}, the largest number of distinct linkage estimates occurs when $S = 1$.
-This makes sense, because the SEI procedure arbitrarily removes large numbers of record labels from consideration, resulting in a noisier estimate of the linkage structure.
-The number of distinct linkage estimates decreases as $S$ increases, with larger values of $S$ providing results more similar to the linkage without SEI.
-In Figure \ref{fig:SEI_eval}, we see similar patterns in precision.
-Setting $S=1$ can arbitrarily remove the index of a true match, leading the Gibbs sampler to concentrate probability on a false match, while larger values of $S$ produce results mirroring implementation with no SEI.
-We note however that even with $S=1$, the loss in precision is small.
-
-Although the figures suggest that $S=2$ is adequate for maintaining linkage performance, we suggest a more conservative value like $S=10$.
-When evaluating the performance of a record linkage algorithm, researchers often examine posterior probabilities.
-By concentrating probability mass on arbitrary nonmatches, low values of $S$ may induce suspiciously high posterior probability for certain record pairs, providing a misleading perception of model performance.
-
-\begin{figure}
-	\begin{minipage}[c]{0.4\linewidth}
-		\begin{center}
-%		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/Z_plot}
-		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
-		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
-		\label{fig:SEI_Z}
-		\end{center}
-	\end{minipage}
-	\hfill
-	\begin{minipage}[c]{0.5\linewidth}
-		\begin{center}
-		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/eval_plot} 
-%	\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
-		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation.}
-		\label{fig:SEI_eval}
-	\end{center}
-	\end{minipage}%
-\end{figure}
-	\section{Case Studies}
-	\label{sec:case-studies}
-
-	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017} and others \citep{lum2013applications, jewell2018accounting}.
-Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors.
-In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study, a larger linkage task that is not feasible in reasonable time under \texttt{BRL} with typical computing setups.
-
-	\subsection{Civilian Casualties from the El Salvadoran Civil War}
-	\label{el_salvador}
-
-	The country of El Salvador was immersed in civil war from 1980 to 1991.
-We are interested in estimating the total number of casualties from the war.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL)\citep{howland2008rescate} and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish)\citep{ball2000salvadoran}.\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization.
-The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
-When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
-Instead, record linkage techniques must be used to merge data files before analyzing the data \citep{lum2013applications}.
-
-	There are several challenges with these data.
-First, both data files have been automatically digitized, which inherently leads to some degree of typographical error.
-Second, the only fields recorded are given name, last name, date of death, and place of death.
-\textcolor{teal}{While it may be common in other data sets for a parent and child to share the same given name, in this data set, it is more common for cousins to share this information, especially in villages, where there may be only four to five extended families.
-In Latin America, an individual receives a surname from their father and their mother, leading to a limited number of surnames in small villages.}
-	%% I believe that Patrick said that this is not common in this data set
-%	It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-
-	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in $n_A = 4420$ records in CHDES and $n_B = 1323$ records in ERTL.
-We standardize names to account for common misspellings and use a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
-Place of death is recorded by municipality and department within that municipality; however, since department is missing in 95\% of records in CHDES and 80\% of records in ERTL, we exclude department from our analysis.
-Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We use uniform priors for the $\bm{m}$ and $\bm{u}$ parameters.
-
-	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in $5^5 \times 3 = 6025$ possible agreement patterns, with 1173 patterns realized in the data.
-However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
-
-	\begin{table}[t]
-		\begin{tabular}[t]{llllll}
-			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
-			\cline{3-6}
-			Fields & Similarity & 1 & 2 & 3 & 4\\
-			\hline
-			First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (.5, 1]\\
-			Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
-			Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
-			Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
-			Municipality & Binary & Agree & Disagree &  & \\
-			\hline
-		\end{tabular}
-		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \cite{sadinle_bayesian_2017}.
-This setup leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
-	\end{table}
-
-	\begin{table}[t]
-	\centering
-	\begin{tabular}[t]{lllll}
-		\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
-		\cline{3-5}
-		Fields & Similarity & 1 & 2 & 3\\
-		\hline
-		First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, 1]\\
-		Year of Death & Binary & Agree & Disagree & \\
-		Month of Death & Binary & Agree & Disagree & \\
-		Day of Death & Absolute Difference & 0 & 1 & 2+\\
-		Municipality & Binary & Agree & Disagree & \\
-		\hline
-	\end{tabular}
-	\caption{Construction of comparison vectors for El Salvador data for increased speed under \texttt{fabl}.
-This setup leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
-\end{table}
-
-	Therefore, we perform our analysis with the agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
-Among the 216 possible agreement patterns, 159 are realized in the data.
-With this revised comparison specification, \texttt{fabl} runs in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
-The estimates of the $\bm{m}$ parameters under each method are similar, as shown in Figure \ref{fig:m-and-u}.
-Estimates of $\bm{u}$ are indistinguishable, and thus omitted.
-Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
-
-	For completeness, we note that linkage with the more detailed comparison vectors requires 240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
-Apparently, the number of patterns is sufficiently many that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
-
-	Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both data files.
-We calculate posterior samples of the size of the overlap across files by finding the number of links in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
-The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate identifies fewer matches than the Gibbs sampler identifies on average.
-This is because a large number of records in ERTL have multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
-We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both data files, and a posterior 95\% credible interval of (211, 244).
-See Figure \ref{fig:overlap-plot} for a visual comparison of the Bayes estimates and posterior credible intervals for the two methods.
-We note that Bayes estimates falling outside of posterior credible intervals has been observed previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
-
-	\begin{figure}[t]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_distribution_smallP_bayes}
-	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
-			\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
-			\label{fig:overlap-plot}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[t]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_posterior_smallP} 
-                  	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
-			\label{fig:m-and-u}
-		\end{center}
-	\end{figure}
-
-	\subsection{National Long Term Care Study}
-	\label{nltcs}
-
-	\begin{table}[t]
-		\centering
-		\begin{tabular}[t]{lllll}
-			\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
-			\cline{3-5}
-			Fields & Similarity & 1 & 2 & 3\\
-			\hline
-			Sex & Binary & Agree & Disagree & \\
-			Year of Birth & Binary & Agree & Disagree & \\
-			Month of Birth & Binary & Agree & Disagree & \\
-			Day of Birth & Binary & Agree & Disagree & \\
-			Location & Custom & Same State and Office & Same State & Otherwise \\
-			\hline
-		\end{tabular}
-		\caption{Construction of comparison vectors for NTLCS data.}\label{Tab:nltcs-comparisons}
-	\end{table}
-
-	The National Long Term Care Study (NLTCS) is a longitudinal study tracking the health outcomes of Medicare recipients \citep{steorts_bayesian_2016}.
-The initial survey began in 1982, with follow-up surveys taken approximately every five years.
-As such, patients are surveyed at most once in a given year, and many patients are surveyed across multiple years.
-In addition, patients can either drop out of the study, pass away, or enter as new patients.
-Hence, the assumptions of our model hold for this study.
-We seek to link records over the $n_A = 20485$ individuals from 1982 to the $n_B = 17466$ individuals from 1989.
-The NLTCS data have longitudinal links, so that in reality one does not need to conduct record linkage.
-However, following the strategy in \cite{guha:reiter:BA}, we break the longitudinal links and treat the data from 1982 and 1989 as stand-alone data files.
-
-	We link records using sex, date of birth, and location using the thresholds shown in Table \ref{Tab:nltcs-comparisons}.
-Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx 400,000,000$ record pairs would require approximately 8GB of memory.
-Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
-{\color{red}{However, through the method described in Section \ref{scaling},  we perform 30 smaller comparison tasks, using $t_A = 1$ and $t_B = 30$.
-We conduct linkage with all record indices recorded and also with SEI using $S=10$, and obtain identical results.
-The hashed $\tilde{\gamma}$ without SEI is about 2.2 GB, and with SEI, it is about 760 MB.
-Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.}}
-
-	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
-Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
-Traceplots do not suggest convergence issues, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
-
-	\begin{figure}[h]
-		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/nltcs/overlap_posterior4}
- 		\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
-			\caption{Posterior distribution and Bayes estimate of overlap across years 1982 and 1989 of NLTCS data.}
-			\label{fig:nltcs-overlap-plot}
-		\end{center}
-	\end{figure}
-
-	\section{Conclusion}
-	\label{discussion}
-
-	In this paper, we have proposed \texttt{fabl}, a Bayesian record linkage method that extends the work of \cite{sadinle_bayesian_2017} to scale to large data sets.
-We have proven that the proposed hashing method and model assumptions allow for a linkage procedure whose computational complexity does not scale with the size of the larger data file.
-This makes \texttt{fabl} computationally advantageous in many linkage scenarios, particularly when one data file is substantially smaller than the other.
-We have also shown that storage efficient indexing, in tandem with hashing, greatly reduces the memory costs required for all-to-all comparisons, giving practitioners an option for larger record linkage tasks potentially even without the use of blocking or indexing.
-We have demonstrated the speed and accuracy of \texttt{fabl} by replicating a simulation study and a case study in \cite{sadinle_bayesian_2017}, and through an additional case study that is computationally impractical under \texttt{BRL}.
-
-	Although the \texttt{fabl} method greatly reduces the memory costs for all-to-all comparisons, computing all $n_A  n_B$ record pairs still can be prohibitive for larger linkage tasks.
-Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
-Due to the quadratic nature of the comparison space, this computation time would grow quickly with the size of the linkage task, and would be infeasibly slow when dealing with millions of records.
-Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task.
-In future work, we seek to extend \texttt{fabl} to account for such deterministic blocking, making the framework amenable to even larger linkage tasks.
-
-	\clearpage
-
-	\bigskip
-
-	\bibliographystyle{jasa}
-	\bibliography{biblio}
-
-	\clearpage
 
@@
 	\hypertarget{app:derivations}{%
-	\section{Derivations of Full Conditionals}\label{app:derivations}}
-\textcolor{red}{Note: Please use s, s+1 consistently throughout.
+	\section{Derivations of Full Conditionals}
+	\label{app:derivations}}
+%\textcolor{red}{Note: Please use s, s+1 consistently throughout.
 Or surpress the notation and tell the reader.
@@
 \end{document}
+
+	\bibliographystyle{jasa}
+	\bibliography{biblio}

### 2023-07-30 ba_resubmit/main.tex
creation of supplement and main to help brian.

--- 2023-07-20 ba_resubmit/kundinger_final.tex
@@
 \endlocaldefs
+
+%%% start code for cross referencing main document with the appendix %%
+
+\usepackage{xr} %%(must be loaded before hyperref) and I found an example of this on Overleaf that was extremely helpful
+
+\makeatletter
+
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+\typeout{(#1)}% latexmk will find this if $recorder=0
+% however, in that case, it will ignore #1 if it is a .aux or 
+% .pdf file etc and it exists! If it doesn't exist, it will appear 
+% in the list of dependents regardless)
+%
+% Write the following if you want it to appear in \listfiles 
+% --- although not really necessary and latexmk doesn't use this
+%
+\@addtofilelist{#1}
+%
+% latexmk will find this message if #1 doesn't exist (yet)
+\IfFileExists{#1}{}{\typeout{No file #1.}}
+}\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+\externaldocument{#1}%
+\addFileDependency{#1.tex}%
+\addFileDependency{#1.aux}%
+}
+%------------End of helper code--------------
+
+% put all the external documents here!
+\myexternaldocument{supplement}
+%% To run this you must run the compiler through pdflatexmk and have the latexmkrc file within the folder
+
+%%% end code for cross referencing main document with the appendix %%
 
@@
 
+\begin{supplement}
+\stitle{Supplementary Material of ``Efficient and Scalable Bipartite Matching with
+Fast Beta Linkage (fabl)." See the .pdf document}
+\sdescription{}
+\end{supplement}
+
 	\clearpage
@@
 	\bibliographystyle{jasa}
-	\bibliography{biblio}
-
-	\clearpage
-
-\appendix
-
-	\hypertarget{app:derivations}{%
-	\section{Derivations of Full Conditionals}\label{app:derivations}}
-\textcolor{red}{Note: Please use s, s+1 consistently throughout.
+	\bibliography{fabl}
+
+%\appendix
+%	
+%		
+%	
+%	\hypertarget{app:derivations}{%
+%	\section{Derivations of Full Conditionals}\label{app:derivations}}
+%\textcolor{red}{Note: Please use s, s+1 consistently throughout.
 Or surpress the notation and tell the reader.
@@
 Perhaps Jerry can state if he would think this is okay as an Editor.}\\
-
-\textcolor{teal}{We provide detailed derivations of the full-conditionals provided in Section~\ref{sec:fast-beta-linkage}.
+%
+%\textcolor{teal}{We provide detailed derivations of the full-conditionals provided in Section~\ref{sec:fast-beta-linkage}.
 In our derivations below, we assume that the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the final update of the full conditional that appears in Section~\ref{sec:fast-beta-linkage}.}
-
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{fl}$, we have
-\begin{align*}
-	\mathcal{L}(m_{fl}|\gamma, \alpha_{fl}, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
-\end{align*}
-where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
+%
+%The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{fl}$, we have
+%\begin{align*}
+%	\mathcal{L}(m_{fl}|\gamma, \alpha_{fl}, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
+%\end{align*}
+%where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
 Analogous procedures lead to the posterior distribution $\mathcal{L}(u_{fl}| \gamma, \beta_{fl}, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
 Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
-
-$$\bm{m}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})),$$
-$$\bm{u}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).$$
-
-Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ and hyperparameters.
+%
+%$$\bm{m}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})),$$
+%$$\bm{u}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).$$
+%
+%Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ and hyperparameters.
 We have the full conditional
-\begin{align*}
-	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-	&\propto \pi^{n_{AB}(\bm{Z})} (1-\pi)^{n_B - n_{AB}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-	&\propto \pi^{n_{AB}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_A - n_{AB}(\bm{Z}) + \beta_{\pi} -1}.
-\end{align*}
-Thus, $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
-
-To express the full conditional for $\bm{Z}$, we consider the cases where $Z_j$ does not have a link in $A$ and where $Z_j$ does have a link in $A$.
+%\begin{align*}
+%	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
+%	&\propto \pi^{n_{AB}(\bm{Z})} (1-\pi)^{n_B - n_{AB}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
+%	&\propto \pi^{n_{AB}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_A - n_{AB}(\bm{Z}) + \beta_{\pi} -1}.
+%\end{align*}
+%Thus, $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
+%
+%To express the full conditional for $\bm{Z}$, we consider the cases where $Z_j$ does not have a link in $A$ and where $Z_j$ does have a link in $A$.
 Because we sample each $Z_j$ independently of all other $Z_{j'}$ (for $j \neq j'$), we need only the full conditional for each individual $Z_j$.
 {\color{red}{Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.}} Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
-\end{align}
-When $Z_j = q$ for some $q\leq n_A$, we have
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
-\end{align}
-We multiply and divide by the $u$ parameters for the matching record pair to obtain
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&= w_{qj}  c_j.
-\end{align}
-We can divide the result of each case by $c_j$ to get
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
-		w_{qj}, & q \leq n_A; \\
-		1, &  q  = n_A + j. \
-	\end{cases}
-\end{align}
-Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
-\begin{align}
-	\label{eqn:z_full_conditional2}
-	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
-	\begin{cases} 
-		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
-		1 - \pi^{(s+1)}, & q  = n_A + j. \
-	\end{cases}
-\end{align}
-
-\textcolor{teal}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
-\begin{align}
-p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi) \propto \prod_q \left(\prod_j
-p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi \right)
-\right).
-\end{align}
-Therefore, we update the full conditional for $\bm{Z}^{(s+1)}$ in parallel.
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+%\end{align}
+%When $Z_j = q$ for some $q\leq n_A$, we have
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+%\end{align}
+%We multiply and divide by the $u$ parameters for the matching record pair to obtain
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+%	&= w_{qj}  c_j.
+%\end{align}
+%We can divide the result of each case by $c_j$ to get
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
+%		w_{qj}, & q \leq n_A; \\
+%		1, &  q  = n_A + j. \
+%	\end{cases}
+%\end{align}
+%Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
+%\begin{align}
+%	\label{eqn:z_full_conditional2}
+%	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+%	\begin{cases} 
+%		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
+%		1 - \pi^{(s+1)}, & q  = n_A + j. \
+%	\end{cases}
+%\end{align}
+%
+%\textcolor{teal}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
+%\begin{align}
+%p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi) \propto \prod_q \left(\prod_j
+%p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi \right)
+%\right).
+%\end{align}
+%Therefore, we update the full conditional for $\bm{Z}^{(s+1)}$ in parallel.
 This contrasts the Gibbs sampler of  \cite{sadinle_bayesian_2017}, where the update to the matching matrix is sequential and accomplished using a partially collapsed Gibbs sampler, where one integrates out $\pi.$ We do not opt for this approach as it would lead to similar sequential updates, which are computationally slow.}
-
-\hypertarget{bayes-estimate}{%
-		\section{Bayes Estimate}
-		\label{bayes-estimate}}
-
-We calculate a Bayes estimate $\hat{\bm{Z}}$ for the linkage parameter $\bm{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
+%
+%
+%
+%\hypertarget{bayes-estimate}{%
+%		\section{Bayes Estimate}
+%		\label{bayes-estimate}}
+%	
+%We calculate a Bayes estimate $\hat{\bm{Z}}$ for the linkage parameter $\bm{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
 We adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which $\hat{Z}_j \in \{1, \ldots, n_A, n_A + j, R\}$, with $R$ representing the option to leave the matching undetermined by the model.
 Specifically, we have
-	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
-		0,  & \text{if } Z_j = \hat{Z_j}; \\
-		\theta_R,  & \text{if } \hat{Z_j} = R; \\
-		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_A + j ; \\
-		\theta_{01},  & \text{if } Z_j = n_A + j,\hat{Z_j} \leq n_A ; \\
-		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
-	\end{cases}\] 
-Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
-
-In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
-	$$\hat{Z}_j =\begin{cases} 
-		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
-		0,  & \text{otherwise}. \\
-	\end{cases}$$
-Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
+%	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
+%		0,  & \text{if } Z_j = \hat{Z_j}; \\
+%		\theta_R,  & \text{if } \hat{Z_j} = R; \\
+%		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_A + j ; \\
+%		\theta_{01},  & \text{if } Z_j = n_A + j,\hat{Z_j} \leq n_A ; \\
+%		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
+%	\end{cases}\] 
+%Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
+%
+%In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
+%	$$\hat{Z}_j =\begin{cases} 
+%		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
+%		0,  & \text{otherwise}. \\
+%	\end{cases}$$
+%Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
 In the event that we have two records $B_j$ and $B_{j'}$ such that both $p(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ p(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
@@
 A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
-
-When we seek a partial estimate of the linkage structure, leaving a portion of record pairs to be classified manually in clerical review, we adopt losses $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, .1)$.
+%
+%When we seek a partial estimate of the linkage structure, leaving a portion of record pairs to be classified manually in clerical review, we adopt losses $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, .1)$.
 For a more in-depth explanation of this function and the induced Bayes estimate, see \cite{sadinle_bayesian_2017}.
-
-	\hypertarget{partial}{%
-	\section{Accuracy under Partial Estimates}\label{partial}}
-
-In this section, we repeat the simulation study in Section \ref{accuracy} of the main text, allowing for clerical review rather than forcing all records to have or not have links.
+%
+%
+%	\hypertarget{partial}{%
+%	\section{Accuracy under Partial Estimates}\label{partial}}
+%
+%In this section, we repeat the simulation study in Section \ref{accuracy} of the main text, allowing for clerical review rather than forcing all records to have or not have links.
 Specifically, by leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
@@
 To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.
-
-In Figure \ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
+%
+%In Figure \ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
 However, with high amounts of error, and thus fewer accurate and discerning fields of information, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
@@
 If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand than would be left under \texttt{BRL}, but the decisions made by each method should have nearly equal accuracy.
-
-%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
+%
+%%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
 If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
-
-\begin{figure}[t]
-	\begin{center}
-%	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
-		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
+%
+%%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
+%
+%\begin{figure}[t]
+%	\begin{center}
+%%	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
+%		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
 We see poorer performance for \texttt{fabl} only in situations with high overlap.}
-		\label{fig:sadinle_simulation_partial}
-	\end{center}
-\end{figure}
-
-	\hypertarget{appendix-sim}{%
-		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
-	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+%		\label{fig:sadinle_simulation_partial}
+%	\end{center}
+%\end{figure}
+%	
+%	
+%	\hypertarget{appendix-sim}{%
+%		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
+%	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
 It is set up with one error across the linkage fields and 50 duplicates across files.
@@
 Note that traceplots for $\bm{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
-
-	\begin{figure}[!h]
-		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
-		\end{center}
-	\end{figure}
-
-\clearpage
-
-	\hypertarget{appendix-speed}{%
-	\section{Additional Speed Simulation Study}\label{app:appendix-speed}}
-	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
+%			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
+%		\end{center}
+%	\end{figure}
+%	
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%%		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
+%			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
+%		\end{center}
+%	\end{figure}
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
+%			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
+%		\end{center}
+%	\end{figure}
+%
+%\clearpage
+%
+%	\hypertarget{appendix-speed}{%
+%	\section{Additional Speed Simulation Study}\label{app:appendix-speed}}
+%	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
 We replicate the speed study of Section \ref{speed} under different settings.
@@
 The $\bm{m}$ and $\bm{u}$ parameters for this simulation are shown Table \ref{Tab:distributions_2}.
-
-		\begin{table}[t]
-		\centering
-		\begin{tabular}{llll|lll}
-			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{$\bm{m}$} & \multicolumn{3}{c}{$\bm{u}$} \\
-			\cline{2-4} \cline{5-7}
-			& Agree & Partial & Disagree & Agree & Partial & Disagree \\
-			\hline
-			Feature 1 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-			Feature 2 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-			Feature 3 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-			Feature 4 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-			\hline
-		\end{tabular}
-		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
-	\end{table}
-
-	\begin{figure}[h!]
-		\begin{center} 
-%		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
+%	
+%		\begin{table}[t]
+%		\centering
+%		\begin{tabular}{llll|lll}
+%			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{$\bm{m}$} & \multicolumn{3}{c}{$\bm{u}$} \\
+%			\cline{2-4} \cline{5-7}
+%			& Agree & Partial & Disagree & Agree & Partial & Disagree \\
+%			\hline
+%			Feature 1 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+%			Feature 2 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+%			Feature 3 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+%			Feature 4 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
+%			\hline
+%		\end{tabular}
+%		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
+%	\end{table}
+%	
+%	\begin{figure}[h!]
+%		\begin{center} 
+%%		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
+%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
 We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\begin{center}
-%		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
-		 \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
+%		\end{center}
+%	\end{figure}
+%	
+%	\begin{figure}[h!]
+%		\begin{center}
+%%		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
+%		 \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
+%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
 We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
-		\end{center}
-	\end{figure}
-
-\clearpage
-
-	\hypertarget{appendix-es}{%
-		\section{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} 
-			\label{fig:m_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} 
-			\label{fig:u_trace}
-		\end{center}
-	\end{figure}
+%		\end{center}
+%	\end{figure}
+%
+%\clearpage
+%	
+%	\hypertarget{appendix-es}{%
+%		\section{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
+%			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
+%		\end{center}
+%	\end{figure}
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
+%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
+%			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} 
+%			\label{fig:m_trace}
+%		\end{center}
+%	\end{figure}
+%	
+%	\begin{figure}[!h]
+%		\begin{center}
+%%		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
+%		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
+%			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} 
+%			\label{fig:u_trace}
+%		\end{center}
+%	\end{figure}
 \end{document}

### 2023-07-30 ba_resubmit/response_to_reviewers3.tex
edits

@@
 
-\textcolor{teal}{We thank the Editor and reviewers for their assessment and constructive feedback.}
+We thank the Editor and Associate Editor for their assessment and constructive feedback.
 
@@
 
-\reply{We have made several revisions to make the paper easier to read, as described in our responses below. 
+\reply{We have made several revisions to the paper, as described in our responses below. 
 }
@@
 \reply{
-To lighten the notation and allow for greater precision, we now refer to the two data files as $A$ and $B$. The updated material now reads:
+We now refer to the two data files as $A$ and $B$. The updated material now reads:
 
@@
 \setcounter{responsectr}{0}
-\section{Reviewer Comments}
+\section{Associate Editor Comments}
 
@@
 
-	\reply{Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
+	\reply{
+		%Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
 For simplicity, we have chosen to use MCAR in the article.
@@
 In the second paragraph of Section 4.3, the H notation includes the matches (Z), in notation and definition that seems to conflict with the statement just below Equation (10).
-This also comes into play in Equation (16)}
+This also comes into play in Equation (16).}
 
@@
 	\reply{Yes, the comparison vectors are created (and therefore, patterns are assigned) without regard to $\bm{Z}$.
-We have removed that sentence, and updated the equations in the Section 4.3 to avoid confusion}
+We have removed that sentence, and updated the equations in the Section 4.3 to avoid confusion.}
 
@@
 
-	\reply{We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
+	\reply{We have revised the text to state ``summary statistics" instead of ``sufficient statistics."
+
+		%We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$.
 However, we cannot express the full likelihood in (14) through these statistics.
@@
 
-	In our revised submission, we have reexpressed the likelihood from (6a) through the perspective of agreement patterns in (14).
+	In our revised submission, we have re-expressed the likelihood from (6a) through the perspective of agreement patterns in (14).
 This form of the likelihood is then used to express contributions to the likelihood for the $\bm{m}$ and $\bm{u}$ parameters in (15) and for $\bm{Z}$ in (17), both of which use the summary statistics.
@@
 
-	The revised text states: ``Then, we conduct hashing, obtain the compressed comparison matrix, $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison matrix, $\gamma^{ab}$, before continuing with the next chunk of data.
-See Section 4.3}
+	The revised text states: ``Then, we conduct hashing, obtain the compressed comparison matrix, $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison matrix, $\gamma^{ab}$, before continuing with the next batch of data.
+See Section 4.3.}
 
@@
 \reply{To make the SEI algorithm more clear, we have revised the ordering of Sections 4, so that we have 1) hashing, 2) posterior inference, 3) chunkwise computation of the comparison matrix, and 4) SEI.
-With this new ordering, and with the newly provided equations (15) through (19), we hope it clearer that that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
+With this new ordering, and with the newly provided equations (15) through (19), we hope it is clearer that that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
 In particular, (16a), (16b), and (18) depend only $\mathcal{N}$.
@@
 	We use $\bm{Z}$ when discussing a random quantity.
-In all instances where we discuss $Z_j$ taking on a particular value, we now use $q$ to avoid confusion and to follow notation used by \cite{sadinle_bayesian_2017}.
+In all instances where we discuss the random variable $Z_j$ taking on a particular value, we now use $q$ to avoid confusion and to follow notation used by \cite{sadinle_bayesian_2017}.
 }
@@
 \reply{%
-	$\gamma$ is a comparison matrix comprised by comparison vectors.
+	The comparison matrix $\gamma$ is comprised of the comparison vectors for all record pairs in $A \times B$.
 All references to $\gamma$ as a set have been removed.
@@
 \reply{%
-	This has been revised; thank you. See Section 4.2
+	This has been revised; thank you. See Section 4.2,
 }

### 2023-07-30 ba_resubmit/kundinger_resubmit3.tex
edits

@@
 %
-\pubyear{0000}
-\volume{00}
-\issue{0}
-\doi{0000}
+\pubyear{2021}
+\volume{TBA}
+\issue{TBA}
+% \doi{0000}
 %\arxiv{}
@@
 \title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
-%\title{A Sample Document\thanksref{T1}}
-%\thankstext{T1}{<thanks text>}
 \runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
@@
 \kwd{data fusion}
-\kwd{distributed computing}
+\kwd{data cleaning}
 \kwd{entity resolution}
@@
 Before conducting data analysis, it is often necessary to identify duplicate records across two data files.
-This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
+This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep[e.g.,][]{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
 In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.
@@
 \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
-We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs $\gamma_{ij}$.
-Additionally, we let $\Gamma_{ij}$ represent a random variable for the comparison vector for $A_i$ and $B_j$, such that $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
+We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$ as the collection of all record pairs $\gamma_{ij}$.
+%Additionally, we let $\Gamma_{ij}$ represent a random variable for the comparison vector for $A_i$ and $B_j$, such that $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
 
@@
 
-%Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.  Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
-
-For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
+Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.
+Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
+For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are independent given the matching status of the record pair, and second, that the matching status of each record pair is independent of the matching status of other pairs.
 Using these independence assumptions, one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.
@@
 When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
-Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
-The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
+Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\Gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\Gamma_{ij}^f = l|\Delta_{ij} = 0)$.
 
@@
 Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
-However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta linkage (\texttt{fabl}) prior below.
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast beta linkage (\texttt{fabl}) prior below.
 For each $Z_j$, we have
@@
 \end{cases} \label{eqn:fast_beta_prior}  \\
-\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \nonumber
+\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}). \nonumber
 %\label{eqn:fast_beta_prior2}.
@@
 This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
-In our preliminary studies, this highly informative prior weighting on matching can result in overly high match rates.
+In our preliminary studies, we have found that this highly informative prior weighting on matching can result in overly high match rates.
 Hence, we prefer (\ref{eqn:fast_beta_prior}).
-We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
+We also note that the flat prior is equivalent to a special case of the fast beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
 
@@
 That is, $\pi$  in (\ref{eqn:fast_beta_prior}) under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in (\ref{eqn:lambda_dist}) in the \cite{fellegi_theory_1969} model estimates the proportion of record pairs that are matches.
-In the bipartite case, we conjecture that some analysts will find $\pi$ to be more a interpretable parameter than $\lambda$.
+In the bipartite case, we conjecture that some analysts will find $\pi$ to be a more interpretable parameter than $\lambda$.
 In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $1/n_A$ and tends towards 0 as the size of the linkage task grows.
@@
 %
-For clarity, we present our full model for \texttt{fabl} below:
+For clarity, we present the full model for \texttt{fabl} below:
 \begin{subequations}
@@
 Denote the current draw of the sampler by $(s)$ and the updated draw by $(s+1)$.
-To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, sample
+To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, we sample
 \begin{subequations}
@@
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields.
-However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
+However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical \cite{fellegi_theory_1969} model, leaving the effect of indexing on Bayesian record linkage models to future work.
 
 With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
-First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors.
+First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of storing the comparison vectors.
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
@@
 We then map the integers in \eqref{eqn:hashing} to sequential integers $\{1, \ldots, P\}$.
-Denote each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+Denote each unique agreement pattern as $h_p = (h_p^1, \ldots, h_p^F)$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
 When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
@@
 	\begin{align}
-		m_p =  p(\Gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(h_p^f = l)I_{obs}(\Gamma_{ij}^f)}\label{eqn:m_p} \\
-		u_p =  p(\Gamma_{ij} = h_p|Z_j \neq i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(h_p^f = l)I_{obs}(\Gamma_{ij}^f)}\label{eqn:u_p}
+		m_p =  p(\Gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(h_p^f = l)I_{obs}(h_p^f)}, \label{eqn:m_p} \\
+		u_p =  p(\Gamma_{ij} = h_p|Z_j \neq i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(h_p^f = l)I_{obs}(h_p^f)}\label{eqn:u_p}
 	\end{align}
@@
 \end{align}
-Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be a concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
+Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
 The terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
@@
 
-We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
-
-Similarly, the posterior update for $\bm{Z}$ depend on the the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
+%We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
+
+Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
 For each pattern $p$, let $w_p =m_p/u_p$.
@@
 
-Let $n_a$ and $n_b$ be the number of records in batches $A^a$ and $B^b$ respectively.
-Instead of storing $n_A n_B$ record labels, with SEI we store at most $t_A \sum_{b = 1}^{t_B} n_b  P  S$ labels.
+%Let $n_a$ and $n_b$ be the number of records in batches $A^a$ and $B^b$ respectively. Instead of storing $n_A n_B$ record labels, with SEI we store at most $t_A \sum_{b = 1}^{t_B} n_b  P  S$ labels.
+
 As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
@@
 			\includegraphics[width=0.99\textwidth]{../notes/figures/sadinle_sim_plot3} 
-			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
+			\caption{Posterior means and credible intervals for accuracy metrics under the replication of the simulation study from \cite{sadinle_bayesian_2017}.
 For each level of overlap and each level of error, we have 100 paired sets of 500 records.
@@
 Finally, our last simulation demonstrates the robustness of \texttt{fabl} to different values of $S$ for the SEI memory reduction procedure.
-We perform record linkage on one set of synthetic data file described in Section \ref{accuracy} with 500 records in each data file, 250 entities in common across data files, and 3 errors present across matching records.
-To achieve more drastic results, we perform SEI without batching the data, that is, $t_A = t_B = 1$.
+We perform record linkage on one set of the synthetic data files described in Section \ref{accuracy} with 500 records in each data file, 250 entities in common across data files, and 3 errors present across matching records.
+We perform SEI without batching the data, that is, $t_A = t_B = 1$.
 In practice, when it is computationally feasible to create and store $\gamma$ without batching, there is no need to reduce the memory of the hashed matrix through SEI.
@@
 Setting $S=1$ can arbitrarily remove the index of a true match, leading the Gibbs sampler to concentrate probability on a false match, while larger values of $S$ produce results mirroring implementation with no SEI.
-We note, however, that even with $S=1$, the loss in precision is small.
+We note, however, that even with $S=1$, the loss in precision is small in these simulations.
 
@@
 		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
-		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
+		\caption{Distinct values of $\hat{Z}$ in the simulations of Section \ref{SEI-sensitivity}.}
 		\label{fig:SEI_Z}
@@
 		\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
-		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation.}
+		\caption{Means and 95\% credible intervals for recall and precision in the simulations of Section \ref{SEI-sensitivity}.}
 		\label{fig:SEI_eval}
@@
 		\includegraphics[width=0.99\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
-		\caption{Posterior distribution and Bayes estimate of overlap across the two files in the El Salvador case study.. We note they are quite similar under both methods.}
+		\caption{Posterior distribution and Bayes estimate of overlap across the two files in the El Salvador case study. We note they are quite similar under both methods.}
 		\label{fig:overlap-plot}
@@
 	We link records using sex, location, and day, month, and year of birth using the criteria shown in Table \ref{Tab:nltcs-comparisons}.
-Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx 400,000,000$ record pairs would require approximately 8GB of memory.
+Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx$ 400 million record pairs would require approximately 8GB of memory.
 Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
@@
 	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+Traceplots do not suggest convergence issues, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}.
+As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure has of 9634 matches, with a 95\% credible interval of (9581, 9740).
 Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
-Traceplots do not suggest convergence issues, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
 
@@
 
-	We provide detailed derivations of the full-conditionals provided in Section \ref{sec:fast-beta-linkage}.
+	We provide detailed derivations of the full-conditionals provided in Section 3 of the main text.
 The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
@@
 	where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
-Analogous procedures lead to the posterior distribution $\mathcal{L}(u_{fl}| \gamma, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
+Analogous procedures lead to $\mathcal{L}(u_{fl}| \gamma, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
 Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
 	\begin{align}
-	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \\
-	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).
+	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s) &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \\
+	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s) &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).
 	\end{align}
@@
 	\end{align}
-	Thus, $\pi^{(s+1)}|\gamma, \bm{Z}^{(s+1)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
+	Thus, $\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
 
 	Due to the independence in the fast beta prior in \eqref{eqn:fast_beta_prior}, we can obtain the full conditional for $\bm{Z}$ through the full conditionals for each individual $Z_j$.
-Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.
+Let $\Gamma_{.j}$ denote the random matrix of $n_A$ comparison vectors relating to an arbitrary record $B_j$, and let $\gamma_{.j}$ be a realization of $\Gamma_{.j}$.
 We have
@@
 
-In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
+In general, we follow \cite{sadinle_bayesian_2017} and set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
 	\begin{align}
@@
 		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
-	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section 5.2 of the main text.
 It is set up with one error across the linkage fields and 50 duplicates across files.
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
+			\caption{Representative traceplot of overlap between files from simulation study in Section 5.2 of the main text.}\label{fig:sim_overlap_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
+			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section 5.2 of the main text.}\label{fig:sim_m_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
+			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section 5.2 of the main text.}\label{fig:sim_u_trace}
 		\end{center}
@@
 
-In this section, we repeat the simulation study in Section \ref{accuracy} of the main text, allowing for clerical review rather than forcing all records to have or not have links.
+In this section, we repeat the simulation study in Section 5.2 of the of the main text, allowing for clerical review rather than forcing all records to have or not have links.
 Specifically, by leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
@@
 	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
-We replicate the speed study of Section \ref{speed} under different settings.
+We replicate the speed study of Section 5.1 of the main text under different settings.
 Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.
-The $\bm{m}$ and $\bm{u}$ parameters for this simulation are shown Table \ref{Tab:distributions_2}.
+The $\bm{m}$ and $\bm{u}$ parameters for this simulation are shown in Table \ref{Tab:distributions_2}.
 
@@
 	\end{subfigure}
-	\caption{Traceplots for parameters of interest in El Salvador case study in Section \ref{el_salvador}}.
+	\caption{Traceplots for parameters of interest in El Salvador case study in Section 5.1 of the main text.}
 	\label{fig:mean and std of nets}

### 2023-07-30 presentations/cpm_proposal.tex

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	filecolor=magenta,      
	urlcolor=cyan,
	pdftitle={Overleaf Example},
	pdfpagemode=FullScreen,
}

\begin{document}
	\title{Coarse Probabilistic Matching for Causal Inference through Bayesian Fellegi Sunter}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger}
	\institute{Duke University}
	\date{\today}

%	\AtBeginSection[]{
%		\begin{frame}
%			\vfill
%			\centering
%			\begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%				\usebeamerfont{title}\insertsectionhead\par%
%			\end{beamercolorbox}
%			\vfill
%		\end{frame}
%	}

\AtBeginSection[]
{
	\begin{frame}
		\frametitle{Table of Contents}
		\tableofcontents[currentsection]
	\end{frame}
}

	\begin{frame}
		\titlepage
	\end{frame}

\section{Stuff I've Been Thinking About}

\begin{frame}{Fellegi Sunter}
	\begin{itemize}
		\item Let datasets $A$ and $B$ have records indexed $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$
		\item For each record pair, compute comparison vector $\gamma_{ij}$, where
		each component $\gamma_{ij}^f$ takes on value $\ell \in \{1, \ldots, L_f\}$ according to predefined distance function
		\begin{itemize}
			\item Categorical: $\gamma_{ij}^f  = 1$ if $(i, j)$ agree on $f$, $\gamma_{ij}^f  = 2$ otherwise
			\item Names: 1 for exact agreement, 2 for close agreement, 3 for no agreement
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Fellegi Sunter}
	\begin{itemize}
		\item Every time I've presented this model, I'm embarassed by this coarsening of data.
		\item Still, I think it might be interesting to see if it could be useful for matching in causal inference
	\end{itemize}
\end{frame}

\begin{frame}{Mahalonobis Distance and Propensity Matching}
	\begin{itemize}
		\item One approach to matching is to use the Mahalonobis distance, which measures distance between units purely in the covariate space, without regard to which features are related to the outcome of interest. It is defined by
		\begin{align*}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1} |X_1 - X_2|}
		\end{align*}
		\item Another method is to estimate propensity scores $p(Z_n = 1 |X_n)$ for each unit in the data, and match units based on this score.
		\item Most matching methods in causal inference then use deterministic algorithms to match treated units to their nearest control unit based on the chosen metric
		\item David King publishes a strong critique of PS matching In any given matching, no guarantee of reducing covariate imbalance
	\end{itemize}
\end{frame}

\begin{frame}{Coarse Exact Matching}
	\begin{itemize}
		\item Coarsen each variable to integer values based on reasonable knowledge (or using automated recommendations)
		\item Each data unit has effectively been transformed to a set of integer values. 
		\item Call treated and control units a match if they match at the coarsened level for \emph{all} variables
		\item This coarsening allows CEM to be Monotonic Imbalance Bounding (MIB). Shown to have have better covariate balance, at expense of sample size
		<2>\item THOUGHT: If coarsening was good in CEM, perhaps it can be good in Fellegi Sunter
	\end{itemize}
\end{frame}

\begin{frame}{Review of Fellegi Sunter}
	Let $\Delta_{ij} \sim \text{Bernoulli}(\lambda)$ be the indicator that  $(A_i, B_j)$ is a match. Let $m_{fl} = p(\gamma_{ij}^f = l | \Delta_{ij} = 1)$, and $u_{fl} = p(\gamma_{ij}^f = l | \Delta_{ij} = 0)$. \linebreak

	The Fellegi Sunter Model is:
	\begin{align*}
		p(\Gamma| m, u, \Delta) = \prod_{i}^{n_A} \prod_j^{n_B} \lambda\left[\prod_{f=1}^F \prod_{l = 1}^{L_f}m_{fl}^{\gamma_{ij}^f = l}\right]  + (1 - \lambda) \left[\prod_{f=1}^F \prod_{l = 1}^{L_f}u_{fl}^{\gamma_{ij}^f = l}\right] 
	\end{align*}
\end{frame}

\section{Fellegi Sunter as Monotone Imbalance Bounding}

%\begin{frame}{EPBR and MIB}
%	A matching method is Equal Percent Bias Reducing (EPBR) if
%	$$E(\bar{X}_{m_T} - \bar{X}_{m_C}) = \gamma(\mu_T - \mu_C)$$
%	Most texts restrict $0 < \gamma < 1$, so I do not think Fellegi-Sunter is EPBR. \linebreak
%	
%	Part of the definition for Monotone Imbalance Bounding (MIB) is
%	$$|\bar{X}_{m_T, f} - \bar{X}_{m_C, f}| \leq \gamma_f|\bar{X}_{T, f} - \bar{X}_{C, f}|$$
%	for $0 \leq \gamma_f \leq 1$ for each feature $f$ separately. I believe that Fellegi-Sunter matching is MIB. 
%\end{frame}

\begin{frame}{MIB for Binary Covariates}
	As defined by \href{https://gking.harvard.edu/files/gking/files/cem_jasa.pdf}{Iacus, King, Porro (2012)}, a matching method is \emph{Monotonic Imbalance Bounding} on a function $f$ with respect to a distance $D(\cdot, \cdot)$, if for a monotonically increasing function $\gamma_{f, D}(\cdot)$ and any $\pi \in \mathbb{R}_{+}^K$, we have
	\begin{align*}
		D(f(\mathcal{X}_{m_T(\pi)}), f(\mathcal{X}_{m_C(\pi)})) \leq \gamma_{f, D}(\pi)
	\end{align*}

Here, $\pi$ is a vector of scalar parameters relating the coarsening of each feature. Also ${m_T(\pi)}$ and ${m_C(\pi)}$ are the sizes of the matched sets, importantly determined by $\pi$, and not determined by the modeller. Lastly, $\mathcal{X}_{m_T}$ and $\mathcal{X}_{m_C}$ are the matched subsets of $T$ and $C$.
\end{frame}

\begin{frame}{MIB for Binary Covariates}
	I've been struggling with the exact best way to interpret this definition, so I've editted my proof a couple times, and I don't have a finished version. But for now, I can show you that they have this flavor: \linebreak

	Let $X_{m_T}$ and $X_{m_C}$ be vectors of binary covariates for units matched through Fellegi-Sunter (and with matchings encoded in $\Delta$)

	\begin{align*}
		|\bar{X}_{m_T} - \bar{X}_{m_C}| &= E[|\bar{X}_{m_T} - \bar{X}_{m_C}|| \Delta] \\
		&\leq E[|X_i - X_j| | \Delta_{ij} = 1] \\
		&= 0 \dot m_{1} + 1 \dot m_{2} \\
		&=m_{2}
	\end{align*}
\end{frame}

\begin{frame}{MIB for Continuous Covariates}
	I can prove this for continuous covariates as well. Suppose $X$ has range $R = X_{\max} - X_{\min}$, and bin the data according to $\epsilon = \frac{R}{K}$, where $\pi_j = K$ is our coarsening parameter. Then the upper bound becomes $\epsilon \sum_{k=1}^K k m_k$. Note that this is increasing in $K$. \linebreak

	Note that we can directly compute this bound, and check that it holds on actual matched samples.
\end{frame}

%\begin{frame}
%	We aim to show Fellegi Sunter is $MIB(f, D)$ for $f$, the sample mean, and $D$, absolute distance. Supposed $X$ has range $R = X_{\max} - X_{\min}$, and bin the data according to $\epsilon = \frac{R}{K}$, where $\pi_j = K$ is our coarsening parameter. 
%\end{frame}

%\begin{frame}{Expected Imbalance Bounding}
%	Tentatively, lets called a matching method \emph{Expected Imbalance Bounding} (EIB) if there exist scalars $0 \leq \gamma_f \leq 1$ for each feature such that
%	$$E(\bar{X}_{m_T, f} - \bar{X}_{m_C, f}) = \gamma_f(\mu_{T, f} - \mu_{C, f})$$
%\end{frame}

%\begin{frame}{FS as Monotone Imbalance Bounding}
%	For clarity, we're dropping the $f$ subscripts. Let $X_T$ and $X_C$ be vectors of binary covariates for treated and control groups. Let $X_{m_T}$ and $X_{m_C}$ be similar vectors for units matched through Fellegi-Sunter (and with matchings encoded in $\Delta$)\linebreak
%	\begin{align*}
%		|\bar{X}_{m_T} - \bar{X}_{m_C}| &= E[|\bar{X}_{m_T} - \bar{X}_{m_C}|| \Delta] \\
%		&\leq E[|X_i - X_j| | \Delta_{ij} = 1] \\
%		&= 0 \dot m_{1} + 1 \dot m_{2} \\
%		&=m_{2}
%	\end{align*}
%\end{frame}
%(The second line is perhaps not an equality. There may be some assumption I can make to smooth it over.)

%
%	For clarity, we're dropping the $f$ subscripts. Let $X_T$ and $X_C$ be vectors of binary covariates of lengths $n_T$ and $n_C$ for treated and control groups. Let $X_{m_T}$ and $X_{m_C}$ be similar vectors units of lengths $n_{m_T}$, and $n_{m_C}$ matched through Fellegi-Sunter (and with matchings encoded in $\Delta$)\linebreak
%\begin{align*}
%	|\bar{X}_{m_T} - \bar{X}_{m_C}| &= E[|\bar{X}_{m_T} - \bar{X}_{m_C}|| \Delta] \\
%	&= \frac{1}{n_{m_T}} E\left[\sum_{i, j= 1}^{n_{m_T}}  |X_i - X_j|| \Delta = 1\right] \\
%	&= E[|X_i - X_j| | \Delta_{ij} = 1] \\
%	&= 0 \dot m_{1} + 1 \dot m_{2} \\
%	&=m_{2}
%\end{align*}

%\begin{frame}{MIB for Binary Covariates}
%	Now the RHS
%	\begin{align*}
%		|\bar{X}_{T} - \bar{X}_{C}| &= E[|\bar{X}_{T} - \bar{X}_{C}||\Delta] \\
%		&=E\left[ E[|X_i - X_j || \Delta_{ij}] \right] \\
%		&= \lambda E[|X_i - X_j| | \Delta_{ij} = 1] + (1 - \lambda) E[|X_i - X_j || \Delta_{ij} = 0] \\
%		&= \lambda m_{2} + (1 - \lambda)(0 \dot u_{1} + 1 \dot u_{2}) \\
%		&= \lambda m_{2} + (1 - \lambda)u_{2}
%	\end{align*}
%
%(Again, the second line is not quite right. But I can maybe argue that there is a constant there that I can end up just dividing out.)
%\end{frame}
%
%\begin{frame}{MIB for Binary Covariates}
%	\begin{align*}
%		|\bar{X}_{m_T} - \bar{X}_{m_C}| - |\bar{X}_{T, f} - \bar{X}_{C, f}| &=
%		m_{2} -\lambda m_2 - (1 - \lambda)u_{2} \\
%		&= (1 - \lambda)m_{2} - (1 - \lambda)u_{2}  \\
%		&= (1 - \lambda)(m_2 - u_{2}) < 0 
%	\end{align*}
%
%Under certain assumptions, trust me that $m_{2} < u_{2}$. \linebreak
%
%Thus $E(\bar{X}_{m_T} - \bar{X}_{m_C}) \leq \gamma	|\bar{X}_{T, f} - \bar{X}_{C, f}|$ for $\gamma = 1$, and Fellegi Sunter is MIB for binary variables.
%\end{frame}

%\begin{frame}{EIB for Binary Covariates}
%	For clarity, we're dropping the $f$ subscripts. Let $X_T$ and $X_C$ be vectors of binary covariates for treated and control groups, and let $X_{m_T}$ and $X_{m_C}$ be such vectors among matched units. \linebreak
%	
%	We start with the LHS:
%	\begin{align*}
%		E(\bar{X}_{m_T} - \bar{X}_{m_C}) &= E(X_i - X_j| \Delta = 1] \\
%		&= 0 \dot m_{1} + 1 \dot m_{2} = m_{2}
%	\end{align*}
%\end{frame}
%
%\begin{frame}{EIB for Binary Covariates}
%	Now the RHS
%	\begin{align*}
%		\mu_T - \mu_C &= E[|X_i - X_j] \\
%		&=E\left[ E[X_i - X_j | \Delta] \right] \\
%		&= \lambda E[X_i - X_j | \Delta = 1] + (1 - \lambda) E[X_i - X_j | \Delta = 0] \\
%		&= \lambda m_{2} + (1 - \lambda)(0 \dot u_{1} + 1 \dot u_{2}) \\
%		&= \lambda m_{2} + (1 - \lambda)u_{2}
%	\end{align*}
%\end{frame}
%
%\begin{frame}{EIB for Binary Covariates}
%\begin{align*}
%	(\mu_T - \mu_C) - E(\bar{X}_{m_T} - \bar{X}_{m_C}) &= \lambda m_{2} + (1 - \lambda)u_{2} - m_{2} \\
%	&= (1 - \lambda)u_{2} + (1 - \lambda)m_{2} \\
%	&= (1 - \lambda)(u_{2} - m_{2}) > 0 
%\end{align*}
%
%Under certain assumptions, trust me that $u_{2} - m_{2} > 0$ holds. \linebreak
%
%Thus $E(\bar{X}_{m_T} - \bar{X}_{m_C}) \leq \gamma(\mu_T - \mu_C)$ for $\gamma = 1$, and Fellegi Sunter is EID for binary variables.
%\end{frame}

%\begin{frame}{MIB for Continuous Covariates}
%	I can prove this for continuous covariates as well, but it requires many more steps and (minor) assumptions \linebreak
%	
%	There are also more parts of the definition of MIB that I don't fully understand yet, but it really feels like Fellegi-Sunter satisfies all of them.
%\end{frame}

\section{Fellegi Sunter as Course Probabilistic Matching}

\begin{frame}{Course Probabilistic Matching}
	In this context, I'd like to propose Course Probabilistic Matching (CPM). Instead coarsening the\emph{distance} between covariates, we will calculate the distance between coarsened covariates. For example, we would have $\gamma_{ij}^f = |C(X_{i, f}) - C(X_{j, f})|$, where $C(\cdot)$ is a coarsening function mapping bins in the covariate space to integer values. \linebreak

	We calculate probabilities exactly as we would under FS:
	\begin{align*}
		P(\Delta_{ij} = 1 | \gamma_{ij}) &= p(\Delta_{ij} = 1)\prod_f p(\gamma_{ij}^f|\Delta_{ij}= 1) \\
		P(\Delta_{ij} = 0 | \gamma_{ij}) &= p(\Delta_{ij} = 0)\prod_f p(\gamma_{ij}^f|\Delta_{ij}= 0)
	\end{align*}
These are just products of relevant $m$ and $u$ probabilities.
\end{frame}

\begin{frame}{CEM}
	Then, CEM can be viewed as a special case of CPM, where comparison vectors are constructed through $\gamma_{ij}^f = I(C(X_{i, f}) = C(X_{j, f}))$, and probabilities are assigned deterministically through \linebreak
\begin{align*}
	P(\Delta_{ij} = 1 | C(X_i) = C(X_j)) &= P(\Delta_{ij} = 1 |\gamma_{ij}^f = 1 , \forall f) = 1 \\
	P(\Delta_{ij} = 1 | C(X_i) \neq C(X_j)) &= 0 \\
\end{align*}
Additionally, if the $m$ parameter gives all its probability mass to exact agreement, then the MIB bound for FS,  $\epsilon \sum_{k=1}^K k m_k$,  reduces down to just $\epsilon$, as provided in the paper.
\end{frame}

\section{Conclusion}
\begin{frame}{Concluding Thoughts}
	\begin{itemize}
		\item Last week, I thought that FS would perform about the same as Mahalanobis distance, and that we would need to modify it a bunch to make it useful for causal inference
		\item Now I suspect the Coarse Probabilistic Matching would outperform Course Exact Matching, right out of the box. It would find all of the coarsened "exact" matches, plus additional near matches, resulting in mild increases in covariate imbalance with substantial increases in sample size after matching
		\item We can measure match quality through posterior probabilities, and discard matches we think are too weak
		\item Since this matching method is fully Bayesian and probabilistic, we can create joint models for the matching and treatment effect estimation, propagating uncertainty between the steps, and averaging over many plausible matchings.
	\end{itemize}
\end{frame}

\begin{frame}{Tons of Ways Forward}
	\begin{itemize}
		\item Establish theory for CPM in the context of other matching methods, compare out of box performance
		\item Incorporate propensity scores in the CPM method (estimate PS through draws of $\beta$, construct a comparison feature based on that score)
		\begin{itemize}
			\item More sophisticated, use scores in a dual latent class model. 
		\end{itemize}
		\item Variable selection prior for propensity score, feeding into a record linkage model on the selected covariates
		\item Empirical Process for deriving bounds on covariate imbalance
	\end{itemize}
\end{frame}

\end{document}

### 2023-07-30 third-submission/suggestions/deprecated/2023-may.tex

--- 2023-05-09 ba_resubmit/third-submission/suggestions/third-revision-review-comments.tex
@@
 \begin{document}
-\title{AE Comments}
+\title{General Comments}
 \author{}
@@
 
-%\begin{enumerate}
-%\item I would revert back to the original notation as I find the A, B to be confusing and it's not easy to extend to multiple files.
-(See Aleshin-Guendel (2022) for reference.
-I'm not sure what Jerry thinks about this though.
-Typically changes should be agreed upon by all authors or some standard.
-%\item Perhaps to make things easy on you, I can propose some changes in the document and help and iterate and then we can ask Jerry regarding his opinions regarding this. 
-%\item Let us also make a plan for outlining your thesis and a schedule so it can move ahead. Would you like suggestions or to propose something yourself? 
-%\end{enumerate}
+\section{Suggestions}
 
+\begin{enumerate}
+\item Let us read through the paper extremely carefully (everyone, including myself and Jerry) to make recommendations and formulate a plan. 
+\item I think it would be helpful to make comments outside the main document before making major changes in the document to make sure we address everything and check that we address every comment that was made by the Editor/AE.
+\item Should try and make clear and crisp answers to them (and cite existing literature to avoid opening any holes or new issues). 
+\item I will do my best to serve as a "reviewer/editor" and find other comments to fix. 
+\item If you find yourself getting frustrated, etc.
+please reach out/let me know.
+One idea is to work on another project that might be more fun or start formulating the rest of your thesis in case this might be weighing you down.
+I'd be happy to sketch this out with you and help you as much as you might like, Brian.
+\end{enumerate}
+
+\section{General Comments}
+
+\begin{enumerate}
+\item The connection between the relaxation of Sadinle (2018) of the bipartite linkage structure may be getting lost due to AE comment 3 as they stated ``they surprised me" regarding how the fabl prior was defined.
+This should be looked at more carefully and introduced with more care regarding the contribution of the prior as a generative (Bayesian) prior with post-processing after due to the computational issues faced with that of Sadinle.
+(This is AE comment 3).
+\item Section 2 could be structured to be more clear to a general reader regarding notation overall, prior work (FS and Sadinle) and then the relaxation of the Sadinle (2018) prior. (Editor comment). 
+\item The overall paper needs extensive checking regarding consistent notation, consistent a, no repetitive comments, no typos, etc. (Editor comment). 
+\item Appendix 8.2 needs to be provided without any typos and all details provided.
+(Don't skip any steps and do not use any short hand notation).
+I think that this was the main point of confusion for the reviewer.
+You may wish to refer to Aleshin and Guendel and Sadinle (2022) and follow their notation as it appears this may be what you want.
+See Appendix B.2 page 27.
+(In short, they use the notation $\mathcal{L}_{ij}$ to make it clear that this is a factor coming from the likelihood).
+I think that you could make your proof more clear for those working on this topic, which make this an opportunity and opening as it wasn't clear to the AE (and it's probably not obvious to others).
+\item Let's walk through each part of AE comment 3. I think it would be helpful to refer to Appendix B.2 of Aleshin and Guendel and Sadinle (2022) regarding this discussion. Page 25--26.
+
+``Thank you for including the details on integrating out pi from the full conditionals.
+They surprised me.
+I had assumed that the authors had integrated out pi in the prior for Z as this is what Sadinle (2017) had done to form the "beta prior for bipartite matchings."
+
+``Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution."
+
+\textcolor{blue}{Yes, as mentioned, our approach is a relaxation of that of Sadinle (2017) as we relax the one-to-one assumption.
+Our algorithm provided is not the same, however, empirically we find the the results are similar due to a pre-processing step that has been utilized in the literature by Jaro (??).}
+
+``Is the presented algorithm justified?"
+
+\textcolor{blue}{It will not be justified unless the full conditionals of the Gibbs sampler contain a sequential update or one proves that this is a justified Gibbs sampler.
+This was the point that Serge mentioned in his email.
+Currently, it is not a valid Gibbs sampler.}
+
+It would need to be something like this, which comes from Aleshin-Guendel and Steorts (2023).
+The partition is equivalent to the co-reference matrix or linkage structure.
+The partition approach is more widely understood by the Bayesian non-parametrics community (or statistics community), which is why we are using this terminology.
+(They will not easily understand the linkage structure or coreference matrix without connections being made for them).
+
+\begin{quote}
+			`` The most common way to update the partition, $\C$, in Bayesian entity resolution is the Gibbs update \citep{Sadinle_2014, Steorts_2015, Sadinle_2017, Tancredi_2020, Marchant_2021, Aleshin-Guendel_2022, Marchant_2023}, which originated in the Bayesian clustering literature \citep[see e.g.][]{Neal_2000}.
+			 We emphasize that this is only referring how the \textit{partition} is being updated.
+
+The Gibbs update works by sampling the cluster assignment for each record $i \in [R]$ sequentially.
+To describe the update we will represent a partition $\C$ by a  vector of labels $\bZ=(Z_1,\dots, Z_R)$, where $Z_i\in\{1,\dots, R\}$ such that $Z_i=Z_j$ if records $i$ and $j$ are in the same cluster in $\C$.
+Suppose that we have already sampled the cluster assignment for the first $i-1$ records.
+Sampling the cluster assignment for record $i$ proceeds as follows:
+			\begin{enumerate}
+				\item Let $\bZ^{[t]}_{-i}=(Z_1^{[t]},\dots,Z_{i-1}^{[t]},Z_{i+1}^{[t-1]},\dots, Z_R^{[t-1]})$ denote the current partition of $[R]$ after removing record $i$. 
+				\item Let $\bc$ denote the union of the unique labels occupied in $\bZ^{[t]}_{-i}$ and one unoccupied label.
+Each of these labels correspond to a cluster to which record $i$ could be assigned, with the unoccupied cluster representing the creation of new cluster with only record $i$.
+				\item Sample record $i$'s cluster assignment, by sampling the cluster $c\in \bc$ with probability proportional to $p(Z_{i}^{[t]}=c\mid f(\bX), \bZ^{[t]}_{-i}, \btheta^{[t]})$.
+			\end{enumerate}
+			"
+
+\end{quote}
+
+\item I have comments about the hashing section, which are largely that I think the AE/Editors are having a difficult time with this section.
+I think their main concerns are with the following \textbf{notation, precise wording, understanding of the paper, correctness of the appendix}.
+I would suggest that we all carefully read through the paper and have a plan, where we are all on the same page regarding our response and revisions.
+
+%Is it not the case that in the setup, one does not need to integrate out $\pi.$ In short, one can write
+%\begin{align}
+%P(Z, \Phi \mid \pi, \gamma_{obs}) \propto L(Z, \Phi \mid \gamma_{obs}) P(\Phi \mid \gamma_{obs}) P(Z \mid \pi) P(\pi)
+%\end{align}
+\end{enumerate}
+
+\section{Language/Writing}
+\begin{enumerate}
+\item There is repetition present in the manuscript, such as the first two lines of the opening paragraph of the introduction.
+For example, analysis, analysts, and analyses are used three times in the first two sentences.
+Tasks is used twice.
+Furthermore, the language could be written to be more precise and condensed.
+``In many applications, users seeks to identify duplications across two data files." (This removes the repetition, and I think the remainder can stay).
+\item This should be checked in an editor checked such as Antidote closer to the time of resubmission. 
+\end{enumerate}
+
+\section{Notation}
+
+\begin{enumerate}
+\item The notation is not clear or consistent throughout. 
+\item The notation is confusing at times when reading through the paper, where one has to flip back and forth between pages to re-reference previously defined notation.
+The authors should more carefully define (perhaps following that of Sadinle (2018) to avoid typos) and use referencing of equations so that the reader doesn't have to constantly flip back and forth throughout the manuscript.
+\item The usage of lower and upper case greek letters should be consistent.
+The usage of subscripts should be consistent.
+It appears that Sadinle (2018) and Aleshin and Guendel and Sadinle (2022) make use of notation consistently, so this could be followed.
+It might be worth noting that they utilize $X_1$ and $X_2$ in order to more effectively extend to any number of databases, so this might be more natural than $A$ and $B.$
+\end{enumerate}
+
+\newpage
 \section{AE Comments with responses}
@@
 \end{quote}
-\item Sections 4.1 and 4.2 (I have gone through this, however, I suggest that we all go through this many times and write out responses very clearly).
-I am doing my best to get through this as quickly as possible and make recommendations.
+
 \end{enumerate}

### 2023-07-30 ba_resubmit_template_files/kundinger_resubmit3.tex
update brian's file.

--- 2023-07-30 ba_resubmit/kundinger_resubmit3.tex
@@
 We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler.
-Appendix \ref{app:derivations} presents derivations for the full conditional distributions.
+Supplement \ref{app:derivations} presents derivations for the full conditional distributions.
 We initialize $\bm{m}$, $\bm{u}$ and $\pi$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
@@
 To obtain a Bayes estimate that fulfills the bipartite requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-See Appendix \ref{bayes-estimate} for details redarding the initial Bayes estimate and this post-processing procedure.
+See Supplement \ref{bayes-estimate} for details redarding the initial Bayes estimate and this post-processing procedure.
 
@@
 This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
-%A more complete example of the one-hot encoding is provided in Appendix \ref{app:ohe}.
+%A more complete example of the one-hot encoding is provided in Supplement \ref{app:ohe}.
 
@@
 The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.
+To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement \ref{app:appendix-speed}.
 
@@
 We run the Gibbs sampler for 1000 iterations, and discard the first 100 as burn-in.
-We calculate Bayes estimates $\hat{\bm{Z}}$ of the linkage structure using the  loss function and post-processing procedure described in Appendix \ref{bayes-estimate}.
-Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
-We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; those results are in Appendix \ref{partial}.
+We calculate Bayes estimates $\hat{\bm{Z}}$ of the linkage structure using the  loss function and post-processing procedure described in Supplement \ref{bayes-estimate}.
+Traceplots for parameters of interest for one example simulation are provided in Supplement \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
+We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; those results are in Supplement \ref{partial}.
 
@@
 Estimates of the $\bm{u}$ parameters are indistinguishable, and thus omitted.
-Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
+Traceplots for parameters of interest are provided in Supplement \ref{app:appendix-es}.
 
@@
 
-	%We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Appendix \ref{partial}.
+	%We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Supplement \ref{partial}.
 Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
@@
 	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-Traceplots do not suggest convergence issues, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}.
+Traceplots do not suggest convergence issues, and are similar to those seen in Supplement \ref{app:appendix-sim} and \ref{app:appendix-es}.
 As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure has of 9634 matches, with a 95\% credible interval of (9581, 9740).
@@
 
-	We provide detailed derivations of the full-conditionals provided in Section 3 of the main text.
+	We provide detailed derivations of the full-conditionals provided in Section \ref{sec:fast-beta-linkage}.
 The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
@@
 		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
-	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section 5.2 of the main text.
+	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
 It is set up with one error across the linkage fields and 50 duplicates across files.
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section 5.2 of the main text.}\label{fig:sim_overlap_trace}
+			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section 5.2 of the main text.}\label{fig:sim_m_trace}
+			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section 5.2 of the main text.}\label{fig:sim_u_trace}
+			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
 		\end{center}
@@
 
-In this section, we repeat the simulation study in Section 5.2 of the of the main text, allowing for clerical review rather than forcing all records to have or not have links.
+In this section, we repeat the simulation study in Section \ref{accuracy}, allowing for clerical review rather than forcing all records to have or not have links.
 Specifically, by leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
@@
 		\includegraphics[width=0.99\textwidth]{../notes/figures/sadinle_sim_plot_partial2} 
-		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
+		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Supplement \ref{partial}.
 We see poorer performance for \texttt{fabl} only in situations with high overlap.}
@@
 	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
-We replicate the speed study of Section 5.1 of the main text under different settings.
+We replicate the speed study of Section \ref{speed} under different settings.
 Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.
@@
 		\end{tabular}
-		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
+		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Supplement \ref{app:appendix-speed}.}\label{Tab:distributions_2}
 	\end{table}
@@
 %		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2_new} 
-%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
+%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Supplement \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
 We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
@@
 %		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB2_new} 
-%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
+%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Supplement \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
 We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
@@
 	\end{subfigure}
-	\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulations described in Appendix \ref{app:appendix-speed}.
+	\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulations described in Supplement \ref{app:appendix-speed}.
 In (\ref{fig:app-speed1}), both $n_A$ and $n_B$ are increasing.
@@
 	\end{subfigure}
-	\caption{Traceplots for parameters of interest in El Salvador case study in Section 5.1 of the main text.}
-	\label{fig:mean and std of nets}
+	\caption{Traceplots for parameters of interest in El Salvador case study in Section \ref{el_salvador}.}
+	\label{fig:el_salvador_trace}
 \end{figure}

### 2023-07-30 SEI_sensitivity/Sadinle_accuracy_cluster.Rmd

@@
 
-fabl_acc_samps <- matrix(NA, nrow = 5, ncol = 3)
-fabl_acc_samps_resolved <- matrix(NA, nrow = 5, ncol = 3)
-sad_acc_samps <- matrix(NA, nrow = 5, ncol = 3)
+fabl_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
+fabl_acc_samps_resolved <- matrix(NA, nrow = 3, ncol = 6)
+sad_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
 
-fabl_acc_samps2 <- matrix(NA, nrow = 6, ncol = 3)
-fabl_acc_samps_resolved2 <- matrix(NA, nrow = 6, ncol = 3)
-sad_acc_samps2 <- matrix(NA, nrow = 6, ncol = 3)
+fabl_acc_samps2 <- matrix(NA, nrow = 3, ncol = 6)
+fabl_acc_samps_resolved2 <- matrix(NA, nrow = 3, ncol = 6)
+sad_acc_samps2 <- matrix(NA, nrow = 3, ncol = 6)
 
@@
   eval_resolved <- GetEvaluations(Zhat_resolved, Ztrue, n1)
-  fabl_acc_samps_resolved[, j] <- c(eval_resolved, elapsed[3], overlap)
-  fabl_acc_samps[, j] <- c(eval, elapsed[3], overlap)
+  fabl_acc_samps_resolved[j, ] <- c(eval_resolved, NA, elapsed[3], overlap)
+  fabl_acc_samps[j, ] <- c(eval, NA, elapsed[3], overlap)
 
@@
   eval[1] <- sum(Zhat == Ztrue & Ztrue > n1) / (sum(Zhat > n1))
-  fabl_acc_samps2[, j] <- c(eval, RR, elapsed[3], overlap)
+  fabl_acc_samps2[j, ] <- c(eval, RR, elapsed[3], overlap)
 
@@
   eval <- GetEvaluations(Zhat, Ztrue, n1)
-  sad_acc_samps[, j] <- c(eval, elapsed[3], overlap)
+  sad_acc_samps[j, ] <- c(eval, NA, elapsed[3], overlap)
 
@@
   eval[1] <- sum(Zhat == Ztrue & Ztrue > n1) / (sum(Zhat > n1))
-  sad_acc_samps2[, j] <- c(eval, RR, elapsed[3], overlap)
+  sad_acc_samps2[j, ] <- c(eval, RR, elapsed[3], overlap)
 
@@
 }
+
+fabl_acc_samps <- data.frame(fabl_acc_samps, "fabl") %>% 
+  unname() %>% 
+  data.frame()
+fabl_acc_samps_resolved <- data.frame(fabl_acc_samps_resolved, "fabl_resolved") %>% 
+  unname() %>% 
+  data.frame()
+fabl_acc_samps2 <- data.frame(fabl_acc_samps2, "fabl_partial") %>% 
+  unname() %>% 
+  data.frame()
+sad_acc_samps <- data.frame(sad_acc_samps, "BRL") %>% 
+  unname() %>% 
+  data.frame()
+sad_acc_samps2 <- data.frame(sad_acc_samps2, "BRL_partial") %>% 
+  unname() %>% 
+  data.frame()
+
+result_df <- rbind(fabl_acc_samps, fabl_acc_samps_resolved, fabl_acc_samps2,
+                   sad_acc_samps, sad_acc_samps2)
 

### 2023-07-30 ba_resubmit_template_files/kundinger_resubmit3.tex
correct complication issue with pi at line 23 and 24; add comma after thus.
update references.
fix authors.

@@
 %
-\pubyear{2021}
+\pubyear{2023}
 \volume{TBA}
@@
 \and
-\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}
+\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr1}\ead[label=e3]{beka@stat.duke.edu}}
 
@@
   \printead{e1}, % print email address of "e1"
-  \printead*{e2}
+  \printead*{e2},
+   \printead*{e3}
 }
 
-\address[addr2]{Departments of Statistical Science and Computer Science,
-  Duke University,
-  P.O.\ Box 90251,
-  Durham, NC 27708, USA
-  \printead{e3}
-}
+%\address[addr2]{Departments of Statistical Science and Computer Science,
+%  Duke University,
+%  P.O.\ Box 90251,
+%  Durham, NC 27708, USA
+%  \printead{e3}
+%}
 
@@
 \begin{abstract}
-%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
 Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.
@@
 
-% Possible citation for bipartite matching. 
-%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_A995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.
-
+\newpage
 The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular for its mathematical simplicity.
@@
 
-%Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
-
 In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
@@
 
-%
-
-%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}}
-
 In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
@@
 
-%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$.
-Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.
-
 Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values.
@@
 We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$ as the collection of all record pairs $\gamma_{ij}$.
-%Additionally, we let $\Gamma_{ij}$ represent a random variable for the comparison vector for $A_i$ and $B_j$, such that $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
 
@@
 The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\Gamma_{ij}^f = l|\Delta_{ij} = 0)$.
-
-%Within this framework, record pairs are independently classified as matches and nonmatches based on the estimated parameters.
-However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results.
-To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching.
-\cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.
-
-%The Fellegi-Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
-This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.
-
-%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
-%\begin{align}
-%     \label{eqn:jaro}
-%     &\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij}
-%     \quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
-%     & \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
-%     & \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
-%\end{align}
-%
-%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs.
-Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.
-
-%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data.
-This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
-%\begin{align}
-%     \label{eqn:wts}
-%     w_{ij} = \frac{p(\gamma_{ij} \mid \Delta_{ij} = 1)}{p(\gamma_{ij} \mid \Delta_{ij} = 0)}
-%\end{align}
-%as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.
-%
-%In practice, $p(\cdot \mid \Delta_{ij} = 1)$ and $p(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated.
-%
-%%is the set of parameters $u_{fl} = p(\gamma_{ij}^f = l|Z_j \neq i)$ for the nonmatching pairs.
-%
-%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review).
-This is a problem in practice as transitive closures can be violated.
-Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.
-
-%\subsection{Beta Record Linkage Model}
-%\label{BRL}
-
-%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.
-
-%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a
 
@@
 For each $Z_j$, we have
-%\begin{subequations}
+
 \begin{align}
@@
 \pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}). \nonumber
-%\label{eqn:fast_beta_prior2}.
-\end{align}
-%\end{subequations}
+\end{align}
+
 We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match.
@@
 
-%\begin{table}[t!]
-%	\centering
-%	\begin{tabular}[t!]{ll}
-%		Symbol & Description \\
-%		\hline
-%		$A, B$ & data files\\
-%		$i \in 1, \ldots, n_A $ & index over records in $A$\\
-%		$j \in 1, \ldots, n_B $ & index over records in $B$\\
-%		$f \in 1, \ldots F$ & index over fields used for comparisons \\
-%		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
-%		$n_{AB}$ & number of matches between $A$ and $B$\\
-%		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
-%		$Z_j = i$ & records $A_i$ and $B_j$ match \\
-%		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
-%		$m_{fl}$ & $p(\gamma_{ij}^f = l | Z_j = i)$ \\
-%		$u_{fl}$ & $p(\gamma_{ij}^f = l | Z_j \neq i)$ \\
-%		$\pi$ & probability that a record in $B$ has a match in $A$ \\
-%		\hline
-%	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
-%\end{table}
-%
 For clarity, we present the full model for \texttt{fabl} below:
@@
 \end{align}
-%\begin{align}
-%\label{eqn:z_full_conditional}
-%p\left(Z_j^{(s+1)}  = z_j|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \bm{Z^{(s)}}\right) \propto
-%\begin{cases} 
-%	w_{z_j, j}^{(s+1)},  & z_j \leq n_A; \\
-%	n_A \frac{n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}}{n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}}, & z_j  = n_A + j, \\
-%\end{cases}
-%\end{align}
 where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
@@
 This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.
-%A more complete example of the one-hot encoding is provided in Supplement \ref{app:ohe}.
 
@@
 
-%We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
-
 Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
@@
 	\begin{cases} 
-		%\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
 		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
@@
 
-		 %For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
-
 		We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$, and delete the memory intensive matrix $\gamma^{ab}$ before continuing with the next batch of data. In detail, we calculate
@@
 We collect these memory reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.
-
-%Let $n_a$ and $n_b$ be the number of records in batches $A^a$ and $B^b$ respectively. Instead of storing $n_A n_B$ record labels, with SEI we store at most $t_A \sum_{b = 1}^{t_B} n_b  P  S$ labels.
 
@@
 
-%	\begin{figure}[t]
-%		\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot_new} 
-%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
-%		\end{center}
-%	\end{figure}
-%
-%	\begin{figure}[h!]
-%	\begin{center} \includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_new} 
-%		\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
-We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
-%	\end{center}
-%\end{figure}
-
 	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}.
@@
 Implementing \texttt{fabl} in C++ with parallelization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
-
-%		\begin{figure}
-%		\begin{minipage}[c]{0.45\linewidth}
-%			\begin{center}
-%				\includegraphics[width=0.8\textwidth]{../notes/figures/sadinle_speed_plot_new}
-%				\caption{Simulations with increasing values of both $n_A$ and $n_B$. We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
-%			\end{center}
-%		\end{minipage}
-%		\hfill
-%		\begin{minipage}[c]{0.45\linewidth}
-%			\begin{center}
-%				\includegraphics[width=0.8\textwidth]{../notes/figures/speed_plot_fixed_nB_new} 
-%				\caption{Simulations with $n_B$ = 500, We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
-%			\end{center}
-%		\end{minipage}%
-%	\end{figure}
 
@@
 In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-	%In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.
 
@@
 
-	%It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-
 	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in $n_A = 4420$ records in CDHES and $n_B = 1323$ records in ERTL.
@@
 This setup leads to 1875 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
-		%	\end{table}
-
-	%	\begin{table}[t]
+
 		\centering
@@
 	\end{center}
-	%	\end{figure}
-
-%	\begin{figure}[t]
+
 	\begin{center}
@@
 We note that Bayes estimates falling outside of posterior credible intervals has been observed previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
-
-	%We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study in Supplement \ref{partial}.
-Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually.
-This means that after clerical review, the number of individuals replicated across data files would fall in the interval (136, 311), encapsulating the posterior credible interval.
-More or fewer records could be identified for clerical review by decreasing or increasing $\theta_R$.
-
-	%and a range of (140, 294) after the partial estimate and clerical review.
 
@@
 	\bibliographystyle{jasa}
-	\bibliography{biblio}
+	\bibliography{fabl}
 
@@
 Analogous procedures lead to $\mathcal{L}(u_{fl}| \gamma, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
-Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
-	\begin{align}
-	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s) &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \\
-	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s) &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).
+Thus, for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
+	\begin{align}
+	\bm{m}_f^{(s+1)} |
+	\gamma, 
+	\bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim 
+	\text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \\
+	\bm{u}_f^{(s+1)} | \gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).
 	\end{align}
@@
 		\end{center}
-%	\end{figure}
-
-%	\begin{figure}[!h]
+
 		\begin{center}
@@
 
-%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
-
 \begin{figure}
@@
 	\end{table}
-
-%	\begin{figure}[h!]
-%		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2_new} 
-%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Supplement \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
-%		\end{center}
-%	\end{figure}
-%	
-%	\begin{figure}[h!]
-%		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB2_new} 
-%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Supplement \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
-We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
-%		\end{center}
-%	\end{figure}
 
@@
 
-%	\begin{figure}[!h]
-%		\begin{center}
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-%			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
-%		\end{center}
-%	\end{figure}
-%
-%	\begin{figure}[!h]
-%	\begin{center}
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/pi_trace} 
-%		\caption{Traceplot for $\pi$ parameter in El Salvador case study.} \label{fig:pi_trace}
-%	\end{center}
-%\end{figure}
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-%			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} \label{fig:m_trace}
-%		\end{center}
-%	\end{figure}
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-%			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} \label{fig:u_trace}
-%		\end{center}
-%	\end{figure}
-
     \begin{figure}[h!]

### 2023-07-30 ba_resubmit_template_files/main.tex
fixed rendering.
working on main and supp.

--- 2023-07-30 ba_resubmit/main.tex
@@
 %
-\pubyear{0000}
-\volume{00}
-\issue{0}
-\doi{0000}
+\pubyear{2023}
+\volume{TBA}
+\issue{TBA}
+% \doi{0000}
 %\arxiv{}
@@
 \usepackage{lineno}
+\usepackage{afterpage}
 \usepackage{amsthm}
@@
 \printead{e1}, % print email address of "e1"
-\printead*{e2}
+\printead*{e2},
+\printead*{e3}
 }
 
-\address[addr2]{Departments of Statistical Science and Computer Science,
-Duke University,
-P.O.\ Box 90251,
-Durham, NC 27708, USA
-\printead{e3}
-}
+%\address[addr2]{Departments of Statistical Science and Computer Science,
+%Duke University,
+%P.O.\ Box 90251,
+%Durham, NC 27708, USA
+%\printead{e3}
+%}
 
@@
 However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.
-To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
-Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, \textcolor{red}{unlike} the sequential updates of \texttt{BRL}.}
-\textcolor{teal}{We ensure \textcolor{red}{\text{fabl} results in a bipartite matching} using decision theoretic post-processing techniques from the existing literature.}
-\textcolor{red}{To reduce computational overhead and memory costs, we use hashing and a new technique that we call storage efficient indexing. }
-%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead.
-Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing.
-%}
-%This allows
-%us to use hashing techniques that reduce computational overhead.
+To address these computational difficulties, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}.
+Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead.
 This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing.
@@
 %% ** Keywords **
-\begin{keyword}
+\begin{keyword}%[class=MSC]
+\kwd{data fusion}
+\kwd{data cleaning}
+\kwd{entity resolution}
+\kwd{hashing}
 \kwd{record linkage}
-\kwd{entity resolution}
-\kwd{data cleaning}
-\kwd{parallelization}
-\kwd{hashing}
 \end{keyword}
@@
 Before conducting data analysis, it is often necessary to identify duplicate records across two data files.
-This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
+This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep[e.g.,][]{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}.
 In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.
 
-Many probabilistic record linkage methods are rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
+Many probabilistic record linkage methods rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}.
 In their approach, the data analyst first creates comparison vectors for each pair of records in the data files.
@@
 
-The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity.
+\newpage
+The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular for its mathematical simplicity.
 However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file.
 Thus, when the procedure results in many-to-one matches, some of these links must be false.
-To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}
+Analysts typically use an additional post-processing step to turn the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.
 
 Alternatively, analysts can embed one-to-one matching requirements into the model specification, at an additional computational cost.
-\cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
+\cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respect one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space.
 \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records.
 \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler.
-Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review.
 \texttt{BRL} has been shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.
@@
 In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability.
-\textcolor{red}{We use independent prior distributions for the matching status of each record, an idea suggested in \cite{wortman2019}.}
-%% Jerry thought the wording below was too vague and made a suggestion that is more direct and clear. 
-%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}}
-%
-\textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, \textcolor{red}{unlike the} sequential updates of \cite{sadinle_bayesian_2017}.
-We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
-This \textcolor{red}{approach} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
+We use independent prior distributions for the matching statuses of each record, and modify the decision theoretic technique of \cite{sadinle_bayesian_2017} to ensure our linkage estimates are bipartite.
+This approach allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
 These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
@@
 In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
-Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching.
+Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and describes the loss function used to calculate the Bayes estimate for the bipartite matching.
 Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}.
@@
 Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values.
-\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define a \emph{comparison vector} $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
-For notational convenience, we define the \emph{comparison matrix} $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.
+\cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$.
+We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$ as the collection of all record pairs $\gamma_{ij}$.
 
@@
 
-%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
+In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$.
 As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values.
-%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
+We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}).
 To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}
+With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
 
@@
 Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.
-For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent.
+For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are independent given the matching status of the record pair, and second, that the matching status of each record pair is independent of the matching status of other pairs.
 Using these independence assumptions, one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.
@@
 When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature.
-Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
-The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.
+Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\Gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$.
+The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\Gamma_{ij}^f = l|\Delta_{ij} = 0)$.
 
 \cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.
-He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.
-\textcolor{red}{However, he replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\mathbf{Z}$ that he calls the ``beta distributions for bipartite matching."}
-%
-%As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  
-This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
-Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
-%
+He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$, and replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\bm{Z}$ that he calls the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
+Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, where $\alpha_{\pi}$ and $\beta_{\pi}$ are known hyperparameters.
 It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$.
@@
 
-\textcolor{red}{For \text{fabl}, we also use (\ref{eqn:fs_model}) and (\ref{eqn:m_dist}), along with Dirichlet prior distributions for the $\mathbf{m}$ and $\mathbf{u}$ parameters.
-However, instead of the prior over $\bm{Z}$}
-%Instead of the prior over $\bm{Z}$ 
-from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, 
-\textcolor{teal}{we propose an alternative, called the fast beta linkage (fabl) prior.}
-%the fast Beta prior as follows. 
+Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$.
+However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast beta linkage (\texttt{fabl}) prior below.
 For each $Z_j$, we have
 
-\begin{subequations}
 \begin{align}
@@
 \end{cases} \label{eqn:fast_beta_prior}  \\
-\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \label{eqn:fast_beta_prior2}.
-\end{align}
-\end{subequations}
+\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}). \nonumber
+\end{align}
+
 We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match.
@@
 
-\begin{rem}
-\textcolor{red}{
 In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$.
 This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$.
-In our preliminary studies, this highly informative prior weighting on matching can result in overly optimistic match rates.
-Hence, \textcolor{red}{we prefer (\ref{eqn:fast_beta_prior}).} We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
-}
-\end{rem}
-
-\begin{rem}
-Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model.
-That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in \textcolor{red}{the \cite{fellegi_theory_1969} model} estimates the proportion of record pairs that are matches.
-In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}).
-In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by \textcolor{red}{$1/n_A$} and tends towards 0 as the size of the linkage task grows.
-Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
+In our preliminary studies, we have found that this highly informative prior weighting on matching can result in overly high match rates.
+Hence, we prefer (\ref{eqn:fast_beta_prior}).
+We also note that the flat prior is equivalent to a special case of the fast beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
+
+Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the \cite{fellegi_theory_1969} model.
+That is, $\pi$  in (\ref{eqn:fast_beta_prior}) under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in (\ref{eqn:lambda_dist}) in the \cite{fellegi_theory_1969} model estimates the proportion of record pairs that are matches.
+In the bipartite case, we conjecture that some analysts will find $\pi$ to be a more interpretable parameter than $\lambda$.
+In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $1/n_A$ and tends towards 0 as the size of the linkage task grows.
+Additionally, while the \cite{fellegi_theory_1969} model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions.
 As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.
-\end{rem}
-
-\textcolor{red}{In many record linkage tasks, the records in $A$ or $B$ have missing fields, so that some $\gamma_{ij}^f$ could be missing.
-To handle this, we follow the approach described in \cite{sadinle_bayesian_2017}.
-Let $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-We assume that any missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
-For likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}, so that we can marginalize over the missing data and perform all computation using the observed data \citep{sadinle_detecting_2014, sadinle_bayesian_2017}.}
-
-%\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors.
-To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing at random (MAR) or missing completely at random (MCAR).
-It is important to note that for likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}.
-Therefore, we can marginalize over the missing data and perform all computation using the observed data, leading to (\ref{eqn:likelihood}).
-For further details on these connections, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
-%}
-
-\textcolor{teal}{We summarize \textcolor{red}{the model for \text{fabl}} as follows:}
-
+
+For clarity, we present the full model for \texttt{fabl} below:
 \begin{subequations}
@@
 	\end{cases} \label{eqn:z}\\
-	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi-prior},
+	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
 \end{align}
 \end{subequations}
-\textcolor{teal}{where recall that \textcolor{red}{$(\alpha_{\pi}, \beta_{\pi})$} are fixed hyper-parameters.}
-
-%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
-%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
-
-%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
-%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,
-
-\textcolor{teal}{We \textcolor{red}{now} present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m},$ $\bm{u},$ and $\pi$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi-prior}).
-%While we could marginalize over $\pi$ and derive a collapsed Gibbs sampler, in the spirit of \cite{sadinle_bayesian_2017}, this leads to sequential updates, which are computationally slow.
-Instead, we proceed with a traditional Gibbs sampler, where the updates to the matching label (or coreference matrix) are parallel, allowing for more efficient updates.
-Refer to Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
-%%
-We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
-\textcolor{blue}{Brian: state how you initialize $\pi$.} \textcolor{teal}{Denote the current draw of the sampler by $s$ and the updated draw by $s+1$.} \textcolor{red}{To update $\mathbf{m}$ and $\mathbf{u}$, for all, $f=1, \ldots, F,$ sample}
-
-%\begin{subequations}
-%\begin{align}
-%	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  \\
-%	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
-%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-%\end{align}
-%\end{subequations}
-
-%\begin{subequations}
-%
-%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
-%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-%\end{align}
-%\end{subequations}
-
-\begin{align}
-	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  
-\end{align}
-and	
-\begin{align}
-	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} 
-\end{align}
-
-where
-
-\begin{align}
-\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \quad \text{and}\\
-\beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
-\end{align}
-
-\textcolor{red}{Next, we sample $\pi$ from its full conditional, given by}
-%Next, the full conditional for $\pi$ is given by
-\begin{align}
-\label{eqn:pi}
-	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
+
+We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler.
+Supplement \ref{app:derivations} presents derivations for the full conditional distributions.
+We initialize $\bm{m}$, $\bm{u}$ and $\pi$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$.
+Denote the current draw of the sampler by $(s)$ and the updated draw by $(s+1)$.
+To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, we sample
+\begin{subequations}
+\begin{align}
+	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
+	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
+	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
+	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
+\end{align}
+\end{subequations}
+Next, we sample $\pi$ from its full conditional, given by
+\begin{align}
+	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
 \end{align}
@@
 \end{align}
-
 where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
@@
 
-\textcolor{teal}{
-The full conditional in (\ref{eqn:z_full_conditional}) does not depend on $\bm{Z}^{(s)},$ implying that 
-\begin{align}
-p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}) \propto \prod_q \left(\prod_j
-p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right)
-\right).
-\end{align}
-}
-
-\textcolor{teal}{Thus, we can make parallel updates to $\bm{Z}$ instead of sequential updates as in \cite{sadinle_bayesian_2017}.
-In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), \textcolor{red}{in contrast to} \textcolor{red}{other} Bayesian record linkage \textcolor{red}{methods} \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}.
-}
-%See Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
-
-%Derivations for the full conditional distributions can be found in Appendix~\ref{app:derivations}.
-
-\textcolor{teal}{Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
+Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}.
 Since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-To obtain a Bayes estimate that fulfills the \textcolor{red}{bipartite} requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-See Appendix \ref{bayes-estimate} for details regarding the initial Bayes estimate and \textcolor{red}{this} post-processing procedure.}
-
-%Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.
+To obtain a Bayes estimate that fulfills the bipartite requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
+See Supplement \ref{bayes-estimate} for details redarding the initial Bayes estimate and this post-processing procedure.
 
@@
 The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains.
-Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking \textcolor{teal}{often}
- undesirable %in some situations 
-\citep{steorts_comparison_2014}.
+Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking often undesirable \citep{steorts_comparison_2014}.
 
@@
 For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields.
-However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the \textcolor{red}{classical} Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.
-
-With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage.
-First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors.
+However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical \cite{fellegi_theory_1969} model, leaving the effect of indexing on Bayesian record linkage models to future work.
+
+With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage.
+First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of storing the comparison vectors.
 Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.
@@
 
-Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. 
-\textcolor{red}{The quantity $P$ is
- bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1),$ which clearly depends only on $F$ and $L_f.$ It does not depend (or scale with) with $n_A$ or $n_B.$}
-% This comment below seems obvious and not needed.  
-% where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.
-
-%is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
+Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$.
+Let $P$ be the number of patterns realized in $\gamma$.
+It is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.
+This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.
 
@@
 
-%\textcolor{red}{Do you mean what I have below or something else? This still is a bit hard to parse.}
-We then map \textcolor{teal}{the integers in (\ref{eqn:hashing})} to sequential integers $\{1, \ldots, P\}$.
-\textcolor{teal}{Denote} each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
-When record pair, \textcolor{teal}{$(i,j)$,} exhibits agreement pattern $p$, \textcolor{teal}{let} $\gamma_{ij} = h_p$.
-
+We then map the integers in \eqref{eqn:hashing} to sequential integers $\{1, \ldots, P\}$.
+Denote each unique agreement pattern as $h_p = (h_p^1, \ldots, h_p^F)$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$.
+When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$.
 In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$.
@@
 
-\begin{example}
-\label{ex:one-hot}
-%% Jerry propose with (want to avoid using with twice in the same sentence :)
-\textcolor{teal}{Consider} five fields with binary agreements and possible missingness (NA).
+For example, consider five fields with binary agreements and possible missingness, denoted $NA$.
 The number of possible patterns is bounded above by $P^{*} = 3^5 = 243$.
-\textcolor{red}{Suppose that} records $A_5$ and $B_7$ exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, \text{NA}, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
-Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
-\textcolor{red}{All records with a hashed value of 1066 map to the integer 42. Then, $\gamma_{5,7} = h_{42}.$}
-%We map the unique values of $h^{*}(\gamma)$ to sequential integers. \textcolor{teal}{Suppose that 1066 maps to 42. This implies $\gamma_{5,7} = h_{42}.$}
-%If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ 
-This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
-\end{example}
+Suppose that records $A_5$ and $B_7$ exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field.
+The expression in  (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$.
+All records with a hashed value of $1066$ map to the integer $42$.
+Thus, $\gamma_{5,7} = h_{42},$ and this agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
 
@@
 We denote this set $r_{p_j} = \{i \in 1, \dots, n_A | \gamma_{ij} = h_p\}$, and collect all such sets in the nested list $\mathcal{R} = \{r_{p_j} | p \in \{1, \ldots, P\}, j \in \{1, \ldots, n_B\} \}$.
-We compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$,
+We compute the number of records in $A$ that share agreement pattern $p$ with record $B_j$, given by
 \begin{align}\label{eqn:N}
@@
 
-The set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for purposes of writing the likelihood function for \texttt{fabl}.
+The set $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N}\}$ fully characterizes the comparison matrix $\gamma$ for writing the likelihood function for \texttt{fabl}.
 To see this, we employ the condensed notation
@@
 	\begin{align}
-		m_p =  \textcolor{red}{p(\Gamma_{ij} = h_p|Z_j = i)} = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:m_p} \\
-		u_p =  \textcolor{red}{p(\Gamma_{ij} = h_p|Z_j \neq i)} = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\label{eqn:u_p}
+		m_p =  p(\Gamma_{ij} = h_p|Z_j = i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(h_p^f = l)I_{obs}(h_p^f)}, \label{eqn:m_p} \\
+		u_p =  p(\Gamma_{ij} = h_p|Z_j \neq i) = \prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(h_p^f = l)I_{obs}(h_p^f)}\label{eqn:u_p}
 	\end{align}
 \end{subequations}
-to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match, in (\ref{eqn:m_p}), and not a match, in (\ref{eqn:u_p}).
-Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) \textcolor{teal}{can be expressed as}
-%is equivalent to
+to express the probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match in (\ref{eqn:m_p}), and not a match in (\ref{eqn:u_p}).
+Viewed through the perspective of agreement patterns, the likelihood in (\ref{eqn:likelihood}) is equivalent to
 
@@
 
-\textcolor{teal}{In our derivations below, we assume the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the updates to each full conditional distribution based upon our computational proposals using hashing.}
-
-%\begin{enumerate}	
-%\item \textcolor{red}{Note that 	$\gamma_{Z_j, j}$ doesn't appear to be defined.
-Also, do you more formally mean something like as follows: Consider $\gamma_{q,q} = h_p \mid Z_j = q$.
-It seems that the conditioning may actually be important here.
-I mention this for two reasons, it appears a bit strange in equation (22).
-When I think through it, it's okay.
-I just wanted to bring it up as a conversation before submission.}
-%\item \textcolor{red}{There appears to be no update for $\pi$ in this section, so I have added one in.}
-%\item Make sure when you give the update for the new efficient full conditional, you give the s, s+1.
-Example, it's missing on the update for $m$ and $u$ so it's not immediately clear how one would draw this directly.
-Wouldn't it just be a vectorized Dirichlet, where the computational savings come from vectorizing the hyperparameters of the Dirichlet?
-%\end{enumerate}
-
-Turning first to the \textcolor{teal}{full conditional updates}
-%posterior updates 
-for $\bm{m}$ and $\bm{u}$, we observe that these parameters only depend on the \textcolor{teal}{observed} data  through quantities that are calculated through $\mathcal{N}$ and $\mathcal{P}$.
+The posterior updates for $\bm{m}$ and $\bm{u}$ depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$.
 Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
-Conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional distribution for $\bm{m}$ and $\bm{u}$ as
+Then, conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
 \begin{align}
@@
 \end{align}
-Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be a concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
-The terms needed for the \textcolor{teal}{full conditional} updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
+Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively.
+The terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
 \begin{subequations}
@@
 \end{subequations}
-%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-%Specifically, the $\ell + \sum_{k=1}^{f-1} L_k$ component of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provides the the posterior updates for level $\ell$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
-However, the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
-\textcolor{teal}{\textcolor{red}{We can} compute the vectorized summations in (\ref{efficient_alpha}) and (\ref{efficient_beta}) more efficiently than the previous updates in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}), which require summing over $n_A n_B$ record pairs for each field and agreement level.}
-
-\textcolor{teal}{Turning to sampling for $\pi,$ it remains the same as in (\ref{eqn:pi}), \textcolor{red}{so we do not repeat it.} }
-
-We also use (\ref{likelihood_efficient}) to sample $\bm{Z}$ efficiently.
-For each pattern $p$, let $w_p = m_p/u_p$.
+Specifically, the $l + \sum_{k=1}^{f-1} L_k$ components of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provide the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}).
+However, the vectorized summations can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
+
+Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
+For each pattern $p$, let $w_p =m_p/u_p$.
 The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
@@
 \end{align}
-This likelihood lends itself to markedly faster posterior computations.
 Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large.
 This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-\textcolor{red}{We use two steps in sampling each $Z_j$.}
-%To speed up computation, we break this sampling step into two.
+To speed up computation, we use \eqref{likelihood_efficient_z} to break this sampling step into two.
 
 We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
-
 \begin{align}
 	\label{eqn:gibbs1}
-	p\left( \Gamma_{Z_j^{(s+1)},j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
-%	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
 	\begin{cases} 
-%		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & \gamma_{Z_j^{(s+1)}, j} = h_p; \\
 		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
@@
 \end{align}
-
 Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
@@
 	\label{eqn:gibbs2}
-%	P\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) 
-P\left(Z_j^{(s+1)} = q \mid \Gamma_j=h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) 
-	= 
-	\begin{cases} 
+	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
 		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
@@
 
-		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint \textcolor{green}{chunks} for more manageable computations.
+		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint batches for more manageable computations.
 Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.
 Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.
-For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records.
-For \textcolor{red}{all $a$ and $b$}, we conduct all-to-all comparisons \textcolor{red}{of the pairs of records in each $(A^a, B^b)$} to construct the comparison matrix $\gamma^{ab}$.
-
-		We conduct hashing, obtain the compressed \textcolor{teal}{comparison matrix} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
-		the 
-\textcolor{teal}{memory-intensive}
-		\textcolor{teal}{comparison matrix} $\gamma^{ab}$ 
-		before continuing with the next chunk of data. 
-		In detail, we calculate
+For each $a$ and $b$, we compute comparison vectors for all records in $A^a \times B^b$ to construct the comparison matrix $\gamma^{ab}$.
+
+		We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$, and delete the memory intensive matrix $\gamma^{ab}$ before continuing with the next batch of data. In detail, we calculate
 \begin{subequations}
 	\begin{align}
-		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
+		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, B_j \in B^b\}, \label{r_{p_j}^{ab}} \\
 		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
@@
 As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses less memory than storing the full comparison matrix $\gamma$.
-However, the memory requirements of each are still quadratic in nature.
-For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome.
-We introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
+However, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome for very large linkage tasks.
+We next introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches.
 This allows all-to-all comparisons for substantially larger linkage tasks.
@@
 Rather than store all of these record labels, we store only a small number $S$.
-For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $\text{SEI}(r_{p_j}^{ab})$.
-We collect these memory reduced lists to form $\text{SEI}(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $\text{SEI}(\mathcal{R})$.
-
-Let $n_{A^a}$ and $n_{B^b}$ be the number of records in \textcolor{green}{chunks} $A^a$ and $B^b$ respectively.
-Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels.
-As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $\text{SEI}(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
-We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
+For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form SEI$(r_{p_j}^{ab})$.
+We collect these memory reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.
+
+As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
+We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.
 
@@
 
-	\begin{figure}[t]
-		\begin{center}
-%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_speed_plot2}  
-	 \includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_speed_plot2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including the hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$, as described in Section \ref{speed}.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:speed1}
-		\end{center}
-	\end{figure}
-
 	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}.
 The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Appendix \ref{app:appendix-speed}.
+To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement \ref{app:appendix-speed}.
 
@@
 Implementing \texttt{fabl} in C++ with parallelization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.
-
-	\begin{figure}[h!]
-		\begin{center} 
-%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides} 
-		\includegraphics[width=0.7\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations, including hashing step for \texttt{fabl}, with $n_B$ fixed at 500, as described in Section \ref{speed}.
-We see near linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:speed2}
-		\end{center}
-	\end{figure}
 
@@
 Under each of these settings, we use 100 pairs of simulated data files in order to obtain uncertainty quantification on our performance metrics.
-We use uniform priors for all $\bm{m}$ and $\bm{u}$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
+We use uniform priors for the $\bm{m}$, $\bm{u}$, and $\pi$ parameters, with $\alpha_{fl} = \beta_{fl} = 1$ for all $f$ and $l$.
 We run the Gibbs sampler for 1000 iterations, and discard the first 100 as burn-in.
-We calculate Bayes estimates $\hat{\bm{Z}}$ of the linkage structure using the  loss function and post-processing procedure described in Appendix \ref{bayes-estimate}.
-Traceplots for parameters of interest for one example simulation are provided in Appendix \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
-We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; those results are in Appendix \ref{partial}.
-
-	\begin{table}[t]
-		\centering
-		\begin{tabular}[t]{llllll}
-
-			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
-			\cline{3-6}
-			Fields & Similarity & 1 & 2 & 3 & 4\\
-			\hline
-			First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (.5, .1]\\
-			Age and Occupation & Binary & Agree & Disagree &  & \\
-			\hline
-		\end{tabular}
-		\caption{Construction of comparison vectors for accuracy study with simulated data files of Section \ref{accuracy}.}
-		\label{Tab:sadinle_simulation_cutoffs}
-	\end{table}
+We calculate Bayes estimates $\hat{\bm{Z}}$ of the linkage structure using the  loss function and post-processing procedure described in Supplement \ref{bayes-estimate}.
+Traceplots for parameters of interest for one example simulation are provided in Supplement \ref{app:appendix-sim}; they show no obvious concern over MCMC convergence.
+We also replicate this simulation allowing \texttt{fabl} to leave some components of the linkage structure undetermined and left for clerical review; those results are in Supplement \ref{partial}.
 
@@
 In Figure \ref{fig:sadinle_simulation}, we see that the two methods have comparable performance at all levels of error and overlap.
-In the specific case of high error and low overlap, widely regarded as the most difficult linkage scenario, we see that \texttt{fabl} performs slightly worse than \texttt{BRL} on average; however, the overall accuracy level remains high.
-
-	\begin{figure}[t]
+
+		\begin{table}
+		\centering
+		\begin{tabular}{llllll}
+
+			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
+			\cline{3-6}
+			Fields & Similarity & 1 & 2 & 3 & 4\\
+			\hline
+			First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (.5, .1]\\
+			Age and Occupation & Binary & Agree & Disagree &  & \\
+			\hline
+		\end{tabular}
+		\caption{Construction of comparison vectors for accuracy study with simulated data files of Section \ref{accuracy}.}
+		\label{Tab:sadinle_simulation_cutoffs}
+	\end{table}
+
+	\begin{figure}
 		\begin{center}
-%		\includegraphics[width=0.7\textwidth]{finalFigures/figures/sadinle_sim_plot2} 
-			\includegraphics[width=0.7\textwidth]{../notes/figures/sadinle_sim_plot2} 
-			\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}.
+			\includegraphics[width=0.99\textwidth]{../notes/figures/sadinle_sim_plot3} 
+			\caption{Posterior means and credible intervals for accuracy metrics under the replication of the simulation study from \cite{sadinle_bayesian_2017}.
 For each level of overlap and each level of error, we have 100 paired sets of 500 records.
@@
 
-Finally, our last simulation demonstrates our method's robustness to different values of $S$ for the SEI memory reduction procedure.
-We perform record linkage on one set of synthetic datafile described in Section \ref{accuracy} with 500 records in each datafile, 250 entities in common across datafiles, and 3 errors present across matching records.
-To achieve more drastic results, we perform SEI without chunking the data, that is, $t_A = t_B = 1$.
-In practice, if it is possible to create and store the comparison matrix for all record pairs at one time, there is no need to reduce the memory of the hashed matrix through SEI.
-For illustration however, it is easier to illustrate the effects of choices of $S$ in this setting.
-
-We preform linkage using SEI with $S = (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler.
-As any particular SEI implementation may improve or worsen linkage performance; if the SEI procedure happens to only remove pairs that are not matches, recall and precision will improve.
+Finally, our last simulation demonstrates the robustness of \texttt{fabl} to different values of $S$ for the SEI memory reduction procedure.
+We perform record linkage on one set of the synthetic data files described in Section \ref{accuracy} with 500 records in each data file, 250 entities in common across data files, and 3 errors present across matching records.
+We perform SEI without batching the data, that is, $t_A = t_B = 1$.
+In practice, when it is computationally feasible to create and store $\gamma$ without batching, there is no need to reduce the memory of the hashed matrix through SEI.
+For illustration, however, it is easier to examine the effects of choices of $S$ in this setting.
+
+We perform linkage using SEI with $S \in (1, 2, 5, 10, 20)$, and without using SEI, always with 500 iterations of the Gibbs sampler.
+Any particular SEI implementation may improve or worsen linkage performance; if the SEI procedure happens to only remove pairs that are not matches, recall and precision will improve.
 Therefore, we perform linkage under each setting 100 times, recording the linkage estimate $\hat{\bm{Z}}$, and recall and precision.
@@
 Setting $S=1$ can arbitrarily remove the index of a true match, leading the Gibbs sampler to concentrate probability on a false match, while larger values of $S$ produce results mirroring implementation with no SEI.
-We note however that even with $S=1$, the loss in precision is small.
+We note, however, that even with $S=1$, the loss in precision is small in these simulations.
 
@@
 		\begin{center}
-%		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/Z_plot}
 		\includegraphics[width=\linewidth]{../SEI_sensitivity/Z_plot}
-		\caption{Distinct values of $\hat{Z}$ in Section \ref{SEI-sensitivity} simulation.}
+		\caption{Distinct values of $\hat{Z}$ in the simulations of Section \ref{SEI-sensitivity}.}
 		\label{fig:SEI_Z}
@@
 		\begin{center}
-		\includegraphics[width=\linewidth]{finalFigures/SEI_sensitivity/eval_plot} 
-%	\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
-		\caption{Means and 95\% credible intervals for recall and precision in Section \ref{SEI-sensitivity} simulation.}
+		\includegraphics[width=\linewidth]{../SEI_sensitivity/eval_plot} 
+		\caption{Means and 95\% credible intervals for recall and precision in the simulations of Section \ref{SEI-sensitivity}.}
 		\label{fig:SEI_eval}
@@
 	The country of El Salvador was immersed in civil war from 1980 to 1991.
-We are interested in estimating the total number of casualties from the war.
-We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL)\citep{howland2008rescate} and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish)\citep{ball2000salvadoran}.\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization.
-The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
-When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
+We are interested in estimating the total number of individuals killed in the war.
+We utilize lists of documented deaths from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises killings reported directly to the organization \citep{howland2008rescate, ball2000salvadoran, green2019civilian}.
+The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CDHES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
+When estimating the total number of individuals killed, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists.
 Instead, record linkage techniques must be used to merge data files before analyzing the data \citep{lum2013applications}.
@@
 There are several challenges with these data.
-First, both data files have been automatically digitized, which inherently leads to some degree of typographical error.
-Second, the only fields recorded are given name, last name, date of death, and place of death.
-\textcolor{teal}{While it may be common in other data sets for a parent and child to share the same given name, in this data set, it is more common for cousins to share this information, especially in villages, where there may be only four to five extended families.
-In Latin America, an individual receives a surname from their father and their mother, leading to a limited number of surnames in small villages.}
-	%% I believe that Patrick said that this is not common in this data set
-%	It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
-
-	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in $n_A = 4420$ records in CHDES and $n_B = 1323$ records in ERTL.
+First, the ERTL data file was automatically digitized, which inherently leads to some degree of typographical error.
+Second, it is common for villages in El Salvador to consist of only four to five extended families.
+This means there are a small number of last names in use, so many individuals have the same first and last name.
+Since the only fields recorded are given name, last name, date of death, and place of death, this leads to several instances in which distinct individuals have identical records.
+These individuals may be distant cousins, or are perhaps entirely unrelated, but would pose challenges for any record linkage method.
+
+	Following \cite{sadinle_bayesian_2017}, we utilize records that have non-missing entries for given and last name, which results in $n_A = 4420$ records in CDHES and $n_B = 1323$ records in ERTL.
 We standardize names to account for common misspellings and use a modified Levenstein distance when comparing names to account for the fact that second names are often omitted in Spanish.
-Place of death is recorded by municipality and department within that municipality; however, since department is missing in 95\% of records in CHDES and 80\% of records in ERTL, we exclude department from our analysis.
+Place of death is recorded by municipality and department within that municipality; however, since department is missing in 95\% of records in CDHES and 80\% of records in ERTL, we exclude department from our analysis.
 Thus, we conduct record linkage using given name, last name, municipality, and day, month, and year of death.
-We use uniform priors for the $\bm{m}$ and $\bm{u}$ parameters.
-
-	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
-This results in $5^5 \times 3 = 6025$ possible agreement patterns, with 1173 patterns realized in the data.
-However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
-
-	\begin{table}[t]
+We use uniform priors for the $\bm{m}$, $\bm{u}$, and $\pi$ parameters.
+
+	\begin{table}[!t]
 		\begin{tabular}[t]{llllll}
@@
 		\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \cite{sadinle_bayesian_2017}.
-This setup leads to 2048 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
-	\end{table}
-
-	\begin{table}[t]
+This setup leads to 1875 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_1}
+
 		\centering
@@
 		\caption{Construction of comparison vectors for El Salvador data for increased speed under \texttt{fabl}.
-This setup leads to 216 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
+This setup leads to 432 possible agreement patterns in total.}\label{Tab:el_salvador_cutoffs_2}
 	\end{table}
 
+	\begin{figure}[p]
+	\begin{center}
+		\includegraphics[width=0.99\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
+		\caption{Posterior distribution and Bayes estimate of overlap across the two files in the El Salvador case study. We note they are quite similar under both methods.}
+		\label{fig:overlap-plot}
+	\end{center}
+
+	\begin{center}
+		\includegraphics[width=0.99\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
+		\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
+		\label{fig:m-and-u}
+	\end{center}
+\end{figure}
+
+	We initially followed the comparison vector constructions set by \cite{sadinle_bayesian_2017}, using four levels of agreement for each field, according to the thresholds provided in Table \ref{Tab:el_salvador_cutoffs_1}.
+This results in 1875 possible agreement patterns, with 1173 patterns realized in the data.
+However, we noticed that the posterior distributions of several levels of the $\bm{m}$ and $\bm{u}$ parameters were nearly identical in an initial run of \texttt{BRL}, suggesting that these levels were unnecessary.
+
 	Therefore, we perform our analysis with the agreement levels for each field according to Table \ref{Tab:el_salvador_cutoffs_2}.
-Among the 216 possible agreement patterns, 159 are realized in the data.
-With this revised comparison specification, \texttt{fabl} runs in 61 seconds, approximately 4 times faster than the \texttt{BRL} run time of 239 seconds.
+Among the 432 possible agreement patterns, 159 are realized in the data.
+With this revised comparison specification, \texttt{fabl} runs in 109 seconds, approximately 3 times faster than the \texttt{BRL} run time of 313 seconds.
 The estimates of the $\bm{m}$ parameters under each method are similar, as shown in Figure \ref{fig:m-and-u}.
-Estimates of $\bm{u}$ are indistinguishable, and thus omitted.
-Traceplots for parameters of interest are provided in Appendix \ref{app:appendix-es}.
-
-	For completeness, we note that linkage with the more detailed comparison vectors requires 240 seconds for \texttt{BRL}, and 261 seconds for \texttt{fabl}.
+Estimates of the $\bm{u}$ parameters are indistinguishable, and thus omitted.
+Traceplots for parameters of interest are provided in Supplement \ref{app:appendix-es}.
+
+	For completeness, we note that linkage with the more detailed comparison vectors requires 945 seconds for \texttt{BRL}, and 1093 seconds for \texttt{fabl}.
 Apparently, the number of patterns is sufficiently many that the computational savings from \texttt{fabl} does not overcome the inherent speed differences of C as opposed to \texttt{R}.
 
-	Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both data files.
+	Through \texttt{fabl}, we arrive at a Bayes estimate of 178 individuals recorded in both data files.
 We calculate posterior samples of the size of the overlap across files by finding the number of links in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching.
-The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate identifies fewer matches than the Gibbs sampler identifies on average.
+The posterior 95\% credible interval for the overlap across files is (205, 236), indicating that the Bayes estimate identifies fewer matches than the Gibbs sampler identifies on average.
 This is because a large number of records in ERTL have multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair as a match in the Bayes estimate.
@@
 
-	\begin{figure}[t]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_distribution_smallP_bayes}
-	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}
-			\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
-			\label{fig:overlap-plot}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[t]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_posterior_smallP} 
-                  	\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
-			\caption{Posterior estimates of $\bm{m}$ parameters with 95\% credible intervals for the El Salvador case study. They are quite similar across the two methods.}\label{fig:m-and-u}
-			\label{fig:m-and-u}
-		\end{center}
-	\end{figure}
-
 	\subsection{National Long Term Care Study}
 	\label{nltcs}
-
-	\begin{table}[t]
-		\centering
-		\begin{tabular}[t]{lllll}
-			\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
-			\cline{3-5}
-			Fields & Similarity & 1 & 2 & 3\\
-			\hline
-			Sex & Binary & Agree & Disagree & \\
-			Year of Birth & Binary & Agree & Disagree & \\
-			Month of Birth & Binary & Agree & Disagree & \\
-			Day of Birth & Binary & Agree & Disagree & \\
-			Location & Custom & Same State and Office & Same State & Otherwise \\
-			\hline
-		\end{tabular}
-		\caption{Construction of comparison vectors for NTLCS data.}\label{Tab:nltcs-comparisons}
-	\end{table}
 
@@
 
-	We link records using sex, date of birth, and location using the thresholds shown in Table \ref{Tab:nltcs-comparisons}.
-Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx 400,000,000$ record pairs would require approximately 8GB of memory.
+	\begin{table}
+		\centering
+		\begin{tabular}[t]{lllll}
+			\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
+			\cline{3-5}
+			Fields & Similarity & 1 & 2 & 3\\
+			\hline
+			Sex & Binary & Agree & Disagree & \\
+			Year of Birth & Binary & Agree & Disagree & \\
+			Month of Birth & Binary & Agree & Disagree & \\
+			Day of Birth & Binary & Agree & Disagree & \\
+			Location & Custom & Same State and Office & Same State & Otherwise \\
+			\hline
+		\end{tabular}
+		\caption{Construction of comparison vectors for NTLCS data.}\label{Tab:nltcs-comparisons}
+	\end{table}
+
+	We link records using sex, location, and day, month, and year of birth using the criteria shown in Table \ref{Tab:nltcs-comparisons}.
+Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx$ 400 million record pairs would require approximately 8GB of memory.
 Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
-{\color{red}{However, through the method described in Section \ref{scaling},  we perform 30 smaller comparison tasks, using $t_A = 1$ and $t_B = 30$.
+However, through the method described in Section \ref{scaling},  we perform 30 smaller comparison tasks, using $t_A = 1$ and $t_B = 30$.
 We conduct linkage with all record indices recorded and also with SEI using $S=10$, and obtain identical results.
 The hashed $\tilde{\gamma}$ without SEI is about 2.2 GB, and with SEI, it is about 760 MB.
-Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.}}
+Constructing the comparisons sequentially took approximately 40 minutes, which could be reduced considerably through parallel computing.
 
 	We run a Gibbs sampler for 1000 iterations, taking about 235 seconds.
-As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure consists of 9634 matches, with a 95\% credible interval of (9581, 9740).
+Traceplots do not suggest convergence issues, and are similar to those seen in Supplement \ref{app:appendix-sim} and \ref{app:appendix-es}.
+As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure has of 9634 matches, with a 95\% credible interval of (9581, 9740).
 Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
-Traceplots do not suggest convergence issues, and are similar to those seen in Appendix \ref{app:appendix-sim} and \ref{app:appendix-es}
-
-	\begin{figure}[h]
+
+	\begin{figure}
 		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/nltcs/overlap_posterior4}
 			\includegraphics[width=0.6\textwidth]{../notes/figures/nltcs/overlap_posterior4}
@@
 
-\begin{supplement}
-\stitle{Supplementary Material of ``Efficient and Scalable Bipartite Matching with
-Fast Beta Linkage (fabl)." See the .pdf document}
-\sdescription{}
-\end{supplement}
-
 	\clearpage
@@
 	\bibliography{fabl}
-
-%\appendix
-%	
-%		
-%	
-%	\hypertarget{app:derivations}{%
-%	\section{Derivations of Full Conditionals}\label{app:derivations}}
-%\textcolor{red}{Note: Please use s, s+1 consistently throughout.
-Or surpress the notation and tell the reader.
-It looks weird to go back and forth.
-I've tried to provide an intermediate solution.
-Perhaps Jerry can state if he would think this is okay as an Editor.}\\
-%
-%\textcolor{teal}{We provide detailed derivations of the full-conditionals provided in Section~\ref{sec:fast-beta-linkage}.
-In our derivations below, we assume that the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the final update of the full conditional that appears in Section~\ref{sec:fast-beta-linkage}.}
-%
-%The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{fl}$, we have
-%\begin{align*}
-%	\mathcal{L}(m_{fl}|\gamma, \alpha_{fl}, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
-%\end{align*}
-%where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
-Analogous procedures lead to the posterior distribution $\mathcal{L}(u_{fl}| \gamma, \beta_{fl}, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
-Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
-%
-%$$\bm{m}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})),$$
-%$$\bm{u}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).$$
-%
-%Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ and hyperparameters.
-We have the full conditional
-%\begin{align*}
-%	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
-%	&\propto \pi^{n_{AB}(\bm{Z})} (1-\pi)^{n_B - n_{AB}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
-%	&\propto \pi^{n_{AB}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_A - n_{AB}(\bm{Z}) + \beta_{\pi} -1}.
-%\end{align*}
-%Thus, $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
-%
-%To express the full conditional for $\bm{Z}$, we consider the cases where $Z_j$ does not have a link in $A$ and where $Z_j$ does have a link in $A$.
-Because we sample each $Z_j$ independently of all other $Z_{j'}$ (for $j \neq j'$), we need only the full conditional for each individual $Z_j$.
-{\color{red}{Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.}} Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
-%\begin{align}
-%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
-%\end{align}
-%When $Z_j = q$ for some $q\leq n_A$, we have
-%\begin{align}
-%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
-%\end{align}
-%We multiply and divide by the $u$ parameters for the matching record pair to obtain
-%\begin{align}
-%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &\propto \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-%	&= w_{qj}  c_j.
-%\end{align}
-%We can divide the result of each case by $c_j$ to get
-%\begin{align}
-%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
-%		w_{qj}, & q \leq n_A; \\
-%		1, &  q  = n_A + j. \
-%	\end{cases}
-%\end{align}
-%Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
-%\begin{align}
-%	\label{eqn:z_full_conditional2}
-%	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
-%	\begin{cases} 
-%		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
-%		1 - \pi^{(s+1)}, & q  = n_A + j. \
-%	\end{cases}
-%\end{align}
-%
-%\textcolor{teal}{Observe that the full conditional in (\ref{eqn:z_full_conditional2}) does not depend on $\bm{Z}^{(s)},$ implying that 
-%\begin{align}
-%p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi) \propto \prod_q \left(\prod_j
-%p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi \right)
-%\right).
-%\end{align}
-%Therefore, we update the full conditional for $\bm{Z}^{(s+1)}$ in parallel.
-This contrasts the Gibbs sampler of  \cite{sadinle_bayesian_2017}, where the update to the matching matrix is sequential and accomplished using a partially collapsed Gibbs sampler, where one integrates out $\pi.$ We do not opt for this approach as it would lead to similar sequential updates, which are computationally slow.}
-%
-%
-%
-%\hypertarget{bayes-estimate}{%
-%		\section{Bayes Estimate}
-%		\label{bayes-estimate}}
-%	
-%We calculate a Bayes estimate $\hat{\bm{Z}}$ for the linkage parameter $\bm{Z}$ by assigning different positive losses to different types of errors, and minimizing posterior expected loss.
-We adopt the loss function proposed in \cite{sadinle_bayesian_2017} in which $\hat{Z}_j \in \{1, \ldots, n_A, n_A + j, R\}$, with $R$ representing the option to leave the matching undetermined by the model.
-Specifically, we have
-%	\[L(\hat{Z_j}, Z_j)=\begin{cases} 
-%		0,  & \text{if } Z_j = \hat{Z_j}; \\
-%		\theta_R,  & \text{if } \hat{Z_j} = R; \\
-%		\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_A + j ; \\
-%		\theta_{01},  & \text{if } Z_j = n_A + j,\hat{Z_j} \leq n_A ; \\
-%		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
-%	\end{cases}\] 
-%Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
-%
-%In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
-%	$$\hat{Z}_j =\begin{cases} 
-%		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
-%		0,  & \text{otherwise}. \\
-%	\end{cases}$$
-%Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
-In the event that we have two records $B_j$ and $B_{j'}$ such that both $p(\hat{Z}_j = i |\gamma) > \frac{1}{2}$ and $ p(\hat{Z}_{j'} = i |\gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match.
-Since each $Z_j$ is independent, this is equivalent to minimizing the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.
-%
-%When we seek a partial estimate of the linkage structure, leaving a portion of record pairs to be classified manually in clerical review, we adopt losses $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, .1)$.
-For a more in-depth explanation of this function and the induced Bayes estimate, see \cite{sadinle_bayesian_2017}.
-%
-%
-%	\hypertarget{partial}{%
-%	\section{Accuracy under Partial Estimates}\label{partial}}
-%
-%In this section, we repeat the simulation study in Section \ref{accuracy} of the main text, allowing for clerical review rather than forcing all records to have or not have links.
-Specifically, by leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
-In this context, we are no longer focused on finding all true matches, but rather protecting against false matches.
-Thus, instead of recall, we use the negative predictive value (NPV), defined as the proportion of non-links that are actual nonmatches.
-Mathematically, $\text{NPV} = \sum_{j=1}^{n_B} I(\hat{Z}_j = Z_j = n_A + j)$/$\sum_{j=1}^{n_B} I(\hat{Z}_j = n_A + j)$.
-We continue to use the precision, which is renamed the positive predictive value (PPV) in this context.
-Lastly, we also examine the rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_B} I(\hat{Z}_j = R)/n_B$.
-To convey this information alongside NPV and PPV, for which values close to 1 indicate strong performance, we report the decision rate (DR), defined as DR = $1 - RR$.
-%
-%In Figure \ref{fig:sadinle_simulation_partial}, we see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings.
-However, with high amounts of error, and thus fewer accurate and discerning fields of information, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV.
-Since \texttt{fabl} does not remove previously matched records from consideration for a new record, posterior probabilities of matches at times can be split across more records; in contrast, \texttt{BRL} is able to maintain higher confidence in matches in this setting.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand than would be left under \texttt{BRL}, but the decisions made by each method should have nearly equal accuracy.
-%
-%%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-%
-%%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
-%
-%\begin{figure}[t]
-%	\begin{center}
-%%	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
-%		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
-We see poorer performance for \texttt{fabl} only in situations with high overlap.}
-%		\label{fig:sadinle_simulation_partial}
-%	\end{center}
-%\end{figure}
-%	
-%	
-%	\hypertarget{appendix-sim}{%
-%		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
-%	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
-It is set up with one error across the linkage fields and 50 duplicates across files.
-Traceplots across other settings exhibit similar behavior.
-Note that traceplots for $\bm{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-%			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
-%		\end{center}
-%	\end{figure}
-%	
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%%		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-%			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
-%		\end{center}
-%	\end{figure}
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-%			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
-%		\end{center}
-%	\end{figure}
-%
-%\clearpage
-%
-%	\hypertarget{appendix-speed}{%
-%	\section{Additional Speed Simulation Study}\label{app:appendix-speed}}
-%	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
-We replicate the speed study of Section \ref{speed} under different settings.
-Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.
-The $\bm{m}$ and $\bm{u}$ parameters for this simulation are shown Table \ref{Tab:distributions_2}.
-%	
-%		\begin{table}[t]
-%		\centering
-%		\begin{tabular}{llll|lll}
-%			\multicolumn{1}{c}{ } & \multicolumn{3}{c}{$\bm{m}$} & \multicolumn{3}{c}{$\bm{u}$} \\
-%			\cline{2-4} \cline{5-7}
-%			& Agree & Partial & Disagree & Agree & Partial & Disagree \\
-%			\hline
-%			Feature 1 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-%			Feature 2 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-%			Feature 3 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-%			Feature 4 & $\frac{9}{10}$ & $\frac{9}{100}$  & $\frac{1}{100}$ & $\frac{1}{100}$ &  $\frac{3}{100}$ & $\frac{96}{100}$ \\ 
-%			\hline
-%		\end{tabular}
-%		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
-%	\end{table}
-%	
-%	\begin{figure}[h!]
-%		\begin{center} 
-%%		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
-%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
-%		\end{center}
-%	\end{figure}
-%	
-%	\begin{figure}[h!]
-%		\begin{center}
-%%		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
-%		 \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
-%			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
-We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
-%		\end{center}
-%	\end{figure}
-%
-%\clearpage
-%	
-%	\hypertarget{appendix-es}{%
-%		\section{Traceplots for El Salvador Case Study}\label{app:appendix-es}}
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-%			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
-%		\end{center}
-%	\end{figure}
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
-%			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-%			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} 
-%			\label{fig:m_trace}
-%		\end{center}
-%	\end{figure}
-%	
-%	\begin{figure}[!h]
-%		\begin{center}
-%%		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
-%		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-%			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} 
-%			\label{fig:u_trace}
-%		\end{center}
-%	\end{figure}
 \end{document}

### 2023-07-30 ba_resubmit_template_files/supplement.tex
create new directory for main and supplement.
fixed rendering.
working on main and supp.

--- 2023-07-30 ba_resubmit/kundinger_supplement_final.tex
@@
-\documentclass[ba]{imsart}
+%\documentclass[ba]{imsart}
+%% use command below for supplement
+ \documentclass[ba,preprint]{imsart}
 %
-\pubyear{0000}
-\volume{00}
-\issue{0}
-\doi{0000}
+\pubyear{2021}
+\volume{TBA}
+\issue{TBA}
+% \doi{0000}
 %\arxiv{}
@@
 \renewcommand{\appendixname}{Supplement}
-\renewcommand{\thefigure}{SI\arabic{figure}} \setcounter{figure}{0}
-\renewcommand{\thetable}{SI\arabic{table}} \setcounter{table}{0}
-\renewcommand{\theequation}{SI\arabic{table}} \setcounter{equation}{0}
 %%%
+
+%%% start code for cross referencing main document with the appendix %%
+
+\usepackage{xr} %%(must be loaded before hyperref) and I found an example of this on Overleaf that was extremely helpful
+
+\makeatletter
+
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+\typeout{(#1)}% latexmk will find this if $recorder=0
+% however, in that case, it will ignore #1 if it is a .aux or 
+% .pdf file etc and it exists! If it doesn't exist, it will appear 
+% in the list of dependents regardless)
+%
+% Write the following if you want it to appear in \listfiles 
+% --- although not really necessary and latexmk doesn't use this
+%
+\@addtofilelist{#1}
+%
+% latexmk will find this message if #1 doesn't exist (yet)
+\IfFileExists{#1}{}{\typeout{No file #1.}}
+}\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+\externaldocument{#1}%
+\addFileDependency{#1.tex}%
+\addFileDependency{#1.aux}%
+}
+%------------End of helper code--------------
+
+% put all the external documents here!
+\myexternaldocument{main}
+%% To run this you must run the compiler through pdflatexmk and have the latexmkrc file within the folder
+
+%%% end code for cross referencing main document with the appendix %%
 
@@
 	\hypertarget{app:derivations}{%
-	\section{Derivations of Full Conditionals}
-	\label{app:derivations}}
-%\textcolor{red}{Note: Please use s, s+1 consistently throughout.
-Or surpress the notation and tell the reader.
-It looks weird to go back and forth.
-I've tried to provide an intermediate solution.
-Perhaps Jerry can state if he would think this is okay as an Editor.}\\
+		\section{Derivations of Full Conditionals}\label{app:derivations}}
 
-\textcolor{teal}{We provide detailed derivations of the full-conditionals provided in Section~\ref{sec:fast-beta-linkage}.
-In our derivations below, we assume that the new update to the Gibbs sampler is $s+1$ and the current state of the sampler is $s.$ As to not overload notation, we suppress the state of the sampler until presenting the final update of the full conditional that appears in Section~\ref{sec:fast-beta-linkage}.}
+	We provide detailed derivations of the full-conditionals provided in Section \ref{sec:fast-beta-linkage}.
+The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
+For a particular $m_{fl}$, we have
+	\begin{align}
+		\mathcal{L}(m_{fl}|\gamma, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
+	\end{align}
+	where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
+Analogous procedures lead to $\mathcal{L}(u_{fl}| \gamma, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i=1}^{n_A}  \sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
+Thus, for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
+	\begin{align}
+	\bm{m}_f^{(s+1)} |
+	\gamma, 
+	\bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim 
+	\text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \\
+	\bm{u}_f^{(s+1)} | \gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).
+	\end{align}
 
-The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions. For a particular $m_{fl}$, we have
-\begin{align*}
-	\mathcal{L}(m_{fl}|\gamma, \alpha_{fl}, \bm{u}, \bm{Z}, \pi) &\propto \prod_{i=1}^{n_A} \prod_{j=1}^{n_B} m_{fl}^{I(Z_j = i) I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}  m_{fl}^{\alpha_{fl} - 1} = m_{fl}^{\alpha_{fl}(\bm{Z}) - 1},
-\end{align*}
-where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j = i)$.
-Analogous procedures lead to the posterior distribution $\mathcal{L}(u_{fl}| \gamma, \beta_{fl}, \bm{m}, \bm{Z}, \pi)$  $\propto u_{fl}^{\beta_{fl}(\bm{Z}) - 1}$, where $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(Z_j \neq i)$.
-Thus for the vectors of parameters $\bm{m}_f$ and $\bm{u}_f$, we have
-
-$$\bm{m}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})),$$
-$$\bm{u}_f^{(s+1)}|\bm{Z}^{(s)}, \gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})).$$
-
-Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j < n_A + j)$ encoded by $\bm{Z}$ and hyperparameters.
+	Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j \leq n_A)$ encoded by $\bm{Z}$ and hyperparameters.
 We have the full conditional
-\begin{align*}
-	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
+	\begin{align}
+		p(\pi |\gamma, \bm{Z}, \bm{m}, \bm{u}) &\propto p(\bm{Z}|\pi)p(\pi) \\
 		&\propto \pi^{n_{AB}(\bm{Z})} (1-\pi)^{n_B - n_{AB}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
 		&\propto \pi^{n_{AB}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_A - n_{AB}(\bm{Z}) + \beta_{\pi} -1}.
-\end{align*}
-Thus, $\pi^{(s+1)}|\bm{Z}^{(s+1)},  \alpha_{\pi}, \beta_{\pi}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
+	\end{align}
+	Thus, $\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}$ has a $\text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi})$ distribution.
 
-To express the full conditional for $\bm{Z}$, we consider the cases where $Z_j$ does not have a link in $A$ and where $Z_j$ does have a link in $A$.
-Because we sample each $Z_j$ independently of all other $Z_{j'}$ (for $j \neq j'$), we need only the full conditional for each individual $Z_j$.
-{\color{red}{Let $\Gamma_{.j}$ denote the matrix of $n_A$ comparison vectors relating to record $B_j$.}} Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+	Due to the independence in the fast beta prior in \eqref{eqn:fast_beta_prior}, we can obtain the full conditional for $\bm{Z}$ through the full conditionals for each individual $Z_j$.
+Let $\Gamma_{.j}$ denote the random matrix of $n_A$ comparison vectors relating to an arbitrary record $B_j$, and let $\gamma_{.j}$ be a realization of $\Gamma_{.j}$.
+We have
+	\begin{align}
+		p(\bm{Z} | \gamma, \bm{m}, \bm{u}, \pi) &= \prod_{j=1}^{n_B} p(Z_j | \gamma_{.j}, \bm{m}, \bm{u}, \pi).
+		%&\propto p(\Gamma| \bm{Z}, \bm{m}, \bm{u}) p(\bm{Z} | \pi) \\
+		%&=\prod_{j=1}^{n_B} p(\Gamma_{.j}| Z_j, \bm{m}, \bm{u}) p(Z_j | \pi) \\
+	\end{align}
+
+Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
 \begin{align}
@@
 		\theta_{11},  & \text{if } Z_j \leq n_A, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
-	\end{cases}\] 
+	\end{cases}
+\end{align}
 Here, $\theta_R$ is the loss from not making a decision on the linkage status, $\theta_{10}$ is the loss from a false nonmatch, $\theta_{01}$ is the loss from a false match, and $\theta_{11}$ is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.
 
-In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
-	$$\hat{Z}_j =\begin{cases} 
+In general, we follow \cite{sadinle_bayesian_2017} and set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the decision rule
+	\begin{align}
+		\hat{Z}_j =\begin{cases} 
 		i,  & \text{if } p(Z_j = i |\gamma) > \frac{1}{2}; \\
 		0,  & \text{otherwise}. \\
-	\end{cases}$$
+	\end{cases}
+\end{align}
 Since \texttt{fabl} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$.
@@
 
+	\hypertarget{appendix-sim}{%
+		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
+	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+It is set up with one error across the linkage fields and 50 duplicates across files.
+Traceplots across other settings exhibit similar behavior.
+Note that traceplots for $\bm{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
+
+	\begin{figure}[h]
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
+			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
+			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
+		\end{center}
+
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
+			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
+		\end{center}
+	\end{figure}
+
+\clearpage
+
 	\hypertarget{partial}{%
@@
 
-In this section, we repeat the simulation study in Section \ref{accuracy} of the main text, allowing for clerical review rather than forcing all records to have or not have links.
+In this section, we repeat the simulation study in Section \ref{accuracy}, allowing for clerical review rather than forcing all records to have or not have links.
 Specifically, by leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
@@
 
-%In practice, it is rare to encounter linkage tasks with 90\% overlap, so we remain confident in \texttt{fabl}'s performance.
-If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeler to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.
-
-%This is reasonable because large numbers of errors blur the distinction between matched and nonmatched pairs, leading to situations where one record in $\bm{X}_2$ has multiple plausible matches in $\bm{X}_1$, so that posterior match probability is split across more records.
-
-\begin{figure}[t]
+\begin{figure}
 	\begin{center}
-%	        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_sim_plot_partial_DR} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
-		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}.
+		\includegraphics[width=0.99\textwidth]{../notes/figures/sadinle_sim_plot_partial2} 
+		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Supplement \ref{partial}.
 We see poorer performance for \texttt{fabl} only in situations with high overlap.}
 		\label{fig:sadinle_simulation_partial}
-	\end{center}
-\end{figure}
-
-	\hypertarget{appendix-sim}{%
-		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
-	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
-It is set up with one error across the linkage fields and 50 duplicates across files.
-Traceplots across other settings exhibit similar behavior.
-Note that traceplots for $\bm{u}$ parameters show very little variation because the overwhelming majority of record pairs are nonmatching.
-
-	\begin{figure}[!h]
-		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_overlap_trace} 
-			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_m_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/figures/sim_u_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
 	\end{center}
@@
 Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.
-The $\bm{m}$ and $\bm{u}$ parameters for this simulation are shown Table \ref{Tab:distributions_2}.
+The $\bm{m}$ and $\bm{u}$ parameters for this simulation are shown in Table \ref{Tab:distributions_2}.
 
-		\begin{table}[t]
+		\begin{table}[h]
 		\centering
@@
 		\end{tabular}
-		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Appendix \ref{app:appendix-speed}.}\label{Tab:distributions_2}
+		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Supplement \ref{app:appendix-speed}.}\label{Tab:distributions_2}
 	\end{table}
 
-	\begin{figure}[h!]
-		\begin{center} 
-%		\includegraphics[width=0.6\textwidth]{finalFigures/figures/sadinle_speed_plot_slides2} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, for increasing values of both $n_A$ and $n_B$.
-We see near quadratic growth in run-time for \texttt{BRL}, and near linear growth for \texttt{fabl}.}\label{fig:app-speed1}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\begin{center}
-%		 \includegraphics[width=0.6\textwidth]{finalFigures/figures/speed_plot_fixed_nB_slides2} 
-		 \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides2} 
-			\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulation study in Appendix \ref{app:appendix-speed}, including hashing step for \texttt{fabl}, with increasing $n_A$ and $n_B$ fixed at 500.
-We see linear growth in run-time for \texttt{BRL}, and near constant run-time for \texttt{fabl}.}\label{fig:app-speed2}
-		\end{center}
+	\begin{figure}[h]
+	\centering
+	\begin{subfigure}{.5\textwidth}
+		\centering
+		\includegraphics[width=0.95\textwidth]{../notes/figures/sadinle_speed_plot2_new}
+		\caption{}
+		\label{fig:app-speed1}
+	\end{subfigure}%
+	\begin{subfigure}{.5\textwidth}
+		\centering
+		\includegraphics[width=0.95\textwidth]{../notes/figures/speed_plot_fixed_nB2_new} 
+		\caption{}
+		\label{fig:app-speed2}
+	\end{subfigure}
+	\caption{Run-time for \texttt{BRL} and \texttt{fabl} to run 1000 Gibbs iterations in simulations described in Supplement \ref{app:appendix-speed}.
+In (\ref{fig:app-speed1}), both $n_A$ and $n_B$ are increasing.
+We see quadratic growth in \texttt{BRL} and linear growth in \texttt{fabl}.
+In (\ref{fig:app-speed2}), only $n_A$ only is increasing.
+We see linear growth in \texttt{BRL} and approximately constant run-time in \texttt{fabl}.}
+	\label{fig:app-speed_sims}
 \end{figure}
@@
 
-	\begin{figure}[!h]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/overlap_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
-			\caption{Traceplot for number of matches found across data files in El Salvador case study.} \label{fig:overlap_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		         \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/m_trace} 
-			\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
-			\caption{Traceplot for $\bm{m}$ parameter in El Salvador case study.} 
+    \begin{figure}[h!]
+	\centering
+	\begin{subfigure}[b]{0.475\textwidth}
+		\centering
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/overlap_trace} 
+		\caption{Traceplot for overlap between files.}    
+		\label{fig:overlap_trace}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}  
+		\centering 
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/pi_trace} 
+		\caption{Traceplot for $\pi$.}  
+		\label{fig:pi_trace}
+	\end{subfigure}
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/m_trace} 
+		\caption{Traceplot for $\bm{m}$ parameters.}   
 		\label{fig:m_trace}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\begin{center}
-%		        \includegraphics[width=0.6\textwidth]{finalFigures/el_salvador/u_trace} 
-		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
-			\caption{Traceplot for $\bm{u}$ parameter in El Salvador case study.} 
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=0.975\textwidth]{../notes/figures/el_salvador/u_trace} 
+		\caption{Traceplot for $\bm{u}$ parameters.}
 		\label{fig:u_trace}
-		\end{center}
+	\end{subfigure}
+	\caption{Traceplots for parameters of interest in El Salvador case study in Section \ref{el_salvador}.}
+	\label{fig:el_salvador_trace}
 \end{figure}
@@
 
+\end{document}
+
 	\bibliographystyle{jasa}
-	\bibliography{biblio}
+	\bibliography{fabl}

### 2023-07-30 ba_resubmit_template_files/main.tex
adding in supp to main file and checking for typos.
fix typos using antidote 11.

@@
 This sparse matrix representation can become cumbersome for large linkage tasks.
-More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
+More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
 \begin{align}
@@
 To obtain a Bayes estimate that fulfills the bipartite requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$.
-See Supplement \ref{bayes-estimate} for details redarding the initial Bayes estimate and this post-processing procedure.
+See Supplement \ref{bayes-estimate} for details regarding the initial Bayes estimate and this post-processing procedure.
 
@@
 There are then trivial computational costs associated with synthesizing summary statistics across these partitions.
-We note that this contribution to computational complexity applies for all versions of \cite{fellegi_theory_1969} algorithms, unless they use blocking to reduce the comparison space.
+We note that this contribution to computational complexity applies for all versions of \cite{fellegi_theory_1969} algorithms unless they use blocking to reduce the comparison space.
 
@@
 Since $P << n_A$ in most applications, we have reduced the complexity of sampling $\bm{Z}$ from $O(F n_A n_B)$ under \texttt{BRL} to $O(n_B P)$ under \texttt{fabl}.
-With parallelization, this complexity is further reduced to $O(n_B P /C)$, and so the entire Gibbs sampler has complexity $O(T n_B P / C)$
+With parallelization, this complexity is further reduced to $O(n_B P /C)$, and so the entire Gibbs sampler has complexity $O(T n_B P / C).$
 	In summary, the total computational complexity for \texttt{fabl} is $O(F n_A n_B / C) + O(T n_B P / C).$
@@
 \end{subequations}
-These can computed sequentially or in parallel. Summary statistics from each pairwise batch comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
+These can be computed sequentially or in parallel.
+Summary statistics from each pairwise batch comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
 \begin{subequations}
@@
 	In Figure \ref{fig:speed2}, where we fix $n_B = 500$, we see near linear growth for the run-time under \texttt{BRL} as $n_A$ increases, and much more static run-time under \texttt{fabl}.
-The slight increases in run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-To illustrate that these trends are generalizeable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement \ref{app:appendix-speed}.
+The slight increases in the run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
+To illustrate that these trends are generalizable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement \ref{app:appendix-speed}.
 
@@
 We are interested in estimating the total number of individuals killed in the war.
-We utilize lists of documented deaths from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises killings reported directly to the organization \citep{howland2008rescate, ball2000salvadoran, green2019civilian}.
+We utilize lists of documented deaths from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL data set comprises digitized denunciations published throughout the conflict, and the CDHES data set comprises killings reported directly to the organization \citep{howland2008rescate, ball2000salvadoran, green2019civilian}.
 The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CDHES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable.
@@
 Among the 432 possible agreement patterns, 159 are realized in the data.
-With this revised comparison specification, \texttt{fabl} runs in 109 seconds, approximately 3 times faster than the \texttt{BRL} run time of 313 seconds.
+With this revised comparison specification, \texttt{fabl} runs in 109 seconds, approximately 3 times faster than the \texttt{BRL} run-time of 313 seconds.
 The estimates of the $\bm{m}$ parameters under each method are similar, as shown in Figure \ref{fig:m-and-u}.
@@
 		\end{tabular}
-		\caption{Construction of comparison vectors for NTLCS data.}\label{Tab:nltcs-comparisons}
+		\caption{Construction of comparison vectors for NLTCS data.}\label{Tab:nltcs-comparisons}
 	\end{table}
@@
 	We link records using sex, location, and day, month, and year of birth using the criteria shown in Table \ref{Tab:nltcs-comparisons}.
-Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx$ 400 million record pairs would require approximately 8GB of memory.
-Standard settings on a 16GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
+Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx$ 400 million record pairs would require approximately 8 GB of memory.
+Standard settings on a 16 GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.
 However, through the method described in Section \ref{scaling},  we perform 30 smaller comparison tasks, using $t_A = 1$ and $t_B = 30$.

### 2023-07-30 ba_resubmit_template_files/main.tex
fix typos using antidote 11.

@@
 
-In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
+In what follows, Section~\ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.
 Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and describes the loss function used to calculate the Bayes estimate for the bipartite matching.
-Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}.
-Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
-Finally, Section \ref{discussion} summarizes our contributions and highlights areas for further research.
+Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}.
+Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.
+Finally, Section~\ref{discussion} summarizes our contributions and highlights areas for further research.
 
@@
 \cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.
-He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$, and replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\bm{Z}$ that he calls the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
+He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$, and replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\bm{Z}$ that he calls the ``beta distribution for bipartite matching.''  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.
 Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, where $\alpha_{\pi}$ and $\beta_{\pi}$ are known hyperparameters.
@@
 After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing. 
-For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields.
+%% antidote 11 flagged as awkward
+For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or 
+%only 
+pairs that exactly match on a specified number of fields.
 However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical \cite{fellegi_theory_1969} model, leaving the effect of indexing on Bayesian record linkage models to future work.
@@
 For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form SEI$(r_{p_j}^{ab})$.
-We collect these memory reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.
+We collect these memory-reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.
 
 As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
-We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.
+We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
 
@@
 The slight increases in the run-time for \texttt{fabl} are due primarily to the hashing step, which again can be run in parallel for large data.
-To illustrate that these trends are generalizable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement \ref{app:appendix-speed}.
+To illustrate that these trends are generalizable to other specifications of the comparison vectors, we have included the run-time results for an additional simulation study, under different comparison vector settings, in Supplement~\ref{app:appendix-speed}.
 
@@
 Traceplots do not suggest convergence issues, and are similar to those seen in Supplement \ref{app:appendix-sim} and \ref{app:appendix-es}.
-As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure has of 9634 matches, with a 95\% credible interval of (9581, 9740).
+As shown in Figure \ref{fig:nltcs-overlap-plot}, the Bayes estimate of the linkage structure has
+	%has of 
+	9634 matches, with a 95\% credible interval of (9581, 9740).
 Since we have access to the true linkage structure, we can calculate recall to be 0.89 and precision to be 0.98, resulting in an F-measure of 0.94.
@@
 Indeed, constructing the comparison vectors for the NLTCS linkage task involving around 40,000 records in Section \ref{nltcs} took around 40 minutes.
-Due to the quadratic nature of the comparison space, this computation time would grow quickly with the size of the linkage task, and would be infeasibly slow when dealing with millions of records.
+Due to the quadratic nature of the comparison space, this computation time would grow quickly with the size of the linkage task, and would be extremely
+	%infeasibly 
+	slow when dealing with millions of records.
 Although it is common to use deterministic blocking to reduce the comparison space and then apply probabilistic record linkage within each block, issues arise when sizes of blocks vary across the linkage task.

### 2023-07-30 ba_resubmit_template_files/supplement.tex
fix typos using antidote 11.

@@
 
-	We provide detailed derivations of the full-conditionals provided in Section \ref{sec:fast-beta-linkage}.
+	We provide detailed derivations of the full conditionals provided in Section \ref{sec:fast-beta-linkage}.
 The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section \ref{accuracy}.}\label{fig:sim_overlap_trace}
+			\caption{Representative traceplot of overlap between files from simulation study in Section~\ref{accuracy}.}\label{fig:sim_overlap_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameter from simulation study in Section \ref{accuracy}.}\label{fig:sim_m_trace}
+			\caption{Representative traceplot of $\bm{m}$ parameters from simulation study in Section~\ref{accuracy}.}\label{fig:sim_m_trace}
 		\end{center}
@@
 		\end{tabular}
-		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Supplement \ref{app:appendix-speed}.}\label{Tab:distributions_2}
+		\caption{Probabilities used for $\bm{m}$ and $\bm{u}$ distributions in simulation study in Supplement~\ref{app:appendix-speed}.}\label{Tab:distributions_2}
 	\end{table}
@@
 
+%\bibliographystyle{jasa}
+%	\bibliography{fabl}
+
 \end{document}
-
-	\bibliographystyle{jasa}
-	\bibliography{fabl}

### 2023-07-30 ba_resubmit_template_files/main.tex
fixd number

@@
 	We link records using sex, location, and day, month, and year of birth using the criteria shown in Table \ref{Tab:nltcs-comparisons}.
-Storing $\gamma$ constructed through three comparison scores for each of $20485 (17466) \approx$ 400 million record pairs would require approximately 8 GB of memory.
+Storing $\gamma$ constructed through three comparison scores for each of the approximately 400 million record pairs would require around 8 GB of memory.
 Standard settings on a 16 GB personal computer do not allow storage of an object of this size, and thus \texttt{BRL} is unable to perform this linkage task on such a machine.

### 2023-07-31 ba_resubmit_template_files/main.tex
final Jerry edits

@@
 Under bipartite matching, the set of matches across $A$ and $B$ can be represented in two equivalent ways.
-First, we may use a matrix $\Delta \in \{0, 1\}^{n_An_B}$, where
+First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
 \begin{align}
@@
 \begin{align}
-&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
-&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
+&\Gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
+&\Gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
 &\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:lambda_dist}
@@
 \begin{align}
-	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
+	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}),
 \end{align}

### 2023-07-31 ba_resubmit_template_files/supplement.tex
final Jerry edits

@@
 
-	We provide detailed derivations of the full conditionals provided in Section \ref{sec:fast-beta-linkage}.
+	We provide detailed derivations of the full conditionals provided in Section \ref{sec:fast-beta-linkage} in the main text.
 The $\bm{m}$ and $\bm{u}$ parameters are updated through standard multinomial-Dirichlet distributions.
@@
 		\section{Traceplots for Simulation Study}\label{app:appendix-sim}}
-	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy}.
+	Figures \ref{fig:sim_overlap_trace}, \ref{fig:sim_m_trace}, and \ref{fig:sim_u_trace} are traceplots for one of the 900 linkage tasks that comprise the simulation in Section \ref{accuracy} in the main text.
 It is set up with one error across the linkage fields and 50 duplicates across files.
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
-			\caption{Representative traceplot of overlap between files from simulation study in Section~\ref{accuracy}.}\label{fig:sim_overlap_trace}
+			\caption{Representative traceplot of overlap between files from simulation study in Section~\ref{accuracy} in the main text.} \label{fig:sim_overlap_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
-			\caption{Representative traceplot of $\bm{m}$ parameters from simulation study in Section~\ref{accuracy}.}\label{fig:sim_m_trace}
+			\caption{Representative traceplot of $\bm{m}$ parameters from simulation study in Section~\ref{accuracy} in the main text.}\label{fig:sim_m_trace}
 		\end{center}
@@
 			\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
-			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy}.}\label{fig:sim_u_trace}
+			\caption{Representative traceplot of $\bm{u}$ parameters from simulation study in Section \ref{accuracy} in the main text.}\label{fig:sim_u_trace}
 		\end{center}
@@
 
-In this section, we repeat the simulation study in Section \ref{accuracy}, allowing for clerical review rather than forcing all records to have or not have links.
+In this section, we repeat the simulation study in Section \ref{accuracy} in the main text, allowing for clerical review rather than forcing all records to have or not have links.
 Specifically, by leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match.
@@
 	Figures \ref{fig:app-speed1} and \ref{fig:app-speed2} illustrate that different constructions of the comparison vectors lead to similar speed gains.
-We replicate the speed study of Section \ref{speed} under different settings.
+We replicate the speed study of Section \ref{speed} in the main text under different settings.
 Here, we use four fields of comparison, each with three possible levels of agreement, resulting in $3^4 = 81$ possible patterns.
@@
 	\end{subfigure}
-	\caption{Traceplots for parameters of interest in El Salvador case study in Section \ref{el_salvador}.}
+	\caption{Traceplots for parameters of interest in El Salvador case study in Section \ref{el_salvador} in the main text.}
 	\label{fig:el_salvador_trace}

### 2023-07-31 ba_resubmit_template_files/response_to_reviewers3.tex
edits.

--- 2023-07-30 ba_resubmit/response_to_reviewers3.tex
@@
 Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.
-Without loss of generality, denote files such that $n_A \geq n_B$.".
+Without loss of generality, denote files such that $n_A \geq n_B$".
 See Section 2.
@@
 	\reply{
-		%Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
+	\textcolor{teal}{The reviewer is correct that the distinction does not matter between MCAR and MAR. We have included the following explanation of missingness in Section 2.}
+%		Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
 For simplicity, we have chosen to use MCAR in the article.
-		We have included the following explanation of missingness in Section 2.
+%		We have included the following explanation of missingness in Section 2.
 
@@
 	We have revised the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-Refer to Appendix A for this extended derivation.
+Refer to \textcolor{teal}{Supplement}~A for this extended derivation.
 	}
@@
 	\reply{Yes, the comparison vectors are created (and therefore, patterns are assigned) without regard to $\bm{Z}$.
-We have removed that sentence, and updated the equations in the Section 4.3 to avoid confusion.}
+We have removed that sentence, and updated the equations in Section 4.3 to avoid confusion.}
 

### 2023-07-31 ba_resubmit_template_files/response_to_reviewers3.tex
updates to paper and response to reviewers.

@@
 
-\section{AE Comments}
+\section{Editor Comments}
 
-\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}
+\pointRaised{E}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}
 
@@
 
-\pointRaised{AE}{Line 80 on p.
+\pointRaised{E}{Line 80 on p.
 3 says: “For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say “record $i \in X_1$” rather than the more compact $x_{1i}$.” Taking this at face value implies that, in any sentence, one could swap “record $i \in X_1$” for “$x_{1i}$,” which clearly is not true.
@@
 	\reply{
-	\textcolor{teal}{The reviewer is correct that the distinction does not matter between MCAR and MAR. We have included the following explanation of missingness in Section 2.}
+	The reviewer is correct that the distinction does not matter between MCAR and MAR. We have included the following explanation of missingness in Section 2.
 %		Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
@@
 	We have revised the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
-Refer to \textcolor{teal}{Supplement}~A for this extended derivation.
+Refer to Supplement~A for this extended derivation.
 	}
@@
 \reply{%
-	This has been revised; thank you. See Section 4.2,
+	This has been revised; thank you. See Section 4.2.
 }

### 2023-07-31 ba_resubmit_template_files/main.tex
fix name on page one.
updates to paper and response to reviewers.

@@
 
-	\subsection{Civilian Casualties from the El Salvadoran Civil War}
+	\subsection{Civilian Killings from the El Salvadoran Civil War}
 	\label{el_salvador}
@@
 \stitle{Supplementary Material of ``Efficient and Scalable Bipartite Matching with
-Fast Beta Linkage (fabl)." See the .pdf document}
+Fast Beta Linkage (fabl).'' See the .pdf document}
 \sdescription{}

### 2023-07-31 ba_resubmit_template_files/response_to_reviewers3.tex
continue pass.

@@
 	\reply{
-	The reviewer is correct that the distinction does not matter between MCAR and MAR. We have included the following explanation of missingness in Section 2.
+	%% Brian: I think it is important to tell the AE he is correct above and we are happy to add a note in later if he/she wants this. In this way, we can close out this question and avoid future ones. 
+	\textcolor{teal}{The reviewer is correct that the distinction does not matter between MCAR and MAR; we opt for using MCAR.
+We have included the following explanation of missingness in Section 2.
+We are happy to add a clarification regarding this point if the AE believes it would be helpful for the general readership.}
 %		Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
@@
 I had assumed that the authors had integrated out pi in the prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite matchings”.
-(1) Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution?
-(2) Is the presented alternative approach justified?}
+Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution?
+Is the presented alternative approach justified?}
 
 \reply{%
-		We realize that the integration of $\pi$ out of the full conditionals is not necessary for our sampler. Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler.
+		We realize that the integration of $\pi$ out of the full conditionals is not necessary for our proposed sampler.
+Therefore, we have updated the paper with the full conditional distributions for a standard Gibbs sampler.
+See Section 3 and Supplement A.
 
@@
 
-\reply{To make the SEI algorithm more clear, we have revised the ordering of Sections 4, so that we have 1) hashing, 2) posterior inference, 3) chunkwise computation of the comparison matrix, and 4) SEI.
+\reply{To make the SEI algorithm more clear, we have revised the ordering of Sections 4, so that we have 1) hashing, 2) posterior inference, 3) batchwise computation of the comparison matrix, and 4) SEI.
 With this new ordering, and with the newly provided equations (15) through (19), we hope it is clearer that that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.

### 2023-08-01 ba_resubmit_template_files/response_to_reviewers3.tex

@@
 	%% Brian: I think it is important to tell the AE he is correct above and we are happy to add a note in later if he/she wants this. In this way, we can close out this question and avoid future ones. 
-	\textcolor{teal}{The reviewer is correct that the distinction does not matter between MCAR and MAR; we opt for using MCAR.
-We have included the following explanation of missingness in Section 2.
-We are happy to add a clarification regarding this point if the AE believes it would be helpful for the general readership.}
+	The reviewer is correct that the distinction does not matter between MCAR and MAR; we opt for using MCAR. We have included the following explanation of missingness in Section 2.
 %		Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here.
@@
 
-	\pointRaised{R}{%
+	\pointRaised{AE}{%
 	Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and think it was a good choice to move the full derivation to an appendix.
@@
 
-	\pointRaised{R}{%
+	\pointRaised{AE}{%
 Appendix 8.2: Thank you for including the details on integrating out pi from the full conditionals.
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are based on observed comparison vectors without consideration of Z (matches).
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Third paragraph of Section 4.1: The authors claim they are computing “sufficient statistics”. What exactly are these statistics sufficient for?
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute the appropriate weight to all records through the summary statistics…” What is meant by the term “weight”?
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Second paragraph of Section 4.2: “and delete those comparison vectors”. Which are “those” vectors?}
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations (13) and (14)?
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Section 4.2: I appreciate the practical advice about choosing S for the SEI method.
@@
 }
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Page 5, 2 sentences before equation (4): I believe the sum should be of $I(Z_j \leq n_1)$, not $I(Z_j \leq n_1 +1)$.}
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Equation (6a): The indices do not match the subscripts in the indicator function in each summand, or their standard meaning in table 1.}
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Generally, the authors seem to arbitrarily use upper and lower case z interchangeably in function definitions.}
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 Equations (8) and (9): Should the weights have superscript (s) (as the Zs do)?}
@@
 
-\pointRaised{R}{%
+\pointRaised{AE}{%
 	Gamma is in some places described as a set and in others as a matrix (particularly in sections 4.2 and 5.1).
@@
 
-\pointRaised{R}{% 
+\pointRaised{AE}{% 
 	Section 4.3, second paragraph, the definitions of the concatenated vectors $\alpha_0$ and $\beta_0$ should have final elements subscripted by $L_f$, where the sub-subscript is capitalized.}

### 2023-08-13 R/NLC_function2.R
vabl

@@ -13,6 +13,7 @@ BKSimple2 <- function(comparisons, m_prior, u_prior,
   # show_progress = set to false to show simulation progress
 
   fields <- length(comparisons[[4]])
+  levels <-
   n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
   indicators_raw <-comparisons[[1]]
 

### 2023-08-13 R/vabl_base.R

@@ -1,0 +1,121 @@
+vabl_base <- function(comparisons, threshold = 1e-5){
+  # Implements bipartite record linkage with BK Sampling Mechanism
+  #
+  # Arguments
+  # comparisons = list calculated from from BRL::compareRe,cords
+  # m.prior = prior distribution for m parameters
+  # u.prior= prior distribution for u parameters
+  # alpha = first parameter of prior for linkage probability
+  # beta = second parameter of prior for linkage probability
+  # S = number of Gibbs iterations
+  # burn = number of iterations to be discarded as burn-in
+  # show_progress = set to false to show simulation progress
+
+  fields <- length(comparisons[[4]])
+  levels <- comparisons[[4]]
+  n1 <- comparisons[[2]]; n2 <- comparisons[[3]]
+  indicators <- data.frame(comparisons[[1]])
+
+  parameter.indicators <- as.vector(unlist(sapply(1:fields, function(x){
+    rep(x, comparisons[[4]][x])
+  })))
+
+  ids <- expand.grid(1:n1, 1:n2)
+  candidates <- 1:n1
+
+  # Storage
+  phi_samps <- matrix(NA, nrow = n2, ncol = n_iter)
+  a_samps <- matrix(NA, nrow = dim(indicators)[2], ncol = n_iter)
+  b_samps <- matrix(NA, nrow = dim(indicators)[2], ncol = n_iter)
+  a_pi_samps <- vector(length = n_iter)
+  b_pi_samps <- vector(length = n_iter)
+
+  # Priors
+  alpha <- rep(1, length(parameter.indicators))
+  beta <- rep(1, length(parameter.indicators))
+  alpha_pi <- 1
+  beta_pi <- 1
+
+  # Initialize
+  a <- rep(1, length(parameter.indicators))
+  b <- rep(1, length(parameter.indicators))
+  a_pi <- 1
+  b_pi <- 1
+
+  # Computational savings
+  # totals <- indicators %>%
+  #   colSums()
+
+for(i in 1:n_iter){
+  a_split <- a %>%
+    split(., parameter.indicators) %>%
+    lapply(., digamma)
+  a_sum <- a %>%
+    split(., parameter.indicators) %>%
+    lapply(., sum) %>%
[…]

### 2023-08-13 R/vabl_hash.R

@@ -1,0 +1,111 @@
+hash_field <- function(L_f, k, Lf_vec){
+  level_seq <- seq_len(L_f)
+  as.numeric(level_seq > 0) * 2 ^ ((level_seq) + (as.numeric(k > 1)  * Lf_vec[k]))
+}
+
+vabl_hash <- function(comparisons, all_patterns = FALSE){
+
+  indicators <- comparisons[[1]]
+  N <- dim(indicators)[1]
+  n1 <- comparisons[[2]]
+  n2 <- comparisons[[3]]
+
+  levels <- comparisons[[4]]
+  fields <- seq_along(comparisons[[4]])
+  field_marker <- sapply(fields, function(x){
+    rep(x, comparisons[[4]][x])
+  }) %>%
+    unlist(.) %>%
+    as.vector(.)
+
+  ids <- expand.grid(1:n1, 1:n2)
+  rec1 <- ids[,1]
+  rec2 <- ids[,2]
+
+  Lf_vec<- (levels) %>%
+    c(0, .) %>%
+    cumsum()
+
+  hash_vals <- imap(comparisons[[4]], ~hash_field(.x, .y, Lf_vec)) %>%
+    unlist()
+
+  hash <- sweep(indicators, 2, hash_vals, "*") %>%
+    rowSums() + 1
+
+  if(all_patterns == TRUE){
+
+    unique_patterns <- GetPossiblePatternsSad_missing(levels)
+    unique_hashed <- sweep(unique_patterns, 2, hash_vals, "*") %>%
+      rowSums() + 1
+    P <- dim(unique_patterns)[1]
+    hash_id <- hash %>%
+      factor(levels = unique_hashed) %>%
+      as.integer() %>%
+      factor(levels = 1:P)
+
+  } else {
+
+    unique_hashed <- unique(hash)
+    P <- length(unique_hashed)
+
+    hash_id <- hash %>%
+      factor(levels = unique_hashed) %>%
+      as.integer() %>%
+      factor(levels = 1:P)
+    unique_patterns <- indicators[!duplicated(hash_id), ]
+  }
+
+  temp <- data.frame(indicators, rec1, rec2, hash_id)
+  pattern_counts <- temp %>%
+    group_by(hash_id, .drop = F) %>%
+    count() %>%
+    pull()
+
+  thing <- split(temp, rec2)
+
+  # This is the computational bottleneck
+  hash_to_file1 <- map(thing, ~ lapply(1:P, function(y){
+    which(.x$hash_id == y)
+  })
+  )
+
[…]

### 2023-08-13 SEI_sensitivity/vabl_accuracy.R

@@ -1,0 +1,167 @@
+library(RecordLinkage)
+library(dplyr)
+library(stringr)
+library(parlrdev)
+library(purrr)
+library(readr)
+library(BRL)
+
+taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+i = taskID
+set.seed(41)
+files <- list.files(path = "data/SimulationDataFiles", full.names = T)
+
+m_prior = 1
+u_prior = 1
+alpha = 1
+beta = 1
+S = 1000
+burn = 100
+show_progress = F
+fast = F
+R = NULL
+all_patterns = TRUE
+threshold = 1e-8
+resolve = T
+
+
+overlap_vals <- c(50, 250, 450)
+
+vabl_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
+
+fabl_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
+fabl_acc_samps_resolved <- matrix(NA, nrow = 3, ncol = 6)
+sad_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
+
+fabl_acc_samps2 <- matrix(NA, nrow = 3, ncol = 6)
+fabl_acc_samps_resolved2 <- matrix(NA, nrow = 3, ncol = 6)
+sad_acc_samps2 <- matrix(NA, nrow = 3, ncol = 6)
+
+for(j in seq_along(overlap_vals)){
+
+  overlap <- overlap_vals[j]
+
+  records <- read_csv(files[i], col_types = cols())
+  records$file <- rep(2:1, length.out = dim(records)[1])
+
+  records <- records %>%
+    janitor::clean_names() %>%
+    mutate(rec_id = as.numeric(str_extract(rec_id, "\\d{3}")) + 1)
+
+  n1 <- 500
+  n2 <- 500
+  #overlap <- n2/2
+
+  Ztrue <- n1 + 1:n2
+  Ztrue[1:overlap] <- 1:overlap
+
+  file1 <- records %>%
+    filter(file ==1,
+           rec_id <= n1) %>%
+    select(-rec_id) %>%
+    as.matrix(.) %>%
+    data.frame(.) %>%
+    mutate(occup = as.numeric(occup))
+
+  file2 <- records %>%
+    filter(file == 2,
+           rec_id %in% c(1:overlap, (n1 +1):(1000 - overlap))) %>%
+    select(-rec_id) %>%
+    as.matrix() %>%
+    data.frame(.) %>%
[…]

### 2023-08-13 R/vabl_efficient.R
vabl_bug

@@ -1,4 +1,4 @@
-vabl_efficient <- function(comparisons, threshold = 1e-8, tmax = 200){
+vabl_efficient <- function(hash, threshold = 1e-8, tmax = 200){
 
   check_every <- 10
 

### 2023-08-13 SEI_sensitivity/vabl_accuracy.R

@@ -77,13 +77,21 @@ for(j in seq_along(overlap_vals)){
   cd[[1]] <- apply(cd[[1]], 2, as.numeric)
 
 
-  hash <- vabl_hash(comparisons, all_patterns)
+  hash <- vabl_hash(cd, all_patterns)
   ptm <- proc.time()
-  out <- vabl_efficient(comparisons, threshold, tmax)
+  out <- vabl_efficient(cd, threshold, tmax)
   elapsed <- proc.time() - ptm
   result <- vabl_estimate_links(out, hash, resolve = T)
   eval <- GetEvaluations(result$Zhat, Ztrue, n1)
   vabl_acc_samps[j, ] <- c(eval, NA, elapsed[3], overlap)
+
+  #parlr method
+  ptm <- proc.time()
+  Zchain <- fabl_gibbs(cd, R = R)
+  elapsed <- proc.time() - ptm
+  Zhat <- LinkRecordsBK(Zchain[[1]], n1, 1, 1, 2, Inf)
+  eval <- GetEvaluations(Zhat[[1]], Ztrue, n1)
+  eval
 
   # Resolution Step
   Zhat_resolved <- ResolveConflicts(Zhat)

### 2024-01-26 ba_resubmit_template_files/response_to_reviewers4.tex
bk_edits

\documentclass[letterpaper, parskip]{scrartcl}

\usepackage{amsmath,amssymb,bm}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=1.4in]{geometry}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{tcolorbox}            % provides shaded box for reviewer comments
\usepackage{enumitem}             % use more compact lists
%\usepackage[parfill]{parskip}     % use skips between paragraphs and no indents
\usepackage{bibentry}             % provides \nobibliography command
\usepackage{xspace}

% Make itemize and enumerate environments more compact
\setlist[itemize]{noitemsep, topsep=0pt, parsep=0pt}
\setlist[enumerate]{noitemsep, topsep=0pt, parsep=0pt}

% -------- Define point raised command --------- %

\definecolor{darkpurple}{HTML}{500050}
\tcbuselibrary{breakable}       % allow tcolorbox to break across pages
\tcbuselibrary{skins}
\tcbset{%
	enhanced,%
	frame hidden,%
	sharp corners,%
	breakable,%
	parbox=false,%
	colback=lightgray!30,%
	enlarge left by=6pt,%
	width=\linewidth-6pt,%
	colback=gray!0,
	left=4pt,%
	right=0pt,%
	top=0pt,%
	bottom=0pt,%
	borderline west={2pt}{0pt}{lightgray},%
	%coltext=darkpurple,%
	before skip=14pt plus 2pt,%
	after skip=14pt plus 2pt}

%\newcommand{\pointRaised}[1]{%
%	\begin{tcolorbox}
%		\itshape #1
%	\end{tcolorbox}
%}
%
%\newcounter{responsectr}[section]     % counter for response resets at each section
%\newcommand{\reply}[2]{%
%	\refstepcounter{responsectr}%
%	\textbf{#1.\theresponsectr:} #2
%}

\newcommand{\pointRaised}[2]{%
	\textbf{#1.\theresponsectr:} #2
}

\newcounter{responsectr}[section]     % counter for response resets at each section
\newcommand{\reply}[1]{%
	\refstepcounter{responsectr}%
		\begin{tcolorbox}
			\itshape #1
		\end{tcolorbox}
}

\newcommand{\todo}{\textcolor{red}{[TODO]}\xspace}

\newcommand{\beka}[1]{{\color{teal}}}

\begin{document}

	% ****************** TITLE ****************************************

	\title{Response to Reviewer Comments}
	\maketitle

We thank the Editor and Associate Editor for their assessment and constructive feedback.

%Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{Editor Comments}

\pointRaised{E}{For Section 4.2, my question starts with the definition of $\Gamma_{Z_{j^{(s+1)}, j}}.$ Is this quantity always defined for possible values of $Z$? That is, if $Z_{j}^{(s+1)} = n_A + j$, what is the meaning of $\Gamma_{n_A+j, j}$?}

\reply{The reviewer is correct that $\Gamma_{n_A+j, j}$ is not defined under this notation. We had attempted to remove all instances in which this notation might be implied, but it seems we did not catch them all. We thank the editor for their thoughtful suggestions in this regard.
}

\pointRaised{E}{This notation first arises in the definition of $n_p(Z)$, which I think could easily be changed to something along the lines of $n_p(Z) = \sum_{j:Z_j \leq n_A} I(\gamma_{Z_j, j} = h_p)$.
}

\reply{Thank you for this suggestion. We have adopted this change.
}

\pointRaised{E}{My bigger concern comes for equation (18), where I am not entirely sure what probability is being computed, or how this sampling procedure combined with equation (19) would result in a “non-match” where $Z_j = n_A + j$. I believe that the idea is to partition the possible values of $Z_j:\left( \{r_{p_j}\}, n_A + j\right)$. Then, in equation (18), the left hand side is intended to be the probability that Z falls into each part of the partition. But, in the revision, I didn’t see where the last “non-match” part of the partition is defined, and so there is no real way to select the non-match from equation (19).
}

\reply{The editor's understanding of the sampler is correct. The editor also correctly identified an error in equation (19), such that there was no real way to select the non-match option. We thank the editor for the recommended change in notation, and have adopted the recommendation. Equations (18) and (19) now read:

	\begin{align*}
		p\left( Z_j^{(s+1)} \in r \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
		\begin{cases} 
			\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & r = r_{p_j}; \\
			1- \pi^{(s+1)} , &   r = n_A + j. \\
		\end{cases}
	\end{align*}
and

\begin{align*}
	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{N_{p_j}}, & r = r_{p_j} \text{ and } q \in r; \\
		1, & r = n_A + j. \\
	\end{cases}
\end{align*} 
}

\pointRaised{E}{Also, it’s not clear to me as currently written that the authors are equating the patterns with that partition so that by definition a “true” match that does not appear in A cannot have a pattern that appears in $\mathcal{P}$, but did not happen to be observed (because the match wasn’t captured in $A$).
}

\reply{We did not quite understand this part of the comment. If record $B_j$ has no match in $A$, then $Z_j = n_A + j$. If there is no matching record, there is no comparison vector (or pattern) for such a non-existent record.

Despite the confusion about this specific sentence, the proposed change in notation more explicitly shows how the sampler handles non-matches, which we think is the editors primary concern. 
}

\pointRaised{E}{Page 9, just below equation (17): Should the reference to equation (9) should be updated to equation (13)?}

\reply{The reference to equation (9) was meant as a comparison with equations (18) and (19), not equation (17). We have revised the paragraph to make this more clear. 
}

\pointRaised{E}{Supplement A, just below equation (3): Should the summation be over index $j$?
}

\reply{Yes. We have made the correction. 
}

\pointRaised{E}{Supplement A, equation (12): I believe that this intermediate result is not described correctly. Because $c_j$ is a function of $\gamma_{ij}$, once the function is divided by $c_j$, the quantity is no longer proportional to the probability. Rather, it is proportional to a likelihood-like function. Because equation (13) is conditional on $\gamma$, this distinction is not important for the end result.
}

\reply{Since $\Gamma_{.j}$ is the data, the expression $p(\Gamma_{.j} | -)$ is intended to be interpreted as a likelihood. To avoid this confusion, and to be consistent with notation in equation (1) of the Supplement, and equations (15) and (17) in the main paper, we have revised equations (8) - (12) of the supplement to use $\mathcal{L}(Z_j|-)$ instead of $p(\Gamma_{.j}|-)$.

{
\color{red}{I added the changed notation in red. However, it relies on fairly ugly (and I think non-standard) uses of likelihood notation (for example, $\mathcal{L}(Z_j = q | -)$). In contrast, Jody's paper stayed with $p(\Gamma_{.j} | -)$ notation. Should we make the switch, or stand by the old notation? The reviewer acknowledged that it practically doesn't make a difference, so I'm perhaps inclined to stick with the version we submitted?
}
}
}

	\clearpage

	\bigskip

	\bibliographystyle{jasa}
	\bibliography{biblio}

\end{document}

### 2024-01-26 ba_resubmit_template_files/main4.tex
bk_edits

--- 2023-07-31 ba_resubmit_template_files/main.tex
@@
 % put all the external documents here!
-\myexternaldocument{supplement}
+\myexternaldocument{supplement4}
 %% To run this you must run the compiler through pdflatexmk and have the latexmkrc file within the folder
@@
 The posterior updates for $\bm{m}$ and $\bm{u}$ depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$.
-Let $n_p(\bm{Z}) = \sum_{j=1}^{n_B} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
+Let $n_p(\bm{Z}) = \sum_{j: Z_j \leq n_A} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs.
 Then, conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
@@
 However, the vectorized summations can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-
+{
+\color{blue}{
 Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
@@
 \begin{align}
-	\label{eqn:gibbs1}
+	%\label{eqn:gibbs1}
 	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
@@
 \begin{align}
-	\label{eqn:gibbs2}
+	%\label{eqn:gibbs2}
 	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
@@
 We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
+}
+}
 
+{
+\color{red}{Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
+For each pattern $p$, let $w_p =m_p/u_p$.
+The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
+\begin{align}
+	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
+\end{align}
+With this form, we break this sampling step into two. We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
+\begin{align}
+	\label{eqn:gibbs1}
+	p\left( Z_j^{(s+1)} \in r \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+	\begin{cases} 
+		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & r = r_{p_j}; \\
+		1- \pi^{(s+1)} , &   r = n_A + j. \\
+	\end{cases}
+\end{align}
+Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
+\begin{align}
+	\label{eqn:gibbs2}
+	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
+		\frac{1}{N_{p_j}}, & r = r_{p_j} \text{ and } q \in r; \\
+		1, & r = n_A + j. \\
+	\end{cases}
+\end{align} 
+Note that sampling from (\ref{eqn:gibbs1}) has complexity $O(P+1)$ and sampling (uniformly) from (\ref{eqn:gibbs2}) has complexity $O(1)$, regardless of the size of $r_{p_j}$.
+In contrast, sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) has complexity $O(n_A)$, because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities.
+
+These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
+We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.}
+}
 \begin{lemma}
@@
 We collect these memory-reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.
-
+{
+\color{blue}{
 As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
 We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
+}
+}
+{
+\color{red}{As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), and (\ref{eqn:gibbs1}), all original record pairs are still accounted for through $\mathcal{N}$.
+We must revise (\ref{eqn:gibbs2}) to account for the reduced number of labels, so instead use
+\begin{align}
+	\label{eqn:gibbs2-SEI}
+	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
+		\frac{1}{|r|}, & r = \text{SEI}(r_{p_j}) \text{ and} \\
+		& q \in r; \\
+		1, & r = n_A + j \\
+	\end{cases}
+\end{align}
+in the second step of the split sampler.
+In this way, we can conduct all posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
+We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
+}
+}
 
@@
 		\begin{tabular}{llllll}
-
 			\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
@@
 
+\textbf{Acknowledgements}: We thank the reviewers for extensive comments and suggestions that immensely improved the quality
+of the article.
+
 \begin{supplement}

### 2024-01-26 ba_resubmit_template_files/supplement4.tex
bk_edits

--- 2023-07-31 ba_resubmit_template_files/supplement.tex
@@
 % put all the external documents here!
-\myexternaldocument{main}
+\myexternaldocument{main4}
 %% To run this you must run the compiler through pdflatexmk and have the latexmkrc file within the folder
@@
 
-	Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{i=1}^{n_B}I(Z_j \leq n_A)$ encoded by $\bm{Z}$ and hyperparameters.
+	Since $\pi$ encodes the rate of matching across the two data files, the full conditional $p(\pi|\gamma, \bm{Z}, \bm{m}, \bm{u}, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B}I(Z_j \leq n_A)$ encoded by $\bm{Z}$ and hyperparameters.
 We have the full conditional
@@
 	\end{align}
-
+{
+\color{blue}{
 Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
@@
 \end{align}
+}
+}
+
+{
+\color{red}{Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+	\begin{align}
+		\mathcal{L}(Z_j = n_A + j| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+	\end{align}
+	When $Z_j = q$ for some $q\leq n_A$, we have
+	\begin{align}
+		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+	\end{align}
+	We multiply and divide by the $u$ parameters for the matching record pair to obtain
+	\begin{align}
+		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+		&= w_{qj}  c_j.
+	\end{align}
+	We can divide the result of each case by $c_j$ to get
+	\begin{align}
+		\mathcal{L}(Z_j | \bm{m}, \bm{u}, \pi, \Gamma_{.j}) \propto \begin{cases} 
+			w_{qj}, & q \leq n_A; \\
+			1, &  q  = n_A + j.
+		\end{cases}
+	\end{align}
+	Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
+	\begin{align}
+		\label{eqn:z_full_conditional2}
+		p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+		\begin{cases} 
+			\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
+			1 - \pi^{(s+1)}, & q  = n_A + j.
+		\end{cases}
+	\end{align}
+}
+}
 

### 2024-01-26 R/CombineHash.R

@@ -59,5 +59,3 @@ CombineHash <- function(hash_list, dfA, dfB){
                    n2 = n2)
 
 }
-
-

### 2024-01-26 R/vabl_efficient.R

@@ -1,4 +1,4 @@
-vabl_efficient <- function(hash, threshold = 1e-8, tmax = 200, fixed_iterations = NULL,
+vabl_efficient <- function(hash, threshold = 1e-10, tmax = 200, fixed_iterations = NULL,
                            a = NULL, b = NULL, a_pi = NULL, b_pi = NULL){
 
   check_every <- 10
@@ -23,13 +23,13 @@ vabl_efficient <- function(hash, threshold = 1e-8, tmax = 200, fixed_iterations 
     a <- rep(1, length(field_marker))
   }
   if(is.null(b)){
-  b <- rep(1, length(field_marker))
+    b <- rep(1, length(field_marker))
   }
   if(is.null(a_pi)){
-  a_pi <- 1
+    a_pi <- 1
   }
   if(is.null(b_pi)){
-  b_pi <- 1
+    b_pi <- 1
   }
 
   t <- 1
@@ -133,13 +133,13 @@ vabl_efficient <- function(hash, threshold = 1e-8, tmax = 200, fixed_iterations 
 
 
     if(is.null(fixed_iterations)){
-    if(t %% check_every == 0){
-      ratio <- abs((elbo_seq[t] - elbo_seq[t - check_every +1])/
-                     elbo_seq[t - check_every +1])
-    }
-    if(ratio < threshold){
-      break
-    }
+      if(t %% check_every == 0){
+        ratio <- abs((elbo_seq[t] - elbo_seq[t - check_every +1])/
+                       elbo_seq[t - check_every +1])
+      }
+      if(ratio < threshold){
+        break
+      }
     }
 
     t <- t + 1
@@ -168,14 +168,3 @@ vabl_efficient <- function(hash, threshold = 1e-8, tmax = 200, fixed_iterations 
 
 }
 
-# exp_log_ratio <- a_chunk - b_chunk
-# weights <- ohe %>%
-#   sweep(., 2, exp_log_ratio, "*") %>%
-#   rowSums()
-
-#elbo[2] <- single * sum((1/C) * log(C))
-#elbo[3] <- total_nonmatch * (log(n1) - log(single)) -log(n1)*n2
-# elbo[4] <- (alpha_pi - a_pi + n2 - total_nonmatch) *
-#   (digamma(a_pi) - digamma(a_pi + b_pi))
-# elbo[5] <- (beta_pi - b_pi + total_nonmatch) *
-#   (digamma(b_pi) - digamma(a_pi + b_pi))

### 2024-01-26 R/vabl_hash.R

@@ -336,53 +336,9 @@ vabl_hash <- function(comparisons,
     unnest(data) %>%
     complete(rec2 = unique(hash_to_file_1$rec2)) %>%
     select(-N) %>%
-    # pull(rec2) %>%
-    # unique() %>%
-    # length() %>%
     setNames(c("rec2", "eligible_patterns", "eligible_records")) %>%
     group_split(rec2, .keep = F)
 
-
-#   thing2 <- thing %>%
-#     filter(is.na(N))
-#
-#
-#   flags <- lapply(thing2, function(x){
-#
-#   eligible_marker <- x$N == 1
-#   eligible_records <- x$data[eligible_marker] %>%
-#     unlist() %>%
-#     unname()
-#
-#   data.frame(eligible_patterns = x$hash_id[eligible_marker],
-#              eligible_records = eligible_records)
-#
-#   })
-# #
-#   flags <- counts %>%
-#     filter(N == 1) %>%
-#     unnest(data) %>%
-#     select(-N) %>%
-#     setNames(c("eligible_patterns", "rec2", "eligible_records"))
-#
-#   flags <- flags %>%
-#    group_split(rec2, .keep = F)
-
-
-  # NOTE: Compare memory costs of hash_to_file_1 and flags
-  # flags <- lapply(hash_to_file_1, function(x){
-  #
-  #   eligible_patterns <- sapply(x, function(y){
-  #     length(y) == 1
-  #   }) %>%
-  #     which()
-  #   eligible_records <- sapply(eligible_patterns, function(y){
-  #     x[[y]]
-  #   })
-  #
-  #   data.frame(eligible_patterns, eligible_records)
-  #
-  # })
 
   if(method == "vabl"){
     hash_to_file_1 <- NULL

### 2024-01-26 SEI_sensitivity/sadine_accuracy_plots.R

@@ -9,13 +9,13 @@ results_df <- results %>%
   select(-RR, -time) %>%
   filter(method %in% c("vabl", "fabl_resolved", "BRL")) %>%
   pivot_longer(cols = 1:3, names_to = "metric") %>%
-  group_by(method, metric, Error, overlap) %>%
+  group_by(method, metric, error, overlap) %>%
   summarize(avg = mean(value),
             lower = quantile(value, .025),
             upper = quantile(value, .975),
             .groups = "drop") %>%
   mutate(metric = factor(metric, c("recall", "precision", "f-Measure")),
-         Error = factor(Error, c("One Error", "Two Errors", "Three Errors")),
+         error = factor(error, c("One Error", "Two Errors", "Three Errors")),
          overlap = case_when(
            overlap == 50 ~ "10% Overlap",
            overlap == 250 ~ "50% Overlap",
@@ -33,7 +33,7 @@ acc_plot <- results_df %>%
              color = method)) +
   geom_pointrange(position = position_dodge2(width = .5),
                   size = .3) +
-  facet_grid(overlap ~ Error) +
+  facet_grid(overlap ~ error) +
   labs(x = NULL,
        y = NULL,
        color = "Method") +
@@ -52,42 +52,43 @@ acc_plot
 
 results_partial_df <- results %>%
   mutate(DR = 1 - RR) %>%
-  select(-RR, -Time_Elapsed) %>%
-  filter(Method %in% c("fabl_partial", "BRL_partial")) %>%
-  pivot_longer(cols = c(1, 2, 3, 7), names_to = "Metric") %>%
-  group_by(Method, Metric, Error, Overlap) %>%
+  select(-RR, -time) %>%
+  filter(method %in% c("fabl_partial", "BRL_partial", "vabl_partial")) %>%
+  pivot_longer(cols = c(1, 2, 3, 7), names_to = "metric") %>%
+  group_by(method, metric, error, overlap) %>%
   summarize(avg = mean(value),
             lower = quantile(value, .025),
             upper = quantile(value, .975),
             .groups = "drop") %>%
-  filter(Metric != "F-Measure") %>%
-  mutate(Error = factor(Error, c("One Error", "Two Errors", "Three Errors")),
-         Overlap = case_when(
-           Overlap == 50 ~ "10% Overlap",
-           Overlap == 250 ~ "50% Overlap",
-           Overlap == 450 ~ "90% Overlap"
+  filter(metric != "f-measure") %>%
[…]

### 2024-01-27 ba_resubmit_template_files/response_to_reviewers4-rcs-revision.tex
update.

--- 2024-01-26 ba_resubmit_template_files/response_to_reviewers4.tex
@@
 
+\section{Editor Comments}
+
+\pointRaised{E}{I received a detailed and careful report from the original associate editor who thinks the paper is now easier to read and understand.
+I still find it intricate to read, but the AE may have a point when saying that the notation can be a difficult challenge in this area of research.}
+\reply{}
+
+\pointRaised{E}{Anyhow, I have decided to follow the AE’s recommendation and would like the authors to carefully consider the comments in the AE report and make the suggested revisions.
+I find the issue raised by the AE regarding the formulas in Section 4.2 particularly concerning for two reasons.
+First, the article cannot be accepted if it contains technical errors and inaccuracies.
+Second, the formulas in question were used to code up the samplers used in the simulations and data analysis examples.
+If the formulas are incorrect (do not make sense, in fact, as the AE points out) how can the computational results be trusted?
+This is a grave concern that must be satisfactorily and convincingly addressed before the manuscript is accepted for publication.}
+\reply{}
+
+\pointRaised{E}{Please prepare a revision addressing all the AE’s comments, explain how the errors in the formulas affected the computations, and revise the computations as needed.}
+\reply{}
 %Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.
 
-\section{Editor Comments}
+\section{Associate Editor Comments}
+
+\textcolor{red}{Brian: change E to AE. Make sure to address the reviewer as AE (many typos below). Address the questions/comments of the Editor above.}
 
@@
 
+\textcolor{red}{Could you add in the new paragraph so it's more clear to the AE and for us to review please?}
+
 \pointRaised{E}{Supplement A, just below equation (3): Should the summation be over index $j$?
@@
 
+\textcolor{red}{I believe the AE is stating that things should be equality and no proportionality.
+If you look at what he/she wrote up this seems to be what he/she is stating to be the case.
+In essence, we cannot drop the constants.
+Please let me know if you agree regarding this Jerry as it's worded a bit strangely to me.}
+
 	\clearpage

### 2024-01-27 ba_resubmit_template_files/response_to_reviewers4-rcs-revision.tex
update.

@@
 
-\textcolor{red}{Brian: change E to AE. Make sure to address the reviewer as AE (many typos below). Address the questions/comments of the Editor above.}
+\textcolor{red}{Brian: change E to AE.
+Make sure to address the reviewer as AE (many typos below).
+Address the questions/comments of the Editor above.
+Please separate out major comments with minor comments as the AE did.}
 
-\pointRaised{E}{For Section 4.2, my question starts with the definition of $\Gamma_{Z_{j^{(s+1)}, j}}.$ Is this quantity always defined for possible values of $Z$?
+\pointRaised{AE}{For Section 4.2, my question starts with the definition of $\Gamma_{Z_{j^{(s+1)}, j}}.$ Is this quantity always defined for possible values of $Z$?
 That is, if $Z_{j}^{(s+1)} = n_A + j$, what is the meaning of $\Gamma_{n_A+j, j}$?}

### 2024-01-27 ba_resubmit_template_files/response_to_reviewers4.tex
bk eddits

@@
 
-\pointRaised{E}{For Section 4.2, my question starts with the definition of $\Gamma_{Z_{j^{(s+1)}, j}}.$ Is this quantity always defined for possible values of $Z$?
+%\pointRaised{E}{I received a detailed and careful report from the original associate editor who thinks the paper is now easier to read and understand.
+I still find it intricate to read, but the AE may have a point when saying that the notation can be a difficult challenge in this area of research.}
+%\reply{}
+
+\pointRaised{E}{Anyhow, I have decided to follow the AE’s recommendation and would like the authors to carefully consider the comments in the AE report and make the suggested revisions.
+I find the issue raised by the AE regarding the formulas in Section 4.2 particularly concerning for two reasons.
+First, the article cannot be accepted if it contains technical errors and inaccuracies.
+Second, the formulas in question were used to code up the samplers used in the simulations and data analysis examples.
+If the formulas are incorrect (do not make sense, in fact, as the AE points out) how can the computational results be trusted?
+This is a grave concern that must be satisfactorily and convincingly addressed before the manuscript is accepted for publication.}
+
+\reply{The Associate Editor made thoughtful suggestions for revising the notation in Section 4.2.
+We have adopted these suggestions as discussed below.
+We assure the Editor that formulas used in all computations is correct, and had already reflected the corrected notation in the previous submission.}
+
+\section{Associate Editor Comments}
+
+\pointRaised{AE}{For Section 4.2, my question starts with the definition of $\Gamma_{Z_{j^{(s+1)}, j}}.$ Is this quantity always defined for possible values of $Z$?
 That is, if $Z_{j}^{(s+1)} = n_A + j$, what is the meaning of $\Gamma_{n_A+j, j}$?}
 
-\reply{The reviewer is correct that $\Gamma_{n_A+j, j}$ is not defined under this notation.
+\reply{The Associate Editor is correct that $\Gamma_{n_A+j, j}$ is not defined under this notation.
 We had attempted to remove all instances in which this notation might be implied, but it seems we did not catch them all.
-We thank the editor for their thoughtful suggestions in this regard.
+We thank the AE for their thoughtful suggestions in this regard.
 }
 
-\pointRaised{E}{This notation first arises in the definition of $n_p(Z)$, which I think could easily be changed to something along the lines of $n_p(Z) = \sum_{j:Z_j \leq n_A} I(\gamma_{Z_j, j} = h_p)$.
+\pointRaised{AE}{This notation first arises in the definition of $n_p(Z)$, which I think could easily be changed to something along the lines of $n_p(Z) = \sum_{j:Z_j \leq n_A} I(\gamma_{Z_j, j} = h_p)$.
 }
@@
 
-\pointRaised{E}{My bigger concern comes for equation (18), where I am not entirely sure what probability is being computed, or how this sampling procedure combined with equation (19) would result in a “non-match” where $Z_j = n_A + j$.
+\pointRaised{AE}{My bigger concern comes for equation (18), where I am not entirely sure what probability is being computed, or how this sampling procedure combined with equation (19) would result in a “non-match” where $Z_j = n_A + j$.
 I believe that the idea is to partition the possible values of $Z_j:\left( \{r_{p_j}\}, n_A + j\right)$.
@@
 
-\reply{The editor's understanding of the sampler is correct.
-The editor also correctly identified an error in equation (19), such that there was no real way to select the non-match option.
-We thank the editor for the recommended change in notation, and have adopted the recommendation.
-Equations (18) and (19) now read:
+\reply{The AE's understanding of the sampler is correct.
+The AE also correctly identified an error in equation (19), such that there was no real way to select the non-match option.
+We thank the AE for the recommended change in notation, and have adopted the recommendation.
+The sampling equations are now provided as follows:
 
-	\begin{align*}
+Define $r$ as an arbitrary set of records. We have
+\begin{align}
+	\label{eqn:gibbs1}
 	p\left( Z_j^{(s+1)} \in r \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
@@
 		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & r = r_{p_j}; \\
-			1- \pi^{(s+1)} , &   r = n_A + j. \\
+		1- \pi^{(s+1)} , &  r = \{n_A + j\}. \\
 	\end{cases}
-	\end{align*}
-and
-
-\begin{align*}
+\end{align}
+Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
+\begin{align}
+	\label{eqn:gibbs2}
 	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
 		\frac{1}{N_{p_j}}, & r = r_{p_j} \text{ and } q \in r; \\
-		1, & r = n_A + j. \\
+		1, & r = \{n_A + j\} \text{ and} \\
+		& q = n_A + j. \\
 	\end{cases}
-\end{align*} 
+\end{align} 
 }
 
-\pointRaised{E}{Also, it’s not clear to me as currently written that the authors are equating the patterns with that partition so that by definition a “true” match that does not appear in A cannot have a pattern that appears in $\mathcal{P}$, but did not happen to be observed (because the match wasn’t captured in $A$).
+\pointRaised{AE}{Also, it’s not clear to me as currently written that the authors are equating the patterns with that partition so that by definition a “true” match that does not appear in A cannot have a pattern that appears in $\mathcal{P}$, but did not happen to be observed (because the match was not captured in $A$).
 }
@@
 
-Despite the confusion about this specific sentence, the proposed change in notation more explicitly shows how the sampler handles non-matches, which we think is the editors primary concern. 
+Despite the confusion about this specific sentence, the proposed change in notation more explicitly shows how the sampler handles non-matches, which we think is the AE's primary concern. 
 }
 
-\pointRaised{E}{Page 9, just below equation (17): Should the reference to equation (9) should be updated to equation (13)?}
+\pointRaised{AE}{Page 9, just below equation (17): Should the reference to equation (9) should be updated to equation (13)?}
 
-\reply{The reference to equation (9) was meant as a comparison with equations (18) and (19), not equation (17). We have revised the paragraph to make this more clear. 
+\reply{The reference to equation (9) was meant as a comparison with equations (18) and (19), not equation (17). We have revised the paragraph to make this more clear. This material now reads:
+
+Note that sampling from (18) has complexity $O(P+1)$ and sampling (uniformly) from (19) has complexity $O(1)$, regardless of the size of $r_{p_j}$.
+In contrast, sampling $Z_j$ from the full conditional provided in (9) has complexity $O(n_A)$, because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities.
 }
 
-\pointRaised{E}{Supplement A, just below equation (3): Should the summation be over index $j$?
+\pointRaised{AE}{Supplement A, just below equation (3): Should the summation be over index $j$?
 }
@@
 
-\pointRaised{E}{Supplement A, equation (12): I believe that this intermediate result is not described correctly.
+\pointRaised{AE}{Supplement A, equation (12): I believe that this intermediate result is not described correctly.
 Because $c_j$ is a function of $\gamma_{ij}$, once the function is divided by $c_j$, the quantity is no longer proportional to the probability.
@@
 
-\reply{Since $\Gamma_{.j}$ is the data, the expression $p(\Gamma_{.j} | -)$ is intended to be interpreted as a likelihood.
-To avoid this confusion, and to be consistent with notation in equation (1) of the Supplement, and equations (15) and (17) in the main paper, we have revised equations (8) - (12) of the supplement to use $\mathcal{L}(Z_j|-)$ instead of $p(\Gamma_{.j}|-)$.
-
-{
-\color{red}{I added the changed notation in red.
-However, it relies on fairly ugly (and I think non-standard) uses of likelihood notation (for example, $\mathcal{L}(Z_j = q | -)$).
-In contrast, Jody's paper stayed with $p(\Gamma_{.j} | -)$ notation.
-Should we make the switch, or stand by the old notation?
-The reviewer acknowledged that it practically doesn't make a difference, so I'm perhaps inclined to stick with the version we submitted?
-}
-}
-}
+\reply{We have removed the intermediate result in equation (12).}
 

### 2024-01-27 ba_resubmit_template_files/main4.tex
bk eddits

@@
 However, the vectorized summations can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
-{
-\color{blue}{
+%{
+%\color{blue}{
+%Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
+For each pattern $p$, let $w_p =m_p/u_p$.
+The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
+%\begin{align}
+%	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
+%\end{align}
+%Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large.
+This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
+To speed up computation, we use \eqref{likelihood_efficient_z} to break this sampling step into two.
+%
+%We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
+%\begin{align}
+%	%\label{eqn:gibbs1}
+%	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+%	\begin{cases} 
+%		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
+%		1- \pi^{(s+1)} , &   \text{otherwise}. \\
+%	\end{cases}
+%\end{align}
+%Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
+%\begin{align}
+%	%\label{eqn:gibbs2}
+%	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
+%		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
+%		0, & \text{otherwise.} \\
+%	\end{cases}
+%\end{align} 
+%
+%These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
+We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
+%}
+%}
+
 Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
@@
 \end{align}
-Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large.
-This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$.
-To speed up computation, we use \eqref{likelihood_efficient_z} to break this sampling step into two.
-
-We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
-\begin{align}
-	%\label{eqn:gibbs1}
-	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
-	\begin{cases} 
-		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
-		1- \pi^{(s+1)} , &   \text{otherwise}. \\
-	\end{cases}
-\end{align}
-Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
-\begin{align}
-	%\label{eqn:gibbs2}
-	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
-		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
-		0, & \text{otherwise.} \\
-	\end{cases}
-\end{align}
-
-These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
-We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
-}
-}
-
-{
-\color{red}{Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$.
-For each pattern $p$, let $w_p =m_p/u_p$.
-The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
-\begin{align}
-	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
-\end{align}
-With this form, we break this sampling step into two. We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
+With this form, we break this sampling step into two.
+We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link.
+Define $r$ as an arbitrary set of records.
+We have
 \begin{align}
@@
 		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & r = r_{p_j}; \\
-		1- \pi^{(s+1)} , &   r = n_A + j. \\
+		1- \pi^{(s+1)} , &  r = \{n_A + j\}. \\
 	\end{cases}
@@
 		\frac{1}{N_{p_j}}, & r = r_{p_j} \text{ and } q \in r; \\
-		1, & r = n_A + j. \\
+		1, & r = \{n_A + j\} \text{ and} \\
+		& q = n_A + j. \\
 	\end{cases}
@@
 These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups.
-We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.}
-}
+We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
+
 \begin{lemma}
@@
 	\begin{align}
-		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
+		r_{p_j} &= \{r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B} \}\text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
 		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab}.\label{N_{p_j}}	\end{align}
@@
 We collect these memory-reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$.
-{
-\color{blue}{
-As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
+%{
+%\color{blue}{
+%As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$.
 We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
-}
-}
-{
-\color{red}{As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), and (\ref{eqn:gibbs1}), all original record pairs are still accounted for through $\mathcal{N}$.
+%}
+%}
+
+As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), and (\ref{eqn:gibbs1}), all original record pairs are still accounted for through $\mathcal{N}$.
 We must revise (\ref{eqn:gibbs2}) to account for the reduced number of labels, so instead use
@@
 		& q \in r; \\
-		1, & r = n_A + j \\
+		1, & r = \{n_A + j\} \text{ and} \\
+		& q = n_A + j \\
 	\end{cases}
@@
 We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
-}
-}
 

### 2024-01-27 ba_resubmit_template_files/supplement4.tex
bk eddits

@@
 	\end{align}
-{
-\color{blue}{
+%{
+%\color{blue}{
+%Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+%\end{align}
+%When $Z_j = q$ for some $q\leq n_A$, we have
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+%\end{align}
+%We multiply and divide by the $u$ parameters for the matching record pair to obtain
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+%	&= w_{qj}  c_j.
+%\end{align}
+%We can divide the result of each case by $c_j$ to get
+%\begin{align}
+%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
+%		w_{qj}, & q \leq n_A; \\
+%		1, &  q  = n_A + j.
+%	\end{cases}
+%\end{align}
+%Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
+%\begin{align}
+%	\label{eqn:z_full_conditional2}
+%	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+%	\begin{cases} 
+%		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
+%		1 - \pi^{(s+1)}, & q  = n_A + j.
+%	\end{cases}
+%\end{align}
+%}
+%}
+
 Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
-\end{align}
-When $Z_j = q$ for some $q\leq n_A$, we have
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
-\end{align}
-We multiply and divide by the $u$ parameters for the matching record pair to obtain
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
-	&= w_{qj}  c_j.
-\end{align}
-We can divide the result of each case by $c_j$ to get
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
-		w_{qj}, & q \leq n_A; \\
-		1, &  q  = n_A + j.
-	\end{cases}
-\end{align}
-Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
-\begin{align}
-	\label{eqn:z_full_conditional2}
-	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
-	\begin{cases} 
-		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
-		1 - \pi^{(s+1)}, & q  = n_A + j.
-	\end{cases}
-\end{align}
-}
-}
-
-{
-\color{red}{Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
 	\begin{align}
@@
 	\end{align}
-	We can divide the result of each case by $c_j$ to get
-	\begin{align}
-		\mathcal{L}(Z_j | \bm{m}, \bm{u}, \pi, \Gamma_{.j}) \propto \begin{cases} 
-			w_{qj}, & q \leq n_A; \\
-			1, &  q  = n_A + j.
-		\end{cases}
-	\end{align}
-	Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
+%	We can divide the result of each case by $c_j$ to get
+%	\begin{align}
+%		\mathcal{L}(Z_j | \bm{m}, \bm{u}, \pi, \Gamma_{.j}) \propto \begin{cases} 
+%			w_{qj}, & q \leq n_A; \\
+%			1, &  q  = n_A + j.
+%		\end{cases}
+%	\end{align}
+	Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) and divide by $c_j$ to obtain the full conditional
 	\begin{align}
@@
 	\end{align}
-}
-}
 

### 2024-01-27 SEI_sensitivity/vabl_accuracy.R

@@ -1,176 +1,0 @@
-library(RecordLinkage)
-library(dplyr)
-library(stringr)
-library(parlrdev)
-library(purrr)
-library(readr)
-library(BRL)
-
-taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
-i = taskID
-set.seed(41)
-files <- list.files(path = "data/SimulationDataFiles", full.names = T)
-
-m_prior = 1
-u_prior = 1
-alpha = 1
-beta = 1
-S = 1000
-burn = 100
-show_progress = F
-fast = F
-R = NULL
-all_patterns = TRUE
-tmax= 200
-threshold = 1e-8
-resolve = T
-
-
-overlap_vals <- c(50, 250, 450)
-
-vabl_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
-vabl_acc_samps2 <- matrix(NA, nrow = 3, ncol = 6)
-
-fabl_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
-fabl_acc_samps_resolved <- matrix(NA, nrow = 3, ncol = 6)
-sad_acc_samps <- matrix(NA, nrow = 3, ncol = 6)
-
-fabl_acc_samps2 <- matrix(NA, nrow = 3, ncol = 6)
-fabl_acc_samps_resolved2 <- matrix(NA, nrow = 3, ncol = 6)
-sad_acc_samps2 <- matrix(NA, nrow = 3, ncol = 6)
-
-for(j in seq_along(overlap_vals)){
-
-  overlap <- overlap_vals[j]
-
-  records <- read_csv(files[i], col_types = cols())
-  records$file <- rep(2:1, length.out = dim(records)[1])
-
-  records <- records %>%
-    janitor::clean_names() %>%
-    mutate(rec_id = as.numeric(str_extract(rec_id, "\\d{3}")) + 1)
-
-  n1 <- 500
-  n2 <- 500
-  #overlap <- n2/2
-
-  Ztrue <- n1 + 1:n2
-  Ztrue[1:overlap] <- 1:overlap
-
-  file1 <- records %>%
-    filter(file ==1,
-           rec_id <= n1) %>%
-    select(-rec_id) %>%
-    as.matrix(.) %>%
-    data.frame(.) %>%
-    mutate(occup = as.numeric(occup))
-
-  file2 <- records %>%
-    filter(file == 2,
[…]

### 2024-02-14 ba_resubmit_template_files/response_to_reviewers4.tex
new stuff

@@
 \usepackage{xspace}
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	linkcolor=blue,
+	filecolor=magenta,      
+	urlcolor=blue,
+	pdftitle={Overleaf Example},
+	pdfpagemode=FullScreen,
+}
+
+\urlstyle{same}
 
@@
 
-We thank the Editor and Associate Editor for their assessment and constructive feedback.
+We thank the Editor and Associate Editor for the continued valuable feedback.
 
@@
 
-\reply{The Associate Editor made thoughtful suggestions for revising the notation in Section 4.2.
-We have adopted these suggestions as discussed below.
-We assure the Editor that formulas used in all computations is correct, and had already reflected the corrected notation in the previous submission.}
+\reply{We assure the Editor that the correct formulas we used in all samplers in the previous submission and match those suggested by the Associate Editor.
+This is evident in our code, which we have made available at \url{github.com/briankundinger/fabl.BA}.
+Simply, our notation in the previous version was imprecise.
+We have now corrected the notation using the Associate Editor's suggestion.}
 
@@
 \reply{The Associate Editor is correct that $\Gamma_{n_A+j, j}$ is not defined under this notation.
-We had attempted to remove all instances in which this notation might be implied, but it seems we did not catch them all.
+We had attempted to remove all instances in which this notation might be implied, but we did not catch them all.
 We thank the AE for their thoughtful suggestions in this regard.
@@
 \reply{The AE's understanding of the sampler is correct.
-The AE also correctly identified an error in equation (19), such that there was no real way to select the non-match option.
+The AE also correctly identified in the way we wrote equation (19), such that there was no real way to select the non-match option.
 We thank the AE for the recommended change in notation, and have adopted the recommendation.
@@
 
-\reply{We did not quite understand this part of the comment.
-If record $B_j$ has no match in $A$, then $Z_j = n_A + j$.
-If there is no matching record, there is no comparison vector (or pattern) for such a non-existent record.
-
-Despite the confusion about this specific sentence, the proposed change in notation more explicitly shows how the sampler handles non-matches, which we think is the AE's primary concern. 
+\reply{We believe that the revised notation addresses this concern.
 }
@@
 
-	\clearpage
-
-	\bigskip
-
-	\bibliographystyle{jasa}
-	\bibliography{biblio}
-
 \end{document}

### 2024-02-14 ba_resubmit_template_files/main4.tex
new stuff

@@
 This approach allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing.
-These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
+These contributions can allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy.
 In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
@@
 To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.
-With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.
+With the MCAR assumption, we can marginalize over the missing data and do all computation simply using the observed data.
 
@@
 This makes the sampler inherently sequential, which can be slow when working on linkage tasks with more than a few thousand records.
+
+%$\{j: I(Z_j \leq n_A); j = 1, \dots, n_B\}$
 
@@
 \end{align}
-With this form, we break this sampling step into two.
+Using (\ref{likelihood_efficient_z}), we break this sampling step into two.
 We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link.
@@
 As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), and (\ref{eqn:gibbs1}), all original record pairs are still accounted for through $\mathcal{N}$.
-We must revise (\ref{eqn:gibbs2}) to account for the reduced number of labels, so instead use
+We must revise (\ref{eqn:gibbs2}) to account for the reduced number of labels, so instead we use
 \begin{align}
@@
 For $n_B/2$ of these records, there is no match in $A$, so we simulate $n_A$ comparison vectors from the $\bm{u}$ probabilities.
-For the other $n_B/2$ of these records, there is one match in $A$, so we simulate $1$ comparison vector from the $\bm{m}$ probabilities, and $n_A - 1$ comparison vectors from the $\bm{u}$ probabilities.
+For the other $n_B/2$ of these records, there is one match in $A$, so we simulate one comparison vector from the $\bm{m}$ probabilities, and $n_A - 1$ comparison vectors from the $\bm{u}$ probabilities.
 We compare the run-time of \texttt{fabl} (with no SEI) against \texttt{BRL} as we increase $n_A$ and $n_B$.

### 2024-02-14 ba_resubmit_template_files/supplement4.tex
new stuff

--- 2023-07-31 ba_resubmit_template_files/supplement.tex
@@
 \end{align}
-We can divide the result of each case by $c_j$ to get
-\begin{align}
-	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
-		w_{qj}, & q \leq n_A; \\
-		1, &  q  = n_A + j.
-	\end{cases}
-\end{align}
-Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
+Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) in the main text, and divide by $c_j$ to obtain the full conditional
 \begin{align}
@@
 \end{align}
+
+%Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
+%	\begin{align}
+%		\mathcal{L}(Z_j = n_A + j| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
+%	\end{align}
+%	When $Z_j = q$ for some $q\leq n_A$, we have
+%	\begin{align}
+%		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
+%	\end{align}
+%	We multiply and divide by the $u$ parameters for the matching record pair to obtain
+%	\begin{align}
+%		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
+%		&= w_{qj}  c_j.
+%	\end{align}
+%	We can divide the result of each case by $c_j$ to get
+%	\begin{align}
+%		\mathcal{L}(Z_j | \bm{m}, \bm{u}, \pi, \Gamma_{.j}) \propto \begin{cases} 
+%			w_{qj}, & q \leq n_A; \\
+%			1, &  q  = n_A + j.
+%		\end{cases}
+%	\end{align}
+%	Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) and divide by $c_j$ to obtain the full conditional
+%	\begin{align}
+%		\label{eqn:z_full_conditional2}
+%		p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
+%		\begin{cases} 
+%			\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
+%			1 - \pi^{(s+1)}, & q  = n_A + j.
+%		\end{cases}
+%	\end{align}
 
